### 2019-11-01 latex/acrobot.tex
Added LaTeX files from Overleaf, to be fixed in Vim later.

%/===============/%
\section{Energy Injection for the Acrobot}

### 2019-11-01 latex/vnhcs.tex
Added LaTeX files from Overleaf, to be fixed in Vim later.

%/===============/%
\section{Virtual Nonholonomic Constraints}

%/----- Motivation -----/%
\subsection{Motivation}
TODO: Why do we bother with Hamiltonian? Why can't we do virtual nonholonomic constraints in Lagrangian?

%/----- Lagrange vs Hamilton -----/%
\subsection{A Hamiltonian Approach to Virtual Nonholonomic Constraints}
Take an underactuated mechanical system 
\begin{equation}\label{eqn:eom_lagrangian}
M(q)\ddot{q} + C(q,\dot{q})\dot{q} + \nabla V(q) = B(q)\tau
\end{equation}
with Lagrangian
\begin{equation}\label{eqn:lagrangian}
\mathcal{L}(q,\dot{q}) = \frac{1}{2} \dot{q}^T M(q) \dot{q} - V(q)
\end{equation}
where \(q = (q_1,\ldots,q_n) \in \mathcal{Q}\) is the configuration vector, the inertia matrix \(M(q) = M^T(q)\) is positive definite for all \(q \in \mathcal{Q}\), \(V(q)\) is the potential of the system, \(B : \mathcal{Q} \rightarrow \mathbb{R}^{n \times m}\) is a \(C^1\) full rank control matrix, and \(\tau \in \mathbb{R}^{m}\) is the control input.

One can compute the Hamiltonian of this system by performing a Legendre transform on the Lagrangian. Define the conjugate of momenta for \(q\) by 
\[
p = \frac{\delta \mathcal{L}}{\delta \dot{q}} \in \mathbb{R}^n
\]
Then, the Legendre transform is performed by taking
\[
\mathcal{H}(q,p) = p^T \dot{q} - \mathcal{L}(q,\dot{q})
\]
It is easy to verify that the Hamiltonian for system (\ref{eqn:lagrangian}) is given by the total mechanical energy (\ref{eqn:hamiltonian}).
\begin{equation}\label{eqn:hamiltonian}
\mathcal{H}(q,p) = E(q,p) = \frac{1}{2} p^T M^{-1}(q) p + V(q)
\end{equation}

The equations of motion for the system in Hamiltonian coordinates is given by
\begin{align}\label{eqn:hamiltionian_eom}
\begin{split}
\dot{q} &= \frac{\delta \mathcal{H}}{\delta p} = M^{-1}(q) p \\
\dot{p} &= -\frac{\delta \mathcal{H}}{\delta q} + B \tau
\end{split}
\end{align}

Suppose the mechanical system has degree of underactuation one, so that coordinates of the system can be split into an unactuated component \(q_u \in [\mathbb{R}]_T, \, T \in \mathbb{R}_{>0}\) which is not influence by control, along with an actuated component \(q_a\); that is, suppose \(B\) is of the form \(B(q) = [0_m, B_1^T(q), \ldots, B_n^T(q)]^T, \, B_i^T(q) \in \mathbb{R}^{n-1}\) and \(\tau \in \mathbb{R}^{n-1}\). In this case, \(q = (q_u, q_a)^T\) and the equations of motion become
\begin{align}\label{eqn:unactuated_actuated_eom}
\begin{split}
\dot{q_u} &= e_1^T M^{-1}(q) p \\
\dot{p_u} &= -p^T\frac{\delta M}{\delta q_u} p - \delta_{q_u}V(q) \\
\dot{q_a} &= 
\begin{bmatrix}
0 \cdots 0 \\
I_{n-1} \\
\end{bmatrix} M^{-1}(q) p \\
\dot{p_a} &= -p^T\frac{\delta M}{\delta q_a} p - \nabla{q_a}V(q) + 
\begin{bmatrix}
B_1(q) \\
\vdots \\
B_n(q)
\end{bmatrix} \tau
\end{split}
\end{align}

Now we can begin to talk about Virtual Nonholonomic Constraints. In a similar fashion to what was defined for VHCs, let us first define the goal of these new virtual constraints.

\begin{defn}
A relation \(h \in C^2\left(\mathcal{Q}\times \mathbb{R}^n ; \mathbb{R}^k\right)\) with \(h(q,p) = 0\) is a \textbf{virtual nonholonomic constraint (VNHC) of order k} if there exists a feedback control \(\tau(q,p)\) which stabilizes the constraint manifold
\[
\Gamma = \left\{(q,p) | h(q,p) = 0, dh_q \dot{q} + dh_p \dot{p} = 0\right\}
\]
\end{defn}
Define the output of the system to be \(e = h(q,p)\). We would like to find \(\tau(q,p)\) which drives \(e\) to zero to stabilize our constraint manifold \(\Gamma\). To accomplish this, we will input-output linearize (\ref{eqn:unactuated_actuated_eom}) to find \(\ddot{e} = -k_p e - k_d \dot{e}\) with \(k_p, k_d \in \mathbb{R}_{> 0}\).

To characterize a certain class of VNHCs, let us make the following assumption.
\begin{assm}\label{assm:vnhc_is_on_qu_pu}
We assume our relation \(h\) is of the form \(h(q,p) = q_a - f(q_u,p_u)\) for some \(f \in C^2\left([\mathbb{R}]_T \times \mathbb{R} ; \mathbb{R}^{n - 1}\right)\).
\end{assm}

Now we solve for \(\tau\) by finding \(\ddot{e}\).
\begin{align*}
    e &= h(q,p) = qa - f(q_u,p_u)\\
    \Rightarrow \dot{e} &= \dot{q_a} - df_{q_u}\dot{q_u} -df_{p_u}\dot{p_u} \\
    &= [-df_{q_u} I_{n-1}]\dot{q} - df_{p_u} \dot{p_u} \\
    &= dh_q M^{-1}(q) p - df_{p_u}\left( -\frac{1}{2}p^T \frac{\delta M^{-1}(q)}{\delta q_u} p - \delta_{q_u}V(q) \right) \\
    &= dh_q M^{-1}(q) p + \frac{1}{2}df_{p_u} p^T \frac{\delta M^{-1}(q)}{\delta q_u} p + df_{p_u}\delta_{q_u}V(q)
\end{align*}
The control input \(\tau\) only appears in \(\dot{p_a}\) (see (\ref{eqn:unactuated_actuated_eom})). To simplify the analysis, terms in \(\ddot{e}\) which do not depend on \(\dot{p_a}\) explicitly are lumped together under the symbol \((*)\):
\begin{align*}
    \ddot{e} &= dh_q M^{-1}(q) \dot{p} + df_{p_u} p^T \frac{\delta M^{-1}(q)}{\delta q_u} \dot{p} + (*) \\
    &= (dh_q M^{-1}(q) + df_{p_u} p^T \frac{\delta M^{-1}(q)}{\delta q_u})B\tau + (*) \\
    &= (dh_q M^{-1}(q) + dh_{p_u} p^T \frac{\delta M^{-1}(q)}{\delta q_u})B\tau + (*)
\end{align*}

From the derivations above, one can solve for \(\tau\) iff the matrix on the left of \(\tau\) is full rank. Thus, for systems with degree of underactuation one we give the following definition.
\begin{defn}
A VNHC \(h(q,p) = 0\) of order \(n - 1\) is \textbf{regular} if \(dh_{p_a} = 0\), \(dh_{q_a} = (1 \ldots 1)^T\), and 
\[
\text{rank}\left\{ (dh_q M^{-1}(q) + dh_{p_u} p^T \frac{\delta M^{-1}(q)}{\delta q_u})B\right\} = n - 1
\]
everywhere on the constraint manifold \(\Gamma\). Equivalently, a VNHC \(h\) of order \(n - 1\) is regular if it satisfies Assumption \ref{assm:vnhc_is_on_qu_pu} and system (\ref{eqn:unactuated_actuated_eom}) with output \(e = h(q,p)\) is of relative degree \(\{2,2,\ldots,2\}\) everywhere on \(\Gamma\).
\end{defn}

In general, \(\dot{e}\) is a function of \(q_u\) and \(p = (p_u,p_a)^T\). Since the purpose of a regular VNHC is to fully parameterize \(\Gamma\) by \((q_u,p_u)\), it is essential that one can solve for \(p_a = p_a(q_u,p_u)\). Unfortunately this often cannot be done, since \(\dot{e}\) contains the quadratic term
\[
\frac{1}{2} df_{p_u} p^T \frac{\delta M^{-1}(q)}{\delta q_u} p 
\]
We can solve for \(p_a\) if this quadratic term does not exist.
\begin{assm}\label{assm:M_is_Mqa}
Assume \(\delta M(q) / \delta q_u = 0 \Leftrightarrow \delta M^{-1}(q) / \delta q_u = 0\)
\end{assm}

Under Assumption \ref{assm:M_is_Mqa}, we get that the rank condition for \(h(q,p)\) to be a regular VNHC reduces to \(\text{rank}\left(dh_q M^{-1} B\right) = n - 1\). This is the same rank condition as required for Virtual Holonomic Constraints.

Now we solve for \(p_a\) on the constraint manifold (when \(e = \dot{e} = 0\)):
\begin{align*}
    \dot{e} = dh_q M^{-1}(q)p + df_{p_u} \delta_{q_u}V(q) &= 0\\
    \Leftrightarrow dh_q M^{-1}(q)e_1 p_u + dh_q M^{-1}(q) \begin{bmatrix}
    0 & \cdots & 0 \\
    & I_{n-1} & \\
    \end{bmatrix} p_a &= -df_{p_u} \delta_{q_u}V(q) \\
\end{align*}
\begin{align*}
    \Leftrightarrow dh_q M^{-1}(q) \begin{bmatrix}
    0 & \cdots & 0 \\
    & I_{n-1} & \\
    \end{bmatrix} p_a = -\left(df_{p_u}\delta_{q_u}V(q) + dh_q M^{-1}(q)e_1 p_u\right)
\end{align*}
One can linearly solve for \(p_a\) if and only if the matrix in front of it is invertible.

This leads us to a natural definition.
\begin{defn}\label{defn:solvability}
A VNHC \(h(q,p)\) is \textbf{solvable} (NOTE: actionable? what's a good name?) if
\[
\text{rank}\left(dh_q M^{-1}(q) \begin{bmatrix}
    0 & \cdots & 0 \\
    & I_{n-1} & \\
    \end{bmatrix}\right) = n - 1
\]
\end{defn}

With this analysis and our new definition in hand, we can solve for the dynamics on the constraint manifold.

\begin{thm}\label{thm:equation_for_pa}
Suppose assumptions \ref{assm:vnhc_is_on_qu_pu} and \ref{assm:M_is_Mqa} hold. If a regular VNHC \(h(q,p) = q_a - f(q_u,p_u)\) is solvable, then the parameterization for \(p_a\) on the constraint manifold is given by
\begin{align*}
p_a &= -\left(dh_q M^{-1}(q) \begin{bmatrix}
    0 & \cdots & 0 \\
    & I_{n-1} & \\
    \end{bmatrix}\right)^{-1}\left( df_{p_u}\delta_{q_u}V(q) + dh_q M^{-1}(q)e_1 p_u\right) \\
    &=: g(q_u,p_u)
\end{align*}
and the constrained dynamics on \(\Gamma\) are given by
\begin{align*}
    \dot{q_u} &= e_1^T M^{-1}(q_a) \begin{bmatrix}
    p_u \\
    p_a
    \end{bmatrix}\mid_{q_a = f(q_u,p_u), p_a = g(q_u,p_u)} \\
    \dot{p_u} &= -\delta_{q_u} V(q_u,q_a) \mid_{q_a = f(q_u,p_u)}
\end{align*}
\end{thm}

Theorem \ref{thm:equation_for_pa} guarantees that, on \(\Gamma\), \(q_a\) is a parameterized completely by \((q_u,p_u)\). Hence, the zero-dynamics on \(\Gamma\) are always two-dimensional regardless of the original dimension \(n\).
%/===============/%

### 2019-11-01 latex/vnhcs.tex
Changed the bibliography style This new version changes the bibliography type to use a different package, and we now compile with lualatex. It helps with error and warning messages, as it actually shows where the message is coming from for proper errors (as opposed to always doing it in main)

@@
-%/===============/%
+%! TEX root = main.tex
+
+%/========== Virtual Nonholonomic Constraints ==========/%
 \section{Virtual Nonholonomic Constraints}
@@
 Take an underactuated mechanical system
+
 \begin{equation}\label{eqn:eom_lagrangian}
@@
 \[
-p = \frac{\delta \mathcal{L}}{\delta \dot{q}} \in \mathbb{R}^n
+p = \frac{\partial \mathcal{L}}{\partial \dot{q}} \in \mathbb{R}^n
 \]
@@
 \begin{split}
-\dot{q} &= \frac{\delta \mathcal{H}}{\delta p} = M^{-1}(q) p \\
-\dot{p} &= -\frac{\delta \mathcal{H}}{\delta q} + B \tau
+\dot{q} &= \frac{\partial \mathcal{H}}{\partial p} = M^{-1}(q) p \\
+\dot{p} &= -\frac{\partial \mathcal{H}}{\partial q} + B \tau
 \end{split}
@@
 \dot{q_u} &= e_1^T M^{-1}(q) p \\
-\dot{p_u} &= -p^T\frac{\delta M}{\delta q_u} p - \delta_{q_u}V(q) \\
+\dot{p_u} &= -p^T\frac{\partial M}{\partial q_u} p - \partial_{q_u}V(q) \\
 \dot{q_a} &= 
@@
 \end{bmatrix} M^{-1}(q) p \\
-\dot{p_a} &= -p^T\frac{\delta M}{\delta q_a} p - \nabla{q_a}V(q) + 
+\dot{p_a} &= -p^T\frac{\partial M}{\partial q_a} p - \nabla{q_a}V(q) + 
 \begin{bmatrix}
@@
     &= [-df_{q_u} I_{n-1}]\dot{q} - df_{p_u} \dot{p_u} \\
-    &= dh_q M^{-1}(q) p - df_{p_u}\left( -\frac{1}{2}p^T \frac{\delta M^{-1}(q)}{\delta q_u} p - \delta_{q_u}V(q) \right) \\
-    &= dh_q M^{-1}(q) p + \frac{1}{2}df_{p_u} p^T \frac{\delta M^{-1}(q)}{\delta q_u} p + df_{p_u}\delta_{q_u}V(q)
+    &= dh_q M^{-1}(q) p - df_{p_u}\left( -\frac{1}{2}p^T \frac{\partial M^{-1}(q)}{\partial q_u} p - \partial_{q_u}V(q) \right) \\
+    &= dh_q M^{-1}(q) p + \frac{1}{2}df_{p_u} p^T \frac{\partial M^{-1}(q)}{\partial q_u} p + df_{p_u}\partial_{q_u}V(q)
 \end{align*}
@@
 \begin{align*}
-    \ddot{e} &= dh_q M^{-1}(q) \dot{p} + df_{p_u} p^T \frac{\delta M^{-1}(q)}{\delta q_u} \dot{p} + (*) \\
-    &= (dh_q M^{-1}(q) + df_{p_u} p^T \frac{\delta M^{-1}(q)}{\delta q_u})B\tau + (*) \\
-    &= (dh_q M^{-1}(q) + dh_{p_u} p^T \frac{\delta M^{-1}(q)}{\delta q_u})B\tau + (*)
+    \ddot{e} &= dh_q M^{-1}(q) \dot{p} + df_{p_u} p^T \frac{\partial M^{-1}(q)}{\partial q_u} \dot{p} + (*) \\
+    &= (dh_q M^{-1}(q) + df_{p_u} p^T \frac{\partial M^{-1}(q)}{\partial q_u})B\tau + (*) \\
+    &= (dh_q M^{-1}(q) + dh_{p_u} p^T \frac{\partial M^{-1}(q)}{\partial q_u})B\tau + (*)
 \end{align*}
@@
 \[
-\text{rank}\left\{ (dh_q M^{-1}(q) + dh_{p_u} p^T \frac{\delta M^{-1}(q)}{\delta q_u})B\right\} = n - 1
+\text{rank}\left\{ (dh_q M^{-1}(q) + dh_{p_u} p^T \frac{\partial M^{-1}(q)}{\partial q_u})B\right\} = n - 1
 \]
@@
 \[
-\frac{1}{2} df_{p_u} p^T \frac{\delta M^{-1}(q)}{\delta q_u} p 
+\frac{1}{2} df_{p_u} p^T \frac{\partial M^{-1}(q)}{\partial q_u} p 
 \]
@@
 \begin{assm}\label{assm:M_is_Mqa}
-Assume \(\delta M(q) / \delta q_u = 0 \Leftrightarrow \delta M^{-1}(q) / \delta q_u = 0\)
+Assume \(\partial M(q) / \partial q_u = 0 \Leftrightarrow \partial M^{-1}(q) / \partial q_u = 0\)
 \end{assm}
@@
 \begin{align*}
-    \dot{e} = dh_q M^{-1}(q)p + df_{p_u} \delta_{q_u}V(q) &= 0\\
+    \dot{e} = dh_q M^{-1}(q)p + df_{p_u} \partial_{q_u}V(q) &= 0\\
     \Leftrightarrow dh_q M^{-1}(q)e_1 p_u + dh_q M^{-1}(q) \begin{bmatrix}
@@
     & I_{n-1} & \\
-    \end{bmatrix} p_a &= -df_{p_u} \delta_{q_u}V(q) \\
+    \end{bmatrix} p_a &= -df_{p_u} \partial_{q_u}V(q) \\
 \end{align*}
@@
     & I_{n-1} & \\
-    \end{bmatrix} p_a = -\left(df_{p_u}\delta_{q_u}V(q) + dh_q M^{-1}(q)e_1 p_u\right)
+    \end{bmatrix} p_a = -\left(df_{p_u}\partial_{q_u}V(q) + dh_q M^{-1}(q)e_1 p_u\right)
 \end{align*}
@@
     & I_{n-1} & \\
-    \end{bmatrix}\right)^{-1}\left( df_{p_u}\delta_{q_u}V(q) + dh_q M^{-1}(q)e_1 p_u\right) \\
+    \end{bmatrix}\right)^{-1}\left( df_{p_u}\partial_{q_u}V(q) + dh_q M^{-1}(q)e_1 p_u\right) \\
     &=: g(q_u,p_u)
@@
     \end{bmatrix}\mid_{q_a = f(q_u,p_u), p_a = g(q_u,p_u)} \\
-    \dot{p_u} &= -\delta_{q_u} V(q_u,q_a) \mid_{q_a = f(q_u,p_u)}
+    \dot{p_u} &= -\partial_{q_u} V(q_u,q_a) \mid_{q_a = f(q_u,p_u)}
 \end{align*}
@@
 Hence, the zero-dynamics on \(\Gamma\) are always two-dimensional regardless of the original dimension \(n\).
-%/===============/%
+%/========== /Virtual Nonholonomic Constraints ==========/%

### 2019-11-05 latex/acrobot.tex
Changed the bibliography style This new version changes the bibliography type to use a different package, and we now compile with lualatex. It helps with error and warning messages, as it actually shows where the message is coming from for proper errors (as opposed to always doing it in main)
Added modelines to all my tex files
Filled out the first draft of acrobot.tex

@@
-%/===============/%
+%! TEX root = main.tex
+
+%/========== Acrobot ==========/%
 \section{Energy Injection for the Acrobot}
+Suppose the acrobot is simple; 
+that is, assume that the acrobot's two links are of equal length \(l\)
+and of equal mass \(m\), with the mass concentrated at the tip of each rod.
+
+Let \(q = (q_u,q_a)\) where \(q_u\) is the shoulder angle and \(q_a\) is the hip angle.
+
+Then the acrobot has hamiltonian given by
+\[
+    \mathcal{H}(q,p) = \frac{1}{2} p^T M^{-1}(q) p + V(q)
+\]
+where
+\[
+    M^{-1}(q) = \frac{1}{m l^2 (2 - \cos^2(q_a))} 
+    \begin{bmatrix}
+        1              & -(1 + \cos(q_a)) \\
+        -(1 + \cos(q_a)) & (3 + 2\cos(q_a))
+    \end{bmatrix}
+\]
+and
+\[
+    V(q) = m g l (3 - 2 \cos(q_u) - \cos(q_u + q_a))
+\]
+
+Notice that the acrobot satisfies \(\partial M^{-1}(q)/ \partial q_u = 0\),
+which means it may be possible to enforce a regular solvable VNHC to inject energy
+into the acrobot.
+One such VNHC is
+\begin{align}\label{eqn:acrobot_vnhc}
+    \begin{split}
+    h(q,p) &= q_a - f(q_u,p_u) \\
+           &= q_a - \bar{q_a}\sin(\arctan(p_u,q_u)) \\
+           &= q_a - \frac{\bar{q_a}p_u}{\sqrt{q_u^2 + p_u^2}}
+    \end{split}
+\end{align}
+where \(\bar{q_a} \in ]0,\pi[\) is a parameter which dictates how much the 
+controlled link can swing back and forth.
+
+To compute the regularity and solvability conditions, first observe that since
+the acrobot has two coordinates and only \(p_a\) is actuatable, the control matrix
+\(B = [0, 1]^T\) is constant and full rank everywhere. This also implies that
+\(h\) is both regular and solvable if \(dh_q M^{-1}(q)B\) is full rank.
+
+Observe that
+\begin{align}
+    \begin{split}
+        dh_q M^{-1}(q)B &= \frac{1}{m l^2 (2 - \cos^2(q_a))}
+\left[ \frac{\bar{q_a}p_uq_u}{(q_u^2 + p_u^2)^{3/2}}, 1\right] 
+    \begin{bmatrix}
+        -(1 + \cos(q_a)) \\
+        (3 + 2\cos(q_a))
+    \end{bmatrix} \\
+                     &= \frac{ (3+2\cos(q_a))(q_u^2 + p_u^2)^{3/2} 
+                     - \bar{q_a}p_uq_u(1+cos(q_a))}
+                     {m l^2 (q_u^2+p_u^2)^{3/2} (2 - \cos^2(q_a))}
+    \end{split}
+\end{align}
+
+TODO: \textbf{Determine if this is full rank everywhere except at \((q_u,p_u) = (0,0)\)}
+
+Solving for \(p_a\) is straightforward. The computation is shown below.
+\begin{multline*}
+    p_a = \left(dh_q M^{-1}(q)B\right)^{-1}  \\
+        \left(\frac{m g l \bar{q_a}p_u q_u^2}{(q_u^2+p_u^2)^{3/2}} 
+            \left(2\sin(q_u) + sin(q_u + q_a) \right)
+        +
+            \frac{1}{m l^2 (2 - \cos^2(q_a))}
+            [ \frac{\bar{q_a}p_uq_u}{(q_u^2+p_u^2)^{3/2}}, 1] 
+            \begin{bmatrix}
+              1 \\
+              -(1 + \cos(q_a))
+          \end{bmatrix} \right)
+\end{multline*}
+\begin{align*}
+    &= -\frac
+    {
+        \frac{ m g l \bar{q_a}q_u^2(2\sin(q_u) + \sin(q_u+q_a))}{(q_u^2+p_u^2)^{3/2}}
+        + \frac{\bar{q_a} p_u^2 q_u}{m l^2 (q_u^2 + p_u^2)^{3/2} (2 -\cos^2(q_a))}
+        - \frac{(1+\cos(q_a)) p_u}{m l^2 (2 - \cos^2(q_a))}
+    }
+    {
+        \frac
+        {
+            (3 + 2\cos(q_a))(q_u^2+p_u^2)^{3/2} - \bar{q_a}(1 + \cos(q_a))q_up_u
+        }
+        {
+            m l^2 (q_u^2+p_u^2)^{3/2} (2 - \cos^2(q_a))
+        }
+    } \\
+    &= \frac
+    {
+        \frac
+        {
+            -m^2 g l^3 \bar{q_a}q_u^2(2\sin(q_u) + \sin(q_u+q_a))(2 - \cos^2(q_a))
+            - \bar{q_a}p_u^2 q_u 
+            + (1 + \cos(q_a))(q_u^2 + p_u^2)^{3/2} p_u
+        }
+        {
+            m l^2 (q_u^2+p_u^2)^{3/2} (2 - \cos^2(q_a))
+        }
+    }
+    {
+        \frac
+        {
+            (3 + 2\cos(q_a))(q_u^2+p_u^2)^{3/2} - \bar{q_a}(1 + \cos(q_a))q_up_u
+        }
+        {
+            m l^2 (q_u^2+p_u^2)^{3/2} (2 - \cos^2(q_a))
+        }
+    }
+\end{align*}
+
+\begin{equation*}
+    \Rightarrow p_a = \frac
+        {
+            (1 + \cos(q_a))(q_u^2 + p_u^2)^{3/2} p_u
+            -m^2 g l^3 \bar{q_a}q_u^2(2\sin(q_u) + \sin(q_u+q_a))(2 - \cos^2(q_a))
+            - \bar{q_a}q_up_u^2
+        }
+        {
+            (3 + 2\cos(q_a))(q_u^2+p_u^2)^{3/2} - \bar{q_a}(1 + \cos(q_a))q_up_u
+        }
+\end{equation*}
+
+With \(p_a\) in hand, it is now possible to solve for the zero dynamics under the constraint. To simplify the notation we define \(s_u := \sin(q_u)\),
+\(s_{ua} := \sin(q_u + q_a)\), 
+and \(c_a := \cos(q_a)\).
+\begin{align*}
+    \dot{q_u} &= \frac{1}{m l^2 (2 - c_a^2)} [ 1, -(1+c_a)]
+    \begin{bmatrix} p_u \\ p_a \end{bmatrix} \\
+    \dot{p_u} &= - m g l (2 s_u + s_{ua})
+\end{align*}
+Expanding out \(\dot{q_u}\) we get
+\begin{multline*}
+    \dot{q_u} = 
+    \frac{p_u}{m l^2 (2 - c_a^2)}\\
+    + \frac
+    {
+        m^2 g l^3 \bar{q_a}q_u^2(2s_u + s_{ua})(2 - c_a^2)(1 + c_a)
+        - (1+c_a)^2(q_u^2+p_u^2)^{3/2}p_u 
+        + \bar{q_a}(1+c_a)q_up_u^2
+    }
+    {
+        ml^2 (2-c_a^2) ((3 + 2c_a)(q_u^2+p_u^2)^{3/2} - \bar{q_a}(1 + c_a)q_up_u)
+    }
+\end{multline*}
+and, putting everything over one denominator,
+\begin{align*}
+    \dot{q_a} &= \frac
+    {
+        (3 + 2c_a)(q_u^2+p_u^2)^{3/2}p_u 
+        - (1+c_a)^2(q_u^2+p_u^2)^{3/2}p_u 
+        +m^2 g l^3 \bar{q_a}q_u^2(2s_u + s_{ua})(2 - c_a^2)(1 + c_a)
+    }
+    {
+        ml^2 (2-c_a^2) ((3 + 2c_a)(q_u^2+p_u^2)^{3/2} - \bar{q_a}(1 + c_a)q_up_u)
+    } \\
+    &= \frac
+    {
+        (2-c_a^2)\left((q_u^2+p_u^2)^{3/2}p_u 
+        +m^2 g l^3 \bar{q_a}q_u^2(2s_u + s_{ua})(1 + c_a)\right)
+    }
+    {
+        ml^2 (2-c_a^2) ((3 + 2c_a)(q_u^2+p_u^2)^{3/2} - \bar{q_a}(1 + c_a)q_up_u)
+    }
+\end{align*}
+Therefore, the zero-dynamics of the acrobot under the constraint 
+\(q_a = \sin(\arctan(p_u,q_u))\) are:
+\begin{align}\label{eqn:acrobot_eom_qp}
+\begin{split}
+    \dot{q_u} &= \frac
+    {
+        (q_u^2+p_u^2)^{3/2}p_u 
+        +m^2 g l^3 \bar{q_a}q_u^2(2s_u + s_{ua})(1 + c_a)
+    }
+    {
+        ml^2 \left((3 + 2c_a)(q_u^2+p_u^2)^{3/2} - \bar{q_a}(1 + c_a)q_up_u\right)
+    } \\
+    \dot{p_u} &= - m g l (2s_u + s_{ua})
+\end{split}
+\end{align}
+
+It may be convenient to represent these equations of motion  in polar coordinates.
+Letting \(r = \sqrt{q_u^2 + p_u^2}\) be the radius and 
+\(\theta = \arctan(p_u,q_u)\) be the angle in the \((q_u,p_u)\) plane and taking
+derivatives, it is easy to show that
+\begin{align}\label{eqn:dr_and_dt_general}
+\begin{split}
+    \dot{r} &= \sin(\theta)]
+    \begin{bmatrix} \dot{q_u}(r,\theta) \\ \dot{p_u}(r,\theta) \end{bmatrix} \\
+    \dot{\theta} &= \frac{1}{r}[-\sin(\theta), \cos(\theta)]
+    \begin{bmatrix} \dot{q_u}(r,\theta) \\ \dot{p_u}(r,\theta) \end{bmatrix}
+\end{split}
+\end{align}
+
+By expanding out (\ref{eqn:dr_and_dt_general}) then substituting \(q_u = r\cos(\theta)\)
+and \(p_u = r\sin(\theta)\)
+it is straightforward computation to find the equations of motion in cylindrical \((r,\theta)\)
+coordinates. To simplify the notation yet again, define
+\(s(g)_\theta := \sin(g(\theta))\) and \(c(g)_\theta := \cos(g(\theta))\) 
+for any function \(g: \mathbb{R} \rightarrow \mathbb{R}\). The final equations
+of motion in \((r,\theta)\) coordinates are
+\begin{align}\label{eqn:acrobot_eom_rt}
+\begin{split}
+    \dot{r} &= \frac
+    {
+        r^2 c_\theta s_\theta 
+        + m^2 g l^3 (2 s(rc)_\theta + s(rc + \bar{q_a}s)_\theta)
+        \left(
+            (1+c(\bar{q_a}s)_\theta)\bar{q_a}c_\theta 
+            - (3 + 2c(\bar{q_a}s)_\theta)s_\theta r
+        \right)
+    }
+    {
+        m l^2 \left((3+2c(\bar{q_a}s)_\theta)r 
+        - \bar{q_a}(1+ c(\bar{q_a}s)_\theta)c_\theta s_\theta\right)
+    }\\
+    \dot{\theta} &= -\frac
+    {
+        r s_\theta^2
+        + m^2 g l^3 (2 s(rc)_\theta + s(rc + \bar{q_a}s)_\theta)
+        (3 + 2c(\bar{q_a}s)_\theta)c_\theta
+    }
+    {
+        m l^2 \left((3+2c(\bar{q_a}s)_\theta)r 
+        - \bar{q_a}(1+ c(\bar{q_a}s)_\theta)c_\theta s_\theta\right)
+    }
+\end{split}
+\end{align}
+
+%/========== /Acrobot ==========/%
+% vim: set ts=4 sw=4 sts=0 et ffs=unix :

### 2019-11-05 latex/vlp.tex
Added an empty file for the variable length pendulum

%! TEX root = main.tex

%/========== Variable Length Pendulum ==========/%
\section{The Variable Length Pendulum}\label{sec:vlp}

%/========== /Variable Length Pendulum ==========/%
% vim: set ts=3 sw=3 sts=0 et ffs=unix :

### 2019-11-06 latex/vnhcs.tex
Added modelines to all my tex files
Filled out the first draft of acrobot.tex
Performed Line Wrapping
This cleans up some lines so I can iterate over them better

@@
 %/========== Virtual Nonholonomic Constraints ==========/%
-\section{Virtual Nonholonomic Constraints}
+\section{Virtual Nonholonomic Constraints}\label{sec:vnhcs}
 
@@
 \end{equation}
-where \(q = (q_1,\ldots,q_n) \in \mathcal{Q}\) is the configuration vector, the inertia matrix \(M(q) = M^T(q)\) is positive definite for all \(q \in \mathcal{Q}\), \(V(q)\) is the potential of the system, \(B : \mathcal{Q} \rightarrow \mathbb{R}^{n \times m}\) is a \(C^1\) full rank control matrix, and \(\tau \in \mathbb{R}^{m}\) is the control input.
+where \(q = (q_1,\ldots,q_n) \in \mathcal{Q}\) is the configuration vector, 
+the inertia matrix \(M(q) = M^T(q)\) is positive definite for all 
+\(q \in \mathcal{Q}\), \(V(q)\) is the potential of the system, 
+\(B : \mathcal{Q} \rightarrow \mathbb{R}^{n \times m}\) is a 
+\(C^1\) full rank control matrix, and \(\tau \in \mathbb{R}^{m}\) is the control 
+input.
 
@@
 \]
-It is easy to verify that the Hamiltonian for system (\ref{eqn:lagrangian}) is given by the total mechanical energy (\ref{eqn:hamiltonian}).
+It is easy to verify that the Hamiltonian for system (\ref{eqn:lagrangian}) is 
+given by the total mechanical energy (\ref{eqn:hamiltonian}).
 \begin{equation}\label{eqn:hamiltonian}
@@
 Suppose assumptions \ref{assm:vnhc_is_on_qu_pu} and \ref{assm:M_is_Mqa} hold.
-If a regular VNHC \(h(q,p) = q_a - f(q_u,p_u)\) is solvable, then the parameterization for \(p_a\) on the constraint manifold is given by
+If a regular VNHC \(h(q,p) = q_a - f(q_u,p_u)\) is solvable, then the
+parameterization for \(p_a\) on the constraint manifold is given by
 \begin{align*}
@@
 
-Theorem \ref{thm:equation_for_pa} guarantees that, on \(\Gamma\), \(q_a\) is a parameterized completely by \((q_u,p_u)\).
-Hence, the zero-dynamics on \(\Gamma\) are always two-dimensional regardless of the original dimension \(n\).
-%/========== /Virtual Nonholonomic Constraints ==========/%
+Theorem \ref{thm:equation_for_pa} guarantees that, on \(\Gamma\), \(q_a\) is a
+parameterized completely by \((q_u,p_u)\). Hence, the zero-dynamics on
+\(\Gamma\) are always two-dimensional regardless of the original dimension
+\(n\).
+%/========== /Virtual Nonholonomic Constraints ==========/% vim: set ts=4 sw=4
+%sts=0 et ffs=unix :

### 2019-11-06 latex/vhcs.tex
Added a VHCs tex file
Started writing up content about virtual holonomic constraints. We will
need to move the lagrangian stuff from vnhcs to here, as well as the
conversion from lagrangian to hamiltonian stuff.

%! TEX root = main.tex

%/========== Virtual Holonomic Constraints ==========/%
\section{Virtual Holonomic Constraints}\label{sec:vhcs}

\subsection{Lagrangian Framework Review}%
\label{sub:lagrangian_vhcs}

\begin{defn}\label{defn:vhc_order_k}
   A \textbf{virtual holonomic constraint (VHC) of order k} is a smooth relation 
   \( h : \mathcal{Q} \rightarrow \mathbb{R}^k\) where
   \( \text{rank}(dh_q) = k \, \forall q \in h^{-1}(0)\) and where there exists
   a smooth feedback controller \(\tau(q,\dot{q})\) such that the constraint
   manifold
   \[
      \Gamma = \left\{ (q,\dot{q}) \in T_q\mathcal{Q} \vert h(q) = 0, \, 
      dh_q \dot{q} = 0\right\}
   \]
   is positively invariant.
\end{defn}

\begin{defn}\label{defn:vhc_regular}
   A VHC is \textbf{regular} if system \textbf{TODO:
   reference to E-L mechanical system} with output \(e = h(q)\) has vector
   relative degree \(\{2,2,\ldots,2\}\).
   Equivalently, a VHC is regular if 
   \[
      dh_q D^{-1}(q)B(q)
   \]
   is nonsingular everywhere on \(\Gamma\).
\end{defn}

%/========== /Virtual Holonomic Constraints ==========/%
% vim: set tw=80 ts=3 sw=3 sts=0 et ffs=unix :

### 2019-11-07 latex/vhcs.tex
Starting to write VHC section
Taking the lagrangian framework review from vnhcs and moving it to vhcs,
as vhcs come first in the thesis

@@
 
-\subsection{Lagrangian Framework Review}%
-\label{sub:lagrangian_vhcs}
+Take an underactuated mechanical system with Lagrangian
+\begin{equation}\label{eqn:lagrangian}
+\mathcal{L}(q,\dot{q}) = \frac{1}{2} \dot{q}^T D(q) \dot{q} - P(q)
+\end{equation}
+where \(q = (q_1,\ldots,q_n) \in \mathcal{Q}\) is the configuration vector, 
+the inertia matrix \(D(q) = D^T(q)\) is positive definite for all 
+\(q \in \mathcal{Q}\), and \(P(q)\) is the potential of the system.
+
+The equations of motion for this system can be computed from 
+the Euler-Lagrange equations (\ref{eqn:el_eqns})
+\begin{equation}\label{eqn:el_eqns}
+   \frac{d}{dt} \left( \frac{\partial \mathcal{L}}{\partial \dot{q}}\right) -
+   \frac{\partial \mathcal{L}}{\partial q} = B(q)\tau
+\end{equation}
+which result in (\ref{eqn:eom_lagrangian}), where
+\(C : T\mathcal{Q} \rightarrow \mathbb{R}^{n \times n}\) represents the
+generalized coriolis and centrifugal forces,
+\(B : \mathcal{Q} \rightarrow \mathbb{R}^{n \times m}\) is a 
+\(C^1\) full rank control matrix, and \(\tau \in \mathbb{R}^{m}\) is the control 
+input. \textbf{TODO: citation for E-L equations}
+
+\begin{equation}\label{eqn:eom_lagrangian}
+D(q)\ddot{q} + C(q,\dot{q})\dot{q} + \nabla P(q) = B(q)\tau
+\end{equation}
+
+Suppose now that one wishes to enforce, through the control input \(\tau\), a
+constraint on the generalized coordinates which depends only on \(q\) and not on
+\(\dot{q}\). This type of constraint is called ``holonomic"; since the
+constraint is enforced by control (as opposed to being enforced by the mechanics
+of the system itself), we call the constraint ``virtual". This intuition leads
+to the following mathematical definition as originally observed in 
+\textbf{TODO: cite VHC definition paper}:
 
@@
 \end{defn}
+\textbf{TODO: cite VHC papers}
 

### 2019-11-07 latex/vnhcs.tex
Starting to write VHC section
Taking the lagrangian framework review from vnhcs and moving it to vhcs,
as vhcs come first in the thesis

@@
 \subsection{A Hamiltonian Approach to Virtual Nonholonomic Constraints}
-Take an underactuated mechanical system
-
-\begin{equation}\label{eqn:eom_lagrangian}
-M(q)\ddot{q} + C(q,\dot{q})\dot{q} + \nabla V(q) = B(q)\tau
-\end{equation}
-with Lagrangian
-\begin{equation}\label{eqn:lagrangian}
-\mathcal{L}(q,\dot{q}) = \frac{1}{2} \dot{q}^T M(q) \dot{q} - V(q)
-\end{equation}
-where \(q = (q_1,\ldots,q_n) \in \mathcal{Q}\) is the configuration vector, 
-the inertia matrix \(M(q) = M^T(q)\) is positive definite for all 
-\(q \in \mathcal{Q}\), \(V(q)\) is the potential of the system, 
-\(B : \mathcal{Q} \rightarrow \mathbb{R}^{n \times m}\) is a 
-\(C^1\) full rank control matrix, and \(\tau \in \mathbb{R}^{m}\) is the control 
-input.
-
-One can compute the Hamiltonian of this system by performing a Legendre transform on the Lagrangian. Define the conjugate of momenta for \(q\) by 
+One can compute the Hamiltonian of the Lagrangian (\ref{eqn:lagrangian}) by 
+performing a Legendre transform on the Lagrangian. 
+First, define the conjugate of momenta for \(q\) by 
 \[

### 2019-11-08 latex/vhcs.tex
Added more to VHCs
This commit adds everything up to the definition of the parameterization
s for VHCs. We still need dynamics of s and the lagrangian structure
before we can move on to the hamiltonian.

@@
 
-Suppose now that one wishes to enforce, through the control input \(\tau\), a
+Suppose one were to use the control input \(\tau\) to enforce a
 constraint on the generalized coordinates which depends only on \(q\) and not on
 \(\dot{q}\). This type of constraint is called ``holonomic"; since the 
-constraint is enforced by control (as opposed to being enforced by the mechanics
-of the system itself), we call the constraint ``virtual". This intuition leads
-to the following mathematical definition as originally observed in 
-\textbf{TODO: cite VHC definition paper}:
+constraint is enforced by control as opposed to being a feature of the mechanics
+of the system, it is a called a ``virtual" constraint. 
+This intuition leads to the following mathematical development, described in 
+\cite{vhcs_for_el_systems}:
 
@@
 
+Typically in the theory of VHCS, one focuses mostly on VHCs of order \((n-1)\).
+This is because \(h^{-1}(0) = \{q \in \mathcal{Q} \vert h(q) = 0\}\) is a
+\((n-k)\)-dimensional submanifold of \(\mathcal{Q}\); if a VHC is of order
+\((n-1)\), then (by the preimage theorem) once the system is on \(\Gamma\), the
+set \(h^{-1}(0)\) is a 1-dimensional curve with no self-intersections which can
+be parameterized by a single configuration variable. In other words, VHCs of
+order \((n-1)\) reduce the dimension of the system to a 2-dimensional state,
+parameterized by some new virtual configuration and its velocity.  If \(k <
+(n-1)\), it becomes much more difficult to analyse the system under constraint
+since it gains more degrees of freedom. Therefore it is assumed for the rest of
+this section that any VHC is of order \((n-1)\).
+
 \begin{defn}\label{defn:vhc_regular}
-   A VHC is \textbf{regular} if system \textbf{TODO:
-   reference to E-L mechanical system} with output \(e = h(q)\) has vector
-   relative degree \(\{2,2,\ldots,2\}\).
+   A VHC of order \((n-1)\) is \textbf{regular} if 
+   system (\ref{eqn:eom_lagrangian}) with output \(e = h(q)\) has vector
+   relative degree \(\{2,2,\ldots,2\}\) on \(\Gamma\).
    Equivalently, a VHC is regular if 
@@
 \end{defn}
-\textbf{TODO: cite VHC papers}
+
+The definition of regularity comes from the following: 
+let \(h\) be a VHC and let \(e = h(q)\) be the
+output of (\ref{eqn:eom_lagrangian}). If \(h\) is regular, then
+it is possible to input-output
+linearize the system to get \(\ddot{e} = -k_p e - k_d \dot{e}\). This
+in turn means \(\Gamma\) becomes control-invariant and asymptotically
+stabilizable.
+
+To show this, we repeat the derivation of
+\textcite{vhcs_for_el_systems,lagrangian_structure_reduced_dynamics_vhcs} and take
+derivatives of \(e\).
+
+\begin{align*}
+   \dot{e} &= dh_q \dot{q} \\
+   \ddot{e} &= dh_q \ddot{q} + \frac{d}{dt}\left(dh_q\right)\dot{q}\\
+   \Rightarrow \ddot{e} &= 
+      -dh_qD^{-1}(q)\left[C(q,\dot{q})\dot{q} + \nabla P(q) \right] 
+      + \begin{bmatrix}
+         \dot{q}^T \nabla^2 h_1(q) \dot{q} \\
+         \vdots \\
+         \dot{q}^T \nabla^2 h_{n-1}(q) \dot{q}
+      \end{bmatrix}
+      + dh_q D^{-1}(q)B\tau \\
+\end{align*}
+
+Let \(\mu(q,\dot{q})\) be given by 
+\begin{equation*}
+   \mu(q,\dot{q}) := -dh_qD^{-1}(q)\left[C(q,\dot{q})\dot{q} + \nabla P(q) \right] 
+      + \begin{bmatrix}
+         \dot{q}^T \nabla^2 h_1(q) \dot{q} \\
+         \vdots \\
+         \dot{q}^T \nabla^2 h_{n-1}(q) \dot{q}
+      \end{bmatrix}
+\end{equation*}
+Then the resulting error dynamics are
+\begin{equation}\label{eqn:vhc_error_dynamics}
+   \ddot{e} = \mu(q,\dot{q}) + dh_q D^{-1}B(q) \tau
+\end{equation}
+
+Since \(h\) is a regular VHC, it is possible to solve for the control input. 
+The controller
+\begin{equation*}
+   \tau(q,\dot{q}) = \left(dh_q D^{-1}B(q)\right)^{-1}\left[-k_p e - k_d \dot{e} -
+   \mu(q,\dot{q})\right]
+\end{equation*}
+will generate the linear error dynamics \(\ddot{e} = -k_p e - k_d\dot{e}\) 
+which make \(\Gamma\) asymptotically stable.
+
+After constructing this controller, a natural question to ask is how system
+(\ref{eqn:eom_lagrangian}) behaves once it has stabilized on the constraint
+manifold \(\Gamma\). For this, one must make the following assumption.
+
+\begin{assm}\label{assm:B_perp_exists}
+   There exists \(B^\perp(q) \in \mathbb{R}^{(n-1) \times n}\) such that
+   \[
+      B^\perp(q) B(q) = 0 \in \mathbb{R}^{(n-1)\times(n-1)} \, \forall q \in
+   \mathcal{Q}
+   \]
+\end{assm}
+
+Given some \(B^\perp\), one can left-multiply the equations of motion to get
+\begin{equation}\label{eqn:eom_lagrangian_with_B_perp}
+   B^\perp(q)D(q)\ddot{q} + B^\perp(q)\left(C(q,\dot{q})\dot{q} \nabla
+   P(q)\right) = 0
+\end{equation}
+
+Recall that, since \(h\) is a regular VHC of order \((n-1)\), the dynamics under
+constraint can be parameterized by a single configuration variable.  If the
+dynamics on \(\Gamma\) are non-periodic, then the curve \(h^{-1}(0)\) is
+diffeomorphic to \(\mathbb{R}\). In this case, define \(\Theta := \mathbb{R}\).
+Otherwise, if the dynamics on \(\Gamma\) are periodic with some period \(T > 0\), 
+then the curve \(h^{-1}(0)\) is diffeomorphic to the circle
+\(\mathbb{S}^1\). In this case, let \(\Theta := [\mathbb{R}]_T\).
+
+As in \cite{lagrangian_structure_reduced_dynamics_vhcs}, 
+now \(s \in \Theta\) be the virtual configuration which parameterizes
+\(h^{-1}(0)\) through the diffeomorphism 
+\begin{align*}
+   \sigma : \Theta &\rightarrow h^{-1}(0) \\
+   s &\mapsto q
+\end{align*}
+Likewise, the tangent bundle \(T\Theta\) is diffeomorphic to \(Gamma\) 
+through the mapping
+\begin{align*}
+   d\sigma : T\Theta &\rightarrow \Gamma \\
+   (s,\dot{s}) &\mapsto (q,\dot{q})
+\end{align*}
+
+Using this new virtual configuration, the state of system
+(\ref{eqn:eom_lagrangian}) on the constraint manifold
+is parameterized by \(q = \sigma(s)\) and \(\dot{q} = \sigma'(s)\dot{s}\).
+Inserting these into (\ref{eqn:eom_lagrangian_with_B_perp}) (and dropping the
+function notation on \(B\),\(D\),\(C\), and \(P\) for simplicity), one gets
+\begin{align*}
+   B^\perp D\left(\sigma'(s)\ddot{s} + \sigma"(s)\dot{s}^2 \right)
+   + B^\perp \left[ C \sigma'(s)\dot{s} + \nabla P \right] = 0
+\end{align*}
 

### 2019-11-08 latex/vhcs.tex
Finished VHC constrained dynamics
Next is hamiltonian

@@
 
+By how \(C(q,\dot{q})\) is defined, 
+one can collect the product \(B^\perp C \dot{q}\) in the following way.
+First, combine the Christoffel symbols  into a matrix
+\(Q_i : \mathcal{Q} \rightarrow \mathbb{R}^{n \times n}\)
+defined element-wise by
+\begin{equation*}
+   [Q_i]_{j,k}(q) = \frac{1}{2}\left(\partial_{q_k} [D]_{i,j}(q)
+      + \partial_{q_j} [D]_{i,k}(q) 
+      + \partial_{q_i} [D]_{k,j}(q)\right)
+\end{equation*}
+
+Then \(B^\perp C \dot{q}\) becomes
+\begin{equation*}
+   \sum \limits_{i=1}^n B_i^\perp \dot{q}^T Q_i
+\end{equation*}
+
+To simplify the notation, define 
+\begin{align*}
+   \Psi_1(s) &:= -\frac{B^\perp \nabla P}{B^\perp D \sigma'} \\
+   \Psi_2(s) &:= -\frac{B^\perp D \sigma" + \sum\limits_{i=1}^n B_i^\perp
+   {\sigma'}^T Q_i \sigma'}{B^\perp D \sigma'} 
+\end{align*}
+
+Hence, the constrained dynamics for any regular VHC of order \((n-1)\) are
+provided entirely by the single, second-order system 
+(\ref{eqn:lagrangian_constrained_dynamics}).
+\begin{equation}\label{eqn:lagrangian_constrained_dynamics}
+   \ddot{s} = \Psi_1(s) + \Psi_2(s)\dot{s}^2
+\end{equation}
+
 %/========== /Virtual Holonomic Constraints ==========/%

### 2019-11-12 latex/vhcs.tex
Finalized derivation of lagrangian for constrained dynamics

@@
 
+\subsection{Review: Lagrangian Virtual Holonomic Constraints}
 Take an underactuated mechanical system with Lagrangian
@@
 
+A natural question to ask is whether these constrained dynamics are also
+Lagrangian. In fact, for VHCs applied to mechanical systems, the dynamics are
+Euler-Lagrange. \cite{lagrangian_structure_reduced_dynamics_vhcs}
+
+To see this, suppose the constrained dynamics had Lagrangian given by
+\begin{equation*}
+   \mathcal{L}_{\Psi}(s,\dot{s}) = \frac{1}{2}
+   D_{\Psi}(s)\dot{s}^2 + V_{\Psi}(s)
+\end{equation*}
+where \(D_{\Psi} : \Theta \rightarrow \mathbb{R}_{>0}\) is the virtual mass
+and \(V_{\Psi} : \Theta \rightarrow \mathbb{R}\) is the virtual potential of the
+system. Then the equations of motion of this system are
+\begin{equation*}
+   D_{\Psi}'(s)\dot{s}^2 + D_{\Psi}(s)\ddot{s} - \frac{1}{2}D_{\Psi}'(s)\dot{s}
+   + V_{\Psi}'(s) = 0
+\end{equation*}
+\begin{equation}\label{eqn:lagrangian_constrained_dynamics_general}
+   \Rightarrow \ddot{s} = -\frac{V_{\Psi}'(s)}{D_{\Psi}(s)} -
+   \frac{D_{\Psi}'(s)}{2 D_{\Psi}(s)}
+\end{equation}
+
+Setting (\ref{eqn:lagrangian_constrained_dynamics}) and
+(\ref{eqn:lagrangian_constrained_dynamics_general}) equal to each other, one
+can compute the values of \(D_{\Psi}\) and \(V_{\Psi}\). First, one sees that
+\(V_{\Psi}(s)\) is defined by \(D_{\Psi}\) and \(\Psi_1\):
+\begin{align*}
+   \Psi_1(s) &= -\frac{V_{\Psi}'(s)}{D_{\Psi}(s)} \\
+   \Rightarrow V_{\Psi}'(s) &= -D_{\Psi}(s) \Psi_1(s) \\
+   \Rightarrow V_{\Psi}(s) &= - \int \limits_{0}^s
+   D_{\Psi}(\tau)\Psi_1(\tau)d\tau
+\end{align*}
+where \(D_{\Psi}\) is given by
+\begin{align*}
+   \Psi_2(s) &= -\frac{D_{\Psi}'(s)}{D_\Psi(s)} \\
+   \Rightarrow D_{\Psi}'(s) &= -D_{\Psi}(s) \Psi_2(s) \\
+   \Rightarrow D_{\Psi}(s) &= \exp\left[ 
+   -2\int \limits_{0}^s \Psi_2(\tau) d\tau \right]
+\end{align*}
+
+\subsection{Review: Hamiltonian Systems}
+\textbf{TODO: What is a Hamiltonian system and why does it matter?}
+
+One can compute the Hamiltonian of the Lagrangian (\ref{eqn:lagrangian}) by 
+performing a Legendre transform on the Lagrangian. 
+First, define the conjugate of momenta for \(q\) by 
+\begin{equation*}
+p = \frac{\partial \mathcal{L}}{\partial \dot{q}} \in \mathbb{R}^n
+\end{equation*}
+Then, the Legendre transform is performed by taking
+\begin{equation*}
+\mathcal{H}(q,p) = p^T \dot{q} - \mathcal{L}(q,\dot{q})
+\end{equation*}
+It is easy to verify that the Hamiltonian for system (\ref{eqn:lagrangian}) is 
+given by the total mechanical energy (\ref{eqn:hamiltonian}).
+\begin{equation}\label{eqn:hamiltonian}
+\mathcal{H}(q,p) = E(q,p) = \frac{1}{2} p^T M^{-1}(q) p + V(q)
+\end{equation}
+
+The equations of motion for the system in Hamiltonian coordinates is given by
+\begin{align}\label{eqn:hamiltionian_eom}
+\begin{split}
+\dot{q} &= \frac{\partial \mathcal{H}}{\partial p} = M^{-1}(q) p \\
+\dot{p} &= -\frac{\partial \mathcal{H}}{\partial q} + B \tau
+\end{split}
+\end{align}
+
+\textbf{Why bother looking at Hamiltonian systems? What is the intuition behind
+these equations of motion?}
+
+\subsection{Hamiltonian Virtual Nonholonomic Constraints}
+
 %/========== /Virtual Holonomic Constraints ==========/%

### 2019-11-12 latex/vnhcs.tex
Finalized derivation of lagrangian for constrained dynamics

@@
 TODO: Why do we bother with Hamiltonian? Why can't we do virtual nonholonomic constraints in Lagrangian?
-
-%/----- Lagrange vs Hamilton -----/%
-\subsection{A Hamiltonian Approach to Virtual Nonholonomic Constraints}
-One can compute the Hamiltonian of the Lagrangian (\ref{eqn:lagrangian}) by 
-performing a Legendre transform on the Lagrangian. 
-First, define the conjugate of momenta for \(q\) by 
-\[
-p = \frac{\partial \mathcal{L}}{\partial \dot{q}} \in \mathbb{R}^n
-\]
-Then, the Legendre transform is performed by taking
-\[
-\mathcal{H}(q,p) = p^T \dot{q} - \mathcal{L}(q,\dot{q})
-\]
-It is easy to verify that the Hamiltonian for system (\ref{eqn:lagrangian}) is 
-given by the total mechanical energy (\ref{eqn:hamiltonian}).
-\begin{equation}\label{eqn:hamiltonian}
-\mathcal{H}(q,p) = E(q,p) = \frac{1}{2} p^T M^{-1}(q) p + V(q)
-\end{equation}
-
-The equations of motion for the system in Hamiltonian coordinates is given by
-\begin{align}\label{eqn:hamiltionian_eom}
-\begin{split}
-\dot{q} &= \frac{\partial \mathcal{H}}{\partial p} = M^{-1}(q) p \\
-\dot{p} &= -\frac{\partial \mathcal{H}}{\partial q} + B \tau
-\end{split}
-\end{align}
 

### 2019-11-13 latex/vhcs.tex
Created minipage for D_psi and V_psi

@@
 To see this, suppose the constrained dynamics had Lagrangian given by
-\begin{equation*}
+\begin{equation}\label{eqn:vhcs_constrained_lagrangian}
    \mathcal{L}_{\Psi}(s,\dot{s}) = \frac{1}{2}
    D_{\Psi}(s)\dot{s}^2 + V_{\Psi}(s)
-\end{equation*}
+\end{equation}
 where \(D_{\Psi} : \Theta \rightarrow \mathbb{R}_{>0}\) is the virtual mass
@@
 (\ref{eqn:lagrangian_constrained_dynamics_general}) equal to each other, one
-can compute the values of \(D_{\Psi}\) and \(V_{\Psi}\). First, one sees that
-\(V_{\Psi}(s)\) is defined by \(D_{\Psi}\) and \(\Psi_1\):
+can compute the values of \(D_{\Psi}\) and \(V_{\Psi}\): 
+% Empty line here required for spacing
+
+\begin{minipage}{0.48\textwidth}
+\begin{align*}
+   \Psi_2(s) &= -\frac{D_{\Psi}'(s)}{D_\Psi(s)} \\
+   \Rightarrow D_{\Psi}'(s) &= -D_{\Psi}(s) \Psi_2(s) \\
+   \Rightarrow D_{\Psi}(s) &= \exp\left[ 
+   -2\int \limits_{0}^s \Psi_2(\tau) d\tau \right]
+\end{align*}
+\end{minipage}
+% No empty lines between these
+\begin{minipage}{0.48\textwidth}
 \begin{align*}
@@
 \end{align*}
-where \(D_{\Psi}\) is given by
-\begin{align*}
-   \Psi_2(s) &= -\frac{D_{\Psi}'(s)}{D_\Psi(s)} \\
-   \Rightarrow D_{\Psi}'(s) &= -D_{\Psi}(s) \Psi_2(s) \\
-   \Rightarrow D_{\Psi}(s) &= \exp\left[ 
-   -2\int \limits_{0}^s \Psi_2(\tau) d\tau \right]
-\end{align*}
+\end{minipage}
+
+Thus, the constrained dynamics for mechanical systems are Euler-Lagrange with
+Lagrangian (\ref{eqn:vhcs_constrained_lagrangian}).
 
@@
 \subsection{Hamiltonian Virtual Nonholonomic Constraints}
+While VHCs are still possible in the Hamiltonian framework, the assumptions
+required to make this work are slightly different. To see this, let us look back
+to the derivation which led to the error dynamics (\ref{eqn:vhc_error_dynamics})
+and attempt to reconstruct this in Hamiltonian coordinates.
 

### 2019-11-13 latex/vhcs.tex
Moved Hamiltonian review to VNHCs again

@@
 
-\subsection{Review: Hamiltonian Systems}
-\textbf{TODO: What is a Hamiltonian system and why does it matter?}
-
-One can compute the Hamiltonian of the Lagrangian (\ref{eqn:lagrangian}) by 
-performing a Legendre transform on the Lagrangian. 
-First, define the conjugate of momenta for \(q\) by 
-\begin{equation*}
-p = \frac{\partial \mathcal{L}}{\partial \dot{q}} \in \mathbb{R}^n
-\end{equation*}
-Then, the Legendre transform is performed by taking
-\begin{equation*}
-\mathcal{H}(q,p) = p^T \dot{q} - \mathcal{L}(q,\dot{q})
-\end{equation*}
-It is easy to verify that the Hamiltonian for system (\ref{eqn:lagrangian}) is 
-given by the total mechanical energy (\ref{eqn:hamiltonian}).
-\begin{equation}\label{eqn:hamiltonian}
-\mathcal{H}(q,p) = E(q,p) = \frac{1}{2} p^T M^{-1}(q) p + V(q)
-\end{equation}
-
-The equations of motion for the system in Hamiltonian coordinates is given by
-\begin{align}\label{eqn:hamiltionian_eom}
-\begin{split}
-\dot{q} &= \frac{\partial \mathcal{H}}{\partial p} = M^{-1}(q) p \\
-\dot{p} &= -\frac{\partial \mathcal{H}}{\partial q} + B \tau
-\end{split}
-\end{align}
-
-\textbf{Why bother looking at Hamiltonian systems? What is the intuition behind
-these equations of motion?}
-
-\subsection{Hamiltonian Virtual Nonholonomic Constraints}
-While VHCs are still possible in the Hamiltonian framework, the assumptions
-required to make this work are slightly different. To see this, let us look back
-to the derivation which led to the error dynamics (\ref{eqn:vhc_error_dynamics})
-and attempt to reconstruct this in Hamiltonian coordinates.
-
 %/========== /Virtual Holonomic Constraints ==========/%

### 2019-11-13 latex/vnhcs.tex
Moved Hamiltonian review to VNHCs again

@@
 \subsection{Motivation}
-TODO: Why do we bother with Hamiltonian? Why can't we do virtual nonholonomic constraints in Lagrangian?
+Suppose one wishes to set a constraint that depends on both the configuration
+and the velocity.
+TODO: Why do we bother with Hamiltonian? Why can't we do virtual nonholonomic 
+constraints in Lagrangian?
+
+\subsection{Review: Hamiltonian Systems}
+\textbf{TODO: What is a Hamiltonian system and why does it matter?}
+
+One can compute the Hamiltonian of the Lagrangian (\ref{eqn:lagrangian}) by 
+performing a Legendre transform on the Lagrangian. 
+First, define the conjugate of momenta for \(q\) by 
+\begin{equation*}
+p = \frac{\partial \mathcal{L}}{\partial \dot{q}} \in \mathbb{R}^n
+\end{equation*}
+Then, the Legendre transform is performed by taking
+\begin{equation*}
+\mathcal{H}(q,p) = p^T \dot{q} - \mathcal{L}(q,\dot{q})
+\end{equation*}
+It is easy to verify that the Hamiltonian for system (\ref{eqn:lagrangian}) is 
+given by the total mechanical energy (\ref{eqn:hamiltonian}).
+\begin{equation}\label{eqn:hamiltonian}
+\mathcal{H}(q,p) = E(q,p) = \frac{1}{2} p^T M^{-1}(q) p + V(q)
+\end{equation}
+
+The equations of motion for the system in Hamiltonian coordinates is given by
+\begin{align}\label{eqn:hamiltionian_eom}
+\begin{split}
+\dot{q} &= \frac{\partial \mathcal{H}}{\partial p} = M^{-1}(q) p \\
+\dot{p} &= -\frac{\partial \mathcal{H}}{\partial q} + B \tau
+\end{split}
+\end{align}
+
+\textbf{Why bother looking at Hamiltonian systems? What is the intuition behind
+these equations of motion?}
+
+\subsection{Hamiltonian Virtual Nonholonomic Constraints}
+While VHCs are still possible in the Hamiltonian framework, the assumptions
+required to make this work are slightly different. To see this, let us look back
+to the derivation which led to the error dynamics (\ref{eqn:vhc_error_dynamics})
+and attempt to reconstruct this in Hamiltonian coordinates.
 
@@
 \(n\).
-%/========== /Virtual Nonholonomic Constraints ==========/% vim: set ts=4 sw=4
-%sts=0 et ffs=unix :
+
+%---------- Hamiltonian VHCS ---------%
+\subsection{Hamiltonian Virtual Holonomic Constraints}
+\textbf{TODO: Why can we not do the standard VHC approach for Hamiltonian? 
+Show how to use the above to make it work}
+
+%/========== /Virtual Nonholonomic Constraints ==========/% 
+% vim: set tw=80 ts=4 sw=4 sts=0 et ffs=unix :

### 2019-11-14 latex/vhcs.tex
Adding more TODOs and renaming subsections
Wrote the motivation of VHCs
Also wrote a bunch more stuff for VNHCs, providing more detail for the
Hamiltonian-from-Lagrangian section.

@@
 
-\subsection{Review: Lagrangian Virtual Holonomic Constraints}
+%---------- Motivation ----------%
+\subsection{Motivation}
+Most classic control architectures achieve one of two goals: equilibrium
+stabilization, or tracking of a signal over time (which is known as
+trajectory tracking). In fact, most trajectory-tracking controllers achieve
+their results by changing a system's coordinates so that the new system is at
+equilibrium if it is on the trajectory; so in fact, most control architectures
+in the past have focused on equilibrium stabilization. \textbf{TODO: citation to
+support this claim?}
+
+While this is a perfectly valid approach for many systems, it is not sufficient
+to provide robust, intricate motion on anything but the simplest of control
+systems \textbf{TODO: citation? Can I suppport this?}. Take the example of an
+autonomous vehicle driving on a road. If the desired path is a time trajectory 
+and the robot experiences a disturbance, it will need to ``catch up" to the
+trajectory it was following. This is a problem, because the vehicle will take
+shortcuts that may take it off its original trajectory, which can cause
+collisions with pedestrians, buildings, or simply cause the robot to speed down
+the road.
+
+\textbf{TODO figure with an example of this robot trajectory tracking issue}
+
+For this reason, mobile roboticists use \textbf{path tracking}:
+they define a constraint on the position of the vehicle, which does not depend on
+time. Rather than catching up to where the robot is ``supposed to be" in time, a
+path-tracking controller drives the robot to the nearest point on the path. This
+is a much more robust control architecture, as it forces the robot to stay
+within the confines of the path, so that it does not accidentally kill an
+unsuspecting pedestrian.
+
+\textbf{TODO: figure showing path tracking}
+
+Take now the example of a walking robot. If the legs are given a trajectory in
+time to follow, then the second the robot stumbles or slips it will try to
+``catch up" to the trajectory and the robot will fall over \textbf{TODO: cite
+grizzle. Does he make this argument too?}.  To fix this issue, one requires a
+path-tracking controller for general robotic systems. An engineer should be able
+to provide a constraint on the robot's configuration which does not depend on
+time, and the controller should keep the robot on that constraint. This is
+precisely what Virtual Holonomic Constraints attempt to accomplish.
+
+%---------- Define VHCs ----------%
+\subsection{Lagrangian Virtual Holonomic Constraints}
 Take an underactuated mechanical system with Lagrangian
@@
 where \(q = (q_1,\ldots,q_n) \in \mathcal{Q}\) is the configuration vector, 
-the inertia matrix \(D(q) = D^T(q)\) is positive definite for all 
-\(q \in \mathcal{Q}\), and \(P(q)\) is the potential of the system.
+the inertia matrix \(D(q) = D^T(q) \mathbb{R}^{n \times n}\) 
+is positive definite for all 
+\(q \in \mathcal{Q}\), and \(P(q) \in \mathbb{R}\) is the potential of the
+system.
 
-The equations of motion for this system can be computed from 
-the Euler-Lagrange equations (\ref{eqn:el_eqns})
+The Euler-Lagrange equations 
 \begin{equation}\label{eqn:el_eqns}
@@
 \end{equation}
-which result in (\ref{eqn:eom_lagrangian}), where
+produce the equations of motion (\ref{eqn:eom_lagrangian}), where
 \(C : T\mathcal{Q} \rightarrow \mathbb{R}^{n \times n}\) represents the
@@
 Suppose one were to use the control input \(\tau\) to enforce a
-constraint on the generalized coordinates which depends only on \(q\) and not on
-\(\dot{q}\). This type of constraint is called ``holonomic"; since the 
-constraint is enforced by control as opposed to being a feature of the mechanics
-of the system, it is a called a ``virtual" constraint. 
-This intuition leads to the following mathematical development, described in 
-\cite{vhcs_for_el_systems}:
+constraint on \(q\), which does not depend on \(\dot{q}\). This type of
+constraint is called ``holonomic". Furthermore, since the constraint is enforced
+by control as opposed to being a feature of the mechanics of the system, it is a
+called a ``virtual" constraint.  This intuition leads to the following
+mathematical development, as was described in \cite{vhcs_for_el_systems}:
 
@@
 \begin{defn}\label{defn:vhc_regular}
-   A VHC of order \((n-1)\) is \textbf{regular} if 
+   A VHC \(h(q)\) of order \((n-1)\) is \textbf{regular} if 
    system (\ref{eqn:eom_lagrangian}) with output \(e = h(q)\) has vector
    relative degree \(\{2,2,\ldots,2\}\) on \(\Gamma\).
-   Equivalently, a VHC is regular if 
+   Equivalently, a VHC is regular if and only if 
    \[
@@
    \]
-   is nonsingular everywhere on \(\Gamma\).
+   is full rank everywhere on \(\Gamma\).
 \end{defn}
@@
 
-To show this, we repeat the derivation of
+To show this, we repeat the analysis by
 \textcite{vhcs_for_el_systems,lagrangian_structure_reduced_dynamics_vhcs} and take
@@
 \end{equation*}
-Then the resulting error dynamics are
+Then the resulting output dynamics are
 \begin{equation}\label{eqn:vhc_error_dynamics}
@@
 \end{equation*}
-will generate the linear error dynamics \(\ddot{e} = -k_p e - k_d\dot{e}\) 
+will generate the linear output dynamics \(\ddot{e} = -k_p e - k_d\dot{e}\) 
 which make \(\Gamma\) asymptotically stable.
 
+%---------- Constrained Dynamics ----------%
+\subsection{Constrained Dynamics for Lagrangian VHCs}
 After constructing this controller, a natural question to ask is how system
@@
 As in \cite{lagrangian_structure_reduced_dynamics_vhcs}, 
-now \(s \in \Theta\) be the virtual configuration which parameterizes
-\(h^{-1}(0)\) through the diffeomorphism 
+let \(s \in \Theta\) be the virtual configuration which parameterizes
+\(h^{-1}(0)\) through a diffeomorphism 
 \begin{align*}
    \sigma : \Theta &\rightarrow h^{-1}(0) \\
-   s &\mapsto q
+   s &\mapsto \sigma(s) = q
 \end{align*}
-Likewise, the tangent bundle \(T\Theta\) is diffeomorphic to \(Gamma\) 
+Likewise, the tangent bundle \(T\Theta\) is diffeomorphic to \(\Gamma\) 
 through the mapping
@@
 Using this new virtual configuration, the state of system
-(\ref{eqn:eom_lagrangian}) on the constraint manifold
+(\ref{eqn:eom_lagrangian}) on \(\Gamma\) 
 is parameterized by \(q = \sigma(s)\) and \(\dot{q} = \sigma'(s)\dot{s}\).
@@
 
-By how \(C(q,\dot{q})\) is defined, 
-one can collect the product \(B^\perp C \dot{q}\) in the following way.
-First, combine the Christoffel symbols  into a matrix
+The matrix \(C(q,\dot{q})\) is defined element-wise by 
+\begin{align*}
+   [C]_{j,k} &= \sum \limits_{i=1}^n c_{i,j,k}\dot{q}_i \\
+   c_{i,j,k} &= ???
+\end{align*}
+where \(c_{i,j,k}\) are the Christoffel symbols.
+As such, one can collect the product \(B^\perp C \dot{q}\) in the following way.
+First, combine the Christoffel symbols  into matrices
 \(Q_i : \mathcal{Q} \rightarrow \mathbb{R}^{n \times n}\)
@@
 \end{equation*}
-
 Then \(B^\perp C \dot{q}\) becomes
@@
 
+%---------- Pros and Cons of VHCs ----------%
+\subsection{Benefits and Drawbacks of VHCs}
+\textbf{TODO: why are VHCs good? Who has used them? What is their downside?}
+
 %/========== /Virtual Holonomic Constraints ==========/%

### 2019-11-14 latex/vnhcs.tex
Adding more TODOs and renaming subsections
Wrote the motivation of VHCs
Also wrote a bunch more stuff for VNHCs, providing more detail for the
Hamiltonian-from-Lagrangian section.

@@
 
-%/----- Motivation -----/%
+%----- Motivation -----%
 \subsection{Motivation}
-Suppose one wishes to set a constraint that depends on both the configuration
-and the velocity.
-TODO: Why do we bother with Hamiltonian? Why can't we do virtual nonholonomic 
-constraints in Lagrangian?
+\textbf{TODO: Why do we bother with Hamiltonian? Why can't we do virtual nonholonomic 
+constraints in Lagrangian? What are some use-cases where VHCs don't work?}
 
+%---------- Hamiltonian from Lagrangian ----------%
 \subsection{Review: Hamiltonian Systems}
@@
 
-One can compute the Hamiltonian of the Lagrangian (\ref{eqn:lagrangian}) by 
-performing a Legendre transform on the Lagrangian. 
+One can compute the Hamiltonian of a system by performing a Legendre transform
+on its Lagrangian \textbf{TODO: citation for legendre transform}.
 First, define the conjugate of momenta for \(q\) by 
@@
 \end{equation*}
-It is easy to verify that the Hamiltonian for system (\ref{eqn:lagrangian}) is 
-given by the total mechanical energy (\ref{eqn:hamiltonian}).
+
+For a mechanical system (\ref{eqn:lagrangian}), the conjugate of momenta for
+\(q\) is given by 
+\begin{equation*}
+    p = M(q)\dot{q}
+\end{equation*}
+which means the Hamiltonian of the system is given by
+the total mechanical energy \(E\) (\ref{eqn:hamiltonian} in 
+\((q,p)\) coordinates.
 \begin{equation}\label{eqn:hamiltonian}
-\mathcal{H}(q,p) = E(q,p) = \frac{1}{2} p^T M^{-1}(q) p + V(q)
+\mathcal{H}(q,p) = E(q,p) = \frac{1}{2} p^T \Minv(q) p + V(q)
 \end{equation}
+Note that \(M(q)\) is the inertia matrix and \(V(q)\) is the potential for
+the Hamiltonian system. This is simply a trick to distinguish them notationally
+from the Lagrangian \(D(q)\) and \(P(q)\); they are, in fact, identical in their
+contents.
 
@@
 \begin{split}
-\dot{q} &= \frac{\partial \mathcal{H}}{\partial p} = M^{-1}(q) p \\
+\dot{q} &= \frac{\partial \mathcal{H}}{\partial p} = \Minv(q) p \\
 \dot{p} &= -\frac{\partial \mathcal{H}}{\partial q} + B \tau
@@
 
+%---------- Hamiltonian VNHCs ----------%
 \subsection{Hamiltonian Virtual Nonholonomic Constraints}
 While VHCs are still possible in the Hamiltonian framework, the assumptions
-required to make this work are slightly different. To see this, let us look back
-to the derivation which led to the error dynamics (\ref{eqn:vhc_error_dynamics})
-and attempt to reconstruct this in Hamiltonian coordinates.
+required to make this work are slightly different. Rather than deriving
+Hamiltonian VHCs directly, we will produce results for nonholonomic constraints
+first as VHCs are a special case of this new framework.
 
@@
 \begin{split}
-\dot{q_u} &= e_1^T M^{-1}(q) p \\
+\dot{q_u} &= e_1^T \Minv(q) p \\
 \dot{p_u} &= -p^T\frac{\partial M}{\partial q_u} p - \partial_{q_u}V(q) \\
@@
 I_{n-1} \\
-\end{bmatrix} M^{-1}(q) p \\
+\end{bmatrix} \Minv(q) p \\
 \dot{p_a} &= -p^T\frac{\partial M}{\partial q_a} p - \nabla{q_a}V(q) + 
@@
     &= [-df_{q_u} I_{n-1}]\dot{q} - df_{p_u} \dot{p_u} \\
-    &= dh_q M^{-1}(q) p - df_{p_u}\left( -\frac{1}{2}p^T \frac{\partial M^{-1}(q)}{\partial q_u} p - \partial_{q_u}V(q) \right) \\
-    &= dh_q M^{-1}(q) p + \frac{1}{2}df_{p_u} p^T \frac{\partial M^{-1}(q)}{\partial q_u} p + df_{p_u}\partial_{q_u}V(q)
+    &= dh_q \Minv(q) p - df_{p_u}\left( -\frac{1}{2}p^T \frac{\partial \Minv(q)}{\partial q_u} p - \partial_{q_u}V(q) \right) \\
+    &= dh_q \Minv(q) p + \frac{1}{2}df_{p_u} p^T \frac{\partial \Minv(q)}{\partial q_u} p + df_{p_u}\partial_{q_u}V(q)
 \end{align*}
@@
 \begin{align*}
-    \ddot{e} &= dh_q M^{-1}(q) \dot{p} + df_{p_u} p^T \frac{\partial M^{-1}(q)}{\partial q_u} \dot{p} + (*) \\
-    &= (dh_q M^{-1}(q) + df_{p_u} p^T \frac{\partial M^{-1}(q)}{\partial q_u})B\tau + (*) \\
-    &= (dh_q M^{-1}(q) + dh_{p_u} p^T \frac{\partial M^{-1}(q)}{\partial q_u})B\tau + (*)
+    \ddot{e} &= dh_q \Minv(q) \dot{p} + df_{p_u} p^T \frac{\partial \Minv(q)}{\partial q_u} \dot{p} + (*) \\
+    &= (dh_q \Minv(q) + df_{p_u} p^T \frac{\partial \Minv(q)}{\partial q_u})B\tau + (*) \\
+    &= (dh_q \Minv(q) + dh_{p_u} p^T \frac{\partial \Minv(q)}{\partial q_u})B\tau + (*)
 \end{align*}
@@
 \[
-\text{rank}\left\{ (dh_q M^{-1}(q) + dh_{p_u} p^T \frac{\partial M^{-1}(q)}{\partial q_u})B\right\} = n - 1
+\text{rank}\left\{ (dh_q \Minv(q) + dh_{p_u} p^T \frac{\partial \Minv(q)}{\partial q_u})B\right\} = n - 1
 \]
@@
 \[
-\frac{1}{2} df_{p_u} p^T \frac{\partial M^{-1}(q)}{\partial q_u} p 
+\frac{1}{2} df_{p_u} p^T \frac{\partial \Minv(q)}{\partial q_u} p 
 \]
@@
 \begin{assm}\label{assm:M_is_Mqa}
-Assume \(\partial M(q) / \partial q_u = 0 \Leftrightarrow \partial M^{-1}(q) / \partial q_u = 0\)
+Assume \(\partial M(q) / \partial q_u = 0 \Leftrightarrow \partial \Minv(q) / \partial q_u = 0\)
 \end{assm}
 
-Under Assumption \ref{assm:M_is_Mqa}, we get that the rank condition for \(h(q,p)\) to be a regular VNHC reduces to \(\text{rank}\left(dh_q M^{-1} B\right) = n - 1\).
+Under Assumption \ref{assm:M_is_Mqa}, we get that the rank condition for \(h(q,p)\) to be a regular VNHC reduces to \(\text{rank}\left(dh_q \Minv B\right) = n - 1\).
 This is the same rank condition as required for Virtual Holonomic Constraints.
@@
 \begin{align*}
-    \dot{e} = dh_q M^{-1}(q)p + df_{p_u} \partial_{q_u}V(q) &= 0\\
-    \Leftrightarrow dh_q M^{-1}(q)e_1 p_u + dh_q M^{-1}(q) \begin{bmatrix}
+    \dot{e} = dh_q \Minv(q)p + df_{p_u} \partial_{q_u}V(q) &= 0\\
+    \Leftrightarrow dh_q \Minv(q)e_1 p_u + dh_q \Minv(q) \begin{bmatrix}
     0 & \cdots & 0 \\
@@
 \begin{align*}
-    \Leftrightarrow dh_q M^{-1}(q) \begin{bmatrix}
+    \Leftrightarrow dh_q \Minv(q) \begin{bmatrix}
     0 & \cdots & 0 \\
     & I_{n-1} & \\
-    \end{bmatrix} p_a = -\left(df_{p_u}\partial_{q_u}V(q) + dh_q M^{-1}(q)e_1 p_u\right)
+    \end{bmatrix} p_a = -\left(df_{p_u}\partial_{q_u}V(q) + dh_q \Minv(q)e_1 p_u\right)
 \end{align*}
@@
 \[
-\text{rank}\left(dh_q M^{-1}(q) \begin{bmatrix}
+\text{rank}\left(dh_q \Minv(q) \begin{bmatrix}
     0 & \cdots & 0 \\
@@
 \begin{align*}
-p_a &= -\left(dh_q M^{-1}(q) \begin{bmatrix}
+p_a &= -\left(dh_q \Minv(q) \begin{bmatrix}
     0 & \cdots & 0 \\
     & I_{n-1} & \\
-    \end{bmatrix}\right)^{-1}\left( df_{p_u}\partial_{q_u}V(q) + dh_q M^{-1}(q)e_1 p_u\right) \\
+    \end{bmatrix}\right)^{-1}\left( df_{p_u}\partial_{q_u}V(q) + dh_q \Minv(q)e_1 p_u\right) \\
     &=: g(q_u,p_u)
@@
 \begin{align*}
-    \dot{q_u} &= e_1^T M^{-1}(q_a) \begin{bmatrix}
+    \dot{q_u} &= e_1^T \Minv(q_a) \begin{bmatrix}
     p_u \\
@@
 %---------- Hamiltonian VHCS ---------%
-\subsection{Hamiltonian Virtual Holonomic Constraints}
+\subsection{Restriction to Hamiltonian VHCs}
 \textbf{TODO: Why can we not do the standard VHC approach for Hamiltonian? 

### 2020-02-28 latex/bcc/main.tex
Added the abstract for the 2020 Biennial Control Conference

%! TEX program = lualatex

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}

%----------- Choose Margins ----------%
\usepackage[margin=1in]{geometry}

%---------- Bibliography Style ----------%
\usepackage[style=ieee,backend=biber]{biblatex}
\addbibresource{bib.bib}

%/========== Preamble ==========/%

%----------- Math Packages -----------%
\usepackage{amsfonts}
\usepackage{mathtools}
\usepackage{amsthm} % allow theorems and definitions as a block

%---------- Remove indenting from paragraphs ----------%
% Also makes each new paragraph start with an vertical space.
\usepackage[parfill]{parskip}

%/========== /Preamble ==========/%

%/========== Main Document ==========/%
\begin{document}
%---------- Title ----------%
\title{Energy Injection for Mechanical Systems via Virtual Nonholonomic Constraints}
\author{Adan Moran-MacDonald}
\maketitle

%---------- Content ----------%
When a gymnast hanging on a bar performs a swing-up routine (a so-called ``giant"),
they do not attempt to track a trajectory over time to start doing backflips.
Likewise, a child on a swing does not swing higher by kicking their legs at
specific time intervals; rather, these natural human motions change one's leg
angle based on the current direction and speed of motion. In doing so, the
gymnast will gain enough energy to start rotating on the bar, 
and the child will reach greater heights on their swing.
This type of motion can be adapted to the control of mechanical systems 
using the method of virtual nonholonomic constraints.

A \textit{virtual nonholonomic constraint} (VNHC) is a smooth relation 
\[
    h : T^\star \mathcal{Q} \rightarrow \mathbb{R}^{n-1}
\]
between the \(n\) configuration variables \(q \in \mathcal{Q}\) and conjugate of
momenta \(p\) of a mechanical system with \(n-1\) control inputs.
Enforcing these types of constraints via feedback control provides
time-independent controllers which can inject or remove energy from the system.

This talk will give sufficient conditions under which a VNHC 
can be enforced by feedback control.
Then, using the motivations of the child on the swing and the gymnast,
we apply the theory to create VNHCs for two classical mechanical systems:
the variable-length pendulum and the two-link acrobot.

For the variable-length pendulum we design and enforce a VNHC which achieves a
time-optimal swingup motion, without tracking a time signal. 
Then, we prove that a system constrained by this VNHC will gain energy and start
rotating.

For the acrobot, we design several VNHCs which achieve human-like motion by
observing the leg motions of human gymnasts. We then apply one of these
VNHCs to a real acrobot and, through experiments, demonstrate the
energy-injection properties of the VNHC.

This talk demonstrates the potential of VNHCs in generating realistic,
biologically-inspired motions which can inject energy into a mechanical system
while forgoing any dependence on time in the control architecture.

%---------- Bibliography ----------%
\newpage
\printbibliography
\end{document}
%/========== /Main Document ==========/%

% vim: set tw=80 ts=4 sw=4 sts=0 et ffs=unix :

### 2020-02-29 latex/bcc/main.tex
Edit the abstract
This commit edits the abstract for the 2020 BCC to have more interesting
and natural language, and removes some errors while rephrasing certain
sections.

@@
 %---------- Bibliography Style ----------%
-\usepackage[style=ieee,backend=biber]{biblatex}
-\addbibresource{bib.bib}
+%\usepackage[style=ieee,backend=biber]{biblatex}
+%\addbibresource{bib.bib}
 
@@
 \usepackage{mathtools}
-\usepackage{amsthm} % allow theorems and definitions as a block
 
@@
 
-%---------- Content ----------%
-When a gymnast hanging on a bar performs a swing-up routine (a so-called ``giant"),
-they do not attempt to track a trajectory over time to start doing backflips.
-Likewise, a child on a swing does not swing higher by kicking their legs at
-specific time intervals; rather, these natural human motions change one's leg
-angle based on the current direction and speed of motion. In doing so, the
-gymnast will gain enough energy to start rotating on the bar, 
+%---------- Motivation ----------%
+When a gymnast is hanging on a bar, they do not attempt to track a trajectory
+over time to perform their swing-up routine (a so-called ``giant");
+likewise, a child on a standing swing does not swing higher by standing or
+squatting at specific time intervals. 
+Rather, these natural human motions involve moving one's legs based on the
+current direction and speed of motion. 
+In doing so, the gymnast will gain enough energy to start rotating on the bar,
 and the child will reach greater heights on their swing. 
-This type of motion can be adapted to the control of mechanical systems 
-using the method of virtual nonholonomic constraints.
+This style of realistic, energy-increasing motion can be achieved when
+controlling mechanical systems by using the method of virtual nonholonomic
+constraints.
 
+%---------- Summary of Ideas ----------%
 A \textit{virtual nonholonomic constraint} (VNHC) is a smooth relation 
@@
 between the \(n\) configuration variables \(q \in \mathcal{Q}\) and conjugate of
-momenta \(p\) of a mechanical system with \(n-1\) control inputs.
+momenta \(p \in T_q^\star\mathcal{Q}\) of a mechanical system with 
+\(n-1\) control inputs.
 Enforcing these types of constraints via feedback control provides
@@
 
-This talk will give sufficient conditions under which a VNHC 
-can be enforced by feedback control.
-Then, using the motivations of the child on the swing and the gymnast,
-we apply the theory to create VNHCs for two classical mechanical systems:
-the variable-length pendulum and the two-link acrobot.
+%---------- Results ----------%
+This talk will give sufficient conditions under which a VNHC can be enforced by
+feedback control.
+Then, using the motivations of the child on the standing swing and the gymnast
+on a bar, we apply the theory to develop energy-injecting VNHCs for two
+classical systems: the variable-length pendulum and the two-link acrobot.
 
 For the variable-length pendulum we design and enforce a VNHC which achieves a
-time-optimal swingup motion, without tracking a time signal. 
-Then, we prove that a system constrained by this VNHC will gain energy and start
-rotating.
+time-optimal swingup motion, without tracking a signal over time. 
+We then prove that anyone swinging while constrained by this VNHC will gain
+enough energy to start rotating completely around the swinging axis.
 
-For the acrobot, we design several VNHCs which achieve human-like motion by
-observing the leg motions of human gymnasts. We then apply one of these
-VNHCs to a real acrobot and, through experiments, demonstrate the
-energy-injection properties of the VNHC.
+For the acrobot, we design several VNHCs (modelled from the leg motions of human
+gymnasts) which achieve human-like ``giant" motion.
+We then apply one of these VNHCs to a real acrobot and, through experiments,
+demonstrate the energy-injection properties of the VNHC as the robot kicks its
+legs and starts doing backflips on the bar.
 
-This talk demonstrates the potential of VNHCs in generating realistic,
-biologically-inspired motions which can inject energy into a mechanical system
-while forgoing any dependence on time in the control architecture.
+%---------- Conclusion ----------%
+In summary, this talk introduces the notion of virtual nonholonomic constraints
+and demonstrates their potential for generating realistic, biologically-inspired
+motions which inject energy into mechanical systems while forgoing any
+dependence on time in the control architecture.
 
 %---------- Bibliography ----------%
-\newpage
-\printbibliography
+%\newpage
+%\printbibliography
 \end{document}

### 2020-03-01 latex/bcc/main.tex
Shorten the abstract
The abstract was over 1 page long, so I cut some stuff out, removed some
newlines, and added a bit of extra detail. It is now exactly one page
long.

@@
 A \textit{virtual nonholonomic constraint} (VNHC) is a smooth relation 
-\[
+\(
     h : T^\star \mathcal{Q} \rightarrow \mathbb{R}^{n-1}
-\]
-between the \(n\) configuration variables \(q \in \mathcal{Q}\) and conjugate of
-momenta \(p \in T_q^\star\mathcal{Q}\) of a mechanical system with 
-\(n-1\) control inputs.
+\) 
+between the configuration variables \(q \in \mathcal{Q}\) and conjugate of
+momenta \(p \in T_q^\star\mathcal{Q}\) of an \(n\)-degree-of-freedom mechanical
+system with \(n-1\) control inputs, such that \(h(q,p) = 0\).
 Enforcing these types of constraints via feedback control provides
@@
 time-optimal swingup motion, without tracking a signal over time. 
-We then prove that anyone swinging while constrained by this VNHC will gain
+We then prove that any pendulum constrained by this VNHC will gain
 enough energy to start rotating completely around the swinging axis.
@@
 demonstrate the energy-injection properties of the VNHC as the robot kicks its
-legs and starts doing backflips on the bar.
+legs until it begins doing backflips on the bar.
 

### 2020-03-21 latex/ogs/main.tex
Starting the OGS application

--- 2020-02-28 latex/bcc/main.tex
@@
 \usepackage[utf8]{inputenc}
+
+%/========== Preamble ==========/%
 
@@
 
-%/========== Preamble ==========/%
-
 %----------- Math Packages -----------%
-\usepackage{amsfonts}
 \usepackage{mathtools}
-\usepackage{amsthm} % allow theorems and definitions as a block
 
@@
 
-%/========== /Preamble ==========/%
-
 %/========== Main Document ==========/%
 \begin{document}
-%---------- Title ----------%
-\title{Energy Injection for Mechanical Systems via Virtual Nonholonomic Constraints}
+
+%---------- Title Page -----------%
+\title{Energy Injection for Mechanical Systems through Virtual Nonholonomic Constraints}
 \author{Adan Moran-MacDonald}
+\date{\today}
 \maketitle
+%---------- Main Content ----------%
 
-%---------- Content ----------%
-When a gymnast hanging on a bar performs a swing-up routine (a so-called ``giant"),
-they do not attempt to track a trajectory over time to start doing backflips.
-Likewise, a child on a swing does not swing higher by kicking their legs at
-specific time intervals; rather, these natural human motions change one's leg
-angle based on the current direction and speed of motion. In doing so, the
-gymnast will gain enough energy to start rotating on the bar, 
-and the child will reach greater heights on their swing.
-This type of motion can be adapted to the control of mechanical systems 
-using the method of virtual nonholonomic constraints.
-
-A \textit{virtual nonholonomic constraint} (VNHC) is a smooth relation 
-\[
-    h : T^\star \mathcal{Q} \rightarrow \mathbb{R}^{n-1}
-\]
-between the \(n\) configuration variables \(q \in \mathcal{Q}\) and conjugate of
-momenta \(p\) of a mechanical system with \(n-1\) control inputs.
-Enforcing these types of constraints via feedback control provides
-time-independent controllers which can inject or remove energy from the system.
-
-This talk will give sufficient conditions under which a VNHC 
-can be enforced by feedback control.
-Then, using the motivations of the child on the swing and the gymnast,
-we apply the theory to create VNHCs for two classical mechanical systems:
-the variable-length pendulum and the two-link acrobot.
-
-For the variable-length pendulum we design and enforce a VNHC which achieves a
-time-optimal swingup motion, without tracking a time signal. 
-Then, we prove that a system constrained by this VNHC will gain energy and start
-rotating.
-
-For the acrobot, we design several VNHCs which achieve human-like motion by
-observing the leg motions of human gymnasts. We then apply one of these
-VNHCs to a real acrobot and, through experiments, demonstrate the
-energy-injection properties of the VNHC.
-
-This talk demonstrates the potential of VNHCs in generating realistic,
-biologically-inspired motions which can inject energy into a mechanical system
-while forgoing any dependence on time in the control architecture.
-
+\cite{hybrid_zero_dynamics_bipedal_nhvcs}
 %---------- Bibliography ----------%
-\newpage
 \printbibliography
 \end{document}
-%/========== /Main Document ==========/%
 

### 2020-04-08 latex/ogs/main.tex
Updated the basic layout to use TNR font with appropriate margins
Copied OGS proposal from 2019

@@
 
-\documentclass[12pt]{article}
-\usepackage[utf8]{inputenc}
+\documentclass[a4paper,12pt]{article}
 
 %/========== Preamble ==========/%
+%----------- Set Main Font ----------%
+\usepackage{times} % Times New Roman
 
 %----------- Choose Margins ----------%
-\usepackage[margin=1in]{geometry}
+\usepackage[a4paper,margin=1.87cm]{geometry}
 
@@
 
-%---------- Title Page -----------%
-\title{Energy Injection for Mechanical Systems through Virtual Nonholonomic Constraints}
+%---------- Title -----------%
+\title{Ontario Graduate Scholarship: Plan of Study}
 \author{Adan Moran-MacDonald}
-\date{\today}
+\date{}
 \maketitle
+
 %---------- Main Content ----------%
+\begin{large}
+\textbf{Background}
+\end{large}
 
-\cite{hybrid_zero_dynamics_bipedal_nhvcs}
+Mechanical systems can have their motion constrained in two ways. First, the
+constraint can be a function of their position, such as how the joints of a
+robotic arm cannot go through each other. This type of constraint is called
+``holonomic”. The second type of constraint involves both position and velocity,
+like the fact that a car cannot normally slide sideways. This type of constraint
+is called ``nonholonomic”.
+
+It has been shown that one can create a holonomic constraint which did not exist
+before through a control algorithm (see for example \cite{vhcs_for_el_systems}).
+These constraints are called ``virtual" holonomic constraints (VHC) since they
+are enforced by the robot’s actuators, as opposed to being enforced by laws of
+physics. VHCs have been used to control walking robots \cite{stable-walking},
+autonomous bicycles \cite{bicycle}, gymnastics robots \cite{xingbo-thesis}, and
+snake robots \cite{snake-robot} among others. Mohammadi \textit{et.al.} showed
+that a single VHC can result in several motion patterns, from which one can be
+chosen if the robot is initialized with a configuration that is ``close enough”
+to the desired motion pattern \cite{manfredi-orbit-stabilization}. However,
+their method does not work for all starting configurations of the robot.
+
+\begin{large} \textbf{Proposal} \end{large}
+
+We propose to investigate a method that can stabilize a desired motion pattern
+for a VHC from any starting configuration of the robot. This involves using
+virtual nonholonomic constraints (VNHC) to maneuver the robot into a position
+where the desired motion pattern can be stabilized.
+
+VNHCs have been studied for application to human-robot cooperation in
+\cite{vnhc-human-robot-coop} and walking robots in \cite{vnhc-biped-robot}. Horn
+\textit{et.al.} have also examined properties of VNHCs for general mechanical
+systems \cite{hybrid_zero_dynamics_bipedal_nhvcs} and then applied their results
+to walking robots. Our approach is novel since we will study VNHCs on general
+mechanical systems with the aim of stabilizing any desired motion pattern on a
+VHC. We will prove our results theoretically, then demonstrate them
+experimentally on the gymnastics robot built by Wang \cite{xingbo-thesis}.
+
+I am currently at the beginning stages of this thesis. We are studying
+mathematical conditions whereby a general mechanical system constrained by a
+VNHC will gain or lose energy over time. This will allow us to create a control
+algorithm to add or remove energy from a robot. In doing so we will position the
+robot configuration to stabilize any desired motion pattern which comes from a
+VHC. After completing this general mathematical formulation, we will apply the
+results to the gymnastics robot created by Wang \cite{xingbo-thesis}. We will
+create virtual constraints that will make the robot perform a backflip routine
+around a bar to confirm that our rigorous mathematical results are applicable to
+real life autonomous systems.
+
+The advancement in mathematics from this thesis will also enable improvements to
+control of other robotic systems. It will provide a more computationally
+feasible method to stabilizing movement patterns than what was created in
+\cite{manfredi-orbit-stabilization}.
+
 %---------- Bibliography ----------%

### 2020-04-08 latex/ogs/main.tex
Rough draft v1 of new proposal

--- 2020-03-21 latex/ogs/main.tex
@@
 
-%---------- Title Page -----------%
-\title{Energy Injection for Mechanical Systems through Virtual Nonholonomic Constraints}
+%---------- Title -----------%
+\title{Ontario Graduate Scholarship: Plan of Study}
 \author{Adan Moran-MacDonald}
-\date{\today}
+\date{}
 \maketitle
+
 %---------- Main Content ----------%
+\begin{large}
+\textbf{Background}
+\end{large}
 
-\cite{hybrid_zero_dynamics_bipedal_nhvcs}
+When a gymnast is performing a swing-up routine on a bar, they do not kick their
+feet at specified time intervals. Rather, a more natural human motion is to
+swing their legs based on their current speed and position. In doing so, they
+are able to generate enough momentum to perform backflips on the bar. When
+controlling mechanical systems, standard practice is to have a robot's actuators
+track a position dependent on time. Much like the gymnast, this tends to
+produce unnatural motion. Furthermore, it is liable to timin errors if the
+motion is interrupted in any way. The more recent techniques of virtual constraints 
+can rectify these issues and produce realistic, biologically inspired motion.
+
+Mechanical systems can have their motion constrained in two ways. Holonomic 
+constraints restrict position, while nonholonomic constraints restrict
+velocities (such as how a car cannot normally slide sideways). Previous research
+has shown that one can enforce holonomic constraints through a robot's
+actuators \cite{vhcs_for_el_systems}. These are called ``virtual holonomic
+constraints" (VHCs). VHCs have been used to control walking robots \cite{stable-walking},
+autonomous bicycles \cite{bicycle}, gymnastics robots \cite{xingbo-thesis}, and
+snake robots \cite{snake-robot} among others. Likewise, in my 
+master's thesis \textbf{TODO: HOW TO CITE?} I developed a general theory for
+enforcing virtual nonholonomic constraints (VNHCs). I designed constraints which
+injected energy into the gymnastics robot created by 
+Wang \cite{xingbo-thesis}, enabling it to perform backflips on a bar.
+
+\begin{large} \textbf{Proposal} \end{large}
+
+Mohammadi \textit{et.al.} showed that a single VHC can result in several motion
+patterns, from which one pattern can be chosen if the robot is initialized with
+a configuration that is ``close enough” to the desired motion pattern
+\cite{manfredi-orbit-stabilization}. However, their method does not work for all
+starting configurations of the robot. We propose to investigate a method to
+stabilize and transition between desired motion patterns from any starting
+configuration of the robot. Rather than using VHCs, we will use VNHCs to
+generate these motions.
+
+VNHCs have been studied for application to human-robot cooperation in
+\cite{vnhc-human-robot-coop} and walking robots in \cite{vnhc-biped-robot}. Horn
+\textit{et.al.} have also examined properties of VNHCs for general mechanical
+systems \cite{hybrid_zero_dynamics_bipedal_nhvcs} and then applied their results
+to walking robots. Our approach is novel since we will study VNHCs on general
+mechanical systems with the aim of transitioning between desired motion patterns.
+
+We will first study mathematical conditions under which on can transition
+between two VNHCs. These conditions will allow us to create a control algorithm
+to switch between motion patterns. After completing this mathematical
+formulation, we will apply the results to the gymanstics robot built by Wang
+\cite{xingbo-thesis}. We will use the theory to create control algorithms which
+enable this robot to perform an entire gymnastics routine by transitioning
+between different VNHCs.
+
+The advancement in mathematics from this research will enable improvements to
+control of other robotic systems. The ability to transition between complex
+constraints will allow for more natural expressive motion in many autonomous
+systems and may become a standard technique for controlling biologically
+inspired robots.
+
 %---------- Bibliography ----------%

### 2020-04-09 latex/ogs/main.tex
Rough Draft update: changed the goal to transitioning between VNHCs

@@
 track a position dependent on time. Much like the gymnast, this tends to
-produce unnatural motion. Furthermore, it is liable to timin errors if the
+produce unnatural motion. Furthermore, it is liable to timing errors if the
 motion is interrupted in any way. The more recent techniques of virtual constraints 
-can rectify these issues and produce realistic, biologically inspired motion.
+can rectify these issues and produce realistic, biologically plausible motion.
 
-Mechanical systems can have their motion constrained in two ways. Holonomic 
-constraints restrict position, while nonholonomic constraints restrict
-velocities (such as how a car cannot normally slide sideways). Previous research
-has shown that one can enforce holonomic constraints through a robot's
-actuators \cite{vhcs_for_el_systems}. These are called ``virtual holonomic
-constraints" (VHCs). VHCs have been used to control walking robots \cite{stable-walking},
-autonomous bicycles \cite{bicycle}, gymnastics robots \cite{xingbo-thesis}, and
-snake robots \cite{snake-robot} among others. Likewise, in my 
-master's thesis \textbf{TODO: HOW TO CITE?} I developed a general theory for
-enforcing virtual nonholonomic constraints (VNHCs). I designed constraints which
-injected energy into the gymnastics robot created by 
-Wang \cite{xingbo-thesis}, enabling it to perform backflips on a bar.
+Mechanical systems can be constrained in two ways. Holonomic constraints
+restrict position (for instance, car tires remain a fixed distance apart), while
+nonholonomic constraints restrict velocities (such as how those tires cannot
+slide sideways). Previous research has shown that one can enforce holonomic
+constraints through a robot's actuators \cite{vhcs_for_el_systems}. These are
+called ``virtual holonomic constraints" (VHCs). VHCs have been used to control
+walking robots \cite{stable-walking}, autonomous bicycles \cite{bicycle},
+gymnastics robots \cite{xingbo-thesis}, and snake robots \cite{snake-robot}
+among others. These VHCs are a special case of the more general virtual
+nonholonomic constraints (VNHCs); in my master's thesis \textbf{TODO: HOW TO
+CITE?} I developed a general theory for enforcing VNHCs and designed
+one which injects energy into the gymnastics robot created by Wang
+\cite{xingbo-thesis}, thereby enabling it to perform backflips on a bar.
 
@@
 Mohammadi \textit{et.al.} showed that a single VHC can result in several motion
-patterns, from which one pattern can be chosen if the robot is initialized with
-a configuration that is ``close enough” to the desired motion pattern
+patterns (also called ``orbits"), from which one specific motion can be chosen
+if the robot is initialized with a configuration that is ``close enough” to it
 \cite{manfredi-orbit-stabilization}. However, their method does not work for all
-starting configurations of the robot. We propose to investigate a method to
-stabilize and transition between desired motion patterns from any starting
-configuration of the robot. Rather than using VHCs, we will use VNHCs to
-generate these motions.
+starting configurations of the robot. We propose to extend this work and 
+investigate a method to stabilize and transition between two VNHCs, which should
+also allow one to stabilize an orbit from any starting configuration.
 
-VNHCs have been studied for application to human-robot cooperation in
-\cite{vnhc-human-robot-coop} and walking robots in \cite{vnhc-biped-robot}. Horn
+VNHCs have been studied for application to human-robot cooperation
+\cite{vnhc-human-robot-coop} and walking robots \cite{vnhc-biped-robot}. Horn
 \textit{et.al.} have also examined properties of VNHCs for general mechanical
-systems \cite{hybrid_zero_dynamics_bipedal_nhvcs} and then applied their results
-to walking robots. Our approach is novel since we will study VNHCs on general
-mechanical systems with the aim of transitioning between desired motion patterns.
+systems \cite{hybrid_zero_dynamics_bipedal_nhvcs} and applied their results
+to stabilizing a chosen gait on walking robots. 
+Our approach is novel since we will develop techniques to transition between  
+VNHCs on general mechanical systems while maintaining safety constraints, 
+rather than stabilizing a single VNHC.
 
@@
 to switch between motion patterns. After completing this mathematical
-formulation, we will apply the results to the gymanstics robot built by Wang
-\cite{xingbo-thesis}. We will use the theory to create control algorithms which
-enable this robot to perform an entire gymnastics routine by transitioning
-between different VNHCs.
+formulation, we will test the theory on the gymanstics robot built by Wang
+\cite{xingbo-thesis} by designing a gymnastics routine which is performed 
+by transitioning between different VNHCs.
 

### 2020-04-10 latex/ogs/main.tex
Added my thesis as a citation, and got to rough draft v3

@@
 
+\textbf{TODO: CITATIONS FOR THIS PARAGRAPH}
 When a gymnast is performing a swing-up routine on a bar, they do not kick their
 feet at specified time intervals. Rather, a more natural human motion is to
-swing their legs based on their current speed and position. In doing so, they
-are able to generate enough momentum to perform backflips on the bar. When
-controlling mechanical systems, standard practice is to have a robot's actuators
-track a position dependent on time. Much like the gymnast, this tends to
-produce unnatural motion. Furthermore, it is liable to timing errors if the
-motion is interrupted in any way. The more recent techniques of virtual constraints 
-can rectify these issues and produce realistic, biologically plausible motion.
+smoothly swing their legs based on their current speed and position
+\cite{pendulum_length_giant_gymnastics}. In doing so, they are able to generate
+enough momentum to perform backflips on the bar. When controlling mechanical
+systems, standard practice is to have a robot's actuators track a function of
+time. Much like the gymnast, it is often more natural to track a
+function of the robot's current speed and position. The more recent techniques
+of virtual constraints offer this ability and produce realistic, biologically
+plausible motion.
 
@@
 slide sideways). Previous research has shown that one can enforce holonomic
-constraints through a robot's actuators \cite{vhcs_for_el_systems}. These are
-called ``virtual holonomic constraints" (VHCs). VHCs have been used to control
+constraints through a robot's actuators, using the method of "virtual holonomic
+constraints" (VHCs) \cite{vhcs_for_el_systems}. VHCs have been used to control
 walking robots \cite{stable-walking}, autonomous bicycles \cite{bicycle},
 gymnastics robots \cite{xingbo-thesis}, and snake robots \cite{snake-robot}
-among others. These VHCs are a special case of the more general virtual
-nonholonomic constraints (VNHCs); in my master's thesis \textbf{TODO: HOW TO
-CITE?} I developed a general theory for enforcing VNHCs and designed
-one which injects energy into the gymnastics robot created by Wang
-\cite{xingbo-thesis}, thereby enabling it to perform backflips on a bar.
+among others. These VHCs are a special case of virtual nonholonomic constraints
+(VNHCs); in my master's thesis \cite{my-thesis} I developed a general theory for
+enforcing VNHCs and designed one which injects energy into the gymnastics robot
+created by Wang \cite{xingbo-thesis}, thereby enabling it to perform backflips
+on a bar.
 
@@
 
-Mohammadi \textit{et.al.} showed that a single VHC can result in several motion
-patterns (also called ``orbits"), from which one specific motion can be chosen
-if the robot is initialized with a configuration that is ``close enough” to it
-\cite{manfredi-orbit-stabilization}. However, their method does not work for all
-starting configurations of the robot. We propose to extend this work and 
-investigate a method to stabilize and transition between two VNHCs, which should
-also allow one to stabilize an orbit from any starting configuration.
+Mohammadi \textit{et.al.} \cite{manfredi-orbit-stabilization} showed that a
+single VHC can result in several motion patterns (known as ``orbits"), from
+which one specific motion can be chosen if the robot is initialized with a
+configuration that is ``close enough" to it. However, their method does not
+work for all starting configurations of the robot. We propose to extend this
+work and investigate a method to transition between VNHCs; doing so should allow
+one to start anywhere and move the configuration close enough to stabilize the
+chosen orbit.
 
-VNHCs have been studied for application to human-robot cooperation
+VNHCs have been studied with applications to human-robot cooperation
 \cite{vnhc-human-robot-coop} and walking robots \cite{vnhc-biped-robot}. Horn
-\textit{et.al.} have also examined properties of VNHCs for general mechanical
-systems \cite{hybrid_zero_dynamics_bipedal_nhvcs} and applied their results
-to stabilizing a chosen gait on walking robots. 
-Our approach is novel since we will develop techniques to transition between  
-VNHCs on general mechanical systems while maintaining safety constraints, 
-rather than stabilizing a single VNHC.
+\textit{et.al.} \cite{hybrid_zero_dynamics_bipedal_nhvcs} have also examined
+properties of VNHCs for general mechanical systems and applied their results to
+stabilizing a chosen gait on walking robots. Our approach is novel since we will
+assume one has already stabilized a VNHC and develop techniques to transition to
+a different VNHC while maintaining pre-specified safety constraints.
 
-We will first study mathematical conditions under which on can transition
+First we will study mathematical conditions under which on can transition
 between two VNHCs. These conditions will allow us to create a control algorithm
 to switch between motion patterns. After completing this mathematical
-formulation, we will test the theory on the gymanstics robot built by Wang
-\cite{xingbo-thesis} by designing a gymnastics routine which is performed 
-by transitioning between different VNHCs.
+formulation, we will test the theory on a gymnastics robot by designing a
+gymnastics routine which is performed by transitioning between different VNHCS.
 

### 2020-04-12 latex/ogs/main.tex
Final version v1 and its PDF

@@
 
-\textbf{TODO: CITATIONS FOR THIS PARAGRAPH}
 When a gymnast is performing a swing-up routine on a bar, they do not kick their
@@
 time. Much like the gymnast, it is often more natural to track a
-function of the robot's current speed and position. The more recent techniques
+function of the robot's current speed and position. The recent techniques
 of virtual constraints offer this ability and produce realistic, biologically
@@
 restrict position (for instance, car tires remain a fixed distance apart), while
-nonholonomic constraints restrict velocities (such as how those tires cannot
+nonholonomic constraints restrict velocities (such as how car tires cannot
 slide sideways). Previous research has shown that one can enforce holonomic
@@
 single VHC can result in several motion patterns (known as ``orbits"), from
-which one specific motion can be chosen if the robot is initialized with a
+which one specific motion pattern can be chosen if the robot is initialized with a
 configuration that is ``close enough" to it. However, their method does not
 work for all starting configurations of the robot. We propose to extend this
-work and investigate a method to transition between VNHCs; doing so should allow
-one to start anywhere and move the configuration close enough to stabilize the
-chosen orbit.
+work by investigating a method for transitioning between VHCs and VNHCs.
 
@@
 properties of VNHCs for general mechanical systems and applied their results to
-stabilizing a chosen gait on walking robots. Our approach is novel since we will
-assume one has already stabilized a VNHC and develop techniques to transition to
-a different VNHC while maintaining pre-specified safety constraints.
+stabilizing a chosen gait on walking robots. Our approach is novel since we 
+assume one has already stabilized a VNHC, and will develop techniques to
+transition to a different VNHC while maintaining pre-specified safety
+constraints. Doing so should allow one to start at any configuration and chain
+together VNHCs to bring the robot ``close enough" to stabilize a desired orbit.
 
@@
 formulation, we will test the theory on a gymnastics robot by designing a
-gymnastics routine which is performed by transitioning between different VNHCS.
+gymnastics routine which is performed by transitioning between several different
+VNHCS.
 
-The advancement in mathematics from this research will enable improvements to
-control of other robotic systems. The ability to transition between complex
-constraints will allow for more natural expressive motion in many autonomous
-systems and may become a standard technique for controlling biologically
-inspired robots.
+Finally, the advancement in mathematics from this research will enable
+improvements to control of other robotic systems. The ability to transition
+between complex constraints will allow for more natural expressive motion in
+many autonomous systems and may become a standard technique for controlling
+biologically inspired robots.
 

### 2020-04-13 latex/ogs/main.tex
Updated to final draft v2

@@
 systems, standard practice is to have a robot's actuators track a function of
-time. Much like the gymnast, it is often more natural to track a
-function of the robot's current speed and position. The recent techniques
-of virtual constraints offer this ability and produce realistic, biologically
-plausible motion.
+time. Much like the gymnast, this can produce unnatural motion.  The recent
+techniques of virtual constraints offer the ability to track a function of speed
+and position to produce realistic, biologically plausible motion.
 
@@
 gymnastics robots \cite{xingbo-thesis}, and snake robots \cite{snake-robot}
-among others. These VHCs are a special case of virtual nonholonomic constraints
-(VNHCs); in my master's thesis \cite{my-thesis} I developed a general theory for
-enforcing VNHCs and designed one which injects energy into the gymnastics robot
-created by Wang \cite{xingbo-thesis}, thereby enabling it to perform backflips
-on a bar.
+among others. These VHCs are a special case of the more general virtual
+nonholonomic constraints (VNHCs); in my master's thesis \cite{my-thesis} I
+developed a theory for enforcing VNHCs and designed one which injects energy
+into the gymnastics robot created by Wang \cite{xingbo-thesis}, thereby enabling
+it to perform backflips on a bar.
 
@@
 assume one has already stabilized a VNHC, and will develop techniques to
-transition to a different VNHC while maintaining pre-specified safety
+transition to a second one while maintaining certain safety
 constraints. Doing so should allow one to start at any configuration and chain
-together VNHCs to bring the robot ``close enough" to stabilize a desired orbit.
+together VNHCs which drive the configuration close enough to a desired orbit for
+stabilization.
 
-First we will study mathematical conditions under which on can transition
+First we will study mathematical conditions under which one can transition
 between two VNHCs. These conditions will allow us to create a control algorithm
@@
 formulation, we will test the theory on a gymnastics robot by designing a
-gymnastics routine which is performed by transitioning between several different
+gymnastics routine which is performed solely by transitioning between several different
 VNHCS.
 
-Finally, the advancement in mathematics from this research will enable
-improvements to control of other robotic systems. The ability to transition
-between complex constraints will allow for more natural expressive motion in
-many autonomous systems and may become a standard technique for controlling
-biologically inspired robots.
+The advancements in mathematics from this research will bring improvements to
+the control of other autonomous systems. The ability to transition between
+complex constraints will allow for more natural expressive motion and may become
+a standard technique for controlling biologically inspired robots.
 

### 2020-04-13 latex/ogs/main.tex
Update to final v3

@@
 systems, standard practice is to have a robot's actuators track a function of
-time. Much like the gymnast, this can produce unnatural motion.  The recent
-techniques of virtual constraints offer the ability to track a function of speed
-and position to produce realistic, biologically plausible motion.
+time. Much like the gymnast, it can be more natural to have the robot track a
+function of speed and position instead. The recent techniques of virtual
+constraints offer this ability to produce realistic, biologically plausible
+motion.
 

### 2020-04-13 latex/ogs/main.tex
Edited some phrasing to get Final v4

@@
 among others. These VHCs are a special case of the more general virtual
-nonholonomic constraints (VNHCs); in my master's thesis \cite{my-thesis} I
+nonholonomic constraints (VNHCs). In my master's thesis \cite{my-thesis} I
 developed a theory for enforcing VNHCs and designed one which injects energy
@@
 work for all starting configurations of the robot. We propose to extend this
-work by investigating a method for transitioning between VHCs and VNHCs.
+work by investigating a method for transitioning between VNHCs.
 
@@
 stabilizing a chosen gait on walking robots. Our approach is novel since we 
-assume one has already stabilized a VNHC, and will develop techniques to
-transition to a second one while maintaining certain safety
-constraints. Doing so should allow one to start at any configuration and chain
-together VNHCs which drive the configuration close enough to a desired orbit for
-stabilization.
+assume one can stabilize two distinct VNHCs independently, and will develop
+techniques to transition between them while maintaining safety constraints.
+Doing so will allow one to chain together VNHCs which drive a robot's
+configuration towards a desired orbit, regardless of the initial configuration.
 
@@
 the control of other autonomous systems. The ability to transition between
-complex constraints will allow for more natural expressive motion and may become
+complex constraints will allow for more expressive motion and may become
 a standard technique for controlling biologically inspired robots.

### 2020-06-04 latex/abstract.tex
Create a basic skeleton
This commit creates the basic skeleton for the preliminary pages -
abstract, dedication, acknowledgements, title, toc, as well as the
bibliography.

Todo is still all the chapters

%! TEX root = main.tex

%% (At most 150 words for M.Sc. or 350 words for Ph.D.)
\begin{abstract}
	\textbf{TODO}: Fill in the abstract
\end{abstract}
% vim: set ts=3 sw=3 sts=0 et ffs=unix :

### 2020-06-08 latex/symbols.tex
Added comments to various files and created a list of symbols page

%! TEX root = main.tex

%/========== Symbols ==========/%

% Set the heading so it appears in the Table of Contents
\addcontentsline{toc}{chapter}{List of Symbols}
\chapter*{List of Symbols}

% Define a table for the symbols using the booktabs convention
\noindent
\begin{table}[h!]
\resizebox{\textwidth}{!}{%
\centering
\begin{tabular}{ll}
   \toprule
   Symbol & Definition \\
   \midrule
   \(\R^n\) & Real numbers in \(n\) dimensions \\
   \(q_u\) & Unactuated coordinates \\
   \(q_a\) & Actuated coordinates \\
   \(p_u\) & Conjugate of momentum to \(q_u\) \\
   \(p_a\) & Conjugate of momentum to \(q_a\) \\
   A & Some really long description to see how this works when we write long text inside the table. \\
   \bottomrule
\end{tabular}

}% resizebox
\end{table}

%/========== /Symbols ==========/%
% vim: set ts=3 sw=3 sts=0 et ffs=unix :

### 2020-06-08 latex/conclusion.tex
Finalize the first skeleton of the thesis
This commit finishes the skeleton of the thesis, with all sections and
headings as desired. There is currently no content here.

%! TEX root = main.tex

%/========== Conclusion ==========/%

\chapter{Conclusion}\label{ch:conclusion}
% TODO: Summarize the work

\section{Limitations of this Work}

\section{Future Research}

%/========== /Conclusion ==========/%
% vim: set ts=3 sw=3 sts=0 et ffs=unix :

### 2020-06-08 latex/introduction.tex
Finalize the first skeleton of the thesis
This commit finishes the skeleton of the thesis, with all sections and
headings as desired. There is currently no content here.

%! TEX root = main.tex

%/========== Introduction ==========/%

\chapter{Introduction}
\section{Literature Review}

\section{Statement of Contributions}

\section{Outline of the Thesis}

%/========== /Introduction ==========/%
% vim: set ts=3 sw=3 sts=0 et ffs=unix :

### 2020-06-08 latex/vnhcs.tex
Finalize the first skeleton of the thesis
This commit finishes the skeleton of the thesis, with all sections and
headings as desired. There is currently no content here.

--- 2019-11-12 latex/vnhcs.tex
@@
 %/========== Virtual Nonholonomic Constraints ==========/%
-\section{Virtual Nonholonomic Constraints}\label{sec:vnhcs}
 
-%/----- Motivation -----/%
-\subsection{Motivation}
-TODO: Why do we bother with Hamiltonian? Why can't we do virtual nonholonomic constraints in Lagrangian?
+\chapter{Development of Virtual Nonholonomic Constraints}
+\section{Mechanical Systems}
+\subsection{Lagrangian Formulation of Mechanical Systems}
+% TODO: Describe the basics of the lagrangian formulation for mechanical systems
+\subsection{Hamiltonian Formulation of Mechanical Systems}
+% TODO: Describe how to get the Hamiltonian from the lagrangian, and show that
+% you can write the hamiltonian directly in certain conditions of the system
+\subsection{Simply Actuated Hamiltonian Systems}
+% TODO: Describe the change of coordinates to get into q_u/q_a mode and show
+% that the actuator directly affects pa but not pu. Use M for simply actuated
+% inertia, D for normal coordinates
 
-Suppose the mechanical system has degree of underactuation one, so that coordinates of the system can be split into an unactuated component \(q_u \in [\mathbb{R}]_T, \, T \in \mathbb{R}_{>0}\) which is not influence by control, along with an actuated component \(q_a\); that is, suppose \(B\) is of the form \(B(q) = [0_m, B_1^T(q), \ldots, B_n^T(q)]^T, \, B_i^T(q) \in \mathbb{R}^{n-1}\) and \(\tau \in \mathbb{R}^{n-1}\).
+\section{Virtual Nonholonomic Constraints}
+% TODO: Perform the full development of VNHCs, including its stabilizing
+% controller.
+%TODO: Special case: when dM/dqu = 0, show that we have a nice form and that we
+%can solve for p_a and the closed-loop dynamics (qu,pu)_dot
+
+%\section{Virtual Nonholonomic Constraints}\label{sec:vnhcs}
+%
+%%----- Motivation -----%
+%\subsection{Motivation}
+%\textbf{TODO: Why do we bother with Hamiltonian? Why can't we do virtual nonholonomic 
+%constraints in Lagrangian? What are some use-cases where VHCs don't work?}
+%
+%%---------- Hamiltonian from Lagrangian ----------%
+%\subsection{Review: Hamiltonian Systems}
+%\textbf{TODO: What is a Hamiltonian system and why does it matter?}
+%
+%One can compute the Hamiltonian of a system by performing a Legendre transform
+%on its Lagrangian \textbf{TODO: citation for legendre transform}.
+%First, define the conjugate of momenta for \(q\) by 
+%\begin{equation*}
+%p = \frac{\partial \mathcal{L}}{\partial \dot{q}} \in \mathbb{R}^n
+%\end{equation*}
+%Then, the Legendre transform is performed by taking
+%\begin{equation*}
+%\mathcal{H}(q,p) = p^T \dot{q} - \mathcal{L}(q,\dot{q})
+%\end{equation*}
+%
+%For a mechanical system (\ref{eqn:lagrangian}), the conjugate of momenta for
+%\(q\) is given by 
+%\begin{equation*}
+%    p = M(q)\dot{q}
+%\end{equation*}
+%which means the Hamiltonian of the system is given by
+%the total mechanical energy \(E\) (\ref{eqn:hamiltonian} in 
+%\((q,p)\) coordinates.
+%\begin{equation}\label{eqn:hamiltonian}
+%\mathcal{H}(q,p) = E(q,p) = \frac{1}{2} p^T \Minv(q) p + V(q)
+%\end{equation}
+%Note that \(M(q)\) is the inertia matrix and \(V(q)\) is the potential for
+%the Hamiltonian system. This is simply a trick to distinguish them notationally
+%from the Lagrangian \(D(q)\) and \(P(q)\); they are, in fact, identical in their
+%contents.
+%
+%The equations of motion for the system in Hamiltonian coordinates is given by
+%\begin{align}\label{eqn:hamiltionian_eom}
+%\begin{split}
+%\dot{q} &= \frac{\partial \mathcal{H}}{\partial p} = \Minv(q) p \\
+%\dot{p} &= -\frac{\partial \mathcal{H}}{\partial q} + B \tau
+%\end{split}
+%\end{align}
+%
+%\textbf{Why bother looking at Hamiltonian systems? What is the intuition behind
+%these equations of motion?}
+%
+%%---------- Hamiltonian VNHCs ----------%
+%\subsection{Hamiltonian Virtual Nonholonomic Constraints}
+%While VHCs are still possible in the Hamiltonian framework, the assumptions
+%required to make this work are slightly different. Rather than deriving
+%Hamiltonian VHCs directly, we will produce results for nonholonomic constraints
+%first as VHCs are a special case of this new framework.
+%
+%Suppose the mechanical system has degree of underactuation one, so that coordinates of the system can be split into an unactuated component \(q_u \in [\mathbb{R}]_T, \, T \in \mathbb{R}_{>0}\) which is not influence by control, along with an actuated component \(q_a\); that is, suppose \(B\) is of the form \(B(q) = [0_m, B_1^T(q), \ldots, B_n^T(q)]^T, \, B_i^T(q) \in \mathbb{R}^{n-1}\) and \(\tau \in \mathbb{R}^{n-1}\).
 In this case, \(q = (q_u, q_a)^T\) and the equations of motion become
-\begin{align}\label{eqn:unactuated_actuated_eom}
-\begin{split}
-\dot{q_u} &= e_1^T M^{-1}(q) p \\
-\dot{p_u} &= -p^T\frac{\partial M}{\partial q_u} p - \partial_{q_u}V(q) \\
-\dot{q_a} &= 
-\begin{bmatrix}
-0 \cdots 0 \\
-I_{n-1} \\
-\end{bmatrix} M^{-1}(q) p \\
-\dot{p_a} &= -p^T\frac{\partial M}{\partial q_a} p - \nabla{q_a}V(q) + 
-\begin{bmatrix}
-B_1(q) \\
-\vdots \\
-B_n(q)
-\end{bmatrix} \tau
-\end{split}
-\end{align}
-
-Now we can begin to talk about Virtual Nonholonomic Constraints. In a similar fashion to what was defined for VHCs, let us first define the goal of these new virtual constraints.
-
-\begin{defn}
-A relation \(h \in C^2\left(\mathcal{Q}\times \mathbb{R}^n ; \mathbb{R}^k\right)\) with \(h(q,p) = 0\) is a \textbf{virtual nonholonomic constraint (VNHC) of order k} if there exists a feedback control \(\tau(q,p)\) which stabilizes the constraint manifold
-\[
-\Gamma = \left\{(q,p) | h(q,p) = 0, dh_q \dot{q} + dh_p \dot{p} = 0\right\}
-\]
-\end{defn}
-Define the output of the system to be \(e = h(q,p)\).
+%\begin{align}\label{eqn:unactuated_actuated_eom}
+%\begin{split}
+%\dot{q_u} &= e_1^T \Minv(q) p \\
+%\dot{p_u} &= -p^T\frac{\partial M}{\partial q_u} p - \partial_{q_u}V(q) \\
+%\dot{q_a} &= 
+%\begin{bmatrix}
+%0 \cdots 0 \\
+%I_{n-1} \\
+%\end{bmatrix} \Minv(q) p \\
+%\dot{p_a} &= -p^T\frac{\partial M}{\partial q_a} p - \nabla{q_a}V(q) + 
+%\begin{bmatrix}
+%B_1(q) \\
+%\vdots \\
+%B_n(q)
+%\end{bmatrix} \tau
+%\end{split}
+%\end{align}
+%
+%Now we can begin to talk about Virtual Nonholonomic Constraints. In a similar fashion to what was defined for VHCs, let us first define the goal of these new virtual constraints.
+%
+%\begin{defn}
+%A relation \(h \in C^2\left(\mathcal{Q}\times \mathbb{R}^n ; \mathbb{R}^k\right)\) with \(h(q,p) = 0\) is a \textbf{virtual nonholonomic constraint (VNHC) of order k} if there exists a feedback control \(\tau(q,p)\) which stabilizes the constraint manifold
+%\[
+%\Gamma = \left\{(q,p) | h(q,p) = 0, dh_q \dot{q} + dh_p \dot{p} = 0\right\}
+%\]
+%\end{defn}
+%Define the output of the system to be \(e = h(q,p)\).
 We would like to find \(\tau(q,p)\) which drives \(e\) to zero to stabilize our constraint manifold \(\Gamma\).
 To accomplish this, we will input-output linearize (\ref{eqn:unactuated_actuated_eom}) to find \(\ddot{e} = -k_p e - k_d \dot{e}\) with \(k_p, k_d \in \mathbb{R}_{> 0}\).
-
-To characterize a certain class of VNHCs, let us make the following assumption.
-\begin{assm}\label{assm:vnhc_is_on_qu_pu}
-We assume our relation \(h\) is of the form \(h(q,p) = q_a - f(q_u,p_u)\) for some \(f \in C^2\left([\mathbb{R}]_T \times \mathbb{R} ; \mathbb{R}^{n - 1}\right)\).
-\end{assm}
-
-Now we solve for \(\tau\) by finding \(\ddot{e}\).
-\begin{align*}
-    e &= h(q,p) = qa - f(q_u,p_u)\\
-    \Rightarrow \dot{e} &= \dot{q_a} - df_{q_u}\dot{q_u} -df_{p_u}\dot{p_u} \\
-    &= [-df_{q_u} I_{n-1}]\dot{q} - df_{p_u} \dot{p_u} \\
-    &= dh_q M^{-1}(q) p - df_{p_u}\left( -\frac{1}{2}p^T \frac{\partial M^{-1}(q)}{\partial q_u} p - \partial_{q_u}V(q) \right) \\
-    &= dh_q M^{-1}(q) p + \frac{1}{2}df_{p_u} p^T \frac{\partial M^{-1}(q)}{\partial q_u} p + df_{p_u}\partial_{q_u}V(q)
-\end{align*}
-The control input \(\tau\) only appears in \(\dot{p_a}\) (see (\ref{eqn:unactuated_actuated_eom})).
+%
+%To characterize a certain class of VNHCs, let us make the following assumption.
+%\begin{assm}\label{assm:vnhc_is_on_qu_pu}
+%We assume our relation \(h\) is of the form \(h(q,p) = q_a - f(q_u,p_u)\) for some \(f \in C^2\left([\mathbb{R}]_T \times \mathbb{R} ; \mathbb{R}^{n - 1}\right)\).
+%\end{assm}
+%
+%Now we solve for \(\tau\) by finding \(\ddot{e}\).
+%\begin{align*}
+%    e &= h(q,p) = qa - f(q_u,p_u)\\
+%    \Rightarrow \dot{e} &= \dot{q_a} - df_{q_u}\dot{q_u} -df_{p_u}\dot{p_u} \\
+%    &= [-df_{q_u} I_{n-1}]\dot{q} - df_{p_u} \dot{p_u} \\
+%    &= dh_q \Minv(q) p - df_{p_u}\left( -\frac{1}{2}p^T \frac{\partial \Minv(q)}{\partial q_u} p - \partial_{q_u}V(q) \right) \\
+%    &= dh_q \Minv(q) p + \frac{1}{2}df_{p_u} p^T \frac{\partial \Minv(q)}{\partial q_u} p + df_{p_u}\partial_{q_u}V(q)
+%\end{align*}
+%The control input \(\tau\) only appears in \(\dot{p_a}\) (see (\ref{eqn:unactuated_actuated_eom})).
 To simplify the analysis, terms in \(\ddot{e}\) which do not depend on \(\dot{p_a}\) explicitly are lumped together under the symbol \((*)\):
-\begin{align*}
-    \ddot{e} &= dh_q M^{-1}(q) \dot{p} + df_{p_u} p^T \frac{\partial M^{-1}(q)}{\partial q_u} \dot{p} + (*) \\
-    &= (dh_q M^{-1}(q) + df_{p_u} p^T \frac{\partial M^{-1}(q)}{\partial q_u})B\tau + (*) \\
-    &= (dh_q M^{-1}(q) + dh_{p_u} p^T \frac{\partial M^{-1}(q)}{\partial q_u})B\tau + (*)
-\end{align*}
-
-From the derivations above, one can solve for \(\tau\) iff the matrix on the left of \(\tau\) is full rank. Thus, for systems with degree of underactuation one we give the following definition.
-\begin{defn}
-A VNHC \(h(q,p) = 0\) of order \(n - 1\) is \textbf{regular} if \(dh_{p_a} = 0\), \(dh_{q_a} = (1 \ldots 1)^T\), and 
-\[
-\text{rank}\left\{ (dh_q M^{-1}(q) + dh_{p_u} p^T \frac{\partial M^{-1}(q)}{\partial q_u})B\right\} = n - 1
-\]
-everywhere on the constraint manifold \(\Gamma\).
+%\begin{align*}
+%    \ddot{e} &= dh_q \Minv(q) \dot{p} + df_{p_u} p^T \frac{\partial \Minv(q)}{\partial q_u} \dot{p} + (*) \\
+%    &= (dh_q \Minv(q) + df_{p_u} p^T \frac{\partial \Minv(q)}{\partial q_u})B\tau + (*) \\
+%    &= (dh_q \Minv(q) + dh_{p_u} p^T \frac{\partial \Minv(q)}{\partial q_u})B\tau + (*)
+%\end{align*}
+%
+%From the derivations above, one can solve for \(\tau\) iff the matrix on the left of \(\tau\) is full rank. Thus, for systems with degree of underactuation one we give the following definition.
+%\begin{defn}
+%A VNHC \(h(q,p) = 0\) of order \(n - 1\) is \textbf{regular} if \(dh_{p_a} = 0\), \(dh_{q_a} = (1 \ldots 1)^T\), and 
+%\[
+%\text{rank}\left\{ (dh_q \Minv(q) + dh_{p_u} p^T \frac{\partial \Minv(q)}{\partial q_u})B\right\} = n - 1
+%\]
+%everywhere on the constraint manifold \(\Gamma\).
 Equivalently, a VNHC \(h\) of order \(n - 1\) is regular if it satisfies Assumption \ref{assm:vnhc_is_on_qu_pu} and system (\ref{eqn:unactuated_actuated_eom}) with output \(e = h(q,p)\) is of relative degree \(\{2,2,\ldots,2\}\) everywhere on \(\Gamma\).
-\end{defn}
-
-In general, \(\dot{e}\) is a function of \(q_u\) and \(p = (p_u,p_a)^T\).
+%\end{defn}
+%
+%In general, \(\dot{e}\) is a function of \(q_u\) and \(p = (p_u,p_a)^T\).
 Since the purpose of a regular VNHC is to fully parameterize \(\Gamma\) by \((q_u,p_u)\), it is essential that one can solve for \(p_a = p_a(q_u,p_u)\).
 Unfortunately this often cannot be done, since \(\dot{e}\) contains the quadratic term
-\[
-\frac{1}{2} df_{p_u} p^T \frac{\partial M^{-1}(q)}{\partial q_u} p 
-\]
-We can solve for \(p_a\) if this quadratic term does not exist.
-\begin{assm}\label{assm:M_is_Mqa}
-Assume \(\partial M(q) / \partial q_u = 0 \Leftrightarrow \partial M^{-1}(q) / \partial q_u = 0\)
-\end{assm}
-
-Under Assumption \ref{assm:M_is_Mqa}, we get that the rank condition for \(h(q,p)\) to be a regular VNHC reduces to \(\text{rank}\left(dh_q M^{-1} B\right) = n - 1\).
+%\[
+%\frac{1}{2} df_{p_u} p^T \frac{\partial \Minv(q)}{\partial q_u} p 
+%\]
+%We can solve for \(p_a\) if this quadratic term does not exist.
+%\begin{assm}\label{assm:M_is_Mqa}
+%Assume \(\partial M(q) / \partial q_u = 0 \Leftrightarrow \partial \Minv(q) / \partial q_u = 0\)
+%\end{assm}
+%
+%Under Assumption \ref{assm:M_is_Mqa}, we get that the rank condition for \(h(q,p)\) to be a regular VNHC reduces to \(\text{rank}\left(dh_q \Minv B\right) = n - 1\).
 This is the same rank condition as required for Virtual Holonomic Constraints.
-
-Now we solve for \(p_a\) on the constraint manifold (when \(e = \dot{e} = 0\)):
-\begin{align*}
-    \dot{e} = dh_q M^{-1}(q)p + df_{p_u} \partial_{q_u}V(q) &= 0\\
-    \Leftrightarrow dh_q M^{-1}(q)e_1 p_u + dh_q M^{-1}(q) \begin{bmatrix}
-    0 & \cdots & 0 \\
-    & I_{n-1} & \\
-    \end{bmatrix} p_a &= -df_{p_u} \partial_{q_u}V(q) \\
-\end{align*}
-\begin{align*}
-    \Leftrightarrow dh_q M^{-1}(q) \begin{bmatrix}
-    0 & \cdots & 0 \\
-    & I_{n-1} & \\
-    \end{bmatrix} p_a = -\left(df_{p_u}\partial_{q_u}V(q) + dh_q M^{-1}(q)e_1 p_u\right)
-\end{align*}
-One can linearly solve for \(p_a\) if and only if the matrix in front of it is invertible.
-
-This leads us to a natural definition.
-\begin{defn}\label{defn:solvability}
-A VNHC \(h(q,p)\) is \textbf{solvable} (NOTE: actionable? what's a good name?) if
-\[
-\text{rank}\left(dh_q M^{-1}(q) \begin{bmatrix}
-    0 & \cdots & 0 \\
-    & I_{n-1} & \\
-    \end{bmatrix}\right) = n - 1
-\]
-\end{defn}
-
-With this analysis and our new definition in hand, we can solve for the dynamics on the constraint manifold.
-
-\begin{thm}\label{thm:equation_for_pa}
-Suppose assumptions \ref{assm:vnhc_is_on_qu_pu} and \ref{assm:M_is_Mqa} hold.
-If a regular VNHC \(h(q,p) = q_a - f(q_u,p_u)\) is solvable, then the
-parameterization for \(p_a\) on the constraint manifold is given by
-\begin{align*}
-p_a &= -\left(dh_q M^{-1}(q) \begin{bmatrix}
-    0 & \cdots & 0 \\
-    & I_{n-1} & \\
-    \end{bmatrix}\right)^{-1}\left( df_{p_u}\partial_{q_u}V(q) + dh_q M^{-1}(q)e_1 p_u\right) \\
-    &=: g(q_u,p_u)
-\end{align*}
-and the constrained dynamics on \(\Gamma\) are given by
-\begin{align*}
-    \dot{q_u} &= e_1^T M^{-1}(q_a) \begin{bmatrix}
-    p_u \\
-    p_a
-    \end{bmatrix}\mid_{q_a = f(q_u,p_u), p_a = g(q_u,p_u)} \\
-    \dot{p_u} &= -\partial_{q_u} V(q_u,q_a) \mid_{q_a = f(q_u,p_u)}
-\end{align*}
-\end{thm}
-
-Theorem \ref{thm:equation_for_pa} guarantees that, on \(\Gamma\), \(q_a\) is a
-parameterized completely by \((q_u,p_u)\). Hence, the zero-dynamics on
-\(\Gamma\) are always two-dimensional regardless of the original dimension
-\(n\).
-%/========== /Virtual Nonholonomic Constraints ==========/% vim: set ts=4 sw=4
-%sts=0 et ffs=unix :
+%
+%Now we solve for \(p_a\) on the constraint manifold (when \(e = \dot{e} = 0\)):
+%\begin{align*}
+%    \dot{e} = dh_q \Minv(q)p + df_{p_u} \partial_{q_u}V(q) &= 0\\
+%    \Leftrightarrow dh_q \Minv(q)e_1 p_u + dh_q \Minv(q) \begin{bmatrix}
+%    0 & \cdots & 0 \\
+%    & I_{n-1} & \\
+%    \end{bmatrix} p_a &= -df_{p_u} \partial_{q_u}V(q) \\
+%\end{align*}
+%\begin{align*}
+%    \Leftrightarrow dh_q \Minv(q) \begin{bmatrix}
+%    0 & \cdots & 0 \\
+%    & I_{n-1} & \\
+%    \end{bmatrix} p_a = -\left(df_{p_u}\partial_{q_u}V(q) + dh_q \Minv(q)e_1 p_u\right)
+%\end{align*}
+%One can linearly solve for \(p_a\) if and only if the matrix in front of it is invertible.
+%
+%This leads us to a natural definition.
+%\begin{defn}\label{defn:solvability}
+%A VNHC \(h(q,p)\) is \textbf{solvable} (NOTE: actionable? what's a good name?) if
+%\[
+%\text{rank}\left(dh_q \Minv(q) \begin{bmatrix}
+%    0 & \cdots & 0 \\
+%    & I_{n-1} & \\
+%    \end{bmatrix}\right) = n - 1
+%\]
+%\end{defn}
+%
+%With this analysis and our new definition in hand, we can solve for the dynamics on the constraint manifold.
+%
+%\begin{thm}\label{thm:equation_for_pa}
+%Suppose assumptions \ref{assm:vnhc_is_on_qu_pu} and \ref{assm:M_is_Mqa} hold.
+%If a regular VNHC \(h(q,p) = q_a - f(q_u,p_u)\) is solvable, then the
+%parameterization for \(p_a\) on the constraint manifold is given by
+%\begin{align*}
+%p_a &= -\left(dh_q \Minv(q) \begin{bmatrix}
+%    0 & \cdots & 0 \\
+%    & I_{n-1} & \\
+%    \end{bmatrix}\right)^{-1}\left( df_{p_u}\partial_{q_u}V(q) + dh_q \Minv(q)e_1 p_u\right) \\
+%    &=: g(q_u,p_u)
+%\end{align*}
+%and the constrained dynamics on \(\Gamma\) are given by
+%\begin{align*}
+%    \dot{q_u} &= e_1^T \Minv(q_a) \begin{bmatrix}
+%    p_u \\
+%    p_a
+%    \end{bmatrix}\mid_{q_a = f(q_u,p_u), p_a = g(q_u,p_u)} \\
+%    \dot{p_u} &= -\partial_{q_u} V(q_u,q_a) \mid_{q_a = f(q_u,p_u)}
+%\end{align*}
+%\end{thm}
+%
+%Theorem \ref{thm:equation_for_pa} guarantees that, on \(\Gamma\), \(q_a\) is a
+%parameterized completely by \((q_u,p_u)\). Hence, the zero-dynamics on
+%\(\Gamma\) are always two-dimensional regardless of the original dimension
+%\(n\).
+%
+%%---------- Hamiltonian VHCS ---------%
+%\subsection{Restriction to Hamiltonian VHCs}
+%\textbf{TODO: Why can we not do the standard VHC approach for Hamiltonian? 
+%Show how to use the above to make it work}
+%
+%%/========== /Virtual Nonholonomic Constraints ==========/% 
+% vim: set tw=80 ts=4 sw=4 sts=0 et ffs=unix :

### 2020-06-10 latex/vlp.tex
Finalize the first skeleton of the thesis
This commit finishes the skeleton of the thesis, with all sections and
headings as desired. There is currently no content here.
First version of the VLP energy-gain theorem

@@
 %/========== Variable Length Pendulum ==========/%
-\section{The Variable Length Pendulum}\label{sec:vlp}
+\chapter{Application of VNHCS: The Variable Length Pendulum}\label{sec:vlp}
+\section{Motivation}
+% TODO: Why is this a good problem to study? What is the problem? What would a human do on a standing swing?
+\section{The VLP Constraint}
+% TODO: Go through the development of the constraint and prove it gains energy
+
+\begin{thm}
+For the variable-length pendulum, define \(\theta := \arctan_2(p,q)\). 
+A VNHC of the form \(l = l(\theta)\) injects energy if there exists \(l_{avg} \in \R_{>0}\) such that
+\[
+   \left(l(\theta) - l_{avg}\right)sin(2\theta) \leq 0 \text{ }\forall \theta \in \mathbb{S}^1
+\]
+with the property that the inequality is strict for almost every \(\theta\).
+\end{thm}
+\begin{proof}
+    Choose, as a candidate anti-Lyapunov function, the energy for the average-length pendulum 
+    \[
+       E_{avg}(q,p) := \frac{1}{2}\frac{p^2}{m l_{avg}^2} 
+                    + m g l_{avg} (1-\cos(q))
+    \]
+
+    which is non-negative and has derivative 
+    \[
+      \dot{E}_{avg} = \frac{-g\sin(q)p \left(l(\theta)^3 - l_{avg}^3\right)}
+                 {l_{avg}^2l(\theta)^2}
+    \]
+
+    We will show that \(E_{avg}\) is increasing.
+
+    % VBOX UNDERFULL WARNING HERE, IGNORE IT
+    Observe that \(\sign{\sin(q)p} = \sign{\sin(2\theta)}\) and, 
+    by Lemma \textbf{TODO: REF LEMMA},
+    \( \sign{l(\theta)^3 - l_{avg}^3} 
+     = \sign{ l(\theta) - l_{avg}}\).
+
+     Then the derivative of \(E_{avg}\) is almost always positive, since
+     \begin{align*}
+        \label{eq:}
+        \sign{\dot{E}_{avg}} &= \sign{-\sin(q)p \left(l(\theta)^3 - l_{avg}^3\right)} \\
+                    &= -\sign{\sin(2\theta) \left(l(\theta) - l_{avg}\right)} \\
+                    &\geq 0 \text{ (by assumption)}
+     \end{align*}
+
+      Hence, \(E_{avg}\) is an anti-Lyapunov function with positive derivative,  
+      so the variable-length pendulum is gaining energy.
+\end{proof}
+
+\section{Simulation Results}
 
 %/========== /Variable Length Pendulum ==========/%
-% vim: set ts=3 sw=3 sts=0 et ffs=unix :
+% vim: set ts=3 sw=3 sts=0 et tw=80 ffs=unix :

### 2020-06-15 latex/vnhcs.tex
Started writing the preliminaries on mechanical systems and analytical mechanics.

@@
 \chapter{Development of Virtual Nonholonomic Constraints}
-\section{Mechanical Systems}
-\subsection{Lagrangian Formulation of Mechanical Systems}
+\section{Preliminaries on Analytical Mechanics}
+In classical mechanics, the position of a point-mass is described using the
+three cartesian coordinates (one for each spatial axis) and its motion with the
+three corresponding equations of motion (EOM). The position is typically
+represented in vector format as a vector in \(\R^3\), while the EOM
+are ordinary differential equations derived from Newton's second law 
+\(F = m a\).
+Generalizing this to a mechanical system with \(N\) point-masses, 
+the position of the system can be represented by a vector in \(\R^{3N}\) and the
+motion by the \(3N\) EOM. These equations, known as the dynamics of the system, 
+are once again derived through Newton's second law \textbf{TODO: SOURCE}.
+
+While the classical approach to mechanics works well for simple systems, solving
+for the EOM of a system with more than two particles can be incredibly
+challenging - especially if the forces on the system are not known explicitly.
+The field of analytical mechanics provides a computational method for finding
+these dynamics using only the kinetic and potential energies of the system.
+
+Most mechanical systems are not just a collection of point masses which can 
+freely move about three-dimensional space; the points are usually restricted
+physically by some equations of constraint (EOC). In a mechanical system with
+\(N\) particles with \(m\) EOC, we define the value \(n = 3N - m\) and say the
+system has \(n\) degrees of freedom.
+Rather than representing the system by one vector in \(\R^{3N}\), it is often
+more desirable to represent the position of the system using \(n\)
+independent scalar-valued variables called the generalized coordinates.
+
+For example, take two planar spheres linked by an axle as in Figure
+\textbf{TODO: draw the points with tikz}. This has 2 point masses at
+\((x_1,y_1,z_1)\) and \((x_2,y_2,z_2)\) along with the three EOC 
+\((x_2 - x_1)^2 + (y_2 - y_1)^2 - 1 = 0\), \(z_1 = 0\), and \(z_2 = 0\).
+The system has 3 degrees of freedom. Instead of using cartesian coordinates, it
+is simpler to represent the system with three generalized coordinates:
+the \((x,y)\) position of the center of mass and the angle
+\(\theta\) between the axle and the horizontal axis.
+
+Generalized coordinates are labelled \(q_i, \, i \in \{1,\ldots,n\}\). 
+Each coordinate takes values in \(\Rt{T_i}\), where \(T_i = \infty\) if 
+\(q_i\) represents a length, and \(T_i = 2\pi\) if \(q_i\) represents an angle.
+It is convention to write \(q = (q_1,\ldots,q_n) \in \mathcal{Q}\) where 
+\(\mathcal{Q} = \Rt{T_1} \times \cdot \times \Rt{T_n}\) is called 
+the configuration space.
+
 % TODO: Describe the basics of the lagrangian formulation for mechanical systems
-\subsection{Hamiltonian Formulation of Mechanical Systems}
 % TODO: Describe how to get the Hamiltonian from the lagrangian, and show that
 % you can write the hamiltonian directly in certain conditions of the system
-\subsection{Simply Actuated Hamiltonian Systems}
+\section{Canonical Control Coordinates for Hamiltonian Systems}
 % TODO: Describe the change of coordinates to get into q_u/q_a mode and show

### 2020-06-17 latex/symbols.tex
Added the R-mod-T symbol and a shorthand for it.
Updated the intro to mech systems
This changes the introduction to mechanical systems, and completes the
description of Lagrangian mechanics (which may be too detailed for the
purposes of this thesis). It also adds new symbols and extra commands
for easily creating partial derivatives.

@@
    \midrule
-   \(\R^n\) & Real numbers in \(n\) dimensions \\
-   \(q_u\) & Unactuated coordinates \\
-   \(q_a\) & Actuated coordinates \\
-   \(p_u\) & Conjugate of momentum to \(q_u\) \\
-   \(p_a\) & Conjugate of momentum to \(q_a\) \\
-   A & Some really long description to see how this works when we write long text inside the table. \\
+   \(\R^n\) & Real numbers in \(n\) dimensions. \\
+   \(\R^{n \times m}\) & Real-valued matrix with \(n\) rows and \(m\) columns. \\
+   \(\Rt{T}\) & Real numbers modulo \(T > 0\), with \(\Rt{\infty} = \R\). \\
+   \(\mathbb{S}^1\) & The unit circle, equivalent to \(\Rt{2\pi}\). \\
+   \([M]_{i,j}\) & The value of row \(i\), column \(j\) for the matrix \(M\). \\
+   \(\dot{x}\) & Derivative of \(x(t,y)\) with respect to time \(t\). \\
+   \(x'\) & Derivative of \(x(t,y)\) with respect to a non-time input \(y\). \\
+   \(\mathcal{Q}\) & The configuration manifold of a mechanical system. \\
+   \(q_u\) & Unactuated coordinates. \\
+   \(q_a\) & Actuated coordinates. \\
+   \(p_u\) & Conjugate of momentum to \(q_u\). \\
+   \(p_a\) & Conjugate of momentum to \(q_a\). \\
    \bottomrule

### 2020-06-17 latex/vnhcs.tex
Updated the intro to mech systems
This changes the introduction to mechanical systems, and completes the
description of Lagrangian mechanics (which may be too detailed for the
purposes of this thesis). It also adds new symbols and extra commands
for easily creating partial derivatives.

--- 2020-06-08 latex/vnhcs.tex
@@
 \chapter{Development of Virtual Nonholonomic Constraints}
-\section{Mechanical Systems}
-\subsection{Lagrangian Formulation of Mechanical Systems}
-% TODO: Describe the basics of the lagrangian formulation for mechanical systems
-\subsection{Hamiltonian Formulation of Mechanical Systems}
+\section{Preliminaries on Analytical Mechanics}
+In classical mechanics, a mechanical system is comprised of \(N\) point masses
+representing the center of mass of a physical body, where the position of each
+mass is described using three cartesian coordinates (one for each spatial axis). 
+The system as a whole is represented by a vector in \(\R^{3N}\), and the
+dynamics of the system are computed by deriving the \(3N\) \textit{equations of
+motion} (EOM) produced by Newton's second law \(F = m a\).  
+While this method works for simple cases, it unfortunately proves impossible
+to apply to a majority of mechanical systems since most of the forces
+are not known explicitly.
+
+Most mechanical systems are not solely a collection of point masses which can 
+freely move about three-dimensional space; the masses are usually restricted
+physically by some equations of constraint (EOC). 
+Rather than representing the system by one vector in \(\R^{3N}\) and \(k\) EOC, 
+it is often feasible to represent the position of the system using \(n\)
+independent scalar-valued variables \(q_1,\ldots,q_n\) called 
+\textit{generalized coordinates}, where \(n = 3N - k\) is the number of
+\textit{degrees of freedom} (DOF) of the system.
+
+Each generalized coordinate \(q_i\) takes values in \(\Rt{T_i}\), where
+\(T_i = \infty\) if \(q_i\) represents a length or \(T_i = 2\pi\) if \(q_i\)
+represents an angle.
+It is convention to write these as a \textit{configuration} vector
+\(q = (q_1,\ldots,q_n) \in \mathcal{Q}\) where the \textit{configuration space}
+\(\mathcal{Q}\) has the following structure:
+\[
+    \mathcal{Q} = \Rt{T_1} \times \cdot \times \Rt{T_n}
+\] 
+The derivative \(\dot{q} = (\dot{q}_1,\ldots,\dot{q}_n)\) of a configuration
+is called a \textit{generalized velocity}, and the joint vector
+\((q,\dot{q})\) is called a \textit{state} of the system.
+
+The field of analytical mechanics provides a computational method for finding
+the EOM of a system in generalized coordinates. The two most common methods for
+modelling robotic systems are \textit{Lagrangian} and \textit{Hamiltonian}
+mechanics.
+
+% ---------- Lagrangian Mechanics ---------- % 
+\subsection{Lagrangian Mechanics}
+
+Lagrangian mechanics uses the kinetic energy \(T(q,\dot{q})\) and potential
+energy \(P(q)\) of the system to define the Lagrangian
+(\ref{eqn:lagrangian_general})
+\begin{equation}\label{eqn:lagrangian_general}
+    \mathcal{L}(q,\dot{q}) = T(q,\dot{q}) - P(q)
+\end{equation}
+In actuated mechanical systems, the EOM are described by \(n\) second-order
+differential equations obtained from the \textit{Euler-Lagrange
+equations} with \textit{input torques} \(\tau \in \R^m\) 
+(\ref{eqn:el_eqns_general}). 
+\begin{equation}\label{eqn:el_eqns_general}
+    \diff{}{t}\left\{ \pdiff{\mathcal{L}}{\dot{q}_i} \right\}
+    - \pdiff{\mathcal{L}}{q_i} = B_i(q) \tau
+\end{equation}
+Here, \(B_i: \mathcal{Q} \rightarrow \R^{1\times m}\) is an input vector 
+and the matrix  \(B: \mathcal{Q} \rightarrow \R^{n \times m}\) given by
+\[
+    B(q) =\begin{bmatrix}
+        - & B_1(q) & - \\
+          & \vdots & \\
+        - & B_n(q) & -
+    \end{bmatrix}
+\]
+is the \textit{input matrix} for the system. If \(m < n\), we say the system is
+\textit{underactuated} with degree of underactuation \(n - m\).
+
+Many actuated mechanical systems have quadratic kinetic energies, so that the
+Lagrangian can be written explicitly as
+\[
+    \mathcal{L}(q,\dot{q}) = \dot{q}\tpose D(q) \dot{q} - P(q)
+\]
+where the \textit{inertia matrix} \(D: \mathcal{Q} \rightarrow \R^{n\times n}\) 
+is symmetric, positive definite matrix for all \(q \in \mathcal{Q}\) and the
+potential function \(P : \mathcal{Q} \rightarrow \R\) is smooth. 
+If this is the case, the Euler-Lagrange equations
+(\ref{eqn:el_eqns_general}) reduce to (\ref{eqn:el_eqns}),
+\begin{equation}\label{eqn:el_eqns}
+    D(q)\ddot{q} + C(q,\dot{q})\dot{q} + \nabla P(q) = B(q)\tau
+\end{equation}
+where the \textit{Coriolis matrix} \(C(q,\dot{q})\) is of the form
+\[
+    [C]_{i,j} = \frac{1}{2}\sum\limits_{k = 1}^n 
+    \left(\pdiff{[D]_{i,j}}{q_k}  +
+     \pdiff{[D]_{i,k}}{q_j} -
+     \pdiff{[D]_{k,j}}{q_i}\right)\dot{q}_k
+\]
+
+% ---------- Hamiltonian Mechanics ---------- %
+\subsection{Hamiltonian Mechanics}
 % TODO: Describe how to get the Hamiltonian from the lagrangian, and show that
 % you can write the hamiltonian directly in certain conditions of the system
-\subsection{Simply Actuated Hamiltonian Systems}
+
+\section{Canonical Control Coordinates for Hamiltonian Systems}
 % TODO: Describe the change of coordinates to get into q_u/q_a mode and show

### 2020-06-23 latex/vnhcs.tex
Finished Hamiltonian (first draft) and started simply actuated systems.

@@
 \section{Preliminaries on Analytical Mechanics}
-In classical mechanics, a mechanical system is comprised of \(N\) point masses
-representing the center of mass of a physical body, where the position of each
-mass is described using three cartesian coordinates (one for each spatial axis). 
-The system as a whole is represented by a vector in \(\R^{3N}\), and the
-dynamics of the system are computed by deriving the \(3N\) \textit{equations of
-motion} (EOM) produced by Newton's second law \(F = m a\).  
-While this method works for simple cases, it unfortunately proves impossible
-to apply to a majority of mechanical systems since most of the forces
-are not known explicitly.
+In classical mechanics, a mechanical system is comprised of \(N\) point masses.
+Each point mass represents the center of mass of a physical body, where the
+position of each mass is described using three cartesian coordinates (one for
+each spatial axis).  The system as a whole is represented by a vector in
+\(\R^{3N}\), and the dynamics of the system are computed by deriving the \(3N\)
+\textit{equations of motion} (EOM) produced by Newton's second law \(F = m a\).
+While this method works for simple cases, it unfortunately proves impossible to
+apply to a majority of mechanical systems since most of the forces are not known
+explicitly.
 
@@
 It is convention to write these as a \textit{configuration} vector 
-\(q = (q_1,\ldots,q_n) \in \mathcal{Q}\) where the \textit{configuration space}
-\(\mathcal{Q}\) has the following structure:
+\(q = (q_1,\ldots,q_n) \in \mathcal{Q}\) 
+where the \textit{configuration manifold} \(\mathcal{Q}\) of the system has the
+following structure:
 \[
@@
 The derivative \(\dot{q} = (\dot{q}_1,\ldots,\dot{q}_n)\) of a configuration
-is called a \textit{generalized velocity}, and the joint vector
+is called a \textit{generalized velocity}, and the combined vector
 \((q,\dot{q})\) is called a \textit{state} of the system.
@@
 energy \(P(q)\) of the system to define the Lagrangian
-(\ref{eqn:lagrangian_general})
-\begin{equation}\label{eqn:lagrangian_general}
+(\ref{eqn:lagrangian-general})
+\begin{equation}\label{eqn:lagrangian-general}
     \mathcal{L}(q,\dot{q}) = T(q,\dot{q}) - P(q)
@@
 In actuated mechanical systems, the EOM are described by \(n\) second-order
-differential equations obtained from the \textit{Euler-Lagrange
-equations} with \textit{input torques} \(\tau \in \R^m\) 
-(\ref{eqn:el_eqns_general}). 
-\begin{equation}\label{eqn:el_eqns_general}
+ordinary differential equations (ODEs) obtained from the \textit{Euler-Lagrange
+equations} with \textit{generalized input forces} \(\tau \in \R^m\) 
+(\ref{eqn:el-eqns-general}). 
+\begin{equation}\label{eqn:el-eqns-general}
     \diff{}{t}\left\{ \pdiff{\mathcal{L}}{\dot{q}_i} \right\}
@@
 \end{equation}
-Here, \(B_i: \mathcal{Q} \rightarrow \R^{1\times m}\) is an input vector 
+Here, \(B_i: \mathcal{Q} \rightarrow \R^{1\times m}\)
 and the matrix  \(B: \mathcal{Q} \rightarrow \R^{n \times m}\) given by
@@
 Lagrangian can be written explicitly as
-\[
-    \mathcal{L}(q,\dot{q}) = \dot{q}\tpose D(q) \dot{q} - P(q)
-\]
+\begin{equation}\label{eqn:lagrangian}
+    \mathcal{L}(q,\dot{q}) = \frac{1}{2} \dot{q}\tpose D(q) \dot{q} - P(q)
+\end{equation}
 where the \textit{inertia matrix} \(D: \mathcal{Q} \rightarrow \R^{n\times n}\) 
-is symmetric, positive definite matrix for all \(q \in \mathcal{Q}\) and the
+is a symmetric, positive definite matrix for all \(q \in \mathcal{Q}\) and the
 potential function \(P : \mathcal{Q} \rightarrow \R\) is smooth. 
 If this is the case, the Euler-Lagrange equations
-(\ref{eqn:el_eqns_general}) reduce to (\ref{eqn:el_eqns}),
-\begin{equation}\label{eqn:el_eqns}
+(\ref{eqn:el-eqns-general}) reduce to (\ref{eqn:el-eqns}),
+\begin{equation}\label{eqn:el-eqns}
     D(q)\ddot{q} + C(q,\dot{q})\dot{q} + \nabla P(q) = B(q)\tau
@@
 % you can write the hamiltonian directly in certain conditions of the system
+Hamiltonian mechanics provides an equivalent representation of the EOM
+by converting the \(n\) second-order ODEs generated by Lagrangian mechanics into
+\(2n\) first-order ODEs.
 
-\section{Canonical Control Coordinates for Hamiltonian Systems}
+To do this, we first define the \textit{conjugate of momentum to \(q_i\)} by
+\begin{equation}\label{eqn:p-i}
+    p_i(q,\dot{q}) = \pdiff{\mathcal{L}}{\dot{q}_i}(q,\dot{q})
+\end{equation}
+To ease notation, we will write \(p = (p_1, \ldots, p_n)\) and say that \(p\) is
+the \textit{conjugate of momenta to \(q\)} Note that each \(p_i\) is a linear
+function of \(\dot{q}\), so one can typically solve for \(\dot{q}(q,p)\) by
+inverting the expressions from (\ref{eqn:p-i}). The combined vector \((q,p)\) is
+called the \textit{phase} of the system.
+
+Next we define the \textit{Hamiltonian} (\ref{eqn:hamiltonian-general}) 
+as the Legendre transform of the Lagrangian \textbf{SOURCE}:
+\begin{equation}\label{eqn:hamiltonian-general}
+    \mathcal{H}(q,p) = p\tpose \dot{q}(q,p) - \mathcal{L}(q,\dot{q}(q,p))
+\end{equation}
+The EOM in this framework can be shown to be the \(2n\)
+first-order equations
+\begin{equation}\label{eqn:hamiltonian-eom-general}
+    \begin{cases}
+        \dot{q} = \pdiff{\mathcal{H}}{p} \\
+        \dot{p} = -\pdiff{\mathcal{H}}{q} + B(q)\tau \\
+    \end{cases}
+\end{equation}
+where \(B(q) \in \R^{n\times m}\) is the input matrix and 
+\(\tau \in \R^m\) is the vector of generalized input forces.
+
+If the Lagrangian is given as in (\ref{eqn:lagrangian}), the conjugate of
+momenta is computed explicitly by
+\[
+    p = D(q)\dot{q}
+\]
+which gives the Hamiltonian (\ref{eqn:hamiltonian})
+\begin{equation}\label{eqn:hamiltonian}
+    \mathcal{H}(q,p) = \frac{1}{2} p\tpose D^{-1}(q) p + P(q)
+\end{equation}
+and the EOM from (\ref{eqn:hamiltonian-eom-general}) reduce to 
+(\ref{eqn:hamiltonian-eom}).
+\begin{equation}\label{eqn:hamiltonian-eom}
+    \begin{cases}
+        \dot{q} = D^{-1}(q)p \\
+        \dot{p_i} = -\frac{1}{2} p\tpose \pdiff{D^{1}}{q_i}(q) p -
+        \pdiff{P}{q_i}(q) + B_i(q) \tau \\
+    \end{cases}
+\end{equation}
+
+\section{Simply Actuated Hamiltonian Systems}
 % TODO: Describe the change of coordinates to get into q_u/q_a mode and show
@@
 % inertia, D for normal coordinates
+Given a mechanical system with phase coordinates \((q,p)\) and Hamiltonian
+(\ref{eqn:hamiltonian}), it is not in general obvious how the input forces
+\(\tau\) are affecting the conjugate of momenta since they are scaled by
+\(B(q)\), which may be quite complicated. Even if \(B(q)\) is constant, multiple
+forces may affect each \(p_i\). This section will demonstrate a useful change of
+coordinates which results in a simpler input matrix.
+
+Throughout the rest of this thesis, we will make the following assumptions.
+\begin{assm}\label{assm:B-const}
+    The input matrix \(B : \mathcal{Q} \rightarrow \R^{n\times k}\) is constant
+    and full rank.
+\end{assm}
+\begin{assm}\label{assm:B-orthogonal}
+    The input matrix is left-semi-orthogonal, whereby \(B\tpose B = \Id{k}\).
+\end{assm}
+\begin{assm}\label{assm:B-perp}
+    There exists a right-semi-orthogonal matrix 
+    \(B^\perp \in \R^{(n-k)\times n}\) for \(B\)
+    which is a left-annihilator for \(B\). That is,
+    \(B^\perp(B^\perp)\tpose = \Id{(n-k)}\) and
+    \(B^\perp B = \Zmat{(n-k) \times k}\) 
+\end{assm}
+
+Observe that Assumption \ref{assm:B-orthogonal} can be taken without loss of
+generality (WLOG). If \(B\) is a constant full-rank matrix, 
+it has a singular-value decomposition 
+\(B = U \Sigma V\tpose\) where \(U^{-1} = U\tpose \in \R^{n \times n}\), 
+\(V^{-1} = V\tpose \in \R^{n \times k}\), and \(\Sigma \in \R^{n \times k}\) is
+defined by
+\[
+    \Sigma = \begin{bmatrix}
+        \sigma_1 & 0 & \cdots & 0 & 0 \\
+        0 & \sigma_2 & \cdots & 0 & 0 \\
+        \vdots & & \ddots &  & \vdots \\
+        0 & 0 & \cdots & 0 & \sigma_k \\
+          & - & \Zmat{(n-k)\times k} & -  &\\
+    \end{bmatrix}
+\]
+where \(\sigma_i \neq 0\).
+\textbf{SOURCE}
+Defining \(T \in \R^{k \times k}\) by
+\[
+    T = \begin{bmatrix}
+        \frac{1}{\sigma_1} & 0 & \cdots & 0 \\
+        0 & \frac{1}{\sigma_2} & \cdots & 0 \\
+    \vdots & & \ddots & \vdots \\
+    0 & 0 & \cdots & \frac{1}{\sigma_k} \\
+    \end{bmatrix}
+\]
+and assigning the input forces to \(\tau = V T \hat{\tau}\), we get a new input
+matrix for \(\hat{\tau}\) given by \(\hat{B} = B V T = U \Sigma T\) 
+which is still constant and full-rank. In particular, 
+\(\hat{B}\tpose \hat{B} = T\tpose \Sigma\tpose \Sigma\tpose T = \Id{k}\).
+
+If \(k = (n-1)\) then if there is a left annihilator \(A\) for
+\(B\), then \(B^\perp = A/\norm{A}\) satisfies Assumption
+\ref{assm:B-perp}.
+
+To perform our coordinate transformation, we define a transformation matrix
+\[
+    \mathbf{B} = 
+    \begin{bmatrix}
+        B^\perp \\
+        B\tpose \\
+    \end{bmatrix}
+\]
+Since both \(B^\perp\) and \(B\tpose\) are right-semi-orthogonal, 
+\(\mathbf{B}\) is orthogonal. That is,
+\(\mathbf{B}\mathbf{B}\tpose = \Id{n}\) which means
+\(\mathbf{B}^{-1} = \mathbf{B}\tpose\).
+
+\begin{defn}
+    We call \(q_u = B^\perp q\) the \textbf{unactuated coordinates} and 
+    \(q_a = B\tpose\) the \textbf{actuated coordinates}, which are obtained
+    through the transformation
+    \[
+        \begin{bmatrix} q_u \\ q_a \end{bmatrix} = \mathbf{B}q
+    \]
+\end{defn}
 

### 2020-06-24 latex/symbols.tex
Added new symbols and commands for identity and zero matrices
Updated symbols again

--- 2020-06-08 latex/symbols.tex
@@
    \midrule
-   \(\R^n\) & Real numbers in \(n\) dimensions \\
-   \(q_u\) & Unactuated coordinates \\
-   \(q_a\) & Actuated coordinates \\
-   \(p_u\) & Conjugate of momentum to \(q_u\) \\
-   \(p_a\) & Conjugate of momentum to \(q_a\) \\
-   A & Some really long description to see how this works when we write long text inside the table. \\
+   \(\R^n\) & Real numbers in \(n\) dimensions. \\
+   \(\Rt{T}\) & Real numbers modulo \(T > 0\), with \(\Rt{\infty} = \R\). \\
+   \(\mathbb{S}^1\) & The unit circle, equivalent to \(\Rt{2\pi}\). \\
+   \(\R^{n \times m}\) & The space of real-valued matrices with \(n\) rows and \(m\) columns. \\
+   \(\Id{n}\) & The \(n \times n\) identity matrix. \\
+   \(\Zmat{n \times m}\) & The \(n \times m\) matrix of all zeros. \\
+   \(M_{i,j}\) & The value of row \(i\), column \(j\) for the matrix \(M\). \\
+   \(A_i\) & If \(A\) is a vector, the \(i\)th element of \(M\). \\
+         & If \(A\) is a matrix, the \(i\)th row of \(M\). \\
+   \([v_i]\) & If \(v_i \in \R\), the column vector with value \(v_i\) at position \(i\). \\
+       & If \(v_i\) is a row vector, the matrix with row \(i\) given by \(v_i\). \\
+   \(\dot{x}\) & Derivative of \(x\) with respect to time \(t\). \\
+   \(\nabla_v F\) & Gradient of \(F)\) with respect to the vector \(v\). \\
    \bottomrule

### 2020-06-24 latex/vnhcs.tex
Continued writing simply actuated mech systems section, almost there.

@@
 and the matrix  \(B: \mathcal{Q} \rightarrow \R^{n \times m}\) given by
+%\[
+%    B(q) =\begin{bmatrix}
+%        - & B_1(q) & - \\
+%          & \vdots & \\
+%        - & B_n(q) & -
+%    \end{bmatrix}
+%\]
 \[
-    B(q) =\begin{bmatrix}
-        - & B_1(q) & - \\
-          & \vdots & \\
-        - & B_n(q) & -
-    \end{bmatrix}
+    B(q) = [B_i(q)]
 \]
@@
     [C]_{i,j} = \frac{1}{2}\sum\limits_{k = 1}^n 
-    \left(\pdiff{[D]_{i,j}}{q_k}  +
-     \pdiff{[D]_{i,k}}{q_j} -
-     \pdiff{[D]_{k,j}}{q_i}\right)\dot{q}_k
+    \left(\pdiff{D_{i,j}}{q_k}  +
+     \pdiff{D_{i,k}}{q_j} -
+     \pdiff{D_{k,j}}{q_i}\right)\dot{q}_k
 \]
@@
 The EOM in this framework can be shown to be the \(2n\)
-first-order equations
-\begin{equation}\label{eqn:hamiltonian-eom-general}
+first-order equations called \textit{Hamilton's equations}
+\begin{equation}\label{eqn:hamiltons-eqns}
     \begin{cases}
-        \dot{q} = \pdiff{\mathcal{H}}{p} \\
-        \dot{p} = -\pdiff{\mathcal{H}}{q} + B(q)\tau \\
+        \dot{q} = \nabla_p\mathcal{H} \\
+        \dot{p} = -\nabla_q\mathcal{H} + B(q)\tau \\
     \end{cases}
@@
 where \(B(q) \in \R^{n\times m}\) is the input matrix and 
-\(\tau \in \R^m\) is the vector of generalized input forces.
+\(\tau \in \R^m\) is the vector of generalized input forces. A pair of
+coordinates \((q,p)\) which satisfy Hamilton's equations are said to be 
+\textit{canonical coordinates for} \(\mathcal{H}\).
 
@@
 \end{equation}
-and the EOM from (\ref{eqn:hamiltonian-eom-general}) reduce to 
+and the EOM from (\ref{eqn:hamiltons-eqns}) reduce to 
 (\ref{eqn:hamiltonian-eom}).
@@
         \dot{q} = D^{-1}(q)p \\
-        \dot{p_i} = -\frac{1}{2} p\tpose \pdiff{D^{1}}{q_i}(q) p -
-        \pdiff{P}{q_i}(q) + B_i(q) \tau \\
+        \dot{p} = -\frac{1}{2} \left[p\tpose \pdiff{D^{-1}}{q_i}(q) p\right] 
+        - \nabla_q P(q) + B(q) \tau \\
     \end{cases}
@@
 % inertia, D for normal coordinates
-Given a mechanical system with phase coordinates \((q,p)\) and Hamiltonian
-(\ref{eqn:hamiltonian}), it is not in general obvious how the input forces
+Given a mechanical system with Lagrangian (\ref{eqn:lagrangian}) yielding 
+Hamiltonian (\ref{eqn:hamiltonian}) and canonical coordinates \((q,p)\), 
+it is not in general obvious how the input forces
 \(\tau\) are affecting the conjugate of momenta since they are scaled by
-\(B(q)\), which may be quite complicated. Even if \(B(q)\) is constant, multiple
+\(B(q)\) (which may be quite complicated). Even if \(B(q)\) is constant, multiple
 forces may affect each \(p_i\). This section will demonstrate a useful change of
-coordinates which results in a simpler input matrix.
+coordinates which results in a new Hamiltonian mechanical system with simpler
+input matrix.
 
@@
 \]
-Since both \(B^\perp\) and \(B\tpose\) are right-semi-orthogonal, 
-\(\mathbf{B}\) is orthogonal. That is,
-\(\mathbf{B}\mathbf{B}\tpose = \Id{n}\) which means
+Since both \(B^\perp\) and \(B\tpose\) are right-semi-orthogonal and
+\(\mathbf{B} \in \R^{n \times n}\) is full rank, \(\mathbf{B}\) is orthogonal. 
+That is, \(\mathbf{B}\mathbf{B}\tpose = \Id{n}\) which means 
 \(\mathbf{B}^{-1} = \mathbf{B}\tpose\).
@@
 \begin{defn}
-    We call \(q_u = B^\perp q\) the \textbf{unactuated coordinates} and 
-    \(q_a = B\tpose\) the \textbf{actuated coordinates}, which are obtained
-    through the transformation
+    Define \(q_u = B^\perp q \in \R^{(n-k)}\) and 
+    \(q_a = B\tpose q \in \R^k\). We say \(q_u\) are the \textit{unactuated
+    coordinates} while \(q_a\) are the \textit{actuated coordinates} of the
+    system. Their corresponding conjugate of momenta are labelled \(p_u\) and
+    \(p_a\).
+\end{defn}
+
+In these new coordinates the input matrix will be modified such
+that \(\tau_i\) only affects \(p_{a,i}\), and all \(p_u\) are unaffected by the
+inputs.
+
+To derive this result, observe that the coordinates 
+\(\hat{q} = (q_u,q_a)\) can be written 
+as the transformation (\ref{eqn:q-transformation}).
+\begin{equation}\label{eqn:q-transformation}
+    \hat{q} = \mathbf{B}q
+\end{equation}
+Since \(\mathbf{B}\) is
+an orthogonal invertible matrix, we can equivalently write this as
+\(q = \mathbf{B}\tpose \hat{q}\). Applying this change of coordinates to the
+Lagrangian (\ref{eqn:lagrangian}) gives the new Lagrangian
 \[
-        \begin{bmatrix} q_u \\ q_a \end{bmatrix} = \mathbf{B}q
+    \hat{\mathcal{L}}(\hat{q},\dot{\hat{q}}) = 
+    \frac{1}{2}\dot{\hat{q}}\tpose \mathbf{B} 
+    D(\mathbf{B}\tpose\hat{q}) 
+    \mathbf{B}\tpose\dot{\hat{q}} - P(\mathbf{B}\hat{q})
 \]
-\end{defn}
+From here it is straightforward to show that the 
+conjugate of momenta \(\hat{p} = (p_u,p_a)\) to \(\hat{q}\) is given by 
+(\ref{eqn:p-transformation}).
+\begin{equation}\label{eqn:p-transformation}
+    \hat{p} = \pdiff{\hat{\mathcal{L}}}{\dot{\hat{q}}}
+            = \mathbf{B}D(q)\dot{q} = \mathbf{B}p
+\end{equation}
+
+For ease of notation, we define
+\(M(\hat{q}) := \mathbf{B}D\left(\mathbf{B}\tpose\hat{q}\right)\mathbf{B}\tpose\)
+and 
+\(V(\hat{q}) := P\left(\mathbf{B}\tpose \hat{q}\right)\)
+Then the Hamiltonian (\ref{eqn:hamiltonian}) is given in
+\((\hat{q},\hat{p})\) coordinates by
+\begin{align*}
+    \hat{\mathcal{H}}(\hat{q},\hat{p})
+    &=
+    \frac{1}{2}\hat{p}\tpose\mathbf{B}
+    D^{-1}\left(\mathbf{B}\tpose\hat{q}\right)
+    \mathbf{B}\tpose\hat{p}
+    + P\left(\mathbf{B}\hat{q}\right) \\
+    &= \hat{p}\tpose \Minv(\hat{q})\hat{p} + V(\hat{q})
+\end{align*}
+
+Now we will show \((\hat{q},\hat{p})\) are canonical coordinates
+for \(\hat{\mathcal{H}}\). From the definition of \(\hat{q}\) in
+(\ref{eqn:q-transformation}), we obtain
+\begin{align*}
+    \dot{\hat{q}} &= \mathbf{B}\dot{q} = \mathbf{B}D^{-1}(q)p \\
+        &= \mathbf{B}D^{-1}(\mathbf{B}\tpose \hat{q})\mathbf{B}\tpose \hat{p} \\
+        &= \Minv(\hat{q})\hat{p} \\
+        &= \nabla_{\hat{p}}\hat{\mathcal{H}}
+\end{align*}
+
+The dynamics of \(\hat{p}\) are given by
+\[
+    \dot{\hat{p}}_i = \mathbf{B}\dot{p}
+                    = -\frac{1}{2}\mathbf{B}
+                    \left[p\tpose \pdiff{D^{-1}}{q_i}p\right]
+                    -\mathbf{B} \nabla_q P(q) + \mathbf{B}B \tau
+\]
+where 
+\[
+    \mathbf{B} \nabla_q P(q)
+    = \nabla_{\hat{q}}(\mathbf{B}\tpose \hat{q})\nabla_q P(q)
+    = \nabla_{\hat{q}}q \nabla_q P(q)
+    = \nabla_{\hat{q}}P(\mathbf{B}\tpose\hat{q}) 
+    = \nabla_{\hat{q}} V(\hat{q}) 
+\]
+
+\textbf{LOOK INTO SYMPLECTIc jacobian method for canonical transformations to
+show this is a canonical transformation}
 

### 2020-06-26 latex/symbols.tex
Added symbols for the kronecker delta.

@@
    \(\Zmat{n \times m}\) & The \(n \times m\) matrix of all zeros. \\
+   \(M_i\) & If \(M\) is a vector, the \(i\)th element of \(M\). \\
+         & If \(M\) is a matrix, the \(i\)th row of \(M\). \\
    \(M_{i,j}\) & The value of row \(i\), column \(j\) for the matrix \(M\). \\
-   \(A_i\) & If \(A\) is a vector, the \(i\)th element of \(M\). \\
-         & If \(A\) is a matrix, the \(i\)th row of \(M\). \\
    \([v_i]\) & If \(v_i \in \R\), the column vector with value \(v_i\) at position \(i\). \\
-       & If \(v_i\) is a row vector, the matrix with row \(i\) given by \(v_i\). \\
+      & If \(v_i \in \R^{1\times n}\), the matrix with row \(i\) given by \(v_i\). \\
    \(\dot{x}\) & Derivative of \(x\) with respect to time \(t\). \\
    \(\nabla_v F\) & Gradient of \(F)\) with respect to the vector \(v\). \\
+   \(\delta_{i,j}\) & The Kronecker delta: 1 if \(i = j\) and 0 otherwise. \\
    \bottomrule
@@
 
+%---------- Nomenclature ----------%
+%\chapter*{Nomenclature}
+%
+%% Define a table for the symbols using the booktabs convention
+%\noindent
+%\begin{table}[h!]
+%\resizebox{\textwidth}{!}{%
+%\centering
+%\begin{tabular}{ll}
+%   \toprule
+%   Acronym & Meaning \\
+%   \midrule
+%   DOF & Degrees of freedom
+%   EOC & Equation of constraint \\
+%   EOM & Equation of motion \\
+%   ODE & Ordinary differential equation \\
+%   VNHC & Virtual nonholonomic constraint \\
+%   WLOG & Without loss of generality \\
+%   \bottomrule
+%\end{tabular}
+%
+%}% resizebox
+%\end{table}
 %/========== /Symbols ==========/%

### 2020-06-26 latex/vnhcs.tex
Finished a base draft of simply actuated mech systems

--- 2020-06-23 latex/vnhcs.tex
@@
 \textit{generalized coordinates}, where \(n = 3N - k\) is the number of
-\textit{degrees of freedom} (DOF) of the system.
+\textit{degrees of freedom} (DOF) of the system \cite{greenwood_dynamics}.
 
@@
 energy \(P(q)\) of the system to define the Lagrangian 
-(\ref{eqn:lagrangian-general})
+(\ref{eqn:lagrangian-general}) \cite{greenwood_dynamics}:
 \begin{equation}\label{eqn:lagrangian-general}
@@
 ordinary differential equations (ODEs) obtained from the \textit{Euler-Lagrange
-equations} with \textit{generalized input forces} \(\tau \in \R^m\) 
+equations} with \textit{generalized input forces} \(\tau \in \R^k\) 
 (\ref{eqn:el-eqns-general}). 
@@
 \end{equation}
-Here, \(B_i: \mathcal{Q} \rightarrow \R^{1\times m}\)
-and the matrix  \(B: \mathcal{Q} \rightarrow \R^{n \times m}\) given by
+Here, \(B_i: \mathcal{Q} \rightarrow \R^{1\times k}\)
+and the matrix  \(B: \mathcal{Q} \rightarrow \R^{n \times k}\) with
 \[
-    B(q) =\begin{bmatrix}
-        - & B_1(q) & - \\
-          & \vdots & \\
-        - & B_n(q) & -
-    \end{bmatrix}
+    B(q) = [B_i(q)]
 \]
-is the \textit{input matrix} for the system. If \(m < n\), we say the system is
-\textit{underactuated} with degree of underactuation \(n - m\).
+is called the \textit{input matrix} for the system. 
+If \(k < n\), we say the system is \textit{underactuated} with degree of
+underactuation \(n - k\).
 
@@
 potential function \(P : \mathcal{Q} \rightarrow \R\) is smooth. 
-If this is the case, the Euler-Lagrange equations
-(\ref{eqn:el-eqns-general}) reduce to (\ref{eqn:el-eqns}),
+If this is the case, the Euler-Lagrange equations reduce to (\ref{eqn:el-eqns}),
 \begin{equation}\label{eqn:el-eqns}
@@
 \subsection{Hamiltonian Mechanics}
-% TODO: Describe how to get the Hamiltonian from the lagrangian, and show that
-% you can write the hamiltonian directly in certain conditions of the system
 Hamiltonian mechanics provides an equivalent representation of the EOM
@@
 
-To do this, we first define the \textit{conjugate of momentum to \(q_i\)} by
+To do this, we first define the \textit{conjugate of momentum \(p_i\) to \(q_i\)} by
 \begin{equation}\label{eqn:p-i}
@@
 Next we define the \textit{Hamiltonian} (\ref{eqn:hamiltonian-general}) 
-as the Legendre transform of the Lagrangian \textbf{SOURCE}:
+as the Legendre transform of the Lagrangian \cite{landau_mechanics}:
 \begin{equation}\label{eqn:hamiltonian-general}
@@
 The EOM in this framework can be shown to be the \(2n\)
-first-order equations
-\begin{equation}\label{eqn:hamiltonian-eom-general}
+first-order equations called \textit{Hamilton's equations}
+\begin{equation}\label{eqn:hamiltons-eqns}
     \begin{cases}
-        \dot{q} = \pdiff{\mathcal{H}}{p} \\
-        \dot{p} = -\pdiff{\mathcal{H}}{q} + B(q)\tau \\
+        \dot{q} = \nabla_p\mathcal{H} \\
+        \dot{p} = -\nabla_q\mathcal{H} + B(q)\tau \\
     \end{cases}
 \end{equation}
-where \(B(q) \in \R^{n\times m}\) is the input matrix and 
-\(\tau \in \R^m\) is the vector of generalized input forces.
+where \(B(q) \in \R^{n\times k}\) is the input matrix and 
+\(\tau \in \R^k\) is the vector of generalized input forces. A pair of
+coordinates \((q,p)\) which satisfy Hamilton's equations are said to be 
+\textit{canonical coordinates for} \(\mathcal{H}\). a change of coordinates
+\((\hat{q}(q,p),\hat{p}(q,p))\) is said to be a \textit{canonical
+transformation} if the EOM \((\dot{\hat{q}},\dot{\hat{p}})\) satisfy Hamilton's
+equations for some other Hamiltonian \(\hat{H}\).
 
-If the Lagrangian is given as in (\ref{eqn:lagrangian}), the conjugate of
-momenta is computed explicitly by
+If the Kinetic energy is quadratic as in (\ref{eqn:lagrangian}), the conjugate of
+momenta to \(q\) can be computed explicitly by
 \[
@@
 \]
-which gives the Hamiltonian (\ref{eqn:hamiltonian})
-\begin{equation}\label{eqn:hamiltonian}
-    \mathcal{H}(q,p) = \frac{1}{2} p\tpose D^{-1}(q) p + P(q)
-\end{equation}
-and the EOM from (\ref{eqn:hamiltonian-eom-general}) reduce to 
-(\ref{eqn:hamiltonian-eom}).
-\begin{equation}\label{eqn:hamiltonian-eom}
-    \begin{cases}
+The resulting Hamiltonian system reduces to (\ref{eqn:hamiltonian}).
+\begin{align}\label{eqn:hamiltonian}
+    \mathcal{H}(q,p) &= \frac{1}{2} p\tpose D^{-1}(q) p + P(q) \\
+                     &\begin{cases}
         \dot{q} = D^{-1}(q)p \\
-        \dot{p_i} = -\frac{1}{2} p\tpose \pdiff{D^{1}}{q_i}(q) p -
-        \pdiff{P}{q_i}(q) + B_i(q) \tau \\
+        \dot{p} = -\frac{1}{2} \left[p\tpose \pdiff{D^{-1}}{q_i}(q) p\right] 
+        - \nabla_q P(q) + B(q) \tau \\
     \end{cases} 
-\end{equation}
+\end{align}
 
@@
 % inertia, D for normal coordinates
-Given a mechanical system with phase coordinates \((q,p)\) and Hamiltonian
-(\ref{eqn:hamiltonian}), it is not in general obvious how the input forces
-\(\tau\) are affecting the conjugate of momenta since they are scaled by
-\(B(q)\), which may be quite complicated. Even if \(B(q)\) is constant, multiple
-forces may affect each \(p_i\). This section will demonstrate a useful change of
-coordinates which results in a simpler input matrix.
+Given an arbitrary Hamiltonian mechanical system (\ref{eqn:hamiltonian}) in
+canonical coordinates \((q,p)\), it is not obvious how the input forces \(\tau\)
+are affecting the conjugate of momenta \(p_i\). This is because \(\tau\) is
+transformed by \(B(q)\), which may be quite complicated.
 
-Throughout the rest of this thesis, we will make the following assumptions.
+For this reason, we will define a new class of Hamiltonian systems with simple
+input matrices, so that the effect of the input forces on the conjugate of
+momenta is obvious. This class of systems will be the backbone for the rest of
+the theory developed in this thesis.
+
+\begin{defn}
+    A Hamiltonian mechanical system (\ref{eqn:hamiltonian}) is said to be 
+    \textit{simply actuated} if the input matrix \(B(q) \in \R^{n \times k}\)
+    is always of the form
+    \[
+        B(q) = \begin{bmatrix}
+            \Zmat{(n-k)\times k} \\
+            \Id{k}
+        \end{bmatrix}
+    \]
+\end{defn}
+
+Under the following mild assumptions on the input matrix, we will show that
+there exists a canonical transformation under which the system is simply
+actuated.
+
 \begin{assm}\label{assm:B-const}
@@
 \end{assm}
-\begin{assm}\label{assm:B-orthogonal}
-    The input matrix is left-semi-orthogonal, whereby \(B\tpose B = \Id{k}\).
-\end{assm}
 \begin{assm}\label{assm:B-perp}
     There exists a right-semi-orthogonal matrix 
-    \(B^\perp \in \R^{(n-k)\times n}\) for \(B\)
+    \(B^\perp \in \R^{(n-k)\times n}\)
     which is a left-annihilator for \(B\). That is,
-    \(B^\perp(B^\perp)\tpose = \Id{(n-k)}\) and
-    \(B^\perp B = \Zmat{(n-k) \times k}\) 
+    \(B^\perp(B^\perp)\tpose = \Id{(n-k)}\) (right semi-orthogonality) and
+    \(B^\perp B = \Zmat{(n-k) \times k}\) (left-annihilation).
 \end{assm}
 
-Observe that Assumption \ref{assm:B-orthogonal} can be taken without loss of
-generality (WLOG). If \(B\) is a constant full-rank matrix, 
+Note that if \(k = (n-1)\), the existence of a left annihilator \(A^0\) for
+\(B\) implies \(B^\perp := A^0/\norm{A^0}\) satisfies Assumption
+\ref{assm:B-perp}.
+
+\begin{assm}\label{assm:B-orthogonal}
+    WLOG, the input matrix \(B\) is left semi-orthogonal. 
+    That is, \(B\tpose B = \Id{k}\).
+\end{assm}
+\begin{proof}
+Since \(B\) is a constant full-rank matrix, 
 it has a singular-value decomposition 
@@
 \(\hat{B}\tpose \hat{B} = T\tpose \Sigma\tpose \Sigma\tpose T = \Id{k}\).
+\end{proof}
 
-If \(k = (n-1)\) then if there is a left annihilator \(A\) for
-\(B\), then \(B^\perp = A/\norm{A}\) satisfies Assumption
-\ref{assm:B-perp}.
-
-To perform our coordinate transformation, we define a transformation matrix
+Let \(\mathbf{B} \in \R^{n\times n}\) be the following matrix:
 \[
@@
 \]
-Since both \(B^\perp\) and \(B\tpose\) are right-semi-orthogonal, 
-\(\mathbf{B}\) is orthogonal. That is,
-\(\mathbf{B}\mathbf{B}\tpose = \Id{n}\) which means
-\(\mathbf{B}^{-1} = \mathbf{B}\tpose\).
+Since \(B^\perp\) is a left annihilator or \(B\) and both \(B^\perp\) and
+\(B\tpose\) are right semi-orthogonal, it is easy to show that \(\mathbf{B}\) is
+orthogonal (i.e. it is full-rank and has inverse equal to its transpose):
+\[
+    \mathbf{B}\mathbf{B}\tpose = 
+    \begin{bmatrix}
+        B^\perp (B^\perp)\tpose & B^\perp B \\
+        (B^\perp B)\tpose & B\tpose B
+    \end{bmatrix} = \Id{n}
+    \Rightarrow
+    \mathbf{B}^{-1} = \mathbf{B}\tpose
+\]
 
 \begin{defn}
-    We call \(q_u = B^\perp q\) the \textbf{unactuated coordinates} and 
-    \(q_a = B\tpose\) the \textbf{actuated coordinates}, which are obtained
-    through the transformation
+    Define the coordinate transformation 
+    \((q_u,q_a) = \mathbf{B}q\), where \(q_u = B^\perp q\)
+    and \(q_a = B\tpose q\).
+    We call \(q_u\) the \textit{unactuated coordinates} and \(q_a\) the
+    \textit{actuated coordinates} of the system. 
+    Their corresponding conjugate of momenta are 
+    \((p_u,p_a) = \mathbf{B}p\).
+\end{defn}
+
+In these new coordinates, the generalized input forces directly affect the
+actuated conjugate of momenta. That is, \(\tau_i\) only affects \(p_{a,i}\), 
+and all \(p_u\) are unaffected by the input forces. We prove this fact through
+the following theorem.
+
+\begin{thm}
+    Under Assumptions \ref{assm:B-const},\ref{assm:B-perp}, and
+    \ref{assm:B-orthogonal}, the change of coordinates 
+    \((Q = (q_u, q_a), P = (p_u,p_a))\) 
+    is a canonical transformation of the
+    original Hamiltonian system (\ref{eqn:hamiltonian}) into
+    the simply actuated Hamiltonian system (\ref{eqn:simple-hamiltonian}),
+    \begin{align}\label{eqn:simple-hamiltonian}
+        \hat{\mathcal{H}}(Q,P) &= 
+        \frac{1}{2} P\tpose \Minv(Q) P + V(Q) \\
+       &\begin{cases}
+            \dot{Q} = \Minv(Q)P \\
+            \dot{P} = -\frac{1}{2}
+                \left[P\tpose \pdiff{\Minv}{Q_i}(Q) P \right] 
+                - \nabla_Q V(Q) + 
+                \begin{bmatrix}
+                    \Zmat{(n-k)\times k} \\
+                    \Id{k} \\
+                \end{bmatrix} \tau
+        \end{cases} \nonumber
+    \end{align}
+    where
+    \begin{align*}
+        \Minv(Q) &= \mathbf{B}D^{-1}(\mathbf{B}\tpose Q)\mathbf{B}\tpose \\
+        V(Q) &= P(\mathbf{B}\tpose Q) \\
+    \end{align*}
+\end{thm}
+\begin{proof}
+    For any constant matrix \(A\), the transformation
+    \((Q = Aq, P = Ap)\) satisfies
+    \(\pdiff{Q_i}{p_s} = \pdiff{P_i}{q_s} = 0\) for all 
+    \(i,s \in \{1,\ldots,n\}\). Hence, the Poisson brackets between the
+    new coordinates are zero:
+    \begin{align*}
+        [Q_i,Q_j] &:= \sum\limits_{s = 1}^n \pdiff{Q_i}{p_s}\pdiff{Q_j}{q_s} - 
+        \pdiff{Q_i}{q_s}\pdiff{Q_j}{p_s} = 0 \\
+        [P_i,P_j] &:= \sum\limits_{s=1}^n \pdiff{P_i}{p_s}\pdiff{P_j}{q_s} -
+        \pdiff{P_i}{q_s}\pdiff{P_j}{p_s} = 0
+    \end{align*}
+    Since the matrix \(A = \mathbf{B}\) is invertible and orthogonal,
+    \(A_i (A\tpose_j)\tpose = A_i (A^{-1}_j)\tpose = \delta_{i,j}\). Using this
+    fact we see that the Poisson brackets between the new momenta and
+    configurations are given by:
+    \begin{align*}
+        [P_i,Q_j] &= \sum\limits_{s=1}^n\pdiff{P_i}{p_s}\pdiff{Q_j}{q_s}
+        - \pdiff{P_i}{q_s}\pdiff{Q_j}{p_s} \\
+                  &= \sum\limits{s=1}^n A_{i,s}A_{j,s} - 0 \\
+                  &= \sum\limits{s=1}^n A_{i,s}A\tpose_{s,j} \\
+                  &= A_i (A\tpose_j)\tpose \\
+                  &= \delta_{i,j}
+    \end{align*}
+
+    By (45.10) in \cite{landau_mechanics}, the coordinate change
+    \((Q = \mathbf{B}q, P = \mathbf{B}p)\) is a canonical transformation with
+    new Hamiltonian 
+    \(\hat{H}(Q,P) = \mathcal{H}(\mathbf{B}\tpose Q,\mathbf{B}\tpose P)\).
+
+    Furthermore, since \(\dot{P} = \mathbf{B}\dot{p}\), the new input matrix is
+    given by 
     \[
-        \begin{bmatrix} q_u \\ q_a \end{bmatrix} = \mathbf{B}q
+        \hat{B} = \mathbf{B}B = \begin{bmatrix}
+            B^\perp B \\
+            B\tpose B \\
+        \end{bmatrix} = 
+        \begin{bmatrix}
+            \Zmat{(n-k)\times k} \\
+            \Id{k}
+        \end{bmatrix}
     \]
-\end{defn}
+    so the new Hamiltonian system is simply actuated, as desired.
+\end{proof}
 

### 2020-06-29 latex/vnhcs.tex
Edited the preliminaries to use B_i a column and rephrased a lot of stuff.

@@
 \section{Preliminaries on Analytical Mechanics}
-In classical mechanics, a mechanical system is comprised of \(N\) point masses.
-Each point mass represents the center of mass of a physical body, where the
-position of each mass is described using three cartesian coordinates (one for
-each spatial axis).  The system as a whole is represented by a vector in
-\(\R^{3N}\), and the dynamics of the system are computed by deriving the \(3N\)
+A mechanical system can be represented by \(N\) point masses, where each point
+represents the center of mass of a physical body, along with \(r\)
+\textit{equations of constraint} (EOC) which model the physical restrictions
+between these masses.
+The position of each point mass is described using three cartesian coordinates (one
+for each spatial axis), so the system as a whole can be described by a vector in
+\(\R^{3N}\) with \(r\) EOC. 
+The dynamics of the system are computed by deriving the \(3N\)
 \textit{equations of motion} (EOM) produced by Newton's second law \(F = m a\).
-While this method works for simple cases, it unfortunately proves impossible to
-apply to a majority of mechanical systems since most of the forces are not known
-explicitly.
+While this technique works for simple systems, it is impossible to
+apply to a majority of mechanical systems since most of the forces 
+on the system are not explicitly known.
 
-Most mechanical systems are not solely a collection of point masses which can 
-freely move about three-dimensional space; the masses are usually restricted
-physically by some equations of constraint (EOC). 
-Rather than representing the system by one vector in \(\R^{3N}\) and \(k\) EOC, 
+Rather than modeling a mechanical system with point masses and constraints,
 it is often feasible to represent the position of the system using \(n\)
 independent scalar-valued variables \(q_1,\ldots,q_n\) called 
-\textit{generalized coordinates}, where \(n = 3N - k\) is the number of
+\textit{generalized coordinates}, where \(n = 3N - m\) is the number of
 \textit{degrees of freedom} (DOF) of the system \cite{greenwood_dynamics}.
@@
 represents an angle.
-It is convention to write these as a \textit{configuration} vector 
+It is convention to collect the coordinates into a \textit{configuration} 
 \(q = (q_1,\ldots,q_n) \in \mathcal{Q}\) 
@@
 The field of analytical mechanics provides a computational method for finding
-the EOM of a system in generalized coordinates. The two most common methods for
-modelling robotic systems are \textit{Lagrangian} and \textit{Hamiltonian}
-mechanics.
+the EOM of a system in generalized coordinates. The two most common analytical
+methods for modelling robotic systems are \textit{Lagrangian} and
+\textit{Hamiltonian} mechanics.
 
@@
 energy \(P(q)\) of the system to define the Lagrangian 
-(\ref{eqn:lagrangian-general}) \cite{greenwood_dynamics}:
+\(\mathcal{L} : \mathcal{Q}\times\R^n \rightarrow \R\) defined by
+(\ref{eqn:lagrangian-general}) \cite{greenwood_dynamics}.
 \begin{equation}\label{eqn:lagrangian-general}
@@
 \end{equation}
-In actuated mechanical systems, the EOM are described by \(n\) second-order
+When the mechanical system is actuated, the EOM are described by \(n\) second-order
 ordinary differential equations (ODEs) obtained from the \textit{Euler-Lagrange
@@
     \diff{}{t}\left\{ \pdiff{\mathcal{L}}{\dot{q}_i} \right\}
-    - \pdiff{\mathcal{L}}{q_i} = B_i(q) \tau
+    - \pdiff{\mathcal{L}}{q_i} = B_i\tpose(q) \tau
 \end{equation}
-Here, \(B_i: \mathcal{Q} \rightarrow \R^{1\times k}\)
-and the matrix  \(B: \mathcal{Q} \rightarrow \R^{n \times k}\) with
+The vector \(B_i\tpose: \mathcal{Q} \rightarrow \R^{1\times k}\) describes how
+the input forces shape the dynamics of \(q_i\).
+The matrix  \(B: \mathcal{Q} \rightarrow \R^{n \times k}\) with
 \[
-    B(q) = [B_i(q)]
+    B(q) = [B_i\tpose(q)]
 \]
@@
 potential function \(P : \mathcal{Q} \rightarrow \R\) is smooth. 
-If this is the case, the Euler-Lagrange equations reduce to (\ref{eqn:el-eqns}),
-\begin{equation}\label{eqn:el-eqns}
-    D(q)\ddot{q} + C(q,\dot{q})\dot{q} + \nabla P(q) = B(q)\tau
-\end{equation}
-where the \textit{Coriolis matrix} \(C(q,\dot{q})\) is of the form
-\[
-    [C]_{i,j} = \frac{1}{2}\sum\limits_{k = 1}^n 
-    \left(\pdiff{D_{i,j}}{q_k}  +
-     \pdiff{D_{i,k}}{q_j} -
-     \pdiff{D_{k,j}}{q_i}\right)\dot{q}_k
-\]
+%If this is the case, the Euler-Lagrange equations reduce to (\ref{eqn:el-eqns}),
+%\begin{equation}\label{eqn:el-eqns}
+%    D(q)\ddot{q} + C(q,\dot{q})\dot{q} + \nabla P(q) = B(q)\tau
+%\end{equation}
+%where the \textit{Coriolis matrix} \(C(q,\dot{q})\) is of the form
+%\[
+%    [C]_{i,j} = \frac{1}{2}\sum\limits_{k = 1}^n 
+%    \left(\pdiff{D_{i,j}}{q_k}  +
+%     \pdiff{D_{i,k}}{q_j} -
+%     \pdiff{D_{k,j}}{q_i}\right)\dot{q}_k
+%\]
 
@@
 To ease notation, we will write \(p = (p_1, \ldots, p_n)\) and say that \(p\) is
-the \textit{conjugate of momenta to \(q\)} Note that each \(p_i\) is a linear
-function of \(\dot{q}\), so one can typically solve for \(\dot{q}(q,p)\) by
-inverting the expressions from (\ref{eqn:p-i}). The combined vector \((q,p)\) is
-called the \textit{phase} of the system.
+the \textit{conjugate of momenta to \(q\)}. 
+Note that each \(p_i\) is a linear function of \(\dot{q}\), so one can typically
+solve for \(\dot{q}(q,p)\) by inverting the expressions from (\ref{eqn:p-i}).
+The combined vector \((q,p)\) is called the \textit{phase} of the system.
 
 Next we define the \textit{Hamiltonian} (\ref{eqn:hamiltonian-general}) 
-as the Legendre transform of the Lagrangian \cite{landau_mechanics}:
+as the Legendre transform of the Lagrangian \cite{landau_mechanics}.
 \begin{equation}\label{eqn:hamiltonian-general}
@@
 \end{equation}
-where \(B(q) \in \R^{n\times k}\) is the input matrix and 
-\(\tau \in \R^k\) is the vector of generalized input forces. A pair of
-coordinates \((q,p)\) which satisfy Hamilton's equations are said to be 
-\textit{canonical coordinates for} \(\mathcal{H}\). a change of coordinates
-\((\hat{q}(q,p),\hat{p}(q,p))\) is said to be a \textit{canonical
-transformation} if the EOM \((\dot{\hat{q}},\dot{\hat{p}})\) satisfy Hamilton's
-equations for some other Hamiltonian \(\hat{H}\).
+where \(B(q) \in \R^{n\times k}\) is the same input matrix used by the
+Lagrangian and \(\tau \in \R^k\) is the vector of generalized input forces.
 
-If the Kinetic energy is quadratic as in (\ref{eqn:lagrangian}), the conjugate of
-momenta to \(q\) can be computed explicitly by
+If the kinetic energy is quadratic as in (\ref{eqn:lagrangian}), the conjugate of
+momenta to \(q\) can be computed explicitly:
 \[
@@
 
+A pair of coordinates \((q,p)\) which satisfy Hamilton's equations 
+under the Hamiltonian \(\mathcal{H}\) are
+said to be \textit{canonical coordinates for} \(\mathcal{H}\).  A change of
+coordinates \((q,p) \rightarrow (Q,P)\) is said to be a \textit{canonical
+transformation} if \((Q,P)\) are canonical coordinates for
+\(H(Q,P) = \mathcal{H}\left(q(Q,P), p(Q,P)\right)\).
+
 \section{Simply Actuated Hamiltonian Systems}
@@
 % inertia, D for normal coordinates
-Given an arbitrary Hamiltonian mechanical system (\ref{eqn:hamiltonian}) in
-canonical coordinates \((q,p)\), it is not obvious how the input forces \(\tau\)
-are affecting the conjugate of momenta \(p_i\). This is because \(\tau\) is
-transformed by \(B(q)\), which may be quite complicated.
+Given a Hamiltonian mechanical system (\ref{eqn:hamiltonian}), it is not obvious
+how the input forces \(\tau\) affect the conjugate of momenta \(p_i\). 
+This is because \(\tau\) is transformed by the input matrix \(B(q)\), which may
+be quite complicated.
 
-For this reason, we will define a new class of Hamiltonian systems with simple
-input matrices, so that the effect of the input forces on the conjugate of
-momenta is obvious. This class of systems will be the backbone for the rest of
-the theory developed in this thesis.
+We will define a new class of Hamiltonian systems where the effect of the input
+forces on the conjugate of momenta is made obvious. This class of systems will
+form the backbone for the rest of the theory developed in this thesis.
 
 \begin{defn}
-    A Hamiltonian mechanical system (\ref{eqn:hamiltonian}) is said to be 
-    \textit{simply actuated} if the input matrix \(B(q) \in \R^{n \times k}\)
-    is always of the form
+    A pair of canonical coordinates \((q,p)\) for \(\mathcal{H}\) are said to be
+    \textit{simply actuated coordinates} if, in these coordinates, the input
+    matrix \(B(q) \in \R^{n \times k}\) is of the form
     \[
@@
 \end{defn}
+\begin{defn}
+    A Hamiltonian system is said to be \textit{simply actuated} if there exists
+    a canonical transformation into simply actuated coordinates for
+    \(\mathcal{H}\).
+\end{defn}
 
-Under the following mild assumptions on the input matrix, we will show that
-there exists a canonical transformation under which the system is simply
-actuated.
+Under the following assumptions on the input matrix, we will show that the
+Hamiltonian system (\ref{eqn:hamiltonian}) is simply actuated.
 
 \begin{assm}\label{assm:B-const}
-    The input matrix \(B : \mathcal{Q} \rightarrow \R^{n\times k}\) is constant
-    and full rank.
+    The input matrix \(B(q) \equiv B \in \R^{n\times k}\) is constant,
+    full rank, and \(k \leq n\).
 \end{assm}
 \begin{assm}\label{assm:B-perp}
-    There exists a right-semi-orthogonal matrix 
+    There exists a matrix 
     \(B^\perp \in \R^{(n-k)\times n}\)
-    which is a left-annihilator for \(B\). That is,
-    \(B^\perp(B^\perp)\tpose = \Id{(n-k)}\) (right semi-orthogonality) and
-    \(B^\perp B = \Zmat{(n-k) \times k}\) (left-annihilation).
+    which is right semi-orthogonal 
+    (\ie \(B^\perp(B^\perp)\tpose = \Id{(n-k)}\))
+    and which is a left-annihilator for \(B\)
+    (\ie \(B^\perp B = \Zmat{(n-k) \times k}\)).
 \end{assm}
@@
 \begin{assm}\label{assm:B-orthogonal}
-    WLOG, the input matrix \(B\) is left semi-orthogonal. 
-    That is, \(B\tpose B = \Id{k}\).
+    Assume without loss of generality that the input matrix \(B\) is left
+    semi-orthogonal.
+    That is, assume \(B\tpose B = \Id{k}\).
 \end{assm}
 \begin{proof}
-Since \(B\) is a constant full-rank matrix, 
+Since \(B\) is a constant matrix, 
 it has a singular-value decomposition 
 \(B = U \Sigma V\tpose\) where \(U^{-1} = U\tpose \in \R^{n \times n}\), 
-\(V^{-1} = V\tpose \in \R^{n \times k}\), and \(\Sigma \in \R^{n \times k}\) is
+\(V^{-1} = V\tpose \in \R^{k \times k}\), and \(\Sigma \in \R^{n \times k}\) is
 defined by
@@
 \]
-where \(\sigma_i \neq 0\).
+where \(\sigma_i \neq 0\) because \(B\) is full-rank.
 \textbf{SOURCE}
@@
 and assigning the input forces to \(\tau = V T \hat{\tau}\), we get a new input
-matrix for \(\hat{\tau}\) given by \(\hat{B} = B V T = U \Sigma T\) 
+matrix for \(\hat{\tau} \in \R^k\) given by \(\hat{B} = B V T = U \Sigma T\) 
 which is still constant and full-rank. In particular, 
@@
 \]
-Since \(B^\perp\) is a left annihilator or \(B\) and both \(B^\perp\) and
+Since \(B^\perp\) is a left annihilator of \(B\) and both \(B^\perp\) and
 \(B\tpose\) are right semi-orthogonal, it is easy to show that \(\mathbf{B}\) is
-orthogonal (i.e. it is full-rank and has inverse equal to its transpose):
+orthogonal:
 \[
@@
 
-In these new coordinates, the generalized input forces directly affect the
-actuated conjugate of momenta. That is, \(\tau_i\) only affects \(p_{a,i}\), 
-and all \(p_u\) are unaffected by the input forces. We prove this fact through
-the following theorem.
+The following theorem shows that these new coordinates are simply actuated
+coordinates, so that \(\tau_i\) only affects \(p_{a,i}\)
+and all \(p_u\) are unaffected by the input forces.
 
@@
     \begin{align}\label{eqn:simple-hamiltonian}
-        \hat{\mathcal{H}}(Q,P) &= 
+        \hat{H}(Q,P) &= 
         \frac{1}{2} P\tpose \Minv(Q) P + V(Q) \\
@@
     \((Q = Aq, P = Ap)\) satisfies
-    \(\pdiff{Q_i}{p_s} = \pdiff{P_i}{q_s} = 0\) for all 
-    \(i,s \in \{1,\ldots,n\}\). Hence, the Poisson brackets between the
+    \(\pdiff{Q_i}{p_m} = \pdiff{P_i}{q_m} = 0\) for all 
+    \(i,m \in \{1,\ldots,n\}\). Hence, the Poisson brackets between the
     new coordinates are zero:
     \begin{align*}
-        [Q_i,Q_j] &:= \sum\limits_{s = 1}^n \pdiff{Q_i}{p_s}\pdiff{Q_j}{q_s} - 
-        \pdiff{Q_i}{q_s}\pdiff{Q_j}{p_s} = 0 \\
-        [P_i,P_j] &:= \sum\limits_{s=1}^n \pdiff{P_i}{p_s}\pdiff{P_j}{q_s} -
-        \pdiff{P_i}{q_s}\pdiff{P_j}{p_s} = 0
+        [Q_i,Q_j] &:= \sum\limits_{m = 1}^n \pdiff{Q_i}{p_m}\pdiff{Q_j}{q_m} - 
+        \pdiff{Q_i}{q_m}\pdiff{Q_j}{p_m} = 0 \\
+        [P_i,P_j] &:= \sum\limits_{m=1}^n \pdiff{P_i}{p_m}\pdiff{P_j}{q_m} -
+        \pdiff{P_i}{q_m}\pdiff{P_j}{p_m} = 0
     \end{align*}
     Since the matrix \(A = \mathbf{B}\) is invertible and orthogonal,
-    \(A_i (A\tpose_j)\tpose = A_i (A^{-1}_j)\tpose = \delta_{i,j}\). Using this
-    fact we see that the Poisson brackets between the new momenta and
-    configurations are given by:
+    \((A_i)\tpose A\tpose_j = (A_i)\tpose (A^{-1})_j = \delta_{i,j}\). Using this
+    fact we see that the Poisson brackets between \(P\) and \(Q\) are given by:
     \begin{align*}
-        [P_i,Q_j] &= \sum\limits_{s=1}^n\pdiff{P_i}{p_s}\pdiff{Q_j}{q_s}
-        - \pdiff{P_i}{q_s}\pdiff{Q_j}{p_s} \\
-                  &= \sum\limits{s=1}^n A_{i,s}A_{j,s} - 0 \\
-                  &= \sum\limits{s=1}^n A_{i,s}A\tpose_{s,j} \\
-                  &= A_i (A\tpose_j)\tpose \\
+        [P_i,Q_j] &= \sum\limits_{m=1}^n\pdiff{P_i}{p_m}\pdiff{Q_j}{q_m}
+        - \pdiff{P_i}{q_m}\pdiff{Q_j}{p_m} \\
+                  &= \sum\limits_{m=1}^n A_{i,m}A_{j,m} - 0 \\
+                  &= \sum\limits_{m=1}^n A_{i,m}A\tpose_{m,j} \\
+                  &= (A_i)\tpose A\tpose_j \\
                   &= \delta_{i,j}
@@
     \[
-        \hat{B} = \mathbf{B}B = \begin{bmatrix}
+        \mathbf{B}B = \begin{bmatrix}
             B^\perp B \\
@@
     \]
-    so the new Hamiltonian system is simply actuated, as desired.
+    so the coordinates \((q_u,q_a,p_u,p_a)\) are simply actuated coordinates for 
+    \(\hat{H}\) as desired.
 \end{proof}

### 2020-06-30 latex/symbols.tex
Updated symbols so that M_i is a column, as is more standard.
Added a bunch of new symbols for derivatives, jacobians, etc.

@@
    \midrule
+   \(\n\) & The index set \(\{1,\ldots,n\}\) of natural numbers up to \(n\). \\
    \(\R^n\) & Real numbers in \(n\) dimensions. \\
@@
    \(M_i\) & If \(M\) is a vector, the \(i\)th element of \(M\). \\
-         & If \(M\) is a matrix, the \(i\)th row of \(M\). \\
+         & If \(M\) is a matrix, the \(i\)th column of \(M\). \\
    \(M_{i,j}\) & The value of row \(i\), column \(j\) for the matrix \(M\). \\
    \([v_i]\) & If \(v_i \in \R\), the column vector with value \(v_i\) at position \(i\). \\
+      & If \(v_i \in \R^n\), the matrix with column \(i\) given by \(v_i\). \\
       & If \(v_i \in \R^{1\times n}\), the matrix with row \(i\) given by \(v_i\). \\
    \(\dot{x}\) & Derivative of \(x\) with respect to time \(t\). \\
-   \(\nabla_v F\) & Gradient of \(F)\) with respect to the vector \(v\). \\
+   \(\nabla_v F\) & Gradient of the \(\R\)-valued function \(F\) with respect to \(v\). \\
+   \(dF_v\) & Total differential (Jacobian) of the function \(F\) with respect to \(v\).\\
+           & If \(F\) is \(\R\)-valued, equivalent to \((\nabla_v F)\tpose\). \\
+   \(\nabla_v dF_w\) & Derivative matrix of the \(\R\)-valued function \(F\), with \((i,j)\) element \(\partial^2 F/(\partial v_i \partial w_j)\). \\
+   \(\nabla_v^2 F\) & Hessian of \(F : \R^n \rightarrow R\), equivalent to \(\nabla_v dF_v\). \\
    \(\delta_{i,j}\) & The Kronecker delta: 1 if \(i = j\) and 0 otherwise. \\

### 2020-07-06 latex/symbols.tex
Updated symbols as per Manfredi's request to change notation.

@@
    \(M_{i,j}\) & The value of row \(i\), column \(j\) for the matrix \(M\). \\
-   \([v_i]\) & If \(v_i \in \R\), the column vector with value \(v_i\) at position \(i\). \\
-      & If \(v_i \in \R^n\), the matrix with column \(i\) given by \(v_i\). \\
-      & If \(v_i \in \R^{1\times n}\), the matrix with row \(i\) given by \(v_i\). \\
    \(\dot{x}\) & Derivative of \(x\) with respect to time \(t\). \\
@@
            & If \(F\) is \(\R\)-valued, equivalent to \((\nabla_v F)\tpose\). \\
-   \(\nabla_v dF_w\) & Derivative matrix of the \(\R\)-valued function \(F\), with \((i,j)\) element \(\partial^2 F/(\partial v_i \partial w_j)\). \\
-   \(\nabla_v^2 F\) & Hessian of \(F : \R^n \rightarrow R\), equivalent to \(\nabla_v dF_v\). \\
+   \(\Hess F\) & Hessian matrix of double derivatives of \(F : \R^n \rightarrow \R\). \\
+   \(\partial_v \partial_w F\) & Derivative matrix of \(F : \R^n \times \R^m \rightarrow \R\), with \((i,j)\) element \(\ppdiff{F}{v_i}{w_j}\). \\
    \(\delta_{i,j}\) & The Kronecker delta: 1 if \(i = j\) and 0 otherwise. \\

### 2020-07-06 latex/vnhcs.tex
STarted the VNHC section, defined a VNHC but haven't done much else.
Edited preliminaries again, changed symbols.

@@
 \chapter{Development of Virtual Nonholonomic Constraints}
+% TODO: Brief introduction before preliminaries
+
 \section{Preliminaries on Analytical Mechanics}
@@
 independent scalar-valued variables \(q_1,\ldots,q_n\) called 
-\textit{generalized coordinates}, where \(n = 3N - m\) is the number of
+\textit{generalized coordinates}, where \(n = 3N - r\) is the number of
 \textit{degrees of freedom} (DOF) of the system \cite{greenwood_dynamics}.
 
-Each generalized coordinate \(q_i\) takes values in \(\Rt{T_i}\), where
+For the mechanical systems of interest in this thesis, we assume that
+each generalized coordinate \(q_i\) represents either the distance or the angle
+between two parts of the system.
+Mathematically, each \(q_i\) takes values in \(\Rt{T_i}\), where
 \(T_i = \infty\) if \(q_i\) represents a length or \(T_i = 2\pi\) if \(q_i\)
@@
 \(q = (q_1,\ldots,q_n) \in \mathcal{Q}\) 
-where the \textit{configuration manifold} \(\mathcal{Q}\) of the system has the
-following structure:
+where the \textit{configuration manifold} \(\mathcal{Q}\) of the system is a
+so-called \textit{generalized cylinder}:
 \[
@@
 The derivative \(\dot{q} = (\dot{q}_1,\ldots,\dot{q}_n)\) of a configuration
-is called a \textit{generalized velocity}, and the combined vector
-\((q,\dot{q})\) is called a \textit{state} of the system.
+is called a \textit{generalized velocity} of the system. For arbitrary systems,
+the space of allowable velocities depends on the current configuration of the
+system.  However, since \(\mathcal{Q}\) is a generalized cylinder,
+\(\dot{q} \in \R^n\).
+The combined vector \((q,\dot{q}) \in \mathcal{Q}\times\R^n\) is called a 
+\textit{state} of the system.
 
@@
 \[
-    B(q) = [B_i\tpose(q)]
+    B(q) = \begin{bmatrix}
+        - & B_1\tpose(q) & - \\
+          & \vdots & \\
+        - & B_n\tpose(q) & - \\
+    \end{bmatrix}
 \]
@@
 \end{equation}
-To ease notation, we will write \(p = (p_1, \ldots, p_n)\) and say that \(p\) is
-the \textit{conjugate of momenta to \(q\)}. 
+To ease notation, we will write \(p = (p_1, \ldots, p_n) \in \R^n\) and say that
+\(p\) is the \textit{conjugate of momenta to \(q\)}.
 Note that each \(p_i\) is a linear function of \(\dot{q}\), so one can typically
 solve for \(\dot{q}(q,p)\) by inverting the expressions from (\ref{eqn:p-i}).
-The combined vector \((q,p)\) is called the \textit{phase} of the system.
+The combined vector \((q,p) \in \mathcal{Q}\times\R^n\) is called a 
+\textit{phase} of the system.
 
-Next we define the \textit{Hamiltonian} (\ref{eqn:hamiltonian-general}) 
-as the Legendre transform of the Lagrangian \cite{landau_mechanics}.
-\begin{equation}\label{eqn:hamiltonian-general}
+The \textit{Hamiltonian} of the system in
+\(\{q,p\}\) coordinates is defined as the Legendre transform 
+(\ref{eqn:hamiltonian-legendre}) of the Lagrangian \cite{landau_mechanics}.
+\begin{equation}\label{eqn:hamiltonian-legendre}
     \mathcal{H}(q,p) = p\tpose \dot{q}(q,p) - \mathcal{L}(q,\dot{q}(q,p))
@@
 
-If the kinetic energy is quadratic as in (\ref{eqn:lagrangian}), the conjugate of
-momenta to \(q\) can be computed explicitly:
-\[
-    p = D(q)\dot{q}
-\]
+If the Lagrangian has quadratic kinetic energy as in (\ref{eqn:lagrangian}), 
+the conjugate of
+momenta to \(q\) can be computed explicitly by \(p = D(q)\dot{q}\).
+
+% TODO: Replace [p D_i p] here with our new notation
 The resulting Hamiltonian system reduces to (\ref{eqn:hamiltonian}).
@@
 
-A pair of coordinates \((q,p)\) which satisfy Hamilton's equations 
-under the Hamiltonian \(\mathcal{H}\) are
-said to be \textit{canonical coordinates for} \(\mathcal{H}\).  A change of
-coordinates \((q,p) \rightarrow (Q,P)\) is said to be a \textit{canonical
-transformation} if \((Q,P)\) are canonical coordinates for
+A set of coordinates \(\{q,p\}\) which satisfy Hamilton's equations 
+under the Hamiltonian \(\mathcal{H}(q,p)\) are
+said to be \textit{canonical coordinates} for the system. A change of
+coordinates \((q,p) \rightarrow (Q,P)\) is a \textit{canonical
+transformation} if \((Q,P)\) are canonical coordinates with 
 \(H(Q,P) = \mathcal{H}\left(q(Q,P), p(Q,P)\right)\).
@@
 \begin{defn}
-    A pair of canonical coordinates \((q,p)\) for \(\mathcal{H}\) are said to be
-    \textit{simply actuated coordinates} if, in these coordinates, the input
-    matrix \(B(q) \in \R^{n \times k}\) is of the form
+    A set of canonical coordinates \(\{q,p\}\) for a Hamiltonian system 
+    \(\mathcal{H}\) are said to be \textit{simply actuated coordinates} if the
+    input matrix \(B(q) \in \R^{n \times k}\) is of the form
     \[
@@
     A Hamiltonian system is said to be \textit{simply actuated} if there exists
-    a canonical transformation into simply actuated coordinates for
-    \(\mathcal{H}\).
+    a canonical transformation from any set of canonical coordinates 
+    \(\{q,p\}\) into simply actuated coordinates for \(\mathcal{H}\).
 \end{defn}
@@
 \begin{defn}
-    Define the coordinate transformation 
+    Let \(\{q,p\}\) be canonical coordinates for a Hamiltonian mechanical system
+    and define the coordinate transformation 
     \((q_u,q_a) = \mathbf{B}q\), where \(q_u = B^\perp q\)
@@
 
+% TODO: Replace the [P M_i P] with new notation
 \begin{thm}
@@
     \ref{assm:B-orthogonal}, the change of coordinates 
-    \((Q = (q_u, q_a), P = (p_u,p_a))\) 
+    \(\{Q = (q_u, q_a), P = (p_u,p_a)\}\) 
     is a canonical transformation of the
@@
 \begin{proof}
+    The Poisson bracket between \(f(q,p)\) and \(g(q,p)\) is defined by
+    \cite{landau_mechanics} as follows:
+    \[
+        [f,g] := \sum \limits_{i=1}^n \pdiff{f}{p_i}\pdiff{g}{q_i} - 
+                \pdiff{f}{q_i}\pdiff{g}{p_i}
+    \]
     For any constant matrix \(A\), the transformation
-    \((Q = Aq, P = Ap)\) satisfies
+    \(\{Q = Aq, P = Ap\}\) satisfies
     \(\pdiff{Q_i}{p_m} = \pdiff{P_i}{q_m} = 0\) for all 
-    \(i,m \in \{1,\ldots,n\}\). Hence, the Poisson brackets between the
-    new coordinates are zero:
+    \(i,m \in \n\).
+    Hence, for our new coordinates \(\{Q = \mathbf{B}q, P = \mathbf{B}p\}\),
     \begin{align*}
@@
     \end{align*}
-    Since the matrix \(A = \mathbf{B}\) is invertible and orthogonal,
-    \((A_i)\tpose A\tpose_j = (A_i)\tpose (A^{-1})_j = \delta_{i,j}\). Using this
+    Since the matrix \(\mathbf{B}\) is orthogonal,
+    \((\mathbf{B}_i)\tpose \mathbf{B}\tpose_j = (\mathbf{B}_i)\tpose (\mathbf{B}^{-1})_j = \delta_{i,j}\). Using this
     fact we see that the Poisson brackets between \(P\) and \(Q\) are given by:
@@
         - \pdiff{P_i}{q_m}\pdiff{Q_j}{p_m} \\
-                  &= \sum\limits_{m=1}^n A_{i,m}A_{j,m} - 0 \\
-                  &= \sum\limits_{m=1}^n A_{i,m}A\tpose_{m,j} \\
-                  &= (A_i)\tpose A\tpose_j \\
+                  &= \sum\limits_{m=1}^n \mathbf{B}_{i,m}\mathbf{B}_{j,m} - 0 \\
+                  &= \sum\limits_{m=1}^n \mathbf{B}_{i,m}\mathbf{B}\tpose_{m,j} \\
+                  &= (\mathbf{B}_i)\tpose \mathbf{B}\tpose_j \\
                   &= \delta_{i,j}
@@
 
-    By (45.10) in \cite{landau_mechanics}, the coordinate change
+    Therefore, by (45.10) in \cite{landau_mechanics}, the coordinate change 
     \((Q = \mathbf{B}q, P = \mathbf{B}p)\) is a canonical transformation with
@@
     \]
-    so the coordinates \((q_u,q_a,p_u,p_a)\) are simply actuated coordinates for 
+    so the coordinates \(\{q_u,q_a,p_u,p_a\}\) are simply actuated coordinates for 
     \(\hat{H}\) as desired.
@@
 \section{Virtual Nonholonomic Constraints}
+% TODO: Motivation for VNHCs, they are an extension of VHCs, etc.
+
 % TODO: Perform the full development of VNHCs, including its stabilizing
 % controller.
+\begin{defn}
+    A \textit{virtual nonholonomic constraint of order \(k\)} is a relation
+    \(h(q,p) = 0\) where \(h : \mathcal{Q}\times\R^n \rightarrow \R^k\) is smooth,
+    \(\rank{\left[ dh_q,\, dh_p \right]} = k\) for all \((q,p) \in h\inv(0)\),
+    and there exists a feedback controller \(\tau(q,p)\) stabilizing the set
+    \[
+        \Gamma = \left\{(q,p) | h(q,p) = 0, dh_q \dot{q} + dh_p \dot{p} = 0\right\}
+    \]
+    which is called the \textit{constraint manifold}.
+\end{defn}
 %TODO: Special case: when dM/dqu = 0, show that we have a nice form and that we

### 2020-07-08 latex/vnhcs.tex
First draft of the intro to VNHC section plus defs and thms

@@
     \[
-        B(q) = \begin{bmatrix}
-            \Zmat{(n-k)\times k} \\
-            \Id{k}
-        \end{bmatrix}
+        B(q) = \simpleB    
     \]
@@
 % TODO: Motivation for VNHCs, they are an extension of VHCs, etc.
+Let us imagine a child on a swing, who wants to go as high as possible. The
+child will push off the ground and start a swing with a very small oscillation,
+then extend and retract their feet appropriately to gain energy. 
+If a roboticist were designing a machine to do this task, they might design a
+control mechanism which makes the legs track a trajectory over time. For
+example, they might tell the robot to extend and retract its legs every two
+seconds. In ideal situtations, this technique would work perfectly because the
+leg motion is synchronized with the swing to gain energy as fast as possible.
+
+Unfortunately for engineers, reality does not provide ideal scenarios very often. 
+Most children have an adult pushing the swing to help them go higher; or perhaps
+the child is swinging on a windy day. In either case, they adjust their leg
+motion accordingly and do not keep track of time when kicking their legs. This
+standard control technique of tracking a function of time, known as
+\textit{trajectory tracking}, would not work because the disturbance affecting
+the swing will desynchronize the motion of the legs with the swing and stop the
+energy-gaining effects.
+
+This is a common issue in the control of many biologically-inspired systems. 
+One method which can provide more realistic, robust motion is the method of
+of virtual holonomic constraints (VHCs). Instead of a robot's actuators tracking
+a trajectory over time, VHCs use the actuators to enforce a relation
+\(h(q) = 0\) of the configuration \cite{vhcs_for_el_systems}. 
+This method has provided incredible results in the development of 
+walking robots \cite{vhc_robotic_walking, vhc_stable_walking}, 
+vehicle motion \cite{vhc_bicycle, vhc_helicopter}, 
+and has even been used to design a snake-like robot that can swim
+\cite{vhc_snake}.
+
+Recent papers have attempted to extend these results to use the robot's
+actuators to enforce a relation 
+\(h(q,\dot{q}) = 0\) of the full state. Since these relations use both
+the configuration and generalized velocity, they are called
+\textit{nonholonomic} constraints. Furthermore, they are called \textit{virtual}
+because the constraint is enforced by the robot's actuators.
+We will give our own characterization of these virtual nonholonomic constraints
+in the Hamiltonian framework, using the phase \((q,p)\) of a Hamiltonian
+system instead of the state \((q,\dot{q})\) of a lagrangian system.
 
@@
 \begin{defn}
-    A \textit{virtual nonholonomic constraint of order \(k\)} is a relation
-    \(h(q,p) = 0\) where \(h : \mathcal{Q}\times\R^n \rightarrow \R^k\) is smooth,
-    \(\rank{\left[ dh_q,\, dh_p \right]} = k\) for all \((q,p) \in h\inv(0)\),
-    and there exists a feedback controller \(\tau(q,p)\) stabilizing the set
+    A \textit{virtual nonholonomic constraint} (VNHC) \textit{of order \(k\)} is a
+    relation \(h(q,p) = 0\) where \(h : \mathcal{Q}\times\R^n \rightarrow \R^k\) is
+    smooth, \(\rank{\left[ dh_q,\, dh_p \right]} = k\) for all 
+    \((q,p) \in h\inv(0)\), and there exists a feedback controller \(\tau(q,p)\)
+    stabilizing the set
     \[
-        \Gamma = \left\{(q,p) | h(q,p) = 0, dh_q \dot{q} + dh_p \dot{p} = 0\right\}
+        \Gamma = \left\{(q,p) \vert h(q,p) = 0, dh_q \dot{q} + dh_p \dot{p} = 0\right\}
     \]
@@
 \end{defn}
+
+Defining the error term  \(e = h(q,p)\), stabilizing \(\Gamma\) is equivalent to 
+solving for \(\tau(q,p)\) which drives \(e \rightarrow 0\)
+and \(\dot{e} \rightarrow 0\).
+Let us imagine that we have no further structure on the VNHC. 
+Then \(\dot{e} = dh_q \dot{q} + dh_p \dot{p}\), so \(\tau\) appears inside
+\(\dot{e}\). To solve for \(\tau\) explicitly, we must have that \(dh_p\) is
+invertible, which means that \(h(q,p)\) is strictly monotonic in \(p\).
+
+This is a very restrictive condition, which many VNHCs will not satisfy. For
+this reason, it is preferable to have the torque \(\tau\) appear after two 
+derivatives of \(e\) so that there is more freedom in the types of constraints
+one can use.
+The terminology for this is to say that \(e\) is of \textit{relative degree}
+\(\{2,2,\ldots,2\}\). We thus define a special type of VNHC which satisfies this
+property.
+
+\begin{defn}
+    A VNHC of order \(k\) \(h(q,p) = 0\) is \textit{regular} if the output 
+    \(e = h(q,p)\) is of relative degree \(\{2,2.\ldots,2\}\) everywhere on the
+    constraint manifold \(\Gamma\).
+\end{defn}
+
+\citeauthor{hybrid_zero_dynamics_bipedal_nhvcs} showed that relations which use only
+the unactuated momentum \(p_u\) result in \(\tau\) appearing only after two
+derivatives of \(e\) \cite{hybrid_zero_dynamics_bipedal_nhvcs}. 
+In the remainder of this chapter, we will formalize and generalize this idea 
+under the following assumption.
+
+% TODO: Write the [p' M p] matrix in our nice nottation
+\begin{assm}\label{assm:H-is-simply-actuated}
+    The mechanical system under consideration is an underactuated 
+    Hamiltonian system of order \((n-k) \geq 0\) in simply
+    actuated coordinates \(\{q = (q_u,q_a), p = (p_u, p_a)\}\):
+    \begin{align*}
+        \mathcal{H}(q,p) &= p\tpose \Minv(q) p + V(q) \\
+         &\begin{cases}
+            \dot{q} = \Minv p \\
+            \dot{p} = -\frac{1}{2} ? - \nabla_q V(q) + \simpleB \tau \\
+        \end{cases}
+    \end{align*}
+\end{assm}
+
+% TODO: write out ??? = [p' dM/(dqu_i)] the matrix of size nxn
+\begin{prop}
+    A VNHC of order \(k\) is regular if and only if \(dh_{p_a} = 0\) 
+    (\ie \(h(q,p) \equiv h(q,p_u)\)) and
+    \[
+        \rank{\left(dh_q \Minv(q) + dh_{p_u} ??? \right)\simpleB} = k
+    \]
+    everywhere on the constraint manifold \(\Gamma\).
+\end{prop}
+\begin{proof}
+\end{proof}
+
 %TODO: Special case: when dM/dqu = 0, show that we have a nice form and that we

### 2020-07-30 latex/symbols.tex
Added Q to the symbols
Major edits to VNHC
Added several new VNHC-related references, updated the whole chapter of
VNHCs, and implemented my new notation using the matrix kronecker
product. Also added further justfication as to why we have our own
definition of VNHCs, which needs more work. Next, we need to update the
proof on why a VNHC is regular iff it is h(q,pu) and satisfies that
matrix rank definition.

@@
    \(\mathbb{S}^1\) & The unit circle, equivalent to \(\Rt{2\pi}\). \\
+   \(\mathcal{Q}\) & The configuration manifold of a system. \\
    \(\R^{n \times m}\) & The space of real-valued matrices with \(n\) rows and \(m\) columns. \\
@@
    \(\dot{x}\) & Derivative of \(x\) with respect to time \(t\). \\
-   \(\nabla_v F\) & Gradient of the \(\R\)-valued function \(F\) with respect to \(v\). \\
-   \(dF_v\) & Total differential (Jacobian) of the function \(F\) with respect to \(v\).\\
-           & If \(F\) is \(\R\)-valued, equivalent to \((\nabla_v F)\tpose\). \\
-   \(\Hess F\) & Hessian matrix of double derivatives of \(F : \R^n \rightarrow \R\). \\
+   \(\nabla_v F\) & If \(F\) is \(\R\)-valued, the gradient of \(F\) with respect to \(v\). \\
+            & If \(F : \R^m \rightarrow \R^{n\times n}\), the block matrix gradient \((\pdiff{F}{v_1},\ldots,\pdiff{F}{v_m}) \in \R^{nm \times n}\). \\
+   \(dF_v\) & Total differential (Jacobian) of \(F\), equivalent to \((\nabla_v F)\tpose\). \\
+   \(\Hess F\) & If \(F : \R^n \rightarrow \R\), the \(n \times n\) Hessian matrix of double derivatives of \(F\). \\
+               & If \(F : \R^n \rightarrow \R^k\), the block matrix \((\Hess F_1, \ldots, \Hess F_k) \in \R^{n \times nk}\). \\
    \(\partial_v \partial_w F\) & Derivative matrix of \(F : \R^n \times \R^m \rightarrow \R\), with \((i,j)\) element \(\ppdiff{F}{v_i}{w_j}\). \\
    \(\delta_{i,j}\) & The Kronecker delta: 1 if \(i = j\) and 0 otherwise. \\
+   \(\otimes\) & The matrix kronecker product (see Appendix \ref{appendix:matrix-kron-prod}). \\
    \bottomrule

### 2020-07-30 latex/vnhcs.tex
Major edits to VNHC
Added several new VNHC-related references, updated the whole chapter of
VNHCs, and implemented my new notation using the matrix kronecker
product. Also added further justfication as to why we have our own
definition of VNHCs, which needs more work. Next, we need to update the
proof on why a VNHC is regular iff it is h(q,pu) and satisfies that
matrix rank definition.

@@
 \section{Preliminaries on Analytical Mechanics}
-A mechanical system can be represented by \(N\) point masses, where each point
+A mechanical system can be represented by \(N\) point masses where each point
 represents the center of mass of a physical body, along with \(r\)
@@
 
-For the mechanical systems of interest in this thesis, we assume that
+For the robotic systems of interest in this thesis, we assume that
 each generalized coordinate \(q_i\) represents either the distance or the angle
@@
 the space of allowable velocities depends on the current configuration of the
-system.  However, since \(\mathcal{Q}\) is a generalized cylinder,
+system.  However, since \(\mathcal{Q}\) is a generalized cylinder, we find that 
 \(\dot{q} \in \R^n\).
@@
                      &\begin{cases}
-        \dot{q} = D^{-1}(q)p \\
-        \dot{p} = -\frac{1}{2} \left[p\tpose \pdiff{D^{-1}}{q_i}(q) p\right] 
+        \dot{q} = D\inv(q)p \\
+        \dot{p} = -\frac{1}{2} (I_n \otimes p\tpose) \nabla_q D\inv(q) p
         - \nabla_q P(q) + B(q) \tau \\
@@
 A set of coordinates \(\{q,p\}\) which satisfy Hamilton's equations 
-under the Hamiltonian \(\mathcal{H}(q,p)\) are
+under the Hamiltonian \(\mathcal{H}\) are
 said to be \textit{canonical coordinates} for the system. A change of
 coordinates \((q,p) \rightarrow (Q,P)\) is a \textit{canonical
-transformation} if \((Q,P)\) are canonical coordinates with 
+transformation} if \(\{Q,P\}\) are canonical coordinates with Hamiltonian
 \(H(Q,P) = \mathcal{H}\left(q(Q,P), p(Q,P)\right)\).
@@
 \begin{defn}
-    A set of canonical coordinates \(\{q,p\}\) for a Hamiltonian system 
-    \(\mathcal{H}\) are said to be \textit{simply actuated coordinates} if the
+    Let \(\mathcal{H}\) be an underactuated Hamiltonian system 
+    \(\mathcal{H}\) with degree of underactuation \((n-k)\geq 0\). 
+    A pair of canonical coordinates \(\{q,p\}\) for this system
+    are said to be \textit{simply actuated coordinates} if the
     input matrix \(B(q) \in \R^{n \times k}\) is of the form
@@
     \]
+    The first \((n-k)\) coordinates \(q_u\) are called the \textit{unactuated
+    coordinates}, while the remaining \(k\) coordinates \(q_a\) are called the
+    \textit{actuated coordinates}. We write \(q = (q_u, q_a)\) and the
+    corresponding conjugate of momenta \(p = (p_u, p_a)\) are the called the
+    unactuated and actuated momenta (respectively).
 \end{defn}
@@
 
-Note that if \(k = (n-1)\), the existence of a left annihilator \(A^0\) for
-\(B\) implies \(B^\perp := A^0/\norm{A^0}\) satisfies Assumption
+Note that if \(k = (n-1)\), the existence of any left annihilator \(A^0\) for
+\(B\) implies the left annihilator \(B^\perp := A^0/\norm{A^0}\) satisfies Assumption
 \ref{assm:B-perp}.
@@
     \Sigma = \begin{bmatrix}
-        \sigma_1 & 0 & \cdots & 0 & 0 \\
-        0 & \sigma_2 & \cdots & 0 & 0 \\
-        \vdots & & \ddots &  & \vdots \\
-        0 & 0 & \cdots & 0 & \sigma_k \\
-          & - & \Zmat{(n-k)\times k} & -  &\\
+        \sigma_1 & 0 & \cdots & 0 \\
+        0 & \sigma_2 & \cdots & 0 \\
+        \vdots & & \ddots & \vdots \\
+        0 & 0 & \cdots & \sigma_k \\
+        - &   & \Zmat{(n-k)\times k} & -  \\
     \end{bmatrix}
 \]
-where \(\sigma_i \neq 0\) because \(B\) is full-rank.
-\textbf{SOURCE}
+where \(\sigma_i \neq 0\) because \(B\) is full-rank \textbf{SOURCE?}.
 Defining \(T \in \R^{k \times k}\) by
@@
 
-\begin{defn}
-    Let \(\{q,p\}\) be canonical coordinates for a Hamiltonian mechanical system
-    and define the coordinate transformation 
-    \((q_u,q_a) = \mathbf{B}q\), where \(q_u = B^\perp q\)
-    and \(q_a = B\tpose q\).
-    We call \(q_u\) the \textit{unactuated coordinates} and \(q_a\) the
-    \textit{actuated coordinates} of the system. 
-    Their corresponding conjugate of momenta are 
-    \((p_u,p_a) = \mathbf{B}p\).
-\end{defn}
-
-The following theorem shows that these new coordinates are simply actuated
-coordinates, so that \(\tau_i\) only affects \(p_{a,i}\)
-and all \(p_u\) are unaffected by the input forces.
+The following theorem shows that \(\mathbf{B}\) provides a canonical
+transformation into simply actuated coordinates, so that all unactuated momenta
+are unaffected by the input forces.
 
@@
     Under Assumptions \ref{assm:B-const},\ref{assm:B-perp}, and
-    \ref{assm:B-orthogonal}, the change of coordinates 
-    \(\{Q = (q_u, q_a), P = (p_u,p_a)\}\) 
-    is a canonical transformation of the
-    original Hamiltonian system (\ref{eqn:hamiltonian}) into
-    the simply actuated Hamiltonian system (\ref{eqn:simple-hamiltonian}),
+    \ref{assm:B-orthogonal}, the Hamiltonian system (\ref{eqn:hamiltonian}) is
+    simply actuated with simply actuated coordinates 
+    \(\{Q = \mathbf{B}q, P = \mathbf{B}p\}\). The resulting dynamics are 
+    given by (\ref{eqn:simple-hamiltonian}),
     \begin{align}\label{eqn:simple-hamiltonian}
@@
             \dot{Q} = \Minv(Q)P \\
-            \dot{P} = -\frac{1}{2}
-                \left[P\tpose \pdiff{\Minv}{Q_i}(Q) P \right] 
-                - \nabla_Q V(Q) + 
-                \begin{bmatrix}
-                    \Zmat{(n-k)\times k} \\
-                    \Id{k} \\
-                \end{bmatrix} \tau
+            \dot{P} = -\frac{1}{2} (I_n \otimes P\tpose) \nabla_Q \Minv(Q) P
+                - \nabla_Q V(Q) + \simpleB \tau
         \end{cases} \nonumber
@@
     \begin{align*}
-        \Minv(Q) &= \mathbf{B}D^{-1}(\mathbf{B}\tpose Q)\mathbf{B}\tpose \\
-        V(Q) &= P(\mathbf{B}\tpose Q) \\
+        \Minv(Q) &:= \mathbf{B}D^{-1}(\mathbf{B}\tpose Q)\mathbf{B}\tpose \\
+        V(Q) &:= P(\mathbf{B}\tpose Q) \\
     \end{align*}
@@
 \begin{proof}
-    The Poisson bracket between \(f(q,p)\) and \(g(q,p)\) is defined by
+    The Poisson bracket between the functions \(f(q,p)\) and \(g(q,p)\) is defined by
     \cite{landau_mechanics} as follows:
@@
     \]
-    so the coordinates \(\{q_u,q_a,p_u,p_a\}\) are simply actuated coordinates for 
+    so the coordinates \(\{Q = (q_u,q_a), P = (p_u,p_a)\}\) are simply actuated coordinates for 
     \(\hat{H}\) as desired.
@@
 Let us imagine a child on a swing, who wants to go as high as possible. The
-child will push off the ground and start a swing with a very small oscillation,
+child will push off the ground and start swinging with small oscillations,
 then extend and retract their feet appropriately to gain energy. 
@@
 
-Unfortunately for engineers, reality does not provide ideal scenarios very often. 
 Most children have an adult pushing the swing to help them go higher; or perhaps
 the child is swinging on a windy day. In either case, they adjust their leg
-motion accordingly and do not keep track of time when kicking their legs. This
+motion accordingly and do not keep track of time when kicking their legs. Te
 standard control technique of tracking a function of time, known as
@@
 vehicle motion \cite{vhc_bicycle, vhc_helicopter}, 
-and has even been used to design a snake-like robot that can swim
+and has even been used to design a snake-like swimming robot
 \cite{vhc_snake}.
 
-Recent papers have attempted to extend these results to use the robot's
-actuators to enforce a relation 
-\(h(q,\dot{q}) = 0\) of the full state. Since these relations use both
-the configuration and generalized velocity, they are called
-\textit{nonholonomic} constraints. Furthermore, they are called \textit{virtual}
-because the constraint is enforced by the robot's actuators.
-We will give our own characterization of these virtual nonholonomic constraints
-in the Hamiltonian framework, using the phase \((q,p)\) of a Hamiltonian
-system instead of the state \((q,\dot{q})\) of a lagrangian system.
+Many authors have attempted to extend these results to enforce a relation 
+\(h(q,\dot{q}) = 0\) of the full state. Since these relations use actuators to
+restrict both the configuration and generalized velocity, they are called
+virtual \textit{nonholonomic} constraints. This idea has been used for
+human-robot interaction
+\cite{vnhc_human_robot_cooperation,psd_based_vnhc_redundant_manipulator,haptic_vnhc},
+error-reduction on time-delayed systems \cite{vnhc_time_delay_teleop},
+and improving bipedal locomotion \cite{nhvc_dynamic_walking,
+hybrid_zero_dynamics_bipedal_nhvcs,nhvc_incline_walking,output_nhvc_bipedal_control}.
+
+Unlike the theory of VHCs, there does not appear to be a standard definition of virtual
+nonholonomic constraints. One issue is that Lagrangian dynamics is not
+particularly suited to finding controllers which stabilize desired constraints. Because
+of this, all the applications listed above use their own definition of a virtual
+nonholonomic constraint, which makes it difficult to compare and analyze their
+work.
+
+This section will provide a new characterization of virtual nonholonomic
+constraints. The goal is to provide a consistent, rigorous foundation for
+designing constraints on a general class of systems.
 
@@
 
-Defining the error term  \(e = h(q,p)\), stabilizing \(\Gamma\) is equivalent to 
-solving for \(\tau(q,p)\) which drives \(e \rightarrow 0\)
-and \(\dot{e} \rightarrow 0\).
+If we define the error term  \(e = h(q,p)\), stabilizing \(\Gamma\) is
+equivalent to solving for \(\tau(q,p)\) which drives \(e \rightarrow 0\) and
+\(\dot{e} \rightarrow 0\).
 Let us imagine that we have no further structure on the VNHC. 
@@
 \(\dot{e}\). To solve for \(\tau\) explicitly, we must have that \(dh_p\) is
-invertible, which means that \(h(q,p)\) is strictly monotonic in \(p\).
+invertible for all \(p\), which is the same as requiring that \(h(q,p)\) is
+strictly monotonic in \(p\).
 
@@
 one can use.
-The terminology for this is to say that \(e\) is of \textit{relative degree}
-\(\{2,2,\ldots,2\}\). We thus define a special type of VNHC which satisfies this
+Mathematically, if \(\tau\) appears only after two derivatives, one says 
+that \(e\) is of \textit{relative degree} \(\{2,2,\ldots,2\}\). 
+We thus define a special type of VNHC which satisfies this
 property.
@@
 
-\citeauthor{hybrid_zero_dynamics_bipedal_nhvcs} showed that relations which use only
-the unactuated momentum \(p_u\) result in \(\tau\) appearing only after two
-derivatives of \(e\) \cite{hybrid_zero_dynamics_bipedal_nhvcs}. 
-In the remainder of this chapter, we will formalize and generalize this idea 
-under the following assumption.
+Modifying the results of
+\cite{nhvc_dynamic_walking,hybrid_zero_dynamics_bipedal_nhvcs,nhvc_incline_walking}
+into the Hamiltonian framework, one observes that
+relations which use only the unactuated momentum \(p_u\) result in \(\tau\)
+appearing only after two derivatives of \(e\).
+To be able to use \(p_u\), we will continue with the following assumption for
+the rest of the chapter.
 
@@
             \dot{q} = \Minv p \\
-            \dot{p} = -\frac{1}{2} ? - \nabla_q V(q) + \simpleB \tau \\
+            \dot{p} = -\frac{1}{2} (I_n \otimes p\tpose) \nabla_q \Minv(q) p
+            - \nabla_q V(q) + \simpleB \tau \\
         \end{cases}
@@
     A VNHC of order \(k\) is regular if and only if \(dh_{p_a} = 0\) 
-    (\ie \(h(q,p) \equiv h(q,p_u)\)) and
+    and
     \[
-        \rank{\left(dh_q \Minv(q) + dh_{p_u} ??? \right)\simpleB} = k
+        \rank{\left(dh_q \Minv(q) + 
+          dh_{p_u} (I_n \otimes p\tpose)\nabla_{q_u}\Minv(q) 
+         \right)\simpleB} = k
     \]

### 2020-08-04 latex/vnhcs.tex
Input the first version of all VNHC theorems
This commit writes out the theorem stating the closed loop dynamics on
Gamma, as well as the corollary showing what happens when the system is
of degree of underactuation one.

Still to do is the proofs of these theorems, as well as the
characterization of the torque phase-feedback controller.

@@
     \[
-        \Gamma = \left\{(q,p) \vert h(q,p) = 0, dh_q \dot{q} + dh_p \dot{p} = 0\right\}
+        \Gamma = \left\{(q,p) \mid h(q,p) = 0, dh_q \dot{q} + dh_p \dot{p} = 0\right\}
     \]
@@
 \begin{assm}\label{assm:H-is-simply-actuated}
-    The mechanical system under consideration is an underactuated 
-    Hamiltonian system of order \((n-k) \geq 0\) in simply
-    actuated coordinates \(\{q = (q_u,q_a), p = (p_u, p_a)\}\):
+    The mechanical system under consideration is a
+    Hamiltonian system with \(n\) degrees of freedom and 
+    \(k < n\) actuators. It is described in simply
+    actuated coordinates \(\{q = (q_u,q_a), p = (p_u, p_a)\}\) and has the
+    following dynamics:
     \begin{align*}
@@
 
-% TODO: write out ??? = [p' dM/(dqu_i)] the matrix of size nxn
-\begin{prop}
+\begin{thm}\label{thm:vnhc-regularity}
     A VNHC of order \(k\) is regular if and only if \(dh_{p_a} = 0\) 
@@
     everywhere on the constraint manifold \(\Gamma\).
-\end{prop}
+\end{thm}
 \begin{proof}
+    % TODO: complete this proof with our notation
 \end{proof}
+
+From the proof for Theorem \ref{thm:vnhc-regularity}, a regular VNHC of order
+\(k\) can be stabilized by the output-linearizing phase-feedback controller
+(\ref{eqn:vnhc-torque-controller}).
+% TODO: Controller for VNHC in closed form
+\begin{equation}\label{eqn:vnhc-torque-controller}
+    \tau(q,p) = ? - k_p e - k_d \dot{e}
+\end{equation}
+where \(k_p, k_d \in \R_{>0}\) can be tuned using standard linear control
+techniques.
+
+Note that one generally cannot measure conjugate of momenta directly, as sensors
+on mechanical systems typically only measure the state \((q,\dot{q})\). To
+implement this controller in practice, one must compute \(p = M(q)\dot{q}\) at
+every iteration. In other words, this controller requires knowledge of the full
+state of the system.
 
@@
 %can solve for p_a and the closed-loop dynamics (qu,pu)_dot
+
+Now that we have found a controller to enforce a regular VNHC of order \(k\), we
+would like to solve for the closed-loop dynamics. Naturally, these dynamics
+should be specified through \((q_u, p_u)\) since \(q_a\) is a function of these
+as specified by \(h(q,p_u) = 0\).
+Unfortunately, \(\dot{q}_u\) depends on \(p_a\), and for general systems one
+cannot solve explicitly for \(p_a\) in terms of \((q_u,p_u)\). This is because
+the \(\dot{p}\) dynamics contains the coupling term 
+\((I_n \otimes p\tpose)\nabla_{q_u}M(q)p\).
+
+We now introduce a class of systems where explicitly solving for the closed-loop
+dynamics is feasible.
+
+\begin{defn}
+    A mechanical system is ------ if \(\nabla_{q_u}M(q) = 0\)
+\end{defn}
+
+\begin{thm}\label{thm:zero-dynamics}
+    Let \(\mathcal{H}\) be a ----- mechanical system satisfying Assumption
+    \ref{assm:H-is-simply-actuated}. Let
+    \(h(q,p_u) = 0\) be a regular VNHC of order \(k\) with constraint manifold
+    \(\Gamma\).
+    Then, on \(\Gamma\), the closed-loop dynamics are given by
+    \begin{equation}\label{eqn:qpu-dynamics}
+        \left.\begin{aligned}
+            \dot{q}_u &= M(q)p \\
+            \dot{p}_u &= -\nabla_{q_u}V(q) \\
+            \end{aligned}\right|_{\begin{array}{c}
+                h(q,p_u) = 0 \\ 
+                pa = g(q_u,p_u) \\
+            \end{array}}
+    \end{equation}
+    where
+    % TODO: This is too wide, shrink it somehow
+    \begin{equation}\label{eqn:g-qpu}
+        g(q_u,p_u) := \left.\left(dh_q \Minv(q) \simpleB \right)\inv 
+        \left(dh_{p_u} \nabla_{q_u}V(q) - dh_q \Minv(q)
+        \begin{bmatrix}
+            I_{n-k} \\
+            \Zmat{k \times (n-k)} \\
+        \end{bmatrix} p_u\right)\right|_{h(q,p_u) = 0}
+    \end{equation}
+\end{thm}
+\begin{proof}
+%TODO: Complete the proof for this theorem
+\end{proof}
+
+The most fascinating part of Theorem \ref{thm:zero-dynamics} is that it shows
+that the dynamics on \(\Gamma\) are entirely described by the \(2(n-k)\)
+unactuated coordinates, regardless of the number of degrees of freedom of the
+system. This means that \(\Gamma\) is a \(2(n-k)\)-dimensional surface from
+which the system cannot escape. This becomes especially important when analyzing
+systems of degree of underactuation one, since it means the constraint is always
+a curve which can be plotted on the plane with axes \(q_u\) and \(p_u\)
+regardless of the original dimension \(n\) of the system.
+
+In future sections of this thesis, we will apply a very specific type of
+constraint. The following corollay applies Theorem \ref{thm:zero-dynamics} to
+these constraints.
+
+\begin{cor}\label{cor:2d-zero-dynamics}
+    Suppose \(\mathcal{H}\) is a ---- mechanical system satisfying Assumption
+    \ref{assm:H-is-simply-actuated} which has degree of underactuation one.
+    Let \(h(q,p_u) = q_a - f(q_u,p_u)\) be a regular VNHC of order \((n-1)\),
+    where \(f\) is a suitably-defined smooth function.
+    Defining \(e_1 := (1,0,\ldots,0) \in \R^n\), the actuated momentum 
+    is given by
+    \begin{equation}\label{eqn:g-qupu}
+        p_a = -\left.\left(dh_q \Minv(q)
+        \begin{bmatrix}
+            \Zmat{1\times (n-1)}\\
+            I_{(n-1)}
+        \end{bmatrix}\right)\inv 
+        \left(\partial_{p_u}f \partial_{q_u}V + dh_q \Minv(q)e_1 p_u\right) 
+            \right|_{q_a =f(q_u,p_u)}
+    \end{equation}
+    Since \(q_u \in \Rt{T}\) for some \(T > 0\), \((q_u(t),p_u(t))\) traces out
+    a curve on the so-called \((q,p)\)-plane \(\Rt{T} \times \R\).
+\end{cor}
 

### 2020-08-11 latex/vnhcs.tex
Wrote the proof of regularity, but it's not compiling so that needs fixing.

@@
         \dot{q} = D\inv(q)p \\
-        \dot{p} = -\frac{1}{2} (I_n \otimes p\tpose) \nabla_q D\inv(q) p
+        \dot{p} = -\frac{1}{2} (\Id{n} \otimes p\tpose) \nabla_q D\inv(q) p
         - \nabla_q P(q) + B(q) \tau \\
@@
             \dot{Q} = \Minv(Q)P \\
-            \dot{P} = -\frac{1}{2} (I_n \otimes P\tpose) \nabla_Q \Minv(Q) P
+            \dot{P} = -\frac{1}{2} (\Id{n} \otimes P\tpose) \nabla_Q \Minv(Q) P
                 - \nabla_Q V(Q) + \simpleB \tau
@@
 
-% TODO: Write the [p' M p] matrix in our nice nottation
 \begin{assm}\label{assm:H-is-simply-actuated}
@@
             \dot{q} = \Minv p \\
-            \dot{p} = -\frac{1}{2} (I_n \otimes p\tpose) \nabla_q \Minv(q) p
-            - \nabla_q V(q) + \simpleB \tau \\
+            \dot{p} = -\frac{1}{2} \pdmat - \nabla_q V(q) + \simpleB \tau \\
         \end{cases}
@@
 \end{assm}
+\begin{notation}
+    We will write \(q_u \in \mathcal{Q}_u\), \(q_a \in \mathcal{Q}_a\) where
+    \(\mathcal{Q}_u \times \mathcal{Q}_a = \mathcal{Q}\). 
+    We also write
+    \(p_u \in \mathcal{P}_u = \R^{n-k}\) and 
+    \(p_a \in \mathcal{P}_a = \R^k\), so that 
+    \(p \in \mathcal{P} := \mathcal{P}_u \times \mathcal{P}_a\). 
+    In this manner, the phase space of our system can be written as
+    \(\mathcal{Q} \times \mathcal{P}\).
+\end{notation}
 
@@
     \[
-        \rank{\left(dh_q \Minv(q) + 
-          dh_{p_u} (I_n \otimes p\tpose)\nabla_{q_u}\Minv(q) 
+        \rank{\left(dh_q \Minv(q) - 
+          dh_{p_u} (\Id{n-k} \otimes p\tpose)\nabla_{q_u}\Minv(q) 
          \right)\simpleB} = k
@@
     % TODO: complete this proof with our notation
+    Let \(e = h(q,p) \in \R^k\). Then 
+    \begin{align*}
+        \dot{e} &= dh_q \dot{q} + dh_p \dot{p} \\
+                &= dh_q \Minv(q)p +{}  \\
+            & \begin{bmatrix} dh_{p_u} & dh_{p_a} \end{bmatrix}
+        \left( -\frac{1}{2} \begin{bmatrix}
+            (\Id{n-k} \otimes p\tpose) \nabla_{q_u}\Minv(q)p \\
+            (\Id{k} \otimes p\tpose) \nabla_{q_a}\Minv(q)p
+            \end{bmatrix} - \begin{bmatrix}
+            \nabla_{q_u}V(q) \\
+            \nabla_{q_a}V(q)
+        \end{bmatrix} + \simpleB \tau\right)
+    \end{align*}
+    If \(dh_{p_a} \neq \Zmat{k \times k}\) for some \((q,p)\) on \(\Gamma\), 
+    then \(\tau\) appears in \(\dot{e}\) and the VNHC is not of relative degree
+    \(\{2,2,\ldots,2\}\).
+    Hence, we must have that \(dh_{p_a} = \Zmat{k \times k}\) if we want the
+    VNHC to be regular. Proceeding with this assumption, we now find that
+    \(h : \mathcal{Q} \times \mathcal{P}_u \rightarrow \R^k\), which means
+    \[
+        \dot{e} = dh_q \Minv(q)p - 
+        dh_{p_u} \left(\frac{1}{2} \pudmat + \nabla_{q_u}V(q)\right)
+    \]
+    Taking one further derivative provides
+    \begin{align*}
+        \ddot{e} &= \diff{}{t}\left\{dh_q\right\}\Minv(q) p + 
+        dh_q \left(\sum\limits_{i=1}^n \pdiff{\Minv}{q_i}(q)\dot{q_i}\right)p + 
+        dh_q \Minv(q) \dot{p} - \\
+         & \diff{}{t}\left\{dh_{p_u}\right\}
+         \left(\frac{1}{2}\pudmat + \nabla_{q_u}V(q)\right) - \\
+         & dh_{p_u}\left(\frac{1}{2}\diff{}{t}\left\{\pudmat\right\} + 
+         \diff{}{t}\left\{\nabla_{q_u}V(q)\right\} \right)
+    \end{align*}
+    We will compute the explicit expression of \(\ddot{e}\) in pieces. 
+    First we begin with \(\diff{}{t}\left\{dh_q\right\}\). 
+    Since \(h = (h^1,\ldots,h^k)\) where 
+    \(h^i : \mathcal{Q} \times \mathcal{P}_u \rightarrow \R\), we have that 
+    \[
+        dh_q = \begin{bmatrix}
+            dh^1_q \\
+            \vdots \\
+            dh^k_q
+        \end{bmatrix} = \begin{bmatrix}
+        \pdiff{h^1}{q_1} & \cdots & \pdiff{h^1}{q_n} \\
+        \vdots & \ddots & \vdots \\
+        \pdiff{h^k}{q_1} & \cdots & \pdiff{h^k}{q_n}
+        \end{bmatrix}
+    \]
+    The derivative is taken element-wise in each matrix, yielding
+    \[
+        \diff{}{t}\left\{dh_q\right\} = \begin{bmatrix}
+            \sum_{j=1}^n \ppdiff{h^1}{q_1}{q_j}\dot{q}_j & \cdots & \sum_{j=1}^n
+            \ppdiff{h^1}{q_n}{q_j}\dot{q}_j \\
+            \vdots & \ddots & \vdots \\
+            \sum_{j=1}^n \ppdiff{h^k}{q_1}{q_j}\dot{q}_j & \cdots & \sum_{j=1}^n
+            \ppdiff{h^k}{q_n}{q_j}\dot{q}_j \\
+        \end{bmatrix} + \begin{bmatrix}
+            \sum_{l=1}^{n-k}\ppdiff{h^1}{q_1}{p_{u_l}}\dot{p}_{u_l} & \cdots &
+            \sum_{l=1}^{n-k}\ppdiff{h^1}{q_n}{p_{u_l}}\dot{p}_{u_l} \\
+            \vdots & \ddots & \vdots \\
+            \sum_{l=1}^{n-k}\ppdiff{h^k}{q_1}{p_{u_l}}\dot{p}_{u_l} & \cdots & 
+            \sum_{l=1}^{n-k}\ppdiff{h^k}{q_n}{p_{u_l}}\dot{p}_{u_l} \\
+        \end{bmatrix}
+    \]
+    It is straightforward computation to confirm that each row of this
+    derivative can be written in vector form as follows:
+    \begin{align*}
+        \diff{}{t}\left\{dh^i_q\right\} &= \begin{bmatrix}
+        \sum_{j=1}^n \ppdiff{h^i}{q_1}{q_j}\dot{q}_j & \cdots & 
+        \sum_{j=1}^n \ppdiff{h^i}{q_n}{q_j}\dot{q}_j 
+        \end{bmatrix} + \begin{bmatrix}
+        \sum_{l=1}^{n-k}\ppdiff{h^i}{q_1}{p_{u_l}}\dot{p}_{u_l} & \cdots &
+        \sum_{l=1}^{n-k}\ppdiff{h^i}{q_n}{p_{u_l}}\dot{p}_{u_l}
+        \end{bmatrix} \\
+       &= \dot{q}\tpose \Hess_q\left\{h^i\right\}\tpose + 
+       \dot{p}_u\tpose \partial_{p_u}\partial_{q}h^i \\
+       &= p\tpose \Minv(q)\Hess_q\left\{h^i\right\}\tpose -
+       \left(\frac{1}{2}p\tpose d\Minv_{q_u}(q)(\Id{n-k}\otimes p) + 
+       dV_{q_u}(q)\right)\partial_{p_u}\partial_{q}h^i
+    \end{align*}
+    This means that the entire matrix is given by
+    \begin{align*}
+        \diff{}{t}\left\{dh_q\right\} &= 
+        \left(\Id{n} \otimes 
+        \left(p\tpose \Minv(q)\right)\right)
+        \Hess_q\left\{h\right\}\tpose -{} \\
+      & \left(\frac{1}{2}\Id{n} \otimes 
+          \left(p\tpose d\Minv_{q_u}(q)(\Id{n-k}\otimes p)\right)
+        + \Id{n}\otimes dV_{q_u}(q)\right)\partial_{p_u}\partial_{q}h
+    \end{align*}
+    For the next segment, let \(e_j \in \R^n\) be the vector of all zeros,
+    except for a single \(1\) at position \(j\). We define
+    \(C_1 : \mathcal{Q} \times \mathcal{P} \rightarrow \R^{n\times n}\) by
+    \[
+        C_1(q,p) := \sum\limits_{j=1}^n \pdiff{\Minv}{q_j}(q)\dot{q}_j 
+        = \sum\limits_{j=1}^n \pdiff{\Minv}{q_j}(q)\left(e_j\tpose\Minv(q)p\right)
+    \]
+    Moving on, one can use a similar approach to \(\diff{}{t}\left\{dh_q\right\}\)
+    to find the derivative of \(dh_{p_u}\) in matrix form. It is given by
+    \begin{align*}
+        \diff{}{t}\left\{dh_{p_u}\right\} &= 
+        \left( \Id{n} \otimes \left(p\tpose \Minv(q)\right)\right) 
+        \partial_q\partial_{p_u}h -{} \\
+      & \left(\frac{1}{2}\Id{n} \otimes \left(\pudmat\right) + 
+      \Id{n} \otimes \nabla_{q_u}V(q)\right)\Hess_{p_u}\left\{h\right\}\tpose
+    \end{align*}
+    Now observe that the \(i^\text{th}\) row of
+    \[
+        \frac{1}{2} \diff{}{t} \left\{ \pudmat \right\}
+    \]
+    is given by
+    \begin{align*}
+        \frac{1}{2} \diff{}{t}\left\{p\tpose \pdiff{\Minv}{q_{u_i}}(q)p\right\}
+        = p\tpose \pdiff{\Minv}{q_{u_i}}(q) \dot{p} + 
+        \frac{1}{2} p\tpose \left( \sum_{j=1}^n \ppdiff{\Minv}{q_{u_i}}{q_j}
+        \dot{q}_j \right) p
+    \end{align*}
+    Defining \(C_2 : \mathcal{Q} \times \mathcal{P} \rightarrow \R^{n(n-k)\times n}\)
+    allows us to write the entire derivative in matrix form:
+    \begin{align*}
+        & \frac{1}{2} \diff{}{t} \left\{ \pudmat \right\} ={} \\
+        & (\Id{n-k} \otimes p\tpose)\nabla_{q_u}\Minv(q)
+        \left(-\frac{1}{2}\pudmat - \nabla_{q_u}V(q) + \simpleB \tau \right) +{}
+        \\
+        & \frac{1}{2}\left(\Id{n-k}\otimes p\tpose\right) C_2(q,p) p
+    \end{align*}
+    Finally, we compute the derivative of \(\nabla_{q_u}V(q)\)
+    \[
+        \diff{}{t}\left\{\nabla_{q_u}V(q)\right\} = 
+        \begin{bmatrix}
+            \sum_{j=1}^n \ppdiff{V}{q_{u_1}}{q_j}\dot{q}_j \\
+            \vdots \\
+            \sum_{j=1}^n \ppdiff{V}{q_{u_{n-k}}}{q_j}\dot{q}_j
+        \end{bmatrix = \partial_{q_u} \partial_q V(q) \dot{q}
+    \]
+    and so we get that
+    \[
+        \diff{}{t}\left\{\nabla_{q_u}V(q)\right\} = 
+        \partial_{q_u} \partial_q V(q) \Minv(q)p
+    \]
+    Putting this all together, we can find the explicit form for \(\ddot{e}\):
+    \begin{align*}
+        \ddot{e} &= \diff{}{t}\left\{dh_q\right\}\Minv(q)p + dh_q C_1(q,p)p -{}\\
+     & \frac{1}{2}dh_q\Minv(q)(\pdmat) + 
+        dh_q\Minv(q)\nabla_qV(q) -{}\\
+     & \diff{}{t}\left\{dh_{p_u}\right\}
+     \left(\frac{1}{2}\pudmat + \nabla_{q_u}V(q)\right) +{} \\
+     & dh_{p_u}(\Id{n-k}\otimes p\tpose)\nabla_{q_u}\Minv(q)\left(
+     \frac{1}{2}\pudmat + \nabla_{q_u}V(q)\right) -{}\\
+     \frac{1}{2}dh_{p_u}(\Id{n-k}\otimes p\tpose)C_2(q,p)p - 
+     dh_{p_u} \partial_{q_u}\partial_q V(q)\Minv(q)p +{} \\
+     \left(dh_q \Minv(q) - dh_{p_u}(\Id{n-k} \otimes
+     p\tpose)\nabla_{q_u}\Minv(q) \right) \simpleB \tau
+    \end{align*}
+    The VNHC \(h\) is regular iff \(e\) is of relative degree
+    \(\{2,\ldots,2\}\), which is true iff \(\tau\) appears in each element of
+    \(e\) on the constraint manifold. 
+    Letting \(\ddot{e} = E(q,p) + H(q,p)\tau\), this is equivalent to requiring
+    that the matrix \(H\) be full rank.
 \end{proof}
 
-From the proof for Theorem \ref{thm:vnhc-regularity}, a regular VNHC of order
-\(k\) can be stabilized by the output-linearizing phase-feedback controller
+Using the expression \(\ddot{e} = E(q,p) + H(q,p)\tau\) from the proof of 
+Theorem \ref{thm:vnhc-regularity}, a regular VNHC of order \(k\) can be
+stabilized by the output-linearizing phase-feedback controller
 (\ref{eqn:vnhc-torque-controller}).
-% TODO: Controller for VNHC in closed form
 \begin{equation}\label{eqn:vnhc-torque-controller}
-    \tau(q,p) = ? - k_p e - k_d \dot{e}
+    \tau(q,p) = -H\inv(q,p)\left(E(q,p) + k_p e + k_d \dot{e}\right)
 \end{equation}
-where \(k_p, k_d \in \R_{>0}\) can be tuned using standard linear control
-techniques.
+where \(k_p, k_d \in \R_{>0}\) are control parameters which can be tuned on the
+resulting linear system \(\ddot{e} = -k_p e - k_d\dot{e}\).
 
@@
 the \(\dot{p}\) dynamics contains the coupling term 
-\((I_n \otimes p\tpose)\nabla_{q_u}M(q)p\).
+\((\Id{n} \otimes p\tpose)\nabla_{q_u}M(q)p\).
 
@@
     where
-    % TODO: This is too wide, shrink it somehow
     \begin{equation}\label{eqn:g-qpu}
-        g(q_u,p_u) := \left.\left(dh_q \Minv(q) \simpleB \right)\inv 
+    \begin{aligned}
+        g(q,p_u) &:= \\
+           &\left.\left(dh_q \Minv(q) \simpleB \right)\inv 
         \left(dh_{p_u} \nabla_{q_u}V(q) - dh_q \Minv(q)
@@
         \end{bmatrix} p_u\right)\right|_{h(q,p_u) = 0}
+    \end{aligned}
     \end{equation}
@@
 \end{cor}
+
+\section{Summary of Results}
+%TODO: summarize the assumptions and results
 

### 2020-08-12 latex/vnhcs.tex
Finished all VNHC proofs
Fixed the errors that were stopping compilation and finished the proofs
in the VNHC chapter. I think I made an error in finding the closed loop
dynamics for general VNHCs, which needs to be discussed with Manfredi.

@@
 
-% TODO: Replace [p D_i p] here with our new notation
 The resulting Hamiltonian system reduces to (\ref{eqn:hamiltonian}).
@@
 \section{Simply Actuated Hamiltonian Systems}
-% TODO: Describe the change of coordinates to get into q_u/q_a mode and show
+% Describe the change of coordinates to get into q_u/q_a mode and show
 % that the actuator directly affects pa but not pu. Use M for simply actuated
@@
 
-% TODO: Replace the [P M_i P] with new notation
 \begin{thm}
@@
 \section{Virtual Nonholonomic Constraints}
-% TODO: Motivation for VNHCs, they are an extension of VHCs, etc.
+% Motivation for VNHCs, they are an extension of VHCs, etc.
 Let us imagine a child on a swing, who wants to go as high as possible. The
@@
 
-% TODO: Perform the full development of VNHCs, including its stabilizing
+% Perform the full development of VNHCs, including its stabilizing
 % controller.
@@
 \begin{proof}
-    % TODO: complete this proof with our notation
     Let \(e = h(q,p) \in \R^k\). Then 
@@
     We will compute the explicit expression of \(\ddot{e}\) in pieces. 
+% d/dt(dh_q)
     First we begin with \(\diff{}{t}\left\{dh_q\right\}\). 
@@
     \end{align*}
-    For the next segment, let \(e_j \in \R^n\) be the vector of all zeros,
+% C_1(q,p)
+    For the next segment, let \(1_j \in \R^n\) be the vector of all zeros,
     except for a single \(1\) at position \(j\). We define
@@
         C_1(q,p) := \sum\limits_{j=1}^n \pdiff{\Minv}{q_j}(q)\dot{q}_j 
-        = \sum\limits_{j=1}^n \pdiff{\Minv}{q_j}(q)\left(e_j\tpose\Minv(q)p\right)
+        = \sum\limits_{j=1}^n \pdiff{\Minv}{q_j}(q)\left(1_j\tpose\Minv(q)p\right)
     \]
+% d/dt(dh_pu)
     Moving on, one can use a similar approach to \(\diff{}{t}\left\{dh_q\right\}\)
@@
     \end{align*}
-    Defining \(C_2 : \mathcal{Q} \times \mathcal{P} \rightarrow \R^{n(n-k)\times n}\)
-    allows us to write the entire derivative in matrix form:
+    Define \(C_2 : \mathcal{Q} \times \mathcal{P} \rightarrow \R^{n(n-k)\times n}\) 
+    by
+    \[
+        C_2(q,p) := \begin{bmatrix}
+            \sum_{j=1}^n \ppdiff{\Minv}{q_{u_1}}{q_j} (1_j\tpose \Minv(q)p) \\
+            \vdots \\
+            \sum_{j=1}^n \ppdiff{\Minv}{q_{u_{n-k}}}{q_j} (1_j\tpose \Minv(q)p) \\
+        \end{bmatrix}
+    \]
+    This allows us to write the entire derivative in matrix form:
     \begin{align*}
@@
     \end{align*}
-    Finally, we compute the derivative of \(\nabla_{q_u}V(q)\)
-    \[
+% d/dt(nabla_qu V)
+    Finally, we compute the derivative of \(\nabla_{q_u}V(q)\):
+    \begin{align*}
         \diff{}{t}\left\{\nabla_{q_u}V(q)\right\} = 
@@
             \sum_{j=1}^n \ppdiff{V}{q_{u_{n-k}}}{q_j}\dot{q}_j
-        \end{bmatrix = \partial_{q_u} \partial_q V(q) \dot{q}
-    \]
-    and so we get that
-    \[
-        \diff{}{t}\left\{\nabla_{q_u}V(q)\right\} = 
-        \partial_{q_u} \partial_q V(q) \Minv(q)p
-    \]
+        \end{bmatrix} = \partial_{q_u} \partial_q V(q) \dot{q} 
+        = \partial_{q_u} \partial_q V(q) \Minv(q)p
+    \end{align*}
+% Full ddot(e)
     Putting this all together, we can find the explicit form for \(\ddot{e}\):
@@
      \frac{1}{2}\pudmat + \nabla_{q_u}V(q)\right) -{}\\
-     \frac{1}{2}dh_{p_u}(\Id{n-k}\otimes p\tpose)C_2(q,p)p - 
+     & \frac{1}{2}dh_{p_u}(\Id{n-k}\otimes p\tpose)C_2(q,p)p - 
      dh_{p_u} \partial_{q_u}\partial_q V(q)\Minv(q)p +{} \\
-     \left(dh_q \Minv(q) - dh_{p_u}(\Id{n-k} \otimes
+     & \left(dh_q \Minv(q) - dh_{p_u}(\Id{n-k} \otimes
      p\tpose)\nabla_{q_u}\Minv(q) \right) \simpleB \tau
     \end{align*}
-    The VNHC \(h\) is regular iff \(e\) is of relative degree
-    \(\{2,\ldots,2\}\), which is true iff \(\tau\) appears in each element of
-    \(e\) on the constraint manifold. 
-    Letting \(\ddot{e} = E(q,p) + H(q,p)\tau\), this is equivalent to requiring
-    that the matrix \(H\) be full rank.
+% End of Proof
+    For shorthand, we'll write \(\ddot{e} = E(q,p) + H(q,p)\tau\) where \(E\)
+    and \(H\) are defined appropriately.
+    From the definition of regularity, the VNHC \(h\) is regular 
+    iff \(e\) is of relative degree \(\{2,\ldots,2\}\), which is true 
+    iff \(\tau\) appears in each row of \(\ddot{e}\) on \(\Gamma\). 
+    This is equivalent to requiring that the matrix \(H\) be full rank
+    everywhere on the constraint manifold, proving the theorem.
 \end{proof}
@@
 
-%TODO: Special case: when dM/dqu = 0, show that we have a nice form and that we
+% Special case: when dM/dqu = 0, show that we have a nice form and that we
 %can solve for p_a and the closed-loop dynamics (qu,pu)_dot
@@
 Now that we have found a controller to enforce a regular VNHC of order \(k\), we
-would like to solve for the closed-loop dynamics. Naturally, these dynamics
-should be specified through \((q_u, p_u)\) since \(q_a\) is a function of these
+would like to solve for the closed-loop dynamics. Intuitively, these dynamics
+should be parameterized by \((q_u, p_u)\) since \(q_a\) is a function of these
 as specified by \(h(q,p_u) = 0\).
@@
     \ref{assm:H-is-simply-actuated}. Let
-    \(h(q,p_u) = 0\) be a regular VNHC of order \(k\) with constraint manifold
-    \(\Gamma\).
-    Then, on \(\Gamma\), the closed-loop dynamics are given by
+    \(h(q,p_u) = 0\) be a regular VNHC of order \(k\) 
+    with constraint manifold \(\Gamma\). Suppose that on \(\Gamma\) one can
+    solve linearly for \(q_a\) as a function of \(q_u,p_u\).
+    Then the closed-loop dynamics are given by
     \begin{equation}\label{eqn:qpu-dynamics}
@@
 \begin{proof}
-%TODO: Complete the proof for this theorem
+    Setting \(e = h(q,p_u)\) and using the fact that 
+    \(\nabla_{q_u}\Minv(q) = 0\), we find that
+    \[
+        \dot{e} = dh_q\Minv(q)p - dh_{p_u}\nabla_{q_u}V(q)
+    \]
+    Observe that
+    \begin{align*}
+        dh_q\Minv(q)p &= dh_q\Minv(q) \begin{bmatrix}p_u \\ p_a \end{bmatrix} \\
+          &= dh_q\Minv(q) \begin{bmatrix}
+              \Id{n-k} & \Zmat{(n-k)\times k} \\
+              \Zmat{k \times(n-k)} & \Id{k} 
+              \end{bmatrix} \begin{bmatrix} p_u \\ p_a \end{bmatrix} \\
+        &= dh_q\Minv(q)\begin{bmatrix} \Id{n-k} \\ \Zmat{k \times
+        (n-k)}\end{bmatrix} p_u + dh_q \Minv(q)\simpleB p_a
+    \end{align*}
+    On the constraint manifold, we have \(e = \dot{e} = 0\), which means
+    \[
+        dh_q\Minv(q)\simpleB p_a = dh_{p_u}\nabla_{q_u}V(q) -
+            dh_q\Minv(q)\begin{bmatrix} 
+            \Id{n-k} \\ \Zmat{k \times (n-k)}\end{bmatrix} p_u
+    \]
+    Since \(h\) is regular and \(\nabla_{q_u}\Minv(q) = 0\), we have that 
+    \[
+        \rank{dh_q\Minv(q)\simpleB} = k
+    \]
+    Solving for \(p_a\) gives
+    \[
+        p_a = \left(dh_q\Minv(q)\simpleB\right)\inv
+        \left(dh_{p_u}\nabla_{q_u}V(q) - 
+            dh_q\Minv(q)\begin{bmatrix} 
+        \Id{n-k} \\ \Zmat{k \times (n-k)}\end{bmatrix} p_u \right)
+    \]
+    This yields a function \(p_a(q,p_u)\). However, on \(\Gamma\) we have 
+    \(h(q,p_u) = 0\) and can solve for \(q_a\) in terms of \((q_u,p_u)\).
+    Hence, on \(\Gamma\) we have \(p_a = g(q_u,p_u)\). Since \(q_a\) and
+    \(p_a\) can be computed directly from \((q_u,p_u)\), the dynamics on
+    \(\Gamma\) are parameterized only by \((\dot{q}_u,\dot{p}_u)\).
 \end{proof}
 
-The most fascinating part of Theorem \ref{thm:zero-dynamics} is that it shows
-that the dynamics on \(\Gamma\) are entirely described by the \(2(n-k)\)
-unactuated coordinates, regardless of the number of degrees of freedom of the
-system. This means that \(\Gamma\) is a \(2(n-k)\)-dimensional surface from
-which the system cannot escape. This becomes especially important when analyzing
-systems of degree of underactuation one, since it means the constraint is always
-a curve which can be plotted on the plane with axes \(q_u\) and \(p_u\)
-regardless of the original dimension \(n\) of the system.
+Theorem \ref{thm:zero-dynamics} shows that, for a particular class of systems
+and constraints, the dynamics on \(\Gamma\) are entirely described by the \(2(n-k)\)
+unactuated coordinates. 
+This is trueregardless of the number of degrees of freedom of the system. 
+This means that \(\Gamma\) is a \(2(n-k)\)-dimensional surface inside 
+\(\mathcal{Q}\times \mathcal{P}\).
 
-In future sections of this thesis, we will apply a very specific type of
-constraint. The following corollay applies Theorem \ref{thm:zero-dynamics} to
-these constraints.
+The following corollay applies Theorem \ref{thm:zero-dynamics} to
+systems with only one unactuated coordinate.
 
@@
     \ref{assm:H-is-simply-actuated} which has degree of underactuation one.
-    Let \(h(q,p_u) = q_a - f(q_u,p_u)\) be a regular VNHC of order \((n-1)\),
+    Let \(h(q,p_u) = 0\) be a regular VNHC of order \((n-1)\) of the form
+    \(h(q,p_u) = q_a - f(q_u,p_u)\)
     where \(f\) is a suitably-defined smooth function.
@@
     \end{equation}
-    Since \(q_u \in \Rt{T}\) for some \(T > 0\), \((q_u(t),p_u(t))\) traces out
-    a curve on the so-called \((q,p)\)-plane \(\Rt{T} \times \R\).
+    Since \(q_u \in \Rt{T}\) for some \(T \in ]0,\infty]\) and \(p_u \in \R\),
+    the orbit \((q_u(t),p_u(t))\) traces out
+    a curve on the 2D-plane \(\Rt{T} \times \R\) which we call the \((q,p)\)-plane.
 \end{cor}

### 2020-08-20 latex/vnhcs.tex
Made minor edits to the wording of a proof

@@
     From the definition of regularity, the VNHC \(h\) is regular 
-    iff \(e\) is of relative degree \(\{2,\ldots,2\}\), which is true 
-    iff \(\tau\) appears in each row of \(\ddot{e}\) on \(\Gamma\). 
-    This is equivalent to requiring that the matrix \(H\) be full rank
-    everywhere on the constraint manifold, proving the theorem.
+    when \(e\) is of relative degree \(\{2,\ldots,2\}\), which is true 
+    if and only if
+    %\(\tau\) appears in each row of \(\ddot{e}\) on \(\Gamma\). 
+    one can solve for \(\tau\) when \(\ddot{e} = 0\).
+    This is equivalent to requiring that the matrix \(H\) be
+    %full rank everywhere on the constraint manifold, 
+    invertible,
+    proving the theorem.
 \end{proof}
@@
 
+% TODO: Come up with a name for this type of system
 \begin{defn}
@@
 
+% TODO: Do we need the assumption that one can solve for qa linearly?
 \begin{thm}\label{thm:zero-dynamics}

### 2020-08-20 latex/vnhcs.tex
Added labels to some sections and theorems

@@
 % ---------- Lagrangian Mechanics ---------- % 
-\subsection{Lagrangian Mechanics}
+\subsection{Lagrangian Mechanics}\label{sec:lagrangian-mechanics}
 
@@
 
-\begin{thm}
+\begin{thm}\label{thm:simply-actuated}
     Under Assumptions \ref{assm:B-const},\ref{assm:B-perp}, and
@@
 \begin{thm}\label{thm:vnhc-regularity}
-    A VNHC of order \(k\) is regular if and only if \(dh_{p_a} = 0\) 
+    A VNHC \(h(q,p) = 0\) of order \(k\) is regular if and only if \(dh_{p_a} = 0\) 
     and
@@
     everywhere on the constraint manifold \(\Gamma\).
+    That is, a regular VNHC is a function with domain 
+    \(\mathcal{Q} \times \mathcal{P}_u\) which satisfies the rank condition.
+    condition.
 \end{thm}

### 2020-08-20 latex/vlp.tex
Wrote the dynamics of the VLP

@@
 \section{Motivation}
+The variable length pendulum (VLP) is modelled as a point with mass \(m\)
+connected to a pivot by a massless rod of length \(l\).
+
 % TODO: Why is this a good problem to study? What is the problem? What would a human do on a standing swing?
+
+Before computing the dynamics of the VLP, we must make it clear that this
+VLP model uses \(l\) directly as the control input to the system, rather than the
+generalized forces \(\tau\) described in Section \ref{sec:lagrangian-mechanics}.
+Since the control input is not considered a configuration variable of the
+system, the configuration of the VLP is
+the angle \(q \in \mathbb{S}^1\) between the rod and the vertical axis. We set
+\(q = 0\) when the mass is at the bottom of its swing.
+
+With this in mind, we can now compute the Hamiltonian dynamics of the system.
+The cartesian position of the mass is given by \(x = (l\sin(q),l\cos(q))\),
+which has velocity 
+\(\dot{x} = (\dot{l}\sin(q) + l\cos{q}\dot{q}, \dot{l}\cos(q) - l\sin{q}\dot{q})\)
+Computing the kinetic energy \(T\) yields
+\[
+   T(q,\dot{q}) = \frac{1}{2}m\norm{\dot{x}}^2 = \frac{1}{2}m\left(\dot{l}^2 +
+   l^2\dot{q}^2\right)
+\]
+The potential energy \(P\) with respect to the pivot (under a gravitational
+acceleration \(g\)) is
+\[
+   P(q) = -mgl\cos(q)
+\]
+which means, the Lagrangian is
+\[
+   \mathcal{L}(q,\dot{q}) = \frac{1}{2}m\left(\dot{l}^2 + l^2\dot{q}^2\right)
+      + mgl\cos(q)
+\]
+Computing the conjugate of momentum to \(q\), we get 
+\[
+   p = \pdiff{\mathcal{L}}{\dot{q}} = ml^2\dot{q}
+\]
+Since \(q\) is not actuated by any generalized force \(\tau\), the system
+is simply actuated. 
+Solving for the Hamiltonian \(\mathcal{H} = \dot{q}p - \mathcal{L}\)
+and applying Theorem \ref{thm:simply-actuated}, we find the dynamics of the VLP
+are given in Hamiltonian form by (\ref{eqn:vlp-hamiltonian}). 
+\begin{align}\label{eqn:vlp-hamiltonian}
+   \mathcal{H}(q,p) &= \frac{p^2}{2ml^2} - \frac{1}{2}m\dot{l}^2 - mgl\cos(q) \\
+     &\begin{cases}
+        \dot{q} = \frac{p}{ml^2} \\
+        \dot{p} = -mgl\sin(q) \\
+   \end{cases} \nonumber
+\end{align}
+
 \section{The VLP Constraint}
 % TODO: Go through the development of the constraint and prove it gains energy
+% TODO: Explain how we came up with the constraint first. Talk about the child
+% on a swing, the paper showing the time-optimal control input, and describe why
+% we want to remove time from the equation. Discuss the lmin and lmax choices.
+% Why do we need to use a VNHC instead of a VHC: we are modelling the
+% time-optimal controller in a time-independent way, which requires knowledge of
+% which quadrant in (q,p)-space we're in. This means using the angle theta,
+% which is inherently a VNHC.
 

### 2020-08-20 latex/vlp.tex
Added the dynamics of the VLP
I managed to explain why we can ignore the dynamics of l, and treat it
as a control input instead of a configuration variable.

--- 2020-06-10 latex/vlp.tex
@@
 \section{Motivation}
-% TODO: Why is this a good problem to study? What is the problem? What would a human do on a standing swing?
+The variable length pendulum (VLP) is a classical dynamical system, which is often
+used to model a the motion of a person on a standing swing
+\cite{dynamics_periodic_vlp,pumping_swing_standing_squatting,how_to_pump_a_swing}.
+% TODO: Why else is it useful? Apparently it models cranes, satellites, etc
+% TODO: Motivate more why this is a good problem to study
+
+The variable length pendulum (VLP) is modelled as a point with mass \(m\)
+connected to a pivot by a massless rod of varying length. Label
+\(q \in \mathbb{S}^1\) as the angle from the vertical line to the rod. In a
+realizable VLP, the rod  has length \(l\) which varies between some minimum
+length \(\underline{l} \geq 0\) and some maximum length 
+\(\overline{l} > \underline{l}\). The configuration of the VLP is the vector
+\(\mathbf{q} := (q,l) \in \mathbb{S}^1 \times [\underline{l},\overline{l}]\).
+
+Using this configuration, we will compute the Hamiltonian dynamics of the system.
+The cartesian position of the mass at the tip of the pendulum
+is given by \(x = (l\sin(q),l\cos(q))\), while its velocity is
+\(\dot{x} = (\dot{l}\sin(q) + l\cos{q}\dot{q}, \dot{l}\cos(q) - l\sin{q}\dot{q})\).
+Computing the kinetic energy \(T\) yields
+\[
+   T(\mathbf{q},\dot{\mathbf{q}}) = 
+   \frac{1}{2}m\norm{\dot{x}}^2 = \frac{1}{2}m\left(\dot{l}^2 + l^2\dot{q}^2\right)
+\]
+The potential energy \(P\) with respect to the pivot (under a gravitational
+acceleration \(g\)) is
+\[
+   P(\mathbf{q}) = -mgl\cos(q)
+\]
+Collecting the kinetic energy into a quadratic form, we get the Lagrangian
+\[
+   \mathcal{L}(\mathbf{q},\dot{\mathbf{q}}) 
+   = \frac{1}{2} \dot{\mathbf{q}}\tpose D(\mathbf{q})\dot{\mathbf{q}} - P(\mathbf{q})
+   = \frac{1}{2}
+   \begin{bmatrix} \dot{q} & \dot{l} \end{bmatrix}
+   \begin{bmatrix}
+      ml^2 & 0 \\
+      0 & m \\
+   \end{bmatrix}
+   \begin{bmatrix} 
+      \dot{q} \\ \dot{l}
+   \end{bmatrix}
+   + mgl\cos(q)
+\]
+Computing the conjugate of momenta to \(\mathbf{q}\), we get 
+\[
+   \mathbf{p} := \begin{bmatrix} p \\ p_l \end{bmatrix} 
+   = \begin{bmatrix} ml^2\dot{q} \\ m\dot{l} \end{bmatrix} 
+\]
+Performing the Legendre transform on \(\mathcal{L}\) and setting
+\(M(\mathbf{q}) := D(\mathbf{q})\), \(V(\mathbf{q}) := P(\mathbf{q})\),
+we find the Hamiltonian is equal to the total mechanical energy of the system:
+\[
+   \mathcal{H} = E = \frac{1}{2}\dot{\mathbf{p}}\tpose \Minv(\mathbf{q})
+   \dot{\mathbf{p}}+V(\mathbf{q})
+\]
+Taking the appropriate derivatives, the dynamics of the VLP in Hamiltonian form
+are described in (\ref{eqn:vlp-hamiltonian-with-pl}). 
+\begin{align}\label{eqn:vlp-hamiltonian-with-pl}
+   \mathcal{H} &= \frac{1}{2} \begin{bmatrix} p & p_l \end{bmatrix}
+      \begin{bmatrix}
+         \frac{1}{ml^2}  & 0 \\
+         0 & \frac{1}{m}
+      \end{bmatrix} \begin{bmatrix} p \\ p_l \end{bmatrix} - mgl\cos(q) \\
+     &\begin{cases}
+        \dot{q} = \frac{p}{ml^2} \\
+        \dot{l} = \frac{p_l}{m} \\
+        \dot{p} = -mgl\sin(q) \\
+        \dot{p}_l = \frac{p^2}{ml^3} + mg\cos(q) + \tau \\
+   \end{cases} \nonumber
+\end{align}
+The control input is a force \(\tau \in \mathbb{R}\) affecting the dynamics of
+\(p_l\), acting colinearly with direction of the rod.
+We assume the force does not affect the dynamics of \(p\) in any way -
+that is, there is no lateral force such as when a person is pushed on a swing.
+In this way, the dynamics of \((q,p)\) and \((l,p_l)\) are decoupled.
+
+This decoupling is extremely useful in simplifying the dynamics. Since one can
+write set \(\dot{p}_l\) to any desired function, we can suppose (with some abuse
+of notation) that \(l\) is tracking some function \(l(t)\). 
+The closed loop dynamics of the system will be described exclusively by
+\((\dot{q},\dot{p})\) with \(l = l(t)\); 
+the fact that \(p_l\) does not appear in these closed-loop dynamics allows us to
+ignore the subdynamics \((\dot{l},\dot{p}_l)\) entirely.
+
+What this means is we can treat \(l\) as the control input directly, rather than
+modelling it as a configuration variable. Re-deriving the Hamiltonian and the
+dynamics with this in mind, we get the system 
+(\ref{eqn:vlp-hamiltonian}) with phase \((q,p) \in \mathbb{S}^1 \times \R\). 
+Note that the control input \(l(t)\) and its derivative \(\dot{l}(t)\) are both
+known variables.
+\begin{align}\label{eqn:vlp-hamiltonian}
+   \mathcal{H}(q,p) &= \frac{p^2}{ml^2} - \frac{1}{2}\dot{l}^2 - mgl\cos(q) \\
+     &\begin{cases}
+        \dot{q} = \frac{p}{ml^2} \\
+        \dot{p} = -mgl\sin(q) \\
+      \end{cases}\nonumber
+\end{align}
+In particular, the Hamiltonian of this simplified model is no longer equal to
+the total mechanical energy of the system, which is given by
+(\ref{eqn:vlp-energy}).
+\begin{equation}\label{eqn:vlp-energy}
+   E(q,p) = \frac{p^2}{ml^2} + \frac{1}{2}\dot{l}^2 - mgl\cos(q)
+\end{equation}
 \section{The VLP Constraint}
 % TODO: Go through the development of the constraint and prove it gains energy
+% TODO: Explain how we came up with the constraint first. Talk about the child
+% on a swing, the paper showing the time-optimal control input, and describe why
+% we want to remove time from the equation. Discuss the lmin and lmax choices.
+% Why do we need to use a VNHC instead of a VHC: we are modelling the
+% time-optimal controller in a time-independent way, which requires knowledge of
+% which quadrant in (q,p)-space we're in. This means using the angle theta,
+% which is inherently a VNHC.
 

### 2020-08-21 latex/vlp.tex
Edited the description of the VLP and added the figure

@@
 The variable length pendulum (VLP) is modelled as a point with mass \(m\)
-connected to a pivot by a massless rod of varying length. Label
-\(q \in \mathbb{S}^1\) as the angle from the vertical line to the rod. In a
-realizable VLP, the rod  has length \(l\) which varies between some minimum
+connected to a pivot by a massless rod of varying length \(l\) and angle 
+\(q \in \mathbb{S}^1\) from the vertical, as is seen in Figure
+\ref{fig:vlp_model}. 
+In a realistic VLP, the rod length \(l\) varies between some minimum
 length \(\underline{l} \geq 0\) and some maximum length 
@@
 \(\mathbf{q} := (q,l) \in \mathbb{S}^1 \times [\underline{l},\overline{l}]\).
+
+\begin{figure}
+   \centering
+   \includestandalone[width=0.4\textwidth]{images/vlp_model}
+   \caption{The variable length pendulum with a mass at the
+   tip.}\label{fig:vlp_model}
+\end{figure}
 

### 2020-08-21 latex/vlp.tex
Better motivaated why VNHCss are important and why we study the VLP

@@
 \section{Motivation}
-The variable length pendulum (VLP) is a classical dynamical system, which is often
-used to model a the motion of a person on a standing swing
-\cite{dynamics_periodic_vlp,pumping_swing_standing_squatting,how_to_pump_a_swing}.
-% TODO: Why else is it useful? Apparently it models cranes, satellites, etc
-% TODO: Motivate more why this is a good problem to study
+The variable length pendulum (VLP) is a classical underactuated dynamical system
+which is often used to model the motion of a person on a swing
+\cite{pumping_swing_standing_squatting,how_to_pump_a_swing}.
+The VLP also represents the swinging of a crane, the (simplified) motion of a
+gymnast on a bar \cite{pendulum_length_giant_gymnastics}, and the
+tuned-mass-damper systems which stabilize skyscrapers
+\cite{vlp_tuned_mass_damper}, among others.
 
-The variable length pendulum (VLP) is modelled as a point with mass \(m\)
-connected to a pivot by a massless rod of varying length \(l\) and angle 
+The motion of the VLP has been well studied (see for instance
+\cite{dynamics_periodic_vlp}, and many control mechanisms exist
+to stabilize trajectories of the system. While many of these control mechanism
+are time-dependent, \cite{vlp_energy_shaping}
+offers a time-independent technique to inject energy into the system and
+stabilize desired energy level sets. The authors designed a controller through a
+technique called \textit{energy shaping}, and proved that their control
+mechanism would allow the VLP to achieve any desired energy level set.
+However, the energy injection mechanism is
+ad-hoc in the sense that it is not derived from any human-like behaviour. In
+this chapter, we will use VNHCs to provably stabilize energy level sets as well;
+the difference is that the control mechanism we derive will stay true to human
+motion and can be implemented by any person on a swing.
+
+\section{Dynamics of the Variable Length Pendulum}
+% Derive the dynamics of the VLP in Hamiltonian form
+We will model the VLP as a point mass \(m\)
+connected to a fixed pivot by a massless rod of varying length \(l\) with angle 
 \(q \in \mathbb{S}^1\) from the vertical, as is seen in Figure
 \ref{fig:vlp_model}. 
+We will also ignore any damping and frictional forces in this model.
 In a realistic VLP, the rod length \(l\) varies between some minimum
@@
    \includestandalone[width=0.4\textwidth]{images/vlp_model}
-   \caption{The variable length pendulum with a mass at the
-   tip.}\label{fig:vlp_model}
+   \caption{The representation of the variable length pendulum as a mass at the
+      tip of a massless rod.}\label{fig:vlp_model}
 \end{figure}
@@
 \end{equation}
+
 \section{The VLP Constraint}

### 2020-08-21 latex/vnhcs.tex
Better motivaated why VNHCss are important and why we study the VLP

@@
 and improving bipedal locomotion \cite{nhvc_dynamic_walking,
-hybrid_zero_dynamics_bipedal_nhvcs,nhvc_incline_walking,output_nhvc_bipedal_control}.
+hybrid_zero_dynamics_bipedal_nhvcs,output_nhvc_bipedal_control}.
+Most interestingly, this nonholonomic approach has been shown to be more robust
+than standard VHCs when applied to bipedal robotics \cite{nhvc_incline_walking}.
+This suggests that virtual nonholonomic constraints are more capable of
+stabilizing specific energy levels while rejecting disturbances, all while
+producing realistic biological motion. It is for this reason we choose to study
+nonholonomic constraints in this thesis.
 

### 2020-08-24 latex/vlp.tex
Fixed a brace which was causing an error
Edited the first chapter and minorly edited the VLP chapter, to send it to Manfredi

@@
 The motion of the VLP has been well studied (see for instance
-\cite{dynamics_periodic_vlp}, and many control mechanisms exist
+\cite{dynamics_periodic_vlp}), and many control mechanisms exist
 to stabilize trajectories of the system. While many of these control mechanism
@@
 However, the energy injection mechanism is
-ad-hoc in the sense that it is not derived from any human-like behaviour. In
-this chapter, we will use VNHCs to provably stabilize energy level sets as well;
-the difference is that the control mechanism we derive will stay true to human
-motion and can be implemented by any person on a swing.
+ad-hoc in the sense that it is not derived from natural behaviour. In
+this chapter we will use VNHCs to inject energy into the VLP in a
+time-independent way;
+the difference is that our control mechanism will maintain the structured motion
+of a human on a swing.
 
@@
 We assume the force does not affect the dynamics of \(p\) in any way -
-that is, there is no lateral force such as when a person is pushed on a swing.
+that is, the control input cannot enact any lateral force on the pendulum.
 In this way, the dynamics of \((q,p)\) and \((l,p_l)\) are decoupled.
 
-This decoupling is extremely useful in simplifying the dynamics. Since one can
-write set \(\dot{p}_l\) to any desired function, we can suppose (with some abuse
+This decoupling is extremely useful for simplifying the dynamics. Since one can
+use \(\tau\) to set \(\dot{p}_l\) to any desired function, we can assume (with abuse
 of notation) that \(l\) is tracking some function \(l(t)\). 

### 2020-08-24 latex/vnhcs.tex
Edited the first chapter and minorly edited the VLP chapter, to send it to Manfredi

@@
 
-Rather than modeling a mechanical system with point masses and constraints,
+Rather than modeling a mechanical system by cartesion positions and constraints,
 it is often feasible to represent the position of the system using \(n\)
@@
 \textit{degrees of freedom} (DOF) of the system \cite{greenwood_dynamics}.
+
+% TODO:
+\textbf{TODO: Image comparing Newtonian to g.c.}
 
@@
 ordinary differential equations (ODEs) obtained from the \textit{Euler-Lagrange
-equations} with \textit{generalized input forces} \(\tau \in \R^k\) 
-(\ref{eqn:el-eqns-general}). 
+equations} (\ref{eqn:el-eqns-general}) with \textit{generalized input forces} 
+\(\tau \in \R^k\) . 
 \begin{equation}\label{eqn:el-eqns-general}
@@
 \(p\) is the \textit{conjugate of momenta to \(q\)}.
-Note that each \(p_i\) is a linear function of \(\dot{q}\), so one can typically
-solve for \(\dot{q}(q,p)\) by inverting the expressions from (\ref{eqn:p-i}).
+Note that each \(p_i\) is a linear function of \(\dot{q}\), and one can typically
+solve for \(\dot{q}(q,p)\) by inverting all the expressions from (\ref{eqn:p-i}).
 The combined vector \((q,p) \in \mathcal{Q}\times\R^n\) is called a 
@@
 \begin{equation}\label{eqn:hamiltonian-legendre}
-    \mathcal{H}(q,p) = p\tpose \dot{q}(q,p) - \mathcal{L}(q,\dot{q}(q,p))
+    \mathcal{H}(q,p) := p\tpose \dot{q}(q,p) - \mathcal{L}(q,\dot{q}(q,p))
 \end{equation}
@@
 
-A set of coordinates \(\{q,p\}\) which satisfy Hamilton's equations 
+Any set of coordinates \(\{q,p\}\) which satisfy Hamilton's equations 
 under the Hamiltonian \(\mathcal{H}\) are
@@
 Given a Hamiltonian mechanical system (\ref{eqn:hamiltonian}), it is not obvious
-how the input forces \(\tau\) affect the conjugate of momenta \(p_i\). 
-This is because \(\tau\) is transformed by the input matrix \(B(q)\), which may
-be quite complicated.
+how any particular input force \(\tau_i\) affects the conjugate of momenta. 
+This is because \(\tau\) is transformed by the input matrix \(B(q)\) before
+entering the EOM.
 
@@
     coordinates}, while the remaining \(k\) coordinates \(q_a\) are called the
-    \textit{actuated coordinates}. We write \(q = (q_u, q_a)\) and the
-    corresponding conjugate of momenta \(p = (p_u, p_a)\) are the called the
+    \textit{actuated coordinates}. We write \(q = (q_u, q_a)\); the
+    corresponding conjugate of momenta \(p = (p_u, p_a)\) are called the
     unactuated and actuated momenta (respectively).
@@
     which is right semi-orthogonal 
-    (\ie \(B^\perp(B^\perp)\tpose = \Id{(n-k)}\))
+    (\(B^\perp(B^\perp)\tpose = \Id{(n-k)}\))
     and which is a left-annihilator for \(B\)
-    (\ie \(B^\perp B = \Zmat{(n-k) \times k}\)).
+    (\(B^\perp B = \Zmat{(n-k) \times k}\)).
 \end{assm}
@@
 \begin{assm}\label{assm:B-orthogonal}
-    Assume without loss of generality that the input matrix \(B\) is left
-    semi-orthogonal.
-    That is, assume \(B\tpose B = \Id{k}\).
+    Assume \textbf{without loss of generality} that the input matrix \(B\) is
+    left semi-orthogonal (\(B\tpose B = \Id{k}\)).
 \end{assm}
@@
 \]
-where \(\sigma_i \neq 0\) because \(B\) is full-rank \textbf{SOURCE?}.
+% TODO: Find a source for SVD
+where \(\sigma_i \neq 0\) because \(B\) is full-rank \textbf{TODO: source?}.
 Defining \(T \in \R^{k \times k}\) by
@@
     Since the matrix \(\mathbf{B}\) is orthogonal,
-    \((\mathbf{B}_i)\tpose \mathbf{B}\tpose_j = (\mathbf{B}_i)\tpose (\mathbf{B}^{-1})_j = \delta_{i,j}\). Using this
-    fact we see that the Poisson brackets between \(P\) and \(Q\) are given by:
+    \((\mathbf{B}_i)\tpose (\mathbf{B}\tpose)_j = (\mathbf{B}_i)\tpose (\mathbf{B}^{-1})_j = \delta_{i,j}\). Using this
+    fact we see that the Poisson brackets between \(P_i\) and \(Q_j\) are given by:
     \begin{align*}
@@
 
-Most children have an adult pushing the swing to help them go higher; or perhaps
-the child is swinging on a windy day. In either case, they adjust their leg
-motion accordingly and do not keep track of time when kicking their legs. Te
+Most children have an adult pushing the swing to help them go higher, or perhaps
+they are swinging on a windy day. In either case, they adjust their leg
+motion accordingly when presented with these external disturbances, without
+keeping track of the time it takes to kick their legs. The
 standard control technique of tracking a function of time, known as
-\textit{trajectory tracking}, would not work because the disturbance affecting
-the swing will desynchronize the motion of the legs with the swing and stop the
-energy-gaining effects.
+\textit{trajectory tracking}, does not work as well as human motion 
+because the disturbance affecting the swing will desynchronize the leg motion
+and the swing, thereby stopping the energy-gaining effects.
 
@@
 Unlike the theory of VHCs, there does not appear to be a standard definition of virtual
-nonholonomic constraints. One issue is that Lagrangian dynamics is not
-particularly suited to finding controllers which stabilize desired constraints. Because
-of this, all the applications listed above use their own definition of a virtual
+nonholonomic constraints: 
+all the applications listed above use their own definition of a virtual
 nonholonomic constraint, which makes it difficult to compare and analyze their
@@
 This section will provide a new characterization of virtual nonholonomic
-constraints. The goal is to provide a consistent, rigorous foundation for
+constraints in the Hamiltonian framework. 
+The goal is to provide a consistent, rigorous foundation for
 designing constraints on a general class of systems.
@@
 Let us imagine that we have no further structure on the VNHC. 
-Then \(\dot{e} = dh_q \dot{q} + dh_p \dot{p}\), so \(\tau\) appears inside
-\(\dot{e}\). To solve for \(\tau\) explicitly, we must have that \(dh_p\) is
-invertible for all \(p\), which is the same as requiring that \(h(q,p)\) is
-strictly monotonic in \(p\).
+Then \(\dot{e} = dh_q \dot{q} + dh_p \dot{p}\), where \(\tau\) appears inside
+\(\dot{p}\). To solve for \(\tau\) explicitly, we must have that \(dh_p\) is
+invertible for all \((q,p)\).
+%which is the same as requiring that \(h(q,p)\) is strictly monotonic in \(p\).
 
@@
 one can use.
-Mathematically, if \(\tau\) appears only after two derivatives, one says 
+Mathematically, if each \(\tau_i\) appears only after two derivatives, one says 
 that \(e\) is of \textit{relative degree} \(\{2,2,\ldots,2\}\). 
@@
 \begin{defn}
-    A VNHC of order \(k\) \(h(q,p) = 0\) is \textit{regular} if the output 
+    A VNHC \(h(q,p) = 0\) of order \(k\) is \textit{regular} if the output 
     \(e = h(q,p)\) is of relative degree \(\{2,2.\ldots,2\}\) everywhere on the
@@
 
-Modifying the results of
+The authors of
 \cite{nhvc_dynamic_walking,hybrid_zero_dynamics_bipedal_nhvcs,nhvc_incline_walking}
-into the Hamiltonian framework, one observes that
-relations which use only the unactuated momentum \(p_u\) result in \(\tau\)
-appearing only after two derivatives of \(e\).
-To be able to use \(p_u\), we will continue with the following assumption for
-the rest of the chapter.
+observed that a relation which uses only the unactuated conjugate of momentum
+\(p_u\) cannot have \(\tau\) appearing after only one derivative. Of course,
+they performed their research in Lagrangian form; 
+for us to be able to use \(p_u\) in its Hamiltonian form, 
+we will continue with the following assumption for the rest of the chapter.
 
@@
     We also write
-    \(p_u \in \mathcal{P}_u = \R^{n-k}\) and 
-    \(p_a \in \mathcal{P}_a = \R^k\), so that 
-    \(p \in \mathcal{P} := \mathcal{P}_u \times \mathcal{P}_a\). 
+    \(p_u \in \mathcal{P}_u := \R^{n-k}\) and 
+    \(p_a \in \mathcal{P}_a := \R^k\), so that 
+    \(p \in \mathcal{P} := \mathcal{P}_u \times \mathcal{P}_a = \R^n\). 
     In this manner, the phase space of our system can be written as
@@
     \(\mathcal{Q} \times \mathcal{P}_u\) which satisfies the rank condition.
-    condition.
 \end{thm}
@@
     \(\{2,2,\ldots,2\}\).
-    Hence, we must have that \(dh_{p_a} = \Zmat{k \times k}\) if we want the
-    VNHC to be regular. Proceeding with this assumption, we now find that
-    \(h : \mathcal{Q} \times \mathcal{P}_u \rightarrow \R^k\), which means
+    Hence, we must have that \(dh_{p_a} = \Zmat{k \times k}\).
+    Proceeding with this assumption, we now find that
+    \(h : \mathcal{Q} \times \mathcal{P}_u \rightarrow \R^k\), which means that
     \[
@@
     \end{align*}
-    We will compute the explicit expression of \(\ddot{e}\) in pieces. 
+    We will compute the explicit expression of \(\ddot{e}\) in pieces to find
+    out where \(\tau\) appears. 
 % d/dt(dh_q)
-    First we begin with \(\diff{}{t}\left\{dh_q\right\}\). 
-    Since \(h = (h^1,\ldots,h^k)\) where 
-    \(h^i : \mathcal{Q} \times \mathcal{P}_u \rightarrow \R\), we have that 
+    We begin with \(\diff{}{t}\left\{dh_q\right\}\). 
+    Let \(h = (h^1,\ldots,h^k)\) with
+    \(h^i : \mathcal{Q} \times \mathcal{P}_u \rightarrow \R\). By definition of
+    the total differential, we have 
     \[
@@
     \]
-    The derivative is taken element-wise in each matrix, yielding
+    The time derivative is taken element-wise, yielding
     \[
@@
 % d/dt(dh_pu)
-    Moving on, one can use a similar approach to \(\diff{}{t}\left\{dh_q\right\}\)
-    to find the derivative of \(dh_{p_u}\) in matrix form. It is given by
+    Moving on to \(\diff{dh_{p_u}}{t}\), a similar approach as
+    \(\diff{}{t}\left\{dh_q\right\}\) yields the derivative in matrix form.
+    It is given by:
     \begin{align*}
@@
     Now observe that the \(i^\text{th}\) row of
-    \[
-        \frac{1}{2} \diff{}{t} \left\{ \pudmat \right\}
-    \]
+    \(\frac{1}{2} \diff{}{t} \left\{ \pudmat \right\}\)
     is given by
@@
 Note that one generally cannot measure conjugate of momenta directly, as sensors
-on mechanical systems typically only measure the state \((q,\dot{q})\). To
+on mechanical systems will only measure the state \((q,\dot{q})\). To
 implement this controller in practice, one must compute \(p = M(q)\dot{q}\) at
@@
 \begin{defn}
-    A mechanical system is ------ if \(\nabla_{q_u}M(q) = 0\)
+    A mechanical system is \textbf{TODO: name this} ------ if 
+    \(\nabla_{q_u}M(q) = \Zmat{n(n-k) \times n}\).
 \end{defn}
@@
     with constraint manifold \(\Gamma\). Suppose that on \(\Gamma\) one can
-    solve linearly for \(q_a\) as a function of \(q_u,p_u\).
+    solve linearly for \(q_a\) as a function of \((q_u,p_u)\).
     Then the closed-loop dynamics are given by
@@
     \[
-        p_a = \left(dh_q\Minv(q)\simpleB\right)\inv
+        p_a(q,p_u) = \left(dh_q\Minv(q)\simpleB\right)\inv
         \left(dh_{p_u}\nabla_{q_u}V(q) - 
@@
     This yields a function \(p_a(q,p_u)\). However, on \(\Gamma\) we have 
-    \(h(q,p_u) = 0\) and can solve for \(q_a\) in terms of \((q_u,p_u)\).
-    Hence, on \(\Gamma\) we have \(p_a = g(q_u,p_u)\). Since \(q_a\) and
+    \(h(q,p_u) = 0\) and can solve for \(q_a\) in terms of \((q_u,p_u)\);
+    hence, we can solve for \(p_a = g(q_u,p_u)\). Since \(q_a\) and
     \(p_a\) can be computed directly from \((q_u,p_u)\), the dynamics on
     \(\Gamma\) are parameterized only by \((\dot{q}_u,\dot{p}_u)\).
+
+    \textbf{MANFREDI: Do we need this assumption that \(q_a\) can be solved in
+        terms of \((q_u,p_u)\) or is there some other way of parameterizing it
+    like you did with \(s\) for VHCs?}
 \end{proof}
@@
 unactuated coordinates. 
-This is trueregardless of the number of degrees of freedom of the system. 
-This means that \(\Gamma\) is a \(2(n-k)\)-dimensional surface inside 
+This is true regardless of the number of degrees of freedom of the system, which
+means that \(\Gamma\) is a \(2(n-k)\)-dimensional surface inside 
 \(\mathcal{Q}\times \mathcal{P}\).
@@
     where \(f\) is a suitably-defined smooth function.
-    Defining \(e_1 := (1,0,\ldots,0) \in \R^n\), the actuated momentum 
-    is given by
+    Then \(dh_q = \begin{bmatrix} -\partial f_{q_u} & \Id{(n-1)}
+    \end{bmatrix}\).
+    Defining \(e_1 := (1,0,\ldots,0) \in \R^n\), the actuated momentum is given
+    by (\ref{eqn:g-qupu}):
     \begin{equation}\label{eqn:g-qupu}
@@
 %TODO: summarize the assumptions and results 
+\textbf{TODO: summarize the assumptions and results }
 

### 2020-08-24 latex/vlp.tex
Small update to begin describing the VLP model

@@
 % which is inherently a VNHC.
+We return once again to the example of a person standing on a swing.
+Condensing them into a center of mass gives a VLP model of their motion, where
+the pendulum shrinks when the person stands and lengthens when they squat
+(as in Figure \ref{fig:child-vlp}).
+\begin{figure}
+   \centering
+   \begin{subfigure}[t]{0.5\textwidth}
+      % TODO: Insert child standing figure + caption
+   \end{subfigure}
+   \begin{subfigure}[t]{0.5\textwidth}
+      % TODO: Insert child squatting figure + captions
+   \end{subfigure}
+   \caption{The variable length pendulum representation of a child on a
+   swing}\label{fig:child-vlp}
+\end{figure}
 

### 2020-08-25 latex/vlp.tex
Wrote the bulk of the VLP VNHC section
Missing only the figures, the energy dissipation, and a brief summary
before moving on to the simulation results

--- 2020-08-24 latex/vlp.tex
@@
 \(q \in \mathbb{S}^1\) from the vertical, as is seen in Figure
-\ref{fig:vlp_model}. 
+\ref{fig:vlp-model}. 
 We will also ignore any damping and frictional forces in this model.
@@
    \caption{The representation of the variable length pendulum as a mass at the
-      tip of a massless rod.}\label{fig:vlp_model}
+      tip of a massless rod.}\label{fig:vlp-model}
 \end{figure}
@@
 % which is inherently a VNHC.
+Since we wish to define a VNHC which injects energy into the VLP in a human-like
+manner,
+we return once again to the example of a person standing on a swing.
+As can be seen in Figure \ref{fig:child-vlp}, condensing the person 
+into a center of mass gives a VLP model of their motion, where
+the pendulum shrinks when the person stands and lengthens when they squat. This
+is equivalent to the VLP model from Figure \ref{fig:vlp-model}.
+\begin{figure}
+   \centering
+   \begin{subfigure}[t]{0.5\textwidth}
+      % TODO: Insert child standing figure + caption
+   \end{subfigure}
+   \begin{subfigure}[t]{0.5\textwidth}
+      % TODO: Insert child squatting figure + captions
+   \end{subfigure}
+   \caption{The variable length pendulum representation of a child on a
+   swing}\label{fig:child-vlp}
+\end{figure}
 
-\begin{thm}
-For the variable-length pendulum, define \(\theta := \arctan_2(p,q)\). 
-A VNHC of the form \(l = l(\theta)\) injects energy if there exists \(l_{avg} \in \R_{>0}\) such that
+The regulation of pendulum length to add energy to the VLP is known as
+``pumping"`. \citeauthor{pumping_swing_standing_squatting} asked whether
+the pumping strategy performed by children is time-optimal, assuming the
+children could squat or stand instantaneously
+\cite{pumping_swing_standing_squatting}. Indeed, they discovered that 
+a child will pump energy into the VLP as fast as is physically possible.
+
+The pumping strategy performed by children is quite 
+straightforward: they stand up at the lowest point of their swing, and squat at
+the highest point. Looking at their VLP representation, the pendulum shortens
+at the bottom of the swing, and lengthens at the top. 
+This pumping strategy makes intuitive sense:
+by shortening the pendulum at the bottom, conservation of angular momentum indicates the
+mass must gain speed to compensate for the decrease in pendulum length
+\cite{how_to_pump_a_swing}.
+Energy is not conserved in this process, and the pendulum gains kinetic energy
+which allows the pendulum to reach a higher point in its swing.
+Lengthening the pendulum when it reaches this point means gravity
+imparts a larger angular momentum to the mass by the time it reaches the bottom
+of its swing.
+By alternating these processes, the pendulum experiences a net gain in
+rotational energy which is generated entirely by gravity.
+
+The time-optimal controller described by \cite{pumping_swing_standing_squatting}
+depends on a state trajectory in time. We will convert this time-based
+controller into a time-independent controller which still uses this time-optimal
+strategy. Notice that the control strategy depends on knowing when the
+system is at the ``bottom" or ``top" of the swing. Since being at the ``bottom"
+is the same as saying the pendulum angle is \(q = 0\) and being at the ``top" is
+equivalent to the pendulum's momentum being \(p = 0\), our method of
+VNHCs is especially well suited to this problem.
+
+Figure \ref{fig:vlp-optimal-controller-qp-plane} shows the time-optimal strategy directly
+on the \((q,p)\)-plane. The axes of this plane define where the pendulum should
+change length. Since the pendulum stays the same length inside each quadrant, we
+will use the angle \(\theta = \arctan_2(p,q)\) in the \((q,p)\)-plane to define
+the optimal controller \(l^\star(\theta)\) (\ref{eqn:vlp-optimal-controller}).
+Figure \ref{fig:vlp-optimal-controller} shows this controller as a function of
+\(\theta\).
+\begin{equation}\label{eqn:vlp-optimal-controller}
+   l^\star(\theta):= \begin{cases}
+      \underbar{l} & \theta \in [-pi, -\frac{\pi}{2}[ \cup [0,\frac{\pi}{2}[ \\
+      \bar{l} & \theta \in [-\frac{\pi}{2},0[ \cup [\frac{\pi}{2}, \pi[ \\
+   \end{cases}
+\end{equation}
+
+\begin{figure}
+   \centering
+   % TODO: Inser the image of the optimal controller in the (q,p) plane
+   \caption{The time-optimal VLP controller from
+   \cite{pumping_swing_standing_squatting} displayed in the
+\((q,p)\)-plane.}\label{fig:vlp-optimal-controller-qp-plane}
+\end{figure}
+
+\begin{figure}
+   \centering
+   %TODO: Insert the image of the optimal controller as a function of theta
+   \caption{The time-optimal VLP controller converted to a discontinuous
+   function \(l(\theta)\).}\label{fig:vlp-optimal-controller}
+\end{figure}
+
+To find a continuously differentiable function which approximates this controller,
+we can attach sinusoids at the transition points of
+\(l^\star(\theta)\) (see Figure \ref{fig:vlp-T-controller}). Let
+\(\Delta l := (\bar{l} - \underbar{l})/2\) and 
+\(l_{\text{avg}} := (\bar{l} + \underbar{l})/2\).
+Providing these newly attached sinusoids with a frequency \(\omega = \frac{pi}{T}\) 
+(where \(T \in ]0,\frac{\pi}{2}]\) is a control parameter), we have a new
+controller \(l_T(\theta)\):
+\begin{equation}\label{eqn:vlp-T-controller}
+   l_T(\theta) = \begin{cases}
+      \underbar{l} & \theta \in \left[-\pi + \frac{T}{2}, -\frac{\pi}{2} - \frac{T}{2}\right] 
+      \cup \left[\frac{T}{2}, \frac{\pi}{2} - \frac{T}{2}\right] \\
+      \bar{l} & \theta \in \left[-\frac{\pi}{2} + \frac{T}{2}, -\frac{T}{2}\right] 
+      \cup \left[\frac{\pi}{2} + \frac{T}{2}, \pi - \frac{T}{2}\right] \\
+      -\Delta l \sin(\omega(\theta + \pi)) + l_{\text{avg}} & \theta \in
+      \left[-\pi,-\pi + \frac{T}{2}\right] \\
+      -\Delta l \sin(\omega(\theta - a)) + l_\text{avg} & 
+      \theta \in \left[a - \frac{T}{2}, a + \frac{T}{2}\right] \text{ for } 
+      a \in \left\{-\frac{\pi}{2}, 0, \frac{\pi}{2}\right\} \\
+      -\Delta l \sin(\omega(\theta-\pi)) & \theta \in \left[\pi - \frac{T}{2},\pi\right] \\
+   \end{cases}
+\end{equation}
+
+\begin{figure}
+   \centering
+   % TODO: Insert an image of the controller with sinusoids attached
+   \caption{}\label{fig:vlp-T-controller}
+\end{figure}
+
+This controller approximates the optimal controller, since 
 \[
-   \left(l(\theta) - l_{avg}\right)sin(2\theta) \leq 0 \text{ }\forall \theta \in \mathbb{S}^1
+   \lim\limits_{T \rightarrow 0} l_T(\theta) = l^\star(\theta)
+\]
+Unfortunately, while \(l_T(\theta)\) is continuously differentiable, it is not
+twice continuously differentiable for most values of \(T\).
+If we wish to use it as a VNHC, we must ensure that either the generalized
+forces \(\tau\) acting on \(p_l\) can be discontinuous, or we must find a value
+of \(T\) where this controller is smooth.
+Thankfully, \(l_{\frac{\pi}{2}}(\theta)\) is smooth; it can be simplified into
+the expression (\ref{eqn:vlp-smoothed-controller}), and is plotted for
+demonstration in Figure \ref{fig:vlp-smoothed-controller}.
+\begin{equation}\label{eqn:vlp-smoothed-controller}
+   l_\frac{\pi}{2}(\theta) = -\Delta l \sin(2\theta) + l_{\text{avg}}
+\end{equation}
+
+\begin{figure}
+   \centering
+   % TODO: Insert the image of the smoothed VLP controller
+   \caption{The smoothed VLP controller attained by setting the parameter 
+      \(T\) to its upper limit
+   \(\frac{\pi}{2}\).}\label{fig:vlp-smoothed-controller}
+\end{figure}
+
+We therefore set our VNHC to be
+\[
+   h(q,p) = l - l_\frac{\pi}{2}\left(\theta(q,p)\right)
+\]
+
+This VNHC is not the same as the optimal controller
+(\ref{eqn:vlp-optimal-controller}), so we need to prove it will still gain
+energy. We will do so by showing the dynamics (\ref{eqn:vlp-hamiltonian}) of the
+VLP diverge from the origin, regardless of the initial condition of the system. 
+To do this, we will use the notion of an anti-Lyapunov function.
+
+\begin{defn}\label{defn:anti-lyapunov}
+   Let \(\dot{x} = f(x) \in R^n\) be an ODE with \(f(0) = 0\). 
+   A positive definite function \(V(x)\) is \textbf{anti-Lyapunov} if 
+   \[
+      \dot{V}(x) = dV(x) f(x) \geq 0
+   \]
+   where the inequality is strict for almost every \(x \in \R^n\).
+\end{defn}
+
+By applying Lyapunov stability theory \cite{lyapunov},
+we see that an anti-Lyapunov function
+\(V\) is simply a Lyapunov function for the negative-time system 
+\(\dot{x} = -f(x)\). In other words, since \(V\) proves that the orbit \(x(-t)\) 
+of the system travelling backwards in time is asymptotically converging to the
+equilibrium, the system must be diverging from the equilibrium in forward time.
+If we can find such a function for the VLP's dynamics under our VNHC, the path
+generated by the VNHC in the \((q,p)\)-plane must be diverging from the origin
+so long as the initial condition is not an equilibrium.
+This in turn tells us the magnitude of the momentum \(p\) is increasing every
+time the path hits the \(p\)-axis, which means the VLP is, on average,
+increasing in energy.
+
+We will show that anti-Lyapunov function exists for our VNHC.
+In order to do so however, we need to prove the following lemma.
+
+\begin{lemma}\label{lemma:sign-of-cube}
+   For any \(x,y \in \R\):
+   \[
+      \sign{x^3 - y^3} = \sign{x-y}
+   \]
+\end{lemma}
+\begin{proof}
+   Observe first that
+   \[
+      x^3 - y^3 =  (x-y)(x^2 + xy + y^2)
+   \]
+   Since \( (x-y)^2 = x^2 - 2xy + y^2) \geq 0 \), 
+   \[
+      \frac{x^2 + y^2}{2} \geq xy
+   \]
+   which means \(x^2 + xy + y^2 \geq 0\), proving the lemma.
+\end{proof}
+
+\begin{thm}\label{thm:vlp-gains-energy}
+   Take the VLP with Hamiltonian dyanmics (\ref{eqn:vlp-hamiltonian}) and define
+   \(\theta := \arctan_2(p,q)\). Suppose the initial condition is not one of the
+   equilibria: \((q(0),p(0)) \not \in \left\{(0,0), (\pi,0)\right\}\).
+   A VNHC of the form \(h(q,p) = l - l(\theta)\) injects energy if there exists 
+   \(l_\text{avg} \in [\underbar{l},\bar{l}]\) such that 
+   \[
+      \left(l(\theta) - l_\text{avg}\right)sin(2\theta) \leq 0 \text{ }\forall \theta \in \mathbb{S}^1
    \]
@@
     \[
-       E_{avg}(q,p) := \frac{1}{2}\frac{p^2}{m l_{avg}^2} 
-                    + m g l_{avg} (1-\cos(q))
+       E_\text{avg}(q,p) := \frac{1}{2}\frac{p^2}{m l_\text{avg}^2} 
+                    + m g l_\text{avg} (1-\cos(q))
     \]
-
-    which is non-negative and has derivative 
+    which is positive definite and has derivative 
     \[
-      \dot{E}_{avg} = \frac{-g\sin(q)p \left(l(\theta)^3 - l_{avg}^3\right)}
-                 {l_{avg}^2l(\theta)^2}
+      \dot{E}_\text{avg} = \frac{-g\sin(q)p \left(l(\theta)^3 - l_\text{avg}^3\right)}
+                 {l_\text{avg}^2l(\theta)^2}
     \]
-
-    We will show that \(E_{avg}\) is increasing.
-
-    % VBOX UNDERFULL WARNING HERE, IGNORE IT
     Observe that \(\sign{\sin(q)p} = \sign{\sin(2\theta)}\) and, 
-    by Lemma \textbf{TODO: REF LEMMA},
-    \( \sign{l(\theta)^3 - l_{avg}^3} 
-     = \sign{ l(\theta) - l_{avg}}\).
-
-     Then the derivative of \(E_{avg}\) is almost always positive, since
+    by Lemma \ref{lemma:sign-of-cube},
+    \[ 
+       \sign{l(\theta)^3 - l_\text{avg}^3} = \sign{ l(\theta) - l_\text{avg}}
+    \]
+    Then the derivative of \(E_\text{avg}\) is almost always positive, since
     \begin{align*}
        \label{eq:}
-        \sign{\dot{E}_{avg}} &= \sign{-\sin(q)p \left(l(\theta)^3 - l_{avg}^3\right)} \\
-                    &= -\sign{\sin(2\theta) \left(l(\theta) - l_{avg}\right)} \\
+       \sign{\dot{E}_\text{avg}} &= \sign{-\sin(q)p \left(l(\theta)^3 - l_\text{avg}^3\right)} \\
+                   &= -\sign{\sin(2\theta) \left(l(\theta) - l_\text{avg}\right)} \\
                    &\geq 0 \text{ (by assumption)}
     \end{align*}
-
-      Hence, \(E_{avg}\) is an anti-Lyapunov function with positive derivative,  
-      so the variable-length pendulum is gaining energy.
+    Hence, \(E_\text{avg}\) is an anti-Lyapunov function
+    Since the initial condition is not an equilibrium,
+    the orbit of the VNHC in the \((q,p)\) plane is diverging away from the
+    origin and the VLP is gaining energy on average.
 \end{proof}
 
+\begin{cor}
+   The VNHC \(h(q,p) = l - l_\frac{\pi}{2}(\theta)\) injects energy into the VLP
+   over time, since 
+   \(\left(l_\frac{\pi}{2}(\theta) - l_\text{avg}\right)\sin(2\theta) = 
+   -\Delta l \sin(2\theta) < 0\) for almost every \(\theta\).
+\end{cor}
+
+% TODO: Describe energy reduction, which comes from switching the sign
+
+Theorem \ref{thm:vlp-gains-energy} provides an alternate explanation for why the
+optimal controller \(l^\star(\theta)\) works so well at injecting energy: it
+maximizes the derivative of \(E_\text{avg}\) under the restriction 
+\(l(\theta) \in [\underbar{l},\bar{l}]\), so that the orbit in the
+\((q,p)\)-plane diverges from the origin as fast as possible. The differentiable
+controllers \(l_T(\theta)\) will also inject energy, albeit at a rate which
+decreases as \(T\) increases. That said, Theorem \ref{thm:vlp-gains-energy} and the
+subsequent Corollary shows that VNHCs are a powerful tool in creating effective energy
+pumping strategies based on natural human motion.
+
+% TODO: How would we stabilize a specific energy level?
+
 \section{Simulation Results}
+% TODO: Show the simulations of the VLP VNHCs (multiple of them). How does the
+% energy injection time of our sin(2theta) VNHC compare to the energy injection
+% time of the optimal one? How does the VNHC version compare in robustness with
+% the time-based one from the pumping paper?
 

### 2020-08-25 latex/vnhcs.tex
Wrote the bulk of the VLP VNHC section
Missing only the figures, the energy dissipation, and a brief summary
before moving on to the simulation results

@@
 
+% TODO: Get some sources here. Show with research that trajectory control is
+% unsafe, and that path tracking is better.
 Most children have an adult pushing the swing to help them go higher, or perhaps
@@
 and the swing, thereby stopping the energy-gaining effects. 
-
-This is a common issue in the control of many biologically-inspired systems. 
-One method which can provide more realistic, robust motion is the method of
-of virtual holonomic constraints (VHCs). Instead of a robot's actuators tracking
+As has been observed in the field of mobile robotics,
+a trajectory tracking controller which has lagged behind due to a disturbance
+will do whatever is required to ``catch up" to
+the trajectory. This can cause major safety issues, which is why mobile
+roboticists prefer to design controllers which drive a vehicle towards a path in
+space rather than a trajectory in time.
+
+This path tracking approach has been extended to biologically-inspired robotics
+in a method known as virtual holonomic constraints (VHCs).
+Instead of a robot's actuators tracking
 a trajectory over time, VHCs use the actuators to enforce a relation
@@
 
-Many authors have attempted to extend these results to enforce a relation 
-\(h(q,\dot{q}) = 0\) of the full state. Since these relations use actuators to
+The downside to VHCs is that they do not use the additional information imparted
+by the generalized velocities of a system.
+Many authors have attempted to extend VHCs to enforce these relations 
+\(h(q,\dot{q}) = 0\) of the full state to account for this drawback. 
+Since these relations use actuators to
 restrict both the configuration and generalized velocity, they are called
@@
 error-reduction on time-delayed systems \cite{vnhc_time_delay_teleop},
-and improving bipedal locomotion \cite{nhvc_dynamic_walking,
+and has shown marked improvements to the field of bipedal locomotion 
+\cite{nhvc_dynamic_walking,
 hybrid_zero_dynamics_bipedal_nhvcs,output_nhvc_bipedal_control}.
-Most interestingly, this nonholonomic approach has been shown to be more robust
+Most interestingly, the nonholonomic approach has been shown to be more robust
 than standard VHCs when applied to bipedal robotics \cite{nhvc_incline_walking}.
@@
 
-Unlike the theory of VHCs, there does not appear to be a standard definition of virtual
-nonholonomic constraints: 
+Unlike the theory of VHCs, there does not appear to be a standard definition of
+virtual nonholonomic constraints: 
 all the applications listed above use their own definition of a virtual

### 2020-08-25 latex/vlp.tex
Fixed a typo

@@
     \end{align*}
-    Hence, \(E_\text{avg}\) is an anti-Lyapunov function
+    Hence, \(E_\text{avg}\) is an anti-Lyapunov function.
     Since the initial condition is not an equilibrium,

### 2020-08-28 latex/vlp.tex
Wrote the energy dissipation and stabilization results
Finished the bulk of th eVLP section, need to just add figures now and
then start on simulation.

@@
 We will show that anti-Lyapunov function exists for our VNHC.
-In order to do so however, we need to prove the following lemma.
+In order to do so, we require the following lemma.
 
@@
 
-\begin{thm}\label{thm:vlp-gains-energy}
+\begin{thm}\label{thm:vlp-energy-stabilization}
    Take the VLP with Hamiltonian dyanmics (\ref{eqn:vlp-hamiltonian}) and define
    \(\theta := \arctan_2(p,q)\). Suppose the initial condition is not one of the
-   equilibria: \((q(0),p(0)) \not \in \left\{(0,0), (\pi,0)\right\}\).
+   equilibria: \((q(0),p(0)) \not \in \left\{(0,0), (\pm\pi,0)\right\}\).
    A VNHC of the form \(h(q,p) = l - l(\theta)\) injects energy if there exists 
    \(l_\text{avg} \in [\underbar{l},\bar{l}]\) such that 
-   \[
+   \begin{equation}\label{eqn:vlp-energy-gain-condition}
       \left(l(\theta) - l_\text{avg}\right)sin(2\theta) \leq 0 \text{ }\forall \theta \in \mathbb{S}^1
-   \]
+   \end{equation}
    with the property that the inequality is strict for almost every \(\theta\).
+   Flipping the inequality of (\ref{eqn:vlp-energy-gain-condition}) leads to
+   energy dissipation.
 \end{thm}
@@
     \]
-    which is positive definite and has derivative 
+    which is positive definite at \((0,0)\) and has derivative 
     \[
@@
     origin and the VLP is gaining energy on average.
+    Flipping the inequality of (\ref{eqn:vlp-energy-gain-condition}) 
+    means \(E_\text{avg}\) is a regular Lyapunov function, since its derivative
+    will be negative almost everywhere.
 \end{proof}
@@
 \begin{cor}
-   The VNHC \(h(q,p) = l - l_\frac{\pi}{2}(\theta)\) injects energy into the VLP
+   The VNHC \(l = l_\frac{\pi}{2}(\theta)\) injects energy into the VLP
    over time, since 
-   \(\left(l_\frac{\pi}{2}(\theta) - l_\text{avg}\right)\sin(2\theta) = 
-   -\Delta l \sin(2\theta) < 0\) for almost every \(\theta\).
+   \[
+      \left(l_\frac{\pi}{2}(\theta) - l_\text{avg}\right)\sin(2\theta) = 
+   -\Delta l \sin^2(2\theta) \geq 0
+   \] 
+   Likewise, let 
+   \(l\inv_\frac{\pi}{2}(\theta) := l_\text{avg} - \Delta l \sin(2\theta)\).
+    The VNHC \(h(q,p) = l - l\inv_\frac{\pi}{2}(\theta)\) which flips 
+    \(l_\frac{\pi}{2}(\theta)\) about \(l_\text{avg}\) will dissipate energy from
+   the VLP, since 
+   \[
+      \left(l\inv_\frac{\pi}{2}(\theta) - l_\text{avg}\right)\sin(2\theta) = 
+      -\Delta l \sin^2(2\theta) \leq 0
+   \] 
 \end{cor}
 
-% TODO: Describe energy reduction, which comes from switching the sign
+The types of VNHCs which satisfy Theorem \ref{thm:vlp-energy-stabilization} are illustrated by Figure
+\ref{fig:vlp-energy-in-out}. To stabilize specific energy level sets, one simple
+approach is to switch between injection and dissipation VNHCs depending on the
+current energy level.
+For the VNHCs we designed in this section, this means toggling
+between \(l_\frac{\pi}{2}(\theta)\) and \(l\inv_\frac{\pi}{2}(\theta)\),
+possibly with some hysteresis to avoid infinite switching.
 
-Theorem \ref{thm:vlp-gains-energy} provides an alternate explanation for why the
-optimal controller \(l^\star(\theta)\) works so well at injecting energy: it
-maximizes the derivative of \(E_\text{avg}\) under the restriction 
-\(l(\theta) \in [\underbar{l},\bar{l}]\), so that the orbit in the
-\((q,p)\)-plane diverges from the origin as fast as possible. The differentiable
-controllers \(l_T(\theta)\) will also inject energy, albeit at a rate which
-decreases as \(T\) increases. That said, Theorem \ref{thm:vlp-gains-energy} and the
-subsequent Corollary shows that VNHCs are a powerful tool in creating effective energy
-pumping strategies based on natural human motion.
+\begin{figure}
+   \centering
+   % TODO: Add a figure showing regions for injection vs dissipation
+   \caption{}\label{fig:vlp-energy-in-out}
+\end{figure}
 
-% TODO: How would we stabilize a specific energy level?
+Theorem \ref{thm:vlp-energy-stabilization} also provides an alternate
+explanation for why the ``optimal" VNHC \(l^\star(\theta)\) works so well at
+injecting energy: it maximizes the derivative of \(E_\text{avg}\) under the
+restriction \(l(\theta) \in [\underbar{l},\bar{l}]\), so that the orbit in the
+\((q,p)\)-plane diverges from the origin as fast as possible.
+
+Let us define \((l^\star)\inv(\theta)\) by swapping the order of 
+\(\underbar{l}\) and \(\bar{l}\) in \(l^\star(\theta)\). Since this 
+\textit{minimizes} the derivative of \(E_\text{avg}\) under the restriction
+\(l(\theta) \in [\underbar{l},\bar{l}]\), we can predict that 
+\((l^\star)\inv(\theta)\) is the VNHC representation of an optimal 
+energy dissipation controller.
+This is, in fact, true: \cite{pumping_swing_standing_squatting} showed 
+that squatting at the lowest point of a swing and standing at the highest
+point (instead of standing and squatting resp.) produces the
+time-optimal trajectory for stopping a standing swing.
+
+All together, these results show that VNHCs can replicate the time-optimal
+strategy performed by humans in a time-independent manner. Furthermore, we see
+that VNHCs are a powerful tool for creating simple energy stabilization
+techniques based on natural human motion.
 

### 2020-09-03 latex/vlp.tex
Updated the VLP text and the underline/overline on l

@@
 tuned-mass-damper systems which stabilize skyscrapers
-\cite{vlp_tuned_mass_damper}, among others.
+\cite{vlp_tuned_mass_damper}.
 
@@
 \cite{dynamics_periodic_vlp}), and many control mechanisms exist
-to stabilize trajectories of the system. While many of these control mechanism
+to stabilize trajectories of the system. While many of these controllers 
 are time-dependent, \cite{vlp_energy_shaping}
@@
 ad-hoc in the sense that it is not derived from natural behaviour. In
-this chapter we will use VNHCs to inject energy into the VLP in a
-time-independent way;
-the difference is that our control mechanism will maintain the structured motion
-of a human on a swing.
+this chapter we will design VNHCs to inject energy into the VLP in a
+time-independent way.
+The difference compared to \cite{vlp_energy_shaping} is that our 
+controller will maintain the structured motion of a human on a swing.
 
@@
    \centering
-   \includestandalone[width=0.4\textwidth]{images/vlp_model}
+   \includestandalone[width=0.5\textwidth]{images/vlp_model}
    \caption{The representation of the variable length pendulum as a mass at the
@@
 that is, the control input cannot enact any lateral force on the pendulum.
-In this way, the dynamics of \((q,p)\) and \((l,p_l)\) are decoupled.
-
-This decoupling is extremely useful for simplifying the dynamics. Since one can
-use \(\tau\) to set \(\dot{p}_l\) to any desired function, we can assume (with abuse
-of notation) that \(l\) is tracking some function \(l(t)\). 
+This is extremely useful for simplifying the dynamics: since one can choose
+\(\tau\) which sets \(\dot{p}_l\) to any desired function, we can assume (with abuse
+of notation) that \(l\) is tracking some known function \(l(t)\). 
 The closed loop dynamics of the system will be described exclusively by
-\((\dot{q},\dot{p})\) with \(l = l(t)\); 
-the fact that \(p_l\) does not appear in these closed-loop dynamics allows us to
+\((\dot{q},\dot{p})\) with \(l \equiv l(t)\).
+The fact that \(p_l\) does not appear in these closed-loop dynamics allows us to
 ignore the subdynamics \((\dot{l},\dot{p}_l)\) entirely.
@@
 modelling it as a configuration variable. Re-deriving the Hamiltonian and the
-dynamics with this in mind, we get the system 
+dynamics from this approach, we get the system 
 (\ref{eqn:vlp-hamiltonian}) with phase \((q,p) \in \mathbb{S}^1 \times \R\). 
@@
    \centering
-   \begin{subfigure}[t]{0.5\textwidth}
-      % TODO: Insert child standing figure + caption
+   \begin{subfigure}[t]{0.45\textwidth}
+      \includegraphics[]{images/child_vlp_standing.png}
+      \caption{A person standing on a swing has their center of mass 
+      close to the pivot.}
    \end{subfigure}
-   \begin{subfigure}[t]{0.5\textwidth}
-      % TODO: Insert child squatting figure + captions
+   \hfill
+   \begin{subfigure}[t]{0.45\textwidth}
+      \includegraphics[]{images/child_vlp_squatting.png}
+      \caption{When a person squats on a swing, their center of mass extends
+      away from the pivot.}
    \end{subfigure}
-   \caption{The variable length pendulum representation of a child on a
-   swing}\label{fig:child-vlp}
+   \caption{The variable length pendulum representation of a person on a
+   standing swing.}\label{fig:child-vlp}
 \end{figure}
 
-The regulation of pendulum length to add energy to the VLP is known as
-``pumping"`. \citeauthor{pumping_swing_standing_squatting} asked whether
+The action of regulating pendulum length to add energy to the VLP is known as
+``pumping". \citeauthor{pumping_swing_standing_squatting} asked whether
 the pumping strategy performed by children is time-optimal, assuming the
@@
 \cite{pumping_swing_standing_squatting}. Indeed, they discovered that 
-a child will pump energy into the VLP as fast as is physically possible.
+a child's pumping strategy will inject energy into the VLP as fast as is 
+physically possible.
 
-The pumping strategy performed by children is quite 
-straightforward: they stand up at the lowest point of their swing, and squat at
-the highest point. Looking at their VLP representation, the pendulum shortens
-at the bottom of the swing, and lengthens at the top. 
-This pumping strategy makes intuitive sense:
-by shortening the pendulum at the bottom, conservation of angular momentum indicates the
-mass must gain speed to compensate for the decrease in pendulum length
+This pumping strategy is straightforward: the child stands up at the lowest point of
+their swing, and squats at the highest point. Looking at their VLP
+representation, the pendulum shortens at the bottom of the swing, and lengthens
+at the top. For an intuitive understanding, conservation of angular momentum
+indicates that shortening the pendulum at the bottom forces the mass to gain
+speed to compensate for the reduced length
 \cite{how_to_pump_a_swing}.
@@
 imparts a larger angular momentum to the mass by the time it reaches the bottom
-of its swing.
+of its swing, which is converted into a higher velocity the next time the
+pendulum is shortened.
 By alternating these processes, the pendulum experiences a net gain in
@@
    l^\star(\theta):= \begin{cases}
-      \underbar{l} & \theta \in [-pi, -\frac{\pi}{2}[ \cup [0,\frac{\pi}{2}[ \\
-      \bar{l} & \theta \in [-\frac{\pi}{2},0[ \cup [\frac{\pi}{2}, \pi[ \\
+      \underline{l} & \theta \in [-pi, -\frac{\pi}{2}[ \cup [0,\frac{\pi}{2}[ \\
+      \overline{l} & \theta \in [-\frac{\pi}{2},0[ \cup [\frac{\pi}{2}, \pi[ \\
    \end{cases}
@@
 \(l^\star(\theta)\) (see Figure \ref{fig:vlp-T-controller}). Let
-\(\Delta l := (\bar{l} - \underbar{l})/2\) and 
-\(l_{\text{avg}} := (\bar{l} + \underbar{l})/2\).
+\(\Delta l := (\overline{l} - \underline{l})/2\) and 
+\(l_{\text{avg}} := (\overline{l} + \underline{l})/2\).
 Providing these newly attached sinusoids with a frequency \(\omega = \frac{pi}{T}\) 
@@
    l_T(\theta) = \begin{cases}
-      \underbar{l} & \theta \in \left[-\pi + \frac{T}{2}, -\frac{\pi}{2} - \frac{T}{2}\right] 
+      \underline{l} & \theta \in \left[-\pi + \frac{T}{2}, -\frac{\pi}{2} - \frac{T}{2}\right] 
       \cup \left[\frac{T}{2}, \frac{\pi}{2} - \frac{T}{2}\right] \\
-      \bar{l} & \theta \in \left[-\frac{\pi}{2} + \frac{T}{2}, -\frac{T}{2}\right] 
+      \overline{l} & \theta \in \left[-\frac{\pi}{2} + \frac{T}{2}, -\frac{T}{2}\right] 
       \cup \left[\frac{\pi}{2} + \frac{T}{2}, \pi - \frac{T}{2}\right] \\
@@
    A VNHC of the form \(h(q,p) = l - l(\theta)\) injects energy if there exists 
-   \(l_\text{avg} \in [\underbar{l},\bar{l}]\) such that 
+   \(l_\text{avg} \in [\underline{l},\overline{l}]\) such that 
    \begin{equation}\label{eqn:vlp-energy-gain-condition}
@@
 injecting energy: it maximizes the derivative of \(E_\text{avg}\) under the
-restriction \(l(\theta) \in [\underbar{l},\bar{l}]\), so that the orbit in the
+restriction \(l(\theta) \in [\underline{l},\overline{l}]\), so that the orbit in the
 \((q,p)\)-plane diverges from the origin as fast as possible.
@@
 Let us define \((l^\star)\inv(\theta)\) by swapping the order of 
-\(\underbar{l}\) and \(\bar{l}\) in \(l^\star(\theta)\). Since this 
+\(\underline{l}\) and \(\overline{l}\) in \(l^\star(\theta)\). Since this 
 \textit{minimizes} the derivative of \(E_\text{avg}\) under the restriction
-\(l(\theta) \in [\underbar{l},\bar{l}]\), we can predict that 
+\(l(\theta) \in [\underline{l},\overline{l}]\), we can predict that 
 \((l^\star)\inv(\theta)\) is the VNHC representation of an optimal 

### 2020-09-03 latex/vlp.tex
added more TODOs and figured out what to do with images

@@
 
+% TODO: Describe the oscillation and rotation motion of a pendulum in the
+% qp-plane, and give a figure showing the orbiting motion for each of these.
+% Then describe what the orbit should do under energy injection (spiral away /
+% begin rotation outward) and energy dissipation (spiral in / rotate inward).
+
 \section{The VLP Constraint}
@@
 The time-optimal controller described by \cite{pumping_swing_standing_squatting}
-depends on a state trajectory in time. We will convert this time-based
-controller into a time-independent controller which still uses this time-optimal
-strategy. Notice that the control strategy depends on knowing when the
+depends on a trajectory in time. We will convert this 
+into a time-independent controller derived from the time-optimal
+strategy. Notice that the optimal controller depends on knowing when the
 system is at the ``bottom" or ``top" of the swing. Since being at the ``bottom"
 is the same as saying the pendulum angle is \(q = 0\) and being at the ``top" is
-equivalent to the pendulum's momentum being \(p = 0\), our method of
-VNHCs is especially well suited to this problem.
+equivalent to the pendulum's momentum being \(p = 0\), the time-independent
+controller will necessarily use the full phase \((q,p)\). While VHCs cannot
+satisfy this requirement, our method of VNHCs is especially well suited to this
+problem.
 
 Figure \ref{fig:vlp-optimal-controller-qp-plane} shows the time-optimal strategy directly
-on the \((q,p)\)-plane. The axes of this plane define where the pendulum should
-change length. Since the pendulum stays the same length inside each quadrant, we
+on the \((q,p)\)-plane.
+Since the VLP must stay the same length inside each quadrant and
+changes length only when it crosses one of the axes, we
 will use the angle \(\theta = \arctan_2(p,q)\) in the \((q,p)\)-plane to define
 the optimal controller \(l^\star(\theta)\) (\ref{eqn:vlp-optimal-controller}).
-Figure \ref{fig:vlp-optimal-controller} shows this controller as a function of
+Figure \ref{fig:vlp-optimal-controller} shows this time-optimal controller as a function of
 \(\theta\).
@@
    \centering
-   % TODO: Inser the image of the optimal controller in the (q,p) plane
+   % TODO: Inser the image of the optimal controller in the (q,p) plane: it
+   % should just be 4 coloured quadrants (no spiral) representing squatting and
+   % standing
    \caption{The time-optimal VLP controller from
@@
    \centering
-   %TODO: Insert the image of the optimal controller as a function of theta
+   %TODO: Insert the image of the optimal controller as a function of theta;
+   %that is, show l(theta) vs theta on ]-pi,pi] and show underline{l} and
+   %overline{l} 
    \caption{The time-optimal VLP controller converted to a discontinuous
@@
    \centering
-   % TODO: Insert the image of the smoothed VLP controller
+   % TODO: Insert the image of the smoothed VLP controller around lavg
    \caption{The smoothed VLP controller attained by setting the parameter 

### 2020-09-04 latex/images/vlp_optimal_controller_qp_plane.tex
Added VLP images for l(theta)

%! TEX program = lualatex

% Use :VimtexToggleMain to compile this file alone with Vimtex
\documentclass[tikz]{standalone}

\usepackage{sty/adantikz}

\begin{document}
% Draw a series of rectangles for the optimal time-varying controller in the
% (q,p) plane
\begin{tikzpicture}
    % Fill each quadrant with a coloured rectangle, where red = standing and
    % blue = squatting
    \filldraw[fill=red!40,draw=red,dashed] (0,0) rectangle (4,4);
    \filldraw[fill=blue!40,draw=blue,dashed] (0,0) rectangle (-4,4);
    \filldraw[fill=red!40,draw=red,dashed] (0,0) rectangle (-4,-4);
    \filldraw[fill=blue!40,draw=blue,dashed] (0,0) rectangle (4,-4);
    % Draw the axes of the (q,p) plane on top of these rectangles
    \draw[draw=black, <->] (-4.5,0) -- (4.5,0) node[right] {$q$};
    \draw[draw=black, <->] (0,-4.5) -- (0,4.5) node[above] {$p$};
    % Draw an arc with the theta label up to some node in the plane
    \draw[draw=black] (0,0) -- (2,1) node[fill=black,inner sep= 1pt] {};
    \draw[draw = black] (1,0) arc (0:26:1) node at (1.2,0.25) {$\theta$};
\end{tikzpicture}
\end{document}
% vim: set tw=80 ts=4 sw=4 sts=0 et ffs=unix :

### 2020-09-04 latex/vlp.tex
Added VLP images for l(theta)

@@
    \centering
-   % TODO: Inser the image of the optimal controller in the (q,p) plane: it
-   % should just be 4 coloured quadrants (no spiral) representing squatting and
-   % standing
-   \caption{The time-optimal VLP controller from
-   \cite{pumping_swing_standing_squatting} displayed in the
-\((q,p)\)-plane.}\label{fig:vlp-optimal-controller-qp-plane}
+   \includestandalone[width=0.5\textwidth]{images/vlp_optimal_controller_qp_plane}
+   \caption{The representation of the time-optimal controller for a standing
+      swing in the \((q,p)\)-plane. As per 
+      \cite{pumping_swing_standing_squatting}, the red region corresponds to
+      squatting and the blue region corresponds to standing.}
+      \label{fig:vlp-optimal-controller-qp-plane}
 \end{figure}
@@
    \centering
-   %TODO: Insert the image of the optimal controller as a function of theta;
-   %that is, show l(theta) vs theta on ]-pi,pi] and show underline{l} and
-   %overline{l} 
+   \includestandalone[width=0.5\textwidth]{images/vlp_optimal_controller}
    \caption{The time-optimal VLP controller converted to a discontinuous
-   function \(l(\theta)\).}\label{fig:vlp-optimal-controller}
+   function \(l(\theta)\), where \(\theta := \arctan_2(p,q)\).}\label{fig:vlp-optimal-controller}
 \end{figure}

### 2020-09-05 latex/vlp.tex
Added the images for vlp controllers, except the smooth one

@@
    \centering
-   \includestandalone[width=0.5\textwidth]{images/vlp_optimal_controller_qp_plane}
-   \caption{The representation of the time-optimal controller for a standing
-      swing in the \((q,p)\)-plane. As per 
-      \cite{pumping_swing_standing_squatting}, the red region corresponds to
-      squatting and the blue region corresponds to standing.}
+   \begin{subfigure}[t]{0.45\textwidth}
+      \includestandalone[width=\linewidth]{images/vlp_optimal_controller_qp_plane}
+      \caption{The time-optimal controller mapped to the \((q,p)\)-plane.}
       \label{fig:vlp-optimal-controller-qp-plane}
-\end{figure}
-
-\begin{figure}
-   \centering
-   \includestandalone[width=0.5\textwidth]{images/vlp_optimal_controller}
-   \caption{The time-optimal VLP controller converted to a discontinuous
-   function \(l(\theta)\), where \(\theta := \arctan_2(p,q)\).}\label{fig:vlp-optimal-controller}
+   \end{subfigure}
+   \hfill
+   \begin{subfigure}[t]{0.45\textwidth}
+      \includestandalone[width=\linewidth]{images/vlp_optimal_controller}
+      \caption{The time-optimal controller converted to a VNHC-style 
+      function \(l^\star(\theta)\), where \(\theta := \arctan_2(p,q)\).}\label{fig:vlp-optimal-controller}
+   \end{subfigure}
+   \caption{The time-optimal controller for a standing swing as derived by
+      \cite{pumping_swing_standing_squatting}. The colour red
+      corresponds to standing and blue to squatting.}
 \end{figure}
@@
       \left[-\pi,-\pi + \frac{T}{2}\right] \\
-      -\Delta l \sin(\omega(\theta - a)) + l_\text{avg} & 
+      -\Delta l \sin(\omega \theta) + l_\text{avg} & \theta \in [-\frac{T}{2},
+      \frac{T}{2}] \\
+      \Delta l \sin(\omega(\theta - a)) + l_\text{avg} & 
       \theta \in \left[a - \frac{T}{2}, a + \frac{T}{2}\right] \text{ for } 
-      a \in \left\{-\frac{\pi}{2}, 0, \frac{\pi}{2}\right\} \\
+      a \in \left\{-\frac{\pi}{2}, \frac{\pi}{2}\right\} \\
       -\Delta l \sin(\omega(\theta-\pi)) & \theta \in \left[\pi - \frac{T}{2},\pi\right] \\
@@
    \centering
-   % TODO: Insert an image of the controller with sinusoids attached
-   \caption{}\label{fig:vlp-T-controller}
+   \includestandalone[width=0.6\textwidth]{images/vlp_T_controller}
+   \caption{The continuous VNHC-style VLP controller \(l_T(\theta)\).}\label{fig:vlp-T-controller}
 \end{figure}

### 2020-09-08 latex/images/vlp_energy_in_out.tex
Added an image showing VNHCs which satisfy the theorem

%! TEX program = lualatex

% Use :VimtexToggleMain to compile this file alone with Vimtex
\documentclass[tikz]{standalone}

\usepackage{sty/adantikz}

\begin{document}
\begin{tikzpicture}
% Set the underline{l} and overline{l} positions
\pgfmathsetmacro{\lu}{2}
\pgfmathsetmacro{\lb}{4}
\pgfmathsetmacro{\dl}{(\lb - \lu)/2}
\pgfmathsetmacro{\lavg}{(\lb + \lu)/2}
% Define the figure
\begin{axis}[ 
 axis line style = thick,
 % Set the axes to be centered nicely
 axis x line = center, 
 axis y line = middle,
 xlabel = $\theta$, 
 x label style={at={(axis description cs:1.08,0)},anchor = east},
 xmin = -3.1415, xmax = 3.1415,
 ylabel = $l(\theta)$,
 y label style={at={(axis description cs:0.5,1.1)},anchor = north}, 
 ymin = 0, ymax = 5,
 % Custom tick labels at specified positions
 xticklabels = {$-\pi$, $-\frac{\pi}{2}$, $0$, $\frac{\pi}{2}$, $\pi$},
 xtick={-3.1415, -1.571, 0, 1.571, 3.1415},
 yticklabels = {$\underline{l}$,$l_{avg}$,$\overline{l}$},
 ytick = {\lu,\lavg,\lb}, % Choose the plot values for underline{l} and overline{l}
 % Move the \overbar{l} label to the right
 yticklabel style={xshift={(\ticknum==1)*(1cm)},yshift={(\ticknum==0)*(-0.25cm)+(\ticknum==1)*(0.25cm) + (\ticknum==2)*(0.25cm)}},
 legend pos=south east,
 axis on top
]
    % Add named boundaries for each of the blocked regions. We use "forget
    % plot" so they don't appear on the legend
    \addplot[draw=none, forget plot, name path global=I1U,domain=-pi:(-pi/2)]{\lu};
    \addplot[draw=none, forget plot, name path global=I2U,domain=(-pi/2):0]{\lu};
    \addplot[draw=none, forget plot, name path global=I3U,domain=0:(pi/2)]{\lu};
    \addplot[draw=none, forget plot, name path global=I4U,domain=(pi/2):pi]{\lu};

    \addplot[draw=none, forget plot, name path global=I1A,domain=-pi:(-pi/2)]{\lavg};
    \addplot[draw=none, forget plot, name path global=I2A,domain=(-pi/2):0]{\lavg};
    \addplot[draw=none, forget plot, name path global=I3A,domain=0:(pi/2)]{\lavg};
    \addplot[draw=none, forget plot, name path global=I4A,domain=(pi/2):pi]{\lavg};

    \addplot[draw=none, forget plot, name path global=I1B,domain=-pi:(-pi/2)]{\lb};
    \addplot[draw=none, forget plot, name path global=I2B,domain=(-pi/2):0]{\lb};
    \addplot[draw=none, forget plot, name path global=I3B,domain=0:(pi/2)]{\lb};
    \addplot[draw=none, forget plot, name path global=I4B,domain=(pi/2):pi]{\lb};
    % Fill in the region where the VNHC needs to be for injection in green
    \addplot[draw=none, forget plot, fill=green!40] fill between[of=I1U and I1A];
    \addplot[draw=none, forget plot, fill=green!40] fill between[of=I2A and I2B];
    \addplot[draw=none, forget plot, fill=green!40] fill between[of=I3U and I3A];
    \addplot[draw=none, fill=green!40] fill between[of=I4A and I4B];
    % Fill in the region where the VNHC needs to be for dissipation in yellow
    \addplot[draw=none, forget plot, fill=yellow!40] fill between[of=I1A and I1B];
    \addplot[draw=none, forget plot, fill=yellow!40] fill between[of=I2U and I2A];
    \addplot[draw=none, forget plot, fill=yellow!40] fill between[of=I3A and I3B];
    \addplot[draw=none, fill=yellow!40] fill between[of=I4U and I4A];

    % Draw the horizontal lines at lu, lb
    \draw[dashed, color=red] (-pi,\lu) -- (pi,\lu);
    \draw[dashed, color=blue] (-pi,\lb) -- (pi,\lb);
    % Draw the horizontal line at lavg
    \draw[dashed, color=black] (-pi,\lavg) -- (pi,\lavg);
    % Draw the vertical lines at +- pi/2
    \draw[dashed, color=gray!40] (-pi/2,0) -- (-pi/2,5);
    \draw[dashed, color=gray!40] (pi/2,0) -- (pi/2,5);

    \legend{injection, dissipation}; 
\end{axis}
\end{tikzpicture}
\end{document}
% vim: set tw=80 ts=4 sw=4 sts=0 et ffs=unix :

### 2020-09-08 latex/vlp.tex
Inserted the vlp T/smooth controllers
Inserted the VNHC injection/dissipation image

@@
 \section{The VLP Constraint}
-% TODO: Go through the development of the constraint and prove it gains energy
-% TODO: Explain how we came up with the constraint first. Talk about the child
-% on a swing, the paper showing the time-optimal control input, and describe why
-% we want to remove time from the equation. Discuss the lmin and lmax choices.
-% Why do we need to use a VNHC instead of a VHC: we are modelling the
-% time-optimal controller in a time-independent way, which requires knowledge of
-% which quadrant in (q,p)-space we're in. This means using the angle theta,
-% which is inherently a VNHC.
+% Go through the development of the constraint and prove it gains energy
 Since we wish to define a VNHC which injects energy into the VLP in a human-like
@@
    \centering
-   \includestandalone[width=0.6\textwidth]{images/vlp_T_controller}
+   \includestandalone[width=0.7\textwidth]{images/vlp_T_controller}
    \caption{The continuous VNHC-style VLP controller \(l_T(\theta)\).}\label{fig:vlp-T-controller}
@@
    \centering
-   % TODO: Insert the image of the smoothed VLP controller around lavg
-   \caption{The smoothed VLP controller attained by setting the parameter 
-      \(T\) to its upper limit
-   \(\frac{\pi}{2}\).}\label{fig:vlp-smoothed-controller}
+   \includestandalone[width=0.8\textwidth]{images/vlp_smoothed_controller}
+   \caption{The smoothed VLP controller which corresponds to the VNHC \(l =
+      l_\frac{\pi}{2}(\theta)\).}\label{fig:vlp-smoothed-controller}
 \end{figure}
@@
    \centering
-   % TODO: Add a figure showing regions for injection vs dissipation
-   \caption{}\label{fig:vlp-energy-in-out}
+   \includestandalone[width=0.5\textwidth]{images/vlp_energy_in_out}
+   \caption{A VNHC of the form \(l(\theta)\) which is entirely contained within
+      the green (yellow) regions will inject (dissipate) energy into the VLP
+      because it will satisfy Theorem \ref{thm:vlp-energy-stabilization}.}
+      \label{fig:vlp-energy-in-out}
 \end{figure}

### 2020-09-11 latex/acrobot.tex
Fixed a typo in acrobot.tex
Performed Line Wrapping
This cleans up some lines so I can iterate over them better
Finalize the first skeleton of the thesis
This commit finishes the skeleton of the thesis, with all sections and
headings as desired. There is currently no content here.
Edited the first chapter and minorly edited the VLP chapter, to send it to Manfredi
Wrote the motivation for the acrobot section

--- 2019-11-01 latex/acrobot.tex
@@
-%/===============/%
-\section{Energy Injection for the Acrobot}
+%! TEX root = main.tex
+
+%/========== Acrobot ==========/%
+
+\chapter{Application of VNHCs: The Acrobot}\label{ch:acrobot}
+\section{Motivation}
+The acrobot is a two-link pendulum, actuated at the center joint (as in Figure
+\ref{fig:acrobot-model}. Since its first description in 1990 
+\cite{nonlinear_controllers_nonintegrable_acrobot}, the
+acrobot has become a benchmark problem in control theory; it is an
+underactuated mechanical system which produces
+complex nonlinear motion from an easy-to-describe model.
+The acrobot models a gymnast on a bar,
+since it represents a torso (top link) and legs (bottom link) with motion
+generated by the swinging of the legs at the hips. It is also one of the
+simplest models for a biped walking robot
+\cite{toward_framework_biped_locomotion}.
+
+\begin{figure}
+    \centering
+    % TODO include acrobot model image
+    \caption{The acrobot model.}%
+    \label{fig:acrobot-model}
+\end{figure}
+
+Controlling the acrobot is a nontrivial task, since it is not feedback
+linearizable \cite{nonlinear_controllers_nonintegrable_acrobot}. Many
+researchers have studied the swing-up problem of driving the acrobot to its
+equilibrium point above the bar using partial feedback linearization
+\cite{swingup_problem_acrobot}, energy-based control
+\cite{swingup_acrobot_pendulum, swingup_acrobot_energy}, and through studying
+human motion \cite{swingup_giant_acrobot, motion_control_gymnastic_skill}.
+
+In gymnastics terminology, a ``giant" is when the gymnast performs full
+rotations around the bar with their body almost fully extended
+\cite{usagym_giant}. We are interested in studying the energy stabilization
+problem for the acrobot by using VNHCs to generate giant motion. The control of
+giant motion for the acrobot has been studied thoroughly 
+\cite{energy_pumping_robotic_swinging, swingup_giant_acrobot,
+control_giant_two_link_gymnastic_robot}, including several studies which use
+virtual (holonomic) constraints to achieve this
+\cite{dynamical_servo_acrobot_vc, control_giant_two_link_gymnastic_robot,
+xingbo_thesis}. However, these controllers are not intuitive nor easy to design:
+\cite{control_giant_two_link_gymnastic_robot} defines a constraint by inverting
+a trajectory in time onto the state space; \cite{dynamical_servo_acrobot_vc}
+requires a cascade controller to stabilize both a constraint and a desired limit
+cycle in the state space; and \cite{xingbo_thesis} enforces the giant by adding
+an extra state to estimate the velocity, which is a crude attempt at using
+VNHCs.
+
+In this chapter we will design a VNHC which can generate giant motion using much
+more intuitive methods. In the process of proving that this VNHC injects energy,
+we will also show a promising method for generating desirable VNHCs in the
+future.
+
+\section{Dynamics of the Acrobot}
+% TODO: Derive the acrobot dynamics in Hamiltonian form
+\section{Previous Approaches}
+% TODO: Give Xingbo's intuition, abstract away to qa = sin(theta) constraint
+% based on VLP but show it doesn't work well in our framework
+% TODO: Describe Xingbo's results and observe that his results were similar to a VNHC
+% except they used VHC tools. We will modify his approach to use VNHCs and prove
+% rigorously that we can stabilize any energy level set on the acrobot.
+
+\section{The Acrobot Constraint}
+% TODO: Give the constraint, state the theorem, and show the outline of the
+% proof (both as a sketch and in a figure)
+\section{Proving the Acrobot Gains Energy}
+% TODO: Prove this in sections
+
+\section{Experimental Results}
+% TODO: This might be a separate chapter
+
+%\section{Energy Injection for the Acrobot}
+%Suppose the acrobot is simple; 
+%that is, assume that the acrobot's two links are of equal length \(l\)
+%and of equal mass \(m\), with the mass concentrated at the tip of each rod.
+%
+%Let \(q = (q_u,q_a)\) where \(q_u\) is the shoulder angle and \(q_a\) is the hip angle.
+%
+%Then the acrobot has hamiltonian given by
+%\[
+%    \mathcal{H}(q,p) = \frac{1}{2} p^T M^{-1}(q) p + V(q)
+%\]
+%where
+%\[
+%    M^{-1}(q) = \frac{1}{m l^2 (2 - \cos^2(q_a))} 
+%    \begin{bmatrix}
+%        1              & -(1 + \cos(q_a)) \\
+%        -(1 + \cos(q_a)) & (3 + 2\cos(q_a))
+%    \end{bmatrix}
+%\]
+%and
+%\[
+%    V(q) = m g l (3 - 2 \cos(q_u) - \cos(q_u + q_a))
+%\]
+%
+%Notice that the acrobot satisfies \(\partial M^{-1}(q)/ \partial q_u = 0\),
+%which means it may be possible to enforce a regular solvable VNHC to inject energy
+%into the acrobot.
+%One such VNHC is
+%\begin{align}\label{eqn:acrobot_vnhc}
+%    \begin{split}
+%    h(q,p) &= q_a - f(q_u,p_u) \\
+%           &= q_a - \bar{q_a}\sin(\arctan(p_u,q_u)) \\
+%           &= q_a - \frac{\bar{q_a}p_u}{\sqrt{q_u^2 + p_u^2}}
+%    \end{split}
+%\end{align}
+%where \(\bar{q_a} \in ]0,\pi[\) is a parameter which dictates how much the 
+%controlled link can swing back and forth.
+%
+%To compute the regularity and solvability conditions, first observe that since
+%the acrobot has two coordinates and only \(p_a\) is actuatable, the control matrix
+%\(B = [0, 1]^T\) is constant and full rank everywhere. This also implies that
+%\(h\) is both regular and solvable if \(dh_q M^{-1}(q)B\) is full rank.
+%
+%Observe that
+%\begin{align}
+%    \begin{split}
+%        dh_q M^{-1}(q)B &= \frac{1}{m l^2 (2 - \cos^2(q_a))}
+%\left[ \frac{\bar{q_a}p_uq_u}{(q_u^2 + p_u^2)^{3/2}}, 1\right] 
+%    \begin{bmatrix}
+%        -(1 + \cos(q_a)) \\
+%        (3 + 2\cos(q_a))
+%    \end{bmatrix} \\
+%                     &= \frac{ (3+2\cos(q_a))(q_u^2 + p_u^2)^{3/2} 
+%                     - \bar{q_a}p_uq_u(1+cos(q_a))}
+%                     {m l^2 (q_u^2+p_u^2)^{3/2} (2 - \cos^2(q_a))}
+%    \end{split}
+%\end{align}
+%
+%TODO: \textbf{Determine if this is full rank everywhere except at \((q_u,p_u) = (0,0)\)}
+%
+%Solving for \(p_a\) is straightforward. The computation is shown below.
+%\begin{multline*}
+%    p_a = \left(dh_q M^{-1}(q)B\right)^{-1}  \\
+%        \left(\frac{m g l \bar{q_a}p_u q_u^2}{(q_u^2+p_u^2)^{3/2}} 
+%            \left(2\sin(q_u) + sin(q_u + q_a) \right)
+%        +
+%            \frac{1}{m l^2 (2 - \cos^2(q_a))}
+%            [ \frac{\bar{q_a}p_uq_u}{(q_u^2+p_u^2)^{3/2}}, 1] 
+%            \begin{bmatrix}
+%              1 \\
+%              -(1 + \cos(q_a))
+%          \end{bmatrix} \right)
+%\end{multline*}
+%\begin{align*}
+%    &= -\frac
+%    {
+%        \frac{ m g l \bar{q_a}q_u^2(2\sin(q_u) + \sin(q_u+q_a))}{(q_u^2+p_u^2)^{3/2}}
+%        + \frac{\bar{q_a} p_u^2 q_u}{m l^2 (q_u^2 + p_u^2)^{3/2} (2 -\cos^2(q_a))}
+%        - \frac{(1+\cos(q_a)) p_u}{m l^2 (2 - \cos^2(q_a))}
+%    }
+%    {
+%        \frac
+%        {
+%            (3 + 2\cos(q_a))(q_u^2+p_u^2)^{3/2} - \bar{q_a}(1 + \cos(q_a))q_up_u
+%        }
+%        {
+%            m l^2 (q_u^2+p_u^2)^{3/2} (2 - \cos^2(q_a))
+%        }
+%    } \\
+%    &= \frac
+%    {
+%        \frac
+%        {
+%            -m^2 g l^3 \bar{q_a}q_u^2(2\sin(q_u) + \sin(q_u+q_a))(2 - \cos^2(q_a))
+%            - \bar{q_a}p_u^2 q_u 
+%            + (1 + \cos(q_a))(q_u^2 + p_u^2)^{3/2} p_u
+%        }
+%        {
+%            m l^2 (q_u^2+p_u^2)^{3/2} (2 - \cos^2(q_a))
+%        }
+%    }
+%    {
+%        \frac
+%        {
+%            (3 + 2\cos(q_a))(q_u^2+p_u^2)^{3/2} - \bar{q_a}(1 + \cos(q_a))q_up_u
+%        }
+%        {
+%            m l^2 (q_u^2+p_u^2)^{3/2} (2 - \cos^2(q_a))
+%        }
+%    }
+%\end{align*}
+%
+%\begin{equation*}
+%    \Rightarrow p_a = \frac
+%        {
+%            (1 + \cos(q_a))(q_u^2 + p_u^2)^{3/2} p_u
+%            -m^2 g l^3 \bar{q_a}q_u^2(2\sin(q_u) + \sin(q_u+q_a))(2 - \cos^2(q_a))
+%            - \bar{q_a}q_up_u^2
+%        }
+%        {
+%            (3 + 2\cos(q_a))(q_u^2+p_u^2)^{3/2} - \bar{q_a}(1 + \cos(q_a))q_up_u
+%        }
+%\end{equation*}
+%
+%With \(p_a\) in hand, it is now possible to solve for the zero dynamics under the constraint. To simplify the notation we define \(s_u := \sin(q_u)\),
+%\(s_{ua} := \sin(q_u + q_a)\), 
+%and \(c_a := \cos(q_a)\).
+%\begin{align*}
+%    \dot{q_u} &= \frac{1}{m l^2 (2 - c_a^2)} [ 1, -(1+c_a)]
+%    \begin{bmatrix} p_u \\ p_a \end{bmatrix} \\
+%    \dot{p_u} &= - m g l (2 s_u + s_{ua})
+%\end{align*}
+%Expanding out \(\dot{q_u}\) we get
+%\begin{multline*}
+%    \dot{q_u} = 
+%    \frac{p_u}{m l^2 (2 - c_a^2)}\\
+%    + \frac
+%    {
+%        m^2 g l^3 \bar{q_a}q_u^2(2s_u + s_{ua})(2 - c_a^2)(1 + c_a)
+%        - (1+c_a)^2(q_u^2+p_u^2)^{3/2}p_u 
+%        + \bar{q_a}(1+c_a)q_up_u^2
+%    }
+%    {
+%        ml^2 (2-c_a^2) ((3 + 2c_a)(q_u^2+p_u^2)^{3/2} - \bar{q_a}(1 + c_a)q_up_u)
+%    }
+%\end{multline*}
+%and, putting everything over one denominator,
+%\begin{align*}
+%    \dot{q_a} &= \frac
+%    {
+%        (3 + 2c_a)(q_u^2+p_u^2)^{3/2}p_u 
+%        - (1+c_a)^2(q_u^2+p_u^2)^{3/2}p_u 
+%        +m^2 g l^3 \bar{q_a}q_u^2(2s_u + s_{ua})(2 - c_a^2)(1 + c_a)
+%    }
+%    {
+%        ml^2 (2-c_a^2) ((3 + 2c_a)(q_u^2+p_u^2)^{3/2} - \bar{q_a}(1 + c_a)q_up_u)
+%    } \\
+%    &= \frac
+%    {
+%        (2-c_a^2)\left((q_u^2+p_u^2)^{3/2}p_u 
+%        +m^2 g l^3 \bar{q_a}q_u^2(2s_u + s_{ua})(1 + c_a)\right)
+%    }
+%    {
+%        ml^2 (2-c_a^2) ((3 + 2c_a)(q_u^2+p_u^2)^{3/2} - \bar{q_a}(1 + c_a)q_up_u)
+%    }
+%\end{align*}
+%Therefore, the zero-dynamics of the acrobot under the constraint 
+%\(q_a = \sin(\arctan(p_u,q_u))\) are:
+%\begin{align}\label{eqn:acrobot_eom_qp}
+%\begin{split}
+%    \dot{q_u} &= \frac
+%    {
+%        (q_u^2+p_u^2)^{3/2}p_u 
+%        +m^2 g l^3 \bar{q_a}q_u^2(2s_u + s_{ua})(1 + c_a)
+%    }
+%    {
+%        ml^2 \left((3 + 2c_a)(q_u^2+p_u^2)^{3/2} - \bar{q_a}(1 + c_a)q_up_u\right)
+%    } \\
+%    \dot{p_u} &= - m g l (2s_u + s_{ua})
+%\end{split}
+%\end{align}
+%
+%It may be convenient to represent these equations of motion  in polar coordinates.
+%Letting \(r = \sqrt{q_u^2 + p_u^2}\) be the radius and 
+%\(\theta = \arctan(p_u,q_u)\) be the angle in the \((q_u,p_u)\) plane and taking
+%derivatives, it is easy to show that
+%\begin{align}\label{eqn:dr_and_dt_general}
+%\begin{split}
+%    \dot{r} &= [\cos(\theta),\sin(\theta)]
+%    \begin{bmatrix} \dot{q_u}(r,\theta) \\ \dot{p_u}(r,\theta) \end{bmatrix} \\
+%    \dot{\theta} &= \frac{1}{r}[-\sin(\theta), \cos(\theta)]
+%    \begin{bmatrix} \dot{q_u}(r,\theta) \\ \dot{p_u}(r,\theta) \end{bmatrix}
+%\end{split}
+%\end{align}
+%
+%By expanding out (\ref{eqn:dr_and_dt_general}) then substituting \(q_u = r\cos(\theta)\)
+%and \(p_u = r\sin(\theta)\)
+%it is straightforward computation to find the equations of motion in cylindrical \((r,\theta)\)
+%coordinates. To simplify the notation yet again, define
+%\(s(g)_\theta := \sin(g(\theta))\) and \(c(g)_\theta := \cos(g(\theta))\) 
+%for any function \(g: \mathbb{R} \rightarrow \mathbb{R}\). The final equations
+%of motion in \((r,\theta)\) coordinates are
+%\begin{align}\label{eqn:acrobot_eom_rt}
+%\begin{split}
+%    \dot{r} &= \frac
+%    {
+%        r^2 c_\theta s_\theta 
+%        + m^2 g l^3 (2 s(rc)_\theta + s(rc + \bar{q_a}s)_\theta)
+%        \left(
+%            (1+c(\bar{q_a}s)_\theta)\bar{q_a}c_\theta 
+%            - (3 + 2c(\bar{q_a}s)_\theta)s_\theta r
+%        \right)
+%    }
+%    {
+%        m l^2 \left((3+2c(\bar{q_a}s)_\theta)r 
+%        - \bar{q_a}(1+ c(\bar{q_a}s)_\theta)c_\theta s_\theta\right)
+%    }\\
+%    \dot{\theta} &= -\frac
+%    {
+%        r s_\theta^2
+%        + m^2 g l^3 (2 s(rc)_\theta + s(rc + \bar{q_a}s)_\theta)
+%        (3 + 2c(\bar{q_a}s)_\theta)c_\theta
+%    }
+%    {
+%        m l^2 \left((3+2c(\bar{q_a}s)_\theta)r 
+%        - \bar{q_a}(1+ c(\bar{q_a}s)_\theta)c_\theta s_\theta\right)
+%    }
+%\end{split}
+%\end{align}
+
+%/========== /Acrobot ==========/%
+% vim: set tw=80 ts=4 sw=4 sts=0 et ffs=unix :

### 2020-09-15 latex/acrobot.tex
UPdated introduction to acrobot and added simple acrobot

@@
     \centering
-    % TODO include acrobot model image
+    \includestandalone[width=0.7\textwidth]{images/acrobot_model}
     \caption{The acrobot model.}%
@@
 control_giant_two_link_gymnastic_robot}, including several studies which use
-virtual (holonomic) constraints to achieve this
+virtual (holonomic) constraints to achieve this behaviour
 \cite{dynamical_servo_acrobot_vc, control_giant_two_link_gymnastic_robot,
-xingbo_thesis}. However, these controllers are not intuitive nor easy to design:
+xingbo_thesis}. However, these controllers are neither intuitive nor easy to design:
 \cite{control_giant_two_link_gymnastic_robot} defines a constraint by inverting
@@
 cycle in the state space; and \cite{xingbo_thesis} enforces the giant by adding
-an extra state to estimate the velocity, which is a crude attempt at using
-VNHCs.
+an extra state to estimate the velocity, which increases the dimensionality of
+the problem as a crude approach to using VNHCs.
 
-In this chapter we will design a VNHC which can generate giant motion using much
-more intuitive methods. In the process of proving that this VNHC injects energy,
-we will also show a promising method for generating desirable VNHCs in the
-future.
+In this chapter we will design a physically-intuitive VNHC which generates giant
+motion. We will prove that this VNHC injects energy into the acrobot;
+in the process of completing this proof,
+we will describe a promising method which may be used to generate desirable 
+VNHCs in the future.
 
@@
 % TODO: Derive the acrobot dynamics in Hamiltonian form
+For the acrobot model displayed in Figure \ref{fig:acrobot-model}, 
+the COM of each link lies somewhere along the link itself. 
+While this is the most general representation of an acrobot, the dynamics are
+somewhat complicated and are difficult to use when proving the results in the 
+rest of this chapter. For simplicity, we will
+assume the acrobot is comprised of two massless rods of equal length with 
+equal masses at the tips, as in Figure \ref{fig:simple-acrobot-model}; we call
+this a \textbf{simple} acrobot.
+
+\begin{figure}
+    \centering
+    \includestandalone[width=0.5\textwidth]{images/simple_acrobot_model}
+    \caption{A simple acrobot with lengths \(l\) and masses of size \(m\).}
+    \label{fig:simple-acrobot-model}
+\end{figure}
+
 \section{Previous Approaches}

### 2020-09-17 latex/vnhcs.tex
Edited phrasing up to and including simply actuated hamiltonian

@@
 \textit{equations of motion} (EOM) produced by Newton's second law \(F = m a\).
-While this technique works for simple systems, it is impossible to
-apply to a majority of mechanical systems since most of the forces 
-on the system are not explicitly known.
-
-Rather than modeling a mechanical system by cartesion positions and constraints,
+While this technique works for simple systems, it is tedius and becomes
+impossible to apply to complex mechanical systems where the forces are not
+explicitly known.
+
+Rather than modeling a mechanical system by cartesian positions and constraints,
 it is often feasible to represent the position of the system using \(n\)
@@
 \textit{generalized coordinates}, where \(n = 3N - r\) is the number of
-\textit{degrees of freedom} (DOF) of the system \cite{greenwood_dynamics}.
-
-% TODO:
-\textbf{TODO: Image comparing Newtonian to g.c.}
+\textit{degrees of freedom} (DOF) of the system \cite{greenwood_dynamics}. For
+instance, Figure \ref{fig:barbell} shows a barbell on a \(2D\)-plane which can
+rotate freely on that plane. 
+The barbell has \(n = 3\) DOF, so it can be described by three
+independent generalized coordinates with no equations of constraint.
+
+\begin{figure}
+   \centering
+   \begin{subfigure}[]{0.75\textwidth}
+      \includestandalone[width=\textwidth]{images/newtonian_barbell}
+      \caption{The Newtonian representation of the barbell 
+      requires all six cartesian positions and the corresponding EOC.}
+   \end{subfigure}
+   \begin{subfigure}[]{0.75\textwidth}
+      \includestandalone[width=\textwidth]{images/gc_barbell}
+      \caption{One possible set of three generalized coordinates is
+       \((x_0,y_0,\theta)\), which represent the position of the 
+       center of the bar and the angle of the barbell in the \(xy\)-plane.}
+   \end{subfigure}
+   \caption{A mechanical system with \(N = 2\) points at \((x_1,y_1,z_1)\)
+   and \((x_2,y_2,z_2)\) separated by a bar with \(r = 3\) EOC given by 
+   \(c = \norm{x_1 - x_2}\), \(z_1 = 0\), and \(z_2 = 0\). This system has \(n =
+   3\) degrees of freedom.}
+   \label{fig:barbell}
+\end{figure}
 
@@
 If \(k < n\), we say the system is \textit{underactuated} with degree of
-underactuation \(n - k\).
+underactuation \((n - k)\).
 
@@
 \subsection{Hamiltonian Mechanics}
-Hamiltonian mechanics provides an equivalent representation of the EOM
-by converting the \(n\) second-order ODEs generated by Lagrangian mechanics into
-\(2n\) first-order ODEs.
+Hamiltonian mechanics converts the \(n\) second-order ODEs generated by
+Lagrangian mechanics into an equivalent set of \(2n\) first-order ODEs.
 
@@
 \begin{equation}\label{eqn:p-i}
-    p_i(q,\dot{q}) = \pdiff{\mathcal{L}}{\dot{q}_i}(q,\dot{q})
+    p_i(q,\dot{q}) := \pdiff{\mathcal{L}}{\dot{q}_i}(q,\dot{q})
 \end{equation}
-To ease notation, we will write \(p = (p_1, \ldots, p_n) \in \R^n\) and say that
-\(p\) is the \textit{conjugate of momenta to \(q\)}.
+To ease notation, we write \(p = (p_1, \ldots, p_n) \in \R^n\) and call
+\(p\) the \textit{conjugate of momenta to \(q\)}.
 Note that each \(p_i\) is a linear function of \(\dot{q}\), and one can typically
@@
 The \textit{Hamiltonian} of the system in
-\(\{q,p\}\) coordinates is defined as the Legendre transform 
+\(\{q,p\}\) coordinates is the ``Legendre transform"
 (\ref{eqn:hamiltonian-legendre}) of the Lagrangian \cite{landau_mechanics}.
@@
 \end{equation}
-The EOM in this framework can be shown to be the \(2n\)
-first-order equations called \textit{Hamilton's equations}
+The EOM in the Hamiltonian framework are the \(2n\)
+first-order equations called \textit{Hamilton's equations}. They are given by:
 \begin{equation}\label{eqn:hamiltons-eqns}
@@
 \end{equation}
-where \(B(q) \in \R^{n\times k}\) is the same input matrix used by the
-Lagrangian and \(\tau \in \R^k\) is the vector of generalized input forces.
-
-If the Lagrangian has quadratic kinetic energy as in (\ref{eqn:lagrangian}), 
-the conjugate of
-momenta to \(q\) can be computed explicitly by \(p = D(q)\dot{q}\).
-
+Here, \(B(q) \in \R^{n\times k}\) is the same input matrix used by the
+Lagrangian framework, with \(\tau \in \R^k\) the same vector of generalized
+input forces.
+
+If the kinetic energy of the system is quadratic as in (\ref{eqn:lagrangian}), 
+the conjugate of momenta becomes \(p = D(q)\dot{q}\). Since \(D(q)\) is symmetric
+and positive definite, it is invertible at each \(q \in \mathcal{Q}\).
 The resulting Hamiltonian system reduces to (\ref{eqn:hamiltonian}).
@@
 coordinates \((q,p) \rightarrow (Q,P)\) is a \textit{canonical
-transformation} if \(\{Q,P\}\) are canonical coordinates with Hamiltonian
-\(H(Q,P) = \mathcal{H}\left(q(Q,P), p(Q,P)\right)\).
+transformation} if \(\{Q,P\}\) preserve the Hamiltonian structure; that is, if
+they are canonical coordinates under the Hamiltonian
+\(\mathcal{H}\left(q(Q,P), p(Q,P)\right)\).
 
 \section{Simply Actuated Hamiltonian Systems}
+\textbf{Manfredi: What term should we use instead of ``simply actuated"?}
+
 % Describe the change of coordinates to get into q_u/q_a mode and show
@@
 % inertia, D for normal coordinates
-Given a Hamiltonian mechanical system (\ref{eqn:hamiltonian}), it is not obvious
-how any particular input force \(\tau_i\) affects the conjugate of momenta. 
-This is because \(\tau\) is transformed by the input matrix \(B(q)\) before
-entering the EOM.
-
-We will define a new class of Hamiltonian systems where the effect of the input
-forces on the conjugate of momenta is made obvious. This class of systems will
-form the backbone for the rest of the theory developed in this thesis.
+Suppose we are given a Hamiltonian mechanical system (\ref{eqn:hamiltonian}).
+Because \(\tau\) is transformed by the input matrix \(B(q)\) before
+entering the EOM, it is not in general clear how any particular input force \(\tau_i\)
+will affect the dynamics of the system. 
+In this section, we define a new class of Hamiltonian systems where the effect
+of the input forces is made obvious. This class of
+systems will form the backbone for the rest of the theory developed in this
+thesis.
 
 \begin{defn}
-    Let \(\mathcal{H}\) be an underactuated Hamiltonian system 
-    \(\mathcal{H}\) with degree of underactuation \((n-k)\geq 0\). 
-    A pair of canonical coordinates \(\{q,p\}\) for this system
+    Let \(\mathcal{H}\) be an \(n\)-DOF Hamiltonian system 
+    with \(k \leq n\) actuators. 
+    A set of canonical coordinates \(\{q,p\}\) for this system
     are said to be \textit{simply actuated coordinates} if the
@@
     \]
-    The first \((n-k)\) coordinates \(q_u\) are called the \textit{unactuated
-    coordinates}, while the remaining \(k\) coordinates \(q_a\) are called the
-    \textit{actuated coordinates}. We write \(q = (q_u, q_a)\); the
-    corresponding conjugate of momenta \(p = (p_u, p_a)\) are called the
-    unactuated and actuated momenta (respectively).
+    The first \((n-k)\) coordinates, labelled \(q_u\), are called the
+    \textit{unactuated coordinates}. The remaining \(k\) coordinates, labelled
+    \(q_a\), are called the \textit{actuated coordinates}. When grouping them
+    together, we will always put them in the order \((q_u, q_a)\) to fit with 
+    the definition. 
+    The corresponding \((p_u, p_a)\) are called the unactuated and actuated
+    momenta, respectively.
 \end{defn}
@@
     A Hamiltonian system is said to be \textit{simply actuated} if there exists
-    a canonical transformation from any set of canonical coordinates 
-    \(\{q,p\}\) into simply actuated coordinates for \(\mathcal{H}\).
+    a canonical transformation from any canonical coordinates 
+    \(\{q,p\}\) into simply actuated coordinates.
 \end{defn}
@@
     The input matrix \(B(q) \equiv B \in \R^{n\times k}\) is constant,
-    full rank, and \(k \leq n\).
+    full rank, and \(k < n\).
 \end{assm}
@@
     which is right semi-orthogonal 
-    (\(B^\perp(B^\perp)\tpose = \Id{(n-k)}\))
-    and which is a left-annihilator for \(B\)
-    (\(B^\perp B = \Zmat{(n-k) \times k}\)).
+    \(\left(B^\perp(B^\perp)\tpose = \Id{(n-k)}\right)\)
+    and which is a left-annihilator for \(B\). 
+    That is, \(B^\perp B = \Zmat{(n-k) \times k}\).
 \end{assm}
 
-Note that if \(k = (n-1)\), the existence of any left annihilator \(A^0\) for
-\(B\) implies the left annihilator \(B^\perp := A^0/\norm{A^0}\) satisfies Assumption
-\ref{assm:B-perp}.
+Note that if \(k = (n-1)\), the existence of any left annihilator 
+\(A^0 \in \R^{1\times n}\) implies the left annihilator 
+\(B^\perp := A^0/\norm{A^0}\) satisfies Assumption \ref{assm:B-perp}.
 
@@
     Assume \textbf{without loss of generality} that the input matrix \(B\) is
-    left semi-orthogonal (\(B\tpose B = \Id{k}\)).
+    left semi-orthogonal. That is, \(B\tpose B = \Id{k}\).
 \end{assm}
@@
 \]
-% TODO: Find a source for SVD
-where \(\sigma_i \neq 0\) because \(B\) is full-rank \textbf{TODO: source?}.
+where \(\sigma_i \neq 0\) because \(B\) is full-rank \cite{calculating_svd}.
 Defining \(T \in \R^{k \times k}\) by
@@
     \end{bmatrix} = \Id{n}
-    \Rightarrow
-    \mathbf{B}^{-1} = \mathbf{B}\tpose
-\]
+\]
+Hence, \(\mathbf{B}\) is invertible with \(\mathbf{B}\inv = \mathbf{B}\tpose\).
 
 The following theorem shows that \(\mathbf{B}\) provides a canonical
-transformation into simply actuated coordinates, so that all unactuated momenta
-are unaffected by the input forces.
+transformation into simply actuated coordinates, so that only the actuated momenta
+are affected by the input forces.
 
@@
     \begin{align}\label{eqn:simple-hamiltonian}
-        \hat{H}(Q,P) &= 
+        \mathcal{H}(Q,P) &= 
         \frac{1}{2} P\tpose \Minv(Q) P + V(Q) \\
@@
 \end{thm}
+\vspace{-4.5em} % For some reason not having this puts a large space before proof
 \begin{proof}
@@
     Therefore, by (45.10) in \cite{landau_mechanics}, the coordinate change 
-    \((Q = \mathbf{B}q, P = \mathbf{B}p)\) is a canonical transformation with
-    new Hamiltonian 
-    \(\hat{H}(Q,P) = \mathcal{H}(\mathbf{B}\tpose Q,\mathbf{B}\tpose P)\).
-
+    \((Q = \mathbf{B}q, P = \mathbf{B}p)\) is a canonical transformation.
     Furthermore, since \(\dot{P} = \mathbf{B}\dot{p}\), the new input matrix is
@@
     so the coordinates \(\{Q = (q_u,q_a), P = (p_u,p_a)\}\) are simply actuated coordinates for 
-    \(\hat{H}\) as desired.
+    \(\mathcal{H}\) as desired.
 \end{proof}
@@
 by the generalized velocities of a system.
-Many authors have attempted to extend VHCs to enforce these relations 
+For the child on a swing, whether they extend or retract their legs
+depends on their direction of motion. This inherently requires knowledge of
+their velocity, which precludes the usage of VHCs. 
+Many authors have attempted to extend the theory of VHCs to enforce relations 
 \(h(q,\dot{q}) = 0\) of the full state to account for this drawback. 
@@
 % d/dt(dh_pu)
-    Moving on to \(\diff{dh_{p_u}}{t}\), a similar approach as
+    Moving on to \(\diff{}{t}\left\{dh_{p_u}\right\}\), a similar approach as
     \(\diff{}{t}\left\{dh_q\right\}\) yields the derivative in matrix form.
@@
     \end{align*}
-    Now observe that the \(i^\text{th}\) row of
+% d/dt(1/2 * pudmat)
+    For the next piece, observe that the \(i^\text{th}\) row of
     \(\frac{1}{2} \diff{}{t} \left\{ \pudmat \right\}\)
@@
     \]
-    This allows us to write the entire derivative in matrix form:
+    In matrix form, this becomes:
     \begin{align*}

### 2020-09-18 latex/vnhcs.tex
First full edit of VNHCs chapter

@@
 % Motivation for VNHCs, they are an extension of VHCs, etc.
-Let us imagine a child on a swing, who wants to go as high as possible. The
-child will push off the ground and start swinging with small oscillations,
-then extend and retract their feet appropriately to gain energy. 
-If a roboticist were designing a machine to do this task, they might design a
-control mechanism which makes the legs track a trajectory over time. For
-example, they might tell the robot to extend and retract its legs every two
-seconds. In ideal situtations, this technique would work perfectly because the
-leg motion is synchronized with the swing to gain energy as fast as possible.
-
-% TODO: Get some sources here. Show with research that trajectory control is
-% unsafe, and that path tracking is better.
-Most children have an adult pushing the swing to help them go higher, or perhaps
+Let us imagine a child on a swing who wants to reach the largest height
+possible. 
+To begin, the child pushes off the ground to imbue the swing with small oscillations.
+What allows them to increase the amplitude of these oscillations is the
+appropriate extension and retraction of their feet.
+If a roboticist were creating a machine to replicate this behaviour, they might
+design a robot whose legs extend and retract at specific time intervals. 
+At first glance, this technique should work perfectly because the leg motion
+would synchronize with the swinging frequency, thereby injecting energy as
+quickly as is physically possible.
+
+Unfortunately, a deeper analysis reveals the flaw with this design.
+Most children are not counting out the time in their head; rather, they observe
+their current position and velocity and adjust their legs as required. 
+For example, many children have an adult pushing the swing, or perhaps
 they are swinging on a windy day. In either case, they adjust their leg
 motion accordingly when presented with these external disturbances, without
-keeping track of the time it takes to kick their legs. The
-standard control technique of tracking a function of time, known as
-\textit{trajectory tracking}, does not work as well as human motion 
-because the disturbance affecting the swing will desynchronize the leg motion
-and the swing, thereby stopping the energy-gaining effects. 
-As has been observed in the field of mobile robotics,
-a trajectory tracking controller which has lagged behind due to a disturbance
-will do whatever is required to ``catch up" to
-the trajectory. This can cause major safety issues, which is why mobile
-roboticists prefer to design controllers which drive a vehicle towards a path in
-space rather than a trajectory in time.
-
-This path tracking approach has been extended to biologically-inspired robotics
-in a method known as virtual holonomic constraints (VHCs).
-Instead of a robot's actuators tracking
-a trajectory over time, VHCs use the actuators to enforce a relation
-\(h(q) = 0\) of the configuration \cite{vhcs_for_el_systems}. 
+keeping track of time. 
+Hence, the standard control technique of tracking a function of time (known as
+\textit{trajectory tracking}) does not truly replicate human behaviour.
+Even if the robot's legs perfectly track a specified trajectory, 
+an external disturbance will desynchronize the leg motion
+with the swing - thereby stopping the amplitude-increasing effects.
+
+%As has been observed in the field of mobile robotics, a trajectory tracking
+%controller which has lagged behind due to a disturbance will do whatever is
+%required to ``catch up" to the trajectory. 
+%This can cause major safety issues, which is why many mobile roboticists prefer to
+%design controllers which drive a vehicle along a gemoetric path in space rather than a
+%trajectory in time.
+
+%This path tracking approach has been extended to biologically-inspired robotics
+%in a method known as \textit{virtual holonomic constraints} (VHCs).
+Rather than tracking a trajectory over time, a more human-like behaviour 
+is to force the robot's legs track a function of the swing's state. 
+One recent control method known as \textit{virtual holonomic constraints} (VHCs)
+uses the actuators to enforce a relation \(h(q) = 0\) of the configuration
+\cite{vhcs_for_el_systems}. 
 This method has provided incredible results in the development of 
@@
 The downside to VHCs is that they do not use the additional information imparted
-by the generalized velocities of a system.
+by the velocity of a system.
 For the child on a swing, whether they extend or retract their legs
-depends on their direction of motion. This inherently requires knowledge of
-their velocity, which precludes the usage of VHCs. 
+depends on their direction of motion. 
+This inherently requires knowledge of their current velocity, which precludes
+the usage of VHCs. 
 Many authors have attempted to extend the theory of VHCs to enforce relations
 \(h(q,\dot{q}) = 0\) of the full state to account for this drawback. 
-Since these relations use actuators to
-restrict both the configuration and generalized velocity, they are called
-virtual \textit{nonholonomic} constraints. This idea has been used for
-human-robot interaction
+Since these relations use actuators to restrict both the configuration and
+velocity of a system, they are called virtual \textit{nonholonomic} constraints.
+This idea has been used for human-robot interaction
 \cite{vnhc_human_robot_cooperation,psd_based_vnhc_redundant_manipulator,haptic_vnhc},
@@
 hybrid_zero_dynamics_bipedal_nhvcs,output_nhvc_bipedal_control}.
-Most interestingly, the nonholonomic approach has been shown to be more robust
+Most interestingly, this nonholonomic approach is more robust
 than standard VHCs when applied to bipedal robotics \cite{nhvc_incline_walking}.
-This suggests that virtual nonholonomic constraints are more capable of
-stabilizing specific energy levels while rejecting disturbances, all while
-producing realistic biological motion. It is for this reason we choose to study
-nonholonomic constraints in this thesis.
+In particular, virtual nonholonomic constraints may be capable of injecting and
+dissipating energy from a system in a robust manner, all while producing
+realistic biological motion. This is what we aim to prove in this thesis.
 
@@
 virtual nonholonomic constraints: 
-all the applications listed above use their own definition of a virtual
-nonholonomic constraint, which makes it difficult to compare and analyze their
-work.
-
-This section will provide a new characterization of virtual nonholonomic
-constraints in the Hamiltonian framework. 
-The goal is to provide a consistent, rigorous foundation for
-designing constraints on a general class of systems.
+all the applications listed above use their own definitions, which makes it
+difficult to compare and generalize their work.
+
+This section will provide a standard characterization of virtual nonholonomic
+constraints using Hamiltonian mechanics. 
+The goal is to provide a consistent, rigorous foundation for designing
+constraints on a general class of systems.
 
@@
 If we define the error term  \(e = h(q,p)\), stabilizing \(\Gamma\) is
-equivalent to solving for \(\tau(q,p)\) which drives \(e \rightarrow 0\) and
-\(\dot{e} \rightarrow 0\).
-Let us imagine that we have no further structure on the VNHC. 
-Then \(\dot{e} = dh_q \dot{q} + dh_p \dot{p}\), where \(\tau\) appears inside
-\(\dot{p}\). To solve for \(\tau\) explicitly, we must have that \(dh_p\) is
-invertible for all \((q,p)\).
+equivalent to solving for the controller \(\tau(q,p)\) which drives 
+\(e \rightarrow 0\) and \(\dot{e} \rightarrow 0\).
+Supposing there is no further structure on the VNHC, 
+then \(\dot{e} = dh_q \dot{q} + dh_p \dot{p}\). Since \(B(q) \tau\) appears inside
+\(\dot{p}\), solving for \(\tau\) explicitly requires 
+\(dh_p(q,p)B(q) \in \R^{k \times k}\) to be invertible for all \((q,p)\).
 %which is the same as requiring that \(h(q,p)\) is strictly monotonic in \(p\).
 
-This is a very restrictive condition, which many VNHCs will not satisfy. For
-this reason, it is preferable to have the torque \(\tau\) appear after two 
-derivatives of \(e\) so that there is more freedom in the types of constraints
-one can use.
+This condition severely restricts the types of constraints one can use. 
+It is preferable instead to have the torque \(\tau\) appear after two
+derivatives of \(e\) to allow for more freedom in choosing a constraint.
 Mathematically, if each \(\tau_i\) appears only after two derivatives, one says 
@@
 observed that a relation which uses only the unactuated conjugate of momentum
-\(p_u\) cannot have \(\tau\) appearing after only one derivative. Of course,
+cannot have \(\tau\) appearing after only one derivative. Of course,
 they performed their research in Lagrangian form; 
-for us to be able to use \(p_u\) in its Hamiltonian form, 
-we will continue with the following assumption for the rest of the chapter.
+for us to be able to use the Hamiltonian coordinate \(p_u\), 
+we must continue with the following assumption for the rest of the chapter.
 
@@
     everywhere on the constraint manifold \(\Gamma\).
-    That is, a regular VNHC is a function with domain 
-    \(\mathcal{Q} \times \mathcal{P}_u\) which satisfies the rank condition.
+    That is, a regular VNHC is a function 
+    \(h : \mathcal{Q} \times \mathcal{P}_u \rightarrow \R^k\) which satisfies
+    the rank condition.
 \end{thm}
@@
     \end{align*}
-    This means that the entire matrix is given by
-    \begin{align*}
+    This means that \(\diff{}{t}\left\{dh_q\right\}\) is given by
+    \begin{align}
         \diff{}{t}\left\{dh_q\right\} &= 
@@
         \left(p\tpose \Minv(q)\right)\right)
-        \Hess_q\left\{h\right\}\tpose -{} \\
+        \Hess_q\left\{h\right\}\tpose -{} 
+        \nonumber \\
       & \left(\frac{1}{2}\Id{n} \otimes 
@@
         + \Id{n}\otimes dV_{q_u}(q)\right)\partial_{p_u}\partial_{q}h 
-    \end{align*}
+    \end{align}
 % C_1(q,p)
@@
     \(C_1 : \mathcal{Q} \times \mathcal{P} \rightarrow \R^{n\times n}\) by
-    \[
+    \begin{equation}
         C_1(q,p) := \sum\limits_{j=1}^n \pdiff{\Minv}{q_j}(q)\dot{q}_j 
         = \sum\limits_{j=1}^n \pdiff{\Minv}{q_j}(q)\left(1_j\tpose\Minv(q)p\right)
-    \]
+    \end{equation}
 % d/dt(dh_pu)
-    Moving on to \(\diff{}{t}\left\{dh_{p_u}\right\}\), a similar approach as
+    Moving on to \(\diff{}{t}\left\{dh_{p_u}\right\}\), a similar approach to
     \(\diff{}{t}\left\{dh_q\right\}\) yields the derivative in matrix form.
     It is given by:
-    \begin{align*}
+    \begin{align}
         \diff{}{t}\left\{dh_{p_u}\right\} &= 
         \left( \Id{n} \otimes \left(p\tpose \Minv(q)\right)\right) 
-        \partial_q\partial_{p_u}h -{} \\
+        \partial_q\partial_{p_u}h -{} 
+        \nonumber \\
       & \left(\frac{1}{2}\Id{n} \otimes \left(\pudmat\right) + 
       \Id{n} \otimes \nabla_{q_u}V(q)\right)\Hess_{p_u}\left\{h\right\}\tpose
-    \end{align*}
+    \end{align}
 % d/dt(1/2 * pudmat)
@@
     \end{align*}
-    Define \(C_2 : \mathcal{Q} \times \mathcal{P} \rightarrow \R^{n(n-k)\times n}\) 
+    Defining \(C_2 : \mathcal{Q} \times \mathcal{P} \rightarrow \R^{n(n-k)\times n}\) 
     by
-    \[
+    \begin{equation}
         C_2(q,p) := \begin{bmatrix}
@@
         \end{bmatrix}
-    \]
-    In matrix form, this becomes:
-    \begin{align*}
-        & \frac{1}{2} \diff{}{t} \left\{ \pudmat \right\} ={} \\
+    \end{equation}
+    allows us to collapse the derivative into:
+    \begin{align}
+        & \frac{1}{2} \diff{}{t} \left\{ \pudmat \right\} ={} 
+        \nonumber \\
         & (\Id{n-k} \otimes p\tpose)\nabla_{q_u}\Minv(q)
         \left(-\frac{1}{2}\pudmat - \nabla_{q_u}V(q) + \simpleB \tau \right) +{}
-        \\
+        \nonumber \\
         & \frac{1}{2}\left(\Id{n-k}\otimes p\tpose\right) C_2(q,p) p
-    \end{align*}
+    \end{align}
 % d/dt(nabla_qu V)
     Finally, we compute the derivative of \(\nabla_{q_u}V(q)\):
-    \begin{align*}
+    \begin{equation}
         \diff{}{t}\left\{\nabla_{q_u}V(q)\right\} = 
@@
         = \partial_{q_u} \partial_q V(q) \Minv(q)p
-    \end{align*}
+    \end{equation}
 % Full ddot(e)
     Putting this all together, we can find the explicit form for \(\ddot{e}\):
-    \begin{align*}
-        \ddot{e} &= \diff{}{t}\left\{dh_q\right\}\Minv(q)p + dh_q C_1(q,p)p -{}\\
+    \begin{align}
+        \ddot{e} &= \diff{}{t}\left\{dh_q\right\}\Minv(q)p + dh_q C_1(q,p)p -{}
+        \nonumber \\
      & \frac{1}{2}dh_q\Minv(q)(\pdmat) + 
-        dh_q\Minv(q)\nabla_qV(q) -{}\\
+        dh_q\Minv(q)\nabla_qV(q) -{} 
+        \nonumber \\
      & \diff{}{t}\left\{dh_{p_u}\right\}
-     \left(\frac{1}{2}\pudmat + \nabla_{q_u}V(q)\right) +{} \\
+     \left(\frac{1}{2}\pudmat + \nabla_{q_u}V(q)\right) +{} 
+     \nonumber \\
      & dh_{p_u}(\Id{n-k}\otimes p\tpose)\nabla_{q_u}\Minv(q)\left(
-     \frac{1}{2}\pudmat + \nabla_{q_u}V(q)\right) -{}\\
+     \frac{1}{2}\pudmat + \nabla_{q_u}V(q)\right) -{}
+     \nonumber \\
      & \frac{1}{2}dh_{p_u}(\Id{n-k}\otimes p\tpose)C_2(q,p)p - 
-     dh_{p_u} \partial_{q_u}\partial_q V(q)\Minv(q)p +{} \\
+     dh_{p_u} \partial_{q_u}\partial_q V(q)\Minv(q)p +{} 
+     \nonumber \\
      & \left(dh_q \Minv(q) - dh_{p_u}(\Id{n-k} \otimes
      p\tpose)\nabla_{q_u}\Minv(q) \right) \simpleB \tau
-    \end{align*}
+    \end{align}
 % End of Proof
@@
         \left.\begin{aligned}
-            \dot{q}_u &= M(q)p \\
+                \dot{q}_u &= \begin{bmatrix}
+                    \Id{(n-k)} & \Zmat{(n-k) \times k}
+                \end{bmatrix}\Minv(q)p \\
             \dot{p}_u &= -\nabla_{q_u}V(q) \\
-            \end{aligned}\right|_{\begin{array}{c}
+            \end{aligned}{}\right|_{\begin{array}{c}
                 h(q,p_u) = 0 \\ 
-                pa = g(q_u,p_u) \\
+                p_a = g(q_u,p_u) \\
             \end{array}}

### 2020-09-18 latex/vlp.tex
Updated some phrasing on the VLP theorem
Updated intro to VLP section

@@
 \cite{pumping_swing_standing_squatting,how_to_pump_a_swing}.
-The VLP also represents the swinging of a crane, the (simplified) motion of a
-gymnast on a bar \cite{pendulum_length_giant_gymnastics}, and the
-tuned-mass-damper systems which stabilize skyscrapers
-\cite{vlp_tuned_mass_damper}.
+The VLP also represents the motion of the load at the end of a crane, 
+the (simplified) motion of a gymnast on a bar
+\cite{pendulum_length_giant_gymnastics}, and the tuned-mass-damper systems which
+stabilize skyscrapers \cite{vlp_tuned_mass_damper}.
 
@@
 to stabilize trajectories of the system. While many of these controllers 
-are time-dependent, \cite{vlp_energy_shaping}
-offers a time-independent technique to inject energy into the system and
-stabilize desired energy level sets. The authors designed a controller through a
-technique called \textit{energy shaping}, and proved that their control
-mechanism would allow the VLP to achieve any desired energy level set.
-However, the energy injection mechanism is
-ad-hoc in the sense that it is not derived from natural behaviour. In
-this chapter we will design VNHCs to inject energy into the VLP in a
+are time-dependent, \citet{vlp_energy_shaping}
+offer a time-independent technique to inject energy into the system. 
+They designed a controller through a technique called \textit{energy
+shaping}, and proved that their control mechanism would allow the VLP to achieve
+any desired energy level set.
+However, the energy injection mechanism is ad-hoc in the sense that it is not
+derived from natural behaviour. 
+In this chapter we will design VNHCs to inject energy into the VLP in a
 time-independent way.
@@
 \begin{proof}
-    Choose, as a candidate anti-Lyapunov function, the energy for the average-length pendulum 
+    Choose, as a candidate anti-Lyapunov function, the energy for the
+    average-length pendulum (with zero potential at the bottom of the swing)
     \[

### 2020-09-22 latex/vlp.tex
Updated intro to VLP chapter

@@
 are time-dependent, \citet{vlp_energy_shaping}
-offer a time-independent technique to inject energy into the system. 
-They designed a controller through a technique called \textit{energy
-shaping}, and proved that their control mechanism would allow the VLP to achieve
-any desired energy level set.
+offer a time-independent technique to inject energy into the VLP. 
+They designed a controller through a technique called \textit{energy shaping},
+and proved that their control mechanism stabilizes any desired energy level set.
 However, the energy injection mechanism is ad-hoc in the sense that it is not
 derived from natural behaviour. 
-In this chapter we will design VNHCs to inject energy into the VLP in a
-time-independent way.
-The difference compared to \cite{vlp_energy_shaping} is that our 
-controller will maintain the structured motion of a human on a swing.
+In this chapter we will design VNHCs to stabilize energy levels of the VLP in a
+time-independent manner, while maintaining the structured motion of a human on a
+swing.
 

### 2020-09-24 latex/vlp.tex
Started editing Chapter 3.3 on the VLP constraint

@@
    \includestandalone[width=0.5\textwidth]{images/vlp_model}
-   \caption{The representation of the variable length pendulum as a mass at the
-      tip of a massless rod.}\label{fig:vlp-model}
+   \caption{The variable length pendulum is a mass attached to the
+      tip of a massless rod which can change length.}\label{fig:vlp-model}
 \end{figure}
@@
 The cartesian position of the mass at the tip of the pendulum
-is given by \(x = (l\sin(q),l\cos(q))\), while its velocity is
-\(\dot{x} = (\dot{l}\sin(q) + l\cos{q}\dot{q}, \dot{l}\cos(q) - l\sin{q}\dot{q})\).
+is given by \(x = (l\sin(q),-l\cos(q))\), while its velocity is
+\(\dot{x} = (\dot{l}\sin(q) + l\cos{q}\dot{q}, -\dot{l}\cos(q) + l\sin{q}\dot{q})\).
 Computing the kinetic energy \(T\) yields
@@
 \[
-   \mathcal{H} = E = \frac{1}{2}\dot{\mathbf{p}}\tpose \Minv(\mathbf{q})
-   \dot{\mathbf{p}}+V(\mathbf{q})
+   \mathcal{H} = E = \frac{1}{2}\mathbf{p}\tpose \Minv(\mathbf{q})
+   \mathbf{p}+V(\mathbf{q})
 \]
@@
 % Go through the development of the constraint and prove it gains energy
-Since we wish to define a VNHC which injects energy into the VLP in a human-like
-manner,
-we return once again to the example of a person standing on a swing.
-As can be seen in Figure \ref{fig:child-vlp}, condensing the person 
-into a center of mass gives a VLP model of their motion, where
-the pendulum shrinks when the person stands and lengthens when they squat. This
-is equivalent to the VLP model from Figure \ref{fig:vlp-model}.
+To motivate why a VNHC could inject energy into the VLP in a human-like
+manner, we will examine the motion of a person standing on a swing.
+As can be seen in Figure \ref{fig:child-vlp}, their center of mass moves closer
+to the swing's pivot when they stand, and moves away from the pivot when they
+squat.
+This is equivalent to the VLP model from Figure \ref{fig:vlp-model}.
 \begin{figure}
@@
    \end{subfigure}
-   \caption{The variable length pendulum representation of a person on a
-   standing swing.}\label{fig:child-vlp}
+   \caption{The VLP representation of a person on a standing swing.}
+   \label{fig:child-vlp}
 \end{figure}
@@
 The action of regulating pendulum length to add energy to the VLP is known as
-``pumping". \citeauthor{pumping_swing_standing_squatting} asked whether
+``pumping". \citet{pumping_swing_standing_squatting} asked whether
 the pumping strategy performed by children is time-optimal, assuming the
-children could squat or stand instantaneously
-\cite{pumping_swing_standing_squatting}. Indeed, they discovered that 
-a child's pumping strategy will inject energy into the VLP as fast as is 
-physically possible.
+children could squat or stand instantaneously. 
+Indeed, they discovered that children increase the height of their swing as fast
+as is physically possible.
 
-This pumping strategy is straightforward: the child stands up at the lowest point of
-their swing, and squats at the highest point. Looking at their VLP
-representation, the pendulum shortens at the bottom of the swing, and lengthens
-at the top. For an intuitive understanding, conservation of angular momentum
-indicates that shortening the pendulum at the bottom forces the mass to gain
-speed to compensate for the reduced length
-\cite{how_to_pump_a_swing}.
-Energy is not conserved in this process, and the pendulum gains kinetic energy
-which allows the pendulum to reach a higher point in its swing.
-Lengthening the pendulum when it reaches this point means gravity
+This optimal pumping strategy is straightforward:
+one stands at the lowest point of the swing, and squats at the highest point.
+Looking at the VLP representation, the pendulum shortens at the bottom of the
+swing, and lengthens at the top. 
+For an intuitive explanation, conservation of angular momentum indicates that
+shortening the pendulum at the bottom forces the mass to gain speed to
+compensate for the reduced length \cite{how_to_pump_a_swing}.
+Energy is not conserved in this process, so the pendulum gains kinetic energy
+and reaches a higher point at the peak of its swing.
+Lengthening the pendulum when it reaches this peak means gravity
 imparts a larger angular momentum to the mass by the time it reaches the bottom
-of its swing, which is converted into a higher velocity the next time the
+of its swing, which in turn is converted to a higher velocity when the
 pendulum is shortened.
 By alternating these processes, the pendulum experiences a net gain in
-rotational energy which is generated entirely by gravity.
+rotational energy on average.
 

### 2020-09-24 latex/vlp.tex
Finished the first edit of VLP

@@
 To motivate why a VNHC could inject energy into the VLP in a human-like
-manner, we will examine the motion of a person standing on a swing.
+manner, we will examine a person standing on a swing.
 As can be seen in Figure \ref{fig:child-vlp}, their center of mass moves closer
@@
 squat.
-This is equivalent to the VLP model from Figure \ref{fig:vlp-model}.
+This is equivalent to the VLP model from Figure \ref{fig:vlp-model}, where
+standing and squatting correspond to shortening and lengthening the pendulum
+(respectively).
 \begin{figure}
@@
 
-This optimal pumping strategy is straightforward:
-one stands at the lowest point of the swing, and squats at the highest point.
+A child's optimal pumping strategy is straightforward:
+they stand at the lowest point of the swing, and squat at the highest point.
 Looking at the VLP representation, the pendulum shortens at the bottom of the
@@
 pendulum is shortened.
-By alternating these processes, the pendulum experiences a net gain in
-rotational energy on average.
+By alternating these processes, the pendulum experiences an average net gain in
+rotational energy.
 
-The time-optimal controller described by \cite{pumping_swing_standing_squatting}
-depends on a trajectory in time. We will convert this 
-into a time-independent controller derived from the time-optimal
-strategy. Notice that the optimal controller depends on knowing when the
-system is at the ``bottom" or ``top" of the swing. Since being at the ``bottom"
-is the same as saying the pendulum angle is \(q = 0\) and being at the ``top" is
-equivalent to the pendulum's momentum being \(p = 0\), the time-independent
-controller will necessarily use the full phase \((q,p)\). While VHCs cannot
-satisfy this requirement, our method of VNHCs is especially well suited to this
-problem.
+Notice that the child's pumping strategy requires knowledge of
+when the system is at the ``bottom" or ``top" of the swing. 
+Since being at the ``bottom" is equivalent to having angle \(q = 0\) and being
+at the ``top" is equivalent to having momentum \(p = 0\), a controller based on
+this strategy will necessarily use the full phase \((q,p)\). 
+This is why we will use VNHCs instead of VHCs.
 
-Figure \ref{fig:vlp-optimal-controller-qp-plane} shows the time-optimal strategy directly
-on the \((q,p)\)-plane.
-Since the VLP must stay the same length inside each quadrant and
-changes length only when it crosses one of the axes, we
-will use the angle \(\theta = \arctan_2(p,q)\) in the \((q,p)\)-plane to define
-the optimal controller \(l^\star(\theta)\) (\ref{eqn:vlp-optimal-controller}).
-Figure \ref{fig:vlp-optimal-controller} shows this time-optimal controller as a function of
-\(\theta\).
+The time-optimal controller from \cite{pumping_swing_standing_squatting} is (in
+our notation)
+\[
+   l_\text{opt}(q,p) := -\sign{qp}
+\]
+which is a piecewise-continuous controller that varies between \(\pm 1\). 
+This looks similar to a VNHC, but it is not smooth and makes the assumption that 
+\(l \in [-1,1]\). 
+Since we want our controller emulate realistic human motion, we need to derive a
+smooth controller where we allow \(l \in [\underbar{l},\overbar{l}]\); then, we
+will prove that it injects energy.
+
+Figure \ref{fig:vlp-optimal-controller-qp-plane} displays \(l_\text{opt}\) 
+directly on the \((q,p)\)-plane.
+Because the length remains constant inside each quadrant and changes only when it
+crosses one of the axes, one can use the angle \(\theta = \arctan_2(p,q)\) to
+redefine the time-optimal controller in a way which is easier to ``smooth out".
+Figure \ref{fig:vlp-optimal-controller} shows the time-optimal pumping strategy
+as a function \(l^\star(\theta)\) (\ref{eqn:vlp-optimal-controller}), where now
+the length varies between \([\underbar{l},\overbar{l}]\) rather than 
+\([-1,1]\).
+
 \begin{equation}\label{eqn:vlp-optimal-controller}
    l^\star(\theta):= \begin{cases}
+      \overline{l} & \theta \in [-\frac{\pi}{2},0[ \cup [\frac{\pi}{2}, \pi[ \\
       \underline{l} & \theta \in [-pi, -\frac{\pi}{2}[ \cup [0,\frac{\pi}{2}[ \\
-      \overline{l} & \theta \in [-\frac{\pi}{2},0[ \cup [\frac{\pi}{2}, \pi[ \\
    \end{cases}
@@
       \includestandalone[width=\linewidth]{images/vlp_optimal_controller_qp_plane}
-      \caption{The time-optimal controller mapped to the \((q,p)\)-plane.}
+      \caption{The time-optimal controller \(l_\text{opt}\) mapped onto the \((q,p)\)-plane.}
       \label{fig:vlp-optimal-controller-qp-plane}
@@
       \includestandalone[width=\linewidth]{images/vlp_optimal_controller}
-      \caption{The time-optimal controller converted to a VNHC-style 
-      function \(l^\star(\theta)\), where \(\theta := \arctan_2(p,q)\).}\label{fig:vlp-optimal-controller}
+      \caption{The time-optimal controller converted to the alternate
+         representation \(l^\star(\theta)\).}
+      \label{fig:vlp-optimal-controller}
    \end{subfigure}
@@
       \cite{pumping_swing_standing_squatting}. The colour red
-      corresponds to standing and blue to squatting.}
+      corresponds to standing, blue to squatting, and 
+      \(\theta := \arctan_2(p,q)\) is the angle of the VLP phase in the
+      \((q,p)\)-plane.}
 \end{figure}
 
-To find a continuously differentiable function which approximates this controller,
-we can attach sinusoids at the transition points of
-\(l^\star(\theta)\) (see Figure \ref{fig:vlp-T-controller}). Let
-\(\Delta l := (\overline{l} - \underline{l})/2\) and 
+We now define a continuous function which
+approximates \(l^\star(\theta)\) by attaching sinusoids at the
+discontinuities (see Figure \ref{fig:vlp-T-controller}). 
+Let \(\Delta l := (\overline{l} - \underline{l})/2\) and 
 \(l_{\text{avg}} := (\overline{l} + \underline{l})/2\).
-Providing these newly attached sinusoids with a frequency \(\omega = \frac{pi}{T}\) 
-(where \(T \in ]0,\frac{\pi}{2}]\) is a control parameter), we have a new
-controller \(l_T(\theta)\):
+Providing these newly attached sinusoids with a frequency \(\omega = \frac{\pi}{T}\) 
+(where \(T \in \, ]0,\frac{\pi}{2}]\) is a control parameter), we get the
+twice-differentiable controller \(l_T(\theta)\) (\ref{eqn:vlp-T-controller}).
 \begin{equation}\label{eqn:vlp-T-controller}
    l_T(\theta) = \begin{cases}
+      \overline{l} & \theta \in \left[-\frac{\pi}{2} + \frac{T}{2}, -\frac{T}{2}\right] 
+      \cup \left[\frac{\pi}{2} + \frac{T}{2}, \pi - \frac{T}{2}\right] \\
       \underline{l} & \theta \in \left[-\pi + \frac{T}{2}, -\frac{\pi}{2} - \frac{T}{2}\right] 
       \cup \left[\frac{T}{2}, \frac{\pi}{2} - \frac{T}{2}\right] \\
-      \overline{l} & \theta \in \left[-\frac{\pi}{2} + \frac{T}{2}, -\frac{T}{2}\right] 
-      \cup \left[\frac{\pi}{2} + \frac{T}{2}, \pi - \frac{T}{2}\right] \\
       -\Delta l \sin(\omega(\theta + \pi)) + l_{\text{avg}} & \theta \in
@@
 
-This controller approximates the optimal controller, since 
+This controller approximates the optimal controller because 
 \[
@@
 \]
-Unfortunately, while \(l_T(\theta)\) is continuously differentiable, it is not
-twice continuously differentiable for most values of \(T\).
+Unfortunately, while \(l_T(\theta)\) is twice-differentiable, its second
+derivative is only piecewise continuous for most values of \(T\).
 If we wish to use it as a VNHC, we must ensure that either the generalized
-forces \(\tau\) acting on \(p_l\) can be discontinuous, or we must find a value
-of \(T\) where this controller is smooth.
-Thankfully, \(l_{\frac{\pi}{2}}(\theta)\) is smooth; it can be simplified into
-the expression (\ref{eqn:vlp-smoothed-controller}), and is plotted for
-demonstration in Figure \ref{fig:vlp-smoothed-controller}.
+forces \(\tau\) acting on \(p_l\) can be discontinuous (which is certainly
+achievable by humans), or we must find a value of \(T\) where this controller is
+smooth.
+Thankfully, \(l_{\frac{\pi}{2}}(\theta)\) is smooth as it can be simplified into
+the expression (\ref{eqn:vlp-smoothed-controller}). 
+It is plotted for demonstration in Figure \ref{fig:vlp-smoothed-controller}.
 \begin{equation}\label{eqn:vlp-smoothed-controller}
@@
    \centering
-   \includestandalone[width=0.8\textwidth]{images/vlp_smoothed_controller}
+   \includestandalone[width=0.6\textwidth]{images/vlp_smoothed_controller}
    \caption{The smoothed VLP controller which corresponds to the VNHC \(l =
@@
 
-We therefore set our VNHC to be
+Because \(l_\frac{\pi}{2}(\theta)\) is smooth and it approximates
+\(l^\star(\theta)\), we set our VNHC to be
 \[
@@
 \]
-
-This VNHC is not the same as the optimal controller
-(\ref{eqn:vlp-optimal-controller}), so we need to prove it will still gain
-energy. We will do so by showing the dynamics (\ref{eqn:vlp-hamiltonian}) of the
-VLP diverge from the origin, regardless of the initial condition of the system. 
-To do this, we will use the notion of an anti-Lyapunov function.
+We still need to prove this VNHC will inject energy into a VLP. 
+We will do so by showing the dynamics (\ref{eqn:vlp-hamiltonian}) 
+diverge from the origin via an ``anti-Lyapunov function".
 
@@
 \end{defn}
+\textbf{Manfredi: Does this notion already exist elsewhere? I couldn't find any
+sources on it, but it seems like it should've been done before.}
 
@@
 \(V\) is simply a Lyapunov function for the negative-time system 
-\(\dot{x} = -f(x)\). In other words, since \(V\) proves that the orbit \(x(-t)\) 
+\(\dot{x} = -f(x)\). In other words, \(V\) proves that the orbit \(x(-t)\) 
 of the system travelling backwards in time is asymptotically converging to the
-equilibrium, the system must be diverging from the equilibrium in forward time.
-If we can find such a function for the VLP's dynamics under our VNHC, the path
-generated by the VNHC in the \((q,p)\)-plane must be diverging from the origin
-so long as the initial condition is not an equilibrium.
-This in turn tells us the magnitude of the momentum \(p\) is increasing every
-time the path hits the \(p\)-axis, which means the VLP is, on average,
-increasing in energy.
+origin; hence, the system must be diverging from the origin in forward time.
 
-We will show that anti-Lyapunov function exists for our VNHC.
-In order to do so, we require the following lemma.
+From Theorem \ref{thm:zero-dynamics}, the dynamics on the constraint manifold of
+the VNHC are given by \((\dot{q},\dot{p})\) with 
+\(l = l_\frac{\pi}{2}(\theta)\).
+Supposing the initial condition is not an equilibrium, the existence of an
+anti-Lyapunov function would indicate that the curve traced out on the
+\((q,p)\)-plane is diverging from the origin.
+This implies that the momentum \(p\) is increasing in magnitude whenever the
+curve hits the \(p\)-axis, which in turn means the VLP is gaining energy on
+average.
+
+Proving the existence of an anti-Lyapunov function for the VLP
+requires the following lemma.
 
@@
    \]
-   Since \( (x-y)^2 = x^2 - 2xy + y^2) \geq 0 \), 
+   The inequality \(x^2 + xy + y^2 \geq 0\) holds because
    \[
-      \frac{x^2 + y^2}{2} \geq xy
+      x^2 + xy + y^2 = \left(x  + \frac{y}{2}\right)^2 + \frac{3y^2}{4} \geq 0
    \]
-   which means \(x^2 + xy + y^2 \geq 0\), proving the lemma.
+   which proves the lemma.
 \end{proof}
@@
 \begin{thm}\label{thm:vlp-energy-stabilization}
-   Take the VLP with Hamiltonian dyanmics (\ref{eqn:vlp-hamiltonian}) and define
-   \(\theta := \arctan_2(p,q)\). Suppose the initial condition is not one of the
-   equilibria: \((q(0),p(0)) \not \in \left\{(0,0), (\pm\pi,0)\right\}\).
+   Take the Hamiltonian VLP (\ref{eqn:vlp-hamiltonian}) with initial condition 
+   \((q_0,p_0) \not \in \left\{(0,0),(\pm\pi,0)\right\}\).
+   Define \(\theta := \arctan_2(p,q)\). 
    A VNHC of the form \(h(q,p) = l - l(\theta)\) injects energy if there exists 
@@
    with the property that the inequality is strict for almost every \(\theta\).
-   Flipping the inequality of (\ref{eqn:vlp-energy-gain-condition}) leads to
-   energy dissipation.
+   If the inequality is flipped, the VNHC dissipates energy. 
 \end{thm}
@@
     Choose, as a candidate anti-Lyapunov function, the energy for the
-    average-length pendulum (with zero potential at the bottom of the swing)
+    average-length pendulum with zero-potential at the bottom of the swing:
     \[
@@
     \]
-    which is positive definite at \((0,0)\) and has derivative 
+    This is positive definite at \((0,0)\) and has derivative 
     \[
@@
     \]
-    Observe that \(\sign{\sin(q)p} = \sign{\sin(2\theta)}\) and, 
-    by Lemma \ref{lemma:sign-of-cube},
+    The sign of \(\sin(q)p\) is dependent only on the quadrant in the
+    \((q,p)\)-plane, so it is equal to the sign of \(\sin(2\theta)\). 
+    Furthermore, by Lemma \ref{lemma:sign-of-cube}:
     \[ 
@@
     \]
-    Then the derivative of \(E_\text{avg}\) is almost always positive, since
+    Thus, the derivative of \(E_\text{avg}\) is almost always positive since
     \begin{align*}
@@
     \end{align*}
-    Hence, \(E_\text{avg}\) is an anti-Lyapunov function.
+    Therefore \(E_\text{avg}\) is an anti-Lyapunov function.
     Since the initial condition is not an equilibrium,
@@
     Flipping the inequality of (\ref{eqn:vlp-energy-gain-condition}) 
-    means \(E_\text{avg}\) is a regular Lyapunov function, since its derivative
-    will be negative almost everywhere.
+    means \(E_\text{avg}\) is a regular Lyapunov function, so the VLP is
+    losing energy while converging to \((0,0)\).
  \end{proof}
@@
 \begin{cor}
+   Recall that
+   \[
+      l_\frac{\pi}{2}(\theta) = -\Delta l \sin(2\theta) + l_\text{avg}
+   \]
+   and define 
+   \[
+      l^{-}_\frac{\pi}{2}(\theta) := \Delta l \sin(2\theta) + l_\text{avg}
+   \]
    The VNHC \(l = l_\frac{\pi}{2}(\theta)\) injects energy into the VLP
@@
       \left(l_\frac{\pi}{2}(\theta) - l_\text{avg}\right)\sin(2\theta) = 
-   -\Delta l \sin^2(2\theta) \geq 0
+      -\Delta l \sin^2(2\theta) \leq 0
    \] 
-   Likewise, let 
-   \(l\inv_\frac{\pi}{2}(\theta) := l_\text{avg} - \Delta l \sin(2\theta)\).
-    The VNHC \(h(q,p) = l - l\inv_\frac{\pi}{2}(\theta)\) which flips 
-    \(l_\frac{\pi}{2}(\theta)\) about \(l_\text{avg}\) will dissipate energy from
-   the VLP, since 
+   The VNHC \(l = l^{-}_\frac{\pi}{2}(\theta)\) dissipates energy
+   from the VLP, since 
    \[
-      \left(l\inv_\frac{\pi}{2}(\theta) - l_\text{avg}\right)\sin(2\theta) = 
-      -\Delta l \sin^2(2\theta) \leq 0
+      \left(l^{-}_\frac{\pi}{2}(\theta) - l_\text{avg}\right)\sin(2\theta) = 
+      \Delta l \sin^2(2\theta) \geq 0
    \] 
@@
 
-The types of VNHCs which satisfy Theorem \ref{thm:vlp-energy-stabilization} are illustrated by Figure
-\ref{fig:vlp-energy-in-out}. To stabilize specific energy level sets, one simple
-approach is to switch between injection and dissipation VNHCs depending on the
-current energy level.
-For the VNHCs we designed in this section, this means toggling
-between \(l_\frac{\pi}{2}(\theta)\) and \(l\inv_\frac{\pi}{2}(\theta)\),
-possibly with some hysteresis to avoid infinite switching.
+The types of VNHCs which satisfy Theorem \ref{thm:vlp-energy-stabilization} are
+illustrated by Figure \ref{fig:vlp-energy-in-out}. 
+To stabilize specific energy level sets, a basic approach is to switch
+between injection and dissipation VNHCs depending on the current energy level.
+For the VNHCs we designed in this chapter, this means toggling
+between \(l_\frac{\pi}{2}(\theta)\) and \(l^{-}_\frac{\pi}{2}(\theta)\),
+with some hysteresis to avoid infinite switching.
 
@@
 
-Let us define \((l^\star)\inv(\theta)\) by swapping the order of 
+Let us define \((l^\star)^{-}(\theta)\) by swapping the order of 
 \(\underline{l}\) and \(\overline{l}\) in \(l^\star(\theta)\). Since this 
@@
 \(l(\theta) \in [\underline{l},\overline{l}]\), we can predict that 
-\((l^\star)\inv(\theta)\) is the VNHC representation of an optimal 
+\((l^\star)^{-}(\theta)\) is the VNHC representation of an optimal 
 energy dissipation controller.
@@
 point (instead of standing and squatting resp.) produces the
-time-optimal trajectory for stopping a standing swing.
+time-optimal trajectory for \textit{stopping} a standing swing.
 
-All together, these results show that VNHCs can replicate the time-optimal
-strategy performed by humans in a time-independent manner. Furthermore, we see
-that VNHCs are a powerful tool for creating simple energy stabilization
-techniques based on natural human motion.
+All together, these theory in this chapter shows that VNHCs can replicate the
+time-optimal pumping strategy performed by humans. 
+Furthermore, we see that VNHCs are a powerful tool for creating simple energy
+stabilization techniques based on natural human motion.
 

### 2020-09-25 latex/vlp.tex
Minor edits to VNHC and VLP to send to Manfredi

@@
 is given by \(x = (l\sin(q),-l\cos(q))\), while its velocity is
-\(\dot{x} = (\dot{l}\sin(q) + l\cos{q}\dot{q}, -\dot{l}\cos(q) + l\sin{q}\dot{q})\).
+\(\dot{x} = (\dot{l}\sin(q) + l\cos(q)\dot{q}, -\dot{l}\cos(q) + l\sin(q)\dot{q})\).
 Computing the kinetic energy \(T\) yields
@@
 This is extremely useful for simplifying the dynamics: since one can choose
-\(\tau\) which sets \(\dot{p}_l\) to any desired function, we can assume (with abuse
-of notation) that \(l\) is tracking some known function \(l(t)\). 
+\(\tau\) which sets \(\dot{p}_l\) to any desired function and \(\dot{l}\)
+depends only on \(p_l\), we can assume (with abuse of notation) that \(l\) is
+tracking some known function \(l(t)\). 
 The closed loop dynamics of the system will be described exclusively by
@@
 
-A child's optimal pumping strategy is straightforward:
+A child's optimal pumping strategy is the following: 
 they stand at the lowest point of the swing, and squat at the highest point.
@@
 (where \(T \in \, ]0,\frac{\pi}{2}]\) is a control parameter), we get the
-twice-differentiable controller \(l_T(\theta)\) (\ref{eqn:vlp-T-controller}).
+continuously-differentiable controller \(l_T(\theta)\) (\ref{eqn:vlp-T-controller}).
 \begin{equation}\label{eqn:vlp-T-controller}
@@
 \]
-Unfortunately, while \(l_T(\theta)\) is twice-differentiable, its second
-derivative is only piecewise continuous for most values of \(T\).
+Unfortunately, while \(l_T(\theta)\) is continuously-differentiable, 
+it is not twice-differentiable for most values of \(T\).
 If we wish to use it as a VNHC, we must ensure that either the generalized
@@
 the expression (\ref{eqn:vlp-smoothed-controller}). 
-It is plotted for demonstration in Figure \ref{fig:vlp-smoothed-controller}.
 \begin{equation}\label{eqn:vlp-smoothed-controller}
@@
 \end{equation}
+It is plotted for demonstration in Figure \ref{fig:vlp-smoothed-controller}.
 
@@
 We will do so by showing the dynamics (\ref{eqn:vlp-hamiltonian}) 
-diverge from the origin via an ``anti-Lyapunov function".
+diverge from the origin via a so-called anti-Lyapunov function.
 
@@
 \begin{thm}\label{thm:vlp-energy-stabilization}
-   Take the Hamiltonian VLP (\ref{eqn:vlp-hamiltonian}) with initial condition 
+   Take the VLP (\ref{eqn:vlp-hamiltonian}) with initial condition 
    \((q_0,p_0) \not \in \left\{(0,0),(\pm\pi,0)\right\}\).
@@
 
-All together, these theory in this chapter shows that VNHCs can replicate the
-time-optimal pumping strategy performed by humans. 
+All together, the theory developed in this chapter shows that VNHCs can
+replicate the time-optimal pumping strategy performed by humans. 
 Furthermore, we see that VNHCs are a powerful tool for creating simple energy

### 2020-09-25 latex/vnhcs.tex
Added a question for Manfredi
Minor edits to VNHC and VLP to send to Manfredi

@@
    \end{subfigure}
-   \caption{A mechanical system with \(N = 2\) points at \((x_1,y_1,z_1)\)
-   and \((x_2,y_2,z_2)\) separated by a bar with \(r = 3\) EOC given by 
-   \(c = \norm{x_1 - x_2}\), \(z_1 = 0\), and \(z_2 = 0\). This system has \(n =
-   3\) degrees of freedom.}
+   \caption{A mechanical system with \(N = 2\) point masses at \(a = (x_1,y_1,z_1)\)
+       and \(b = (x_2,y_2,z_2)\), separated by a bar of length \(c\). 
+       There are \(r = 3\) EOC given by \(\norm{a - b} = c\), \(z_1 = 0\),
+       and \(z_2 = 0\). This system has \(n = 3\) degrees of freedom.}
    \label{fig:barbell}
@@
 \end{defn}
+\textbf{Manfredi: Do we require \(C^\infty\), or can we relax it to \(C^2\)?}
 
@@
     everywhere on the constraint manifold \(\Gamma\).
-    That is, a regular VNHC is a function 
-    \(h : \mathcal{Q} \times \mathcal{P}_u \rightarrow \R^k\) which satisfies
-    the rank condition.
 \end{thm}
@@
     where \(f\) is a suitably-defined smooth function.
-    Then \(dh_q = \begin{bmatrix} -\partial f_{q_u} & \Id{(n-1)}
+    Then \(dh_q = \begin{bmatrix} -\partial_{q_u}f & \Id{(n-1)}
     \end{bmatrix}\).

### 2020-09-28 latex/acrobot.tex
Finished writing the first version of the acrobot dynamics

@@
     \includestandalone[width=0.7\textwidth]{images/acrobot_model}
-    \caption{The acrobot model.}%
+    \caption{The general acrobot model, represented by two weighted rods
+    differing in both length and mass.}%
     \label{fig:acrobot-model}
@@
 \section{Dynamics of the Acrobot}
-% TODO: Derive the acrobot dynamics in Hamiltonian form
-For the acrobot model displayed in Figure \ref{fig:acrobot-model}, 
-the COM of each link lies somewhere along the link itself. 
+Suppose we are given an acrobot as in Figure \ref{fig:acrobot-model} modelling a
+gymnast hanging on a horizontal bar, where the ``torso" has moment of
+inertia \(J_u\) and the ``leg" link has moment of inertia \(J_a\) (each with respect
+to their own center of mass).
+Let \(q_u \in \mathbb{S}^1\) be the shoulder angle and \(q_a \in \mathbb{S}^1\) 
+be the hip angle, where only \(q_a\) is actuated. 
+Collecting them together provides the configuration
+\(q = (q_u,q_a) \in \mathbb{S}^1 \times \mathbb{S}^1\). 
+The acrobot has inertia matrix \(D\), potential function \(P\) (with respect to
+the horizontal bar), and input matrix \(B\)  given as
+follows \cite{xingbo_thesis}:
+\begin{align}\label{eqn:general-acrobot-inertia}
+    D(q) &= \begin{bmatrix}
+      m_al_u^2 + 2m_a\cos(q_a)l_u l_{c_a} + m_al_{c_a}^2 + m_ul_{c_u}^2 + J_u + J_a &
+      m_al_{c_a}^2 + m_al_ul_{c_a}\cos(q_a) + J_a \\
+      m_al_{c_a}^2 + m_al_ul_{c_a}\cos(q_a) + J_a &
+      m_al_{c_a}^2 + J_a
+    \end{bmatrix} \\
+    \label{eqn:general-acrobot-potential}
+    P(q) &= g\left(m_al_{c_a}(1 - \cos(q_u+q_a)) + 
+        (m_al_u + m_ul_{c_u})(1-\cos(q_u))\right) \\
+    B(q) &= \begin{bmatrix} 0 \\ 1 \end{bmatrix}
+\end{align}
+
 While this is the most general representation of an acrobot, the dynamics are
-somewhat complicated and are difficult to use when proving the results in the 
-rest of this chapter. For simplicity, we will
-assume the acrobot is comprised of two massless rods of equal length with 
-equal masses at the tips, as in Figure \ref{fig:simple-acrobot-model}; we call
-this a \textbf{simple} acrobot.
+unwieldy to the point where it becomes difficult to prove the results in this
+chapter analytically.
+We will therefore simplify the dynamics by assuming the acrobot is comprised of
+two massless rods of equal length \(l\), with equal point masses \(m\) at the
+tips (as in Figure \ref{fig:simple-acrobot-model}).
+We call this type of acrobot a \textbf{simple} acrobot. We will also ignore any
+frictional forces at both the hips and shoulder. Finally, it is important to note
+that a real gymnast cannot swing their hips in full circles; for this reason, we
+assume that \(q_a \in [-\bar{q}_a, \bar{q}_a]\) where 
+\(\bar{q}_a \in ]-\pi, \pi[\). On mechanical acrobots, this can be taken as a
+control parameter.
 
@@
     \centering
-    \includestandalone[width=0.5\textwidth]{images/simple_acrobot_model}
-    \caption{A simple acrobot with lengths \(l\) and masses of size \(m\).}
+    \includestandalone[width=0.7\textwidth]{images/simple_acrobot_model}
+    \caption{A simple acrobot has massless rods of equal length \(l\) and 
+    equal masses \(m\) at the tips.}
     \label{fig:simple-acrobot-model}
@@
 
-\section{Previous Approaches}
+Given the assumptions listed, the moments of inertia \(J_u\) and \(J_a\) of the
+rods vanish. Furthermore, \(l_{c_u} = l_{c_a} = l_u = l_a = l\) and
+\(m_u = m_a = m\). Reducing
+(\ref{eqn:general-acrobot-inertia}-\ref{eqn:general-acrobot-potential}) with this
+knowledge yields the following ``simple" inertia matrix and potential function:
+\begin{align}
+    D(q) &= \begin{bmatrix}
+        ml^2\left(3+2\cos(q_a)\right) & 
+        ml^2\left(1+\cos(q_a)\right) \\
+        ml^2\left(1+\cos(q_a)\right) &
+        ml^2
+    \end{bmatrix} \\
+    P(q) &= -mgl\left(2\cos(q_u)+\cos(q_u+q_a)\right)
+\end{align}
+\begin{notation}
+    For shorthand, we write \(c_u := \cos(q_u)\), \(c_a := \cos(q_a)\), and 
+    \(c_{ua} := \cos(q_u + q_a)\). Likewise, \(s_u := \sin(q_u)\), 
+    \(s_a := \sin(q_a)\), and \(s_{ua} := \sin(q_u + q_a)\).
+\end{notation}
+
+Defining \(M(q) := D(q)\) and \(V(q) := P(q)\), we find the conjugate of momenta 
+is \(p = (p_u, p_a) = M(q)\dot{q}\).
+The dynamics in \((q,p)\) coordinates are given by
+\begin{align}\label{eqn:acrobot-hamiltonian}
+    \mathcal{H}(q,p) &= \frac{1}{2}p\tpose \Minv(q) p -
+    mgl\left(2 c_u + c_{ua}\right)\\
+     &\begin{cases}
+        \dot{q} = \Minv(q) p \\
+        \dot{p}_u = -mgl\left(2s_u + s_{ua}\right) \\
+        \dot{p}_a =-\frac{1}{2}p\tpose \nabla_{q_a}\Minv(q) p
+        - mgl s_{ua} + \tau
+    \end{cases} \nonumber
+\end{align}
+where the inverse inertia matrix is
+\begin{equation}\label{eqn:Minv}
+    \Minv(q) = \frac{1}{ml^2\left(2-c_a^2\right)}
+    \begin{bmatrix}
+        1 &
+        -\left(1+c_a\right) \\
+        -\left(1+c_a\right) &
+        3+2c_a
+    \end{bmatrix}
+\end{equation}
+
+\section{Previous Constraint Approaches}
 % TODO: Give Xingbo's intuition, abstract away to qa = sin(theta) constraint
@@
 
-%\section{Energy Injection for the Acrobot}
-%Suppose the acrobot is simple; 
-%that is, assume that the acrobot's two links are of equal length \(l\)
-%and of equal mass \(m\), with the mass concentrated at the tip of each rod.
-%
-%Let \(q = (q_u,q_a)\) where \(q_u\) is the shoulder angle and \(q_a\) is the hip angle.
-%
-%Then the acrobot has hamiltonian given by
-%\[
-%    \mathcal{H}(q,p) = \frac{1}{2} p^T M^{-1}(q) p + V(q)
-%\]
-%where
-%\[
-%    M^{-1}(q) = \frac{1}{m l^2 (2 - \cos^2(q_a))} 
-%    \begin{bmatrix}
-%        1              & -(1 + \cos(q_a)) \\
-%        -(1 + \cos(q_a)) & (3 + 2\cos(q_a))
-%    \end{bmatrix}
-%\]
-%and
-%\[
-%    V(q) = m g l (3 - 2 \cos(q_u) - \cos(q_u + q_a))
-%\]
-%
-%Notice that the acrobot satisfies \(\partial M^{-1}(q)/ \partial q_u = 0\),
-%which means it may be possible to enforce a regular solvable VNHC to inject energy
-%into the acrobot.
-%One such VNHC is
-%\begin{align}\label{eqn:acrobot_vnhc}
-%    \begin{split}
-%    h(q,p) &= q_a - f(q_u,p_u) \\
-%           &= q_a - \bar{q_a}\sin(\arctan(p_u,q_u)) \\
-%           &= q_a - \frac{\bar{q_a}p_u}{\sqrt{q_u^2 + p_u^2}}
-%    \end{split}
-%\end{align}
-%where \(\bar{q_a} \in ]0,\pi[\) is a parameter which dictates how much the 
-%controlled link can swing back and forth.
-%
-%To compute the regularity and solvability conditions, first observe that since
-%the acrobot has two coordinates and only \(p_a\) is actuatable, the control matrix
-%\(B = [0, 1]^T\) is constant and full rank everywhere. This also implies that
-%\(h\) is both regular and solvable if \(dh_q M^{-1}(q)B\) is full rank.
-%
-%Observe that
-%\begin{align}
-%    \begin{split}
-%        dh_q M^{-1}(q)B &= \frac{1}{m l^2 (2 - \cos^2(q_a))}
-%\left[ \frac{\bar{q_a}p_uq_u}{(q_u^2 + p_u^2)^{3/2}}, 1\right] 
-%    \begin{bmatrix}
-%        -(1 + \cos(q_a)) \\
-%        (3 + 2\cos(q_a))
-%    \end{bmatrix} \\
-%                     &= \frac{ (3+2\cos(q_a))(q_u^2 + p_u^2)^{3/2} 
-%                     - \bar{q_a}p_uq_u(1+cos(q_a))}
-%                     {m l^2 (q_u^2+p_u^2)^{3/2} (2 - \cos^2(q_a))}
-%    \end{split}
-%\end{align}
-%
-%TODO: \textbf{Determine if this is full rank everywhere except at \((q_u,p_u) = (0,0)\)}
-%
-%Solving for \(p_a\) is straightforward. The computation is shown below.
-%\begin{multline*}
-%    p_a = \left(dh_q M^{-1}(q)B\right)^{-1}  \\
-%        \left(\frac{m g l \bar{q_a}p_u q_u^2}{(q_u^2+p_u^2)^{3/2}} 
-%            \left(2\sin(q_u) + sin(q_u + q_a) \right)
-%        +
-%            \frac{1}{m l^2 (2 - \cos^2(q_a))}
-%            [ \frac{\bar{q_a}p_uq_u}{(q_u^2+p_u^2)^{3/2}}, 1] 
-%            \begin{bmatrix}
-%              1 \\
-%              -(1 + \cos(q_a))
-%          \end{bmatrix} \right)
-%\end{multline*}
-%\begin{align*}
-%    &= -\frac
-%    {
-%        \frac{ m g l \bar{q_a}q_u^2(2\sin(q_u) + \sin(q_u+q_a))}{(q_u^2+p_u^2)^{3/2}}
-%        + \frac{\bar{q_a} p_u^2 q_u}{m l^2 (q_u^2 + p_u^2)^{3/2} (2 -\cos^2(q_a))}
-%        - \frac{(1+\cos(q_a)) p_u}{m l^2 (2 - \cos^2(q_a))}
-%    }
-%    {
-%        \frac
-%        {
-%            (3 + 2\cos(q_a))(q_u^2+p_u^2)^{3/2} - \bar{q_a}(1 + \cos(q_a))q_up_u
-%        }
-%        {
-%            m l^2 (q_u^2+p_u^2)^{3/2} (2 - \cos^2(q_a))
-%        }
-%    } \\
-%    &= \frac
-%    {
-%        \frac
-%        {
-%            -m^2 g l^3 \bar{q_a}q_u^2(2\sin(q_u) + \sin(q_u+q_a))(2 - \cos^2(q_a))
-%            - \bar{q_a}p_u^2 q_u 
-%            + (1 + \cos(q_a))(q_u^2 + p_u^2)^{3/2} p_u
-%        }
-%        {
-%            m l^2 (q_u^2+p_u^2)^{3/2} (2 - \cos^2(q_a))
-%        }
-%    }
-%    {
-%        \frac
-%        {
-%            (3 + 2\cos(q_a))(q_u^2+p_u^2)^{3/2} - \bar{q_a}(1 + \cos(q_a))q_up_u
-%        }
-%        {
-%            m l^2 (q_u^2+p_u^2)^{3/2} (2 - \cos^2(q_a))
-%        }
-%    }
-%\end{align*}
-%
-%\begin{equation*}
-%    \Rightarrow p_a = \frac
-%        {
-%            (1 + \cos(q_a))(q_u^2 + p_u^2)^{3/2} p_u
-%            -m^2 g l^3 \bar{q_a}q_u^2(2\sin(q_u) + \sin(q_u+q_a))(2 - \cos^2(q_a))
-%            - \bar{q_a}q_up_u^2
-%        }
-%        {
-%            (3 + 2\cos(q_a))(q_u^2+p_u^2)^{3/2} - \bar{q_a}(1 + \cos(q_a))q_up_u
-%        }
-%\end{equation*}
-%
-%With \(p_a\) in hand, it is now possible to solve for the zero dynamics under the constraint. To simplify the notation we define \(s_u := \sin(q_u)\),
-%\(s_{ua} := \sin(q_u + q_a)\), 
-%and \(c_a := \cos(q_a)\).
-%\begin{align*}
-%    \dot{q_u} &= \frac{1}{m l^2 (2 - c_a^2)} [ 1, -(1+c_a)]
-%    \begin{bmatrix} p_u \\ p_a \end{bmatrix} \\
-%    \dot{p_u} &= - m g l (2 s_u + s_{ua})
-%\end{align*}
-%Expanding out \(\dot{q_u}\) we get
-%\begin{multline*}
-%    \dot{q_u} = 
-%    \frac{p_u}{m l^2 (2 - c_a^2)}\\
-%    + \frac
-%    {
-%        m^2 g l^3 \bar{q_a}q_u^2(2s_u + s_{ua})(2 - c_a^2)(1 + c_a)
-%        - (1+c_a)^2(q_u^2+p_u^2)^{3/2}p_u 
-%        + \bar{q_a}(1+c_a)q_up_u^2
-%    }
-%    {
-%        ml^2 (2-c_a^2) ((3 + 2c_a)(q_u^2+p_u^2)^{3/2} - \bar{q_a}(1 + c_a)q_up_u)
-%    }
-%\end{multline*}
-%and, putting everything over one denominator,
-%\begin{align*}
-%    \dot{q_a} &= \frac
-%    {
-%        (3 + 2c_a)(q_u^2+p_u^2)^{3/2}p_u 
-%        - (1+c_a)^2(q_u^2+p_u^2)^{3/2}p_u 
-%        +m^2 g l^3 \bar{q_a}q_u^2(2s_u + s_{ua})(2 - c_a^2)(1 + c_a)
-%    }
-%    {
-%        ml^2 (2-c_a^2) ((3 + 2c_a)(q_u^2+p_u^2)^{3/2} - \bar{q_a}(1 + c_a)q_up_u)
-%    } \\
-%    &= \frac
-%    {
-%        (2-c_a^2)\left((q_u^2+p_u^2)^{3/2}p_u 
-%        +m^2 g l^3 \bar{q_a}q_u^2(2s_u + s_{ua})(1 + c_a)\right)
-%    }
-%    {
-%        ml^2 (2-c_a^2) ((3 + 2c_a)(q_u^2+p_u^2)^{3/2} - \bar{q_a}(1 + c_a)q_up_u)
-%    }
-%\end{align*}
-%Therefore, the zero-dynamics of the acrobot under the constraint 
-%\(q_a = \sin(\arctan(p_u,q_u))\) are:
-%\begin{align}\label{eqn:acrobot_eom_qp}
-%\begin{split}
-%    \dot{q_u} &= \frac
-%    {
-%        (q_u^2+p_u^2)^{3/2}p_u 
-%        +m^2 g l^3 \bar{q_a}q_u^2(2s_u + s_{ua})(1 + c_a)
-%    }
-%    {
-%        ml^2 \left((3 + 2c_a)(q_u^2+p_u^2)^{3/2} - \bar{q_a}(1 + c_a)q_up_u\right)
-%    } \\
-%    \dot{p_u} &= - m g l (2s_u + s_{ua})
-%\end{split}
-%\end{align}
-%
-%It may be convenient to represent these equations of motion  in polar coordinates.
-%Letting \(r = \sqrt{q_u^2 + p_u^2}\) be the radius and 
-%\(\theta = \arctan(p_u,q_u)\) be the angle in the \((q_u,p_u)\) plane and taking
-%derivatives, it is easy to show that
-%\begin{align}\label{eqn:dr_and_dt_general}
-%\begin{split}
-%    \dot{r} &= [\cos(\theta),\sin(\theta)]
-%    \begin{bmatrix} \dot{q_u}(r,\theta) \\ \dot{p_u}(r,\theta) \end{bmatrix} \\
-%    \dot{\theta} &= \frac{1}{r}[-\sin(\theta), \cos(\theta)]
-%    \begin{bmatrix} \dot{q_u}(r,\theta) \\ \dot{p_u}(r,\theta) \end{bmatrix}
-%\end{split}
-%\end{align}
-%
-%By expanding out (\ref{eqn:dr_and_dt_general}) then substituting \(q_u = r\cos(\theta)\)
-%and \(p_u = r\sin(\theta)\)
-%it is straightforward computation to find the equations of motion in cylindrical \((r,\theta)\)
-%coordinates. To simplify the notation yet again, define
-%\(s(g)_\theta := \sin(g(\theta))\) and \(c(g)_\theta := \cos(g(\theta))\) 
-%for any function \(g: \mathbb{R} \rightarrow \mathbb{R}\). The final equations
-%of motion in \((r,\theta)\) coordinates are
-%\begin{align}\label{eqn:acrobot_eom_rt}
-%\begin{split}
-%    \dot{r} &= \frac
-%    {
-%        r^2 c_\theta s_\theta 
-%        + m^2 g l^3 (2 s(rc)_\theta + s(rc + \bar{q_a}s)_\theta)
-%        \left(
-%            (1+c(\bar{q_a}s)_\theta)\bar{q_a}c_\theta 
-%            - (3 + 2c(\bar{q_a}s)_\theta)s_\theta r
-%        \right)
-%    }
-%    {
-%        m l^2 \left((3+2c(\bar{q_a}s)_\theta)r 
-%        - \bar{q_a}(1+ c(\bar{q_a}s)_\theta)c_\theta s_\theta\right)
-%    }\\
-%    \dot{\theta} &= -\frac
-%    {
-%        r s_\theta^2
-%        + m^2 g l^3 (2 s(rc)_\theta + s(rc + \bar{q_a}s)_\theta)
-%        (3 + 2c(\bar{q_a}s)_\theta)c_\theta
-%    }
-%    {
-%        m l^2 \left((3+2c(\bar{q_a}s)_\theta)r 
-%        - \bar{q_a}(1+ c(\bar{q_a}s)_\theta)c_\theta s_\theta\right)
-%    }
-%\end{split}
-%\end{align}
-
 %/========== /Acrobot ==========/%

### 2020-09-30 latex/vnhcs.tex
Started updating VNHCs as per Manfredi's comments
This is the beginning of updating the VNHC chapter according to
Manfredi's comments. I've moved some proofs around and started adding
punctuation to equations, but there is still much to do. Refer to the
draft_MMComments.pdf that I've annotated.

@@
     \mathcal{Q} = \Rt{T_1} \times \cdot \times \Rt{T_n}
+    .
 \] 
@@
 \(\mathcal{L} : \mathcal{Q}\times\R^n \rightarrow \R\) defined by
-(\ref{eqn:lagrangian-general}) \cite{greenwood_dynamics}.
+(\ref{eqn:lagrangian-general}) \cite{greenwood_dynamics}:
 \begin{equation}\label{eqn:lagrangian-general}
     \mathcal{L}(q,\dot{q}) = T(q,\dot{q}) - P(q)
+    .
 \end{equation}
@@
 equations} (\ref{eqn:el-eqns-general}) with \textit{generalized input forces} 
-\(\tau \in \R^k\) . 
+\(\tau \in \R^k\)
 \begin{equation}\label{eqn:el-eqns-general}
@@
     - \pdiff{\mathcal{L}}{q_i} = B_i\tpose(q) \tau
+    .
 \end{equation}
@@
     \end{bmatrix}
+    ,
 \]
@@
     \mathcal{L}(q,\dot{q}) = \frac{1}{2} \dot{q}\tpose D(q) \dot{q} - P(q)
+    ,
 \end{equation}
@@
     p_i(q,\dot{q}) := \pdiff{\mathcal{L}}{\dot{q}_i}(q,\dot{q})
+    .
 \end{equation}
@@
 \(\{q,p\}\) coordinates is the ``Legendre transform"
-(\ref{eqn:hamiltonian-legendre}) of the Lagrangian \cite{landau_mechanics}.
+(\ref{eqn:hamiltonian-legendre}) of the Lagrangian \cite{landau_mechanics},
 \begin{equation}\label{eqn:hamiltonian-legendre}
     \mathcal{H}(q,p) := p\tpose \dot{q}(q,p) - \mathcal{L}(q,\dot{q}(q,p))
+    .
 \end{equation}
 The EOM in the Hamiltonian framework are the \(2n\)
-first-order equations called \textit{Hamilton's equations}. They are given by:
+first-order equations called \textit{Hamilton's equations}. They are given by
 \begin{equation}\label{eqn:hamiltons-eqns}
@@
         \dot{q} = \nabla_p\mathcal{H} \\
-        \dot{p} = -\nabla_q\mathcal{H} + B(q)\tau \\
+        \dot{p} = -\nabla_q\mathcal{H} + B(q)\tau
+        .
     \end{cases}
@@
 and positive definite, it is invertible at each \(q \in \mathcal{Q}\).
-The resulting Hamiltonian system reduces to (\ref{eqn:hamiltonian}).
+The Legendre transform (\ref{eqn:hamiltonian-legendre}) can be explicitly
+written as
+\[
+    \mathcal{H}(q,p) = p\tpose D\inv(q)p - \frac{1}{2}p\tpose D\inv(q) p + V(q)
+    ,
+\]
+which reduces to (\ref{eqn:hamiltonian}). 
+Finding the derivative of each conjugate of momentum yields
+\[
+    \dot{p}_i = -\frac{1}{2} p\tpose \nabla_{q_i}D\inv(q) p 
+        - \pdiff{P}{q_i}(q) + B_i\tpose(q) \tau
+    ,
+\]
+which can be collected into the vector form (\ref{eqn:hamiltonian-full-dynamics}) 
+by using the matrix Kronecker product.
+In sum, the Hamiltonian system reduces to
 \begin{align}\label{eqn:hamiltonian}
-    \mathcal{H}(q,p) &= \frac{1}{2} p\tpose D^{-1}(q) p + P(q) \\
+    \mathcal{H}(q,p) &= \frac{1}{2} p\tpose D^{-1}(q) p + P(q)
+    , \\
      &\begin{cases}
@@
         \dot{p} = -\frac{1}{2} (\Id{n} \otimes p\tpose) \nabla_q D\inv(q) p
-        - \nabla_q P(q) + B(q) \tau \\
-    \end{cases} 
+        - \nabla_q P(q) + B(q) \tau
+        . \\
+    \end{cases} \label{eqn:hamiltonian-full-dynamics}
 \end{align}
 
-Any set of coordinates \(\{q,p\}\) which satisfy Hamilton's equations 
+Any set of coordinates \((q,p)\) which satisfy Hamilton's equations 
 under the Hamiltonian \(\mathcal{H}\) are
@@
 coordinates \((q,p) \rightarrow (Q,P)\) is a \textit{canonical
-transformation} if \(\{Q,P\}\) preserve the Hamiltonian structure; that is, if
+transformation} if \((Q,P)\) preserve the Hamiltonian structure; that is, if
 they are canonical coordinates under the Hamiltonian
@@
 
+\citet{landau_mechanics} provides a useful result for showing whether a given
+change of coordinates \((q,p) \to (Q,P)\) is a canonical transformation.
+\begin{defn}
+    The \textit{Poisson bracket} between the functions \(f(q,p)\) and \(g(q,p)\)
+    is defined as
+    \begin{equation}\label{eqn:poisson-bracket}
+        [f,g] := \sum \limits_{i=1}^n \pdiff{f}{p_i}\pdiff{g}{q_i} - 
+            \pdiff{f}{q_i}\pdiff{g}{p_i}
+        .
+    \end{equation}
+\end{defn}
+\begin{thm}\label{thm:canonical-transformations}
+    A change of coordinates 
+    \((q,p) \to (Q,P)\) is a canonical transformation if
+    \begin{align*}
+        [Q_i, Q_j] &= 0
+        , \\
+        [P_i, P_j] &= 0
+        , \\
+        [P_i, Q_j] &= \delta_{i,j}
+        , 
+    \end{align*}
+    for all \(i,j \in \n\).
+\end{thm}
+\begin{proof}
+    See (45.10) in \cite{landau_mechanics}.
+\end{proof}
+
+Later in this chapter we will define a particular type of coordinate change.
+The following Lemma allows us to prove it is a canonical transformation.
+
+\begin{lemma}\label{lemma:linear-is-canonical}
+    Let \(\mathcal{H}\) be a Hamiltonian system in canonical coordinates
+    \((q,p)\).
+    Let \(A \in \R^{n\times n}\) be an invertible matrix such that
+    \(A\inv = A\tpose\).
+    The change of coordinates \((q,p) \to (Q = Aq, P = Ap)\) is a canonical
+    transformation.
+\end{lemma}
+\begin{proof}
+    For any constant matrix \(A\), the transformation
+    \((Q = Aq, P = Ap)\) satisfies
+    \(\pdiff{Q_i}{p_m} = \pdiff{P_i}{q_m} = 0\) for all 
+    \(i,m \in \n\).
+    Hence, 
+    \begin{align*}
+        [Q_i,Q_j] &:= \sum\limits_{m = 1}^n \pdiff{Q_i}{p_m}\pdiff{Q_j}{q_m} - 
+        \pdiff{Q_i}{q_m}\pdiff{Q_j}{p_m} = 0
+        , \\
+        [P_i,P_j] &:= \sum\limits_{m=1}^n \pdiff{P_i}{p_m}\pdiff{P_j}{q_m} -
+        \pdiff{P_i}{q_m}\pdiff{P_j}{p_m} = 0
+        .
+    \end{align*}
+    Note that \((A_i)\tpose (A\tpose)_j = (A_i)\tpose (A\inv)_j = \delta_{i,j}\). 
+    Using this fact we see that the Poisson brackets between \(P_i\) and \(Q_j\)
+    are given by
+    \begin{align*}
+        [P_i,Q_j] &= \sum\limits_{m=1}^n\pdiff{P_i}{p_m}\pdiff{Q_j}{q_m}
+        - \pdiff{P_i}{q_m}\pdiff{Q_j}{p_m} \\
+                  &= \sum\limits_{m=1}^n A_{i,m}A_{j,m} - 0 \\
+                  &= \sum\limits_{m=1}^n A_{i,m}A\tpose_{m,j} \\
+                  &= (A_i)\tpose A\tpose_j \\
+                  &= \delta_{i,j}
+        .
+    \end{align*}
+    Therefore, by Theorem \ref{thm:canonical-transformations}, the coordinate
+    change \((Q = Aq, P = Ap)\) is a canonical transformation.
+\end{proof}
+
 \section{Simply Actuated Hamiltonian Systems}
-\textbf{Manfredi: What term should we use instead of ``simply actuated"?}
-
 % Describe the change of coordinates to get into q_u/q_a mode and show
@@
     with \(k \leq n\) actuators. 
-    A set of canonical coordinates \(\{q,p\}\) for this system
+    A set of canonical coordinates \((q,p)\) for this system
     are said to be \textit{simply actuated coordinates} if the
@@
         B(q) = \simpleB    
+        .
     \]
@@
     the definition. 
-    The corresponding \((p_u, p_a)\) are called the unactuated and actuated
-    momenta, respectively.
+    The corresponding \((p_u, p_a)\) are called the \textit{unactuated} and 
+    \textit{actuated momenta}, respectively.
 \end{defn}
+% TODO: delete this next definition and rephrase the text accordingly
 \begin{defn}
@@
 
-Note that if \(k = (n-1)\), the existence of any left annihilator 
+Assumption \ref{assm:B-perp} requires the rows of \(B^\perp\) to be unit vectors
+that are mutually orthogonal. 
+In the case that \(k = (n-1)\), the existence of any left annihilator 
 \(A^0 \in \R^{1\times n}\) implies the left annihilator 
-\(B^\perp := A^0/\norm{A^0}\) satisfies Assumption \ref{assm:B-perp}.
+\(B^\perp := A^0/\norm{A^0}\) will satisfy Assumption \ref{assm:B-perp}.
 
@@
     \end{bmatrix}
-\]
-and assigning the input forces to \(\tau = V T \hat{\tau}\), we get a new input
-matrix for \(\hat{\tau} \in \R^k\) given by \(\hat{B} = B V T = U \Sigma T\) 
-which is still constant and full-rank. In particular, 
+    ,
+\]
+and defining the regular feedback transformation\(\tau = V T \hat{\tau}\), 
+we get a new input matrix for \(\hat{\tau} \in \R^k\) given by 
+\[
+    \hat{B} = B V T = 
+    U \begin{bmatrix}
+        \Id{k} \\ \Zmat{(n-k)\times k}
+    \end{bmatrix}
+    ,
+\]
+which is still constant and full-rank. 
+In particular, 
 \(\hat{B}\tpose \hat{B} = T\tpose \Sigma\tpose \Sigma\tpose T = \Id{k}\).
@@
     \end{bmatrix}
+    .
 \]
@@
 orthogonal:
+\begin{proof}
 \[
@@
     \end{bmatrix} = \Id{n}
+    .
 \]
 Hence, \(\mathbf{B}\) is invertible with \(\mathbf{B}\inv = \mathbf{B}\tpose\).
+\end{proof}
 
@@
         \mathcal{H}(Q,P) &= 
-        \frac{1}{2} P\tpose \Minv(Q) P + V(Q) \\
+        \frac{1}{2} P\tpose \Minv(Q) P + V(Q)
+        , \\
        &\begin{cases}
@@
                 - \nabla_Q V(Q) + \simpleB \tau
+            ,
         \end{cases} \nonumber
@@
     \begin{align*}
-        \Minv(Q) &:= \mathbf{B}D^{-1}(\mathbf{B}\tpose Q)\mathbf{B}\tpose \\
-        V(Q) &:= P(\mathbf{B}\tpose Q) \\
+        \Minv(Q) &:= \mathbf{B}D^{-1}(\mathbf{B}\tpose Q)\mathbf{B}\tpose
+        , \\
+        V(Q) &:= P(\mathbf{B}\tpose Q)
+        .
     \end{align*}
@@
 \begin{proof}
-    The Poisson bracket between the functions \(f(q,p)\) and \(g(q,p)\) is defined by
-    \cite{landau_mechanics} as follows:
-    \[
-        [f,g] := \sum \limits_{i=1}^n \pdiff{f}{p_i}\pdiff{g}{q_i} - 
-                \pdiff{f}{q_i}\pdiff{g}{p_i}
-    \]
-    For any constant matrix \(A\), the transformation
-    \(\{Q = Aq, P = Ap\}\) satisfies
-    \(\pdiff{Q_i}{p_m} = \pdiff{P_i}{q_m} = 0\) for all 
-    \(i,m \in \n\).
-    Hence, for our new coordinates \(\{Q = \mathbf{B}q, P = \mathbf{B}p\}\),
-    \begin{align*}
-        [Q_i,Q_j] &:= \sum\limits_{m = 1}^n \pdiff{Q_i}{p_m}\pdiff{Q_j}{q_m} - 
-        \pdiff{Q_i}{q_m}\pdiff{Q_j}{p_m} = 0 \\
-        [P_i,P_j] &:= \sum\limits_{m=1}^n \pdiff{P_i}{p_m}\pdiff{P_j}{q_m} -
-        \pdiff{P_i}{q_m}\pdiff{P_j}{p_m} = 0
-    \end{align*}
-    Since the matrix \(\mathbf{B}\) is orthogonal,
-    \((\mathbf{B}_i)\tpose (\mathbf{B}\tpose)_j = (\mathbf{B}_i)\tpose (\mathbf{B}^{-1})_j = \delta_{i,j}\). Using this
-    fact we see that the Poisson brackets between \(P_i\) and \(Q_j\) are given by:
-    \begin{align*}
-        [P_i,Q_j] &= \sum\limits_{m=1}^n\pdiff{P_i}{p_m}\pdiff{Q_j}{q_m}
-        - \pdiff{P_i}{q_m}\pdiff{Q_j}{p_m} \\
-                  &= \sum\limits_{m=1}^n \mathbf{B}_{i,m}\mathbf{B}_{j,m} - 0 \\
-                  &= \sum\limits_{m=1}^n \mathbf{B}_{i,m}\mathbf{B}\tpose_{m,j} \\
-                  &= (\mathbf{B}_i)\tpose \mathbf{B}\tpose_j \\
-                  &= \delta_{i,j}
-    \end{align*}
-
-    Therefore, by (45.10) in \cite{landau_mechanics}, the coordinate change 
-    \((Q = \mathbf{B}q, P = \mathbf{B}p)\) is a canonical transformation.
+    The change of coordinates \((Q = \mathbf{B}q, P = \mathbf{B}p)\) is a
+    canonical transformation because it satisfies Lemma
+    \ref{lemma:linear-is-canonical}.
     Furthermore, since \(\dot{P} = \mathbf{B}\dot{p}\), the new input matrix is
@@
         \end{bmatrix}
-    \]
-    so the coordinates \(\{Q = (q_u,q_a), P = (p_u,p_a)\}\) are simply actuated coordinates for 
-    \(\mathcal{H}\) as desired.
+        ,
+    \]
+    so \(\left(Q = (q_u,q_a), P = (p_u,p_a)\right)\) are simply actuated
+    coordinates for \(\mathcal{H}\) as desired.
 \end{proof}

### 2020-10-01 latex/vlp.tex
Added KL and restated proof of VLP theorem

@@
 We still need to prove this VNHC will inject energy into a VLP. 
-We will do so by showing the dynamics (\ref{eqn:vlp-hamiltonian}) 
-diverge from the origin via a so-called anti-Lyapunov function.
-
-\begin{defn}\label{defn:anti-lyapunov}
-   Let \(\dot{x} = f(x) \in R^n\) be an ODE with \(f(0) = 0\). 
-   A positive definite function \(V(x)\) is \textbf{anti-Lyapunov} if 
-   \[
-      \dot{V}(x) = dV(x) f(x) \geq 0
-   \]
-   where the inequality is strict for almost every \(x \in \R^n\).
-\end{defn}
-\textbf{Manfredi: Does this notion already exist elsewhere? I couldn't find any
-sources on it, but it seems like it should've been done before.}
-
-By applying Lyapunov stability theory \cite{lyapunov},
-we see that an anti-Lyapunov function
-\(V\) is simply a Lyapunov function for the negative-time system 
-\(\dot{x} = -f(x)\). In other words, \(V\) proves that the orbit \(x(-t)\) 
-of the system travelling backwards in time is asymptotically converging to the
-origin; hence, the system must be diverging from the origin in forward time.
-
-From Theorem \ref{thm:zero-dynamics}, the dynamics on the constraint manifold of
-the VNHC are given by \((\dot{q},\dot{p})\) with 
-\(l = l_\frac{\pi}{2}(\theta)\).
-Supposing the initial condition is not an equilibrium, the existence of an
-anti-Lyapunov function would indicate that the curve traced out on the
-\((q,p)\)-plane is diverging from the origin.
-This implies that the momentum \(p\) is increasing in magnitude whenever the
-curve hits the \(p\)-axis, which in turn means the VLP is gaining energy on
-average.
-
-Proving the existence of an anti-Lyapunov function for the VLP
-requires the following lemma.
+Before we can do this, we require the following lemma.
 
@@
 
+Supposing the initial condition is not an equilibrium, we will show that the
+constrained dynamics trace out a curve on the \((q,p)\)-plane which is diverging
+from the origin.  
+This implies that the momentum \(p\) is increasing in magnitude whenever the
+curve hits the \(p\)-axis, which in turn means the VLP is gaining energy on
+average.
+
 \begin{thm}\label{thm:vlp-energy-stabilization}
@@
    \end{equation}
-   with the property that the inequality is strict for almost every \(\theta\).
+   with the property that the inequality is strict except at the
+   coordinate axes.
    If the inequality is flipped, the VNHC dissipates energy. 
@@
 \begin{proof}
-    Choose, as a candidate anti-Lyapunov function, the energy for the
+   Let us examine the negative-time system
+   \begin{equation}\label{eqn:negative-time-vlp}
+      f(q,p) = 
+      \begin{bmatrix}
+         \dot{q} \\ \dot{p}
+      \end{bmatrix} = \begin{bmatrix}
+         -\frac{p}{ml(\theta)^2} \\
+         mgl(\theta)\sin(q)
+      \end{bmatrix}
+      ,
+   \end{equation}
+   with \(D = (\mathcal{Q}\times \mathcal{P}) \backslash \{(\pm \pi, 0)\}\)
+   the domain of interest on which we study this system.
+   Choose, as a candidate Lyapunov-like function, the energy for the
    average-length pendulum with zero-potential at the bottom of the swing:
@@
    \]
-    This is positive definite at \((0,0)\) and has derivative 
+   This is positive definite at \((0,0)\) and has compact sublevel
+   sets on \(D\). The derivative with respect to (\ref{eqn:negative-time-vlp})
+   is
    \[
-      \dot{E}_\text{avg} = \frac{-g\sin(q)p \left(l(\theta)^3 - l_\text{avg}^3\right)}
+      \dot{E}_\text{avg} = \frac{g\sin(q)p \left(l(\theta)^3 - l_\text{avg}^3\right)}
               {l_\text{avg}^2l(\theta)^2}
+      .
    \]
@@
    \((q,p)\)-plane, so it is equal to the sign of \(\sin(2\theta)\). 
-    Furthermore, by Lemma \ref{lemma:sign-of-cube}:
+   Furthermore, by Lemma \ref{lemma:sign-of-cube} we have
    \[ 
        \sign{l(\theta)^3 - l_\text{avg}^3} = \sign{ l(\theta) - l_\text{avg}}
+       .
    \]
-    Thus, the derivative of \(E_\text{avg}\) is almost always positive since
+   The derivative of \(E_\text{avg}\) is nonnegative since
    \begin{align*}
-       \label{eq:}
-       \sign{\dot{E}_\text{avg}} &= \sign{-\sin(q)p \left(l(\theta)^3 - l_\text{avg}^3\right)} \\
-                   &= -\sign{\sin(2\theta) \left(l(\theta) - l_\text{avg}\right)} \\
-                   &\geq 0 \text{ (by assumption)}
+    \sign{\dot{E}_\text{avg}} &= \sign{\sin(q)p \left(l(\theta)^3 - l_\text{avg}^3\right)} \\
+                &= \sign{\sin(2\theta) \left(l(\theta) - l_\text{avg}\right)} \\
+                &\leq 0 \text{ (by assumption)}
+      ,
    \end{align*}
-    Therefore \(E_\text{avg}\) is an anti-Lyapunov function.
-    Since the initial condition is not an equilibrium,
-    the orbit of the VNHC in the \((q,p)\) plane is diverging away from the
-    origin and the VLP is gaining energy on average.
+   which means \(E_\text{avg}\) is a Lyapunov function.
+   Thus, the negative-time system is stable \cite{lyapunov}; all that remains is
+   to show that (\ref{eqn:negative-time-vlp}) asymptotically converges to
+   \((0,0)\). 
+   For this, we will use a Krazovskii-LaSalle argument
+   \cite{krazovskii_lasalle}.
+   First, we define 
+   \(Z = \left{(q,p) \in \mathbb{S}^1\times \R \mid 
+      \dot{E}_\text{avg}(q,p) = 0 \right}\).
+   It is easy to show that \(Z\) is the union of the coordinate axes
+   \[
+      Z= \mathbb{S}^1 \times \left{0\right} \cup 
+         \left{0} \times \R\right}
+      .
+   \]
+   Then, since \(f(q,p) = (0,0)\) only when \((q,p) = (0,0)\),
+   Krazovskii-LaSalle tells us that the origin is asymptotically stable on
+   \(D\) for the negative-time system.
+   Therefore, the orbit of the VNHC in the \((q,p)\) plane is
+   diverging away from the origin in positive time, which means
+   the VLP is gaining energy on average.
    Flipping the inequality of (\ref{eqn:vlp-energy-gain-condition}) 
-    means \(E_\text{avg}\) is a regular Lyapunov function, so the VLP is
-    losing energy while converging to \((0,0)\).
+   means \(E_\text{avg}\) is a Lyapunov function for the positive time system,
+   so by the same argument the VLP is losing energy while converging
+   asymptotically to \((0,0)\).
  \end{proof}

### 2020-10-01 latex/vlp.tex
Fixed errors that stopped compilation

@@
    First, we define 
-   \(Z = \left{(q,p) \in \mathbb{S}^1\times \R \mid 
-      \dot{E}_\text{avg}(q,p) = 0 \right}\).
+   \[
+      Z = \left\{(q,p) \in \mathbb{S}^1\times \R \mid 
+      \dot{E}_\text{avg}(q,p) = 0 \right\}
+      .
+   \]
    It is easy to show that \(Z\) is the union of the coordinate axes
    \[
-      Z= \mathbb{S}^1 \times \left{0\right} \cup 
-         \left{0} \times \R\right}
+      Z= \mathbb{S}^1 \times \left\{0\right\} \cup 
+         \left\{0\right\} \times \R
       .
    \]
-   Then, since \(f(q,p) = (0,0)\) only when \((q,p) = (0,0)\),
-   Krazovskii-LaSalle tells us that the origin is asymptotically stable on
-   \(D\) for the negative-time system.
-   Therefore, the orbit of the VNHC in the \((q,p)\) plane is
-   diverging away from the origin in positive time, which means
-   the VLP is gaining energy on average.
+   Krazovskii-LaSalle tells us solutions to the negative-time ODE converge to
+   the largest invariant set in \(Z\).
+   Since \(f(q,p) = (0,0)\) only when \((q,p) = (0,0)\),
+   the origin is asymptotically stable on \(D\) for this negative-time system.
+   Therefore, the positive-time system is diverging away from the origin 
+   in the \((q,p)\) plane, which means the VLP is gaining energy on average.
    Flipping the inequality of (\ref{eqn:vlp-energy-gain-condition}) 

### 2020-10-02 latex/vnhcs.tex
Finished the first edit of VNHCs with Manfredi's notes

@@
 which can be collected into the vector form (\ref{eqn:hamiltonian-full-dynamics}) 
-by using the matrix Kronecker product.
-In sum, the Hamiltonian system reduces to
+% TODO: Update Appendix info?
+by using the matrix Kronecker product (see Appendix ???).
+In sum, when the kinetic energy is quadratic the Hamiltonian system reduces to
 \begin{align}\label{eqn:hamiltonian}
@@
                   &= \sum\limits_{m=1}^n A_{i,m}A\tpose_{m,j} \\
-                  &= (A_i)\tpose A\tpose_j \\
+                  &= (A_i)\tpose (A\tpose)_j \\
                   &= \delta_{i,j}
@@
 
-\section{Simply Actuated Hamiltonian Systems}
+\section{Simply Actuated Hamiltonian Systems}\label{sec:simply-actuated}
 % Describe the change of coordinates to get into q_u/q_a mode and show
@@
 \end{defn}
-% TODO: delete this next definition and rephrase the text accordingly
-\begin{defn}
-    A Hamiltonian system is said to be \textit{simply actuated} if there exists
-    a canonical transformation from any canonical coordinates 
-    \(\{q,p\}\) into simply actuated coordinates.
-\end{defn}
-
-Under the following assumptions on the input matrix, we will show that the
-Hamiltonian system (\ref{eqn:hamiltonian}) is simply actuated.
+
+Under the following assumptions on the input matrix, we will show that there is
+a canonical transformation of (\ref{eqn:hamiltonian}) into simply actuated
+coordinates.
 
@@
 
-\begin{assm}\label{assm:B-orthogonal}
-    Assume \textbf{without loss of generality} that the input matrix \(B\) is
-    left semi-orthogonal. That is, \(B\tpose B = \Id{k}\).
-\end{assm}
+\begin{lemma}\label{lemma:B-orthogonal}
+    Suppose Assumption \ref{assm:B-const} holds. Then
+    there exist nonsingular matrices \(V,T \in \R^{k \times k}\) 
+    so that the regular feedback transformation 
+    \[
+        \tau = V T \hat{\tau}
+    \] 
+    has a new input matrix \(\hat{B}\) for \(\hat{\tau}\) which is left
+    semi-orthogonal.  
+    That is, \(\hat{B}\tpose \hat{B} = \Id{k}\).
+\end{lemma}
 \begin{proof}
@@
 
-Let \(\mathbf{B} \in \R^{n\times n}\) be the following matrix:
+In light of Lemma \ref{lemma:B-orthogonal}, there is no loss of generality
+making the following assumption.
+\begin{assm}\label{assm:B-orthogonal}
+    Assume that the input matrix \(B\) is
+    left semi-orthogonal. 
+\end{assm}
+
+Let now \(\mathbf{B} \in \R^{n\times n}\) be the following matrix:
 \[
@@
     Under Assumptions \ref{assm:B-const},\ref{assm:B-perp}, and
-    \ref{assm:B-orthogonal}, the Hamiltonian system (\ref{eqn:hamiltonian}) is
-    simply actuated with simply actuated coordinates 
+    \ref{assm:B-orthogonal}, the Hamiltonian system (\ref{eqn:hamiltonian})
+    has simply actuated canonical coordinates 
     \(\{Q = \mathbf{B}q, P = \mathbf{B}p\}\). The resulting dynamics are 
@@
 \begin{proof}
-    The change of coordinates \((Q = \mathbf{B}q, P = \mathbf{B}p)\) is a
-    canonical transformation because it satisfies Lemma
-    \ref{lemma:linear-is-canonical}.
+    The change of coordinates \((Q = \mathbf{B}q, P = \mathbf{B}p)\)
+    satisfies Lemma \ref{lemma:linear-is-canonical}, making it a canonical
+    transformation.
     Furthermore, since \(\dot{P} = \mathbf{B}\dot{p}\), the new input matrix is
@@
     \]
-    so \(\left(Q = (q_u,q_a), P = (p_u,p_a)\right)\) are simply actuated
-    coordinates for \(\mathcal{H}\) as desired.
+    which means \(\left(Q = (q_u,q_a), P = (p_u,p_a)\right)\) are simply
+    actuated coordinates for \(\mathcal{H}\) as desired.
 \end{proof}
@@
 
-The downside to VHCs is that they do not use the additional information imparted
-by the velocity of a system.
+The downside to VHCs is that they only depend on the configuration of a
+mechanical system, and not its generalized velocity.
 For the child on a swing, whether they extend or retract their legs
@@
 the usage of VHCs. 
-Many authors have attempted to extend the theory of VHCs to enforce relations
+A few authors have attempted to extend the theory of VHCs to enforce relations
 \(h(q,\dot{q}) = 0\) of the full state to account for this drawback. 
@@
     relation \(h(q,p) = 0\) where \(h : \mathcal{Q}\times\R^n \rightarrow \R^k\) is
-    smooth, \(\rank{\left[ dh_q,\, dh_p \right]} = k\) for all 
+    \(C^2\), \(\rank{\left[ dh_q,\, dh_p \right]} = k\) for all 
     \((q,p) \in h\inv(0)\), and there exists a feedback controller \(\tau(q,p)\)
-    stabilizing the set
+    stabilizing the \textit{constraint manifold}
     \[
         \Gamma = \left\{(q,p) \mid h(q,p) = 0, dh_q \dot{q} + dh_p \dot{p} = 0\right\}
-    \]
-    which is called the \textit{constraint manifold}.
+        .
+    \]
 \end{defn}
-\textbf{Manfredi: Do we require \(C^\infty\), or can we relax it to \(C^2\)?}
-
-If we define the error term  \(e = h(q,p)\), stabilizing \(\Gamma\) is
-equivalent to solving for the controller \(\tau(q,p)\) which drives 
-\(e \rightarrow 0\) and \(\dot{e} \rightarrow 0\).
-Supposing there is no further structure on the VNHC, 
-then \(\dot{e} = dh_q \dot{q} + dh_p \dot{p}\). Since \(B(q) \tau\) appears inside
-\(\dot{p}\), solving for \(\tau\) explicitly requires 
-\(dh_p(q,p)B(q) \in \R^{k \times k}\) to be invertible for all \((q,p)\).
-%which is the same as requiring that \(h(q,p)\) is strictly monotonic in \(p\).
-
-This condition severely restricts the types of constraints one can use. 
-It is preferable instead to have the torque \(\tau\) appear after two
-derivatives of \(e\) to allow for more freedom in choosing a constraint.
-Mathematically, if each \(\tau_i\) appears only after two derivatives, one says 
-that \(e\) is of \textit{relative degree} \(\{2,2,\ldots,2\}\). 
-We thus define a special type of VNHC which satisfies this
-property.
+
+From the definition of VNHCs, one finds that the constraint manifold \(\Gamma\)
+is a \(2(n-k)\)-dimensional surface inside \(\mathcal{Q} \times \R^n\). 
+The next obvious question is the following: when does a feedback controller
+stabilizing \(\Gamma\) exist?
+
+One approach to answering this question is to
+define the error term \(e = h(q,p)\). 
+Finding \(\tau(q,p)\) driving \(e \to 0\) and \(\dot{e} \to 0\)
+is sufficient for stabilizing \(\Gamma\). 
+If \(\tau\) appears after only \(r_i \geq 0\) derivatives of \(e_i\), 
+then that \(e_i\) is said to have \textit{relative degree} \(r_i\).
+We would then say \(e\) has relative degree \(\{r_1,\ldots,r_k\}\).
+
+One important note is that having relative degree \(1\) for any \(e_i\) means we
+may not be able to stabilize \(\Gamma\); we can guarantee \(h(q,p) \to 0\) 
+but not \(\dot{h}(q,p) = 0\). 
+Requiring that \(e\) have relative degree \(\{2,\ldots,2\}\) is much more
+useful, since it allows us to easily solve for a \(\tau\) which stabilizes
+\(\Gamma\). 
+This is a common requirementa in the VHC literature, so we will make use of it
+here by defining a special type of VNHC that satisfies this property.
 
@@
 cannot have \(\tau\) appearing after only one derivative. Of course,
-they performed their research in Lagrangian form; 
-for us to be able to use the Hamiltonian coordinate \(p_u\), 
-we must continue with the following assumption for the rest of the chapter.
-
-\begin{assm}\label{assm:H-is-simply-actuated}
-    The mechanical system under consideration is a
-    Hamiltonian system with \(n\) degrees of freedom and 
-    \(k < n\) actuators. It is described in simply
-    actuated coordinates \(\{q = (q_u,q_a), p = (p_u, p_a)\}\) and has the
-    following dynamics:
-    \begin{align*}
-        \mathcal{H}(q,p) &= p\tpose \Minv(q) p + V(q) \\
+they performed their research in Lagrangian form; we will be using
+the Hamiltonian formulation from Chapter \ref{sec:simply-actuated}.
+As a reminder, our system is described in \((q,p)\) coordinates with 
+\(q = (q_u,q_a)\) and \(p = (p_u, p_a)\) and has the dynamics
+\begin{align}\label{eqn:simply-actuated-dynamics}
+    \mathcal{H}(q,p) &= p\tpose \Minv(q) p + V(q)
+    , \\
     &\begin{cases}
        \dot{q} = \Minv p \\
-            \dot{p} = -\frac{1}{2} \pdmat - \nabla_q V(q) + \simpleB \tau \\
+       \dot{p} = -\frac{1}{2} \pdmat - \nabla_q V(q) + \simpleB \tau 
+       .\\
     \end{cases}
-    \end{align*}
-\end{assm}
+\end{align}
+
 \begin{notation}
@@
          \right)\simpleB} = k
+         ,
     \]
@@
         \end{bmatrix} + \simpleB \tau\right)
+        .
     \end{align*}
@@
         dh_{p_u} \left(\frac{1}{2} \pudmat + \nabla_{q_u}V(q)\right)
+        .
     \]
@@
     \begin{align*}
-        \ddot{e} &= \diff{}{t}\left\{dh_q\right\}\Minv(q) p + 
+        \ddot{e} &= \diff{}{t}\left[dh_q\right]\Minv(q) p + 
         dh_q \left(\sum\limits_{i=1}^n \pdiff{\Minv}{q_i}(q)\dot{q_i}\right)p + 
         dh_q \Minv(q) \dot{p} - \\
-         & \diff{}{t}\left\{dh_{p_u}\right\}
+         & \diff{}{t}\left[dh_{p_u}\right]
          \left(\frac{1}{2}\pudmat + \nabla_{q_u}V(q)\right) - \\
-         & dh_{p_u}\left(\frac{1}{2}\diff{}{t}\left\{\pudmat\right\} + 
-         \diff{}{t}\left\{\nabla_{q_u}V(q)\right\} \right)
+         & dh_{p_u}\left(\frac{1}{2}\diff{}{t}\left[\pudmat\right] + 
+         \diff{}{t}\left[\nabla_{q_u}V(q)\right] \right)
+         .
     \end{align*}
-    We will compute the explicit expression of \(\ddot{e}\) in pieces to find
-    out where \(\tau\) appears. 
-% d/dt(dh_q)
-    We begin with \(\diff{}{t}\left\{dh_q\right\}\). 
-    Let \(h = (h^1,\ldots,h^k)\) with
-    \(h^i : \mathcal{Q} \times \mathcal{P}_u \rightarrow \R\). By definition of
-    the total differential, we have 
-    \[
-        dh_q = \begin{bmatrix}
-            dh^1_q \\
-            \vdots \\
-            dh^k_q
-        \end{bmatrix} = \begin{bmatrix}
-        \pdiff{h^1}{q_1} & \cdots & \pdiff{h^1}{q_n} \\
-        \vdots & \ddots & \vdots \\
-        \pdiff{h^k}{q_1} & \cdots & \pdiff{h^k}{q_n}
-        \end{bmatrix}
-    \]
-    The time derivative is taken element-wise, yielding
-    \[
-        \diff{}{t}\left\{dh_q\right\} = \begin{bmatrix}
-            \sum_{j=1}^n \ppdiff{h^1}{q_1}{q_j}\dot{q}_j & \cdots & \sum_{j=1}^n
-            \ppdiff{h^1}{q_n}{q_j}\dot{q}_j \\
-            \vdots & \ddots & \vdots \\
-            \sum_{j=1}^n \ppdiff{h^k}{q_1}{q_j}\dot{q}_j & \cdots & \sum_{j=1}^n
-            \ppdiff{h^k}{q_n}{q_j}\dot{q}_j \\
-        \end{bmatrix} + \begin{bmatrix}
-            \sum_{l=1}^{n-k}\ppdiff{h^1}{q_1}{p_{u_l}}\dot{p}_{u_l} & \cdots &
-            \sum_{l=1}^{n-k}\ppdiff{h^1}{q_n}{p_{u_l}}\dot{p}_{u_l} \\
-            \vdots & \ddots & \vdots \\
-            \sum_{l=1}^{n-k}\ppdiff{h^k}{q_1}{p_{u_l}}\dot{p}_{u_l} & \cdots & 
-            \sum_{l=1}^{n-k}\ppdiff{h^k}{q_n}{p_{u_l}}\dot{p}_{u_l} \\
-        \end{bmatrix}
-    \]
-    It is straightforward computation to confirm that each row of this
-    derivative can be written in vector form as follows:
-    \begin{align*}
-        \diff{}{t}\left\{dh^i_q\right\} &= \begin{bmatrix}
-        \sum_{j=1}^n \ppdiff{h^i}{q_1}{q_j}\dot{q}_j & \cdots & 
-        \sum_{j=1}^n \ppdiff{h^i}{q_n}{q_j}\dot{q}_j 
-        \end{bmatrix} + \begin{bmatrix}
-        \sum_{l=1}^{n-k}\ppdiff{h^i}{q_1}{p_{u_l}}\dot{p}_{u_l} & \cdots &
-        \sum_{l=1}^{n-k}\ppdiff{h^i}{q_n}{p_{u_l}}\dot{p}_{u_l}
-        \end{bmatrix} \\
-       &= \dot{q}\tpose \Hess_q\left\{h^i\right\}\tpose + 
-       \dot{p}_u\tpose \partial_{p_u}\partial_{q}h^i \\
-       &= p\tpose \Minv(q)\Hess_q\left\{h^i\right\}\tpose -
-       \left(\frac{1}{2}p\tpose d\Minv_{q_u}(q)(\Id{n-k}\otimes p) + 
-       dV_{q_u}(q)\right)\partial_{p_u}\partial_{q}h^i
-    \end{align*}
-    This means that \(\diff{}{t}\left\{dh_q\right\}\) is given by
-    \begin{align}
-        \diff{}{t}\left\{dh_q\right\} &= 
-        \left(\Id{n} \otimes 
-        \left(p\tpose \Minv(q)\right)\right)
-        \Hess_q\left\{h\right\}\tpose -{} 
-        \nonumber \\
-      & \left(\frac{1}{2}\Id{n} \otimes 
-          \left(p\tpose d\Minv_{q_u}(q)(\Id{n-k}\otimes p)\right)
-        + \Id{n}\otimes dV_{q_u}(q)\right)\partial_{p_u}\partial_{q}h 
-    \end{align}
-% C_1(q,p)
-    For the next segment, let \(1_j \in \R^n\) be the vector of all zeros,
-    except for a single \(1\) at position \(j\). We define
-    \(C_1 : \mathcal{Q} \times \mathcal{P} \rightarrow \R^{n\times n}\) by
-    \begin{equation}
-        C_1(q,p) := \sum\limits_{j=1}^n \pdiff{\Minv}{q_j}(q)\dot{q}_j 
-        = \sum\limits_{j=1}^n \pdiff{\Minv}{q_j}(q)\left(1_j\tpose\Minv(q)p\right)
-    \end{equation}
-% d/dt(dh_pu)
-    Moving on to \(\diff{}{t}\left\{dh_{p_u}\right\}\), a similar approach to
-    \(\diff{}{t}\left\{dh_q\right\}\) yields the derivative in matrix form.
-    It is given by:
-    \begin{align}
-        \diff{}{t}\left\{dh_{p_u}\right\} &= 
-        \left( \Id{n} \otimes \left(p\tpose \Minv(q)\right)\right) 
-        \partial_q\partial_{p_u}h -{} 
-        \nonumber \\
-      & \left(\frac{1}{2}\Id{n} \otimes \left(\pudmat\right) + 
-      \Id{n} \otimes \nabla_{q_u}V(q)\right)\Hess_{p_u}\left\{h\right\}\tpose
-    \end{align}
+    Most of these terms do not involve \(\dot{p}\) and hence do not contain 
+    \(\tau\), so we shorten this to 
+    \[
+        \ddot{e} = (\star) - 
+        dh_{p_u}\left(\frac{1}{2}\diff{}{t}\left[\pudmat\right] +
+        dh_q \Minv(q) \dot{p}\right)
+        .
+    \]
 % d/dt(1/2 * pudmat)
-    For the next piece, observe that the \(i^\text{th}\) row of
-    \(\frac{1}{2} \diff{}{t} \left\{ \pudmat \right\}\)
+    Observe that the \(i^\text{th}\) row of
+    \(\frac{1}{2} \diff{}{t} \left[ \pudmat \right]\)
     is given by
     \begin{align*}
-        \frac{1}{2} \diff{}{t}\left\{p\tpose \pdiff{\Minv}{q_{u_i}}(q)p\right\}
+        \frac{1}{2} \diff{}{t}\left[p\tpose \pdiff{\Minv}{q_{u_i}}(q)p\right]
         = p\tpose \pdiff{\Minv}{q_{u_i}}(q) \dot{p} + 
@@
         \dot{q}_j \right) p
+        .
     \end{align*}
-    Defining \(C_2 : \mathcal{Q} \times \mathcal{P} \rightarrow \R^{n(n-k)\times n}\) 
-    by
-    \begin{equation}
-        C_2(q,p) := \begin{bmatrix}
-            \sum_{j=1}^n \ppdiff{\Minv}{q_{u_1}}{q_j} (1_j\tpose \Minv(q)p) \\
-            \vdots \\
-            \sum_{j=1}^n \ppdiff{\Minv}{q_{u_{n-k}}}{q_j} (1_j\tpose \Minv(q)p) \\
-        \end{bmatrix}
-    \end{equation}
-    allows us to collapse the derivative into:
-    \begin{align}
-        & \frac{1}{2} \diff{}{t} \left\{ \pudmat \right\} ={} 
-        \nonumber \\
-        & (\Id{n-k} \otimes p\tpose)\nabla_{q_u}\Minv(q)
-        \left(-\frac{1}{2}\pudmat - \nabla_{q_u}V(q) + \simpleB \tau \right) +{}
-        \nonumber \\
-        & \frac{1}{2}\left(\Id{n-k}\otimes p\tpose\right) C_2(q,p) p
-    \end{align}
-% d/dt(nabla_qu V)
-    Finally, we compute the derivative of \(\nabla_{q_u}V(q)\):
-    \begin{equation}
-        \diff{}{t}\left\{\nabla_{q_u}V(q)\right\} = 
-        \begin{bmatrix}
-            \sum_{j=1}^n \ppdiff{V}{q_{u_1}}{q_j}\dot{q}_j \\
-            \vdots \\
-            \sum_{j=1}^n \ppdiff{V}{q_{u_{n-k}}}{q_j}\dot{q}_j
-        \end{bmatrix} = \partial_{q_u} \partial_q V(q) \dot{q} 
-        = \partial_{q_u} \partial_q V(q) \Minv(q)p
-    \end{equation}
+    Highlighting only the term containing \(\tau\), we get the vector form
+    \begin{align*}
+        \frac{1}{2} \diff{}{t} \left[ \pudmat \right] =
+        (\star) +
+        (\Id{n-k} \otimes p\tpose)\nabla_{q_u}\Minv(q)\simpleB \tau
+        .
+    \end{align*}
 % Full ddot(e)
-    Putting this all together, we can find the explicit form for \(\ddot{e}\):
-    \begin{align}
-        \ddot{e} &= \diff{}{t}\left\{dh_q\right\}\Minv(q)p + dh_q C_1(q,p)p -{}
-        \nonumber \\
-     & \frac{1}{2}dh_q\Minv(q)(\pdmat) + 
-        dh_q\Minv(q)\nabla_qV(q) -{} 
-        \nonumber \\
-     & \diff{}{t}\left\{dh_{p_u}\right\}
-     \left(\frac{1}{2}\pudmat + \nabla_{q_u}V(q)\right) +{} 
-     \nonumber \\
-     & dh_{p_u}(\Id{n-k}\otimes p\tpose)\nabla_{q_u}\Minv(q)\left(
-     \frac{1}{2}\pudmat + \nabla_{q_u}V(q)\right) -{}
-     \nonumber \\
-     & \frac{1}{2}dh_{p_u}(\Id{n-k}\otimes p\tpose)C_2(q,p)p - 
-     dh_{p_u} \partial_{q_u}\partial_q V(q)\Minv(q)p +{} 
-     \nonumber \\
-     & \left(dh_q \Minv(q) - dh_{p_u}(\Id{n-k} \otimes
+    Plugging this into \(\ddot{e}\) reveals that
+    \begin{align*}
+        \ddot{e} = (\star) +
+     \left(dh_q \Minv(q) - dh_{p_u}(\Id{n-k} \otimes
      p\tpose)\nabla_{q_u}\Minv(q) \right) \simpleB \tau
-    \end{align}
+     ,
+    \end{align*}
+    where \((\star)\) is a function of \(q\) and \(p\).
 % End of Proof
@@
     when \(e\) is of relative degree \(\{2,\ldots,2\}\), which is true 
-    if and only if
-    %\(\tau\) appears in each row of \(\ddot{e}\) on \(\Gamma\). 
-    one can solve for \(\tau\) when \(\ddot{e} = 0\).
-    This is equivalent to requiring that the matrix \(H\) be
-    %full rank everywhere on the constraint manifold, 
-    invertible,
-    proving the theorem.
+    if and only if the matrix premultiplying \(\tau\) is nonsingular, and hence
+    that \(H\) be invertible, proving the theorem.
 \end{proof}
@@
 stabilized by the output-linearizing phase-feedback controller
-(\ref{eqn:vnhc-torque-controller}).
 \begin{equation}\label{eqn:vnhc-torque-controller}
     \tau(q,p) = -H\inv(q,p)\left(E(q,p) + k_p e + k_d \dot{e}\right)
+    ,
 \end{equation}
@@
 Now that we have found a controller to enforce a regular VNHC of order \(k\), we
-would like to solve for the closed-loop dynamics. Intuitively, these dynamics
-should be parameterized by \((q_u, p_u)\) since \(q_a\) is a function of these
-as specified by \(h(q,p_u) = 0\).
+would like to determine the dynamics on the constraint manifold \(\Gamma\). 
+Intuitively, these dynamics should be parameterized by \((q_u, p_u)\) since
+\(q_a\) is a function of these as specified by \(h(q,p_u) = 0\).
 Unfortunately, \(\dot{q}_u\) depends on \(p_a\), and for general systems one
-cannot solve explicitly for \(p_a\) in terms of \((q_u,p_u)\). This is because
-the \(\dot{p}\) dynamics contains the coupling term 
+cannot solve explicitly for \(p_a\) in terms of \((q_u,p_u)\). 
+This is because the \(\dot{p}\) dynamics contains the coupling term 
 \((\Id{n} \otimes p\tpose)\nabla_{q_u}M(q)p\).
 
-We now introduce a class of systems where explicitly solving for the closed-loop
-dynamics is feasible.
-
-% TODO: Come up with a name for this type of system
-\begin{defn}
-    A mechanical system is \textbf{TODO: name this} ------ if 
-    \(\nabla_{q_u}M(q) = \Zmat{n(n-k) \times n}\).
-\end{defn}
-
-% TODO: Do we need the assumption that one can solve for qa linearly?
+We now introduce an assumption so we can solve explicitly for the constrained
+dynamics.
+
+\begin{assm}\label{assm:inertially-actuated}
+    The Hamiltonian system has an inertia matrix that does not depend on the
+    unactuated coordinates:
+    \[
+        \nabla_{q_u}M(q) = \Zmat{n(n-k) \times n}
+        .
+    \]
+\end{assm}
+
 \begin{thm}\label{thm:zero-dynamics}
-    Let \(\mathcal{H}\) be a ----- mechanical system satisfying Assumption
-    \ref{assm:H-is-simply-actuated}. Let
-    \(h(q,p_u) = 0\) be a regular VNHC of order \(k\) 
-    with constraint manifold \(\Gamma\). Suppose that on \(\Gamma\) one can
-    solve linearly for \(q_a\) as a function of \((q_u,p_u)\).
-    Then the closed-loop dynamics are given by
+    Let \(\mathcal{H}\) be a mechanical system in simply actuated
+    coordinates satisfying Assumption \ref{assm:inertially-actuated}. 
+    Let \(h(q,p_u) = 0\) be a regular VNHC of order \(k\) with constraint
+    manifold \(\Gamma\). Suppose that on \(\Gamma\) one can solve linearly for
+    \(q_a\) as a function of \((q_u,p_u)\).
+    Then the constrained dynamics are given by
     \begin{equation}\label{eqn:qpu-dynamics}
@@
             \end{array}}
+            ,
     \end{equation}
@@
     \begin{aligned}
-        g(q,p_u) &:= \\
+        g(q_u,p_u) &:= \\
            &\left.\left(dh_q \Minv(q) \simpleB \right)\inv 
@@
         \end{bmatrix} p_u\right)\right|_{h(q,p_u) = 0}
+        .
     \end{aligned}
@@
         \dot{e} = dh_q\Minv(q)p - dh_{p_u}\nabla_{q_u}V(q)
+        .
     \]
@@
         (n-k)}\end{bmatrix} p_u + dh_q \Minv(q)\simpleB p_a
+        .
     \end{align*}
@@
             \Id{n-k} \\ \Zmat{k \times (n-k)}\end{bmatrix} p_u
+        .
     \]
@@
         \rank{dh_q\Minv(q)\simpleB} = k
+        .
     \]
@@
         \Id{n-k} \\ \Zmat{k \times (n-k)}\end{bmatrix} p_u \right)
+        .
     \]
@@
     \(\Gamma\) are parameterized only by \((\dot{q}_u,\dot{p}_u)\).
-
-    \textbf{MANFREDI: Do we need this assumption that \(q_a\) can be solved in
-        terms of \((q_u,p_u)\) or is there some other way of parameterizing it
-    like you did with \(s\) for VHCs?}
 \end{proof}
@@
 unactuated coordinates.
-This is true regardless of the number of degrees of freedom of the system, which
-means that \(\Gamma\) is a \(2(n-k)\)-dimensional surface inside 
-\(\mathcal{Q}\times \mathcal{P}\).
+This is true regardless of the number of degrees of freedom of the system.
 
@@
 \begin{cor}\label{cor:2d-zero-dynamics}
-    Suppose \(\mathcal{H}\) is a ---- mechanical system satisfying Assumption
-    \ref{assm:H-is-simply-actuated} which has degree of underactuation one.
+    Suppose \(\mathcal{H}\) is a ---- mechanical system in simply actuated
+    coordinates which has degree of underactuation one.
     Let \(h(q,p_u) = 0\) be a regular VNHC of order \((n-1)\) of the form

### 2020-10-05 latex/vlp.tex
Edited intro to VLP chapter

@@
 offer a time-independent technique to inject energy into the VLP. 
-They designed a controller through a technique called \textit{energy shaping},
-and proved that their control mechanism stabilizes any desired energy level set.
+They design a controller through a technique called \textit{energy shaping},
+and prov that their control mechanism stabilizes any desired energy level set.
 However, the energy injection mechanism is ad-hoc in the sense that it is not
-derived from natural behaviour. 
-In this chapter we will design VNHCs to stabilize energy levels of the VLP in a
-time-independent manner, while maintaining the structured motion of a human on a
-swing.
+derived from human behaviour; their technique is tailored to this specific
+problem, and cannot be generalized easily.
+In this chapter we will make use of the general VNHC framework developed in Chapter
+\ref{ch:vnhcs} to stabilize energy levels of the VLP in a
+time-independent manner. 
+These VNHCs will add and remove energy from the VLP while maintaining the
+structured motion of a human on a swing.
 
@@
    \frac{1}{2}m\norm{\dot{x}}^2 = \frac{1}{2}m\left(\dot{l}^2 + l^2\dot{q}^2\right)
+   .
 \]
@@
    P(\mathbf{q}) = -mgl\cos(q)
+   .
 \]
@@
    + mgl\cos(q)
+   .
 \]
@@
    = \begin{bmatrix} ml^2\dot{q} \\ m\dot{l} \end{bmatrix} 
+   .
 \]
@@
 \(M(\mathbf{q}) := D(\mathbf{q})\), \(V(\mathbf{q}) := P(\mathbf{q})\),
-we find the Hamiltonian is equal to the total mechanical energy of the system:
-\[
-   \mathcal{H} = E = \frac{1}{2}\mathbf{p}\tpose \Minv(\mathbf{q})
-   \mathbf{p}+V(\mathbf{q})
-\]
+we get the Hamiltonian (\ref{eqn:simply-actuated-hamiltonian}).
 Taking the appropriate derivatives, the dynamics of the VLP in Hamiltonian form
@@
 that is, the control input cannot enact any lateral force on the pendulum.
-This is extremely useful for simplifying the dynamics: since one can choose
-\(\tau\) which sets \(\dot{p}_l\) to any desired function and \(\dot{l}\)
-depends only on \(p_l\), we can assume (with abuse of notation) that \(l\) is
-tracking some known function \(l(t)\). 
-The closed loop dynamics of the system will be described exclusively by
-\((\dot{q},\dot{p})\) with \(l \equiv l(t)\).
-The fact that \(p_l\) does not appear in these closed-loop dynamics allows us to
-ignore the subdynamics \((\dot{l},\dot{p}_l)\) entirely.
+This is extremely useful because it means the VLP is described in simply
+actuated coordinates, which allows us to apply the theory of VNHCs we
+developed in Chapter \ref{ch:vnhcs}.
+% TODO: Replace this with a description that the dynamics of (q,p) are the
+% constrained dynamics, and that we are using a VNHC o f the form h = l -
+% L(q,p). We will simplify notation by saying that l = L(q,p) to mean that we
+% are stabilizing h = l - L = 0
+Let us define a ``VNHC" of the form \(l = L(q,p)\) (by this we mean we are
+actually defining the VNHC \(h(q,p) = l - L(q,p) = 0\)).
+Assuming this is a regular VNHC, the constraint manifold \(\Gamma\) is
+\(\mathbb{S}^1 \times \R\), and it is parameterized by the unactuated phase
+\((q,p)\).
+The constrained dynamics are described entirely by
+\((\dot{q},\dot{p})\) with \(l\) replaced by \(L(q,p)\) (omitting the inputs to
+\(L\) for clarity):
+\begin{equation}\label{eqn:vlp-dynamics}
+   \begin{cases}
+      \dot{q} = \frac{p}{m L^2} \\
+      \dot{p} = -mgL\sin(q)
+      . \\
+   \end{cases}
+\end{equation}
 
-What this means is we can treat \(l\) as the control input directly, rather than
-modelling it as a configuration variable. Re-deriving the Hamiltonian and the
-dynamics from this approach, we get the system 
-(\ref{eqn:vlp-hamiltonian}) with phase \((q,p) \in \mathbb{S}^1 \times \R\). 
-Note that the control input \(l(t)\) and its derivative \(\dot{l}(t)\) are both
-known variables.
-\begin{align}\label{eqn:vlp-hamiltonian}
-   \mathcal{H}(q,p) &= \frac{p^2}{ml^2} - \frac{1}{2}\dot{l}^2 - mgl\cos(q) \\
-     &\begin{cases}
-        \dot{q} = \frac{p}{ml^2} \\
-        \dot{p} = -mgl\sin(q) \\
-      \end{cases}\nonumber
-\end{align}
-In particular, the Hamiltonian of this simplified model is no longer equal to
-the total mechanical energy of the system, which is given by
-(\ref{eqn:vlp-energy}).
+The total mechanical energy of the system restricted to the constraint manifold
+is given by (\ref{eqn:vlp-energy}), where \(L(q,p)\) and \(\dot{L}(q,p)\) are
+known.
 \begin{equation}\label{eqn:vlp-energy}
-   E(q,p) = \frac{p^2}{ml^2} + \frac{1}{2}\dot{l}^2 - mgl\cos(q)
+   E(q,p) = \frac{p^2}{mL^2} + \frac{1}{2}\dot{L}^2 - mgL\cos(q)
 \end{equation}
+In the rest of this chapter, we will derive the function \(L(q,p)\) which
+allows the constrained dynamics to inject energy into the VLP over time.
 

### 2020-10-06 latex/vlp.tex
Fully edited the intro to VLP

@@
 offer a time-independent technique to inject energy into the VLP. 
-They design a controller through a technique called \textit{energy shaping},
-and prov that their control mechanism stabilizes any desired energy level set.
-However, the energy injection mechanism is ad-hoc in the sense that it is not
-derived from human behaviour; their technique is tailored to this specific
-problem, and cannot be generalized easily.
+They design a controller through a technique called \textit{energy shaping}
+and prove that it stabilizes any desired energy level set.
+However, their control input depends on a pre-specified target energy and requires
+knowledge of the current total energy of the VLP.
+This makes their energy injection mechanism ``ad-hoc" in the sense that it is
+tailored very specifically to the VLP, and is not generalizable to a larger 
+methodology.
+
+It may be better to base the control design on natural biological behaviour.
 In this chapter we will make use of the general VNHC framework developed in Chapter
-\ref{ch:vnhcs} to stabilize energy levels of the VLP in a
-time-independent manner. 
-These VNHCs will add and remove energy from the VLP while maintaining the
-structured motion of a human on a swing.
+\ref{ch:vnhcs} to add and remove energy from the VLP in a time-independent
+manner.
+We'll show that, unlike energy shaping, VNHCs can be used to stabilize energy
+levels while maintaining the structured motion of a human on a swing.
 
@@
 connected to a fixed pivot by a massless rod of varying length \(l\) with angle 
-\(q \in \mathbb{S}^1\) from the vertical, as is seen in Figure
+\(q \in \mathbb{S}^1\) from the vertical, as in Figure
 \ref{fig:vlp-model}. 
-We will also ignore any damping and frictional forces in this model.
+We will ignore any damping and frictional forces in this model.
 In a realistic VLP, the rod length \(l\) varies between some minimum
@@
 \(M(\mathbf{q}) := D(\mathbf{q})\), \(V(\mathbf{q}) := P(\mathbf{q})\),
-we get the Hamiltonian (\ref{eqn:simply-actuated-hamiltonian}).
-Taking the appropriate derivatives, the dynamics of the VLP in Hamiltonian form
-are described in (\ref{eqn:vlp-hamiltonian-with-pl}). 
+we get the Hamiltonian (\ref{eqn:simply-actuated-hamiltonian}) whose
+dynamics (\ref{eqn:simply-actuated-dynamics}) resolve to  
 \begin{align}\label{eqn:vlp-hamiltonian-with-pl}
@@
          0 & \frac{1}{m}
-      \end{bmatrix} \begin{bmatrix} p \\ p_l \end{bmatrix} - mgl\cos(q) \\
+      \end{bmatrix} \begin{bmatrix} p \\ p_l \end{bmatrix} - mgl\cos(q)
+      , \\
      &\begin{cases}
@@
         \dot{p} = -mgl\sin(q) \\
-        \dot{p}_l = \frac{p^2}{ml^3} + mg\cos(q) + \tau \\
+        \dot{p}_l = \frac{p^2}{ml^3} + mg\cos(q) + \tau
+        . \\
    \end{cases} \nonumber
@@
 The control input is a force \(\tau \in \mathbb{R}\) affecting the dynamics of
-\(p_l\), acting colinearly with direction of the rod.
+\(p_l\), acting colinearly with the rod.
 We assume the force does not affect the dynamics of \(p\) in any way -
@@
 developed in Chapter \ref{ch:vnhcs}.
-% TODO: Replace this with a description that the dynamics of (q,p) are the
-% constrained dynamics, and that we are using a VNHC o f the form h = l -
-% L(q,p). We will simplify notation by saying that l = L(q,p) to mean that we
-% are stabilizing h = l - L = 0
-Let us define a ``VNHC" of the form \(l = L(q,p)\) (by this we mean we are
-actually defining the VNHC \(h(q,p) = l - L(q,p) = 0\)).
-Assuming this is a regular VNHC, the constraint manifold \(\Gamma\) is
-\(\mathbb{S}^1 \times \R\), and it is parameterized by the unactuated phase
-\((q,p)\).
-The constrained dynamics are described entirely by
-\((\dot{q},\dot{p})\) with \(l\) replaced by \(L(q,p)\) (omitting the inputs to
-\(L\) for clarity):
-\begin{equation}\label{eqn:vlp-dynamics}
-   \begin{cases}
+
+Let us define the VNHC \(l = L(q,p)\) -- by which we mean we are
+actually defining the VNHC \(h(q,p) = l - L(q,p) = 0\).
+Assuming this is a regular VNHC, the constraint manifold is
+\(\Gamma = \mathbb{S}^1 \times \R\), which is parameterized by the unactuated
+phase \((q,p)\).
+The constrained dynamics are Hamiltonian, and Theorem \ref{thm:zero-dynamics}
+tells us they are described entirely by \((\dot{q},\dot{p})\) with \(l\)
+replaced by \(L(q,p)\):
+\begin{align}\label{eqn:vlp-hamiltonian}
+   \mathcal{H}(q,p) &= \frac{1}{2} \frac{p^2}{mL^2} - \frac{1}{2}\dot{L}^2 -
+      mgL\cos(q)
+   , \\
+  &\begin{cases}
       \dot{q} = \frac{p}{m L^2} \\
@@
       . \\ 
-   \end{cases}
-\end{equation}
+   \end{cases} \nonumber
+\end{align}
+Note that we suppress the function notation for \(L(q,p)\) and
+\(\dot{L}(q,p)\) for clarity.
 
@@
 \begin{equation}\label{eqn:vlp-energy}
-   E(q,p) = \frac{p^2}{mL^2} + \frac{1}{2}\dot{L}^2 - mgL\cos(q)
+   E(q,p) = \frac{1}{2} \frac{p^2}{mL^2} + \frac{1}{2}\dot{L}^2 - mgL\cos(q)
 \end{equation}
-In the rest of this chapter, we will derive the function \(L(q,p)\) which
-allows the constrained dynamics to inject energy into the VLP over time.
+In the rest of this chapter, we will derive a \(C^2\) function \(L(q,p)\) 
+which produces constrained dynamics that inject energy into the VLP in a
+human-like manner.
 
@@
 manner, we will examine a person standing on a swing.
-As can be seen in Figure \ref{fig:child-vlp}, their center of mass moves closer
+As can be seen in Figure \ref{fig:child-vlp}, a person's center of mass moves closer
 to the swing's pivot when they stand, and moves away from the pivot when they
@@
 this strategy will necessarily use the full phase \((q,p)\). 
-This is why we will use VNHCs instead of VHCs.
+This is why we must use VNHCs instead of VHCs to perform this maneuver.
 
@@
 \[
-   l_\text{opt}(q,p) := -\sign{qp}
+   L^\star(q,p) := -\sign{qp}
+   ,
 \]
 which is a piecewise-continuous controller that varies between \(\pm 1\). 
-This looks similar to a VNHC, but it is not smooth and makes the assumption that 
-\(l \in [-1,1]\). 
-Since we want our controller emulate realistic human motion, we need to derive a
-smooth controller where we allow \(l \in [\underbar{l},\overbar{l}]\); then, we
-will prove that it injects energy.
+We could set our constraint to be \(l = L^\star(q,p)\), but this is not a VNHC
+because it is not \(C^2\).
+It also makes the assumption that \(l \in [-1,1]\) and can switch
+instantaneously.
+Since need to enforce the constraint using the physical input \(\tau\) (which
+ideally would emulate realistic human motion), we cannot use \(L^\star\) as our
+VNHC.
+We will instead find an alternate representation of \(L^\star\) which can be
+converted into a VNHC, and which allows 
+\(l \in [\underbar{l},\overbar{l}]\).
 
-Figure \ref{fig:vlp-optimal-controller-qp-plane} displays \(l_\text{opt}\) 
-directly on the \((q,p)\)-plane.
+Figure \ref{fig:vlp-optimal-controller-qp-plane} displays \(L^\star(q,p)\) 
+on the \((q,p)\)-plane.
 Because the length remains constant inside each quadrant and changes only when it
 crosses one of the axes, one can use the angle \(\theta = \arctan_2(p,q)\) to
-redefine the time-optimal controller in a way which is easier to ``smooth out".
-Figure \ref{fig:vlp-optimal-controller} shows the time-optimal pumping strategy
-as a function \(l^\star(\theta)\) (\ref{eqn:vlp-optimal-controller}), where now
-the length varies between \([\underbar{l},\overbar{l}]\) rather than 
+redefine the time-optimal pumping strategy as a function of \(\theta\).
+Abusing notation, we denote this by \(L^\star(\theta)\), which is defined in
+(\ref{eqn:vlp-optimal-controller}).
+Figure \ref{fig:vlp-optimal-controller} shows the graph of \(L^\star(\theta)\), 
+where now the length varies between \([\underbar{l},\overbar{l}]\) rather than 
 \([-1,1]\).
@@
 \begin{equation}\label{eqn:vlp-optimal-controller}
-   l^\star(\theta):= \begin{cases}
+   L^\star(\theta):= \begin{cases}
       \overline{l} & \theta \in [-\frac{\pi}{2},0[ \cup [\frac{\pi}{2}, \pi[ \\
@@
    \end{cases}
+   .
 \end{equation}
@@
       \includestandalone[width=\linewidth]{images/vlp_optimal_controller_qp_plane}
-      \caption{The time-optimal controller \(l_\text{opt}\) mapped onto the \((q,p)\)-plane.}
+      \caption{The time-optimal controller \(L^\star(q,p)\) mapped onto the
+      \((q,p)\)-plane. Here, red corresponds to \(L^\star(q,p) = -1\) and blue
+      to \(L^\star(q,p) = 1\).}
       \label{fig:vlp-optimal-controller-qp-plane}
@@
       \caption{The time-optimal controller converted to the alternate
-         representation \(l^\star(\theta)\).}
+      representation \(L^\star(\theta)\).}
       \label{fig:vlp-optimal-controller}
@@
 We now define a continuous function which
-approximates \(l^\star(\theta)\) by attaching sinusoids at the
+approximates \(L^\star(\theta)\) by attaching sinusoids at the
 discontinuities (see Figure \ref{fig:vlp-T-controller}). 
@@
 (where \(T \in \, ]0,\frac{\pi}{2}]\) is a control parameter), we get the
-continuously-differentiable controller \(l_T(\theta)\) (\ref{eqn:vlp-T-controller}).
+continuously-differentiable controller \(L_T(\theta)\) (\ref{eqn:vlp-T-controller}).
 \begin{equation}\label{eqn:vlp-T-controller}
-   l_T(\theta) = \begin{cases}
+   L_T(\theta) = \begin{cases}
       \overline{l} & \theta \in \left[-\frac{\pi}{2} + \frac{T}{2}, -\frac{T}{2}\right] 
@@
    \includestandalone[width=0.7\textwidth]{images/vlp_T_controller}
-   \caption{The continuous VNHC-style VLP controller \(l_T(\theta)\).}\label{fig:vlp-T-controller}
+   \caption{The continuous VLP constraint \(L_T(\theta)\).}\label{fig:vlp-T-controller}
 \end{figure}
 
-This controller approximates the optimal controller because 
+This family of constraints approximates \(L^\star(\theta)\) because 
 \[
-   \lim\limits_{T \rightarrow 0} l_T(\theta) = l^\star(\theta)
+   \lim\limits_{T \rightarrow 0} L_T(\theta) = L^\star(\theta)
+   .
 \]
-Unfortunately, while \(l_T(\theta)\) is continuously-differentiable, 
+Unfortunately, while \(L_T(\theta)\) is continuously-differentiable, 
 it is not twice-differentiable for most values of \(T\).
@@
 forces \(\tau\) acting on \(p_l\) can be discontinuous (which is certainly
-achievable by humans), or we must find a value of \(T\) where this controller is
-smooth.
-Thankfully, \(l_{\frac{\pi}{2}}(\theta)\) is smooth as it can be simplified into
-the expression (\ref{eqn:vlp-smoothed-controller}). 
+achievable by humans), or we must find a value of \(T\) where this constraint is
+at least \(C^2\).
+Thankfully, setting \(T = \frac{\pi}{2}\) yields the smooth function 
+\(L_{\frac{\pi}{2}}(\theta)\), which can be simplified into
 \begin{equation}\label{eqn:vlp-smoothed-controller}
-   l_\frac{\pi}{2}(\theta) = -\Delta l \sin(2\theta) + l_{\text{avg}}
+   L_\frac{\pi}{2}(\theta) = -\Delta l \sin(2\theta) + l_{\text{avg}}
+   .
 \end{equation}
-It is plotted for demonstration in Figure \ref{fig:vlp-smoothed-controller}.
+This smooth constraint is plotted for demonstration in Figure
+\ref{fig:vlp-smoothed-controller}.
 
@@
    \includestandalone[width=0.6\textwidth]{images/vlp_smoothed_controller}
-   \caption{The smoothed VLP controller which corresponds to the VNHC \(l =
-      l_\frac{\pi}{2}(\theta)\).}\label{fig:vlp-smoothed-controller}
+   \caption{The smoothed VLP constraint which corresponds to the VNHC 
+      \(l = L_\frac{\pi}{2}(\theta)\).}\label{fig:vlp-smoothed-controller}
 \end{figure}
 
-Because \(l_\frac{\pi}{2}(\theta)\) is smooth and it approximates
-\(l^\star(\theta)\), we set our VNHC to be
+Because \(L_\frac{\pi}{2}(\theta)\) is smooth and it approximates
+\(L^\star(\theta)\), we set our VNHC to be
 \[
-   h(q,p) = l - l_\frac{\pi}{2}\left(\theta(q,p)\right)
+   h(q,p) = l - L_\frac{\pi}{2}\left(\theta(q,p)\right)
+   .
 \]
-We still need to prove this VNHC will inject energy into a VLP. 
+We still need to prove this will inject energy into the VLP. 
 Before we can do this, we require the following lemma.
@@
    \((q_0,p_0) \not \in \left\{(0,0),(\pm\pi,0)\right\}\).
-   Define \(\theta := \arctan_2(p,q)\). 
-   A VNHC of the form \(h(q,p) = l - l(\theta)\) injects energy if there exists 
+   Define \(\theta := \arctan_2(p,q)\) and let 
+   \(L : \mathbb{S}^1 \rightarrow [\underline{l},\overline{l}]\) be \(C^2\).
+   A VNHC of the form \(h(q,p) = l - L(\theta)\) injects energy into the VLP
+   if there exists 
    \(l_\text{avg} \in [\underline{l},\overline{l}]\) such that 
    \begin{equation}\label{eqn:vlp-energy-gain-condition}
-      \left(l(\theta) - l_\text{avg}\right)sin(2\theta) \leq 0 \text{ }\forall \theta \in \mathbb{S}^1
+      \left(l_\text{avg} - L(\theta)\right)sin(2\theta) \geq 0 \text{ }\forall \theta \in \mathbb{S}^1
+      ,
    \end{equation}
@@
 \begin{proof}
-   Let us examine the negative-time system
+   Let us examine the negative-time system for the constrained dynamics
    \begin{equation}\label{eqn:negative-time-vlp}
       f(q,p) = 
-      \begin{bmatrix}
+      -\begin{bmatrix}
          \dot{q} \\ \dot{p}
       \end{bmatrix} = \begin{bmatrix}
-         -\frac{p}{ml(\theta)^2} \\
-         mgl(\theta)\sin(q)
+         -\frac{p}{mL(\theta)^2} \\
+         mgL(\theta)\sin(q)
       \end{bmatrix}
@@
    This is positive definite at \((0,0)\) and has compact sublevel
-   sets on \(D\). The derivative with respect to (\ref{eqn:negative-time-vlp})
-   is
+   sets on \(D\). 
+   The derivative with respect to (\ref{eqn:negative-time-vlp}) is
    \[
-      \dot{E}_\text{avg} = \frac{g\sin(q)p \left(l(\theta)^3 - l_\text{avg}^3\right)}
-              {l_\text{avg}^2l(\theta)^2}
+      \dot{E}_\text{avg} = \frac{g\sin(q)p \left(L(\theta)^3 - l_\text{avg}^3\right)}
+              {l_\text{avg}^2L(\theta)^2}
       .
@@
    \[ 
-       \sign{l(\theta)^3 - l_\text{avg}^3} = \sign{ l(\theta) - l_\text{avg}}
+       \sign{L(\theta)^3 - l_\text{avg}^3} = \sign{ L(\theta) - l_\text{avg}}
        .
@@
    \begin{align*}
-    \sign{\dot{E}_\text{avg}} &= \sign{\sin(q)p \left(l(\theta)^3 - l_\text{avg}^3\right)} \\
-                &= \sign{\sin(2\theta) \left(l(\theta) - l_\text{avg}\right)} \\
+    \sign{\dot{E}_\text{avg}} &= \sign{\sin(q)p \left(L(\theta)^3 - l_\text{avg}^3\right)} \\
+                &= \sign{\sin(2\theta) \left(L(\theta) - l_\text{avg}\right)} \\
+                &= -\sign{(l_\text{avg} - L(\theta))\sin(2\theta)}
                 &\leq 0 \text{ (by assumption)}
@@
    \end{align*}
-   which means \(E_\text{avg}\) is a Lyapunov function.
+   which means \(E_\text{avg}\) is a Lyapunov function for the negative-time
+   system.
    Thus, the negative-time system is stable \cite{lyapunov}; all that remains is
@@
    \[
-      l_\frac{\pi}{2}(\theta) = -\Delta l \sin(2\theta) + l_\text{avg}
+      L_\frac{\pi}{2}(\theta) = -\Delta l \sin(2\theta) + l_\text{avg}
    \]
@@
    \[
-      l^{-}_\frac{\pi}{2}(\theta) := \Delta l \sin(2\theta) + l_\text{avg}
+      L^{-}_\frac{\pi}{2}(\theta) := \Delta l \sin(2\theta) + l_\text{avg}
    \]
-   The VNHC \(l = l_\frac{\pi}{2}(\theta)\) injects energy into the VLP
+   The VNHC \(l = L_\frac{\pi}{2}(\theta)\) injects energy into the VLP
    over time, since 
    \[
-      \left(l_\frac{\pi}{2}(\theta) - l_\text{avg}\right)\sin(2\theta) = 
-      -\Delta l \sin^2(2\theta) \leq 0
+      \left(l_\text{avg} - L_\frac{\pi}{2}(\theta)\right)\sin(2\theta) = 
+      \Delta l \sin^2(2\theta) \geq 0
    \] 
-   The VNHC \(l = l^{-}_\frac{\pi}{2}(\theta)\) dissipates energy
+   The VNHC \(l = L^{-}_\frac{\pi}{2}(\theta)\) dissipates energy
    from the VLP, since 
    \[
-      \left(l^{-}_\frac{\pi}{2}(\theta) - l_\text{avg}\right)\sin(2\theta) = 
-      \Delta l \sin^2(2\theta) \geq 0
+      \left(l_\text{avg} - L^{-}_\frac{\pi}{2}(\theta)\right)\sin(2\theta) = 
+      - \Delta l \sin^2(2\theta) \leq 0
    \] 
@@
 For the VNHCs we designed in this chapter, this means toggling
-between \(l_\frac{\pi}{2}(\theta)\) and \(l^{-}_\frac{\pi}{2}(\theta)\),
+between \(L_\frac{\pi}{2}(\theta)\) and \(L^{-}_\frac{\pi}{2}(\theta)\),
 with some hysteresis to avoid infinite switching.
@@
    \includestandalone[width=0.5\textwidth]{images/vlp_energy_in_out}
-   \caption{A VNHC of the form \(l(\theta)\) which is entirely contained within
+   \caption{A VNHC of the form \(l = L(\theta)\) where \(L(\theta)\)
+      is entirely contained within
       the green (yellow) regions will inject (dissipate) energy into the VLP

### 2020-10-06 latex/vlp.tex
Fuly eddited VLP up to L_T definition

@@
 \end{align}
-Note that we suppress the function notation for \(L(q,p)\) and
+Note that we suppress the function notation of \(L(q,p)\) and
 \(\dot{L}(q,p)\) for clarity.
@@
 this strategy will necessarily use the full phase \((q,p)\). 
-This is why we must use VNHCs instead of VHCs to perform this maneuver.
+This is why we must use VNHCs instead of other methods (such as VHCs) 
+to perform this maneuver.
 
@@
 because it is not \(C^2\).
-It also makes the assumption that \(l \in [-1,1]\) and can switch
-instantaneously.
-Since need to enforce the constraint using the physical input \(\tau\) (which
-ideally would emulate realistic human motion), we cannot use \(L^\star\) as our
+Additionally, it would force us to assume that \(l \in \{-1,0,1\}\) and that one
+can switch \(l\) instantaneously.
+Since we need to enforce the constraint using the physical input \(\tau\) (which
+would ideally emulate realistic human motion), we cannot use \(L^\star\) as our
 VNHC.
@@
 on the \((q,p)\)-plane.
-Because the length remains constant inside each quadrant and changes only when it
-crosses one of the axes, one can use the angle \(\theta = \arctan_2(p,q)\) to
-redefine the time-optimal pumping strategy as a function of \(\theta\).
+Note that the length remains constant inside each quadrant and changes only when it
+crosses one of the axes. Using this fact, we can redefine
+the time-optimal pumping strategy as a function of 
+\(\theta := \arctan_2(p,q)\).
 Abusing notation, we denote this by \(L^\star(\theta)\), which is defined in
@@
 where now the length varies between \([\underbar{l},\overbar{l}]\) rather than 
-\([-1,1]\).
+\(\{-1,0,1\}\).
 
@@
       \overline{l} & \theta \in [-\frac{\pi}{2},0[ \cup [\frac{\pi}{2}, \pi[ \\
-      \underline{l} & \theta \in [-pi, -\frac{\pi}{2}[ \cup [0,\frac{\pi}{2}[ \\
+      \underline{l} & \theta \in [-pi, -\frac{\pi}{2}[ \cup [0,\frac{\pi}{2}[ 
+      .
    \end{cases}
-   .
 \end{equation}
@@
 We now define a continuous function which
-approximates \(L^\star(\theta)\) by attaching sinusoids at the
-discontinuities (see Figure \ref{fig:vlp-T-controller}). 
+approximates \(L^\star(\theta)\). 
 Let \(\Delta l := (\overline{l} - \underline{l})/2\) and 
 \(l_{\text{avg}} := (\overline{l} + \underline{l})/2\).
-Providing these newly attached sinusoids with a frequency \(\omega = \frac{\pi}{T}\) 
-(where \(T \in \, ]0,\frac{\pi}{2}]\) is a control parameter), we get the
-continuously-differentiable controller \(L_T(\theta)\) (\ref{eqn:vlp-T-controller}).
+Let \(T \in \, ]0,\frac{\pi}{2}]\) be a parameter of our choosing.
+By intelligently attaching sinusoids of frequency \(\omega = \frac{\pi}{T}\) to
+\(L^\star(\theta)\) (see Figure \ref{fig:vlp-T-controller}), we get a family of
+\(C^1\) constraints \(L_T(\theta)\) (\ref{eqn:vlp-T-controller}) parameterized
+by \(T\).
+
 \begin{equation}\label{eqn:vlp-T-controller}
@@
       a \in \left\{-\frac{\pi}{2}, \frac{\pi}{2}\right\} \\
-      -\Delta l \sin(\omega(\theta-\pi)) & \theta \in \left[\pi - \frac{T}{2},\pi\right] \\
+      -\Delta l \sin(\omega(\theta-\pi)) & \theta \in \left[\pi - \frac{T}{2},\pi\right] 
+      .
    \end{cases}
-   .
 \end{equation}

### 2020-10-07 latex/vlp.tex
Fully edited VLP as per Manfredi's suggestions

@@
 that is, the control input cannot enact any lateral force on the pendulum.
-This is extremely useful because it means the VLP is described in simply
-actuated coordinates, which allows us to apply the theory of VNHCs we
+This makes the VLP into an underactuated mechanical system with degree of
+underactuation 1. 
+It is also a useful assumption because it means \((\mathbf{q},\mathbf{p})\) 
+are simply actuated coordinates, which allows us to apply the theory of VNHCs we
 developed in Chapter \ref{ch:vnhcs}.
@@
 Let us define the VNHC \(l = L(q,p)\) -- by which we mean we are
-actually defining the VNHC \(h(q,p) = l - L(q,p) = 0\).
-Assuming this is a regular VNHC, the constraint manifold is
+actually defining the VNHC \(h(\mathbf{q},\mathbf{p}) = l - L(q,p) = 0\) of
+order 1.
+The VLP satisfies
+\(\nabla_q \Minv(\mathbf{q}) = \Zmat{2 \times 2}\);
+therefore, \(h(\mathbf{q},\mathbf{p})\) is a regular VNHC whenever 
+\(L(q,p)\) is \(C^2\) because
+\[
+   dh_{\mathbf{q}} \Minv(\mathbf{q})B = 
+   \begin{bmatrix}
+      -\pdiff{L}{q} & 1
+   \end{bmatrix}
+   \begin{bmatrix}
+      \frac{1}{ml^2}  & 0 \\
+      0 & \frac{1}{m}
+   \end{bmatrix} 
+   \begin{bmatrix}
+      0 \\ 
+      1
+   \end{bmatrix}
+   = \frac{1}{m}
+\]
+is always full rank.
+The constraint manifold is
 \(\Gamma = \mathbb{S}^1 \times \R\), which is parameterized by the unactuated
@@
 \end{equation}
-In the rest of this chapter, we will derive a \(C^2\) function \(L(q,p)\) 
-which produces constrained dynamics that inject energy into the VLP in a
-human-like manner.
+In the rest of this chapter, we will derive a \(C^2\) function \(L(q,p)\) based
+on natural human motion.
+This function will produce constrained dynamics that inject energy into the VLP.
 
@@
 standing and squatting correspond to shortening and lengthening the pendulum
-(respectively).
+respectively.
 \begin{figure}
@@
 
-The action of regulating pendulum length to add energy to the VLP is known as
+The action of regulating pendulum length to inject energy into the VLP is known as
 ``pumping". \citet{pumping_swing_standing_squatting} asked whether
@@
 when the system is at the ``bottom" or ``top" of the swing. 
-Since being at the ``bottom" is equivalent to having angle \(q = 0\) and being
-at the ``top" is equivalent to having momentum \(p = 0\), a controller based on
-this strategy will necessarily use the full phase \((q,p)\). 
+Since being at the bottom is equivalent to having angle \(q = 0\) and being
+at the top is equivalent to having momentum \(p = 0\), a controller based on
+this strategy will necessarily involve the full unactuated phase \((q,p)\). 
 This is why we must use VNHCs instead of other methods (such as VHCs) 
@@
 
-The time-optimal controller from \cite{pumping_swing_standing_squatting} is (in
-our notation)
+The time-optimal controller from \cite{pumping_swing_standing_squatting} is, in
+our notation,
 \[
@@
 \(L^\star(\theta)\) (see Figure \ref{fig:vlp-T-controller}), we get a family of
-\(C^1\) constraints \(L_T(\theta)\) (\ref{eqn:vlp-T-controller}) parameterized
-by \(T\).
-
+\(C^1\) constraints \(L_T(\theta)\) parameterized
+by \(T\): 
 \begin{equation}\label{eqn:vlp-T-controller}
-   L_T(\theta) = \begin{cases}
+   L_T(\theta) := \begin{cases}
       \overline{l} & \theta \in \left[-\frac{\pi}{2} + \frac{T}{2}, -\frac{T}{2}\right] 
@@
    \includestandalone[width=0.7\textwidth]{images/vlp_T_controller}
-   \caption{The continuous VLP constraint \(L_T(\theta)\).}\label{fig:vlp-T-controller}
+   \caption{The continuous VLP constraint \(l = L_T(\theta)\).}
+   \label{fig:vlp-T-controller}
 \end{figure}
@@
 Thankfully, setting \(T = \frac{\pi}{2}\) yields the smooth function 
-\(L_{\frac{\pi}{2}}(\theta)\), which can be simplified into
+\(L_{\frac{\pi}{2}}(\theta)\), which can be simplified from
+(\ref{eqn:vlp-T-controller}) into
 \begin{equation}\label{eqn:vlp-smoothed-controller}
@@
    \includestandalone[width=0.6\textwidth]{images/vlp_smoothed_controller}
-   \caption{The smoothed VLP constraint which corresponds to the VNHC 
-      \(l = L_\frac{\pi}{2}(\theta)\).}\label{fig:vlp-smoothed-controller}
+   \caption{The smoothed VLP constraint \(l = L_\frac{\pi}{2}(\theta)\).}
+   \label{fig:vlp-smoothed-controller}
 \end{figure}
@@
 \[
-   h(q,p) = l - L_\frac{\pi}{2}\left(\theta(q,p)\right)
+   h(\mathbf{q},\mathbf{p}) = l - L_\frac{\pi}{2}\left(\theta(q,p)\right)
    .
 \]
-We still need to prove this will inject energy into the VLP. 
-Before we can do this, we require the following lemma.
+What remains is to prove this injects energy into the VLP. 
+As part of the proof, we will require the following lemma.
 
 \begin{lemma}\label{lemma:sign-of-cube}
-   For any \(x,y \in \R\):
+   For any \(x,y \in \R\),
    \[
@@
 \begin{proof}
-   Observe first that
-   \[
-      x^3 - y^3 =  (x-y)(x^2 + xy + y^2)
-   \]
+   Observe that \(x^3 - y^3 =  (x-y)(x^2 + xy + y^2)\).
    The inequality \(x^2 + xy + y^2 \geq 0\) holds because
@@
       x^2 + xy + y^2 = \left(x  + \frac{y}{2}\right)^2 + \frac{3y^2}{4} \geq 0
+      ,
    \]
@@
 
-Supposing the initial condition is not an equilibrium, we will show that the
+Supposing the initial condition of the VLP is not an equilibrium, we will show that the
 constrained dynamics trace out a curve on the \((q,p)\)-plane which is diverging
@@
 \begin{thm}\label{thm:vlp-energy-stabilization}
-   Take the VLP (\ref{eqn:vlp-hamiltonian}) with initial condition 
+   Define \(\theta := \arctan_2(p,q)\).
+   Let \(L : \mathbb{S}^1 \rightarrow [\underline{l},\overline{l}]\) be a
+   \(C^2\) function of \(\theta\).
+   Suppose the initial condition of the VLP's unactuated variables is
    \((q_0,p_0) \not \in \left\{(0,0),(\pm\pi,0)\right\}\).
-   Define \(\theta := \arctan_2(p,q)\) and let 
-   \(L : \mathbb{S}^1 \rightarrow [\underline{l},\overline{l}]\) be \(C^2\).
-   A VNHC of the form \(h(q,p) = l - L(\theta)\) injects energy into the VLP
+   A VNHC of the form 
+   \(h(\mathbf{q},\mathbf{p}) = l - L(\theta)\) injects energy into the VLP
    if there exists 
@@
    \begin{equation}\label{eqn:vlp-energy-gain-condition}
-      \left(l_\text{avg} - L(\theta)\right)sin(2\theta) \geq 0 \text{ }\forall \theta \in \mathbb{S}^1
+      \left(l_\text{avg} - L(\theta)\right)sin(2\theta) \geq 0 
+      \text{ }\forall \theta \in \mathbb{S}^1
       ,
@@
 \begin{proof}
-   Let us examine the negative-time system for the constrained dynamics
+   Let \(\Gamma\) be the constraint manifold for \(h\).
+   In negative-time, the dynamics on \(\Gamma\) are
    \begin{equation}\label{eqn:negative-time-vlp}
@@
       \end{bmatrix}
-      ,
+      .
    \end{equation}
-   with \(D = (\mathcal{Q}\times \mathcal{P}) \backslash \{(\pm \pi, 0)\}\)
-   the domain of interest on which we study this system.
+   Let \(D = \Gamma \backslash \{(\pm \pi, 0)\}\)
+   be the domain of interest on which we study this negative-time system.
    Choose, as a candidate Lyapunov-like function, the energy for the
@@
    sets on \(D\). 
-   The derivative with respect to (\ref{eqn:negative-time-vlp}) is
+   The derivative of \(E_\text{avg}\) along (\ref{eqn:negative-time-vlp}) is
    \[
@@
    The sign of \(\sin(q)p\) is dependent only on the quadrant in the
-   \((q,p)\)-plane, so it is equal to the sign of \(\sin(2\theta)\). 
+   \((q,p)\)-plane, so one can show it is equal to the sign of
+   \(\sin(2\theta)\). 
    Furthermore, by Lemma \ref{lemma:sign-of-cube} we have
    \[ 
-       \sign{L(\theta)^3 - l_\text{avg}^3} = \sign{ L(\theta) - l_\text{avg}}
+       \sign{l_\text{avg}^3 - L(\theta)^3} = \sign{l_\text{avg} - L(\theta)}
        .
@@
                 &= \sign{\sin(2\theta) \left(L(\theta) - l_\text{avg}\right)} \\
-                &= -\sign{(l_\text{avg} - L(\theta))\sin(2\theta)}
+                &= -\sign{(l_\text{avg} - L(\theta))\sin(2\theta)} \\
                 &\leq 0 \text{ (by assumption)}
-      ,
+      .
    \end{align*}
-   which means \(E_\text{avg}\) is a Lyapunov function for the negative-time
-   system.
-   Thus, the negative-time system is stable \cite{lyapunov}; all that remains is
-   to show that (\ref{eqn:negative-time-vlp}) asymptotically converges to
-   \((0,0)\). 
-   For this, we will use a Krazovskii-LaSalle argument
-   \cite{krazovskii_lasalle}.
-   First, we define 
+   Thus, \(E_\text{avg}\) is a Lyapunov function and the negative-time system is
+   stable \cite{lyapunov}. 
+   The last step of the proof is to show that (\ref{eqn:negative-time-vlp})
+   asymptotically converges to \((0,0)\). 
+   For this, we will use a Krasovskii-LaSalle argument
+   \cite{krasovskii_lasalle}.
+   First, we define the zero-set
    \[
-      Z = \left\{(q,p) \in \mathbb{S}^1\times \R \mid 
+      Z = \left\{(q,p) \in \Gamma \mid 
       \dot{E}_\text{avg}(q,p) = 0 \right\}
@@
    \]
-   Krazovskii-LaSalle tells us solutions to the negative-time ODE converge to
-   the largest invariant set in \(Z\).
+   Krasovskii-LaSalle tells us that solutions to (\ref{eqn:negative-time-vlp})
+   converge to the largest invariant set in \(Z \cap D\).
    Since \(f(q,p) = (0,0)\) only when \((q,p) = (0,0)\),
-   the origin is asymptotically stable on \(D\) for this negative-time system.
-   Therefore, the positive-time system is diverging away from the origin 
+   the origin is asymptotically stable on \(D\) for the negative-time system.
+   Hence, the positive-time system is diverging away from the origin 
    in the \((q,p)\) plane, which means the VLP is gaining energy on average.
    Flipping the inequality of (\ref{eqn:vlp-energy-gain-condition}) 
-   means \(E_\text{avg}\) is a Lyapunov function for the positive time system,
-   so by the same argument the VLP is losing energy while converging
+   means \(E_\text{avg}\) is a Lyapunov function for the positive time system;
+   by the same arguments presented here, the VLP loses energy while converging
    asymptotically to \((0,0)\).
@@
    \[
-      L_\frac{\pi}{2}(\theta) = -\Delta l \sin(2\theta) + l_\text{avg}
+      L_\frac{\pi}{2}(\theta) := -\Delta l \sin(2\theta) + l_\text{avg}
+      ,
    \]
@@
       L^{-}_\frac{\pi}{2}(\theta) := \Delta l \sin(2\theta) + l_\text{avg}
+      .
    \]
-   The VNHC \(l = L_\frac{\pi}{2}(\theta)\) injects energy into the VLP
-   over time, since 
+   The VNHC \(l = L_\frac{\pi}{2}(\theta)\) injects energy into the VLP,
+   while \(l = L^{-}_\frac{\pi}{2}(\theta)\) dissipates energy.
+\end{cor}
+\begin{proof}
+   Both VNHCs satisfy Theorem \ref{thm:vlp-energy-stabilization} because
    \[
@@
       \Delta l \sin^2(2\theta) \geq 0
+      ,
    \] 
-   The VNHC \(l = L^{-}_\frac{\pi}{2}(\theta)\) dissipates energy
-   from the VLP, since 
+   and
    \[
@@
       - \Delta l \sin^2(2\theta) \leq 0
+      ,
    \] 
-\end{cor}
+   where the inequalities are strict everywhere except the coordinate axes.
+\end{proof}
 
-The types of VNHCs which satisfy Theorem \ref{thm:vlp-energy-stabilization} are
-illustrated by Figure \ref{fig:vlp-energy-in-out}. 
-To stabilize specific energy level sets, a basic approach is to switch
-between injection and dissipation VNHCs depending on the current energy level.
+The class of VNHCs which satisfies Theorem \ref{thm:vlp-energy-stabilization} is
+illustrated graphically in Figure \ref{fig:vlp-energy-in-out}. 
+To stabilize specific energy level sets, one simple approach is to switch
+between injection and dissipation VNHCs when the momentum \(p\) reaches
+a pre-determined value at the bottom of the swing.
 For the VNHCs we designed in this chapter, this means toggling
@@
    \includestandalone[width=0.5\textwidth]{images/vlp_energy_in_out}
-   \caption{A VNHC of the form \(l = L(\theta)\) where \(L(\theta)\)
+   \caption{Any VNHC of the form \(l = L(\theta)\) where \(L(\theta)\)
       is entirely contained within
-      the green (yellow) regions will inject (dissipate) energy into the VLP
-      because it will satisfy Theorem \ref{thm:vlp-energy-stabilization}.}
+      the green (yellow) regions will inject (dissipate) energy.}
       \label{fig:vlp-energy-in-out}
@@
 
-Theorem \ref{thm:vlp-energy-stabilization} also provides an alternate
-explanation for why the ``optimal" VNHC \(l^\star(\theta)\) works so well at
-injecting energy: it maximizes the derivative of \(E_\text{avg}\) under the
-restriction \(l(\theta) \in [\underline{l},\overline{l}]\), so that the orbit in the
-\((q,p)\)-plane diverges from the origin as fast as possible.
+Theorem \ref{thm:vlp-energy-stabilization} provides an alternate
+explanation for why the optimal pumping strategy \(L^\star(\theta)\) works
+so well at injecting energy: it maximizes the derivative of \(E_\text{avg}\)
+under the restriction \(l \in [\underline{l},\overline{l}]\), so that the orbit
+in the \((q,p)\)-plane diverges from the origin as fast as possible.
 
-Let us define \((l^\star)^{-}(\theta)\) by swapping the order of 
-\(\underline{l}\) and \(\overline{l}\) in \(l^\star(\theta)\). Since this 
-\textit{minimizes} the derivative of \(E_\text{avg}\) under the restriction
-\(l(\theta) \in [\underline{l},\overline{l}]\), we can predict that 
-\((l^\star)^{-}(\theta)\) is the VNHC representation of an optimal 
-energy dissipation controller.
-This is, in fact, true: \cite{pumping_swing_standing_squatting} showed 
+Let us define \((L^\star)^{-}(\theta)\) by swapping the order of 
+\(\underline{l}\) and \(\overline{l}\) in \(L^\star(\theta)\): 
+\[
+   (L^\star)^-(\theta) := \begin{cases}
+      \underline{l} & \theta \in [-\frac{\pi}{2},0[ \cup [\frac{\pi}{2}, \pi[ \\
+      \overline{l} & \theta \in [-pi, -\frac{\pi}{2}[ \cup [0,\frac{\pi}{2}[ 
+      .
+   \end{cases}
+\]
+Since this function \textit{minimizes} the derivative of \(E_\text{avg}\) under
+the restriction \(l \in [\underline{l},\overline{l}]\), one might predict that 
+\((L^\star)^{-}(\theta)\) is the optimal energy dissipation strategy for the VLP.
+This is, in fact, true. \citet{pumping_swing_standing_squatting} showed 
 that squatting at the lowest point of a swing and standing at the highest
-point (instead of standing and squatting resp.) produces the
+(instead of standing and squatting, resp.) produces the
 time-optimal trajectory for \textit{stopping} a standing swing.
@@
 All together, the theory developed in this chapter shows that VNHCs can
-replicate the time-optimal pumping strategy performed by humans. 
+replicate the time-optimal pumping/dissipation strategies performed by humans on
+swings.
 Furthermore, we see that VNHCs are a powerful tool for creating simple energy

### 2020-10-07 latex/vnhcs.tex
Added new labels for the VNHC chapter
Removed comments and added summary of results

@@
 
-\chapter{Development of Virtual Nonholonomic Constraints}
+\chapter{Development of Virtual Nonholonomic Constraints}\label{ch:vnhcs}
 % TODO: Brief introduction before preliminaries
@@
 \(q = (q_u,q_a)\) and \(p = (p_u, p_a)\) and has the dynamics
-\begin{align}\label{eqn:simply-actuated-dynamics}
+\begin{align}\label{eqn:simply-actuated-hamiltonian}
     \mathcal{H}(q,p) &= p\tpose \Minv(q) p + V(q)
     , \\
+    \label{eqn:simply-actuated-dynamics}
     &\begin{cases}
@@
 \section{Summary of Results}
-%TODO: summarize the assumptions and results 
-\textbf{TODO: summarize the assumptions and results }
+In this chapter, we developed the framework of virtual nonholonomic constraints
+for underactuated Hamiltonian mechanical systems.
+We made the following assumptions:
+\begin{enumerate}
+    \item The input matrix \(B(q) \equiv B \in \R^{n \times k}\) is constant and
+        full rank.
+    \item The input matrix has a left-annihilator 
+        \(B^\perp \in \R^{(n-k)\times n}\). 
+    \item The annihilator matrix \(B^\perp\) is right semi-orthogonal.
+\end{enumerate}
+These assumptions allowed us to define a canonical change of coordinates
+into the simply actuated coordinates 
+\((q,p) \in \mathcal{Q} \times \mathcal{P}\),
+where \(q = (q_u,q_a)\) and \(p = (p_u,p_a)\).
 
-%\section{Virtual Nonholonomic Constraints}\label{sec:vnhcs}
-%
-%%----- Motivation -----%
-%\subsection{Motivation}
-%\textbf{TODO: Why do we bother with Hamiltonian? Why can't we do virtual nonholonomic 
-%constraints in Lagrangian? What are some use-cases where VHCs don't work?}
-%
-%%---------- Hamiltonian from Lagrangian ----------%
-%\subsection{Review: Hamiltonian Systems}
-%\textbf{TODO: What is a Hamiltonian system and why does it matter?}
-%
-%One can compute the Hamiltonian of a system by performing a Legendre transform
-%on its Lagrangian \textbf{TODO: citation for legendre transform}.
-%First, define the conjugate of momenta for \(q\) by 
-%\begin{equation*}
-%p = \frac{\partial \mathcal{L}}{\partial \dot{q}} \in \mathbb{R}^n
-%\end{equation*}
-%Then, the Legendre transform is performed by taking
-%\begin{equation*}
-%\mathcal{H}(q,p) = p^T \dot{q} - \mathcal{L}(q,\dot{q})
-%\end{equation*}
-%
-%For a mechanical system (\ref{eqn:lagrangian}), the conjugate of momenta for
-%\(q\) is given by 
-%\begin{equation*}
-%    p = M(q)\dot{q}
-%\end{equation*}
-%which means the Hamiltonian of the system is given by
-%the total mechanical energy \(E\) (\ref{eqn:hamiltonian} in 
-%\((q,p)\) coordinates.
-%\begin{equation}\label{eqn:hamiltonian}
-%\mathcal{H}(q,p) = E(q,p) = \frac{1}{2} p^T \Minv(q) p + V(q)
-%\end{equation}
-%Note that \(M(q)\) is the inertia matrix and \(V(q)\) is the potential for
-%the Hamiltonian system. This is simply a trick to distinguish them notationally
-%from the Lagrangian \(D(q)\) and \(P(q)\); they are, in fact, identical in their
-%contents.
-%
-%The equations of motion for the system in Hamiltonian coordinates is given by
-%\begin{align}\label{eqn:hamiltionian_eom}
-%\begin{split}
-%\dot{q} &= \frac{\partial \mathcal{H}}{\partial p} = \Minv(q) p \\
-%\dot{p} &= -\frac{\partial \mathcal{H}}{\partial q} + B \tau
-%\end{split}
-%\end{align}
-%
-%\textbf{Why bother looking at Hamiltonian systems? What is the intuition behind
-%these equations of motion?}
-%
-%%---------- Hamiltonian VNHCs ----------%
-%\subsection{Hamiltonian Virtual Nonholonomic Constraints}
-%While VHCs are still possible in the Hamiltonian framework, the assumptions
-%required to make this work are slightly different. Rather than deriving
-%Hamiltonian VHCs directly, we will produce results for nonholonomic constraints
-%first as VHCs are a special case of this new framework.
-%
-%Suppose the mechanical system has degree of underactuation one, so that coordinates of the system can be split into an unactuated component \(q_u \in [\mathbb{R}]_T, \, T \in \mathbb{R}_{>0}\) which is not influence by control, along with an actuated component \(q_a\); that is, suppose \(B\) is of the form \(B(q) = [0_m, B_1^T(q), \ldots, B_n^T(q)]^T, \, B_i^T(q) \in \mathbb{R}^{n-1}\) and \(\tau \in \mathbb{R}^{n-1}\).
-In this case, \(q = (q_u, q_a)^T\) and the equations of motion become
-%\begin{align}\label{eqn:unactuated_actuated_eom}
-%\begin{split}
-%\dot{q_u} &= e_1^T \Minv(q) p \\
-%\dot{p_u} &= -p^T\frac{\partial M}{\partial q_u} p - \partial_{q_u}V(q) \\
-%\dot{q_a} &= 
-%\begin{bmatrix}
-%0 \cdots 0 \\
-%I_{n-1} \\
-%\end{bmatrix} \Minv(q) p \\
-%\dot{p_a} &= -p^T\frac{\partial M}{\partial q_a} p - \nabla{q_a}V(q) + 
-%\begin{bmatrix}
-%B_1(q) \\
-%\vdots \\
-%B_n(q)
-%\end{bmatrix} \tau
-%\end{split}
-%\end{align}
-%
-%Now we can begin to talk about Virtual Nonholonomic Constraints. In a similar fashion to what was defined for VHCs, let us first define the goal of these new virtual constraints.
-%
-%\begin{defn}
-%A relation \(h \in C^2\left(\mathcal{Q}\times \mathbb{R}^n ; \mathbb{R}^k\right)\) with \(h(q,p) = 0\) is a \textbf{virtual nonholonomic constraint (VNHC) of order k} if there exists a feedback control \(\tau(q,p)\) which stabilizes the constraint manifold
-%\[
-%\Gamma = \left\{(q,p) | h(q,p) = 0, dh_q \dot{q} + dh_p \dot{p} = 0\right\}
-%\]
-%\end{defn}
-%Define the output of the system to be \(e = h(q,p)\).
-We would like to find \(\tau(q,p)\) which drives \(e\) to zero to stabilize our constraint manifold \(\Gamma\).
-To accomplish this, we will input-output linearize (\ref{eqn:unactuated_actuated_eom}) to find \(\ddot{e} = -k_p e - k_d \dot{e}\) with \(k_p, k_d \in \mathbb{R}_{> 0}\).
-%
-%To characterize a certain class of VNHCs, let us make the following assumption.
-%\begin{assm}\label{assm:vnhc_is_on_qu_pu}
-%We assume our relation \(h\) is of the form \(h(q,p) = q_a - f(q_u,p_u)\) for some \(f \in C^2\left([\mathbb{R}]_T \times \mathbb{R} ; \mathbb{R}^{n - 1}\right)\).
-%\end{assm}
-%
-%Now we solve for \(\tau\) by finding \(\ddot{e}\).
-%\begin{align*}
-%    e &= h(q,p) = qa - f(q_u,p_u)\\
-%    \Rightarrow \dot{e} &= \dot{q_a} - df_{q_u}\dot{q_u} -df_{p_u}\dot{p_u} \\
-%    &= [-df_{q_u} I_{n-1}]\dot{q} - df_{p_u} \dot{p_u} \\
-%    &= dh_q \Minv(q) p - df_{p_u}\left( -\frac{1}{2}p^T \frac{\partial \Minv(q)}{\partial q_u} p - \partial_{q_u}V(q) \right) \\
-%    &= dh_q \Minv(q) p + \frac{1}{2}df_{p_u} p^T \frac{\partial \Minv(q)}{\partial q_u} p + df_{p_u}\partial_{q_u}V(q)
-%\end{align*}
-%The control input \(\tau\) only appears in \(\dot{p_a}\) (see (\ref{eqn:unactuated_actuated_eom})).
-To simplify the analysis, terms in \(\ddot{e}\) which do not depend on \(\dot{p_a}\) explicitly are lumped together under the symbol \((*)\):
-%\begin{align*}
-%    \ddot{e} &= dh_q \Minv(q) \dot{p} + df_{p_u} p^T \frac{\partial \Minv(q)}{\partial q_u} \dot{p} + (*) \\
-%    &= (dh_q \Minv(q) + df_{p_u} p^T \frac{\partial \Minv(q)}{\partial q_u})B\tau + (*) \\
-%    &= (dh_q \Minv(q) + dh_{p_u} p^T \frac{\partial \Minv(q)}{\partial q_u})B\tau + (*)
-%\end{align*}
-%
-%From the derivations above, one can solve for \(\tau\) iff the matrix on the left of \(\tau\) is full rank. Thus, for systems with degree of underactuation one we give the following definition.
-%\begin{defn}
-%A VNHC \(h(q,p) = 0\) of order \(n - 1\) is \textbf{regular} if \(dh_{p_a} = 0\), \(dh_{q_a} = (1 \ldots 1)^T\), and 
-%\[
-%\text{rank}\left\{ (dh_q \Minv(q) + dh_{p_u} p^T \frac{\partial \Minv(q)}{\partial q_u})B\right\} = n - 1
-%\]
-%everywhere on the constraint manifold \(\Gamma\).
-Equivalently, a VNHC \(h\) of order \(n - 1\) is regular if it satisfies Assumption \ref{assm:vnhc_is_on_qu_pu} and system (\ref{eqn:unactuated_actuated_eom}) with output \(e = h(q,p)\) is of relative degree \(\{2,2,\ldots,2\}\) everywhere on \(\Gamma\).
-%\end{defn}
-%
-%In general, \(\dot{e}\) is a function of \(q_u\) and \(p = (p_u,p_a)^T\).
-Since the purpose of a regular VNHC is to fully parameterize \(\Gamma\) by \((q_u,p_u)\), it is essential that one can solve for \(p_a = p_a(q_u,p_u)\).
-Unfortunately this often cannot be done, since \(\dot{e}\) contains the quadratic term
-%\[
-%\frac{1}{2} df_{p_u} p^T \frac{\partial \Minv(q)}{\partial q_u} p 
-%\]
-%We can solve for \(p_a\) if this quadratic term does not exist.
-%\begin{assm}\label{assm:M_is_Mqa}
-%Assume \(\partial M(q) / \partial q_u = 0 \Leftrightarrow \partial \Minv(q) / \partial q_u = 0\)
-%\end{assm}
-%
-%Under Assumption \ref{assm:M_is_Mqa}, we get that the rank condition for \(h(q,p)\) to be a regular VNHC reduces to \(\text{rank}\left(dh_q \Minv B\right) = n - 1\).
-This is the same rank condition as required for Virtual Holonomic Constraints.
-%
-%Now we solve for \(p_a\) on the constraint manifold (when \(e = \dot{e} = 0\)):
-%\begin{align*}
-%    \dot{e} = dh_q \Minv(q)p + df_{p_u} \partial_{q_u}V(q) &= 0\\
-%    \Leftrightarrow dh_q \Minv(q)e_1 p_u + dh_q \Minv(q) \begin{bmatrix}
-%    0 & \cdots & 0 \\
-%    & I_{n-1} & \\
-%    \end{bmatrix} p_a &= -df_{p_u} \partial_{q_u}V(q) \\
-%\end{align*}
-%\begin{align*}
-%    \Leftrightarrow dh_q \Minv(q) \begin{bmatrix}
-%    0 & \cdots & 0 \\
-%    & I_{n-1} & \\
-%    \end{bmatrix} p_a = -\left(df_{p_u}\partial_{q_u}V(q) + dh_q \Minv(q)e_1 p_u\right)
-%\end{align*}
-%One can linearly solve for \(p_a\) if and only if the matrix in front of it is invertible.
-%
-%This leads us to a natural definition.
-%\begin{defn}\label{defn:solvability}
-%A VNHC \(h(q,p)\) is \textbf{solvable} (NOTE: actionable? what's a good name?) if
-%\[
-%\text{rank}\left(dh_q \Minv(q) \begin{bmatrix}
-%    0 & \cdots & 0 \\
-%    & I_{n-1} & \\
-%    \end{bmatrix}\right) = n - 1
-%\]
-%\end{defn}
-%
-%With this analysis and our new definition in hand, we can solve for the dynamics on the constraint manifold.
-%
-%\begin{thm}\label{thm:equation_for_pa}
-%Suppose assumptions \ref{assm:vnhc_is_on_qu_pu} and \ref{assm:M_is_Mqa} hold.
-%If a regular VNHC \(h(q,p) = q_a - f(q_u,p_u)\) is solvable, then the
-%parameterization for \(p_a\) on the constraint manifold is given by
-%\begin{align*}
-%p_a &= -\left(dh_q \Minv(q) \begin{bmatrix}
-%    0 & \cdots & 0 \\
-%    & I_{n-1} & \\
-%    \end{bmatrix}\right)^{-1}\left( df_{p_u}\partial_{q_u}V(q) + dh_q \Minv(q)e_1 p_u\right) \\
-%    &=: g(q_u,p_u)
-%\end{align*}
-%and the constrained dynamics on \(\Gamma\) are given by
-%\begin{align*}
-%    \dot{q_u} &= e_1^T \Minv(q_a) \begin{bmatrix}
-%    p_u \\
-%    p_a
-%    \end{bmatrix}\mid_{q_a = f(q_u,p_u), p_a = g(q_u,p_u)} \\
-%    \dot{p_u} &= -\partial_{q_u} V(q_u,q_a) \mid_{q_a = f(q_u,p_u)}
-%\end{align*}
-%\end{thm}
-%
-%Theorem \ref{thm:equation_for_pa} guarantees that, on \(\Gamma\), \(q_a\) is a
-%parameterized completely by \((q_u,p_u)\). Hence, the zero-dynamics on
-%\(\Gamma\) are always two-dimensional regardless of the original dimension
-%\(n\).
-%
-%%---------- Hamiltonian VHCS ---------%
-%\subsection{Restriction to Hamiltonian VHCs}
-%\textbf{TODO: Why can we not do the standard VHC approach for Hamiltonian? 
-%Show how to use the above to make it work}
-%
+We defined a virtual nonholonomic
+constraint as a \(C^2\) function 
+\(h : \mathcal{Q} \times \mathcal{P} \rightarrow \R^k\) which has no singular
+points on its constraint manifold
+\[
+    \Gamma = \left\{(q,p) \mid h(q,p) = 0, dh_q \dot{q} + dh_p \dot{p} = 0\right\}
+    .
+\]
+
+We then showed that a VNHC 
+\(h : \mathcal{Q} \times \mathcal{P}_u \rightarrow \R^k\) is regular if and only
+if the square matrix
+\[
+    \left(
+    dh_q \Minv(q) - 
+    dh_{p_u} (\Id{n-k} \otimes p\tpose)\nabla_{q_u}\Minv(q) 
+    \right)
+    \simpleB
+\]
+is invertible on \(\Gamma\).
+
+To find the explicit equations for the dynamics on \(\Gamma\) of a regular VNHC, 
+we made the following assumptions:
+\begin{itemize}
+    \item The inertia matrix satisfies 
+        \(\nabla_{q_u}M(q) = \Zmat{n(n-k) \times n}\).
+    \item On \(\Gamma\), one can solve linearly for \(q_a\) as a function of
+        \((q_u,p_u)\).
+\end{itemize}
+If these assumptions hold, one can solve for \(p_a = g(q_u,p_u)\) on \(\Gamma\).
+The constrained dynamics are then given by
+\((\dot{q_u}, \dot{p_u})\) subject to \(h(q,p_u) = 0\) and \(p_a = g(q_u,p_u)\).
+
+Finally, the benefit of using VNHCs is that they reduce the
+dimensionality of the system from \(2n\) equations of motion to \(2(n-k)\) of
+them.
+This can significantly reduce the complexity of analyzing the evolution of the
+system.
+
 %%/========== /Virtual Nonholonomic Constraints ==========/% 

### 2020-10-08 latex/vlp.tex
Edited VNHC up to definition of VNHCs

@@
 tells us they are described entirely by \((\dot{q},\dot{p})\) with \(l\)
-replaced by \(L(q,p)\):
+replaced by \(L(q,p)\),
 \begin{align}\label{eqn:vlp-hamiltonian}

### 2020-10-08 latex/vnhcs.tex
Edited VNHC up to definition of VNHCs

@@
 The \textit{Hamiltonian} of the system in
-\(\{q,p\}\) coordinates is the ``Legendre transform"
+\((q,p)\) coordinates is the ``Legendre transform"
 (\ref{eqn:hamiltonian-legendre}) of the Lagrangian \cite{landau_mechanics},
@@
 and positive definite, it is invertible at each \(q \in \mathcal{Q}\).
-The Legendre transform (\ref{eqn:hamiltonian-legendre}) can be explicitly
-written as
-\[
-    \mathcal{H}(q,p) = p\tpose D\inv(q)p - \frac{1}{2}p\tpose D\inv(q) p + V(q)
-    ,
-\]
-which reduces to (\ref{eqn:hamiltonian}). 
-Finding the derivative of each conjugate of momentum yields
+The Legendre transform (\ref{eqn:hamiltonian-legendre}) becomes
+\begin{align*}
+    \mathcal{H}(q,p) &= p\tpose D\inv(q)p - \frac{1}{2}p\tpose D\inv(q) p +
+    P(q) \\
+         &= \frac{1}{2} p\tpose D^{-1}(q) p + P(q)
+    .
+\end{align*}
+Finding the derivative of each momentum coordinate yields
 \[
@@
 \]
-which can be collected into the vector form (\ref{eqn:hamiltonian-full-dynamics}) 
+which can be collected into vector form
 % TODO: Update Appendix info?
@@
 
-\citet{landau_mechanics} provides a useful result for showing whether a given
+\citet{landau_mechanics} provide a useful result for showing whether a given
 change of coordinates \((q,p) \to (Q,P)\) is a canonical transformation.
@@
 
-Later in this chapter we will define a particular type of coordinate change.
+Later in this chapter we will define a particular change of coordinates.
 The following Lemma allows us to prove it is a canonical transformation.
@@
                   &= \sum\limits_{m=1}^n A_{i,m}A_{j,m} - 0 \\
-                  &= \sum\limits_{m=1}^n A_{i,m}A\tpose_{m,j} \\
+                  &= \sum\limits_{m=1}^n A_{i,m}(A\tpose)_{m,j} \\
                   &= (A_i)\tpose (A\tpose)_j \\
@@
 \(B\tpose\) are right semi-orthogonal, it is easy to show that \(\mathbf{B}\) is
-orthogonal:
+orthogonal.
 \begin{proof}
@@
 \end{thm}
-\vspace{-4.5em} % For some reason not having this puts a large space before proof
 \begin{proof}
@@
 define the error term \(e = h(q,p)\). 
-Finding \(\tau(q,p)\) driving \(e \to 0\) and \(\dot{e} \to 0\)
-is sufficient for stabilizing \(\Gamma\). 
-If \(\tau\) appears after only \(r_i \geq 0\) derivatives of \(e_i\), 
-then that \(e_i\) is said to have \textit{relative degree} \(r_i\).
-We would then say \(e\) has relative degree \(\{r_1,\ldots,r_k\}\).
+If there exists some \(\tau(q,p)\) driving \(e \to 0\) and \(\dot{e} \to 0\),
+then \(\tau\) stabilizes \(\Gamma\). 
+If \(\tau\) first appears after \(r_i > 0\) derivatives of \(e_i\), 
+then \(e_i\) is said to have \textit{relative degree} \(r_i\).
+We then say \(e\) has relative degree \(\{r_1,\ldots,r_k\}\).
 
-One important note is that having relative degree \(1\) for any \(e_i\) means we
-may not be able to stabilize \(\Gamma\); we can guarantee \(h(q,p) \to 0\) 
-but not \(\dot{h}(q,p) = 0\). 
-Requiring that \(e\) have relative degree \(\{2,\ldots,2\}\) is much more
+With no further structure on \(h(q,p)\), the control input \(\tau\) usually
+appears after one derivative. 
+Unfortunately, if any \(e_i\) has relative degree \(1\) we may not in general
+be able to stabilize \(\Gamma\); we could guarantee \(h(q,p) \to 0\), but not
+necessarily \(\dot{h}(q,p) \to 0\). 
+Requiring \(e\) to have relative degree \(\{2,\ldots,2\}\) is much more
 useful, since it allows us to easily solve for a \(\tau\) which stabilizes
 \(\Gamma\). 
-This is a common requirementa in the VHC literature, so we will make use of it
+This is a common requirement in the VHC literature, so we will make use of it
 here by defining a special type of VNHC that satisfies this property.

### 2020-10-13 latex/vnhcs.tex
Fully completed / edited VNHCs, we are now at v1.

@@
 \chapter{Development of Virtual Nonholonomic Constraints}\label{ch:vnhcs}
-% TODO: Brief introduction before preliminaries
 
@@
 where the \textit{configuration manifold} \(\mathcal{Q}\) of the system is a
-so-called \textit{generalized cylinder}:
+so-called \textit{generalized cylinder}
 \[
@@
 \(\mathcal{L} : \mathcal{Q}\times\R^n \rightarrow \R\) defined by
-(\ref{eqn:lagrangian-general}) \cite{greenwood_dynamics}:
+(\ref{eqn:lagrangian-general}) \cite{greenwood_dynamics},
 \begin{equation}\label{eqn:lagrangian-general}
@@
 \begin{align*}
-    \mathcal{H}(q,p) &= p\tpose D\inv(q)p - \frac{1}{2}p\tpose D\inv(q) p +
-    P(q) \\
+    \mathcal{H}(q,p) &= p\tpose D\inv(q)p - \left( \frac{1}{2}p\tpose D\inv(q) p -
+    P(q) \right) \\
          &= \frac{1}{2} p\tpose D^{-1}(q) p + P(q)
@@
 which can be collected into vector form
-% TODO: Update Appendix info?
-by using the matrix Kronecker product (see Appendix ???).
+by using the matrix Kronecker product (see Appendix \ref{apx:kron-prod}).
 In sum, when the kinetic energy is quadratic the Hamiltonian system reduces to
@@
     The \textit{Poisson bracket} between the functions \(f(q,p)\) and \(g(q,p)\)
-    is defined as
+    is
     \begin{equation}\label{eqn:poisson-bracket}
@@
     A change of coordinates 
-    \((q,p) \to (Q,P)\) is a canonical transformation if
+    \((q,p) \to (Q,P)\) is a canonical transformation if and only if
     \begin{align*}
@@
     \begin{align*}
-        [Q_i,Q_j] &:= \sum\limits_{m = 1}^n \pdiff{Q_i}{p_m}\pdiff{Q_j}{q_m} - 
+        [Q_i,Q_j] &= \sum\limits_{m = 1}^n \pdiff{Q_i}{p_m}\pdiff{Q_j}{q_m} - 
         \pdiff{Q_i}{q_m}\pdiff{Q_j}{p_m} = 0
         , \\
-        [P_i,P_j] &:= \sum\limits_{m=1}^n \pdiff{P_i}{p_m}\pdiff{P_j}{q_m} -
+        [P_i,P_j] &= \sum\limits_{m=1}^n \pdiff{P_i}{p_m}\pdiff{P_j}{q_m} -
         \pdiff{P_i}{q_m}\pdiff{P_j}{p_m} = 0
@@
 \(A^0 \in \R^{1\times n}\) implies the left annihilator 
-\(B^\perp := A^0/\norm{A^0}\) will satisfy Assumption \ref{assm:B-perp}.
+\(B^\perp := A^0/\norm{A^0}\) will be a unit vector satisfying Assumption \ref{assm:B-perp}.
 
@@
 \]
-and defining the regular feedback transformation\(\tau = V T \hat{\tau}\), 
+and assigning the regular feedback transformation\(\tau = V T \hat{\tau}\), 
 we get a new input matrix for \(\hat{\tau} \in \R^k\) given by 
@@
 In particular, 
-\(\hat{B}\tpose \hat{B} = T\tpose \Sigma\tpose \Sigma\tpose T = \Id{k}\).
+\(\hat{B}\tpose \hat{B} = T\tpose \Sigma\tpose \Sigma T = \Id{k}\).
 \end{proof}
@@
 In light of Lemma \ref{lemma:B-orthogonal}, there is no loss of generality
-making the following assumption.
+by making the following assumption.
 \begin{assm}\label{assm:B-orthogonal}
@@
 define the error term \(e = h(q,p)\). 
-If there exists some \(\tau(q,p)\) driving \(e \to 0\) and \(\dot{e} \to 0\),
-then \(\tau\) stabilizes \(\Gamma\). 
+If there exists some controller \(\tau(q,p)\) driving \(e \to 0\) and 
+\(\dot{e} \to 0\), then the same \(\tau\) will necessarily stabilize \(\Gamma\).
+
 If \(\tau\) first appears after \(r_i > 0\) derivatives of \(e_i\), 
 then \(e_i\) is said to have \textit{relative degree} \(r_i\).
-We then say \(e\) has relative degree \(\{r_1,\ldots,r_k\}\).
+Collecting the \(e_i\) together, we say that \(e\) is of relative degree
+\(\{r_1,\ldots,r_k\}\).
+With no further structure on \(h(q,p)\), the control input \(\tau\) usually
+appears after one derivative of \(e\); 
+unfortunately, if any \(e_i\) has relative degree \(r_i = 1\) we may not
+be able to stabilize \(\Gamma\). 
+We could guarantee \(h(q,p) \to 0\), but we could not in general guarantee 
+\(\dot{h}(q,p) \to 0\).
 
-With no further structure on \(h(q,p)\), the control input \(\tau\) usually
-appears after one derivative. 
-Unfortunately, if any \(e_i\) has relative degree \(1\) we may not in general
-be able to stabilize \(\Gamma\); we could guarantee \(h(q,p) \to 0\), but not
-necessarily \(\dot{h}(q,p) \to 0\). 
 Requiring \(e\) to have relative degree \(\{2,\ldots,2\}\) is much more
-useful, since it allows us to easily solve for a \(\tau\) which stabilizes
-\(\Gamma\). 
-This is a common requirement in the VHC literature, so we will make use of it
-here by defining a special type of VNHC that satisfies this property.
+useful, since it allows us to easily solve for a controller stabilizing \(\Gamma\).
+This kind of relative degree requirement is already common in the VHC
+literature.
+Taking advantage of that precedent, we define a special type of VNHC that
+satisfies this property.
 
@@
     \begin{align*}
-        \ddot{e} &= \diff{}{t}\left[dh_q\right]\Minv(q) p + 
+        \ddot{e} &= \left(\diff{}{t}dh_q\right)\Minv(q) p + 
         dh_q \left(\sum\limits_{i=1}^n \pdiff{\Minv}{q_i}(q)\dot{q_i}\right)p + 
         dh_q \Minv(q) \dot{p} - \\
-         & \diff{}{t}\left[dh_{p_u}\right]
+         & \left(\diff{}{t}dh_{p_u}\right)
          \left(\frac{1}{2}\pudmat + \nabla_{q_u}V(q)\right) - \\
-         & dh_{p_u}\left(\frac{1}{2}\diff{}{t}\left[\pudmat\right] + 
-         \diff{}{t}\left[\nabla_{q_u}V(q)\right] \right)
+         & dh_{p_u}\left(\frac{1}{2}\diff{}{t}\left(\pudmat\right) + 
+         \left(\diff{}{t}\nabla_{q_u}V(q)\right) \right)
          .
@@
         \ddot{e} = (\star) - 
-        dh_{p_u}\left(\frac{1}{2}\diff{}{t}\left[\pudmat\right] +
-        dh_q \Minv(q) \dot{p}\right)
+        dh_{p_u}\left(\frac{1}{2}\diff{}{t}\left(\pudmat\right)\right) +
+        dh_q \Minv(q) \simpleB \tau
         .
@@
     Observe that the \(i^\text{th}\) row of
-    \(\frac{1}{2} \diff{}{t} \left[ \pudmat \right]\)
+    \(\frac{1}{2} \diff{}{t} \left( \pudmat \right)\)
     is given by
     \begin{align*}
-        \frac{1}{2} \diff{}{t}\left[p\tpose \pdiff{\Minv}{q_{u_i}}(q)p\right]
+        \frac{1}{2} \diff{}{t}\left(p\tpose \pdiff{\Minv}{q_{u_i}}(q)p\right)
         = p\tpose \pdiff{\Minv}{q_{u_i}}(q) \dot{p} + 
@@
     \begin{align*}
-        \frac{1}{2} \diff{}{t} \left[ \pudmat \right] =
+        \frac{1}{2} \diff{}{t} \left( \pudmat \right) =
         (\star) +
@@
     \end{align*}
-    where \((\star)\) is a function of \(q\) and \(p\).
+    where \((\star)\) is a continuous function of \(q\) and \(p\).
 % End of Proof
-    For shorthand, we'll write \(\ddot{e} = E(q,p) + H(q,p)\tau\) where \(E\)
-    and \(H\) are defined appropriately.
+    For shorthand, we'll write 
+    \[
+        \ddot{e} = E(q,p) + H(q,p)\tau
+        ,
+    \] 
+    where \(E\) and \(H\) are defined appropriately.
     From the definition of regularity, the VNHC \(h\) is regular 
@@
     if and only if the matrix premultiplying \(\tau\) is nonsingular, and hence
-    that \(H\) be invertible, proving the theorem.
+    that \(H\) is invertible. This proves the theorem.
 \end{proof}
@@
     Let \(h(q,p_u) = 0\) be a regular VNHC of order \(k\) with constraint
-    manifold \(\Gamma\). Suppose that on \(\Gamma\) one can solve linearly for
+    manifold \(\Gamma\). Suppose that on \(\Gamma\) one can solve for
     \(q_a\) as a function of \((q_u,p_u)\).
@@
     \begin{aligned}
-        g(q_u,p_u) &:= \\
+        &g(q_u,p_u) := \\
            &\left.\left(dh_q \Minv(q) \simpleB \right)\inv 
@@
 \begin{cor}\label{cor:2d-zero-dynamics}
-    Suppose \(\mathcal{H}\) is a ---- mechanical system in simply actuated
-    coordinates which has degree of underactuation one.
+    Suppose \(\mathcal{H}\) is a mechanical system in simply actuated
+    coordinates satisfying Assumption \ref{assm:inertially-actuated} 
+    which has degree of underactuation one.
     Let \(h(q,p_u) = 0\) be a regular VNHC of order \((n-1)\) of the form
-    \(h(q,p_u) = q_a - f(q_u,p_u)\)
-    where \(f\) is a suitably-defined smooth function.
+    \(h(q,p_u) = q_a - f(q_u,p_u)\),
+    where \(f\) is a suitably defined \(C^2\) function.
     Then \(dh_q = \begin{bmatrix} -\partial_{q_u}f & \Id{(n-1)}
     \end{bmatrix}\).
-    Defining \(e_1 := (1,0,\ldots,0) \in \R^n\), the actuated momentum is given
-    by (\ref{eqn:g-qupu}):
+    Defining \(e_1 := (1,0,\ldots,0) \in \R^n\), the actuated momentum is
     \begin{equation}\label{eqn:g-qupu}
@@
             \right|_{q_a =f(q_u,p_u)}
+        .
     \end{equation}
-    Since \(q_u \in \Rt{T}\) for some \(T \in ]0,\infty]\) and \(p_u \in \R\),
+    Since \(q_u \in \Rt{T_u}\) for some \(T_u \in ]0,\infty]\) and \(p_u \in \R\),
     the orbit \((q_u(t),p_u(t))\) traces out
-    a curve on the 2D-plane \(\Rt{T} \times \R\) which we call the \((q,p)\)-plane.
+    a curve on the 2D-plane \(\Rt{T_u} \times \R\) which we call the \((q,p)\)-plane.
 \end{cor}
@@
 We defined a virtual nonholonomic
-constraint as a \(C^2\) function 
-\(h : \mathcal{Q} \times \mathcal{P} \rightarrow \R^k\) which has no singular
-points on its constraint manifold
+constraint as a function 
+\(h \in C^2(\mathcal{Q}\times\mathcal{P}; \R^k)\) 
+which has no singular points on its constraint manifold
 \[
@@
 
-To find the explicit equations for the dynamics on \(\Gamma\) of a regular VNHC, 
+To find the explicit equations for constrained dynamics of a regular VNHC, 
 we made the following assumptions:
@@
         \(\nabla_{q_u}M(q) = \Zmat{n(n-k) \times n}\).
-    \item On \(\Gamma\), one can solve linearly for \(q_a\) as a function of
+    \item On \(\Gamma\), one can solve for \(q_a\) as a function of
         \((q_u,p_u)\).
@@
 
-Finally, the benefit of using VNHCs is that they reduce the
-dimensionality of the system from \(2n\) equations of motion to \(2(n-k)\) of
-them.
-This can significantly reduce the complexity of analyzing the evolution of the
-system.
+Finally, we saw the benefit of using VNHCs is that they reduce the
+dimensionality of the system from \(2n\) equations of motion to \(2(n-k)\)
+equations significantly reduces the complexity of analyzing large systems.
+In particular, if the system has degree of underactuation one, the dynamics
+reduce to a \(2\)D system on the ``\((q,p)\)-plane" \(\Rt{T}\times\R\).
 

### 2020-10-13 latex/vnhcs.tex
Fixed a typo iin the summary of VNHCs

@@
 dimensionality of the system from \(2n\) equations of motion to \(2(n-k)\)
-equations significantly reduces the complexity of analyzing large systems.
+equations, which significantly reduces the complexity of analyzing large systems.
 In particular, if the system has degree of underactuation one, the dynamics

### 2020-10-13 latex/vlp.tex
Edited VLP up to version 1, not including sims

--- 2020-10-07 latex/vlp.tex
@@
 This makes the VLP into an underactuated mechanical system with degree of
-underactuation 1. 
+underactuation one. 
 It is also a useful assumption because it means \((\mathbf{q},\mathbf{p})\) 
@@
 
-Let us define the VNHC \(l = L(q,p)\) -- by which we mean we are
+Let us define the VNHC \(l = L(q,p)\), by which we mean we are
 actually defining the VNHC \(h(\mathbf{q},\mathbf{p}) = l - L(q,p) = 0\) of
@@
 The VLP satisfies
-\(\nabla_q \Minv(\mathbf{q}) = \Zmat{2 \times 2}\);
-therefore, \(h(\mathbf{q},\mathbf{p})\) is a regular VNHC whenever 
+\(\nabla_q \Minv(\mathbf{q}) = \Zmat{2 \times 2}\).
+By Theorem \ref{thm:vnhc-regularity},
+\(h(\mathbf{q},\mathbf{p})\) is a regular VNHC whenever 
 \(L(q,p)\) is \(C^2\) because
@@
       \sign{x^3 - y^3} = \sign{x-y}
+      .
    \]
@@
            + m g l_\text{avg} (1-\cos(q))
+      .
    \]
@@
    \]
-   The sign of \(\sin(q)p\) is dependent only on the quadrant in the
-   \((q,p)\)-plane, so one can show it is equal to the sign of
-   \(\sin(2\theta)\). 
+   The sign of \(\sin(q)p\) depends only on the quadrant in the
+   \((q,p)\)-plane. In fact, it is easy to show that
+   \[
+      \sign{sin(q)p} = \sign{\sin(2\theta)}
+      .
+   \]
    Furthermore, by Lemma \ref{lemma:sign-of-cube} we have
    \[ 
-       \sign{l_\text{avg}^3 - L(\theta)^3} = \sign{l_\text{avg} - L(\theta)}
+       \sign{L(\theta)^3 - l_\text{avg}^3} = \sign{L(\theta) - l_\text{avg}}
        .
@@
    \end{align*}
-   Thus, \(E_\text{avg}\) is a Lyapunov function and the negative-time system is
+   Thus, \(E_\text{avg}\) is a Lyapunov function, which means the negative-time system is
    stable \cite{lyapunov}. 
@@
 that squatting at the lowest point of a swing and standing at the highest
-(instead of standing and squatting, resp.) produces the
+(instead of standing and squatting, respectively) produces the
 time-optimal trajectory for \textit{stopping} a standing swing.

### 2020-10-13 latex/images/vlp_energy_in_out.tex
Updated labels on all VLP images

@@
  xmin = -3.1415, xmax = 3.1415,
- ylabel = $l(\theta)$,
+ ylabel = $l$,
  y label style={at={(axis description cs:0.5,1.1)},anchor = north}, 

### 2020-10-13 latex/acrobot.tex
Edited acrobot up to dynamics

@@
 The acrobot is a two-link pendulum, actuated at the center joint (as in Figure
-\ref{fig:acrobot-model}. Since its first description in 1990 
-\cite{nonlinear_controllers_nonintegrable_acrobot}, the
-acrobot has become a benchmark problem in control theory; it is an
-underactuated mechanical system which produces
-complex nonlinear motion from an easy-to-describe model.
-The acrobot models a gymnast on a bar,
-since it represents a torso (top link) and legs (bottom link) with motion
-generated by the swinging of the legs at the hips. It is also one of the
-simplest models for a biped walking robot
+\ref{fig:acrobot-model}). 
+Since its first description in 1990
+\cite{nonlinear_controllers_nonintegrable_acrobot}, the acrobot has become a
+benchmark problem in control theory; 
+it is an underactuated mechanical system which produces complex nonlinear motion
+from an easy-to-describe model.
+The acrobot models a gymnast on a bar, since it represents a torso (top link)
+and legs (bottom link) with motion generated by the swinging of the legs at the
+hips. 
+It is also one of the simplest models for a biped walking robot
 \cite{toward_framework_biped_locomotion}.
@@
 Controlling the acrobot is a nontrivial task, since it is not feedback
-linearizable \cite{nonlinear_controllers_nonintegrable_acrobot}. Many
-researchers have studied the swing-up problem of driving the acrobot to its
-equilibrium point above the bar using partial feedback linearization
+linearizable \cite{nonlinear_controllers_nonintegrable_acrobot}. 
+Several researchers have studied the swing-up problem of driving the acrobot to
+its equilibrium point above the bar using partial feedback linearization
 \cite{swingup_problem_acrobot}, energy-based control
@@
 
-In gymnastics terminology, a ``giant" is when the gymnast performs full
-rotations around the bar with their body almost fully extended
-\cite{usagym_giant}. We are interested in studying the energy stabilization
-problem for the acrobot by using VNHCs to generate giant motion. The control of
-giant motion for the acrobot has been studied thoroughly 
+In gymnastics terminology, a ``giant" is the motion a gymnast performs to
+achieve full rotations around the bar with their body almost fully extended
+\cite{usagym_giant}. 
+We are interested in studying the energy stabilization problem for the acrobot
+by using VNHCs to generate giant motion. 
+The control of giant motion for the acrobot has been studied thoroughly
 \cite{energy_pumping_robotic_swinging, swingup_giant_acrobot,
@@
 \cite{dynamical_servo_acrobot_vc, control_giant_two_link_gymnastic_robot,
-xingbo_thesis}. However, these controllers are neither intuitive nor easy to design:
+xingbo_thesis}. 
+However, these controllers are neither intuitive nor easy to
+design:
 \cite{control_giant_two_link_gymnastic_robot} defines a constraint by inverting
-a trajectory in time onto the state space; \cite{dynamical_servo_acrobot_vc}
-requires a cascade controller to stabilize both a constraint and a desired limit
-cycle in the state space; and \cite{xingbo_thesis} enforces the giant by adding
-an extra state to estimate the velocity, which increases the dimensionality of
-the problem as a crude approach to using VNHCs.
+a trajectory in time onto the state space; 
+\cite{dynamical_servo_acrobot_vc} requires a cascade controller to stabilize
+both a constraint and a desired limit cycle in the state space; 
+and \cite{xingbo_thesis} enforces the giant by adding an extra state to estimate
+the velocity, which increases the dimensionality of the problem in a crude
+approach to using VNHCs.
 
@@
 motion. We will prove that this VNHC injects energy into the acrobot;
-in the process of completing this proof,
-we will describe a promising method which may be used to generate desirable 
-VNHCs in the future.
+in the process of completing this proof, we will arrive at a promising method
+which might be useful for generating desirable VNHCs in the future.
 
@@
 gymnast hanging on a horizontal bar, where the ``torso" has moment of
-inertia \(J_u\) and the ``leg" link has moment of inertia \(J_a\) (each with respect
+inertia \(J_u\) and the ``leg" has moment of inertia \(J_a\) (each with respect
 to their own center of mass).
@@
       m_al_{c_a}^2 + J_a
-    \end{bmatrix} \\
+    \end{bmatrix} 
+    , \\
     \label{eqn:general-acrobot-potential}
     P(q) &= g\left(m_al_{c_a}(1 - \cos(q_u+q_a)) + 
-        (m_al_u + m_ul_{c_u})(1-\cos(q_u))\right) \\
+        (m_al_u + m_ul_{c_u})(1-\cos(q_u))\right) 
+    , \\
     B(q) &= \begin{bmatrix} 0 \\ 1 \end{bmatrix}
+    .
 \end{align}
 
-While this is the most general representation of an acrobot, the dynamics are
-unwieldy to the point where it becomes difficult to prove the results in this
-chapter analytically.
+While this is the most general representation of an acrobot, the Hamiltonian
+dynamics become too unwieldy for analytically proving results in this chapter.
 We will therefore simplify the dynamics by assuming the acrobot is comprised of
@@
 We call this type of acrobot a \textbf{simple} acrobot. We will also ignore any
-frictional forces at both the hips and shoulder. Finally, it is important to note
-that a real gymnast cannot swing their hips in full circles; for this reason, we
-assume that \(q_a \in [-\bar{q}_a, \bar{q}_a]\) where 
-\(\bar{q}_a \in ]-\pi, \pi[\). On mechanical acrobots, this can be taken as a
-control parameter.
+frictional forces at both the hips and shoulder. 
+Finally, it is important to note that a real gymnast cannot swing their hips in
+full circles; 
+for this reason, we assume that \(q_a \in [-\bar{q}_a, \bar{q}_a]\) where 
+\(\bar{q}_a \in ]-\pi, \pi[\). 
+If the acrobot is a robot as opposed to the model of a gymnast, 
+\(\bar{q}_a\) can be used as a control parameter when designing VNHCs.
 
@@
 
-Given the assumptions listed, the moments of inertia \(J_u\) and \(J_a\) of the
-rods vanish. Furthermore, \(l_{c_u} = l_{c_a} = l_u = l_a = l\) and
-\(m_u = m_a = m\). Reducing
-(\ref{eqn:general-acrobot-inertia}-\ref{eqn:general-acrobot-potential}) with this
-knowledge yields the following ``simple" inertia matrix and potential function:
+Taking these assumptions on the model of a simple acrobot, 
+the moments of inertia \(J_u\) and \(J_a\) of the rods vanish. 
+Furthermore, \(l_{c_u} = l_{c_a} = l_u = l_a = l\) and
+\(m_u = m_a = m\). 
+Reducing
+(\ref{eqn:general-acrobot-inertia}-\ref{eqn:general-acrobot-potential})
+yields the simplified inertia matrix \(D_s\) and potential function \(P_s\),
+where
 \begin{align}
-    D(q) &= \begin{bmatrix}
+    D_s(q) &= \begin{bmatrix}
         ml^2\left(3+2\cos(q_a)\right) & 
@@
         ml^2
-    \end{bmatrix} \\
-    P(q) &= -mgl\left(2\cos(q_u)+\cos(q_u+q_a)\right)
+    \end{bmatrix} 
+    , \\
+    P_s(q) &= -mgl\left(2\cos(q_u)+\cos(q_u+q_a)\right)
+    .
 \end{align}
@@
 
-Defining \(M(q) := D(q)\) and \(V(q) := P(q)\), we find the conjugate of momenta 
+Defining \(M(q) := D_s(q)\) and \(V(q) := P_s(q)\), we find the conjugate of momenta 
 is \(p = (p_u, p_a) = M(q)\dot{q}\).

### 2020-10-14 latex/acrobot.tex
Added the section on previous constraints in Acrobot chapter

@@
 % rigorously that we can stabilize any energy level set on the acrobot.
+As we saw at the beginning of this chapter, there are some existing approaches
+to generating giant motion for the acrobot.
+We will examine a few of these approaches and examine whether they are viable
+candidates on which to base a VNHC, akin to how we designed the VLP constraint
+in Chapter \ref{ch:vlp}.
+
+An instinctive approach to controlling the acrobot is to model it as a
+variable-length pendulum as in Figure \ref{fig:acrobot-ecm}, 
+since the length of its equivalent center of mass (ECM) changes with the leg motion. 
+For instance, \citet{swingup_giant_acrobot} design a desired trajectory for the
+ECM, then determine which leg angles \(q_a(t)\) are required to generate that
+trajectory.
+
+\begin{figure}
+    \centering
+    \includestandalone[width=0.5\textwidth]{images/acrobot_ecm}
+    \caption{A simple acrobot modelled as a VLP with equivalent center of mass \(2m\). 
+        The length of the VLP changes according to \(q_a\).}
+    \label{fig:acrobot-ecm}
+\end{figure}
+
+Unfortunately, the VLP is not a true representation of the acrobot.
+The effective length of the ECM is 
+\[
+    l_e(q_a) := l\sqrt{\frac{5}{4} + cos(q_a)}
+    ,
+\]
+and its effective angle is
+\[
+    q_e := \arctan_2(s_u + \frac{1}{2}s_{ua}, -c_u - \frac{1}{2}c_{ua})
+    .
+\]
+There are two things of note here. 
+First, Figure \ref{fig:acrobot-vlp-symmetry} shows there are two configurations
+of the acrobot which give the same effective angle and length.
+This means the acrobot and the VLP are not equivalent representations;
+designing a VNHC that injects energy using the ECM may not produce human-like
+leg motion on the acrobot.
+Second, the torque input \(\tau\) applied to \(q_a\) affects both
+\(l_e\) and \(q_e\). 
+In the VLP model from Chapter \ref{ch:vlp}, the control input only
+affects the dynamics of the length variable.
+If we wanted to design a VNHC to provably inject energy into the acrobot, we
+would need to fully redesign the VLP constraint with a model where the control
+input has a direct effect on the angle.
+
+\begin{figure}
+    \centering
+    % TODO: Generate a figure with the ECM at one position, and two mirrored
+    % acrobots (dashed) around it to show how the VLP and acrobot are not
+    % equivalent.
+%    \includestandalone[width=0.5\textwidth]{}
+    \caption{The equivalent center of mass of the acrobot generally has two configurations
+        which correspond to the same effective length and angle. These
+        configurations are symmetric about the line connecting the pivot to the
+        ECM.}
+    \label{fig:acrobot-vlp-symmetry}
+\end{figure}
+
+Since the VLP representation of the acrobot requires redesigning a VLP
+controller, and the resulting motion may not even be realistic giant motion,
+this model is clearly ineffective for our purposes.
+
+We turn to the thesis of \citet{xingbo_thesis}, who designs a VHC to enforce a
+so-called ``tap" motion, with the purpose of injecting energy into the acrobot. 
+First, he defines a compensator variable \(s\) which tracks \(\dot{q}_u\), so
+that he can use the theory of VHCs with the extended configuration 
+\((q_u,q_a,s)\).
+He then finds \(h_1, h_2 \in \R_{>0}\) to define the
+normalized radius \(\rho\) and normalized angle \(\xi\) in the
+\((q_u, s)\)-plane.
+These normalized variables are given by
+\begin{align*}
+    \rho &:= \sqrt{h_1 q_u^2 + h_2 s^2}
+    , \\
+    \xi &:= \arctan_2(h_2 s, h_1 q_u)
+    . 
+\end{align*}
+He then sets the VHC to be \(h(q) = q_a - f_\text{rad}(\rho)f_\text{ang}(\xi)\)
+with the control parameters \(\bar{q}_u\) and \(\rho_0\), where
+\begin{align*}
+    f_\text{rad}(\rho) &:= \tanh^2(\rho/\rho_0)
+    , \\
+    f_\text{ang}(\xi) &:= 
+    \begin{cases}
+        0 & -\pi < \xi \leq 0 \\
+        \bar{q}_u \exp\left(1 - \frac{1}{1-(\frac{4\xi}{\pi} - 1)^2}\right) 
+          & 0 < \xi \leq \frac{\pi}{2} \\
+        0 & \frac{\pi}{2} < \xi \leq \pi
+        .
+    \end{cases}
+\end{align*}
+
+While his VHC shows promising experimental results, \citeauthor{xingbo_thesis}
+does not provide an analytical proof that the acrobot gains energy.
+His lack of analysis is tied to the fact that the constrained
+(Lagrangian) dynamics are incredibly complicated.
+In fact, it is even challenging to show that the constraint is regular.
+While we could very easily convert his VHC into a VNHC by replacing \(s\) with
+\(p_u\), we would run into the same problem. 
+Since we want our constraint to \textit{provably} inject energy, we choose to
+forgo this type of constraint.
 

### 2020-10-15 latex/acrobot.tex
More writing for acrobot 4.1-4.3

@@
 dynamics become too unwieldy for analytically proving results in this chapter.
-We will therefore simplify the dynamics by assuming the acrobot is comprised of
+To make the analysis tractable, we begin by assuming the acrobot is comprised of
 two massless rods of equal length \(l\), with equal point masses \(m\) at the
 tips (as in Figure \ref{fig:simple-acrobot-model}).
-We call this type of acrobot a \textbf{simple} acrobot. We will also ignore any
-frictional forces at both the hips and shoulder. 
-Finally, it is important to note that a real gymnast cannot swing their hips in
+We call this type of acrobot a \textbf{simple} acrobot. 
+We will also ignore any frictional forces at both the hips and shoulder. 
+Finally, it is important to note that a real gymnast cannot swing their legs in
 full circles; 
@@
     \mathcal{H}(q,p) &= \frac{1}{2}p\tpose \Minv(q) p -
-    mgl\left(2 c_u + c_{ua}\right)\\
+    mgl\left(2 c_u + c_{ua}\right)
+    , \\
      &\begin{cases}
@@
         \dot{p}_a =-\frac{1}{2}p\tpose \nabla_{q_a}\Minv(q) p
-        - mgl s_{ua} + \tau
+        - mgl s_{ua} + \tau,
     \end{cases} \nonumber
@@
     \end{bmatrix}
+    .
 \end{equation}
+The control input is a force \(\tau \in \R\) affecting only the dynamics of
+\(p_a\), representing a torque acting on the hip joint.
+This means \((q,p)\) are simply actuated coordinates, allowing us once again to
+apply the theory of VNHCs from Chapter \ref{ch:vnhcs}.
+
+Let us define the VNHC \(h(q,p) = q_a - f(q_u,p_u)\) of order 1, where
+\(f \in C^2\left(\mathbb{S}^1 \times \R; \R\right)\).
+To enforce regularity of \(h(q,p)\), we note that 
+\(\nabla_{q_u}\Minv(q) = \Zmat{2\times 2}\).
+Examining Theorem \ref{thm:vnhc-regularity}, our VNHC will be regular
+if the regularity matrix
+\[
+    dh_q \Minv(q) \begin{bmatrix}0 \\ 1 \end{bmatrix}
+    ,
+\]
+is of full rank on the constraint manifold \(\Gamma\).
+Since 
+\[
+    dh_q = \begin{bmatrix}
+        -\partial_{q_u} f & 1
+    \end{bmatrix}
+    ,
+\]
+the regularity matrix evaluates to the scalar equation
+\[
+    \frac{(1+c_a)\partial_{q_u}f(q_u,p_u) + (3+2c_a)}{ml^2(2-c_a^2)}
+\]
+which is full rank if and only if the numerator does not
+change sign.
+A sufficient condition for this is when \(f\) is a function solely of \(p_u\),
+so that \(\partial_{q_u}f = 0\).
+If this is the case, the regularity matrix is always positive and the VNHC 
+\(h(q,p)\) is regular.
+
+The acrobot is a noticeably more complex system than the VLP, as
+the dynamics are coupled because \(\Minv(q)\) is not diagonal.
+Because of this, the constrained dynamics of an arbitrary VNHC may not be easy
+to determine analytically, which will impact our ability to prove energy
+injection analytically.
+In the rest of this chapter, our goal is to design the function \(f(q_u,p_u)\)
+based on the natural human motion of a gymnast, with the caveat that the
+resulting VNHC be regular and that we can prove the constrained dynamics will
+inject energy into the acrobot.
 
@@
 % rigorously that we can stabilize any energy level set on the acrobot.
-As we saw at the beginning of this chapter, there are some existing approaches
-to generating giant motion for the acrobot.
-We will examine a few of these approaches and examine whether they are viable
-candidates on which to base a VNHC, akin to how we designed the VLP constraint
-in Chapter \ref{ch:vlp}.
+There are a few existing approaches to generating giant motion for the acrobot
+which may be viable candidates on which to base a VNHC, akin to how we designed
+the VLP constraint in Chapter \ref{ch:vlp}.
 
-An instinctive approach to controlling the acrobot is to model it as a
-variable-length pendulum as in Figure \ref{fig:acrobot-ecm}, 
-since the length of its equivalent center of mass (ECM) changes with the leg motion. 
-For instance, \citet{swingup_giant_acrobot} design a desired trajectory for the
-ECM, then determine which leg angles \(q_a(t)\) are required to generate that
-trajectory.
+An initial approach to controlling the acrobot is to model it as a
+variable-length pendulum by collapsing the two rods and masses into one
+equivalent center of mass (ECM), as in Figure \ref{fig:acrobot-ecm}.
+This seems a reasonable model reduction, since the length from the pivot to the
+ECM changes depending on the angle \(q_a\) of the leg.
+Indeed, \citet{swingup_giant_acrobot} use this approach to design a trajectory
+for the ECM, then determine which leg angles \(q_a(t)\) are required to generate
+that trajectory.
 
@@
 \]
-There are two things of note here. 
-First, Figure \ref{fig:acrobot-vlp-symmetry} shows there are two configurations
+There are two important thing to note here. 
+First, Figure \ref{fig:acrobot-vlp-symmetry} shows that for each pose of the
+VLP representation, there are two configurations
 of the acrobot which give the same effective angle and length.
@@
 leg motion on the acrobot.
-Second, the torque input \(\tau\) applied to \(q_a\) affects both
-\(l_e\) and \(q_e\). 
+
+Second, if we were to compute the conjugate of momenta
+\(p_{l_e}\) to \(l_e\) and \(p_e\) to \(q_e\), we would see the torque input
+\(\tau\) appearing in both of their dynamic equations.
 In the VLP model from Chapter \ref{ch:vlp}, the control input only
 affects the dynamics of the length variable.
-If we wanted to design a VNHC to provably inject energy into the acrobot, we
-would need to fully redesign the VLP constraint with a model where the control
-input has a direct effect on the angle.
+If we want to design a VNHC for this system, we cannot use any of the results
+from Chapter \ref{ch:vlp} because the VLP models don't match.
 
@@
 Since the VLP representation of the acrobot requires redesigning a VLP
-controller, and the resulting motion may not even be realistic giant motion,
+VNHC, and the resulting ECM motion may not produce realistic leg motion,
 this model is clearly ineffective for our purposes.
 
-We turn to the thesis of \citet{xingbo_thesis}, who designs a VHC to enforce a
-so-called ``tap" motion, with the purpose of injecting energy into the acrobot. 
+Let us turn to the thesis of \citet{xingbo_thesis}, who designs a VHC to enforce a
+so-called ``tap" motion with the purpose of injecting energy into the acrobot. 
 First, he defines a compensator variable \(s\) which tracks \(\dot{q}_u\), so
@@
 
-While his VHC shows promising experimental results, \citeauthor{xingbo_thesis}
-does not provide an analytical proof that the acrobot gains energy.
+While this constraint shows promising experimental results and it accurately
+emulates true human motion, \citeauthor{xingbo_thesis}
+does not provide an analytical proof that the acrobot will gain energy.
 His lack of analysis is tied to the fact that the constrained
-(Lagrangian) dynamics are incredibly complicated.
-In fact, it is even challenging to show that the constraint is regular.
+dynamics are incredibly complicated.
+In fact, just showing the constraint is regular is a challenging task.
 While we could very easily convert his VHC into a VNHC by replacing \(s\) with
 \(p_u\), we would run into the same problem. 
-Since we want our constraint to \textit{provably} inject energy, we choose to
-forgo this type of constraint.
+Since we want our constraint to \textit{provably} inject energy, we must forgo
+this type of constraint in favour of something less complex.
 

### 2020-10-15 latex/acrobot.tex
Edits to describe the phase space explicitly

@@
 \(p_a\), representing a torque acting on the hip joint.
-This means \((q,p)\) are simply actuated coordinates, allowing us once again to
+This means \((q,p)\) are simply actuated coordinates inside the phase space
+\(\mathcal{Q} \times \mathcal{P}\) where
+\[
+    \mathcal{Q} = \mathcal{Q}_u \times \mathcal{Q}_a = 
+    \mathbb{S}^ \times \mathbb{S}^1
+    ,
+\]
+and
+\[
+    \mathcal{P} = \mathcal{P}_u \times \mathcal{P}_a
+    = \R \times \R
+    .
+\]
+This allowg us once again to
 apply the theory of VNHCs from Chapter \ref{ch:vnhcs}.
@@
 Let us define the VNHC \(h(q,p) = q_a - f(q_u,p_u)\) of order 1, where
-\(f \in C^2\left(\mathbb{S}^1 \times \R; \R\right)\).
+\(f \in C^2\left(\mathcal{Q}_u \times \mathcal{P}_u; \mathcal{Q}_a\right)\).
 To enforce regularity of \(h(q,p)\), we note that 
@@
 the regularity matrix evaluates to the scalar equation
-\[
+\begin{equation}\label{eqn:regularity-matrix-acrobot}
     \frac{(1+c_a)\partial_{q_u}f(q_u,p_u) + (3+2c_a)}{ml^2(2-c_a^2)}
-\]
+    ,
+\end{equation}
 which is full rank if and only if the numerator does not
 change sign.
-A sufficient condition for this is when \(f\) is a function solely of \(p_u\),
-so that \(\partial_{q_u}f = 0\).
-If this is the case, the regularity matrix is always positive and the VNHC 
-\(h(q,p)\) is regular.
 
-The acrobot is a noticeably more complex system than the VLP, as
-the dynamics are coupled because \(\Minv(q)\) is not diagonal.
+A sufficient condition for regularity is when \(f\) is a function solely of \(p_u\),
+because \(\partial_{q_u}f = 0\)
+and (\ref{eqn:regularity-matrix-acrobot}) becomes strictly positive for all
+values of \(q_a\).
+
+The acrobot is noticeably more complex than the VLP, as
+the dynamics of \((q_u,p_u)\) and \((q_a,p_a)\) are coupled through \(\Minv(q)\).
 Because of this, the constrained dynamics of an arbitrary VNHC may not be easy
-to determine analytically, which will impact our ability to prove energy
-injection analytically.
+to write out.
 In the rest of this chapter, our goal is to design the function \(f(q_u,p_u)\)

### 2020-10-16 latex/acrobot.tex
Edited up to Previous Acrobot Constraints

@@
 In gymnastics terminology, a ``giant" is the motion a gymnast performs to
-achieve full rotations around the bar with their body almost fully extended
-\cite{usagym_giant}. 
-We are interested in studying the energy stabilization problem for the acrobot
-by using VNHCs to generate giant motion. 
-The control of giant motion for the acrobot has been studied thoroughly
-\cite{energy_pumping_robotic_swinging, swingup_giant_acrobot,
-control_giant_two_link_gymnastic_robot}, including several studies which use
-virtual (holonomic) constraints to achieve this behaviour
+achieve full rotations around the bar \cite{usagym_giant}. 
+We are interested in using VNHCs to generate giant motion, with the aim of
+stabilizing desired energy levels.
+The control of giant motion for the acrobot has been studied in
+\cite{energy_pumping_robotic_swinging, swingup_giant_acrobot}, 
+and some authors have used virtual holonomic constraints to achieve this
+behaviour
 \cite{dynamical_servo_acrobot_vc, control_giant_two_link_gymnastic_robot,
@@
 and \cite{xingbo_thesis} enforces the giant by adding an extra state to estimate
-the velocity, which increases the dimensionality of the problem in a crude
+velocity, which increases the dimensionality of the problem in a crude
 approach to using VNHCs.
@@
 In this chapter we will design a physically-intuitive VNHC which generates giant
-motion. We will prove that this VNHC injects energy into the acrobot;
-in the process of completing this proof, we will arrive at a promising method
-which might be useful for generating desirable VNHCs in the future.
+motion and prove the acrobot gains energy. 
+In the process of completing this proof, we will arrive at a promising method
+which might one day be useful for generating energy-injecting VNHCs on arbitrary
+mechanical systems.
 
@@
 
-While this is the most general representation of an acrobot, the Hamiltonian
-dynamics become too unwieldy for analytically proving results in this chapter.
-To make the analysis tractable, we begin by assuming the acrobot is comprised of
-two massless rods of equal length \(l\), with equal point masses \(m\) at the
-tips (as in Figure \ref{fig:simple-acrobot-model}).
-We call this type of acrobot a \textbf{simple} acrobot. 
-We will also ignore any frictional forces at both the hips and shoulder. 
+While this is the most general representation of an acrobot, the dynamics
+become unwieldy.
+To make rigorous analysis of these dynamics more tractable, we begin by assuming
+the acrobot is comprised of two massless rods of equal length \(l\), with equal
+point masses \(m\) at the tips.
+We call this a \textit{simple} acrobot, which is displayed in Figure
+\ref{fig:simple-acrobot-model}.
+We will also ignore any frictional forces at both the hip and shoulder joints. 
 Finally, it is important to note that a real gymnast cannot swing their legs in
@@
 \(\bar{q}_a \in ]-\pi, \pi[\).
-If the acrobot is a robot as opposed to the model of a gymnast, 
-\(\bar{q}_a\) can be used as a control parameter when designing VNHCs.
 
@@
 
-Taking these assumptions on the model of a simple acrobot, 
-the moments of inertia \(J_u\) and \(J_a\) of the rods vanish. 
-Furthermore, \(l_{c_u} = l_{c_a} = l_u = l_a = l\) and
+Since we are now working with a simple acrobot, 
+we have \(l_{c_u} = l_{c_a} = l_u = l_a = l\) and
 \(m_u = m_a = m\). 
+On top of this, the moments of inertia \(J_u\) and \(J_a\) of the rods vanish.
 Reducing
@@
 \[
-    \mathcal{Q} = \mathcal{Q}_u \times \mathcal{Q}_a = 
-    \mathbb{S}^ \times \mathbb{S}^1
+    \mathcal{Q} = \mathcal{Q}_u \times \mathcal{Q}_a := 
+    \mathbb{S}^1 \times \mathbb{S}^1
     ,
@@
     \mathcal{P} = \mathcal{P}_u \times \mathcal{P}_a
-    = \R \times \R
+    := \R \times \R
     .
 \]
-This allowg us once again to
-apply the theory of VNHCs from Chapter \ref{ch:vnhcs}.
+This allows us once again to apply the theory of VNHCs from Chapter
+\ref{ch:vnhcs}.
 
@@
 \(f \in C^2\left(\mathcal{Q}_u \times \mathcal{P}_u; \mathcal{Q}_a\right)\).
-To enforce regularity of \(h(q,p)\), we note that 
-\(\nabla_{q_u}\Minv(q) = \Zmat{2\times 2}\).
-Examining Theorem \ref{thm:vnhc-regularity}, our VNHC will be regular
+Since \(\nabla_{q_u}\Minv(q) = \Zmat{2\times 2}\),
+Theorem \ref{thm:vnhc-regularity} tells us that this VNHC will be regular
 if the regularity matrix
@@
 is of full rank on the constraint manifold \(\Gamma\).
-Since 
+Then, since 
 \[
@@
     \frac{(1+c_a)\partial_{q_u}f(q_u,p_u) + (3+2c_a)}{ml^2(2-c_a^2)}
-    ,
+    .
 \end{equation}
-which is full rank if and only if the numerator does not
+This is full rank on \(\Gamma\) if and only if the numerator does not
 change sign.
@@
 A sufficient condition for regularity is when \(f\) is a function solely of \(p_u\),
-because \(\partial_{q_u}f = 0\)
-and (\ref{eqn:regularity-matrix-acrobot}) becomes strictly positive for all
-values of \(q_a\).
+because then \(\partial_{q_u}f = 0\) and (\ref{eqn:regularity-matrix-acrobot})
+is strictly positive for all values of \(q_a\). 
+This will be useful later, as we will not need to check regularity if we design
+a function of the unactuated momentum.
 
@@
 In the rest of this chapter, our goal is to design the function \(f(q_u,p_u)\)
-based on the natural human motion of a gymnast, with the caveat that the
-resulting VNHC be regular and that we can prove the constrained dynamics will
-inject energy into the acrobot.
+based on the natural human motion of a gymnast, with one caveat: 
+we must be able to prove the constrained dynamics will inject energy into the
+acrobot.
 
@@
 % rigorously that we can stabilize any energy level set on the acrobot.
-There are a few existing approaches to generating giant motion for the acrobot
-which may be viable candidates on which to base a VNHC, akin to how we designed
-the VLP constraint in Chapter \ref{ch:vlp}.
+Let us examine some of the existing approaches to generating giant motion for
+the acrobot, since these may be viable candidates on which to base a VNHC.
 
-An initial approach to controlling the acrobot is to model it as a
+One initial approach to controlling the acrobot is to model it as a
 variable-length pendulum by collapsing the two rods and masses into one
@@
 that trajectory.
-
+Following in their footsteps, we might consider using the results
+from Chapter \ref{ch:vlp} to find the leg angles that allow the ECM to gain
+energy. 
+Then we could apply Theorem \ref{thm:vlp-energy-stabilization} to prove
+the acrobot is gaining energy too.
 \begin{figure}
@@
 \[
-    l_e(q_a) := l\sqrt{\frac{5}{4} + cos(q_a)}
+    l_e(q_a) := l\sqrt{\frac{5}{4} + c_a}
     ,
@@
 \[
-    q_e := \arctan_2(s_u + \frac{1}{2}s_{ua}, -c_u - \frac{1}{2}c_{ua})
+    q_e := \arctan_2\left(s_u + \frac{1}{2}s_{ua}, -c_u - \frac{1}{2}c_{ua}\right)
     .
 \]
-There are two important thing to note here. 
+There are two important notes to consider based on these equations. 
 First, Figure \ref{fig:acrobot-vlp-symmetry} shows that for each pose of the
 VLP representation, there are two configurations
-of the acrobot which give the same effective angle and length.
+of the acrobot which give the same effective length and angle.
 This means the acrobot and the VLP are not equivalent representations;
@@
     \centering
-    % TODO: Generate a figure with the ECM at one position, and two mirrored
-    % acrobots (dashed) around it to show how the VLP and acrobot are not
-    % equivalent.
-%    \includestandalone[width=0.5\textwidth]{}
+    \includestandalone[width=0.3\textwidth]{images/acrobot_vlp_symmetry}
     \caption{The equivalent center of mass of the acrobot generally has two configurations
@@
 
-Since the VLP representation of the acrobot requires redesigning a VLP
-VNHC, and the resulting ECM motion may not produce realistic leg motion,
-this model is clearly ineffective for our purposes.
+Since we cannot apply the results of Chapter \ref{ch:vlp} to simplify the proof
+of energy injection, and the resulting ECM motion may not even produce
+realistic leg motion, this model reduction is ineffective for our purposes.
 
-Let us turn to the thesis of \citet{xingbo_thesis}, who designs a VHC to enforce a
+Let us turn next to the thesis of \citet{xingbo_thesis}, who designs a VHC to enforce a
 so-called ``tap" motion with the purpose of injecting energy into the acrobot. 

### 2020-10-20 latex/vlp.tex
Updated VLP chapter label
Added more to the VLP proof about escaping compact sets. NEeds a lot of editing.

@@
 %/========== Variable Length Pendulum ==========/%
-\chapter{Application of VNHCS: The Variable Length Pendulum}\label{sec:vlp}
+\chapter{Application of VNHCS: The Variable Length Pendulum}\label{ch:vlp}
 \section{Motivation}
@@
 
+Before stating our theorem, we must rigorously what we mean by ``energy
+injection" for the VLP. 
+First, we want the phase of the VLP to move away from the origin; 
+second, we want the system to reach any desired momentum in finite time.
+These are captured mathematically by the following definitions.
+
+\begin{defn}\label{defn:energy-injection}
+   Let \(\dot{x} = f(x)\) be a planar ODE defined on \(D \subset \R^2\) and that
+   \(f(0) = (0,0)\).
+   The ODE experiences \textit{energy injection on \(D\)} if
+   \begin{enumerate}
+      \item The origin is a repeller 
+         (or equivalently, it is asymptotically stable in negative time).
+      \item For all compact sets \(K \subset D\), there exists \(T > 0\) where,
+         for all \(t > T\), \(x(t) \not \in K\).
+   \end{enumerate}
+\end{defn}
+
+\begin{defn}
+   We say that a regular VNHC \(h\) \textit{injects energy into the VLP} if the
+   constrained dynamics experience energy injection everywhere on the
+   constraint manifold, except for possibly a set of measure zero.
+\end{defn}
+
 Supposing the initial condition of the VLP is not an equilibrium, we will show that the
@@
    \(C^2\) function of \(\theta\).
-   Suppose the initial condition of the VLP's unactuated variables is
-   \((q_0,p_0) \not \in \left\{(0,0),(\pm\pi,0)\right\}\).
-   A VNHC of the form 
+   A regular VNHC of the form 
    \(h(\mathbf{q},\mathbf{p}) = l - L(\theta)\) injects energy into the VLP
@@
 \begin{proof}
-   Let \(\Gamma\) be the constraint manifold for \(h\).
+   Let \(\Gamma = \mathbb{S}^1 \times \R \) be the constraint manifold for \(h\).
    In negative-time, the dynamics on \(\Gamma\) are
@@
    \end{equation}
-   Let \(D = \Gamma \backslash \{(\pm \pi, 0)\}\)
-   be the domain of interest on which we study this negative-time system.
+   We will first study the upright equilibrium \((\pm \pi, 0)\). 
+   The linearization of (\ref{eqn:negative-time-vlp}) at this
+   equilibrium is given by
+   \[
+      df_{(\pm \pi,0)} = \begin{bmatrix}
+         0 & -\frac{1}{mL(\pi)^2} \\
+         -m g L(\pi) & 0 \\
+      \end{bmatrix}
+      ,
+   \]
+   which has eigenvalues at \(\lambda = \pm \frac{g}{L(\pi)}\).
+   Since this equilibrium has only one negative eigenvalue, the stable manifold
+   theorem tells us that the set of initial conditions on \(\Gamma\) which
+   converge to \((\pm \pi,0)\) is one-dimensional \textbf{SOURCE?}.
+   Label this set of initial conditions
+   \[
+      \Pi_0 := \left\{ (q_0,p_0) \mid \lim_{t \to \infty} (q(t),p(t)) = (\pm pi, 0) \right\}
+      .
+   \]
+   Since \(\Gamma\) is 2D, \(\Pi_0\) is a set of measure zero.
+   We therefore assume the system is not initialized on \(\Pi_0\), and will show
+   that the origin is almost everywherethat globally asymptotically stable.
+   % TODO: show (0,0) is a repeller and that all compact sets are left, forever,
+   % in finite time
+%   Let \(D = \Gamma \backslash \{(\pm \pi, 0)\}\)
+%   be the domain of interest on which we study this negative-time system.
    Choose, as a candidate Lyapunov-like function, the energy for the
@@
    This is positive definite at \((0,0)\) and has compact sublevel
-   sets on \(D\). 
+   sets on \(\Gamma\). 
    The derivative of \(E_\text{avg}\) along (\ref{eqn:negative-time-vlp}) is
@@
    The derivative of \(E_\text{avg}\) is nonnegative since
-   \begin{align*}
+   \begin{align}\label{eqn:vlp-proof-neg-inv}
     \sign{\dot{E}_\text{avg}} &= \sign{\sin(q)p \left(L(\theta)^3 - l_\text{avg}^3\right)} \\
@@
       .
-   \end{align*}
+   \end{align}
    Thus, \(E_\text{avg}\) is a Lyapunov function, which means the negative-time system is
    stable \cite{lyapunov}. 
-   The last step of the proof is to show that (\ref{eqn:negative-time-vlp})
-   asymptotically converges to \((0,0)\). 
-   For this, we will use a Krasovskii-LaSalle argument
-   \cite{krasovskii_lasalle}.
+   Define \(D := \Gamma \backslash \Pi_0\) to be the domain of viable initial
+   conditions.
+   Using a Krasovskii-LaSalle argument \cite{krasovskii_lasalle}, 
+   we will prove that (\ref{eqn:negative-time-vlp}) asymptotically converges to
+   \((0,0)\) if it is initialized on \(D\).
    First, we define the zero-set
@@
    Hence, the positive-time system is diverging away from the origin 
-   in the \((q,p)\) plane, which means the VLP is gaining energy on average.
+   in the \((q,p)\) plane, which means the origin is a repeller.
+   This proves the first property of Definition \ref{defn:energy-injection}
+   holds for the positive-time system; we need only show the second property
+   also holds to complete the proof.
+   We will prove the second property by contradiction.
+   Let \(x(t)\) be the orbit for the positive-time system and assume
+   \(x(0) \in K\backslash \Pi_0\).
+   Suppose there exists a compact set \(K \subset \Gamma\) so that, for all
+   \(T > 0\), there exists \(t > T\) where \(x(t) \in K\).
+   Let 
+   \[
+      k := \sup\left\{E_\text{avg}(q,p) \mid (q,p) \in K \right\}
+      ,
+   \] 
+   be the largest energy value of \(K\), and let
+   \[
+      E_k := \left\{ (q,p) \in \Gamma \mid E_\text{avg}(q,p) \leq k \right\}
+      ,
+   \]
+   be the largest energy sublevel set surrounding \(K\).
+   Note that \(E_k\) is compact and negatively invariant (it is positively
+   invariant for the negative-time system).
+   Because \(K \subset E_k\), it follows that \(x(t) \in E_k\). 
+   Then, since \(E_k\) is negatively invariant, the half orbit
+   \(x([0,t])\) is contained fully within \(E_k\).
+   In particular, \(x(T) \in E_k\); 
+   but \(T\) was chosen arbitrarily, which implies \(x(t) \in E_k\) for all
+   \(t \in [0,\infty[\).
+   Since \(E_k\) is compact, \(x(t)\) has a positive limit set in
+   \(E_k\) which (by Poincar\'{e}-Bendixson) is either an equilibrium or a
+   closed orbit \textbf{SOURCE?}.
+   The system cannot have closed orbits, since this would violate asymptotic
+   stability of \((0,0)\) in negative time. 
+   Hence, the positive limit set must be one of the equilibria. 
+   The origin is a repeller, which means \(x(t)\) must be converging to 
+   \((\pm \pi,0)\).
+   This is a contradiction, since we assumed \(x(0) \not \in \Pi_0\).
+   \[
+      \Rightarrow \Leftarrow
+   \]
+   We have shown that orbits of the constrained dynamics will leave compact sets in
+   finite time, including sublevel sets of \(E_\text{avg}\).
+   Thus, the VLP must be gaining energy because its momentum is increasing.
    Flipping the inequality of (\ref{eqn:vlp-energy-gain-condition}) 

### 2020-10-22 latex/vlp.tex
First edit of the new proof of VLP energy injection

--- 2020-10-08 latex/vlp.tex
@@
 tells us they are described entirely by \((\dot{q},\dot{p})\) with \(l\)
-replaced by \(L(q,p)\),
+replaced by \(L(q,p)\):
 \begin{align}\label{eqn:vlp-hamiltonian}
@@
 \]
-What remains is to prove this injects energy into the VLP. 
+
+Before continuing, we must rigorously define what it means for a VNHC to inject
+energy into the VLP.
+We want the the VLP to behave in
+such a way that momentum increases over time;
+this intuition is captured mathematically by the following definitions.
+
+\begin{defn}\label{defn:energy-injection}
+   Let \(\dot{x} = f(x)\) be a planar ODE defined on \(D \subset \R^2\), with
+   \(f(0) = (0,0)\).
+   This system \textit{gains energy on \(D\)} if 
+   \begin{enumerate}
+      \item The origin is a repeller 
+         (or equivalently, the origin is asymptotically stable in negative time).
+      \item For all compact sets \(K \subset D\) and for almost every initial
+         condition \(x(0) \in K\), there exists \(T > 0\) such
+         that \(x(t) \not \in K \, (\forall t > T)\).
+   \end{enumerate}
+\end{defn}
+
+\begin{defn}
+   A regular VNHC \(h(\mathbf{q},\mathbf{p}) = 0\) 
+   \textit{injects energy into the VLP} if the constrained dynamics gain energy
+   everywhere on the constraint manifold, except possibly on a set of measure
+   zero.
+\end{defn}
+
+We can now prove our VNHC injects energy into the VLP. 
 As part of the proof, we will require the following lemma.
@@
    \end{equation}
-   Let \(D = \Gamma \backslash \{(\pm \pi, 0)\}\)
-   be the domain of interest on which we study this negative-time system.
-   Choose, as a candidate Lyapunov-like function, the energy for the
+   We begin by analyzing the upright equilibria \((\pm \pi, 0)\). 
+   The linearization of (\ref{eqn:negative-time-vlp}) at these
+   equilibria is given by
+   \[
+      df_{(\pm \pi,0)} = \begin{bmatrix}
+         0 & -\frac{1}{mL(\pi)^2} \\
+         -m g L(\pi) & 0 \\
+      \end{bmatrix}
+      ,
+   \]
+   which has eigenvalues at \(\lambda = \pm \sqrt{\frac{g}{L(\pi)}}\).
+   The positive-time system has the same eigenvalues at these upright
+   equilibria.
+   We label the set of initial conditions converging to \((\pm \pi, 0)\) in
+   positive time by
+   \[
+      \Pi_0^+ := \left\{ (q_0,p_0) \in \Gamma \mid \lim_{t \to \infty} (q(t),p(t)) = (\pm \pi, 0) \right\}
+      ,
+   \]
+   and the set of initial conditions converging to \((\pm \pi,0)\)
+   in negative time by \(\Pi_0^-\).
+   Since only one eigenvalue is negative, the stable manifold theorem tells us
+   that both \(\Pi_0^+\) and \(\Pi_0^-\) are one-dimensional \textbf{SOURCE?},
+   which means they have measure zero within \(\Gamma\).
+
+   To prove the origin of the positive-time system is a repeller, we will show
+   the origin of the negative-time system is almost globally
+   asymptotically stable.
+
+   Choose, as a candidate Lyapunov function, the energy for the
    average-length pendulum with zero-potential at the bottom of the swing:
@@
            + m g l_\text{avg} (1-\cos(q))
+      .
    \]
    This is positive definite at \((0,0)\) and has compact sublevel
-   sets on \(D\). 
+   sets on \(\Gamma\). 
    The derivative of \(E_\text{avg}\) along (\ref{eqn:negative-time-vlp}) is
@@
    \]
-   The sign of \(\sin(q)p\) is dependent only on the quadrant in the
-   \((q,p)\)-plane, so one can show it is equal to the sign of
-   \(\sin(2\theta)\). 
+   The sign of \(\sin(q)p\) depends only on the quadrant in the
+   \((q,p)\)-plane. In fact, it is easy to show that
+   \[
+      \sign{\sin(q)p} = \sign{\sin(2\theta)}
+      .
+   \]
    Furthermore, by Lemma \ref{lemma:sign-of-cube} we have
    \[ 
-       \sign{l_\text{avg}^3 - L(\theta)^3} = \sign{l_\text{avg} - L(\theta)}
+       \sign{L(\theta)^3 - l_\text{avg}^3} = \sign{L(\theta) - l_\text{avg}}
        .
    \]
-   The derivative of \(E_\text{avg}\) is nonnegative since
-   \begin{align*}
+   The derivative of \(E_\text{avg}\) is never positive since
+   \begin{align}\label{eqn:vlp-proof-neg-inv}
     \sign{\dot{E}_\text{avg}} &= \sign{\sin(q)p \left(L(\theta)^3 - l_\text{avg}^3\right)} \\
@@
       .
-   \end{align*}
-   Thus, \(E_\text{avg}\) is a Lyapunov function and the negative-time system is
+   \end{align}
+   Thus, \(E_\text{avg}\) is a Lyapunov function, which means the negative-time system is
    stable \cite{lyapunov}. 
-   The last step of the proof is to show that (\ref{eqn:negative-time-vlp})
-   asymptotically converges to \((0,0)\). 
-   For this, we will use a Krasovskii-LaSalle argument
-   \cite{krasovskii_lasalle}.
+   Define \(D := \Gamma \backslash \Pi_0^-\) to be the domain of interest.
+   Using a Krasovskii-LaSalle argument \cite{krasovskii_lasalle}, 
+   we will prove that (\ref{eqn:negative-time-vlp}) asymptotically converges to
+   \((0,0)\) if it is initialized on \(D\).
    First, we define the zero-set
@@
    Hence, the positive-time system is diverging away from the origin 
-   in the \((q,p)\) plane, which means the VLP is gaining energy on average.
+   in the \((q,p)\) plane, which means the origin is a repeller.
+
+   We now prove the system escapes compact sets in finite time.
+   Let \(x(t) = (q(t),p(t))\) be the orbit for the positive-time system.
+   Let \(K \subset \Gamma\) be compact and assume \(x(0) \in K\backslash \Pi_0^+\).
+   For a proof by contradiction, suppose that for all \(T > 0\) there exists \(t > T\)
+   where \(x(t) \in K\).
+   Let 
+   \[
+      k := \max\limits_{(q,p) \in K} E_\text{avg}(q,p)
+      ,
+   \] 
+   be the largest energy value attained in \(K\), and let
+   \[
+      E_k := \left\{ (q,p) \in \Gamma \mid E_\text{avg}(q,p) \leq k \right\}
+      ,
+   \]
+   be the energy sublevel set with that largest energy. 
+   Note that \(E_k\) is compact, contains \(K\), and is negatively invariant (it
+   is positively invariant for the negative-time system).
+   Because \(K \subset E_k\), it follows that \(x(t) \in E_k\). 
+   Then, since \(E_k\) is negatively invariant, the half orbit
+   \(x([0,t])\) is contained fully within \(E_k\).
+   In particular, \(x(T) \in E_k\); 
+   but \(T\) was chosen arbitrarily, which implies \(x(t) \in E_k\) for all
+   \(t \in [0,\infty[\).
+   Since \(E_k\) is compact, \(x(t)\) has a positive limit set in
+   \(E_k\) which (by Poincar\'{e}-Bendixson) is either an equilibrium or a
+   closed orbit \textbf{SOURCE?}.
+   The system cannot have closed orbits, since this would violate asymptotic
+   stability of \((0,0)\) in negative time. 
+   Hence, the positive limit set must be one of the equilibria. 
+   The origin is a repeller, which means \(x(t)\) must be converging to 
+   \((\pm \pi,0)\).
+   This is a contradiction, since we assumed \(x(0) \not \in \Pi_0^+\).
+
+   We have shown that orbits of the constrained dynamics will leave compact sets in
+   finite time, including all sublevel sets of \(E_\text{avg}\). 
+   The orbit \((q(t),p(t))\) is spiraling away from the origin and (for almost
+   all initial conditions) will not hit the upper equilibrium.
+   Thus, the VLP is gaining energy on average because its momentum is increasing.
    Flipping the inequality of (\ref{eqn:vlp-energy-gain-condition}) 
    means \(E_\text{avg}\) is a Lyapunov function for the positive time system;
-   by the same arguments presented here, the VLP loses energy while converging
-   asymptotically to \((0,0)\).
+   by the Krasovskii-LaSalle arguments presented earlier, the VLP loses energy
+   while converging asymptotically to \((0,0)\).
  \end{proof}

### 2020-10-23 latex/vlp.tex
Second edit of the VLP proof, much better now.
Time to move on to editing what Manfredi sent me on Oct 20

@@
 
-Before continuing, we must rigorously define what it means for a VNHC to inject
-energy into the VLP.
-We want the the VLP to behave in
-such a way that momentum increases over time;
-this intuition is captured mathematically by the following definitions.
+Before proving the VLP will gain energy under this VNHC, 
+we must rigorously define what that means.
+If any initial condition of the constrained dynamics converges to the point
+\((q,p) = (0,0)\), the momentum must be decreasing over time, so the VLP will
+clearly be losing energy.
+A notion of energy injection must therefore require that the origin repels all
+solutions.
+We need to avoid closed orbits or limit cycles in the \((q,p)\)-plane,
+so that the system can always reach higher speeds. 
+Finally, we want the VLP's momentum to increase on average, regardless of
+its initial value.
+This intuition is captured mathematically by the following definition.
 
 \begin{defn}\label{defn:energy-injection}
-   Let \(\dot{x} = f(x)\) be a planar ODE defined on \(D \subset \R^2\), with
-   \(f(0) = (0,0)\).
-   This system \textit{gains energy on \(D\)} if 
+   Let \(D \subset \R^2\) with \(0 \in D\).
+   Let \(f : D \rightarrow \R^2\) be locally Lipschitz with \(f(0) = 0\).
+   The system described by \(\dot{x} = f(x)\) 
+   \textit{gains energy on \(D\)} if 
    \begin{enumerate}
@@
    \end{enumerate}
+   The system \textit{loses energy on \(D\)} if it gains energy in
+   negative-time.
 \end{defn}
+
+The second property of Definition \ref{defn:energy-injection} allows the system
+to have unstable equilibria, but not limit cycles nor closed orbits.
+The next definition ties this notion of energy gain to VNHCs.
 
@@
    A regular VNHC \(h(\mathbf{q},\mathbf{p}) = 0\) 
-   \textit{injects energy into the VLP} if the constrained dynamics gain energy
-   everywhere on the constraint manifold, except possibly on a set of measure
-   zero.
+   \textit{injects (dissipates) energy} if the constrained dynamics gain (lose)
+   energy everywhere on the constraint manifold, except possibly on a set of
+   measure zero.
 \end{defn}
@@
 
-Supposing the initial condition of the VLP is not an equilibrium, we will show that the
-constrained dynamics trace out a curve on the \((q,p)\)-plane which is diverging
-from the origin.  
+We will show that the constrained dynamics trace out a curve on the
+\((q,p)\)-plane which is diverging from the origin.  
 This implies that the momentum \(p\) is increasing in magnitude whenever the
@@
 \begin{proof}
-   Let \(\Gamma = \mathbb{S}^1 \times \R \) be the constraint manifold for \(h\).
+   The constraint manifold for \(h\) is the cylinder \(\Gamma = \mathbb{S}^1 \times \R \). 
    In negative-time, the dynamics on \(\Gamma\) are
@@
    \end{equation}
-   We begin by analyzing the upright equilibria \((\pm \pi, 0)\). 
-   The linearization of (\ref{eqn:negative-time-vlp}) at these
-   equilibria is given by
-   \[
-      df_{(\pm \pi,0)} = \begin{bmatrix}
-         0 & -\frac{1}{mL(\pi)^2} \\
-         -m g L(\pi) & 0 \\
-      \end{bmatrix}
-      ,
-   \]
-   which has eigenvalues at \(\lambda = \pm \sqrt{\frac{g}{L(\pi)}}\).
-   The positive-time system has the same eigenvalues at these upright
-   equilibria.
-   We label the set of initial conditions converging to \((\pm \pi, 0)\) in
-   positive time by
-   \[
-      \Pi_0^+ := \left\{ (q_0,p_0) \in \Gamma \mid \lim_{t \to \infty} (q(t),p(t)) = (\pm \pi, 0) \right\}
-      ,
-   \]
-   and the set of initial conditions converging to \((\pm \pi,0)\)
-   in negative time by \(\Pi_0^-\).
-   Since only one eigenvalue is negative, the stable manifold theorem tells us
-   that both \(\Pi_0^+\) and \(\Pi_0^-\) are one-dimensional \textbf{SOURCE?},
-   which means they have measure zero within \(\Gamma\).
-
-   To prove the origin of the positive-time system is a repeller, we will show
-   the origin of the negative-time system is almost globally
-   asymptotically stable.
+   This has equilibria at \((q,p) = (0,0)\) and \((q,p) = (\pi,0)\) (the
+   third ``equilibrium" \((-\pi,0)\) is the same as \((\pi,0)\) because 
+   \(q \in \mathbb{S}^1\)).
+   To prove the origin of the positive-time system (\ref{eqn:vlp-hamiltonian})
+   is a repeller, we will show the origin of the negative-time system 
+   (\ref{eqn:negative-time-vlp}) is almost everywhere globally asymptotically
+   stable.
 
@@
    \begin{align}\label{eqn:vlp-proof-neg-inv}
-    \sign{\dot{E}_\text{avg}} &= \sign{\sin(q)p \left(L(\theta)^3 - l_\text{avg}^3\right)} \\
-                &= \sign{\sin(2\theta) \left(L(\theta) - l_\text{avg}\right)} \\
-                &= -\sign{(l_\text{avg} - L(\theta))\sin(2\theta)} \\
+    \sign{\dot{E}_\text{avg}} &= \sign{\sin(q)p \left(L(\theta)^3 - l_\text{avg}^3\right)}
+          \nonumber \\
+          &= \sign{\sin(2\theta) \left(L(\theta) - l_\text{avg}\right)}
+          \nonumber \\
+          &= -\sign{(l_\text{avg} - L(\theta))\sin(2\theta)} 
+          \nonumber \\
           &\leq 0 \text{ (by assumption)}
@@
    stable \cite{lyapunov}.
+
+   The linearization of (\ref{eqn:negative-time-vlp}) at the upright equilibrium
+   \((\pi, 0)\) is given by
+   \[
+      df_{(\pi,0)} = \begin{bmatrix}
+         0 & -\frac{1}{mL(0)^2} \\
+         -m g L(0) & 0 \\
+      \end{bmatrix}
+      ,
+   \]
+   which has eigenvalues at \(\lambda = \pm \sqrt{\frac{g}{L(0)}}\).
+   The positive-time system (\ref{eqn:vlp-hamiltonian}) has the same eigenvalues
+   at this equilibrium.
+   Label the set of initial conditions converging to \((\pi, 0)\) in
+   positive time by
+   \[
+      \Pi_0^+ := \left\{ (q_0,p_0) \in \Gamma \mid \lim_{t \to \infty} (q(t),p(t)) = (\pi, 0) \right\}
+      ,
+   \]
+   and the set of initial conditions converging to \((\pi,0)\)
+   in negative time by \(\Pi_0^-\).
+   Since only one eigenvalue is negative, the stable manifold theorem tells us
+   that both \(\Pi_0^+\) and \(\Pi_0^-\) are one-dimensional.
+   While this implies the origin is not globally asymptotically stable for
+   (\ref{eqn:negative-time-vlp}), we can still show it is asymptotically stable
+   for almost every initial condition because \(\Pi_0^+\) and \(\Pi_0^-\) are of
+   measure zero within \(\Gamma\).
+
    Define \(D := \Gamma \backslash \Pi_0^-\) to be the domain of interest.
@@
    converge to the largest invariant set in \(Z \cap D\).
-   Since \(f(q,p) = (0,0)\) only when \((q,p) = (0,0)\),
+   Since \(f(q,p) = (0,0)\) on \(D\) only when \((q,p) = (0,0)\),
    the origin is asymptotically stable on \(D\) for the negative-time system.
@@
    Let \(K \subset \Gamma\) be compact and assume \(x(0) \in K\backslash \Pi_0^+\).
-   For a proof by contradiction, suppose that for all \(T > 0\) there exists \(t > T\)
-   where \(x(t) \in K\).
+   For a proof by contradiction, suppose that for all \(T > 0\) there exists
+   \(t' > T\) where \(x(t') \in K\).
    Let 
@@
    \] 
-   be the largest energy value attained in \(K\), and let
+   be the largest energy value attained in \(K\), and let its sublevel set be
    \[
       E_k := \left\{ (q,p) \in \Gamma \mid E_\text{avg}(q,p) \leq k \right\}
-      ,
+      .
    \]
-   be the energy sublevel set with that largest energy. 
-   Note that \(E_k\) is compact, contains \(K\), and is negatively invariant (it
-   is positively invariant for the negative-time system).
-   Because \(K \subset E_k\), it follows that \(x(t) \in E_k\). 
+   All sublevel sets of \(E_\text{avg}\) are compact, negatively
+   invariant (by (\ref{eqn:vlp-proof-neg-inv})), and \(E_k\) contains \(K\).
+   We assumed \(x(t') \in K\), so it follows that \(x(t') \in E_k\). 
    Then, since \(E_k\) is negatively invariant, the half orbit
-   \(x([0,t])\) is contained fully within \(E_k\).
+   \(x([0,t'])\) is contained fully within \(E_k\).
    In particular, \(x(T) \in E_k\);
@@
    The origin is a repeller, which means \(x(t)\) must be converging to 
-   \((\pm \pi,0)\).
+   \((\pi,0)\).
    This is a contradiction, since we assumed \(x(0) \not \in \Pi_0^+\).
@@
    finite time, including all sublevel sets of \(E_\text{avg}\). 
-   The orbit \((q(t),p(t))\) is spiraling away from the origin and (for almost
-   all initial conditions) will not hit the upper equilibrium.
+   The orbit \((q(t),p(t))\) is spiraling away from the origin and, for almost
+   all initial conditions, will not converge to the upper equilibrium.
    Thus, the VLP is gaining energy on average because its momentum is increasing.
@@
    means \(E_\text{avg}\) is a Lyapunov function for the positive time system;
-   by the Krasovskii-LaSalle arguments presented earlier, the VLP loses energy
-   while converging asymptotically to \((0,0)\).
+   by the same arguments presented here, the VLP loses energy while
+   converging asymptotically to \((0,0)\).
  \end{proof}

### 2020-10-25 latex/symbols.tex
Better motivaated why VNHCss are important and why we study the VLP
Made Theorem numbers as chapter.num
Fully completed / edited VNHCs, we are now at v1.
Made all edits suggested by Manfredi on Oct 19

@@
    \(\mathcal{Q}\) & The configuration manifold of a system. \\
+   \(C^r(X;Y)\) & The space of \(r\)-times continuously differentiable functions from \(X\) to \(Y\). \\
+         & If \(r = \infty\), the space of smooth functions from \(X\) to \(Y\). \\
    \(\R^{n \times m}\) & The space of real-valued matrices with \(n\) rows and \(m\) columns. \\
@@
    \(\nabla_v F\) & If \(F\) is \(\R\)-valued, the gradient of \(F\) with respect to \(v\). \\
-            & If \(F : \R^m \rightarrow \R^{n\times n}\), the block matrix gradient \((\pdiff{F}{v_1},\ldots,\pdiff{F}{v_m}) \in \R^{nm \times n}\). \\
+            & If \(F\) is \(\R^{n\times n}\)-valued and \(v \in \R^m\), the block gradient \(\left[\nabla_{v_1}F\tpose \ldots \nabla_{v_m}F\tpose\right]\tpose \in \R^{nm \times n}\). \\
    \(dF_v\) & Total differential (Jacobian) of \(F\), equivalent to \((\nabla_v F)\tpose\). \\
@@
    \(\delta_{i,j}\) & The Kronecker delta: 1 if \(i = j\) and 0 otherwise. \\
-   \(\otimes\) & The matrix kronecker product (see Appendix \ref{appendix:matrix-kron-prod}). \\
+   \(\otimes\) & The matrix kronecker product. \\ 
+   \(\norm{\cdot}\) & The euclidean norm on \(\R^n\). \\
    \bottomrule

### 2020-10-25 latex/vnhcs.tex
Made all edits suggested by Manfredi on Oct 19

@@
         - \pdiff{P}{q_i}(q) + B_i\tpose(q) \tau
-    ,
+    .
 \]
-which can be collected into vector form
-by using the matrix Kronecker product (see Appendix \ref{apx:kron-prod}).
+Recall the Kronecker product of matrices \cite{kronprod}, defined as follows.
+\begin{defn}
+    Let \(A \in \R^{n \times m}\) and \(B \in \R^{r \times s}\).
+    The Kronecker product \(A \otimes B \in \R^{nr \times ms}\) is the  matrix
+    \[
+        A \otimes B = \begin{bmatrix}
+            a_{1,1}B & \cdots & a_{1,m} B \\
+            \vdots & \ddots & \vdots \\
+            a_{n,1} B & \cdots & a_{n,m} B
+        \end{bmatrix} 
+        .
+    \]
+\end{defn}
+Using the kronecker product, one can collect the derivative of momentum into
+the vector form
+\[
+    \dot{p} = -\frac{1}{2} (\Id{n} \otimes p\tpose) \nabla_q D\inv(q) p
+    - \nabla_q P(q) + B(q) \tau
+    .
+\]
 In sum, when the kinetic energy is quadratic the Hamiltonian system reduces to
@@
     \((q,p)\).
-    Let \(A \in \R^{n\times n}\) be an invertible matrix such that
-    \(A\inv = A\tpose\).
+    Let \(A \in \R^{n\times n}\) be an orthogonal matrix.
     The change of coordinates \((q,p) \to (Q = Aq, P = Ap)\) is a canonical
@@
     Suppose Assumption \ref{assm:B-const} holds. Then
-    there exist nonsingular matrices \(V,T \in \R^{k \times k}\) 
+    there exists a nonsingular matrix \(\hat{T} \in \R^{k \times k}\) 
     so that the regular feedback transformation 
     \[
-        \tau = V T \hat{\tau}
+        \tau = \hat{T} \hat{\tau}
     \] 
@@
 where \(\sigma_i \neq 0\) because \(B\) is full-rank \cite{calculating_svd}.
-Defining \(T \in \R^{k \times k}\) by
+Define \(T \in \R^{k \times k}\) by
 \[
@@
     \end{bmatrix}
-    ,
+    .
 \]
-and assigning the regular feedback transformation\(\tau = V T \hat{\tau}\), 
+Letting \(\hat{T} = V T\) and assigning the regular feedback
+transformation \(\tau = \hat{T} \hat{\tau}\), 
 we get a new input matrix for \(\hat{\tau} \in \R^k\) given by 
 \[
-    \hat{B} = B V T = 
+    \hat{B} = B \hat{T} = B V T = 
     U \begin{bmatrix}
@@
     Assume that the input matrix \(B\) is
-    left semi-orthogonal. 
+    left semi-orthogonal, \ie, \(B\tpose B = I_k\). 
 \end{assm}
@@
     has simply actuated canonical coordinates 
-    \(\{Q = \mathbf{B}q, P = \mathbf{B}p\}\). The resulting dynamics are 
+    \(\left(Q = \mathbf{B}q, P = \mathbf{B}p\right)\). The resulting dynamics are 
     given by (\ref{eqn:simple-hamiltonian}),
@@
     \((q,p) \in h\inv(0)\), and there exists a feedback controller \(\tau(q,p)\)
-    stabilizing the \textit{constraint manifold}
+    rendering the \textit{constraint manifold} \(\Gamma\) invariant
+    where
     \[
@@
 From the definition of VNHCs, one finds that the constraint manifold \(\Gamma\)
-is a \(2(n-k)\)-dimensional surface inside \(\mathcal{Q} \times \R^n\). 
+is a \(2(n-k)\)-dimensional closed embedded submanifold of \(\mathcal{Q} \times \R^n\). 
 The next obvious question is the following: when does a feedback controller
@@
 If there exists some controller \(\tau(q,p)\) driving \(e \to 0\) and 
-\(\dot{e} \to 0\), then the same \(\tau\) will necessarily stabilize \(\Gamma\).
+\(\dot{e} \to 0\), then the same \(\tau\) will necessarily stabilize \(\Gamma\)
+(under additional mild conditions, see \cite{vhcs_for_el_systems})
+\footnote{The type of constraints we consider in this thesis are graphs of
+functions, and so meet these mild assumptions.}.
 
-If \(\tau\) first appears after \(r_i > 0\) derivatives of \(e_i\), 
-then \(e_i\) is said to have \textit{relative degree} \(r_i\).
-Collecting the \(e_i\) together, we say that \(e\) is of relative degree
-\(\{r_1,\ldots,r_k\}\).
+We say that \(e\) is of \textit{relative degree} \(\{r_1,\ldots,r_k\}\) if
+\begin{enumerate}
+    \item Some component of \(\tau\) appears at the \(r_i^\text{th}\) derivative
+        of \(e_i\) for \(i \in \{1,\ldots,k\}\).
+    \item In the vector of time-derivatives
+        \((e_1^{(r_1)}, \ldots, e_k^{(r_k)})\), the vector \(\tau\) is
+        premultiplied by a nonsingular matrix.
+\end{enumerate}
 With no further structure on \(h(q,p)\), the control input \(\tau\) usually
@@
 be able to stabilize \(\Gamma\). 
-We could guarantee \(h(q,p) \to 0\), but we could not in general guarantee 
-\(\dot{h}(q,p) \to 0\).
+We could guarantee \(e_i(t) \to 0\), but we could not in general guarantee 
+\(\dot{e}_i(t) \to 0\).
 
@@
 \begin{thm}\label{thm:vnhc-regularity}
-    A VNHC \(h(q,p) = 0\) of order \(k\) is regular if and only if \(dh_{p_a} = 0\) 
+    A relation \(h(q,p) = 0\) for system (\ref{eqn:simply-actuated-hamiltonian})
+    is a regular VNHC of order \(k\) if and only if \(dh_{p_a} = 0\) 
     and
@@
 Theorem \ref{thm:vnhc-regularity}, a regular VNHC of order \(k\) can be
-stabilized by the output-linearizing phase-feedback controller
+stabilized by\footnote{Under additional mild conditions
+\cite{vhcs_for_el_systems}} the output-linearizing phase-feedback controller
 \begin{equation}\label{eqn:vnhc-torque-controller}
@@
     manifold \(\Gamma\). Suppose that on \(\Gamma\) one can solve for
-    \(q_a\) as a function of \((q_u,p_u)\).
+    \(q_a\) as a function \(q_a = f(q_u,p_u)\).
     Then the constrained dynamics are given by
@@
             \end{aligned}{}\right|_{\begin{array}{c}
-                h(q,p_u) = 0 \\ 
+                q_a = f(q_u,p_u) \\ 
                 p_a = g(q_u,p_u) \\
@@
             \Zmat{k \times (n-k)} \\
-        \end{bmatrix} p_u\right)\right|_{h(q,p_u) = 0}
+        \end{bmatrix} p_u\right)\right|_{q_a = f(q_u,p_u)}
         .
@@
     This yields a function \(p_a(q,p_u)\). However, on \(\Gamma\) we have 
-    \(h(q,p_u) = 0\) and can solve for \(q_a\) in terms of \((q_u,p_u)\);
-    hence, we can solve for \(p_a = g(q_u,p_u)\). Since \(q_a\) and
-    \(p_a\) can be computed directly from \((q_u,p_u)\), the dynamics on
-    \(\Gamma\) are parameterized only by \((\dot{q}_u,\dot{p}_u)\).
+    \(q_a = f(q_u,p_u)\), which we use to solve for \(p_a = g(q_u,p_u)\). 
+    Since \(q_a\) and \(p_a\) can be computed directly from \((q_u,p_u)\), the
+    dynamics on \(\Gamma\) are parameterized only by \((\dot{q}_u,\dot{p}_u)\).
 \end{proof}
@@
 \begin{cor}\label{cor:2d-zero-dynamics}
-    Suppose \(\mathcal{H}\) is a mechanical system in simply actuated
-    coordinates satisfying Assumption \ref{assm:inertially-actuated} 
-    which has degree of underactuation one.
+    Let \(\mathcal{H}\) be the system (\ref{eqn:simple-hamiltonian}) with degree
+    of underactuation one.
+    Suppose \(\mathcal{H}\) satisfies Assumption \ref{assm:inertially-actuated}.
     Let \(h(q,p_u) = 0\) be a regular VNHC of order \((n-1)\) of the form

### 2020-10-26 latex/vlp.tex
Made all edits suggested by Manfredi on Oct 19
Added references to P-B and stable manifold
Also edited more of the VLP proof, it's not quite done but it's coming
along nicely.

@@
    E(q,p) = \frac{1}{2} \frac{p^2}{mL^2} + \frac{1}{2}\dot{L}^2 - mgL\cos(q)
+   .
 \end{equation}
@@
    \end{equation}
-   This has equilibria at \((q,p) = (0,0)\) and \((q,p) = (\pi,0)\) (the
-   third ``equilibrium" \((-\pi,0)\) is the same as \((\pi,0)\) because 
-   \(q \in \mathbb{S}^1\)).
+   Because \(q \in \mathbb{S}^1\), the only equilibria of these dynamics are 
+   at \((q,p) = (0,0)\) and \((q,p) = (\pi,0)\).
    To prove the origin of the positive-time system (\ref{eqn:vlp-hamiltonian})
@@
    \((\pi, 0)\) is given by
-   \[
+   \begin{equation}\label{eqn:linearized-vlp-neg-time}
       df_{(\pi,0)} = \begin{bmatrix}
@@
       ,
-   \]
+   \end{equation}
    which has eigenvalues at \(\lambda = \pm \sqrt{\frac{g}{L(0)}}\).
@@
    in negative time by \(\Pi_0^-\).
-   Since only one eigenvalue is negative, the stable manifold theorem tells us
-   that both \(\Pi_0^+\) and \(\Pi_0^-\) are one-dimensional.
-   While this implies the origin is not globally asymptotically stable for
-   (\ref{eqn:negative-time-vlp}), we can still show it is asymptotically stable
+   Taking the sublevel set
+   \[
+      E_{(\pi,0)} := \left\{ (q,p) \in \Gamma \mid E_\text{avg}(q,p) <
+         E_\text{avg}(\pi,0) \right\}
+      ,
+   \] 
+   it is clear form (\ref{eqn:vlp-proof-neg-inv}) that 
+   \(\Pi_0^+ \subset E_{(\pi,0)}\)
+   and 
+   \(\Pi_0^- \subset \left(\Gamma \backslash E_{(\pi,0)} \right) \cup (\pi,0)\).
+   Since only one eigenvalue of (\ref{eqn:linearized-vlp-neg-time})
+   is negative, the stable manifold theorem tells us that both \(\Pi_0^+\) and
+   \(\Pi_0^-\) are one-dimensional \cite{stable_manifold}.
+   This implies the origin cannot be globally asymptotically stable for
+   (\ref{eqn:negative-time-vlp}); however, we can still show it is asymptotically stable
    for almost every initial condition because \(\Pi_0^+\) and \(\Pi_0^-\) are of
@@
    the origin is asymptotically stable on \(D\) for the negative-time system.
+   In particular, the origin is asymptotically stable for every initial
+   condition in the interior of \(E_{(\pi,0)}\).
    Hence, the positive-time system is diverging away from the origin 
@@
    \] 
-   be the largest energy value attained in \(K\), and let its sublevel set be
-   \[
-      E_k := \left\{ (q,p) \in \Gamma \mid E_\text{avg}(q,p) \leq k \right\}
-      .
-   \]
+   be the largest energy value attained in \(K\), and let \(E_k\) be its
+   sublevel set.
    All sublevel sets of \(E_\text{avg}\) are compact, negatively
@@
    \(E_k\) which (by Poincar\'{e}-Bendixson) is either an equilibrium or a
-   closed orbit \textbf{SOURCE?}.
+   closed orbit \cite{poincare_bendixson}.
    The system cannot have closed orbits, since this would violate asymptotic

### 2020-10-27 latex/acrobot.tex
Started writing the acrobot constraint section
Need to describe the constrained dynamics in (qu,pu) and explain why we
can't use any form of Eavg before moving onto the proof.

@@
 full circles; 
-for this reason, we assume that \(q_a \in [-\bar{q}_a, \bar{q}_a]\) where 
-\(\bar{q}_a \in ]-\pi, \pi[\).
+for this reason, we assume that \(q_a \in [-Q_a, Q_a]\) where 
+\(Q_a \in ]-\pi, \pi[\).
 
@@
 with the control parameters \(\bar{q}_u\) and \(\rho_0\), where
-\begin{align*}
+\begin{align}
+    \label{eqn:xingbo-frad}
     f_\text{rad}(\rho) &:= \tanh^2(\rho/\rho_0)
     , \\
+    \label{eqn:xingbo-fang}
     f_\text{ang}(\xi) &:= 
@@
     \end{cases}
-\end{align*}
+\end{align}
 
@@
 % proof (both as a sketch and in a figure)
+One may be tempted to design a constraint \(q_a = f(\theta)\) 
+with \(\theta := \arctan_2(p_u,q_u)\), since it was so effective for the VLP in
+Chapter \ref{ch:vlp}.
+It can be challenging to show this type of constraint is regular.
+Recall that all constraints of the form \(h(q,p) = q_a - f(p_u)\)
+are regular, since the regularity equation (\ref{eqn:regularity-matrix-acrobot})
+becomes strictly positive.
+A momentum-based constraint approximates human motion may still be likely to
+inject energy into the acrobot.
+
+To design the acrobot constraint, let us imagine what a human does on
+a seated swing.
+When they are moving forward they extend their legs, and when they are moving
+backward they retract them. 
+In reality, the motion is slightly more complicated than that.:
+as the swing gains speed, the person leans their body back while
+extending their legs.
+This allows them to bring their legs higher, shortening the distance
+from their center of mass to the pivot and adding more energy to the swing.
+When the swing moves backward, they sit up and fully retract their legs
+underneath them.
+
+Now imagine we have affixed the torso of the acrobot to the swing's rope, so 
+that the legs can pivot both forwards and backwards.  
+In this case, the legs would pivot in the direction of motion
+proportionally to the swing's speed.
+One VNHC which emulates this process is \(q_a = \bar{q}_a\arctan_2( I p_u)\), as
+in Figure \ref{fig:qa-arctan}.
+Here, \(\bar{q}_a \in ]0,\frac{2 Q_a}{\pi}]\) so that 
+\(q_a \in [-Q_a, Q_a]\) and \(I > 0 \in \R\) is a fixed control parameter.
+
+This constraint is not a perfect emulation of a gymnast's movement -- during
+a giant the gymnast's legs are almost completely extended, while the constraint
+retracts the legs partially during rotations.
+However, the motion looks similar enough that this constraint should provide a
+decent foundation for injecting energy into the acrobot.
+It is for this reason that we choose our acrobot's constraint to be
+\begin{equation}\label{eqn:acrobot-constraint}
+    h(q,p) = q_a - \bar{q}_a \arctan_2(I p_u)
+    .
+\end{equation}
+
+\begin{figure}
+    \centering
+    \includestandalone[width=0.5\textwidth]{images/qa_arctan}
+    \caption{The acrobot constraint \(q_a = \bar{q}_a \arctan_2(I p_u)\).}
+    \label{fig:qa-arctan}
+\end{figure}
+
+Let us now compute the constrained dynamics for this
+(\ref{eqn:acrobot-constraint}).
+% TODO: should I do this for a generic h(q_a,p_u)? That's how I did it in my
+% notes...
+Note that \(dh_q = \begin{bmatrix}0 & 1\end{bmatrix}\tpose\), while
+\[
+    dh_{p_u} = \frac{-\bar{q}_a I}{1 + I^2 p_u^2}
+    .
+\]
+Applying these to (\ref{eqn:g-qupu}), we get the solution for \(p_a\) on the
+constraint manifold as
+\[
+    p_a(q_u,p_u) = \frac{
+        (1+c_a)(1+I^2 p_u^2)p_u - m^2gl^3\bar{q}_a I (2-c_a^2)(2s_u + s_{ua})
+    }{ml^2(3+2c_a)(1+I^2 p_u^2)}
+    .
+\]
+
 \section{Proving the Acrobot Gains Energy}

### 2020-10-27 latex/vlp.tex
Major edits to the VLP chapter
The proof was lacking some detail, so I added it according to manfredi's
suggestions. It is now in a slightly better format.

@@
 phase \((q,p)\).
-The constrained dynamics are Hamiltonian, and Theorem \ref{thm:zero-dynamics}
-tells us they are described entirely by \((\dot{q},\dot{p})\) with \(l\)
+By Theorem \ref{thm:zero-dynamics}, the constrained dynamics
+are described entirely by \((\dot{q},\dot{p})\) with \(l\)
 replaced by \(L(q,p)\):
-\begin{align}\label{eqn:vlp-hamiltonian}
-   \mathcal{H}(q,p) &= \frac{1}{2} \frac{p^2}{mL^2} - \frac{1}{2}\dot{L}^2 -
-      mgL\cos(q)
-   , \\
-  &\begin{cases}
+\begin{equation}\label{eqn:vlp-hamiltonian}
+   \begin{cases}
       \dot{q} = \frac{p}{m L^2} \\
@@
       . \\ 
-   \end{cases} \nonumber
-\end{align}
-Note that we suppress the function notation of \(L(q,p)\) and
-\(\dot{L}(q,p)\) for clarity.
+   \end{cases}
+\end{equation}
+Note that we suppress the function notation of \(L(q,p)\) for clarity.
 
@@
 \begin{defn}\label{defn:energy-injection}
-   Let \(D \subset \R^2\) with \(0 \in D\).
+   Let \(D \subset \R^2\) be open with \(0 \in D\).
    Let \(f : D \rightarrow \R^2\) be locally Lipschitz with \(f(0) = 0\).
@@
    sets on \(\Gamma\). 
-   The derivative of \(E_\text{avg}\) along (\ref{eqn:negative-time-vlp}) is
-   \[
-      \dot{E}_\text{avg} = \frac{g\sin(q)p \left(L(\theta)^3 - l_\text{avg}^3\right)}
+   The derivative of \(E_\text{avg}\) along (\ref{eqn:vlp-hamiltonian}) is
+   \begin{equation}\label{eqn:eavg-dot-pos}
+      \dot{E}_\text{avg}^+ = -\frac{g\sin(q)p \left(L(\theta)^3 - l_\text{avg}^3\right)}
               {l_\text{avg}^2L(\theta)^2}
+      ,
+   \end{equation}
+   and the derivative of \(E_\text{avg}\) along (\ref{eqn:negative-time-vlp}) is
+   \begin{equation}\label{eqn:eavg-dot-neg}
+      \dot{E}_\text{avg}^- = - \dot{E}_\text{avg}^+
       .
-   \]
+   \end{equation}
    The sign of \(\sin(q)p\) depends only on the quadrant in the
@@
    \]
-   The derivative of \(E_\text{avg}\) is never positive since
+   The positive-time derivative (\ref{eqn:eavg-dot-pos}) is nonnegative since
    \begin{align}\label{eqn:vlp-proof-neg-inv}
-    \sign{\dot{E}_\text{avg}} &= \sign{\sin(q)p \left(L(\theta)^3 - l_\text{avg}^3\right)}
+    \sign{\dot{E}_\text{avg}^+} &= \sign{-\sin(q)p \left(L(\theta)^3 - l_\text{avg}^3\right)}
           \nonumber \\
-          &= \sign{\sin(2\theta) \left(L(\theta) - l_\text{avg}\right)}
+          &= \sign{-\sin(2\theta) \left(L(\theta) - l_\text{avg}\right)}
           \nonumber \\
-          &= -\sign{(l_\text{avg} - L(\theta))\sin(2\theta)} 
+          &= \sign{(l_\text{avg} - L(\theta))\sin(2\theta)} 
           \nonumber \\
-          &\leq 0 \text{ (by assumption)}
+          &\geq 0 \text{ (by assumption)}
       .
    \end{align}
-   Thus, \(E_\text{avg}\) is a Lyapunov function, which means the negative-time system is
+   Likewise, the negative-time derivative (\ref{eqn:eavg-dot-neg}) is never
+   positive, so \(E_\text{avg}\) is a Lyapunov function and the negative-time system is
    stable \cite{lyapunov}.
 
-   The linearization of (\ref{eqn:negative-time-vlp}) at the upright equilibrium
-   \((\pi, 0)\) is given by
-   \begin{equation}\label{eqn:linearized-vlp-neg-time}
-      df_{(\pi,0)} = \begin{bmatrix}
-         0 & -\frac{1}{mL(0)^2} \\
-         -m g L(0) & 0 \\
-      \end{bmatrix}
-      ,
-   \end{equation}
-   which has eigenvalues at \(\lambda = \pm \sqrt{\frac{g}{L(0)}}\).
-   The positive-time system (\ref{eqn:vlp-hamiltonian}) has the same eigenvalues
-   at this equilibrium.
+   Let us briefly analyze the upright equilibrium.
    Label the set of initial conditions converging to \((\pi, 0)\) in
@@
    in negative time by \(\Pi_0^-\).
-   Taking the sublevel set
+   Take now the sublevel set
    \[
-      E_{(\pi,0)} := \left\{ (q,p) \in \Gamma \mid E_\text{avg}(q,p) <
+      E_{(\pi,0)} := \left\{ (q,p) \in \Gamma \mid E_\text{avg}(q,p) \leq
          E_\text{avg}(\pi,0) \right\}
+      .
+   \] 
+   From (\ref{eqn:vlp-proof-neg-inv}), all sublevel sets of \(E_\text{avg}\) are
+   negatively invariant in positive-time. 
+   Hence, the complement \(\left(E_{(\pi,0)}\right)^C = \Gamma \backslash E_{(\pi,0)}\) is
+   positively invariant.
+   If the system is initialized on \(\left(E_{(\pi,0)}\right)^C\), it cannot
+   posibly converge to \((\pi,0)\);
+   hence, \(\Pi_0^+ \subset E_{(\pi,0)}\).
+   Similarly, positive invariance of \(E_{(\pi,0)}\) in negative-time allows us
+   to show that \(\Pi_0^- \subset \left(E_{(\pi,0)}\right)^C \cup (\pi,0)\).
+
+   The linearization of (\ref{eqn:vlp-hamiltonian}) at \((\pi, 0)\) is given by
+   \begin{equation}\label{eqn:linearized-vlp}
+      \begin{bmatrix}
+         0 & \frac{1}{mL(0)^2} \\
+         m g L(0) & 0 \\
+      \end{bmatrix}
       ,
-   \] 
-   it is clear form (\ref{eqn:vlp-proof-neg-inv}) that 
-   \(\Pi_0^+ \subset E_{(\pi,0)}\)
-   and 
-   \(\Pi_0^- \subset \left(\Gamma \backslash E_{(\pi,0)} \right) \cup (\pi,0)\).
-   Since only one eigenvalue of (\ref{eqn:linearized-vlp-neg-time})
-   is negative, the stable manifold theorem tells us that both \(\Pi_0^+\) and
-   \(\Pi_0^-\) are one-dimensional \cite{stable_manifold}.
+   \end{equation}
+   which has eigenvalues at \(\lambda = \pm \sqrt{\frac{g}{L(0)}}\).
+   Linearizing (\ref{eqn:negative-time-vlp}) at this point yields the same eigenvalues.
+   Since one eigenvalue of (\ref{eqn:linearized-vlp})
+   is positive and one is negative, the stable manifold theorem tells
+   us that both \(\Pi_0^+\) and \(\Pi_0^-\) are one-dimensional
+   \cite{stable_manifold}.
    This implies the origin cannot be globally asymptotically stable for
@@
 
-   Define \(D := \Gamma \backslash \Pi_0^-\) to be the domain of interest.
-   Using a Krasovskii-LaSalle argument \cite{krasovskii_lasalle}, 
-   we will prove that (\ref{eqn:negative-time-vlp}) asymptotically converges to
+   We cannot linearize the dynamics at the origin to prove it is a repeller,
+   because \(\theta\) is not defined there.
+   Instead, let \(D := \Gamma \backslash \Pi_0^-\).
+   We will prove that (\ref{eqn:negative-time-vlp}) asymptotically converges to
    \((0,0)\) if it is initialized on \(D\).
-   First, we define the zero-set
+   First, define the zero-set
    \[
@@
    Krasovskii-LaSalle tells us that solutions to (\ref{eqn:negative-time-vlp})
-   converge to the largest invariant set in \(Z \cap D\).
+   converge to the largest invariant set in \(Z \cap D\)
+   \cite{krasovskii_lasalle}.
    Since \(f(q,p) = (0,0)\) on \(D\) only when \((q,p) = (0,0)\),
@@
    Let \(x(t) = (q(t),p(t))\) be the orbit for the positive-time system.
-   Let \(K \subset \Gamma\) be compact and assume \(x(0) \in K\backslash \Pi_0^+\).
+   Let \(K \subset \Gamma\) be compact and assume 
+   \(x(0) \in K\backslash \Pi_0^+\) and \(x(0) \neq (0,0)\).
    For a proof by contradiction, suppose that for all \(T > 0\) there exists
@@
    All sublevel sets of \(E_\text{avg}\) are compact, negatively
-   invariant (by (\ref{eqn:vlp-proof-neg-inv})), and \(E_k\) contains \(K\).
+   invariant, and \(E_k\) contains \(K\).
    We assumed \(x(t') \in K\), so it follows that \(x(t') \in E_k\). 
@@
    Since \(E_k\) is compact, \(x(t)\) has a positive limit set in
-   \(E_k\) which (by Poincar\'{e}-Bendixson) is either an equilibrium or a
+   \(E_k\) which, by Poincar\'{e}-Bendixson, is either an equilibrium or a
    closed orbit \cite{poincare_bendixson}.
-   The system cannot have closed orbits, since this would violate asymptotic
-   stability of \((0,0)\) in negative time. 
-   Hence, the positive limit set must be one of the equilibria. 
-   The origin is a repeller, which means \(x(t)\) must be converging to 
-   \((\pi,0)\).
+
+   Suppose the system has a closed orbit. It would necessarily be contained in a
+   level set of \(E_\text{avg}\) (since this is a non-decreasing function and
+   closed orbits are periodic).
+   Thus, \(E_\text{avg}\) is constant along the orbit, which means the
+   orbit is contained in one of the axes because that is the only region where
+   \(\dot{E}_\text{avg}^+ = 0\).
+   Since the only invariant subsets of the axes are equilibria, the system cannot
+   contain closed orbits.
+
+   Hence, the positive limit set of \(x(t)\) must be one of the equilibria. 
+   The origin is a repeller and \(x(0) \neq (0,0)\), which means \(x(t)\) is
+   converging to \((\pi,0)\).
    This is a contradiction, since we assumed \(x(0) \not \in \Pi_0^+\).
 
-   We have shown that orbits of the constrained dynamics will leave compact sets in
-   finite time, including all sublevel sets of \(E_\text{avg}\). 
-   The orbit \((q(t),p(t))\) is spiraling away from the origin and, for almost
-   all initial conditions, will not converge to the upper equilibrium.
-   Thus, the VLP is gaining energy on average because its momentum is increasing.
-   Flipping the inequality of (\ref{eqn:vlp-energy-gain-condition}) 
-   means \(E_\text{avg}\) is a Lyapunov function for the positive time system;
-   by the same arguments presented here, the VLP loses energy while
-   converging asymptotically to \((0,0)\).
+   We conclude that the VNHC is injecting energy into the VLP.
+   By flipping the inequality of (\ref{eqn:vlp-energy-gain-condition})
+   we find the VLP is gaining energy in negative-time, so the VNHC would be
+   dissipating energy.
 \end{proof}

### 2020-10-28 latex/acrobot.tex
Added comments on what the proof should look like for the acrobot
First completion draft of acrobot constraint chapter

@@
 Finally, it is important to note that a real gymnast cannot swing their legs in
-full circles; 
+full circles, though they are usually flexible enough to raise them parallel to
+the floor; 
 for this reason, we assume that \(q_a \in [-Q_a, Q_a]\) where 
-\(Q_a \in ]-\pi, \pi[\).
+\(Q_a \in [\frac{\pi}{2}, \pi[\).
 
@@
 % proof (both as a sketch and in a figure)
-One may be tempted to design a constraint \(q_a = f(\theta)\) 
-with \(\theta := \arctan_2(p_u,q_u)\), since it was so effective for the VLP in
-Chapter \ref{ch:vlp}.
-It can be challenging to show this type of constraint is regular.
-Recall that all constraints of the form \(h(q,p) = q_a - f(p_u)\)
-are regular, since the regularity equation (\ref{eqn:regularity-matrix-acrobot})
-becomes strictly positive.
-A momentum-based constraint approximates human motion may still be likely to
-inject energy into the acrobot.
+One may be tempted to design a constraint of the form \(q_a = \bar{q}_a\sin(\theta)\) 
+with \(\theta := \arctan_2(p_u,q_u)\), since a similar approach 
+was so effective for the VLP in Chapter \ref{ch:vlp}.
+Unfortunately, this constraint is not regular, and it is difficult to find any
+VNHC of the form \(q_a = f(\theta)\) where it is easy to prove regularity
+analytically.
+Instead, we will develop a constraint \(h(q,p) = q_a - f(p_u)\) which,
+as we saw earlier, is always regular and may still be likely to inject energy.
 
-To design the acrobot constraint, let us imagine what a human does on
+To design this constraint, let us imagine what a human does on
 a seated swing.
@@
 backward they retract them. 
-In reality, the motion is slightly more complicated than that.:
+In reality, the motion is slightly more complicated than that:
 as the swing gains speed, the person leans their body back while
@@
 that the legs can pivot both forwards and backwards.
-In this case, the legs would pivot in the direction of motion
-proportionally to the swing's speed.
+The legs cannot bend, so in an attempt to emulate motion the legs should pivot
+in the direction of motion proportionally to the swing's speed.
 One VNHC which emulates this process is \(q_a = \bar{q}_a\arctan_2( I p_u)\), as
@@
 Here, \(\bar{q}_a \in ]0,\frac{2 Q_a}{\pi}]\) so that 
-\(q_a \in [-Q_a, Q_a]\) and \(I > 0 \in \R\) is a fixed control parameter.
+\(q_a \in [-Q_a, Q_a]\) and \(I \in \R_{>0}\) is a fixed control parameter.
 
 This constraint is not a perfect emulation of a gymnast's movement -- during
-a giant the gymnast's legs are almost completely extended, while the constraint
+a giant the gymnast's legs are almost completely extended, while our constraint
 retracts the legs partially during rotations.
@@
 
-Let us now compute the constrained dynamics for this
+Let us now compute the constrained dynamics under
 (\ref{eqn:acrobot-constraint}).
@@
 \]
+The dynamics for \(p_u\) do not contain \(p_a\), so they remain unchanged.
+The constrained dynamics for \(q_u\) are given by 
+\begin{equation*}
+    \dot{q}_u = e_1\tpose \Minv(q) \begin{bmatrix}
+                    p_u \\ p_a(q_u,p_u)
+                \end{bmatrix} %\\
+\end{equation*}
+%        &= \frac{1}{ml^2(2-c_a^2)} 
+%            \left( p_u - 
+%                (1+c_a)\frac{(1+c_a)(1+I^2 p_u^2)p_u - m^2gl^3\bar{q}_a I (2-c_a^2)(2s_u + s_{ua})
+%                }{ml^2(3+2c_a)(1+I^2 p_u^2)}
+%            \right) 
+%        ,
+%\end{align*}
+which one can simplify into 
+\begin{equation*}
+    \dot{q_u} = \frac{(1+I^2 p_u^2)p_u - m^2gl^3\bar{q}_a I(2s_u + s_{ua})(1+c_a) }{ml^2(1+I^2 p_u^2)(3+2c_a)}
+    .
+\end{equation*}
+Hence, the constrained dynamics for the acrobot under
+(\ref{eqn:acrobot-constraint}) are
+\begin{equation}\label{eqn:acrobot-constrained-dynamics}
+    \begin{cases}
+        \dot{q}_u = \frac{(1+I^2 p_u^2)p_u - m^2gl^3\bar{q}_a I(2s_u + s_{ua})(1+c_a) }
+            {ml^2(1+I^2 p_u^2)(3+2c_a)}
+        , \\
+        \dot{p}_u = - m g l (2s_u + s_{ua})
+        .
+    \end{cases}
+\end{equation}
 
-\section{Proving the Acrobot Gains Energy}
+These dynamics do not always gain energy -- we need some additional conditions
+on the control parameters to guarantee this is true. 
+With these conditions, (\ref{eqn:acrobot-constraint}) injects energy into the
+acrobot, as is explained by the following Theorem.
+
+\begin{thm}\label{thm:acrobot-energy-stabilization}
+    Consider the simple acrobot (\ref{eqn:acrobot-hamiltonian}) constrained by
+    the regular VNHC \(h(q,p_u) = q_a - \bar{q}_a \arctan(I p_u)\).
+    Suppose \(\bar{q}_a \in ]0,1]\).
+    Then there exists a small enough\footnotemark \(I \in \R_{>0}\)
+    such that the VNHC injects energy. 
+    Setting \(I \in \R_{<0}\) allows the VNHC to dissipate energy.
+\end{thm}
+\footnotetext{In fact, we can be more precise than this. 
+    Letting \(\epsilon > 0\), we need 
+    \(I \leq \min\left\{I_1(\epsilon),I_2\right\}\) where
+    \(I_1(\epsilon) = \frac{\tan(\epsilon)}{(\pi-\epsilon)}
+    \left(\frac{m^2gl^3(\cos(\epsilon)-2\sin(\epsilon))}{\pi-\epsilon}
+    \left(\frac{2\bar{q}_a\tan(\epsilon)}{(1+\tan(\epsilon)^2)(\pi-\epsilon)}+5\right) 
+    \right)^{-1/2}\)
+    and \(I_2 = \sqrt{\frac{3}{10 m^2gl^3\bar{q}_a^2}}\).
+    See Apendix \ref{apx:I-bounds} for details.
+}
+
+Proving Theorem \ref{thm:acrobot-energy-stabilization} requires a change of
+coordinates and the use of perturbation theory.
+The full proof is given in Chapter \ref{sec:acrobot-proof}. 
+We will conclude this section with a high level sketch of the proof.
+
+The energy-based proof used in Chapter \ref{ch:vlp} does not work for the
+acrobot, since no ``average energy" function has positive derivative everywhere
+for these dynamics.
+We instead use a perturbation analysis. 
+When \(I = 0\), the acrobot behaves like a simple pendulum with two masses. 
+There is a suitable change of coordinates into a pseudo-radius 
+\(\mu \in \R_{>0}\) and pseudo-angle \(\alpha \in \mathbb{S}^1\) where \(\mu\)
+is constant along orbits of the pendulum and \(\alpha\) is strictly increasing.
+We can show there exists \(I > 0\) small enough such that
+the orbit of the acrobot in \((\mu,\alpha)\) coordinates
+tracks the pendulum motion to order \(I\).
+What's more, we can show that \(\mu\) is increasing on average, which means the
+acrobot is gaining energy.
+
+\section{Proving the Acrobot Gains Energy}\label{sec:acrobot-proof}
 % TODO: Prove this in sections
+% TODO: You don't have to use the proof of theta_dot being positive. Stick to
+% Manfredi's proof, and when you say "for I small enough" put in a footnote that
+% says "in fact, we can be more precise and set bounds for I given as I <
+% min{...}. See appendix for more detail."
+% Then, we can have an appendix where we derive the bounds on I so that
+% theta_dot is always positive. This will help the reader keep track and not
+% bother them too much.
 

### 2020-10-28 latex/vlp.tex
First completion draft of acrobot constraint chapter

@@
 
-\begin{defn}\label{defn:energy-injection}
+\begin{defn}\label{defn:energy-gain}
    Let \(D \subset \R^2\) be open with \(0 \in D\).
@@
 
-The second property of Definition \ref{defn:energy-injection} allows the system
+The second property of Definition \ref{defn:energy-gain} allows the system
 to have unstable equilibria, but not limit cycles nor closed orbits.
@@
 
-\begin{defn}
+\begin{defn}\label{defn:energy-injection}
    A regular VNHC \(h(\mathbf{q},\mathbf{p}) = 0\) 

### 2020-11-03 latex/vlp.tex
VLP proof updated, sending to Manfredi for comments.

@@
    \end{align}
-   Likewise, the negative-time derivative (\ref{eqn:eavg-dot-neg}) is never
-   positive, so \(E_\text{avg}\) is a Lyapunov function and the negative-time system is
-   stable \cite{lyapunov}.
+   Using (\ref{eqn:eavg-dot-neg}) and (\ref{eqn:vlp-proof-neg-inv}) 
+   we can determine that \(\sign{\dot{E}_\text{avg}^-} \leq 0\). 
+   Thus, \(E_\text{avg}\) is a Lyapunov function for
+   (\ref{eqn:negative-time-vlp}), which means the negative-time system is stable
+   \cite{lyapunov}.
 
@@
    From (\ref{eqn:vlp-proof-neg-inv}), all sublevel sets of \(E_\text{avg}\) are
-   negatively invariant in positive-time. 
-   Hence, the complement \(\left(E_{(\pi,0)}\right)^C = \Gamma \backslash E_{(\pi,0)}\) is
-   positively invariant.
-   If the system is initialized on \(\left(E_{(\pi,0)}\right)^C\), it cannot
-   posibly converge to \((\pi,0)\);
-   hence, \(\Pi_0^+ \subset E_{(\pi,0)}\).
-   Similarly, positive invariance of \(E_{(\pi,0)}\) in negative-time allows us
-   to show that \(\Pi_0^- \subset \left(E_{(\pi,0)}\right)^C \cup (\pi,0)\).
+   negatively invariant in positive-time, which means
+   complements of sublevel sets are positively invariant.
+   Hence, the system cannot possibly converge to \((\pi,0)\) if it is
+   initialized on 
+   \(\left(E_{(\pi,0)}\right)^C = \Gamma \backslash E_{(\pi,0)}\).
+   This implies that \(\Pi_0^+ \subset E_{(\pi,0)}\).
+   Similarly, the interior of \(E_{(\pi,0)}\) is positively
+   invariant for (\ref{eqn:negative-time-vlp}), 
+   which implies \(\Pi_0^- \subset \left(E_{(\pi,0)}\right)^C \cup (\pi,0)\).
 
@@
 
-   We cannot linearize the dynamics at the origin to prove it is a repeller,
-   because \(\theta\) is not defined there.
+   We cannot linearize (\ref{eqn:vlp-hamiltonian}) at \((0,0)\) to prove 
+   the origin is a repeller, because \(\theta(0,0)\) is not well-defined. 
    Instead, let \(D := \Gamma \backslash \Pi_0^-\).
@@
    Let \(K \subset \Gamma\) be compact and assume 
-   \(x(0) \in K\backslash \Pi_0^+\) and \(x(0) \neq (0,0)\).
+   \(x(0) \in K\backslash \left(\Pi_0^+ \cup \left\{(0,0)\right\}\right)\).
    For a proof by contradiction, suppose that for all \(T > 0\) there exists
@@
    sublevel set.
-   All sublevel sets of \(E_\text{avg}\) are compact, negatively
-   invariant, and \(E_k\) contains \(K\).
-   We assumed \(x(t') \in K\), so it follows that \(x(t') \in E_k\). 
-   Then, since \(E_k\) is negatively invariant, the half orbit
+
+   Since \(E_k\) contains \(K\) and \(x(t') \in K\), it follows that 
+   \(x(t') \in E_k\). 
+   Furthermore, negative invariance of \(E_k\) means the half orbit
    \(x([0,t'])\) is contained fully within \(E_k\).
-   In particular, \(x(T) \in E_k\);
-   but \(T\) was chosen arbitrarily, which implies \(x(t) \in E_k\) for all
+   In particular, \(x(T) \in E_k\).
+   However, \(T\) was chosen arbitrarily, which implies \(x(t) \in E_k\) for all
    \(t \in [0,\infty[\).
@@
    closed orbit \cite{poincare_bendixson}.
+   We will show this positive limit set must be an equilibrium and thereby
+   reach our contradiction.
 
-   Suppose the system has a closed orbit. It would necessarily be contained in a
-   level set of \(E_\text{avg}\) (since this is a non-decreasing function and
-   closed orbits are periodic).
+   Suppose the system had a closed orbit. 
+   The orbit would necessarily be contained in a
+   level set of \(E_\text{avg}\), since this is a non-decreasing function and
+   closed orbits are periodic.
    Thus, \(E_\text{avg}\) is constant along the orbit, which means the
-   orbit is contained in one of the axes because that is the only region where
-   \(\dot{E}_\text{avg}^+ = 0\).
-   Since the only invariant subsets of the axes are equilibria, the system cannot
-   contain closed orbits.
+   orbit is an invariant subset of the set \(Z\) consisting of the two axes.
+   However, the only invariant subsets of \(Z\) are equilibria, so 
+   the system cannot contain closed orbits.
 
-   Hence, the positive limit set of \(x(t)\) must be one of the equilibria. 
+   The lack of closed orbits means the positive limit set of \(x(t)\) must 
+   be one of the equilibria. 
    The origin is a repeller and \(x(0) \neq (0,0)\), which means \(x(t)\) is

### 2020-11-03 latex/vlp.tex
Updated a typo before sending VLP proof to Manfredi.

@@
    By flipping the inequality of (\ref{eqn:vlp-energy-gain-condition})
-   we find the VLP is gaining energy in negative-time, so the VNHC would be
+   we find the VLP is gaining energy in negative-time, so the VNHC is
    dissipating energy.

### 2020-11-03 latex/acrobot.tex
Updated proof chapter with perturbation and began oscillations.

@@
 Instead, we will develop a constraint \(h(q,p) = q_a - f(p_u)\) which,
-as we saw earlier, is always regular and may still be likely to inject energy.
+as we saw earlier, is always regular and is likely to inject energy.
 
-To design this constraint, let us imagine what a human does on
-a seated swing.
-When they are moving forward they extend their legs, and when they are moving
-backward they retract them. 
-In reality, the motion is slightly more complicated than that:
-as the swing gains speed, the person leans their body back while
+To design this constraint, let us begin by examining a person on a seated swing.
+The person extends their legs when the swing moves forwards, and retracts their
+legs when the swing is moving backwards.
+As the swing gains speed, the person leans their body back while
 extending their legs.
@@
 When the swing moves backward, they sit up and fully retract their legs
-underneath them.
+underneath them \cite{how_to_pump_a_swing}.
 
-Now imagine we have affixed the torso of the acrobot to the swing's rope, so 
-that the legs can pivot both forwards and backwards.
-The legs cannot bend, so in an attempt to emulate motion the legs should pivot
-in the direction of motion proportionally to the swing's speed.
+Now imagine the person's torso is affixed to the swing's rope so that they are
+always upright. 
+Imagine further that the swing has no seat at all, allowing the person to extend
+their legs beneath them. 
+This position is identical to that of a gymnast on a bar, which is why we can
+use the leg motion from seated swinging to design a controller for the acrobot.
+
+The acrobot's legs are rigid rods which cannot retract, so we emulate the person
+on a swing by pivoting the legs in the direction of motion. 
+To account for how a person leans back at higher speeds, the legs should pivot to an
+angle proportional to the swing's speed.
 One VNHC which emulates this process is \(q_a = \bar{q}_a\arctan_2( I p_u)\), as
 in Figure \ref{fig:qa-arctan}.
-Here, \(\bar{q}_a \in ]0,\frac{2 Q_a}{\pi}]\) so that 
-\(q_a \in [-Q_a, Q_a]\) and \(I \in \R_{>0}\) is a fixed control parameter.
+Here, \(\bar{q}_a \in ]0,\frac{2 Q_a}{\pi}]\)
+and \(I \in \R_{>0}\) is a fixed control parameter.
 
-This constraint is not a perfect emulation of a gymnast's movement -- during
-a giant the gymnast's legs are almost completely extended, while our constraint
-retracts the legs partially during rotations.
+This constraint does not perfectly recreate a gymnast's motion -- during
+a giant the gymnast's legs are almost completely extended, while this constraint
+pivots the legs partially during rotations.
 However, the motion looks similar enough that this constraint should provide a
@@
 (\ref{eqn:acrobot-constraint}).
-% TODO: should I do this for a generic h(q_a,p_u)? That's how I did it in my
-% notes...
-Note that \(dh_q = \begin{bmatrix}0 & 1\end{bmatrix}\tpose\), while
+Note that \(dh_q = \begin{bmatrix}0 & 1\end{bmatrix}\), while
 \[
@@
 \]
-Applying these to (\ref{eqn:g-qupu}), we get the solution for \(p_a\) on the
-constraint manifold as
+Inserting these to (\ref{eqn:g-qupu}), we get the solution for \(p_a\) on the
+constraint manifold:
 \[
@@
                 \end{bmatrix} %\\
+    ,
 \end{equation*}
@@
 %\end{align*}
-which one can simplify into 
+which can be simplified to 
 \begin{equation*}
@@
 
-These dynamics do not always gain energy -- we need some additional conditions
+These dynamics do not always gain energy -- we need additional conditions
 on the control parameters to guarantee this is true.
-With these conditions, (\ref{eqn:acrobot-constraint}) injects energy into the
-acrobot, as is explained by the following Theorem.
 
+% TODO: This needs modification. The energy injection is provable for any system
+% only during oscillations. For rotations, we need to do a numerical check of an
+% integral that depends on m,g,l. If this integral is positive, then the system
+% is gaining energy up to a point - it sems the integral decays to 0 so that you
+% can reach a limiting speed, which means you don't get full energy injection
+% everywhere on Gamma.
 \begin{thm}\label{thm:acrobot-energy-stabilization}
     Consider the simple acrobot (\ref{eqn:acrobot-hamiltonian}) constrained by
-    the regular VNHC \(h(q,p_u) = q_a - \bar{q}_a \arctan(I p_u)\).
-    Suppose \(\bar{q}_a \in ]0,1]\).
-    Then there exists a small enough\footnotemark \(I \in \R_{>0}\)
-    such that the VNHC injects energy. 
-    Setting \(I \in \R_{<0}\) allows the VNHC to dissipate energy.
+    (\ref{eqn:acrobot-constraint}) with \(\bar{q}_a \in ]0,1]\).
+    There exists a small enough\footnotemark \(I^\star \in \R_{>0}\)
+    such that the VNHC injects energy whenever \( I \in ]0,I^\star[\). 
+    If \(I \in ]-I^\star,0[\), the VNHC dissipates energy.
 \end{thm}
@@
 
-Proving Theorem \ref{thm:acrobot-energy-stabilization} requires a change of
-coordinates and the use of perturbation theory.
-The full proof is given in Chapter \ref{sec:acrobot-proof}. 
-We will conclude this section with a high level sketch of the proof.
+Unfortunately, the energy-based proof used in Chapter \ref{ch:vlp} does not
+readily transfer to the acrobot -- there is no obvious Lyapunov function for
+this constrained system.
+Proving Theorem \ref{thm:acrobot-energy-stabilization} requires an intelligent
+change of coordinates and the use of perturbation theory from
+\cite{khalil_nonlinear}.
+The full proof is given in Chapter \ref{sec:acrobot-proof}, so 
+we conclude this section with a high level sketch of the proof.
 
-The energy-based proof used in Chapter \ref{ch:vlp} does not work for the
-acrobot, since no ``average energy" function has positive derivative everywhere
-for these dynamics.
-We instead use a perturbation analysis. 
-When \(I = 0\), the acrobot behaves like a simple pendulum with two masses. 
 There is a suitable change of coordinates into a pseudo-radius 
-\(\mu \in \R_{>0}\) and pseudo-angle \(\alpha \in \mathbb{S}^1\) where \(\mu\)
-is constant along orbits of the pendulum and \(\alpha\) is strictly increasing.
-We can show there exists \(I > 0\) small enough such that
-the orbit of the acrobot in \((\mu,\alpha)\) coordinates
-tracks the pendulum motion to order \(I\).
-What's more, we can show that \(\mu\) is increasing on average, which means the
-acrobot is gaining energy.
+\(\mu \in \R_{>0}\) and pseudo-angle \(\alpha \in \mathbb{S}^1\) on the
+\((q_u,p_u)\)-plane where \(\dot{\mu} = 0\) and \(\dot{\alpha} > 0\) when 
+we set \(I = 0\).
+Perturbation theory shows that, for small enough \(I\), the radius \(\mu\)
+increases on average along \(\alpha\).
+Hence, the acrobot is gaining energy since orbits in the \((q_u,p_u)\)-plane
+spiral away from the origin, and the acrobot always begins performing rotations.
 
@@
 % bother them too much.
+Proving Theorem \ref{thm:acrobot-energy-stabilization} is not a simple task.
+In an effort to make the proof as clear as possible, we will break it down into
+the following segments:
+\begin{enumerate}
+    \item Background on perturbation theory and averaging.
+    \item Perturbation analysis for oscillations.
+    \item Perturbation analysis for rotations.
+\end{enumerate}
+
+When \(I = 0\), the constrained acrobot behaves like a single
+pendulum with masses at a distance \(l\) and \(2l\) from the pivot 
+(Figure \ref{fig:acrobot-I0}) whose energy
+\[
+    E(q_u,p_u) = \frac{p_u^2}{10ml^2} + 3(1 - \cos(q_u))
+    ,
+\]
+is conserved.
+Level sets of \(E\) are ellipses when \(E \leq E(\pi,0)\), which we call
+``oscillations"; and they are open curves when \(E > E(\pi,0)\), which we call
+``rotations". 
+Examples of these can be seen in Figure \ref{fig:pendulum-level-sets}.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[t]{0.45\textwidth}
+        % TODO: figure of acrobot with I = 0
+        %\includestandalone[]{images/acrobot_I0}
+        \caption{The simple pendulum with two masses.}
+        \label{fig:acrobot-I0}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.45\textwidth}
+        % TODO: figure showing oscillations and rotations as level curves of E
+        %\includestandalone[]{images/pendulum_level_sets}
+        \caption{Level sets of the pendulum energy function.
+            The red ellipse represents oscillations while the purple curve
+            represents rotations.}
+        \label{fig:pendulum-level-sets}
+    \end{subfigure}
+    \caption{Our constrained acrobot is a simple pendulum when \(I = 0\).}
+\end{figure}
+
+Drawing inspiration from \citet{dynamic_vhcs_stabilize_closed_orbits}, we want
+to find a change of coordinates \((q_u,p_u) \to (\alpha, \mu)\) with the
+following properties:
+\begin{itemize}
+    \item All oscillations in \((\alpha,\mu)\) coordinates are represented by
+        \(\mu = \text{const}\).
+    \item \(\alpha\) is a pseudo-angle living in \(\mathbb{S}^1\) which is
+        always increasing.
+\end{itemize}
+We also want to find a second set of coordinates where the same properties hold for
+rotations.
+Once we have these coordinates in hand, we can use perturbation theory to prove
+that \(\mu\) increases on average. 
+If this is true, the acrobot must be gaining energy since the momentum is
+increasing over time.
+
+\subsection*{Background on Perturbation Theory}
+Perturbation theory allows one to understand how solutions of nonlinear systems
+behave by studying a nominal system which can be analyzed analytically.
+The solutions of nonlinear systems are approximated by a Taylor-like expansion
+around this nominal system.
+
+\citet{khalil_nonlinear} considers a system of the form
+\begin{equation}\label{eqn:khalil-setup}
+    \begin{cases}
+        \dot{x} = f(t,x,\epsilon), \\
+        x(t_0) = \eta(\epsilon),
+    \end{cases}
+\end{equation}
+where 
+\(f : [t_0,t_1] \times D \times [-\epsilon_0,\epsilon_0] \rightarrow \R^n\) is
+sufficiently smooth.
+In the context of the acrobot, \(x\) is the phase in whichever coordinates make
+our analysis convenient;
+\(\epsilon\) is our control parameter \(I\);
+the function \(f\) is the constrained dynamics in \(x\)-coordinates;
+and \(\eta(\epsilon) \equiv \eta_0\) is a constant initial condition.
+To keep a notational reminder that we are applying perturbation theory to
+the acrobot, we will henceforth use \(I\) in place of \(\epsilon\).
+
+Letting \(I = 0\) we get the nominal system
+\begin{equation}\label{eqn:khalil-perturbation-nominal}
+    \begin{cases}
+        \dot{x}_0 = f(t,x,0) ,\\
+        x_0(t_0) = \eta_0 ,
+    \end{cases}
+\end{equation}
+which we need to solve for the explicit solution \(x_0(t)\) on \([t_0,t_1]\).
+One can use \(x_0(t)\) to compute a first-order approximation of
+the nonlinear system
+\begin{equation}\label{eqn:khalil-perturbation-firstorder}
+    \begin{cases}
+        \dot{x}_1 = \pdiff{f}{x}(t,x_0(t),0)x_1 + \pdiff{f}{I}(t,x_0(t),0)
+        , \\
+        x_1(t_0) = 0, \\
+    \end{cases}
+\end{equation}
+which is a linear time-varying system with solution \(x_1(t)\).
+
+We now paraphrase Khalil's Theorem 10.1 \cite{khalil_nonlinear} on the
+accuracy of perturbation analysis.
+\begin{thm}
+    Suppose \(f : [t_0,t_1] \times D \times [-I_0,I_0] \rightarrow \R^2\) is
+    \(C^1\), and that the nominal problem (\ref{eqn:khalil-perturbation-nominal}) has a
+    unique solution \(x_0(t) \in D\) on \([t_0,t_1]\).
+    Then there exists \(I^\star > 0\) such that for all \(|I| < I^\star\), the
+    solution \(x(t,I)\) to (\ref{eqn:khalil-setup}) satisfies
+    \[
+        \norm{x(t,I) - \left(x_0(t) + I x_1(t)\right)} \leq k |I^2|
+    \]
+    for some \(k > 0\).
+\end{thm}
+
+The above Theorem tells us that we can approximate the solution of the nonlinear
+system by the Taylor approximation \(x_0(t) + I x_1(t)\).
+In other words, when \(I\) is small enough solutions of the nonlinear system and
+the Taylor approximation are the same up to order \(I^2\) along compact time
+intervals.
+Since we know that the acrobot behaves like a pendulum at \(I = 0\), we can
+use this theory to more easily study the behaviour of orbits and prove energy
+injection properties of our VNHC.
+
+\subsection*{Perturbation Analysis for Oscillations}
+An oscillation of the nominal pendulum (when \(I = 0\)) is uniquely determined
+by its intersection \(\mu\) with the \(q_u\)-axis, as in Figure
+\ref{fig:mu-intersection}. Since \(\mu\) is to be a pseudo-radius and
+\(q_u \in ]-\pi,\pi]\), we have \(\mu \in ]0,\pi]\). 
+The corresponding energy level set is \(E(q_u,p_u) = 3mgl(1- \cos(\mu))\), which
+gives the relationship
+\begin{equation}\label{eqn:oscillation-pu2}
+    p_u^2 = 30m^2gl^3(\cos(q_u) - \cos(\mu))
+    .
+\end{equation}
+
+\begin{figure}
+    \centering
+    % TODO: Add a figure with one oscillation whose we mark the distance mu from
+    % the origin to the intersection on the q-axis
+    %\includestandalone[]{images/mu_intersection}
+    \caption{The pseudo-radius \(\mu\) for oscillations is within \(]0,\pi]\).}
+    \label{fig:mu-intersection}
+\end{figure}
+
+On this level set, \(q_u\) ranges between \([-\mu,\mu]\) and can be uniquely
+parameterized by \(q_u = \mu \cos(\alpha)\), where \(\alpha\) is our
+pseudo-angle.
+Substituting this into (\ref{eqn:oscillation-pu2}), we get
+\[
+    p_u^2 = 30m^2gl^3(\cos(\mu \cos(\alpha)) - \cos(\mu))
+    .
+\]
+We want to find \(p_u\) as a function of \((\alpha,\mu)\); noting that we can
+determine the sign of \(p_u\) from the sign of \(\sin(\alpha)\), we get the
+(clockwise) parameterization
+\begin{equation}\label{eqn:oscillation-pu}
+    p_u = -\sign{\sin(\alpha)} \sqrt{30m^2gl^3 \left(\cos(\mu c_\alpha) - c_\mu\right)}
+    ,
+\end{equation}
+which is, in fact, continuously differentiable.
+
+We have thus found a mapping \((\alpha,\mu) \mapsto (q_u,p_u)\) for which we
+need the inverse mapping.
+Notice from (\ref{eqn:oscillation-pu2}) that
+\[
+    \cos(\mu) = -\frac{p_u^2}{30m^2gl^3} + \cos(q_u) =: C_\mu(q_u,p_u)
+    .
+\]
+Since \(\mu \in ]0,\pi]\), we can uniquely express \(\mu\) by
+\begin{equation}\label{eqn:mu-qupu}
+    \mu = \arccos\left(C_\mu(q_u,p_u)\right)
+    .
+\end{equation}
+
+\subsection*{Perturbation Analysis for Rotations}
 

### 2020-11-04 latex/acrobot.tex
Updated the acrobot stuff again, still on oscillations.

@@
 (Figure \ref{fig:acrobot-I0}) whose energy
-\[
+\begin{equation}\label{eqn:acrobot-nominal-E}
     E(q_u,p_u) = \frac{p_u^2}{10ml^2} + 3(1 - \cos(q_u))
     ,
-\]
+\end{equation}
 is conserved.
-Level sets of \(E\) are ellipses when \(E \leq E(\pi,0)\), which we call
+Level sets of \(E\) are ellipses when \(E < E(\pi,0)\), which we call
 ``oscillations"; and they are open curves when \(E > E(\pi,0)\), which we call
@@
 \subsection*{Perturbation Analysis for Oscillations}
-An oscillation of the nominal pendulum (when \(I = 0\)) is uniquely determined
-by its intersection \(\mu\) with the \(q_u\)-axis, as in Figure
-\ref{fig:mu-intersection}. Since \(\mu\) is to be a pseudo-radius and
-\(q_u \in ]-\pi,\pi]\), we have \(\mu \in ]0,\pi]\). 
-The corresponding energy level set is \(E(q_u,p_u) = 3mgl(1- \cos(\mu))\), which
-gives the relationship
+The nominal pendulum (\ie when \(I = 0\)) oscillates whenever the
+nominal energy (\ref{eqn:acrobot-nominal-E}) is less than \(E(\pi,0)\).
+Hence, the domain of oscillations on \(\Gamma\) is given by
+\[
+    D := \left\{ (q_u,p_u) \in \Gamma | E(q_u,p_u) < E(\pi,0)\right\}
+    .
+\]
+We wish to find a transformation \(T(q_u,p_u)\) on \(D\) into
+\((\alpha,\mu)\)-coordinates where \(\alpha \in \mathbb{S}^1\) is a pseudo-angle
+and \(\mu > 0\) is a pseudo-radius.
+Since the energy of oscillation can be determined by the orbit's intersection
+with the \(q_u\)-axis and \(q_u \in ]-\pi,\pi[\) on \(D\), 
+we set \(\mu \in ]0,\pi[\) as in Figure \ref{fig:mu-intersection}.
+The transformation we want is therefore a diffeomorphism of the form
+\begin{align*}
+    T : D\backslash \{(0,0)\} &\rightarrow \mathbb{S}^1 \times \, ]0,\pi], \\
+    (q_u, p_u) &\mapsto (\alpha,\mu)
+    .
+\end{align*}
+
+\begin{figure}
+    \centering
+    % TODO: Add a figure with one oscillation mark at a distance mu from
+    % the origin to the intersection on the q-axis. We also draw the boundary of
+    % D and shade the interior
+    %\includestandalone[]{images/mu_intersection}
+    \caption{The domain \(D\) where a pendulum oscillates.
+    The pseudo-radius coordinate \(\mu\) corresponds to the
+    intersection of an orbit of oscillation with the \(q_u\)-axis.}
+    \label{fig:mu-intersection}
+\end{figure}
+
+The energy level set corresponding to the intersection \((q_u,p_u) = (\mu,0)\)
+is 
+\[
+    \left\{(q_u,p_u) \in \Gamma \mid E(q_u,p_u) = 3mgl(1- \cos(\mu))\right\}
+    ,
+\]
+which gives the relationship
 \begin{equation}\label{eqn:oscillation-pu2}
@@
 \end{equation}
-
-\begin{figure}
-    \centering
-    % TODO: Add a figure with one oscillation whose we mark the distance mu from
-    % the origin to the intersection on the q-axis
-    %\includestandalone[]{images/mu_intersection}
-    \caption{The pseudo-radius \(\mu\) for oscillations is within \(]0,\pi]\).}
-    \label{fig:mu-intersection}
-\end{figure}
-
 On this level set, \(q_u\) ranges between \([-\mu,\mu]\) and can be uniquely
@@
 \end{equation}
-which is, in fact, continuously differentiable.
+which is, in fact, smooth.
 
-We have thus found a mapping \((\alpha,\mu) \mapsto (q_u,p_u)\) for which we
-need the inverse mapping.
+We have thus found a transformation \(T\inv(\alpha,\mu) = (q_u,p_u)\).
+We need the inverse of this map to get our diffeomorphism \(T(q_u,p_u)\).
 Notice from (\ref{eqn:oscillation-pu2}) that
@@
 \end{equation}
+Next we need to find \(\alpha\). 
+Recall that 
+\[
+    \cos(\alpha) = \frac{q_u}{\mu}
+    ,
+\] 
+which means 
+\[
+    \sin(\alpha) = \pm \sqrt{1 - \frac{q_u^2}{\mu^2}}
+    .
+\]
+Notice from (\ref{eqn:oscillation-pu}) that 
+\(\sign{\sin(\alpha)} = -\sign{p_u}\).
+We use these facts to deduce that
+\begin{equation}\label{eqn:alpha-qupu}
+    \alpha = \left.
+        \arctan_2\left( -\sign{p_u}\sqrt{1 - \frac{q_u^2}{\mu^2}}, \frac{q_u}{\mu}\right)
+        \right|_{\mu = \arccos(C_\mu(q_u,p_u))}
+    ,
+\end{equation}
+and thus have our transformation \(T(q_u,p_u)\) into
+\((\alpha,\mu)\)-coordinates on \(D\).
+
+The acrobot's constrained dynamics in
+\((\alpha,\mu)\)-coordinates can be computed by evaluating
+\[
+    \left.\begin{cases}
+        \dot{\mu} = 
+        \begin{bmatrix} \pdiff{\mu}{q_u} & \pdiff{\mu}{p_u} \end{bmatrix}
+        \begin{bmatrix} \dot{q}_u \\ \dot{p}_u \end{bmatrix}
+        , \\
+        \dot{\alpha} = \begin{bmatrix} \pdiff{\alpha}{q_u} & \pdiff{\alpha}{p_u} \end{bmatrix}
+        \begin{bmatrix} \dot{q}_u \\ \dot{p}_u \end{bmatrix}
+        .
+    \end{cases} \right|_{(q_u,p_u) = T\inv(\mu,\alpha)}
+\]
+These dynamics are incredibly complicated and difficult to express, so we will
+write them simply as
+\[
+    \begin{cases}
+        \dot{\mu} = f_\mu(\alpha,\mu,I)
+        ,\\
+        \dot{\alpha} = f_\alpha(\alpha,\mu,I)
+        .
+    \end{cases}
+\]
+
+To find the nominal system we set \(I = 0\); 
+MATLAB's symbolic toolbox operations tell us that
+\begin{equation}\label{eqn:acrobot-nominal-ma-dynamics}
+    \begin{cases}
+        \dot{\mu} = 0
+        , \\
+        \dot{alpha} = \sqrt{\frac{6g}{5l}} 
+        \sqrt{\frac{\cos(\mu\cos(\alpha)) - \cos(\mu)}{\mu^2 \sin(\alpha)^2}}
+        .
+    \end{cases}
+\end{equation}
 

### 2020-11-05 matlab/thesis_sims/vlp_l_T.m
Added a bunch of MATLAB code for thesis-related sims

@@ -1,0 +1,51 @@
+%% VLP Continuous Controller $l_T(\theta)$
+% This function defines the continuous version of the time-optimal
+% controller for the VLP. It attaches sinusoids of the correct frequency to
+% the original time-optimal controller from
+% from the paper "Pumping a Swing by Standing and Squatting".
+%
+% Inputs:
+% * theta = atan2(p,q)
+% * lu = lower actuator limit for l(theta). Domain is ]0,lb[
+% * lb = upper actuator limit for l(theta). Domain is ]lu,inf[
+% * T = amount by which we truncate each part of the time-optimal
+% controller. Domain is [0,pi/2].
+%
+% Output:
+% * l_T(theta). If T = 0, we return vlp_l_optimal. If T = pi/2, we return
+% vlp_l_smooth.
+function l = vlp_l_T(theta,lu,lb,T)
+% Return vlp_l_optimal if T = 0
+if T == 0
+    l = vlp_l_optimal(theta,lu,lb);
+    return;
+elseif T == pi/2
+     l = vlp_l_smooth(theta,lu,lb);
+    return;
+end
+% Set the frequency of the sinusoid
+w = pi/T;
+% Compute Deltal and lavg
+dl = (lb - lu)/2;
+lavg = (lb + lu)/2;
+% If theta is within [-pi,-pi+T/2[ we use a neg. sinusoid shifted by -pi
+if theta < -pi + T/2
+    l = -dl*sin(w*(theta+pi)) + lavg;
+% If it's within ]pi-T/2,pi] we have a neg. sin shifted by pi
+elseif theta > pi-T/2
+    l = -dl*sin(w*(theta-pi)) + lavg;
+% If it's within T/2 of 0, it's a neg. sin with no shift
+elseif -T/2 < theta && theta < T/2
+    l = -dl*sin(w*theta) + lavg;
+% If it's around +- pi/2, it's a positive sin with a shift of +- pi/2
+elseif -pi/2 - T/2 < theta && theta < -pi/2 + T/2
+    l = dl*sin(w*(theta+pi/2)) + lavg;
+elseif pi/2 - T/2 < theta && theta < pi/2 + T/2
+    l = dl*sin(w*(theta-pi/2)) + lavg;
+% If it's none of these cases, then we are within the domain of the optimal
+% controller, so we return the optimal result.
+else
+    l = vlp_l_optimal(theta,lu,lb);
+end
+end
+

### 2020-11-05 matlab/thesis_sims/vlp_l_optimal.m

@@ -1,0 +1,21 @@
+%% VLP Optimal Controller $l^\star(\theta)$
+% This function defines the time-optimal controller for the VLP, based on
+% the results from the paper "Pumping a Swing by Standing and Squatting".
+%
+% Inputs:
+% * theta = atan2(p,q)
+% * lu = lower actuator limit for l(theta). Domain is ]0,lb[
+% * lb = upper actuator limit for l(theta). Domain is ]lu,inf[
+%
+% Output:
+% * l^\star(theta), the time-optimal controller.
+function l = vlp_l_optimal(theta,lu,lb)
+% If theta is within ]-pi,-pi/2] or [0,pi/2], we stand.
+if theta <= -pi/2 || (0 <= theta && theta <= pi/2)
+    l = lu;
+% Otherwise, if theta is in ]-pi/2,0[ or ]pi/2,pi], we squat.
+else 
+    l = lb;
+end
+end
+

### 2020-11-05 matlab/thesis_sims/vlp_l_smooth.m

@@ -1,0 +1,20 @@
+%% VLP Smooth Controller $l(\theta)$
+% This function defines the smoothed version of the time-optimal
+% controller for the VLP. It converts the results
+% from the paper "Pumping a Swing by Standing and Squatting" into a
+% sinusoid around the average length of the pendulum.
+%
+% Inputs:
+% * theta = atan2(p,q)
+% * lu = lower actuator limit for l(theta). Domain is ]0,lb[
+% * lb = upper actuator limit for l(theta). Domain is ]lu,inf[
+%
+% Output:
+% * l(theta) = -Delta * sin(2*theta) + lavg
+function l = vlp_l_smooth(theta,lu,lb)
+% Compute Delta_l and lavg
+dl = (lb - lu)/2;
+lavg = (lb + lu)/2;
+% Return the smoothed controller
+l = -dl*sin(2*theta) + lavg;
+end

### 2020-11-05 latex/acrobot.tex
Updated a lot of the proof, almost done oscillations for acrobot

@@
 \begin{equation*}
-    \dot{q_u} = \frac{(1+I^2 p_u^2)p_u - m^2gl^3\bar{q}_a I(2s_u + s_{ua})(1+c_a) }{ml^2(1+I^2 p_u^2)(3+2c_a)}
+    \dot{q_u} = \frac{(1+I^2 p_u^2)p_u + m^2gl^3\bar{q}_a I(2s_u + s_{ua})(1+c_a) }{ml^2(1+I^2 p_u^2)(3+2c_a)}
     .
@@
 (\ref{eqn:acrobot-constraint}) are
-\begin{equation}\label{eqn:acrobot-constrained-dynamics}
-    \begin{cases}
-        \dot{q}_u = \frac{(1+I^2 p_u^2)p_u - m^2gl^3\bar{q}_a I(2s_u + s_{ua})(1+c_a) }
+\begin{align}\label{eqn:acrobot-constrained-dynamics}
+    \dot{q}_u &= \frac{(1+I^2 p_u^2)p_u + m^2gl^3\bar{q}_a I(2s_u + s_{ua})(1+c_a) }
             {ml^2(1+I^2 p_u^2)(3+2c_a)}
         , \\
-        \dot{p}_u = - m g l (2s_u + s_{ua})
-        .
-    \end{cases}
-\end{equation}
+    \dot{p}_u &= - m g l (2s_u + s_{ua})
+    . \nonumber
+\end{align}
 
@@
 \end{equation}
-which we need to solve for the explicit solution \(x_0(t)\) on \([t_0,t_1]\).
-One can use \(x_0(t)\) to compute a first-order approximation of
+which we need to solve for the explicit solution \(x_0(t,\eta_0)\) on \([t_0,t_1]\).
+One can use \(x_0(t,\eta_0)\) to compute the first-order approximation of
 the nonlinear system 
+\(x(t,\eta_0) = x_0(t,\eta_0) + I x_1(t,\eta_0) + R(t,I)\), where the remainder
+term \(R(t,I)\) satisfies
+\[
+    \lim \limits_{I \to 0} \frac{R(t,I)}{I} = 0
+    ,
+\]
+and \(x_1(t,\eta_0)\) is the solution to
 \begin{equation}\label{eqn:khalil-perturbation-firstorder}
     \begin{cases}
-        \dot{x}_1 = \pdiff{f}{x}(t,x_0(t),0)x_1 + \pdiff{f}{I}(t,x_0(t),0)
+        \dot{x}_1 = \pdiff{f}{x}(t,x_0(t,\eta_0),0)x_1 + \pdiff{f}{I}(t,x_0(t,\eta_0),0)
         , \\
-        x_1(t_0) = 0, \\
+        x_1(t_0,\eta_0) = 0. \\
     \end{cases}
 \end{equation}
-which is a linear time-varying system with solution \(x_1(t)\).
 
@@
     \(C^1\), and that the nominal problem (\ref{eqn:khalil-perturbation-nominal}) has a
-    unique solution \(x_0(t) \in D\) on \([t_0,t_1]\).
+    unique solution \(x_0(t,\eta_0) \in D\) on \([t_0,t_1]\).
     Then there exists \(I^\star > 0\) such that for all \(|I| < I^\star\), the
-    solution \(x(t,I)\) to (\ref{eqn:khalil-setup}) satisfies
+    solution \(x(t,I,\eta_0)\) to (\ref{eqn:khalil-setup}) satisfies
     \[
-        \norm{x(t,I) - \left(x_0(t) + I x_1(t)\right)} \leq k |I^2|
+        \norm{x(t,I,\eta_0) - \left(x_0(t,\eta_0) + I x_1(t,\eta_))\right)} \leq k |I^2|
     \]
@@
 The above Theorem tells us that we can approximate the solution of the nonlinear
-system by the Taylor approximation \(x_0(t) + I x_1(t)\).
+system by the Taylor approximation 
+\(x(t,\eta_0) \approx x_0(t,\eta_0) + I x_1(t,\eta_0)\).
 In other words, when \(I\) is small enough solutions of the nonlinear system and
@@
 \((\alpha,\mu)\)-coordinates can be computed by evaluating
+\begin{align*}
+    \dot{\mu} &= \left.\pdiff{\mu(q_u,p_u)}{q}_u\dot{q_u} +
+         \pdiff{\mu(q_u,p_u)}{p_u}\dot{p}_u
+         \right|_{(q_u,p_u) = T\inv(\mu,\alpha)}
+    , \\
+    \dot{\alpha} &= \left.\pdiff{\alpha(q_u,p_u)}{q_u}\dot{q}_u + 
+        \pdiff{\alpha(q_u,p_u)}{p_u}\dot{p}_u
+        \right|_{(q_u,p_u) = T\inv(\mu,\alpha)}
+    .
+\end{align*}
+with \((\dot{q}_u,\dot{p}_u)\) the given by 
+ (\ref{eqn:acrobot-constrained-dynamics}).
+
+These dynamics are much too large to write out, so we simply denote them by
+\begin{align}\label{eqn:acrobot-mu-dot}
+    \dot{\mu} &= f_\mu(\alpha,\mu,I)
+    ,\\
+    \label{eqn:acrobot-alpha-dot}
+    \dot{\alpha} &= f_\alpha(\alpha,\mu,I)
+    .
+\end{align}
+To find the nominal system we set \(I = 0\);
+MATLAB's symbolic toolbox evaluates these expressions as
+\begin{align}\label{eqn:acrobot-mu-dot-nom}
+    \dot{\mu} &= 0
+    , \\
+    \label{eqn:acrobot-alpha-dot-nom}
+    \dot{\alpha} &= \sqrt{\frac{6g}{5l}} 
+        \sqrt{\frac{\cos(\mu\cos(\alpha)) - \cos(\mu)}{\mu^2 \sin(\alpha)^2}}
+    .
+\end{align}
+One can confirm that \(\dot{\alpha} > 0\) for every \(\mu \in ]0,\pi[\) -- this
+is expected, since the nominal system is a single pendulum.
+By continuity of (\ref{eqn:acrobot-mu-dot}) with respect to \(I\), 
+\(\dot{\alpha}\) remains positive for \(I\) small enough on the domain \(D\).
+
+Suppose at time \(t = 0\) we have initialized the system at 
+\((q_u,p_u) = (\mu_0,0)\), which corresponds to \((\alpha,\mu) = (0,\mu_0)\).
+Since \(\alpha\) is continuously increasing, we know that the orbit will do one
+of two things: it will either leave \(D\) so that the acrobot is now rotating,
+or it will remain in \(D\) and hit the \(q_u\)-axis at some point 
+\((\alpha,\mu) = (\pi,m)\) as is illustrated in Figure
+\ref{fig:acrobot-possible-orbits}.
+From there, the acrobot could escape \(D\) along the second half of the orbit or
+return to some point \((\alpha,\mu) = (0,\tilde{m})\).
+If \(\tilde{m} < \mu_0\) then the acrobot has lost energy; 
+if \(\tilde{m} > \mu_0\) the acrobot has gained energy; 
+and if \(\tilde{m} = \mu_0\), the acrobot is on a closed orbit.
+Which situation occurs might depend entirely on the choice of
+\(\mu_0\).
+
+\begin{figure}
+    \centering 
+    % TODO: Draw the domain D, an initial point (q0,p0) = (mu0,0), and show the
+    % 4 clockwise half orbits: when the orbit exits D on the bottom left corner;
+    % when the orbit hits the left q-axis at (-mu0-x,0); when it hits (-mu0,0);
+    % and when it hits (-mu0+x,0) for some x > 0.
+    %includestandalone[width=0.5\textwdith]{images/acrobot_possible_orbits}
+    \caption{If the acrobot is initialized at \((q_u,p_u) = (\mu_0,0)\), then
+        either the orbit escapes \(D\) into the rotation
+        zone (purple), or the orbit hits the other half of the \(q_u\)-axis 
+        with a higher (red), equal (black), or lower (blue) angle than when it
+        started.}
+    \label{fig:acrobot-possible-orbits}
+\end{figure}
+
+Since \(\dot{\alpha} > 0\), we could theoretically find some time scaling 
+\(t = \tau(\alpha)\) and use \(\alpha\) as a new ``time" variable.
+Setting \(\hat{\mu}(\alpha) := \mu(\tau(\alpha))\) allows us to ignore the
+dynamics of \(\alpha\) and study the evolution of \(\mu\) as a 
+function of \(\alpha\) rather than as a function of \(t\) (assuming the acrobot
+does not leave \(D\)).
+The dynamics of \(\hat{\mu}\) are
 \[
-    \left.\begin{cases}
-        \dot{\mu} = 
-        \begin{bmatrix} \pdiff{\mu}{q_u} & \pdiff{\mu}{p_u} \end{bmatrix}
-        \begin{bmatrix} \dot{q}_u \\ \dot{p}_u \end{bmatrix}
+    \diff{\hat{\mu}}{\alpha} = 
+    \left.\diff{\mu}{t} \diff{t}{\alpha} \right|_{t =\tau(\alpha)}
+    ,
+\] 
+yielding the ``time"-varying scalar ODE
+\begin{equation}\label{eqn:muhat-dot}
+    \begin{cases}
+        \diff{\hat{\mu}}{\alpha} 
+        = \frac{f_\mu(\tau(\alpha),\mu,I)}{f_\alpha(\tau(\alpha),\mu,I)}
+        =: g(\alpha,\mu,I)
         , \\
-        \dot{\alpha} = \begin{bmatrix} \pdiff{\alpha}{q_u} & \pdiff{\alpha}{p_u} \end{bmatrix}
-        \begin{bmatrix} \dot{q}_u \\ \dot{p}_u \end{bmatrix}
+        \hat{\mu}(0) = \mu_0
         .
-    \end{cases} \right|_{(q_u,p_u) = T\inv(\mu,\alpha)}
+    \end{cases}
+\end{equation}
+
+In the remainder of this section we will find a first-order
+approximation of \(\hat{\mu}\). 
+We will show the dynamics of this approximation are \(\pi\)-periodic in
+\(\alpha\), so that we only need to study whether \(\hat{\mu}\) increases or
+decreases along a half-orbit \(\alpha \in [0,\pi]\).
+To complete the proof of energy gain for oscillations, we will demonstrate that
+\(\hat{\mu}(\pi) > \hat{\mu}(0)\).
+This will imply that the the radius of orbits increases
+until all orbits within \(D\) eventually leave \(D\) in finite time.
+
+In the spirit of perturbation analysis, we expand our solution of
+(\ref{eqn:muhat-dot}) into
+\[
+    \hat{\mu}(\alpha) = \hat{\mu}_0(\alpha) + I \hat{\mu}_1(\alpha)
+    + R(\alpha,I)
+    ,
 \]
-These dynamics are incredibly complicated and difficult to express, so we will
-write them simply as
+where (\ref{eqn:khalil-perturbation-nominal}) tells us that
+\(\hat{\mu}_0\) is the solution to the nominal system
 \[
 \begin{cases}
-        \dot{\mu} = f_\mu(\alpha,\mu,I)
-        ,\\
-        \dot{\alpha} = f_\alpha(\alpha,\mu,I)
+    \dot{\hat{\mu}}_0 = g(\alpha,\mu,0)
+    , \\
+    \hat{\mu}_0(0) = \mu_0
     .
@@
 \]
-
-To find the nominal system we set \(I = 0\); 
-MATLAB's symbolic toolbox operations tell us that
-\begin{equation}\label{eqn:acrobot-nominal-ma-dynamics}
+Since \(g(\alpha,\mu,0) = 0\) by 
+(\ref{eqn:acrobot-mu-dot-nom})-(\ref{eqn:acrobot-alpha-dot-nom}), we find that
+\(\hat{\mu}_0(\alpha) = \mu_0\) for all \(\alpha\).
+Likewise, (\ref{eqn:khalil-perturbation-firstorder}) tells us that 
+\(\hat{\mu}_1\) is the solution to
+\begin{equation}\label{eqn:acrobot-mu1-dot}
     \begin{cases}
-        \dot{\mu} = 0
+        \dot{\hat{\mu}}_1 = a(\alpha,\mu_0)\hat{\mu}_1 + b(\alpha,\mu_0)
         , \\
-        \dot{alpha} = \sqrt{\frac{6g}{5l}} 
-        \sqrt{\frac{\cos(\mu\cos(\alpha)) - \cos(\mu)}{\mu^2 \sin(\alpha)^2}}
+        \hat{\mu}_1(0) = 0
         .
@@
 \end{equation}
+where
+\begin{align*}
+    a(\alpha,\mu_0) &= \pdiff{g}{\hat{\mu}}(\alpha,\mu_0,0)
+    , \\
+    b(\alpha, \mu_0) &= \pdiff{g}{I}(\alpha, \mu_0, 0)
+    .
+\end{align*}
+These are difficult to compute by hand, so we resort to MATLAB symbolic
+computation to reveal that
+\begin{align*}
+    a(\alpha,\mu_0) &= 0
+    , \\
+    b(\alpha,\mu_0) &= K \hat{b}(\alpha,\mu_0)
+\end{align*}
+where we define
+\begin{align*}
+    K &= \frac{\bar{q}_a \sqrt{30m^2g l^3}}{15}
+    , \\
+    \hat{b}(\alpha,\mu_0) &= \frac{
+        \mu_0 |\sin(\alpha)| \left(
+        5 c_{\mu_0} \cos(\mu_0 c_\alpha) - 8 \cos(\mu_0c_\alpha)^2 + 3
+    \right)
+    }{
+    \sin(\mu_0)\sqrt{\cos(\mu_0c_\alpha) - c_{\mu_0}}
+    }
+    .
+\end{align*}
+Hence, the solution to (\ref{eqn:acrobot-mu1-dot}) is given by
+\[
+    \hat{\mu}_1(\alpha,\mu_0) =
+    K \int \limits_0^\alpha \hat{b}(\sigma,\mu_0)d\sigma
+    .
+\]
 

### 2020-11-06 latex/acrobot.tex
Finished oscillations and started on rotations

@@
 which we need to solve for the explicit solution \(x_0(t,\eta_0)\) on \([t_0,t_1]\).
-One can use \(x_0(t,\eta_0)\) to compute the first-order approximation of
-the nonlinear system 
-\(x(t,\eta_0) = x_0(t,\eta_0) + I x_1(t,\eta_0) + R(t,I)\), where the remainder
-term \(R(t,I)\) satisfies
-\[
-    \lim \limits_{I \to 0} \frac{R(t,I)}{I} = 0
-    ,
-\]
-and \(x_1(t,\eta_0)\) is the solution to
+One can use this to compute the solution \(x_1(t,\eta_0)\) of
 \begin{equation}\label{eqn:khalil-perturbation-firstorder}
@@
         , \\
-        x_1(t_0,\eta_0) = 0. \\
+        x_1(t_0,\eta_0) = 0
+        .
     \end{cases}
 \end{equation}
-
+Finally, one can compute the first-order approximation
+\(x(t,\eta_0) = x_0(t,\eta_0) + I x_1(t,\eta_0) + R(t,I)\), where the remainder
+term \(R(t,I)\) is \(O(I^2)\);
+that is,
+\[
+    \lim \limits_{I \to 0} \frac{R(t,I)}{I} = 0
+    .
+\]
 We now paraphrase Khalil's Theorem 10.1 \cite{khalil_nonlinear} on the
@@
 \subsection*{Perturbation Analysis for Oscillations}
-The nominal pendulum (\ie when \(I = 0\)) oscillates whenever the
+When \(I = 0\), the acrobot becomes a nominal pendulum that oscillates whenever the
 nominal energy (\ref{eqn:acrobot-nominal-E}) is less than \(E(\pi,0)\).
@@
 \end{align}
-To find the nominal system we set \(I = 0\);
-MATLAB's symbolic toolbox evaluates these expressions as
+The nominal system \((\alpha_n,\mu_n)\) comes from setting \(I = 0\).
+MATLAB's symbolic toolbox evaluates the nominal dynamics as
 \begin{align}\label{eqn:acrobot-mu-dot-nom}
-    \dot{\mu} &= 0
+    \dot{\mu}_n &= 0
     , \\
     \label{eqn:acrobot-alpha-dot-nom}
-    \dot{\alpha} &= \sqrt{\frac{6g}{5l}} 
-        \sqrt{\frac{\cos(\mu\cos(\alpha)) - \cos(\mu)}{\mu^2 \sin(\alpha)^2}}
+    \dot{\alpha}_n &= \sqrt{\frac{6g}{5l}} 
+        \sqrt{\frac{\cos(\mu_n\cos(\alpha_n)) - \cos(\mu_n)}
+            {\mu_n^2 \sin(\alpha_n)^2}}
     .
 \end{align}
-One can confirm that \(\dot{\alpha} > 0\) for every \(\mu \in ]0,\pi[\) -- this
+One can confirm that \(\dot{\alpha}_n > 0\) for every \(\mu_n \in ]0,\pi[\) -- this
 is expected, since the nominal system is a single pendulum.
@@
 (\ref{eqn:muhat-dot}) into
-\[
+\begin{equation}\label{eqn:acrobot-muhat-approx}
     \hat{\mu}(\alpha) = \hat{\mu}_0(\alpha) + I \hat{\mu}_1(\alpha)
@@
     ,
-\]
-where (\ref{eqn:khalil-perturbation-nominal}) tells us that
+\end{equation}
+where \(R(\alpha,I)\) is \(O(I^2)\).
+From (\ref{eqn:khalil-perturbation-nominal}) we know that 
 \(\hat{\mu}_0\) is the solution to the nominal system
@@
     \hat{\mu}_0(0) = \mu_0
-    .
+    ,
 \end{cases}
@@
 \(\hat{\mu}_0(\alpha) = \mu_0\) for all \(\alpha\).
+
 Likewise, (\ref{eqn:khalil-perturbation-firstorder}) tells us that 
@@
     b(\alpha,\mu_0) &= K \hat{b}(\alpha,\mu_0)
+    ,
 \end{align*}
@@
 \begin{align*}
-    K &= \frac{\bar{q}_a \sqrt{30m^2g l^3}}{15}
+    K &:= \frac{\bar{q}_a \sqrt{30m^2g l^3}}{15}
     , \\
-    \hat{b}(\alpha,\mu_0) &= \frac{
+    \hat{b}(\alpha,\mu_0) &:= \frac{
         \mu_0 |\sin(\alpha)| \left(
@@
 
+To show that the first-order approximation (\ref{eqn:acrobot-muhat-approx}) is
+gaining energy, we define the Poincar\'{e} section 
+\(P(\mu_0) := \hat{\mu}(\pi,\mu_0)\), which expands into
+\[
+    P(\mu_0) = \mu_0 + I \hat{\mu}_1(\pi,\mu_0) + R(\pi,I)
+    .
+\]
+Let us define
+\[
+    Q(\mu_0) := \int \limits_0^\pi \hat{b}(\sigma,\mu_0)d\sigma
+    ,
+\]
+so that \(\hat{\mu}_1(\pi,\mu_0) = K Q(\mu_0)\).
+Note that \(K\) is a positive constant that takes into account the acrobot's
+mass and length, while \(\hat{b}(\alpha,\mu)\) is adimensional;
+that is, it is the same for every possible choice of acrobot.
+Hence, \(Q(\mu_0)\) is the same for every acrobot.
+
+We numerically compute \(Q(\mu_0)\) for \(\mu_0 \in [10^{-5}, \pi - 10^{-5}]\)
+in Figure \ref{fig:acrobot-Q} and we see that \(Q(\mu_0)\) is strictly positive,
+with an asymptote at \(\mu_0 = \pi\). 
+Simulations with smaller \(\mu_0\) vanish in MATLAB, so we believe that 
+\(Q(\mu_0)\) is in fact positive for all \(\mu_0\).
+
+\begin{figure}
+    \centering
+    % TODO: The plot of Q(mu) from MATLAB
+    %\includegraphics[width=0.5\textwidth]{images/Qmu.png}
+    \caption{The plot of \(Q(\mu_0)\).}
+    \label{fig:acrobot-Q}
+\end{figure}
+
+We conclude that, for all \(m\), \(g\), \(l\), \(\bar{q}_a\), and for small
+enough \(I\), the Poincar\'{e} section \(P(\mu_0)\) is positive, monotonically
+increasing, and \(\lim \limits_{\mu_0 \to \pi} P(\mu_0) = \infty\).
+Thus, there are no closed orbits and the origin must be a repeller.
+Furthermore, for any non-zero initial condition of the acrobot and for small
+enough \(I\), the \((q_u,p_u)\) phase of the acrobot will eventually exit \(D\)
+and the acrobot will begin rotating.
+The acrobot's orbit must leave all compact sets in \(D\) in finite time, which
+means the acrobot is gaining energy on \(D\).
+
+If we have \(I < 0\), it is equivalent to using the constraint 
+\(q_a = -\bar{q}_a \arctan(I p_u)\). 
+By the same analysis, this injects energy on \(D\) in negative-time, so the
+constraint dissipates energy on \(D\).
+
+A hybrid supervisor that flips the sign of \(I\) based on the current
+value of \(\mu\) will stabilize any energy level of oscillation.
+As was the case in the VLP, this supervisor should implement some kind of
+hysteresis to avoid infinite switching.
+% Empty proof environment to put QED square on the right of the page
+\begin{proof}[\unskip\nopunct]
+\end{proof}
+
 \subsection*{Perturbation Analysis for Rotations}
+Recall that the domain of oscillations for the nominal pendulum of the acrobot
+is 
+\[
+    D := \left\{ (q_u,p_u) \in \Gamma | E(q_u,p_u) < E(\pi,0)\right\}
+    .
+\]
+Outside of \(D\), the acrobot is experiencing rotations.
+We wish to find a new set of coordinates \((\alpha,\mu)\) and perform a similar
+analysis.
+During oscillations we showed that the energy of oscillations is determined by
+the intersection with the \(q_u\)-axis; 
+when the acrobot is rotating, the energy is instead determined by the
+intersection on the \(p_u\)-axis.
+Hence, we choose \(\mu\) to be this intersection point.
+Note that \(D\) intersects the \(p_u\) axis when 
+\(\mu^2 = 60m^2 g l^3\), so we must have 
+\(\mu > \sqrt{60 m^2 g l^3}\) if the rotation intersects at \(p_u > 0\) 
+or \(\mu < -\sqrt{60 m^2 g l^3}\) if the rotation intersects at \(p_u < 0\).
+The energy level set associated with \(\mu\) is
+\[
+    \left\{(q_u,p_u) \in \Gamma \mid E(q_u,p_u) = \frac{\mu^2}{10ml^2}\right\}
+    ,
+\]
+which gives the relationship
+\begin{equation}\label{eqn:rotation-pu2}
+    \frac{p_u^2}{10m l^2} + 30mgl(1 - c_u) = \frac{\mu^2}{10 ml^2}
+    .
+\end{equation}
+On this level set, \(q_u\) ranges fully between \(]-pi,\pi]\), so our angle of
+rotation is uniquely parameterized by \(\alpha = q_u\).
+Since \(\mu\) does not change sign along the rotation, we have the smooth
+relationship
+\begin{align}\label{eqn:rotation-Tinv}
+    q_u &= \alpha
+    , \\
+    p_u &= \sign{\mu}\sqrt{\mu^2 - 30 m^2 g l^3\left(1 - c_\alpha \right)}
+    .
+\end{align}
+
+We have thus found a transformation \(T\inv(\alpha,\mu) = (q_u,p_u)\), with
+inverse 
+\begin{align}\label{eqn:rotation-T}
+    \alpha &= q_u
+    , \\
+    \mu &= \sign{p_u}\sqrt{p_u^2 + 30 m^2 g l^3 \left(1 - c_u \right)}
+    .
+\end{align}
+
+From this point on, we do the exact same procedure as with oscillations.
+MATLAB symbolic operations evaluate the nominal system's dynamics as
+\begin{align}\label{eqn:acrobot-rot-mu-dot-nom}
+    \dot{\mu}_n &= 0
+    , \\
+    \label{eqn:acrobot-rot-alpha-dot-nom}
+    \dot{\alpha}_n &=  what
+    .
+\end{align}
+Once again, \(\dot{\alpha}_n > 0\) so for small enough \(I\) we must also have
+\(\dot{\alpha} > 0\).
+We take the time-scaled system \(\hat{\mu}(\alpha,I)\) with dynamics 
+\(\dot{\hat{\mu}}(\alpha) = \frac{\dot{\mu}}{\dot{\alpha}}\) with initial
+\(\mu_0\).
+We expand the solution of this into the same form as
+(\ref{eqn:acrobot-muhat-approx}) where once again we have
+\(\hat{\mu}_0(\alpha) \equiv \mu_0\).
+The dynamics for \(\hat{\mu}_1\) are
+\begin{equation}\label{eqn:acrobot-rot-mu1-dot}
+    \dot{\hat{\mu}}_1 =
+    \sign{\mu_0}\frac{5m^2 g l^3 \bar{q}_a \left(
+        m^2gl^3\left(18s_\alpha^2 + 30c_\alpha(1 - c_\alpha)\right)
+        - c_\alpha\mu_0^2
+    \right)}{
+    |\mu_0|\sqrt{\mu_0^2 - 30m^2gl^3(1 - c_\alpha)}
+    }
+    .
+\end{equation}
 

### 2020-11-08 latex/acrobot.tex
Edited oscillation section to be more clear
Added alpha_dot for rotations, as per results from simulation.

@@
 \]
+Notice that \(\hat{b}(\alpha,\mu_0)\) is both even and 
+\(\pi\)-periodic in \(\alpha\), so that the following properties hold:
+\begin{equation}\label{eqn:bhat-integral-properties}
+    \begin{split}
+    \int \limits_0 ^ {N \pi} \hat{b}(\sigma,\mu_0) d\sigma
+    &= N \int \limits_0^\pi \hat{b}(\sigma,\mu_0) d\sigma \, 
+    \left(\forall N \in \mathbb{Z}_{> 0}\right)
+    , \\
+    \int \limits_{-\pi}^0 \hat{b}(\sigma,\mu_0) d\sigma
+    &= \int \limits_0^\pi \hat{b}(\sigma,\mu_0)d\sigma
+    .
+    \end{split}
+\end{equation}
+Suppose we start the system at \((\alpha,\mu) = (0,\mu_0)\). 
+Then the new pseudo-radius after a half orbit where \(\alpha\) goes from \(0\)
+to \(\pi\) (that is, when the orbit hits the negative \(q_u\)-axis)
+is \(\rho = \hat{\mu}(\pi,\mu_0)\).
+Hence, the change in pseudo-radius is \(\rho - \mu_0 = I\hat{\mu}_1(\pi,\mu_0)\).
+Now look at the full orbit where \(\alpha\) goes from \(0\) to \(2\pi\). 
+This lands on the positive \(q_u\)-axis with pseudo-radius
+\(\tilde{\rho} = \mu_0 + I\hat{\mu}_1(2\pi,\mu_0)\).
+As we know from (\ref{eqn:bhat-integral-properties}), we have
+\(\hat{\mu}_1(2\pi,\mu_0) = 2 \hat{\mu}_1(\pi,\mu_0)\). 
+This means the change in pseudo-radius after one entire oscillation is 
+\(\tilde{\rho} - \mu_0 = 2I\hat{\mu}1(\pi,\mu_0) = 2 (\rho - \mu_0)\).
+Hence, if we can show that \(\hat{\mu}_1(\pi,\cdot)\) is always positive, we
+will guarantee that the pseudo-radius increases after every half oscillation.
+This in turn guarantees \(q_u\) is increasing after every half oscillation, so
+that eventually we will reach a rotation.
 
@@
 Hence, \(Q(\mu_0)\) is the same for every acrobot.
+If \(Q(\mu_0)\) is strictly positive, then \(\hat{\mu}_1(\pi,\mu_0)\) will be
+positive no matter the values of \(m\), \(g\), \(l\), or \(\bar{q}_a\).
 
 We numerically compute \(Q(\mu_0)\) for \(\mu_0 \in [10^{-5}, \pi - 10^{-5}]\)
-in Figure \ref{fig:acrobot-Q} and we see that \(Q(\mu_0)\) is strictly positive,
-with an asymptote at \(\mu_0 = \pi\). 
+in Figure \ref{fig:acrobot-Q}. 
+We see that \(Q(\mu_0)\) is strictly positive and monotonically increasing, with
+an asymptote at \(\mu_0 = \pi\). 
 Simulations with smaller \(\mu_0\) vanish in MATLAB, so we believe that 
@@
     \label{eqn:acrobot-rot-alpha-dot-nom}
-    \dot{\alpha}_n &=  what
+    \dot{\alpha}_n &=  \sign(\mu_0) 
+    \frac{\sqrt{\mu^2 - 30m^2gl^3(1 - c_\alpha)}}{5ml^2}
     .
 \end{align}
-Once again, \(\dot{\alpha}_n > 0\) so for small enough \(I\) we must also have
-\(\dot{\alpha} > 0\).
-We take the time-scaled system \(\hat{\mu}(\alpha,I)\) with dynamics 
-\(\dot{\hat{\mu}}(\alpha) = \frac{\dot{\mu}}{\dot{\alpha}}\) with initial
+As is expected, \(\dot{\alpha}_n\) does not change sign because the acrobot is
+rotating;
+if \(\mu_0 > 0\), the orbit traces a curve going from \(\alpha = -\pi\) to 
+\(\alpha = \pi\), and if \(\mu_0 < 0\) the curve goes in the reverse direction.
+For small enough \(I > 0\) we must also have 
+\(\dot{\alpha} = \sign{\mu_0}\) and can hence use \(\alpha\) as a new time
+variable.
+
+Take the time-scaled system \(\hat{\mu}(\alpha,I)\) with dynamics 
+\(\dot{\hat{\mu}}(\alpha) = \frac{\dot{\mu}}{\dot{\alpha}}\) and initial
 \(\mu_0\).
@@
     \dot{\hat{\mu}}_1 =
-    \sign{\mu_0}\frac{5m^2 g l^3 \bar{q}_a \left(
+    \frac{5m^2 g l^3 \bar{q}_a \left(
         m^2gl^3\left(18s_\alpha^2 + 30c_\alpha(1 - c_\alpha)\right)
@@
 \end{equation}
+This is an even, \(2\pi\)-periodic function in \(\alpha\).
 

### 2020-11-09 latex/vlp.tex
Replaced overbar/underbar with overline/underline for consistency

@@
 converted into a VNHC, and which allows 
-\(l \in [\underbar{l},\overbar{l}]\).
+\(l \in [\underline{l},\overline{l}]\).
 
@@
 Figure \ref{fig:vlp-optimal-controller} shows the graph of \(L^\star(\theta)\), 
-where now the length varies between \([\underbar{l},\overbar{l}]\) rather than 
+where now the length varies between \([\underline{l},\overline{l}]\) rather than 
 \(\{-1,0,1\}\).

### 2020-11-10 latex/acrobot.tex
Finished the first writeup of rotations, added another section for full energy gain

@@
 Furthermore, for any non-zero initial condition of the acrobot and for small
-enough \(I\), the \((q_u,p_u)\) phase of the acrobot will eventually exit \(D\)
-and the acrobot will begin rotating.
-The acrobot's orbit must leave all compact sets in \(D\) in finite time, which
-means the acrobot is gaining energy on \(D\).
+enough \(I\), \((q_u,p_u)\) will eventually exit \(D\).
+This implies the orbit will escape any compact set in \(D\) in finite time. 
+By Definition \ref{defn:energy-gain}, the acrobot is gaining energy on \(D\).
 
@@
 \(\dot{\hat{\mu}}(\alpha) = \frac{\dot{\mu}}{\dot{\alpha}}\) and initial
-\(\mu_0\).
+state \(\mu_0\).
 We expand the solution of this into the same form as
@@
     }
-    .
-\end{equation}
-This is an even, \(2\pi\)-periodic function in \(\alpha\).
+    =: b(\alpha,\mu_0)
+    .
+\end{equation}
+The solution to (\ref{eqn:acrobot-rot-mu1-dot}) is given by
+\[
+    \hat{\mu}_1(\alpha,\mu_0) = \int \limits_0^\alpha b(\sigma,\mu_0)d\sigma
+    .
+\]
+Since one full rotation amounts to \(\alpha\) traversing \(2\pi\)rad in a clockwise
+direction (\ie \(\alpha\) going from \(0\) to \(\sign{\mu_0}2\pi\)), 
+we define a Poincar\'{e} section as the point reached on the \(p_u\) axis after
+every rotation:
+\[
+    P(\mu_0) = \mu_0 + I \hat{\mu}_1(\sign{\mu_0}2\pi, \mu_0)
+    .
+\]
+If \(|P(\mu_0)| > |\mu_0|\), then the rotation has
+driven the orbit further from the origin.
+In other words, \(|p_u|\) has increased across the rotation.
+If this is true for any \(\mu_0\), the acrobot is gaining energy.
+
+The function \(b(\alpha,\mu_0)\) is even in both its inputs, and is 
+\(2\pi\)-periodic in \(\alpha\).
+Hence, we have 
+\begin{align*}
+    \hat{\mu}_1(\sign{\mu_0}2\pi, \mu_0) &=
+    \int \limits_0^{\sign{\mu_0}2\pi} b(\sigma,\mu_0)d\sigma
+    , \\
+     &=\sign{\mu_0}\int\limits_0^{2\pi}b(\sigma,\mu_0)d\sigma
+     , \\
+     &= \sign{\mu_0} \hat{\mu}_1(2\pi,\mu_0)
+     .
+\end{align*}
+Furthermore, \(\hat{\mu}_1(2\pi,\mu_0) = \hat{\mu}_1(2\pi,|\mu_0|)\).
+This implies that
+\[
+    |P(\mu_0)| = \left|\sign{\mu_0}\left(
+        |\mu_0| + I \hat{\mu}_1(2\pi,|\mu_0|)\right)
+    \right|
+    = P(|\mu_0|)
+    .
+\]
+Therefore, if \(P(|\mu_0|) > |\mu_0|\), then the acrobot is gaining energy.
+Showing that \(\hat{\mu}_1(2\pi,\mu_0) > 0\) for \(\mu_0 > \sqrt{60m^2gl^3}\) is
+sufficient for this condition to hold.
+
+Unlike in the case of oscillations, we cannot adimensionalize
+\(b(\alpha,\mu_0)\) and prove that this condition will be true for all acrobots. 
+Let \(U = ]\sqrt{60m^2gl^3},\bar{\mu}[\), and assume that 
+\(\hat{\mu}_1(2\pi,\mu_0) > 0 \) for all \(\mu_0 \in U\).
+If this is true, then by perturbation analysis we can claim that there exists a
+small enough \(I\) such that the Poincar\'{e} section \(P(\mu_0)\) is monotonic
+on \(U\).
+
+Define the set of points in \(\Gamma\) which have pseudo-radius in \(U\) by
+\[
+    W := \left\{(q_u,p_u) \in \Gamma \mid |\mu(q_u,p_u)| \in U \right\}
+    ,
+\]
+as in Figure \ref{fig:acrobot-rot-W}.
+Monotonicity of \(P(\mu_0)\) means there are no closed orbits of 
+\((q_u,p_u)\) on \(W\). 
+Furthermore, \(|p_u|\) is increasing every time the orbit hits the
+\(p_u\)-axis.
+
+\begin{figure}
+    \centering
+    % TODO: Display the set W, which is the two rotation bands at the top and
+    % bottom with outer limit +- \bar{mu}  and inner limit +- sqrt{60m^2gl^3}.
+    % Show also that the inner limit connects to D
+    %\includestandalone[width=0.5\textwidth]{images/acrobot_rot_W}
+    \caption{The set \(W\) of points which have pseudo-radius 
+        \(|\mu| \in ]\sqrt{60m^2gl^3}, \bar{\mu}[\). The inner boundary of \(W\)
+        is \((q_u,p_u) = (\pi,0)\), which is also the boundary of the
+        oscillation domain \(D\).}
+    \label{fig:acrobot-rot-W}
+\end{figure}
+
+We conclude that \(|\mu(q_u,p_u)|\) will eventually reach \(\bar{\mu}\). 
+We cannot claim this fits the definition  of energy injection 
+(since \((0,0) \not \in W\)), but we can still say that \((q_u,p_u)\) will
+eventually depart \(W\) through the outer boundary in finite time.
+The acrobot is thereby gaining energy in the literal sense that its momentum
+is increasing on \(W\).
+
+If we set \(I < 0\), we have \(P(|\mu_0|) < |\mu_0|\).
+Hence, \(|p_u|\) is decreasing every time the orbit hits the \(p_u\)-axis, so
+the acrobot will eventually depart \(W\) through the inner boundary. 
+The acrobot is losing momentum, and therefore it is losing energy on \(W\).
+
+\subsection*{Connecting Oscillations and Rotations}
+To complete this proof, we will prove the acrobot gains energy on the set 
+\[
+    E_\bar{\mu} := 
+    \left\{ (q_u,p_u) \in \Gamma \mid E(q_u,p_u) < E(0,\bar{\mu}) \right\}
+    .
+\]
+% TODO: W and D's boundary is two elements: the point (pi,0) and the curve of the
+% homoclinic orbit connecting (pi,0) to (-pi,0). We need to show that almost
+% every initial condition starting in \(D\) will NOT converge to (pi,0), and
+% therefore exit D and enter W.
 

### 2020-11-11 latex/vlp.tex
Incorporating Manfredi's changes to VLP proof

@@
 Using this configuration, we will compute the Hamiltonian dynamics of the system.
-The cartesian position of the mass at the tip of the pendulum
+The Cartesian position of the mass at the tip of the pendulum
 is given by \(x = (l\sin(q),-l\cos(q))\), while its velocity is
@@
 The control input is a force \(\tau \in \mathbb{R}\) affecting the dynamics of
-\(p_l\), acting colinearly with the rod.
+\(p_l\), acting collinearly with the rod.
 We assume the force does not affect the dynamics of \(p\) in any way -
@@
 \begin{defn}\label{defn:energy-injection}
-   A regular VNHC \(h(\mathbf{q},\mathbf{p}) = 0\) 
-   \textit{injects (dissipates) energy} if the constrained dynamics gain (lose)
-   energy everywhere on the constraint manifold, except possibly on a set of
-   measure zero.
+   A regular VNHC \(h(\mathbf{q},\mathbf{p}) = 0\) \textit{injects (dissipates)
+     energy} if the constrained dynamics in~\eqref{eqn:vlp-hamiltonian} gain
+   (lose) energy everywhere on the constraint manifold, except possibly on a set
+   of measure zero.
 \end{defn}
@@
    \end{equation}
-   with the property that the inequality is strict except at the
-   coordinate axes.
-   If the inequality is flipped, the VNHC dissipates energy. 
+   with the property that the inequality is strict except at $\theta \in
+   \{0,\pi/2,\pi,3 \pi/2\}$.  If instead 
+\[
+ \left(l_\text{avg} - L(\theta)\right)sin(2\theta) \leq 0 
+      \text{ }\forall \theta \in \mathbb{S}^1
+      ,
+\]
+the VNHC dissipates energy.
 \end{thm}
 \begin{proof}
-   The constraint manifold for \(h\) is the cylinder \(\Gamma = \mathbb{S}^1 \times \R \). 
-   In negative-time, the dynamics on \(\Gamma\) are
+   We will first show that, under the given assumptions, the origin of the
+   constrained dynamics~\eqref{eqn:vlp-hamiltonian} is a repeller, and for that
+   we consider the negative time system obtained
+   from~\eqref{eqn:vlp-hamiltonian} by reversing the sign of the vector field:
 \begin{equation}\label{eqn:negative-time-vlp}
-      f(q,p) = 
-      -\begin{bmatrix}
-         \dot{q} \\ \dot{p}
-      \end{bmatrix} = \begin{bmatrix}
-         -\frac{p}{mL(\theta)^2} \\
-         mgL(\theta)\sin(q)
-      \end{bmatrix}
-      .
+   \begin{cases}
+      \dot{q} = -\frac{p}{m L^2} \\
+      \dot{p} = mgL\sin(q)
+      . \\ 
+   \end{cases}
 \end{equation}
-   Because \(q \in \mathbb{S}^1\), the only equilibria of these dynamics are 
-   at \((q,p) = (0,0)\) and \((q,p) = (\pi,0)\).
-   To prove the origin of the positive-time system (\ref{eqn:vlp-hamiltonian})
-   is a repeller, we will show the origin of the negative-time system 
-   (\ref{eqn:negative-time-vlp}) is almost everywhere globally asymptotically
-   stable.
+The state space of~\eqref{eqn:negative-time-vlp} is \(\mathbb{S}^1
+\times \R\), and the system has two equilibria, \((q,p)=(0,0)\) and
+\((q,p)=(\pi,0)\). We need to show that the equilibrium
+\((q,p)=(0,0)\) is asymptotically stable
+for~\eqref{eqn:negative-time-vlp}.
 
-   Choose, as a candidate Lyapunov function, the energy for the
-   average-length pendulum with zero-potential at the bottom of the swing:
+Consider the Lyapunov function candidate
    \[
@@
            + m g l_\text{avg} (1-\cos(q))
-      .
+      ,
    \]
-   This is positive definite at \((0,0)\) and has compact sublevel
-   sets on \(\Gamma\). 
-   The derivative of \(E_\text{avg}\) along (\ref{eqn:vlp-hamiltonian}) is
-   \begin{equation}\label{eqn:eavg-dot-pos}
-      \dot{E}_\text{avg}^+ = -\frac{g\sin(q)p \left(L(\theta)^3 - l_\text{avg}^3\right)}
-              {l_\text{avg}^2L(\theta)^2}
-      ,
-   \end{equation}
-   and the derivative of \(E_\text{avg}\) along (\ref{eqn:negative-time-vlp}) is
+which corresponds to the energy of VLP when the pendulum length is fixed at
+\(l_\text{avg}\).  The function \(E_\text{avg}\) is positive definite at
+\((0,0)\), and has compact sublevel sets on \(\mathbb{S}^1 \times \R\).  The
+derivative of \(E_\text{avg}\) along~\eqref{eqn:negative-time-vlp} is
+   %% \begin{equation}\label{eqn:eavg-dot-pos}
+   %%    \dot{E}_\text{avg}^+ = -\frac{g\sin(q)p \left(L(\theta)^3 - l_\text{avg}^3\right)}
+   %%            {l_\text{avg}^2L(\theta)^2}
+   %%    ,
+   %% \end{equation}
+   %% and the derivative of \(E_\text{avg}\) along (\ref{eqn:negative-time-vlp}) is
    \begin{equation}\label{eqn:eavg-dot-neg}
-      \dot{E}_\text{avg}^- = - \dot{E}_\text{avg}^+
+      \dot{E}_\text{avg} = \frac{g\sin(q)p \left(L(\theta)^3 -
+        l_\text{avg}^3\right)} {l_\text{avg}^2L(\theta)^2}
       .
    \end{equation}
-   The sign of \(\sin(q)p\) depends only on the quadrant in the
-   \((q,p)\)-plane. In fact, it is easy to show that
+It is easy to show that
    \[
@@
        \sign{L(\theta)^3 - l_\text{avg}^3} = \sign{L(\theta) - l_\text{avg}}
+       ,
+   \]
+and therefore,
+   \begin{align}\label{eqn:vlp-proof-neg-inv}
+    \sign{\dot{E}_\text{avg}} &= \sign{\sin(q)p \left(L(\theta)^3 - l_\text{avg}^3\right)}
+          \nonumber \\
+          &= \sign{\sin(2\theta) \left(L(\theta) - l_\text{avg}\right)}
+          \nonumber \\
+          &= -\sign{(l_\text{avg} - L(\theta))\sin(2\theta)} 
+          \nonumber \\
+          &\leq 0 \text{ (by assumption)}
+      .
+   \end{align}
+We have thus shown that \(\dot{E}_\text{avg} \leq 0\), and therefore
+the equilibrium \((q,p)=(0,0)\) is stable for~\eqref{eqn:negative-time-vlp}. We
+now apply the Krasovskii-LaSalle invariance principle, and consider the largest
+subset of \( Z = \left\{(q,p) \in \mathbb{S}^1 \times \R \mid
+\dot{E}_\text{avg}(q,p) = 0 \right\} \). We see from~\eqref{eqn:eavg-dot-neg}
+that \(\dot{E}_\text{avg}(q,p)\) is zero when $p=0$, \(\sin(q)=0\), or
+$L(\theta) = l_\text{avg}$. This latter condition by assumption is met when
+\(\theta \in \{0,\pi/2,\pi, 3\pi/2\}\), or equivalently when either $q=0$ or
+$p=0$. All in all, we have that
+   \[
+      Z= \left\{(q,p) \in \mathbb{S}^1 \times \R \mid q =0\right\} \cup \left\{(q,p) \in \mathbb{S}^1 \times \R \mid p =0\right\} 
       .
    \]
-   The positive-time derivative (\ref{eqn:eavg-dot-pos}) is nonnegative since
-   \begin{align}\label{eqn:vlp-proof-neg-inv}
-    \sign{\dot{E}_\text{avg}^+} &= \sign{-\sin(q)p \left(L(\theta)^3 - l_\text{avg}^3\right)}
-          \nonumber \\
-          &= \sign{-\sin(2\theta) \left(L(\theta) - l_\text{avg}\right)}
-          \nonumber \\
-          &= \sign{(l_\text{avg} - L(\theta))\sin(2\theta)} 
-          \nonumber \\
-          &\geq 0 \text{ (by assumption)}
-      .
-   \end{align}
-   Using (\ref{eqn:eavg-dot-neg}) and (\ref{eqn:vlp-proof-neg-inv}) 
-   we can determine that \(\sign{\dot{E}_\text{avg}^-} \leq 0\). 
-   Thus, \(E_\text{avg}\) is a Lyapunov function for
-   (\ref{eqn:negative-time-vlp}), which means the negative-time system is stable
-   \cite{lyapunov}.
-
-   Let us briefly analyze the upright equilibrium.
-   Label the set of initial conditions converging to \((\pi, 0)\) in
-   positive time by
+It is easily seen that the largest invariant subset of \(Z\) is the union of the two equilibria
 \[
-      \Pi_0^+ := \left\{ (q_0,p_0) \in \Gamma \mid \lim_{t \to \infty} (q(t),p(t)) = (\pi, 0) \right\}
-      ,
-   \]
-   and the set of initial conditions converging to \((\pi,0)\)
-   in negative time by \(\Pi_0^-\).
-   Take now the sublevel set
-   \[
-      E_{(\pi,0)} := \left\{ (q,p) \in \Gamma \mid E_\text{avg}(q,p) \leq
-         E_\text{avg}(\pi,0) \right\}
+\Omega =  \{(0,0)\}\cup \{(\pi,0)\}
 .
 \]  
-   From (\ref{eqn:vlp-proof-neg-inv}), all sublevel sets of \(E_\text{avg}\) are
-   negatively invariant in positive-time, which means
-   complements of sublevel sets are positively invariant.
-   Hence, the system cannot possibly converge to \((\pi,0)\) if it is
-   initialized on 
-   \(\left(E_{(\pi,0)}\right)^C = \Gamma \backslash E_{(\pi,0)}\).
-   This implies that \(\Pi_0^+ \subset E_{(\pi,0)}\).
-   Similarly, the interior of \(E_{(\pi,0)}\) is positively
-   invariant for (\ref{eqn:negative-time-vlp}), 
-   which implies \(\Pi_0^- \subset \left(E_{(\pi,0)}\right)^C \cup (\pi,0)\).
+Since all sublevel sets of \(E_\text{avg}\) are compact, the
+Krasovskii-LaSalle invariance principle implies that the set
+\(\Omega\) is globally attractive. Since the two equilibria are
+isolated, and since \((q,p)=(0,0)\) is stable, we deduce that the
+equilibrium \((q,p)=(0,0)\) is asymptotically stable for the negative
+time system~\eqref{eqn:negative-time-vlp}, and thus it is a repeller
+for the constrained dynamics in~\eqref{eqn:vlp-hamiltonian}.
 
-   The linearization of (\ref{eqn:vlp-hamiltonian}) at \((\pi, 0)\) is given by
+We now prove that almost all solutions of the reduced
+dynamics~\eqref{eqn:vlp-hamiltonian} escape compact sets in finite
+time, i.e., that for each compact subset \(K\) of \(\mathbb{S}^1
+\times \R\), and for almost every condition \( (q_0,p_0) \in K \),
+there exists \( T > 0\) such that \( (q(t),p(t)) \not \in K \) for all
+\( t > T \), where \( (q(t),p(t)) \) denotes the solution
+of~\eqref{eqn:vlp-hamiltonian} with \( (q(0),p(0)) = (q_0,p_0)\). In
+what follows, we denote by \( x = (q,p) \) the state of the reduced
+dynamics~\eqref{eqn:vlp-hamiltonian}. Also, we denote by \(\Pi^-\) the
+stable manifold of the equilibrium \((\pi,0)\) for
+system~\eqref{eqn:vlp-hamiltonian}, defined as
+ \[
+      \Pi^- := \left\{ x(0) \in \mathbb{S}^1 \times \R \mid
+      \lim_{t \to \infty} x(t) = (\pi, 0) \right\}.
+   \]
+The linearization of system~\eqref{eqn:vlp-hamiltonian} at the
+equilibrium \((\pi,0)\) has a system matrix given by
   \begin{equation}\label{eqn:linearized-vlp}
       \begin{bmatrix}
-         0 & \frac{1}{mL(0)^2} \\
-         m g L(0) & 0 \\
+         0 & -\frac{1}{mL(0)^2} \\
+         -m g L(0) & 0 \\
       \end{bmatrix}
@@
    \end{equation}
-   which has eigenvalues at \(\lambda = \pm \sqrt{\frac{g}{L(0)}}\).
-   Linearizing (\ref{eqn:negative-time-vlp}) at this point yields the same eigenvalues.
-   Since one eigenvalue of (\ref{eqn:linearized-vlp})
-   is positive and one is negative, the stable manifold theorem tells
-   us that both \(\Pi_0^+\) and \(\Pi_0^-\) are one-dimensional
-   \cite{stable_manifold}.
-   This implies the origin cannot be globally asymptotically stable for
-   (\ref{eqn:negative-time-vlp}); however, we can still show it is asymptotically stable
-   for almost every initial condition because \(\Pi_0^+\) and \(\Pi_0^-\) are of
-   measure zero within \(\Gamma\).
+whose spectrum is \(\{\pm \sqrt{{g}/{L(0)}}\}\).  Since one eigenvalue
+of (\ref{eqn:linearized-vlp}) is positive and one is negative, the
+stable manifold theorem implies that the stable manifold \( \Pi^- \)
+is a one-dimensional immersed submanifold of \(\mathbb{S}^1 \times
+\R\), and thus a set of zero measure.
 
-   We cannot linearize (\ref{eqn:vlp-hamiltonian}) at \((0,0)\) to prove 
-   the origin is a repeller, because \(\theta(0,0)\) is not well-defined. 
-   Instead, let \(D := \Gamma \backslash \Pi_0^-\).
-   We will prove that (\ref{eqn:negative-time-vlp}) asymptotically converges to
-   \((0,0)\) if it is initialized on \(D\).
-   First, define the zero-set
+Let \( K \subset \mathbb{S}^1 \times \R\) be an arbitrary compact set
+and let \( x(0) \in K \backslash \left(\Pi^- \cup
+\left\{(0,0)\right\}\right)\). Since \( x(0) \not \in \Pi^- \), we have that
+\begin{equation}\label{eqn:vlp-x-does-not-converge}
+x(t) \not \to_{t \to \infty} (\pi,0).
+\end{equation}
+Suppose, by way of contradiction, that for each \(T
+> 0\) there exists \(t' > T\) such that \(x(t') \in K\).  Letting
    \[
-      Z = \left\{(q,p) \in \Gamma \mid 
-      \dot{E}_\text{avg}(q,p) = 0 \right\}
-      .
-   \]
-   It is easy to show that \(Z\) is the union of the coordinate axes
-   \[
-      Z= \mathbb{S}^1 \times \left\{0\right\} \cup 
-         \left\{0\right\} \times \R
-      .
-   \]
-   Krasovskii-LaSalle tells us that solutions to (\ref{eqn:negative-time-vlp})
-   converge to the largest invariant set in \(Z \cap D\)
-   \cite{krasovskii_lasalle}.
-   Since \(f(q,p) = (0,0)\) on \(D\) only when \((q,p) = (0,0)\),
-   the origin is asymptotically stable on \(D\) for the negative-time system.
-   In particular, the origin is asymptotically stable for every initial
-   condition in the interior of \(E_{(\pi,0)}\).
-   Hence, the positive-time system is diverging away from the origin 
-   in the \((q,p)\) plane, which means the origin is a repeller.
-
-   We now prove the system escapes compact sets in finite time.
-   Let \(x(t) = (q(t),p(t))\) be the orbit for the positive-time system.
-   Let \(K \subset \Gamma\) be compact and assume 
-   \(x(0) \in K\backslash \left(\Pi_0^+ \cup \left\{(0,0)\right\}\right)\).
-   For a proof by contradiction, suppose that for all \(T > 0\) there exists
-   \(t' > T\) where \(x(t') \in K\).
-   Let 
-   \[
-      k := \max\limits_{(q,p) \in K} E_\text{avg}(q,p)
+      k := \max\limits_{x \in K} E_\text{avg}(x)
       ,
    \] 
-   be the largest energy value attained in \(K\), and let \(E_k\) be its
-   sublevel set.
+and \(E_k =\{ x \mid E_\text{avg}(x) \leq k\} \), we have that \( K
+\subset E_k \), and thus \(x(t') \in E_k\). We have shown earlier that
+the function \( E_\text{avg} \) is nonincreasing along solutions
+of~\eqref{eqn:negative-time-vlp}, which implies that \(
+E_\text{avg}(x(t)) \) is nondecreasing for solutions of the reduced
+dynamics~\eqref{eqn:vlp-hamiltonian}, and therefore the half orbit
+\(x([0,t'])\) is contained fully within \(E_k\).  In particular,
+\(x(T) \in E_k\). Since this is true for each \( T >0 \), we deduce
+that \(x(t) \in E_k\) for all \(t \in [0,\infty[\).  Since \(E_k\) is
+    compact, \(x(t)\) has a positive limit set in \(E_k\) which, by
+    the Poincar\'{e}-Bendixson theorem, is either an equilibrium or a
+    closed orbit \cite{poincare_bendixson}.  We will show this
+    positive limit set must be an equilibrium and thereby reach a
+    contradiction.
 
-   Since \(E_k\) contains \(K\) and \(x(t') \in K\), it follows that 
-   \(x(t') \in E_k\). 
-   Furthermore, negative invariance of \(E_k\) means the half orbit
-   \(x([0,t'])\) is contained fully within \(E_k\).
-   In particular, \(x(T) \in E_k\).
-   However, \(T\) was chosen arbitrarily, which implies \(x(t) \in E_k\) for all
-   \(t \in [0,\infty[\).
-   Since \(E_k\) is compact, \(x(t)\) has a positive limit set in
-   \(E_k\) which, by Poincar\'{e}-Bendixson, is either an equilibrium or a
-   closed orbit \cite{poincare_bendixson}.
-   We will show this positive limit set must be an equilibrium and thereby
-   reach our contradiction.
+In order to rule out closed orbits, suppose that the reduced
+dynamics~\eqref{eqn:vlp-hamiltonian} have a periodic solution \(
+z(t)\). Since \( E_\text{avg}(z(t)) \) is nondecreasing and \(z(t)\)
+is periodic, it must be that \( E_\text{avg}(z(t)) \) is constant, or
+\(\dot E_\text{avg}(z(t)) \equiv 0\). We have shown earlier
+that\footnote{We have shown it for the negative time
+  system~\eqref{eqn:negative-time-vlp}, but changing the sign of the
+  vector field does not change the set where \(\dot E_\text{avg}
+  =0\) nor the invariant sets.}
+\[
+\{x \mid \dot E_\text{avg}(x)=0\} = \left\{(q,p) \mid q =0\right\}
+\cup \left\{(q,p) \mid p =0\right\},
+\]
+and the only invariant subset of this set is the union of two disjoint
+equilibria. Since the orbit \(z(\R)\) is an invariant set which is
+connected, \(z(\R)\) must be an equilibrium and cannot be a nontrivial
+closed orbit.
 
-   Suppose the system had a closed orbit. 
-   The orbit would necessarily be contained in a
-   level set of \(E_\text{avg}\), since this is a non-decreasing function and
-   closed orbits are periodic.
-   Thus, \(E_\text{avg}\) is constant along the orbit, which means the
-   orbit is an invariant subset of the set \(Z\) consisting of the two axes.
-   However, the only invariant subsets of \(Z\) are equilibria, so 
-   the system cannot contain closed orbits.
+Returning to the Poincar\`e-Bendixson theorem, the positive limit set
+of \( x(t)\) must be an equilibrium and,
+by~\eqref{eqn:vlp-x-does-not-converge}, this equilibrium must be \(
+(0,0)\). Since we have chosen \(x(0) \neq (0,0)\), and since \( (q,p) = (0,0) \) is a
+repeller, \( x(t) \) cannot converge to \( (0,0)\), which gives a
+contradiction.
 
-   The lack of closed orbits means the positive limit set of \(x(t)\) must 
-   be one of the equilibria. 
-   The origin is a repeller and \(x(0) \neq (0,0)\), which means \(x(t)\) is
-   converging to \((\pi,0)\).
-   This is a contradiction, since we assumed \(x(0) \not \in \Pi_0^+\).
-
-   We conclude that the VNHC is injecting energy into the VLP.
+We conclude that the VNHC injects energy into the VLP.
    By flipping the inequality of (\ref{eqn:vlp-energy-gain-condition})

### 2020-11-11 latex/acrobot.tex
Edits to Manfredi's improvements of VLP proof
Changing regular ref to eqref for equations

@@
 Reducing
-(\ref{eqn:general-acrobot-inertia}-\ref{eqn:general-acrobot-potential})
+~\eqref{eqn:general-acrobot-inertia}-\ref{eqn:general-acrobot-potential}
 yields the simplified inertia matrix \(D_s\) and potential function \(P_s\),
@@
 A sufficient condition for regularity is when \(f\) is a function solely of \(p_u\),
-because then \(\partial_{q_u}f = 0\) and (\ref{eqn:regularity-matrix-acrobot})
+because then \(\partial_{q_u}f = 0\) and ~\eqref{eqn:regularity-matrix-acrobot}
 is strictly positive for all values of \(q_a\). 
@@
 Let us now compute the constrained dynamics under
-(\ref{eqn:acrobot-constraint}).
+~\eqref{eqn:acrobot-constraint}.
 Note that \(dh_q = \begin{bmatrix}0 & 1\end{bmatrix}\), while
@@
 \]
-Inserting these to (\ref{eqn:g-qupu}), we get the solution for \(p_a\) on the
+Inserting these to ~\eqref{eqn:g-qupu}, we get the solution for \(p_a\) on the
 constraint manifold:
@@
 Hence, the constrained dynamics for the acrobot under
-(\ref{eqn:acrobot-constraint}) are
+~\eqref{eqn:acrobot-constraint} are
 \begin{align}\label{eqn:acrobot-constrained-dynamics}
@@
 \begin{thm}\label{thm:acrobot-energy-stabilization}
-    Consider the simple acrobot (\ref{eqn:acrobot-hamiltonian}) constrained by
-    (\ref{eqn:acrobot-constraint}) with \(\bar{q}_a \in ]0,1]\).
+    Consider the simple acrobot ~\eqref{eqn:acrobot-hamiltonian} constrained by
+    ~\eqref{eqn:acrobot-constraint} with \(\bar{q}_a \in ]0,1]\).
     There exists a small enough\footnotemark \(I^\star \in \R_{>0}\)
@@
     Suppose \(f : [t_0,t_1] \times D \times [-I_0,I_0] \rightarrow \R^2\) is
-    \(C^1\), and that the nominal problem (\ref{eqn:khalil-perturbation-nominal}) has a
+    \(C^1\), and that the nominal problem ~\eqref{eqn:khalil-perturbation-nominal} has a
     unique solution \(x_0(t,\eta_0) \in D\) on \([t_0,t_1]\).
     Then there exists \(I^\star > 0\) such that for all \(|I| < I^\star\), the
-    solution \(x(t,I,\eta_0)\) to (\ref{eqn:khalil-setup}) satisfies
+    solution \(x(t,I,\eta_0)\) to ~\eqref{eqn:khalil-setup} satisfies
     \[
@@
 When \(I = 0\), the acrobot becomes a nominal pendulum that oscillates whenever the
-nominal energy (\ref{eqn:acrobot-nominal-E}) is less than \(E(\pi,0)\).
+nominal energy ~\eqref{eqn:acrobot-nominal-E} is less than \(E(\pi,0)\).
 Hence, the domain of oscillations on \(\Gamma\) is given by
@@
 pseudo-angle.
-Substituting this into (\ref{eqn:oscillation-pu2}), we get
+Substituting this into ~\eqref{eqn:oscillation-pu2}, we get
 \[
@@
 We need the inverse of this map to get our diffeomorphism \(T(q_u,p_u)\).
-Notice from (\ref{eqn:oscillation-pu2}) that
+Notice from ~\eqref{eqn:oscillation-pu2} that
 \[
@@
 \]
-Notice from (\ref{eqn:oscillation-pu}) that 
+Notice from ~\eqref{eqn:oscillation-pu} that 
 \(\sign{\sin(\alpha)} = -\sign{p_u}\).
@@
 with \((\dot{q}_u,\dot{p}_u)\) the given by 
- (\ref{eqn:acrobot-constrained-dynamics}).
+ ~\eqref{eqn:acrobot-constrained-dynamics}.
 
@@
 is expected, since the nominal system is a single pendulum.
-By continuity of (\ref{eqn:acrobot-mu-dot}) with respect to \(I\), 
+By continuity of ~\eqref{eqn:acrobot-mu-dot} with respect to \(I\), 
 \(\dot{\alpha}\) remains positive for \(I\) small enough on the domain \(D\).
@@
 In the spirit of perturbation analysis, we expand our solution of
-(\ref{eqn:muhat-dot}) into
+~\eqref{eqn:muhat-dot} into
 \begin{equation}\label{eqn:acrobot-muhat-approx}
@@
 where \(R(\alpha,I)\) is \(O(I^2)\).
-From (\ref{eqn:khalil-perturbation-nominal}) we know that 
+From ~\eqref{eqn:khalil-perturbation-nominal} we know that 
 \(\hat{\mu}_0\) is the solution to the nominal system
@@
 Since \(g(\alpha,\mu,0) = 0\) by 
-(\ref{eqn:acrobot-mu-dot-nom})-(\ref{eqn:acrobot-alpha-dot-nom}), we find that
+~\eqref{eqn:acrobot-mu-dot-nom})-(\ref{eqn:acrobot-alpha-dot-nom}, we find that
 \(\hat{\mu}_0(\alpha) = \mu_0\) for all \(\alpha\).
 
-Likewise, (\ref{eqn:khalil-perturbation-firstorder}) tells us that 
+Likewise, ~\eqref{eqn:khalil-perturbation-firstorder} tells us that 
 \(\hat{\mu}_1\) is the solution to
@@
 \end{align*}
-Hence, the solution to (\ref{eqn:acrobot-mu1-dot}) is given by
+Hence, the solution to ~\eqref{eqn:acrobot-mu1-dot} is given by
 \[
@@
 \(\tilde{\rho} = \mu_0 + I\hat{\mu}_1(2\pi,\mu_0)\).
-As we know from (\ref{eqn:bhat-integral-properties}), we have
+As we know from ~\eqref{eqn:bhat-integral-properties}, we have
 \(\hat{\mu}_1(2\pi,\mu_0) = 2 \hat{\mu}_1(\pi,\mu_0)\). 
@@
 
-To show that the first-order approximation (\ref{eqn:acrobot-muhat-approx}) is
+To show that the first-order approximation ~\eqref{eqn:acrobot-muhat-approx} is
 gaining energy, we define the Poincar\'{e} section
@@
 We expand the solution of this into the same form as
-(\ref{eqn:acrobot-muhat-approx}) where once again we have
+~\eqref{eqn:acrobot-muhat-approx} where once again we have
 \(\hat{\mu}_0(\alpha) \equiv \mu_0\).
@@
 \end{equation}
-The solution to (\ref{eqn:acrobot-rot-mu1-dot}) is given by
+The solution to ~\eqref{eqn:acrobot-rot-mu1-dot} is given by
 \[
@@
 We cannot claim this fits the definition  of energy injection 
-(since \((0,0) \not \in W\)), but we can still say that \((q_u,p_u)\) will
+(since \((0,0) \notin W\)), but we can still say that \((q_u,p_u)\) will
 eventually depart \(W\) through the outer boundary in finite time.
@@
 \[
-    E_\bar{\mu} := 
+    E_{\bar{\mu}} := 
     \left\{ (q_u,p_u) \in \Gamma \mid E(q_u,p_u) < E(0,\bar{\mu}) \right\}

### 2020-11-11 latex/vlp.tex
Edits to Manfredi's improvements of VLP proof
Changing regular ref to eqref for equations

@@
 \(M(\mathbf{q}) := D(\mathbf{q})\), \(V(\mathbf{q}) := P(\mathbf{q})\),
-we get the Hamiltonian (\ref{eqn:simply-actuated-hamiltonian}) whose
-dynamics (\ref{eqn:simply-actuated-dynamics}) resolve to  
+we get the Hamiltonian ~\eqref{eqn:simply-actuated-hamiltonian} whose
+dynamics ~\eqref{eqn:simply-actuated-dynamics} resolve to  
 \begin{align}\label{eqn:vlp-hamiltonian-with-pl}
@@
 The total mechanical energy of the system restricted to the constraint manifold
-is given by (\ref{eqn:vlp-energy}), where \(L(q,p)\) and \(\dot{L}(q,p)\) are
+is given by ~\eqref{eqn:vlp-energy}, where \(L(q,p)\) and \(\dot{L}(q,p)\) are
 known.
@@
 Abusing notation, we denote this by \(L^\star(\theta)\), which is defined in
-(\ref{eqn:vlp-optimal-controller}).
+~\eqref{eqn:vlp-optimal-controller}.
 Figure \ref{fig:vlp-optimal-controller} shows the graph of \(L^\star(\theta)\), 
@@
 \(L_{\frac{\pi}{2}}(\theta)\), which can be simplified from
-(\ref{eqn:vlp-T-controller}) into
+~\eqref{eqn:vlp-T-controller} into
 \begin{equation}\label{eqn:vlp-smoothed-controller}
@@
 \]
-
-Before proving the VLP will gain energy under this VNHC, 
-we must rigorously define what that means.
-If any initial condition of the constrained dynamics converges to the point
-\((q,p) = (0,0)\), the momentum must be decreasing over time, so the VLP will
-clearly be losing energy.
-A notion of energy injection must therefore require that the origin repels all
-solutions.
-We need to avoid closed orbits or limit cycles in the \((q,p)\)-plane,
-so that the system can always reach higher speeds. 
-Finally, we want the VLP's momentum to increase on average, regardless of
-its initial value.
-This intuition is captured mathematically by the following definition.
-
-\begin{defn}\label{defn:energy-gain}
-   Let \(D \subset \R^2\) be open with \(0 \in D\).
-   Let \(f : D \rightarrow \R^2\) be locally Lipschitz with \(f(0) = 0\).
-   The system described by \(\dot{x} = f(x)\) 
-   \textit{gains energy on \(D\)} if 
-   \begin{enumerate}
-      \item The origin is a repeller 
-         (or equivalently, the origin is asymptotically stable in negative time).
-      \item For all compact sets \(K \subset D\) and for almost every initial
-         condition \(x(0) \in K\), there exists \(T > 0\) such
-         that \(x(t) \not \in K \, (\forall t > T)\).
-   \end{enumerate}
-   The system \textit{loses energy on \(D\)} if it gains energy in
-   negative-time.
-\end{defn}
-
-The second property of Definition \ref{defn:energy-gain} allows the system
-to have unstable equilibria, but not limit cycles nor closed orbits.
-The next definition ties this notion of energy gain to VNHCs.
-
-\begin{defn}\label{defn:energy-injection}
-   A regular VNHC \(h(\mathbf{q},\mathbf{p}) = 0\) \textit{injects (dissipates)
-     energy} if the constrained dynamics in~\eqref{eqn:vlp-hamiltonian} gain
-   (lose) energy everywhere on the constraint manifold, except possibly on a set
-   of measure zero.
-\end{defn}
 
@@
    \end{equation}
-   with the property that the inequality is strict except at $\theta \in
-   \{0,\pi/2,\pi,3 \pi/2\}$.  If instead 
+   with the property that the inequality is strict except at 
+   \(\theta \in \{0, \frac{\pi}{2}, \frac{\pi}, \frac{3\pi}{2}\}\).  
+   If instead 
    \[
@@
 \end{equation}
-The state space of~\eqref{eqn:negative-time-vlp} is \(\mathbb{S}^1
-\times \R\), and the system has two equilibria, \((q,p)=(0,0)\) and
-\((q,p)=(\pi,0)\). We need to show that the equilibrium
-\((q,p)=(0,0)\) is asymptotically stable
-for~\eqref{eqn:negative-time-vlp}.
+The state space of~\eqref{eqn:negative-time-vlp} is 
+\(\mathbb{S}^1 \times \R\), and the system has two equilibria,
+\((q,p) = (0,0)\) and \((q,p) = (\pi,0)\). 
+We need to show that the equilibrium \((q,p) = (0,0)\) is
+asymptotically stable for~\eqref{eqn:negative-time-vlp}.
 
@@
 which corresponds to the energy of VLP when the pendulum length is fixed at
-\(l_\text{avg}\).  The function \(E_\text{avg}\) is positive definite at
-\((0,0)\), and has compact sublevel sets on \(\mathbb{S}^1 \times \R\).  The
-derivative of \(E_\text{avg}\) along~\eqref{eqn:negative-time-vlp} is
-   %% \begin{equation}\label{eqn:eavg-dot-pos}
-   %%    \dot{E}_\text{avg}^+ = -\frac{g\sin(q)p \left(L(\theta)^3 - l_\text{avg}^3\right)}
-   %%            {l_\text{avg}^2L(\theta)^2}
-   %%    ,
-   %% \end{equation}
-   %% and the derivative of \(E_\text{avg}\) along (\ref{eqn:negative-time-vlp}) is
+\(l_\text{avg}\). 
+The function \(E_\text{avg}\) is positive definite at \((0,0)\), and has compact
+sublevel sets on \(\mathbb{S}^1 \times \R\).  
+The derivative of \(E_\text{avg}\) along~\eqref{eqn:negative-time-vlp} is
 \begin{equation}\label{eqn:eavg-dot-neg}
@@
 \begin{align}\label{eqn:vlp-proof-neg-inv}
-    \sign{\dot{E}_\text{avg}} &= \sign{\sin(q)p \left(L(\theta)^3 - l_\text{avg}^3\right)}
+    \sign{\dot{E}_\text{avg}} &= 
+       \sign{\sin(q)p \left(L(\theta)^3 - l_\text{avg}^3\right)}
     \nonumber \\
@@
 We have thus shown that \(\dot{E}_\text{avg} \leq 0\), and therefore
-the equilibrium \((q,p)=(0,0)\) is stable for~\eqref{eqn:negative-time-vlp}. We
-now apply the Krasovskii-LaSalle invariance principle, and consider the largest
-subset of \( Z = \left\{(q,p) \in \mathbb{S}^1 \times \R \mid
-\dot{E}_\text{avg}(q,p) = 0 \right\} \). We see from~\eqref{eqn:eavg-dot-neg}
-that \(\dot{E}_\text{avg}(q,p)\) is zero when $p=0$, \(\sin(q)=0\), or
-$L(\theta) = l_\text{avg}$. This latter condition by assumption is met when
-\(\theta \in \{0,\pi/2,\pi, 3\pi/2\}\), or equivalently when either $q=0$ or
-$p=0$. All in all, we have that
+the equilibrium \((q,p)=(0,0)\) is stable for~\eqref{eqn:negative-time-vlp}.
+We now apply the Krasovskii-LaSalle invariance principle, and consider the
+largest subset of 
+\( Z = \left\{(q,p) \in \mathbb{S}^1 \times \R \mid \dot{E}_\text{avg}(q,p) = 0
+\right\} \). 
+We see from~\eqref{eqn:eavg-dot-neg}
+that \(\dot{E}_\text{avg}(q,p)\) is zero when \(p = 0\), \(\sin(q) = 0\), or
+\(L(\theta) = l_\text{avg}\). 
+This latter condition by assumption is met when 
+\(\theta \in \{0,\pi/2,\pi, 3\pi/2\}\), or equivalently when either 
+\(q = 0\) or \(p = 0\). 
+All in all, we have that
 \[
-      Z= \left\{(q,p) \in \mathbb{S}^1 \times \R \mid q =0\right\} \cup \left\{(q,p) \in \mathbb{S}^1 \times \R \mid p =0\right\} 
+   Z= \left\{(q,p) \in \mathbb{S}^1 \times \R \mid q =0\right\} 
+   \cup \left\{(q,p) \in \mathbb{S}^1 \times \R \mid p =0\right\} 
    .
 \]
-It is easily seen that the largest invariant subset of \(Z\) is the union of the two equilibria
+It is easily seen that the largest invariant subset of \(Z\) is the union of the
+two equilibria
 \[
@@
 Krasovskii-LaSalle invariance principle implies that the set
-\(\Omega\) is globally attractive. Since the two equilibria are
-isolated, and since \((q,p)=(0,0)\) is stable, we deduce that the
-equilibrium \((q,p)=(0,0)\) is asymptotically stable for the negative
-time system~\eqref{eqn:negative-time-vlp}, and thus it is a repeller
+\(\Omega\) is globally attractive \cite{krasovskii_lasalle}. 
+Since the two equilibria are isolated, and since \((q,p)=(0,0)\) is stable, we
+deduce that the equilibrium \((q,p)=(0,0)\) is asymptotically stable for the
+negative time system~\eqref{eqn:negative-time-vlp}, and thus it is a repeller
 for the constrained dynamics in~\eqref{eqn:vlp-hamiltonian}.
@@
 dynamics~\eqref{eqn:vlp-hamiltonian} escape compact sets in finite
-time, i.e., that for each compact subset \(K\) of \(\mathbb{S}^1
-\times \R\), and for almost every condition \( (q_0,p_0) \in K \),
-there exists \( T > 0\) such that \( (q(t),p(t)) \not \in K \) for all
-\( t > T \), where \( (q(t),p(t)) \) denotes the solution
-of~\eqref{eqn:vlp-hamiltonian} with \( (q(0),p(0)) = (q_0,p_0)\). In
-what follows, we denote by \( x = (q,p) \) the state of the reduced
-dynamics~\eqref{eqn:vlp-hamiltonian}. Also, we denote by \(\Pi^-\) the
-stable manifold of the equilibrium \((\pi,0)\) for
-system~\eqref{eqn:vlp-hamiltonian}, defined as
+time, i.e., that for each compact subset \(K\) of 
+\(\mathbb{S}^1 \times \R\), and for almost every condition 
+\( (q_0,p_0) \in K \), there exists \( T > 0\) such that
+\( (q(t),p(t)) \notin K \) for all \( t > T \), where \( (q(t),p(t)) \)
+denotes the solution of~\eqref{eqn:vlp-hamiltonian} with 
+\( (q(0),p(0)) = (q_0,p_0)\). 
+In what follows, we denote by \( x = (q,p) \) the state of the reduced
+dynamics~\eqref{eqn:vlp-hamiltonian}. 
+Also, we denote by \(\Pi^-\) the stable manifold of the equilibrium \((\pi,0)\)
+for system~\eqref{eqn:vlp-hamiltonian}, defined as
 \[
@@
 \end{equation}
-whose spectrum is \(\{\pm \sqrt{{g}/{L(0)}}\}\).  Since one eigenvalue
-of (\ref{eqn:linearized-vlp}) is positive and one is negative, the
-stable manifold theorem implies that the stable manifold \( \Pi^- \)
-is a one-dimensional immersed submanifold of \(\mathbb{S}^1 \times
-\R\), and thus a set of zero measure.
+whose spectrum is \(\{\pm \sqrt{{g}/{L(0)}}\}\).  
+Since one eigenvalue of ~\eqref{eqn:linearized-vlp} is positive and one is
+negative, the stable manifold theorem implies that the stable manifold 
+\( \Pi^- \) is a one-dimensional immersed submanifold of
+\(\mathbb{S}^1 \times \R\), and thus is a set of measure zero
+\cite{stable_manifold}.
 
-Let \( K \subset \mathbb{S}^1 \times \R\) be an arbitrary compact set
-and let \( x(0) \in K \backslash \left(\Pi^- \cup
-\left\{(0,0)\right\}\right)\). Since \( x(0) \not \in \Pi^- \), we have that
+Let \( K \subset \mathbb{S}^1 \times \R\) be an arbitrary compact set and let 
+\( x(0) \in K \backslash \left(\Pi^- \cup \left\{(0,0)\right\}\right)\).
+Since \( x(0) \notin \Pi^- \), we have that
 \begin{equation}\label{eqn:vlp-x-does-not-converge}
-x(t) \not \to_{t \to \infty} (\pi,0).
+   x(t) \nrightarrow_{t \to \infty} (\pi,0).
 \end{equation}
-Suppose, by way of contradiction, that for each \(T
-> 0\) there exists \(t' > T\) such that \(x(t') \in K\).  Letting
+Suppose, by way of contradiction, that for each \(T > 0\) there exists 
+\(t' > T\) such that \(x(t') \in K\).  
+Letting
 \[
@@
 \] 
-and \(E_k =\{ x \mid E_\text{avg}(x) \leq k\} \), we have that \( K
-\subset E_k \), and thus \(x(t') \in E_k\). We have shown earlier that
-the function \( E_\text{avg} \) is nonincreasing along solutions
-of~\eqref{eqn:negative-time-vlp}, which implies that \(
-E_\text{avg}(x(t)) \) is nondecreasing for solutions of the reduced
-dynamics~\eqref{eqn:vlp-hamiltonian}, and therefore the half orbit
-\(x([0,t'])\) is contained fully within \(E_k\).  In particular,
-\(x(T) \in E_k\). Since this is true for each \( T >0 \), we deduce
-that \(x(t) \in E_k\) for all \(t \in [0,\infty[\).  Since \(E_k\) is
-    compact, \(x(t)\) has a positive limit set in \(E_k\) which, by
-    the Poincar\'{e}-Bendixson theorem, is either an equilibrium or a
-    closed orbit \cite{poincare_bendixson}.  We will show this
-    positive limit set must be an equilibrium and thereby reach a
+and \(E_k =\{ x \mid E_\text{avg}(x) \leq k\} \), we have that 
+\( K \subset E_k \), and thus \(x(t') \in E_k\).
+We have shown earlier that the function \( E_\text{avg} \) is nonincreasing
+along solutions of~\eqref{eqn:negative-time-vlp}, which implies that
+\( E_\text{avg}(x(t)) \) is nondecreasing for solutions of the reduced
+dynamics~\eqref{eqn:vlp-hamiltonian}, and therefore the half orbit \(x([0,t'])\)
+is contained fully within \(E_k\).  
+In particular, \(x(T) \in E_k\). Since this is true for each \( T >0 \), we
+deduce that \(x(t) \in E_k\) for all \(t \in [0,\infty[\).  
+Since \(E_k\) is compact, \(x(t)\) has a positive limit set in \(E_k\) which, by
+the Poincar\'{e}-Bendixson theorem, is either an equilibrium or a closed orbit
+\cite{poincare_bendixson}.  
+We will show this positive limit set must be an equilibrium and thereby reach a
 contradiction.
@@
 In order to rule out closed orbits, suppose that the reduced
-dynamics~\eqref{eqn:vlp-hamiltonian} have a periodic solution \(
-z(t)\). Since \( E_\text{avg}(z(t)) \) is nondecreasing and \(z(t)\)
+dynamics~\eqref{eqn:vlp-hamiltonian} have a periodic solution \( z(t)\). 
+Since \( E_\text{avg}(z(t)) \) is nondecreasing and \(z(t)\)
 is periodic, it must be that \( E_\text{avg}(z(t)) \) is constant, or
-\(\dot E_\text{avg}(z(t)) \equiv 0\). We have shown earlier
+\(\dot E_\text{avg}(z(t)) \equiv 0\). 
+We have shown earlier
 that\footnote{We have shown it for the negative time
 system~\eqref{eqn:negative-time-vlp}, but changing the sign of the
-  vector field does not change the set where \(\dot E_\text{avg}
-  =0\) nor the invariant sets.}
+vector field does not change the set where 
+\(\dot E_\text{avg} =0\) nor the invariant sets.}
 \[
-\{x \mid \dot E_\text{avg}(x)=0\} = \left\{(q,p) \mid q =0\right\}
+   \{x \mid \dot E_\text{avg}(x) = 0\} = \left\{(q,p) \mid q = 0\right\}
    \cup \left\{(q,p) \mid p =0\right\},
@@
 and the only invariant subset of this set is the union of two disjoint
-equilibria. Since the orbit \(z(\R)\) is an invariant set which is
+equilibria. 
+Since the orbit \(z(\R)\) is an invariant set which is
 connected, \(z(\R)\) must be an equilibrium and cannot be a nontrivial
@@
 of \( x(t)\) must be an equilibrium and,
-by~\eqref{eqn:vlp-x-does-not-converge}, this equilibrium must be \(
-(0,0)\). Since we have chosen \(x(0) \neq (0,0)\), and since \( (q,p) = (0,0) \) is a
-repeller, \( x(t) \) cannot converge to \( (0,0)\), which gives a
-contradiction.
+by~\eqref{eqn:vlp-x-does-not-converge}, this equilibrium must be 
+\( (0,0)\). 
+Since we have chosen \(x(0) \neq (0,0)\), and since \( (q,p) = (0,0) \) is a
+repeller, \( x(t) \) cannot converge to \( (0,0)\), which gives a contradiction.
 
 We conclude that the VNHC injects energy into the VLP.
-   By flipping the inequality of (\ref{eqn:vlp-energy-gain-condition})
+By flipping the inequality of ~\eqref{eqn:vlp-energy-gain-condition}
 we find the VLP is gaining energy in negative-time, so the VNHC is

### 2020-11-11 latex/vnhcs.tex
Edits to Manfredi's improvements of VLP proof
Changing regular ref to eqref for equations

@@
 \(\mathcal{L} : \mathcal{Q}\times\R^n \rightarrow \R\) defined by
-(\ref{eqn:lagrangian-general}) \cite{greenwood_dynamics},
+~\eqref{eqn:lagrangian-general} \cite{greenwood_dynamics},
 \begin{equation}\label{eqn:lagrangian-general}
@@
 ordinary differential equations (ODEs) obtained from the \textit{Euler-Lagrange
-equations} (\ref{eqn:el-eqns-general}) with \textit{generalized input forces} 
+equations} ~\eqref{eqn:el-eqns-general} with \textit{generalized input forces} 
 \(\tau \in \R^k\)
@@
 potential function \(P : \mathcal{Q} \rightarrow \R\) is smooth. 
-%If this is the case, the Euler-Lagrange equations reduce to (\ref{eqn:el-eqns}),
+%If this is the case, the Euler-Lagrange equations reduce to ~\eqref{eqn:el-eqns},
 %\begin{equation}\label{eqn:el-eqns}
@@
 Note that each \(p_i\) is a linear function of \(\dot{q}\), and one can typically
-solve for \(\dot{q}(q,p)\) by inverting all the expressions from (\ref{eqn:p-i}).
+solve for \(\dot{q}(q,p)\) by inverting all the expressions from ~\eqref{eqn:p-i}.
 The combined vector \((q,p) \in \mathcal{Q}\times\R^n\) is called a 
@@
 \((q,p)\) coordinates is the ``Legendre transform"
-(\ref{eqn:hamiltonian-legendre}) of the Lagrangian \cite{landau_mechanics},
+~\eqref{eqn:hamiltonian-legendre} of the Lagrangian \cite{landau_mechanics},
 \begin{equation}\label{eqn:hamiltonian-legendre}
@@
 
-If the kinetic energy of the system is quadratic as in (\ref{eqn:lagrangian}), 
+If the kinetic energy of the system is quadratic as in ~\eqref{eqn:lagrangian}, 
 the conjugate of momenta becomes \(p = D(q)\dot{q}\). Since \(D(q)\) is symmetric
 and positive definite, it is invertible at each \(q \in \mathcal{Q}\).
-The Legendre transform (\ref{eqn:hamiltonian-legendre}) becomes
+The Legendre transform ~\eqref{eqn:hamiltonian-legendre} becomes
 \begin{align*}
@@
 % inertia, D for normal coordinates
-Suppose we are given a Hamiltonian mechanical system (\ref{eqn:hamiltonian}).
+Suppose we are given a Hamiltonian mechanical system ~\eqref{eqn:hamiltonian}.
 Because \(\tau\) is transformed by the input matrix \(B(q)\) before
@@
 Under the following assumptions on the input matrix, we will show that there is
-a canonical transformation of (\ref{eqn:hamiltonian}) into simply actuated
+a canonical transformation of ~\eqref{eqn:hamiltonian} into simply actuated
 coordinates.
@@
     Under Assumptions \ref{assm:B-const},\ref{assm:B-perp}, and
-    \ref{assm:B-orthogonal}, the Hamiltonian system (\ref{eqn:hamiltonian})
+    \ref{assm:B-orthogonal}, the Hamiltonian system ~\eqref{eqn:hamiltonian}
     has simply actuated canonical coordinates 
     \(\left(Q = \mathbf{B}q, P = \mathbf{B}p\right)\). The resulting dynamics are 
-    given by (\ref{eqn:simple-hamiltonian}),
+    given by ~\eqref{eqn:simple-hamiltonian},
     \begin{align}\label{eqn:simple-hamiltonian}
@@
 \begin{thm}\label{thm:vnhc-regularity}
-    A relation \(h(q,p) = 0\) for system (\ref{eqn:simply-actuated-hamiltonian})
+    A relation \(h(q,p) = 0\) for system ~\eqref{eqn:simply-actuated-hamiltonian}
     is a regular VNHC of order \(k\) if and only if \(dh_{p_a} = 0\) 
@@
 \begin{cor}\label{cor:2d-zero-dynamics}
-    Let \(\mathcal{H}\) be the system (\ref{eqn:simple-hamiltonian}) with degree
+    Let \(\mathcal{H}\) be the system ~\eqref{eqn:simple-hamiltonian} with degree
     of underactuation one.
@@
     the orbit \((q_u(t),p_u(t))\) traces out
-    a curve on the 2D-plane \(\Rt{T_u} \times \R\) which we call the \((q,p)\)-plane.
+    a curve on the 2D constraint manifold 
+    \(\Gamma = \Rt{T_u} \times \R\) which we call the \((q_u,p_u)\)-plane.
 \end{cor}
+
+We conclude this chapter by formalizing the notion of energy injection for
+VNHCs.
+To glean some intuition for this idea, suppose our mechanical system satisfies
+the above corollary so that the constraint manifold \(\Gamma\) is the
+\((q_u,p_u)\)-plane.
+If any initial condition of the constrained dynamics converges to the origin, 
+the system must be losing energy because the unactuated momentum
+\(p_u\) is decreasing.
+Any notion of energy injection must therefore require that the origin of 
+\(\Gamma\) repels all solutions.
+Furthermore, periodic solutions prevent the system from attaining higher speeds,
+so the constrained dynamics should not have closed orbits or limit cycles.
+Finally, we want the unactuated momentum to increase on average, regardless of
+its initial value. 
+This intuition is generalized to ODEs on manifolds by the following definition.
+
+% TODO: Change this, 0 is not an element of the manifold. Do we need other
+% notions, like a riemannian metric? Should we just use Gamma?
+\begin{defn}\label{defn:energy-gain}
+    Let \(M\) be an \(n\)-dimensional manifold and \(D \subset M\) be open.
+    Let \(f : D \rightarrow \R^n\) be smooth with \(f(0) = 0\).
+    The system described by \(\dot{x} = f(x)\) 
+    \textit{gains energy on \(D\)} if 
+    \begin{enumerate}
+        \item If \(0 \in D\), then \(0\) is a repeller 
+         (or equivalently, it is asymptotically stable in negative time).
+      \item For all compact sets \(K \subset D\) and for almost every initial
+         condition \(x(0) \in K\), there exists \(T > 0\) such
+         that \(x(t) \notin K \, (\forall t > T)\).
+    \end{enumerate}
+    The system \textit{loses energy on \(D\)} if it gains energy in
+    negative-time.
+\end{defn}
+
+The second property of Definition \ref{defn:energy-gain} allows the system
+to have unstable equilibria, but not limit cycles nor closed orbits.
+The next definition ties this notion of energy gain to VNHCs.
+
+\begin{defn}\label{defn:energy-injection}
+    A regular VNHC \(h(q, p) = 0\) with constraint manifold \(\Gamma\)
+    \textit{injects (dissipates) energy} on \(D \subset \Gamma\) if the
+    constrained dynamics gain (lose) energy everywhere on \(D\), except possibly
+    on a set of measure zero.
+\end{defn}
 

### 2020-11-12 latex/acrobot.tex
Edited acrobot constraint and improved theorem

@@
 \(\mathcal{Q} \times \mathcal{P}\) where
-\[
-    \mathcal{Q} = \mathcal{Q}_u \times \mathcal{Q}_a := 
-    \mathbb{S}^1 \times \mathbb{S}^1
-    ,
-\]
-and
-\[
-    \mathcal{P} = \mathcal{P}_u \times \mathcal{P}_a
-    := \R \times \R
-    .
-\]
-This allows us once again to apply the theory of VNHCs from Chapter
+\(\mathcal{Q} = \mathcal{Q}_u \times \mathcal{Q}_a 
+:= \mathbb{S}^1 \times \mathbb{S}^1\), and
+\(\mathcal{P} = \mathcal{P}_u \times \mathcal{P}_a
+:= \R \times \R\).
+This allows us to apply the theory of VNHCs from Chapter
 \ref{ch:vnhcs}.
@@
 Theorem \ref{thm:vnhc-regularity} tells us that this VNHC will be regular
-if the regularity matrix
+when the regularity matrix
 \[
@@
 \]
-is of full rank on the constraint manifold \(\Gamma\).
-Then, since 
-\[
-    dh_q = \begin{bmatrix}
-        -\partial_{q_u} f & 1
-    \end{bmatrix}
-    ,
-\]
+is of full rank 1 on the constraint manifold \(\Gamma\).
+Given that
+\(dh_q = \begin{bmatrix} -\partial_{q_u} f & 1 \end{bmatrix}\),
 the regularity matrix evaluates to the scalar equation
@@
 change sign.
-
-A sufficient condition for regularity is when \(f\) is a function solely of \(p_u\),
-because then \(\partial_{q_u}f = 0\) and ~\eqref{eqn:regularity-matrix-acrobot}
-is strictly positive for all values of \(q_a\). 
-This will be useful later, as we will not need to check regularity if we design
-a function of the unactuated momentum.
+The following proposition provides a sufficient condition for this to happen.
+
+\begin{prop}\label{prop:acrobot-fpu-regular}
+    A relation \(h(q,p) = q_a - f(p_u) = 0\) for ~\eqref{eqn:acrobot-hamiltonian} 
+    with \(f \in C^2\left(\mathcal{P}_u; \mathcal{Q}_a\right)\) is a regular
+    VNHC of order 1.
+\end{prop}
+\begin{proof}
+    Since \(\partial_{q_u} f = 0\), the regularity equation
+    ~\eqref{eqn:regularity-matrix-acrobot} is strictly positive for all values
+    of \(q_a\), and hence is of full rank 1.
+    By Theorem \ref{thm:vnhc-regularity}, \(h\) is a regular VNHC of order 1.
+\end{proof}
+Proposition \ref{prop:acrobot-fpu-regular} will be useful later, as we will not
+need to check regularity if we design a function of the unactuated momentum.
 
@@
 Then we could apply Theorem \ref{thm:vlp-energy-stabilization} to prove
-the acrobot is gaining energy too.
+the acrobot is gaining energy.
 \begin{figure}
@@
 If we want to design a VNHC for this system, we cannot use any of the results
-from Chapter \ref{ch:vlp} because the VLP models don't match.
+from Chapter \ref{ch:vlp} because the VLP models do not match.
 
@@
 Unfortunately, this constraint is not regular, and it is difficult to find any
-VNHC of the form \(q_a = f(\theta)\) where it is easy to prove regularity
-analytically.
-Instead, we will develop a constraint \(h(q,p) = q_a - f(p_u)\) which,
-as we saw earlier, is always regular and is likely to inject energy.
+VNHC of the form \(q_a = f(\theta)\) where regularity can be proven easily.
+Instead, we will develop a constraint \(h(q,p) = q_a - f(p_u)\) because these
+constraints are always regular (as per Proposition \ref{prop:acrobot-fpu-regular}).
 
@@
 This position is identical to that of a gymnast on a bar, which is why we can
-use the leg motion from seated swinging to design a controller for the acrobot.
+use leg motion from the seated swing to design a controller for the acrobot.
 
@@
 angle proportional to the swing's speed.
-One VNHC which emulates this process is \(q_a = \bar{q}_a\arctan_2( I p_u)\), as
+One VNHC which emulates this process is \(q_a = \bar{q}_a\arctan( I p_u)\), as
 in Figure \ref{fig:qa-arctan}.
@@
 \begin{equation}\label{eqn:acrobot-constraint}
-    h(q,p) = q_a - \bar{q}_a \arctan_2(I p_u)
+    h(q,p) = q_a - \bar{q}_a \arctan(I p_u)
     .
@@
     \includestandalone[width=0.5\textwidth]{images/qa_arctan}
-    \caption{The acrobot constraint \(q_a = \bar{q}_a \arctan_2(I p_u)\).}
+    \caption{The acrobot constraint \(q_a = \bar{q}_a \arctan(I p_u)\).}
     \label{fig:qa-arctan}
@@
 
-These dynamics do not always gain energy -- we need additional conditions
-on the control parameters to guarantee this is true.
-
-% TODO: This needs modification. The energy injection is provable for any system
-% only during oscillations. For rotations, we need to do a numerical check of an
-% integral that depends on m,g,l. If this integral is positive, then the system
-% is gaining energy up to a point - it sems the integral decays to 0 so that you
-% can reach a limiting speed, which means you don't get full energy injection
-% everywhere on Gamma.
+These dynamics do not always gain energy -- we need \(|I|\) to be small enough
+to guarantee this is true.
+
 \begin{thm}\label{thm:acrobot-energy-stabilization}
     Consider the simple acrobot ~\eqref{eqn:acrobot-hamiltonian} constrained by
-    ~\eqref{eqn:acrobot-constraint} with \(\bar{q}_a \in ]0,1]\).
-    There exists a small enough\footnotemark \(I^\star \in \R_{>0}\)
-    such that the VNHC injects energy whenever \( I \in ]0,I^\star[\). 
-    If \(I \in ]-I^\star,0[\), the VNHC dissipates energy.
+    the VNHC ~\eqref{eqn:acrobot-constraint}, whose
+    constraint manifold is \(\Gamma = \mathbb{S}^1 \times \R\).
+    Let \(E\) be the nominal energy when \(I = 0\):
+    \[
+        E(q_u,p_u) := \frac{p_u^2}{10ml^2} + 3(1 - \cos(q_u))
+        .
+    \]
+\begin{enumerate}
+    \item There always exists \(I \in \R_{>0}\) small enough that 
+    ~\eqref{eqn:acrobot-constraint} injects energy into the acrobot on the set
+    \[
+        \mathcal{O} := \left\{(q_u,p_u) \in \Gamma 
+            \mid E(q_u,p_u) < E(\pi,0) \right\}
+        .
+    \]
+    If instead \(I \in \R_{<0}\), the VNHC dissipates energy on \(\mathcal{O}\).
+\item Define \(b : \mathbb{S}^1 \times \R_{> 0} \rightarrow \R\) by
+    \[
+        b(\alpha,\mu_0) := 
+        \frac{5m^2 g l^3 \bar{q}_a \left(
+            m^2gl^3\left(18s_\alpha^2 + 30c_\alpha(1 - c_\alpha)\right)
+            - c_\alpha\mu_0^2
+        \right)}{
+        |\mu_0|\sqrt{\mu_0^2 - 30m^2gl^3(1 - c_\alpha)}
+        }
+        ,
+    \]
+    and fix \(\bar{\mu} > 60m^2gl^3\).
+    Suppose \(\int \limits_{0}^{2\pi} b(\sigma,\mu_0)d\sigma > 0\) for all 
+    \(\mu_0 \in ]60m^2gl^3, \bar{\mu}[\).
+    Then there exists \(I \in \R_{> 0}\) small enough that
+    ~\eqref{eqn:acrobot-constraint} injects energy into the acrobot on
+    \[
+        \Omega := \left\{(q_u,p_u) \in \Gamma 
+            \mid E(q_u,p_u) < E(0,\bar{\mu})\right\}
+        .
+    \]
+    If instead \(I \in \R_{<0}\), the VNHC dissipates energy on \(\Omega\).
+\end{enumerate}
 \end{thm}
-\footnotetext{In fact, we can be more precise than this. 
-    Letting \(\epsilon > 0\), we need 
-    \(I \leq \min\left\{I_1(\epsilon),I_2\right\}\) where
-    \(I_1(\epsilon) = \frac{\tan(\epsilon)}{(\pi-\epsilon)}
-    \left(\frac{m^2gl^3(\cos(\epsilon)-2\sin(\epsilon))}{\pi-\epsilon}
-    \left(\frac{2\bar{q}_a\tan(\epsilon)}{(1+\tan(\epsilon)^2)(\pi-\epsilon)}+5\right) 
-    \right)^{-1/2}\)
-    and \(I_2 = \sqrt{\frac{3}{10 m^2gl^3\bar{q}_a^2}}\).
-    See Apendix \ref{apx:I-bounds} for details.
-}
+%\footnotetext{In fact, we can be more precise than this. 
+%    Letting \(\epsilon > 0\), we need 
+%    \(I \leq \min\left\{I_1(\epsilon),I_2\right\}\) where
+%    \(I_1(\epsilon) = \frac{\tan(\epsilon)}{(\pi-\epsilon)}
+%    \left(\frac{m^2gl^3(\cos(\epsilon)-2\sin(\epsilon))}{\pi-\epsilon}
+%    \left(\frac{2\bar{q}_a\tan(\epsilon)}{(1+\tan(\epsilon)^2)(\pi-\epsilon)}+5\right) 
+%    \right)^{-1/2}\)
+%    and \(I_2 = \sqrt{\frac{3}{10 m^2gl^3\bar{q}_a^2}}\).
+%    See Apendix \ref{apx:I-bounds} for details.
+%}
+The region \(\mathcal{O}\) is the oscillation domain of the simple pendulum
+obtained by setting \(q_a = 0\);
+\(\Omega\) includes \(\mathcal{O}\), alongside all rotations of this 
+pendulum with momentum less than \(\bar{\mu}\). 
+Hence, the first result of Theorem \ref{thm:acrobot-energy-stabilization} states
+that the acrobot will gain enough energy to begin performing giant motion;
+the second result guarantees can achieve giants with momentum
+\(|p_u| < \bar{\mu}\).
 
@@
 \(\mu \in \R_{>0}\) and pseudo-angle \(\alpha \in \mathbb{S}^1\) on the
-\((q_u,p_u)\)-plane where \(\dot{\mu} = 0\) and \(\dot{\alpha} > 0\) when 
-we set \(I = 0\).
-Perturbation theory shows that, for small enough \(I\), the radius \(\mu\)
+\((q_u,p_u)\)-plane.
+These coordinates have the property that \(\dot{\alpha} > 0\) when
+\(I\) is small enough.
+Perturbation theory shows that (for possibly smaller \(I\)) the radius \(\mu\)
 increases on average along \(\alpha\).
-Hence, the acrobot is gaining energy since orbits in the \((q_u,p_u)\)-plane
-spiral away from the origin, and the acrobot always begins performing rotations.
+Hence, orbits in the \((q_u,p_u)\)-plane spiral away from the origin, 
+which means the acrobot is gaining energy.
 

### 2020-11-13 latex/vlp.tex
Slight modifications to VLP phrasing and definition of energy injection

@@
    A regular VNHC of the form 
-   \(h(\mathbf{q},\mathbf{p}) = l - L(\theta)\) injects energy into the VLP
-   if there exists 
-   \(l_\text{avg} \in [\underline{l},\overline{l}]\) such that 
+   \(h(\mathbf{q},\mathbf{p}) = l - L(\theta)\) for the VLP injects energy 
+   on the constraint manifold \(\Gamma = \mathbb{S}^1 \times \R\) if there
+   exists \(l_\text{avg} \in [\underline{l},\overline{l}]\) such that 
    \begin{equation}\label{eqn:vlp-energy-gain-condition}
@@
    \]
-   the VNHC dissipates energy.
+   the VNHC dissipates energy on \(\Gamma\).
 \end{thm}
@@
 dynamics~\eqref{eqn:vlp-hamiltonian} escape compact sets in finite
-time, i.e., that for each compact subset \(K\) of 
+time, \ie, that for each compact subset \(K\) of 
 \(\mathbb{S}^1 \times \R\), and for almost every condition 
@@
 
-We conclude that the VNHC injects energy into the VLP.
+We conclude that the VNHC injects energy into the VLP on 
+\(\mathbb{S}^1 \times \R\).
 By flipping the inequality of ~\eqref{eqn:vlp-energy-gain-condition}

### 2020-11-13 latex/vnhcs.tex
Slight modifications to VLP phrasing and definition of energy injection

@@
 \begin{defn}\label{defn:energy-gain}
-    Let \(M\) be an \(n\)-dimensional manifold and \(D \subset M\) be open.
-    Let \(f : D \rightarrow \R^n\) be smooth with \(f(0) = 0\).
+    Let \(M = \Rt{T_1} \times \cdots \times \Rt{T_n}\) 
+    (where \(T_1, \ldots, T_n \in \{2\pi, \infty\}\)) be an
+    \(n\)-dimensional generalized cylinder.
+    Let \(f : M \rightarrow \R^n\) be smooth with \(f(0) = 0\)
+    and let \(D \subset M\) be open.
     The system described by \(\dot{x} = f(x)\) 
-    \textit{gains energy on \(D\)} if 
+    \textit{gains energy on \(D\)} if these conditions are satisfied:
     \begin{enumerate}
@@
     A regular VNHC \(h(q, p) = 0\) with constraint manifold \(\Gamma\)
-    \textit{injects (dissipates) energy} on \(D \subset \Gamma\) if the
+    \textit{injects (dissipates) energy on \(D \subset \Gamma\)} if the
     constrained dynamics gain (lose) energy everywhere on \(D\), except possibly

### 2020-11-13 latex/acrobot.tex
First editing pass of oscillation analysis

@@
 \(q = (q_u,q_a) \in \mathbb{S}^1 \times \mathbb{S}^1\). 
-The acrobot has inertia matrix \(D\), potential function \(P\) (with respect to
+The acrobot has inertia matrix \(mathcal{O}\), potential function \(P\) (with respect to
 the horizontal bar), and input matrix \(B\)  given as
@@
 is conserved.
-Level sets of \(E\) are ellipses when \(E < E(\pi,0)\), which we call
-``oscillations"; and they are open curves when \(E > E(\pi,0)\), which we call
+Level sets of \(E\) are ellipses when \(E(q_u,p_u) < E(\pi,0)\), which we call
+``oscillations"; and they are open curves when \(E(q_u,p_u) > E(\pi,0)\), which we call
 ``rotations". 
@@
 
-Drawing inspiration from \citet{dynamic_vhcs_stabilize_closed_orbits}, we want
-to find a change of coordinates \((q_u,p_u) \to (\alpha, \mu)\) with the
+Using a method developed by \citet{dynamic_vhcs_stabilize_closed_orbits},
+we can find a change of coordinates \((q_u,p_u) \to (\alpha, \mu)\) with the
 following properties:
 \begin{itemize}
-    \item All oscillations in \((\alpha,\mu)\) coordinates are represented by
-        \(\mu = \text{const}\).
+    \item The energy of oscillation in \((\alpha,\mu)\) coordinates is 
+        uniquely defined by \(\mu\), which remains constant along oscillations
+        of the simple pendulum.
     \item \(\alpha\) is a pseudo-angle living in \(\mathbb{S}^1\) which is
@@
 \end{itemize}
-We also want to find a second set of coordinates where the same properties hold for
-rotations.
 Once we have these coordinates in hand, we can use perturbation theory to prove
-that \(\mu\) increases on average. 
-If this is true, the acrobot must be gaining energy since the momentum is
-increasing over time.
+that \(\mu\) increases on average for the acrobot when \(I\) is small enough.
+Likewise, we will find a second set of coordinates \((\beta, \rho)\) where
+\(\beta \in \mathbb{S}^1\) is always increasing, \(\rho\) is constant along
+rotations of the simple pendulum, and \(\rho\) increases for the acrobot when
+\(I\) is small enough.
+Putting these together, we will show the acrobot is gaining energy.
 
 \subsection*{Background on Perturbation Theory}
-Perturbation theory allows one to understand how solutions of nonlinear systems
-behave by studying a nominal system which can be analyzed analytically.
-The solutions of nonlinear systems are approximated by a Taylor-like expansion
-around this nominal system.
+Nonlinear systems like the acrobot are difficult (or impossible) to solve
+analytically.
+Perturbation theory allows one to understand the behaviour of the nonlinear system
+by studying a simpler nominal system. 
+Solutions of the nonlinear system can often be approximated by taking a Taylor
+expansion around this nominal system.
 
@@
 sufficiently smooth.
-In the context of the acrobot, \(x\) is the phase in whichever coordinates make
+In the context of the acrobot, \(x\) is the set of coordinates which make
 our analysis convenient;
@@
 
-Letting \(I = 0\) we get the nominal system
+Setting \(I = 0\) we get the nominal system
 \begin{equation}\label{eqn:khalil-perturbation-nominal}
@@
 \end{equation}
-Finally, one can compute the first-order approximation
+Finally, one can compute the first-order Taylor approximation
 \(x(t,\eta_0) = x_0(t,\eta_0) + I x_1(t,\eta_0) + R(t,I)\), where the remainder
-term \(R(t,I)\) is \(O(I^2)\);
-that is,
+term \(R(t,I)\) is \(O(I^2)\), \ie,
 \[
@@
 accuracy of perturbation analysis.
-\begin{thm}
+\begin{thm}\label{thm:khalil-perturbation}
     Suppose \(f : [t_0,t_1] \times D \times [-I_0,I_0] \rightarrow \R^2\) is
@@
     unique solution \(x_0(t,\eta_0) \in D\) on \([t_0,t_1]\).
-    Then there exists \(I^\star > 0\) such that for all \(|I| < I^\star\), the
+    Then there exists \(I^\star > 0\) such that for all \(I,\, |I| < I^\star\), the
     solution \(x(t,I,\eta_0)\) to ~\eqref{eqn:khalil-setup} satisfies
     \[
-        \norm{x(t,I,\eta_0) - \left(x_0(t,\eta_0) + I x_1(t,\eta_))\right)} \leq k |I^2|
+        \norm{x(t,I,\eta_0) - \left(x_0(t,\eta_0) + I x_1(t,\eta_0)\right)} \leq k |I^2|
     \]
@@
 
-The above Theorem tells us that we can approximate the solution of the nonlinear
-system by the Taylor approximation 
+Theorem \ref{thm:khalil-perturbation} tells us that we can approximate the
+solution of the nonlinear system by the Taylor approximation 
 \(x(t,\eta_0) \approx x_0(t,\eta_0) + I x_1(t,\eta_0)\).
-In other words, when \(I\) is small enough solutions of the nonlinear system and
+When \(I\) is small enough, solutions of the nonlinear system and
 the Taylor approximation are the same up to order \(I^2\) along compact time
@@
 Since we know that the acrobot behaves like a pendulum at \(I = 0\), we can
-use this theory to more easily study the behaviour of orbits and prove energy
-injection properties of our VNHC.
+use this theory to prove energy injection properties on the acrobot by studying
+the simple pendulum.
 
 \subsection*{Perturbation Analysis for Oscillations}
-When \(I = 0\), the acrobot becomes a nominal pendulum that oscillates whenever the
+As we have seen, the acrobot becomes a nominal pendulum that oscillates whenever the
 nominal energy ~\eqref{eqn:acrobot-nominal-E} is less than \(E(\pi,0)\).
-Hence, the domain of oscillations on \(\Gamma\) is given by
-\[
-    D := \left\{ (q_u,p_u) \in \Gamma | E(q_u,p_u) < E(\pi,0)\right\}
-    .
-\]
-We wish to find a transformation \(T(q_u,p_u)\) on \(D\) into
+Hence, the domain of oscillations for the nominal pendulum is given by
+\[
+    \mathcal{O} := \left\{ (q_u,p_u) \in \Gamma \mid E(q_u,p_u) < E(\pi,0)\right\}
+    ,
+\]
+where \(\Gamma = \mathbb{S}^1\times\R\) is the consraint manifold of the
+acrobot.
+We wish to find a transformation \(T(q_u,p_u)\) on \(\mathcal{O}\) into
 \((\alpha,\mu)\)-coordinates where \(\alpha \in \mathbb{S}^1\) is a pseudo-angle
@@
 Since the energy of oscillation can be determined by the orbit's intersection
-with the \(q_u\)-axis and \(q_u \in ]-\pi,\pi[\) on \(D\), 
-we set \(\mu \in ]0,\pi[\) as in Figure \ref{fig:mu-intersection}.
+with the \(q_u\)-axis (as in Figure \ref{fig:mu-intersection}) 
+and \(q_u \in ]-\pi,\pi[\) on \(\mathcal{O}\), we set \(\mu \in ]0,\pi[\).
 The transformation we want is therefore a diffeomorphism of the form
 \begin{align*}
-    T : D\backslash \{(0,0)\} &\rightarrow \mathbb{S}^1 \times \, ]0,\pi], \\
+    T : D\backslash \{(0,0)\} &\rightarrow \mathbb{S}^1 \times \, ]0,\pi[, \\
     (q_u, p_u) &\mapsto (\alpha,\mu)
@@
     %\includestandalone[]{images/mu_intersection}
-    \caption{The domain \(D\) where a pendulum oscillates.
+    \caption{The domain \(\mathcal{O}\) where a pendulum oscillates.
     The pseudo-radius coordinate \(\mu\) corresponds to the
@@
 parameterized by \(q_u = \mu \cos(\alpha)\), where \(\alpha\) is our
-pseudo-angle.
+desired pseudo-angle.
 Substituting this into ~\eqref{eqn:oscillation-pu2}, we get
@@
 \end{equation}
-which is, in fact, smooth.
+which is smooth for all \(\mu \in ]0,\pi[\).
 
@@
 \]
-Since \(\mu \in ]0,\pi]\), we can uniquely express \(\mu\) by
+Since \(\mu \in ]0,\pi[\), we can uniquely express \(\mu\) by
 \begin{equation}\label{eqn:mu-qupu}
@@
 Recall that 
-\[
+\begin{equation}\label{eqn:acrobot-cosalpha-qu}
     \cos(\alpha) = \frac{q_u}{\mu}
     ,
-\] 
+\end{equation}
 which means 
-\[
+\begin{equation}\label{eqn:acrobot-sinalpha-qu}
     \sin(\alpha) = \pm \sqrt{1 - \frac{q_u^2}{\mu^2}}
     .
-\]
-Notice from ~\eqref{eqn:oscillation-pu} that 
+\end{equation}
+Using ~\eqref{eqn:oscillation-pu}, we determine that
 \(\sign{\sin(\alpha)} = -\sign{p_u}\).
-We use these facts to deduce that
+Putting together ~\eqref{eqn:acrobot-cosalpha-qu} and
+~\eqref{eqn:acrobot-sinalpha-qu}, we deduce that
 \begin{equation}\label{eqn:alpha-qupu}
@@
         \right|_{\mu = \arccos(C_\mu(q_u,p_u))}
-    ,
-\end{equation}
-and thus have our transformation \(T(q_u,p_u)\) into
-\((\alpha,\mu)\)-coordinates on \(D\).
+    .
+\end{equation}
+Thus, ~\eqref{eqn:mu-qupu} and ~\eqref{eqn:alpha-qupu} define 
+our transformation into \((\alpha,\mu)\)-coordinates on \(\mathcal{O}\).
 
@@
 \end{align*}
-with \((\dot{q}_u,\dot{p}_u)\) the given by 
+with \((\dot{q}_u,\dot{p}_u)\) given by 
  ~\eqref{eqn:acrobot-constrained-dynamics}.
@@
 \end{align}
-The nominal system \((\alpha_n,\mu_n)\) comes from setting \(I = 0\).
+The dynamics of the nominal pendulum arise from evaluating
+\(f_\mu\) and \(f_\alpha\) at \(I = 0\).
 MATLAB's symbolic toolbox evaluates the nominal dynamics as
@@
 \end{align}
-One can confirm that \(\dot{\alpha}_n > 0\) for every \(\mu_n \in ]0,\pi[\) -- this
-is expected, since the nominal system is a single pendulum.
-By continuity of ~\eqref{eqn:acrobot-mu-dot} with respect to \(I\), 
-\(\dot{\alpha}\) remains positive for \(I\) small enough on the domain \(D\).
-
-Suppose at time \(t = 0\) we have initialized the system at 
+Note that we write the nominal coordinates as \((\alpha_n,\mu_n)\) to
+distinguish them from the acrobot's coordinates.
+One can confirm that \(\dot{\alpha}_n > 0\) for every \(\mu_n \in ]0,\pi[\).
+This is expected of an oscillating single pendulum.
+By continuity of ~\eqref{eqn:acrobot-mu-dot}, 
+\(\dot{\alpha}\) remains positive on \(\mathcal{O}\) for \(I\) small enough.
+
+Suppose at time \(t = 0\) we have initialized the acrobot at 
 \((q_u,p_u) = (\mu_0,0)\), which corresponds to \((\alpha,\mu) = (0,\mu_0)\).
 Since \(\alpha\) is continuously increasing, we know that the orbit will do one
-of two things: it will either leave \(D\) so that the acrobot is now rotating,
-or it will remain in \(D\) and hit the \(q_u\)-axis at some point 
-\((\alpha,\mu) = (\pi,m)\) as is illustrated in Figure
-\ref{fig:acrobot-possible-orbits}.
-From there, the acrobot could escape \(D\) along the second half of the orbit or
+of two things: it will either leave \(\mathcal{O}\) so that the acrobot is now rotating,
+or it will remain in \(\mathcal{O}\) and hit the \(q_u\)-axis at some point 
+\((\alpha,\mu) = (\pi,m)\) (as illustrated in Figure
+\ref{fig:acrobot-possible-orbits}).
+From there, the acrobot could escape \(\mathcal{O}\) along the second half of the orbit or
 return to some point \((\alpha,\mu) = (0,\tilde{m})\).
@@
     \caption{If the acrobot is initialized at \((q_u,p_u) = (\mu_0,0)\), then
-        either the orbit escapes \(D\) into the rotation
+        either the orbit escapes \(\mathcal{O}\) into the rotation
         zone (purple), or the orbit hits the other half of the \(q_u\)-axis 
@@
 function of \(\alpha\) rather than as a function of \(t\) (assuming the acrobot
-does not leave \(D\)).
+does not leave \(\mathcal{O}\)).
 The dynamics of \(\hat{\mu}\) are
@@
 \end{equation}
-
-In the remainder of this section we will find a first-order
-approximation of \(\hat{\mu}\). 
-We will show the dynamics of this approximation are \(\pi\)-periodic in
-\(\alpha\), so that we only need to study whether \(\hat{\mu}\) increases or
-decreases along a half-orbit \(\alpha \in [0,\pi]\).
-To complete the proof of energy gain for oscillations, we will demonstrate that
-\(\hat{\mu}(\pi) > \hat{\mu}(0)\).
-This will imply that the the radius of orbits increases
-until all orbits within \(D\) eventually leave \(D\) in finite time.
-
-In the spirit of perturbation analysis, we expand our solution of
+with solution \(\hat{\mu}(\alpha,\mu_0,I)\).
+
+%Using perturbation theory, we will find a first-order
+%approximation of \(\hat{\mu}\). 
+%We will show the dynamics of this approximation are \(\pi\)-periodic in
+%\(\alpha\), so that we only need to study whether \(\hat{\mu}\) increases or
+%decreases along a half-orbit \(\alpha \in [0,\pi]\).
+%To complete the proof of energy gain for oscillations, we will demonstrate that
+%\(\hat{\mu}(\pi) > \hat{\mu}(0)\).
+%This will imply that the the radius of orbits increases
+%until all orbits within \(\mathcal{O}\) eventually leave \(\mathcal{O}\) in finite time.
+
+In the spirit of perturbation analysis, we expand the solution of
 ~\eqref{eqn:muhat-dot} into
 \begin{equation}\label{eqn:acrobot-muhat-approx}
-    \hat{\mu}(\alpha) = \hat{\mu}_0(\alpha) + I \hat{\mu}_1(\alpha)
-    + R(\alpha,I)
-    ,
-\end{equation}
-where \(R(\alpha,I)\) is \(O(I^2)\).
+    \hat{\mu}(\alpha,\mu_0,I) = \hat{\mu}_0(\alpha,\mu_0) + I
+    \hat{\mu}_1(\alpha,\mu_0)
+    + R(\alpha,\mu_0,I)
+    ,
+\end{equation}
+where \(R(\alpha,\mu_0,I)\) is \(O(I^2)\).
 From ~\eqref{eqn:khalil-perturbation-nominal} we know that 
@@
 Since \(g(\alpha,\mu,0) = 0\) by 
-~\eqref{eqn:acrobot-mu-dot-nom})-(\ref{eqn:acrobot-alpha-dot-nom}, we find that
-\(\hat{\mu}_0(\alpha) = \mu_0\) for all \(\alpha\).
+~\eqref{eqn:acrobot-mu-dot-nom}-\eqref{eqn:acrobot-alpha-dot-nom}, we find that
+\(\hat{\mu}_0(\alpha,\mu_0) \equiv \mu_0\) for all \(\alpha\).
 
@@
 \end{equation}
-Suppose we start the system at \((\alpha,\mu) = (0,\mu_0)\). 
-Then the new pseudo-radius after a half orbit where \(\alpha\) goes from \(0\)
-to \(\pi\) (that is, when the orbit hits the negative \(q_u\)-axis)
-is \(\rho = \hat{\mu}(\pi,\mu_0)\).
-Hence, the change in pseudo-radius is \(\rho - \mu_0 = I\hat{\mu}_1(\pi,\mu_0)\).
-Now look at the full orbit where \(\alpha\) goes from \(0\) to \(2\pi\). 
-This lands on the positive \(q_u\)-axis with pseudo-radius
-\(\tilde{\rho} = \mu_0 + I\hat{\mu}_1(2\pi,\mu_0)\).
-As we know from ~\eqref{eqn:bhat-integral-properties}, we have
+Suppose we initialize the acrobot at \((\alpha,\mu) = (0,\mu_0)\). 
+After half an orbit, when \(\alpha = \pi\), the acrobot will have a new
+pseudo-radius \(m = \hat{\mu}(\pi,\mu_0)\).
+Hence, the change in pseudo-radius is 
+\(m - \mu_0 \approx I\hat{\mu}_1(\pi,\mu_0)\).
+After an additional half orbit, the acrobot lands on the positive \(q_u\)-axis
+with pseudo-radius
+\(\tilde{m} \approx \mu_0 + I\hat{\mu}_1(2\pi,\mu_0)\).
+The properties outlined in ~\eqref{eqn:bhat-integral-properties} imply that
 \(\hat{\mu}_1(2\pi,\mu_0) = 2 \hat{\mu}_1(\pi,\mu_0)\). 
 This means the change in pseudo-radius after one entire oscillation is 
-\(\tilde{\rho} - \mu_0 = 2I\hat{\mu}1(\pi,\mu_0) = 2 (\rho - \mu_0)\).
-Hence, if we can show that \(\hat{\mu}_1(\pi,\cdot)\) is always positive, we
-will guarantee that the pseudo-radius increases after every half oscillation.
-This in turn guarantees \(q_u\) is increasing after every half oscillation, so
-that eventually we will reach a rotation.
-
-To show that the first-order approximation ~\eqref{eqn:acrobot-muhat-approx} is
-gaining energy, we define the Poincar\'{e} section
+\(\tilde{m} - \mu_0 \approx 2I\hat{\mu}_1(\pi,\mu_0) \approx 2 (m - \mu_0)\).
+Hence, if \(\hat{\mu}_1(\pi,\cdot)\) is always positive,
+\(\hat{\mu}\) must be increasing every half oscillation.
+This in turn would guarantee \(q_u\) is increasing, which means the acrobot
+will eventually reach a rotation.
+
+To prove this intuitive argument rigorously, we define the Poincar\'{e} section
 \(P(\mu_0) := \hat{\mu}(\pi,\mu_0)\), which expands into
@@
 increasing, and \(\lim \limits_{\mu_0 \to \pi} P(\mu_0) = \infty\).
-Thus, there are no closed orbits and the origin must be a repeller.
-Furthermore, for any non-zero initial condition of the acrobot and for small
-enough \(I\), \((q_u,p_u)\) will eventually exit \(D\).
-This implies the orbit will escape any compact set in \(D\) in finite time. 
-By Definition \ref{defn:energy-gain}, the acrobot is gaining energy on \(D\).
+
+Because \(\dot{\alpha} > 0\), any orbit which begins and remains in \(\mathcal{O}\)
+must spiral clockwise around the origin until it intersect the \(q_u\) axis.
+Once it does, the Poincar\'{e} argument demonstrates it must spiral away
+from the origin; hence, there are no closed orbits and the origin is a repeller.
+
+Once the orbit reaches an intersection point \(\mu_0\) on the \(q_u\)-axis such
+that \(P(\mu_0) > \pi\), the orbit will no longer return to the \(q_u\)-axis; it
+will escape \(\mathcal{O}\) entirely.
+This must always happen in finite time, which implies the orbit must also escape
+any compact set in \(mathcal{O}\) in finite time. 
+By Definition \ref{defn:energy-gain}, the VNHC is injecting energy into the
+acrobot on \(\mathcal{O}\).
 
@@
 \(q_a = -\bar{q}_a \arctan(I p_u)\). 
-By the same analysis, this injects energy on \(D\) in negative-time, so the
-constraint dissipates energy on \(D\).
+By the same analysis, this injects energy on \(mathcal{O}\) in negative-time, so the
+constraint dissipates energy on \(mathcal{O}\).
 
@@
 \]
-Outside of \(D\), the acrobot is experiencing rotations.
+Outside of \(mathcal{O}\), the acrobot is experiencing rotations.
 We wish to find a new set of coordinates \((\alpha,\mu)\) and perform a similar
@@
 Hence, we choose \(\mu\) to be this intersection point.
-Note that \(D\) intersects the \(p_u\) axis when 
+Note that \(mathcal{O}\) intersects the \(p_u\) axis when 
 \(\mu^2 = 60m^2 g l^3\), so we must have 
@@
         is \((q_u,p_u) = (\pi,0)\), which is also the boundary of the
-        oscillation domain \(D\).}
+        oscillation domain \(mathcal{O}\).}
     \label{fig:acrobot-rot-W}
@@
 % homoclinic orbit connecting (pi,0) to (-pi,0). We need to show that almost
-% every initial condition starting in \(D\) will NOT converge to (pi,0), and
+% every initial condition starting in \(mathcal{O}\) will NOT converge to (pi,0), and
 % therefore exit D and enter W.

### 2020-11-15 latex/acrobot.tex
First editing pass of rotation for acrobot

@@
 To prove this intuitive argument rigorously, we define the Poincar\'{e} section
-\(P(\mu_0) := \hat{\mu}(\pi,\mu_0)\), which expands into
-\[
-    P(\mu_0) = \mu_0 + I \hat{\mu}_1(\pi,\mu_0) + R(\pi,I)
+\(P_\mathcal{O}(\mu_0) := \hat{\mu}(\pi,\mu_0)\), which expands into
+\[
+    P_\mathcal{O}(\mu_0) = \mu_0 + I \hat{\mu}_1(\pi,\mu_0) + R(\pi,I)
     .
@@
 We conclude that, for all \(m\), \(g\), \(l\), \(\bar{q}_a\), and for small
-enough \(I\), the Poincar\'{e} section \(P(\mu_0)\) is positive, monotonically
-increasing, and \(\lim \limits_{\mu_0 \to \pi} P(\mu_0) = \infty\).
-
-Because \(\dot{\alpha} > 0\), any orbit which begins and remains in \(\mathcal{O}\)
+enough \(I\), the Poincar\'{e} section \(P_\mathcal{O}(\mu_0)\) is positive,
+monotonically increasing, and 
+\(\lim \limits_{\mu_0 \to \pi} P_\mathcal{O}(\mu_0) = \infty\).
+
+% Empty proof environment to put QED square on the right of the page
+\begin{proof}[\unskip\nopunct]
+\end{proof}
+
+\subsection*{Perturbation Analysis for Rotations}
+The rotation region for the nominal pendulum is the set
+\[
+    \mathcal{R} := \left\{ (q_u,p_u) \in \Gamma \mid E(q_u,p_u) > E(\pi,0)\right\}
+    .
+\]
+We wish to find a new set of coordinates \((\beta,\rho)\) for the acrobot
+and perform a similar analysis to what we did for oscillations.
+In the oscillation region \(\mathcal{O}\) we showed that the nominal pendulum's
+energy (and hence, its orbit) is uniquely determined by the intersection point
+\(\mu\) with the \(q_u\)-axis; 
+when the nominal pendulum is rotating, the energy is instead determined by the
+intersection \(\rho\) on the \(p_u\)-axis.
+Note that the boundary of \(\mathcal{O}\) intersects the \(p_u\) axis when 
+\(p_u^2 = 60m^2 g l^3\); 
+hence, for a rotation we must have \(\rho > \sqrt{60 m^2 g l^3}\) if the
+orbit has momentum \(p_u > 0\) 
+or \(\rho < -\sqrt{60 m^2 g l^3}\) if the orbit has momentum \(p_u < 0\).
+The energy level set associated with \(\rho\) is
+\[
+    \left\{(q_u,p_u) \in \Gamma \mid E(q_u,p_u) = \frac{\rho^2}{10ml^2}\right\}
+    ,
+\]
+which gives the relationship
+\begin{equation}\label{eqn:rotation-pu2}
+    \frac{p_u^2}{10m l^2} + 30mgl(1 - c_u) = \frac{\rho^2}{10 ml^2}
+    .
+\end{equation}
+On this level set, \(q_u\) takes all values on \(\mathbb{S}^1\), so our angle of
+rotation is uniquely parameterized by \(\beta = q_u\).
+Since \(\rho\) does not change sign along the rotation, we have the smooth
+relationship
+\begin{align}\label{eqn:rotation-Tinv}
+    q_u &= \beta
+    , \\
+    p_u &= \sign{\rho}\sqrt{\rho^2 - 30 m^2 g l^3\left(1 - c_\beta \right)}
+    .
+\end{align}
+
+Inverting this relationship gives our (smooth) transformation
+\begin{equation*}
+    T : \mathbb{S}^1 \times \R \rightarrow 
+    \mathbb{S}^1 \times ]-\infty,-60m^2gl^3[\, \cup \, ]60m^2gl^3,\infty[ 
+    ,
+\end{equation*}
+where \(T(q_u,p_u) = (\beta,\rho)\) is defined by
+\begin{align}\label{eqn:rotation-T}
+    \beta &= q_u
+    , \\
+    \rho &= \sign{p_u}\sqrt{p_u^2 + 30 m^2 g l^3 \left(1 - c_u \right)}
+    .
+\end{align}
+
+Computing the acrobot's constrained dynamics in \((\beta,\rho)\)-coordinates and
+setting \(I = 0\) yields the dynamics of the nominal pendulum.
+MATLAB evaluates those dynamics as
+\begin{align}\label{eqn:acrobot-rot-mu-dot-nom}
+    \dot{\rho}_n &= 0
+    , \\
+    \label{eqn:acrobot-rot-alpha-dot-nom}
+    \dot{\beta}_n &=  \sign{\rho} 
+    \frac{\sqrt{\rho^2 - 30m^2gl^3(1 - c_\beta)}}{5ml^2}
+    .
+\end{align}
+As expected, \(\dot{\beta}_n\) does not change sign on \(\mathcal{R}\) because
+a rotating pendulum does not change direction:
+if \(\rho_n > 0\), the orbit traces a curve going from \(\beta_n = -\pi\) to 
+\(\beta_n = \pi\); if \(\rho_n < 0\) the curve goes in the reverse direction.
+By continuity of the acrobot's constrained dynamics 
+\((\dot{\beta},\dot{\rho})\), we must also have 
+\(\sign{\dot{\beta}} = \sign{\rho_0}\) for small enough \(I\).
+Hence, we can use \(\beta\) as a new ``time" variable on \(\mathcal{R}\) 
+by performing a time transformation \(t = \tau(\beta)\).
+This produces the time-scaled system \(\hat{\rho}(\beta) = \rho(\tau(\beta))\)
+with dynamics
+\[
+    \begin{cases}
+        \dot{\hat{\rho}} = \frac{\dot{\rho}}{\dot{\beta}}
+        , \\
+        \hat{\rho}(0) = \rho_0
+        .
+    \end{cases}
+\]
+For a perturbation approach, be expand the solution of these dynamics into
+\begin{equation}\label{eqn:acrobot-rhohat-approx}
+    \hat{\rho}(\beta,\rho_0,I) = \hat{\rho}_0(\beta,\rho_0) + I
+    \hat{\rho}_1(\beta,\rho_0) + R(\beta,\rho_0,I)
+    ,
+\end{equation}
+where \(R(\beta,\rho_0,I)\) is \(O(I^2)\).
+
+From ~\eqref{eqn:khalil-perturbation-nominal} we know
+\(\hat{\rho}_0\) is the solution to the nominal system
+\[
+    \begin{cases} 
+        \dot{\hat{\rho}}_0 = 0
+        , \\
+        \hat{\mu}_0(0) = \rho_0
+        ,
+    \end{cases}
+\]
+which means \(\hat{\mu}_0(\beta,\rho_0) \equiv \rho_0\) for all \(\beta\).
+Likewise, ~\eqref{eqn:khalil-perturbation-firstorder} 
+gives us the following dynamics for \(\hat{\rho}_1\):
+\begin{equation}\label{eqn:acrobot-rho1-dot}
+  \begin{cases}
+    \dot{\hat{\rho}}_1 =
+    \frac{5m^2 g l^3 \bar{q}_a \left(
+        m^2gl^3\left(18s_\beta^2 + 30c_\beta(1 - c_\beta)\right)
+        - c_\beta\rho_0^2
+    \right)}{
+    |\rho_0|\sqrt{\rho_0^2 - 30m^2gl^3(1 - c_\beta)}
+    }
+    =: b(\beta,\rho_0)
+     , \\
+     \hat{\rho}_1(0) = 0
+     .
+ \end{cases}
+\end{equation}
+The solution to ~\eqref{eqn:acrobot-rho1-dot} is given by
+\[
+    \hat{\rho}_1(\beta,\rho_0) = \int \limits_0^\beta b(\sigma,\rho_0)d\sigma
+    .
+\]
+
+Supposing we initialize the acrobot at \((\beta,\rho) = (0,\rho_0)\),
+then one full rotation amounts to \(\beta\) traversing \(2\pi\)rad in a clockwise
+direction (\ie \(\beta\) going from \(0\) to \(\sign{\rho_0}2\pi\)).
+We define a Poincar\'{e} section as the point reached on the \(p_u\) axis after
+every rotation:
+\[
+    P(\rho_0) = \rho_0 + I \hat{\rho}_1(\sign{\rho_0}2\pi, \rho_0)
+      + R(2\pi,I)
+    .
+\]
+If \(|P(\rho_0)| > |\rho_0|\), then the rotation has
+driven the orbit further from the origin.
+In other words, \(|p_u|\) has increased across the rotation.
+If this is true for any \(\rho_0\), the acrobot is gaining momentum, which in
+turn means it is gaining energy.
+
+The function \(b(\beta,\rho_0)\) is even in both its inputs, and is 
+\(2\pi\)-periodic in \(\beta\).
+Hence, we have 
+\begin{align*}
+    \hat{\rho}_1(\sign{\rho_0}2\pi, \rho_0) &=
+    \int \limits_0^{\sign{\rho_0}2\pi} b(\sigma,\rho_0)d\sigma
+    , \\
+     &=\sign{\rho_0}\int\limits_0^{2\pi}b(\sigma,\rho_0)d\sigma
+     , \\
+     &= \sign{\rho_0} \hat{\rho}_1(2\pi,\rho_0)
+     .
+\end{align*}
+Furthermore, \(\hat{\rho}_1(2\pi,\rho_0) = \hat{\rho}_1(2\pi,|\rho_0|)\).
+This implies that
+\[
+    |P(\rho_0)| = \left|\sign{\rho_0}\left(
+        |\rho_0| + I \hat{\rho}_1(2\pi,|\rho_0|)\right)
+    \right|
+    = P(|\rho_0|)
+    .
+\]
+Hence, if \(P(|\rho_0|) > |\rho_0|\), then a single rotation starting at \(\rho_0\)
+will drive the acrobot's orbit further from the origin.
+Equivalently, one need simply guarantee that \(\hat{\rho}_1(2\pi,\rho_0) > 0\)
+for all \(\rho_0 > \sqrt{60m^2gl^3}\).
+
+Unfortunately, we cannot adimensionalize
+\(b(\beta,\rho_0)\) like we did for oscillations. 
+Hence, we let \(U = ]\sqrt{60m^2gl^3},\bar{\rho}[\) and assume that 
+\(\hat{\rho}_1(2\pi,\rho_0) > 0 \) for all \(\rho_0 \in U\).
+We conclude that, for small enough \(I\), the Poincar\'{e} section 
+\(P(\rho_0)\) is monotonically diverging from 0 on \(U\)
+and the acrobot is gaining momentum.
+
+\subsection*{Energy Gain}
+We will prove the acrobot gains energy on the set 
+\[
+    E_{\bar{\mu}} := 
+    \left\{ (q_u,p_u) \in \Gamma \mid E(q_u,p_u) < E(0,\bar{\rho}) \right\}
+    .
+\]
+% Can we use the jacobian? The spectrum is hard to find.
+
+Recall that on \(\mathcal{O}\) we have \(\dot{\alpha} > 0\). 
+Hence, any orbit which begins and remains in \(\mathcal{O}\)
 must spiral clockwise around the origin until it intersect the \(q_u\) axis.
-Once it does, the Poincar\'{e} argument demonstrates it must spiral away
-from the origin; hence, there are no closed orbits and the origin is a repeller.
+Once it does, the Poincar\'{e} section \(P_\mathcal{O}(\cdot)\) 
+demonstrates the orbit must spiral away from the origin; 
+this implies there cannot be any closed orbits on \(\mathcal{O}\), and that the
+origin is a repeller.
 
 Once the orbit reaches an intersection point \(\mu_0\) on the \(q_u\)-axis such
-that \(P(\mu_0) > \pi\), the orbit will no longer return to the \(q_u\)-axis; it
-will escape \(\mathcal{O}\) entirely.
-This must always happen in finite time, which implies the orbit must also escape
-any compact set in \(mathcal{O}\) in finite time. 
+that \(P_\mathcal{O}(\mu_0) > \pi\), the orbit will no longer return to the
+\(q_u\)-axis. 
+Instead, it will escape \(\mathcal{O}\) entirely.
+This must always happen in a finite number of iterations of the 
+Poincar\'{e} section, which means the orbit escapes \(\mathcal{O}\) in finite
+time., which implies the orbit must also escape
+any compact set in \(\mathcal{O}\) in finite time. 
 By Definition \ref{defn:energy-gain}, the VNHC is injecting energy into the
@@
 \(q_a = -\bar{q}_a \arctan(I p_u)\). 
-By the same analysis, this injects energy on \(mathcal{O}\) in negative-time, so the
-constraint dissipates energy on \(mathcal{O}\).
+By the same analysis, this injects energy on \(\mathcal{O}\) in negative-time, so the
+constraint dissipates energy on \(\mathcal{O}\).
 
@@
 hysteresis to avoid infinite switching.
-% Empty proof environment to put QED square on the right of the page
-\begin{proof}[\unskip\nopunct]
-\end{proof}
-
-\subsection*{Perturbation Analysis for Rotations}
-Recall that the domain of oscillations for the nominal pendulum of the acrobot
-is 
-\[
-    D := \left\{ (q_u,p_u) \in \Gamma | E(q_u,p_u) < E(\pi,0)\right\}
-    .
-\]
-Outside of \(mathcal{O}\), the acrobot is experiencing rotations.
-We wish to find a new set of coordinates \((\alpha,\mu)\) and perform a similar
-analysis.
-During oscillations we showed that the energy of oscillations is determined by
-the intersection with the \(q_u\)-axis; 
-when the acrobot is rotating, the energy is instead determined by the
-intersection on the \(p_u\)-axis.
-Hence, we choose \(\mu\) to be this intersection point.
-Note that \(mathcal{O}\) intersects the \(p_u\) axis when 
-\(\mu^2 = 60m^2 g l^3\), so we must have 
-\(\mu > \sqrt{60 m^2 g l^3}\) if the rotation intersects at \(p_u > 0\) 
-or \(\mu < -\sqrt{60 m^2 g l^3}\) if the rotation intersects at \(p_u < 0\).
-The energy level set associated with \(\mu\) is
-\[
-    \left\{(q_u,p_u) \in \Gamma \mid E(q_u,p_u) = \frac{\mu^2}{10ml^2}\right\}
-    ,
-\]
-which gives the relationship
-\begin{equation}\label{eqn:rotation-pu2}
-    \frac{p_u^2}{10m l^2} + 30mgl(1 - c_u) = \frac{\mu^2}{10 ml^2}
-    .
-\end{equation}
-On this level set, \(q_u\) ranges fully between \(]-pi,\pi]\), so our angle of
-rotation is uniquely parameterized by \(\alpha = q_u\).
-Since \(\mu\) does not change sign along the rotation, we have the smooth
-relationship
-\begin{align}\label{eqn:rotation-Tinv}
-    q_u &= \alpha
-    , \\
-    p_u &= \sign{\mu}\sqrt{\mu^2 - 30 m^2 g l^3\left(1 - c_\alpha \right)}
-    .
-\end{align}
-
-We have thus found a transformation \(T\inv(\alpha,\mu) = (q_u,p_u)\), with
-inverse 
-\begin{align}\label{eqn:rotation-T}
-    \alpha &= q_u
-    , \\
-    \mu &= \sign{p_u}\sqrt{p_u^2 + 30 m^2 g l^3 \left(1 - c_u \right)}
-    .
-\end{align}
-
-From this point on, we do the exact same procedure as with oscillations.
-MATLAB symbolic operations evaluate the nominal system's dynamics as
-\begin{align}\label{eqn:acrobot-rot-mu-dot-nom}
-    \dot{\mu}_n &= 0
-    , \\
-    \label{eqn:acrobot-rot-alpha-dot-nom}
-    \dot{\alpha}_n &=  \sign(\mu_0) 
-    \frac{\sqrt{\mu^2 - 30m^2gl^3(1 - c_\alpha)}}{5ml^2}
-    .
-\end{align}
-As is expected, \(\dot{\alpha}_n\) does not change sign because the acrobot is
-rotating;
-if \(\mu_0 > 0\), the orbit traces a curve going from \(\alpha = -\pi\) to 
-\(\alpha = \pi\), and if \(\mu_0 < 0\) the curve goes in the reverse direction.
-For small enough \(I > 0\) we must also have 
-\(\dot{\alpha} = \sign{\mu_0}\) and can hence use \(\alpha\) as a new time
-variable.
-
-Take the time-scaled system \(\hat{\mu}(\alpha,I)\) with dynamics 
-\(\dot{\hat{\mu}}(\alpha) = \frac{\dot{\mu}}{\dot{\alpha}}\) and initial
-state \(\mu_0\).
-We expand the solution of this into the same form as
-~\eqref{eqn:acrobot-muhat-approx} where once again we have
-\(\hat{\mu}_0(\alpha) \equiv \mu_0\).
-The dynamics for \(\hat{\mu}_1\) are
-\begin{equation}\label{eqn:acrobot-rot-mu1-dot}
-    \dot{\hat{\mu}}_1 =
-    \frac{5m^2 g l^3 \bar{q}_a \left(
-        m^2gl^3\left(18s_\alpha^2 + 30c_\alpha(1 - c_\alpha)\right)
-        - c_\alpha\mu_0^2
-    \right)}{
-    |\mu_0|\sqrt{\mu_0^2 - 30m^2gl^3(1 - c_\alpha)}
-    }
-    =: b(\alpha,\mu_0)
-    .
-\end{equation}
-The solution to ~\eqref{eqn:acrobot-rot-mu1-dot} is given by
-\[
-    \hat{\mu}_1(\alpha,\mu_0) = \int \limits_0^\alpha b(\sigma,\mu_0)d\sigma
-    .
-\]
-Since one full rotation amounts to \(\alpha\) traversing \(2\pi\)rad in a clockwise
-direction (\ie \(\alpha\) going from \(0\) to \(\sign{\mu_0}2\pi\)), 
-we define a Poincar\'{e} section as the point reached on the \(p_u\) axis after
-every rotation:
-\[
-    P(\mu_0) = \mu_0 + I \hat{\mu}_1(\sign{\mu_0}2\pi, \mu_0)
-    .
-\]
-If \(|P(\mu_0)| > |\mu_0|\), then the rotation has
-driven the orbit further from the origin.
-In other words, \(|p_u|\) has increased across the rotation.
-If this is true for any \(\mu_0\), the acrobot is gaining energy.
-
-The function \(b(\alpha,\mu_0)\) is even in both its inputs, and is 
-\(2\pi\)-periodic in \(\alpha\).
-Hence, we have 
-\begin{align*}
-    \hat{\mu}_1(\sign{\mu_0}2\pi, \mu_0) &=
-    \int \limits_0^{\sign{\mu_0}2\pi} b(\sigma,\mu_0)d\sigma
-    , \\
-     &=\sign{\mu_0}\int\limits_0^{2\pi}b(\sigma,\mu_0)d\sigma
-     , \\
-     &= \sign{\mu_0} \hat{\mu}_1(2\pi,\mu_0)
-     .
-\end{align*}
-Furthermore, \(\hat{\mu}_1(2\pi,\mu_0) = \hat{\mu}_1(2\pi,|\mu_0|)\).
-This implies that
-\[
-    |P(\mu_0)| = \left|\sign{\mu_0}\left(
-        |\mu_0| + I \hat{\mu}_1(2\pi,|\mu_0|)\right)
-    \right|
-    = P(|\mu_0|)
-    .
-\]
-Therefore, if \(P(|\mu_0|) > |\mu_0|\), then the acrobot is gaining energy.
-Showing that \(\hat{\mu}_1(2\pi,\mu_0) > 0\) for \(\mu_0 > \sqrt{60m^2gl^3}\) is
-sufficient for this condition to hold.
-
-Unlike in the case of oscillations, we cannot adimensionalize
-\(b(\alpha,\mu_0)\) and prove that this condition will be true for all acrobots. 
-Let \(U = ]\sqrt{60m^2gl^3},\bar{\mu}[\), and assume that 
-\(\hat{\mu}_1(2\pi,\mu_0) > 0 \) for all \(\mu_0 \in U\).
-If this is true, then by perturbation analysis we can claim that there exists a
-small enough \(I\) such that the Poincar\'{e} section \(P(\mu_0)\) is monotonic
-on \(U\).
+
+% TODO: W and D's boundary is two elements: the point (pi,0) and the curve of the
+% homoclinic orbit connecting (pi,0) to (-pi,0). We need to show that almost
+% every initial condition starting in \(mathcal{O}\) will NOT converge to (pi,0), and
+% therefore exit D and enter W.
 
@@
 \[
-    W := \left\{(q_u,p_u) \in \Gamma \mid |\mu(q_u,p_u)| \in U \right\}
+    W := \left\{(q_u,p_u) \in \Gamma \mid |\rho(q_u,p_u)| \in U \right\}
     ,
@@
 as in Figure \ref{fig:acrobot-rot-W}.
-Monotonicity of \(P(\mu_0)\) means there are no closed orbits of 
+Monotonicity of \(P(\rho_0)\) means there are no closed orbits of 
 \((q_u,p_u)\) on \(W\). 
@@
     \caption{The set \(W\) of points which have pseudo-radius 
-        \(|\mu| \in ]\sqrt{60m^2gl^3}, \bar{\mu}[\). The inner boundary of \(W\)
+        \(|\rho| \in ]\sqrt{60m^2gl^3}, \bar{\rho}[\). The inner boundary of \(W\)
         is \((q_u,p_u) = (\pi,0)\), which is also the boundary of the
@@
 
-We conclude that \(|\mu(q_u,p_u)|\) will eventually reach \(\bar{\mu}\). 
+We conclude that \(|\rho(q_u,p_u)|\) will eventually reach \(\bar{\rho}\). 
 We cannot claim this fits the definition  of energy injection 
@@
 
-If we set \(I < 0\), we have \(P(|\mu_0|) < |\mu_0|\).
+If we set \(I < 0\), we have \(P(|\rho_0|) < |\rho_0|\).
 Hence, \(|p_u|\) is decreasing every time the orbit hits the \(p_u\)-axis, so
@@
 
-\subsection*{Connecting Oscillations and Rotations}
-To complete this proof, we will prove the acrobot gains energy on the set 
-\[
-    E_{\bar{\mu}} := 
-    \left\{ (q_u,p_u) \in \Gamma \mid E(q_u,p_u) < E(0,\bar{\mu}) \right\}
-    .
-\]
-% TODO: W and D's boundary is two elements: the point (pi,0) and the curve of the
-% homoclinic orbit connecting (pi,0) to (-pi,0). We need to show that almost
-% every initial condition starting in \(mathcal{O}\) will NOT converge to (pi,0), and
-% therefore exit D and enter W.
-
 \section{Experimental Results}

### 2020-11-15 latex/acrobot.tex
Another edit of the acrobot proof

@@
 \end{equation*}
-%        &= \frac{1}{ml^2(2-c_a^2)} 
-%            \left( p_u - 
-%                (1+c_a)\frac{(1+c_a)(1+I^2 p_u^2)p_u - m^2gl^3\bar{q}_a I (2-c_a^2)(2s_u + s_{ua})
-%                }{ml^2(3+2c_a)(1+I^2 p_u^2)}
-%            \right) 
-%        ,
-%\end{align*}
-which can be simplified to 
+which can be simplified into 
 \begin{equation*}
@@
     constraint manifold is \(\Gamma = \mathbb{S}^1 \times \R\).
-    Let \(E\) be the nominal energy when \(I = 0\):
+    Let 
     \[
         E(q_u,p_u) := \frac{p_u^2}{10ml^2} + 3(1 - \cos(q_u))
-        .
-    \]
+        ,
+    \]
+    be the energy of the simple pendulum obtained by setting \(I = 0\).
 \begin{enumerate}
-    \item There always exists \(I \in \R_{>0}\) small enough that 
+    \item There always exists \(I > 0\) small enough that 
     ~\eqref{eqn:acrobot-constraint} injects energy into the acrobot on the set
@@
     \]
-    If instead \(I \in \R_{<0}\), the VNHC dissipates energy on \(\mathcal{O}\).
+    If instead \(I < 0\), the VNHC dissipates energy on \(\mathcal{O}\).
 \item Define \(b : \mathbb{S}^1 \times \R_{> 0} \rightarrow \R\) by
     \[
-        b(\alpha,\mu_0) := 
+        b(\alpha,\rho_0) := 
         \frac{5m^2 g l^3 \bar{q}_a \left(
             m^2gl^3\left(18s_\alpha^2 + 30c_\alpha(1 - c_\alpha)\right)
-            - c_\alpha\mu_0^2
+            - c_\alpha\rho_0^2
         \right)}{
-        |\mu_0|\sqrt{\mu_0^2 - 30m^2gl^3(1 - c_\alpha)}
+        |\rho_0|\sqrt{\rho_0^2 - 30m^2gl^3(1 - c_\alpha)}
         }
@@
     \]
-    and fix \(\bar{\mu} > 60m^2gl^3\).
+    and fix \(\bar{\rho} > \sqrt{60m^2gl^3}\).
     Suppose \(\int \limits_{0}^{2\pi} b(\sigma,\mu_0)d\sigma > 0\) for all 
-    \(\mu_0 \in ]60m^2gl^3, \bar{\mu}[\).
-    Then there exists \(I \in \R_{> 0}\) small enough that
+    \(\rho_0 \in ]\sqrt{60m^2gl^3}, \bar{\rho}[\).
+    Then there exists \(I > 0\) small enough that
     ~\eqref{eqn:acrobot-constraint} injects energy into the acrobot on
@@
         \Omega := \left\{(q_u,p_u) \in \Gamma 
-            \mid E(q_u,p_u) < E(0,\bar{\mu})\right\}
-        .
-    \]
-    If instead \(I \in \R_{<0}\), the VNHC dissipates energy on \(\Omega\).
+            \mid E(q_u,p_u) < E(0,\bar{\rho})\right\}
+        .
+    \]
+    If instead \(I < 0\), the VNHC dissipates energy on \(\Omega\).
 \end{enumerate}
@@
 
+Suppose we initialize the acrobot at \((q_0,p_0) \in \mathcal{O}\).
+Since on \(\mathcal{O}\) we have \(\dot{\alpha} > 0\), the orbit of the acrobot
+in \(\Gamma\) spirals clockwise around the origin; 
+if the spiral does not exit \(\mathcal{O}\), it must intersect the \(q_u\) axis.
+Once it does, the Poincar\'{e} section \(P_\mathcal{O}(\cdot)\) tells us 
+the orbit must spiral away from the origin. 
+This implies there cannot be any closed orbits in \(\mathcal{O}\) and that the
+origin is a repeller.
+
+Once the orbit reaches an intersection point \((\mu_0,0)\) on the \(q_u\)-axis such
+that \(P_\mathcal{O}(\mu_0) > \pi\), the orbit will no longer return to the
+\(q_u\)-axis. 
+Instead, it will escape \(\mathcal{O}\) entirely, which means it must 
+escape any compact subset of \(\mathcal{O}\) in
+finite time. 
+We conclude that the acrobot is gaining energy on \(\mathcal{O}\).
+
+Setting \(I < 0\), the Poincar\'{e} section \(P_\mathcal{O}(\mu_0)\) is now
+decreasing.
+Hence, the orbit in \(\mathcal{O}\) spirals towards the origin, which means the
+constraint is dissipating energy on \(\mathcal{O}\).
+
 % Empty proof environment to put QED square on the right of the page
@@
 \[
-    P(\rho_0) = \rho_0 + I \hat{\rho}_1(\sign{\rho_0}2\pi, \rho_0)
+    P_\mathcal{R}(\rho_0) = \rho_0 + I \hat{\rho}_1(\sign{\rho_0}2\pi, \rho_0)
       + R(2\pi,I)
@@
 \]
-If \(|P(\rho_0)| > |\rho_0|\), then the rotation has
+If \(|P_\mathcal{R}(\rho_0)| > |\rho_0|\), then the rotation has
 driven the orbit further from the origin.
@@
 \[
-    |P(\rho_0)| = \left|\sign{\rho_0}\left(
+    |P_\mathcal{R}(\rho_0)| = \left|\sign{\rho_0}\left(
         |\rho_0| + I \hat{\rho}_1(2\pi,|\rho_0|)\right)
     \right|
-    = P(|\rho_0|)
-    .
-\]
-Hence, if \(P(|\rho_0|) > |\rho_0|\), then a single rotation starting at \(\rho_0\)
-will drive the acrobot's orbit further from the origin.
-Equivalently, one need simply guarantee that \(\hat{\rho}_1(2\pi,\rho_0) > 0\)
-for all \(\rho_0 > \sqrt{60m^2gl^3}\).
+    = P_\mathcal{R}(|\rho_0|)
+    .
+\]
+Hence, if \(P_\mathcal{R}(|\rho_0|) > |\rho_0|\) (or equivalently if
+\(\hat{\rho}_1(2\pi,|\rho_0|) > 0\))
+the phase \((q_u(t),p_u(t))\) initialized at \((0,\rho_0)\) 
+will intersect the \(p_u\) axis at a point \(p > \rho_0\).
+If we could guarantee that \(\hat{\rho}_1(2\pi,\rho_0) > 0\)
+for all \(\rho_0 > \sqrt{60m^2gl^3}\), then the acrobot would be spiraling away
+from the origin on \(\mathcal{R}\).
 
@@
 \(b(\beta,\rho_0)\) like we did for oscillations. 
-Hence, we let \(U = ]\sqrt{60m^2gl^3},\bar{\rho}[\) and assume that 
-\(\hat{\rho}_1(2\pi,\rho_0) > 0 \) for all \(\rho_0 \in U\).
+Using the assumption , we have that 
+\(\hat{\rho}_1(2\pi,\rho_0) > 0 \) for all 
+\(\rho_0 \in ]\sqrt{60m^2gl^3},\rho_0[\).
 We conclude that, for small enough \(I\), the Poincar\'{e} section 
-\(P(\rho_0)\) is monotonically diverging from 0 on \(U\)
-and the acrobot is gaining momentum.
-
-\subsection*{Energy Gain}
+\(P_\mathcal{R}(\rho_0)\) is monotonically increasing when 
+\(0 < \rho_0 < \bar{\rho}\) and
+monotonically decreasing when \(-\bar{\rho} < \rho_0 < 0\).
+
+\subsection*{Proving Energy Gain}
 We will prove the acrobot gains energy on the set 
 \[
-    E_{\bar{\mu}} := 
-    \left\{ (q_u,p_u) \in \Gamma \mid E(q_u,p_u) < E(0,\bar{\rho}) \right\}
+    \Omega := \left\{ (q_u,p_u) \in \Gamma 
+     \mid E(q_u,p_u) < E(0,\bar{\rho}) \right\}
     .
@@
 % Can we use the jacobian? The spectrum is hard to find.
-
-Recall that on \(\mathcal{O}\) we have \(\dot{\alpha} > 0\). 
-Hence, any orbit which begins and remains in \(\mathcal{O}\)
-must spiral clockwise around the origin until it intersect the \(q_u\) axis.
-Once it does, the Poincar\'{e} section \(P_\mathcal{O}(\cdot)\) 
-demonstrates the orbit must spiral away from the origin; 
-this implies there cannot be any closed orbits on \(\mathcal{O}\), and that the
-origin is a repeller.
-
-Once the orbit reaches an intersection point \(\mu_0\) on the \(q_u\)-axis such
-that \(P_\mathcal{O}(\mu_0) > \pi\), the orbit will no longer return to the
-\(q_u\)-axis. 
-Instead, it will escape \(\mathcal{O}\) entirely.
-This must always happen in a finite number of iterations of the 
-Poincar\'{e} section, which means the orbit escapes \(\mathcal{O}\) in finite
-time., which implies the orbit must also escape
-any compact set in \(\mathcal{O}\) in finite time. 
-By Definition \ref{defn:energy-gain}, the VNHC is injecting energy into the
-acrobot on \(\mathcal{O}\).
-
-If we have \(I < 0\), it is equivalent to using the constraint 
-\(q_a = -\bar{q}_a \arctan(I p_u)\). 
-By the same analysis, this injects energy on \(\mathcal{O}\) in negative-time, so the
-constraint dissipates energy on \(\mathcal{O}\).
-
-A hybrid supervisor that flips the sign of \(I\) based on the current
-value of \(\mu\) will stabilize any energy level of oscillation.
-As was the case in the VLP, this supervisor should implement some kind of
-hysteresis to avoid infinite switching.
-
 % TODO: W and D's boundary is two elements: the point (pi,0) and the curve of the
@@
 
-Define the set of points in \(\Gamma\) which have pseudo-radius in \(U\) by
+Define the set of points in \(\mathcal{R}\) which have pseudo-radius in \(U\) by
 \[
@@
 
+A hybrid supervisor that flips the sign of \(I\) based on the current
+value of \(\mu\) will stabilize any energy level of oscillation.
+As was the case in the VLP, this supervisor should implement some kind of
+hysteresis to avoid infinite switching.
+
 \section{Experimental Results}

### 2020-11-17 latex/acrobot.tex
Minor edits to wording up to theorem statement of acrobot

@@
 
-Controlling the acrobot is a nontrivial task, since it is not feedback
+Controlling the acrobot is a nontrivial task because it is not feedback
 linearizable \cite{nonlinear_controllers_nonintegrable_acrobot}. 
@@
 \(q = (q_u,q_a) \in \mathbb{S}^1 \times \mathbb{S}^1\). 
-The acrobot has inertia matrix \(mathcal{O}\), potential function \(P\) (with respect to
+The acrobot has inertia matrix \(D\), potential function \(P\) (with respect to
 the horizontal bar), and input matrix \(B\)  given as
@@
 While this is the most general representation of an acrobot, the dynamics
-become unwieldy.
+are unwieldy.
 To make rigorous analysis of these dynamics more tractable, we begin by assuming
@@
 \]
-is of full rank 1 on the constraint manifold \(\Gamma\).
+is of full rank \(1\) on the constraint manifold \(\Gamma\).
 Given that
@@
 \end{equation}
-This is full rank on \(\Gamma\) if and only if the numerator does not
+This is full rank if and only if the numerator does not
 change sign.
-The following proposition provides a sufficient condition for this to happen.
+The following proposition provides a sufficient condition for regularity.
 
@@
     ~\eqref{eqn:regularity-matrix-acrobot} is strictly positive for all values
-    of \(q_a\), and hence is of full rank 1.
+    of \(q_a\), and hence is full rank everywhere on the constraint manifold.
     By Theorem \ref{thm:vnhc-regularity}, \(h\) is a regular VNHC of order 1.
@@
 emulates true human motion, \citeauthor{xingbo_thesis}
-does not provide an analytical proof that the acrobot will gain energy.
+does not provide analytical proof that the acrobot will gain energy.
 His lack of analysis is tied to the fact that the constrained
@@
 
-To design this constraint, let us begin by examining a person on a seated swing.
+To design this constraint, let us begin (perhaps unexpectedly) by examining a person on
+a seated swing.
 The person extends their legs when the swing moves forwards, and retracts their
-legs when the swing is moving backwards.
+legs when the swing moves backwards.
 As the swing gains speed, the person leans their body back while
@@
 
-The acrobot's legs are rigid rods which cannot retract, so we emulate the person
-on a swing by pivoting the legs in the direction of motion. 
+The acrobot's legs are rigid rods which cannot retract; so we emulate the person
+on a swing by pivoting the legs toward the direction of motion. 
 To account for how a person leans back at higher speeds, the legs should pivot to an
 angle proportional to the swing's speed.
-One VNHC which emulates this process is \(q_a = \bar{q}_a\arctan( I p_u)\), as
-in Figure \ref{fig:qa-arctan}.
-Here, \(\bar{q}_a \in ]0,\frac{2 Q_a}{\pi}]\)
-and \(I \in \R_{>0}\) is a fixed control parameter.
-
-This constraint does not perfectly recreate a gymnast's motion -- during
-a giant the gymnast's legs are almost completely extended, while this constraint
-pivots the legs partially during rotations.
-However, the motion looks similar enough that this constraint should provide a
+Since the direction of motion is entirely determined by \(p_u\), 
+one such VNHC which emulates this process is \(q_a = \bar{q}_a\arctan( I p_u)\),
+displayed in Figure \ref{fig:qa-arctan}.
+Here, \(\bar{q}_a \in ]0,\frac{2 Q_a}{\pi}]\) and \(I \in \R_{>0}\) is a fixed
+control parameter.
+
+This constraint does not perfectly recreate giant motion, during which
+the gymnast's legs are almost completely extended -- it instead pivots the legs
+partially during rotations.
+However, the behaviour looks similar enough that the constraint should provide a
 decent foundation for injecting energy into the acrobot.
@@
 ~\eqref{eqn:acrobot-constraint} are
-\begin{align}\label{eqn:acrobot-constrained-dynamics}
+\begin{equation}\label{eqn:acrobot-constrained-dynamics}
+\left.\begin{cases}
     \dot{q}_u &= \frac{(1+I^2 p_u^2)p_u + m^2gl^3\bar{q}_a I(2s_u + s_{ua})(1+c_a) }
             {ml^2(1+I^2 p_u^2)(3+2c_a)}
-        , \\
+        \\
     \dot{p}_u &= - m g l (2s_u + s_{ua})
-    . \nonumber
-\end{align}
-
-These dynamics do not always gain energy -- we need \(|I|\) to be small enough
-to guarantee this is true.
+    .
+    \end{cases} \right|_{q_a = \bar{q}_a\arctan(Ip_u)}
+\end{equation}
+
+These dynamics do not always gain energy -- we need certain conditions on our
+control parameter \(I\) to guarantee this is true.
 
@@
     \[
-        b(\alpha,\rho_0) := 
+        b(\beta,\rho_0) := 
         \frac{5m^2 g l^3 \bar{q}_a \left(
-            m^2gl^3\left(18s_\alpha^2 + 30c_\alpha(1 - c_\alpha)\right)
-            - c_\alpha\rho_0^2
+            m^2gl^3\left(18s_\beta^2 + 30c_\beta(1 - c_\beta)\right)
+            - c_\beta\rho_0^2
         \right)}{
-        |\rho_0|\sqrt{\rho_0^2 - 30m^2gl^3(1 - c_\alpha)}
+        |\rho_0|\sqrt{\rho_0^2 - 30m^2gl^3(1 - c_\beta)}
         }
@@
     Suppose \(\int \limits_{0}^{2\pi} b(\sigma,\mu_0)d\sigma > 0\) for all 
-    \(\rho_0 \in ]\sqrt{60m^2gl^3}, \bar{\rho}[\).
+    \(\rho_0 \in \, \left]\sqrt{60m^2gl^3}, \bar{\rho}\right[\).
     Then there exists \(I > 0\) small enough that
@@
 %    Letting \(\epsilon > 0\), we need 
-%    \(I \leq \min\left\{I_1(\epsilon),I_2\right\}\) where
+%    \(I \leq \min\left\{I_1(\epsilon),I_2\}\) where
 %    \(I_1(\epsilon) = \frac{\tan(\epsilon)}{(\pi-\epsilon)}

### 2020-11-18 latex/vnhcs.tex
Updated energy gain defn and made shorthands in sty

@@
     a curve on the 2D constraint manifold 
-    \(\Gamma = \Rt{T_u} \times \R\) which we call the \((q_u,p_u)\)-plane.
+    \(\Gamma \simeq \Rt{T_u} \times \R\) which we call the \((q_u,p_u)\)-plane.
 \end{cor}
@@
 \begin{defn}\label{defn:energy-gain}
-    Let \(M = \Rt{T_1} \times \cdots \times \Rt{T_n}\) 
-    (where \(T_1, \ldots, T_n \in \{2\pi, \infty\}\)) be an
+    Let \(\mathcal{Q}\) be an
     \(n\)-dimensional generalized cylinder.
-    Let \(f : M \rightarrow \R^n\) be smooth with \(f(0) = 0\)
-    and let \(D \subset M\) be open.
+    Let \(f : \mathcal{Q} \rightarrow \mathcal{Q}\times\R^n\) be a smoth vector
+    field and let \(D \subset M\) be open.
     The system described by \(\dot{x} = f(x)\) 
-    \textit{gains energy on \(D\)} if these conditions are satisfied:
-    \begin{enumerate}
-        \item If \(0 \in D\), then \(0\) is a repeller 
-         (or equivalently, it is asymptotically stable in negative time).
-      \item For all compact sets \(K \subset D\) and for almost every initial
+    \textit{gains energy on \(D\)} if, 
+    for all compact sets \(K \subset D\) and for almost every initial
     condition \(x(0) \in K\), there exists \(T > 0\) such
     that \(x(t) \notin K \, (\forall t > T)\).
-    \end{enumerate}
     The system \textit{loses energy on \(D\)} if it gains energy in
@@
 
-The second property of Definition \ref{defn:energy-gain} allows the system
-to have unstable equilibria, but not limit cycles nor closed orbits.
+Any system satisfying Definition \ref{defn:energy-gain}
+can have unstable equilibria on \(D\), but not limit cycles nor closed orbits.
 The next definition ties this notion of energy gain to VNHCs.

### 2020-11-18 latex/acrobot.tex
Updated \mathbb{S}^1 to \Sone, made Gamma \simeq \SxR instead of eq.

@@
 to their own center of mass).
-Let \(q_u \in \mathbb{S}^1\) be the shoulder angle and \(q_a \in \mathbb{S}^1\) 
+Let \(q_u \in \Sone\) be the shoulder angle and \(q_a \in \Sone\) 
 be the hip angle, where only \(q_a\) is actuated. 
 Collecting them together provides the configuration
-\(q = (q_u,q_a) \in \mathbb{S}^1 \times \mathbb{S}^1\). 
+\(q = (q_u,q_a) \in \Sone \times \Sone\). 
 The acrobot has inertia matrix \(D\), potential function \(P\) (with respect to
@@
 \(\mathcal{Q} = \mathcal{Q}_u \times \mathcal{Q}_a 
-:= \mathbb{S}^1 \times \mathbb{S}^1\), and
+:= \Sone \times \Sone\), and
 \(\mathcal{P} = \mathcal{P}_u \times \mathcal{P}_a
@@
     the VNHC ~\eqref{eqn:acrobot-constraint}, whose
-    constraint manifold is \(\Gamma = \mathbb{S}^1 \times \R\).
+    constraint manifold is \(\Gamma \simeq \SxR\).
     Let 
@@
     If instead \(I < 0\), the VNHC dissipates energy on \(\mathcal{O}\).
-\item Define \(b : \mathbb{S}^1 \times \R_{> 0} \rightarrow \R\) by
+\item Define \(b : \SxR_{> 0} \rightarrow \R\) by
     \[
@@
 There is a suitable change of coordinates into a pseudo-radius 
-\(\mu \in \R_{>0}\) and pseudo-angle \(\alpha \in \mathbb{S}^1\) on the
+\(\mu \in \R_{>0}\) and pseudo-angle \(\alpha \in \Sone\) on the
 \((q_u,p_u)\)-plane.
@@
         of the simple pendulum.
-    \item \(\alpha\) is a pseudo-angle living in \(\mathbb{S}^1\) which is
+    \item \(\alpha\) is a pseudo-angle living in \(\Sone\) which is
         always increasing.
@@
 Likewise, we will find a second set of coordinates \((\beta, \rho)\) where
-\(\beta \in \mathbb{S}^1\) is always increasing, \(\rho\) is constant along
+\(\beta \in \Sone\) is always increasing, \(\rho\) is constant along
 rotations of the simple pendulum, and \(\rho\) increases for the acrobot when
@@
 \]
-where \(\Gamma = \mathbb{S}^1\times\R\) is the consraint manifold of the
+where \(\Gamma \simeq \Sone\times\R\) is the constraint manifold of the
 acrobot.
 We wish to find a transformation \(T(q_u,p_u)\) on \(\mathcal{O}\) into
-\((\alpha,\mu)\)-coordinates where \(\alpha \in \mathbb{S}^1\) is a pseudo-angle
+\((\alpha,\mu)\)-coordinates where \(\alpha \in \Sone\) is a pseudo-angle
 and \(\mu > 0\) is a pseudo-radius.
@@
 \begin{align*}
-    T : D\backslash \{(0,0)\} &\rightarrow \mathbb{S}^1 \times \, ]0,\pi[, \\
+    T : D\backslash \{(0,0)\} &\rightarrow \Sone \times \, ]0,\pi[, \\
     (q_u, p_u) &\mapsto (\alpha,\mu)
@@
 \end{equation}
-On this level set, \(q_u\) takes all values on \(\mathbb{S}^1\), so our angle of
+On this level set, \(q_u\) takes all values on \(\Sone\), so our angle of
 rotation is uniquely parameterized by \(\beta = q_u\).
@@
 \begin{equation*}
-    T : \mathbb{S}^1 \times \R \rightarrow 
-    \mathbb{S}^1 \times ]-\infty,-60m^2gl^3[\, \cup \, ]60m^2gl^3,\infty[ 
+    T : \SxR \rightarrow 
+    \Sone \times ]-\infty,-60m^2gl^3[\, \cup \, ]60m^2gl^3,\infty[ 
     ,

### 2020-11-18 latex/vlp.tex
Updated \mathbb{S}^1 to \Sone, made Gamma \simeq \SxR instead of eq.

@@
 connected to a fixed pivot by a massless rod of varying length \(l\) with angle 
-\(q \in \mathbb{S}^1\) from the vertical, as in Figure
+\(q \in \Sone\) from the vertical, as in Figure
 \ref{fig:vlp-model}. 
@@
 \(\overline{l} > \underline{l}\). The configuration of the VLP is the vector
-\(\mathbf{q} := (q,l) \in \mathbb{S}^1 \times [\underline{l},\overline{l}]\).
+\(\mathbf{q} := (q,l) \in \Sone \times [\underline{l},\overline{l}]\).
 
@@
 \end{align}
-The control input is a force \(\tau \in \mathbb{R}\) affecting the dynamics of
+The control input is a force \(\tau \in \R\) affecting the dynamics of
 \(p_l\), acting collinearly with the rod.
@@
 is always full rank.
-The constraint manifold is
-\(\Gamma = \mathbb{S}^1 \times \R\), which is parameterized by the unactuated
+The constraint manifold \(\Gamma\) is diffeomorphic to 
+\(\SxR\), which is parameterized by the unactuated
 phase \((q,p)\).
@@
    Define \(\theta := \arctan_2(p,q)\).
-   Let \(L : \mathbb{S}^1 \rightarrow [\underline{l},\overline{l}]\) be a
+   Let \(L : \Sone \rightarrow [\underline{l},\overline{l}]\) be a
    \(C^2\) function of \(\theta\).
@@
    \(h(\mathbf{q},\mathbf{p}) = l - L(\theta)\) for the VLP injects energy 
-   on the constraint manifold \(\Gamma = \mathbb{S}^1 \times \R\) if there
+   on the constraint manifold \(\Gamma \simeq \SxR\) if there
    exists \(l_\text{avg} \in [\underline{l},\overline{l}]\) such that 
@@
       \left(l_\text{avg} - L(\theta)\right)sin(2\theta) \geq 0 
-      \text{ }\forall \theta \in \mathbb{S}^1
+      \text{ }\forall \theta \in \Sone
       ,
@@
     \left(l_\text{avg} - L(\theta)\right)sin(2\theta) \leq 0 
-         \text{ }\forall \theta \in \mathbb{S}^1
+         \text{ }\forall \theta \in \Sone
          ,
@@
 The state space of~\eqref{eqn:negative-time-vlp} is 
-\(\mathbb{S}^1 \times \R\), and the system has two equilibria,
+\(\SxR\), and the system has two equilibria,
 \((q,p) = (0,0)\) and \((q,p) = (\pi,0)\). 
@@
 The function \(E_\text{avg}\) is positive definite at \((0,0)\), and has compact
-sublevel sets on \(\mathbb{S}^1 \times \R\).  
+sublevel sets on \(\SxR\).  
 The derivative of \(E_\text{avg}\) along~\eqref{eqn:negative-time-vlp} is
@@
 largest subset of 
-\( Z = \left\{(q,p) \in \mathbb{S}^1 \times \R \mid \dot{E}_\text{avg}(q,p) = 0
+\( Z = \left\{(q,p) \in \SxR \mid \dot{E}_\text{avg}(q,p) = 0
 \right\} \). 
@@
 \[
-   Z= \left\{(q,p) \in \mathbb{S}^1 \times \R \mid q =0\right\} 
-   \cup \left\{(q,p) \in \mathbb{S}^1 \times \R \mid p =0\right\} 
+   Z= \left\{(q,p) \in \SxR \mid q =0\right\} 
+   \cup \left\{(q,p) \in \SxR \mid p =0\right\} 
    .
@@
 time, \ie, that for each compact subset \(K\) of 
-\(\mathbb{S}^1 \times \R\), and for almost every condition 
+\(\SxR\), and for almost every condition 
 \( (q_0,p_0) \in K \), there exists \( T > 0\) such that
@@
 \[
-   \Pi^- := \left\{ x(0) \in \mathbb{S}^1 \times \R \mid
+   \Pi^- := \left\{ x(0) \in \SxR \mid
    \lim_{t \to \infty} x(t) = (\pi, 0) \right\}.
@@
 \( \Pi^- \) is a one-dimensional immersed submanifold of
-\(\mathbb{S}^1 \times \R\), and thus is a set of measure zero
+\(\SxR\), and thus is a set of measure zero
 \cite{stable_manifold}.
 
-Let \( K \subset \mathbb{S}^1 \times \R\) be an arbitrary compact set and let 
+Let \( K \subset \SxR\) be an arbitrary compact set and let 
 \( x(0) \in K \backslash \left(\Pi^- \cup \left\{(0,0)\right\}\right)\).
@@
 We conclude that the VNHC injects energy into the VLP on 
-\(\mathbb{S}^1 \times \R\).
+\(\SxR\).
 By flipping the inequality of ~\eqref{eqn:vlp-energy-gain-condition}
@@
 \end{proof}
+
+% TODO: Add a remark:
+% Our controller not only injects energy, but makes the lower equilibrium a
+% repeller which means the "bad" set of x(0) does not contain a nbhd of (0,0),
+% so anything near (0,0) will start to gain eenrgy aka ANY slight deviation from
+% 0 will gain energy.
 

### 2020-11-18 latex/vlp.tex
Updated VLP according to new defn of energy gain

@@
 This function will produce constrained dynamics that inject energy into the VLP.
-
-% TODO: Describe the oscillation and rotation motion of a pendulum in the
-% qp-plane, and give a figure showing the orbiting motion for each of these.
-% Then describe what the orbit should do under energy injection (spiral away /
-% begin rotation outward) and energy dissipation (spiral in / rotate inward).
 
@@
 \begin{proof}
-We will first show that, under the given assumptions, the origin of the
-constrained dynamics~\eqref{eqn:vlp-hamiltonian} is a repeller, and for that
-we consider the negative time system obtained
+Since a system that gains energy per Definition \ref{defn:energy-gain}
+cannot have stable equilibria,
+we will first show that, under the given assumptions, the origin of the
+constrained dynamics~\eqref{eqn:vlp-hamiltonian} is a repeller. 
+For that we consider the negative time system obtained
 from~\eqref{eqn:vlp-hamiltonian} by reversing the sign of the vector field:
@@
 % 0 will gain energy.
+\begin{remark}
+   Theorem \ref{thm:vlp-energy-stabilization} not only guarantees that our VNHC
+   only injects energy, but that the lower equilibrium is a repeller. 
+   If the VLP is initialized at the origin, any slight push along
+   \(q\) will allow it to start gaining energy.
+\end{remark}
 

### 2020-11-18 latex/acrobot.tex
First writeup of acrobot energy gain proof for oscillations.

@@
 \begin{enumerate}
-    \item There always exists \(I > 0\) small enough that 
+    \item For all \(\epsilon > 0\), there exists \(I > 0\) small enough that 
     ~\eqref{eqn:acrobot-constraint} injects energy into the acrobot on the set
     \[
-        \mathcal{O} := \left\{(q_u,p_u) \in \Gamma 
-            \mid E(q_u,p_u) < E(\pi,0) \right\}
-        .
-    \]
-    If instead \(I < 0\), the VNHC dissipates energy on \(\mathcal{O}\).
+        \mathcal{O}_\epsilon := \left\{(q_u,p_u) \in \Gamma 
+        \mid E(\epsilon,0) < E(q_u,p_u) < E(\pi,0) \right\}
+        .
+    \]
+    If instead \(I < 0\), the VNHC dissipates energy on \(\mathcal{O}_\epsilon\).
 \item Define \(b : \SxR_{> 0} \rightarrow \R\) by
@@
     and fix \(\bar{\rho} > \sqrt{60m^2gl^3}\).
-    Suppose \(\int \limits_{0}^{2\pi} b(\sigma,\mu_0)d\sigma > 0\) for all 
+    Suppose \(\int \limits_{0}^{2\pi} b(\sigma,\rho_0)d\sigma > 0\) for all 
     \(\rho_0 \in \, \left]\sqrt{60m^2gl^3}, \bar{\rho}\right[\).
@@
 \begin{enumerate}
-    \item Background on perturbation theory and averaging.
+    \item Background on perturbation theory.
     \item Perturbation analysis for oscillations.
+    \item Energy gain for oscillations.
     \item Perturbation analysis for rotations.
+    \item Energy gain for rotations.
 \end{enumerate}
@@
 \citet{khalil_nonlinear} considers a system of the form
+
 \begin{equation}\label{eqn:khalil-setup}
     \begin{cases}
-        \dot{x} = f(t,x,\epsilon), \\
-        x(t_0) = \eta(\epsilon),
+        \dot{x} = f(t,x,I), \\
+        x(t_0) = \eta(I),
     \end{cases}
@@
 where 
-\(f : [t_0,t_1] \times D \times [-\epsilon_0,\epsilon_0] \rightarrow \R^n\) is
-sufficiently smooth.
+\(f : [t_0,t_1] \times D \times [-I_0,I_0] \rightarrow \R^n\) is smooth
+\footnote{Khalil actually considers \(f(t,x,\epsilon)\) which is ``sufficiently
+    smooth". We replace \(\epsilon\) with \(I\) and assume smoothness to
+    more easily connect his theory to the acrobot constraint.}.
 In the context of the acrobot, \(x\) is the set of coordinates which make
 our analysis convenient;
-\(\epsilon\) is our control parameter \(I\);
+\(I\) is a parameter of the system;
 the function \(f\) is the constrained dynamics in \(x\)-coordinates;
-and \(\eta(\epsilon) \equiv \eta_0\) is a constant initial condition.
-To keep a notational reminder that we are applying perturbation theory to
-the acrobot, we will henceforth use \(I\) in place of \(\epsilon\).
+and \(\eta(I) \equiv \eta_0\) is a constant initial condition.
 
@@
 Finally, one can compute the first-order Taylor approximation
-\(x(t,\eta_0) = x_0(t,\eta_0) + I x_1(t,\eta_0) + R(t,I)\), where the remainder
-term \(R(t,I)\) is \(O(I^2)\), \ie,
+\(x(t,\eta_0) = x_0(t,\eta_0) + I x_1(t,\eta_0) + R(t,\eta_0,I)\), where the remainder
+term \(R(t,\eta_0,I)\) is smooth and \(O(I^2)\), \ie,
 \[
@@
 \subsection*{Perturbation Analysis for Oscillations}
-As we have seen, the acrobot becomes a nominal pendulum that oscillates whenever the
-nominal energy ~\eqref{eqn:acrobot-nominal-E} is less than \(E(\pi,0)\).
+As we have seen, setting \(I = 0\) turns the acrobot into a nominal pendulum. 
+This pendulum oscillates whenever the nominal energy
+~\eqref{eqn:acrobot-nominal-E} is less than \(E(\pi,0)\). 
 Hence, the domain of oscillations for the nominal pendulum is given by
@@
 \]
-where \(\Gamma \simeq \Sone\times\R\) is the constraint manifold of the
+where \(\Gamma \simeq \SxR\) is the constraint manifold of the
 acrobot.
-We wish to find a transformation \(T(q_u,p_u)\) on \(\mathcal{O}\) into
-\((\alpha,\mu)\)-coordinates where \(\alpha \in \Sone\) is a pseudo-angle
-and \(\mu > 0\) is a pseudo-radius.
-Since the energy of oscillation can be determined by the orbit's intersection
+Recall from Figure \ref{fig:pendulum-level-sets} that orbits of a pendulum 
+are level sets of \(E\). 
+Oscillations of a pendulum look like ellipses contained in \(\mathcal{O}\).
+Since perturbation theory requires us to analyze the nominal pendulum, we will
+make our analysis convenient by changing coordinates into an
+``action" \(\mu > 0\) which remains constant on level sets of \(E\), 
+along with an ``angle" \(\alpha \in \Sone\) satisfying \(\dot{\alpha} > 0\) on
+\(\mathcal{O}\).
+
+The energy of oscillation can be determined by the orbit's intersection
 with the \(q_u\)-axis (as in Figure \ref{fig:mu-intersection}) 
-and \(q_u \in ]-\pi,\pi[\) on \(\mathcal{O}\), we set \(\mu \in ]0,\pi[\).
+and \(q_u \in ]-\pi,\pi[\) on \(\mathcal{O}\), so we set \(\mu \in ]0,\pi[\).
 The transformation we want is therefore a diffeomorphism of the form
 \begin{align*}
-    T : D\backslash \{(0,0)\} &\rightarrow \Sone \times \, ]0,\pi[, \\
+    T : \mathcal{O} \backslash \{(0,0)\} &\rightarrow \Sone \times \, ]0,\pi[, \\
     (q_u, p_u) &\mapsto (\alpha,\mu)
@@
 \begin{align}\label{eqn:acrobot-mu-dot-nom}
-    \dot{\mu}_n &= 0
+    \dot{\mu} &= 0
     , \\
     \label{eqn:acrobot-alpha-dot-nom}
-    \dot{\alpha}_n &= \sqrt{\frac{6g}{5l}} 
-        \sqrt{\frac{\cos(\mu_n\cos(\alpha_n)) - \cos(\mu_n)}
-            {\mu_n^2 \sin(\alpha_n)^2}}
+    \dot{\alpha} &= \sqrt{\frac{6g}{5l}} 
+        \sqrt{\frac{\cos(\mu\cos(\alpha)) - \cos(\mu)}
+            {\mu^2 \sin(\alpha)^2}}
     .
 \end{align}
-Note that we write the nominal coordinates as \((\alpha_n,\mu_n)\) to
-distinguish them from the acrobot's coordinates.
-One can confirm that \(\dot{\alpha}_n > 0\) for every \(\mu_n \in ]0,\pi[\).
-This is expected of an oscillating single pendulum.
+One can confirm that \(\dot{\alpha} > 0\) for every \(\mu \in ]0,\pi[\).
 By continuity of ~\eqref{eqn:acrobot-mu-dot}, 
@@
 
-%Using perturbation theory, we will find a first-order
-%approximation of \(\hat{\mu}\). 
-%We will show the dynamics of this approximation are \(\pi\)-periodic in
-%\(\alpha\), so that we only need to study whether \(\hat{\mu}\) increases or
-%decreases along a half-orbit \(\alpha \in [0,\pi]\).
-%To complete the proof of energy gain for oscillations, we will demonstrate that
-%\(\hat{\mu}(\pi) > \hat{\mu}(0)\).
-%This will imply that the the radius of orbits increases
-%until all orbits within \(\mathcal{O}\) eventually leave \(\mathcal{O}\) in finite time.
-
 In the spirit of perturbation analysis, we expand the solution of
@@
 \end{equation}
-where \(R(\alpha,\mu_0,I)\) is \(O(I^2)\).
+where \(R(\alpha,\mu_0,I)\) is smooth and \(O(I^2)\).
 From ~\eqref{eqn:khalil-perturbation-nominal} we know that 
@@
 \[
-    P_\mathcal{O}(\mu_0) = \mu_0 + I \hat{\mu}_1(\pi,\mu_0) + R(\pi,I)
+    P_\mathcal{O}(\mu_0) = \mu_0 + I \hat{\mu}_1(\pi,\mu_0) + R(\pi,\mu_0,I)
     .
@@
 
-We conclude that, for all \(m\), \(g\), \(l\), \(\bar{q}_a\), and for small
-enough \(I\), the Poincar\'{e} section \(P_\mathcal{O}(\mu_0)\) is positive,
-monotonically increasing, and 
+We can therefore state that, for each \(\mu_0 \in \, ]0,\pi[\), there exists
+\(I\) such that
+\[
+    P_\mathcal{O}(\mu_0) = \mu_0\left(1+ IKQ(\mu_0)\right) + R(\pi,\mu_0,I) > 0
+    .
+\]
+Unfortunately, Khalil's perturbation theory does not guarantee this \(I\) is the
+same for all \(\mu_0\).
+
+We deal with this issue by taking the compact set
+\[
+    \bar{\mathcal{O}} := \left\{(q_u,p_u) \in \Gamma 
+    \mid E(q_u,p_u) \leq E(\pi,0) \right\}
+    .
+\]
+The remainder term \(R(\pi,\mu_0,I)\) is smooth and of order \(I^2\),
+which means it can be written in the form 
+\(R(\pi,\mu_0,I) = I^2\tilde{R}(\mu_0,I)\)
+for some \(\tilde{R}\) a smooth function.
+Since \(\bar{O}\) is compact, \(\tilde{R}\) attains its minimum
+\(\underbar{r}\) on \(\bar{O}\).
+Hence, \(R(\pi,\mu_0,I)\) is bounded below by the parabola
+\(\underbar{r}I^2\), which means
+\begin{equation}\label{eqn:acrobot-Po-lowerbound}
+    P_\mathcal{O}(\mu_0) \geq \mu_0\left(1 + IKQ(\mu_0)\right) + I^2\underbar{r}
+    .
+\end{equation}
+Because \(I\) is assumed to be small, the term \(I^2\) shrinks rapidly.
+If \(\underbar{r} < 0\), we can find a value of \(I\) which overcomes
+the effect of \(I^2\underbar{r}\) and make \(P_\mathcal{O}\) positive.
+Furthermore, ~\eqref{eqn:acrobot-Po-lowerbound} implies that, for fixed \(I\),
+\(P_\mathcal{O}(\mu_0)\) is monotonically increasing in \(\mu_0\).
+
+We conclude that, for all \(m\), \(g\), \(l\), \(\bar{q}_a\), and for
+any \(\epsilon > 0\), there exists \(I(\epsilon)\) small enough so that
+the Poincar\'{e} section \(P_\mathcal{O}(\mu_0)\) is positive for all
+\(\mu_0 \in ]\epsilon,\pi[\) and 
 \(\lim \limits_{\mu_0 \to \pi} P_\mathcal{O}(\mu_0) = \infty\).
 
-Suppose we initialize the acrobot at \((q_0,p_0) \in \mathcal{O}\).
-Since on \(\mathcal{O}\) we have \(\dot{\alpha} > 0\), the orbit of the acrobot
-in \(\Gamma\) spirals clockwise around the origin; 
-if the spiral does not exit \(\mathcal{O}\), it must intersect the \(q_u\) axis.
-Once it does, the Poincar\'{e} section \(P_\mathcal{O}(\cdot)\) tells us 
-the orbit must spiral away from the origin. 
-This implies there cannot be any closed orbits in \(\mathcal{O}\) and that the
-origin is a repeller.
-
-Once the orbit reaches an intersection point \((\mu_0,0)\) on the \(q_u\)-axis such
-that \(P_\mathcal{O}(\mu_0) > \pi\), the orbit will no longer return to the
+\subsection*{Energy gain on \(\mathcal{O}_\epsilon\)}
+To simplify notation, denote by \(x(t) := (q_u(t),p_u(t))\) the orbit of the
+acrobot with initial condition \(x(0)\).
+Let \(\epsilon > 0\) and define
+\[
+    \mathcal{O}_\epsilon := \left\{(q_u,p_u) \in \Gamma
+    \mid E(\epsilon,0) < E(q_u,p_u) < E(\pi,0)\right\}
+    ,
+\]
+with \(\bar{\mathcal{O}}_\epsilon\) its (compact) closure.
+
+Since \(\bar{\mathcal{O}}_\epsilon \subset \mathcal{O}\), we still have 
+\(\dot{\alpha} > 0\), which means \(x(t)\) spirals clockwise around the origin.
+If this orbit remains in \(\mathcal{O}\), it must intersect
+the \(q_u\) axis at some radius \(\mu = \delta\)
+which depends on the initial condition \(x(0)\).
+Take 
+\[
+    \underbar{\delta} = 
+    \min\limits_{x(0) \in \bar{\mathcal{O}}_\epsilon} \delta(x(0))
+    ,
+\] 
+which exists because \(\bar{\mathcal{O}}_\epsilon\) is compact.
+From the previous perturbation analysis,
+there exists \(I\) small enough that \(P_\mathcal{O}(\mu_0)\) 
+is positive for all \(\mu_0 \in ]\underbar{\delta},\pi[\).
+
+Suppose \(x(0) \in \mathcal{O}_\epsilon\), and the orbit
+intersects the \(q_u\)-axis at some point \(x(t_0) = (m_0, 0)\).
+The Poincar\'{e} section \(P_\mathcal{O}(|m_0|)\) is positive because 
+\(|m_0| > \underbar{\delta}\).
+As the acrobot continues its oscillation, it might exit
+\(\mathcal{O}\) if it has gained enough energy;
+if not, it will intersect the \(q_u\) axis again at
+a new point \(x(t_1) = (m_1,0)\).
+Theorem \ref{thm:khalil-perturbation} guarantees that \(|m_1|\) differs
+from \(P_\mathcal{O}(|m_0|)\) only up to order \(I^2\), 
+which means \(|m_1| > |m_0|\).
+From here, we can apply Theorem \ref{thm:khalil-perturbation} iteratively:
+initializing the acrobot at \(x(t_i) = (m_i, 0)\), the
+next intersection with the \(q_u\) axis will be at
+\(x(t_{i+1}) = (m_{i+1},0)\) with \(|m_{i+1}| > |m_i|\).
+
+Because the Poincar\'{e} section is monotonically increasing, 
+the orbit will eventually reach an intersection point \(x(t_N) = (m_N,0)\)
+such that \(P_\mathcal{O}(|m_N|) > \pi\). 
+At this point, the orbit will no longer return to the
 \(q_u\)-axis.
-Instead, it will escape \(\mathcal{O}\) entirely, which means it must 
-escape any compact subset of \(\mathcal{O}\) in
-finite time. 
-We conclude that the acrobot is gaining energy on \(\mathcal{O}\).
+Instead, it will escape \(\mathcal{O}\) entirely; 
+in doing so, it also escapes \(\mathcal{O}_\epsilon\).
+Hence, the orbit must escape any compact subset of \(\mathcal{O}\) in finite
+time, which means the acrobot is gaining energy on \(\mathcal{O}_\epsilon\).
 
@@
 decreasing.
-Hence, the orbit in \(\mathcal{O}\) spirals towards the origin, which means the
-constraint is dissipating energy on \(\mathcal{O}\).
+Hence, the orbit in \(\mathcal{O}_\epsilon\) spirals clockwise towards the origin, 
+which means the constraint is dissipating energy.
 
@@
 
-\subsection*{Proving Energy Gain}
+\subsection*{Energy Gain on \(\Omega\)}
 We will prove the acrobot gains energy on the set 

### 2020-11-20 latex/acrobot.tex
Edited half of oscillation perturbation

@@
 
-Now imagine the person's torso is affixed to the swing's rope so that they are
+Now imagine the person's torso is affixed to the swing's rope so they are
 always upright. 
@@
 
-The acrobot's legs are rigid rods which cannot retract; so we emulate the person
+The acrobot's legs are rigid rods which cannot retract, so we emulate the person
 on a swing by pivoting the legs toward the direction of motion. 
@@
 \]
-Inserting these to ~\eqref{eqn:g-qupu}, we get the solution for \(p_a\) on the
+Inserting these into ~\eqref{eqn:g-qupu}, we get the solution for \(p_a\) on the
 constraint manifold:
@@
     \dot{p}_u &= - m g l (2s_u + s_{ua})
-    .
     \end{cases} \right|_{q_a = \bar{q}_a\arctan(Ip_u)}
+    .
 \end{equation}
@@
     \(\rho_0 \in \, \left]\sqrt{60m^2gl^3}, \bar{\rho}\right[\).
-    Then there exists \(I > 0\) small enough that
+    Then for all \(\epsilon > 0\) there exists \(I > 0\) small enough that
     ~\eqref{eqn:acrobot-constraint} injects energy into the acrobot on
     \[
-        \Omega := \left\{(q_u,p_u) \in \Gamma 
-            \mid E(q_u,p_u) < E(0,\bar{\rho})\right\}
+        \Omega_\epsilon := \left\{(q_u,p_u) \in \Gamma 
+        \mid E(\epsilon,0) < E(q_u,p_u) < E(0,\bar{\rho})\right\}
         .
@@
 %}
-The region \(\mathcal{O}\) is the oscillation domain of the simple pendulum
-obtained by setting \(q_a = 0\);
-\(\Omega\) includes \(\mathcal{O}\), alongside all rotations of this 
-pendulum with momentum less than \(\bar{\mu}\). 
-Hence, the first result of Theorem \ref{thm:acrobot-energy-stabilization} states
-that the acrobot will gain enough energy to begin performing giant motion;
-the second result guarantees can achieve giants with momentum
-\(|p_u| < \bar{\mu}\).
-
-Unfortunately, the energy-based proof used in Chapter \ref{ch:vlp} does not
-readily transfer to the acrobot -- there is no obvious Lyapunov function for
-this constrained system.
+The first result of Theorem \ref{thm:acrobot-energy-stabilization} states
+that all acrobots constrained by ~\eqref{eqn:acrobot-constraint} 
+will gain enough energy to begin rotating around the bar.
+An acrobot satisfying the second result will additionally achieve giants with momentum
+\(|p_u| < \bar{\rho}\).
+
+Unfortunately, the proof used in Chapter \ref{ch:vlp} does not
+readily transfer to the acrobot.
 Proving Theorem \ref{thm:acrobot-energy-stabilization} requires an intelligent
-change of coordinates and the use of perturbation theory from
-\cite{khalil_nonlinear}.
-The full proof is given in Chapter \ref{sec:acrobot-proof}, so 
-we conclude this section with a high level sketch of the proof.
-
-There is a suitable change of coordinates into a pseudo-radius 
-\(\mu \in \R_{>0}\) and pseudo-angle \(\alpha \in \Sone\) on the
-\((q_u,p_u)\)-plane.
-These coordinates have the property that \(\dot{\alpha} > 0\) when
-\(I\) is small enough.
-Perturbation theory shows that (for possibly smaller \(I\)) the radius \(\mu\)
-increases on average along \(\alpha\).
-Hence, orbits in the \((q_u,p_u)\)-plane spiral away from the origin, 
-which means the acrobot is gaining energy.
+change of coordinates and the use Khalil's perturbation
+theory \cite{khalil_nonlinear}.
+We provide the full proof of this theorem in Chapter \ref{sec:acrobot-proof}.
+%and conclude this section with a high level sketch of the proof.
+%
+%There is a suitable change of coordinates into a pseudo-radius 
+%\(\mu \in \R_{>0}\) and pseudo-angle \(\alpha \in \Sone\) on
+%\(\mathcal{O}_\epsilon\).
+%These coordinates have the property that \(\dot{\alpha} > 0\) when
+%\(I\) is small enough.
+%Perturbation theory shows that the radius \(\mu\)
+%increases on average along \(\alpha(t)\).
+%Hence, orbits in the \((q_u,p_u)\)-plane spiral away from the origin, 
+%which means the acrobot is gaining energy.
+%A similar approach can be used for \(\Omega_\epsilon\).
 
@@
 is conserved.
-Level sets of \(E\) are ellipses when \(E(q_u,p_u) < E(\pi,0)\), which we call
-``oscillations"; and they are open curves when \(E(q_u,p_u) > E(\pi,0)\), which we call
+Level sets of \(E\) are ellipses on the \((q_u,p_u)\)-plane
+when \(E(q_u,p_u) < E(\pi,0)\), which we call ``oscillations";
+and they are open curves when \(E(q_u,p_u) > E(\pi,0)\), which we call
 ``rotations". 
@@
         %\includestandalone[]{images/acrobot_I0}
-        \caption{The simple pendulum with two masses.}
+        \caption{A simple pendulum with two masses.}
         \label{fig:acrobot-I0}
@@
         \caption{Level sets of the pendulum energy function.
-            The red ellipse represents oscillations while the purple curve
+            The blue ellipse represents oscillations while the red curve
             represents rotations.}
@@
 Once we have these coordinates in hand, we can use perturbation theory to prove
-that \(\mu\) increases on average for the acrobot when \(I\) is small enough.
+that \(\mu\) increases on \(\mathcal{O}_\epsilon\) when \(I > 0\) is small enough.
 Likewise, we will find a second set of coordinates \((\beta, \rho)\) where
@@
 rotations of the simple pendulum, and \(\rho\) increases for the acrobot when
-\(I\) is small enough.
-Putting these together, we will show the acrobot is gaining energy.
+\(I > 0\) is small enough.
+Putting these together, we will show the acrobot is gaining energy on
+\(\Omega_\epsilon\).
 
@@
     smooth". We replace \(\epsilon\) with \(I\) and assume smoothness to
-    more easily connect his theory to the acrobot constraint.}.
-In the context of the acrobot, \(x\) is the set of coordinates which make
+    more easily connect his theory to the acrobot constraint.}
+on a domain \(D \subset \R^n\).
+In the context of the acrobot, \(x\) is the any coordinates which make
 our analysis convenient;
-\(I\) is a parameter of the system;
+\(I\) is a control parameter of the system;
 the function \(f\) is the constrained dynamics in \(x\)-coordinates;
@@
 Finally, one can compute the first-order Taylor approximation
-\(x(t,\eta_0) = x_0(t,\eta_0) + I x_1(t,\eta_0) + R(t,\eta_0,I)\), where the remainder
+\(x(t,\eta_0,I) = x_0(t,\eta_0) + I x_1(t,\eta_0) + R(t,\eta_0,I)\), where the remainder
 term \(R(t,\eta_0,I)\) is smooth and \(O(I^2)\), \ie,
@@
     unique solution \(x_0(t,\eta_0) \in D\) on \([t_0,t_1]\).
-    Then there exists \(I^\star > 0\) such that for all \(I,\, |I| < I^\star\), the
-    solution \(x(t,I,\eta_0)\) to ~\eqref{eqn:khalil-setup} satisfies
-    \[
-        \norm{x(t,I,\eta_0) - \left(x_0(t,\eta_0) + I x_1(t,\eta_0)\right)} \leq k |I^2|
+    Then there exists \(I^\star > 0\) such that, for all \(I\) with \(|I| < I^\star\), the
+    solution \(x(t,\eta_0,I)\) to ~\eqref{eqn:khalil-setup} satisfies
+    \[
+        \norm{x(t,\eta_0,I) - \left(x_0(t,\eta_0) + I x_1(t,\eta_0)\right)} \leq k |I^2|
     \]
@@
 solution of the nonlinear system by the Taylor approximation 
-\(x(t,\eta_0) \approx x_0(t,\eta_0) + I x_1(t,\eta_0)\).
+\(x(t,\eta_0,I) \approx x_0(t,\eta_0) + I x_1(t,\eta_0)\).
 When \(I\) is small enough, solutions of the nonlinear system and
@@
 acrobot.
-Recall from Figure \ref{fig:pendulum-level-sets} that orbits of a pendulum 
-are level sets of \(E\). 
-Oscillations of a pendulum look like ellipses contained in \(\mathcal{O}\).
+Recall that orbits of a pendulum are level sets of \(E\). 
+From Figure \ref{fig:pendulum-level-sets}, we know oscillations of a pendulum
+look like ellipses contained in \(\mathcal{O}\).
 Since perturbation theory requires us to analyze the nominal pendulum, we will
-make our analysis convenient by changing coordinates into an
-``action" \(\mu > 0\) which remains constant on level sets of \(E\), 
+make our analysis convenient by changing coordinates into a
+``radius" \(\mu > 0\) which remains constant on level sets of \(E\), 
 along with an ``angle" \(\alpha \in \Sone\) satisfying \(\dot{\alpha} > 0\) on
@@
 The energy of oscillation can be determined by the orbit's intersection
-with the \(q_u\)-axis (as in Figure \ref{fig:mu-intersection}) 
-and \(q_u \in ]-\pi,\pi[\) on \(\mathcal{O}\), so we set \(\mu \in ]0,\pi[\).
+\(\mu\) with the \(q_u\)-axis (as in Figure \ref{fig:mu-intersection}).
+Since \(q_u \in ]-\pi,\pi[\) on \(\mathcal{O}\), we set \(\mu \in ]0,\pi[\).
 The transformation we want is therefore a diffeomorphism of the form
@@
     % the origin to the intersection on the q-axis. We also draw the boundary of
-    % D and shade the interior
+    % mathcalO and shade the interior, and draw a line to the oscillation with
+    % clockwise angle alpha
     %\includestandalone[]{images/mu_intersection}
     \caption{The domain \(\mathcal{O}\) where a pendulum oscillates.
-    The pseudo-radius coordinate \(\mu\) corresponds to the
-    intersection of an orbit of oscillation with the \(q_u\)-axis.}
+    The pseudo-radius \(\mu\) corresponds to the
+    intersection of an orbit of oscillation with the \(q_u\)-axis.
+    The psuedo-angle \(\alpha\) is taken to be clockwise positive because 
+    oscillations of a pendulum move clockwise on \(\mathcal{O}\).}
     \label{fig:mu-intersection}
@@
 \begin{equation}\label{eqn:oscillation-pu2}
-    p_u^2 = 30m^2gl^3(\cos(q_u) - \cos(\mu))
+    p_u^2 = 30m^2gl^3\left(\cos(q_u) - \cos(\mu)\right)
     .
@@
 \[
-    p_u^2 = 30m^2gl^3(\cos(\mu \cos(\alpha)) - \cos(\mu))
+    p_u^2 = 30m^2gl^3\left(\cos(\mu \cos(\alpha)) - \cos(\mu)\right)
     .
@@
 \((\alpha,\mu)\)-coordinates can be computed by evaluating
-\begin{align*}
-    \dot{\mu} &= \left.\pdiff{\mu(q_u,p_u)}{q}_u\dot{q_u} +
+\begin{equation*}
+\left.\begin{cases}
+    \dot{\mu} &= \pdiff{\mu(q_u,p_u)}{q_u}\dot{q_u} +
          \pdiff{\mu(q_u,p_u)}{p_u}\dot{p}_u
-         \right|_{(q_u,p_u) = T\inv(\mu,\alpha)}
-    , \\
-    \dot{\alpha} &= \left.\pdiff{\alpha(q_u,p_u)}{q_u}\dot{q}_u + 
+    \\
+    \dot{\alpha} &= \pdiff{\alpha(q_u,p_u)}{q_u}\dot{q}_u + 
         \pdiff{\alpha(q_u,p_u)}{p_u}\dot{p}_u
-        \right|_{(q_u,p_u) = T\inv(\mu,\alpha)}
-    .
-\end{align*}
+    \end{cases}\right|_{(q_u,p_u) = T\inv(\mu,\alpha)}
+    .
+\end{equation*}
 with \((\dot{q}_u,\dot{p}_u)\) given by 
@@
 One can confirm that \(\dot{\alpha} > 0\) for every \(\mu \in ]0,\pi[\).
-By continuity of ~\eqref{eqn:acrobot-mu-dot}, 
+By continuity of ~\eqref{eqn:acrobot-alpha-dot}, 
 \(\dot{\alpha}\) remains positive on \(\mathcal{O}\) for \(I\) small enough.
 
-Suppose at time \(t = 0\) we have initialized the acrobot at 
-\((q_u,p_u) = (\mu_0,0)\), which corresponds to \((\alpha,\mu) = (0,\mu_0)\).
+Suppose at time \(t = 0\) we have initialized the acrobot at \((\alpha,\mu) = (0,\mu_0)\).
 Since \(\alpha\) is continuously increasing, we know that the orbit will do one
@@
 or it will remain in \(\mathcal{O}\) and hit the \(q_u\)-axis at some point 
-\((\alpha,\mu) = (\pi,m)\) (as illustrated in Figure
+\((\alpha,\mu) = (\pi,m)\).
+From there, the acrobot could escape \(\mathcal{O}\) along the second half of the orbit or
+return to some point \((\alpha,\mu) = (0,\tilde{m})\) (as illustrated in Figure
 \ref{fig:acrobot-possible-orbits}).
-From there, the acrobot could escape \(\mathcal{O}\) along the second half of the orbit or
-return to some point \((\alpha,\mu) = (0,\tilde{m})\).
 If \(\tilde{m} < \mu_0\) then the acrobot has lost energy; 
-if \(\tilde{m} > \mu_0\) the acrobot has gained energy; 
-and if \(\tilde{m} = \mu_0\), the acrobot is on a closed orbit.
+if \(\tilde{m} = \mu_0\), the acrobot is on a closed orbit;
+and if \(\tilde{m} > \mu_0\) the acrobot has gained energy.
 Which situation occurs might depend entirely on the choice of
@@
     \centering 
-    % TODO: Draw the domain D, an initial point (q0,p0) = (mu0,0), and show the
-    % 4 clockwise half orbits: when the orbit exits D on the bottom left corner;
-    % when the orbit hits the left q-axis at (-mu0-x,0); when it hits (-mu0,0);
-    % and when it hits (-mu0+x,0) for some x > 0.
+    % TODO: Draw the domain mathcalO, an initial point (q0,p0) = (mu0,0), and show the
+    % half orbit hitting the left q-axis at (-m,0).
+    % Then show the 
+    % 4 clockwise half orbits starting at (-m,0): 
+    % one orbit exits mathcalO on the top right corner;
+    % one orbit hits the right q-axis at (mu0+x,0); 
+    % one hits (mu0,0);
+    % and one hits (mu0-x,0) for some x > 0.
     %includestandalone[width=0.5\textwdith]{images/acrobot_possible_orbits}
@@
 
+We can simplify the analysis of where orbits intersect the \(q_u\)-axis 
+by studying how the system behaves on average.
 Since \(\dot{\alpha} > 0\), we could theoretically find some time scaling 
@@
 Setting \(\hat{\mu}(\alpha) := \mu(\tau(\alpha))\) allows us to ignore the
-dynamics of \(\alpha\) and study the evolution of \(\mu\) as a 
-function of \(\alpha\) rather than as a function of \(t\) (assuming the acrobot
-does not leave \(\mathcal{O}\)).
+dynamics of \(\alpha\) (assuming the acrobot remains in \(\mathcal{O}\)).
+This allows us to study the evolution of \(\mu\) as a 
+function of \(\alpha\) rather than as a function of time.
 The dynamics of \(\hat{\mu}\) are
@@
 \]
+%TODO: EDIT FROM HERE
 Notice that \(\hat{b}(\alpha,\mu_0)\) is both even and 

### 2020-11-21 latex/acrobot.tex
Finished first edit of oscillation

@@
 \]
-%TODO: EDIT FROM HERE
-Notice that \(\hat{b}(\alpha,\mu_0)\) is both even and 
-\(\pi\)-periodic in \(\alpha\), so that the following properties hold:
-\begin{equation}\label{eqn:bhat-integral-properties}
-    \begin{split}
-    \int \limits_0 ^ {N \pi} \hat{b}(\sigma,\mu_0) d\sigma
-    &= N \int \limits_0^\pi \hat{b}(\sigma,\mu_0) d\sigma \, 
-    \left(\forall N \in \mathbb{Z}_{> 0}\right)
-    , \\
-    \int \limits_{-\pi}^0 \hat{b}(\sigma,\mu_0) d\sigma
-    &= \int \limits_0^\pi \hat{b}(\sigma,\mu_0)d\sigma
-    .
-    \end{split}
-\end{equation}
+
 Suppose we initialize the acrobot at \((\alpha,\mu) = (0,\mu_0)\). 
-After half an orbit, when \(\alpha = \pi\), the acrobot will have a new
-pseudo-radius \(m = \hat{\mu}(\pi,\mu_0)\).
+After half an oscillation, when \(\alpha = \pi\), the orbit will land on the
+negative \(q_u\)-axis with pseudo-radius 
+\(m = \hat{\mu}(\pi,\mu_0) \approx \mu_0 + I\hat{\mu}_1(\pi,\mu_0)\).
 Hence, the change in pseudo-radius is 
 \(m - \mu_0 \approx I\hat{\mu}_1(\pi,\mu_0)\).
-After an additional half orbit, the acrobot lands on the positive \(q_u\)-axis
+If the orbit remains in \(\mathcal{O}\) for the full oscillation, 
+it will land on the positive \(q_u\)-axis
 with pseudo-radius
 \(\tilde{m} \approx \mu_0 + I\hat{\mu}_1(2\pi,\mu_0)\).
-The properties outlined in ~\eqref{eqn:bhat-integral-properties} imply that
-\(\hat{\mu}_1(2\pi,\mu_0) = 2 \hat{\mu}_1(\pi,\mu_0)\). 
-This means the change in pseudo-radius after one entire oscillation is 
-\(\tilde{m} - \mu_0 \approx 2I\hat{\mu}_1(\pi,\mu_0) \approx 2 (m - \mu_0)\).
-Hence, if \(\hat{\mu}_1(\pi,\cdot)\) is always positive,
-\(\hat{\mu}\) must be increasing every half oscillation.
-This in turn would guarantee \(q_u\) is increasing, which means the acrobot
-will eventually reach a rotation.
-
-To prove this intuitive argument rigorously, we define the Poincar\'{e} section
-\(P_\mathcal{O}(\mu_0) := \hat{\mu}(\pi,\mu_0)\), which expands into
+
+Notice that \(\hat{b}(\alpha,\mu_0)\) is even and 
+\(\pi\)-periodic in \(\alpha\), so that
+\begin{equation}\label{eqn:bhat-integral-properties}
+    \int \limits_{-\pi}^\pi \hat{b}(\sigma,\mu_0) d\sigma
+    = \int \limits_0 ^ {2 \pi} \hat{b}(\sigma,\mu_0) d\sigma
+    = 2 \int \limits_0^\pi \hat{b}(\sigma,\mu_0) d\sigma
+    .
+\end{equation}
+This implies that 
+\(\hat{\mu}_1(2\pi,\mu_0) = 2 \hat{\mu}_1(\pi,\mu_0)\),
+which means the change in pseudo-radius after a full oscillation is 
+\(\tilde{m} - \mu_0 \approx 2I\hat{\mu}_1(\pi,\mu_0)\).
+If \(\hat{\mu}_1(\pi,\mu_0) > 0\), 
+then both \(m - \mu_0 > 0\) and \(\tilde{m} - \mu_0 > 0\).
+In other words, we can study the average behaviour of \(\hat{\mu}\) by examining
+the behaviour over each half oscillation.
+
+We define the Poincar\'{e} section 
+\(P_\mathcal{O}(\mu_0) := \hat{\mu}(\pi,\mu_0)\)
+corresponding to the change in \(\hat{\mu}\) after half an oscillation.
+The section expands into
 \[
@@
 so that \(\hat{\mu}_1(\pi,\mu_0) = K Q(\mu_0)\).
-Note that \(K\) is a positive constant that takes into account the acrobot's
-mass and length, while \(\hat{b}(\alpha,\mu)\) is adimensional;
-that is, it is the same for every possible choice of acrobot.
-Hence, \(Q(\mu_0)\) is the same for every acrobot.
+Note that \(K\) is a positive constant which contains the acrobot's physical
+parameters, while \(\hat{b}(\alpha,\mu)\) is adimensional. 
+This means \(Q(\mu_0)\) is identical for every acrobot.
 If \(Q(\mu_0)\) is strictly positive, then \(\hat{\mu}_1(\pi,\mu_0)\) will be
@@
 in Figure \ref{fig:acrobot-Q}. 
-We see that \(Q(\mu_0)\) is strictly positive and monotonically increasing, with
+We see that it is strictly positive and monotonically increasing, with
 an asymptote at \(\mu_0 = \pi\). 
@@
 
-We can therefore state that, for each \(\mu_0 \in \, ]0,\pi[\), there exists
-\(I\) such that
-\[
-    P_\mathcal{O}(\mu_0) = \mu_0\left(1+ IKQ(\mu_0)\right) + R(\pi,\mu_0,I) > 0
-    .
-\]
-Unfortunately, Khalil's perturbation theory does not guarantee this \(I\) is the
-same for all \(\mu_0\).
-
+This fact, coupled with knowledge that \(R(\pi,\mu_0,I)\) is \(O(I^2)\), 
+implies there exists \(I > 0\) small enough that
+\[
+    P_\mathcal{O}(\mu_0) = \mu_0 + IKQ(\mu_0) + R(\pi,\mu_0,I) > 0
+    .
+\]
+Unfortunately, this \(I\) may not be the
+same for all \(\mu_0 \in \, ]0,\pi[\), since
+the value of \(R(\pi,\mu_0,I)\) could become more negative
+for other initial conditions.
 We deal with this issue by taking the compact set
@@
 \]
-The remainder term \(R(\pi,\mu_0,I)\) is smooth and of order \(I^2\),
+The remainder term \(R(\alpha,\mu_0,I)\) is smooth and of order \(I^2\),
 which means it can be written in the form 
-\(R(\pi,\mu_0,I) = I^2\tilde{R}(\mu_0,I)\)
+\(R(\alpha,\mu_0,I) = I^2\tilde{R}(\alpha,\mu_0,I)\)
 for some \(\tilde{R}\) a smooth function.
-Since \(\bar{O}\) is compact, \(\tilde{R}\) attains its minimum
-\(\underbar{r}\) on \(\bar{O}\).
-Hence, \(R(\pi,\mu_0,I)\) is bounded below by the parabola
-\(\underbar{r}I^2\), which means
+Since \(\bar{\mathcal{O}}\) is compact, \(\tilde{R}\) attains its minimum
+\({r}\) on \(\bar{\mathcal{O}}\).
+Hence, \(R(\alpha,\mu_0,I)\) is bounded below on \(\mathcal{O}\) by
+\(I^2{r}\), which implies 
 \begin{equation}\label{eqn:acrobot-Po-lowerbound}
-    P_\mathcal{O}(\mu_0) \geq \mu_0\left(1 + IKQ(\mu_0)\right) + I^2\underbar{r}
-    .
-\end{equation}
-Because \(I\) is assumed to be small, the term \(I^2\) shrinks rapidly.
-If \(\underbar{r} < 0\), we can find a value of \(I\) which overcomes
-the effect of \(I^2\underbar{r}\) and make \(P_\mathcal{O}\) positive.
-Furthermore, ~\eqref{eqn:acrobot-Po-lowerbound} implies that, for fixed \(I\),
+    P_\mathcal{O}(\mu_0) \geq \mu_0 + IKQ(\mu_0) + I^2{r}
+    .
+\end{equation}
+This equation implies that, for fixed \(I\),
 \(P_\mathcal{O}(\mu_0)\) is monotonically increasing in \(\mu_0\).
+Thus, for every \(\epsilon > 0\) and \(\mu_0 \in \, [\epsilon,\pi[\) we have
+\[
+    P_\mathcal{O}(\mu_0) \geq P_\mathcal{O}(\epsilon)
+    = \epsilon + IKQ(\epsilon) + I^2{r}
+    .
+\]
+Since \(I\) is assumed to be small, \(I^2\) shrinks rapidly.
+Therefore, we can always find a small enough \(I > 0\) making 
+\(\epsilon + I^2{r} \geq 0\), so that
+\[
+    P_\mathcal{O}(\epsilon) \geq IKQ(\epsilon) > 0
+    .
+\]
 
@@
 the Poincar\'{e} section \(P_\mathcal{O}(\mu_0)\) is positive for all
-\(\mu_0 \in ]\epsilon,\pi[\) and 
+\(\mu_0 \in [\epsilon,\pi[\) and 
 \(\lim \limits_{\mu_0 \to \pi} P_\mathcal{O}(\mu_0) = \infty\).
@@
 
-Since \(\bar{\mathcal{O}}_\epsilon \subset \mathcal{O}\), we still have 
-\(\dot{\alpha} > 0\), which means \(x(t)\) spirals clockwise around the origin.
+Let \(x(0) \in \mathcal{O}_\epsilon \subset \mathcal{O}\).
+Since \(\dot{\alpha} > 0\), the orbit \(x(t)\) spirals clockwise around
+the origin.
 If this orbit remains in \(\mathcal{O}\), it must intersect
@@
 there exists \(I\) small enough that \(P_\mathcal{O}(\mu_0)\) 
-is positive for all \(\mu_0 \in ]\underbar{\delta},\pi[\).
+is positive for all \(\mu_0 \in [\underbar{\delta},\pi[\).
 
@@
 The Poincar\'{e} section \(P_\mathcal{O}(|m_0|)\) is positive because 
-\(|m_0| > \underbar{\delta}\).
-As the acrobot continues its oscillation, it might exit
-\(\mathcal{O}\) if it has gained enough energy;
-if not, it will intersect the \(q_u\) axis again at
-a new point \(x(t_1) = (m_1,0)\).
-Theorem \ref{thm:khalil-perturbation} guarantees that \(|m_1|\) differs
-from \(P_\mathcal{O}(|m_0|)\) only up to order \(I^2\), 
-which means \(|m_1| > |m_0|\).
+\(|m_0| \geq \underbar{\delta}\).
+As the acrobot continues its oscillation, it will intersect the \(q_u\) axis
+again at a new point \(x(t_1) = (m_1,0)\).
+
+Setting \(p_1 := P_\mathcal{O}(|m_0|)\),
+Theorem \ref{thm:khalil-perturbation} guarantees that
+\[
+    \norm{m_1 - p_1} \leq cI^2
+    ,
+\]
+for some \(c > 0\).
+%TODO 
+\textbf{SHOW THIS IMPLIES |m1| > |m0|}
+
 From here, we can apply Theorem \ref{thm:khalil-perturbation} iteratively:
@@
 Setting \(I < 0\), the Poincar\'{e} section \(P_\mathcal{O}(\mu_0)\) is now
-decreasing.
-Hence, the orbit in \(\mathcal{O}_\epsilon\) spirals clockwise towards the origin, 
+decreasing instead of increasing,
 which means the constraint is dissipating energy.

### 2020-11-22 latex/acrobot.tex
Completed oscillation energy proof
Also completed first edit of perturbation section for oscillations. Needs one more pass
before sending to Manfredi.

@@
     \centering
-    \includestandalone[width=0.7\textwidth]{images/simple_acrobot_model}
+    \includestandalone[width=0.5\textwidth]{images/simple_acrobot_model}
     \caption{A simple acrobot has massless rods of equal length \(l\) and 
@@
 \begin{prop}\label{prop:acrobot-fpu-regular}
-    A relation \(h(q,p) = q_a - f(p_u) = 0\) for ~\eqref{eqn:acrobot-hamiltonian} 
+    A relation \(h(q,p) = q_a - f(p_u) = 0\) 
     with \(f \in C^2\left(\mathcal{P}_u; \mathcal{Q}_a\right)\) is a regular
-    VNHC of order 1.
+    VNHC of order 1 for the simple acrobot.
 \end{prop}
@@
 \section{Previous Constraint Approaches}
-% TODO: Give Xingbo's intuition, abstract away to qa = sin(theta) constraint
-% based on VLP but show it doesn't work well in our framework
-% TODO: Describe Xingbo's results and observe that his results were similar to a VNHC
-% except they used VHC tools. We will modify his approach to use VNHCs and prove
-% rigorously that we can stabilize any energy level set on the acrobot.
 Let us examine some of the existing approaches to generating giant motion for
@@
 \section{The Acrobot Constraint}
-% TODO: Give the constraint, state the theorem, and show the outline of the
-% proof (both as a sketch and in a figure)
 One may be tempted to design a constraint of the form \(q_a = \bar{q}_a\sin(\theta)\) 
@@
 displayed in Figure \ref{fig:qa-arctan}.
-Here, \(\bar{q}_a \in ]0,\frac{2 Q_a}{\pi}]\) and \(I \in \R_{>0}\) is a fixed
+Here, \(\bar{q}_a \in ]0,\frac{2 Q_a}{\pi}]\) and \(I \in \R\) is a fixed
 control parameter.
@@
 This constraint does not perfectly recreate giant motion, during which
-the gymnast's legs are almost completely extended -- it instead pivots the legs
-partially during rotations.
+the gymnast's legs are almost completely extended \cite{usagym_giant}
+-- it instead pivots the legs partially during rotations.
 However, the behaviour looks similar enough that the constraint should provide a
@@
 \section{Proving the Acrobot Gains Energy}\label{sec:acrobot-proof}
-% TODO: Prove this in sections
-% TODO: You don't have to use the proof of theta_dot being positive. Stick to
-% Manfredi's proof, and when you say "for I small enough" put in a footnote that
-% says "in fact, we can be more precise and set bounds for I given as I <
-% min{...}. See appendix for more detail."
-% Then, we can have an appendix where we derive the bounds on I so that
-% theta_dot is always positive. This will help the reader keep track and not
-% bother them too much.
 Proving Theorem \ref{thm:acrobot-energy-stabilization} is not a simple task.
@@
 \subsection*{Background on Perturbation Theory}
-Nonlinear systems like the acrobot are difficult (or impossible) to solve
+Nonlinear systems like the acrobot are difficult (or even impossible) to solve
 analytically.
-Perturbation theory allows one to understand the behaviour of the nonlinear system
+Perturbation theory allows one to understand the behaviour of nonlinear systems
 by studying a simpler nominal system. 
@@
 \citet{khalil_nonlinear} considers a system of the form
-
 \begin{equation}\label{eqn:khalil-setup}
@@
 where 
-\(f : [t_0,t_1] \times D \times [-I_0,I_0] \rightarrow \R^n\) is smooth 
-\footnote{Khalil actually considers \(f(t,x,\epsilon)\) which is ``sufficiently
-    smooth". We replace \(\epsilon\) with \(I\) and assume smoothness to
-    more easily connect his theory to the acrobot constraint.}
+\(f : [t_0,t_1] \times D \times [-I_0,I_0] \rightarrow \R^n\) is  
+smooth\footnote{Khalil actually considers \(f(t,x,\epsilon)\) which is ``sufficiently
+    smooth". To more easily connect his theory to the acrobot constraint, we
+    assume smoothness of \(f\) and replace \(\epsilon\) with \(I\).}
 on a domain \(D \subset \R^n\).
-In the context of the acrobot, \(x\) is the any coordinates which make
+In the context of the acrobot, \(x\) is any pair of coordinates which make
 our analysis convenient;
-\(I\) is a control parameter of the system;
+\(I\) is the control parameter of the system;
 the function \(f\) is the constrained dynamics in \(x\)-coordinates;
@@
     Suppose \(f : [t_0,t_1] \times D \times [-I_0,I_0] \rightarrow \R^2\) is
-    \(C^1\), and that the nominal problem ~\eqref{eqn:khalil-perturbation-nominal} has a
+    \(C^1\), and that the nominal system ~\eqref{eqn:khalil-perturbation-nominal} has a
     unique solution \(x_0(t,\eta_0) \in D\) on \([t_0,t_1]\).
@@
 Recall that orbits of a pendulum are level sets of \(E\). 
-From Figure \ref{fig:pendulum-level-sets}, we know oscillations of a pendulum
+From Figure \ref{fig:pendulum-level-sets}, we know that oscillation level sets
 look like ellipses contained in \(\mathcal{O}\).
@@
 or it will remain in \(\mathcal{O}\) and hit the \(q_u\)-axis at some point 
-\((\alpha,\mu) = (\pi,m)\).
+\((\alpha,\mu) = (\pi,u)\).
 From there, the acrobot could escape \(\mathcal{O}\) along the second half of the orbit or
-return to some point \((\alpha,\mu) = (0,\tilde{m})\) (as illustrated in Figure
+return to some point \((\alpha,\mu) = (0,\tilde{u})\) (as illustrated in Figure
 \ref{fig:acrobot-possible-orbits}).
-If \(\tilde{m} < \mu_0\) then the acrobot has lost energy; 
-if \(\tilde{m} = \mu_0\), the acrobot is on a closed orbit;
-and if \(\tilde{m} > \mu_0\) the acrobot has gained energy.
+If \(\tilde{u} < \mu_0\) then the acrobot has lost energy; 
+if \(\tilde{u} = \mu_0\), the acrobot is on a closed orbit;
+and if \(\tilde{u} > \mu_0\) the acrobot has gained energy.
 Which situation occurs might depend entirely on the choice of
@@
         either the orbit escapes \(\mathcal{O}\) into the rotation
-        zone (purple), or the orbit hits the other half of the \(q_u\)-axis 
-        with a higher (red), equal (black), or lower (blue) angle than when it
+        zone (purple), or the orbit returns to the \(q_u\)-axis 
+        with a higher (red), equal (black), or lower (blue) pseudo-radius than when it
         started.}
@@
 
-We can simplify the analysis of where orbits intersect the \(q_u\)-axis 
-by studying how the system behaves on average.
+We can simplify our analysis of the acrobot by studying how its orbits
+intersect the \(q_u\)-axis on average.
 Since \(\dot{\alpha} > 0\), we could theoretically find some time scaling 
@@
 Setting \(\hat{\mu}(\alpha) := \mu(\tau(\alpha))\) allows us to ignore the
-dynamics of \(\alpha\) (assuming the acrobot remains in \(\mathcal{O}\)).
-This allows us to study the evolution of \(\mu\) as a 
-function of \(\alpha\) rather than as a function of time.
+dynamics of \(\alpha\) by observing the evolution of \(\mu\) as a function of
+\(\alpha\) rather than a function of time (assuming the acrobot remains in
+\(\mathcal{O}\)).
 The dynamics of \(\hat{\mu}\) are
@@
 \end{equation}
-with solution \(\hat{\mu}(\alpha,\mu_0,I)\).
 
@@
     \begin{cases}
-        \dot{\hat{\mu}}_1 = a(\alpha,\mu_0)\hat{\mu}_1 + b(\alpha,\mu_0)
+        \dot{\hat{\mu}}_1 = 
+        \pdiff{g}{\hat{\mu}}(\alpha,\mu_0,0)\hat{\mu}_1 + \pdiff{g}{I}(\alpha, \mu_0, 0)
         , \\
@@
 \end{equation}
-where
-\begin{align*}
-    a(\alpha,\mu_0) &= \pdiff{g}{\hat{\mu}}(\alpha,\mu_0,0)
-    , \\
-    b(\alpha, \mu_0) &= \pdiff{g}{I}(\alpha, \mu_0, 0)
-    .
-\end{align*}
-These are difficult to compute by hand, so we resort to MATLAB symbolic
+These dynamics are difficult to compute by hand, so we resort to MATLAB symbolic
 computation to reveal that
 \begin{align*}
-    a(\alpha,\mu_0) &= 0
-    , \\
-    b(\alpha,\mu_0) &= K \hat{b}(\alpha,\mu_0)
+    \pdiff{g}{\hat{\mu}}(\alpha,\mu_0,0) &= 0
+    , \\
+    \pdiff{g}{I}(\alpha, \mu_0, 0) &= K b(\alpha,\mu_0)
     ,
@@
     , \\
-    \hat{b}(\alpha,\mu_0) &:= \frac{
+    b(\alpha,\mu_0) &:= \frac{
         \mu_0 |\sin(\alpha)| \left(
@@
     \hat{\mu}_1(\alpha,\mu_0) =
-    K \int \limits_0^\alpha \hat{b}(\sigma,\mu_0)d\sigma
+    K \int \limits_0^\alpha b(\sigma,\mu_0)d\sigma
     .
@@
 negative \(q_u\)-axis with pseudo-radius 
-\(m = \hat{\mu}(\pi,\mu_0) \approx \mu_0 + I\hat{\mu}_1(\pi,\mu_0)\).
+\(u = \hat{\mu}(\pi,\mu_0) \approx \mu_0 + I\hat{\mu}_1(\pi,\mu_0)\).
 Hence, the change in pseudo-radius is 
-\(m - \mu_0 \approx I\hat{\mu}_1(\pi,\mu_0)\).
+\(u - \mu_0 \approx I\hat{\mu}_1(\pi,\mu_0)\).
 If the orbit remains in \(\mathcal{O}\) for the full oscillation, 
@@
 with pseudo-radius
-\(\tilde{m} \approx \mu_0 + I\hat{\mu}_1(2\pi,\mu_0)\).
-
-Notice that \(\hat{b}(\alpha,\mu_0)\) is even and 
+\(\tilde{u} \approx \mu_0 + I\hat{\mu}_1(2\pi,\mu_0)\).
+
+Notice that \(b(\alpha,\mu_0)\) is even and 
 \(\pi\)-periodic in \(\alpha\), so that
 \begin{equation}\label{eqn:bhat-integral-properties}
-    \int \limits_{-\pi}^\pi \hat{b}(\sigma,\mu_0) d\sigma
-    = \int \limits_0 ^ {2 \pi} \hat{b}(\sigma,\mu_0) d\sigma
-    = 2 \int \limits_0^\pi \hat{b}(\sigma,\mu_0) d\sigma
+    \int \limits_{-\pi}^\pi b(\sigma,\mu_0) d\sigma
+    = \int \limits_0 ^ {2 \pi} b(\sigma,\mu_0) d\sigma
+    = 2 \int \limits_0^\pi b(\sigma,\mu_0) d\sigma
     .
@@
 which means the change in pseudo-radius after a full oscillation is 
-\(\tilde{m} - \mu_0 \approx 2I\hat{\mu}_1(\pi,\mu_0)\).
+\(\tilde{u} - \mu_0 \approx 2I\hat{\mu}_1(\pi,\mu_0)\).
 If \(\hat{\mu}_1(\pi,\mu_0) > 0\), 
-then both \(m - \mu_0 > 0\) and \(\tilde{m} - \mu_0 > 0\).
+then both \(u - \mu_0 > 0\) and \(\tilde{u} - \mu_0 > 0\).
 In other words, we can study the average behaviour of \(\hat{\mu}\) by examining
-the behaviour over each half oscillation.
+the behaviour at each half oscillation.
 
@@
 \[
-    Q(\mu_0) := \int \limits_0^\pi \hat{b}(\sigma,\mu_0)d\sigma
+    Q(\mu_0) := \int \limits_0^\pi b(\sigma,\mu_0)d\sigma
     ,
@@
 Note that \(K\) is a positive constant which contains the acrobot's physical
-parameters, while \(\hat{b}(\alpha,\mu)\) is adimensional. 
+parameters, while \(b(\alpha,\mu)\) is adimensional. 
 This means \(Q(\mu_0)\) is identical for every acrobot.
@@
 Since \(\bar{\mathcal{O}}\) is compact, \(\tilde{R}\) attains its minimum
-\({r}\) on \(\bar{\mathcal{O}}\).
+\({r}\) somewhere on \(\bar{\mathcal{O}}\).
 Hence, \(R(\alpha,\mu_0,I)\) is bounded below on \(\mathcal{O}\) by
@@
 \end{equation}
-This equation implies that, for fixed \(I\),
+This equation further implies that, for fixed \(I\),
 \(P_\mathcal{O}(\mu_0)\) is monotonically increasing in \(\mu_0\).
@@
 We conclude that, for all \(m\), \(g\), \(l\), \(\bar{q}_a\), and for
-any \(\epsilon > 0\), there exists \(I(\epsilon)\) small enough so that
+any \(\epsilon > 0\), there exists \(I^\prime(\epsilon)\) small enough so that,
+for all \(0 < I \leq I^\prime(\epsilon)\),
 the Poincar\'{e} section \(P_\mathcal{O}(\mu_0)\) is positive for all
@@
 which exists because \(\bar{\mathcal{O}}_\epsilon\) is compact.
+Fruthermore, \(\underbar{\delta} \leq \epsilon\).
 From the previous perturbation analysis,
-there exists \(I\) small enough that \(P_\mathcal{O}(\mu_0)\) 
-is positive for all \(\mu_0 \in [\underbar{\delta},\pi[\).
+there exists \(I^\prime(\underbar{\delta})\) small enough that
+\(P_\mathcal{O}(\mu_0)\) is positive for all 
+\(\mu_0 \in [\underbar{\delta},\pi[\).
 
 Suppose \(x(0) \in \mathcal{O}_\epsilon\), and the orbit
-intersects the \(q_u\)-axis at some point \(x(t_0) = (m_0, 0)\).
-The Poincar\'{e} section \(P_\mathcal{O}(|m_0|)\) is positive because 
-\(|m_0| \geq \underbar{\delta}\).
+intersects the \(q_u\)-axis at some point \(x(t_0) = (u_0, 0)\).
+The Poincar\'{e} section \(P_\mathcal{O}(|u_0|)\) is positive because 
+\(|u_0| \geq \underbar{\delta}\).
 As the acrobot continues its oscillation, it will intersect the \(q_u\) axis
-again at a new point \(x(t_1) = (m_1,0)\).
-
-Setting \(p_1 := P_\mathcal{O}(|m_0|)\),
-Theorem \ref{thm:khalil-perturbation} guarantees that
-\[
-    \norm{m_1 - p_1} \leq cI^2
-    ,
-\]
-for some \(c > 0\).
-%TODO 
-\textbf{SHOW THIS IMPLIES |m1| > |m0|}
-
-From here, we can apply Theorem \ref{thm:khalil-perturbation} iteratively:
-initializing the acrobot at \(x(t_i) = (m_i, 0)\), the
+again at a new point \(x(t_1) = (u_1,0)\).
+
+Setting \(p_1 := P_\mathcal{O}(|u_0|)\),
+Theorem \ref{thm:khalil-perturbation} guarantees that,
+for all \(I \in ]0,I^\prime(\underbar{\delta})]\),
+\[
+    \abs{|u_1| - p_1} \leq cI^2
+    ,
+\]
+where \(c > 0\) is an independent constant.
+This implies that \(|u_1| \geq p_1 - cI^2\). 
+We also know that
+\[
+    p_1 - |u_0| \geq IKQ(|u_0|) \geq IKQ(\underbar{\delta}) > 0
+    .
+\]
+There exists \(I^\star \in ]0,I^\prime(\underbar{\delta})]\) where
+\(c{I^\star}^2 < I^\star KQ(\underbar{\delta})\), which means
+\begin{align*}
+    |u_1| - |u_0| 
+    &= \left(|u_1| - p_1\right) + \left(p_1 - |u_0|\right)
+    \\
+    &\geq \left(p_1 - c{I^\star}^2 - p_1\right) + I^\star KQ(\underbar{\delta})
+    \\
+    &= I^\star KQ(\underbar{\delta}) - c{I^\star}^2
+    \\
+    &> 0
+    .
+\end{align*}
+
+Since \(u_0 \in [\underbar{\delta},\pi[\) was arbitrary, 
+we can apply Theorem \ref{thm:khalil-perturbation} iteratively:
+initializing the acrobot at \(x(t_i) = (u_i, 0)\), the
 next intersection with the \(q_u\) axis will be at
-\(x(t_{i+1}) = (m_{i+1},0)\) with \(|m_{i+1}| > |m_i|\).
+\(x(t_{i+1}) = (u_{i+1},0)\) with \(|u_{i+1}| > |u_i|\).
 
 Because the Poincar\'{e} section is monotonically increasing, 
-the orbit will eventually reach an intersection point \(x(t_N) = (m_N,0)\)
-such that \(P_\mathcal{O}(|m_N|) > \pi\). 
+the orbit will eventually reach an intersection point \(x(t_N) = (u_N,0)\)
+such that \(P_\mathcal{O}(|u_N|) > \pi + c{I^\star}^2\). 
 At this point, the orbit will no longer return to the
@@
 Hence, the orbit must escape any compact subset of \(\mathcal{O}\) in finite
-time, which means the acrobot is gaining energy on \(\mathcal{O}_\epsilon\).
+time. 
+We conclude that the acrobot is gaining energy on \(\mathcal{O}_\epsilon\).
 
 Setting \(I < 0\), the Poincar\'{e} section \(P_\mathcal{O}(\mu_0)\) is now
-decreasing instead of increasing,
-which means the constraint is dissipating energy.
+monotonically decreasing in \(\mu_0\).
+By the same arguments presented in this section, the constraint 
+must be dissipating energy.
 

### 2020-11-22 latex/acrobot.tex
Small edit of the oscillation proof, and updated the draft pdf

@@
 which exists because \(\bar{\mathcal{O}}_\epsilon\) is compact.
-Fruthermore, \(\underbar{\delta} \leq \epsilon\).
+Furthermore, \(\underbar{\delta} \leq \epsilon\) because
+\(x(0) = (\epsilon,0) \in \bar{\mathcal{O}}_\epsilon\).
 From the previous perturbation analysis,
@@
 
-Suppose \(x(0) \in \mathcal{O}_\epsilon\), and the orbit
-intersects the \(q_u\)-axis at some point \(x(t_0) = (u_0, 0)\).
-The Poincar\'{e} section \(P_\mathcal{O}(|u_0|)\) is positive because 
-\(|u_0| \geq \underbar{\delta}\).
+Suppose \(x(0) \in \mathcal{O}_\epsilon\) and that the orbit
+intersects the \(q_u\)-axis at some point \(x(t_0) = (u_0, 0)\), which
+has pseudo-radius \(\mu = \abs{u_0}\).
+The Poincar\'{e} section \(P_\mathcal{O}(\abs{u_0})\) is positive because 
+\(\abs{u_0} \geq \underbar{\delta}\).
 As the acrobot continues its oscillation, it will intersect the \(q_u\) axis
@@
 
-Setting \(p_1 := P_\mathcal{O}(|u_0|)\),
+Setting \(p_1 := P_\mathcal{O}(\abs{u_0})\),
 Theorem \ref{thm:khalil-perturbation} guarantees that,
@@
 \[
-    \abs{|u_1| - p_1} \leq cI^2
+    \abs{\abs{u_1} - p_1} \leq cI^2
     ,
@@
 where \(c > 0\) is an independent constant.
-This implies that \(|u_1| \geq p_1 - cI^2\). 
+This implies that \(\abs{u_1} \geq p_1 - cI^2\). 
 We also know that
 \[
-    p_1 - |u_0| \geq IKQ(|u_0|) \geq IKQ(\underbar{\delta}) > 0
+    p_1 - \abs{u_0} \geq IKQ(\abs{u_0}) \geq IKQ(\underbar{\delta}) > 0
     .
@@
 \begin{align*}
-    |u_1| - |u_0| 
-    &= \left(|u_1| - p_1\right) + \left(p_1 - |u_0|\right)
+    \abs{u_1} - \abs{u_0} 
+    &= \left(\abs{u_1} - p_1\right) + \left(p_1 - \abs{u_0}\right)
     \\
@@
 next intersection with the \(q_u\) axis will be at
-\(x(t_{i+1}) = (u_{i+1},0)\) with \(|u_{i+1}| > |u_i|\).
+\(x(t_{i+1}) = (u_{i+1},0)\) with \(\abs{u_{i+1}} > \abs{u_i}\).
 
@@
 the orbit will eventually reach an intersection point \(x(t_N) = (u_N,0)\)
-such that \(P_\mathcal{O}(|u_N|) > \pi + c{I^\star}^2\). 
+such that \(P_\mathcal{O}(\abs{u_N}) > \pi + c{I^\star}^2\). 
 At this point, the orbit will no longer return to the
 \(q_u\)-axis.
-Instead, it will escape \(\mathcal{O}\) entirely; 
-in doing so, it also escapes \(\mathcal{O}_\epsilon\).
-Hence, the orbit must escape any compact subset of \(\mathcal{O}\) in finite
-time. 
+Instead, it will escape \(\mathcal{O}\) entirely, which means
+it also escapes \(\mathcal{O}_\epsilon\).
+Hence, the orbit must escape any compact subset of 
+\(\mathcal{O}_\epsilon\) in finite time. 
 We conclude that the acrobot is gaining energy on \(\mathcal{O}_\epsilon\).
@@
 By the same arguments presented in this section, the constraint 
-must be dissipating energy.
+must be dissipating energy on \(\mathcal{O}_\epsilon\).
 

### 2020-11-23 latex/acrobot.tex
First rewrite of rotation proof

@@
     Suppose \(\int \limits_{0}^{2\pi} b(\sigma,\rho_0)d\sigma > 0\) for all 
-    \(\rho_0 \in \, \left]\sqrt{60m^2gl^3}, \bar{\rho}\right[\).
+    \(\rho_0 \in \, \left]\sqrt{60m^2gl^3}, \bar{\rho}\right]\).
     Then for all \(\epsilon > 0\) there exists \(I > 0\) small enough that
@@
     \]
-    If instead \(I < 0\), the VNHC dissipates energy on \(\Omega\).
+    If instead \(I < 0\), the VNHC dissipates energy on \(\Omega_\epsilon\).
 \end{enumerate}
@@
     \hat{\mu}_0(0) = \mu_0
-    ,
+   .,
 \end{cases}
@@
 We define the Poincar\'{e} section 
-\(P_\mathcal{O}(\mu_0) := \hat{\mu}(\pi,\mu_0)\)
+\(P_\mathcal{O}(\mu_0) := \hat{\mu}(\pi,\mu_0,I)\)
 corresponding to the change in \(\hat{\mu}\) after half an oscillation.
@@
 \]
-We wish to find a new set of coordinates \((\beta,\rho)\) for the acrobot
-and perform a similar analysis to what we did for oscillations.
+Figure \ref{fig:pendulum-level-sets} shows that orbits of the 
+pendulum on \(\mathcal{R}\) look like open curves.
+Performing a similar process to what we did for oscillations,
+we wish to find a new set of coordinates \((\beta,\rho)\)
+where \(\rho\) remains constant on level sets of \(E\)
+and \(\beta \in \Sone\) is always increasing on \(\mathcal{R}\).
+
 In the oscillation region \(\mathcal{O}\) we showed that the nominal pendulum's
@@
 \(\mu\) with the \(q_u\)-axis;
-when the nominal pendulum is rotating, the energy is instead determined by the
-intersection \(\rho\) on the \(p_u\)-axis.
-Note that the boundary of \(\mathcal{O}\) intersects the \(p_u\) axis when 
-\(p_u^2 = 60m^2 g l^3\); 
-hence, for a rotation we must have \(\rho > \sqrt{60 m^2 g l^3}\) if the
+likewise, on \(\mathcal{R}\) the energy is determined by the
+intersection \(\rho\) on the \(p_u\)-axis, as in Figure
+\ref{fig:rho-intersection}.
+
+\begin{figure}
+    \centering
+    %TODO: Add a figure with one rotation mark a distance rho away from the
+    %origin along the p-axis. Draw the lower boundary of R and shade the
+    %interior, and draw a line to the rotation with angle beta.
+    %\includestandalone[]{images/rho_intersection}
+    \caption{The domain \(\mathcal{R}\) where a pendulum rotates. The
+        pseudo-radius \(\rho\) corresponds to the intersection of an orbit of
+        rotation with the \(p_u\)-axis. The pseudo-angle \(\beta\) is taken to
+        be clockwise positive, because oscillations of a pendulum move clockwise
+        on \(\mathcal{R}\).}
+    \label{fig:rho-intersection}
+\end{figure}
+
+Note that the boundary of \(\mathcal{R}\) intersects the \(p_u\) axis when 
+\(p_u^2 = 60m^2 g l^3\). 
+Hence, we must have \(\rho > \sqrt{60 m^2 g l^3}\) if the
 orbit has momentum \(p_u > 0\) 
-or \(\rho < -\sqrt{60 m^2 g l^3}\) if the orbit has momentum \(p_u < 0\).
+and \(\rho < -\sqrt{60 m^2 g l^3}\) if the orbit has momentum \(p_u < 0\).
 The energy level set associated with \(\rho\) is 
@@
 
-Inverting this relationship gives our (smooth) transformation
-\begin{equation*}
-    T : \SxR \rightarrow 
-    \Sone \times ]-\infty,-60m^2gl^3[\, \cup \, ]60m^2gl^3,\infty[ 
-    ,
-\end{equation*}
-where \(T(q_u,p_u) = (\beta,\rho)\) is defined by
+Inverting this relationship gives the smooth transformation
 \begin{align}\label{eqn:rotation-T}
@@
 \begin{align}\label{eqn:acrobot-rot-mu-dot-nom}
-    \dot{\rho}_n &= 0
+    \dot{\rho} &= 0
     , \\
     \label{eqn:acrobot-rot-alpha-dot-nom}
-    \dot{\beta}_n &=  \sign{\rho} 
+    \dot{\beta} &=  \sign{\rho} 
     \frac{\sqrt{\rho^2 - 30m^2gl^3(1 - c_\beta)}}{5ml^2}
@@
 \end{align}
-As expected, \(\dot{\beta}_n\) does not change sign on \(\mathcal{R}\) because
-a rotating pendulum does not change direction:
-if \(\rho_n > 0\), the orbit traces a curve going from \(\beta_n = -\pi\) to 
-\(\beta_n = \pi\); if \(\rho_n < 0\) the curve goes in the reverse direction.
-By continuity of the acrobot's constrained dynamics 
-\((\dot{\beta},\dot{\rho})\), we must also have 
-\(\sign{\dot{\beta}} = \sign{\rho_0}\) for small enough \(I\).
-Hence, we can use \(\beta\) as a new ``time" variable on \(\mathcal{R}\) 
+As expected, \(\dot{\beta}\) does not change sign on \(\mathcal{R}\) because the
+orbits always flow clockwise.
+If \(\rho > 0\), the rotation curve goes from \(\beta = -\pi\) to 
+\(\beta = \pi\); 
+if \(\rho < 0\), it goes from \(\beta = \pi\) to \(\beta = -\pi\).
+
+We can use \(\beta\) as a new ``time" variable on \(\mathcal{R}\) 
 by performing a time transformation \(t = \tau(\beta)\).
@@
 \]
-For a perturbation approach, be expand the solution of these dynamics into
+Using the perturbation method, we expand the solution of these dynamics into
 \begin{equation}\label{eqn:acrobot-rhohat-approx}
-    \hat{\rho}(\beta,\rho_0,I) = \hat{\rho}_0(\beta,\rho_0) + I
-    \hat{\rho}_1(\beta,\rho_0) + R(\beta,\rho_0,I)
-    ,
-\end{equation}
-where \(R(\beta,\rho_0,I)\) is \(O(I^2)\).
+    \hat{\rho}(\beta,\rho_0,I) = \hat{\rho}_0(\beta,\rho_0) +
+    I\hat{\rho}_1(\beta,\rho_0) + R(\beta,\rho_0,I)
+    ,
+\end{equation}
+where \(R(\beta,\rho_0,I)\) is smooth and \(O(I^2)\).
 
@@
 
-Supposing we initialize the acrobot at \((\beta,\rho) = (0,\rho_0)\),
-then one full rotation amounts to \(\beta\) traversing \(2\pi\)rad in a clockwise
+Suppose we initialize the acrobot at \((\beta,\rho) = (0,\rho_0)\).
+One full rotation amounts to \(\beta\) traversing \(2\pi\)rad in a clockwise
 direction (\ie \(\beta\) going from \(0\) to \(\sign{\rho_0}2\pi\)).
-We define a Poincar\'{e} section as the point reached on the \(p_u\) axis after
-every rotation:
+We define the Poincar\'{e} section
+\(P_\mathcal{R}(\rho_0) := \hat{\rho}(\sign{\rho_0}2\pi,\rho_0,I)\),
+which is the intersection point on the \(p_u\)-axis after a single rotation.
+This expands into
 \[
     P_\mathcal{R}(\rho_0) = \rho_0 + I \hat{\rho}_1(\sign{\rho_0}2\pi, \rho_0)
-      + R(2\pi,I)
-    .
-\]
-If \(|P_\mathcal{R}(\rho_0)| > |\rho_0|\), then the rotation has
-driven the orbit further from the origin.
-In other words, \(|p_u|\) has increased across the rotation.
-If this is true for any \(\rho_0\), the acrobot is gaining momentum, which in
-turn means it is gaining energy.
-
-The function \(b(\beta,\rho_0)\) is even in both its inputs, and is 
-\(2\pi\)-periodic in \(\beta\).
-Hence, we have 
+      + R(2\pi,\rho_0,I)
+    .
+\]
+
+The function \(b(\beta,\rho_0)\) is even and is 
+\(2\pi\)-periodic in \(\beta\), which implies that
 \begin{align*}
@@
 \end{align*}
-Furthermore, \(\hat{\rho}_1(2\pi,\rho_0) = \hat{\rho}_1(2\pi,|\rho_0|)\).
+Furthermore, \(b(\beta,\rho_0)\) is also even in \(\rho_0\), 
+so \(\hat{\rho}_1(2\pi,\rho_0) = \hat{\rho}_1(2\pi,\abs{\rho_0})\).
 This implies that
-\[
-    |P_\mathcal{R}(\rho_0)| = \left|\sign{\rho_0}\left(
-        |\rho_0| + I \hat{\rho}_1(2\pi,|\rho_0|)\right)
-    \right|
-    = P_\mathcal{R}(|\rho_0|)
-    .
-\]
-Hence, if \(P_\mathcal{R}(|\rho_0|) > |\rho_0|\) (or equivalently if
-\(\hat{\rho}_1(2\pi,|\rho_0|) > 0\))
-the phase \((q_u(t),p_u(t))\) initialized at \((0,\rho_0)\) 
-will intersect the \(p_u\) axis at a point \(p > \rho_0\).
-If we could guarantee that \(\hat{\rho}_1(2\pi,\rho_0) > 0\)
-for all \(\rho_0 > \sqrt{60m^2gl^3}\), then the acrobot would be spiraling away
-from the origin on \(\mathcal{R}\).
-
-Unfortunately, we cannot adimensionalize
-\(b(\beta,\rho_0)\) like we did for oscillations. 
-Using the assumption , we have that 
-\(\hat{\rho}_1(2\pi,\rho_0) > 0 \) for all 
-\(\rho_0 \in ]\sqrt{60m^2gl^3},\rho_0[\).
-We conclude that, for small enough \(I\), the Poincar\'{e} section 
-\(P_\mathcal{R}(\rho_0)\) is monotonically increasing when 
-\(0 < \rho_0 < \bar{\rho}\) and
-monotonically decreasing when \(-\bar{\rho} < \rho_0 < 0\).
+\begin{align*}
+    \abs{P_\mathcal{R}(\rho_0)} &= \left|\sign{\rho_0}\left(
+        \abs{\rho_0} + I \hat{\rho}_1(2\pi,\abs{\rho_0})
+        + \sign{\rho_0}R(2\pi,\rho_0,I)\right)\right|
+    \\
+    &= \left|\abs{\rho_0} + I \hat{\rho}_1(2\pi,\abs{\rho_0})
+        + \sign{\rho_0}R(2\pi,\rho_0,I)\right|
+    .
+\end{align*}
+
+Let now \(\bar{\rho} > \sqrt{60m^2gl^3}\) and take the compact set
+\[
+    \bar{\mathcal{R}} := \left\{(q_u,p_u) \in \Gamma 
+    \mid E(\pi,0) \leq E(q_u,p_u) \leq E(0,\bar{\rho}) \right\}
+    .
+\]
+The (modified) remainder term \(\sign{\rho_0}{R}(\beta,\rho_0,I)\) 
+is smooth and \(O(I^2)\), which means
+it attains its minimum \(I^2\hat{r}\) somewhere \(\bar{\mathcal{R}}\).
+Likewise, \(\hat{\rho}_1(2\pi,\abs{\rho_0})\) also attains its minimum, which we
+denote by \(\underbar{\rho}_1\).
+We cannot adimensionalize \(b(\beta,\rho_0)\), so refer to one of the 
+assumptions of Theorem \ref{thm:acrobot-energy-stabilization}
+to suppose  that \(\hat{\rho}_1(2\pi,\rho_0) > 0 \) for all 
+\(\rho_0 \in ]\sqrt{60m^2gl^3},\bar{\rho}]\).
+In particular, this means \(\underbar{\rho}_1 > 0\).
+
+There exists \(I > 0\) making 
+\(I\underbar{\rho}_1 + I^2\hat{r} > 0\).
+Hence,
+\begin{align*}
+    \abs{\rho_0} + I\hat{\rho}_1(2\pi,\abs{\rho_0}) + \sign{\rho_0}R(2\pi,\rho_0,I)
+    &\geq \sqrt{60m^2gl^3} + I\underbar{\rho}_1 + I^2\hat{r}
+    \\
+    &> \sqrt{60m^2gl^3}
+    \\
+    &> 0
+    ,
+\end{align*}
+which implies that
+\begin{align*}
+    \abs{P_\mathcal{R}(\rho_0)} - \abs{\rho_0}
+    &= \abs{ \abs{\rho_0} + I\hat{\rho}_1(2\pi,\rho_0) +
+    \sign{\rho_0}R(2\pi,\rho_0,I)} - \abs{\rho_0}
+    \\
+    &= \abs{\rho_0} + I\hat{\rho}_1(2\pi,\rho_0) +
+    \sign{\rho_0}R(2\pi,\rho_0,I) - \abs{\rho_0}
+    \\
+    &= I\hat{\rho}_1(2\pi,\rho_0) + \sign{\rho_0}R(2\pi,\rho_0,I)
+    \\
+    &> 0
+    .
+\end{align*}
+
+We conclude that there exists \(I^\prime(\bar{\rho})\)
+small enough that, for all \( 0 < I \leq I^\prime(\bar{\rho})\), 
+the Poincar\'{e} section \(P_\mathcal{R}(\rho_0)\) is monotonically increasing
+when \(\sqrt{60m^2gl^3} < \rho_0 < \bar{\rho}\) and
+monotonically decreasing when \(-\bar{\rho} < \rho_0 < -\sqrt{60m^2gl^3}\).
 

### 2020-11-25 latex/acrobot.tex
Partial update of oscillation proof

@@
 However, the behaviour looks similar enough that the constraint should provide a
-decent foundation for injecting energy into the acrobot.
+reasonable foundation for injecting energy into the acrobot.
 It is for this reason that we choose our acrobot's constraint to be
@@
     \[
-        E(q_u,p_u) := \frac{p_u^2}{10ml^2} + 3(1 - \cos(q_u))
+        E(q_u,p_u) := \frac{p_u^2}{10ml^2} + 3mgl(1 - \cos(q_u))
         ,
@@
 \end{thm}
-%\footnotetext{In fact, we can be more precise than this. 
-%    Letting \(\epsilon > 0\), we need 
-%    \(I \leq \min\left\{I_1(\epsilon),I_2\}\) where
-%    \(I_1(\epsilon) = \frac{\tan(\epsilon)}{(\pi-\epsilon)}
-%    \left(\frac{m^2gl^3(\cos(\epsilon)-2\sin(\epsilon))}{\pi-\epsilon}
-%    \left(\frac{2\bar{q}_a\tan(\epsilon)}{(1+\tan(\epsilon)^2)(\pi-\epsilon)}+5\right) 
-%    \right)^{-1/2}\)
-%    and \(I_2 = \sqrt{\frac{3}{10 m^2gl^3\bar{q}_a^2}}\).
-%    See Apendix \ref{apx:I-bounds} for details.
-%}
+
 The first result of Theorem \ref{thm:acrobot-energy-stabilization} states
@@
 will gain enough energy to begin rotating around the bar.
-An acrobot satisfying the second result will additionally achieve giants with momentum
-\(|p_u| < \bar{\rho}\).
+Setting \(\epsilon > 0\) arbitrarily small, a fixed value
+\(I(\epsilon)\) will enable the acrobot to gain energy and escape any compact
+subset of \(\mathcal{O}_\epsilon\) in finite time. 
+What's more, the acrobot's phase will flow towards the heteroclinic orbit of the
+nominal pendulum, as seen in Figure \ref{fig:O-epsilon}.
+
+If we want to guarantee the acrobot can reach rotations beyond a given level set
+of \(E\), 
+An acrobot satisfying the second result of Theorem
+\ref{thm:acrobot-energy-stabilization} is guaranteed to achieve giants with
+momentum \(|p_u| < \bar{\rho}\).
+This is true because the acrobot escapes all compact sets of \(\Omega_\epsilon\)
+in finite time, as seen in Figure \ref{fig:Omega-epsilon}.
+Note that, for this result, the parameter \(I\) depend on both \(\epsilon\) and
+the target momentum \(\bar{\rho}\).
+
+\begin{figure}
+    \centering
+    % TODO: show an image of two subfigures: one is O_epsilon, with a spiraling
+    % orbit heading towards the heteroclinic orbit E(pi,0).
+    % The second is Omega_epsilon, with a spiral leaving O_epsilon and becoming
+    % a set of rotations that approaches E(0,rho_bar)
+    \begin{subfigure}[t]{0.45\textwidth}
+        % TODO: 
+        %includestandalone[]{images/O_epsilon}
+        \caption{The set \(\mathcal{O}_\epsilon\). Orbits starting in this set
+            will reach the level set \(E(\pi,0)\) of the nominal pendulum in
+            finite time.}
+        \label{fig:O-epsilon}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.45\textwidth}
+        %TODO:
+        %includestandalone[]{images/Omega_epsilon}
+        \caption{The set \(\Omega_\epsilon\). Orbits starting in this set will
+            reach the level set \(E(0,\bar{\rho})\) of the nominal pendulum in
+            finite time.}
+        \label{fig:Omega-epsilon}
+    \end{subfigure}
+    \caption{The sets defined by Theorem \ref{thm:acrobot-energy-stabilization}
+        on which the acrobot gains energy.}
+\end{figure}
 
@@
 Proving Theorem \ref{thm:acrobot-energy-stabilization} requires an intelligent
-change of coordinates and the use Khalil's perturbation
+change of coordinates and the use perturbation
 theory \cite{khalil_nonlinear}.
 We provide the full proof of this theorem in Chapter \ref{sec:acrobot-proof}.
-%and conclude this section with a high level sketch of the proof.
-%
-%There is a suitable change of coordinates into a pseudo-radius 
-%\(\mu \in \R_{>0}\) and pseudo-angle \(\alpha \in \Sone\) on
-%\(\mathcal{O}_\epsilon\).
-%These coordinates have the property that \(\dot{\alpha} > 0\) when
-%\(I\) is small enough.
-%Perturbation theory shows that the radius \(\mu\)
-%increases on average along \(\alpha(t)\).
-%Hence, orbits in the \((q_u,p_u)\)-plane spiral away from the origin, 
-%which means the acrobot is gaining energy.
-%A similar approach can be used for \(\Omega_\epsilon\).
-
-\section{Proving the Acrobot Gains Energy}\label{sec:acrobot-proof}
-Proving Theorem \ref{thm:acrobot-energy-stabilization} is not a simple task.
-In an effort to make the proof as clear as possible, we will break it down into
-the following segments:
+
+\section{Proof of Theorem \ref{thm:acrobot-energy-stabilization}}\label{sec:acrobot-proof}
+In an effort to make the proof of Theorem \ref{thm:acrobot-energy-stabilization}
+as clear as possible, we will break it down into the following segments:
 \begin{enumerate}
@@
     \item Perturbation analysis for oscillations.
-    \item Energy gain for oscillations.
     \item Perturbation analysis for rotations.
-    \item Energy gain for rotations.
 \end{enumerate}
@@
 \begin{equation}\label{eqn:acrobot-nominal-E}
-    E(q_u,p_u) = \frac{p_u^2}{10ml^2} + 3(1 - \cos(q_u))
+    E(q_u,p_u) = \frac{p_u^2}{10ml^2} + 3mgl(1 - \cos(q_u))
     ,
@@
         \caption{Level sets of the pendulum energy function.
-            The blue ellipse represents oscillations while the red curve
-            represents rotations.}
+            The blue ellipse represents oscillations; the red curve
+            represents rotations; and the black ellipse is the heteroclinic
+            orbit with energy \(E(\pi,0)\) forming the boundary of the
+            oscillation and rotation regions.}
         \label{fig:pendulum-level-sets}
@@
     \item \(\alpha\) is a pseudo-angle living in \(\Sone\) which is
-        always increasing.
+        always increasing along solutions.
 \end{itemize}
+The coordinates \((\alpha,\mu)\) can be thought of as deformed polar coordinates
+adapted to level sets of \(E\) below the value \(E(\pi,0)\), 
+where \(\mu\) specifies the level set and \(\alpha\) identifies a point on that
+level set.
 Once we have these coordinates in hand, we can use perturbation theory to prove
 that \(\mu\) increases on \(\mathcal{O}_\epsilon\) when \(I > 0\) is small enough.
+
 Likewise, we will find a second set of coordinates \((\beta, \rho)\) where
@@
 
-\subsection*{Background on Perturbation Theory}
+\subsection{Background on Perturbation Theory}
 Nonlinear systems like the acrobot are difficult (or even impossible) to solve
@@
 Solutions of the nonlinear system can often be approximated by taking a Taylor
-expansion around this nominal system.
+expansion around a solution of the nominal system.
 
@@
 which we need to solve for the explicit solution \(x_0(t,\eta_0)\) on \([t_0,t_1]\).
-One can use this to compute the solution \(x_1(t,\eta_0)\) of
+
+Now consider the solution \(x(t,\eta_0,I)\) to ~\eqref{eqn:khalil-setup}.
+Performing a first-order Taylor series expansion at \(I = 0\), we get
+\(x(t,\eta_0,I) = x_0(t,\eta_0) + I x_1(t,\eta_0) + R(t,\eta_0,I)\), where the
+remainder term \(R(t,\eta_0,I)\) is smooth and \(O(I^2)\), \ie,
+\[
+    \lim \limits_{I \to 0} \frac{R(t,I)}{I} = 0
+    .
+\]
+One can show \(x_1(t,\eta_0)\) is the solution to the scalar time-varying ODE
 \begin{equation}\label{eqn:khalil-perturbation-firstorder}
@@
 \end{equation}
-Finally, one can compute the first-order Taylor approximation 
-\(x(t,\eta_0,I) = x_0(t,\eta_0) + I x_1(t,\eta_0) + R(t,\eta_0,I)\), where the remainder
-term \(R(t,\eta_0,I)\) is smooth and \(O(I^2)\), \ie,
-\[
-    \lim \limits_{I \to 0} \frac{R(t,I)}{I} = 0
-    .
-\]
+
 We now paraphrase Khalil's Theorem 10.1 \cite{khalil_nonlinear} on the
@@
 \begin{thm}\label{thm:khalil-perturbation}
+    Let \(t_0 < t_1 \in \mathbb{R}\).
     Suppose \(f : [t_0,t_1] \times D \times [-I_0,I_0] \rightarrow \R^2\) is
-    \(C^1\), and that the nominal system ~\eqref{eqn:khalil-perturbation-nominal} has a
-    unique solution \(x_0(t,\eta_0) \in D\) on \([t_0,t_1]\).
-    Then there exists \(I^\star > 0\) such that, for all \(I\) with \(|I| < I^\star\), the
-    solution \(x(t,\eta_0,I)\) to ~\eqref{eqn:khalil-setup} satisfies
+    \(C^1\), and that the nominal system
+    ~\eqref{eqn:khalil-perturbation-nominal} has a unique solution
+    \(x_0(t,\eta_0) \in D\) on \([t_0,t_1]\).
+    Then there exists \(I^\star > 0\) such that, for all \(I\) with 
+    \(|I| < I^\star\), the solution \(x(t,\eta_0,I)\) to
+    ~\eqref{eqn:khalil-setup} satisfies
     \[
@@
 
-\subsection*{Perturbation Analysis for Oscillations}
+\subsection{Perturbation Analysis for Oscillations}
 As we have seen, setting \(I = 0\) turns the acrobot into a nominal pendulum. 
@@
 \(\mathcal{O}\).
+
+\subsubsection*{Pseudo-Polar Coordinates}
 
@@
 One can confirm that \(\dot{\alpha} > 0\) for every \(\mu \in ]0,\pi[\).
-By continuity of ~\eqref{eqn:acrobot-alpha-dot}, 
-\(\dot{\alpha}\) remains positive on \(\mathcal{O}\) for \(I\) small enough.
-
-Suppose at time \(t = 0\) we have initialized the acrobot at \((\alpha,\mu) = (0,\mu_0)\).
-Since \(\alpha\) is continuously increasing, we know that the orbit will do one
-of two things: it will either leave \(\mathcal{O}\) so that the acrobot is now rotating,
-or it will remain in \(\mathcal{O}\) and hit the \(q_u\)-axis at some point 
-\((\alpha,\mu) = (\pi,u)\).
-From there, the acrobot could escape \(\mathcal{O}\) along the second half of the orbit or
-return to some point \((\alpha,\mu) = (0,\tilde{u})\) (as illustrated in Figure
-\ref{fig:acrobot-possible-orbits}).
-If \(\tilde{u} < \mu_0\) then the acrobot has lost energy; 
-if \(\tilde{u} = \mu_0\), the acrobot is on a closed orbit;
-and if \(\tilde{u} > \mu_0\) the acrobot has gained energy.
-Which situation occurs might depend entirely on the choice of
-\(\mu_0\).
-
-\begin{figure}
-    \centering 
-    % TODO: Draw the domain mathcalO, an initial point (q0,p0) = (mu0,0), and show the
-    % half orbit hitting the left q-axis at (-m,0).
-    % Then show the 
-    % 4 clockwise half orbits starting at (-m,0): 
-    % one orbit exits mathcalO on the top right corner;
-    % one orbit hits the right q-axis at (mu0+x,0); 
-    % one hits (mu0,0);
-    % and one hits (mu0-x,0) for some x > 0.
-    %includestandalone[width=0.5\textwdith]{images/acrobot_possible_orbits}
-    \caption{If the acrobot is initialized at \((q_u,p_u) = (\mu_0,0)\), then
-        either the orbit escapes \(\mathcal{O}\) into the rotation
-        zone (purple), or the orbit returns to the \(q_u\)-axis 
-        with a higher (red), equal (black), or lower (blue) pseudo-radius than when it
-        started.}
-    \label{fig:acrobot-possible-orbits}
-\end{figure}
-
-We can simplify our analysis of the acrobot by studying how its orbits
-intersect the \(q_u\)-axis on average.
-Since \(\dot{\alpha} > 0\), we could theoretically find some time scaling 
-\(t = \tau(\alpha)\) and use \(\alpha\) as a new ``time" variable.
-Setting \(\hat{\mu}(\alpha) := \mu(\tau(\alpha))\) allows us to ignore the
-dynamics of \(\alpha\) by observing the evolution of \(\mu\) as a function of
-\(\alpha\) rather than a function of time (assuming the acrobot remains in
-\(\mathcal{O}\)).
-The dynamics of \(\hat{\mu}\) are
+By continuity of ~\eqref{eqn:acrobot-alpha-dot}, there exists \(I_1 > 0\) small
+enough that \(\dot{\alpha}\) remains positive on \(\mathcal{O}\) for 
+\(I \in [-I_1,I_1]\).
+
+\subsubsection*{Time Scaling}
+
+%Suppose at time \(t = 0\) we have initialized the acrobot at \((\alpha,\mu) = (0,\mu_0)\).
+%Since \(\alpha\) is continuously increasing, we know that the orbit will do one
+%of two things: it will either leave \(\mathcal{O}\) so that the acrobot is now rotating,
+%or it will remain in \(\mathcal{O}\) and hit the \(q_u\)-axis at some point 
+%\((\alpha,\mu) = (\pi,q_1)\).
+%From there, the acrobot could escape \(\mathcal{O}\) along the second half of the orbit or
+%return to some point \((\alpha,\mu) = (0,\tilde{u})\) (as illustrated in Figure
+%\ref{fig:acrobot-possible-orbits}).
+%If \(\tilde{u} < \mu_0\) then the acrobot has lost energy; 
+%if \(\tilde{u} = \mu_0\), the acrobot is on a closed orbit;
+%and if \(\tilde{u} > \mu_0\) the acrobot has gained energy.
+%Which situation occurs might depend entirely on the choice of
+%\(\mu_0\).
+%
+%\begin{figure}
+%    \centering 
+%    % TODO: Draw the domain mathcalO, an initial point (q0,p0) = (mu0,0), and show the
+%    % half orbit hitting the left q-axis at (-m,0).
+%    % Then show the 
+%    % 4 clockwise half orbits starting at (-m,0): 
+%    % one orbit exits mathcalO on the top right corner;
+%    % one orbit hits the right q-axis at (mu0+x,0); 
+%    % one hits (mu0,0);
+%    % and one hits (mu0-x,0) for some x > 0.
+%    %includestandalone[width=0.5\textwdith]{images/acrobot_possible_orbits}
+%    \caption{If the acrobot is initialized at \((q_u,p_u) = (\mu_0,0)\), then
+%        either the orbit escapes \(\mathcal{O}\) into the rotation
+%        zone (purple), or the orbit returns to the \(q_u\)-axis 
+%        with a higher (red), equal (black), or lower (blue) pseudo-radius than when it
+%        started.}
+%    \label{fig:acrobot-possible-orbits}
+%\end{figure}
+
+We have observed that \(\dot{\alpha} > 0\) on \(\mathcal{O}\) for small enough \(I\). 
+Hence, we can use \(\alpha\) as our time variable by reparameterizing \(t\) as a
+function \(t = \tau(\alpha)\).
+This allows us to study the evolution of \(\mu\) as a function of \(\alpha\)
+rather than a function of time.
+Setting \(\hat{\mu}(\alpha) := \mu(\tau(\alpha))\) yields the dynamics
 \[
     \diff{\hat{\mu}}{\alpha} = 
-    \left.\diff{\mu}{t} \diff{t}{\alpha} \right|_{t =\tau(\alpha)}
-    ,
-\] 
-yielding the ``time"-varying scalar ODE
+    \diff{\mu}{t} \diff{t}{\alpha}
+    .
+\] 
+This reduces the system \((\dot{\alpha},\dot{\mu})\) into the time-varying
+scalar ODE 
 \begin{equation}\label{eqn:muhat-dot}
@@
 
-In the spirit of perturbation analysis, we expand the solution of
-~\eqref{eqn:muhat-dot} into
+\subsubsection*{Perturbation Analysis of the Time Scaled System}
+
+In the spirit of perturbation analysis, we find the nominal system 
+\(\hat{\mu}_0(\alpha,\mu_0)\) which,
+by ~\eqref{eqn:khalil-perturbation-nominal}, is the solution to 
+\[
+\begin{cases}
+    \dot{\hat{\mu}}_0 = g(\alpha,\hat{\mu}_0,0)
+    , \\
+    \hat{\mu}_0(0) = \mu_0
+   .
+\end{cases}
+\]
+Since \(g(\alpha,\mu,0) = 0\) by 
+~\eqref{eqn:acrobot-mu-dot-nom}-\eqref{eqn:acrobot-alpha-dot-nom}, we find that
+\(\hat{\mu}_0(\alpha,\mu_0) \equiv \mu_0\) for all \(\alpha\).
+
+We now take a first-order Taylor approximation of \(\hat{\mu}(\alpha,\mu_0,I)\)
+approximation around the nominal system \(\hat{\mu}_0(\alpha,\mu_0)\), which 
+expands into 
 \begin{equation}\label{eqn:acrobot-muhat-approx}
@@
 where \(R(\alpha,\mu_0,I)\) is smooth and \(O(I^2)\).
-From ~\eqref{eqn:khalil-perturbation-nominal} we know that 
-\(\hat{\mu}_0\) is the solution to the nominal system
-\[
-\begin{cases}
-    \dot{\hat{\mu}}_0 = g(\alpha,\mu,0)
-    , \\
-    \hat{\mu}_0(0) = \mu_0
-   .,
-\end{cases}
-\]
-Since \(g(\alpha,\mu,0) = 0\) by 
-~\eqref{eqn:acrobot-mu-dot-nom}-\eqref{eqn:acrobot-alpha-dot-nom}, we find that
-\(\hat{\mu}_0(\alpha,\mu_0) \equiv \mu_0\) for all \(\alpha\).
-
-Likewise, ~\eqref{eqn:khalil-perturbation-firstorder} tells us that 
-\(\hat{\mu}_1\) is the solution to
+
+By ~\eqref{eqn:khalil-perturbation-firstorder}, we know the function
+\(\hat{\mu}_1(\alpha,\mu_0)\) is the solution to the linear time-varying scalar
+system
 \begin{equation}\label{eqn:acrobot-mu1-dot}
@@
 \]
-
-Suppose we initialize the acrobot at \((\alpha,\mu) = (0,\mu_0)\). 
-After half an oscillation, when \(\alpha = \pi\), the orbit will land on the
-negative \(q_u\)-axis with pseudo-radius 
-\(u = \hat{\mu}(\pi,\mu_0) \approx \mu_0 + I\hat{\mu}_1(\pi,\mu_0)\).
-Hence, the change in pseudo-radius is 
-\(u - \mu_0 \approx I\hat{\mu}_1(\pi,\mu_0)\).
-If the orbit remains in \(\mathcal{O}\) for the full oscillation, 
-it will land on the positive \(q_u\)-axis
-with pseudo-radius
-\(\tilde{u} \approx \mu_0 + I\hat{\mu}_1(2\pi,\mu_0)\).
-
-Notice that \(b(\alpha,\mu_0)\) is even and 
-\(\pi\)-periodic in \(\alpha\), so that
-\begin{equation}\label{eqn:bhat-integral-properties}
-    \int \limits_{-\pi}^\pi b(\sigma,\mu_0) d\sigma
-    = \int \limits_0 ^ {2 \pi} b(\sigma,\mu_0) d\sigma
-    = 2 \int \limits_0^\pi b(\sigma,\mu_0) d\sigma
-    .
-\end{equation}
-This implies that 
-\(\hat{\mu}_1(2\pi,\mu_0) = 2 \hat{\mu}_1(\pi,\mu_0)\),
-which means the change in pseudo-radius after a full oscillation is 
-\(\tilde{u} - \mu_0 \approx 2I\hat{\mu}_1(\pi,\mu_0)\).
-If \(\hat{\mu}_1(\pi,\mu_0) > 0\), 
-then both \(u - \mu_0 > 0\) and \(\tilde{u} - \mu_0 > 0\).
-In other words, we can study the average behaviour of \(\hat{\mu}\) by examining
-the behaviour at each half oscillation.
-
-We define the Poincar\'{e} section 
-\(P_\mathcal{O}(\mu_0) := \hat{\mu}(\pi,\mu_0,I)\)
-corresponding to the change in \(\hat{\mu}\) after half an oscillation.
-The section expands into
-\[
-    P_\mathcal{O}(\mu_0) = \mu_0 + I \hat{\mu}_1(\pi,\mu_0) + R(\pi,\mu_0,I)
+Therefore, the first-order Taylor approximation of \(\hat{\mu}(\alpha,\mu_0,I)\) is
+\begin{equation}\label{eqn:acrobot-muhat-expanded}
+    \hat{\mu}(\alpha,\mu_0,I)
+    = \mu_0 + I K \int \limits_0^\alpha b(\sigma,\mu_0)d\sigma +
+    R(\alpha,\mu_0,I)
+\end{equation}
+
+\subsubsection{Poincar\'{e} Analysis}
+
+%Suppose we initialize the acrobot at \((\alpha,\mu) = (0,\mu_0)\). 
+%After half an oscillation, when \(\alpha = \pi\), the orbit will land on the
+%negative \(q_u\)-axis with pseudo-radius 
+%\(u = \hat{\mu}(\pi,\mu_0) \approx \mu_0 + I\hat{\mu}_1(\pi,\mu_0)\).
+%Hence, the change in pseudo-radius is 
+%\(u - \mu_0 \approx I\hat{\mu}_1(\pi,\mu_0)\).
+%If the orbit remains in \(\mathcal{O}\) for the full oscillation, 
+%it will land on the positive \(q_u\)-axis
+%with pseudo-radius
+%\(\tilde{u} \approx \mu_0 + I\hat{\mu}_1(2\pi,\mu_0)\).
+%
+%Notice that \(b(\alpha,\mu_0)\) is even and 
+%\(\pi\)-periodic in \(\alpha\), so that
+%\begin{equation}\label{eqn:bhat-integral-properties}
+%    \int \limits_{-\pi}^\pi b(\sigma,\mu_0) d\sigma
+%    = \int \limits_0 ^ {2 \pi} b(\sigma,\mu_0) d\sigma
+%    = 2 \int \limits_0^\pi b(\sigma,\mu_0) d\sigma
+%    .
+%\end{equation}
+%This implies that 
+%\(\hat{\mu}_1(2\pi,\mu_0) = 2 \hat{\mu}_1(\pi,\mu_0)\),
+%which means the change in pseudo-radius after a full oscillation is 
+%\(\tilde{u} - \mu_0 \approx 2I\hat{\mu}_1(\pi,\mu_0)\).
+%If \(\hat{\mu}_1(\pi,\mu_0) > 0\), 
+%then both \(u - \mu_0 > 0\) and \(\tilde{u} - \mu_0 > 0\).
+%In other words, we can study the average behaviour of \(\hat{\mu}\) by examining
+%the behaviour at each half oscillation.
+
+Since \(\dot{\alpha} > 0\), there is a well-defined Poincar\'{e} map 
+showing how \(\mu\) changes every time the orbit of \((q_u(t),p_u(t))\) hits the
+\(q_u\)-axis.
+This intersection with the \(q_u\)-axis is represented by \(\alpha = n\pi\) for
+\(n \in \mathbb{Z}_{\geq 0}\).
+Using our time-scaled system, we can define this Poincar\'{e} map as the change
+in \(\hat{\mu}\) at each of these intersections: 
+\begin{align*}
+    P_\mathcal{O} : ]0,\pi[ &\rightarrow \R_+,
+    \\
+    \mu_0 \mapsto \hat{\mu}(\pi,\mu_0,I)
+    .
+\end{align*}
+From our perturbation analysis, the section expands into
+\[
+    P_\mathcal{O}(\mu_0) = \mu_0 + I K \int\limits_0^\pi b(\sigma,\mu_0)d\sigma
+    + R(\pi,\mu_0,I)
     .
@@
 
-We numerically compute \(Q(\mu_0)\) for \(\mu_0 \in [10^{-5}, \pi - 10^{-5}]\)
+We numerically compute \(Q(\mu_0)\) for \(\mu_0 \in [10^{-10}, \pi - 10^{-3}]\)
 in Figure \ref{fig:acrobot-Q}. 
@@
 an asymptote at \(\mu_0 = \pi\). 
-Simulations with smaller \(\mu_0\) vanish in MATLAB, so we believe that 
-\(Q(\mu_0)\) is in fact positive for all \(\mu_0\).
+Simulations with smaller \(\mu_0\) result in an infinite integration error,
+so we believe that \(Q(\mu_0)\) is in fact positive for all \(\mu_0\).
 
@@
 
+\subsubsection*{Energy Gain on \(\mathcal{O}\)}
+
+The Poincar\'{e} section \(P_\mathcal{O}\) allows us understand the evolution of
+\(\hat{\mu}(\alpha,\mu_0,I)\) by studying the evolution of a discrete time system;
+abusing notation, we set the state of this discrete time system to be
+\(\hat{\mu}(n)\) for \(n \in \mathbb{Z}_{\geq 0}\). 
+The dynamics of \(\hat{\mu}(n)\) are
+\begin{equation}\label{eqn:muhat-discrete}
+    \begin{cases}
+        \hat{\mu}(n+1) = P_\mathcal{O}(\hat{\mu}(n)) 
+        = \hat{\mu}(n) + I K Q(\hat{\mu}(n)) + R(\pi,\hat{\mu}(n),I)
+        , \\
+        \hat{\mu}(0) = \mu_0
+        .
+    \end{cases}
+\end{equation}
+
+To show energy gain on \(\mathcal{O}\), it is sufficient to 
+show this discrete time system is non-decreasing.
+That is, we wish to show there exists \(I^\star \in ]0, I_1]\) where, for all 
+\(\delta > 0\) and \(\mu_0 \in ]0,\pi-\delta]\), there exists \(N > 0\) so that
+for all \(n \geq N\), \(\hat{\mu}(n) \notin ]0,\pi-\delta]\).
+
+% TODO: Replace from here
 This fact, coupled with knowledge that \(R(\pi,\mu_0,I)\) is \(O(I^2)\), 

### 2020-11-25 latex/acrobot.tex
New proof of oscillation energy gain, needs edits next

@@
     \item For all \(\epsilon > 0\), there exists \(I > 0\) small enough that 
-    ~\eqref{eqn:acrobot-constraint} injects energy into the acrobot on the set
-    \[
-        \mathcal{O}_\epsilon := \left\{(q_u,p_u) \in \Gamma 
-        \mid E(\epsilon,0) < E(q_u,p_u) < E(\pi,0) \right\}
-        .
-    \]
-    If instead \(I < 0\), the VNHC dissipates energy on \(\mathcal{O}_\epsilon\).
+    ~\eqref{eqn:acrobot-constraint} injects energy into the acrobot on
+    \[
+        \mathcal{O} := \left\{(q_u,p_u) \in \Gamma 
+        \mid E(q_u,p_u) < E(\pi,0) \right\}
+        .
+    \]
+    If instead \(I < 0\), the VNHC dissipates energy on \(\mathcal{O}\).
 \item Define \(b : \SxR_{> 0} \rightarrow \R\) by
@@
     \(\rho_0 \in \, \left]\sqrt{60m^2gl^3}, \bar{\rho}\right]\).
-    Then for all \(\epsilon > 0\) there exists \(I > 0\) small enough that
+    Then there exists \(I > 0\) small enough that
     ~\eqref{eqn:acrobot-constraint} injects energy into the acrobot on
     \[
-        \Omega_\epsilon := \left\{(q_u,p_u) \in \Gamma 
-        \mid E(\epsilon,0) < E(q_u,p_u) < E(0,\bar{\rho})\right\}
-        .
-    \]
-    If instead \(I < 0\), the VNHC dissipates energy on \(\Omega_\epsilon\).
+        \Omega := \left\{(q_u,p_u) \in \Gamma 
+        \mid E(q_u,p_u) < E(0,\bar{\rho})\right\}
+        .
+    \]
+    If instead \(I < 0\), the VNHC dissipates energy on \(\Omega\).
 \end{enumerate}
@@
 remainder term \(R(t,\eta_0,I)\) is smooth and \(O(I^2)\), \ie,
-\[
-    \lim \limits_{I \to 0} \frac{R(t,I)}{I} = 0
+there exists \(c > 0\) such that 
+\[
+    \norm{ R(t,\eta_0,I) } \leq c \abs{I^2}
+    ,
+\]
+for all \((t,\eta_0,I) \in [t_0,t_1] \times D \times [-I_0,I_0]\), and
+\[
+    \lim \limits_{I \to 0} \frac{R(t,\eta_0,I)}{I} = 0
     .
@@
 When \(I\) is small enough, solutions of the nonlinear system and
-the Taylor approximation are the same up to order \(I^2\) along compact time
+the Taylor approximation are the same up to order \(O(I^2)\) along compact time
 intervals.
@@
 
+This fact, coupled with knowledge that \(R(\pi,\mu_0,I)\) is \(O(I^2)\), 
+implies there exists \(I > 0\) small enough that
+\[
+    P_\mathcal{O}(\mu_0) = \mu_0 + IKQ(\mu_0) + R(\pi,\mu_0,I) > 0
+    .
+\]
+Unfortunately, perturbation analysis does not guarantee 
+that \(I\) is the same for all \(\mu_0 \in \, ]0,\pi[\), since the value of
+\(R(\pi,\mu_0,I)\) could become more negative for other initial conditions.
+We will deal with this next.
+
 \subsubsection*{Energy Gain on \(\mathcal{O}\)}
@@
 abusing notation, we set the state of this discrete time system to be
-\(\hat{\mu}(n)\) for \(n \in \mathbb{Z}_{\geq 0}\). 
-The dynamics of \(\hat{\mu}(n)\) are
+\(\mu(n)\) for \(n \in \mathbb{Z}_{\geq 0}\). 
+The dynamics of \(\mu(n)\) are
 \begin{equation}\label{eqn:muhat-discrete}
     \begin{cases}
-        \hat{\mu}(n+1) = P_\mathcal{O}(\hat{\mu}(n)) 
-        = \hat{\mu}(n) + I K Q(\hat{\mu}(n)) + R(\pi,\hat{\mu}(n),I)
-        , \\
-        \hat{\mu}(0) = \mu_0
+        \mu(n+1) = P_\mathcal{O}(\mu(n)) 
+        = \mu(n) + I K Q(\mu(n)) + R(\pi,\mu(n),I)
+        , \\
+        \mu(0) = \mu_0
         .
@@
 To show energy gain on \(\mathcal{O}\), it is sufficient to 
-show this discrete time system is non-decreasing.
-That is, we wish to show there exists \(I^\star \in ]0, I_1]\) where, for all 
+show this discrete time system is strictly increasing.
+This is because \(\mu(n)\) represents the distance along the \(q_u\)-axis
+when an orbit intersects the \(q_u\)-axis for the \(n\)th time, assuming
+the orbit was initialized at \((q_u,p_u) = (\mu_0,0)\).
+
+Formally, we wish to show there exists \(I^\star \in ]0, I_1]\) where, for all 
 \(\delta > 0\) and \(\mu_0 \in ]0,\pi-\delta]\), there exists \(N > 0\) so that
-for all \(n \geq N\), \(\hat{\mu}(n) \notin ]0,\pi-\delta]\).
-
-% TODO: Replace from here
-This fact, coupled with knowledge that \(R(\pi,\mu_0,I)\) is \(O(I^2)\), 
-implies there exists \(I > 0\) small enough that
-\[
-    P_\mathcal{O}(\mu_0) = \mu_0 + IKQ(\mu_0) + R(\pi,\mu_0,I) > 0
-    .
-\]
-Unfortunately, this \(I\) may not be the
-same for all \(\mu_0 \in \, ]0,\pi[\), since
-the value of \(R(\pi,\mu_0,I)\) could become more negative
-for other initial conditions.
-We deal with this issue by taking the compact set
-\[
-    \bar{\mathcal{O}} := \left\{(q_u,p_u) \in \Gamma 
-    \mid E(q_u,p_u) \leq E(\pi,0) \right\}
-    .
-\]
-The remainder term \(R(\alpha,\mu_0,I)\) is smooth and of order \(I^2\),
-which means it can be written in the form 
-\(R(\alpha,\mu_0,I) = I^2\tilde{R}(\alpha,\mu_0,I)\)
-for some \(\tilde{R}\) a smooth function.
-Since \(\bar{\mathcal{O}}\) is compact, \(\tilde{R}\) attains its minimum
-\({r}\) somewhere on \(\bar{\mathcal{O}}\).
-Hence, \(R(\alpha,\mu_0,I)\) is bounded below on \(\mathcal{O}\) by
-\(I^2{r}\), which implies 
-\begin{equation}\label{eqn:acrobot-Po-lowerbound}
-    P_\mathcal{O}(\mu_0) \geq \mu_0 + IKQ(\mu_0) + I^2{r}
-    .
-\end{equation}
-This equation further implies that, for fixed \(I\),
-\(P_\mathcal{O}(\mu_0)\) is monotonically increasing in \(\mu_0\).
-Thus, for every \(\epsilon > 0\) and \(\mu_0 \in \, [\epsilon,\pi[\) we have
-\[
-    P_\mathcal{O}(\mu_0) \geq P_\mathcal{O}(\epsilon)
-    = \epsilon + IKQ(\epsilon) + I^2{r}
-    .
-\]
-Since \(I\) is assumed to be small, \(I^2\) shrinks rapidly.
-Therefore, we can always find a small enough \(I > 0\) making 
-\(\epsilon + I^2{r} \geq 0\), so that
-\[
-    P_\mathcal{O}(\epsilon) \geq IKQ(\epsilon) > 0
-    .
-\]
-
-We conclude that, for all \(m\), \(g\), \(l\), \(\bar{q}_a\), and for
-any \(\epsilon > 0\), there exists \(I^\prime(\epsilon)\) small enough so that,
-for all \(0 < I \leq I^\prime(\epsilon)\),
-the Poincar\'{e} section \(P_\mathcal{O}(\mu_0)\) is positive for all
-\(\mu_0 \in [\epsilon,\pi[\) and 
-\(\lim \limits_{\mu_0 \to \pi} P_\mathcal{O}(\mu_0) = \infty\).
-
-\subsection*{Energy gain on \(\mathcal{O}_\epsilon\)}
-To simplify notation, denote by \(x(t) := (q_u(t),p_u(t))\) the orbit of the
-acrobot with initial condition \(x(0)\).
-Let \(\epsilon > 0\) and define
-\[
-    \mathcal{O}_\epsilon := \left\{(q_u,p_u) \in \Gamma
-    \mid E(\epsilon,0) < E(q_u,p_u) < E(\pi,0)\right\}
-    ,
-\]
-with \(\bar{\mathcal{O}}_\epsilon\) its (compact) closure.
-
-Let \(x(0) \in \mathcal{O}_\epsilon \subset \mathcal{O}\).
-Since \(\dot{\alpha} > 0\), the orbit \(x(t)\) spirals clockwise around
-the origin.
-If this orbit remains in \(\mathcal{O}\), it must intersect
-the \(q_u\) axis at some radius \(\mu = \delta\)
-which depends on the initial condition \(x(0)\).
-Take 
-\[
-    \underbar{\delta} = 
-    \min\limits_{x(0) \in \bar{\mathcal{O}}_\epsilon} \delta(x(0))
-    ,
-\] 
-which exists because \(\bar{\mathcal{O}}_\epsilon\) is compact.
-Furthermore, \(\underbar{\delta} \leq \epsilon\) because
-\(x(0) = (\epsilon,0) \in \bar{\mathcal{O}}_\epsilon\).
-From the previous perturbation analysis,
-there exists \(I^\prime(\underbar{\delta})\) small enough that
-\(P_\mathcal{O}(\mu_0)\) is positive for all 
-\(\mu_0 \in [\underbar{\delta},\pi[\).
-
-Suppose \(x(0) \in \mathcal{O}_\epsilon\) and that the orbit
-intersects the \(q_u\)-axis at some point \(x(t_0) = (u_0, 0)\), which
-has pseudo-radius \(\mu = \abs{u_0}\).
-The Poincar\'{e} section \(P_\mathcal{O}(\abs{u_0})\) is positive because 
-\(\abs{u_0} \geq \underbar{\delta}\).
-As the acrobot continues its oscillation, it will intersect the \(q_u\) axis
-again at a new point \(x(t_1) = (u_1,0)\).
-
-Setting \(p_1 := P_\mathcal{O}(\abs{u_0})\),
-Theorem \ref{thm:khalil-perturbation} guarantees that,
-for all \(I \in ]0,I^\prime(\underbar{\delta})]\),
-\[
-    \abs{\abs{u_1} - p_1} \leq cI^2
-    ,
-\]
-where \(c > 0\) is an independent constant.
-This implies that \(\abs{u_1} \geq p_1 - cI^2\). 
-We also know that
-\[
-    p_1 - \abs{u_0} \geq IKQ(\abs{u_0}) \geq IKQ(\underbar{\delta}) > 0
-    .
-\]
-There exists \(I^\star \in ]0,I^\prime(\underbar{\delta})]\) where
-\(c{I^\star}^2 < I^\star KQ(\underbar{\delta})\), which means
-\begin{align*}
-    \abs{u_1} - \abs{u_0} 
-    &= \left(\abs{u_1} - p_1\right) + \left(p_1 - \abs{u_0}\right)
-    \\
-    &\geq \left(p_1 - c{I^\star}^2 - p_1\right) + I^\star KQ(\underbar{\delta})
-    \\
-    &= I^\star KQ(\underbar{\delta}) - c{I^\star}^2
-    \\
-    &> 0
-    .
-\end{align*}
-
-Since \(u_0 \in [\underbar{\delta},\pi[\) was arbitrary, 
-we can apply Theorem \ref{thm:khalil-perturbation} iteratively:
-initializing the acrobot at \(x(t_i) = (u_i, 0)\), the
-next intersection with the \(q_u\) axis will be at
-\(x(t_{i+1}) = (u_{i+1},0)\) with \(\abs{u_{i+1}} > \abs{u_i}\).
-
-Because the Poincar\'{e} section is monotonically increasing, 
-the orbit will eventually reach an intersection point \(x(t_N) = (u_N,0)\)
-such that \(P_\mathcal{O}(\abs{u_N}) > \pi + c{I^\star}^2\). 
-At this point, the orbit will no longer return to the
-\(q_u\)-axis.
-Instead, it will escape \(\mathcal{O}\) entirely, which means
-it also escapes \(\mathcal{O}_\epsilon\).
-Hence, the orbit must escape any compact subset of 
-\(\mathcal{O}_\epsilon\) in finite time. 
-We conclude that the acrobot is gaining energy on \(\mathcal{O}_\epsilon\).
+for all \(n \geq N\), \(\mu(n) \notin ]0,\pi-\delta]\).
+
+Linearizing ~\eqref{eqn:muhat-discrete} at \(\mu = 0\) yields
+\[
+    P^\prime_\mathcal{O}(0) = 1 + I K Q^\prime(0) +
+    \pdiff{R}{\mu}(\pi,0,I)
+    .
+\]
+If there is some value of \(I \in ]0,I_1]\) for which this term is greater than
+\(1\), then \(0\) is a repeller of the discrete time system.
+To show such a value of \(I\) exists, we first compute 
+\[
+    Q^\prime(0) = \int \limits_0^\pi \pdiff{b}{\mu}(\sigma,0) d\sigma
+    .
+\]
+MATLAB reveals that 
+\[
+    \lim \limits_{\mu_0 \to 0^+}
+    \pdiff{b}{\mu}(\alpha,\mu_0) 
+    = -\frac{\sqrt{2}}{2} \left(11\sin(\alpha)^2 - 6\right)
+    ,
+\]
+which means 
+\[
+    Q^\prime(0)  = \frac{\pi}{2\sqrt{2}} > 0
+    .
+\]
+Since the remainder term \(R(\alpha,\mu_0,I)\) is \(O(I^2)\), 
+its partial derivative \(\pdiff{R}{\mu}(\pi,\mu_0,I)\) is also \(O(I^2)\).
+Hence, it can be written in the form
+\(I^2 \tilde{R}^\prime(\mu_0,I)\) where \(\tilde{R}^\prime(\mu_0,I)\) is smooth and
+zero at \(I = 0\).
+Since \(I\) is assumed to be small, \(I^2\) shrinks rapidly;
+thus, there exists \(I_2 \in ]0,I_1]\) where 
+\(I_2 K Q^\prime(0) + I_2^2 \tilde{R}^\prime(0,I) > 0\).
+Hence,
+\[
+    P^\prime_\mathcal{O}(0) \geq 1 + IKQ^\prime(0) + I^2 \tilde{R}^\prime(0,I) 
+    > 1
+    ,
+\] 
+for all \(I \in ]0, I_2]\), .
+
+This proves \(0\) is a repeller of the discrete time system, which means
+there exists some (unknown) \(\epsilon > 0\) where the interval
+\(]0,\epsilon[\) is negatively invariant for ~\eqref{eqn:muhat-discrete}.
+What's more, all solutions starting in this interval will flow towards the value
+\(\mu = \epsilon\).
+In other words, we have found a value \(I_2 > 0\) small enough
+that \(\mu(n)\) is strictly increasing on \([0,\epsilon[\).
+
+Next, recall that \(R(\pi,\mu_0,I)\) is smooth and \(O(I^2)\)
+and is therefore bounded below on \(\mu \in [\epsilon,\pi]\) by some value
+\(-I^2r\).
+Furhtermore, \(Q(\mu_0)\) is a strictly increasing function, so
+for\(\mu_0 \in [\epsilon,\pi[\) we have
+\[
+    P_\mathcal{O}(\mu_0) \geq \mu_0 + I K Q(\epsilon) - I^2 r
+    .
+\]
+There clearly exists a small enough \(I^\star \in ]0,I_2]\) making
+\(I^\star K Q(\epsilon) - (I^\star)^2 r > 0\), which means
+\[
+    P_\mathcal{O}(\mu_0) > \mu_0
+    ,
+\]
+for all \(\mu_0 \in [\epsilon,\pi[\).
+
+We have thus shown that \(P_\mathcal{O}(\mu_0) > \mu_0\) for all
+\(\mu_0 \in ]0,\pi[\).
+Looking at ~\eqref{eqn:muhat-discrete}, this means
+\(\mu(n+1) > \mu(n)\) and all solutions of the discrete time system will flow
+towards \(\mu = \pi\).
+Hence, orbits of the constrained dynamics will eventually escape
+\(\mathcal{O}\).
+
+We conclude that, for all \(m\), \(g\), \(l\), \(\bar{q}_a\), 
+there exists \(I > 0)\) small enough that the constraint 
+~\eqref{eqn:acrobot-constraint} is injecting energy on \(\mathcal{O}\).
 
@@
 By the same arguments presented in this section, the constraint 
-must be dissipating energy on \(\mathcal{O}_\epsilon\).
+must be dissipating energy on \(\mathcal{O}\).
 
@@
 
-\subsection*{Perturbation Analysis for Rotations}
+\subsection{Perturbation Analysis for Rotations}
 The rotation region for the nominal pendulum is the set

### 2020-11-26 latex/acrobot.tex
Full edit v1 of oscillation proof, time for round twoooooo

@@
 \begin{enumerate}
-    \item For all \(\epsilon > 0\), there exists \(I > 0\) small enough that 
+    \item There exists \(I > 0\) small enough that 
     ~\eqref{eqn:acrobot-constraint} injects energy into the acrobot on
@@
 
-The first result of Theorem \ref{thm:acrobot-energy-stabilization} states
+A small enough fixed value \(I > 0\) will enable the acrobot to 
+escape any compact subset of \(\mathcal{O}\) in finite time. 
+What's more, the acrobot's phase will flow towards the heteroclinic orbit
+corresponding to \(E(\pi,0)\), as in Figure
+\ref{fig:acrobot-oscillation-domain}.
+Hence, the first result of Theorem \ref{thm:acrobot-energy-stabilization} states
 that all acrobots constrained by ~\eqref{eqn:acrobot-constraint} 
 will gain enough energy to begin rotating around the bar.
-Setting \(\epsilon > 0\) arbitrarily small, a fixed value
-\(I(\epsilon)\) will enable the acrobot to gain energy and escape any compact
-subset of \(\mathcal{O}_\epsilon\) in finite time. 
-What's more, the acrobot's phase will flow towards the heteroclinic orbit of the
-nominal pendulum, as seen in Figure \ref{fig:O-epsilon}.
-
-If we want to guarantee the acrobot can reach rotations beyond a given level set
-of \(E\), 
-An acrobot satisfying the second result of Theorem
-\ref{thm:acrobot-energy-stabilization} is guaranteed to achieve giants with
-momentum \(|p_u| < \bar{\rho}\).
-This is true because the acrobot escapes all compact sets of \(\Omega_\epsilon\)
-in finite time, as seen in Figure \ref{fig:Omega-epsilon}.
-Note that, for this result, the parameter \(I\) depend on both \(\epsilon\) and
-the target momentum \(\bar{\rho}\).
+In the worst case, the acrobot will at least perform a swing-up routine to reach
+the \((q_u,p_u) = (\pi,0)\) equilibrium.
+
+If we want the acrobot to achieve giants with energy
+\(E(0,\bar{\rho})\), the acrobot needs to satisfy the assumption on the
+integral of \(b(\beta,\rho_0)\).
+In this case, orbits will escape compact subsets of \(\Omega\) in finite time
+until the acrobot has reached a momentum of at least \(\bar{\rho}\).
+One example of this behaviour is displayed in Figure \ref{fig:acrobot-omega}.
+Note that the value of \(I\) will depend on \(\bar{\rho}\).
 
@@
     \centering
-    % TODO: show an image of two subfigures: one is O_epsilon, with a spiraling
+    % TODO: show an image of two subfigures: one is O, with a spiraling
     % orbit heading towards the heteroclinic orbit E(pi,0).
-    % The second is Omega_epsilon, with a spiral leaving O_epsilon and becoming
+    % The second is Omega, with a spiral leaving O and becoming
     % a set of rotations that approaches E(0,rho_bar)
@@
         %includestandalone[]{images/O_epsilon}
-        \caption{The set \(\mathcal{O}_\epsilon\). Orbits starting in this set
+        \caption{The set \(\mathcal{O}\). Orbits starting in this set
             will reach the level set \(E(\pi,0)\) of the nominal pendulum in
             finite time.}
-        \label{fig:O-epsilon}
+        \label{fig:acrobot-oscillation-domain}
     \end{subfigure}
@@
         %includestandalone[]{images/Omega_epsilon}
-        \caption{The set \(\Omega_\epsilon\). Orbits starting in this set will
+        \caption{The set \(\Omega\). Orbits starting in this set will
             reach the level set \(E(0,\bar{\rho})\) of the nominal pendulum in
             finite time.}
-        \label{fig:Omega-epsilon}
+            \label{fig:acrobot-omega}
     \end{subfigure}
@@
     \item \(\alpha\) is a pseudo-angle living in \(\Sone\) which is
-        always increasing along solutions.
+        always increasing along oscillations.
 \end{itemize}
@@
 Once we have these coordinates in hand, we can use perturbation theory to prove
-that \(\mu\) increases on \(\mathcal{O}_\epsilon\) when \(I > 0\) is small enough.
-
-Likewise, we will find a second set of coordinates \((\beta, \rho)\) where
-\(\beta \in \Sone\) is always increasing, \(\rho\) is constant along
-rotations of the simple pendulum, and \(\rho\) increases for the acrobot when
-\(I > 0\) is small enough.
-Putting these together, we will show the acrobot is gaining energy on
-\(\Omega_\epsilon\).
+that \(\mu\) increases on \(\mathcal{O}\) when \(I > 0\) is small enough.
+
+Likewise, we will find a second set of coordinates \((\beta, \rho)\) for
+rotations, where \(\beta \in \Sone\) is always increasing along solutions,
+\(\rho\) is constant along rotations of the simple pendulum, and \(\rho\)
+increases for the acrobot when \(I > 0\) is small enough.
+Using both \((\alpha,\mu)\) and \((\beta,\rho)\), we will show the acrobot is
+gaining energy on \(\Omega\).
 
@@
 \end{equation}
-which we need to solve for the explicit solution \(x_0(t,\eta_0)\) on \([t_0,t_1]\).
+which we need to solve for the explicit solution \(x_0(t,\eta_0)\) on
+\([t_0,t_1]\).
+We assume this solution is contained in \(D\).
 
@@
 Performing a first-order Taylor series expansion at \(I = 0\), we get
-\(x(t,\eta_0,I) = x_0(t,\eta_0) + I x_1(t,\eta_0) + R(t,\eta_0,I)\), where the
-remainder term \(R(t,\eta_0,I)\) is smooth and \(O(I^2)\), \ie,
-there exists \(c > 0\) such that 
-\[
-    \norm{ R(t,\eta_0,I) } \leq c \abs{I^2}
-    ,
-\]
-for all \((t,\eta_0,I) \in [t_0,t_1] \times D \times [-I_0,I_0]\), and
+\(x(t,\eta_0,I) = x_0(t,\eta_0) + I x_1(t,\eta_0) + R(t,\eta_0,I)\),
+where the remainder term \(R(t,\eta_0,I)\) is smooth and \(O(I^2)\), \ie, 
 \[
@@
 \]
-One can show \(x_1(t,\eta_0)\) is the solution to the scalar time-varying ODE
+One can recover \(x_1(t,\eta_0)\) as the solution to the scalar
+time-varying ODE
 \begin{equation}\label{eqn:khalil-perturbation-firstorder}
@@
 \begin{thm}\label{thm:khalil-perturbation}
-    Let \(t_0 < t_1 \in \mathbb{R}\).
+    Fix \(t_0, t_1 \in \mathbb{R}\) with \(t_0 < t_1\).
     Suppose \(f : [t_0,t_1] \times D \times [-I_0,I_0] \rightarrow \R^2\) is
@@
 When \(I\) is small enough, solutions of the nonlinear system and
-the Taylor approximation are the same up to order \(O(I^2)\) along compact time
+the Taylor approximation are the same up to order \(I^2\) along compact time
 intervals.
+This is equivalent to saying that the remainder term \(R(t,\eta_0,I)\) is
+bounded, because 
+\[
+    \norm{x(t,\eta_0,I) - \left(x_0(t,\eta_0) + I x_1(t,\eta_0)\right)} 
+    = \norm{R(t,\eta_0,I)} \leq k |I^2|
+    .
+\]
 Since we know that the acrobot behaves like a pendulum at \(I = 0\), we can
@@
 make our analysis convenient by changing coordinates into a
-``radius" \(\mu > 0\) which remains constant on level sets of \(E\), 
-along with an ``angle" \(\alpha \in \Sone\) satisfying \(\dot{\alpha} > 0\) on
+pseudo-radius \(\mu > 0\) which remains constant on level sets of \(E\), 
+along with a pseudo-angle \(\alpha \in \Sone\) satisfying \(\dot{\alpha} > 0\) on
 \(\mathcal{O}\).
@@
 Hence, we can use \(\alpha\) as our time variable by reparameterizing \(t\) as a
-function \(t = \tau(\alpha)\).
+function \(t(\alpha)\).
 This allows us to study the evolution of \(\mu\) as a function of \(\alpha\)
 rather than a function of time.
-Setting \(\hat{\mu}(\alpha) := \mu(\tau(\alpha))\) yields the dynamics
+Setting \(\hat{\mu}(\alpha) := \mu(t(\alpha))\) yields the dynamics
 \[
@@
 \] 
-This reduces the system \((\dot{\alpha},\dot{\mu})\) into the time-varying
-scalar ODE 
+This reduces the system \((\dot{\alpha},\dot{\mu})\) into the scalar
+time-varying ODE 
 \begin{equation}\label{eqn:muhat-dot}
@@
 
-In the spirit of perturbation analysis, we find the nominal system 
-\(\hat{\mu}_0(\alpha,\mu_0)\) which,
-by ~\eqref{eqn:khalil-perturbation-nominal}, is the solution to 
+In the spirit of perturbation analysis, we will expand our time-scaled system
+\(\hat{\mu}(\alpha,\mu_0,I)\).
+We begin by setting \(I = 0\) to find the nominal system
 \[
@@
 \]
-Since \(g(\alpha,\mu,0) = 0\) by 
-~\eqref{eqn:acrobot-mu-dot-nom}-\eqref{eqn:acrobot-alpha-dot-nom}, we find that
+Equations ~\eqref{eqn:acrobot-mu-dot-nom}-\eqref{eqn:acrobot-alpha-dot-nom})
+imply that \(g(\alpha,\mu,0) = 0\),
+so the solution to this nominal system is 
 \(\hat{\mu}_0(\alpha,\mu_0) \equiv \mu_0\) for all \(\alpha\).
@@
 We now take a first-order Taylor approximation of \(\hat{\mu}(\alpha,\mu_0,I)\)
-approximation around the nominal system \(\hat{\mu}_0(\alpha,\mu_0)\), which 
+around the nominal system \(\hat{\mu}_0(\alpha,\mu_0)\), which 
 expands into 
@@
     R(\alpha,\mu_0,I)
+    .
 \end{equation}
@@
 Since \(\dot{\alpha} > 0\), there is a well-defined Poincar\'{e} map 
-showing how \(\mu\) changes every time the orbit of \((q_u(t),p_u(t))\) hits the
-\(q_u\)-axis.
-This intersection with the \(q_u\)-axis is represented by \(\alpha = n\pi\) for
-\(n \in \mathbb{Z}_{\geq 0}\).
+describing how \(\mu\) changes each time the orbit of \((q_u(t),p_u(t))\)
+intersects the \(q_u\)-axis, \ie, when \(\alpha = 0\) or \(\alpha = \pi\).
 Using our time-scaled system, we can define this Poincar\'{e} map as the change
@@
 \begin{align*}
-    P_\mathcal{O} : ]0,\pi[ &\rightarrow \R_+,
+    P_\mathcal{O} : \,]0,\pi[ &\rightarrow \R_+,
     \\
-    \mu_0 \mapsto \hat{\mu}(\pi,\mu_0,I)
+    \mu_0 &\mapsto \hat{\mu}(\pi,\mu_0,I)
     .
@@
 \[
-    P_\mathcal{O}(\mu_0) = \mu_0 + IKQ(\mu_0) + R(\pi,\mu_0,I) > 0
+    P_\mathcal{O}(\mu_0) = \mu_0 + IKQ(\mu_0) + R(\pi,\mu_0,I) > \mu_0
     .
@@
 The Poincar\'{e} section \(P_\mathcal{O}\) allows us understand the evolution of
-\(\hat{\mu}(\alpha,\mu_0,I)\) by studying the evolution of a discrete time system;
-abusing notation, we set the state of this discrete time system to be
-\(\mu(n)\) for \(n \in \mathbb{Z}_{\geq 0}\). 
-The dynamics of \(\mu(n)\) are
+\(\hat{\mu}(\alpha,\mu_0,I)\) by studying the evolution of the discrete time
 \begin{equation}\label{eqn:muhat-discrete}
@@
 \end{equation}
-
-To show energy gain on \(\mathcal{O}\), it is sufficient to 
-show this discrete time system is strictly increasing.
-This is because \(\mu(n)\) represents the distance along the \(q_u\)-axis
-when an orbit intersects the \(q_u\)-axis for the \(n\)th time, assuming
-the orbit was initialized at \((q_u,p_u) = (\mu_0,0)\).
-
-Formally, we wish to show there exists \(I^\star \in ]0, I_1]\) where, for all 
+Tolerating the abuse of notation, \(\mu(n)\) represents the distance along the
+\(q_u\)-axis when an orbit of the constrained dynamics intersects the
+\(q_u\)-axis for the \(n\)th time, assuming the orbit was initialized at
+\((q_u,p_u) = (\mu_0,0)\).
+
+Show energy gain on \(\mathcal{O}\) is equivalent to 
+finding some \(I^\star \in ]0, I_1]\) where, for all 
 \(\delta > 0\) and \(\mu_0 \in ]0,\pi-\delta]\), there exists \(N > 0\) so that
 for all \(n \geq N\), \(\mu(n) \notin ]0,\pi-\delta]\).
+We will show \(P_\mathcal{O}(\mu_0) > \mu_0\) for all \(\mu_0 \in ]0,\pi[\),
+which implies this characterization of energy gain holds true.
 
@@
 
-This proves \(0\) is a repeller of the discrete time system, which means
+We have shown \(0\) is a repeller of the discrete time system, which means
 there exists some (unknown) \(\epsilon > 0\) where the interval
 \(]0,\epsilon[\) is negatively invariant for ~\eqref{eqn:muhat-discrete}.
-What's more, all solutions starting in this interval will flow towards the value
+All solutions starting in this interval will flow towards the value
 \(\mu = \epsilon\).
@@
 
-Next, recall that \(R(\pi,\mu_0,I)\) is smooth and \(O(I^2)\)
-and is therefore bounded below on \(\mu \in [\epsilon,\pi]\) by some value
-\(-I^2r\).
-Furhtermore, \(Q(\mu_0)\) is a strictly increasing function, so
+Next, recall that \(R(\pi,\mu_0,I)\) is smooth in all its parameters as well as
+being \(O(I^2)\).
+Therefore, it is bounded below on the compact set \(\mu_0 \in [\epsilon,\pi]\)
+by some value \(R(\pi,\underbar{\mu},I)\). 
+Theorem \ref{thm:khalil-perturbation} asserts that there exists some
+\(I_3 > 0\) and \(r > 0\) so that, for all \(I \in ]0, I_3]\),
+\[
+    R(\pi,\mu_0,I) \geq R(\pi, \underbar{\mu}, I) > -I^2 r
+    .
+\]
+Furthermore, \(Q(\mu_0)\) is a strictly increasing function, so
 for\(\mu_0 \in [\epsilon,\pi[\) we have
 \[
-    P_\mathcal{O}(\mu_0) \geq \mu_0 + I K Q(\epsilon) - I^2 r
-    .
-\]
-There clearly exists a small enough \(I^\star \in ]0,I_2]\) making
+    P_\mathcal{O}(\mu_0) > \mu_0 + I K Q(\epsilon) - I^2 r
+    .
+\]
+There clearly exists a small enough \(I^\star \in ]0,I_3]\) making
 \(I^\star K Q(\epsilon) - (I^\star)^2 r > 0\), which means

### 2020-11-27 latex/acrobot.tex
Completed the edit of oscillation proof

@@
 Equations ~\eqref{eqn:acrobot-mu-dot-nom}-\eqref{eqn:acrobot-alpha-dot-nom})
-imply that \(g(\alpha,\mu,0) = 0\),
+reveal that \(g(\alpha,\mu,0) = 0\),
 so the solution to this nominal system is 
@@
     , \\
-    \pdiff{g}{I}(\alpha, \mu_0, 0) &= K b(\alpha,\mu_0)
+    \pdiff{g}{I}(\alpha, \mu_0, 0) &= K a(\alpha,\mu_0)
     ,
@@
     , \\
-    b(\alpha,\mu_0) &:= \frac{
+    a(\alpha,\mu_0) &:= \frac{
         \mu_0 |\sin(\alpha)| \left(
@@
     \hat{\mu}_1(\alpha,\mu_0) =
-    K \int \limits_0^\alpha b(\sigma,\mu_0)d\sigma
+    K \int \limits_0^\alpha a(\sigma,\mu_0)d\sigma
     .
@@
     \hat{\mu}(\alpha,\mu_0,I)
-    = \mu_0 + I K \int \limits_0^\alpha b(\sigma,\mu_0)d\sigma +
+    = \mu_0 + I K \int \limits_0^\alpha a(\sigma,\mu_0)d\sigma +
     R(\alpha,\mu_0,I)
@@
 
-%Suppose we initialize the acrobot at \((\alpha,\mu) = (0,\mu_0)\). 
-%After half an oscillation, when \(\alpha = \pi\), the orbit will land on the
-%negative \(q_u\)-axis with pseudo-radius 
-%\(u = \hat{\mu}(\pi,\mu_0) \approx \mu_0 + I\hat{\mu}_1(\pi,\mu_0)\).
-%Hence, the change in pseudo-radius is 
-%\(u - \mu_0 \approx I\hat{\mu}_1(\pi,\mu_0)\).
-%If the orbit remains in \(\mathcal{O}\) for the full oscillation, 
-%it will land on the positive \(q_u\)-axis
-%with pseudo-radius
-%\(\tilde{u} \approx \mu_0 + I\hat{\mu}_1(2\pi,\mu_0)\).
-%
-%Notice that \(b(\alpha,\mu_0)\) is even and 
-%\(\pi\)-periodic in \(\alpha\), so that
-%\begin{equation}\label{eqn:bhat-integral-properties}
-%    \int \limits_{-\pi}^\pi b(\sigma,\mu_0) d\sigma
-%    = \int \limits_0 ^ {2 \pi} b(\sigma,\mu_0) d\sigma
-%    = 2 \int \limits_0^\pi b(\sigma,\mu_0) d\sigma
-%    .
-%\end{equation}
-%This implies that 
-%\(\hat{\mu}_1(2\pi,\mu_0) = 2 \hat{\mu}_1(\pi,\mu_0)\),
-%which means the change in pseudo-radius after a full oscillation is 
-%\(\tilde{u} - \mu_0 \approx 2I\hat{\mu}_1(\pi,\mu_0)\).
-%If \(\hat{\mu}_1(\pi,\mu_0) > 0\), 
-%then both \(u - \mu_0 > 0\) and \(\tilde{u} - \mu_0 > 0\).
-%In other words, we can study the average behaviour of \(\hat{\mu}\) by examining
-%the behaviour at each half oscillation.
-
 Since \(\dot{\alpha} > 0\), there is a well-defined Poincar\'{e} map 
-describing how \(\mu\) changes each time the orbit of \((q_u(t),p_u(t))\)
+describing how the pseudo-radius \(\mu\) changes each time the orbit of \((q_u(t),p_u(t))\)
 intersects the \(q_u\)-axis, \ie, when \(\alpha = 0\) or \(\alpha = \pi\).
 Using our time-scaled system, we can define this Poincar\'{e} map as the change
-in \(\hat{\mu}\) at each of these intersections:
+in \(\hat{\mu}\) when \(\alpha\) increases by \(\pi\):
 \begin{align*}
@@
 \end{align*}
-From our perturbation analysis, the section expands into
-\[
-    P_\mathcal{O}(\mu_0) = \mu_0 + I K \int\limits_0^\pi b(\sigma,\mu_0)d\sigma
+From our perturbation analysis, the Poincar\'{e} map -- also known as a
+Poincar\'{e} section --  expands into
+\[
+    P_\mathcal{O}(\mu_0) = \mu_0 + I K \int\limits_0^\pi a(\sigma,\mu_0)d\sigma
     + R(\pi,\mu_0,I)
@@
 \[
-    Q(\mu_0) := \int \limits_0^\pi b(\sigma,\mu_0)d\sigma
+    Q(\mu_0) := \int \limits_0^\pi a(\sigma,\mu_0)d\sigma
     ,
@@
 Note that \(K\) is a positive constant which contains the acrobot's physical
-parameters, while \(b(\alpha,\mu)\) is adimensional. 
+parameters, while \(a(\alpha,\mu)\) is adimensional. 
 This means \(Q(\mu_0)\) is identical for every acrobot.
@@
 an asymptote at \(\mu_0 = \pi\). 
-Simulations with smaller \(\mu_0\) result in an infinite integration error,
-so we believe that \(Q(\mu_0)\) is in fact positive for all \(\mu_0\).
+Simulations with smaller \(\mu_0\) result in an infinite integral error due
+to a division by zero, so we believe that \(Q(\mu_0)\) is in fact positive for
+all \(\mu_0\).
 
@@
 \(\hat{\mu}(\alpha,\mu_0,I)\) by studying the evolution of the discrete time
+system
 \begin{equation}\label{eqn:muhat-discrete}
     \begin{cases}
-        \mu(n+1) = P_\mathcal{O}(\mu(n)) 
+        \mu(n+1) := P_\mathcal{O}(\mu(n)) 
         = \mu(n) + I K Q(\mu(n)) + R(\pi,\mu(n),I)
@@
 
-Show energy gain on \(\mathcal{O}\) is equivalent to 
-finding some \(I^\star \in ]0, I_1]\) where, for all 
+Show energy gain on \(\mathcal{O}\) is equivalent to showing \(\mu(n)\)
+eventually reaches \(\pi\).
+That is, we want to find \(I^\star \in ]0, I_1]\) where, for all 
 \(\delta > 0\) and \(\mu_0 \in ]0,\pi-\delta]\), there exists \(N > 0\) so that
-for all \(n \geq N\), \(\mu(n) \notin ]0,\pi-\delta]\).
-We will show \(P_\mathcal{O}(\mu_0) > \mu_0\) for all \(\mu_0 \in ]0,\pi[\),
-which implies this characterization of energy gain holds true.
+for all \(n \geq N\), \(\mu(n) \notin [0,\pi-\delta]\).
+A sufficient condition for this characterization is to show 
+\(P_\mathcal{O}(\mu_0) > \mu_0\) for all \(\mu_0 \in ]0,\pi[\),
 
@@
 \[
-    Q^\prime(0) = \int \limits_0^\pi \pdiff{b}{\mu}(\sigma,0) d\sigma
+    Q^\prime(0) = \int \limits_0^\pi \pdiff{a}{\mu}(\sigma,0) d\sigma
     .
@@
     \lim \limits_{\mu_0 \to 0^+}
-    \pdiff{b}{\mu}(\alpha,\mu_0) 
+    \pdiff{a}{\mu}(\alpha,\mu_0) 
     = -\frac{\sqrt{2}}{2} \left(11\sin(\alpha)^2 - 6\right)
@@
 Hence, it can be written in the form
-\(I^2 \tilde{R}^\prime(\mu_0,I)\) where \(\tilde{R}^\prime(\mu_0,I)\) is smooth and
+\(I^2 \tilde{R}(\mu_0,I)\) where \(\tilde{R}(\mu_0,I)\) is smooth and
 zero at \(I = 0\).
@@
 thus, there exists \(I_2 \in ]0,I_1]\) where 
-\(I_2 K Q^\prime(0) + I_2^2 \tilde{R}^\prime(0,I) > 0\).
+\(I_2 K Q^\prime(0) + (I_2)^2 \tilde{R}(0,I) > 0\).
 Hence,
 \[
-    P^\prime_\mathcal{O}(0) \geq 1 + IKQ^\prime(0) + I^2 \tilde{R}^\prime(0,I) 
+    P^\prime_\mathcal{O}(0) \geq 1 + IKQ^\prime(0) + I^2 \tilde{R}(0,I) 
     > 1
@@
 \(]0,\epsilon[\) is negatively invariant for ~\eqref{eqn:muhat-discrete}.
-All solutions starting in this interval will flow towards the value
+What's more, all solutions starting in this interval will flow towards the value
 \(\mu = \epsilon\).
 In other words, we have found a value \(I_2 > 0\) small enough
-that \(\mu(n)\) is strictly increasing on \([0,\epsilon[\).
+that \(\mu(n)\) is strictly increasing on \(]0,\epsilon[\).
 
@@
 Theorem \ref{thm:khalil-perturbation} asserts that there exists some
-\(I_3 > 0\) and \(r > 0\) so that, for all \(I \in ]0, I_3]\),
+\(I_3 > 0\) and \(r > 0\) so that, for all \(I \in [-I_3, I_3]\),
 \[
@@
 \]
-There clearly exists a small enough \(I^\star \in ]0,I_3]\) making
+There exists a small enough \(I^\star \in ]0,I_3]\) making
 \(I^\star K Q(\epsilon) - (I^\star)^2 r > 0\), which means
@@
 We conclude that, for all \(m\), \(g\), \(l\), \(\bar{q}_a\), 
-there exists \(I > 0)\) small enough that the constraint 
+there exists \(I > 0\) small enough that the constraint 
 ~\eqref{eqn:acrobot-constraint} is injecting energy on \(\mathcal{O}\).
 
-Setting \(I < 0\), the Poincar\'{e} section \(P_\mathcal{O}(\mu_0)\) is now
-monotonically decreasing in \(\mu_0\).
-By the same arguments presented in this section, the constraint 
-must be dissipating energy on \(\mathcal{O}\).
+Setting \(I < 0\), the Poincar\'{e} section now satisfies
+\[
+    P_\mathcal{O}(\mu_0) < \mu_0
+    ,
+\]
+thanks to the same arguments presented in this section. 
+The constraint is therefore dissipating energy on \(\mathcal{O}\).
 

### 2020-11-28 latex/acrobot.tex
Completed the edit of oscillation proof, ready to send to Manfredi

@@
 escape any compact subset of \(\mathcal{O}\) in finite time. 
-What's more, the acrobot's phase will flow towards the heteroclinic orbit
+What's more, the acrobot's phase will flow towards the homoclinic orbit
 corresponding to \(E(\pi,0)\), as in Figure
@@
 In this case, orbits will escape compact subsets of \(\Omega\) in finite time
-until the acrobot has reached a momentum of at least \(\bar{\rho}\).
+until the acrobot eventually reaches a momentum of at least \(\bar{\rho}\).
 One example of this behaviour is displayed in Figure \ref{fig:acrobot-omega}.
@@
     \centering
-    % TODO: show an image of two subfigures: one is O, with a spiraling
-    % orbit heading towards the heteroclinic orbit E(pi,0).
-    % The second is Omega, with a spiral leaving O and becoming
-    % a set of rotations that approaches E(0,rho_bar)
     \begin{subfigure}[t]{0.45\textwidth}
-        % TODO: 
-        %includestandalone[]{images/O_epsilon}
-        \caption{The set \(\mathcal{O}\). Orbits starting in this set
+        \includestandalone[width=\linewidth]{images/acrobot_oscillation_domain}
+        \caption{The set \(\mathcal{O}\). An orbit starting in this set (blue)
             will reach the level set \(E(\pi,0)\) of the nominal pendulum in
@@
     \begin{subfigure}[t]{0.45\textwidth}
-        %TODO:
-        %includestandalone[]{images/Omega_epsilon}
+        \includestandalone[width=\linewidth]{images/acrobot_omega}
         \caption{The set \(\Omega\). Orbits starting in this set will
@@
     \begin{subfigure}[t]{0.45\textwidth}
-        % TODO: figure of acrobot with I = 0
-        %\includestandalone[]{images/acrobot_I0}
+        \includestandalone[]{images/acrobot_I_zero}
         \caption{A simple pendulum with two masses.}
@@
     \hfill
-    \begin{subfigure}[t]{0.45\textwidth}
-        % TODO: figure showing oscillations and rotations as level curves of E
-        %\includestandalone[]{images/pendulum_level_sets}
-        \caption{Level sets of the pendulum energy function.
-            The blue ellipse represents oscillations; the red curve
-            represents rotations; and the black ellipse is the heteroclinic
-            orbit with energy \(E(\pi,0)\) forming the boundary of the
-            oscillation and rotation regions.}
+    \begin{subfigure}[t]{0.5\textwidth}
+        \includestandalone[]{images/pendulum_level_sets}
+        \caption{Level sets of \(E\).
+            The blue ellipse is an oscillation; the red curves
+            are rotations; and the black ellipse is the homoclinic
+            orbit with energy \(E(\pi,0)\).}
         \label{fig:pendulum-level-sets}
@@
     \centering
-    % TODO: Add a figure with one oscillation mark at a distance mu from
-    % the origin to the intersection on the q-axis. We also draw the boundary of
-    % mathcalO and shade the interior, and draw a line to the oscillation with
-    % clockwise angle alpha
-    %\includestandalone[]{images/mu_intersection}
-    \caption{The domain \(\mathcal{O}\) where a pendulum oscillates.
+    \includestandalone[width=0.5\textwidth]{images/mu_intersection}
+    \caption{The domain \(\mathcal{O}\) (blue) where a pendulum oscillates.
     The pseudo-radius \(\mu\) corresponds to the
-    intersection of an orbit of oscillation with the \(q_u\)-axis.
+    intersection of an oscillation (red) with the \(q_u\)-axis.
     The psuedo-angle \(\alpha\) is taken to be clockwise positive because 
@@
 
-%Suppose at time \(t = 0\) we have initialized the acrobot at \((\alpha,\mu) = (0,\mu_0)\).
-%Since \(\alpha\) is continuously increasing, we know that the orbit will do one
-%of two things: it will either leave \(\mathcal{O}\) so that the acrobot is now rotating,
-%or it will remain in \(\mathcal{O}\) and hit the \(q_u\)-axis at some point 
-%\((\alpha,\mu) = (\pi,q_1)\).
-%From there, the acrobot could escape \(\mathcal{O}\) along the second half of the orbit or
-%return to some point \((\alpha,\mu) = (0,\tilde{u})\) (as illustrated in Figure
-%\ref{fig:acrobot-possible-orbits}).
-%If \(\tilde{u} < \mu_0\) then the acrobot has lost energy; 
-%if \(\tilde{u} = \mu_0\), the acrobot is on a closed orbit;
-%and if \(\tilde{u} > \mu_0\) the acrobot has gained energy.
-%Which situation occurs might depend entirely on the choice of
-%\(\mu_0\).
-%
-%\begin{figure}
-%    \centering 
-%    % TODO: Draw the domain mathcalO, an initial point (q0,p0) = (mu0,0), and show the
-%    % half orbit hitting the left q-axis at (-m,0).
-%    % Then show the 
-%    % 4 clockwise half orbits starting at (-m,0): 
-%    % one orbit exits mathcalO on the top right corner;
-%    % one orbit hits the right q-axis at (mu0+x,0); 
-%    % one hits (mu0,0);
-%    % and one hits (mu0-x,0) for some x > 0.
-%    %includestandalone[width=0.5\textwdith]{images/acrobot_possible_orbits}
-%    \caption{If the acrobot is initialized at \((q_u,p_u) = (\mu_0,0)\), then
-%        either the orbit escapes \(\mathcal{O}\) into the rotation
-%        zone (purple), or the orbit returns to the \(q_u\)-axis 
-%        with a higher (red), equal (black), or lower (blue) pseudo-radius than when it
-%        started.}
-%    \label{fig:acrobot-possible-orbits}
-%\end{figure}
-
 We have observed that \(\dot{\alpha} > 0\) on \(\mathcal{O}\) for small enough \(I\). 
@@
     + R(\alpha,\mu_0,I)
-    ,
-\end{equation}
-where \(R(\alpha,\mu_0,I)\) is smooth and \(O(I^2)\).
+    .
+\end{equation}
 
@@
     \centering
-    % TODO: The plot of Q(mu) from MATLAB
-    %\includegraphics[width=0.5\textwidth]{images/Qmu.png}
+    \includegraphics[width=0.8\textwidth]{images/Qmu.png}
     \caption{The plot of \(Q(\mu_0)\).}
@@
 
-Show energy gain on \(\mathcal{O}\) is equivalent to showing \(\mu(n)\)
-eventually reaches \(\pi\).
-That is, we want to find \(I^\star \in ]0, I_1]\) where, for all 
+Proving the constrained dynamics gain energy on \(\mathcal{O}\) is equivalent to
+showing \(\mu(n)\) eventually reaches \(\pi\).
+That is, we want to find \(I^\star \in\, ]0, I_1]\) where, for all 
 \(\delta > 0\) and \(\mu_0 \in ]0,\pi-\delta]\), there exists \(N > 0\) so that
@@
 A sufficient condition for this characterization is to show 
-\(P_\mathcal{O}(\mu_0) > \mu_0\) for all \(\mu_0 \in ]0,\pi[\),
+\(P_\mathcal{O}(\mu_0) > \mu_0\) for all \(\mu_0 \in\, ]0,\pi[\),
 
@@
 What's more, all solutions starting in this interval will flow towards the value
-\(\mu = \epsilon\).
-In other words, we have found a value \(I_2 > 0\) small enough
-that \(\mu(n)\) is strictly increasing on \(]0,\epsilon[\).
+\(\mu = \epsilon\), so we must have \(P_\mathcal{O}(\mu_0) > \mu_0\) for all 
+\(\mu_0 \in \, ]0,\epsilon[\).
 
@@
 \]
-There exists a small enough \(I^\star \in ]0,I_3]\) making
-\(I^\star K Q(\epsilon) - (I^\star)^2 r > 0\), which means
-\[
-    P_\mathcal{O}(\mu_0) > \mu_0
-    ,
-\]
+Choosing \(I^\star \in\, ]0,I_3]\) so that
+\(I^\star K Q(\epsilon) - (I^\star)^2 r > 0\)
+means that
+\( P_\mathcal{O}(\mu_0) > \mu_0\)
 for all \(\mu_0 \in [\epsilon,\pi[\).
@@
 \(\mathcal{O}\).
-
 We conclude that, for all \(m\), \(g\), \(l\), \(\bar{q}_a\), 
@@
 
-Setting \(I < 0\), the Poincar\'{e} section now satisfies
-\[
-    P_\mathcal{O}(\mu_0) < \mu_0
-    ,
-\]
-thanks to the same arguments presented in this section. 
-The constraint is therefore dissipating energy on \(\mathcal{O}\).
+By the same arguments presented in this section, the Poincar\'{e} section
+satisfies \( P_\mathcal{O}(\mu_0) < \mu_0 \) when \(I < 0\). 
+In this case, the constraint dissipates energy on \(\mathcal{O}\).
 

### 2020-11-30 latex/acrobot.tex
Partial edit of rotations, working on the new proof.

@@
 \]
-Choosing \(I^\star \in\, ]0,I_3]\) so that
+Choosing \(I^\star \in\, ]0,\min\left\{I_2,I_3\right\}]\) so that
 \(I^\star K Q(\epsilon) - (I^\star)^2 r > 0\)
@@
 \subsection{Perturbation Analysis for Rotations}
-The rotation region for the nominal pendulum is the set
+Figure \ref{fig:pendulum-level-sets} reminds us that rotations of the 
+nominal pendulum obtained by setting \(I = 0\) look like open curves on some
+rotation region \(\mathcal{R} \subset \Gamma\).
+One can compute this rotation domain to be the set
 \[
@@
 \]
-Figure \ref{fig:pendulum-level-sets} shows that orbits of the 
-pendulum on \(\mathcal{R}\) look like open curves.
 Performing a similar process to what we did for oscillations,
 we wish to find a new set of coordinates \((\beta,\rho)\)
-where \(\rho\) remains constant on level sets of \(E\)
-and \(\beta \in \Sone\) is always increasing on \(\mathcal{R}\).
+where the pseudo-radius \(\rho \in \mathbb{R}\) remains constant on level sets
+of \(E\) and the pseudo-angle \(\beta \in \Sone\) is always increasing on
+\(\mathcal{R}\).
+
+\subsubsection*{Pseudo-Polar Coordinates}
 
@@
 \(\mu\) with the \(q_u\)-axis;
-likewise, on \(\mathcal{R}\) the energy is determined by the
-intersection \(\rho\) on the \(p_u\)-axis, as in Figure
-\ref{fig:rho-intersection}.
+likewise, on \(\mathcal{R}\) the energy is determined by the intersection
+\(\rho\) on the \(p_u\)-axis, as in Figure \ref{fig:rho-intersection}.
 
@@
     \centering
-    %TODO: Add a figure with one rotation mark a distance rho away from the
-    %origin along the p-axis. Draw the lower boundary of R and shade the
-    %interior, and draw a line to the rotation with angle beta.
-    %\includestandalone[]{images/rho_intersection}
-    \caption{The domain \(\mathcal{R}\) where a pendulum rotates. The
+    \includestandalone[]{images/rho_intersection}
+    \caption{The domain \(\mathcal{R}\) (blue) where a pendulum rotates. The
         pseudo-radius \(\rho\) corresponds to the intersection of an orbit of
-        rotation with the \(p_u\)-axis. The pseudo-angle \(\beta\) is taken to
-        be clockwise positive, because oscillations of a pendulum move clockwise
-        on \(\mathcal{R}\).}
+        rotation with the \(p_u\)-axis. The pseudo-angle \(\beta\) selects a
+        point on the rotation.}
     \label{fig:rho-intersection}
@@
 \(p_u^2 = 60m^2 g l^3\). 
-Hence, we must have \(\rho > \sqrt{60 m^2 g l^3}\) if the
-orbit has momentum \(p_u > 0\) 
-and \(\rho < -\sqrt{60 m^2 g l^3}\) if the orbit has momentum \(p_u < 0\).
+The level set of this boundary is \(E(\pi,0)\), the homoclinic orbit for the
+upright equilibrium of the pendulum.
+Hence, we must have \(\rho > \sqrt{60 m^2 g l^3}\) if a
+rotation has momentum \(p_u > 0\) 
+and \(\rho < -\sqrt{60 m^2 g l^3}\) if it has momentum \(p_u < 0\).
 The energy level set associated with \(\rho\) is 
@@
 if \(\rho < 0\), it goes from \(\beta = \pi\) to \(\beta = -\pi\).
-
-We can use \(\beta\) as a new ``time" variable on \(\mathcal{R}\) 
-by performing a time transformation \(t = \tau(\beta)\).
-This produces the time-scaled system \(\hat{\rho}(\beta) = \rho(\tau(\beta))\)
-with dynamics
+By continuity of ~\eqref{eqn:rotation-T}, there exists \(I_1 > 0\) small enough
+that \(\dot{\beta}\) is non-zero and does not change sign on \(\mathcal{R}\) for
+all \(I \in [-I_1,I_1]\).
+
+\subsubsection*{Time Scaling}
+
+Using \(\beta\) as our new time variable produces the time-scaled system
+\(\hat{\rho}(\beta) = \rho(\tau(\beta))\) with dynamics
 \[
@@
 \]
-Using the perturbation method, we expand the solution of these dynamics into
-\begin{equation}\label{eqn:acrobot-rhohat-approx}
-    \hat{\rho}(\beta,\rho_0,I) = \hat{\rho}_0(\beta,\rho_0) +
-    I\hat{\rho}_1(\beta,\rho_0) + R(\beta,\rho_0,I)
-    ,
-\end{equation}
-where \(R(\beta,\rho_0,I)\) is smooth and \(O(I^2)\).
-
-From ~\eqref{eqn:khalil-perturbation-nominal} we know
-\(\hat{\rho}_0\) is the solution to the nominal system
+
+\subsubsection*{Perturbation Analysis of the Time Scaled System}
+
+In the spirit of perturbation analysis, we expand the time-scaled system 
+\(\hat{\rho}(\beta,\rho_0,I)\). 
+From ~\eqref{eqn:khalil-perturbation-nominal} we know the nominal system at
+\(I = 0\) is
 \[
@@
 \]
-which means \(\hat{\mu}_0(\beta,\rho_0) \equiv \rho_0\) for all \(\beta\).
+with solution \(\hat{\rho}_0(\beta,\rho_0) \equiv \rho_0\) for all \(\beta\).
 Likewise, ~\eqref{eqn:khalil-perturbation-firstorder} 
-gives us the following dynamics for \(\hat{\rho}_1\):
+tells us that \(\hat{\rho}_1(\beta,\rho_0)\) is the solution to the linear
+time-varying scalar system
 \begin{equation}\label{eqn:acrobot-rho1-dot}
@@
      \hat{\rho}_1(0) = 0
-     .
+     ,
  \end{cases}
 \end{equation}
-The solution to ~\eqref{eqn:acrobot-rho1-dot} is given by
+whose solution is
 \[
@@
 \]
+Therefore, the first-order Taylor approximation of
+\(\hat{\rho}(\beta,\rho_0,I)\) is
+\begin{equation}\label{eqn:acrobot-rhohat-approx}
+    \hat{\rho}(\beta,\rho_0,I) = \hat{\rho}_0(\beta,\rho_0) +
+    I\hat{\rho}_1(\beta,\rho_0) + R(\beta,\rho_0,I)
+    ,
+\end{equation}
+where \(R(\beta,\rho_0,I)\) is smooth and \(O(I^2)\).
+
+\subsubsection*{Poincar\'{e} Analysis}
 
@@
 direction (\ie \(\beta\) going from \(0\) to \(\sign{\rho_0}2\pi\)).
-We define the Poincar\'{e} section
+Since \(\dot{\beta}\) does not change sign on \(\mathcal{R}\),
+there is a well-defined Poincar\'{e} map describing how the pseudo-radius
+\(\rho\) changes each time the orbit \((q(t),p(t))\) hits the \(p_u\) axis, \ie,
+every time \(\beta\) changes by \(2\pi\).
+We define this Poincar\'{e} map as
 \(P_\mathcal{R}(\rho_0) := \hat{\rho}(\sign{\rho_0}2\pi,\rho_0,I)\),
-which is the intersection point on the \(p_u\)-axis after a single rotation.
-This expands into
+which expands into
 \[
     P_\mathcal{R}(\rho_0) = \rho_0 + I \hat{\rho}_1(\sign{\rho_0}2\pi, \rho_0)
-      + R(2\pi,\rho_0,I)
-    .
-\]
-
-The function \(b(\beta,\rho_0)\) is even and is 
+    + R(\sign{\rho_0}2\pi,\rho_0,I)
+    .
+\]
+
+The function \(b(\beta,\rho_0)\) is even and
 \(2\pi\)-periodic in \(\beta\), which implies that
@@
 \end{align*}
-Furthermore, \(b(\beta,\rho_0)\) is also even in \(\rho_0\), 
-so \(\hat{\rho}_1(2\pi,\rho_0) = \hat{\rho}_1(2\pi,\abs{\rho_0})\).
-This implies that
-\begin{align*}
-    \abs{P_\mathcal{R}(\rho_0)} &= \left|\sign{\rho_0}\left(
-        \abs{\rho_0} + I \hat{\rho}_1(2\pi,\abs{\rho_0})
-        + \sign{\rho_0}R(2\pi,\rho_0,I)\right)\right|
-    \\
-    &= \left|\abs{\rho_0} + I \hat{\rho}_1(2\pi,\abs{\rho_0})
-        + \sign{\rho_0}R(2\pi,\rho_0,I)\right|
-    .
-\end{align*}
-
+Thus, we have that
+\[
+    P_\mathcal{R}(\rho_0) = \rho_0 + I\sign{\rho_0}\hat{\rho}_1(2\pi,\rho_0) 
+    + R(\sign{\rho_0}2\pi,\rho_0,I)
+    .
+\]
+
+Note that \(b(\beta,\rho_0)\) is even in \(\rho_0\), which means
+\(\hat{\rho}_1(2\pi,\rho_0) = \hat{\rho}_0(2\pi,\abs{\rho_0})\).
+Recalling the assumption that there exists \(\bar{\rho} > 0\) so that
+\(\int \limits_{0}^{2\pi} b(\sigma,\rho_0)d\sigma > 0\) for all 
+\(\rho_0 \in \,]\sqrt{60m^2gl^3},\bar{\rho}]\),
+we have that \(\hat{\rho}_1(2\pi,\rho_0) > 0\)
+for every \(\rho_0\) with \(\sqrt{60m^2gl^3} < \abs{\rho} \leq \bar{rho}\).
+Finally, \(R(\sign{\rho_0}2\pi,\rho_0,I)\) is \(O(I^2)\). 
+Defining 
+\[
+    D := [-\bar{\rho},\sqrt{60m^2gl^3}[ \, \cup \, ]\sqrt{60m^2gl^3},\bar{\rho}]
+    ,
+\]
+these facts all together imply that, for
+each \(\rho_0 \in D\),
+there exists \(I > 0\) small enough that 
+\[
+    \begin{cases}
+        P_\mathcal{R}(\rho_0) > \rho_0 & \rho_0 > 0
+        ,\\
+        P_\mathcal{R}(\rho_0) < \rho_0 & \rho_0 < 0
+        .\\
+    \end{cases}
+\]
+
+\subsubsection*{Energy Gain on \(\mathcal{R}\)}
+
+The Poincar\'{e} section \(P_\mathcal{R}\) allows us understand the evolution of
+\(\hat{\rho}(\beta,\rho_0,I)\) by studying the evolution of the discrete time
+system
+\begin{equation}\label{eqn:rhohat-discrete}
+    \begin{cases}
+        \rho(n+1) := P_\mathcal{R}(\rho(n)) 
+        = \rho(n) + I\sign{\rho_0}\hat{\rho}_1(2\pi,\rho(n)) + 
+            R(\sign{\rho_0}2\pi,\rho(n),I)
+        , \\
+        \rho(0) = \rho_0
+        .
+    \end{cases}
+\end{equation}
+Tolerating the abuse of notation, \(\rho(n)\) represents the distance along the
+\(p_u\)-axis when an orbit of the constrained dynamics intersects the
+\(p_u\)-axis for the \(n\)th time, assuming the orbit was initialized at
+\((q_u,p_u) = (0,\rho_0)\).
+
+We wish to show \(\abs{\rho(n)}\) eventually reaches \(\bar{\rho}\).
+That is, we wish to find some \(I^\star \in \,]0,I_1]\) making
+\(\abs{P_\mathcal{R}(\rho_0)} > \abs{\rho_0}\) for all 
+\(\rho_0 \in D\)
+We will restrict ourselves, without loss of generality, to the case where
+\(\rho_0 > 0\).
+
+Recall that \(R(\beta,\rho_0,I)\) is \(O(I^2)\) and smooth in all
+its parameters.
+Therefore, it is bounded below on the compact set
+\(\rho_0 \in [\sqrt{60m^2gl^3},\bar{\rho}]\) by some value
+\(R(2\pi,\underbar{\rho},I)\).
+Theorem \ref{thm:khalil-perturbation} asserts that there exists some 
+\(I_2 > 0\) and \(r > 0\) so that, for all \(I \in [-I_2,I_2]\),
+\[
+    R(2\pi,\rho_0,I) \geq R(2\pi,\underbar{\rho},I) > -I^2r
+    .
+\]
+
+%TODO: replace this with an analysis of poincare section
 Let now \(\bar{\rho} > \sqrt{60m^2gl^3}\) and take the compact set

### 2020-12-02 latex/acrobot.tex
Finished first version of rotation and omega proof

@@
     \]
-    and fix \(\bar{\rho} > \sqrt{60m^2gl^3}\).
-    Suppose \(\int \limits_{0}^{2\pi} b(\sigma,\rho_0)d\sigma > 0\) for all 
-    \(\rho_0 \in \, \left]\sqrt{60m^2gl^3}, \bar{\rho}\right]\).
+    and set \(S(\rho_0) := \int \limits_{0}^{2\pi} b(\sigma,\rho_0)d\sigma\).
+    Fix \(\bar{\rho} > \sqrt{60m^2gl^3}\).
+    Suppose there exists \(\epsilon > 0\) so that \(S(\rho_0) > \epsilon\) for
+    all \(\rho_0 \in \, \left]\sqrt{60m^2gl^3}, \bar{\rho}\right]\).
     Then there exists \(I > 0\) small enough that
@@
 Choosing \(I^\star \in\, ]0,\min\left\{I_2,I_3\right\}]\) so that
-\(I^\star K Q(\epsilon) - (I^\star)^2 r > 0\)
+\(I^\star K Q(\epsilon) - (I^\star)^2 r \geq 0\)
 means that
@@
 \end{align*}
-Thus, we have that
-\[
-    P_\mathcal{R}(\rho_0) = \rho_0 + I\sign{\rho_0}\hat{\rho}_1(2\pi,\rho_0) 
+Defining \(S(\rho_0) := \hat{\rho}_1(2\pi,\rho_0)\),
+we get
+\[
+    P_\mathcal{R}(\rho_0) = \rho_0 + I\sign{\rho_0}S(\rho_0) 
     + R(\sign{\rho_0}2\pi,\rho_0,I)
@@
 Note that \(b(\beta,\rho_0)\) is even in \(\rho_0\), which means
-\(\hat{\rho}_1(2\pi,\rho_0) = \hat{\rho}_0(2\pi,\abs{\rho_0})\).
-Recalling the assumption that there exists \(\bar{\rho} > 0\) so that
-\(\int \limits_{0}^{2\pi} b(\sigma,\rho_0)d\sigma > 0\) for all 
-\(\rho_0 \in \,]\sqrt{60m^2gl^3},\bar{\rho}]\),
-we have that \(\hat{\rho}_1(2\pi,\rho_0) > 0\)
-for every \(\rho_0\) with \(\sqrt{60m^2gl^3} < \abs{\rho} \leq \bar{rho}\).
+\(S(\rho_0) = S(\abs{\rho_0})\).
+Recall the assumption that there exist \(\bar{\rho} > 0\) and
+\(\epsilon > 0\) where \(S(\rho_0) \geq \epsilon\) for all 
+\(\rho_0 \in \,]\sqrt{60m^2gl^3},\bar{\rho}]\).
+Hence, we have that \(S(\rho_0) \geq \epsilon\)
+for every \(\rho_0\) with \(\sqrt{60m^2gl^3} < \abs{\rho} \leq \bar{\rho}\).
 Finally, \(R(\sign{\rho_0}2\pi,\rho_0,I)\) is \(O(I^2)\). 
@@
 \]
-
-\subsubsection*{Energy Gain on \(\mathcal{R}\)}
+As with oscillations, this \(I\) may not work for all \(\rho_0 \in D\). We deal
+with this issue next.
+
+\subsubsection*{Energy Gain on a Subset of \(\mathcal{R}\)}
 
@@
         \rho(n+1) := P_\mathcal{R}(\rho(n)) 
-        = \rho(n) + I\sign{\rho_0}\hat{\rho}_1(2\pi,\rho(n)) + 
+        = \rho(n) + I\sign{\rho_0}S(\rho(n)) + 
             R(\sign{\rho_0}2\pi,\rho(n),I)
@@
 \(\rho_0 \in D\)
-We will restrict ourselves, without loss of generality, to the case where
-\(\rho_0 > 0\).
-
+
+We begin with examining the case where \(\rho_0 > 0\).
 Recall that \(R(\beta,\rho_0,I)\) is \(O(I^2)\) and smooth in all
@@
 \]
-
-%TODO: replace this with an analysis of poincare section
-Let now \(\bar{\rho} > \sqrt{60m^2gl^3}\) and take the compact set
-\[
-    \bar{\mathcal{R}} := \left\{(q_u,p_u) \in \Gamma 
-    \mid E(\pi,0) \leq E(q_u,p_u) \leq E(0,\bar{\rho}) \right\}
-    .
-\]
-The (modified) remainder term \(\sign{\rho_0}{R}(\beta,\rho_0,I)\) 
-is smooth and \(O(I^2)\), which means
-it attains its minimum \(I^2\hat{r}\) somewhere \(\bar{\mathcal{R}}\).
-Likewise, \(\hat{\rho}_1(2\pi,\abs{\rho_0})\) also attains its minimum, which we
-denote by \(\underbar{\rho}_1\).
-We cannot adimensionalize \(b(\beta,\rho_0)\), so refer to one of the 
-assumptions of Theorem \ref{thm:acrobot-energy-stabilization}
-to suppose  that \(\hat{\rho}_1(2\pi,\rho_0) > 0 \) for all 
-\(\rho_0 \in ]\sqrt{60m^2gl^3},\bar{\rho}]\).
-In particular, this means \(\underbar{\rho}_1 > 0\).
-
-There exists \(I > 0\) making 
-\(I\underbar{\rho}_1 + I^2\hat{r} > 0\).
+Since \(S(\rho_0) \geq \epsilon\), we find that
+\[
+    P_\mathcal{R}(\rho_0) > \rho_0 + I\epsilon -I^2r
+    ,
+\]
+for all \(\rho_0 \in ]\sqrt{60m^2gl^3},\bar{\rho}]\).
+Choosing \(I_3 \in ]0,I_2]\) so that 
+\(I_3\epsilon - (I_3)^2 r \geq 0\) means that
+\(P_\mathcal{R}(\rho_0) > \rho_0\)
+for all \(\rho_0 \in ]\sqrt{60m^2gl^3},\bar{\rho}]\)
+
+We now examine the case when \(\rho_0 < 0\), where we wish to show that 
+\(P_\mathcal{R}(\rho_0) < \rho_0\).
+The remainder \(R(\beta,\rho_0,I)\) is bounded above on the compact set 
+\(\rho_0 \in [-\bar{\rho},-\sqrt{60m^2gl^3}]\), and 
+Theorem \ref{thm:khalil-perturbation} asserts that there exists some 
+\(I_4 > 0\) and \(r^\prime > 0\) so that, for all \(I \in [-I_4,I_4]\), 
+\[
+    R(-2\pi,\rho_0,I) < I^2 r^\prime
+    .
+\]
 Hence,
 \begin{align*}
-    \abs{\rho_0} + I\hat{\rho}_1(2\pi,\abs{\rho_0}) + \sign{\rho_0}R(2\pi,\rho_0,I)
-    &\geq \sqrt{60m^2gl^3} + I\underbar{\rho}_1 + I^2\hat{r}
-    \\
-    &> \sqrt{60m^2gl^3}
-    \\
-    &> 0
-    ,
+    P_\mathcal{R}(\rho_0)
+    &= \rho_0 - IS(\rho_0) + R(-2\pi,\rho_0,I)
+    &< \rho_0 - IS(\epsilon) + I^2r^\prime
+    .
 \end{align*}
-which implies that
-\begin{align*}
-    \abs{P_\mathcal{R}(\rho_0)} - \abs{\rho_0}
-    &= \abs{ \abs{\rho_0} + I\hat{\rho}_1(2\pi,\rho_0) +
-    \sign{\rho_0}R(2\pi,\rho_0,I)} - \abs{\rho_0}
-    \\
-    &= \abs{\rho_0} + I\hat{\rho}_1(2\pi,\rho_0) +
-    \sign{\rho_0}R(2\pi,\rho_0,I) - \abs{\rho_0}
-    \\
-    &= I\hat{\rho}_1(2\pi,\rho_0) + \sign{\rho_0}R(2\pi,\rho_0,I)
-    \\
-    &> 0
-    .
-\end{align*}
-
-We conclude that there exists \(I^\prime(\bar{\rho})\)
-small enough that, for all \( 0 < I \leq I^\prime(\bar{\rho})\), 
-the Poincar\'{e} section \(P_\mathcal{R}(\rho_0)\) is monotonically increasing
-when \(\sqrt{60m^2gl^3} < \rho_0 < \bar{\rho}\) and
-monotonically decreasing when \(-\bar{\rho} < \rho_0 < -\sqrt{60m^2gl^3}\).
-
-\subsection*{Energy Gain on \(\Omega\)}
-We will prove the acrobot gains energy on the set 
-\[
-    \Omega := \left\{ (q_u,p_u) \in \Gamma 
-     \mid E(q_u,p_u) < E(0,\bar{\rho}) \right\}
-    .
-\]
-% Can we use the jacobian? The spectrum is hard to find.
-% TODO: W and D's boundary is two elements: the point (pi,0) and the curve of the
-% homoclinic orbit connecting (pi,0) to (-pi,0). We need to show that almost
-% every initial condition starting in \(mathcal{O}\) will NOT converge to (pi,0), and
-% therefore exit D and enter W.
-
-Define the set of points in \(\mathcal{R}\) which have pseudo-radius in \(U\) by
-\[
-    W := \left\{(q_u,p_u) \in \Gamma \mid |\rho(q_u,p_u)| \in U \right\}
-    ,
-\]
-as in Figure \ref{fig:acrobot-rot-W}.
-Monotonicity of \(P(\rho_0)\) means there are no closed orbits of 
-\((q_u,p_u)\) on \(W\). 
-Furthermore, \(|p_u|\) is increasing every time the orbit hits the
-\(p_u\)-axis.
+Choosing \(I_5 \in ]0,I_4]\) so that 
+\((I_5)^2r^\prime - I_5 S(\epsilon) \leq 0\) makes
+\(P_\mathcal{R}(\rho_0) < \rho_0\) for all 
+\(\rho_0 \in [-\bar{\rho},-\sqrt{60m^2gl^3}[\).
+
+Finally, choosing \(I^\star \in ]0,\min{I_3,I_5}]\) yields that
+\(\abs{P_\mathcal{R}(\rho_0)} > \abs{\rho_0}\) for all \(\rho_0 \in D\), as
+desired.
+Looking at ~\eqref{eqn:rhohat-discrete}, this means all solutions of the
+discrete-time system will flow away from the origin until it reaches a magnitude
+\(\abs{\rho(n)} = \bar{\rho}\).
+Hence, the constrained dynamics will eventually escape the set
+\(\omega \subset \mathcal{R}\) defined by
+\[
+    \omega := \left\{(q_u,p_u) \in \Gamma
+        \mid E(\pi,0) < E(q_u,p_u) < E(0,\bar{\rho} \right\}
+    ,
+\]
+which is displayed in Figure \ref{fig:acrobot-little-omega}.
+
+We conclude that there exists \(I > 0\) small enough that the constraint
+~\eqref{eqn:acrobot-constraint} is injecting energy on \(\omega\).
 
@@
     \centering
-    % TODO: Display the set W, which is the two rotation bands at the top and
-    % bottom with outer limit +- \bar{mu}  and inner limit +- sqrt{60m^2gl^3}.
-    % Show also that the inner limit connects to D
-    %\includestandalone[width=0.5\textwidth]{images/acrobot_rot_W}
-    \caption{The set \(W\) of points which have pseudo-radius 
-        \(|\rho| \in ]\sqrt{60m^2gl^3}, \bar{\rho}[\). The inner boundary of \(W\)
-        is \((q_u,p_u) = (\pi,0)\), which is also the boundary of the
-        oscillation domain \(mathcal{O}\).}
-    \label{fig:acrobot-rot-W}
+    \includestandalone[width=0.5\textwidth]{images/acrobot_little_omega}
+    \caption{The set \(\omega \subset \mathcal{R}\) on which the acrobot will gain energy.
+    }
+    \label{fig:acrobot-little-omega}
 \end{figure}
 
-We conclude that \(|\rho(q_u,p_u)|\) will eventually reach \(\bar{\rho}\). 
-We cannot claim this fits the definition  of energy injection 
-(since \((0,0) \notin W\)), but we can still say that \((q_u,p_u)\) will
-eventually depart \(W\) through the outer boundary in finite time.
-The acrobot is thereby gaining energy in the literal sense that its momentum
-is increasing on \(W\).
-
-If we set \(I < 0\), we have \(P(|\rho_0|) < |\rho_0|\).
-Hence, \(|p_u|\) is decreasing every time the orbit hits the \(p_u\)-axis, so
-the acrobot will eventually depart \(W\) through the inner boundary. 
-The acrobot is losing momentum, and therefore it is losing energy on \(W\).
-
-A hybrid supervisor that flips the sign of \(I\) based on the current
-value of \(\mu\) will stabilize any energy level of oscillation.
-As was the case in the VLP, this supervisor should implement some kind of
-hysteresis to avoid infinite switching.
+By the same arguments presented in this section, the Poincar\'{e} section
+satisfies \(\abs{P_\mathcal{R}(\rho_0)} < \abs{\rho_0}\) when \(I < 0\).
+In this case, the constraint dissipates energy on \(\omega\).
+
+\subsection{Energy Gain on \(\Omega\)}
+Define \(E_\pi\) to be the energy level set associated with \(E(pi,0)\).
+We will prove the acrobot gains energy on the set 
+\begin{align*}
+    \Omega &= \left\{(q_u,p_u) \in \Gamma
+    \mid E(q_u,p_u) < E(0,\bar{\rho}) \right\}
+    ,\\
+    &= \mathcal{O} \cup E_\pi \cup \omega
+    .
+\end{align*}
+There exists some control input \(I_1 > 0\) for which the acrobot is gaining
+energy on \(\mathcal{O}\). 
+Hence, there are no closed orbits in \(\mathcal{O}\) and almost all orbits flow
+towards \(E_\pi\).
+Likewise, there exists some \(I_2 > 0\)for which the acrobot is gaining energy
+on \(\omega\), so there are no closed orbits in \(\omega\) and almost every
+orbit is flowing away from \(E_\pi\).
+Choosing \(I^\star \in ]0,\min{I_1,I_2}]\) means the acrobot will gain energy on
+both \(\mathcal{O}\) as well as \(\omega\).
+
+We can prove energy gain on \(\Omega\) by showing two properties of the system.
+First, that the set \(\Pi^+\) of initial conditions converging to  \((\pi,0)\)
+is of measure zero in \(\Gamma\);
+second, that \(E_\pi\) is not an invariant set and therefore is not equal to 
+\(\Pi^+\).
+If both these conditions are true, then orbits flowing from \(\mathcal{O}\) into
+\(E_\pi\) will generally swing past \(E_\pi\) into \(\omega\).
+
+To show the first condition, we will simply prove that \((\pi,0)\) is an
+unstable equilibrium of the constrained dynamics.
+Taking the jacobian of ~\eqref{eqn:acrobot-constrained-dynamics} at \((\pi,0)\)
+yields
+\[
+    J = \begin{bmatrix}
+        -\frac{6mgl\bar{q}_aI}{5} & \frac{1 - 2m^2gl^3\bar{q}_a^2 I^2}{5ml^2} \\
+        3mgl & mgl\bar{q}_aI
+    \end{bmatrix}
+    ,
+\]
+which has the characteristic polynomial
+\[
+    \det\left(\lambda \Id{2} - J\right)
+    = \lambda^2 + \frac{mgl\bar{q}_a I}{5} \lambda - 3g
+    .
+\]
+By Descartes' rule of signs, this polynomial has one root with positive real
+part. 
+Hence, the equilibrium \((\pi,0)\) is unstable.
+
+To show the second condition, we evaluate \(\dot{E}(q_u,p_u)\) at the point
+\((0,\pm\sqrt{60m^2gl^3}) \in E_\pi\).
+MATLAB returns the value
+\[
+    \dot{E}(0,\pm\sqrt{60m^2gl^3}) 
+    = \mp\frac{g}{5l}
+    \sqrt{60m^2gl^3} \sin\left(\bar{q}_a \arctan(I \sqrt{60m^2gl^3})\right)
+    .
+\]
+This is zero only if \(\bar{q}_a \arctan(I\sqrt{60m^2gl^3}) = \frac{\pi}{2}\), a
+situation which occurs only when \(\bar{q}_a \geq 1\) at a single value of
+\(I\).
+Avoiding this value of \(I\) (if it exists) makes \(E_\pi\) a non-invariant set
+as desired.
+
+We know that almost all orbits starting in \(\omega\) will escape \(\Omega\).
+Non-invariance of \(E_\pi\) and instability of \((\pi,0)\) implies that 
+almost all orbits initialized in \(\mathcal{O}\) will reach \(E_\pi\) and
+eventually swing into \(\omega\), rather than stay in \(E_\pi\) and converge to
+\((\pi,0)\). 
+We conclude that almost all orbits will depart the energy level set associated with
+\(E(0,\bar{\rho})\), thereby escaping \(\Omega\) all together.
+Hence, the acrobot is gaining energy on \(\Omega\).
+
+If \(I < 0\), orbits in \(\mathcal{O}\) will converge to the origin and orbits in
+\(\omega\) will flow towards \(E_\pi\).
+However, the upright equilibrium \((\pi,0)\) will still be unstable, and
+\(E_\pi\) will remain a non-invariant set. 
+This means that almost all orbits in \(\omega\) will flow past \(E_\pi\) into
+\(\mathcal{O}\), so the constrained system will be gaining energy in
+negative-time. Thus, the constraint is dissipating energy on \(\Omega\).
 

### 2020-12-05 latex/acrobot.tex
Updated the rotation section, ready to send it to Manfredi.

@@
 
-Using \(\beta\) as our new time variable produces the time-scaled system
-\(\hat{\rho}(\beta) = \rho(\tau(\beta))\) with dynamics
+Using \(\beta\) as our new time variable (via a time reparameterization
+\(t = t(\beta)\)) produces the time-scaled system
+\(\hat{\rho}(\beta,\rho_0,I) = \rho(t(\beta),\rho_0,I)\) with dynamics
 \[
     \begin{cases}
-        \dot{\hat{\rho}} = \frac{\dot{\rho}}{\dot{\beta}}
+        \diff{\hat{\rho}}{\beta} = \frac{\dot{\rho}}{\dot{\beta}}
         , \\
@@
 Suppose we initialize the acrobot at \((\beta,\rho) = (0,\rho_0)\).
-One full rotation amounts to \(\beta\) traversing \(2\pi\)rad in a clockwise
-direction (\ie \(\beta\) going from \(0\) to \(\sign{\rho_0}2\pi\)).
+One full rotation amounts to \(\beta\) traversing \(2\pi\) rad in a clockwise
+direction; that is, when \(\beta\) goes from \(0\) to \(\sign{\rho_0}2\pi\)).
 Since \(\dot{\beta}\) does not change sign on \(\mathcal{R}\),
@@
 
-Note that \(b(\beta,\rho_0)\) is even in \(\rho_0\), which means
-\(S(\rho_0) = S(\abs{\rho_0})\).
 Recall the assumption that there exist \(\bar{\rho} > 0\) and
@@
 \(\rho_0 \in \,]\sqrt{60m^2gl^3},\bar{\rho}]\).
-Hence, we have that \(S(\rho_0) \geq \epsilon\)
-for every \(\rho_0\) with \(\sqrt{60m^2gl^3} < \abs{\rho} \leq \bar{\rho}\).
-Finally, \(R(\sign{\rho_0}2\pi,\rho_0,I)\) is \(O(I^2)\). 
-Defining 
-\[
-    D := [-\bar{\rho},\sqrt{60m^2gl^3}[ \, \cup \, ]\sqrt{60m^2gl^3},\bar{\rho}]
-    ,
-\]
-these facts all together imply that, for
-each \(\rho_0 \in D\),
+Let us define
+\[
+    D := \left[-\bar{\rho},\sqrt{60m^2gl^3}\right[ \, 
+    \cup \, 
+    \left]\sqrt{60m^2gl^3},\bar{\rho}\right]
+    .
+\]
+Since \(b(\beta,\rho_0)\) is even in \(\rho_0\),
+\(S(\rho_0) = S(\abs{\rho_0})\);
+hence, \(S(\rho_0) \geq \epsilon\) for every \(\rho_0 \in D\).
+
+Finally, recall that \(R(\sign{\rho_0}2\pi,\rho_0,I)\) is \(O(I^2)\), so 
+for each fixed \(\rho_0 \in D\)
 there exists \(I > 0\) small enough that 
+\( IS(\rho_0) + R(\sign{\rho_0}2\pi,\rho_0,I) > 0\).
+For this value of \(I\), we have that
 \[
@@
 \]
-As with oscillations, this \(I\) may not work for all \(\rho_0 \in D\). We deal
+Unfortunately, this \(I\) may not work for all \(\rho_0 \in D\). We deal
 with this issue next.
@@
     &= \rho_0 - IS(\rho_0) + R(-2\pi,\rho_0,I)
+    , \\
     &< \rho_0 - IS(\epsilon) + I^2r^\prime
@@
 
-Finally, choosing \(I^\star \in ]0,\min{I_3,I_5}]\) yields that
+Finally, choosing \(I^\star \in\, ]0,\min\left\{I_3,I_5\right\}]\) yields that
 \(\abs{P_\mathcal{R}(\rho_0)} > \abs{\rho_0}\) for all \(\rho_0 \in D\), as
 desired.
-Looking at ~\eqref{eqn:rhohat-discrete}, this means all solutions of the
-discrete-time system will flow away from the origin until it reaches a magnitude
-\(\abs{\rho(n)} = \bar{\rho}\).
+Looking at the discrete-time system ~\eqref{eqn:rhohat-discrete}, 
+a solution of the system will flow away from the origin until it reaches a
+magnitude \(\abs{\rho(n)} = \bar{\rho}\).
 Hence, the constrained dynamics will eventually escape the set
@@
     \omega := \left\{(q_u,p_u) \in \Gamma
-        \mid E(\pi,0) < E(q_u,p_u) < E(0,\bar{\rho} \right\}
+    \mid E(\pi,0) < E(q_u,p_u) < E(0,\bar{\rho}) \right\}
     ,
@@
 \subsection{Energy Gain on \(\Omega\)}
-Define \(E_\pi\) to be the energy level set associated with \(E(pi,0)\).
+Define \(E_\pi\) to be the energy level set associated with \(E(\pi,0)\).
 We will prove the acrobot gains energy on the set 
@@
 energy on \(\mathcal{O}\). 
-Hence, there are no closed orbits in \(\mathcal{O}\) and almost all orbits flow
-towards \(E_\pi\).
-Likewise, there exists some \(I_2 > 0\)for which the acrobot is gaining energy
+Hence, there are no closed orbits in \(\mathcal{O}\) and almost all orbits 
+hit \(E_\pi\) in finite time.
+Likewise, there exists some \(I_2 > 0\) for which the acrobot is gaining energy
 on \(\omega\), so there are no closed orbits in \(\omega\) and almost every
 orbit is flowing away from \(E_\pi\).
-Choosing \(I^\star \in ]0,\min{I_1,I_2}]\) means the acrobot will gain energy on
+Choosing \(I^\star \in ]0,\min\left\{I_1,I_2\right\}]\) means the acrobot will gain energy on
 both \(\mathcal{O}\) as well as \(\omega\).
@@
 is of measure zero in \(\Gamma\);
-second, that \(E_\pi\) is not an invariant set and therefore is not equal to 
-\(\Pi^+\).
+second, that \(E_\pi\) is not a positively-invariant set.
 If both these conditions are true, then orbits flowing from \(\mathcal{O}\) into
@@
 unstable equilibrium of the constrained dynamics.
-Taking the jacobian of ~\eqref{eqn:acrobot-constrained-dynamics} at \((\pi,0)\)
-yields
+Taking the Jacobian of the constrained dynamics
+~\eqref{eqn:acrobot-constrained-dynamics} at \((\pi,0)\) yields
 \[
@@
 part. 
-Hence, the equilibrium \((\pi,0)\) is unstable.
-
-To show the second condition, we evaluate \(\dot{E}(q_u,p_u)\) at the point
-\((0,\pm\sqrt{60m^2gl^3}) \in E_\pi\).
-MATLAB returns the value
-\[
-    \dot{E}(0,\pm\sqrt{60m^2gl^3}) 
-    = \mp\frac{g}{5l}
-    \sqrt{60m^2gl^3} \sin\left(\bar{q}_a \arctan(I \sqrt{60m^2gl^3})\right)
-    .
-\]
-This is zero only if \(\bar{q}_a \arctan(I\sqrt{60m^2gl^3}) = \frac{\pi}{2}\), a
-situation which occurs only when \(\bar{q}_a \geq 1\) at a single value of
-\(I\).
-Avoiding this value of \(I\) (if it exists) makes \(E_\pi\) a non-invariant set
-as desired.
-
-We know that almost all orbits starting in \(\omega\) will escape \(\Omega\).
+The equilibrium \((\pi,0)\) is therefore unstable, so the set \(\Pi^+\) of
+initial conditions converging to \((\pi,0)\) is one-dimensional, and hence is
+of measure zero in \(\Gamma\).
+
+For the second condition, let us seek a contradiction by 
+supposing that \(E_\pi\) is a positively-invariant set. 
+Let \(x(t)\) be an orbit of the constrained dynamics initialized at 
+\(x(0) \in \mathal{O}\).
+Since the constrained dynamics gain energy on \(\mathcal{O}\), there exists a
+time \(T > 0\) such that \(x(T) \in E_\pi\).
+By the assumed positive invariance of \(E_\pi\), this means 
+\(x(t) \in E_\pi\) for all \(t \in [T,\infty[\).
+Since \(E_\pi\) is a closed curve containing the equilibrium \((\pi,0)\) and no
+other equilibria, \(x(t)\) must converge to \((\pi,0)\).
+The initial condition \(x(0) \in \mathcal{O}\) was arbitrary, so this implies
+that \(\mathcal{O} \subset \Pi^+\).
+But \(\Pi^+\) is a set of measure zero, while \(\mathcal{O}\) has positive
+measure. 
+This is a contradiction: \(E_\pi\) cannot be positively-invariant.
+
 Non-invariance of \(E_\pi\) and instability of \((\pi,0)\) implies that 
@@
 \((\pi,0)\). 
-We conclude that almost all orbits will depart the energy level set associated with
-\(E(0,\bar{\rho})\), thereby escaping \(\Omega\) all together.
+We know that almost all orbits starting in \(\omega\) will reach 
+\(E(0,\bar{\rho})\) in finite time;
+we conclude that almost all orbits starting in \(\Omega\) will depart the energy
+escaping \(\Omega\) all together.
 Hence, the acrobot is gaining energy on \(\Omega\).

### 2020-12-11 latex/acrobot.tex
Major Update to acrobot proof
Added a bunch of sections that Manfredi recommended, and fixed a few
errors causing compilation failure. Now we need  abit more editing and
some restructuring of the rotation section, as well as part of the
poincare map section of oscillations.

@@
 
-These dynamics do not always gain energy -- we need certain conditions on our
-control parameter \(I\) to guarantee this is true.
+These dynamics do not gain energy everywhere; we need to define some sets where
+this energy gain is possible.
+First, let 
+\[
+    E(q_u,p_u) := \frac{p_u^2}{10ml^2} + 3mgl(1 - \cos(q_u))
+    ,
+\]
+be the energy function obtained by setting \(I = 0\).
+As we will see later, this is the mechanical energy of a simple pendulum with
+two masses.
+Next, define the set
+\begin{equation}\label{eqn:oscillation-domain}
+    \mathcal{O}_1 := \left\{(q_u,p_u) \in \SxR 
+    \mid E(q_u,p_u) < E(\pi,0) \right\}
+    ,
+\end{equation}
+which is seen in Figure \ref{fig:acrobot-oscillation-domain}.
+Finally, let \(\bar{\rho} > \sqrt{60m^2gl^3}\) and define the set
+\begin{equation}\label{eqn:o-rhobar}
+    \mathcal{O}_2(\bar{\rho}) := \left\{(q_u,p_u) \in \SxR
+        \mid E(q_u,p_u) < E(0,\bar{\rho}) \right\}
+    ,
+\end{equation}
+as in Figure \ref{fig:acrobot-o2}
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[t]{0.45\textwidth}
+        \includestandalone[width=\linewidth]{images/acrobot_oscillation_domain}
+        \caption{The set \(\mathcal{O}\). An orbit starting in this set (blue)
+            will reach the level set \(E(\pi,0)\) of the nominal pendulum.}
+        \label{fig:acrobot-oscillation-domain}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.45\textwidth}
+        \includestandalone[width=\linewidth]{images/acrobot_omega}
+        \caption{The set \(\mathcal{O}_2(\bar{\rho})\). Orbits starting in this set will
+            reach the level set \(E(0,\bar{\rho})\) of the nominal pendulum.}
+            \label{fig:acrobot-o2}
+    \end{subfigure}
+    \caption{The sets on which Theorem \ref{thm:acrobot-energy-stabilization}
+        claims the acrobot gains energy.}
+\end{figure}
+
+For small enough \(I\), our VNHC will inject energy on the open set
+\(\mathcal{O}_1\) and (provided a suitable condition is met) on the open set
+\(\mathcal{O}_2(\bar{\rho})\).
 
@@
     constraint manifold is \(\Gamma \simeq \SxR\).
-    Let 
-    \[
-        E(q_u,p_u) := \frac{p_u^2}{10ml^2} + 3mgl(1 - \cos(q_u))
-        ,
-    \]
-    be the energy of the simple pendulum obtained by setting \(I = 0\).
 \begin{enumerate}
@@
     ~\eqref{eqn:acrobot-constraint} injects energy into the acrobot on
-    \[
-        \mathcal{O} := \left\{(q_u,p_u) \in \Gamma 
-        \mid E(q_u,p_u) < E(\pi,0) \right\}
-        .
-    \]
-    If instead \(I < 0\), the VNHC dissipates energy on \(\mathcal{O}\).
+    \(\mathcal{O}_1\).
+    If instead \(I < 0\), the VNHC dissipates energy on this set.
 \item Define \(b : \SxR_{> 0} \rightarrow \R\) by
@@
     ~\eqref{eqn:acrobot-constraint} injects energy into the acrobot on
-    \[
-        \Omega := \left\{(q_u,p_u) \in \Gamma 
-        \mid E(q_u,p_u) < E(0,\bar{\rho})\right\}
-        .
-    \]
-    If instead \(I < 0\), the VNHC dissipates energy on \(\Omega\).
+    \(\mathcal{O}_2(\bar{\rho})\).
+    If instead \(I < 0\), the VNHC dissipates energy on this set.
 \end{enumerate}
@@
 
-A small enough fixed value \(I > 0\) will enable the acrobot to 
-escape any compact subset of \(\mathcal{O}\) in finite time. 
+Notice that \(\mathcal{O}_1 \subset \mathcal{O}_2(\bar{\rho})\), yet
+Theorem \ref{thm:acrobot-energy-stabilization} considers these sets separately.
+This separation is advantageous: 
+regardless of one's choice of acrobot, 
+a small enough fixed value \(I > 0\) will enable the acrobot to 
+escape any compact subset of \(\mathcal{O}_1\) in finite time. 
 What's more, the acrobot's phase will flow towards the homoclinic orbit
@@
 \ref{fig:acrobot-oscillation-domain}.
+
 Hence, the first result of Theorem \ref{thm:acrobot-energy-stabilization} states
@@
 In the worst case, the acrobot will at least perform a swing-up routine to reach
-the \((q_u,p_u) = (\pi,0)\) equilibrium.
-
+the unstable equilibrium at \((q_u,p_u) = (\pi,0)\).
+
+The second result of Theorem
+\ref{thm:acrobot-energy-stabilization} pertains to energy injection once the
+acrobot has started rotating - that is, when its phase lies in 
+\(\mathcal{O}_2(\bar{\rho}) \backslash \mathcal{O}_1\).
 If we want the acrobot to achieve giants with energy
@@
 integral of \(b(\beta,\rho_0)\).
-In this case, orbits will escape compact subsets of \(\Omega\) in finite time
-until the acrobot eventually reaches a momentum of at least \(\bar{\rho}\).
-One example of this behaviour is displayed in Figure \ref{fig:acrobot-omega}.
+In this case, orbits will escape compact subsets of
+\(\mathcal{O}_2(\bar{\rho})\) in finite time until the acrobot eventually
+reaches a momentum of at least \(\bar{\rho}\).
+One example of this behaviour is displayed in Figure \ref{fig:acrobot-o2}.
 Note that the value of \(I\) will depend on \(\bar{\rho}\).
-
-\begin{figure}
-    \centering
-    \begin{subfigure}[t]{0.45\textwidth}
-        \includestandalone[width=\linewidth]{images/acrobot_oscillation_domain}
-        \caption{The set \(\mathcal{O}\). An orbit starting in this set (blue)
-            will reach the level set \(E(\pi,0)\) of the nominal pendulum in
-            finite time.}
-        \label{fig:acrobot-oscillation-domain}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.45\textwidth}
-        \includestandalone[width=\linewidth]{images/acrobot_omega}
-        \caption{The set \(\Omega\). Orbits starting in this set will
-            reach the level set \(E(0,\bar{\rho})\) of the nominal pendulum in
-            finite time.}
-            \label{fig:acrobot-omega}
-    \end{subfigure}
-    \caption{The sets defined by Theorem \ref{thm:acrobot-energy-stabilization}
-        on which the acrobot gains energy.}
-\end{figure}
 
@@
 Once we have these coordinates in hand, we can use perturbation theory to prove
-that \(\mu\) increases on \(\mathcal{O}\) when \(I > 0\) is small enough.
+that \(\mu\) increases on \(\mathcal{O}_1\) when \(I > 0\) is small enough.
 
@@
 Using both \((\alpha,\mu)\) and \((\beta,\rho)\), we will show the acrobot is
-gaining energy on \(\Omega\).
+gaining energy on \(\mathcal{O}_2(\bar{\rho})\).
 
@@
 \[
-    \mathcal{O} := \left\{ (q_u,p_u) \in \Gamma \mid E(q_u,p_u) < E(\pi,0)\right\}
-    ,
-\]
-where \(\Gamma \simeq \SxR\) is the constraint manifold of the
-acrobot.
+    \mathcal{O}_1 := \left\{ (q_u,p_u) \in \SxR \mid E(q_u,p_u) < E(\pi,0)\right\}
+    .
+\]
 Recall that orbits of a pendulum are level sets of \(E\). 
@@
 
-The energy of oscillation can be determined by the orbit's intersection
-\(\mu\) with the \(q_u\)-axis (as in Figure \ref{fig:mu-intersection}).
-Since \(q_u \in ]-\pi,\pi[\) on \(\mathcal{O}\), we set \(\mu \in ]0,\pi[\).
+Once again, to define pseudo-polar coordinates, we consider the dynamics of the
+pendulum obtained by setting \(I = 0\) in
+~\eqref{eqn:acrobot-constrained-dynamics}, resulting in
+\begin{equation}\label{eqn:acrobot-nominal-dynamics}
+    \begin{cases}
+        \dot{q}_u = \frac{p_u}{5ml^2}
+        , \\
+        \dot{p}_u = -3mgl\sin(q_u)
+        . 
+    \end{cases}
+\end{equation}
+
+The above is a simple mechanical system whose total energy is precisely the
+function \(E(q_u,p_u)\) defined in ~\eqref{eqn:acrobot-nominal-E}.
+The energy of oscillation can be determined by the intersection
+\(\mu\) of an orbit of ~\eqref{eqn:acrobot-nominal-dynamics}
+with the \(q_u\)-axis (as in Figure \ref{fig:mu-intersection}).
+Since \(q_u \in ]-\pi,\pi[\) on \(\mathcal{O}_1\), we set \(\mu \in ]0,\pi[\).
 The transformation we want is therefore a diffeomorphism of the form
 \begin{align*}
-    T : \mathcal{O} \backslash \{(0,0)\} &\rightarrow \Sone \times \, ]0,\pi[, \\
+    T : \mathcal{O}_1 \backslash \{(0,0)\} &\rightarrow \Sone \times \, ]0,\pi[, \\
     (q_u, p_u) &\mapsto (\alpha,\mu)
@@
     \includestandalone[width=0.5\textwidth]{images/mu_intersection}
-    \caption{The domain \(\mathcal{O}\) (blue) where a pendulum oscillates.
+    \caption{The domain \(\mathcal{O}_1\) (blue) where a pendulum oscillates.
     The pseudo-radius \(\mu\) corresponds to the
@@
     The psuedo-angle \(\alpha\) is taken to be clockwise positive because 
-    oscillations of a pendulum move clockwise on \(\mathcal{O}\).}
+    oscillations of a pendulum move clockwise on \(\mathcal{O}_1\).}
     \label{fig:mu-intersection}
@@
 \[
-    \left\{(q_u,p_u) \in \Gamma \mid E(q_u,p_u) = 3mgl(1- \cos(\mu))\right\}
+    \left\{(q_u,p_u) \in \SxR \mid E(q_u,p_u) = 3mgl(1- \cos(\mu))\right\}
     ,
@@
 Since \(\mu \in ]0,\pi[\), we can uniquely express \(\mu\) by
-\begin{equation}\label{eqn:mu-qupu}
+\begin{equation*}
     \mu = \arccos\left(C_\mu(q_u,p_u)\right)
     .
-\end{equation}
+\end{equation*}
 Next we need to find \(\alpha\). 
@@
 ~\eqref{eqn:acrobot-sinalpha-qu}, we deduce that
-\begin{equation}\label{eqn:alpha-qupu}
+\begin{equation*}
     \alpha = \left.
@@
     .
-\end{equation}
-Thus, ~\eqref{eqn:mu-qupu} and ~\eqref{eqn:alpha-qupu} define 
-our transformation into \((\alpha,\mu)\)-coordinates on \(\mathcal{O}\).
+\end{equation*}
+Thus, our transformation into \((\alpha,\mu)\)-coordinates is
+\begin{align}
+        \label{eqn:mu-qupu}
+        \mu &= \arccos\left(C_\mu(q_u,p_u)\right), \\
+        \label{eqn:alpha-qupu}
+        \alpha &= \left.
+        \arctan_2\left( -\sign{p_u}\sqrt{1 - \frac{q_u^2}{\mu^2}}, \frac{q_u}{\mu}\right)
+        \right|_{\mu = \arccos(C_\mu(q_u,p_u))}
+        .
+\end{align}
 
@@
 \end{align}
-One can confirm that \(\dot{\alpha} > 0\) for every \(\mu \in ]0,\pi[\).
+Note that ~\eqref{eqn:acrobot-alpha-dot-nom} has a removable singularity at
+\(\alpha \in \{0,\pi\}\): 
+taking the limit as \(\alpha\) approaches these points yields
+\[
+    \lim \limits_{\alpha \to 0}\dot{\alpha} 
+    = \lim \limits_{\alpha \to \pi} \dot{\alpha}
+    = \sqrt{\frac{6g \sin(\mu)}{10l \mu}}
+    ,
+\]
+which is smooth and well-defined for all \(\mu \in ]0,\pi[\).
+Taking this removable singularity into account,
+one can confirm that \(\dot{\alpha} > 0\) for every \(\mu \in ]0,\pi[\).
 By continuity of ~\eqref{eqn:acrobot-alpha-dot}, there exists \(I_1 > 0\) small
-enough that \(\dot{\alpha}\) remains positive on \(\mathcal{O}\) for 
+enough that \(\dot{\alpha}\) remains positive on \(\mathcal{O}_1\) for 
 \(I \in [-I_1,I_1]\).
 
+These pseudo-polar coordinates can be extended smoothly to 
+\(\bar{\mathcal{O}}_1\), the closure of \(\mathcal{O}_1\).
+Notice that, for each \(\alpha \in \Sone\),
+\[
+    \lim \limits_{\mu \to \pi} \dot{\alpha}
+    = \sqrt{\frac{6g}{5l}} 
+        \sqrt{\frac{\cos(\pi\cos(\alpha)) + 1}
+            {\pi^2 \sin(\alpha)^2}}
+    \geq 0
+    ,
+\]
+with equality if and only if \(\alpha \in \{0,\pi\}\).
+In other words, the coordinates \(\mu = \pi, \alpha\) dictate the behaviour of
+the nominal pendulum on the homoclinic orbit with energy \(E(\pi,0)\).
+Hence, for small enough \(I\) we have \(\dot{\alpha} \geq 0\) everywhere on the 
+closed set \(\bar{\mathcal{O}}_1\), with equality only at the upright
+equilibrium \((q_u,p_u) = (\pi,0)\).
+We will use this later to prove that all orbits starting in
+\(\mathcal{O}_1\) cross the boundary and enter the rotation region.
+
 \subsubsection*{Time Scaling}
 
-We have observed that \(\dot{\alpha} > 0\) on \(\mathcal{O}\) for small enough \(I\). 
+We have observed in ~\eqref{eqn:acrobot-mu-dot}-\eqref{eqn:acrobot-alpha-dot}
+that \(\dot{\alpha} > 0\) on \(\mathcal{O}_1\) for small enough \(I\). 
 Hence, we can use \(\alpha\) as our time variable by reparameterizing \(t\) as a
@@
 \begin{cases}
-    \dot{\hat{\mu}}_0 = g(\alpha,\hat{\mu}_0,0)
+    \pdiff{\hat{\mu}_0}{\alpha} = g(\alpha,\hat{\mu}_0,0)
     , \\
@@
 We now take a first-order Taylor approximation of \(\hat{\mu}(\alpha,\mu_0,I)\)
-around the nominal system \(\hat{\mu}_0(\alpha,\mu_0)\), which 
-expands into 
+around the nominal \(\hat{\mu}_0(\alpha,\mu_0)\),
 \begin{equation}\label{eqn:acrobot-muhat-approx}
@@
     \begin{cases}
-        \dot{\hat{\mu}}_1 = 
+        \pdiff{\hat{\mu}_1}{\alpha} = 
         \pdiff{g}{\hat{\mu}}(\alpha,\mu_0,0)\hat{\mu}_1 + \pdiff{g}{I}(\alpha, \mu_0, 0)
@@
 \end{align*}
-where we define
+where 
 \begin{align*}
@@
 \end{align*}
-Hence, the solution to ~\eqref{eqn:acrobot-mu1-dot} is given by
+Equation ~\eqref{eqn:acrobot-mu1-dot} can be solved by quadrature, giving
 \[
@@
 intersects the \(q_u\)-axis, \ie, when \(\alpha = 0\) or \(\alpha = \pi\).
-Using our time-scaled system, we can define this Poincar\'{e} map as the change
-in \(\hat{\mu}\) when \(\alpha\) increases by \(\pi\):
+Using our time-scaled system ~\eqref{eqn:muhat-dot}, we can define this
+Poincar\'{e} map as the change in \(\hat{\mu}\) when \(\alpha\) increases by
+\(\pi\):
 \begin{align*}
-    P_\mathcal{O} : \,]0,\pi[ &\rightarrow \R_+,
+    P_{\mathcal{O}_1} : \,]0,\pi[ &\rightarrow \R_+,
     \\
@@
 \end{align*}
-From our perturbation analysis, the Poincar\'{e} map -- also known as a
-Poincar\'{e} section --  expands into
-\[
-    P_\mathcal{O}(\mu_0) = \mu_0 + I K \int\limits_0^\pi a(\sigma,\mu_0)d\sigma
+The Poincar\'{e} map expands into
+\[
+    P_{\mathcal{O}_1}(\mu_0) = \mu_0 + I K \int\limits_0^\pi a(\sigma,\mu_0)d\sigma
     + R(\pi,\mu_0,I)
@@
     Q(\mu_0) := \int \limits_0^\pi a(\sigma,\mu_0)d\sigma
-    ,
-\]
-so that \(\hat{\mu}_1(\pi,\mu_0) = K Q(\mu_0)\).
+    .
+\]
 Note that \(K\) is a positive constant which contains the acrobot's physical
-parameters, while \(a(\alpha,\mu)\) is adimensional. 
+parameters \(m\), \(g\), \(l\), and \(\bar{q}_a\), 
+while \(a(\alpha,\mu)\) is adimensional. 
 This means \(Q(\mu_0)\) is identical for every acrobot.
 If \(Q(\mu_0)\) is strictly positive, then \(\hat{\mu}_1(\pi,\mu_0)\) will be
-positive no matter the values of \(m\), \(g\), \(l\), or \(\bar{q}_a\).
+positive for any acrobot.
 
@@
 
+% TODO: Replace or remove this as per Manfredi's suggestions
 This fact, coupled with knowledge that \(R(\pi,\mu_0,I)\) is \(O(I^2)\),
@@
 \[
-    P_\mathcal{O}(\mu_0) = \mu_0 + IKQ(\mu_0) + R(\pi,\mu_0,I) > \mu_0
+    P_{\mathcal{O}_1}(\mu_0) = \mu_0 + IKQ(\mu_0) + R(\pi,\mu_0,I) > \mu_0
     .
@@
 
-\subsubsection*{Energy Gain on \(\mathcal{O}\)}
-
-The Poincar\'{e} section \(P_\mathcal{O}\) allows us understand the evolution of
+\subsubsection*{Energy Gain on \(\mathcal{O}_1\)}
+
+The Poincar\'{e} map \(P_{\mathcal{O}_1}\) allows us understand the evolution of
 \(\hat{\mu}(\alpha,\mu_0,I)\) by studying the evolution of the discrete time
@@
     \begin{cases}
-        \mu(n+1) := P_\mathcal{O}(\mu(n)) 
-        = \mu(n) + I K Q(\mu(n)) + R(\pi,\mu(n),I)
+        \mu_{n+1} := P_{\mathcal{O}_1}(\mu_n) 
+        = \mu_n + I K Q(\mu_n) + R(\pi,\mu_n,I)
         , \\
@@
 \end{equation}
-Tolerating the abuse of notation, \(\mu(n)\) represents the distance along the
+Here, \(\mu_n\) represents the distance along the
 \(q_u\)-axis when an orbit of the constrained dynamics intersects the
@@
 
-Proving the constrained dynamics gain energy on \(\mathcal{O}\) is equivalent to
-showing \(\mu(n)\) eventually reaches \(\pi\).
+Proving the constrained dynamics gain energy on \(\mathcal{O}_1\) is equivalent to
+showing \(\mu_n\) eventually reaches \(\pi\).
 That is, we want to find \(I^\star \in\, ]0, I_1]\) where, for all 
 \(\delta > 0\) and \(\mu_0 \in ]0,\pi-\delta]\), there exists \(N > 0\) so that
-for all \(n \geq N\), \(\mu(n) \notin [0,\pi-\delta]\).
+for all \(n \geq N\), \(\mu_n \notin [0,\pi-\delta]\).
+% TODO: Show two sufficient conditions, explain how we are using each, etc
 A sufficient condition for this characterization is to show 
-\(P_\mathcal{O}(\mu_0) > \mu_0\) for all \(\mu_0 \in\, ]0,\pi[\),
+\(P_{\mathcal{O}_1}(\mu_0) > \mu_0\) for all \(\mu_0 \in\, ]0,\pi[\),
 
@@
 \[
-    P^\prime_\mathcal{O}(0) = 1 + I K Q^\prime(0) +
+    P^\prime_{\mathcal{O}_1}(0) = 1 + I K Q^\prime(0) +
     \pdiff{R}{\mu}(\pi,0,I)
-    .
-\]
+    ,
+\]
+where prime denotes differentiation with respect to \(\mu\).
 If there is some value of \(I \in ]0,I_1]\) for which this term is greater than
@@
 \]
-MATLAB reveals that 
+Numerical computations reveal that 
 \[
@@
 \]
-Since the remainder term \(R(\alpha,\mu_0,I)\) is \(O(I^2)\), 
-its partial derivative \(\pdiff{R}{\mu}(\pi,\mu_0,I)\) is also \(O(I^2)\).
+Since the remainder term \(R(\pi,0,I)\) is \(O(I^2)\), 
+its partial derivative \(R^\prime(\pi,0,I)\) is also \(O(I^2)\).
 Hence, it can be written in the form
-\(I^2 \tilde{R}(\mu_0,I)\) where \(\tilde{R}(\mu_0,I)\) is smooth and
+\(I^2 \tilde{R}(I)\) where \(\tilde{R}(I)\) is smooth and
 zero at \(I = 0\).
-Since \(I\) is assumed to be small, \(I^2\) shrinks rapidly;
-thus, there exists \(I_2 \in ]0,I_1]\) where 
-\(I_2 K Q^\prime(0) + (I_2)^2 \tilde{R}(0,I) > 0\).
+Thus, there exists \(I_2 \in ]0,I_1]\) where 
+\(I_2 K Q^\prime(0) + (I_2)^2 \tilde{R}(I) > 0\).
 Hence,
 \[
-    P^\prime_\mathcal{O}(0) \geq 1 + IKQ^\prime(0) + I^2 \tilde{R}(0,I) 
+    P^\prime_{\mathcal{O}_1}(0) \geq 1 + IKQ^\prime(0) + I^2 \tilde{R}(I) 
     > 1
@@
 What's more, all solutions starting in this interval will flow towards the value
-\(\mu = \epsilon\), so we must have \(P_\mathcal{O}(\mu_0) > \mu_0\) for all 
-\(\mu_0 \in \, ]0,\epsilon[\).
+\(\mu = \epsilon\).
 
@@
 Therefore, it is bounded below on the compact set \(\mu_0 \in [\epsilon,\pi]\)
-by some value \(R(\pi,\underbar{\mu},I)\). 
+by some value \(\underbar{R}(I)\). 
 Theorem \ref{thm:khalil-perturbation} asserts that there exists some
@@
 \[
-    R(\pi,\mu_0,I) \geq R(\pi, \underbar{\mu}, I) > -I^2 r
+    R(\pi,\mu_0,I) \geq \underbar{R}(I) > -I^2 r
     .
@@
 \[
-    P_\mathcal{O}(\mu_0) > \mu_0 + I K Q(\epsilon) - I^2 r
-    .
-\]
-Choosing \(I^\star \in\, ]0,\min\left\{I_2,I_3\right\}]\) so that
-\(I^\star K Q(\epsilon) - (I^\star)^2 r \geq 0\)
+    P_{\mathcal{O}_1}(\mu_0) > \mu_0 + I K Q(\epsilon) - I^2 r
+    .
+\]
+Picking a small \(\lambda > 0\) and choosing 
+\(I^\star \in\, ]0,\min\left\{I_2,I_3\right\}]\) so that
+\(I^\star K Q(\epsilon) - (I^\star)^2 r \geq \lambda > 0\)
 means that
-\( P_\mathcal{O}(\mu_0) > \mu_0\)
+\( P_{\mathcal{O}_1}(\mu_0) > \mu_0 + \lambda\)
 for all \(\mu_0 \in [\epsilon,\pi[\).
 
-We have thus shown that \(P_\mathcal{O}(\mu_0) > \mu_0\) for all
-\(\mu_0 \in ]0,\pi[\).
 Looking at ~\eqref{eqn:muhat-discrete}, this means
-\(\mu(n+1) > \mu(n)\) and all solutions of the discrete time system will flow
+\(\mu_{n+1} > \mu_n + \lambda\), and all solutions of the discrete time system will flow
 towards \(\mu = \pi\).
 Hence, orbits of the constrained dynamics will eventually escape
-\(\mathcal{O}\).
+\(\mathcal{O}_1\).
 We conclude that, for all \(m\), \(g\), \(l\), \(\bar{q}_a\), 
 there exists \(I > 0\) small enough that the constraint 
-~\eqref{eqn:acrobot-constraint} is injecting energy on \(\mathcal{O}\).
+~\eqref{eqn:acrobot-constraint} is injecting energy on \(\mathcal{O}_1\).
 
 By the same arguments presented in this section, the Poincar\'{e} section
-satisfies \( P_\mathcal{O}(\mu_0) < \mu_0 \) when \(I < 0\). 
-In this case, the constraint dissipates energy on \(\mathcal{O}\).
+satisfies \( P_{\mathcal{O}_1}(\mu_0) < \mu_0 \) when \(I < 0\). 
+In this case, the constraint dissipates energy on \(\mathcal{O}_1\).
+
+\subsubsection*{Summary}
+
+We have proven the first part of Theorem \ref{thm:acrobot-energy-stabilization},
+which claims there exists \(I > 0\) small enough that 
+~\eqref{eqn:acrobot-constraint} injects energy into the acrobot on
+\(\mathcal{O}_1\).
+
+Here is a summary of the proof:
+\begin{enumerate}
+    \item We defined pseudo-polar coordinates \((\mu,\alpha)\) adapted to level
+        sets of the nominal energy function ~\eqref{eqn:acrobot-nominal-E}.
+    \item We showed there exists a value \(I_1\) small enough where 
+        \(\dot{\alpha} > 0\) everywhere on \(\mathcal{O}_1\).
+    \item We used time scaling to find \(\hat{\mu}\) and
+        expanded it into ~\eqref{eqn:acrobot-muhat-approx} using
+        perturbation theory.
+    \item Taking this expanded solution, we defined and expanded the
+        Poincar\'{e} map \(P_{\mathcal{O}_1}\) to analyze the discrete-time system
+        ~\eqref{eqn:muhat-discrete}.
+    \item We found a value \(I_2 \leq I_1\) making 0 a repeller of the
+        ~\eqref{eqn:muhat-discrete}, and a value \(I_3 \leq I_2\) which drives
+        all its orbits towards \(\mu = \pi\).
+\end{enumerate}
+Hence, choosing \(I \in [0,I_3]\) guarantees energy injection on
+\(\mathcal{O}_1\) because all orbits approach the pseudo-radius \(\mu = \pi\).
+Likewise, choosing \(I \in [-I_3,0]\) guarantees energy dissipation on
+\(\mathcal{O}_1\).
 
@@
 nominal pendulum obtained by setting \(I = 0\) look like open curves on some
-rotation region \(\mathcal{R} \subset \Gamma\).
+rotation region \(\mathcal{R} \subset \SxR\).
 One can compute this rotation domain to be the set
 \[
-    \mathcal{R} := \left\{ (q_u,p_u) \in \Gamma \mid E(q_u,p_u) > E(\pi,0)\right\}
+    \mathcal{R} := \left\{ (q_u,p_u) \in \SxR \mid E(q_u,p_u) > E(\pi,0)\right\}
     .
@@
 
-In the oscillation region \(\mathcal{O}\) we showed that the nominal pendulum's
+In the oscillation region \(\mathcal{O}_1\) we showed that the nominal pendulum's
 energy (and hence, its orbit) is uniquely determined by the intersection point
@@
 \[
-    \left\{(q_u,p_u) \in \Gamma \mid E(q_u,p_u) = \frac{\rho^2}{10ml^2}\right\}
+    \left\{(q_u,p_u) \in \SxR \mid E(q_u,p_u) = \frac{\rho^2}{10ml^2}\right\}
     ,
@@
 Hence, the constrained dynamics will eventually escape the set
-\(\omega \subset \mathcal{R}\) defined by
-\[
-    \omega := \left\{(q_u,p_u) \in \Gamma
+\(\mathcal{O}_2(\bar{\rho}) \subset \mathcal{R}\) defined by
+\[
+    \mathcal{O}_2(\bar{\rho}) := \left\{(q_u,p_u) \in \SxR
     \mid E(\pi,0) < E(q_u,p_u) < E(0,\bar{\rho}) \right\}
@@
 We conclude that there exists \(I > 0\) small enough that the constraint
-~\eqref{eqn:acrobot-constraint} is injecting energy on \(\omega\).
+~\eqref{eqn:acrobot-constraint} is injecting energy on \(\mathcal{O}_2(\bar{\rho})\).
 
@@
     \includestandalone[width=0.5\textwidth]{images/acrobot_little_omega}
-    \caption{The set \(\omega \subset \mathcal{R}\) on which the acrobot will gain energy.
+    \caption{The set \(\mathcal{O}_2(\bar{\rho}) \subset \mathcal{R}\) on which the acrobot will gain energy.
     }
@@
 satisfies \(\abs{P_\mathcal{R}(\rho_0)} < \abs{\rho_0}\) when \(I < 0\).
-In this case, the constraint dissipates energy on \(\omega\).
-
-\subsection{Energy Gain on \(\Omega\)}
+In this case, the constraint dissipates energy on \(\mathcal{O}_2(\bar{\rho})\).
+
+\subsection{Energy Gain on \(\mathcal{O}_2(\bar{\rho})\)}
 Define \(E_\pi\) to be the energy level set associated with \(E(\pi,0)\).
@@
 \begin{align*}
-    \Omega &= \left\{(q_u,p_u) \in \Gamma
+    \mathcal{O}_2(\bar{\rho}) &= \left\{(q_u,p_u) \in \SxR
     \mid E(q_u,p_u) < E(0,\bar{\rho}) \right\}
     ,\\
-    &= \mathcal{O} \cup E_\pi \cup \omega
+    &= \mathcal{O}_1 \cup E_\pi \cup \mathcal{O}_2(\bar{\rho})
     .
@@
 There exists some control input \(I_1 > 0\) for which the acrobot is gaining
-energy on \(\mathcal{O}\). 
-Hence, there are no closed orbits in \(\mathcal{O}\) and almost all orbits 
+energy on \(\mathcal{O}_1\). 
+Hence, there are no closed orbits in \(\mathcal{O}_1\) and almost all orbits 
 hit \(E_\pi\) in finite time.
 Likewise, there exists some \(I_2 > 0\) for which the acrobot is gaining energy
-on \(\omega\), so there are no closed orbits in \(\omega\) and almost every
+on \(\mathcal{O}_2(\bar{\rho})\), so there are no closed orbits in \(\mathcal{O}_2(\bar{\rho})\) and almost every
 orbit is flowing away from \(E_\pi\).
 Choosing \(I^\star \in ]0,\min\left\{I_1,I_2\right\}]\) means the acrobot will gain energy on
-both \(\mathcal{O}\) as well as \(\omega\).
-
-We can prove energy gain on \(\Omega\) by showing two properties of the system.
+both \(\mathcal{O}_1\) as well as \(\mathcal{O}_2(\bar{\rho})\).
+
+We can prove energy gain on \(\mathcal{O}_2(\bar{\rho})\) by showing two properties of the system.
 First, that the set \(\Pi^+\) of initial conditions converging to  \((\pi,0)\)
-is of measure zero in \(\Gamma\);
+is of measure zero in \(\SxR\);
 second, that \(E_\pi\) is not a positively-invariant set.
-If both these conditions are true, then orbits flowing from \(\mathcal{O}\) into
-\(E_\pi\) will generally swing past \(E_\pi\) into \(\omega\).
+If both these conditions are true, then orbits flowing from \(\mathcal{O}_1\) into
+\(E_\pi\) will generally swing past \(E_\pi\) into \(\mathcal{O}_2(\bar{\rho})\).
 
@@
 initial conditions converging to \((\pi,0)\) is one-dimensional, and hence is
-of measure zero in \(\Gamma\).
+of measure zero in \(\SxR\).
 
@@
 Let \(x(t)\) be an orbit of the constrained dynamics initialized at 
-\(x(0) \in \mathal{O}\).
-Since the constrained dynamics gain energy on \(\mathcal{O}\), there exists a
+\(x(0) \in \mathcal{O}_1\).
+Since the constrained dynamics gain energy on \(\mathcal{O}_1\), there exists a
 time \(T > 0\) such that \(x(T) \in E_\pi\).
@@
 other equilibria, \(x(t)\) must converge to \((\pi,0)\).
-The initial condition \(x(0) \in \mathcal{O}\) was arbitrary, so this implies
-that \(\mathcal{O} \subset \Pi^+\).
-But \(\Pi^+\) is a set of measure zero, while \(\mathcal{O}\) has positive
+The initial condition \(x(0) \in \mathcal{O}_1\) was arbitrary, so this implies
+that \(\mathcal{O}_1 \subset \Pi^+\).
+But \(\Pi^+\) is a set of measure zero, while \(\mathcal{O}_1\) has positive
 measure. 
@@
 Non-invariance of \(E_\pi\) and instability of \((\pi,0)\) implies that 
-almost all orbits initialized in \(\mathcal{O}\) will reach \(E_\pi\) and
-eventually swing into \(\omega\), rather than stay in \(E_\pi\) and converge to
+almost all orbits initialized in \(\mathcal{O}_1\) will reach \(E_\pi\) and
+eventually swing into \(\mathcal{O}_2(\bar{\rho})\), rather than stay in \(E_\pi\) and converge to
 \((\pi,0)\). 
-We know that almost all orbits starting in \(\omega\) will reach 
+We know that almost all orbits starting in \(\mathcal{O}_2(\bar{\rho})\) will reach 
 \(E(0,\bar{\rho})\) in finite time;
-we conclude that almost all orbits starting in \(\Omega\) will depart the energy
-escaping \(\Omega\) all together.
-Hence, the acrobot is gaining energy on \(\Omega\).
-
-If \(I < 0\), orbits in \(\mathcal{O}\) will converge to the origin and orbits in
-\(\omega\) will flow towards \(E_\pi\).
+we conclude that almost all orbits starting in \(\mathcal{O}_2(\bar{\rho})\) will depart the energy
+escaping \(\mathcal{O}_2(\bar{\rho})\) all together.
+Hence, the acrobot is gaining energy on \(\mathcal{O}_2(\bar{\rho})\).
+
+If \(I < 0\), orbits in \(\mathcal{O}_1\) will converge to the origin and orbits in
+\(\mathcal{O}_2(\bar{\rho})\) will flow towards \(E_\pi\).
 However, the upright equilibrium \((\pi,0)\) will still be unstable, and
 \(E_\pi\) will remain a non-invariant set. 
-This means that almost all orbits in \(\omega\) will flow past \(E_\pi\) into
-\(\mathcal{O}\), so the constrained system will be gaining energy in
-negative-time. Thus, the constraint is dissipating energy on \(\Omega\).
+This means that almost all orbits in \(\mathcal{O}_2(\bar{\rho})\) will flow past \(E_\pi\) into
+\(\mathcal{O}_1\), so the constrained system will be gaining energy in
+negative-time. Thus, the constraint is dissipating energy on \(\mathcal{O}_2(\bar{\rho})\).
 

### 2020-12-12 latex/acrobot.tex
Full edit of oscillation, including expansion from O1 into R

@@
 
-These dynamics do not gain energy everywhere; we need to define some sets where
-this energy gain is possible.
+These dynamics do not always gain energy; rather, energy gain is only
+guaranteed under certain conditions on the size of \(I\).
 First, let 
@@
         \includestandalone[width=\linewidth]{images/acrobot_oscillation_domain}
-        \caption{The set \(\mathcal{O}\). An orbit starting in this set (blue)
-            will reach the level set \(E(\pi,0)\) of the nominal pendulum.}
+        \caption{The set \(\mathcal{O}_1\). An orbit starting in this set (blue)
+            will pass through the level set \(E(\pi,0)\) of the nominal pendulum.}
         \label{fig:acrobot-oscillation-domain}
@@
         \includestandalone[width=\linewidth]{images/acrobot_omega}
-        \caption{The set \(\mathcal{O}_2(\bar{\rho})\). Orbits starting in this set will
-            reach the level set \(E(0,\bar{\rho})\) of the nominal pendulum.}
+        \caption{The set \(\mathcal{O}_2(\bar{\rho})\). Orbits starting in this
+            set are guaranteed to reach the level set \(E(0,\bar{\rho})\) of the
+            nominal pendulum.}
             \label{fig:acrobot-o2}
     \end{subfigure}
-    \caption{The sets on which Theorem \ref{thm:acrobot-energy-stabilization}
-        claims the acrobot gains energy.}
+    \caption{The sets on which the acrobot gains energy, according to Theorem
+        \ref{thm:acrobot-energy-stabilization}.}
 \end{figure}
@@
     \(\mathcal{O}_1\).
+    What's more, almost every orbit will escape the closure of 
+    \(\mathcal{O}_1\) and begin rotating.
     If instead \(I < 0\), the VNHC dissipates energy on this set.
@@
 escape any compact subset of \(\mathcal{O}_1\) in finite time. 
-What's more, the acrobot's phase will flow towards the homoclinic orbit
+What's more, the acrobot's phase will flow towards (and through) the homoclinic orbit
 corresponding to \(E(\pi,0)\), as in Figure
@@
 From Figure \ref{fig:pendulum-level-sets}, we know that oscillation level sets
-look like ellipses contained in \(\mathcal{O}\).
+look like ellipses contained in \(\mathcal{O}_1\).
 Since perturbation theory requires us to analyze the nominal pendulum, we will
@@
 \begin{align}
-        \label{eqn:mu-qupu}
-        \mu &= \arccos\left(C_\mu(q_u,p_u)\right), \\
         \label{eqn:alpha-qupu}
@@
         \right|_{\mu = \arccos(C_\mu(q_u,p_u))}
+        ,\\
+        \label{eqn:mu-qupu}
+        \mu &= \arccos\left(C_\mu(q_u,p_u)\right).
         .
@@
 \left.\begin{cases}
+    \dot{\alpha} &= \pdiff{\alpha(q_u,p_u)}{q_u}\dot{q}_u + 
+        \pdiff{\alpha(q_u,p_u)}{p_u}\dot{p}_u
+    \\
     \dot{\mu} &= \pdiff{\mu(q_u,p_u)}{q_u}\dot{q_u} +
          \pdiff{\mu(q_u,p_u)}{p_u}\dot{p}_u
-    \\
-    \dot{\alpha} &= \pdiff{\alpha(q_u,p_u)}{q_u}\dot{q}_u + 
-        \pdiff{\alpha(q_u,p_u)}{p_u}\dot{p}_u
-    \end{cases}\right|_{(q_u,p_u) = T\inv(\mu,\alpha)}
+    \end{cases}\right|_{(q_u,p_u) = T\inv(\alpha,\mu)}
     .
@@
 These dynamics are much too large to write out, so we simply denote them by
-\begin{align}\label{eqn:acrobot-mu-dot}
-    \dot{\mu} &= f_\mu(\alpha,\mu,I)
-    ,\\
+\begin{align}
     \label{eqn:acrobot-alpha-dot}
     \dot{\alpha} &= f_\alpha(\alpha,\mu,I)
+    ,\\
+    \label{eqn:acrobot-mu-dot}
+    \dot{\mu} &= f_\mu(\alpha,\mu,I)
     .
 \end{align}
-The dynamics of the nominal pendulum arise from evaluating
-\(f_\mu\) and \(f_\alpha\) at \(I = 0\).
-MATLAB's symbolic toolbox evaluates the nominal dynamics as
-\begin{align}\label{eqn:acrobot-mu-dot-nom}
-    \dot{\mu} &= 0
-    , \\
+MATLAB's symbolic toolbox evaluates the nominal dynamics (when \(I = 0\)) as
+\begin{align}
     \label{eqn:acrobot-alpha-dot-nom}
@@
             {\mu^2 \sin(\alpha)^2}}
+    , \\
+    \label{eqn:acrobot-mu-dot-nom}
+    \dot{\mu} &= 0
     .
 \end{align}
-Note that ~\eqref{eqn:acrobot-alpha-dot-nom} has a removable singularity at
-\(\alpha \in \{0,\pi\}\): 
-taking the limit as \(\alpha\) approaches these points yields
+Note that ~\eqref{eqn:acrobot-alpha-dot-nom} has removable singularities at
+\(\alpha \in \{0,\pi\}\).
+Taking the limit as \(\alpha\) approaches these points yields
 \[
@@
 which is smooth and well-defined for all \(\mu \in ]0,\pi[\).
-Taking this removable singularity into account,
-one can confirm that \(\dot{\alpha} > 0\) for every \(\mu \in ]0,\pi[\).
+Taking these removable singularities into account,
+one can verify that \(\dot{\alpha} > 0\) for every \(\mu \in ]0,\pi[\).
 By continuity of ~\eqref{eqn:acrobot-alpha-dot}, there exists \(I_1 > 0\) small
@@
 These pseudo-polar coordinates can be extended smoothly to 
-\(\bar{\mathcal{O}}_1\), the closure of \(\mathcal{O}_1\).
+the boundary of \(\mathcal{O}_1\).
 Notice that, for each \(\alpha \in \Sone\),
-\[
+\begin{equation}\label{eqn:alpha-dot-boundary}
     \lim \limits_{\mu \to \pi} \dot{\alpha}
@@
     ,
-\]
+\end{equation}
 with equality if and only if \(\alpha \in \{0,\pi\}\).
-In other words, the coordinates \(\mu = \pi, \alpha\) dictate the behaviour of
+In other words, setting \(\mu = \pi\) dictates the behaviour of
 the nominal pendulum on the homoclinic orbit with energy \(E(\pi,0)\).
 Hence, for small enough \(I\) we have \(\dot{\alpha} \geq 0\) everywhere on the 
-closed set \(\bar{\mathcal{O}}_1\), with equality only at the upright
+closure of \(\mathcal{O}_1\), with equality only at the upright
 equilibrium \((q_u,p_u) = (\pi,0)\).
 We will use this later to prove that all orbits starting in
-\(\mathcal{O}_1\) cross the boundary and enter the rotation region.
+\(\mathcal{O}_1\) will eventually begin rotating.
 
@@
 \]
-Equations ~\eqref{eqn:acrobot-mu-dot-nom}-\eqref{eqn:acrobot-alpha-dot-nom})
+Equations ~\eqref{eqn:acrobot-mu-dot-nom}-\eqref{eqn:acrobot-alpha-dot-nom}
 reveal that \(g(\alpha,\mu,0) = 0\),
@@
 \begin{align*}
-    P_{\mathcal{O}_1} : \,]0,\pi[ &\rightarrow \R_+,
+    P_\mathcal{O} : \,]0,\pi[ &\rightarrow \R_+,
     \\
@@
 \[
-    P_{\mathcal{O}_1}(\mu_0) = \mu_0 + I K \int\limits_0^\pi a(\sigma,\mu_0)d\sigma
+    P_\mathcal{O}(\mu_0) = \mu_0 + I K \int\limits_0^\pi a(\sigma,\mu_0)d\sigma
     + R(\pi,\mu_0,I)
@@
 
-% TODO: Replace or remove this as per Manfredi's suggestions
-This fact, coupled with knowledge that \(R(\pi,\mu_0,I)\) is \(O(I^2)\),
-implies there exists \(I > 0\) small enough that
-\[
-    P_{\mathcal{O}_1}(\mu_0) = \mu_0 + IKQ(\mu_0) + R(\pi,\mu_0,I) > \mu_0
-    .
-\]
-Unfortunately, perturbation analysis does not guarantee 
-that \(I\) is the same for all \(\mu_0 \in \, ]0,\pi[\), since the value of
-\(R(\pi,\mu_0,I)\) could become more negative for other initial conditions.
-We will deal with this next.
-
 \subsubsection*{Energy Gain on \(\mathcal{O}_1\)}
 
-The Poincar\'{e} map \(P_{\mathcal{O}_1}\) allows us understand the evolution of
+The Poincar\'{e} map \(P_\mathcal{O}\) allows us understand the evolution of
 \(\hat{\mu}(\alpha,\mu_0,I)\) by studying the evolution of the discrete time
@@
 \begin{equation}\label{eqn:muhat-discrete}
-    \begin{cases}
-        \mu_{n+1} := P_{\mathcal{O}_1}(\mu_n) 
+        \mu_{n+1} := P_\mathcal{O}(\mu_n) 
         = \mu_n + I K Q(\mu_n) + R(\pi,\mu_n,I)
-        , \\
-        \mu(0) = \mu_0
-        .
-    \end{cases}
-\end{equation}
+        ,
+\end{equation}
+with initial condition \(\mu_0\).
 Here, \(\mu_n\) represents the distance along the
@@
 for all \(n \geq N\), \(\mu_n \notin [0,\pi-\delta]\).
-% TODO: Show two sufficient conditions, explain how we are using each, etc
-A sufficient condition for this characterization is to show 
-\(P_{\mathcal{O}_1}(\mu_0) > \mu_0\) for all \(\mu_0 \in\, ]0,\pi[\),
-
-Linearizing ~\eqref{eqn:muhat-discrete} at \(\mu = 0\) yields
-\[
-    P^\prime_{\mathcal{O}_1}(0) = 1 + I K Q^\prime(0) +
-    \pdiff{R}{\mu}(\pi,0,I)
+
+A sufficient conditions for this characterization is to prove that 
+\(P_\mathcal{O}(\mu_0) \geq \mu_0 + \lambda\) for some \(\lambda > 0\).
+We will separate our analysis of the Poincar\'{e} map into two sections.
+First, we will prove that orbits near the origin flow away from it by showing the
+origin is a repeller of ~\eqref{eqn:muhat-discrete}.
+Then, we will use the sufficient condition to guarantee all orbits reach 
+\(\mu = \pi\).
+
+Linearizing ~\eqref{eqn:muhat-discrete} at \(\mu_0 = 0\) yields
+\[
+    P^\prime_\mathcal{O}(0) = 1 + I K Q^\prime(0) +
+    R^\prime(\pi,0,I)
     ,
@@
 where prime denotes differentiation with respect to \(\mu\).
-If there is some value of \(I \in ]0,I_1]\) for which this term is greater than
+If there is some value of \(I \in \,]0,I_1]\) for which this term is greater than
 \(1\), then \(0\) is a repeller of the discrete time system.
@@
 zero at \(I = 0\).
-Thus, there exists \(I_2 \in ]0,I_1]\) where 
+Thus, there exists \(I_2 \in \,]0,I_1]\) where 
 \(I_2 K Q^\prime(0) + (I_2)^2 \tilde{R}(I) > 0\).
@@
 \[
-    P^\prime_{\mathcal{O}_1}(0) \geq 1 + IKQ^\prime(0) + I^2 \tilde{R}(I) 
+    P^\prime_\mathcal{O}(0) \geq 1 + IKQ^\prime(0) + I^2 \tilde{R}(I) 
     > 1
@@
 \] 
-for all \(I \in ]0, I_2]\), .
+for all \(I \in \,]0, I_2]\).
 
@@
 
-Next, recall that \(R(\pi,\mu_0,I)\) is smooth in all its parameters as well as
-being \(O(I^2)\).
-Therefore, it is bounded below on the compact set \(\mu_0 \in [\epsilon,\pi]\)
+To complete this proof, recall that \(R(\pi,\mu_0,I)\) is smooth in all its
+parameters as well as being \(O(I^2)\).
+It is therefore bounded below on the compact set \(\mu_0 \in [\epsilon,\pi]\)
 by some value \(\underbar{R}(I)\). 
@@
 \]
+Note that we can assume \(I_3 \leq I_2\) without loss of generality.
 Furthermore, \(Q(\mu_0)\) is a strictly increasing function, so
-for\(\mu_0 \in [\epsilon,\pi[\) we have
-\[
-    P_{\mathcal{O}_1}(\mu_0) > \mu_0 + I K Q(\epsilon) - I^2 r
+for \(\mu_0 \in [\epsilon,\pi[\),
+\[
+    P_\mathcal{O}(\mu_0) > \mu_0 + I K Q(\epsilon) - I^2 r
     .
@@
 Picking a small \(\lambda > 0\) and choosing 
-\(I^\star \in\, ]0,\min\left\{I_2,I_3\right\}]\) so that
-\(I^\star K Q(\epsilon) - (I^\star)^2 r \geq \lambda > 0\)
-means that
-\( P_{\mathcal{O}_1}(\mu_0) > \mu_0 + \lambda\)
+\(I^\star \in\, ]0, I_3]\) so that
+\[
+    I^\star K Q(\epsilon) - (I^\star)^2 r \geq \lambda > 0
+    ,
+\]
+means that \( P_\mathcal{O}(\mu_0) \geq \mu_0 + \lambda\)
 for all \(\mu_0 \in [\epsilon,\pi[\).
 
-Looking at ~\eqref{eqn:muhat-discrete}, this means
-\(\mu_{n+1} > \mu_n + \lambda\), and all solutions of the discrete time system will flow
-towards \(\mu = \pi\).
-Hence, orbits of the constrained dynamics will eventually escape
-\(\mathcal{O}_1\).
+Looking at ~\eqref{eqn:muhat-discrete} we find that
+\(\mu_{n+1} \geq \mu_n + \lambda\). 
+This implies that all solutions of the discrete time system will flow towards
+\(\mu = \pi\).
+
 We conclude that, for all \(m\), \(g\), \(l\), \(\bar{q}_a\), 
@@
 By the same arguments presented in this section, the Poincar\'{e} section
-satisfies \( P_{\mathcal{O}_1}(\mu_0) < \mu_0 \) when \(I < 0\). 
+satisfies \( P_\mathcal{O}(\mu_0) \leq \mu_0 - \lambda\) when \(I < 0\). 
 In this case, the constraint dissipates energy on \(\mathcal{O}_1\).
@@
 \begin{enumerate}
-    \item We defined pseudo-polar coordinates \((\mu,\alpha)\) adapted to level
-        sets of the nominal energy function ~\eqref{eqn:acrobot-nominal-E}.
+    \item We defined pseudo-polar coordinates \((\alpha,\mu)\) adapted to level
+        sets of the nominal pendulum's mechanical energy.
     \item We showed there exists a value \(I_1\) small enough where 
@@
     \item Taking this expanded solution, we defined and expanded the
-        Poincar\'{e} map \(P_{\mathcal{O}_1}\) to analyze the discrete-time system
+        Poincar\'{e} map \(P_\mathcal{O}\) to analyze the discrete-time system
         ~\eqref{eqn:muhat-discrete}.
@@
 \end{enumerate}
-Hence, choosing \(I \in [0,I_3]\) guarantees energy injection on
+Hence, choosing \(I \in \,]0,I_3]\) guarantees energy injection on
 \(\mathcal{O}_1\) because all orbits approach the pseudo-radius \(\mu = \pi\).
-Likewise, choosing \(I \in [-I_3,0]\) guarantees energy dissipation on
+Likewise, choosing \(I \in [-I_3,0[\) guarantees energy dissipation on
 \(\mathcal{O}_1\).
@@
 \end{proof}
+
+\subsection{Energy Gain on \(\bar{\mathcal{O}}_1\)}
+Before moving on to the rotation analysis, we must first confirm that any
+acrobot constrained by ~\eqref{eqn:acrobot-constraint} will eventually start
+rotating.
+The definition of energy gain states that all orbits initialized in
+\(\mathcal{O}_1\) will escape compact sets of \(\mathcal{O}_1\) in finite time. 
+This is not enough to prove that all orbits will escape the boundary of
+\(\mathcal{O}_1\).
+
+Indeed, let \(E_\pi\) be the level set with energy \(E(\pi,0)\), 
+which forms the boundary of \(\mathcal{O}_1\). 
+Define \(\bar{\mathcal{O}}_1 = \mathcal{O}_1 \cup E_\pi\) to be the closure of
+\(\mathcal{O}_1\).
+An orbit of the acrobot will only begin rotating if it escapes 
+\(\bar{\mathcal{O}}_1\) by crossing through \(E_\pi\).
+It is possible for orbits starting in \(\mathcal{O}_1\) to always approach
+\(E_\pi\) without ever crossing into the rotation zone.
+We will prove this does not happen in general.
+
+To begin, let us analyze the upright equilibrium.
+Taking the Jacobian of the constrained dynamics
+~\eqref{eqn:acrobot-constrained-dynamics} at \((q_u,p_u) = (\pi,0)\) yields
+\[
+    J = \begin{bmatrix}
+        -\frac{6mgl\bar{q}_aI}{5} & \frac{1 - 2m^2gl^3\bar{q}_a^2 I^2}{5ml^2} \\
+        3mgl & mgl\bar{q}_aI
+    \end{bmatrix}
+    ,
+\]
+which has characteristic polynomial
+\[
+    \det\left(\lambda \Id{2} - J\right)
+    = \lambda^2 + \frac{mgl\bar{q}_a I}{5} \lambda - 3g
+    .
+\]
+By Descartes' rule of signs, this polynomial has one root with positive real
+part. 
+The equilibrium \((\pi,0)\) is therefore unstable, so the set \(\Pi^+\) of
+initial conditions converging to \((\pi,0)\) is one-dimensional, and hence is
+of measure zero in \(\SxR\).
+
+Using \(x(t) := (q_u(t),p_u(t))\) as shorthand, let
+\(x(0) \in \mathcal{O}_1\) be a nonzero initial condition of the acrobot.
+Suppose the orbit \(x(\R)\) does not cross through \(E_\pi\)
+into the rotation zone.
+Since the acrobot gains energy on \(\mathcal{O}_1\), the positive limit set of
+\(x(t)\) must be \(E_\pi\). 
+By the Birkhoff Theorem, \(x(t)\) converges to an invariant subset of \(E_\pi\)
+\cite{birkhoff}.
+From our discussion on pseudo-polar coordinates (in particular the extension
+~\eqref{eqn:alpha-dot-boundary} of \(\dot{\alpha}\) to \(\bar{\mathcal{O}}_1\)),
+we know that \(\dot{\alpha} \geq 0\) on \(E_\pi\), with equality if and only if 
+\(\alpha \in \{0,\pi\}\).
+In other words, the only invariant subset of \(E_\pi\) is the upright equilibrium.
+Hence, \(x(t)\) must converge to \((\pi,0)\), which means \(x(0) \in \Pi^+\).
+Since we know that \(\Pi^+\) is a set of measure zero in \(\SxR\), almost every
+other orbit initialized in \(\mathcal{O}_1\) must escape through the boundary
+\(E_\pi\) into the rotation domain.
 
@@
 
-\subsubsection*{Energy Gain on a Subset of \(\mathcal{R}\)}
-
+\subsubsection*{Energy Gain on \(\mathcal{O}_2(\bar{\rho})\)}
+%TODO: Rework this whole section
 The Poincar\'{e} section \(P_\mathcal{R}\) allows us understand the evolution of
@@
 
-\subsection{Energy Gain on \(\mathcal{O}_2(\bar{\rho})\)}
-Define \(E_\pi\) to be the energy level set associated with \(E(\pi,0)\).
-We will prove the acrobot gains energy on the set 
-\begin{align*}
-    \mathcal{O}_2(\bar{\rho}) &= \left\{(q_u,p_u) \in \SxR
-    \mid E(q_u,p_u) < E(0,\bar{\rho}) \right\}
-    ,\\
-    &= \mathcal{O}_1 \cup E_\pi \cup \mathcal{O}_2(\bar{\rho})
-    .
-\end{align*}
-There exists some control input \(I_1 > 0\) for which the acrobot is gaining
-energy on \(\mathcal{O}_1\). 
-Hence, there are no closed orbits in \(\mathcal{O}_1\) and almost all orbits 
-hit \(E_\pi\) in finite time.
-Likewise, there exists some \(I_2 > 0\) for which the acrobot is gaining energy
-on \(\mathcal{O}_2(\bar{\rho})\), so there are no closed orbits in \(\mathcal{O}_2(\bar{\rho})\) and almost every
-orbit is flowing away from \(E_\pi\).
-Choosing \(I^\star \in ]0,\min\left\{I_1,I_2\right\}]\) means the acrobot will gain energy on
-both \(\mathcal{O}_1\) as well as \(\mathcal{O}_2(\bar{\rho})\).
-
-We can prove energy gain on \(\mathcal{O}_2(\bar{\rho})\) by showing two properties of the system.
-First, that the set \(\Pi^+\) of initial conditions converging to  \((\pi,0)\)
-is of measure zero in \(\SxR\);
-second, that \(E_\pi\) is not a positively-invariant set.
-If both these conditions are true, then orbits flowing from \(\mathcal{O}_1\) into
-\(E_\pi\) will generally swing past \(E_\pi\) into \(\mathcal{O}_2(\bar{\rho})\).
-
-To show the first condition, we will simply prove that \((\pi,0)\) is an
-unstable equilibrium of the constrained dynamics.
-Taking the Jacobian of the constrained dynamics
-~\eqref{eqn:acrobot-constrained-dynamics} at \((\pi,0)\) yields
-\[
-    J = \begin{bmatrix}
-        -\frac{6mgl\bar{q}_aI}{5} & \frac{1 - 2m^2gl^3\bar{q}_a^2 I^2}{5ml^2} \\
-        3mgl & mgl\bar{q}_aI
-    \end{bmatrix}
-    ,
-\]
-which has the characteristic polynomial
-\[
-    \det\left(\lambda \Id{2} - J\right)
-    = \lambda^2 + \frac{mgl\bar{q}_a I}{5} \lambda - 3g
-    .
-\]
-By Descartes' rule of signs, this polynomial has one root with positive real
-part. 
-The equilibrium \((\pi,0)\) is therefore unstable, so the set \(\Pi^+\) of
-initial conditions converging to \((\pi,0)\) is one-dimensional, and hence is
-of measure zero in \(\SxR\).
-
-For the second condition, let us seek a contradiction by 
-supposing that \(E_\pi\) is a positively-invariant set. 
-Let \(x(t)\) be an orbit of the constrained dynamics initialized at 
-\(x(0) \in \mathcal{O}_1\).
-Since the constrained dynamics gain energy on \(\mathcal{O}_1\), there exists a
-time \(T > 0\) such that \(x(T) \in E_\pi\).
-By the assumed positive invariance of \(E_\pi\), this means 
-\(x(t) \in E_\pi\) for all \(t \in [T,\infty[\).
-Since \(E_\pi\) is a closed curve containing the equilibrium \((\pi,0)\) and no
-other equilibria, \(x(t)\) must converge to \((\pi,0)\).
-The initial condition \(x(0) \in \mathcal{O}_1\) was arbitrary, so this implies
-that \(\mathcal{O}_1 \subset \Pi^+\).
-But \(\Pi^+\) is a set of measure zero, while \(\mathcal{O}_1\) has positive
-measure. 
-This is a contradiction: \(E_\pi\) cannot be positively-invariant.
-
-Non-invariance of \(E_\pi\) and instability of \((\pi,0)\) implies that 
-almost all orbits initialized in \(\mathcal{O}_1\) will reach \(E_\pi\) and
-eventually swing into \(\mathcal{O}_2(\bar{\rho})\), rather than stay in \(E_\pi\) and converge to
-\((\pi,0)\). 
-We know that almost all orbits starting in \(\mathcal{O}_2(\bar{\rho})\) will reach 
-\(E(0,\bar{\rho})\) in finite time;
-we conclude that almost all orbits starting in \(\mathcal{O}_2(\bar{\rho})\) will depart the energy
-escaping \(\mathcal{O}_2(\bar{\rho})\) all together.
-Hence, the acrobot is gaining energy on \(\mathcal{O}_2(\bar{\rho})\).
-
-If \(I < 0\), orbits in \(\mathcal{O}_1\) will converge to the origin and orbits in
-\(\mathcal{O}_2(\bar{\rho})\) will flow towards \(E_\pi\).
-However, the upright equilibrium \((\pi,0)\) will still be unstable, and
-\(E_\pi\) will remain a non-invariant set. 
-This means that almost all orbits in \(\mathcal{O}_2(\bar{\rho})\) will flow past \(E_\pi\) into
-\(\mathcal{O}_1\), so the constrained system will be gaining energy in
-negative-time. Thus, the constraint is dissipating energy on \(\mathcal{O}_2(\bar{\rho})\).
-
 \section{Experimental Results}

### 2020-12-13 latex/acrobot.tex
First full edit of latest acrobot proof

@@
 \end{equation}
-which is seen in Figure \ref{fig:acrobot-oscillation-domain}.
+as in Figure \ref{fig:acrobot-oscillation-domain}.
 Finally, let \(\bar{\rho} > \sqrt{60m^2gl^3}\) and define the set
@@
     What's more, almost every orbit will escape the closure of 
-    \(\mathcal{O}_1\) and begin rotating.
-    If instead \(I < 0\), the VNHC dissipates energy on this set.
+    \(\mathcal{O}_1\).
+    If instead \(I < 0\), the VNHC dissipates energy.
 \item Define \(b : \SxR_{> 0} \rightarrow \R\) by
@@
     \]
-    and set \(S(\rho_0) := \int \limits_{0}^{2\pi} b(\sigma,\rho_0)d\sigma\).
+    and define \(S(\rho_0) := \int \limits_{0}^{2\pi} b(\sigma,\rho_0)d\sigma\).
     Fix \(\bar{\rho} > \sqrt{60m^2gl^3}\).
-    Suppose there exists \(\epsilon > 0\) so that \(S(\rho_0) > \epsilon\) for
+    Suppose there exists \(\epsilon > 0\) so that \(S(\rho_0) \geq \epsilon\) for
     all \(\rho_0 \in \, \left]\sqrt{60m^2gl^3}, \bar{\rho}\right]\).
@@
     \(\mathcal{O}_2(\bar{\rho})\).
-    If instead \(I < 0\), the VNHC dissipates energy on this set.
+    If instead \(I < 0\), the VNHC dissipates energy.
 \end{enumerate}
@@
 Theorem \ref{thm:acrobot-energy-stabilization} considers these sets separately.
-This separation is advantageous: 
-regardless of one's choice of acrobot, 
+This separation is advantageous because the first result holds for any
+\(m\), \(g\), \(l\), and \(\bar{q}_a\). 
+That is, regardless of one's choice of acrobot, 
 a small enough fixed value \(I > 0\) will enable the acrobot to 
@@
 
-Hence, the first result of Theorem \ref{thm:acrobot-energy-stabilization} states
-that all acrobots constrained by ~\eqref{eqn:acrobot-constraint} 
-will gain enough energy to begin rotating around the bar.
+In other words, the first result of Theorem
+\ref{thm:acrobot-energy-stabilization} states that all acrobots constrained by
+~\eqref{eqn:acrobot-constraint} will gain enough energy to begin rotating around
+the bar.
 In the worst case, the acrobot will at least perform a swing-up routine to reach
@@
 \ref{thm:acrobot-energy-stabilization} pertains to energy injection once the
-acrobot has started rotating - that is, when its phase lies in 
-\(\mathcal{O}_2(\bar{\rho}) \backslash \mathcal{O}_1\).
-If we want the acrobot to achieve giants with energy
-\(E(0,\bar{\rho})\), the acrobot needs to satisfy the assumption on the
-integral of \(b(\beta,\rho_0)\).
-In this case, orbits will escape compact subsets of
+acrobot has started rotating.
+For the acrobot to achieve giants with energy
+\(E(0,\bar{\rho})\), it must satisfy the assumption on the integral of
+\(b(\beta,\rho_0)\).
+The value of this integral depends on the acrobot's physical parameters.
+If the assumption holds, orbits will escape compact subsets of
 \(\mathcal{O}_2(\bar{\rho})\) in finite time until the acrobot eventually
@@
 One example of this behaviour is displayed in Figure \ref{fig:acrobot-o2}.
-Note that the value of \(I\) will depend on \(\bar{\rho}\).
-
-Unfortunately, the proof used in Chapter \ref{ch:vlp} does not
+Note that the size of \(I\) will depend on \(\bar{\rho}\).
+
+The proof used in Chapter \ref{ch:vlp} does not
 readily transfer to the acrobot.
@@
 \section{Proof of Theorem \ref{thm:acrobot-energy-stabilization}}\label{sec:acrobot-proof}
-In an effort to make the proof of Theorem \ref{thm:acrobot-energy-stabilization}
-as clear as possible, we will break it down into the following segments:
+To make the proof of Theorem \ref{thm:acrobot-energy-stabilization}
+as clear as possible, we break it down into the following segments:
 \begin{enumerate}
@@
 
-\subsection{Perturbation Analysis for Oscillations}
+\subsection{Perturbation Analysis for Oscillations}\label{sec:acrobot-proof-o1}
 As we have seen, setting \(I = 0\) turns the acrobot into a nominal pendulum. 
@@
 along with a pseudo-angle \(\alpha \in \Sone\) satisfying \(\dot{\alpha} > 0\) on
-\(\mathcal{O}\).
+\(\mathcal{O}_1\).
 
@@
         \label{eqn:mu-qupu}
-        \mu &= \arccos\left(C_\mu(q_u,p_u)\right).
+        \mu &= \arccos\left(C_\mu(q_u,p_u)\right)
         .
@@
 Let us define
-\[
+\begin{equation}\label{eqn:acrobot-Qmu}
     Q(\mu_0) := \int \limits_0^\pi a(\sigma,\mu_0)d\sigma
     .
-\]
+\end{equation}
 Note that \(K\) is a positive constant which contains the acrobot's physical
@@
 A sufficient conditions for this characterization is to prove that 
-\(P_\mathcal{O}(\mu_0) \geq \mu_0 + \lambda\) for some \(\lambda > 0\).
-We will separate our analysis of the Poincar\'{e} map into two sections.
-First, we will prove that orbits near the origin flow away from it by showing the
-origin is a repeller of ~\eqref{eqn:muhat-discrete}.
-Then, we will use the sufficient condition to guarantee all orbits reach 
+\(P_\mathcal{O}(\mu_0) \geq \mu_0 + \gamma\) for some \(\gamma > 0\).
+We will perform our analysis in two sections.
+First, we will show the origin is a repeller of ~\eqref{eqn:muhat-discrete}.
+This implies that orbits near the origin flow away from it.
+Then, we will use the sufficient condition above to guarantee all orbits reach
 \(\mu = \pi\).
@@
 zero at \(I = 0\).
-Thus, there exists \(I_2 \in \,]0,I_1]\) where 
-\(I_2 K Q^\prime(0) + (I_2)^2 \tilde{R}(I) > 0\).
+Thus, there exists \(I_2 \in \,]0,I_1]\) such that
+\[
+    I_2 K Q^\prime(0) + (I_2)^2 \tilde{R}(I) > 0
+    .
+\]
 Hence,
@@
 
-To complete this proof, recall that \(R(\pi,\mu_0,I)\) is smooth in all its
+To complete the proof, recall that \(R(\pi,\mu_0,I)\) is smooth in all its
 parameters as well as being \(O(I^2)\).
@@
 \]
-Picking a small \(\lambda > 0\) and choosing 
+Picking a small \(\gamma > 0\) and choosing 
 \(I^\star \in\, ]0, I_3]\) so that
 \[
-    I^\star K Q(\epsilon) - (I^\star)^2 r \geq \lambda > 0
-    ,
-\]
-means that \( P_\mathcal{O}(\mu_0) \geq \mu_0 + \lambda\)
+    I^\star K Q(\epsilon) - (I^\star)^2 r \geq \gamma > 0
+    ,
+\]
+means that \( P_\mathcal{O}(\mu_0) \geq \mu_0 + \gamma\)
 for all \(\mu_0 \in [\epsilon,\pi[\).
@@
 Looking at ~\eqref{eqn:muhat-discrete} we find that
-\(\mu_{n+1} \geq \mu_n + \lambda\). 
+\(\mu_{n+1} \geq \mu_n + \gamma\). 
 This implies that all solutions of the discrete time system will flow towards
@@
 By the same arguments presented in this section, the Poincar\'{e} section
-satisfies \( P_\mathcal{O}(\mu_0) \leq \mu_0 - \lambda\) when \(I < 0\). 
+satisfies \( P_\mathcal{O}(\mu_0) \leq \mu_0 - \gamma\) when \(I < 0\). 
 In this case, the constraint dissipates energy on \(\mathcal{O}_1\).
 
-\subsubsection*{Summary}
-
-We have proven the first part of Theorem \ref{thm:acrobot-energy-stabilization},
-which claims there exists \(I > 0\) small enough that 
-~\eqref{eqn:acrobot-constraint} injects energy into the acrobot on
-\(\mathcal{O}_1\).
-
-Here is a summary of the proof:
-\begin{enumerate}
-    \item We defined pseudo-polar coordinates \((\alpha,\mu)\) adapted to level
-        sets of the nominal pendulum's mechanical energy.
-    \item We showed there exists a value \(I_1\) small enough where 
-        \(\dot{\alpha} > 0\) everywhere on \(\mathcal{O}_1\).
-    \item We used time scaling to find \(\hat{\mu}\) and
-        expanded it into ~\eqref{eqn:acrobot-muhat-approx} using
-        perturbation theory.
-    \item Taking this expanded solution, we defined and expanded the
-        Poincar\'{e} map \(P_\mathcal{O}\) to analyze the discrete-time system
-        ~\eqref{eqn:muhat-discrete}.
-    \item We found a value \(I_2 \leq I_1\) making 0 a repeller of the
-        ~\eqref{eqn:muhat-discrete}, and a value \(I_3 \leq I_2\) which drives
-        all its orbits towards \(\mu = \pi\).
-\end{enumerate}
-Hence, choosing \(I \in \,]0,I_3]\) guarantees energy injection on
-\(\mathcal{O}_1\) because all orbits approach the pseudo-radius \(\mu = \pi\).
-Likewise, choosing \(I \in [-I_3,0[\) guarantees energy dissipation on
-\(\mathcal{O}_1\).
-
-% Empty proof environment to put QED square on the right of the page
-\begin{proof}[\unskip\nopunct]
-\end{proof}
-
-\subsection{Energy Gain on \(\bar{\mathcal{O}}_1\)}
+\subsubsection*{Energy Gain on \(\bar{\mathcal{O}}_1\)}
 Before moving on to the rotation analysis, we must first confirm that any
@@
 The definition of energy gain states that all orbits initialized in
-\(\mathcal{O}_1\) will escape compact sets of \(\mathcal{O}_1\) in finite time. 
+\(\mathcal{O}_1\) will escape compact subsets of \(\mathcal{O}_1\) in finite time. 
 This is not enough to prove that all orbits will escape the boundary of
@@
 \(x(t)\) must be \(E_\pi\). 
-By the Birkhoff Theorem, \(x(t)\) converges to an invariant subset of \(E_\pi\)
-\cite{birkhoff}.
+By the Birkhoff Theorem \cite{birkhoff}, \(x(t)\) converges to an invariant
+subset of \(E_\pi\).
 From our discussion on pseudo-polar coordinates (in particular the extension
@@
 
+\subsubsection*{Summary}
+
+We have proven the first part of Theorem \ref{thm:acrobot-energy-stabilization},
+which claims there exists \(I > 0\) small enough that 
+~\eqref{eqn:acrobot-constraint} injects energy into the acrobot on
+\(\mathcal{O}_1\).
+Enough energy is injected that orbits will exit the closure of \(\mathcal{O}_1\)
+and enter the rotation domain.
+
+Here is a summary of the proof:
+\begin{enumerate}
+    \item We found pseudo-polar coordinates \((\alpha,\mu)\) adapted to level
+        sets of the nominal pendulum's mechanical energy on \(\mathcal{O}_1\).
+    \item We showed there exists a value \(I_1\) small enough where 
+        \(\dot{\alpha} > 0\) everywhere on \(\mathcal{O}_1\).
+    \item We found the time-scaled pseudo-radius \(\hat{\mu}\) and
+        expanded it into ~\eqref{eqn:acrobot-muhat-approx} using
+        perturbation theory.
+    \item Taking this expanded solution, we defined and expanded the
+        Poincar\'{e} map \(P_\mathcal{O}\) to analyze the discrete-time system
+        ~\eqref{eqn:muhat-discrete}.
+    \item We found a value \(I_2 \leq I_1\) making the origin a repeller of 
+        ~\eqref{eqn:muhat-discrete}, and a value \(I_3 \leq I_2\) which drives
+        \(\mu_n\) towards \(\mu = \pi\).
+    \item We showed that the upright equilibrium \((\pi,0)\) is unstable, and
+        used this fact to prove that almost every orbit crosses through \(E_\pi\)
+        into the rotation zone.
+\end{enumerate}
+Choosing \(I \in \,]0,I_3]\) guarantees energy injection on
+\(\mathcal{O}_1\) because all orbits approach the pseudo-radius \(\mu = \pi\).
+Likewise, choosing \(I \in [-I_3,0[\) guarantees energy dissipation on
+\(\mathcal{O}_1\).
+
+% Empty proof environment to put QED square on the right of the page
+\begin{proof}[\unskip\nopunct]
+\end{proof}
+
 \subsection{Perturbation Analysis for Rotations}
@@
 nominal pendulum obtained by setting \(I = 0\) look like open curves on some
-rotation region \(\mathcal{R} \subset \SxR\).
-One can compute this rotation domain to be the set
+rotation domain \(\mathcal{R} \subset \SxR\), where
 \[
@@
 
-In the oscillation region \(\mathcal{O}_1\) we showed that the nominal pendulum's
-energy (and hence, its orbit) is uniquely determined by the intersection point
-\(\mu\) with the \(q_u\)-axis;
-likewise, on \(\mathcal{R}\) the energy is determined by the intersection
-\(\rho\) on the \(p_u\)-axis, as in Figure \ref{fig:rho-intersection}.
+In the oscillation region \(\mathcal{O}_1\) we showed that the energy of the
+nominal pendulum \eqref{eqn:acrobot-nominal-dynamics} is uniquely determined by
+an orbit's intersection point \(\mu\) with the \(q_u\)-axis.
+Likewise, on \(\mathcal{R}\) the energy is determined by the orbit's intersection
+point \(\rho\) with the \(p_u\)-axis, as in Figure \ref{fig:rho-intersection}.
 
@@
 \(p_u^2 = 60m^2 g l^3\). 
-The level set of this boundary is \(E(\pi,0)\), the homoclinic orbit for the
-upright equilibrium of the pendulum.
-Hence, we must have \(\rho > \sqrt{60 m^2 g l^3}\) if a
-rotation has momentum \(p_u > 0\) 
+This boundary is the homoclinic orbit for the
+upright equilibrium of the pendulum, which is the level set with energy
+\(E(\pi,0)\).
+Hence, we must have \(\rho > \sqrt{60 m^2 g l^3}\), if a
+rotation has momentum \(p_u > 0\), 
 and \(\rho < -\sqrt{60 m^2 g l^3}\) if it has momentum \(p_u < 0\).
@@
 
-Inverting this relationship gives the smooth transformation
+Inverting this relationship gives our pseudo-polar coordinates
 \begin{align}\label{eqn:rotation-T}
@@
 Using \(\beta\) as our new time variable (via a time reparameterization
-\(t = t(\beta)\)) produces the time-scaled system
-\(\hat{\rho}(\beta,\rho_0,I) = \rho(t(\beta),\rho_0,I)\) with dynamics
-\[
+\(t = t(\beta)\)) produces the time-scaled pseudo-radius
+\(\hat{\rho}(\beta) := \rho(t(\beta))\).
+This reduces the system \((\dot{\beta},\dot{\rho})\) into the scalar time-varying
+ODE
+\begin{equation}\label{eqn:rhohat-dot}
     \begin{cases}
@@
     \end{cases}
-\]
+\end{equation}
 
@@
     \begin{cases} 
-        \dot{\hat{\rho}}_0 = 0
-        , \\
-        \hat{\mu}_0(0) = \rho_0
+        \diff{\hat{\rho}_0}{\beta} = 0
+        , \\
+        \hat{\rho}_0(0) = \rho_0
         ,
@@
 \]
-with solution \(\hat{\rho}_0(\beta,\rho_0) \equiv \rho_0\) for all \(\beta\).
-Likewise, ~\eqref{eqn:khalil-perturbation-firstorder} 
-tells us that \(\hat{\rho}_1(\beta,\rho_0)\) is the solution to the linear
-time-varying scalar system
+which has solution \(\hat{\rho}_0(\beta,\rho_0) \equiv \rho_0\).
+
+We take a first-order Taylor approximation of \(\hat{\rho}(\beta,\rho_0,I)\)
+around \(\hat{\rho}_0(\beta,\rho_0)\) to get
+\begin{equation}\label{eqn:acrobot-rhohat-approx}
+    \hat{\rho}(\beta,\rho_0,I) = \hat{\rho}_0(\beta,\rho_0) +
+    I\hat{\rho}_1(\beta,\rho_0) + R(\beta,\rho_0,I)
+    ,
+\end{equation}
+where \(R(\beta,\rho_0,I)\) is smooth and \(O(I^2)\).
+
+Using MATLAB's symbolic toolbox and 
+~\eqref{eqn:khalil-perturbation-firstorder}, 
+we discover that \(\hat{\rho}_1(\beta,\rho_0)\) is the solution to the linear
+time-varying scalar ODE
 \begin{equation}\label{eqn:acrobot-rho1-dot}
   \begin{cases}
-    \dot{\hat{\rho}}_1 =
+      \diff{\hat{\rho}_1}{\beta} =
     \frac{5m^2 g l^3 \bar{q}_a \left(
@@
 \]
-Therefore, the first-order Taylor approximation of
-\(\hat{\rho}(\beta,\rho_0,I)\) is
-\begin{equation}\label{eqn:acrobot-rhohat-approx}
-    \hat{\rho}(\beta,\rho_0,I) = \hat{\rho}_0(\beta,\rho_0) +
-    I\hat{\rho}_1(\beta,\rho_0) + R(\beta,\rho_0,I)
-    ,
-\end{equation}
-where \(R(\beta,\rho_0,I)\) is smooth and \(O(I^2)\).
 
@@
 One full rotation amounts to \(\beta\) traversing \(2\pi\) rad in a clockwise
-direction; that is, when \(\beta\) goes from \(0\) to \(\sign{\rho_0}2\pi\)).
+direction; that is, when \(\beta\) goes from \(0\) to \(\sign{\rho_0}2\pi\).
 Since \(\dot{\beta}\) does not change sign on \(\mathcal{R}\),
@@
 
-The function \(b(\beta,\rho_0)\) is even and
+Let us pause here for a moment to remember what we are trying to accomplish.
+The second part of Theorem \ref{thm:acrobot-energy-stabilization} states that,
+under suitable conditions on the integral of \(b(\beta,\rho_0)\), 
+the acrobot will gain energy on 
+\[
+    \mathcal{O}_2(\bar{\rho}) := \left\{ (q_u,p_u) \in \SxR
+    \mid E(q_u,p_u) < E(0,\bar{\rho}) \right\}
+    ,
+\]
+for some \(\bar{\rho} > \sqrt{60m^2gl^3}\).
+Recall from Section \ref{sec:acrobot-proof-o1} that any acrobot initialized in
+\(\mathcal{O}_1\) will eventually enter the rotation region \(\mathcal{R}\).
+What remains, then, is to prove that the acrobot's orbits will also gain
+energy on
+\[
+    \mathcal{R}_{\bar{\rho}} := \mathcal{R} \cap \mathcal{O}_2(\bar{\rho})
+    .
+\]
+This can be proven by analyzing the Poincar\'{e} map \(P_\mathcal{R}(\rho_0)\).
+Looking at Figure \ref{fig:acrobot-rhobar-regions}, it is clear that
+\(P_\mathcal{R}\) has the domain
+\[
+    D := \left[-\bar{\rho},\sqrt{60m^2gl^3}\right[ \, 
+    \cup \, 
+    \left]\sqrt{60m^2gl^3},\bar{\rho}\right]
+    ,
+\]
+which is the intersection of \(\mathcal{R}_{\bar{\rho}}\) with the \(p_u\)-axis.
+We label the part of \(D\) contained in the negative \(p_u\)-axis as
+\[
+    D^- :=\left[-\bar{\rho},\sqrt{60m^2gl^3}\right[
+    ,
+\]
+and the part contained in the positive \(p_u\)-axis as
+\[
+    D^+ := \left]\sqrt{60m^2gl^3},\bar{\rho}\right]
+    ,
+\]
+so that \(D = D^- \cup D^+\).
+If \(P_\mathcal{R}(\rho_0)\) is always further from the origin than 
+\(\rho_0\) (\ie~ it is ``expanding" on \(D\)), the acrobot will be gaining energy
+on \(\mathcal{O}_2(\bar{\rho})\).
+
+\begin{figure}
+    \centering
+    \includestandalone[width=0.5\textwidth]{images/acrobot_rhobar_regions}
+    \caption{The region \(\mathcal{R}_{\bar{\rho}}\) of rotations in
+        \(\mathcal{O}_2(\bar{\rho})\) is coloured
+        in blue, while the domain \(D = D^- \cup D^+\) of the Poincar\'{e} map 
+        is coloured in red.}
+    \label{fig:acrobot-rhobar-regions}
+\end{figure}
+
+Mirroring the Poincar\'{e} analysis for oscillations, let us find a
+function \(S(\rho_0)\) which is positive on \(D\).
+This will be the rotational analogue to \(Q(\mu_0)\) from
+~\eqref{eqn:acrobot-Qmu}.
+
+Notice that the function \(b(\beta,\rho_0)\) is even and
 \(2\pi\)-periodic in \(\beta\), which implies that
@@
 Defining \(S(\rho_0) := \hat{\rho}_1(2\pi,\rho_0)\),
-we get
-\[
+the Poincar\'{e} map becomes
+\begin{equation}\label{eqn:acrobot-poincare-r}
     P_\mathcal{R}(\rho_0) = \rho_0 + I\sign{\rho_0}S(\rho_0) 
@@
     .
-\]
-
-Recall the assumption that there exist \(\bar{\rho} > 0\) and
+\end{equation}
+
+Proving expansion of \(P_\mathcal{R}(\rho_0)\) becomes easier if
+\(S(\rho_0)\) is positive on \(D\).
+Recall the assumption that there exists
 \(\epsilon > 0\) where \(S(\rho_0) \geq \epsilon\) for all 
-\(\rho_0 \in \,]\sqrt{60m^2gl^3},\bar{\rho}]\).
-Let us define
-\[
-    D := \left[-\bar{\rho},\sqrt{60m^2gl^3}\right[ \, 
-    \cup \, 
-    \left]\sqrt{60m^2gl^3},\bar{\rho}\right]
-    .
-\]
+\(\rho_0 \in D^+\).
 Since \(b(\beta,\rho_0)\) is even in \(\rho_0\),
-\(S(\rho_0) = S(\abs{\rho_0})\);
-hence, \(S(\rho_0) \geq \epsilon\) for every \(\rho_0 \in D\).
-
-Finally, recall that \(R(\sign{\rho_0}2\pi,\rho_0,I)\) is \(O(I^2)\), so 
-for each fixed \(\rho_0 \in D\)
-there exists \(I > 0\) small enough that 
-\( IS(\rho_0) + R(\sign{\rho_0}2\pi,\rho_0,I) > 0\).
-For this value of \(I\), we have that
-\[
-    \begin{cases}
-        P_\mathcal{R}(\rho_0) > \rho_0 & \rho_0 > 0
-        ,\\
-        P_\mathcal{R}(\rho_0) < \rho_0 & \rho_0 < 0
-        .\\
-    \end{cases}
-\]
-Unfortunately, this \(I\) may not work for all \(\rho_0 \in D\). We deal
-with this issue next.
+its integral \(S(\rho_0)\) is also even in \(\rho_0\),
+which means \(S(\rho_0) \geq \epsilon\) for every \(\rho_0 \in D\).
 
 \subsubsection*{Energy Gain on \(\mathcal{O}_2(\bar{\rho})\)}
-%TODO: Rework this whole section
 The Poincar\'{e} section \(P_\mathcal{R}\) allows us understand the evolution of
 \(\hat{\rho}(\beta,\rho_0,I)\) by studying the evolution of the discrete time
-system
+ODE
 \begin{equation}\label{eqn:rhohat-discrete}
-    \begin{cases}
-        \rho(n+1) := P_\mathcal{R}(\rho(n)) 
-        = \rho(n) + I\sign{\rho_0}S(\rho(n)) + 
-            R(\sign{\rho_0}2\pi,\rho(n),I)
-        , \\
-        \rho(0) = \rho_0
-        .
-    \end{cases}
-\end{equation}
-Tolerating the abuse of notation, \(\rho(n)\) represents the distance along the
+    \rho_{n+1} := P_\mathcal{R}(\rho_n) 
+    = \rho_n + I\sign{\rho_n}S(\rho_n) + 
+        R(\sign{\rho_n}2\pi,\rho_n,I)
+    .
+\end{equation}
+Here, \(\rho_n\) represents the distance along the
 \(p_u\)-axis when an orbit of the constrained dynamics intersects the
@@
 
-We wish to show \(\abs{\rho(n)}\) eventually reaches \(\bar{\rho}\).
-That is, we wish to find some \(I^\star \in \,]0,I_1]\) making
-\(\abs{P_\mathcal{R}(\rho_0)} > \abs{\rho_0}\) for all 
+To prove energy gain on \(\mathcal{R}_{\bar{\rho}}\), it is
+enough to prove that \(\abs{\rho_n}\) eventually reaches \(\bar{\rho}\).
+This is true if there exists some \(\gamma > 0\)
+whereby \(\abs{P_\mathcal{R}(\rho_0)} \geq \abs{\rho_0} + \gamma\) for all 
 \(\rho_0 \in D\)
 
-We begin with examining the case where \(\rho_0 > 0\).
+We begin with \(\rho_0 \in D^+\).
 Recall that \(R(\beta,\rho_0,I)\) is \(O(I^2)\) and smooth in all
@@
 Therefore, it is bounded below on the compact set
-\(\rho_0 \in [\sqrt{60m^2gl^3},\bar{\rho}]\) by some value
-\(R(2\pi,\underbar{\rho},I)\).
+\(\rho_0 \in \left[\sqrt{60m^2gl^3},\bar{\rho}\right]\) by some value
+\(\underbar{R}(I)\).
 Theorem \ref{thm:khalil-perturbation} asserts that there exists some 
@@
 \[
-    R(2\pi,\rho_0,I) \geq R(2\pi,\underbar{\rho},I) > -I^2r
+    R(2\pi,\rho_0,I) \geq \underbar{R}(I) > -I^2r
     .
@@
 \]
-for all \(\rho_0 \in ]\sqrt{60m^2gl^3},\bar{\rho}]\).
-Choosing \(I_3 \in ]0,I_2]\) so that 
-\(I_3\epsilon - (I_3)^2 r \geq 0\) means that
-\(P_\mathcal{R}(\rho_0) > \rho_0\)
-for all \(\rho_0 \in ]\sqrt{60m^2gl^3},\bar{\rho}]\)
-
-We now examine the case when \(\rho_0 < 0\), where we wish to show that 
-\(P_\mathcal{R}(\rho_0) < \rho_0\).
+for all \(\rho_0 \in D^+\).
+Choosing \(I_3 \in \,]0,I_2]\) so that
+\[
+    I_3\epsilon - (I_3)^2 r \geq \gamma
+    ,
+\] 
+makes \(P_\mathcal{R}(\rho_0) \geq \rho_0 + \gamma\) for all 
+\(\rho_0 \in D^+\)
+
+Now take \(\rho_0 \in D^-\), where we wish to show that 
+\(P_\mathcal{R}(\rho_0) \leq \rho_0 - \gamma\).
 The remainder \(R(\beta,\rho_0,I)\) is bounded above on the compact set 
-\(\rho_0 \in [-\bar{\rho},-\sqrt{60m^2gl^3}]\), and 
+\(\rho_0 \in \left[-\bar{\rho},-\sqrt{60m^2gl^3}\right]\), and 
 Theorem \ref{thm:khalil-perturbation} asserts that there exists some 
@@
     , \\
-    &< \rho_0 - IS(\epsilon) + I^2r^\prime
+    &\leq \rho_0 - I\epsilon + I^2r^\prime
     .
@@
 Choosing \(I_5 \in ]0,I_4]\) so that 
-\((I_5)^2r^\prime - I_5 S(\epsilon) \leq 0\) makes
-\(P_\mathcal{R}(\rho_0) < \rho_0\) for all 
-\(\rho_0 \in [-\bar{\rho},-\sqrt{60m^2gl^3}[\).
-
-Finally, choosing \(I^\star \in\, ]0,\min\left\{I_3,I_5\right\}]\) yields that
-\(\abs{P_\mathcal{R}(\rho_0)} > \abs{\rho_0}\) for all \(\rho_0 \in D\), as
-desired.
-Looking at the discrete-time system ~\eqref{eqn:rhohat-discrete}, 
-a solution of the system will flow away from the origin until it reaches a
-magnitude \(\abs{\rho(n)} = \bar{\rho}\).
-Hence, the constrained dynamics will eventually escape the set
-\(\mathcal{O}_2(\bar{\rho}) \subset \mathcal{R}\) defined by
-\[
-    \mathcal{O}_2(\bar{\rho}) := \left\{(q_u,p_u) \in \SxR
-    \mid E(\pi,0) < E(q_u,p_u) < E(0,\bar{\rho}) \right\}
-    ,
-\]
-which is displayed in Figure \ref{fig:acrobot-little-omega}.
-
-We conclude that there exists \(I > 0\) small enough that the constraint
-~\eqref{eqn:acrobot-constraint} is injecting energy on \(\mathcal{O}_2(\bar{\rho})\).
-
-\begin{figure}
-    \centering
-    \includestandalone[width=0.5\textwidth]{images/acrobot_little_omega}
-    \caption{The set \(\mathcal{O}_2(\bar{\rho}) \subset \mathcal{R}\) on which the acrobot will gain energy.
-    }
-    \label{fig:acrobot-little-omega}
-\end{figure}
+\[
+    I_5 \epsilon - (I_5)^2r^\prime \geq \gamma
+    ,
+\] 
+makes \(P_\mathcal{R}(\rho_0) \leq \rho_0 - \gamma\) for all 
+\(\rho_0 \in D^-\).
+
+To connect the two subregions of \(D\) together, choose
+\(I^\star \in\, ]0,\min\left\{I_3,I_5\right\}]\).
+Then \(\abs{P_\mathcal{R}(\rho_0)} \geq \abs{\rho_0} + \gamma\) for all
+\(\rho_0 \in D\), so the constrained dynamics are gaining energy on 
+\(\mathcal{R}_{\bar{\rho}}\).
 
 By the same arguments presented in this section, the Poincar\'{e} section
-satisfies \(\abs{P_\mathcal{R}(\rho_0)} < \abs{\rho_0}\) when \(I < 0\).
-In this case, the constraint dissipates energy on \(\mathcal{O}_2(\bar{\rho})\).
+satisfies \(\abs{P_\mathcal{R}(\rho_0)} \leq \abs{\rho_0} - \gamma\) 
+when \(I < 0\).
+
+\subsubsection*{Conclusion}
+We have proven the second part of Theorem
+\ref{thm:acrobot-energy-stabilization}: if the acrobot satisfies the assumptions
+on \(S(\rho_0)\), there exists \(I > 0\) small enough that the constraint
+~\eqref{eqn:acrobot-constraint} injects energy into the acrobot on
+\(\mathcal{O}_2(\bar{\rho})\).
+
+Here is a summary of the proof:
+\begin{enumerate}
+    \item We found pseudo-polar coordinates \((\beta,\rho)\) adapted to level
+        sets of the nominal pendulum's mechanical energy on \(\mathcal{R}\).
+    \item We showed there exists a value \(I_1\) small enough where
+        \(\dot{\beta} > 0\) everywhere on \(\mathcal{R}\).
+    \item We found the time-scaled pseudo-radius \(\hat{\rho}\) and expanded it
+        into ~\eqref{eqn:acrobot-rhohat-approx} using perturbation theory.
+    \item Taking this expanded solution, we defined and expanded the
+        Poincar\'{e} map \(P_\mathcal{R}\) on the domain 
+        \(D = D^- \cup D^+\) which shows how the acrobot's orbits behave on
+        \(\mathcal{R}_{\bar{\rho}}\).
+    \item Using the Poincar\'{e} map to analyze the discrete-time system
+        ~\eqref{eqn:rhohat-discrete}, we found values \(I_2,I_4\) bounding the
+        remainder term \(R(\beta,\rho_0,I)\) on \(D^+\) and \(D^-\)
+        (respectively), then found values \(I_3, I_5\) which drive all orbits of
+        ~\eqref{eqn:rhohat-discrete} towards \(\abs{\rho} = \bar{\rho}\).
+\end{enumerate}
+Recall from Section \ref{sec:acrobot-proof-o1} that there exists \(I^\star > 0\)
+making orbits escape \(\bar{\mathcal{O}}_1\) into 
+\(\mathcal{R}\). 
+Choosing \(I \in \,]0,\min\{I_3,I_5,I^\star\}]\)
+guarantees energy injection on \(\mathcal{O}_2(\bar{\rho})\).
+Likewise, choosing \(I \in [-\min\{I_3,I_5,I^\star\},0[\) guarantees energy
+dissipation on \(\mathcal{O}_2(\bar{\rho})\).
+
+% Draw the proof square only
+\begin{proof}[\unskip\nopunct]
+\end{proof}
 

### 2020-12-14 sugar_bot/arduino/libraries/SUGAR_VNHC_Library/include/VNHC.h
Added the ArctanVNHC

@@ -259,6 +259,72 @@ private:
 
 }; // class SinuVNHC
 
+/**
+ * @brief: A class for the acrobot VNHC of the type qa = Arctan(pu), which is a
+ * regular VNHC that successfully injects energy into the acrobot in simulation.
+ */
+class ArctanVNHC : public AcrobotVNHC
+{
+public:
+    /**
+     * @brief: Constructs a VNHC of the type qa = qmax*2/pi*Arctan(pu). Note that qmax
+     * must be wtihin [0, pi]. If not, it is truncated to be within that
+     * value.
+     *
+     * @param: const Acrobot&
+     * @param: ActuatorLimit
+     */
+    ArctanVNHC(const Acrobot& acrobot, ActuatorLimit qmax);
+
+    /**
+     * @brief: Constructs a VNHC of the type qa = qmax*2/pi*Arctan(scale*pu). Note that
+     * qmax must be within [0, pi]. If not, it is truncated to be within that
+     * value.
+     *
+     * @param: const Acrobot&
+     * @param: ActuatorLimit
+     * @param: ScalingFactor
+     */
+    ArctanVNHC(const Acrobot& acrobot, ActuatorLimit qmax, ScalingFactor scale);
+
+    ~ArctanVNHC();
+
+    /**
+     * @brief: Implementation of AcrobotVNHC's qa function, which gives qa =
+     * qmax*2/pi*Arctan(scale*pu)
+     *
+     * @param: const UnactuatedPhase& 
+     * @return: double
+     */
+    auto qa(const UnactuatedPhase& qpu) const -> double;
+
+    /**
+    * @brief: Implementation of AcrobotVNHC's dqu function, whch gives 
+    * dh/dqu = 0 since d/dqu arctan(pu) = 0.
+    *
+    * @param: const UnactuatedPhase&
+    * @return: double
+    */
+    auto dqu(const UnactuatedPhase& qpu) const -> double;
+
+    /**
+    * @brief: Implementation of AcrobotVNHC's dpu function, which gives dh/dpu =
+    * -d/dpu qmax*2/pi*arctan(scale*pu)
+    *
+    * @param: const UnactuatedPhase& qpu
+    * @return: double 
+    */
+    auto dpu(const UnactuatedPhase& qpu) const -> double;
+
+private:
+    // Store the scaling factor
+    ScalingFactor scale_;
+    // Store the value for 2/pi
[…]

### 2020-12-14 sugar_bot/arduino/libraries/SUGAR_VNHC_Library/src/VNHC.cpp

@@ -164,6 +164,40 @@ auto SinuVNHC::dpu(const UnactuatedPhase& qpu) const -> double
     return -qmax()*qpu.qu*qpu.qu/qpu_norm3;
 }
 
+////////////////
+// ArctanVNHC //
+////////////////
+ArctanVNHC::ArctanVNHC(const Acrobot& acrobot, ActuatorLimit qmax)
+: ArctanVNHC(acrobot, qmax, ScalingFactor(1.0))
+{}
+
+ArctanVNHC::ArctanVNHC(const Acrobot& acrobot, ActuatorLimit qmax, ScalingFactor scale)
+: AcrobotVNHC(acrobot,qmax),
+  scale_(scale)
+{}
+
+ArctanVNHC::~ArctanVNHC()
+{}
+
+auto ArctanVNHC::qa(const UnactuatedPhase& qpu) const -> double
+{
+    // This should return qmax*Arctan(scale*pu)
+    return qmax()*pib2inv_*atan(scale_*qpu.pu);
+}
+
+auto ArctanVNHC::dqu(const UnactuatedPhase& qpu) const -> double
+{
+    // This should return -d/dqu Arctan(pu) = 0
+    return 0.0;
+}
+
+auto ArctanVNHC::dpu(const UnactuatedPhase& qpu) const -> double
+{
+    // This returns the derivative dh/dpu = -qmax* d/dpu Arctan(scale*pu) =
+    // -qmax*scale/(1 + scale^2*pu^2)
+    return -qmax()*pib2inv_*scale_/(1 + scale_*scale_*qpu.pu*qpu.pu);
+}
+
 //-------------------//
 // Private Functions //
 //-------------------//

### 2020-12-14 sugar_bot/arduino/libraries/SUGAR_VNHC_Library/src/test/VNHCTest.cpp

@@ -61,7 +61,8 @@ public:
         g_),
       zero_(xingbot_),
       tanhVNHC_(xingbot_, qmax_),
-      sinuVNHC_(xingbot_, qmax_)
+      sinuVNHC_(xingbot_, qmax_),
+      arctanVNHC_(xingbot_,qmax_)
     {}
 protected:
     void SetUp() override 
@@ -99,6 +100,7 @@ protected:
     // Define our concrete VNHCs
     TanhVNHC tanhVNHC_;
     SinuVNHC sinuVNHC_;
+    ArctanVNHC arctanVNHC_;
 }; // class VNHCTest
 
 // Test that dh/dqa is always 1 and dh/dpa = 0 for all inputs
@@ -220,6 +222,21 @@ TEST_F(VNHCTest, CONCRETE_PA_MATCHES_SOLVABILITY_EQUATIONS)
     qpu_.qu = sqrt(3); qpu_.pu = 1;
     auto pa_pb6 = 0.348085485082814; 
     EXPECT_PRED3(Within, pVNHC_->pa(qpu_), pa_pb6, eps_);
+
+    // Since we're getting the correct output for this with the previous two
+    // constraints, let's just check at pu = 0 for the arctan constraint.
+    pVNHC_ = &arctanVNHC_;
+
+    //At pu = 0, we get pa = -2.939607210240234e-04*sin(qu) - the same as
+    //tanh*2/pi.
+    qpu_.pu = 0;
+    for(int i = -10; i <= 10; ++i)
+    {
+        auto qu = i*M_PI/10;
+        qpu_.qu = qu; 
+        auto expected_pa = -2.939607210240234e-04*sin(qu);
+        ASSERT_PRED3(Within, pVNHC_->pa(qpu_), expected_pa, eps_);
+    }
 }
 
 // Test for the TanhVNHC, to make sure its outputs are valid
@@ -279,6 +296,32 @@ TEST_F(VNHCTest, SINU_OUTPUTS_VALID)
         }
     }
 }
+
+// Test for the ArctanVNHC, to make sure its outputs are valid
+TEST_F(VNHCTest, ARCTAN_OUTPUTS_VALID)
+{
+    for (int i = -10; i <= 10; ++i)
+    {
+        qpu_.qu = i*M_PI/10.0;
+        qpu_.qu = i/2.0;
+        for(int j = -5; j <= 5; ++j)
+        {
+            qpu_.pu = j*10/5.0;
+
+            // qa is only a function of pu
[…]

### 2020-12-16 latex/acrobot.tex
Final edit of acrobot draft

@@
 \begin{enumerate}
-    \item There exists \(I > 0\) small enough that 
+    \item There exists \(I^\star > 0\) such that, for all 
+        \(I \in \,]0,I^\star]\), 
     ~\eqref{eqn:acrobot-constraint} injects energy into the acrobot on
     \(\mathcal{O}_1\).
-    What's more, almost every orbit will escape the closure of 
-    \(\mathcal{O}_1\).
-    If instead \(I < 0\), the VNHC dissipates energy.
+    Moreover, almost every orbit will escape the closure of 
+    \(\mathcal{O}_1\) in finite time.
+    If instead \(I \in [-I^\star,0[\), the VNHC dissipates energy.
 \item Define \(b : \SxR_{> 0} \rightarrow \R\) by
@@
     all \(\rho_0 \in \, \left]\sqrt{60m^2gl^3}, \bar{\rho}\right]\).
-    Then there exists \(I > 0\) small enough that
+    Then there exists \(I^\star > 0\) such that, for all 
+    \(I \in \, ]0,I^\star]\)
     ~\eqref{eqn:acrobot-constraint} injects energy into the acrobot on
     \(\mathcal{O}_2(\bar{\rho})\).
-    If instead \(I < 0\), the VNHC dissipates energy.
+    If instead \(I \in [-I^\star,0[\), the VNHC dissipates energy.
 \end{enumerate}
@@
 escape any compact subset of \(\mathcal{O}_1\) in finite time.
-What's more, the acrobot's phase will flow towards (and through) the homoclinic orbit
-corresponding to \(E(\pi,0)\), as in Figure
-\ref{fig:acrobot-oscillation-domain}.
+Therefore, the acrobot will display an oscillatory behaviour whereby the
+amplitude of oscillation increases with time.
+In so doing, the acrobot's orbit will exit in finite time the closure of
+\(\mathcal{O}_1\) as in Figure \ref{fig:acrobot-oscillation-domain}.
 
@@
 part. 
-The equilibrium \((\pi,0)\) is therefore unstable, so the set \(\Pi^+\) of
+The equilibrium \((\pi,0)\) is therefore unstable, so the stable manifold \(\Pi^+\) of
 initial conditions converging to \((\pi,0)\) is one-dimensional, and hence is
@@
 \(x(0) \in \mathcal{O}_1\) be a nonzero initial condition of the acrobot.
-Suppose the orbit \(x(\R)\) does not cross through \(E_\pi\)
-into the rotation zone.
+Suppose the orbit \(x(\R)\) is confined within 
+\(\bar{\mathcal{O}}_1\), and does not exit through \(E_\pi\) into the rotation
+zone.
+Since \(\bar{\mathcal{O}}_1\) is compact, the Birkhoff Theorem \cite{birkhoff}
+implies:
+\begin{itemize}
+    \item The positive limit set \(L_+\) of \(x(t)\) is non-empty, compact, and
+        invariant.
+    \item The solution \(x(t)\) asymptotically tends to \(L_+\).
+\end{itemize}
 Since the acrobot gains energy on \(\mathcal{O}_1\), the positive limit set of
-\(x(t)\) must be \(E_\pi\). 
-By the Birkhoff Theorem \cite{birkhoff}, \(x(t)\) converges to an invariant
-subset of \(E_\pi\).
+\(x(t)\) must be the largest invariant subset of \(E_\pi\). 
 From our discussion on pseudo-polar coordinates (in particular the extension
@@
 \(\alpha \in \{0,\pi\}\).
-In other words, the only invariant subset of \(E_\pi\) is the upright equilibrium.
-Hence, \(x(t)\) must converge to \((\pi,0)\), which means \(x(0) \in \Pi^+\).
+There are two possibilities: either \(E_\pi\) itself is invariant, or the
+largest invariant subset of \(E_\pi\) is \(\{(\pi,0)\}\).
+To rule out the first possibility, take the derivative of \(E(q_u,p_u)\) at the
+\(p_u\)-axis to get
+\[
+    \dot{E}(0,p_u) = -\frac{g}{5l} \sin(q_a)p_u
+    .
+\]
+This is non-zero everywhere except at the origin, which means it is non-zero on
+\(E_\pi\). 
+This means \(E_\pi\) is not invariant, so the positive limit set of \(x(t)\)
+must be the set \(\{(\pi,0)\}\).
+Hence, \(x(t)\) converges to \((\pi,0)\), which means \(x(0) \in \Pi^+\).
 Since we know that \(\Pi^+\) is a set of measure zero in \(\SxR\), almost every
@@
 We have proven the first part of Theorem \ref{thm:acrobot-energy-stabilization},
-which claims there exists \(I > 0\) small enough that 
-~\eqref{eqn:acrobot-constraint} injects energy into the acrobot on
-\(\mathcal{O}_1\).
+which claims there exists a control value \(I^\star > 0\) such that, for 
+\(I \in \, ]0,I^\star]\), ~\eqref{eqn:acrobot-constraint} injects energy into
+the acrobot on \(\mathcal{O}_1\).
 Enough energy is injected that orbits will exit the closure of \(\mathcal{O}_1\)
@@
     \item We found pseudo-polar coordinates \((\alpha,\mu)\) adapted to level
-        sets of the nominal pendulum's mechanical energy on \(\mathcal{O}_1\).
+        sets of the nominal pendulum's mechanical energy on \(\mathcal{O}_1\),
+        where \(\alpha \in \Sone\) is a pseudo-angle and \(\mu \in \, ]0,\pi[\)
+        is a pseudo-radius.
     \item We showed there exists a value \(I_1\) small enough where 
         \(\dot{\alpha} > 0\) everywhere on \(\mathcal{O}_1\).
-    \item We found the time-scaled pseudo-radius \(\hat{\mu}\) and
-        expanded it into ~\eqref{eqn:acrobot-muhat-approx} using
-        perturbation theory.
+    \item Using \(\alpha\) as a time variable, we found the time-scaled
+        pseudo-radius \(\hat{\mu}\) and expanded it into
+        ~\eqref{eqn:acrobot-muhat-approx} using perturbation theory.
     \item Taking this expanded solution, we defined and expanded the
@@
     \item We showed that the upright equilibrium \((\pi,0)\) is unstable, and
-        used this fact to prove that almost every orbit crosses through \(E_\pi\)
-        into the rotation zone.
+        used this fact to prove that almost every orbit crosses through
+        \(E_\pi\), thereby exiting the closure of \(\mathcal{O}_1\) in finite
+        time.
 \end{enumerate}
@@
 \(\mathcal{O}_1\) because all orbits approach the pseudo-radius \(\mu = \pi\).
+In fact, almost all orbits escape \(\bar{\mathcal{O}}_1\).
 Likewise, choosing \(I \in [-I_3,0[\) guarantees energy dissipation on
@@
     \item We found pseudo-polar coordinates \((\beta,\rho)\) adapted to level
-        sets of the nominal pendulum's mechanical energy on \(\mathcal{R}\).
+        sets of the nominal pendulum's mechanical energy on \(\mathcal{R}\),
+        where \(\beta \in \Sone\) is a pseudo-angle and
+        \(\rho \in D\) is a pseudo-radius.
     \item We showed there exists a value \(I_1\) small enough where
         \(\dot{\beta} > 0\) everywhere on \(\mathcal{R}\).
-    \item We found the time-scaled pseudo-radius \(\hat{\rho}\) and expanded it
-        into ~\eqref{eqn:acrobot-rhohat-approx} using perturbation theory.
+    \item Using \(\beta\) as a time variable, we found the evolution of the
+        time-scaled pseudo-radius \(\hat{\rho}\) and expanded it into
+        ~\eqref{eqn:acrobot-rhohat-approx} using perturbation theory.
     \item Taking this expanded solution, we defined and expanded the
@@
 
+This analysis, together with that of section \ref{sec:acrobot-proof-o1},
+completes the proof of Theorem \ref{thm:acrobot-energy-stabilization}.
+
 % Draw the proof square only

### 2020-12-17 latex/introduction.tex
Added a setion header in intro for notation
Started the introduction

@@
 \chapter{Introduction}
+% TODO: Introduce the thesis with a short, sweet, intuitive approach.
+% No math, just whet the reader's appetite for what we are doing. What are the motivating examples? Why do we study what we study? Why choose VNHCs over other methods?
+Consider a child who is standing on a swing.
+Naturally, the child pushes off the ground and starts bending and extending
+their knees; they swing higher and faster over time, until they 
+have reached their express goal of going as high and fast as possible.
+
+Now imagine that the child is a robot, and their creator is teaching them to
+swing like a real boy.
+The first thing the roboticist might do is plot the leg motion of a human child
+as a function of time and have the robot synchronize with this plot.
+In an ideal world, this technique would work perfectly and the robot boy would
+swing high, experiencing the same soaring feeling as his human counterpart.
+
+Unfortunately, this behaviour is entirely unnatural. 
+Humans on swings do not bend or extend their knees at specific times; rather,
+they adjust their leg position based on the current angle of the swing along
+with their direction of motion, as in Figure \ref{fig:swing-pos-vel}.
+This position-velocity adjustment allows humans to correct for external
+disturbances (like strong winds or overly enthusiastic parents).
+Even if the robot could perfectly track a time-based trajectory, an external
+disturbance may desynchronize the leg motion with the swing to the point that
+the swing will slow down rather than speed up.
+
+\begin{figure}
+    \centering
+    %\includegraphis[width=0.5\textwidth]{images/swing_pos_vel.png}
+    \caption{A person on a standing swing will squat as they approach the
+        bottom of the swing path, and stand after they pass it until they hit
+        the peak of their swing. Image modified from one in
+    \cite{pumping_swing_standing_squatting}.}
+    \label{fig:swing-pos-vel}
+\end{figure}
+
+One recent control technique known as \textit{virtual holonomic constraints}
+forces a robot's actuators to track a function of position rather than time.
+Sadly, this still won't do for our robot boy because a real child also bends
+their knees according to their direction of motion.
+In other words, the robot's legs need to move as a function not only of
+position, but also of velocity.
+
+These are the types of controllers we study in this thesis: they are called
+\textit{virtual nonholonomic constraints}.
+
 \section{Literature Review}
 
+% TODO: write a brief lit review on energy injection and VNHCs.
+
 \section{Statement of Contributions}
+Here are the contributions of this thesis.
+\begin{itemize}[label={}]
+   \item \textbf{Chapter \ref{ch:vnhcs}} The development of the framework of
+      virtual nonholonomic constraints.
+      This includes the definition of simply actuated mechanical systems, virtual
+      nonholonomic constraints, regular constraints, and energy injection.
+      Theorem \ref{thm:vnhc-regularity} yields a characterization of regularity, 
+      while Theorem \ref{thm:zero-dynamics} provides the constrained dynamics
+      for a certain class of systems.
+   \item \textbf{Chapter \ref{ch:vlp}} An application of virtual nonholonomic
+      constraints to the variable-length pendulum, based on a pumping technique
+      used by children on standing swings.
+      The chapter culminates in Theorem \ref{thm:vlp-energy-stabilization},
+      which guarantees that a certain class of VNHCs will inject energy into
+      this system.
+   \item \textbf{Chapter \label{ch:acrobot}} An application of virtual
+      nonholonomic constraints to the acrobot, where we develop a constraint
+      by examining the intuitive motion performed by human gymnasts.
+      The chapter ends with Theorem \ref{thm:acrobot-energy-stabilization},
+      which proves that the constraint injects energy into the acrobot.
+\end{itemize}
 
-\section{Outline of the Thesis}
+\section{Organization of the Thesis}
+The thesis is laid out as such: 
+in Chapter \ref{ch:vnhcs} we cover the requisite background on Hamiltonian
+mechanics, after which we develop the main theory of virtual nonholonomic
+constraints;
+in Chapter \ref{ch:vlp} we reformulate a time-optimal energy-injection strategy
+for the variable-length pendulum as a virtual nonholonomic constraint, and
+prove that the pendulum will gain enough energy to rotate around the bar with
+arbitrary speed;
+and in Chapter \ref{ch:acrobot} we find a virtual nonholonomic constraint 
+which enables the acrobot to kick its legs like a gymnast until it is
+performing backflips on a horizontal bar.
+Finally, we show experimental results for the acrobot which confirm that the
+theory works in the real world.
 
 %/========== /Introduction ==========/%
-% vim: set ts=3 sw=3 sts=0 et ffs=unix :
+% vim: set tw=80 ts=4 sw=4 sts=0 et ffs=unix :

### 2020-12-17 latex/introduction.tex
Update to the intro with a new image

@@
     \centering
-    %\includegraphis[width=0.5\textwidth]{images/swing_pos_vel.png}
+    \includegraphics[width=0.75\textwidth]{images/swing_pos_vel.png}
     \caption{A person on a standing swing will squat as they approach the
         bottom of the swing path, and stand after they pass it until they hit
-        the peak of their swing. Image modified from one in
+        the peak of their swing. Image taken (with modifications) from
     \cite{pumping_swing_standing_squatting}.}
@@
       this system.
-   \item \textbf{Chapter \label{ch:acrobot}} An application of virtual
+   \item \textbf{Chapter \ref{ch:acrobot}} An application of virtual
       nonholonomic constraints to the acrobot, where we develop a constraint

### 2020-12-19 latex/introduction.tex
First edit of the intro, pre-lit review

@@
 % No math, just whet the reader's appetite for what we are doing. What are the motivating examples? Why do we study what we study? Why choose VNHCs over other methods?
-Consider a child who is standing on a swing.
-Naturally, the child pushes off the ground and starts bending and extending
-their knees; they swing higher and faster over time, until they 
-have reached their express goal of going as high and fast as possible.
+Consider a child on a standing swing.
+Naturally, they push off the ground and proceed to alternate between standing
+and squatting. 
+This allows them to swing higher and faster over time, until they have reached
+their express goal of going as high and fast as possible.
 
 Now imagine that the child is a robot, and their creator is teaching them to
-swing like a real boy.
-The first thing the roboticist might do is plot the leg motion of a human child
-as a function of time and have the robot synchronize with this plot.
-In an ideal world, this technique would work perfectly and the robot boy would
-swing high, experiencing the same soaring feeling as his human counterpart.
-
-Unfortunately, this behaviour is entirely unnatural. 
-Humans on swings do not bend or extend their knees at specific times; rather,
-they adjust their leg position based on the current angle of the swing along
-with their direction of motion, as in Figure \ref{fig:swing-pos-vel}.
+behave like a real boy.
+If the roboticist had studied classical control theory, they would plot the leg
+motion of a human child as a trajectory over time and tell the robot to
+synchronize with this trajectory.
+In an ideal world, this technique would work perfectly; unfortunately, this
+control approach is entirely unnatural. 
+Humans on swings do not have an internal stopwatch.
+Rather, they adjust their leg position based on the current angle of the swing
+along with their direction of motion, as in Figure \ref{fig:swing-pos-vel}.
 This position-velocity adjustment allows humans to correct for external
 disturbances (like strong winds or overly enthusiastic parents).
+A person will squat at the peak of the swing and stand at the bottom.
 Even if the robot could perfectly track a time-based trajectory, an external
-disturbance may desynchronize the leg motion with the swing to the point that
-the swing will slow down rather than speed up.
+disturbance may swap the order of squatting and standing to the point that
+the swing slows down rather than speeding up.
 
@@
         bottom of the swing path, and stand after they pass it until they hit
-        the peak of their swing. Image taken (with modifications) from
+        the peak of their swing. Image taken and modified from
     \cite{pumping_swing_standing_squatting}.}
@@
 forces a robot's actuators to track a function of position rather than time.
-Sadly, this still won't do for our robot boy because a real child also bends
-their knees according to their direction of motion.
-In other words, the robot's legs need to move as a function not only of
-position, but also of velocity.
+Sadly, this will not work because one needs velocity to determine when
+they have reached the peak of their swing.
+In other words, the robot needs to move according to a function of both position
+and velocity if it has any hope of behaving like a real person.
 
-These are the types of controllers we study in this thesis: they are called
+These position-velocity controllers are the focus of this thesis: they are called
 \textit{virtual nonholonomic constraints}.

### 2020-12-19 latex/conclusion.tex
First version of Intro and Conclusion
I commented out the rest of the thesis so that I could focus on editing
these. Make sure to put the rest of them back in!

@@
 \chapter{Conclusion}\label{ch:conclusion}
-% TODO: Summarize the work
 
-\section{Limitations of this Work}
+This thesis has shown the utility of VNHCs as a means of injecting energy into
+mechanical systems.
+In Chapter \ref{ch:vnhcs} we developed the framework of VNHCs for underactuated
+mechanical systems, with a focus on VNHCs for simply actuated Hamiltonian
+systems.
+We then applied this framework to two benchmark systems: the variable-length
+pendulum and the acrobot.
+In Chapter \ref{ch:vlp} we proved that a certain class of VNHCs will always
+inject energy into the VLP, while in Chapter \ref{ch:acrobot} we designed a
+constraint inspired by gymnastics and proved it injects energy into the
+acrobot.
+In the end, we demonstrated that virtual nonholonomic constraints are capable of
+injecting and dissipating energy in a robust manner, all while producing
+realistic biological motion.
 
-\section{Future Research}
+\section{Limitations and Future Research}
+Our VNHC framework is relies on the following assumptions:
+\begin{enumerate}
+    \item The input matrix \(B(q) \equiv B \in \R^{n \times k}\) is constant and
+        full rank.
+    \item The input matrix has a left-annihilator 
+        \(B^\perp \in \R^{(n-k)\times n}\). 
+    \item The annihilator matrix \(B^\perp\) is right semi-orthogonal.
+    \item The inertia matrix of the system satisfies 
+        \(\nabla_{q_u}M(q) = \Zmat{n(n-k) \times n}\).
+    \item On the constraint manifold \(\Gamma\), one can solve for \(q_a\) as a
+        function of \((q_u,p_u)\).
+\end{enumerate}
+If any of these assumptions are not satisfied, one may not be able to find the
+constrained dynamics in simply actuated coordinates, and the results may not
+hold.
+In particular, the theoretical guarantees of this thesis do not hold for systems
+with physical nonholonomic constraints like friction, nor does it apply to vehicles
+with wheels.
+
+This guides us to the following research directions. 
+First, one might consider relaxing the assumptions on the input and inertia
+matrices to widen the class of systems to which VNHCs can be applied.
+Second, one might forgo the assumption that \(q_a\) is solvable on the
+constraint manifold; instead, one could borrow from the VHC literature
+and parameterize the constraint manifold with a different set of
+coordinates than \((q_u,p_u)\).
+Third, one might consider abstract mechanical systems with dynamics
+\[
+    \begin{cases}
+        \dot{q} = \Minv(q) p
+        , \\
+        \dot{p} = -\pdmat - \nabla_q V(q) + Q_{nh}(q,p) + B(q)\tau
+        ,
+    \end{cases}
+\]
+where \(Q_{nh}(q,p) \in \R^n\) are the equations for physical nonholonomic
+constraints like friction.
+Fourth, one can investigate a smoother mechanism for energy stabilization, where
+a VNHC is explicitly designed to inject energy below some energy level, and
+dissipate energy above it.
+For example, this might involve a mechanism for smoothly transferring between 
+VNHCs while maintaining certain safety requirements.
+
+Finally, the proof of Theorem \ref{thm:acrobot-energy-stabilization} 
+in Chapter \ref{sec:acrobot-proof} opens a door to the idea of VNHC generation. 
+One might search for suitable conditions on a VNHC whereby the Poincar\'{e}
+sections for oscillations and rotations are increasing.
+Then, one can extend Otsason's ``virtual constraint
+generator" \cite{vcg} with the goal of generating regular VNHCs which satisfy
+these energy injection/dissipation conditions.
+In this way, one could automatically generate VNHCs which stabilize a
+desired energy level.
 
 %/========== /Conclusion ==========/%
-% vim: set ts=3 sw=3 sts=0 et ffs=unix :
+% vim: set tw=80 ts=4 sw=4 sts=0 et ffs=unix :

### 2020-12-19 latex/introduction.tex
First version of Intro and Conclusion
I commented out the rest of the thesis so that I could focus on editing
these. Make sure to put the rest of them back in!

@@
 \chapter{Introduction}
-% TODO: Introduce the thesis with a short, sweet, intuitive approach.
-% No math, just whet the reader's appetite for what we are doing. What are the motivating examples? Why do we study what we study? Why choose VNHCs over other methods?
+
 Consider a child on a standing swing.
-Naturally, they push off the ground and proceed to alternate between standing
-and squatting. 
+Naturally, they begin by pushing off the ground and proceed to alternate between
+standing and squatting. 
 This allows them to swing higher and faster over time, until they have reached
@@
 Humans on swings do not have an internal stopwatch.
-Rather, they adjust their leg position based on the current angle of the swing
-along with their direction of motion, as in Figure \ref{fig:swing-pos-vel}.
+Rather, they adjust how much they bend their knees based on the current angle of
+the swing along with their direction of motion.
 This position-velocity adjustment allows humans to correct for external
 disturbances (like strong winds or overly enthusiastic parents).
-A person will squat at the peak of the swing and stand at the bottom.
+Figure \ref{fig:swing-pos-vel} shows this usual adjustment:
+a person will squat at the peak of the swing and stand as they reach the fastest
+point at the bottom \cite{pumping_swing_standing_squatting}.
 Even if the robot could perfectly track a time-based trajectory, an external
@@
 One recent control technique known as \textit{virtual holonomic constraints}
-forces a robot's actuators to track a function of position rather than time.
-Sadly, this will not work because one needs velocity to determine when
-they have reached the peak of their swing.
+\cite{vhcs_for_el_systems} forces a robot's actuators to track a function of
+position rather than time.
+Sadly, this will not work for the swing-bot: one needs to know their velocity to
+determine when they have reached the peak of their swing.
 In other words, the robot needs to move according to a function of both position
 and velocity if it has any hope of behaving like a real person.
+These position-velocity controllers are known as 
+\textit{virtual nonholonomic constraints}, and they are the focus of this
+thesis.
 
-These position-velocity controllers are the focus of this thesis: they are called
-\textit{virtual nonholonomic constraints}.
+Throughout this thesis we will refer to two examples to build intuition.
+The first is the variable-length pendulum, which models the child on a swing.
+The second is the acrobot, which models a gymnast hanging on a horizontal bar.
+For each of these mechanical systems, we will develop virtual nonholonomic
+constraints which replicate human behaviours. 
+Finally, we will rigorously prove that these constraints inject energy into
+their respective robots.
 
 \section{Literature Review}
+Here we review some of the previous work on energy injection and virtual
+nonholonomic constraints.
 
-% TODO: write a brief lit review on energy injection and VNHCs.
+Energy injection for mechanical systems is often performed by passivity-based
+control and energy shaping --- one views the robot and its inputs as energy
+transformations which are moulded to force the system to behave as desired.
+Much of the theoretical work in this field has been performed by Ortega, van der
+Schaft, and others (see for example \cite{ida_pbc_underactuation_one,
+ida_pbc_acrobot_example,energy_shaping_revisited}).
+Energy shaping has been applied to both the variable-length
+pendulum \cite{vlp_energy_shaping} and the acrobot
+\cite{swingup_acrobot_energy,swingup_giant_acrobot}.
+This technique is useful, but it does not necessarily produce the structured
+human-like motion we wish to generate using virtual nonholonomic constraints.
+
+The modern concept of virtual nonholonomic constraints was described by
+\citet{nhvc_dynamic_walking} in 2015, though there are references to more
+primitive versions going back as early as the the year 2000
+\cite{vnhc_human_robot_cooperation}.
+Virtual nonholonomic constraints have been of most notable use in the area of
+bipedal locomotion \cite{nhvc_incline_walking,output_nhvc_bipedal_conrol},
+where they have shown marked improvements in the robustness of walking gaits
+when compared to previous control technqiues \cite{nhvc_incline_walking}.
+They have also been used for error-reduction in time-delayed teleoperation
+\cite{vnhc_time_delay_teleop} and human-robot interaction
+\cite{psd_based_vnhc_redundant_manipulator,haptic_vnhc}.
+All these fields use different formulations of virtual nonholonomic constraints;
+\citet{hybrid_zero_dynamics_bipedal_nhvcs} have derived the constrained dynamics
+of virtual nonholonomic constraints for hybrid mechanical systems,
+but to the best of our knowledge no one has attempted to generalize the
+concept to abstract mechanical systems.
+
+Many of our contributions in Chapter \ref{ch:vnhcs} extend
+concepts from the field of virtual \textit{holonomic} constraints. 
+Of note is the theoretical framework devised by Mohammadi, Maggiore, and
+Consolini, among others
+\cite{vhcs_for_el_systems,dynamic_vhcs_stabilize_closed_orbits,lagrangian_structure_reduced_dynamics_vhcs,xingbo_thesis}.
 
@@
       nonholonomic constraints, regular constraints, and energy injection.
-      Theorem \ref{thm:vnhc-regularity} yields a characterization of regularity, 
-      while Theorem \ref{thm:zero-dynamics} provides the constrained dynamics
-      for a certain class of systems.
+      Theorem \ref{thm:vnhc-regularity} yields a computational characterization
+      of regularity, while Theorem \ref{thm:zero-dynamics} explicitly finds the
+      constrained dynamics for a certain class of systems.
    \item \textbf{Chapter \ref{ch:vlp}} An application of virtual nonholonomic
@@
    \item \textbf{Chapter \ref{ch:acrobot}} An application of virtual
-      nonholonomic constraints to the acrobot, where we develop a constraint
-      by examining the intuitive motion performed by human gymnasts.
+      nonholonomic constraints to the acrobot, based on the motion performed by
+      human gymnasts.
       The chapter ends with Theorem \ref{thm:acrobot-energy-stabilization},
-      which proves that the constraint injects energy into the acrobot.
+      which proves that the constraint we design does in fact inject energy into
+      the acrobot.
 \end{itemize}
@@
 \section{Organization of the Thesis}
-The thesis is laid out as such: 
-in Chapter \ref{ch:vnhcs} we cover the requisite background on Hamiltonian
+The thesis is laid out as follows: 
+in Chapter \ref{ch:vnhcs} we cover the requisite background on analytical 
 mechanics, after which we develop the main theory of virtual nonholonomic
@@
 performing backflips on a horizontal bar.
-Finally, we show experimental results for the acrobot which confirm that the
-theory works in the real world.
+Finally, we show experimental results on a physical acrobot which confirm
+that the theory works in the real world.
 

### 2020-12-19 latex/conclusion.tex
Added a better intro, edited symbols and concusion

@@
 \section{Limitations and Future Research}
-Our VNHC framework is relies on the following assumptions:
+Our VNHC framework relies on the following assumptions:
 \begin{enumerate}
@@
 If any of these assumptions are not satisfied, one may not be able to find the
-constrained dynamics in simply actuated coordinates, and the results may not
-hold.
-In particular, the theoretical guarantees of this thesis do not hold for systems
-with physical nonholonomic constraints like friction, nor does it apply to vehicles
+constrained dynamics in simply actuated coordinates, and the results of this
+thesis may not hold.
+In particular, the theoretical guarantees of this thesis do not apply to systems
+with physical nonholonomic constraints (\eg friction), nor do they apply to vehicles
 with wheels.
 
-This guides us to the following research directions. 
-First, one might consider relaxing the assumptions on the input and inertia
-matrices to widen the class of systems to which VNHCs can be applied.
+These limitations guide us toward the following research directions. 
+First, one might relax the assumptions on the input and inertia
+matrices, thereby widening the class of systems to which VNHCs can be applied.
 Second, one might forgo the assumption that \(q_a\) is solvable on the
@@
 coordinates than \((q_u,p_u)\).
-Third, one might consider abstract mechanical systems with dynamics
+Third, one might consider more general mechanical systems with dynamics given by
 \[
@@
 \]
-where \(Q_{nh}(q,p) \in \R^n\) are the equations for physical nonholonomic
+where \(Q_{nh}(q,p) \in \R^n\) are terms corresponding to physical nonholonomic
 constraints like friction.
@@
 Finally, the proof of Theorem \ref{thm:acrobot-energy-stabilization} 
-in Chapter \ref{sec:acrobot-proof} opens a door to the idea of VNHC generation. 
-One might search for suitable conditions on a VNHC whereby the Poincar\'{e}
+in Chapter \ref{sec:acrobot-proof} opens a door to the possibility of VNHC generation. 
+One might search for suitable conditions on VNHCs where the Poincar\'{e}
 sections for oscillations and rotations are increasing.
 Then, one can extend Otsason's ``virtual constraint
-generator" \cite{vcg} with the goal of generating regular VNHCs which satisfy
+generator" \cite{vcg} to generate regular VNHCs which satisfy
 these energy injection/dissipation conditions.
-In this way, one could automatically generate VNHCs which stabilize a
+In this way, one could automatically generate regular VNHCs which stabilize a
 desired energy level.

### 2020-12-19 latex/introduction.tex
Added a better intro, edited symbols and concusion

@@
 
-Consider a child on a standing swing.
-Naturally, they begin by pushing off the ground and proceed to alternate between
-standing and squatting. 
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=0.6\textwidth,angle=270]{images/kiiking.jpg}
+    \caption{Kiiking, which translates to ``swing", involves a person standing
+        and squatting on a giant swing until they are rotating around the bar.
+    Image taken from \cite{kiiking-img}.}
+    \label{fig:kiiking}
+\end{figure}
+
+The Estonian sport of ``Kiiking" (Figure \ref{fig:kiiking}) is the adult
+equivalent of a child playing on a standing swing. 
+Here, an athlete attempts to make a giant solid steel swing gain momentum.
+Naturally, the kiiker begins by pushing off the ground and proceeds to alternate
+between standing and squatting. 
 This allows them to swing higher and faster over time, until they have reached
-their express goal of going as high and fast as possible.
+their express goal of spinning at high speed around the bar.
 
-Now imagine that the child is a robot, and their creator is teaching them to
+Now imagine that the kiiker is a robot, and their creator is teaching them to
 behave like a real boy.
-If the roboticist had studied classical control theory, they would plot the leg
-motion of a human child as a trajectory over time and tell the robot to
+If the roboticist had studied classical control theory, they would plot 
+a human's squat depth as a trajectory over time and tell the robot to
 synchronize with this trajectory.
@@
 control approach is entirely unnatural. 
-Humans on swings do not have an internal stopwatch.
+After all, humans on swings do not have an internal stopwatch telling them when
+to squat.
 Rather, they adjust how much they bend their knees based on the current angle of
@@
 This position-velocity adjustment allows humans to correct for external
-disturbances (like strong winds or overly enthusiastic parents).
+disturbances (like strong winds or overly enthusiastic pushers). 
 Figure \ref{fig:swing-pos-vel} shows this usual adjustment:
@@
 disturbance may swap the order of squatting and standing to the point that
-the swing slows down rather than speeding up.
+the swing can never begin rotating.
 
@@
         bottom of the swing path, and stand after they pass it until they hit
-        the peak of their swing. Image taken and modified from
+        the peak of their swing. Figure modified from
     \cite{pumping_swing_standing_squatting}.}
@@
 position rather than time.
-Sadly, this will not work for the swing-bot: one needs to know their velocity to
-determine when they have reached the peak of their swing.
+Sadly, this will not work for the swinging robot since one needs to know their
+velocity to determine when they have reached the peak of their swing.
 In other words, the robot needs to move according to a function of both position
@@
 
-Throughout this thesis we will refer to two examples to build intuition.
-The first is the variable-length pendulum, which models the child on a swing.
-The second is the acrobot, which models a gymnast hanging on a horizontal bar.
-For each of these mechanical systems, we will develop virtual nonholonomic
-constraints which replicate human behaviours. 
+Our goal throughout this thesis is to develop rigorous proof that virtual
+nonholonomic constraints are capable of increasing a robot's momentum through a
+process known as \textit{energy injection}.
+We will use two benchmark robotic systems to build intuition and test
+our theory.
+The first system is the variable-length pendulum, which models a person
+on a swing.
+The second is the acrobot, an actuated double pendulum which models a gymnast
+hanging on a horizontal bar.
+For each of these examples, we will create virtual nonholonomic
+constraints which model human motion to make the robot rotate around a bar.
 Finally, we will rigorously prove that these constraints inject energy into
@@
 \section{Literature Review}
-Here we review some of the previous work on energy injection and virtual
+Let us review some of the previous work on energy injection and virtual
 nonholonomic constraints.
@@
 \cite{swingup_acrobot_energy,swingup_giant_acrobot}.
-This technique is useful, but it does not necessarily produce the structured
-human-like motion we wish to generate using virtual nonholonomic constraints.
+This technique is useful and widely applicable, but it does not
+produce the structured human-like motion we wish to generate using virtual
+nonholonomic constraints.
 
@@
 All these fields use different formulations of virtual nonholonomic constraints;
-\citet{hybrid_zero_dynamics_bipedal_nhvcs} have derived the constrained dynamics
-of virtual nonholonomic constraints for hybrid mechanical systems,
-but to the best of our knowledge no one has attempted to generalize the
-concept to abstract mechanical systems.
+\citet{hybrid_zero_dynamics_bipedal_nhvcs} have derived the so-called
+constrained dynamics for hybrid mechanical systems under virtual nonholonomic
+constraints, but to the best of our knowledge no one has attempted to generalize
+the concept to abstract mechanical systems.
 
-Many of our contributions in Chapter \ref{ch:vnhcs} extend
-concepts from the field of virtual \textit{holonomic} constraints. 
-Of note is the theoretical framework devised by Mohammadi, Maggiore, and
+Many of our contributions are extensions of concepts from the field of virtual
+holonomic constraints. 
+Of note is the framework devised by Mohammadi, Maggiore, and
 Consolini, among others

### 2020-12-19 latex/symbols.tex
Updated \mathbb{S}^1 to \Sone, made Gamma \simeq \SxR instead of eq.
Added a better intro, edited symbols and concusion

@@
    \midrule
-   \(\n\) & The index set \(\{1,\ldots,n\}\) of natural numbers up to \(n\). \\
    \(\R^n\) & Real numbers in \(n\) dimensions. \\
-   \(\Rt{T}\) & Real numbers modulo \(T > 0\), with \(\Rt{\infty} = \R\). \\
-   \(\mathbb{S}^1\) & The unit circle, equivalent to \(\Rt{2\pi}\). \\
+   \(\Rt{T}\) & Real numbers modulo \(T > 0\), with \(\Rt{\infty} := \R\). \\
+   \(\mathbb{S}^1\) & The unit circle, diffeomorphic to \(\Rt{2\pi}\). \\
    \(\mathcal{Q}\) & The configuration manifold of a system. \\
    \(C^r(X;Y)\) & The space of \(r\)-times continuously differentiable functions from \(X\) to \(Y\). \\
-         & If \(r = \infty\), the space of smooth functions from \(X\) to \(Y\). \\
    \(\R^{n \times m}\) & The space of real-valued matrices with \(n\) rows and \(m\) columns. \\
@@
    \(dF_v\) & Total differential (Jacobian) of \(F\), equivalent to \((\nabla_v F)\tpose\). \\
-   \(\Hess F\) & If \(F : \R^n \rightarrow \R\), the \(n \times n\) Hessian matrix of double derivatives of \(F\). \\
-               & If \(F : \R^n \rightarrow \R^k\), the block matrix \((\Hess F_1, \ldots, \Hess F_k) \in \R^{n \times nk}\). \\
-   \(\partial_v \partial_w F\) & Derivative matrix of \(F : \R^n \times \R^m \rightarrow \R\), with \((i,j)\) element \(\ppdiff{F}{v_i}{w_j}\). \\
+   \(\partial_w F\) & Partial differential of \(F\) with respect to \(w\), with \(i^\text{th}\) element \(\pdiff{F}{w_i}\). \\
    \(\delta_{i,j}\) & The Kronecker delta: 1 if \(i = j\) and 0 otherwise. \\
@@
    \(\norm{\cdot}\) & The euclidean norm on \(\R^n\). \\
+   \(\simeq\) & Equivalent under diffeomorphism. \\
    \bottomrule

### 2020-12-19 latex/abstract.tex
Added comments to various files and created a list of symbols page
Wrote the abstract and did another edit of the intro

@@
 
+%/========== Abstract ==========/%
 %% (At most 150 words for M.Sc. or 350 words for Ph.D.)
 \begin{abstract}
-	\textbf{TODO}: Fill in the abstract
+   This thesis develops a theoretical foundation for the development of virtual
+   nonholonomic constraints, which are relations between the generalized
+   coordinates and generalized momenta of a mechanical system that can be
+   enforced via feedback control.
+   The theory is applied towards energy injection for two standard underactuated
+   systems: the variable-length pendulum and the acrobot. 
+   Virtual nonholonomic constraints are designed for each system by examining
+   human motion, and energy injection properties of these constraints are proven
+   rigorously.
+   The acrobot constraint is tested on a real-world acrobot, demonstrating
+   highly effective energy-injection properties and robustness to a variety of
+   external disturbances.
 \end{abstract}
-% vim: set ts=3 sw=3 sts=0 et ffs=unix :
+
+%/========== /Abstract ==========/%
+% vim: set tw=80 ts=4 sw=4 sts=0 et ffs=unix :

### 2020-12-19 latex/introduction.tex
Wrote the abstract and did another edit of the intro

@@
 disturbances (like strong winds or overly enthusiastic pushers). 
-Figure \ref{fig:swing-pos-vel} shows this usual adjustment:
-a person will squat at the peak of the swing and stand as they reach the fastest
-point at the bottom \cite{pumping_swing_standing_squatting}.
+Figure \ref{fig:swing-pos-vel} shows this adjustment process:
+a person will squat when they are stopped at the peak of the swing arc, and
+stand when they reach the fastest point at the bottom
+\cite{pumping_swing_standing_squatting}.
 Even if the robot could perfectly track a time-based trajectory, an external
@@
 For each of these examples, we will create virtual nonholonomic
-constraints which model human motion to make the robot rotate around a bar.
-Finally, we will rigorously prove that these constraints inject energy into
-their respective robots.
+constraints which model human motion to increase the robot's momentum.
+Finally, we will rigorously prove that these constraints inject energy.
 
 \section{Literature Review}
-Let us review some of the previous work on energy injection and virtual
+Let us review some of the existing research on energy injection and virtual
 nonholonomic constraints.
@@
 control and energy shaping --- one views the robot and its inputs as energy
-transformations which are moulded to force the system to behave as desired.
+transformations which are moulded to enforce desired behaviour.
 Much of the theoretical work in this field has been performed by Ortega, van der
@@
 \cite{vnhc_human_robot_cooperation}.
-Virtual nonholonomic constraints have been of most notable use in the area of
-bipedal locomotion \cite{nhvc_incline_walking,output_nhvc_bipedal_conrol},
+Virtual nonholonomic constraints have been of most notable use in
+bipedal locomotion \cite{nhvc_incline_walking,output_nhvc_bipedal_control},
 where they have shown marked improvements in the robustness of walking gaits
-when compared to previous control technqiues \cite{nhvc_incline_walking}.
+when compared to previous control technqiues.
 They have also been used for error-reduction in time-delayed teleoperation
-\cite{vnhc_time_delay_teleop} and human-robot interaction
+\cite{vnhc_time_delay_teleop} and in the field of human-robot interaction
 \cite{psd_based_vnhc_redundant_manipulator,haptic_vnhc}.
-All these fields use different formulations of virtual nonholonomic constraints;
+All these research areas use different formulations of virtual nonholonomic constraints;
 \citet{hybrid_zero_dynamics_bipedal_nhvcs} have derived the so-called
-constrained dynamics for hybrid mechanical systems under virtual nonholonomic
+``constrained dynamics" for hybrid mechanical systems under virtual nonholonomic
 constraints, but to the best of our knowledge no one has attempted to generalize
@@
 
-Many of our contributions are extensions of concepts from the field of virtual
-holonomic constraints. 
+Many of the contributions in this thesis are extensions of concepts from the
+field of virtual holonomic constraints. 
 Of note is the framework devised by Mohammadi, Maggiore, and
@@
 and in Chapter \ref{ch:acrobot} we find a virtual nonholonomic constraint 
-which enables the acrobot to kick its legs like a gymnast until it is
-performing backflips on a horizontal bar.
+which enables the acrobot to kick its legs like a gymnast until it
+performs backflips on a horizontal bar.
 Finally, we show experimental results on a physical acrobot which confirm

### 2020-12-20 latex/introduction.tex
Changed some wording and the location of introduction images

@@
 \chapter{Introduction}
-
-\begin{figure}[t]
-    \centering
-    \includegraphics[width=0.6\textwidth,angle=270]{images/kiiking.jpg}
-    \caption{Kiiking, which translates to ``swing", involves a person standing
-        and squatting on a giant swing until they are rotating around the bar.
-    Image taken from \cite{kiiking-img}.}
-    \label{fig:kiiking}
-\end{figure}
 
@@
 
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.6\textwidth,angle=270]{images/kiiking.jpg}
+    \caption{Kiiking, which translates to ``swing", involves a person standing
+        and squatting on a giant swing until they are rotating around the bar.
+    Image taken from \cite{kiiking-img}.}
+    \label{fig:kiiking}
+\end{figure}
+
 Now imagine that the kiiker is a robot, and their creator is teaching them to
-behave like a real boy.
+behave like a real person.
 If the roboticist had studied classical control theory, they would plot 
@@
 synchronize with this trajectory.
-In an ideal world, this technique would work perfectly; unfortunately, this
-control approach is entirely unnatural. 
+In an ideal world, this technique would work perfectly; unfortunately, it is
+also entirely unnatural. 
 After all, humans on swings do not have an internal stopwatch telling them when
@@
 Figure \ref{fig:swing-pos-vel} shows this adjustment process:
-a person will squat when they are stopped at the peak of the swing arc, and
+a person will squat when they are stopped at the peak of their swing, and
 stand when they reach the fastest point at the bottom
@@
 One recent control technique known as \textit{virtual holonomic constraints}
-\cite{vhcs_for_el_systems} forces a robot's actuators to track a function of
-position rather than time.
-Sadly, this will not work for the swinging robot since one needs to know their
+forces a robot's actuators to track a function of
+position rather than time \cite{vhcs_for_el_systems}.
+Sadly, this will not help the swinging robot since one needs
 velocity to determine when they have reached the peak of their swing.
@@
 For each of these examples, we will create virtual nonholonomic
-constraints which model human motion to increase the robot's momentum.
-Finally, we will rigorously prove that these constraints inject energy.
+constraints which model human motion and will rigorously prove that these
+constraints inject energy.
 

### 2020-12-20 matlab/thesis_sims/energyEvent.m
Saving all the new matlab files that I've worked on over the past few months, for experiments and simulations. We also update the sugarbot and sugarbox arduino code to use the arctan function, and update the comments.

@@ -1,0 +1,26 @@
+%% Energy Level Events
+% Triggers an event in ode45 when the inputted energy E reaches a specific
+% value.
+% You use this in ode45 by setting the events function handle to
+% @(t,qp) energyEvent(t,qp,E,Eval).
+%
+% Inputs:
+% * t = current time value of simulation
+% * qp = current value of the phase on the (qu,pu)-plane
+% * E = function handle to the energy of the system
+% * Eval = energy value at which to trigger the event. This will typically
+% be E(pi,0), but it can be any energy value.
+
+function [value,isterminal,direction] = energyEvent(t,qp,E,Eval)
+    % Trigger when E(q,p) = Eval. The value increases as E -> Eval from
+    % below.
+    value = E(qp(1),qp(2))-Eval;
+    % Don't stop the integration, just record the event
+    isterminal = 0;
+    % Track only values where the energy is increasing past Eval, not those
+    % where you come back down. If you want it to track only decreasing,
+    % set direction to -1. If you want both increasing and decreasing, set
+    % it to 0.
+    direction = 1;
+end
+

### 2020-12-20 matlab/thesis_sims/vlp_l_optimal_gain.m
Added the simulations images for the VLP

@@ -9,7 +9,7 @@
 %
 % Output:
 % * l^\star(theta), the time-optimal controller.
-function l = vlp_l_optimal(theta,lu,lb)
+function l = vlp_l_optimal_gain(theta,lu,lb)
 % If theta is within ]-pi,-pi/2] or [0,pi/2], we stand.
 if theta <= -pi/2 || (0 <= theta && theta <= pi/2)
     l = lu;

### 2020-12-20 latex/vlp.tex
Finished the VLP simulations version 1

@@
 \section{Simulation Results}
-% TODO: Show the simulations of the VLP VNHCs (multiple of them). How does the
-% energy injection time of our sin(2theta) VNHC compare to the energy injection
-% time of the optimal one? How does the VNHC version compare in robustness with
-% the time-based one from the pumping paper?
+In this section, we compare the energy injection properties of our smooth VNHC
+\(L_\frac{\pi}{2}(\theta)\), the time-optimal VNHC \(L^\star(\theta)\) from
+Equation \eqref{eqn:vlp-optimal-controller}, and the differentiable
+VNHC \(L_T(\theta)\) from Equation \eqref{eqn:vlp-T-controller} with 
+\(T = 0.01\).
+To do this comparison, we simulate an adult Estonian performing a Kiiking
+routine.
+
+The average weight of a 25 year old Estonian man is \(80.17\)kg
+\cite{estonian-body-data}, so we set the VLP mass \(m\) to this value.
+To find our upper and lower bounds on length, we first need to choose the size
+of a swing.
+The common model of a Kiiking swing is between \(3-4\)m tall
+\cite{kiiking-sale}, so we take the length to be \(3\)m.
+The average 25 year old Estonian is \(1.8\)m tall \cite{estonian-body-data}, and
+we assume their squat height is half that at \(0.9\)m tall.
+We further assume that their center of mass (COM) is always located at half
+their current height, which means their standing COM is \(0.9\)m from the ground
+while their squatting COM is \(0.45\)m from the ground.
+This means the distance from their standing COM to the pivot is \(2.1\)m, while
+the distance from their squatting COM to the pivot is \(2.55\)m.
+Hence, we set \(\underbar{l} = 2.1\) and \(\overbar{l} = 2.55\).
+We initialize a VLP with these parameters at 
+\((q,p) = \left(\frac{\pi}{32},0\right)\) and plot the resulting orbits
+in Figure \ref{fig:vlp-sim-orbit}.
+
+\begin{figure}
+   \centering
+   \includegraphics[width=0.8\textwidth]{images/kiiking_orbits.png}
+   \caption{The orbits of the time-optimal controller \(L^\star(\theta)\)
+      (black), the differentiable controller \(L_T(\theta)\) (red),
+      and the smooth VNHC \(L_\frac{\pi}{2}(\theta)\) (cyan) employed by a 25
+      year old Estonian male on a 3m tall Kiiking swing. All the controllers are
+      initialized at \(q(0) = \frac{\pi}{32}\), and they each steadily gain
+      energy until they reach a rotation.}
+   \label{fig:vlp-sim-orbit}
+\end{figure}
+
+The time-optimal controller reaches a rotation within 18 seconds, the
+differentiable controller rotates in 41 seconds, and
+the smooth VNHC takes 4691 seconds (78 minutes).
+Despite the fact that the smooth VNHC is several orders of magnitude slower than
+the other controllers, it still succeeds at injecting energy into the
+system.
+This is especially clear when looking at Figure \ref{fig:vlp-sim-energy}, which
+shows the smooth VNHC steadily gaining energy over time.
+
+\begin{figure}
+   \centering
+   \includegraphics[width=0.8\textwidth]{images/E_kiiking_ltheta.png}
+   \caption{The energy \(E_\text{avg}(q,p)\) of a 25 year old Estonian male on a
+      3m tall Kiiking swing who employs the smooth VNHC
+   \(L_\frac{\pi}{2}(\theta)\).}
+   \label{fig:vlp-sim-energy}
+\end{figure}
+
+Overall, this simulation confirms that VNHCs are capable of injecting energy
+into the VLP. 
+Moreover, if one can use the differentiable controller \(L_T(\theta)\) (which
+humans certainly can), one is able to approximate the time-optimal controller
+arbitrarily well.
+We conclude that VNHCs are powerful tools for injecting energy into the VLP
+while maintaining the natural structure of human motion.
 

### 2020-12-20 latex/introduction.tex
Edits to abstract and intro

@@
 The Estonian sport of ``Kiiking" (Figure \ref{fig:kiiking}) is the adult
-equivalent of a child playing on a standing swing. 
-Here, an athlete attempts to make a giant solid steel swing gain momentum.
+equivalent of a child playing on a standing swing, though here
+an athlete attempts to make a giant solid steel swing gain momentum.
 Naturally, the kiiker begins by pushing off the ground and proceeds to alternate
@@
 Now imagine that the kiiker is a robot, and their creator is teaching them to
-behave like a real person.
+swing like a human.
 If the roboticist had studied classical control theory, they would plot 
-a human's squat depth as a trajectory over time and tell the robot to
+a kiiker's squat depth as a trajectory over time and tell the robot to
 synchronize with this trajectory.
@@
 This position-velocity adjustment allows humans to correct for external
-disturbances (like strong winds or overly enthusiastic pushers). 
+disturbances (like strong winds or enthusiastic swing pushers). 
 Figure \ref{fig:swing-pos-vel} shows this adjustment process:
-a person will squat when they are stopped at the peak of their swing, and
+a person will squat when they reach the peak of their swing, and
 stand when they reach the fastest point at the bottom
@@
 Even if the robot could perfectly track a time-based trajectory, an external
-disturbance may swap the order of squatting and standing to the point that
-the swing can never begin rotating.
+disturbance may affect where squatting and standing occurs to the point that
+the swing actually slows down.
 
@@
     \includegraphics[width=0.75\textwidth]{images/swing_pos_vel.png}
-    \caption{A person on a standing swing will squat as they approach the
-        bottom of the swing path, and stand after they pass it until they hit
-        the peak of their swing. Figure modified from
+    \caption{A person on a standing swing will stand at the bottom of the swing
+        arc, and squat at the top. Figure modified from
         \cite{pumping_swing_standing_squatting}.}
@@
 
-One recent control technique known as \textit{virtual holonomic constraints}
-forces a robot's actuators to track a function of
-position rather than time \cite{vhcs_for_el_systems}.
-Sadly, this will not help the swinging robot since one needs
-velocity to determine when they have reached the peak of their swing.
+To fix this issue, a clever roboticist might use \textit{virtual holonomic
+constraints}, which are controllers that track a function of position rather
+than time \cite{vhcs_for_el_systems}.
+Sadly, this still fails since one needs velocity to determine when they
+have reached the peak of their swing.
 In other words, the robot needs to move according to a function of both position
@@
 
-Our goal throughout this thesis is to develop rigorous proof that virtual
-nonholonomic constraints are capable of increasing a robot's momentum through a
+Our goal throughout this thesis is to prove that virtual
+nonholonomic constraints can increase a robot's momentum through a
 process known as \textit{energy injection}.

### 2020-12-21 matlab/thesis_sims/vlp_l_T.m
Adding the acrobot figures as simulated by acrobot_sims.mlx

@@ -17,7 +17,7 @@
 function l = vlp_l_T(theta,lu,lb,T)
 % Return vlp_l_optimal if T = 0
 if T == 0
-    l = vlp_l_optimal(theta,lu,lb);
+    l = vlp_l_optimal_gain(theta,lu,lb);
     return;
 elseif T == pi/2
      l = vlp_l_smooth(theta,lu,lb);
@@ -45,7 +45,7 @@ elseif pi/2 - T/2 < theta && theta < pi/2 + T/2
 % If it's none of these cases, then we are within the domain of the optimal
 % controller, so we return the optimal result.
 else
-    l = vlp_l_optimal(theta,lu,lb);
+    l = vlp_l_optimal_gain(theta,lu,lb);
 end
 end
 

### 2020-12-21 latex/abstract.tex

@@
 \begin{abstract}
-   This thesis develops a theoretical foundation for the development of virtual
+   This thesis develops a theoretical foundation for the framework of virtual
    nonholonomic constraints, which are relations between the generalized

### 2020-12-21 latex/acrobot.tex
Finished the acrobot experiments section

@@
 \section{The Acrobot Constraint}
-One may be tempted to design a constraint of the form \(q_a = \bar{q}_a\sin(\theta)\) 
+One may be tempted to design a constraint of the form 
+\(q_a = \bar{q}_a\sin(\theta)\),
 with \(\theta := \arctan_2(p_u,q_u)\), since a similar approach 
@@
 By the same arguments presented in this section, the Poincar\'{e} section
-satisfies \( P_\mathcal{O}(\mu_0) \leq \mu_0 - \gamma\) when \(I < 0\). 
+satisfies
+\[
+    P_\mathcal{O}(\mu_0) \leq \mu_0 - \gamma
+    ,
+\] 
+when \(I < 0\). 
 In this case, the constraint dissipates energy on \(\mathcal{O}_1\).
@@
 Since we know that \(\Pi^+\) is a set of measure zero in \(\SxR\), almost every
-other orbit initialized in \(\mathcal{O}_1\) must escape through the boundary
+orbit initialized in \(\mathcal{O}_1\) must escape through the boundary
 \(E_\pi\) into the rotation domain.
+
+Escaping once into the rotation domain does not guarantee the acrobot
+orbit will not return to \(\mathcal{O}_1\).
+Indeed, the orbit could escape through \(E_\pi\) and return to \(\mathcal{O}_1\)
+several times.
+However, the orbit must eventually remain outside the closure of 
+\(\mathcal{O}_1\). 
+If not, the positive limit set would once again be
+\(E_\pi\); by the same argument as above, the orbit must converge to 
+\((\pi,0)\).
+
+We conclude that all orbits beginning in \(\mathcal{O}_1\) will, in finite time,
+escape the closure of \(\mathcal{O}_1\) and remain in the rotation domain
+forever after.
 
@@
 We define this Poincar\'{e} map as
-\(P_\mathcal{R}(\rho_0) := \hat{\rho}(\sign{\rho_0}2\pi,\rho_0,I)\),
+\[
+    P_\mathcal{R}(\rho_0) := \hat{\rho}\left(\sign{\rho_0}2\pi,\rho_0,I\right)
+    ,
+\]
 which expands into
@@
 By the same arguments presented in this section, the Poincar\'{e} section
-satisfies \(\abs{P_\mathcal{R}(\rho_0)} \leq \abs{\rho_0} - \gamma\) 
+satisfies 
+\[
+    \abs{P_\mathcal{R}(\rho_0)} \leq \abs{\rho_0} - \gamma
+    ,
+\]
 when \(I < 0\).
@@
 \section{Experimental Results}
-% TODO: This might be a separate chapter
+In this section, we test our VNHC on the physical acrobot built by
+\citet{xingbo_thesis} (Figure \ref{fig:xingbo-acrobot}).
+Since it is not a simple acrobot, this experiment will test whether VNHCs are
+robust to model uncertainties.
+As we will see, our VNHC can inject energy on the general acrobot model.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\linewidth]{images/xingbo_acrobot.png}
+    \caption{The acrobot built by Wang \cite{xingbo_thesis}.}
+    \label{fig:xingbo-acrobot}
+\end{figure}
+
+Wang's acrobot  has inertia matrix as in ~\eqref{eqn:general-acrobot-inertia}
+and potential function as in ~\eqref{eqn:general-acrobot-potential}.
+The physical parameters are written explicitly in Table
+\ref{tab:acrobot-parameters}.
+
+\begin{table}
+    \centering
+    \caption{Physical parameters for the real acrobot, as measured by
+    \citet{xingbo_thesis}.}
+    \label{tab:acrobot-parameters}
+    \begin{tabular}{ccccccccc}
+        \toprule
+        $m_t$ & $m_l$ & $d_t$ & $d_l$ & $l_t$ & $l_l$ & $J_t$ & $J_l$ & $g$ \\
+        \midrule
+        0.2112 & 0.1979 & % mass
+        0.148 & 0.145 & % total length
+        0.073 & 0.083 & % COM distance
+        0.00129 & 0.00075 & % moment of inertia
+        9.81 \\ % gravity
+        \bottomrule
+    \end{tabular}
+\end{table}
+
+This acrobot's actuator is a servo motor with a built-in PID controller, so we
+can set \(q_a = \arctan\left(I p_u\right)\) directly without needing to solve
+for the VNHC's torque input \(\tau\).
+The control parameter \(I\) needs to be small, but it also needs to be large
+enough to overcome friction in the servo motor.
+Since the acrobot's mass and length values are small, its momentum is
+also small.
+Through various experiments, we determined that \(I = 10\) is sufficient to
+ensure the actuator can fully rotate in its allowable range
+\(q_a \in \left[ -\frac{\pi}{2}, \frac{\pi}{2}\right]\).
+
+\subsection{Simulations}
+
+To compute the energy of the nominal pendulum, we evaluate the system at the
+VNHC \(q_a = 0\), which yields
+\[
+    E(q_u,p_u) \approx 396.5501 p_u^2 + 0.5997(1 - \cos(q_u))
+    .
+\]
+Hence, the level set of \(E_\pi\) of the energy \(E(\pi,0)\) is parameterized by
+\begin{align*}
+    p_u &\approx \pm \sqrt{\frac{1.1994}{793.1001}(2 - (1 - \cos(q_u))}
+    ,
+\end{align*}
+for each \(q_u \in ]-\pi,\pi]\).
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=0.8\linewidth]{images/acrobot_orbit.png}
+    \caption{A simulation of the acrobot from \cite{xingbo_thesis}.}
+    \label{fig:acrobot-orbit}
+\end{figure}
+
+We initialize the simulation for this acrobot at 
+\((q_u,p_u) = \left(\frac{\pi}{32},0 \right)\) and plot the resulting orbit in 
+Figure \ref{fig:acrobot-orbit}.
+The nominal energy level set \(E_\pi\) is outlined in black. 
+Because this acrobot is not simple, \(E_\pi\) is not the boundary between
+rotations and oscillations.
+The oscillation domain is actually much larger: orbits rotate once they hit the
+\(p_u\)-axis at \(\abs{p_u} \approx 0.15\), while \(E_\pi\) intersects the
+\(p_u\)-axis at \(\abs{p_u} \approx 0.055\).
+Despite this, we see that eventually the acrobot remains outside \(E_\pi\).
+The points where the orbit exits \(E_\pi\) are highlighted with black stars,
+with the final departure highlighted in red.
+At this final departure, the acrobot begins rotating and continues to gain
+energy over time.
+
+To verify that the acrobot will consistently begin rotating, we use a Monte
+Carlo method \cite{montecarlo} to simulate this
+acrobot \(1000\) times.
+At each iteration, we initialize the acrobot randomly inside the sublevel set
+\[
+    \left\{(q_u,p_u) \in \SxR \mid
+    E(q_u,p_u) \leq E\left(\frac{\pi}{32},0\right)\right\}
+    ,
+\] 
+and measure how long it takes the acrobot to begin rotating.
+The results are plotted in Figure \ref{fig:acrobot-mc}.
+The acrobot always rotates within 10--35 seconds.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\textwidth]{images/acrobot_mc.png}
+    \caption{Monte Carlo simulation results for the acrobot from
+    \cite{xingbo_thesis}.}
+    \label{fig:acrobot-mc}
+\end{figure}
+
+\subsection{Physical Experiments}
+To complete this chapter, we perform a four experiments with the physical
+acrobot:
+\begin{enumerate}
+    \item An undisturbed test to validate that the acrobot reaches rotations.
+    \item A test where we stop the acrobot once it has started rotating.
+    \item A test where we push the acrobot in its direction of motion after it has
+        started rotating.
+    \item A test where we push the acrobot against its direction of motion.
+\end{enumerate}
+
+The acrobot has an encoder at the pivot which measures \(q_u\) and
+\(\dot{q}_u\), and the servo provides a measurement of \(q_a\).
+We estimate \(\dot{q}_a\) through sequential values of \(q_a\) and
+compute \(p_u\) through \(p_u = e_1^T M(q) \dot{q}\).
+The actuator value at iteration \(k \in \mathbb{Z}_{> 0}\) is assigned through
+the equation
+\[
+    q_a^{k} = \arctan(I p_u^{k-1})
+    .
+\]
+
+\subsubsection*{Test 1: No Disturbances}
+For this test we initialize the acrobot at 
+\((q_u,p_u) \approx \left(\frac{\pi}{8},0\right)\). 
+The resulting orbit is shown in Figure \ref{fig:acrobot-unperturbed-orbit},
+which is clearly gaining energy over time.
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=0.5\linewidth]{images/acrobot_unperturbed_orbit.png}
+    \caption{The orbit of the physical acrobot during an unperturbed test.}
+    \label{fig:acrobot-unperturbed-orbit}
+\end{figure}
+
+\subsubsection*{Test 2: Stopping the Acrobot}
+For this test we initialize the acrobot at 
+\((q_u,p_u) \approx \left(\pi,0\right)\) and let it run for 15
+seconds, then stop the acrobot as it reaches the bottom of its arc.
+The resulting orbit is shown in Figure \ref{fig:acrobot-stopped-orbit}.
+The blue curves correspond to the orbit before the disturbance, while the red
+spiral shows that the acrobot continues to oscillate after it is stopped.
+Despite the disturbance, it gains energy and eventually reaches a rotation
+again.
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=0.5\linewidth]{images/acrobot_stopped_orbit.png}
+    \caption{The orbit of the physical acrobot before (blue) and after (red) it
+    is stopped.}
+    \label{fig:acrobot-stopped-orbit}
+\end{figure}
+
+\subsubsection*{Test 3: Pushing the Acrobot Forwards}
+To see how the acrobot responds when pushed in its direction of motion, we allow
+the acrobot to rotate freely for 15 seconds and then give it a push in its
+direction of motion.
+The orbit in Figure \ref{fig:acrobot-fpush-orbit} shows that the acrobot speeds
+up to rotate with energy \(E(0,0.22)\), but then slows down until it reaches a
+stable rotation with energy \(E(0,-0.195)\).
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=0.5\textwidth]{images/acrobot_fpush_orbit.png}
+    \caption{The orbit of the physical acrobot before (blue) and after (red) it
+    is pushed in its current direction of motion.}
+    \label{fig:acrobot-fpush-orbit}
+\end{figure}
+
+\subsubsection*{Test 4: Pushing the Acrobot Backwards}
+In this final experiment, we test whether the acrobot can easily change
+directions when pushed against its current direction of motion.
+We again allow the acrobot to rotate freely for 15 seconds, then push it the
+opposite way.
+The orbit in Figure \ref{fig:acrobot-rpush-orbit} demonstrates that the acrobot
+responds by readily changing direction, and quickly achieves its maximum speed
+with energy \(E(0,0.195)\).
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=0.5\textwidth]{images/acrobot_rpush_orbit.png}
+    \caption{The orbit of the physical acrobot before (blue) and after (red) it
+    is pushed against its current direction of motion.}
+    \label{fig:acrobot-rpush-orbit}
+\end{figure}
+
+\subsection{Summary of Experimental Results}
+The simulations and experiments in this section show that VNHCs are excellent
+tools for injecting energy even in non-simple acrobots. 
+Furthermore, the experimental results show that the energy gain is robust
+against a variety of disturbances.
+Finally, the two push tests suggest that Wang's acrobot constrained by our VNHC
+will gain energy on \(\mathcal{O}_2(0.195)\).
 

### 2020-12-21 latex/conclusion.tex
Fixed overfull hbox errors

@@
 acrobot.
+At the end of each chapter we performed simulations and experiments which
+validated the theory and showed the utility of VNHCs. 
 In the end, we demonstrated that virtual nonholonomic constraints are capable of
@@
 
-These limitations guide us toward the following research directions. 
+These limitations guide us to the following research directions. 
 First, one might relax the assumptions on the input and inertia
@@
 constraints like friction.
-Fourth, one can investigate a smoother mechanism for energy stabilization, where
+Fourth, one can investigate a smoother method for energy stabilization, where
 a VNHC is explicitly designed to inject energy below some energy level, and
 dissipate energy above it.
-For example, this might involve a mechanism for smoothly transferring between 
-VNHCs while maintaining certain safety requirements.
+For example, one might devise a mechanism for smoothly transferring
+between VNHCs while maintaining certain safety requirements.
 

### 2020-12-21 latex/vlp.tex
Fixed overfull hbox errors

@@
 The action of regulating pendulum length to inject energy into the VLP is known as
-``pumping". \citet{pumping_swing_standing_squatting} asked whether
-the pumping strategy performed by children is time-optimal, assuming the
-children could squat or stand instantaneously. 
+``pumping". 
+\citet{pumping_swing_standing_squatting} asked whether the pumping strategy
+performed by children is time-optimal, assuming they can squat or
+stand instantaneously.
 Indeed, they discovered that children increase the height of their swing as fast
@@
    \caption{The time-optimal controller for a standing swing as derived by
-      \cite{pumping_swing_standing_squatting}. The colour red
-      corresponds to standing, blue to squatting, and 
+      \cite{pumping_swing_standing_squatting}. 
+      Red corresponds to standing, blue to squatting, and 
       \(\theta := \arctan_2(p,q)\) is the angle of the VLP phase in the

### 2020-12-22 latex/acknowledgements.tex
Create a basic skeleton
This commit creates the basic skeleton for the preliminary pages -
abstract, dedication, acknowledgements, title, toc, as well as the
bibliography.

Todo is still all the chapters
Added comments to various files and created a list of symbols page
First version of Manfredi's acknowledgement, not done yet

--- 2020-06-04 latex/abstract.tex
@@
 
-%% (At most 150 words for M.Sc. or 350 words for Ph.D.)
-\begin{abstract}
-	\textbf{TODO}: Fill in the abstract
-\end{abstract}
-% vim: set ts=3 sw=3 sts=0 et ffs=unix :
+%/========== Acknowledgements ==========/%
+
+\begin{acknowledgements}
+    Manfredi, you are \textit{without a doubt} the greatest supervisor, the
+    best mentor, and one of the most wonderful people in the world.
+    Knowing you, you'll likely respond to that by saying 
+    ``thank you Adan, you're too kind", and then proceed to laugh uncomfortably before
+    changing the topic.
+    But it's true, and I don't know how else to tell you how lucky and grateful I am to
+    have you in my life.
+
+    These two years of doing research together were the most fun I've ever
+    had; yet they've also been fraught with death, sorrow, and the most
+    devastating heartbreak I've ever experienced.
+    Through it all you've been nothing but kind, honest, generous,
+    compassionate, understanding, and so incredibly humble.
+    You work unbelievably hard in the background to ensure your students succeed, and you
+    ask for so little in return.
+    My name is on the title page of this thesis, but if not for the required
+    academic formality I would change the authorship to read
+    ``Adan Moran-MacDonald, who couldn't have done any of this without the love
+    and support of his supervisor and dear friend, Manfredi Maggiore."
+
+    Manfredi: you are my hero.
+
+    % TODO: do the rest of the acknowledgements: parents, SCG friends, other
+    % friends, committee
+
+\end{acknowledgements}
+
+%/========== /Acknowledgements ==========/%
+% vim: set tw=80 ts=4 sw=4 sts=0 et ffs=unix :

### 2020-12-22 latex/acknowledgements.tex
Final edit, she be complete!

@@
 
-    These two years of doing research together were the most fun I've ever
-    had; yet they've also been fraught with death, sorrow, and the most
+    These two years of research together were the most fun I've ever
+    had; they were also fraught with death, sorrow, and the most
     devastating heartbreak I've ever experienced.
@@
     compassionate, understanding, and so incredibly humble.
-    You work unbelievably hard in the background to ensure your students succeed, and you
-    ask for so little in return.
+    You work tirelessly in the background to ensure your students
+    succeed, yet ask for so little in return.
     My name is on the title page of this thesis, but if not for the required
@@
 
-    Manfredi: you are my hero.
+    Manfredi: you're my hero.
 
-    % TODO: do the rest of the acknowledgements: parents, SCG friends, other
-    % friends, committee
+    \bigskip % big line break to separate
+
+    My gratitude also goes to Lacra Pavel, Raymond Kwong, and Stark Draper for
+    being on my committee. I hope you enjoy reading this thesis. Thank you for
+    your hard work.
+
+    Dear friends in the Systems Control Group,
+    you remind me what it means to have a zest for life. 
+    Every moment with you is wild, crazy, never-ending fun. Thank you
+    for the memes, deep chats, pub games, and bean juice; the distractions
+    always came when I needed them most.
+
+    Oh family of mine, your unconditional support has shaped every word I've
+    written in this thesis, and you deserve more appreciation than I can put
+    into words. Thank you for driving me to be myself. I love you all.
+
+    Finally, to Mallory, Leela, and Fry: our life together may have ended, but
+    you filled my heart and helped me become who I am today. 
+    Thank you for the happiest years of my life.
 

### 2020-12-22 latex/acrobot.tex
Final edit, she be complete!

@@
      &\begin{cases}
-        \dot{q} = \Minv(q) p \\
-        \dot{p}_u = -mgl\left(2s_u + s_{ua}\right) \\
+        \dot{q} = \Minv(q) p 
+        ,\\
+        \dot{p}_u = -mgl\left(2s_u + s_{ua}\right) 
+        ,\\
         \dot{p}_a =-\frac{1}{2}p\tpose \nabla_{q_a}\Minv(q) p
@@
 \]
-Equations ~\eqref{eqn:acrobot-mu-dot-nom}-\eqref{eqn:acrobot-alpha-dot-nom}
+Equations ~\eqref{eqn:acrobot-alpha-dot-nom}-\eqref{eqn:acrobot-mu-dot-nom}
 reveal that \(g(\alpha,\mu,0) = 0\),
@@
 By the same arguments presented in this section, the Poincar\'{e} section
-satisfies
-\[
-    P_\mathcal{O}(\mu_0) \leq \mu_0 - \gamma
-    ,
-\] 
-when \(I < 0\). 
+satisfies \(P_\mathcal{O}(\mu_0) \leq \mu_0 - \gamma\) when \(I < 0\). 
 In this case, the constraint dissipates energy on \(\mathcal{O}_1\).
@@
 An orbit of the acrobot will only begin rotating if it escapes
-\(\bar{\mathcal{O}}_1\) by crossing through \(E_\pi\).
+\(\bar{\mathcal{O}}_1\) by crossing through \(E_\pi\), and eventually remains
+outside this set.
 It is possible for orbits starting in \(\mathcal{O}_1\) to always approach
@@
 Escaping once into the rotation domain does not guarantee the acrobot
-orbit will not return to \(\mathcal{O}_1\).
+orbit will never return to \(\mathcal{O}_1\).
 Indeed, the orbit could escape through \(E_\pi\) and return to \(\mathcal{O}_1\)
 several times.
-However, the orbit must eventually remain outside the closure of 
+However, almost every orbit must eventually remain outside the closure of 
 \(\mathcal{O}_1\). 
 If not, the positive limit set would once again be
-\(E_\pi\); by the same argument as above, the orbit must converge to 
-\((\pi,0)\).
+\(E_\pi\); by the same argument as above, that orbit would be in \(\Pi^+\).
 
@@
 \begin{equation}\label{eqn:rotation-pu2}
-    \frac{p_u^2}{10m l^2} + 30mgl(1 - c_u) = \frac{\rho^2}{10 ml^2}
+    \frac{p_u^2}{10m l^2} + 3mgl(1 - c_u) = \frac{\rho^2}{10 ml^2}
     .
@@
 MATLAB evaluates those dynamics as
-\begin{align}\label{eqn:acrobot-rot-mu-dot-nom}
-    \dot{\rho} &= 0
-    , \\
+\begin{align}
     \label{eqn:acrobot-rot-alpha-dot-nom}
@@
     \frac{\sqrt{\rho^2 - 30m^2gl^3(1 - c_\beta)}}{5ml^2}
+    , \\
+    \label{eqn:acrobot-rot-mu-dot-nom}
+    \dot{\rho} &= 0
     .
@@
 Recall from Section \ref{sec:acrobot-proof-o1} that any acrobot initialized in
-\(\mathcal{O}_1\) will eventually enter the rotation region \(\mathcal{R}\).
+\(\mathcal{O}_1\) will eventually remain in the rotation region \(\mathcal{R}\).
 What remains, then, is to prove that the acrobot's orbits will also gain
@@
 \[
-    D := \left[-\bar{\rho},\sqrt{60m^2gl^3}\right[ \, 
+    D := \left[-\bar{\rho},-\sqrt{60m^2gl^3}\right[ \, 
     \cup \, 
@@
 \[
-    D^- :=\left[-\bar{\rho},\sqrt{60m^2gl^3}\right[
+    D^- :=\left[-\bar{\rho},-\sqrt{60m^2gl^3}\right[
     ,
@@
 \(p_u\)-axis when an orbit of the constrained dynamics intersects the
-\(p_u\)-axis for the \(n\)th time, assuming the orbit was initialized at
+\(p_u\)-axis for the \(n^\text{th}\) time, assuming the orbit was initialized at
 \((q_u,p_u) = (0,\rho_0)\).
@@
 makes \(P_\mathcal{R}(\rho_0) \geq \rho_0 + \gamma\) for all 
-\(\rho_0 \in D^+\)
+\(\rho_0 \in D^+\).
 
@@
 \begin{align*}
-    P_\mathcal{R}(\rho_0)
-    &= \rho_0 - IS(\rho_0) + R(-2\pi,\rho_0,I)
-    , \\
-    &\leq \rho_0 - I\epsilon + I^2r^\prime
+    P_\mathcal{R}(\rho_0) < \rho_0 - I\epsilon + I^2r^\prime
     .
@@
 satisfies 
-\[
-    \abs{P_\mathcal{R}(\rho_0)} \leq \abs{\rho_0} - \gamma
-    ,
-\]
+\(\abs{P_\mathcal{R}(\rho_0)} \leq \abs{\rho_0} - \gamma\)
 when \(I < 0\).
@@
 Since it is not a simple acrobot, this experiment will test whether VNHCs are
-robust to model uncertainties.
-As we will see, our VNHC can inject energy on the general acrobot model.
+robust to model mismatch.
+As we will see, our VNHC does inject energy on the general acrobot model.
 
@@
         \toprule
-        $m_t$ & $m_l$ & $d_t$ & $d_l$ & $l_t$ & $l_l$ & $J_t$ & $J_l$ & $g$ \\
+        $m_u$ & $m_a$ & $l_u$ & $l_a$ & $l_{c_u}$ & $l_{c_a}$ & $J_u$ & $J_a$ & $g$ \\
         \midrule
@@
 \begin{align*}
-    p_u &\approx \pm \sqrt{\frac{1.1994}{793.1001}(2 - (1 - \cos(q_u))}
+    p_u &\approx \pm \sqrt{\frac{1.1994}{793.1001}(1 + \cos(q_u))}
     ,
@@
 \subsection{Physical Experiments}
-To complete this chapter, we perform a four experiments with the physical
+To complete this chapter, we perform four experiments with the physical
 acrobot:
@@
 The blue curves correspond to the orbit before the disturbance, while the red
-spiral shows that the acrobot continues to oscillate after it is stopped.
-Despite the disturbance, it gains energy and eventually reaches a rotation
+spiral shows that the acrobot begins oscillating after it is stopped.
+Despite the disturbance, it gains energy and eventually starts rotating
 again.
@@
 To see how the acrobot responds when pushed in its direction of motion, we allow
-the acrobot to rotate freely for 15 seconds and then give it a push in its
+the acrobot to rotate undisturbed for 15 seconds and then give it a push in its
 direction of motion.
@@
 directions when pushed against its current direction of motion.
-We again allow the acrobot to rotate freely for 15 seconds, then push it the
-opposite way.
+We again allow the acrobot to rotate undisturbed for 15 seconds, then push it
+the opposite way.
 The orbit in Figure \ref{fig:acrobot-rpush-orbit} demonstrates that the acrobot

### 2020-12-22 latex/conclusion.tex
Final edit, she be complete!

@@
 At the end of each chapter we performed simulations and experiments which
-validated the theory and showed the utility of VNHCs. 
+validated the theory, and which showed the efficacy and robustness of VNHC-driven motion. 
 In the end, we demonstrated that virtual nonholonomic constraints are capable of
@@
 \begin{enumerate}
+    \item The Hamiltonian of the system is of the form 
+        \[
+            H(q,p) = \frac{1}{2}p\tpose \Minv(q) p + V(q)
+            .
+        \]
     \item The input matrix \(B(q) \equiv B \in \R^{n \times k}\) is constant and

### 2020-12-22 latex/introduction.tex
Fixed overfull hbox errors
Final edit, she be complete!

@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
     \centering
@@
 All these research areas use different formulations of virtual nonholonomic constraints;
-\citet{hybrid_zero_dynamics_bipedal_nhvcs} have derived the so-called
-``constrained dynamics" for hybrid mechanical systems under virtual nonholonomic
+\citet{hybrid_zero_dynamics_bipedal_nhvcs} have derived the 
+zero dynamics for hybrid mechanical systems under virtual nonholonomic
 constraints, but to the best of our knowledge no one has attempted to generalize
@@
 \begin{itemize}[label={}]
-   \item \textbf{Chapter \ref{ch:vnhcs}} The development of the framework of
+   \item \textbf{Chapter \ref{ch:vnhcs}} The development of 
       virtual nonholonomic constraints.

### 2020-12-22 latex/vlp.tex
Final edit, she be complete!

@@
      &\begin{cases}
-        \dot{q} = \frac{p}{ml^2} \\
-        \dot{l} = \frac{p_l}{m} \\
-        \dot{p} = -mgl\sin(q) \\
+        \dot{q} = \frac{p}{ml^2} 
+        ,\\
+        \dot{l} = \frac{p_l}{m} 
+        ,\\
+        \dot{p} = -mgl\sin(q) 
+        ,\\
         \dot{p}_l = \frac{p^2}{ml^3} + mg\cos(q) + \tau
@@
    = \frac{1}{m}
+   ,
 \]
@@
    \begin{cases}
-      \dot{q} = \frac{p}{m L^2} \\
+      \dot{q} = \frac{p}{m L^2} 
+      , \\
       \dot{p} = -mgL\sin(q)
@@
       \overline{l} & \theta \in [-\frac{\pi}{2},0[ \cup [\frac{\pi}{2}, \pi[ \\
-      \underline{l} & \theta \in [-pi, -\frac{\pi}{2}[ \cup [0,\frac{\pi}{2}[ 
+      \underline{l} & \theta \in [-\pi, -\frac{\pi}{2}[ \cup [0,\frac{\pi}{2}[ 
       .
@@
    with the property that the inequality is strict except at 
-   \(\theta \in \{0, \frac{\pi}{2}, \frac{\pi}, \frac{3\pi}{2}\}\).  
+   \(\theta \in \left\{0, \frac{\pi}{2}, \pi, \frac{3\pi}{2}\right\}\).  
    If instead 
@@
    \begin{cases}
-      \dot{q} = -\frac{p}{m L^2} \\
+      \dot{q} = -\frac{p}{m L^2} 
+      ,\\
       \dot{p} = mgL\sin(q)
@@
 dynamics~\eqref{eqn:vlp-hamiltonian}. 
-Also, we denote by \(\Pi^-\) the stable manifold of the equilibrium \((\pi,0)\)
+Also, we denote by \(\Pi^+\) the stable manifold of the equilibrium \((\pi,0)\)
 for system~\eqref{eqn:vlp-hamiltonian}, defined as
 \[
-   \Pi^- := \left\{ x(0) \in \SxR \mid
+   \Pi^+ := \left\{ x(0) \in \SxR \mid
    \lim_{t \to \infty} x(t) = (\pi, 0) \right\}.
@@
 negative, the stable manifold theorem implies that the stable manifold 
-\( \Pi^- \) is a one-dimensional immersed submanifold of
+\( \Pi^+ \) is a one-dimensional immersed submanifold of
 \(\SxR\), and thus is a set of measure zero
@@
 Let \( K \subset \SxR\) be an arbitrary compact set and let 
-\( x(0) \in K \backslash \left(\Pi^- \cup \left\{(0,0)\right\}\right)\).
-Since \( x(0) \notin \Pi^- \), we have that
+\( x(0) \in K \backslash \left(\Pi^+ \cup \left\{(0,0)\right\}\right)\).
+Since \( x(0) \notin \Pi^+ \), we have that
 \begin{equation}\label{eqn:vlp-x-does-not-converge}
@@
 \[
-   \{x \mid \dot E_\text{avg}(x) = 0\} = \left\{(q,p) \mid q = 0\right\}
+   \left\{x \mid \dot E_\text{avg}(x) = 0\right\} = \left\{(q,p) \mid q = 0\right\}
    \cup \left\{(q,p) \mid p =0\right\},
@@
 
-Returning to the Poincar\`e-Bendixson theorem, the positive limit set
+Returning to the Poincar\'e-Bendixson theorem, the positive limit set
 of \( x(t)\) must be an equilibrium and,
@@
       \underline{l} & \theta \in [-\frac{\pi}{2},0[ \cup [\frac{\pi}{2}, \pi[ \\
-      \overline{l} & \theta \in [-pi, -\frac{\pi}{2}[ \cup [0,\frac{\pi}{2}[ 
+      \overline{l} & \theta \in [-\pi, -\frac{\pi}{2}[ \cup [0,\frac{\pi}{2}[ 
       .
@@
    \caption{The orbits of the time-optimal controller \(L^\star(\theta)\)
-      (black), the differentiable controller \(L_T(\theta)\) (red),
+      (black), the differentiable controller \(L_T(\theta)\) (magenta),
       and the smooth VNHC \(L_\frac{\pi}{2}(\theta)\) (cyan) employed by a 25
@@
    \centering
-   \includegraphics[width=0.8\textwidth]{images/E_kiiking_ltheta.png}
+   \includegraphics[width=0.9\textwidth]{images/E_kiiking_ltheta.png}
    \caption{The energy \(E_\text{avg}(q,p)\) of a 25 year old Estonian male on a

### 2020-12-22 latex/vnhcs.tex
Update to the intro with a new image
Added a better intro, edited symbols and concusion
Final edit, she be complete!

@@
 \textit{equations of motion} (EOM) produced by Newton's second law \(F = m a\).
-While this technique works for simple systems, it is tedius and becomes
+While this technique works for simple systems, it is tedious and becomes
 impossible to apply to complex mechanical systems where the forces are not
@@
 equations} ~\eqref{eqn:el-eqns-general} with \textit{generalized input forces} 
-\(\tau \in \R^k\)
+\(\tau \in \R^k\),
 \begin{equation}\label{eqn:el-eqns-general}
@@
 potential function \(P : \mathcal{Q} \rightarrow \R\) is smooth.
-%If this is the case, the Euler-Lagrange equations reduce to ~\eqref{eqn:el-eqns},
-%\begin{equation}\label{eqn:el-eqns}
-%    D(q)\ddot{q} + C(q,\dot{q})\dot{q} + \nabla P(q) = B(q)\tau
-%\end{equation}
-%where the \textit{Coriolis matrix} \(C(q,\dot{q})\) is of the form
-%\[
-%    [C]_{i,j} = \frac{1}{2}\sum\limits_{k = 1}^n 
-%    \left(\pdiff{D_{i,j}}{q_k}  +
-%     \pdiff{D_{i,k}}{q_j} -
-%     \pdiff{D_{k,j}}{q_i}\right)\dot{q}_k
-%\]
 
@@
     \begin{cases}
-        \dot{q} = \nabla_p\mathcal{H} \\
+        \dot{q} = \nabla_p\mathcal{H} 
+        , \\
         \dot{p} = -\nabla_q\mathcal{H} + B(q)\tau
@@
         A \otimes B = \begin{bmatrix}
-            a_{1,1}B & \cdots & a_{1,m} B \\
+            A_{1,1}B & \cdots & A_{1,m} B \\
             \vdots & \ddots & \vdots \\
-            a_{n,1} B & \cdots & a_{n,m} B
+            A_{n,1} B & \cdots & A_{n,m} B
         \end{bmatrix} 
@@
     \end{align*}
-    for all \(i,j \in \n\).
+    for all \(i,j \in \{1,\ldots,n\}\).
 \end{thm}
@@
     \end{align*}
-    Note that \((A_i)\tpose (A\tpose)_j = (A_i)\tpose (A\inv)_j = \delta_{i,j}\). 
+    Since \(A\) is orthogonal, 
+    \((A_i)\tpose (A\tpose)_j = (A_i)\tpose (A\inv)_j = \delta_{i,j}\). 
     Using this fact we see that the Poisson brackets between \(P_i\) and \(Q_j\)
@@
     has simply actuated canonical coordinates 
-    \(\left(Q = \mathbf{B}q, P = \mathbf{B}p\right)\). The resulting dynamics are 
-    given by ~\eqref{eqn:simple-hamiltonian},
+    \(\left(\tilde{q} = \mathbf{B}q, \tilde{p} = \mathbf{B}p\right)\). 
+    The resulting dynamics are given by 
     \begin{align}\label{eqn:simple-hamiltonian}
-        \mathcal{H}(Q,P) &= 
-        \frac{1}{2} P\tpose \Minv(Q) P + V(Q)
+        \mathcal{H}(\tilde{q},\tilde{p}) &= 
+        \frac{1}{2} \tilde{p}\tpose \Minv(\tilde{q}) \tilde{p} + V(\tilde{q})
         , \\
        &\begin{cases}
-            \dot{Q} = \Minv(Q)P \\
-            \dot{P} = -\frac{1}{2} (\Id{n} \otimes P\tpose) \nabla_Q \Minv(Q) P
-                - \nabla_Q V(Q) + \simpleB \tau
+           \dot{\tilde{q}} = \Minv(\tilde{q})\tilde{p}
+           , \\
+           \dot{\tilde{p}} = -\frac{1}{2} (\Id{n} \otimes \tilde{p}\tpose)
+           \nabla_{\tilde{q}} \Minv(\tilde{q}) \tilde{p}
+           - \nabla_{\tilde{q}} V(\tilde{q}) + \simpleB \tau
             ,
@@
     \begin{align*}
-        \Minv(Q) &:= \mathbf{B}D^{-1}(\mathbf{B}\tpose Q)\mathbf{B}\tpose
+        \Minv(\tilde{q}) &:= 
+        \mathbf{B}D^{-1}(\mathbf{B}\tpose \tilde{q})\mathbf{B}\tpose
         , \\
-        V(Q) &:= P(\mathbf{B}\tpose Q)
+        V(\tilde{q}) &:= P(\mathbf{B}\tpose \tilde{q})
         .
@@
 \begin{proof}
-    The change of coordinates \((Q = \mathbf{B}q, P = \mathbf{B}p)\)
+    The change of coordinates \((\tilde{q} = \mathbf{B}q, \tilde{p} = \mathbf{B}p)\)
     satisfies Lemma \ref{lemma:linear-is-canonical}, making it a canonical
     transformation.
-    Furthermore, since \(\dot{P} = \mathbf{B}\dot{p}\), the new input matrix is
+    Furthermore, since \(\dot{\tilde{p}} = \mathbf{B}\dot{p}\), the new input matrix is
     given by 
@@
     \]
-    which means \(\left(Q = (q_u,q_a), P = (p_u,p_a)\right)\) are simply
+    which means \(\left(\tilde{q} = (q_u,q_a), \tilde{p} = (p_u,p_a)\right)\) are simply
     actuated coordinates for \(\mathcal{H}\) as desired.
@@
 
-%As has been observed in the field of mobile robotics, a trajectory tracking
-%controller which has lagged behind due to a disturbance will do whatever is
-%required to ``catch up" to the trajectory. 
-%This can cause major safety issues, which is why many mobile roboticists prefer to
-%design controllers which drive a vehicle along a gemoetric path in space rather than a
-%trajectory in time.
-
-%This path tracking approach has been extended to biologically-inspired robotics
-%in a method known as \textit{virtual holonomic constraints} (VHCs).
 Rather than tracking a trajectory over time, a more human-like behaviour 
@@
     \((q,p) \in h\inv(0)\), and there exists a feedback controller \(\tau(q,p)\)
-    rendering the \textit{constraint manifold} \(\Gamma\) invariant
+    rendering the \textit{constraint manifold} \(\Gamma\) invariant,
     where
@@
 is a \(2(n-k)\)-dimensional closed embedded submanifold of \(\mathcal{Q} \times \R^n\). 
-The next obvious question is the following: when does a feedback controller
-stabilizing \(\Gamma\) exist?
+The next obvious question is the following: when does there exists a feedback
+controller stabilizing \(\Gamma\)?
 
@@
     &\begin{cases}
-       \dot{q} = \Minv p \\
+       \dot{q} = \Minv p 
+       ,\\
        \dot{p} = -\frac{1}{2} \pdmat - \nabla_q V(q) + \simpleB \tau 
@@
 stabilized by\footnote{Under additional mild conditions
-\cite{vhcs_for_el_systems}} the output-linearizing phase-feedback controller
+\cite{vhcs_for_el_systems}.} the output-linearizing phase-feedback controller
 \begin{equation}\label{eqn:vnhc-torque-controller}
@@
     \end{bmatrix}\).
-    Defining \(e_1 := (1,0,\ldots,0) \in \R^n\), the actuated momentum is
+    Defining \(e_1 := (1,0,\ldots,0)\tpose \in \R^n\), the actuated momentum is
     \begin{equation}\label{eqn:g-qupu}

### 2020-12-24 latex/acrobot.tex
Added more detail to lit review and made edits as per Manfredi's suggestions

@@
 \(x(0) \in \mathcal{O}_1\) be a nonzero initial condition of the acrobot.
-Suppose the orbit \(x(\R)\) is confined within 
+Suppose by way of contradiction that the orbit \(x(\R)\) is confined within 
 \(\bar{\mathcal{O}}_1\), and does not exit through \(E_\pi\) into the rotation
@@
 In this section, we test our VNHC on the physical acrobot built by
-\citet{xingbo_thesis} (Figure \ref{fig:xingbo-acrobot}).
-Since it is not a simple acrobot, this experiment will test whether VNHCs are
-robust to model mismatch.
+Xingbo Wang \cite{xingbo_thesis} (Figure \ref{fig:xingbo-acrobot}).
+This acrobot is not simple, that is, the torso and leg rods have unequal mass
+and differ in length. 
+Furthermore, the centers of mass are not located at the tip of each rod.
+These experiments will therefore test whether VNHCs are robust to model
+mismatch, since our VNHC is only proven to inject energy on simple acrobots.
 As we will see, our VNHC does inject energy on the general acrobot model.

### 2020-12-24 latex/introduction.tex
Added more detail to lit review and made edits as per Manfredi's suggestions

@@
 
-The Estonian sport of ``Kiiking" (Figure \ref{fig:kiiking}) is the adult
+The Estonian sport of \textit{Kiiking} (Figure \ref{fig:kiiking}) is the adult
 equivalent of a child playing on a standing swing, though here
@@
 Virtual nonholonomic constraints have been of most notable use in
-bipedal locomotion \cite{nhvc_incline_walking,output_nhvc_bipedal_control},
+bipedal locomotion 
+\cite{nhvc_gait_optimization,output_nhvc_bipedal_control},
 where they have shown marked improvements in the robustness of walking gaits
-when compared to previous control technqiues.
+when compared to previous control techniques.
 They have also been used for error-reduction in time-delayed teleoperation
@@
 \cite{psd_based_vnhc_redundant_manipulator,haptic_vnhc}.
-All these research areas use different formulations of virtual nonholonomic constraints;
 \citet{hybrid_zero_dynamics_bipedal_nhvcs} have derived the 
 zero dynamics for hybrid mechanical systems under virtual nonholonomic
-constraints, but to the best of our knowledge no one has attempted to generalize
-the concept to abstract mechanical systems.
+constraints.
+In particular, they study a class of constraints formed by taking an affine
+sum of holonomic functions with nonholonomic B\'{e}zier polynomials.
+They widen this class of constraints in \cite{nhvc_incline_walking} to improve
+bipedal locomotion on variable-slope terrain, and generate constraints through
+a specialized optimization routine.
+
+Each of these research areas use different formulations of virtual
+nonholonomic constraints.
+Furthermore, the most recent literature surrounding virtual nonholonomic constraints
+has focused on enforcing stable periodic orbits for hybrid mechanical systems
+with impacts.
+To the best of our knowledge, no one has attempted to create a single unified
+framework of virtual nonholonomic constraints for general mechanical systems.
+Moreover, no one has studied how virtual nonholonomic constraints can be used for energy
+injection, where periodic motion is not desirable.
+This thesis attempts to fill these gaps in the research.
 

### 2020-12-24 latex/vnhcs.tex
Added more detail to lit review and made edits as per Manfredi's suggestions

@@
     the orbit \((q_u(t),p_u(t))\) traces out
-    a curve on the 2D constraint manifold 
+    a curve on the constraint manifold 
     \(\Gamma \simeq \Rt{T_u} \times \R\) which we call the \((q_u,p_u)\)-plane.
@@
 
-%%/========== /Virtual Nonholonomic Constraints ==========/% 
+%/========== /Virtual Nonholonomic Constraints ==========/% 
 % vim: set tw=80 ts=4 sw=4 sts=0 et ffs=unix :

### 2020-12-25 latex/introduction.tex
More edits to VNHC and intro to clarify the differences between us and Alireza

@@
 
-Each of these research areas use different formulations of virtual
-nonholonomic constraints.
-Furthermore, the most recent literature surrounding virtual nonholonomic constraints
+The most recent literature surrounding virtual nonholonomic constraints
 has focused on enforcing stable periodic orbits for hybrid mechanical systems
 with impacts.
-To the best of our knowledge, no one has attempted to create a single unified
-framework of virtual nonholonomic constraints for general mechanical systems.
-Moreover, no one has studied how virtual nonholonomic constraints can be used for energy
-injection, where periodic motion is not desirable.
-This thesis attempts to fill these gaps in the research.
+To the best of our knowledge, no one has studied how virtual nonholonomic
+constraints can be used for energy injection, where periodic motion is not
+desirable.
+This thesis attempts to fill this gap in the research.
+For more detail on the differences between this thesis and the existing research
+on virtual nonholonomic constraints, see Chapter \ref{ch:vnhc-compare}.
 

### 2020-12-25 latex/vnhcs.tex
More edits to VNHC and intro to clarify the differences between us and Alireza

@@
 
+\section{Comparison with Existing VNHC Literature}\label{ch:vnhc-compare}
+
+Griffin and Grizzle were the first to use the unactuated conjugate of momentum
+to define relative degree \(\{2,\ldots,2\}\) VNHCs in
+\cite{nhvc_dynamic_walking} and \cite{nhvc_gait_optimization}.
+\citet{hybrid_zero_dynamics_bipedal_nhvcs} used this idea to define
+the unactuated coordinate \(q_u = B^\perp q\), derive the unactuated momentum in
+Lagrangian coordinates, and specify that a VNHC must use this unactauted
+momentum.
+These ideas are what inspired us to develop the theory of simply actuated
+Hamiltonian systems, which is more suited to this area of research than
+Lagrangian mechanics. 
+Furthermore, our definition of a VNHC is more general than that of Griffin and
+Grizzle.
+They defined a VNHC as a relation \(h(q,\sigma_u)\) (where \(\sigma_u\) is the
+Lagrangian mechanics proxy for \(p_u\)), while our definition of VNHCs allows
+for relations that use the full phase.
+Moreover, we are the first to provide the definition of regular VNHCs, and to
+provide the characterization in Theorem \ref{thm:vnhc-regularity}.
+
+\citet{hybrid_zero_dynamics_bipedal_nhvcs} were the first to derive the
+constrained dynamics for VNHCs of the form \(h(q,\sigma_u)\).
+In our derivation of the constrained dynamics, we use the same assumptions as
+they do, with the additional assumption that one can solve for 
+\(q_a = f(q_u,p_u)\) on \(\Gamma\).
+While this means the constrained dynamics in
+\cite{hybrid_zero_dynamics_bipedal_nhvcs} hold for a larger class of VNHCs, 
+their derivations are terse and unclear.
+By making this additional assumption, we were able to find a the constrained
+dynamics in simply actuated coordinates, which are much more physically
+intuitive.
+Having the constrained dynamics in simply actuated coordinates will also
+allow us to more easily study energy injection properties of VNHCs in the next
+two chapters.
+
 %/========== /Virtual Nonholonomic Constraints ==========/% 

### 2020-12-26 latex/vnhcs.tex
Edit of the comparison with existing VNHC literature, plus an addition to the conclusion

--- 2020-12-24 latex/vnhcs.tex
@@
 
+\section{Comparison with Existing VNHC Literature}\label{ch:vnhc-compare}
+
+Griffin and Grizzle were the first to use the unactuated conjugate of momentum
+to define relative degree \(\{2,\ldots,2\}\) VNHCs in
+\cite{nhvc_dynamic_walking} and \cite{nhvc_gait_optimization}.
+\citet{hybrid_zero_dynamics_bipedal_nhvcs} extended their work by defining
+the unactuated coordinate \(q_u = B^\perp q\) and specifying that a
+VNHC must use the conjugate of momentum to \(q_u\).
+This idea is what inspired us to use Hamiltonian mechanics in simply actuated
+coordinates, which are more suitable for the study of VNHCs than
+Lagrangian mechanics.
+
+Our definition of a VNHC is more general than that of Griffin and
+Grizzle.
+They define a VNHC as a relation \(h(q,\sigma_u)\), where \(\sigma_u\) is the
+Lagrangian mechanics proxy for \(p_u\).
+By contrast, our definition of a VNHC allows for relations \(h(q,p)\)
+that use the full phase.
+Moreover, we are the first to provide the definition of regular VNHCs, and to
+provide the characterization in Theorem \ref{thm:vnhc-regularity}.
+
+\citet{hybrid_zero_dynamics_bipedal_nhvcs} derive the
+constrained dynamics for VNHCs of the form \(h(q,\sigma_u)\).
+For our derivation of the constrained dynamics in Theorem
+\ref{thm:zero-dynamics}, we use the same assumptions as they do, with the
+additional assumption that one can solve for \(q_a = f(q_u,p_u)\) on \(\Gamma\).
+While this means Horn \etal's constrained dynamics hold for a larger class of VNHCs, 
+their derivations are obscured by a need to use more complicated machinery
+to find the closed form of the dynamics. 
+By making the aforementioned additional assumption, we were able to find a the constrained
+dynamics as functions of the unactuated coordinates.
+Hence, our constrained dynamics are thus more physically intuitive than those of
+Horn \etal~% no period because etal has one 
+Finally, the additional structure of our constrained dynamics will allow us to
+more easily study energy injection properties of VNHCs in the next two chapters.
+
 %/========== /Virtual Nonholonomic Constraints ==========/% 

### 2020-12-27 latex/vnhcs.tex
Final edit of VNHC chapter for the lit review

--- 2020-12-24 latex/vnhcs.tex
@@
 
+\section{Comparison with Existing VNHC Literature}\label{ch:vnhc-compare}
+
+Griffin and Grizzle were the first to define relative degree \(\{2,\ldots,2\}\)
+VNHCs (see \cite{nhvc_dynamic_walking} and \cite{nhvc_gait_optimization}). 
+They required that the first coordinate be unactuated, so 
+\citet{hybrid_zero_dynamics_bipedal_nhvcs} extended their work to more general
+mechanical systems by defining the unactuated coordinate \(q_u = B^\perp q\).
+These ideas are what inspired us to use simply actuated Hamiltonian systems,
+which are more suitable for the study of VNHCs than Lagrangian mechanics.
+
+Our definition of a VNHC is more general than that of Griffin and
+Grizzle.
+They define a VNHC as a relation \(h(q,\sigma_u)\), where \(\sigma_u\) is the
+Lagrangian mechanics proxy for \(p_u\).
+Their definition carries an implicit assumption that VNHCs are of
+relative degree \(\{2,\ldots,2\}\), which are what we call regular VNHCs.
+By contrast, our definition of a VNHC allows for relations \(h(q,p)\)
+that use the full phase.
+We are the first to distinguish these from regular VNHCs, and to
+provide the characterization of regularity in Theorem \ref{thm:vnhc-regularity}.
+
+\citet{hybrid_zero_dynamics_bipedal_nhvcs} derive the
+constrained dynamics for VNHCs of the form \(h(q,\sigma_u)\).
+Our derivation of the constrained dynamics in Theorem
+\ref{thm:zero-dynamics} uses the same assumptions they require, with the
+additional assumption that one can solve for \(q_a = f(q_u,p_u)\) on the
+constraint manifold.
+While this means Horn \etal's constrained dynamics hold for a larger class of
+VNHCs, their derivations are obscured by a complicated change of coordinates
+when finding the closed form of the dynamics. 
+By making the aforementioned additional assumption, we are instead able to find
+the constrained dynamics in terms of the unactuated coordinates.
+This makes our constrained dynamics more physically intuitive, since they
+describe the behaviour of a specific part of the mechanical system.
+This additional structure will facilitate our analysis of the energy injection
+properties of VNHCs across the next two chapters.
+
 %/========== /Virtual Nonholonomic Constraints ==========/% 

### 2020-12-28 latex/acrobot.tex
Fixed some typos caught by mother

@@
 \(x(t,\eta_0,I) = x_0(t,\eta_0) + I x_1(t,\eta_0) + R(t,\eta_0,I)\),
-where the remainder term \(R(t,\eta_0,I)\) is smooth and \(O(I^2)\), \ie, 
+where the remainder term \(R(t,\eta_0,I)\) is smooth and \(O(I^2)\), \ie 
 \[
@@
 describing how the pseudo-radius \(\mu\) changes each time the orbit of \((q_u(t),p_u(t))\)
-intersects the \(q_u\)-axis, \ie, when \(\alpha = 0\) or \(\alpha = \pi\).
+intersects the \(q_u\)-axis, \ie when \(\alpha = 0\) or \(\alpha = \pi\).
 Using our time-scaled system ~\eqref{eqn:muhat-dot}, we can define this
@@
 there is a well-defined Poincar\'{e} map describing how the pseudo-radius
-\(\rho\) changes each time the orbit \((q(t),p(t))\) hits the \(p_u\) axis, \ie,
+\(\rho\) changes each time the orbit \((q(t),p(t))\) hits the \(p_u\) axis, \ie
 every time \(\beta\) changes by \(2\pi\).
@@
 for some \(\bar{\rho} > \sqrt{60m^2gl^3}\).
-Recall from Section \ref{sec:acrobot-proof-o1} that any acrobot initialized in
+Recall from Chapter \ref{sec:acrobot-proof-o1} that any acrobot initialized in
 \(\mathcal{O}_1\) will eventually remain in the rotation region \(\mathcal{R}\).
@@
 If \(P_\mathcal{R}(\rho_0)\) is always further from the origin than 
-\(\rho_0\) (\ie~ it is ``expanding" on \(D\)), the acrobot will be gaining energy
+\(\rho_0\) (\ie it is ``expanding" on \(D\)), the acrobot will be gaining energy
 on \(\mathcal{O}_2(\bar{\rho})\).
@@
 \end{enumerate}
-Recall from Section \ref{sec:acrobot-proof-o1} that there exists \(I^\star > 0\)
+Recall from Chapter \ref{sec:acrobot-proof-o1} that there exists \(I^\star > 0\)
 making orbits escape \(\bar{\mathcal{O}}_1\) into 
@@
 
-This analysis, together with that of section \ref{sec:acrobot-proof-o1},
+This analysis, together with that of Chapter \ref{sec:acrobot-proof-o1},
 completes the proof of Theorem \ref{thm:acrobot-energy-stabilization}.

### 2020-12-28 latex/vlp.tex
Fixed some typos caught by mother

@@
 dynamics~\eqref{eqn:vlp-hamiltonian} escape compact sets in finite
-time, \ie, that for each compact subset \(K\) of 
+time, \ie that for each compact subset \(K\) of 
 \(\SxR\), and for almost every condition 
@@
 
-The average weight of a 25 year old Estonian man is \(80.17\)kg
+The average weight of a 25-year-old Estonian man is \(80.17\)kg
 \cite{estonian-body-data}, so we set the VLP mass \(m\) to this value.
@@
 \cite{kiiking-sale}, so we take the length to be \(3\)m.
-The average 25 year old Estonian is \(1.8\)m tall \cite{estonian-body-data}, and
+The average 25-year-old Estonian is \(1.8\)m tall \cite{estonian-body-data}, and
 we assume their squat height is half that at \(0.9\)m tall.
@@
       (black), the differentiable controller \(L_T(\theta)\) (magenta),
-      and the smooth VNHC \(L_\frac{\pi}{2}(\theta)\) (cyan) employed by a 25
-      year old Estonian male on a 3m tall Kiiking swing. All the controllers are
-      initialized at \(q(0) = \frac{\pi}{32}\), and they each steadily gain
+      and the smooth VNHC \(L_\frac{\pi}{2}(\theta)\) (cyan) employed by a
+      25-year-old Estonian male on a 3m tall Kiiking swing. All the controllers
+      are initialized at \(q(0) = \frac{\pi}{32}\), and they each steadily gain
       energy until they reach a rotation.}
@@
    \includegraphics[width=0.9\textwidth]{images/E_kiiking_ltheta.png}
-   \caption{The energy \(E_\text{avg}(q,p)\) of a 25 year old Estonian male on a
+   \caption{The energy \(E_\text{avg}(q,p)\) of a 25-year-old Estonian male on a
       3m tall Kiiking swing who employs the smooth VNHC

### 2021-02-02 latex/paper/main.tex
Initial setup of the paper for VNHC + Acrobot, without the template ready to compile yet.

--- 2020-02-28 latex/bcc/main.tex
@@
 
-\documentclass[12pt]{article}
-\usepackage[utf8]{inputenc}
-
-%----------- Choose Margins ----------%
-\usepackage[margin=1in]{geometry}
+\documentclass[]{}
+%/========== Preamble ==========/%
+\usepackage{adanstyle}
 
 %---------- Bibliography Style ----------%
-\usepackage[style=ieee,backend=biber]{biblatex}
-\addbibresource{bib.bib}
-
-%/========== Preamble ==========/%
-
-%----------- Math Packages -----------%
-\usepackage{amsfonts}
-\usepackage{mathtools}
-\usepackage{amsthm} % allow theorems and definitions as a block
-
-%---------- Remove indenting from paragraphs ----------%
-% Also makes each new paragraph start with an vertical space.
-\usepackage[parfill]{parskip}
 
@@
 \begin{document}
-%---------- Title ----------%
-\title{Energy Injection for Mechanical Systems via Virtual Nonholonomic Constraints}
-\author{Adan Moran-MacDonald}
-\maketitle
 
-%---------- Content ----------%
-When a gymnast hanging on a bar performs a swing-up routine (a so-called ``giant"),
-they do not attempt to track a trajectory over time to start doing backflips.
-Likewise, a child on a swing does not swing higher by kicking their legs at
-specific time intervals; rather, these natural human motions change one's leg
-angle based on the current direction and speed of motion. In doing so, the
-gymnast will gain enough energy to start rotating on the bar, 
-and the child will reach greater heights on their swing.
-This type of motion can be adapted to the control of mechanical systems 
-using the method of virtual nonholonomic constraints.
-
-A \textit{virtual nonholonomic constraint} (VNHC) is a smooth relation 
-\[
-    h : T^\star \mathcal{Q} \rightarrow \mathbb{R}^{n-1}
-\]
-between the \(n\) configuration variables \(q \in \mathcal{Q}\) and conjugate of
-momenta \(p\) of a mechanical system with \(n-1\) control inputs.
-Enforcing these types of constraints via feedback control provides
-time-independent controllers which can inject or remove energy from the system.
-
-This talk will give sufficient conditions under which a VNHC 
-can be enforced by feedback control.
-Then, using the motivations of the child on the swing and the gymnast,
-we apply the theory to create VNHCs for two classical mechanical systems:
-the variable-length pendulum and the two-link acrobot.
-
-For the variable-length pendulum we design and enforce a VNHC which achieves a
-time-optimal swingup motion, without tracking a time signal. 
-Then, we prove that a system constrained by this VNHC will gain energy and start
-rotating.
-
-For the acrobot, we design several VNHCs which achieve human-like motion by
-observing the leg motions of human gymnasts. We then apply one of these
-VNHCs to a real acrobot and, through experiments, demonstrate the
-energy-injection properties of the VNHC.
-
-This talk demonstrates the potential of VNHCs in generating realistic,
-biologically-inspired motions which can inject energy into a mechanical system
-while forgoing any dependence on time in the control architecture.
-
-%---------- Bibliography ----------%
-\newpage
-\printbibliography
 \end{document}
 %/========== /Main Document ==========/%
-
 % vim: set tw=80 ts=4 sw=4 sts=0 et ffs=unix :

### 2021-02-02 latex/paper/main.tex
First empty setup of the paper

@@
 
-\documentclass[]{}
+\documentclass[journal,twoside,web]{ieeecolor}
+\usepackage{generic}
 %/========== Preamble ==========/%
-\usepackage{adanstyle}
-
+\usepackage{cite}
+\usepackage{amsmath,amssymb,amsfonts}
+\usepackage{algorithmic}
+\usepackage{graphicx}
+\usepackage{textcomp}
 %---------- Bibliography Style ----------%
-
+\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
+    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+%---------- Header ----------% 
+\newcommand*{\Title}{VNHC and Acrobot Title}
+\markboth{\journalname, VOL. XX, NO. XX, XXXX 2021}
+{Moran-MacDonald \MakeLowercase{\textit{et al.}}: \Title}
 %/========== /Preamble ==========/%
@@
 \begin{document}
+\title{\Title}
+\author{Adan Moran-MacDonald, \IEEEmembership{Member, IEEE}, Manfredi Maggiore
+\IEEEmembership{Fellow, IEEE}*, and Xingbo Wang
+\thanks{Manuscript submitted for review on \today.}
+\thanks{A. Moran-MacDonald is with the Department of Electrical and Computer
+    Engineering, University of Toronto, Toronto, ON, Canada (e-mail:
+adan.moran@mail.utoronto.ca).}
+\thanks{M. Maggiore is with the Department of Electrical and Computer
+Engineering, University of Toronto, ON, Canada (e-mail:
+maggiore@control.utoronto.ca).}
+\thanks{X. Wang is with ??? (e-mail: ???).}
+} %/author
+
+\maketitle
+
+%/========== Abstract ==========/%
+\begin{abstract}
+TODO: Abstract here.
+\end{abstract}
+
+\begin{IEEEkeywords}
+TODO: keywords in alphabetical order, separated by commas.
+\end{IEEEkeywords}
+
+%/========== Introduction ==========/%
+\section{Introduction}\label{sec:introduction}
+\IEEEPARstart{T}ODO: intro.
+
+\subsection{Notation}
+
+%/========== Problem Formulation ==========/%
+\section{Problem Formulation}\label{sec:problem-formulation}
+TODO: motivate injection/dissipation for energy regulation in an acrobot. Use
+this to lead into VNHCs. e.g. represent torso as chain of a swing, pivot as
+seat, and legs as human legs. Replicate fig3.3 from thesis but for acrobot: take
+(qu,pu)-plane and show how a person moves their legs wrt theta, qu, or pu. Leave
+the gymnast model until the end, or ignore it entirely.
+
+Alternatively, ff human movement journal says legs move wrt qu, we can use that
+as a foundation.
+
+%/========== VNHC ==========/%
+\section{Theory of VNHCs}\label{sec:vnhc}
+Just like in the thesis, cover the theory of VNHCs. Justify our results compared
+to those of Horn. Make it terse but legible. Cover simply actuated systems,
+vnhcs, regularity, the theorems, etc.
+
+%/========== Acrobot ==========/%
+\section{The Acrobot VNHC}\label{sec:acrobot}
+specialize the theory of VNHCs to the acrobot. consider constraints which depend
+only on pu, and show arctan vnhc and reduced dynamics. conclude with our
+theorem. Make this section short and sweet.
+
+%/========== Simulation ==========/%
+\section{Simulation Results}\label{sec:simulations}
+
+%/========== Experiments ==========/%
+\section{Experimental Results}\label{sec:experiments}
+
+%/========== Proof ==========/%
+\section{Proof of Theorem \textbf{TODO: ref theorem}}\label{sec:proof}
+
+%/========== Conclusion ==========/%
+\section{Conclusion}\label{sec:conclusion}
 

### 2021-02-03 latex/paper/main.tex
Started the paper. For some reason citations are not numbers.

@@
 \usepackage{textcomp}
+
 %---------- Bibliography Style ----------%
-\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
-    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+\bibliographystyle{IEEEtranN}
+
 %---------- Header ----------% 
@@
 {Moran-MacDonald \MakeLowercase{\textit{et al.}}: \Title}
+
+%---------- Special Commands ----------%
+\newcommand*{\tpose}{^\mathsf{T}} 
+\newcommand*{\inv}{^\mathsf{-1}}
+\newcommand*{\Rt}[1]{[\R]_{#1}}
+\newcommand*{\R}{\mathbb{R}}
+\newcommand*{\n}{\mathbf{n}}
+\newcommand*{\N}{\mathbb{N}}
+\newcommand*{\Sone}{\mathbb{S}^1}
+\newcommand*{\SxR}{\Sone \times \R}
+\newcommand*{\Minv}{M^\mathsf{-1}}
+\newcommand*{\Id}[1]{I_{#1}}
+\newcommand*{\Zmat}[1]{\bm{0}_{#1}}
+\newcommand*{\diff}[2]{\frac{d #1}{d #2}}
+\newcommand*{\ddiff}[3]{\frac{d^2 #1}{d #2 d #3}}
+\newcommand*{\pdiff}[2]{\frac{\partial #1}{\partial #2}}
 %/========== /Preamble ==========/%
@@
 \subsection{Notation}
+We use the following notation in this paper. 
+For \(T > 0\), the set of real numbers modulo is denote \(\Rt{T}\), with
+\(\Rt{\infty} := \R\).
+The gradient of the matrix-valued function 
+\(A : \R^m \rightarrow \R^{n\times n}\) is the matrix of stacked partial
+derivatives, 
+\(\nabla_xA := (\pdiff{A}{x_1},\ldots,\pdiff{A}{x_m})\tpose \in \R^{nm \times n}\).
+Given two matrices \(A \in \R^{n \times m}\) and \(B \in \R^{r \times s}\), the
+Kronecker product \cite{kronprod} \(A \otimes B \in \R^{nr \times ms}\) is the
+matrix
+\[
+    A \otimes B = \begin{bmatrix}
+        A_{1,1}B & \cdots & A_{1,m} B \\
+        \vdots & \ddots & \vdots \\
+        A_{n,1} B & \cdots & A_{n,m} B
+    \end{bmatrix} 
+    .
+\]
 
@@
 \section{Theory of VNHCs}\label{sec:vnhc}
-Just like in the thesis, cover the theory of VNHCs. Justify our results compared
+TODO: Just like in the thesis, cover the theory of VNHCs. Justify our results compared
 to those of Horn. Make it terse but legible. Cover simply actuated systems,
 vnhcs, regularity, the theorems, etc.
+
+\subsection{Simply Actuated Hamiltonian Systems}
+Take an underactuated mechanical system modelled with generalized coordinates 
+\(q = (q_1, \ldots, q_n)\) on the configuration manifold
+\(\mathcal{Q} = \Rt{T_1} \times \cdot \Rt{T_n}\), where
+\(T_i = 2\pi\) if \(q_i\) is an angle and \(T_i = \infty\) if \(q_i\) is a
+length. The corresponding velocities are 
+\(\dot{q} = (\dot{q}_1,\ldots,\dot{q}_n) \in \R^n\).
+
+Suppose this system has Lagrangian
+\(\mathcal{L}(q,\dot{q}) = 1/2 \dot{q}^T D(q) \dot{q} - P(q)\),
+where \(P : \mathcal{Q} \rightarrow \mathbb{R}\) is the potential energy, and
+the inertia matrix \(D : \mathcal{Q} \rightarrow \mathbb{R}^{n \times n}\)
+is positive definite for all \(q \in \mathcal{Q}\).
+By defining the \textit{conjugate of momentum} 
+\(p_i := \pdiff{\mathcal{L}}{\dot{q}_i}\) so that the vector 
+\(p = (p_1,\ldots,p_n) \in \R^n\) is given by \(p = D(q)\dot{q}\),
+we find the \textit{Hamiltonian} of the system in \((q,p)\) coordinates is
+\(\mathcal{H}(q,p) = 1/2 p\tpose D(q) p + P(q)\) \cite{landau_mechanics}.
+The dynamics of the system in Hamiltonian form are
+\begin{equation}\label{eqn:hamiltonian-eom-general}
+    \begin{cases}
+        \dot{q} = \nabla_p\mathcal{H} 
+        , \\
+        \dot{p} = -\nabla_q\mathcal{H} + \tau
+        ,
+    \end{cases}
+\end{equation}
+where \(\tau \in \R^k\) is a vector of generalized input forces and the input
+matrix \(B \in \R^{n \times k}\) is assumed to be full rank \(k < n\).
+
+It is easy to show that \eqref{eqn:hamiltonian-eom-general} reduces to
+\begin{equation}\label{eqn:hamiltonian-full-dynamics}
+     \begin{cases}
+        \dot{q} = D\inv(q)p \\
+        \dot{p} = -\frac{1}{2} (\Id{n} \otimes p\tpose) \nabla_q D\inv(q) p
+        - \nabla_q P(q) + B(q) \tau
+        . \\
+    \end{cases}
+\end{equation}
 
@@
 
+%---------- Bibliography ----------%
+\bibliography{bib}
 \end{document}

### 2021-02-10 latex/paper/main.tex
Updated the bib style so that it works well
Started filling in VNHC section

@@
 \documentclass[journal,twoside,web]{ieeecolor}
+
 \usepackage{generic}
 %/========== Preamble ==========/%
-\usepackage{cite}
-\usepackage{amsmath,amssymb,amsfonts}
+% Packages required by ieeecolor
+\usepackage{amsmath}
+\usepackage{amsfonts}
+\usepackage{amssymb}
 \usepackage{algorithmic}
@@
 \usepackage{textcomp}
+\usepackage{cite}
+
+%---------- More packages ----------%
+% NOTE: cannot use amsthm for some reason
+\usepackage{mathtools}
+\usepackage{bm}
 
 %---------- Bibliography Style ----------%
-\bibliographystyle{IEEEtranN}
+\bibliographystyle{IEEEtran}
 
@@
 
+%---------- Theorems ----------%
+\newtheorem{thm}{Theorem}% Uncomment: reset numbering at each chapter
+%\newtheorem{prop}[thm]{Proposition} % Propositions depend on Theorem number
+\newtheorem{lemma}[thm]{Lemma} % Same with lemmas
+\newtheorem{defn}[thm]{Definition} % Definitions depend on theorem number
+\newtheorem{assm}{Assumption} % Assumptions do not reset
+
 %---------- Special Commands ----------%
+\DeclarePairedDelimiter{\norm}{\lVert}{\rVert}
+\DeclareMathOperator{\Rank}{rank}
+\DeclareMathOperator{\Sign}{sgn}
+\newcommand*{\rank}[1]{\Rank\left(#1\right)}
+\newcommand*{\sign}[1]{\Sign\left(#1\right)}
+
 \newcommand*{\tpose}{^\mathsf{T}} 
@@
 \newcommand*{\pdiff}[2]{\frac{\partial #1}{\partial #2}}
+
+\newcommand*{\simpleB}{\begin{bmatrix}\Zmat{(n-k)\times k}\\ \Id{k}\end{bmatrix}}
+\newcommand*{\pdmat}{(\Id{n} \otimes p\tpose)\nabla_q\Minv(q)p}
+\newcommand*{\pudmat}{(\Id{n-k} \otimes p\tpose)\nabla_{q_u}\Minv(q)p}
 %/========== /Preamble ==========/%
@@
 \author{Adan Moran-MacDonald, \IEEEmembership{Member, IEEE}, Manfredi Maggiore
-\IEEEmembership{Fellow, IEEE}*, and Xingbo Wang
+\IEEEmembership{Member, IEEE}*, and Xingbo Wang
 \thanks{Manuscript submitted for review on \today.}
@@
 \(p = (p_1,\ldots,p_n) \in \R^n\) is given by \(p = D(q)\dot{q}\),
-we find the \textit{Hamiltonian} of the system in \((q,p)\) coordinates is
-\(\mathcal{H}(q,p) = 1/2 p\tpose D(q) p + P(q)\) \cite{landau_mechanics}.
+we find the \textit{Hamiltonian} of the system in \((q,p)\) coordinates
+is\cite{landau_mechanics}
+\begin{equation}\label{eqn:hamiltonian}
+    \mathcal{H}(q,p) = 1/2 p\tpose D(q) p + P(q).
+\end{equation}
 The dynamics of the system in Hamiltonian form are
@@
         , \\
-        \dot{p} = -\nabla_q\mathcal{H} + \tau
+        \dot{p} = -\nabla_q\mathcal{H} + B(q) \tau
         ,
@@
 where \(\tau \in \R^k\) is a vector of generalized input forces and the input
-matrix \(B \in \R^{n \times k}\) is assumed to be full rank \(k < n\).
+matrix \(B : \mathcal{Q} \rightarrow \R^{n \times k}\) is full rank for all 
+\(q \in \mathcal{Q}\).
 
-It is easy to show that \eqref{eqn:hamiltonian-eom-general} reduces to
+It is easy to show using the matrix Kronecker product that
+\eqref{eqn:hamiltonian-eom-general} reduces to
 \begin{equation}\label{eqn:hamiltonian-full-dynamics}
@@
 \end{equation}
+
+Because \(\tau\) is transformed by \(B(q)\), it is not in general clear how any
+particular input force \(\tau_i\) will affect the dynamics of the system. 
+The following definition provides a class of systems where the effect of input
+forces is made explicit.
+
+\begin{defn}
+    Let \(\mathcal{H}\) be an \(n\)-DOF Hamiltonian system 
+    with \(k \leq n\) actuators. 
+    A set of canonical coordinates \((q,p)\) for this system
+    are said to be \textit{simply actuated coordinates} if the
+    input matrix \(B(q) \in \R^{n \times k}\) is of the form
+    \[
+        B(q) = \simpleB    
+        .
+    \]
+
+\end{defn}
+
+The first \((n-k)\) coordinates, labelled \(q_u\), are called the
+\textit{unactuated coordinates}. The remaining \(k\) coordinates, labelled
+\(q_a\), are called the \textit{actuated coordinates}. When grouping them
+together, we will always put them in the order \((q_u, q_a)\) to fit with 
+the definition. 
+The corresponding \((p_u, p_a)\) are called the \textit{unactuated} and 
+\textit{actuated momenta}, respectively.
+
+Under the following assumptions on the input matrix, we will show that there is
+a canonical transformation of ~\eqref{eqn:hamiltonian} into simply actuated
+coordinates.
+
+\begin{assm}\label{assm:B-const}
+    The input matrix \(B(q) \equiv B \in \R^{n\times k}\) is constant,
+    full rank, and \(k < n\).
+\end{assm}
+
+\begin{assm}\label{assm:B-perp}
+    There exists a matrix 
+    \(B^\perp \in \R^{(n-k)\times n}\)
+    which is right semi-orthogonal 
+    \(\left(B^\perp(B^\perp)\tpose = \Id{(n-k)}\right)\)
+    and which is a left-annihilator for \(B\). 
+    That is, \(B^\perp B = \Zmat{(n-k) \times k}\).
+\end{assm}
+
+Assumption \ref{assm:B-perp} requires the rows of \(B^\perp\) to be unit vectors
+that are mutually orthogonal. 
+In the case that \(k = (n-1)\), the existence of any left annihilator 
+\(A^0 \in \R^{1\times n}\) implies the left annihilator 
+\(B^\perp := A^0/\norm{A^0}\) will be a unit vector satisfying Assumption \ref{assm:B-perp}.
+
+\begin{lemma}\label{lemma:B-orthogonal}
+    Suppose Assumption \ref{assm:B-const} holds. Then
+    there exists a nonsingular matrix \(\hat{T} \in \R^{k \times k}\) 
+    so that the regular feedback transformation 
+    \[
+        \tau = \hat{T} \hat{\tau}
+    \] 
+    has a new input matrix \(\hat{B}\) for \(\hat{\tau}\) which is left
+    semi-orthogonal.  
+    That is, \(\hat{B}\tpose \hat{B} = \Id{k}\).
+\end{lemma}
+\begin{proof}
+    \textbf{TODO: prove this in a terse way}.
+\end{proof}
+
+In light of Lemma \ref{lemma:B-orthogonal}, there is no loss of generality
+by making the following assumption.
+\begin{assm}\label{assm:B-orthogonal}
+    Assume that the input matrix \(B\) is
+    left semi-orthogonal, i.e.~\(B\tpose B = I_k\). 
+\end{assm}
+
+Let now \(\mathbf{B} \in \R^{n\times n}\) be the following matrix:
+\[
+    \mathbf{B} = 
+    \begin{bmatrix}
+        B^\perp \\
+        B\tpose \\
+    \end{bmatrix}
+    .
+\]
+Since \(B^\perp\) is a left annihilator of \(B\) and both \(B^\perp\) and
+\(B\tpose\) are right semi-orthogonal, it is easy to show that \(\mathbf{B}\) is
+orthogonal.
+\begin{proof}
+\[
+    \mathbf{B}\mathbf{B}\tpose = 
+    \begin{bmatrix}
+        B^\perp (B^\perp)\tpose & B^\perp B \\
+        (B^\perp B)\tpose & B\tpose B
+    \end{bmatrix} = \Id{n}
+    .
+\]
+Hence, \(\mathbf{B}\) is invertible with \(\mathbf{B}\inv = \mathbf{B}\tpose\).
+\end{proof}
+
+The following theorem shows that \(\mathbf{B}\) provides a canonical
+transformation into simply actuated coordinates, so that only the actuated momenta
+are affected by the input forces.
+
+\begin{thm}\label{thm:simply-actuated}
+    Under Assumptions \ref{assm:B-const},\ref{assm:B-perp}, and
+    \ref{assm:B-orthogonal}, the Hamiltonian system ~\eqref{eqn:hamiltonian}
+    has simply actuated canonical coordinates 
+    \(\left(\tilde{q} = \mathbf{B}q, \tilde{p} = \mathbf{B}p\right)\). 
+    The resulting dynamics are given by 
+    \begin{gather}\label{eqn:simple-hamiltonian}
+        \mathcal{H}(\tilde{q},\tilde{p}) = 
+        \frac{1}{2} \tilde{p}\tpose \Minv(\tilde{q}) \tilde{p} + V(\tilde{q})
+        , \\
+       \begin{cases}
+           \dot{\tilde{q}} = \Minv(\tilde{q})\tilde{p}
+           , \\
+           \dot{\tilde{p}} = -\frac{1}{2} (\Id{n} \otimes \tilde{p}\tpose)
+           \nabla_{\tilde{q}} \Minv(\tilde{q}) \tilde{p}
+           - \nabla_{\tilde{q}} V(\tilde{q}) + \simpleB \tau
+            ,
+        \end{cases} \nonumber
+    \end{gather}
+    where 
+    \(\Minv(\tilde{q}) := \mathbf{B}D^{-1}(\mathbf{B}\tpose
+    \tilde{q})\mathbf{B}\tpose\)
+    and
+    \(V(\tilde{q}) := P(\mathbf{B}\tpose \tilde{q})\).
+\end{thm}
+\begin{proof}
+    \textbf{TODO: prove this in a terse way.}
+\end{proof}
+Throughout the rest of this paper, we will assume the system we are studying is
+in simply actuated coordinates.
+
+\subsection{Virtual Nonholonomic Constraints}
+
+\begin{defn}
+    A \textit{virtual nonholonomic constraint} (VNHC) \textit{of order \(k\)} is a
+    relation \(h(q,p) = 0\) where \(h : \mathcal{Q}\times\R^n \rightarrow \R^k\) is
+    \(C^2\), \(\rank{\left[ dh_q,\, dh_p \right]} = k\) for all 
+    \((q,p) \in h\inv(0)\), and there exists a feedback controller \(\tau(q,p)\)
+    rendering the \textit{constraint manifold} \(\Gamma\) invariant,
+    where
+    \[
+        \Gamma = \left\{(q,p) \mid h(q,p) = 0, dh_q \dot{q} + dh_p \dot{p} = 0\right\}
+        .
+    \]
+\end{defn}
+
+From the definition of VNHCs, one finds that the constraint manifold \(\Gamma\)
+is a \(2(n-k)\)-dimensional closed embedded submanifold of 
+\(\mathcal{Q} \times \R^n\).
+
+%TODO: tie this together
+
+\begin{thm}\label{thm:vnhc-regularity}
+    A relation \(h(q,p) = 0\) for system ~\eqref{eqn:simply-actuated-hamiltonian}
+    is a regular VNHC of order \(k\) if and only if \(dh_{p_a} = 0\) 
+    and
+    \[
+        \rank{\left(dh_q \Minv(q) - 
+          dh_{p_u} (\Id{n-k} \otimes p\tpose)\nabla_{q_u}\Minv(q) 
+         \right)\simpleB} = k
+         ,
+    \]
+    everywhere on the constraint manifold \(\Gamma\).
+\end{thm}
+\begin{proof}
+    \textbf{TODO: prove this in a terse way}.
+\end{proof}
+
+The constrained dynamics for a general VNHC were derived by 
+Horn et. al \cite{nhvc_dynamic_walking}.
+In our case, we will be enforcing regular constraints where the actuators of the
+acrobot track a function of the unactuated variables. 
+Intuitively, the dynamics should be parameterized by \((q_u, p_u)\).
+Unfortunately, \(\dot{q}_u\) depends on \(p_a\), and for general systems one
+cannot solve explicitly for \(p_a\) in terms of \((q_u,p_u)\). 
+This is because the \(\dot{p}\) dynamics contains the coupling term 
+\((\Id{n} \otimes p\tpose)\nabla_{q_u}M(q)p\).
+
+We now introduce an assumption so we can solve explicitly for the constrained
+dynamics.
+
+\begin{assm}\label{assm:inertially-actuated}
+    The Hamiltonian system has an inertia matrix that does not depend on the
+    unactuated coordinates:
+    \[
+        \nabla_{q_u}M(q) = \Zmat{n(n-k) \times n}
+        .
+    \]
+\end{assm}
+
+\begin{thm}\label{thm:zero-dynamics}
+    Let \(\mathcal{H}\) be a mechanical system in simply actuated
+    coordinates satisfying Assumption \ref{assm:inertially-actuated}. 
+    Let \(h(q,p_u) = 0\) be a regular VNHC of order \(k\) with constraint
+    manifold \(\Gamma\). Suppose that on \(\Gamma\) one can solve for
+    \(q_a\) as a function \(q_a = f(q_u,p_u)\).
+    Then the constrained dynamics are given by
+    \begin{equation}\label{eqn:qpu-dynamics}
+        \left.\begin{aligned}
+                \dot{q}_u &= \begin{bmatrix}
+                    \Id{(n-k)} & \Zmat{(n-k) \times k}
+                \end{bmatrix}\Minv(q)p \\
+            \dot{p}_u &= -\nabla_{q_u}V(q) \\
+            \end{aligned}{}\right|_{\begin{array}{c}
+                q_a = f(q_u,p_u) \\ 
+                p_a = g(q_u,p_u) \\
+            \end{array}}
+            ,
+    \end{equation}
+    where
+    \begin{equation}\label{eqn:g-qpu}
+    \begin{aligned}
+        &g(q_u,p_u) := \\
+           &\left(dh_q \Minv(q) \simpleB \right)\inv 
+               \\
+           &\left.\left(dh_{p_u} \nabla_{q_u}V(q) - dh_q \Minv(q)
+        \begin{bmatrix}
+            I_{n-k} \\
+            \Zmat{k \times (n-k)} \\
+    \end{bmatrix} p_u\right)\right|_{qa = f(q_u,p_u)}
+        .
+    \end{aligned}
+    \end{equation}
+\end{thm}
+\begin{proof}
+    This is a special case of Theorem 1 in \cite{nhvc_dynamic_walking} using
+    \(\theta_1 = q_u\) and \(\theta_2 = p_u\).
+    The proof is ommitted.
+\end{proof}
 

### 2021-02-22 latex/paper/main.tex
Edits to the VNHC section

@@
 \subsection{Notation}
-We use the following notation in this paper. 
-For \(T > 0\), the set of real numbers modulo is denote \(\Rt{T}\), with
+We use the following notation.
+For \(T > 0\), the set of real numbers modulo \(T\) is denoted \(\Rt{T}\), with
 \(\Rt{\infty} := \R\).
@@
 \(T_i = 2\pi\) if \(q_i\) is an angle and \(T_i = \infty\) if \(q_i\) is a
-length. The corresponding velocities are 
+displacement. The corresponding velocities are 
 \(\dot{q} = (\dot{q}_1,\ldots,\dot{q}_n) \in \R^n\).
@@
 \(\mathcal{L}(q,\dot{q}) = 1/2 \dot{q}^T D(q) \dot{q} - P(q)\),
-where \(P : \mathcal{Q} \rightarrow \mathbb{R}\) is the potential energy, and
-the inertia matrix \(D : \mathcal{Q} \rightarrow \mathbb{R}^{n \times n}\)
+where the potential energy 
+\(P : \mathcal{Q} \rightarrow \mathbb{R}\) 
+is smooth, and the inertia matrix 
+\(D : \mathcal{Q} \rightarrow \mathbb{R}^{n \times n}\)
 is positive definite for all \(q \in \mathcal{Q}\).
-By defining the \textit{conjugate of momentum} 
-\(p_i := \pdiff{\mathcal{L}}{\dot{q}_i}\) so that the vector 
-\(p = (p_1,\ldots,p_n) \in \R^n\) is given by \(p = D(q)\dot{q}\),
-we find the \textit{Hamiltonian} of the system in \((q,p)\) coordinates
+Define the \textit{conjugate of momentum} 
+\(p_i := \partial\mathcal{L}/\partial\dot{q}_i\) so that the vector 
+\(p = (p_1,\ldots,p_n) \in \R^n\) is given by \(p = D(q)\dot{q}\).
+The \textit{Hamiltonian} of the system in \((q,p)\) coordinates
 is\cite{landau_mechanics}
@@
 \end{equation}
-The dynamics of the system in Hamiltonian form are
+with dynamics
 \begin{equation}\label{eqn:hamiltonian-eom-general}
@@
 particular input force \(\tau_i\) will affect the dynamics of the system. 
-The following definition provides a class of systems where the effect of input
-forces is made explicit.
-
-\begin{defn}
-    Let \(\mathcal{H}\) be an \(n\)-DOF Hamiltonian system 
-    with \(k \leq n\) actuators. 
-    A set of canonical coordinates \((q,p)\) for this system
-    are said to be \textit{simply actuated coordinates} if the
-    input matrix \(B(q) \in \R^{n \times k}\) is of the form
-    \[
-        B(q) = \simpleB    
-        .
-    \]
-
-\end{defn}
-
-The first \((n-k)\) coordinates, labelled \(q_u\), are called the
-\textit{unactuated coordinates}. The remaining \(k\) coordinates, labelled
-\(q_a\), are called the \textit{actuated coordinates}. When grouping them
-together, we will always put them in the order \((q_u, q_a)\) to fit with 
-the definition. 
-The corresponding \((p_u, p_a)\) are called the \textit{unactuated} and 
-\textit{actuated momenta}, respectively.
-
-Under the following assumptions on the input matrix, we will show that there is
-a canonical transformation of ~\eqref{eqn:hamiltonian} into simply actuated
-coordinates.
+Using the following assumptions on the input matrix, we will show there is
+a canonical coordinate transformation of ~\eqref{eqn:hamiltonian} 
+which makes explicit the effect of input forces.
 
@@
 
-In light of Lemma \ref{lemma:B-orthogonal}, there is no loss of generality
-by making the following assumption.
-\begin{assm}\label{assm:B-orthogonal}
-    Assume that the input matrix \(B\) is
-    left semi-orthogonal, i.e.~\(B\tpose B = I_k\). 
-\end{assm}
-
+In light of Lemma \ref{lemma:B-orthogonal}, there is no loss of generality in
+assuming that the input matrix is left semi-orthogonal.
 Let now \(\mathbf{B} \in \R^{n\times n}\) be the following matrix:
@@
 orthogonal.
-\begin{proof}
-\[
-    \mathbf{B}\mathbf{B}\tpose = 
-    \begin{bmatrix}
-        B^\perp (B^\perp)\tpose & B^\perp B \\
-        (B^\perp B)\tpose & B\tpose B
-    \end{bmatrix} = \Id{n}
-    .
-\]
-Hence, \(\mathbf{B}\) is invertible with \(\mathbf{B}\inv = \mathbf{B}\tpose\).
-\end{proof}
-
-The following theorem shows that \(\mathbf{B}\) provides a canonical
-transformation into simply actuated coordinates, so that only the actuated momenta
-are affected by the input forces.
 
 \begin{thm}\label{thm:simply-actuated}
-    Under Assumptions \ref{assm:B-const},\ref{assm:B-perp}, and
-    \ref{assm:B-orthogonal}, the Hamiltonian system ~\eqref{eqn:hamiltonian}
-    has simply actuated canonical coordinates 
-    \(\left(\tilde{q} = \mathbf{B}q, \tilde{p} = \mathbf{B}p\right)\). 
+    Take the Hamiltonian system ~\eqref{eqn:hamiltonian}.
+    Under Assumptions \ref{assm:B-const} and \ref{assm:B-perp},
+    the coordinate transformation
+    \(\left(\tilde{q} = \mathbf{B}q, \tilde{p} = \mathbf{B}p\right)\)
+    is a canonical transformation.
     The resulting dynamics are given by 
@@
 \end{proof}
+These new coordinates are called \textit{simply actuated coordinates}.
+The first \((n-k)\) coordinates of \(\tilde{q}\), labelled \(q_u\), are called
+the \textit{unactuated coordinates}. 
+The remaining \(k\) coordinates, labelled \(q_a\), are called the
+\textit{actuated coordinates}.
+The corresponding \((p_u, p_a)\) are called the \textit{unactuated} and 
+\textit{actuated momenta}, respectively.
+
 Throughout the rest of this paper, we will assume the system we are studying is
@@
 
-\begin{defn}
+\begin{defn}\label{defn:vnhc}
     A \textit{virtual nonholonomic constraint} (VNHC) \textit{of order \(k\)} is a
@@
 
-From the definition of VNHCs, one finds that the constraint manifold \(\Gamma\)
+Definition \ref{defn:vnhc} implies that the constraint manifold \(\Gamma\)
 is a \(2(n-k)\)-dimensional closed embedded submanifold of 
@@
 
+Define the error output \(e = h(q,p)\) 
 %TODO: tie this together
@@
 \begin{thm}\label{thm:vnhc-regularity}
-    A relation \(h(q,p) = 0\) for system ~\eqref{eqn:simply-actuated-hamiltonian}
+    A relation \(h(q,p) = 0\) for system ~\eqref{eqn:simple-hamiltonian}
     is a regular VNHC of order \(k\) if and only if \(dh_{p_a} = 0\) 

### 2021-02-22 latex/paper/main.tex
First pass of the simply actuated section
This commit contains the first fully written intro to simply actuated
hamiltonian systems, including a terse proof that the transformation we
define is indeed canonical.

@@
 \DeclareMathOperator{\Sign}{sgn}
+\DeclareMathOperator{\Diag}{diag}
 \newcommand*{\rank}[1]{\Rank\left(#1\right)}
 \newcommand*{\sign}[1]{\Sign\left(#1\right)}
+\newcommand*{\diag}[1]{\Diag\left(#1\right)}
 
@@
 \subsection{Notation}
-We use the following notation.
+We use the following notation in this article.
+For \(A \in \R^{n\times m}\) and \(B \in \R^{p \times m}\),
+\((A,B)\tpose \in \R^{(n+p)\times m}\) is the matrix obtained by stacking \(A\)
+on top of \(B\). 
 For \(T > 0\), the set of real numbers modulo \(T\) is denoted \(\Rt{T}\), with
@@
 The gradient of the matrix-valued function 
-\(A : \R^m \rightarrow \R^{n\times n}\) is the matrix of stacked partial
+\(A : \R^m \rightarrow \R^{n\times n}\) is the block matrix of stacked partial
 derivatives, 
@@
 matrix
-\[
+\begin{equation}\label{eqn:kronprod}
     A \otimes B = \begin{bmatrix}
@@
     .
-\]
+\end{equation}
+The Poisson bracket \cite{landau_mechanics} between the functions
+\(f(q,p)\) and \(g(q,p)\) is
+\begin{equation}\label{eqn:poisson-bracket}
+    [f,g] := \sum \limits_{i=1}^n \pdiff{f}{p_i}\pdiff{g}{q_i} - 
+        \pdiff{f}{q_i}\pdiff{g}{p_i}
+    .
+\end{equation}
+Finally, the Kronecker delta \(\delta_i^j = 1\) if \(i = j\) and \(0\)
+otherwise.
 
@@
 \begin{proof}
-    \textbf{TODO: prove this in a terse way}.
+    Since \(B\) is constant and full rank, it has a singular value decomposition 
+    \(B = U\tpose \Sigma V\) where 
+    \(\Sigma = (\diag{\sigma_1,\ldots,\sigma_k}, \Zmat{(n-k)\times k})\tpose\),
+    \(\sigma_i > 0\), and \(U \in R^{n \times n}\),
+    \(V \in \R^{k \times k}\) are unitary matrices \cite{calculating_svd}.
+    Defining \(T = \diag{1/\sigma_1^2,\ldots,1/\sigma_n^2}\) and assigning the
+    regular feedback transformation \(\tau = V T \hat{\tau}\) yields a new input
+    matrix \(\hat{B} = B V T\) for \(\hat{\tau}\) such that
+    \(\hat{B}\tpose \hat{B} = T\tpose \Sigma\tpose \Sigma T = \Id{k}\).
 \end{proof}
@@
 \begin{proof}
-    \textbf{TODO: prove this in a terse way.}
+    Since \(\mathbf{B}\) is constant, this transformation satisfies
+    \(\pdiff{\tilde{q}_i}{p_m} = \pdiff{\tilde{p}_i}{q_m} = 0\) for all 
+    \(i,m \in \{1,\ldots,n\}\).
+    Hence, \([\tilde{q}_i, \tilde{q}_j] = [\tilde{p}_i,\tilde{p}_j] = 0\).
+    Then, since \(\mathbf{B}\) is orthogonal, 
+    \((\mathbf{B}_i)\tpose (\mathbf{B}\tpose)_j = \delta_i^j\).
+    Hence, the Poisson bracket between \(\tilde{p}_i\) and \(\tilde{q}_j\) is 
+    \begin{align*}
+        [\tilde{p}_i, \tilde{q}_j] &= \sum\limits_{m=1}^n A_{i,m}A_{j,m} 
+                                   = (\mathbf{B}_i)\tpose (\mathbf{B}\tpose)_j
+                                   = \delta_i^j
+                                   .
+    \end{align*}
+    By (45.10) in \cite{landau_mechanics}, this is a canonical transformation,
+    so the new Hamiltonian is
+    \(\mathcal{H}(\mathbf{B}\tpose q, \mathbf{B}\tpose p)\).
+    Finally, since \(\dot{\tilde{p}} = \mathbf{B}\tpose \dot{p}\), the input
+    matrix for the system in \((\tilde{q},\tilde{p})\) coordinates is
+    \[
+        \mathbf{B}B = \simpleB
+        ,
+    \]
+    which proves the theorem.
 \end{proof}
-These new coordinates are called \textit{simply actuated coordinates}.
-The first \((n-k)\) coordinates of \(\tilde{q}\), labelled \(q_u\), are called
-the \textit{unactuated coordinates}. 
-The remaining \(k\) coordinates, labelled \(q_a\), are called the
+These new coordinates are called 
+\textit{simply actuated coordinates}.
+The first \((n-k)\) coordinates, labelled \(q_u\), are named
+the \textit{unactuated coordinates}; 
+the remaining \(k\) coordinates, labelled \(q_a\), are the
 \textit{actuated coordinates}.
-The corresponding \((p_u, p_a)\) are called the \textit{unactuated} and 
+The corresponding \((p_u, p_a)\) are the \textit{unactuated} and 
 \textit{actuated momenta}, respectively.
 
-Throughout the rest of this paper, we will assume the system we are studying is
-in simply actuated coordinates.
+Throughout the rest of this article, we will assume the system of inquiry is
+expressed in simply actuated coordinates.
 

### 2021-02-23 latex/paper/main.tex
First draft of the VNHC proof in terse mode

@@
            \dot{\tilde{p}} = -\frac{1}{2} (\Id{n} \otimes \tilde{p}\tpose)
-           \nabla_{\tilde{q}} \Minv(\tilde{q}) \tilde{p}
-           - \nabla_{\tilde{q}} V(\tilde{q}) + \simpleB \tau
+           \nabla_{\tilde{q}} \Minv(\tilde{q}) \tilde{p} \\
+           \phantom{---} - \nabla_{\tilde{q}} V(\tilde{q}) + \simpleB \tau
             ,
@@
 
-Throughout the rest of this article, we will assume the system of inquiry is
-expressed in simply actuated coordinates.
+\subsection{Virtual Nonholonomic Constraints}
 
-\subsection{Virtual Nonholonomic Constraints}
+In this section, we will take the system of inquiry to be a Hamiltonian
+mechanical system in simply actuated coordinates.
+We wish to find a controller which constraints the system according to a
+function of both the configuration \(q \in \mathcal{Q}\) and the conjugate of
+momentum \(p \in \R^n\).
 
@@
 
-Definition \ref{defn:vnhc} implies that the constraint manifold \(\Gamma\)
-is a \(2(n-k)\)-dimensional closed embedded submanifold of 
-\(\mathcal{Q} \times \R^n\).
+The nonholonomic constraint \(h(q,p)\) is \textit{virtual} because it is
+enforced by the controller \(\tau(q,p)\), rather than arising from the physics
+of the system.
+The constraint manifold \(\Gamma\) is a \(2(n-k)\)-dimensional
+closed embedded submanifold of \(\mathcal{Q} \times \R^n\).
+Hence, the VNHC reduces the dimensionality of the dynamics from \(2n\) to
+\(2(n-k)\). 
+In particular, if \(k = (n-1)\), the constraint manifold is always
+2-dimensional.
 
-Define the error output \(e = h(q,p)\) 
-%TODO: tie this together
+We often want to stabilize the constraint within a neighbourhood of \(\Gamma\).
+To see when this is possible, let us define the error output \(e = h(q,p)\).
+Notice that if any component of \(e_i\) has relative degree 1, we may not be able
+to stabilize \(\Gamma\) -- we can always guarantee \(e_i \to 0\), but not
+necessarily \(\dot{e}_i \to 0\).
+It is for this reason that we define the following special type of VNHC.
+
+\begin{defn}
+    A VNHC \(h(q,p) = 0\) of order \(k\) is \textit{regular} if the output 
+    \(e = h(q,p)\) is of relative degree \(\{2,2.\ldots,2\}\) everywhere on the
+    constraint manifold \(\Gamma\).
+\end{defn}
+
+The authors of
+\cite{nhvc_dynamic_walking,hybrid_zero_dynamics_bipedal_nhvcs,nhvc_incline_walking}
+observed that a relation which uses only the unactuated conjugate of momentum
+will have vector relative degree \(\{2,\ldots,2\}\).
+In fact, this is a hard requirement: any VNHC which depends on the actuated
+momentum cannot be regular. 
+Before providing a characterization of regularity which proves this statement,
+we define the following matrices as shorthand:
+\begin{align}
+    \mathcal{M}(q,p) &:= (\Id{n-k} \otimes p\tpose)\nabla_{q_u}\Minv(q) 
+        ,\\
+    \mathcal{D}(q,p_u) &:= dh_q(q,p_u) \Minv(q) 
+    .
+\end{align}
 
@@
     \[
-        \rank{\left(dh_q \Minv(q) - 
-          dh_{p_u} (\Id{n-k} \otimes p\tpose)\nabla_{q_u}\Minv(q) 
-         \right)\simpleB} = k
+        \rank{\left(\mathcal{D}(q,p_u) - dh_{p_u}\mathcal{M}(q,p)\right)\simpleB} = k
          ,
@@
 \begin{proof}
-    \textbf{TODO: prove this in a terse way}.
+    Let \(e = h(q,p) \in \R^k\).
+    If \(dh_{p_a} \neq \Zmat{k\times k}\) for some \((q,p) \in \Gamma\), 
+    then \(\tau\) appears in \(\dot{e}\) and the VNHC is not of relative degree
+    \(\{2,\ldots,2\}\). Suppose now that \(dh_{p_a} = \Zmat{k\times k}\).
+    Then 
+    \[
+        \dot{e} = \mathcal{D}(q,p_u)p - 
+        dh_{p_u}\left(\frac{1}{2}\mathcal{M}(q,p)p + \nabla_{q_u}V(q)\right)
+        .
+    \]
+    Taking one further derivative provides
+    \begin{align*}
+        \ddot{e} &= (\star) - 
+        dh_{p_u}\left(\frac{1}{2}\diff{}{t}\left(\mathcal{M}(q,p)p\right)\right) 
+        \\
+                 &+\mathcal{D}(q,p_u)\simpleB \tau
+    \end{align*}
+    where \((\star)\) is a continuous function of \(q\) and \(p\).
+    One can show that
+    \begin{align*}
+        dh_{p_u}\left(\frac{1}{2}\diff{}{t}\left(\mathcal{M}(q,p)p\right)\right)
+        &= (\star) +\\
+        &dh_{p_u} \mathcal{M}(q,p)\simpleB \tau
+        .
+    \end{align*}
+    Hence,
+    \[
+       \ddot{e} = (\star) +
+       \left(\mathcal{D}(q,p_u) - dh_{p_u}\mathcal{M}(q,p)\right) \simpleB \tau
+        ,
+    \]
+    which we write as \( \ddot{e} = E(q,p) + H(q,p)\tau\) for appropriate \(E\)
+    and \(H\).
+    From the definition of regularity, the VNHC \(h\) is regular 
+    when \(e\) is of relative degree \(\{2,\ldots,2\}\), which is true 
+    if and only if the matrix premultiplying \(\tau\) is nonsingular, and hence
+    that \(H\) is invertible. This proves the theorem.
 \end{proof}
@@
     The Hamiltonian system has an inertia matrix that does not depend on the
-    unactuated coordinates:
-    \[
-        \nabla_{q_u}M(q) = \Zmat{n(n-k) \times n}
-        .
-    \]
+    unactuated coordinates, so that 
+    \(\nabla_{q_u}M(q) = \Zmat{n(n-k) \times n}\).
 \end{assm}
@@
         &g(q_u,p_u) := \\
-           &\left(dh_q \Minv(q) \simpleB \right)\inv 
+        &\left(\mathcal{D}(q,p_u) \simpleB \right)\inv 
                \\
-           &\left.\left(dh_{p_u} \nabla_{q_u}V(q) - dh_q \Minv(q)
+        &\left.\left(dh_{p_u} \nabla_{q_u}V(q) - \mathcal{D}(q,p_u)
         \begin{bmatrix}

### 2021-03-01 latex/paper/main.tex
Slight update of notation for matrix stacking

@@
 For \(A \in \R^{n\times m}\) and \(B \in \R^{p \times m}\),
-\((A,B)\tpose \in \R^{(n+p)\times m}\) is the matrix obtained by stacking \(A\)
+define \([A;B] \in \R^{(n+p)\times m}\) as the matrix obtained by stacking \(A\)
 on top of \(B\). 
@@
 derivatives, 
-\(\nabla_xA := (\pdiff{A}{x_1},\ldots,\pdiff{A}{x_m})\tpose \in \R^{nm \times n}\).
+\(\nabla_xA := [\pdiff{A}{x_1};\ldots;\pdiff{A}{x_m}] \in \R^{nm \times n}\).
 Given two matrices \(A \in \R^{n \times m}\) and \(B \in \R^{r \times s}\), the
@@
     \(B = U\tpose \Sigma V\) where 
-    \(\Sigma = (\diag{\sigma_1,\ldots,\sigma_k}, \Zmat{(n-k)\times k})\tpose\),
+    \(\Sigma = [\diag{\sigma_1,\ldots,\sigma_k}; \Zmat{(n-k)\times k}]\),
     \(\sigma_i > 0\), and \(U \in R^{n \times n}\),
     \(V \in \R^{k \times k}\) are unitary matrices \cite{calculating_svd}.
-    Defining \(T = \diag{1/\sigma_1^2,\ldots,1/\sigma_n^2}\) and assigning the
+    Defining \(T = \diag{1/\sigma_1^2,\ldots,1/\sigma_k^2}\) and assigning the
     regular feedback transformation \(\tau = V T \hat{\tau}\) yields a new input
@@
 assuming that the input matrix is left semi-orthogonal.
-Let now \(\mathbf{B} \in \R^{n\times n}\) be the following matrix:
-\[
-    \mathbf{B} = 
-    \begin{bmatrix}
-        B^\perp \\
-        B\tpose \\
-    \end{bmatrix}
-    .
-\]
+Let now \(\mathbf{B} = [B^\perp; B^\tpose]\).
 Since \(B^\perp\) is a left annihilator of \(B\) and both \(B^\perp\) and
@@
     where 
-    \(\Minv(\tilde{q}) := \mathbf{B}D^{-1}(\mathbf{B}\tpose
-    \tilde{q})\mathbf{B}\tpose\)
+    \(\Minv(\tilde{q}) := 
+    \mathbf{B}D^{-1}(\mathbf{B}\tpose \tilde{q})\mathbf{B}\tpose\)
     and
@@
     \(i,m \in \{1,\ldots,n\}\).
-    Hence, \([\tilde{q}_i, \tilde{q}_j] = [\tilde{p}_i,\tilde{p}_j] = 0\).
+    This implies the Poisson brackets \([\tilde{q}_i, \tilde{q}_j]\)
+    and \([\tilde{p}_i,\tilde{p}_j]\) are both zero.
     Then, since \(\mathbf{B}\) is orthogonal, 
     \((\mathbf{B}_i)\tpose (\mathbf{B}\tpose)_j = \delta_i^j\).
-    Hence, the Poisson bracket between \(\tilde{p}_i\) and \(\tilde{q}_j\) is 
-    \begin{align*}
-        [\tilde{p}_i, \tilde{q}_j] &= \sum\limits_{m=1}^n A_{i,m}A_{j,m} 
+    Hence, 
+    \begin{equation*}
+        [\tilde{p}_i, \tilde{q}_j]
+            = \sum\limits_{m=1}^n \mathbf{B}_{i,m}\mathbf{B}_{j,m} 
             = (\mathbf{B}_i)\tpose (\mathbf{B}\tpose)_j
@@
             .
-    \end{align*}
-    By (45.10) in \cite{landau_mechanics}, this is a canonical transformation,
-    so the new Hamiltonian is
+    \end{equation*}
+    By (45.10) in \cite{landau_mechanics}, this is a canonical transformation
+    with new Hamiltonian
     \(\mathcal{H}(\mathbf{B}\tpose q, \mathbf{B}\tpose p)\).
@@
 
-The nonholonomic constraint \(h(q,p)\) is \textit{virtual} because it is
-enforced by the controller \(\tau(q,p)\), rather than arising from the physics
-of the system.
+The nonholonomic constraint \(h(q,p)\) is called \textit{virtual} because it is
+enforced by the controller \(\tau(q,p)\), rather than being a constraint which
+arises from the physics of the system.
 The constraint manifold \(\Gamma\) is a \(2(n-k)\)-dimensional
@@
 To see when this is possible, let us define the error output \(e = h(q,p)\).
-Notice that if any component of \(e_i\) has relative degree 1, we may not be able
+If any component of \(e_i\) has relative degree 1, we may not be able
 to stabilize \(\Gamma\) -- we can always guarantee \(e_i \to 0\), but not
@@
 \cite{nhvc_dynamic_walking,hybrid_zero_dynamics_bipedal_nhvcs,nhvc_incline_walking}
-observed that a relation which uses only the unactuated conjugate of momentum
-will have vector relative degree \(\{2,\ldots,2\}\).
-In fact, this is a hard requirement: any VNHC which depends on the actuated
-momentum cannot be regular. 
-Before providing a characterization of regularity which proves this statement,
+observed that only relations which use the unactuated conjugate of momentum
+have vector relative degree \(\{2,\ldots,2\}\).
+However, this is not enough to guarantee the
+
+We now provide a computational characterization of regularity
 we define the following matrices as shorthand:

### 2021-03-02 latex/paper/main.tex
First full draft of the VNHC section of the paper

@@
 assuming that the input matrix is left semi-orthogonal.
-Let now \(\mathbf{B} = [B^\perp; B^\tpose]\).
+Let now \(\mathbf{B} = [B^\perp; B\tpose]\).
 Since \(B^\perp\) is a left annihilator of \(B\) and both \(B^\perp\) and
@@
 \cite{nhvc_dynamic_walking,hybrid_zero_dynamics_bipedal_nhvcs,nhvc_incline_walking}
-observed that only relations which use the unactuated conjugate of momentum
-have vector relative degree \(\{2,\ldots,2\}\).
-However, this is not enough to guarantee the
+observed that relations which use only the unactuated conjugate of momentum
+often have vector relative degree \(\{2,\ldots,2\}\).
+Indeed, we shall now provide a characterization of regularity which shows that
+regular constraints cannot use the actuated momentum at all.
 
-We now provide a computational characterization of regularity
-we define the following matrices as shorthand:
+To ease notation in the rest of this section, we use the following shorthand:
 \begin{align}
@@
     \begin{align*}
-        \ddot{e} &= (\star) - 
+        \ddot{e} 
+            &= (\star) - 
             dh_{p_u}\left(\frac{1}{2}\diff{}{t}\left(\mathcal{M}(q,p)p\right)\right) 
                 \\
-                 &+\mathcal{D}(q,p_u)\simpleB \tau
+            &+ \mathcal{D}(q,p_u)\simpleB \tau
     \end{align*}
@@
 
-The constrained dynamics for a general VNHC were derived by 
-Horn et. al \cite{nhvc_dynamic_walking}.
-In our case, we will be enforcing regular constraints where the actuators of the
-acrobot track a function of the unactuated variables. 
-Intuitively, the dynamics should be parameterized by \((q_u, p_u)\).
+Under additional mild conditions (see \cite{vhcs_for_el_systems}), a regular VNHC of
+order \(k\) can be stabilized by the output-linearizing phase-feedback
+controller
+\begin{equation}
+    \tau(q,p) = -H\inv(q,p)\left(E(q,p) + k_p e + k_d \dot{e}\right)
+    ,
+\end{equation}
+where \(k_p, k_d > 0\) are control parameters which can be tuned on the
+resulting error dynamices \(\ddot{e} = -k_p e - k_d \dot{e}\).
+
+In Section \ref{sec:acrobot} we will enforce a regular constraint on the
+acrobot of the form \(h(q,p) = q_a - f(q_u,p_u)\), where the actuators track a
+function of the unactuated variables.
+Intuitively then, the constrained dynamics should be parameterized by \((q_u, p_u)\).
 Unfortunately, \(\dot{q}_u\) depends on \(p_a\), and for general systems one
-cannot solve explicitly for \(p_a\) in terms of \((q_u,p_u)\). 
-This is because the \(\dot{p}\) dynamics contains the coupling term 
+cannot solve explicitly for \(p_a\) in terms of \((q_u,p_u)\) because
+the \(\dot{p}\) dynamics contains the coupling term 
 \((\Id{n} \otimes p\tpose)\nabla_{q_u}M(q)p\).
 
-We now introduce an assumption so we can solve explicitly for the constrained
+We now introduce an assumption which allows us to solve for \(p_a\) as a
+function of \((q_u,p_u)\), which in turn allows us to find the constrained
 dynamics.
@@
 \begin{assm}\label{assm:inertially-actuated}
-    The Hamiltonian system has an inertia matrix that does not depend on the
-    unactuated coordinates, so that 
+The inertia matrix does not depend on the unactuated coordinates, so that 
 \(\nabla_{q_u}M(q) = \Zmat{n(n-k) \times n}\).
@@
 \begin{proof}
-    This is a special case of Theorem 1 in \cite{nhvc_dynamic_walking} using
-    \(\theta_1 = q_u\) and \(\theta_2 = p_u\).
-    The proof is ommitted.
+    Setting \(e = h(q,p_u)\) and using Assumption
+    \ref{assm:inertially-actuated}, we find that
+    \(\dot{e} = \mathcal{D}(q,p_u)p - dh_{p_u}\nabla_{q_u}V(q)\).
+    Notice that
+    \(\mathcal{D}(q,p_u)p = \mathcal{D}(q,p_u)[\Zmat{(n-k)\times k}; \Id{k}]p_a
+    + \mathcal{D}(q,p_u)[\Id{n-k};\Zmat{k \times (n-k)}] p_u\).
+    Since \(h(q,p_u)\) is regular, \(\mathcal{D}(q,p_u)\) is invertible.
+    Taking \(e = \dot{e} = 0\), solving for \(p_a\), and setting 
+    \(q_a = f(q_u,p_u)\) completes the proof.
 \end{proof}
+
+\textbf{Comparison with existing literature}: Horn et.al. provide constrained
+dynamics for VNHCs in \cite{nhvc_incline_walking}.
+In fact, our constrained dynamics \eqref{eqn:qpu-dynamics} coincide with their
+system (17) when one chooses the special case \(\theta_1 = q_u\) and 
+\(\theta_2 = p_u\).
+Other than this distinction, our assumptions are identical to theirs.
+In particular, their assumption \textbf{H2} is what we call
+regularity.
+The only real distinction is that our constrained dynamics
+are explicit functions of the Hamiltonian phase coordinates
+\((q_u,p_u)\).
+This representation will be beneficial when we apply the theory of VNHCs to the
+acrobot.
 

### 2021-03-02 latex/paper/main.tex
First full editing pass of simply actuated section

@@
 \subsection{Simply Actuated Hamiltonian Systems}
-Take an underactuated mechanical system modelled with generalized coordinates 
+Take a mechanical system modelled with generalized coordinates 
 \(q = (q_1, \ldots, q_n)\) on the configuration manifold
@@
 \(T_i = 2\pi\) if \(q_i\) is an angle and \(T_i = \infty\) if \(q_i\) is a
-displacement. The corresponding velocities are 
+displacement. The corresponding generalized velocities are 
 \(\dot{q} = (\dot{q}_1,\ldots,\dot{q}_n) \in \R^n\).
@@
 is positive definite for all \(q \in \mathcal{Q}\).
-Define the \textit{conjugate of momentum} 
+The \textit{conjugate of momentum} to \(q\) is the vector
 \(p_i := \partial\mathcal{L}/\partial\dot{q}_i\) so that the vector 
-\(p = (p_1,\ldots,p_n) \in \R^n\) is given by \(p = D(q)\dot{q}\).
-The \textit{Hamiltonian} of the system in \((q,p)\) coordinates
-is\cite{landau_mechanics}
+\(p = D(q)\dot{q} \in \R^n\).
+As per \cite{landau_mechanics}, 
+the \textit{Hamiltonian} of the system in \((q,p)\) coordinates
+is
 \begin{equation}\label{eqn:hamiltonian}
-    \mathcal{H}(q,p) = 1/2 p\tpose D(q) p + P(q).
+    \mathcal{H}(q,p) = 1/2 p\tpose D(q) p + P(q)
+    ,
 \end{equation}
@@
 \(q \in \mathcal{Q}\).
+If \(k < n\), we say the system is \textit{underactuated} with degree of
+underactuation \((n-k)\).
 
 It is easy to show using the matrix Kronecker product that
-\eqref{eqn:hamiltonian-eom-general} reduces to
+\eqref{eqn:hamiltonian-eom-general} expands to
 \begin{equation}\label{eqn:hamiltonian-full-dynamics}
@@
 
-Because \(\tau\) is transformed by \(B(q)\), it is not in general clear how any
-particular input force \(\tau_i\) will affect the dynamics of the system. 
-Using the following assumptions on the input matrix, we will show there is
-a canonical coordinate transformation of ~\eqref{eqn:hamiltonian} 
-which makes explicit the effect of input forces.
+Because \(\tau\) is transformed by \(B(q)\), it is not obvious how any
+particular input force \(\tau_i\) affects the system.
+To address this problem, we make the following assumptions.
 
@@
 \begin{assm}\label{assm:B-perp}
-    There exists a matrix 
+    There exists a right semi-orthogonal matrix 
     \(B^\perp \in \R^{(n-k)\times n}\)
-    which is right semi-orthogonal 
-    \(\left(B^\perp(B^\perp)\tpose = \Id{(n-k)}\right)\)
-    and which is a left-annihilator for \(B\). 
-    That is, \(B^\perp B = \Zmat{(n-k) \times k}\).
+    which is a left-annihilator for \(B\). 
 \end{assm}
 
-Assumption \ref{assm:B-perp} requires the rows of \(B^\perp\) to be unit vectors
+Note that Assumption \ref{assm:B-perp} requires the rows of \(B^\perp\) be unit vectors
 that are mutually orthogonal. 
@@
 \(A^0 \in \R^{1\times n}\) implies the left annihilator 
-\(B^\perp := A^0/\norm{A^0}\) will be a unit vector satisfying Assumption \ref{assm:B-perp}.
+\(B^\perp := A^0/\norm{A^0}\) will be a unit vector satisfying this assumption.
+
+The above assumptions allow us to define a
+canonical coordinate transformation of ~\eqref{eqn:hamiltonian} 
+which decouples the input forces.
+To define this transformation we will make use of the following lemma.
 
@@
     Since \(\mathbf{B}\) is constant, this transformation satisfies
-    \(\pdiff{\tilde{q}_i}{p_m} = \pdiff{\tilde{p}_i}{q_m} = 0\) for all 
+    \(\partial\tilde{q}_i/\partial p_m = \partial\tilde{p}_i/\partial q_m = 0\) for all 
     \(i,m \in \{1,\ldots,n\}\).
@@
     Then, since \(\mathbf{B}\) is orthogonal, 
-    \((\mathbf{B}_i)\tpose (\mathbf{B}\tpose)_j = \delta_i^j\).
-    Hence, 
-    \begin{equation*}
-        [\tilde{p}_i, \tilde{q}_j]
-            = \sum\limits_{m=1}^n \mathbf{B}_{i,m}\mathbf{B}_{j,m} 
-            = (\mathbf{B}_i)\tpose (\mathbf{B}\tpose)_j
-            = \delta_i^j
-            .
-    \end{equation*}
-    By (45.10) in \cite{landau_mechanics}, this is a canonical transformation
-    with new Hamiltonian
-    \(\mathcal{H}(\mathbf{B}\tpose q, \mathbf{B}\tpose p)\).
-    Finally, since \(\dot{\tilde{p}} = \mathbf{B}\tpose \dot{p}\), the input
+    \([\tilde{p}_i, \tilde{q}_j] = (\mathbf{B}_i)\tpose (\mathbf{B}\tpose)_j
+        = \delta_i^j\).
+    By (45.10) in \cite{landau_mechanics}, this transformation is canonical and
+    the new Hamiltonian is
+    \(\mathcal{H}(\mathbf{B}\tpose \tilde{q}, \mathbf{B}\tpose \tilde{p})\).
+    Finally, since \(\dot{\tilde{p}} = \mathbf{B} \dot{p}\), the input
     matrix for the system in \((\tilde{q},\tilde{p})\) coordinates is
-    \[
-        \mathbf{B}B = \simpleB
-        ,
-    \]
-    which proves the theorem.
+    \(\mathbf{B}B = [\Zmat{(n-k)\times k}; \Id{k}]\), which proves the theorem.
 \end{proof}
+
 These new coordinates are called 
 \textit{simply actuated coordinates}.
-The first \((n-k)\) coordinates, labelled \(q_u\), are named
+The first \((n-k)\) coordinates, labelled \(q_u\), are
 the \textit{unactuated coordinates}; 

### 2021-03-03 latex/paper/main.tex
First revision of the VNHC section, which is now down to 4 columns.

@@
 Take a mechanical system modelled with generalized coordinates 
-\(q = (q_1, \ldots, q_n)\) on the configuration manifold
+\(q = (q_1, \ldots, q_n)\) on a configuration manifold
 \(\mathcal{Q} = \Rt{T_1} \times \cdot \Rt{T_n}\), where
@@
 Suppose this system has Lagrangian
-\(\mathcal{L}(q,\dot{q}) = 1/2 \dot{q}^T D(q) \dot{q} - P(q)\),
+\(\mathcal{L}(q,\dot{q}) = 1/2~\dot{q}^T D(q) \dot{q} - P(q)\),
 where the potential energy 
@@
 \(D : \mathcal{Q} \rightarrow \mathbb{R}^{n \times n}\)
-is positive definite for all \(q \in \mathcal{Q}\).
+is smooth and positive definite for all \(q \in \mathcal{Q}\).
 The \textit{conjugate of momentum} to \(q\) is the vector
-\(p_i := \partial\mathcal{L}/\partial\dot{q}_i\) so that the vector 
-\(p = D(q)\dot{q} \in \R^n\).
+\(p := \partial\mathcal{L}/\partial\dot{q} = D(q) \dot{q} \in \R^n\).
 As per \cite{landau_mechanics}, 
@@
 \begin{equation}\label{eqn:hamiltonian}
-    \mathcal{H}(q,p) = 1/2 p\tpose D(q) p + P(q)
+    \mathcal{H}(q,p) = \frac{1}{2} p\tpose D\inv(q) p + P(q)
     ,
@@
     semi-orthogonal.  
-    That is, \(\hat{B}\tpose \hat{B} = \Id{k}\).
 \end{lemma}
@@
 assuming that the input matrix is left semi-orthogonal.
-Let now \(\mathbf{B} = [B^\perp; B\tpose]\).
+Now, let \(\mathbf{B} := [B^\perp; B\tpose]\).
 Since \(B^\perp\) is a left annihilator of \(B\) and both \(B^\perp\) and
-\(B\tpose\) are right semi-orthogonal, it is easy to show that \(\mathbf{B}\) is
+\(B\tpose\) are right semi-orthogonal, one can easily show that \(\mathbf{B}\) is
 orthogonal.
@@
     Since \(\mathbf{B}\) is constant, this transformation satisfies
-    \(\partial\tilde{q}_i/\partial p_m = \partial\tilde{p}_i/\partial q_m = 0\) for all 
-    \(i,m \in \{1,\ldots,n\}\).
+    \(\partial\tilde{q}_i/\partial p_j = \partial\tilde{p}_i/\partial q_j = 0\) for all 
+    \(i,j \in \{1,\ldots,n\}\).
     This implies the Poisson brackets \([\tilde{q}_i, \tilde{q}_j]\)
@@
 
-These new coordinates are called 
+We call these new coordinates 
 \textit{simply actuated coordinates}.
-The first \((n-k)\) coordinates, labelled \(q_u\), are
+The first \((n-k)\) configuration variables, labelled \(q_u\), are
 the \textit{unactuated coordinates}; 
-the remaining \(k\) coordinates, labelled \(q_a\), are the
+the remaining \(k\) configuration variables, labelled \(q_a\), are the
 \textit{actuated coordinates}.
@@
 
-In this section, we will take the system of inquiry to be a Hamiltonian
+The notion of an enforcable virtual holonomic constraint was first described by
+Griffin and Grizzle in \cite{nhvc_dynamic_walking}.
+Horn et. al later extended their results in
+\cite{hybrid_zero_dynamics_bipedal_nhvcs} and
+derived the dynamics of a constrained system.
+In this section, We will reformulate their work in the Hamiltonian framework
+because the results play nicely with unactuated and actuated momenta.
+For this reason, we take the system of inquiry to be a Hamiltonian
 mechanical system in simply actuated coordinates.
-We wish to find a controller which constraints the system according to a
-function of both the configuration \(q \in \mathcal{Q}\) and the conjugate of
-momentum \(p \in \R^n\).
 
@@
     where
-    \[
+    \begin{equation}
         \Gamma = \left\{(q,p) \mid h(q,p) = 0, dh_q \dot{q} + dh_p \dot{p} = 0\right\}
         .
-    \]
+    \end{equation}
 \end{defn}
 
-The nonholonomic constraint \(h(q,p)\) is called \textit{virtual} because it is
-enforced by the controller \(\tau(q,p)\), rather than being a constraint which
-arises from the physics of the system.
-The constraint manifold \(\Gamma\) is a \(2(n-k)\)-dimensional
+The constraint manifold is a \(2(n-k)\)-dimensional
 closed embedded submanifold of \(\mathcal{Q} \times \R^n\).
-Hence, the VNHC reduces the dimensionality of the dynamics from \(2n\) to
-\(2(n-k)\). 
-In particular, if \(k = (n-1)\), the constraint manifold is always
-2-dimensional.
+A VNHC thereby allows us to study a reduced-order model of the system: it reduces
+the original \(2n\) differential equations to \(2(n-k)\) equations.
+In particular, if \(k = (n-1)\), the constraint manifold is \textit{always}
+2-dimensional and its dynamics can easily be plotted on a plane.
 
-We often want to stabilize the constraint within a neighbourhood of \(\Gamma\).
+We often want to stabilize a constraint within some neighbourhood of \(\Gamma\).
 To see when this is possible, let us define the error output \(e = h(q,p)\).
@@
 The authors of
-\cite{nhvc_dynamic_walking,hybrid_zero_dynamics_bipedal_nhvcs,nhvc_incline_walking}
+\cite{nhvc_dynamic_walking,hybrid_zero_dynamics_bipedal_nhvcs}
 observed that relations which use only the unactuated conjugate of momentum
@@
 \begin{align}
+    \mathcal{A}(q,p_u) &:= dh_q(q,p_u) \Minv(q) 
+        ,\\
     \mathcal{M}(q,p) &:= (\Id{n-k} \otimes p\tpose)\nabla_{q_u}\Minv(q) 
-        ,\\
-    \mathcal{D}(q,p_u) &:= dh_q(q,p_u) \Minv(q) 
     .
@@
     A relation \(h(q,p) = 0\) for system ~\eqref{eqn:simple-hamiltonian}
-    is a regular VNHC of order \(k\) if and only if \(dh_{p_a} = 0\) 
+    is a regular VNHC of order \(k\) if and only if 
+    \(dh_{p_a} = \Zmat{k \times k}\) 
     and
     \[
-        \rank{\left(\mathcal{D}(q,p_u) - dh_{p_u}\mathcal{M}(q,p)\right)\simpleB} = k
+        \rank{\left(\mathcal{A}(q,p_u) - dh_{p_u}\mathcal{M}(q,p)\right)\simpleB} = k
          ,
@@
     Then 
-    \[
-        \dot{e} = \mathcal{D}(q,p_u)p - 
-        dh_{p_u}\left(\frac{1}{2}\mathcal{M}(q,p)p + \nabla_{q_u}V(q)\right)
-        .
-    \]
+    \(\dot{e} = \mathcal{A}(q,p_u)p - 
+     dh_{p_u}\left(1/2~\mathcal{M}(q,p)p + \nabla_{q_u}V(q)\right)\).
     Taking one further derivative provides
-    \begin{align*}
-        \ddot{e} 
-            &= (\star) - 
-            dh_{p_u}\left(\frac{1}{2}\diff{}{t}\left(\mathcal{M}(q,p)p\right)\right) 
-                \\
-            &+ \mathcal{D}(q,p_u)\simpleB \tau
-    \end{align*}
+    \( \ddot{e} = (\star) - 
+        dh_{p_u}\left(1/2~d/dt~\left(\mathcal{M}(q,p)p\right)\right) 
+        + \mathcal{A}(q,p_u)[\Zmat{(n-k)\times k};\Id{k}] \tau\),
     where \((\star)\) is a continuous function of \(q\) and \(p\).
-    One can show that
-    \begin{align*}
-        dh_{p_u}\left(\frac{1}{2}\diff{}{t}\left(\mathcal{M}(q,p)p\right)\right)
-        &= (\star) +\\
-        &dh_{p_u} \mathcal{M}(q,p)\simpleB \tau
-        .
-    \end{align*}
+    One can further show that
+    \(dh_{p_u}\left(1/2~d/dt~\left(\mathcal{M}(q,p)p\right)\right)
+        = (\star) + dh_{p_u}\mathcal{M}(q,p)[\Zmat{(n-k)\times k};
+        \Id{k}]\tau\).
     Hence,
@@
        \ddot{e} = (\star) +
-       \left(\mathcal{D}(q,p_u) - dh_{p_u}\mathcal{M}(q,p)\right) \simpleB \tau
+       \left(\mathcal{A}(q,p_u) - dh_{p_u}\mathcal{M}(q,p)\right) \simpleB \tau
         ,
@@
         \left.\begin{aligned}
-                \dot{q}_u &= \begin{bmatrix}
-                    \Id{(n-k)} & \Zmat{(n-k) \times k}
-                \end{bmatrix}\Minv(q)p \\
+                \dot{q}_u &= 
+                \left[\Id{(n-k)} ~ \Zmat{(n-k) \times k}\right]
+                \Minv(q)p \\
             \dot{p}_u &= -\nabla_{q_u}V(q) \\
@@
     where
-    \begin{equation}\label{eqn:g-qpu}
-    \begin{aligned}
-        &g(q_u,p_u) := \\
-        &\left(\mathcal{D}(q,p_u) \simpleB \right)\inv 
+    \begin{align}\label{eqn:g-qpu}
+        &g(q_u,p_u) := 
+        \left(\mathcal{A}(q,p_u)[\Zmat{(n-k)\times k};\Id{k}] \right)\inv 
+        (dh_{p_u} \nabla_{q_u}V(q) \nonumber
         \\
-        &\left.\left(dh_{p_u} \nabla_{q_u}V(q) - \mathcal{D}(q,p_u)
-        \begin{bmatrix}
-            I_{n-k} \\
-            \Zmat{k \times (n-k)} \\
-    \end{bmatrix} p_u\right)\right|_{qa = f(q_u,p_u)}
+        &- \mathcal{A}(q,p_u)[\Id{(n-k)};\Zmat{k \times(n-k)}]p_u
+        \left.)\right|_{qa = f(q_u,p_u)}
         .
-    \end{aligned}
-    \end{equation}
+    \end{align}
 \end{thm}
@@
     \ref{assm:inertially-actuated}, we find that
-    \(\dot{e} = \mathcal{D}(q,p_u)p - dh_{p_u}\nabla_{q_u}V(q)\).
+    \(\dot{e} = \mathcal{A}(q,p_u)p - dh_{p_u}\nabla_{q_u}V(q)\).
     Notice that
-    \(\mathcal{D}(q,p_u)p = \mathcal{D}(q,p_u)[\Zmat{(n-k)\times k}; \Id{k}]p_a
-    + \mathcal{D}(q,p_u)[\Id{n-k};\Zmat{k \times (n-k)}] p_u\).
-    Since \(h(q,p_u)\) is regular, \(\mathcal{D}(q,p_u)\) is invertible.
+    \(\mathcal{A}(q,p_u)p = \mathcal{A}(q,p_u)[\Zmat{(n-k)\times k}; \Id{k}]p_a
+    + \mathcal{A}(q,p_u)[\Id{n-k};\Zmat{k \times (n-k)}] p_u\).
+    Since \(h(q,p_u)\) is regular, \(\mathcal{A}(q,p_u)\) is invertible.
     Taking \(e = \dot{e} = 0\), solving for \(p_a\), and setting 

### 2021-03-04 latex/paper/main.tex
Final edit of VNHC section before sending it to Manfredi

@@
 \(\Rt{\infty} := \R\).
-The gradient of the matrix-valued function 
+The gradient of a matrix-valued function 
 \(A : \R^m \rightarrow \R^{n\times n}\) is the block matrix of stacked partial
@@
 \section{Theory of VNHCs}\label{sec:vnhc}
-TODO: Just like in the thesis, cover the theory of VNHCs. Justify our results compared
-to those of Horn. Make it terse but legible. Cover simply actuated systems,
-vnhcs, regularity, the theorems, etc.
 
@@
      \begin{cases}
-        \dot{q} = D\inv(q)p \\
+        \dot{q} = D\inv(q)p 
+        , \\
         \dot{p} = -\frac{1}{2} (\Id{n} \otimes p\tpose) \nabla_q D\inv(q) p
@@
 particular input force \(\tau_i\) affects the system.
-To address this problem, we make the following assumptions.
+As a first step in addressing this problem, we make the following assumptions.
 
@@
 \(B\tpose\) are right semi-orthogonal, one can easily show that \(\mathbf{B}\) is
-orthogonal.
+an orthogonal matrix.
 
@@
 
-The notion of an enforcable virtual holonomic constraint was first described by
-Griffin and Grizzle in \cite{nhvc_dynamic_walking}.
+The notion of a nonholonomic constraint which can be stabilized via state feedback
+was first described by Griffin and Grizzle in \cite{nhvc_dynamic_walking}.
 Horn et. al later extended their results in
 \cite{hybrid_zero_dynamics_bipedal_nhvcs} and
-derived the dynamics of a constrained system.
-In this section, We will reformulate their work in the Hamiltonian framework
-because the results play nicely with unactuated and actuated momenta.
+derived the dynamics of constrained systems.
+In this section we reformulate these ideas for the Hamiltonian framework,
+because the theory is cleaner when using unactuated and actuated momenta.
 For this reason, we take the system of inquiry to be a Hamiltonian
@@
 In particular, if \(k = (n-1)\), the constraint manifold is \textit{always}
-2-dimensional and its dynamics can easily be plotted on a plane.
+2-dimensional and its dynamics can be plotted on a plane.
 
@@
 the \(\dot{p}\) dynamics contains the coupling term 
-\((\Id{n} \otimes p\tpose)\nabla_{q_u}M(q)p\).
+\((\Id{n} \otimes p\tpose)\nabla_{q}M(q)p\).
 
@@
         &- \mathcal{A}(q,p_u)[\Id{(n-k)};\Zmat{k \times(n-k)}]p_u
-        \left.)\right|_{qa = f(q_u,p_u)}
+        \left.)\right|_{q_a = f(q_u,p_u)}
         .
@@
 
-\textbf{Comparison with existing literature}: Horn et.al. provide constrained
+\textbf{Comparison with existing literature}: Horn et.al. provide the constrained
 dynamics for VNHCs in \cite{nhvc_incline_walking}.
+Their assumption \textbf{H2} is what we call regularity, and our requirement
+that one can solve for \(q_a = f(q_u,p_u)\) on \(\Gamma\) implies their
+assumption \textbf{H3} holds true.
+The only real distinction between this section and their work
+is that our constrained dynamics are explicit functions of the Hamiltonian phase
+coordinates \((q_u,p_u)\).
 In fact, our constrained dynamics \eqref{eqn:qpu-dynamics} coincide with their
@@
 \(\theta_2 = p_u\).
-Other than this distinction, our assumptions are identical to theirs.
-In particular, their assumption \textbf{H2} is what we call
-regularity.
-The only real distinction is that our constrained dynamics
-are explicit functions of the Hamiltonian phase coordinates
-\((q_u,p_u)\).
-This representation will be beneficial when we apply the theory of VNHCs to the
-acrobot.
+This explicit representation will be beneficial when we apply the theory of
+VNHCs to the acrobot.
 
@@
 specialize the theory of VNHCs to the acrobot. consider constraints which depend
-only on pu, and show arctan vnhc and reduced dynamics. conclude with our
+only on pu, and show arctan vnhc and reduced dynamics. 
+give definition of energy injection/dissipation. conclude with our
 theorem. Make this section short and sweet.

### 2021-03-04 latex/paper/main.tex
Copied over acrobot's intertia, potential, etc to SecIV

@@
 
+TODO: after writing S2, re-introduce what we want to do with acrobot (energy
+injection for giant stabilization)
+
+For our model of the acrobot, we assume the acrobot iscomprised of two massless
+rods of equal length \(l\), with equal point masses \(m\) at the tips.
+We call this a \textit{simple} acrobot, which is displayed in Figure
+\ref{fig:simple-acrobot-model}.
+We will also ignore any frictional forces at both the hip and shoulder joints. 
+Finally, it is important to note that a real gymnast cannot swing their legs in
+full circles, though they are usually flexible enough to raise them parallel to
+the floor; 
+for this reason, we assume that \(q_a \in [-Q_a, Q_a]\) where 
+\(Q_a \in [\frac{\pi}{2}, \pi[\).
+
+\begin{figure}
+    \centering
+    %\includegraphics[width=0.5\textwidth]{images/simple_acrobot_model}
+    \caption{A simple acrobot has massless rods of equal length \(l\) and 
+    equal masses \(m\) at the tips.}
+    \label{fig:simple-acrobot-model}
+\end{figure}
+
+This system has inertia matrix 
+\(D\), potential function \(P\) (with respect to
+the horizontal bar), and input matrix \(B\) given as follows:
+\begin{align}\label{eqn:acrobot-inertia}
+    D_s(q) &= \begin{bmatrix}
+        ml^2\left(3+2\cos(q_a)\right) & 
+        ml^2\left(1+\cos(q_a)\right) \\
+        ml^2\left(1+\cos(q_a)\right) &
+        ml^2
+    \end{bmatrix} 
+    , \\
+    \label{eqn:acrobot-potential}
+    P_s(q) &= -mgl\left(2\cos(q_u)+\cos(q_u+q_a)\right)
+    , \\
+    \label{eqn:acrobot-B}
+    B &= [0;1]
+    .
+\end{align}
 %/========== Simulation ==========/%

### 2021-03-16 latex/paper/main.tex
First draft of the problem formulation section

@@
 \newtheorem{thm}{Theorem}% Uncomment: reset numbering at each chapter
-%\newtheorem{prop}[thm]{Proposition} % Propositions depend on Theorem number
+\newtheorem{prop}[thm]{Proposition} % Propositions depend on Theorem number
 \newtheorem{lemma}[thm]{Lemma} % Same with lemmas
@@
 \section{Problem Formulation}\label{sec:problem-formulation}
-TODO: motivate injection/dissipation for energy regulation in an acrobot. Use
-this to lead into VNHCs. e.g. represent torso as chain of a swing, pivot as
-seat, and legs as human legs. Replicate fig3.3 from thesis but for acrobot: take
-(qu,pu)-plane and show how a person moves their legs wrt theta, qu, or pu. Leave
-the gymnast model until the end, or ignore it entirely.
+%TODO: motivate injection/dissipation for energy regulation in an acrobot. Use
+%this to lead into VNHCs. e.g. represent torso as chain of a swing, pivot as
+%seat, and legs as human legs. Replicate fig3.3 from thesis but for acrobot: take
+%(qu,pu)-plane and show how a person moves their legs wrt theta, qu, or pu. Leave
+%the gymnast model until the end, or ignore it entirely.
+%
+%Alternatively, ff human movement journal says legs move wrt qu, we can use that
+%as a foundation.
+%Ideas: Human movement says appropriate regulation of pendulum allows energy to
+%be pumped into system. Pendulum length rapidly increased on downward swing, then
+%rapidly decreased on upward swing (I think these mean backward and forward).
+%They plot pendulum length rdot/r vs "body angle" theta to determine energy
+%injection, which hints at some relationship between (r,rdot) and (theta,thetad).
+%Indeed, the peak value of rdot/r did not occur at a fixed value of theta, nor
+%once anyone reached a particular speed; it occured at the same thetadot/theta for all
+%participants of the study. This perfectly motivates VNHCs for robots, because we
+%can shape (r,rtdot) as functions of (theta,thetad).
+%Start motivation from energy injection - how do humans inject energy? How does
+%acrobot model gymnast? use human movement to motivate tracking functions of both
+%position and velocity rather than time.
 
-Alternatively, ff human movement journal says legs move wrt qu, we can use that
-as a foundation.
+In gymnastics terminology, a ``giant" is the motion a gymnast performs to
+achieve full rotations around a horizontal bar \cite{usagym_giant}. 
+A gymnast will begin (at rest) by hanging on a bar, then swing their legs
+appropriately to gain energy over time.
+The authors of \cite{pendulum_length_giant_gymnastics} modelled the gymnast as a
+variable length pendulum, and studied how the pendulum's length changes as a
+function of the orientation of a gymnast's joints.
+Labeling the pendulum length by \(r\) and the gymnast's body orientation
+by \(\theta\), they discovered experimentally that the value \(\dot{r}/r\) has
+the highest impact on the magintude of energy injection; they also discovered
+that the peak value of \(\dot{r}/r\) occured at the same value of
+\(\dot{\theta}/\theta\) for all gymnasts.
+In other words, gymnasts were moving their legs as a function of their body's
+angle and their angular velocity around the bar, not as functions of time.
+
+Typically as control theorists we like our system's actuatiors to track a fixed
+trajectory over time; but if we were attempting design a gymnastics robot that
+could replicate human behaviour, the results of
+\cite{pendulum_length_giant_gymnastics} indicate that we should not use this
+standard approach.
+Instead, we should have our robot's legs follow a function of its
+body angle and angular velocity.
+That is exactly what we set out to do in this article.
+
+The simplest way to model the gymnast is as a variable-length pendulum, but a
+more realistic model is that of a two-link acrobot (Figure \ref{fig:acrobot}).
+The acrobot's first link represents the gymnast's torso, the bottom link represents
+their legs, and it is actuated at the center joint (the hips).
+
+\begin{figure}
+    \centering
+    %\includegraphics[width=\textwidth]{images/acrobot.png}
+    \caption{The two-link acrobot.}
+    \label{fig:acrobot}
+\end{figure}
+
+In this article, our goal is to design a leg controller for the acrobot so that
+the resulting motion mimics that of a human performing a giant.
+Previous approaches to solve this problem used
+trajectory tracking, set-point tracking, or energy
+pumping regulation (see
+\cite{energy_pumping_robotic_swinging,swingup_giant_acrobot,dynamical_servo_acrobot_vc,control_giant_two_link_gymnastic_robot}
+).
+While they all result in the acrobot performing rotations on the bar, none of
+the are particularly human-like.
+The most realistic controller to date was designed by Wang \cite{xingbo_thesis}; it
+forces the hips to track a function of the torso angle.
+However, a truly human-like controller would force
+the hips to track a function involving the torso velocity.
+These kinds of controllers are known as ``virtual nonholonomic constraints", and
+they have been used for human-robot interaction
+\cite{vnhc_human_robot_cooperation,psd_based_vnhc_redundant_manipulator,haptic_vnhc},
+error-reduction on time-delayed systems \cite{vnhc_time_delay_teleop},
+and they have shown marked improvements to the field of bipedal locomotion 
+\cite{nhvc_dynamic_walking,
+hybrid_zero_dynamics_bipedal_nhvcs,output_nhvc_bipedal_control}.
+Indeed, virtual nonholonomic constraints produce more robust bipedal motion than
+controllers which do not depend on velocity \cite{nhvc_incline_walking}.
+In particular, virtual nonholonomic constraints may be capable of injecting and
+dissipating energy from a system in a robust manner, all while producing
+realistic biological motion. 
+Hence, we will design one of these to achieve stable giant motion on the
+acrobot.
+
+Here then is the problem we will solve: design a virtual nonholonomic constraint
+for the acrobot, so that the acrobot provably gains energy over time
+(as per Definition \ref{defn:energy-gain}).
+
+\begin{defn}\label{defn:energy-gain}
+    Let \(\mathcal{Q}\) be an
+    \(n\)-dimensional generalized cylinder.
+    Let \(f : \mathcal{Q} \rightarrow \mathcal{Q}\times\R^n\) be a smoth vector
+    field and let \(D \subset M\) be open.
+    The system described by \(\dot{x} = f(x)\) 
+    \textit{gains energy on \(D\)} if, 
+    for all compact sets \(K \subset D\) and for almost every initial
+    condition \(x(0) \in K\), there exists \(T > 0\) such
+    that \(x(t) \notin K \, (\forall t > T)\).
+    The system \textit{loses energy on \(D\)} if it gains energy in
+    negative-time.
+\end{defn}
 
@@
 This system has inertia matrix 
-\(D\), potential function \(P\) (with respect to
+\(M\), potential function \(V\) (with respect to
 the horizontal bar), and input matrix \(B\) given as follows:
 \begin{align}\label{eqn:acrobot-inertia}
-    D_s(q) &= \begin{bmatrix}
+    M(q) &= \begin{bmatrix}
         ml^2\left(3+2\cos(q_a)\right) & 
@@
     \label{eqn:acrobot-potential}
-    P_s(q) &= -mgl\left(2\cos(q_u)+\cos(q_u+q_a)\right)
+    V(q) &= -mgl\left(2\cos(q_u)+\cos(q_u+q_a)\right)
     , \\
@@
 \end{align}
+The conjugate of momenta is \(p = (p_u,p_a) = M(q)\dot{q}\).
+The dynamics of the acrobot in \((q,p)\) coordinates are given in
+~\eqref{eqn:acrobot-hamiltonian}, where
+for shorthand, we write \(c_u := \cos(q_u)\), \(c_a := \cos(q_a)\), and 
+\(c_{ua} := \cos(q_u + q_a)\); likewise, \(s_u := \sin(q_u)\), 
+\(s_a := \sin(q_a)\), and \(s_{ua} := \sin(q_u + q_a)\).
+\begin{align}\label{eqn:acrobot-hamiltonian}
+    \mathcal{H}(q,p) &= \frac{1}{2}p\tpose \Minv(q) p -
+    mgl\left(2 c_u + c_{ua}\right)
+    , \\
+     &\begin{cases}
+        \dot{q} = \Minv(q) p 
+        ,\\
+        \dot{p}_u = -mgl\left(2s_u + s_{ua}\right) 
+        ,\\
+        \dot{p}_a =-\frac{1}{2}p\tpose \nabla_{q_a}\Minv(q) p
+        - mgl s_{ua} + \tau,
+    \end{cases} \nonumber
+\end{align}
+The control input is a force \(\tau \in \R\) affecting only the dynamics of
+\(p_a\), representing a torque acting on the hip joint.
+This means \((q,p)\) are simply actuated coordinates inside the phase space
+\(\mathcal{Q} \times \mathcal{P}\) where
+\(\mathcal{Q} = \mathcal{Q}_u \times \mathcal{Q}_a 
+:= \Sone \times \Sone\), and
+\(\mathcal{P} = \mathcal{P}_u \times \mathcal{P}_a
+:= \R \times \R\).
+Hence, we can apply the theory of VNHCs from Section \ref{sec:vnhc} to design an
+energy-injecting controller.
+Since we need our VNHC to be regular, the following proposition will prove
+useful.
+\begin{prop}\label{prop:acrobot-fpu-regular}
+    A relation \(h(q,p) = q_a - f(p_u) = 0\) 
+    with \(f \in C^2\left(\mathcal{P}_u; \mathcal{Q}_a\right)\) is a regular
+    VNHC of order 1 for the simple acrobot.
+\end{prop}
+\begin{proof}
+    The regularity matrix for this VNHC evaluates to
+    \(\frac{(1+c_a)\partial_{q_u}f(q_u,p_u) + (3+2c_a)}{ml^2(2-c_a^2)}\).
+    Since \(\partial_{q_u} f = 0\), this matrix is strictly positive for all values
+    of \(q_a\), and hence is full rank everywhere on the constraint manifold.
+\end{proof}
+
 %/========== Simulation ==========/%

### 2021-03-17 latex/paper/main.tex
Edits of first section and VNHC section
Made changes suggested by Manfredi and edited the Problem Formulation to
get the secnd draft.

@@
 \subsection{Notation}
-We use the following notation in this article.
+We use the following notation and terminology in this article.
+A matrix \(A \in \R^{n \times m}\) is \textit{right semi-orthogonal} if
+\(A A\tpose = \Id{m}\) and is \textit{left semi-orthogonal} if 
+\(A\tpose A = \Id{n}\).
 For \(A \in \R^{n\times m}\) and \(B \in \R^{p \times m}\),
-define \([A;B] \in \R^{(n+p)\times m}\) as the matrix obtained by stacking \(A\)
+we define \([A;B] \in \R^{(n+p)\times m}\) as the matrix obtained by stacking \(A\)
 on top of \(B\). 
@@
 achieve full rotations around a horizontal bar \cite{usagym_giant}. 
-A gymnast will begin (at rest) by hanging on a bar, then swing their legs
+A gymnast will begin by hanging at rest, then swing their legs
 appropriately to gain energy over time.
@@
 \(\dot{\theta}/\theta\) for all gymnasts.
-In other words, gymnasts were moving their legs as a function of their body's
-angle and their angular velocity around the bar, not as functions of time.
+In other words, gymnasts appear to move their legs as a function of their body
+angle and velocity; this is the mechanism by which they gain enough energy to
+rotate around the bar.
 
-Typically as control theorists we like our system's actuatiors to track a fixed
-trajectory over time; but if we were attempting design a gymnastics robot that
+Control theorists typically want a system's actuators to track a fixed
+trajectory over time. 
+If we were attempting design a gymnastics robot which
 could replicate human behaviour, the results of
@@
     \centering
-    %\includegraphics[width=\textwidth]{images/acrobot.png}
-    \caption{The two-link acrobot.}
+    \includegraphics[width=\linewidth]{acrobot_gymnast.png}
+    \caption{The two-link acrobot as a model for a gymnast.
+    Image modified from \cite{xingbo_thesis}.}
     \label{fig:acrobot}
@@
 the resulting motion mimics that of a human performing a giant.
-Previous approaches to solve this problem used
+Previous attempts at solving this problem have involved
 trajectory tracking, set-point tracking, or energy
@@
 ).
-While they all result in the acrobot performing rotations on the bar, none of
-the are particularly human-like.
-The most realistic controller to date was designed by Wang \cite{xingbo_thesis}; it
-forces the hips to track a function of the torso angle.
-However, a truly human-like controller would force
-the hips to track a function involving the torso velocity.
-These kinds of controllers are known as ``virtual nonholonomic constraints", and
-they have been used for human-robot interaction
+While these approaches all result in the acrobot performing rotations on the
+bar, none of the are particularly human-like.
+Wang designed a controller which
+forces the hips to track a function of the torso angle using virtual holonomic
+constraints \cite{xingbo_thesis};
+however, truly human-like hip motion would also take into account the torso velocity.
+This style of control action is captured by the method of virtual nonholonomic
+constraints.
+
+Virtual nonholonomic constraints (VNHCs) have been used for human-robot interaction
 \cite{vnhc_human_robot_cooperation,psd_based_vnhc_redundant_manipulator,haptic_vnhc},
@@
 hybrid_zero_dynamics_bipedal_nhvcs,output_nhvc_bipedal_control}.
-Indeed, virtual nonholonomic constraints produce more robust bipedal motion than
+Indeed, they produce more robust bipedal motion than
 controllers which do not depend on velocity \cite{nhvc_incline_walking}.
-In particular, virtual nonholonomic constraints may be capable of injecting and
+In particular, VNHCs may be capable of injecting and
 dissipating energy from a system in a robust manner, all while producing
 realistic biological motion.
-Hence, we will design one of these to achieve stable giant motion on the
-acrobot.
 
 Here then is the problem we will solve: design a virtual nonholonomic constraint
-for the acrobot, so that the acrobot provably gains energy over time
-(as per Definition \ref{defn:energy-gain}).
+so that the acrobot performs giant-like motion.
+We will rigorously prove the acrobot ``gains energy" according to the following
+definition.
 
@@
 \end{defn}
+
+Before embarking on this design problem, let us summarize the relevant theory of
+VNHCs.
 
@@
 that are mutually orthogonal. 
-In the case that \(k = (n-1)\), the existence of any left annihilator 
-\(A^0 \in \R^{1\times n}\) implies the left annihilator 
-\(B^\perp := A^0/\norm{A^0}\) will be a unit vector satisfying this assumption.
+When \(k = (n-1)\), Assumption \ref{assm:B-perp} can be removed because it is
+automatically implied by Assumption \ref{assm:B-const}.
 
@@
 \begin{thm}\label{thm:simply-actuated}
-    Take the Hamiltonian system ~\eqref{eqn:hamiltonian}.
-    Under Assumptions \ref{assm:B-const} and \ref{assm:B-perp},
-    the coordinate transformation
+    Take the Hamiltonian system ~\eqref{eqn:hamiltonian} and suppose
+    Assumptions \ref{assm:B-const} and \ref{assm:B-perp} hold.
+    The coordinate transformation
     \(\left(\tilde{q} = \mathbf{B}q, \tilde{p} = \mathbf{B}p\right)\)
-    is a canonical transformation.
-    The resulting dynamics are given by 
+    is a canonical transformation and the resulting dynamics are given by 
     \begin{gather}\label{eqn:simple-hamiltonian}
@@
 
-We call these new coordinates 
+We call the \((\tilde{q},\tilde{p})\) coordinates
 \textit{simply actuated coordinates}.
-The first \((n-k)\) configuration variables, labelled \(q_u\), are
-the \textit{unactuated coordinates}; 
+The first \((n-k)\) configuration variables in \(\tilde{q}\), labelled \(q_u\),
+are the \textit{unactuated coordinates}; 
 the remaining \(k\) configuration variables, labelled \(q_a\), are the
 \textit{actuated coordinates}.
-The corresponding \((p_u, p_a)\) are the \textit{unactuated} and 
-\textit{actuated momenta}, respectively.
+The corresponding \((p_u, p_a)\) in \(\tilde{p}\) are the \textit{unactuated}
+and \textit{actuated momenta}, respectively.
 
@@
 For this reason, we take the system of inquiry to be a Hamiltonian
-mechanical system in simply actuated coordinates.
+mechanical system in simply actuated coordinates, as in
+\eqref{eqn:simple-hamiltonian}.
+For simplicity of notation, we relabel \((\tilde{q},\tilde{p})\) as \((q,p)\),
+where we now have \((q,p) = (q_u,q_a,p_u,p_a)\).
 
@@
 \begin{proof}
+    \textbf{TODO: should we make this a ``proof sketch"?}
     Let \(e = h(q,p) \in \R^k\).

### 2021-03-17 latex/paper/main.tex
Draft 3 of Problem Formulation

@@
 by \(\theta\), they discovered experimentally that the value \(\dot{r}/r\) has
-the highest impact on the magintude of energy injection; they also discovered
-that the peak value of \(\dot{r}/r\) occured at the same value of
-\(\dot{\theta}/\theta\) for all gymnasts.
+the highest impact on the magintude of energy injection. 
+After testing several gymnasts under a variety of experimental conditions, 
+they discovered that the peak value of \(\dot{r}/r\) occured at the same fixed
+value of \(\dot{\theta}/\theta\) for all gymnasts.
 In other words, gymnasts appear to move their legs as a function of their body
-angle and velocity; this is the mechanism by which they gain enough energy to
-rotate around the bar.
+angle and velocity when performing giants, not as functions of time. 
+Doing so allows them to gain energy and rotate around the bar.
 
-Control theorists typically want a system's actuators to track a fixed
-trajectory over time. 
-If we were attempting design a gymnastics robot which
-could replicate human behaviour, the results of
-\cite{pendulum_length_giant_gymnastics} indicate that we should not use this
-standard approach.
-Instead, we should have our robot's legs follow a function of its
-body angle and angular velocity.
-That is exactly what we set out to do in this article.
-
-The simplest way to model the gymnast is as a variable-length pendulum, but a
+The simplest way to model a gymnast is as a variable-length pendulum, but a
 more realistic model is that of a two-link acrobot (Figure \ref{fig:acrobot}).
@@
 their legs, and it is actuated at the center joint (the hips).
+In this article, our goal is to design a leg controller which provably injects
+energy into the acrobot, so that the resulting motion mimics that of a human
+performing a giant.
 
@@
 
-In this article, our goal is to design a leg controller for the acrobot so that
-the resulting motion mimics that of a human performing a giant.
 Previous attempts at solving this problem have involved
-trajectory tracking, set-point tracking, or energy
-pumping regulation (see
+trajectory tracking, or other feedback methods to stabilize
+certain energy levels (see
 \cite{energy_pumping_robotic_swinging,swingup_giant_acrobot,dynamical_servo_acrobot_vc,control_giant_two_link_gymnastic_robot}
 ).
-While these approaches all result in the acrobot performing rotations on the
-bar, none of the are particularly human-like.
-Wang designed a controller which
-forces the hips to track a function of the torso angle using virtual holonomic
-constraints \cite{xingbo_thesis};
-however, truly human-like hip motion would also take into account the torso velocity.
-This style of control action is captured by the method of virtual nonholonomic
-constraints.
+While all these approaches result in the acrobot rotating around the
+bar, none of them are particularly human-like.
+After all, the results of \cite{pendulum_length_giant_gymnastics} indicate that
+a robot's legs should track a function of its body angle and velocity.
+In 2016, Wang attempted to design a controller with this in mind \cite{xingbo_thesis}
+but was unable to prove the acrobot would gain energy over time.
+His approach was a rudimentary version of a recent technique know as the method
+of virtual nonholonomic constraint.
 
@@
 
-Here then is the problem we will solve: design a virtual nonholonomic constraint
-so that the acrobot performs giant-like motion.
-We will rigorously prove the acrobot ``gains energy" according to the following
-definition.
+Here then is the problem we will solve in this article: design a virtual
+nonholonomic constraint so that the acrobot provably gains energy while
+performing giant-like motion.
+The following is a rigorous definition of ``energy gain".
 
@@
 
-Before embarking on this design problem, let us summarize the relevant theory of
+Before embarking on our design problem, let us summarize the relevant theory of
 VNHCs.
@@
 \begin{proof}
-    \textbf{TODO: should we make this a ``proof sketch"?}
     Let \(e = h(q,p) \in \R^k\).

### 2021-03-17 latex/paper/main.tex
Fixing some typos

@@
     Let \(\mathcal{Q}\) be an
-    \(n\)-dimensional generalized cylinder.
-    Let \(f : \mathcal{Q} \rightarrow \mathcal{Q}\times\R^n\) be a smoth vector
+    \(n\)-dimensional smooth manifold.
+    Let \(f : \mathcal{Q} \rightarrow \mathcal{Q}\times\R^n\) be a smooth vector
     field and let \(D \subset M\) be open.
@@
 For simplicity of notation, we relabel \((\tilde{q},\tilde{p})\) as \((q,p)\),
-where we now have \((q,p) = (q_u,q_a,p_u,p_a)\).
+where we now have \((q,p) = (q_u,q_a,p_u,p_a) \in \mathcal{Q}\times \R^n\).
 
@@
 \section{The Acrobot VNHC}\label{sec:acrobot}
-specialize the theory of VNHCs to the acrobot. consider constraints which depend
-only on pu, and show arctan vnhc and reduced dynamics. 
-give definition of energy injection/dissipation. conclude with our
-theorem. Make this section short and sweet.
+%specialize the theory of VNHCs to the acrobot. consider constraints which depend
+%only on pu, and show arctan vnhc and reduced dynamics. 
+%give definition of energy injection/dissipation. conclude with our
+%theorem. Make this section short and sweet.
+%
+%TODO: after writing S2, re-introduce what we want to do with acrobot (energy
+%injection for giant stabilization)
 
-TODO: after writing S2, re-introduce what we want to do with acrobot (energy
-injection for giant stabilization)
-
-For our model of the acrobot, we assume the acrobot iscomprised of two massless
+For our model of the acrobot, we assume the acrobot is comprised of two massless
 rods of equal length \(l\), with equal point masses \(m\) at the tips.

### 2021-03-19 latex/paper/main.tex
Revision of Problem Formulation, final draft 1

@@
 A matrix \(A \in \R^{n \times m}\) is \textit{right semi-orthogonal} if
-\(A A\tpose = \Id{m}\) and is \textit{left semi-orthogonal} if 
-\(A\tpose A = \Id{n}\).
+\(A A\tpose = \Id{n}\) and is \textit{left semi-orthogonal} if 
+\(A\tpose A = \Id{m}\).
 For \(A \in \R^{n\times m}\) and \(B \in \R^{p \times m}\),
@@
 variable length pendulum, and studied how the pendulum's length changes as a
-function of the orientation of a gymnast's joints.
+function of the gymnast's limb angle.
 Labeling the pendulum length by \(r\) and the gymnast's body orientation
-by \(\theta\), they discovered experimentally that the value \(\dot{r}/r\) has
+by \(\theta\), they observed experimentally that the value \(\dot{r}/r\) has
 the highest impact on the magintude of energy injection. 
@@
 In other words, gymnasts appear to move their legs as a function of their body
-angle and velocity when performing giants, not as functions of time. 
-Doing so allows them to gain energy and rotate around the bar.
+angle and velocity when performing giants; 
+doing so allows them to gain energy and rotate around the bar.
 
-The simplest way to model a gymnast is as a variable-length pendulum, but a
-more realistic model is that of a two-link acrobot (Figure \ref{fig:acrobot}).
-The acrobot's first link represents the gymnast's torso, the bottom link represents
-their legs, and it is actuated at the center joint (the hips).
-In this article, our goal is to design a leg controller which provably injects
-energy into the acrobot, so that the resulting motion mimics that of a human
-performing a giant.
+While the simplest model of a gymnast is the variable-length pendulum, a
+more realistic model is the two-link acrobot (Figure \ref{fig:acrobot}).
+Here, the top link represents the torso, while the bottom link represents
+the legs. 
+The acrobot is actuated exclusively at the center joint (the hips)
+Controlling the acrobot is a nontrivial task because it is not feedback
+linearizable \cite{nonlinear_controllers_nonintegrable_acrobot}. 
+To solve the swingup problem, one might begin by designing a leg controller
+which provably injects energy into the acrobot, so that the resulting motion
+mimics that of a human performing a giant.
 
@@
 
-Previous attempts at solving this problem have involved
-trajectory tracking, or other feedback methods to stabilize
-certain energy levels (see
+Previous attempts at acrobot giant generation have involved
+trajectory tracking, partial feedback linearization, or other energy-based
+methods
+( see
 \cite{energy_pumping_robotic_swinging,swingup_giant_acrobot,dynamical_servo_acrobot_vc,control_giant_two_link_gymnastic_robot}
 ).
-While all these approaches result in the acrobot rotating around the
-bar, none of them are particularly human-like.
-After all, the results of \cite{pendulum_length_giant_gymnastics} indicate that
-a robot's legs should track a function of its body angle and velocity.
+While all these approaches succeed at making the acrobot rotate around the
+bar, none of them use the results of \cite{pendulum_length_giant_gymnastics}.
+That is, none of these leg controllers track a function of the acrobot's body
+angle and velocity.
 In 2016, Wang attempted to design a controller with this in mind \cite{xingbo_thesis}
@@
 His approach was a rudimentary version of a recent technique know as the method
-of virtual nonholonomic constraint.
+of virtual nonholonomic constraints.
 
@@
 
-Here then is the problem we will solve in this article: design a virtual
-nonholonomic constraint so that the acrobot provably gains energy while
-performing giant-like motion.
-The following is a rigorous definition of ``energy gain".
-
-\begin{defn}\label{defn:energy-gain}
-    Let \(\mathcal{Q}\) be an
-    \(n\)-dimensional smooth manifold.
-    Let \(f : \mathcal{Q} \rightarrow \mathcal{Q}\times\R^n\) be a smooth vector
-    field and let \(D \subset M\) be open.
-    The system described by \(\dot{x} = f(x)\) 
-    \textit{gains energy on \(D\)} if, 
-    for all compact sets \(K \subset D\) and for almost every initial
-    condition \(x(0) \in K\), there exists \(T > 0\) such
-    that \(x(t) \notin K \, (\forall t > T)\).
-    The system \textit{loses energy on \(D\)} if it gains energy in
-    negative-time.
-\end{defn}
+In this article, we will design a virtual
+nonholonomic constraint which provably injects energy into the acrobot through 
+human-like giant motion.
 
@@
 For our model of the acrobot, we assume the acrobot is comprised of two massless
-rods of equal length \(l\), with equal point masses \(m\) at the tips.
-We call this a \textit{simple} acrobot, which is displayed in Figure
+rods of equal length \(l\), with equal point masses \(m\) at the tips, as in Figure
 \ref{fig:simple-acrobot-model}.
-We will also ignore any frictional forces at both the hip and shoulder joints. 
-Finally, it is important to note that a real gymnast cannot swing their legs in
-full circles, though they are usually flexible enough to raise them parallel to
-the floor; 
+Note that a real gymnast cannot swing their legs in full circles, though they
+are usually flexible enough to raise them parallel to the floor; 
 for this reason, we assume that \(q_a \in [-Q_a, Q_a]\) where 
 \(Q_a \in [\frac{\pi}{2}, \pi[\). 
+We also ignore any frictional forces at both the hip and shoulder joints.
 
@@
 
+\textbf{TODO: Where does this go?}
+\begin{defn}\label{defn:energy-gain}
+    Let \(\mathcal{Q}\) be an
+    \(n\)-dimensional smooth manifold.
+    Let \(f : \mathcal{Q} \rightarrow \mathcal{Q}\times\R^n\) be a smooth vector
+    field and let \(D \subset M\) be open.
+    The system described by \(\dot{x} = f(x)\) 
+    \textit{gains energy on \(D\)} if, 
+    for all compact sets \(K \subset D\) and for almost every initial
+    condition \(x(0) \in K\), there exists \(T > 0\) such
+    that \(x(t) \notin K \, (\forall t > T)\).
+    The system \textit{loses energy on \(D\)} if it gains energy in
+    negative-time.
+\end{defn}
+
 %/========== Simulation ==========/%

### 2021-03-19 latex/paper/main.tex
Put the definition of energy injection into VNHC section

@@
 
+We conclude this section by formalizing the notion of energy injection for
+VNHCs.
+
+\begin{defn}\label{defn:energy-gain}
+    Let \(\mathcal{Q}\) be an
+    \(n\)-dimensional smooth manifold.
+    Let \(f : \mathcal{Q} \rightarrow \mathcal{Q}\times\R^n\) be a smooth vector
+    field and let \(D \subset M\) be open.
+    The system described by \(\dot{x} = f(x)\) 
+    \textit{gains energy on \(D\)} if, 
+    for all compact sets \(K \subset D\) and for almost every initial
+    condition \(x(0) \in K\), there exists \(T > 0\) such
+    that \(x(t) \notin K \, (\forall t > T)\).
+    The system \textit{loses energy on \(D\)} if it gains energy in
+    negative-time.
+\end{defn}
+
+Any system satisfying Definition \ref{defn:energy-gain}
+can have unstable equilibria on \(D\), but not limit cycles nor closed orbits.
+The next definition ties this notion of energy gain to VNHCs.
+
+\begin{defn}\label{defn:energy-injection}
+    A regular VNHC \(h(q, p) = 0\) with constraint manifold \(\Gamma\)
+    \textit{injects (dissipates) energy on \(D \subset \Gamma\)} if the
+    constrained dynamics gain (lose) energy everywhere on \(D\), except possibly
+    on a set of measure zero.
+\end{defn}
+
 \textbf{Comparison with existing literature}: Horn et.al. provide the constrained
@@
     \centering
-    %\includegraphics[width=0.5\textwidth]{images/simple_acrobot_model}
+    \includegraphics[width=0.7\linewidth]{simple_acrobot_model.pdf}
     \caption{A simple acrobot has massless rods of equal length \(l\) and 
@@
 
-\textbf{TODO: Where does this go?}
-\begin{defn}\label{defn:energy-gain}
-    Let \(\mathcal{Q}\) be an
-    \(n\)-dimensional smooth manifold.
-    Let \(f : \mathcal{Q} \rightarrow \mathcal{Q}\times\R^n\) be a smooth vector
-    field and let \(D \subset M\) be open.
-    The system described by \(\dot{x} = f(x)\) 
-    \textit{gains energy on \(D\)} if, 
-    for all compact sets \(K \subset D\) and for almost every initial
-    condition \(x(0) \in K\), there exists \(T > 0\) such
-    that \(x(t) \notin K \, (\forall t > T)\).
-    The system \textit{loses energy on \(D\)} if it gains energy in
-    negative-time.
-\end{defn}
-
 %/========== Simulation ==========/%

### 2021-03-19 latex/paper/main.tex
Added VNHC description to Acrobot section

@@
 the legs. 
-The acrobot is actuated exclusively at the center joint (the hips)
+The acrobot is actuated exclusively at the center joint (the hips).
 Controlling the acrobot is a nontrivial task because it is not feedback
@@
 hybrid_zero_dynamics_bipedal_nhvcs,output_nhvc_bipedal_control}.
-Indeed, they produce more robust bipedal motion than
-controllers which do not depend on velocity \cite{nhvc_incline_walking}.
+Indeed, they produce more robust walking motion in biped robots than
+other controllers which do not depend on velocity \cite{nhvc_incline_walking}.
 In particular, VNHCs may be capable of injecting and
@@
 \eqref{eqn:simple-hamiltonian}.
-For simplicity of notation, we relabel \((\tilde{q},\tilde{p})\) as \((q,p)\),
-where we now have \((q,p) = (q_u,q_a,p_u,p_a) \in \mathcal{Q}\times \R^n\).
+For simplicity of notation, we relabel \((\tilde{q},\tilde{p})\) as \((q,p)\).
 
@@
     Let \(\mathcal{Q}\) be an
-    \(n\)-dimensional smooth manifold.
+    \(n\)-dimensional smooth generalized cylinder.
     Let \(f : \mathcal{Q} \rightarrow \mathcal{Q}\times\R^n\) be a smooth vector
@@
 %injection for giant stabilization)
-
-For our model of the acrobot, we assume the acrobot is comprised of two massless
-rods of equal length \(l\), with equal point masses \(m\) at the tips, as in Figure
+Recall that we wish to design a VNHC which injects energy into the acrobot
+through giant-like motion.
+For our model of the acrobot, we assume the torso and leg rods are of equal
+length \(l\) with equal point masses \(m\) at the tips, as in Figure
 \ref{fig:simple-acrobot-model}.
@@
 \(\mathcal{Q} \times \mathcal{P}\) where
-\(\mathcal{Q} = \mathcal{Q}_u \times \mathcal{Q}_a 
-:= \Sone \times \Sone\), and
-\(\mathcal{P} = \mathcal{P}_u \times \mathcal{P}_a
-:= \R \times \R\).
+\(\mathcal{Q} = \Sone \times \Sone\), and
+\(\mathcal{P} = \R \times \R\).
 Hence, we can apply the theory of VNHCs from Section \ref{sec:vnhc} to design an
@@
     A relation \(h(q,p) = q_a - f(p_u) = 0\) 
-    with \(f \in C^2\left(\mathcal{P}_u; \mathcal{Q}_a\right)\) is a regular
+    with \(f \in C^2\left(\R; \Sone\right)\) is a regular
     VNHC of order 1 for the simple acrobot.
@@
 \end{proof}
+
+To design our VNHC, we begin by examining a person on a
+seated swing.
+The person extends their legs when the swing moves forwards, and retracts them
+when the swing moves backwards.
+As the swing gains speed, the person leans their body back while
+extending their legs higher, thereby shortening the distance
+from their center of mass to the pivot and adding more energy to the swing
+\cite{how_to_pump_a_swing}.
+
+Now imagine the person's torso is affixed to the swing's rope so they are
+always upright. 
+Imagine further that the swing has no seat at all, allowing the person to extend
+their legs beneath them. 
+This position is identical to that of a gymnast on a bar, which is why we can
+use leg motion from the seated swing to design a controller for the acrobot.
+
+The acrobot's legs are rigid rods which cannot retract, so we emulate the person
+on a swing by pivoting the legs toward the direction of motion. 
+To account for how a person leans back at higher speeds, the legs should pivot to an
+angle proportional to the swing's speed.
+Since the direction of motion is entirely determined by \(p_u\), 
+one such VNHC which emulates this process is \(q_a = \bar{q}_a\arctan( I p_u)\),
+displayed in Figure \ref{fig:qa-arctan}.
+Here, \(\bar{q}_a \in ]0,\frac{2 Q_a}{\pi}]\) and \(I \in \R\) is a fixed
+control parameter.
+
+This constraint does not perfectly recreate giant motion, during which
+the gymnast's legs are almost completely extended \cite{usagym_giant}
+-- it instead pivots the legs partially during rotations.
+However, the behaviour looks similar enough that the constraint should provide a
+reasonable foundation for injecting energy into the acrobot.
+Hence, our VNHC is
+\begin{equation}\label{eqn:acrobot-constraint}
+    h(q,p) = q_a - \bar{q}_a \arctan(I p_u)
+    .
+\end{equation}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.7\linewidth]{qa_arctan.pdf}
+    \caption{The acrobot constraint \(q_a = \bar{q}_a \arctan(I p_u)\).}
+    \label{fig:qa-arctan}
+\end{figure}
+
+It is not hard to show that the constrained dynamics for the acrobot under
+~\eqref{eqn:acrobot-constraint} are
+\begin{equation}\label{eqn:acrobot-constrained-dynamics}
+    \begin{cases}
+    \dot{q}_u &= \frac{(1+I^2 p_u^2)p_u + m^2gl^3\bar{q}_a I(2s_u + s_{ua})(1+c_a) }
+            {ml^2(1+I^2 p_u^2)(3+2c_a)}
+    ,    \\
+    \dot{p}_u &= - m g l (2s_u + s_{ua})
+    ,
+    \end{cases}
+\end{equation}
+subject to \(q_a = \bar{q}_a \arctan(I p_u)\).
 

### 2021-03-23 latex/paper/main.tex
Initial draft of the acrobot section.
I've written the theorem statement, the lead-up to it, and described the
regions O_1 and O_2(rhobar).
There are some overfull hboxes to deal with and the whole thing needs a
phrasing edit.

@@
 \usepackage{graphicx}
+\usepackage{subcaption}
 \usepackage{textcomp}
@@
 by \(\theta\), they observed experimentally that the value \(\dot{r}/r\) has
-the highest impact on the magintude of energy injection. 
+the biggest impact on the magnitude of energy injection. 
 After testing several gymnasts under a variety of experimental conditions, 
@@
 angle and velocity.
-In 2016, Wang attempted to design a controller with this in mind \cite{xingbo_thesis}
+In 2016, Wang attempted to design a controller which does this \cite{xingbo_thesis}
 but was unable to prove the acrobot would gain energy over time.
@@
 displayed in Figure \ref{fig:qa-arctan}.
-Here, \(\bar{q}_a \in ]0,\frac{2 Q_a}{\pi}]\) and \(I \in \R\) is a fixed
+Here, \(\bar{q}_a \in ]0,2 Q_a/\pi]\) and \(I \in \R\) is a fixed
 control parameter.
@@
 
-It is not hard to show that the constrained dynamics for the acrobot under
-~\eqref{eqn:acrobot-constraint} are
+Recall that \((q_u, p_u)\) denote the angle and momentum of the acrobot's torso. 
+Hence, the constrained dynamics are the torso dynamics when one swings the legs
+according to \eqref{eqn:acrobot-constraint}.
+It is not hard to show that the constrained dynamics are
 \begin{equation}\label{eqn:acrobot-constrained-dynamics}
@@
 subject to \(q_a = \bar{q}_a \arctan(I p_u)\).
+
+One can study the constrained system by examining solely the motion of 
+\((q_u,p_u) \in \SxR\).
+Suppose for a moment that \(I = 0\) -- that is, the legs stay fully extended.
+The acrobot becomes a pendulum with two masses, whose total mechanical
+energy is
+\begin{equation}\label{eqn:acrobot-energy}
+    E(q_u,p_u) := \frac{p_u^2}{10ml^2} + 3mgl(1 - \cos(q_u))
+    .
+\end{equation}
+If the pendulum is oscillating, its phase lies in the domain
+\begin{equation}\label{eqn:oscillation-domain}
+    \mathcal{O}_1 := \left\{(q_u,p_u) \in \SxR 
+    \mid E(q_u,p_u) < E(\pi,0) \right\}
+    ,
+\end{equation}
+which is shown on the \((q_u,p_u)\)-plane in Figure 
+\ref{fig:acrobot-oscillation-domain}.
+If for some \(I \neq 0\) our VNHC injects energy into the acrobot on 
+\(\mathcal{O}_1\), and additionally the constrained dynamics escape
+\(\mathcal{O}_1\) in finite time, then we can guarantee the acrobot will perform
+giant-like motion and begin rotating around the bar.
+
+The upright equilibrium of the pendulum is at \((q_u,p_u) = (\pi,0)\).
+If the pendulum were initialized at \(q_u = 0\), the minimum initial momentum
+required to reach the upright equilibrium would be
+\(p_u = \sqrt{60m^2gl^3}\), since \(E(\pi,0) = E(0,\sqrt{60m^2gl^3})\).
+Hence, if the pendulum is rotating with momentum 
+\(\sqrt{60m^2gl^3} < p_u < \bar{\rho}\),
+the pendulum phase must lie inside
+\begin{equation}\label{eqn:rotation-domain}
+    \mathcal{R}(\rho) := \left\{
+        (q_u,p_u) \in \SxR \mid E(\pi,0) < E(q_u,p_u) < E(0,\rho)
+    \right\}
+    .
+\end{equation}
+Connecting the regions \eqref{eqn:oscillation-domain} and
+\eqref{eqn:rotation-domain} yields the set
+\begin{equation}\label{eqn:o-rhobar}
+    \mathcal{O}_2(\bar{\rho}) := \left\{(q_u,p_u) \in \SxR
+        \mid E(q_u,p_u) < E(0,\bar{\rho}) \right\}
+    ,
+\end{equation}
+as in Figure \ref{fig:acrobot-o2}
+If the VNHC injects energy on \(\mathcal{O}_2(\bar{\rho})\) for some 
+\(I \neq 0\), then the acrobot will swing up, begin rotating, and eventually
+rotate with a momentum of at least \(\bar{\rho}\).
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[t]{0.45\linewidth}
+        \includegraphics[width=\linewidth]{acrobot_oscillation_domain.pdf}
+        \caption{The set \(\mathcal{O}_1\). An orbit starting in this set (blue)
+            will pass through the level set \(E(\pi,0)\) of the nominal pendulum.}
+        \label{fig:acrobot-oscillation-domain}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.45\linewidth}
+        \includegraphics[width=\linewidth]{acrobot_omega.pdf}
+        \caption{The set \(\mathcal{O}_2(\bar{\rho})\). Orbits starting in this
+            set are guaranteed to reach the level set \(E(0,\bar{\rho})\) of the
+            nominal pendulum.}
+            \label{fig:acrobot-o2}
+    \end{subfigure}
+    \caption{The sets on which the acrobot gains energy, according to Theorem
+        \ref{thm:acrobot-energy-stabilization}.}
+\end{figure}
+
+Unfortunately, the constrained dynamics do not always gain energy on
+\(\mathcal{O}_1\) and \(\mathcal{O}_2(\bar{\rho})\);
+energy gain is only guaranteed under certain conditions on the size of \(I\).
+If \(I\) is too large, the leg controller saturates (i.e. the legs swing up too
+quickly) and the body does not swing faster nor higher. 
+However, choosing \(I\) small enough guarantees the legs will synchronize
+properly with the body, and the acrobot will begin doing backflips on the bar.
+The following theorem provides conditions under which such an \(I\) exists.
+
+\begin{thm}\label{thm:acrobot-energy-stabilization}
+    Consider the acrobot ~\eqref{eqn:acrobot-hamiltonian}.
+\begin{enumerate}
+    \item For any \(m\), \(g\), \(l\), \(\bar{q}_a\), there exists
+        \(I^\star > 0\) such that, for all \(I \in \, ]0,I^\star]\), the VNHC
+        ~\eqref{eqn:acrobot-constraint} injects energy into the acrobot on
+        \(\mathcal{O}_1\).
+        Moreover, almost every orbit will escape the closure of
+        \(\mathcal{O}_1\) in finite time.
+        If instead \(I \in [-I^\star,0[\), the VNHC dissipates energy.
+    \item Let \(C = m^2gl^3\) and 
+        define \(b : \SxR_{> 0} \rightarrow \R\) by
+    \[
+        b(\beta,\rho_0) := 
+        \frac{5C\bar{q}_a \left(
+            C\left(18s_\beta^2 + 30c_\beta(1 - c_\beta)\right)
+            - c_\beta\rho_0^2
+        \right)}{
+        |\rho_0|\sqrt{\rho_0^2 - 30m^2gl^3(1 - c_\beta)}
+        }
+        .
+    \]
+        Define 
+        \(S(\rho_0) := \int \limits_{0}^{2\pi} b(\sigma,\rho_0)d\sigma\).
+        Fix \(\bar{\rho} > \sqrt{60m^2gl^3}\) and
+        suppose there exists \(\epsilon > 0\) so that 
+        \(S(\rho_0) \geq \epsilon\) for all 
+        \(\rho_0 \in \, \left]\sqrt{60m^2gl^3}, \bar{\rho}\right]\).
+        Then there exists \(I^{\star\star} \in \, ]0, I^\star]\) such that, for all 
+        \(I \in \, ]0,I^{\star\star}]\), the VNHC
+        ~\eqref{eqn:acrobot-constraint} injects energy into the acrobot on
+        \(\mathcal{O}_2(\bar{\rho})\).
+        If instead \(I \in [-I^{\star\star},0[\), the VNHC dissipates energy.
+\end{enumerate}
+\end{thm}
 

### 2021-03-26 latex/paper/main.tex
First draft of the full acrobot section, need some editing still.

@@
 \DeclarePairedDelimiter{\norm}{\lVert}{\rVert}
+\DeclarePairedDelimiter{\abs}{\lvert}{\rvert}
 \DeclareMathOperator{\Rank}{rank}
@@
 %injection for giant stabilization)
-Recall that we wish to design a VNHC which injects energy into the acrobot
-through giant-like motion.
-For our model of the acrobot, we assume the torso and leg rods are of equal
-length \(l\) with equal point masses \(m\) at the tips, as in Figure
-\ref{fig:simple-acrobot-model}.
-Note that a real gymnast cannot swing their legs in full circles, though they
+Our goal in this article isto design a VNHC which injects energy into the
+acrobot through giant-like motion.
+We use the simplified acrobot model acrobot in Figure
+\ref{fig:simple-acrobot-model}, where we assume the torso and leg rods
+are of equal length \(l\) with equal point masses \(m\) at the tips.
+A real gymnast cannot swing their legs in full circles, though they
 are usually flexible enough to raise them parallel to the floor;
-for this reason, we assume that \(q_a \in [-Q_a, Q_a]\) where 
+hence, we assume the leg angle \(q_a\) lies in \([-Q_a, Q_a]\) for some
 \(Q_a \in [\frac{\pi}{2}, \pi[\). 
-We also ignore any frictional forces at both the hip and shoulder joints.
+We also ignore any frictional forces.
 
@@
 
-This system has inertia matrix 
+The acrobot has inertia matrix 
 \(M\), potential function \(V\) (with respect to
@@
 \(\mathcal{P} = \R \times \R\).
-Hence, we can apply the theory of VNHCs from Section \ref{sec:vnhc} to design an
-energy-injecting controller.
-Since we need our VNHC to be regular, the following proposition will prove
+We can therefore apply the theory of VNHCs from Section \ref{sec:vnhc} to design
+a constraint of the form \(q_a = f(q_u,p_u)\). 
+This constraint must inject energy into the acrobot according to
+Definition \ref{defn:energy-injection}.
+Since we need the VNHC to be regular, the following proposition will prove
 useful.
@@
 seated swing.
-The person extends their legs when the swing moves forwards, and retracts them
-when the swing moves backwards.
+The person extends their legs when the swing moves forwards, and retracts their
+legs when the swing moves backwards.
 As the swing gains speed, the person leans their body back while
@@
 their legs beneath them. 
-This position is identical to that of a gymnast on a bar, which is why we can
-use leg motion from the seated swing to design a controller for the acrobot.
+This position is identical to that of a gymnast on a bar.
 
@@
 on a swing by pivoting the legs toward the direction of motion. 
-To account for how a person leans back at higher speeds, the legs should pivot to an
-angle proportional to the swing's speed.
-Since the direction of motion is entirely determined by \(p_u\), 
-one such VNHC which emulates this process is \(q_a = \bar{q}_a\arctan( I p_u)\),
+Since a person lifts their legs higher at higher speeds, the acrobot's legs should
+pivot to an angle proportional to the swing's speed.
+Because the direction of motion is entirely determined by \(p_u\), 
+one VNHC which emulates this process is \(q_a = \bar{q}_a\arctan( I p_u)\),
 displayed in Figure \ref{fig:qa-arctan}.
-Here, \(\bar{q}_a \in ]0,2 Q_a/\pi]\) and \(I \in \R\) is a fixed
+Here, \(\bar{q}_a \in \, ]0,2 Q_a/\pi]\) is constant and \(I \in \R\) is a fixed
 control parameter.
@@
 This constraint does not perfectly recreate giant motion, during which
-the gymnast's legs are almost completely extended \cite{usagym_giant}
--- it instead pivots the legs partially during rotations.
-However, the behaviour looks similar enough that the constraint should provide a
-reasonable foundation for injecting energy into the acrobot.
-Hence, our VNHC is
+the gymnast's legs are almost completely extended \cite{usagym_giant}:
+it instead pivots the legs partially during rotations.
+However, the behaviour looks similar enough that this constraint should 
+still inject energy into the acrobot.
+Our final VNHC is
 \begin{equation}\label{eqn:acrobot-constraint}
@@
 Recall that \((q_u, p_u)\) denote the angle and momentum of the acrobot's torso.
-Hence, the constrained dynamics are the torso dynamics when one swings the legs
-according to \eqref{eqn:acrobot-constraint}.
-It is not hard to show that the constrained dynamics are
+By Theorem \ref{thm:zero-dynamics}, we can study the constrained system by
+examining solely the motion of \((q_u,p_u) \in \SxR\).
+It is easy to show that the constrained dynamics are
 \begin{equation}\label{eqn:acrobot-constrained-dynamics}
@@
 subject to \(q_a = \bar{q}_a \arctan(I p_u)\).
+These are just the torso dynamics when one swings the legs
+according to \eqref{eqn:acrobot-constraint}.
 
-One can study the constrained system by examining solely the motion of 
-\((q_u,p_u) \in \SxR\).
-Suppose for a moment that \(I = 0\) -- that is, the legs stay fully extended.
-The acrobot becomes a pendulum with two masses, whose total mechanical
+Suppose for a moment that \(I = 0\), i.e.~that the legs stay fully extended.
+The acrobot becomes a nominal pendulum with two masses, whose total mechanical
 energy is
@@
 \end{equation}
-If the pendulum is oscillating, its phase lies in the domain
+The upright equilibrium of this pendulum is located at \((q_u,p_u) = (\pi,0)\).
+Imagine the pendulum hits the bottom of the swing arc with momentum 
+\(p_u \neq 0\). 
+To reach the upright equilibrium, this momentum must be
+\(p_u = \pm\sqrt{60m^2gl^3}\) because \(E(\pi,0) = E(0,\pm\sqrt{60m^2gl^3})\).
+If the momentum is smaller in magnitude, the acrobot will oscillate; 
+if it is larger, the pendulum will rotate around the bar.
+
+When the pendulum is oscillating, its phase \((q_u,p_u)\) lies in the set
 \begin{equation}\label{eqn:oscillation-domain}
@@
 \end{equation}
-which is shown on the \((q_u,p_u)\)-plane in Figure 
-\ref{fig:acrobot-oscillation-domain}.
+which is shown in Figure \ref{fig:acrobot-oscillation-domain}.
 If for some \(I \neq 0\) our VNHC injects energy into the acrobot on 
-\(\mathcal{O}_1\), and additionally the constrained dynamics escape
-\(\mathcal{O}_1\) in finite time, then we can guarantee the acrobot will perform
+\(\mathcal{O}_1\), and the constrained dynamics escape
+\(\mathcal{O}_1\) in finite time, then the acrobot is guaranteed to perform
 giant-like motion and begin rotating around the bar.
 
-The upright equilibrium of the pendulum is at \((q_u,p_u) = (\pi,0)\).
-If the pendulum were initialized at \(q_u = 0\), the minimum initial momentum
-required to reach the upright equilibrium would be
-\(p_u = \sqrt{60m^2gl^3}\), since \(E(\pi,0) = E(0,\sqrt{60m^2gl^3})\).
-Hence, if the pendulum is rotating with momentum 
-\(\sqrt{60m^2gl^3} < p_u < \bar{\rho}\),
-the pendulum phase must lie inside
-\begin{equation}\label{eqn:rotation-domain}
+When the pendulum is rotating with bounded momentum
+\(\abs{p_u} < \bar{\rho}\) (for some chosen \(\bar{\rho}\)),
+the phase must lie inside the rotation domain
+\begin{multline}\label{eqn:rotation-domain}
     \mathcal{R}(\rho) := \left\{
-        (q_u,p_u) \in \SxR \mid E(\pi,0) < E(q_u,p_u) < E(0,\rho)
+        (q_u,p_u) \in \SxR \mid\right.
+        \\
+        \left.E(\pi,0) < E(q_u,p_u) < E(0,\rho)
     \right\}
     .
-\end{equation}
+\end{multline}
 Connecting the regions \eqref{eqn:oscillation-domain} and
@@
 \end{equation}
-as in Figure \ref{fig:acrobot-o2}
+shown in Figure \ref{fig:acrobot-o2}.
 If the VNHC injects energy on \(\mathcal{O}_2(\bar{\rho})\) for some 
-\(I \neq 0\), then the acrobot will swing up, begin rotating, and eventually
-rotate with a momentum of at least \(\bar{\rho}\).
+\(I \neq 0\), then the acrobot must necessarily swing up, begin rotating, 
+and eventually rotate with a momentum of at least \(\bar{\rho}\).
 
@@
 
-Unfortunately, the constrained dynamics do not always gain energy on
-\(\mathcal{O}_1\) and \(\mathcal{O}_2(\bar{\rho})\);
-energy gain is only guaranteed under certain conditions on the size of \(I\).
+Unfortunately, our VNHC does not always inject energy on \(\mathcal{O}_1\) and
+\(\mathcal{O}_2(\bar{\rho})\).
 If \(I\) is too large, the leg controller saturates (i.e. the legs swing up too
-quickly) and the body does not swing faster nor higher. 
+quickly) and the body may not swing higher. 
 However, choosing \(I\) small enough guarantees the legs will synchronize
@@
 \begin{thm}\label{thm:acrobot-energy-stabilization}
-    Consider the acrobot ~\eqref{eqn:acrobot-hamiltonian}.
+    Consider the acrobot with Hamiltonian dynamics, as in ~\eqref{eqn:acrobot-hamiltonian}.
 \begin{enumerate}
@@
 \end{thm}
+\begin{proof}
+    See Section \ref{sec:proof}.
+\end{proof}
+
+Notice that \(\mathcal{O}_1 \subset \mathcal{O}_2(\bar{\rho})\), yet
+Theorem \ref{thm:acrobot-energy-stabilization} considers these sets separately.
+This separation is advantageous because the first result holds for any
+\(m\), \(g\), \(l\), and \(\bar{q}_a\). 
+In other words, the first result of Theorem
+\ref{thm:acrobot-energy-stabilization} states that all acrobots constrained by
+~\eqref{eqn:acrobot-constraint} will gain enough energy to begin rotating around
+the bar.
+In the worst case, the acrobot will at least perform a swing-up routine to reach
+the unstable equilibrium at \((q_u,p_u) = (\pi,0)\).
+
+For the acrobot to achieve giants with energy
+\(E(0,\bar{\rho})\), it must satisfy the assumption on the integral
+of \(b(\beta,\rho_0)\). 
+The value of this integral depends on the acrobot's physical parameters.
+If the assumption holds, there is some control value \(I > 0\)
+(which depends on \(\bar{\rho}\)) for which the acrobot will 
+achieve rotations with a momentum of at least \(\bar{\rho}\).
 
@@
 %/========== Proof ==========/%
-\section{Proof of Theorem \textbf{TODO: ref theorem}}\label{sec:proof}
+\section{Proof of Theorem \ref{thm:acrobot-energy-stabilization}}\label{sec:proof}
 

### 2021-03-26 latex/paper/main.tex
Began the experimental section

@@
         If instead \(I \in [-I^\star,0[\), the VNHC dissipates energy.
-    \item Let \(C = m^2gl^3\) and 
+    \item Let \(C = m^2gl^3\bar{q}_a\) and 
         define \(b : \SxR_{> 0} \rightarrow \R\) by
@@
         b(\beta,\rho_0) := 
-        \frac{5C\bar{q}_a \left(
-            C\left(18s_\beta^2 + 30c_\beta(1 - c_\beta)\right)
+        \frac{5C \left(
+        \frac{C}{\bar{q}_a}\left(18s_\beta^2 + 30c_\beta(1 - c_\beta)\right)
             - c_\beta\rho_0^2
@@
 \section{Simulation Results}\label{sec:simulations}
-
+\textbf{Run a sim of a simplfied acrobot model of a human}
 %/========== Experiments ==========/%
 \section{Experimental Results}\label{sec:experiments}
+In this section, we test our VNHC on the physical acrobot built by
+Xingbo Wang \cite{xingbo_thesis} (Figure \ref{fig:xingbo-acrobot}).
+The torso and leg rods on this robot have unequal mass and differ in length, so
+its inertia matrix and potential function are not described correctly by 
+~\eqref{eqn:acrobot-inertia}-\eqref{eqn:acrobot-potential}.
+These experiments will therefore show that VNHCs are robust to model mismatch.
 
+\begin{figure}
+    \centering
+    %\includegraphics[width=0.5\linewidth]{}
+    \caption{\textbf{TODO: take a square photo of the acrobot}The acrobot built by Wang \cite{xingbo_thesis}.}
+    \label{fig:xingbo-acrobot}
+\end{figure}
+
+The leg actuator is a servo motor with a built-in PID controller, so we
+can directly assign \(q_a = \arctan\left(I p_u\right)\).
+The control parameter \(I\) must be small to satisfy Theorem
+\ref{thm:acrobot-energy-stabilization}, but it must also be large enough to
+overcome friction in the servo motor.
+We experimentally determined that \(I = 10\) is sufficient to
+ensure the actuator can fully rotate in its allowable range
+\(q_a \in \left[ -\frac{\pi}{2}, \frac{\pi}{2}\right]\).
 %/========== Proof ==========/%

### 2021-03-29 latex/paper/main.tex
First rough draft of the experiments section

@@
 
-%/========== Simulation ==========/%
-\section{Simulation Results}\label{sec:simulations}
-\textbf{Run a sim of a simplfied acrobot model of a human}
 %/========== Experiments ==========/%
@@
 
-The leg actuator is a servo motor with a built-in PID controller, so we
-can directly assign \(q_a = \arctan\left(I p_u\right)\).
 The control parameter \(I\) must be small to satisfy Theorem
@@
 \(q_a \in \left[ -\frac{\pi}{2}, \frac{\pi}{2}\right]\).
+
+\subsection{Simulations Results}
+Evaluating the true mechanical energy of Wang's acrobot at the VNHC
+\(q_a = 0\) yields the energy of the nominal pendulum,
+\[
+    E(q_u,p_u) \approx 396.5501 p_u^2 + 0.5997(1 - \cos(q_u))
+    .
+\]
+We simulated Wang's acrobot starting at 
+\((q_u,p_u) = \left(\frac{\pi}{32},0 \right)\) and plotted the resulting orbit
+in Figure \ref{fig:acrobot-orbit}.
+
+The set \(E_\pi := E(\pi,0)\) is outlined in black. 
+Because of model mismatch, \(E_\pi\) is no longer the boundary between
+rotations and oscillations.
+The oscillation domain is actually much larger: orbits rotate once they hit the
+\(p_u\)-axis at \(\abs{p_u} \approx 0.15\), while \(E_\pi\) intersects the
+\(p_u\)-axis at \(\abs{p_u} \approx 0.055\).
+Despite this, the VNHC still injects energy into the acrobot.
+The points where the orbit exits \(E_\pi\) are marked with black stars,
+with the final departure marked with a red star.
+After this final departure, the acrobot begins rotating and continues to gain
+energy over time.
+
+\begin{figure}[h]
+    \centering
+    \begin{subfigure}[t]{0.45\linewidth}
+        \includegraphics[width=\linewidth]{acrobot_orbit.png}
+        \caption{Orbit initialized at \((q_u,p_u) = (\pi/32,0)\).}
+        \label{fig:acrobot-orbit}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.45\linewidth}
+        \includegraphics[width=\linewidth]{acrobot_mc.png}
+        \caption{Monte Carlo results.}
+        \label{fig:acrobot-mc}
+    \end{subfigure}
+    \caption{Simulations using Wang's acrobot \cite{xingbo_thesis}.}
+\end{figure}
+
+To verify that the acrobot would always achieve rotations, we ran a
+Monte-Carlo simulation \cite{montecarlo}:
+we initialized the acrobot randomly inside the sublevel set
+\[
+    \left\{(q_u,p_u) \in \SxR \mid
+    E(q_u,p_u) \leq E\left(\frac{\pi}{32},0\right)\right\}
+    ,
+\] 
+and measured the how long it took to begin rotating.
+The results in Figure \ref{fig:acrobot-mc} show that
+the acrobot always rotated within 10--35 seconds.
+
+\subsection{Physical Experiments}
+Wang's acrobot has an encoder at the pivot which measures \(q_u\) and
+\(\dot{q}_u\).
+The leg actuator is a servo motor with a built-in PID controller, which provides
+measurements of \(q_a\).
+We estimate \(\dot{q}_a\) through sequential values of \(q_a\) and
+compute \(p_u\) through \(p_u = e_1^T M(q) \dot{q}\).
+We then assign the actuator value at iteration \(k \in \mathbb{Z}_{> 0}\)
+via \(q_a^{k} = \arctan(I p_u^{k-1})\).
+
+\textbf{Baseline Test:} 
+for this test, we initialized the acrobot at 
+\((q_u,p_u) \approx \left(\frac{\pi}{8},0\right)\). 
+The resulting orbit is shown in Figure \ref{fig:acrobot-unperturbed-orbit};
+the acrobot is clearly gaining energy over time.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[t]{0.45\linewidth}
+        \includegraphics[width=\linewidth]{acrobot_unperturbed_orbit.png}
+        \caption{The baseline test.}
+        \label{fig:acrobot-unperturbed-orbit}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.45\linewidth}
+        \includegraphics[width=\linewidth]{acrobot_stopped_orbit.png}
+        \caption{Acrobot orbit before (blue) and after (red) stopping.}
+        \label{fig:acrobot-stopped-orbit}
+    \end{subfigure}
+    \caption{The physical acrobot's orbit during baseline and stop tests.}
+\end{figure}
+
+\textbf{Stop Test:}
+for this test we initialized the acrobot at 
+\((q_u,p_u) \approx \left(\pi,0\right)\), let it run for 15
+seconds, then stopped it at the bottom of its arc.
+The resulting orbit is shown in Figure \ref{fig:acrobot-stopped-orbit}.
+The blue curves correspond to the orbit before the disturbance, while the red
+spiral confirms that the acrobot begins oscillating after it is stopped.
+Despite the disturbance, it gains energy and eventually starts rotating
+again.
+
+\textbf{Push Test:}
+to see how the acrobot responds when pushed, we allowed
+the acrobot to rotate undisturbed for 15 seconds and then pushed it in its
+direction of motion.
+The orbit in Figure \ref{fig:acrobot-fpush-orbit} shows that the acrobot speeds
+up to rotate with energy \(E(0,0.22)\), but then slows down until it reaches a
+stable rotation with energy \(E(0,-0.195)\).
+We repeated this test by pushing the acrobot against its direction of motion.
+The orbit in Figure \ref{fig:acrobot-rpush-orbit} demonstrates that the acrobot
+responds by readily changing direction, and quickly achieves its maximum speed
+with energy \(E(0,0.195)\).
+
+These push tests suggest that our VNHC injects energy into Wang's acrobot on
+\(\mathcal{O}_2(0.195)\).
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[ht]{0.45\linewidth}
+        \includegraphics[width=\linewidth]{acrobot_fpush_orbit.png}
+        \caption{The forwards push test.}
+        \label{fig:acrobot-fpush-orbit}
+    \end{subfigure}
+    \begin{subfigure}[ht]{0.45\linewidth}
+        \includegraphics[width=\linewidth]{acrobot_rpush_orbit.png}
+        \caption{The reverse push test.}
+        \label{fig:acrobot-rpush-orbit}
+    \end{subfigure}
+    \caption{The acrobot orbit before (blue) and after (red) pushing.}
+\end{figure}
+
+\subsection{Summary of Experiments}
+These simulations and experiments demonstrate that VNHCs are excellent
+tools for injecting energy, even with significant model mismatch.
+In particular, energy injection through VNHCs appears to be robust against a
+variety of disturbances.
+
 %/========== Proof ==========/%
@@
 \section{Conclusion}\label{sec:conclusion}
+In this article we summarized the existing theory of virtual nonholonomic
+constraints (VNHCs) ........
+\textbf{TODO conclude nicely}
+In the end, we demonstrated that virtual nonholonomic constraints are capable of
+injecting and dissipating energy in a robust manner, all while producing
+realistic biological motion.
 

### 2021-04-01 latex/paper/main.tex
Edit of acrobot and experiment section

@@
 %injection for giant stabilization)
-Our goal in this article isto design a VNHC which injects energy into the
+Our goal in this article is to design a VNHC which injects energy into the
 acrobot through giant-like motion.
-We use the simplified acrobot model acrobot in Figure
-\ref{fig:simple-acrobot-model}, where we assume the torso and leg rods
-are of equal length \(l\) with equal point masses \(m\) at the tips.
+We will use the simplified acrobot model in Figure
+\ref{fig:simple-acrobot-model}, where we assume the torso and leg rods are of
+equal length \(l\) with equal point masses \(m\) at the tips.
 A real gymnast cannot swing their legs in full circles, though they
@@
 \(p_a\), representing a torque acting on the hip joint.
-This means \((q,p)\) are simply actuated coordinates inside the phase space
+This means \((q,p)\) are simply actuated coordinates in the phase space
 \(\mathcal{Q} \times \mathcal{P}\) where
@@
 \(\mathcal{P} = \R \times \R\).
-We can therefore apply the theory of VNHCs from Section \ref{sec:vnhc} to design
-a constraint of the form \(q_a = f(q_u,p_u)\). 
-This constraint must inject energy into the acrobot according to
-Definition \ref{defn:energy-injection}.
+We can therefore apply the theory from Section \ref{sec:vnhc} to design
+a VNHC of the form \(q_a = f(q_u,p_u)\). 
 Since we need the VNHC to be regular, the following proposition will prove
@@
     with \(f \in C^2\left(\R; \Sone\right)\) is a regular
-    VNHC of order 1 for the simple acrobot.
+    VNHC of order 1 for the acrobot.
 \end{prop}
 \begin{proof}
-    The regularity matrix for this VNHC evaluates to
-    \(\frac{(1+c_a)\partial_{q_u}f(q_u,p_u) + (3+2c_a)}{ml^2(2-c_a^2)}\).
+    The regularity matrix for the acrobot evaluates to
+    \(((1+c_a)\partial_{q_u}f(q_u,p_u) + (3+2c_a))/(ml^2(2-c_a^2))\).
     Since \(\partial_{q_u} f = 0\), this matrix is strictly positive for all values
@@
 legs when the swing moves backwards.
-As the swing gains speed, the person leans their body back while
+As the swing gains speed, the person leans their body while
 extending their legs higher, thereby shortening the distance
@@
 Recall that \((q_u, p_u)\) denote the angle and momentum of the acrobot's torso.
-By Theorem \ref{thm:zero-dynamics}, we can study the constrained system by
-examining solely the motion of \((q_u,p_u) \in \SxR\).
+By Theorem \ref{thm:zero-dynamics}, the constrained acrobot is parameterized
+fully by \((q_u,p_u) \in \SxR\).
 It is easy to show that the constrained dynamics are
+just the torso dynamics obtained when one swings the legs
+according to \eqref{eqn:acrobot-constraint}; that is, they are
 \begin{equation}\label{eqn:acrobot-constrained-dynamics}
@@
 subject to \(q_a = \bar{q}_a \arctan(I p_u)\).
-These are just the torso dynamics when one swings the legs
-according to \eqref{eqn:acrobot-constraint}.
 
@@
 \(\mathcal{O}_2(\bar{\rho})\).
-If \(I\) is too large, the leg controller saturates (i.e. the legs swing up too
-quickly) and the body may not swing higher. 
-However, choosing \(I\) small enough guarantees the legs will synchronize
-properly with the body, and the acrobot will begin doing backflips on the bar.
+If \(I\) is too large, the leg controller saturates and the body oscillates
+without gaining energy.
+Choosing \(I\) small enough guarantees the legs will synchronize
+properly with the body, and the acrobot will begin rotating around the bar.
 The following theorem provides conditions under which such an \(I\) exists.
@@
 \begin{thm}\label{thm:acrobot-energy-stabilization}
-    Consider the acrobot with Hamiltonian dynamics, as in ~\eqref{eqn:acrobot-hamiltonian}.
+    Consider the acrobot with Hamiltonian dynamics as in ~\eqref{eqn:acrobot-hamiltonian}.
 \begin{enumerate}
@@
 The value of this integral depends on the acrobot's physical parameters.
-If the assumption holds, there is some control value \(I > 0\)
+If the assumption holds, there is some control value
 (which depends on \(\bar{\rho}\)) for which the acrobot will 
@@
 \section{Experimental Results}\label{sec:experiments}
-In this section, we test our VNHC on the physical acrobot built by
+We tested our VNHC on the physical acrobot built by
 Xingbo Wang \cite{xingbo_thesis} (Figure \ref{fig:xingbo-acrobot}).
-The torso and leg rods on this robot have unequal mass and differ in length, so
-its inertia matrix and potential function are not described correctly by 
+The torso and legs on this robot have differing masses and lengths, so
+its inertia matrix and potential function are \textbf{not} described by 
 ~\eqref{eqn:acrobot-inertia}-\eqref{eqn:acrobot-potential}.
-These experiments will therefore show that VNHCs are robust to model mismatch.
+If our VNHC injects energy into this acrobot, then we can infer that
+VNHCs are robust to model mismatch.
 
@@
 
-The control parameter \(I\) must be small to satisfy Theorem
-\ref{thm:acrobot-energy-stabilization}, but it must also be large enough to
-overcome friction in the servo motor.
-We experimentally determined that \(I = 10\) is sufficient to
-ensure the actuator can fully rotate in its allowable range
-\(q_a \in \left[ -\frac{\pi}{2}, \frac{\pi}{2}\right]\).
+For our VNHC to inject energy, we know the control parameter \(I\) must be small
+(as per Theorem \ref{thm:acrobot-energy-stabilization}), but it must also be
+large enough to overcome friction in leg motor.
+Setting \(\bar{q}_a = 1\), we experimentally determined that \(I = 10\) is
+sufficient to ensure the actuator can fully rotate in its allowable range \(q_a
+\in \left[ -\frac{\pi}{2}, \frac{\pi}{2}\right]\). 
+Rhis is the value of \(I\) we used in all simulations and experiments.
 
@@
 \]
-We simulated Wang's acrobot starting at 
-\((q_u,p_u) = \left(\frac{\pi}{32},0 \right)\) and plotted the resulting orbit
+We simulated Wang's acrobot with initial condition
+\((q_u,p_u) = \left(\pi/32,0 \right)\) and plotted the resulting orbit
 in Figure \ref{fig:acrobot-orbit}.
@@
 The set \(E_\pi := E(\pi,0)\) is outlined in black. 
-Because of model mismatch, \(E_\pi\) is no longer the boundary between
+Due to the model mismatch, \(E_\pi\) is no longer the boundary between
 rotations and oscillations.
@@
 
-\begin{figure}[h]
+\begin{figure}[]
     \centering
-    \begin{subfigure}[t]{0.45\linewidth}
-        \includegraphics[width=\linewidth]{acrobot_orbit.png}
-        \caption{Orbit initialized at \((q_u,p_u) = (\pi/32,0)\).}
+    \includegraphics[width=0.7\linewidth]{acrobot_orbit.png}
+    \caption{A simulated orbit of Wang's acrobot.}
     \label{fig:acrobot-orbit}
-    \end{subfigure}
-    \begin{subfigure}[t]{0.45\linewidth}
-        \includegraphics[width=\linewidth]{acrobot_mc.png}
-        \caption{Monte Carlo results.}
-        \label{fig:acrobot-mc}
-    \end{subfigure}
-    \caption{Simulations using Wang's acrobot \cite{xingbo_thesis}.}
 \end{figure}
-
-To verify that the acrobot would always achieve rotations, we ran a
-Monte-Carlo simulation \cite{montecarlo}:
-we initialized the acrobot randomly inside the sublevel set
+To verify numerically that the acrobot would consistently achieve rotations, we
+ran a Monte-Carlo simulation \cite{montecarlo} where we initialized the acrobot
+randomly inside the sublevel set
 \[
@@
 
+\begin{figure}[]
+    \centering
+    \includegraphics[width=0.7\linewidth]{acrobot_mc.png}
+    \caption{Monte Carlo simulation results.}
+    \label{fig:acrobot-mc}
+\end{figure}
+
 \subsection{Physical Experiments}
-Wang's acrobot has an encoder at the pivot which measures \(q_u\) and
+
+Wang's acrobot has an encoder which measures \(q_u\) and
 \(\dot{q}_u\).
-The leg actuator is a servo motor with a built-in PID controller, which provides
-measurements of \(q_a\).
-We estimate \(\dot{q}_a\) through sequential values of \(q_a\) and
-compute \(p_u\) through \(p_u = e_1^T M(q) \dot{q}\).
-We then assign the actuator value at iteration \(k \in \mathbb{Z}_{> 0}\)
-via \(q_a^{k} = \arctan(I p_u^{k-1})\).
-
-\textbf{Baseline Test:} 
-for this test, we initialized the acrobot at 
+Additionally, the leg actuator is a servo motor with a built-in PID controller
+which provides measurements of \(q_a\).
+To implement our VNHC, we estimated \(\dot{q}_a\) through sequential values of
+\(q_a\) and computed \(p_u\) through \(p_u = e_1^T M(q) \dot{q}\).
+We directly assigned the actuator setpoint at iteration \(k \in \mathbb{Z}_{> 0}\)
+via \(q_a^{k} = \arctan(10 p_u^{k-1})\).
+We then performed the following tests.
+
+\begin{enumerate}
+    \item \textbf{Baseline Test:} 
+    we initialized the acrobot at 
     \((q_u,p_u) \approx \left(\frac{\pi}{8},0\right)\). 
@@
 
-\textbf{Stop Test:}
-for this test we initialized the acrobot at 
+\item \textbf{Stop Test:}
+    we initialized the acrobot at 
     \((q_u,p_u) \approx \left(\pi,0\right)\), let it run for 15
@@
 
-\textbf{Push Test:}
+\item \textbf{Push Test:}
     to see how the acrobot responds when pushed, we allowed
@@
     up to rotate with energy \(E(0,0.22)\), but then slows down until it reaches a
-stable rotation with energy \(E(0,-0.195)\).
+    rotation with energy \(E(0,-0.195)\).
     We repeated this test by pushing the acrobot against its direction of motion.
@@
     \(\mathcal{O}_2(0.195)\).
+\end{enumerate}
 
@@
 
-\subsection{Summary of Experiments}
+\subsection{Summary of Results} 
 These simulations and experiments demonstrate that VNHCs are excellent
-tools for injecting energy, even with significant model mismatch.
+tools for injecting energy.
 In particular, energy injection through VNHCs appears to be robust against a
-variety of disturbances.
+variety of disturbances, including significant model mismatch.
 

### 2021-04-01 latex/paper/main.tex
Added the physical acrobot photo to the paper
Added a comma to the authors list

@@
 \title{\Title}
-\author{Adan Moran-MacDonald, \IEEEmembership{Member, IEEE}, Manfredi Maggiore
+\author{Adan Moran-MacDonald, \IEEEmembership{Member, IEEE}, Manfredi Maggiore,
 \IEEEmembership{Member, IEEE}*, and Xingbo Wang
@@
     \centering
-    %\includegraphics[width=0.5\linewidth]{}
-    \caption{\textbf{TODO: take a square photo of the acrobot}The acrobot built by Wang \cite{xingbo_thesis}.}
+    \includegraphics[width=0.6\linewidth]{xingbo_acrobot.jpg}
+    \caption{The acrobot built by Wang \cite{xingbo_thesis}.}
     \label{fig:xingbo-acrobot}

### 2021-04-09 latex/ogs/main.tex
OGS 2021 update

@@
 
-When a gymnast is performing a swing-up routine on a bar, they do not kick their
-feet at specified time intervals. Rather, a more natural human motion is to
-smoothly swing their legs based on their current speed and position
-\cite{pendulum_length_giant_gymnastics}. In doing so, they are able to generate
-enough momentum to perform backflips on the bar. When controlling mechanical
-systems, standard practice is to have a robot's actuators track a function of
-time. Much like the gymnast, it can be more natural to have the robot track a
-function of speed and position instead. The recent techniques of virtual
-constraints offer this ability to produce realistic, biologically plausible
-motion.
+When a gymnast on a horizontal bar wants to swing up to do backflips, they
+typically perform a technique called a ``giant".
+This technique involves the gymnast swinging their legs based on their current
+speed and position, thereby generating enough momentum to rotate around the bar
+\cite{pendulum_length_giant_gymnastics}.
 
-Mechanical systems can be constrained in two ways. Holonomic constraints
-restrict position (for instance, car tires remain a fixed distance apart), while
-nonholonomic constraints restrict velocities (such as how car tires cannot
-slide sideways). Previous research has shown that one can enforce holonomic
-constraints through a robot's actuators, using the method of "virtual holonomic
-constraints" (VHCs) \cite{vhcs_for_el_systems}. VHCs have been used to control
-walking robots \cite{stable-walking}, autonomous bicycles \cite{bicycle},
-gymnastics robots \cite{xingbo-thesis}, and snake robots \cite{snake-robot}
-among others. These VHCs are a special case of the more general virtual
-nonholonomic constraints (VNHCs). In my master's thesis \cite{my-thesis} I
-developed a theory for enforcing VNHCs and designed one which injects energy
-into the gymnastics robot created by Wang \cite{xingbo-thesis}, thereby enabling
-it to perform backflips on a bar.
+Standard practice in robotics begins with finding a trajectory over time
+replicating a desired behaviour, then using the robot's actuators to track this
+time-based trajectory. 
+While this often works for simple systems, it can cause timing delays, it is not
+robust to external disturbances, and it is not what humans do.
+A more natural approach is to constrain the robot's actuators according to
+some function of speed and position.
+This type of behaviour is well described by the method of virtual constraints.
+
+There are two types of constraints on a mechanical system:
+\textit{holonomic} constraints restrict position (e.g. a snake robot can slither
+but not fly),
+while \textit{nonholonomic} constraints restrict position and
+velocity (e.g. cars cannot slide sideways).
+One can design a specific constraint for a robot and enforce it using the
+actuators.
+Since this constraint is not enforced by physics, it is known as a ``virtual"
+constraint
+\cite{vhcs_for_el_systems}.
+Virtual holonomic constraints (VHCs) have been used to control walking robots
+\cite{stable-walking}, autonomous bicycles \cite{bicycle},
+gymnastics robots \cite{xingbo-thesis}, and snake robots \cite{snake-robot}.
+Virtual nonholonomic constraints (VNHCs) have been used for human-robot
+cooperation \cite{vnhc-human-robot-coop} and for improving robustness of walking
+robots \cite{vnhc-biped-robot,hybrid_zero_dynamics_bipedal_nhvcs}, among other
+applications.
 
@@
 
-Mohammadi \textit{et.al.} \cite{manfredi-orbit-stabilization} showed that a
-single VHC can result in several motion patterns (known as ``orbits"), from
-which one specific motion pattern can be chosen if the robot is initialized with a
-configuration that is ``close enough" to it. However, their method does not
-work for all starting configurations of the robot. We propose to extend this
-work by investigating a method for transitioning between VNHCs.
+Virtual constraints have proven useful for generating biologically-inspired
+behaviour in robotic systems.
+In my master's thesis, I demonstrated that VNHCs can also inject energy into
+robots:
+I designed a VNHC which enabled a gymnastics robot to gain energy
+and perform backflips on a horizontal bar \cite{my-thesis}.
+Using VNHCs for \textit{energy injection} is beneficial because it allows robots to
+safely increase their momentum.
+Likewise, designing VNHCs for \textit{energy dissipation} allows robots to
+safely slow down in a realistic and reliable manner.
 
-VNHCs have been studied with applications to human-robot cooperation
-\cite{vnhc-human-robot-coop} and walking robots \cite{vnhc-biped-robot}. Horn
-\textit{et.al.} \cite{hybrid_zero_dynamics_bipedal_nhvcs} have also examined
-properties of VNHCs for general mechanical systems and applied their results to
-stabilizing a chosen gait on walking robots. Our approach is novel since we 
-assume one can stabilize two distinct VNHCs independently, and will develop
-techniques to transition between them while maintaining safety constraints.
-Doing so will allow one to chain together VNHCs which drive a robot's
-configuration towards a desired orbit, regardless of the initial configuration.
-
-First we will study mathematical conditions under which one can transition
-between two VNHCs. These conditions will allow us to create a control algorithm
-to switch between motion patterns. After completing this mathematical
-formulation, we will test the theory on a gymnastics robot by designing a
-gymnastics routine which is performed solely by transitioning between several different
-VNHCS.
+We propose to study the energy injection and dissipation properties of VNHCs.
+First, we will find mathematical conditions under which VNHCs are guaranteed to
+inject/dissipate energy in abstract mechanical systems.
+Then we will design a means of safely transitioning between two different VNHCs.
+Finally, we will test the theory on a gymnastics robot by designing a gymnastics
+routine which is performed solely by transitioning between different virtual
+constraints.
 
 The advancements in mathematics from this research will bring improvements to
-the control of other autonomous systems. The ability to transition between
-complex constraints will allow for more expressive motion and may become
-a standard technique for controlling biologically inspired robots.
+the control of autonomous systems. 
+The ability to safely transition between complex constraints by adding or
+removing energy will allow for more expressive motion, and may become a standard
+technique for controlling biologically inspired robots.
 

### 2021-04-19 latex/paper/main.tex
Edit before sending to Manfredi

@@
 Given two matrices \(A \in \R^{n \times m}\) and \(B \in \R^{r \times s}\), the
-Kronecker product \cite{kronprod} \(A \otimes B \in \R^{nr \times ms}\) is the
-matrix
+Kronecker product \(A \otimes B \in \R^{nr \times ms}\) (see \cite{kronprod}) is
+the matrix
 \begin{equation}\label{eqn:kronprod}
@@
 angle and velocity.
-In 2016, Wang attempted to design a controller which does this \cite{xingbo_thesis}
+In 2016, Wang attempted to design such a controller \cite{xingbo_thesis}
 but was unable to prove the acrobot would gain energy over time.
@@
 human-like giant motion.
+
+\textbf{MANFREDI: This section feels more like an intro than a problem
+formulation. Thoughts?}
 
@@
     \(V \in \R^{k \times k}\) are unitary matrices \cite{calculating_svd}.
-    Defining \(T = \diag{1/\sigma_1^2,\ldots,1/\sigma_k^2}\) and assigning the
+    Defining \(T = \diag{1/\sigma_1,\ldots,1/\sigma_k}\) and assigning the
     regular feedback transformation \(\tau = V T \hat{\tau}\) yields a new input
@@
         = \delta_i^j\).
-    By (45.10) in \cite{landau_mechanics}, this transformation is canonical and
-    the new Hamiltonian is
+    By (45.10) in \cite{landau_mechanics}, this is a canonical transformation
+    and the new Hamiltonian is
     \(\mathcal{H}(\mathbf{B}\tpose \tilde{q}, \mathbf{B}\tpose \tilde{p})\).
@@
 
-The notion of a nonholonomic constraint which can be stabilized via state feedback
-was first described by Griffin and Grizzle in \cite{nhvc_dynamic_walking}.
-Horn et. al later extended their results in
-\cite{hybrid_zero_dynamics_bipedal_nhvcs} and
-derived the dynamics of constrained systems.
-In this section we reformulate these ideas for the Hamiltonian framework,
-because the theory is cleaner when using unactuated and actuated momenta.
-For this reason, we take the system of inquiry to be a Hamiltonian
-mechanical system in simply actuated coordinates, as in
+Griffin and Grizzle \cite{nhvc_dynamic_walking} were the first to define
+relative degree two nonholonomic constraints which can be enforced
+through state feedback.
+Horn et. al later extended their results
+\cite{hybrid_zero_dynamics_bipedal_nhvcs} to derive the constrained dynamics for
+a certain class of mechanical systems.
+
+These researchers made use of the unactuated conjugate of momentum, but they
+developed their results in the Lagrangian framework.
+In particular, they focused on Lagrangian systems with degree of underactuation
+one.
+In this section we reformulate their ideas into the Hamiltonian framework and
+extend the definition of a virtual nonholonomic constraint to systems with
+higher degree of underactuation.
+This reformulation will provide us with better intuition and a more explicit form of the
+constrained dynamics.
+
+For the rest of this section we take the system of inquiry to be a
+Hamiltonian mechanical system in simply actuated coordinates, as in
 \eqref{eqn:simple-hamiltonian}.
-For simplicity of notation, we relabel \((\tilde{q},\tilde{p})\) as \((q,p)\).
+For simplicity of notation, we relabel \((\tilde{q},\tilde{p})\) to \((q,p)\).
 
@@
 a VNHC of the form \(q_a = f(q_u,p_u)\). 
-Since we need the VNHC to be regular, the following proposition will prove
+Since we need the VNHC to be regular, the following proposition will be
 useful.
@@
 This constraint does not perfectly recreate giant motion, during which
-the gymnast's legs are almost completely extended \cite{usagym_giant}:
-it instead pivots the legs partially during rotations.
-However, the behaviour looks similar enough that this constraint should 
+the gymnast's legs are almost completely extended \cite{usagym_giant}.
+It instead pivots the legs partially during rotations.
+However, the behaviour looks similar enough that this constraint will
 still inject energy into the acrobot.
@@
 just the torso dynamics obtained when one swings the legs
-according to \eqref{eqn:acrobot-constraint}; that is, they are
+according to \eqref{eqn:acrobot-constraint}:
 \begin{equation}\label{eqn:acrobot-constrained-dynamics}
@@
 the bar.
-In the worst case, the acrobot will at least perform a swing-up routine to reach
-the unstable equilibrium at \((q_u,p_u) = (\pi,0)\).
+%In the worst case, the acrobot will at least perform a swing-up routine to reach
+%the unstable equilibrium at \((q_u,p_u) = (\pi,0)\).
 
@@
 The value of this integral depends on the acrobot's physical parameters.
-If the assumption holds, there is some control value
+If the assumption holds, there is some control value \(I\)
 (which depends on \(\bar{\rho}\)) for which the acrobot will 
@@
 \in \left[ -\frac{\pi}{2}, \frac{\pi}{2}\right]\). 
-Rhis is the value of \(I\) we used in all simulations and experiments.
+This is the value of \(I\) we used in all simulations and experiments.
 
@@
 The points where the orbit exits \(E_\pi\) are marked with black stars,
-with the final departure marked with a red star.
+with the final departure marked by a red star.
 After this final departure, the acrobot begins rotating and continues to gain
@@
     The orbit in Figure \ref{fig:acrobot-fpush-orbit} shows that the acrobot speeds
-    up to rotate with energy \(E(0,0.22)\), but then slows down until it reaches a
+    up to rotate with energy \(E(0,-0.22)\), but then slows down until it reaches a
     rotation with energy \(E(0,-0.195)\).

### 2021-04-26 sugar_bot/arduino/libraries/SUGAR_VNHC_Library/include/Supervisor.h
Created the base code for the supervisor, including CMake and a test

@@ -1,0 +1,35 @@
+/*******************************************************************************
+* File:	        Supervisor.h
+*		 
+* Author:       Adan Moran-MacDonald
+* Created:      26/Apr/21
+*
+* Systems Control Group
+* Department of Electrical and Computer Engineering
+* University of Toronto
+*
+* Last Modified: 26/Apr/21
+* Last Editor:   Adan Moran-MacDonald
+* Description:   A simple supervisor for switching between injection and
+* dissipation VNHCs.
+*******************************************************************************/
+#ifndef __SUPERVISOR_H__
+#define __SUPERVISOR_H__
+
+namespace SUGAR
+{
+
+class Supervisor
+{
+public:
+	Supervisor (); 
+	virtual ~Supervisor ();
+
+private:
+	/* data */
+}; // class Supervisor
+
+} // namespace SUGAR
+
+#endif // __SUPERVISOR_H__
+/* vim: set tw=80 ts=4 sw=4 sts=0 et ffs=unix : */

### 2021-04-26 sugar_bot/arduino/libraries/SUGAR_VNHC_Library/src/Supervisor.cpp

@@ -1,0 +1,28 @@
+/*******************************************************************************
+* File:	        Supervisor.cpp
+*		 
+* Author:       Adan Moran-MacDonald
+* Created:      26/Apr/21
+*
+* Systems Control Group
+* Department of Electrical and Computer Engineering
+* University of Toronto
+*
+* Last Modified: 26/Apr/21
+* Last Editor:   Adan Moran-MacDonald
+* Description:   Implementation of Supervisor.h
+*******************************************************************************/
+
+#include "include/Supervisor.h"
+
+namespace SUGAR
+{
+
+Supervisor::Supervisor()
+{}
+
+Supervisor::~Supervisor()
+{}
+
+}; // namespace SUGAR
+/* vim: set tw=80 ts=4 sw=4 sts=0 et ffs=unix : */

### 2021-04-26 sugar_bot/arduino/libraries/SUGAR_VNHC_Library/src/test/SupervisorTest.cpp

@@ -1,0 +1,97 @@
+/*******************************************************************************
+* File:	        SupervisorTest.cpp
+*		 
+* Author:       Adan Moran-MacDonald
+* Created:      26/Apr/21
+*
+* Systems Control Group
+* Department of Electrical and Computer Engineering
+* University of Toronto
+*
+* Last Modified: 26/Apr/21
+* Last Editor:   Adan Moran-MacDonald
+* Description:   Test suite for Supervisor.h
+*******************************************************************************/
+
+// Includes for our code
+#include "StateSpace.h"
+#include "AcrobotDynamics.h"
+#include "Phase.h"
+#include "VNHC.h"
+#include "Supervisor.h"
+// Include predicates
+#include "test/Predicates.h"
+// Includes for M_PI, cosine, etc
+#define _USE_MATH_DEFINES
+#include <math.h>
+#undef _USE_MATH_DEFINES
+// Includes for std library
+#include <iostream>
+// Includes for gtest
+#include "gtest/gtest.h"
+
+using namespace SUGAR;
+using SUGAR::predicates::Within;
+
+class SupervisorTest : public ::testing::Test
+{
+public:
+    SupervisorTest()
+    : xingbot_(
+        mt_, ml_,
+        dt_, dl_,
+        lt_, ll_,
+        Jt_, Jl_,
+        g_),
+      in_(xingbot_,qmax_,I_),
+      diss_(xingbot_,qmax_,-I_),
+      sup_() // Todo: implement this correctly
+    {}
+protected:
+    void SetUp() override 
+    {
+
+    }
+
+    void TearDown() override
+    {}
+protected:
+    // constants for Xingbo's acrobot
+    double mt_ = 0.2112;  double ml_ = 0.1979;
+    double dt_ = 0.148;   double dl_ = 0.145;
+    double lt_ = 0.073;   double ll_ = 0.083;
+    double Jt_ = 0.00075; double Jl_ = 0.00129;
+    double g_ = 9.81;
+
+    // Constant control input for the VNHC
+    double I_ = 10;
+    
+    // error margin for predicates
+    double eps_ = 10e-15;
+
+    // Dynamics for Xingbo's bot
+    Acrobot xingbot_;
+
+    // Define a pointer to the VNHC
+    AcrobotVNHC* pVNHC_;
+
+    // Define an unactuatedPhase for testing the objects
[…]

### 2021-04-26 sugar_bot/arduino/libraries/SUGAR_VNHC_Library/include/Supervisor.h
Initialized the stabilize() function
This function will stabilize the desired value using the injection or
dissipation VNHCs.

Wrote the first test to ensure that the stabilization mechanism returns
qa = 0 whenever your value is within the hysteresis of the desired
value.

@@ -16,17 +16,40 @@
 #ifndef __SUPERVISOR_H__
 #define __SUPERVISOR_H__
 
+#include "VNHC.h"
+
 namespace SUGAR
 {
 
 class Supervisor
 {
 public:
-	Supervisor (); 
+    /**
+     * @brief: Constructor which produces a supervisor that toggles between the
+     * given injection and dissipation VNHCs. These do not have to be the same.
+     */
+	Supervisor(const AcrobotVNHC& in, const AcrobotVNHC& diss); 
 	virtual ~Supervisor ();
 
+    /**
+     * @brief: Function to stabilize a given nonnegative value.
+     *
+     * @param: double the current value of the acrobot that you are trying to
+     * stabilize (examples are qu, pu, or E)
+     * @param: double the desired value of the acrobot which you want the first
+     * input to reach. This must be >= 0.
+     * @param: double the hysteresis; the supervisor will set qa = 0 whenever
+     * abs(abs(value) - value_des) <= hys.
+     *
+     * @return: double the value of qa associated with the correct VNHC
+     */
+    auto stabilize(double val, double des, double hys) const -> double;
+
 private:
-	/* data */
+    // The injection and dissipation VNHCs
+    const AcrobotVNHC& in_;
+    const AcrobotVNHC& diss_;
+
 }; // class Supervisor
 
 } // namespace SUGAR

### 2021-04-26 sugar_bot/arduino/libraries/SUGAR_VNHC_Library/src/Supervisor.cpp

@@ -14,15 +14,28 @@
 *******************************************************************************/
 
 #include "include/Supervisor.h"
+#include <math.h>
 
 namespace SUGAR
 {
 
-Supervisor::Supervisor()
+Supervisor::Supervisor(const AcrobotVNHC& in, const AcrobotVNHC& diss)
+: in_(in),
+  diss_(diss)
 {}
 
 Supervisor::~Supervisor()
 {}
+
+auto Supervisor::stabilize(double value, double des, double hys) const -> double
+{
+    if(fabs(fabs(value)-des) <= hys)
+    {
+        return 0;
+    }
+
+    return 1.0;
+}
 
 }; // namespace SUGAR
 /* vim: set tw=80 ts=4 sw=4 sts=0 et ffs=unix : */

### 2021-04-26 sugar_bot/arduino/libraries/SUGAR_VNHC_Library/src/test/SupervisorTest.cpp

@@ -45,7 +45,7 @@ public:
         g_),
       in_(xingbot_,qmax_,I_),
       diss_(xingbot_,qmax_,-I_),
-      sup_() // Todo: implement this correctly
+      sup_(in_,diss_) // Todo: implement this correctly
     {}
 protected:
     void SetUp() override 
@@ -87,6 +87,22 @@ protected:
     // Define our supervisor, which takes in the two VNHCs
     Supervisor sup_;
 }; // class VNHCTest
+
+// Make sure if abs(value - des) <= hys, that the supervisor returns qa = 0
+TEST_F(SupervisorTest, ZERO_WITHIN_HYS)
+{
+    double des = M_PI/2;
+    double hys = 0.2;
+    // check that if value is in [des-hys, des+hys] that we get qa = 0
+    EXPECT_EQ(sup_.stabilize(des-hys,des,hys),0);
+    EXPECT_EQ(sup_.stabilize(des,des,hys),0);
+    EXPECT_EQ(sup_.stabilize(des+hys,des,hys),0);
+
+    // Check that if value is in [-des-hys,-des+hys] that we get qa = 0
+    EXPECT_EQ(sup_.stabilize(-des-hys,des,hys),0);
+    EXPECT_EQ(sup_.stabilize(-des,des,hys),0);
+    EXPECT_EQ(sup_.stabilize(-des+hys,des,hys),0);
+}
 
 int main(int argc, char** argv)
 {

### 2021-04-26 sugar_bot/arduino/libraries/SUGAR_VNHC_Library/include/Supervisor.h
Completed the Supervisor class and its tests, now we need oscillation/rotation/energy supervisors

@@ -32,18 +32,24 @@ public:
 	virtual ~Supervisor ();
 
     /**
-     * @brief: Function to stabilize a given nonnegative value.
+     * @brief: Function to stabilize a given nonnegative value to a desired
+     * nonnegative value, within some hysteresis.
      *
+     * @param: const UnactuatedPhase&. The current unactuated phase for the
+     * acrobot.
      * @param: double the current value of the acrobot that you are trying to
      * stabilize (examples are qu, pu, or E)
      * @param: double the desired value of the acrobot which you want the first
      * input to reach. This must be >= 0.
-     * @param: double the hysteresis; the supervisor will set qa = 0 whenever
-     * abs(abs(value) - value_des) <= hys.
+     * @param: double the hysteresis. This must be >= 0. the supervisor will set
+     * qa = 0 whenever abs(abs(value) - value_des) <= hys.
      *
      * @return: double the value of qa associated with the correct VNHC
      */
-    auto stabilize(double val, double des, double hys) const -> double;
+    auto stabilize( const UnactuatedPhase& qpu, 
+                    double value, 
+                    double des, 
+                    double hys) const -> double;
 
 private:
     // The injection and dissipation VNHCs

### 2021-04-26 sugar_bot/arduino/libraries/SUGAR_VNHC_Library/src/Supervisor.cpp

@@ -27,14 +27,28 @@ Supervisor::Supervisor(const AcrobotVNHC& in, const AcrobotVNHC& diss)
 Supervisor::~Supervisor()
 {}
 
-auto Supervisor::stabilize(double value, double des, double hys) const -> double
+auto Supervisor::stabilize(const UnactuatedPhase& qpu, 
+                           double value, 
+                           double des, 
+                           double hys) const -> double
 {
-    if(fabs(fabs(value)-des) <= hys)
-    {
-        return 0;
-    }
+    // Make sure everything is positive.
+    des = fabs(des);
+    value = fabs(value);
+    hys = fabs(hys);
 
-    return 1.0;
+    // injection
+    if (value < des - hys)
+    {
+        return in_.qa(qpu);
+    }
+    // dissipation
+    else if (value > des + hys)
+    {
+        return diss_.qa(qpu);
+    }
+    // stability
+    return 0;
 }
 
 }; // namespace SUGAR

### 2021-04-26 sugar_bot/arduino/libraries/SUGAR_VNHC_Library/src/test/SupervisorTest.cpp

@@ -86,6 +86,8 @@ protected:
 
     // Define our supervisor, which takes in the two VNHCs
     Supervisor sup_;
+    // A default unactuated phase for testing
+    UnactuatedPhase qpu;
 }; // class VNHCTest
 
 // Make sure if abs(value - des) <= hys, that the supervisor returns qa = 0
@@ -94,14 +96,79 @@ TEST_F(SupervisorTest, ZERO_WITHIN_HYS)
     double des = M_PI/2;
     double hys = 0.2;
     // check that if value is in [des-hys, des+hys] that we get qa = 0
-    EXPECT_EQ(sup_.stabilize(des-hys,des,hys),0);
-    EXPECT_EQ(sup_.stabilize(des,des,hys),0);
-    EXPECT_EQ(sup_.stabilize(des+hys,des,hys),0);
+    EXPECT_EQ(sup_.stabilize(qpu_,des-hys,des,hys),0);
+    EXPECT_EQ(sup_.stabilize(qpu_,des,des,hys),0);
+    EXPECT_EQ(sup_.stabilize(qpu_,des+hys,des,hys),0);
 
     // Check that if value is in [-des-hys,-des+hys] that we get qa = 0
-    EXPECT_EQ(sup_.stabilize(-des-hys,des,hys),0);
-    EXPECT_EQ(sup_.stabilize(-des,des,hys),0);
-    EXPECT_EQ(sup_.stabilize(-des+hys,des,hys),0);
+    EXPECT_EQ(sup_.stabilize(qpu_,-des-hys,des,hys),0);
+    EXPECT_EQ(sup_.stabilize(qpu_,-des,des,hys),0);
+    EXPECT_EQ(sup_.stabilize(qpu_,-des+hys,des,hys),0);
+
+    // If the desried value is negative, it should be the same as it being
+    // positive.
+    EXPECT_EQ(sup_.stabilize(qpu_,des-hys,-des,hys),0);
+    EXPECT_EQ(sup_.stabilize(qpu_,des,-des,hys),0);
+    EXPECT_EQ(sup_.stabilize(qpu_,des+hys,-des,hys),0);
+
+    // If the hysteresis is negative, it should be the same as it being
+    // positive.
+    EXPECT_EQ(sup_.stabilize(qpu_,des-hys,des,-hys),0);
+    EXPECT_EQ(sup_.stabilize(qpu_,des,des,-hys),0);
[…]

### 2021-04-26 sugar_bot/arduino/libraries/SUGAR_VNHC_Library/include/Supervisor.h
Created the Oscillation Supervisor.

@@ -58,6 +58,49 @@ private:
 
 }; // class Supervisor
 
+/*****************************************************************************/
+
+class OscillationSupervisor
+{
+public:
+    /**
+     * @brief: Create a meta-supervisor which tracks the value qu = qudes using the given supervisor.
+     */
+    OscillationSupervisor(const Supervisor& sup);
+
+    /**
+      * @brief: Set qa to stabilize a desired qu in ]-pi,pi[. This updates an
+      * internal state which keeps track of the injection/dissipation process;
+      * if this fails, simply input a (qu,pu)=(0,0) into the unactuated phase to
+      * reset.
+      * @param: const UnactuatedPhase&. The current phase of the acrobot.
+      * @param: double. The desired angle qu, in ]-pi,pi[. If it is out of
+      * range, we set the desired angle to -pi+hys or pi-hys depending on the
+      * sign of the input.
+      * @param: double. The hysteresis.
+      *
+      * @return: double. The value of qa.
+      */
+    auto stabilize(const UnactuatedPhase& qpu,
+                   double qudes,
+                   double hys) const -> double;
+
+private:
+    // The sign function
+    auto sign(double x) const -> double
+    {
+        return (x > 0) - (x < 0);
+    }
+    // The supervisor we are using
+    const Supervisor& sup_;
+
+    // The last peak value of qpu that we reached.
+    // this is modified by stabilize(), which is a const-correct function, so we
+    // define this as mutable.
+    mutable  UnactuatedPhase prevPeak_;
+
+}; // class OscillationSupervisor
+
 } // namespace SUGAR
 
 #endif // __SUPERVISOR_H__

### 2021-04-26 sugar_bot/arduino/libraries/SUGAR_VNHC_Library/src/Supervisor.cpp

@@ -51,5 +51,50 @@ auto Supervisor::stabilize(const UnactuatedPhase& qpu,
     return 0;
 }
 
+///////////////////////////
+// OscillationSupervisor //
+///////////////////////////
+
+OscillationSupervisor::OscillationSupervisor(const Supervisor& sup)
+: sup_(sup)
+{}
+
+auto OscillationSupervisor::stabilize(const UnactuatedPhase& qpu,
+                                      double qudes,
+                                      double hys) const -> double
+{
+    // Make sure qudes and hys are positive.
+    qudes = fabs(qudes);
+    hys = fabs(hys);
+
+    // Extract qu. We don't care about its sign.
+    double qu = fabs(qpu.qu);
+
+    // If we are within range of qudes, then we update our "peak"
+    // (even though this is not a peak) so that we maintain our qa = 0.
+    // This means that dissipation only occurs when we are outside the range;
+    // as soon as we hit the range again, we go back to qa = 0.
+    // TODO: If this doesn't work in experiments, remove it.
+    if (fabs(qudes - qu) <= hys)
+    {
+        prevPeak_ = qpu;
+    }
+
+    // Check if pu has changed sign since last time. If so we reached a peak,
+    // which we take as this qu.
+
+    // NOTE: if pu = 0 we save this as a peak, but the next position we measure
+    // where pu != 0 will also be taken as a peak. If you have fast enough
+    // measurements this should not be a problem.
+    double pu = qpu.pu;
+    if (pu == 0 || (sign(pu) != sign(prevPeak_.pu)))
+    {
+        prevPeak_ = qpu;
+    }
+    
+    // Return the appropriate injection mechanism for this value qpu, based on the previous peak.
+    return sup_.stabilize(qpu,prevPeak_.qu,qudes,hys);
+}
+
 }; // namespace SUGAR
 /* vim: set tw=80 ts=4 sw=4 sts=0 et ffs=unix : */

### 2021-04-26 sugar_bot/arduino/libraries/SUGAR_VNHC_Library/src/test/SupervisorTest.cpp

@@ -19,8 +19,6 @@
 #include "Phase.h"
 #include "VNHC.h"
 #include "Supervisor.h"
-// Include predicates
-#include "test/Predicates.h"
 // Includes for M_PI, cosine, etc
 #define _USE_MATH_DEFINES
 #include <math.h>
@@ -31,8 +29,10 @@
 #include "gtest/gtest.h"
 
 using namespace SUGAR;
-using SUGAR::predicates::Within;
 
+////////////////
+// Supervisor //
+////////////////
 class SupervisorTest : public ::testing::Test
 {
 public:
@@ -45,7 +45,7 @@ public:
         g_),
       in_(xingbot_,qmax_,I_),
       diss_(xingbot_,qmax_,-I_),
-      sup_(in_,diss_) // Todo: implement this correctly
+      sup_(in_,diss_)
     {}
 protected:
     void SetUp() override 
@@ -72,9 +72,6 @@ protected:
     // Dynamics for Xingbo's bot
     Acrobot xingbot_;
 
-    // Define a pointer to the VNHC
-    AcrobotVNHC* pVNHC_;
-
     // Define an unactuatedPhase for testing the objects
     UnactuatedPhase qpu_;
 
@@ -84,11 +81,9 @@ protected:
     ArctanVNHC in_;
     ArctanVNHC diss_;
 
-    // Define our supervisor, which takes in the two VNHCs
+    // Define our supervisors, which take in the two VNHCs
     Supervisor sup_;
-    // A default unactuated phase for testing
-    UnactuatedPhase qpu;
-}; // class VNHCTest
+}; // class SupervisorTest
 
 // Make sure if abs(value - des) <= hys, that the supervisor returns qa = 0
 TEST_F(SupervisorTest, ZERO_WITHIN_HYS)
@@ -171,6 +166,271 @@ TEST_F(SupervisorTest, DISSIPATION)
     EXPECT_EQ(sup_.stabilize(qpu_, -des-0.1,des,0),diss_.qa(qpu_));
 }
 
+/*****************************************************************************/
+
+
+////////////////////////////
+// Oscillation Supervisor //
+////////////////////////////
[…]

### 2021-04-27 sugar_bot/arduino/libraries/SUGAR_VNHC_Library/include/Supervisor.h
Created the initial version of the rotation supervisor and fixed up the oscillation supervisor. The latter now has a flag which we can use to specify whether we should set qa = 0 only on peaks, or anytime we go through the desired range. The former has been tested for injection, but it cannot do dissipation yet.

@@ -17,6 +17,7 @@
 #define __SUPERVISOR_H__
 
 #include "VNHC.h"
+#include <math.h> // get M_PI
 
 namespace SUGAR
 {
@@ -69,10 +70,7 @@ public:
     OscillationSupervisor(const Supervisor& sup);
 
     /**
-      * @brief: Set qa to stabilize a desired qu in ]-pi,pi[. This updates an
-      * internal state which keeps track of the injection/dissipation process;
-      * if this fails, simply input a (qu,pu)=(0,0) into the unactuated phase to
-      * reset.
+      * @brief: stabilizes and sets qa = 0 any time we go in range of qudes.
       * @param: const UnactuatedPhase&. The current phase of the acrobot.
       * @param: double. The desired angle qu, in ]-pi,pi[. If it is out of
       * range, we set the desired angle to -pi+hys or pi-hys depending on the
@@ -84,6 +82,27 @@ public:
     auto stabilize(const UnactuatedPhase& qpu,
                    double qudes,
                    double hys) const -> double;
+
+    /**
+      * @brief: Set qa to stabilize a desired qu in ]-pi,pi[. This updates an
+      * internal state which keeps track of the injection/dissipation process;
+      * if this fails, simply input a (qu,pu)=(0,0) into the unactuated phase to
+      * reset.
+      * @param: const UnactuatedPhase&. The current phase of the acrobot.
+      * @param: double. The desired angle qu, in ]-pi,pi[. If it is out of
+      * range, we set the desired angle to -pi+hys or pi-hys depending on the
+      * sign of the input.
+      * @param: double. The hysteresis.
+      * @param: bool. If true, we set qa = 0 whenever we go in range of qudes.
+      * If false, we do not switch to qa = 0 until we hit a peak within range of
+      * qudes.
+      *
+      * @return: double. The value of qa.
+      */
+    auto stabilize(const UnactuatedPhase& qpu,
+                   double qudes,
+                   double hys,
+                   bool qaZeroInRange) const -> double;
 
 private:
     // The sign function
@@ -100,6 +119,46 @@ private:
     mutable  UnactuatedPhase prevPeak_;
 
 }; // class OscillationSupervisor
+
+/*****************************************************************************/
+class RotationSupervisor
+{
+public:
+    /**
+     * @brief: Create a meta-supervisor which tracks the value pu = pudes using the given supervisor.
+     */
+    RotationSupervisor(const Supervisor& sup);
+
[…]

### 2021-04-27 sugar_bot/arduino/libraries/SUGAR_VNHC_Library/src/Supervisor.cpp

@@ -15,6 +15,8 @@
 
 #include "include/Supervisor.h"
 #include <math.h>
+// TODO: remove
+#include <iostream>
 
 namespace SUGAR
 {
@@ -51,6 +53,8 @@ auto Supervisor::stabilize(const UnactuatedPhase& qpu,
     return 0;
 }
 
+/*****************************************************************************/
+
 ///////////////////////////
 // OscillationSupervisor //
 ///////////////////////////
@@ -63,9 +67,23 @@ auto OscillationSupervisor::stabilize(const UnactuatedPhase& qpu,
                                       double qudes,
                                       double hys) const -> double
 {
+    return stabilize(qpu,qudes,hys,true);
+}
+
+auto OscillationSupervisor::stabilize(const UnactuatedPhase& qpu,
+                                      double qudes,
+                                      double hys,
+                                      bool qaZeroInRange) const -> double
+{
     // Make sure qudes and hys are positive.
     qudes = fabs(qudes);
     hys = fabs(hys);
+
+    // Make sure qudes <= M_PI
+    if (qudes > M_PI)
+    {
+        qudes = M_PI;
+    }
 
     // Extract qu. We don't care about its sign.
     double qu = fabs(qpu.qu);
@@ -74,8 +92,7 @@ auto OscillationSupervisor::stabilize(const UnactuatedPhase& qpu,
     // (even though this is not a peak) so that we maintain our qa = 0.
     // This means that dissipation only occurs when we are outside the range;
     // as soon as we hit the range again, we go back to qa = 0.
-    // TODO: If this doesn't work in experiments, remove it.
-    if (fabs(qudes - qu) <= hys)
+    if (qaZeroInRange && fabs(qudes - qu) <= hys)
     {
         prevPeak_ = qpu;
     }
@@ -95,6 +112,53 @@ auto OscillationSupervisor::stabilize(const UnactuatedPhase& qpu,
     // Return the appropriate injection mechanism for this value qpu, based on the previous peak.
     return sup_.stabilize(qpu,prevPeak_.qu,qudes,hys);
 }
+
+/*****************************************************************************/
+
+/////////////////////////
+// Rotation Supervisor //
+/////////////////////////
+
+RotationSupervisor::RotationSupervisor(const Supervisor& sup)
+: sup_(sup)
[…]

### 2021-04-27 sugar_bot/arduino/libraries/SUGAR_VNHC_Library/src/test/SupervisorTest.cpp

@@ -213,7 +213,67 @@ protected:
     double hys_ = 0.1;
 }; // class OscillationSupervisorTest
 
- 
+/**
+ * The maximum qudes is M_PI-hys, which is then reset to M_PI-hys if it is above
+ * this value.
+ */
+TEST_F(OscillationSupervisorTest, QUDES_BELOW_PI)
+{
+    // If qudes > M_PI, we should get qa = 0 if 
+    // |qu| in [M_PI - hys, M_PI]
+    // i.e. setting qudes > M_PI resets it to M_PI.
+
+    double qudes = M_PI;
+    qpu_.qu = M_PI-hys_;
+    EXPECT_EQ(osup_.stabilize(qpu_,qudes,hys_),0);
+
+    // Increase qudes above pi
+    qudes = M_PI + 2*hys_;
+    qpu_.qu = M_PI-hys_;
+    EXPECT_EQ(osup_.stabilize(qpu_,qudes,hys_),0);
+    qpu_.qu = M_PI;
+    EXPECT_EQ(osup_.stabilize(qpu_,qudes,hys_),0);
+    qpu_.qu = -M_PI + hys_;
+    EXPECT_EQ(osup_.stabilize(qpu_,qudes,hys_),0);
+    qpu_.qu = -M_PI;
+    EXPECT_EQ(osup_.stabilize(qpu_,qudes,hys_),0);
+
+    // Increase qudes well above pi
+    qudes = 3*M_PI/2;
+    qpu_.qu = M_PI-hys_;
+    EXPECT_EQ(osup_.stabilize(qpu_,qudes,hys_),0);
+    qpu_.qu = M_PI;
+    EXPECT_EQ(osup_.stabilize(qpu_,qudes,hys_),0);
+    qpu_.qu = -M_PI + hys_;
+    EXPECT_EQ(osup_.stabilize(qpu_,qudes,hys_),0);
+    qpu_.qu = -M_PI;
+    EXPECT_EQ(osup_.stabilize(qpu_,qudes,hys_),0);
+
+    // negative qudes shouldn't affect this
+    qudes = -M_PI - 2*hys_;
+    qpu_.qu = M_PI-hys_;
+    EXPECT_EQ(osup_.stabilize(qpu_,qudes,hys_),0);
+    qpu_.qu = M_PI;
+    EXPECT_EQ(osup_.stabilize(qpu_,qudes,hys_),0);
+    qpu_.qu = -M_PI + hys_;
[…]

### 2021-04-27 sugar_bot/arduino/libraries/SUGAR_VNHC_Library/src/Supervisor.cpp
Fixed some errors in the rotation mechanism, completed the code for now.

@@ -135,10 +135,6 @@ auto RotationSupervisor::stabilize(
         prevBottom_ = qpu;
     }
 
-//    std::cout << "pu = " << prevBottom_.pu;
-//    std::cout << "| pudes = " << pudes;
-//    std::cout << "| stab = " << sup_.stabilize(qpu,prevBottom_.pu,pudes,hys) << std::endl;
-
     // If we are within range of pudes, fix qa to the value of the injection
     // mechanism at (0,pudes).
     // Otherwise, update qa using the supervisor.
@@ -147,15 +143,16 @@ auto RotationSupervisor::stabilize(
     // Make sure everything is in absolutes
     pudes = fabs(pudes);
     hys = fabs(hys);
+    double prevPu = fabs(prevBottom_.pu);
     // If we are in range, we actually want qa to be fixed as the inejction for
     // (qu,pu) = (0,pudes)
-    if (fabs(pudes - prevBottom_.pu) <= hys)
+    if (fabs(pudes - prevPu) <= hys)
     {
         UnactuatedPhase qpudes;
         qpudes.pu = pudes;
         qa = sup_.stabilize(qpudes,0,pudes,hys);
     }
-    
+
     return qa;
 }
 

### 2021-04-27 sugar_bot/arduino/libraries/SUGAR_VNHC_Library/src/test/SupervisorTest.cpp

@@ -611,45 +611,137 @@ TEST_F(RotationSupervisorTest, INJECTION)
 
 TEST_F(RotationSupervisorTest, DISSIPATE)
 {
-    EXPECT_FALSE(true);
     // Initialized at (0,pu), should dissipate if |pu| > pudes + hys
-    
+    qpu_.pu = pudes_+2*hys_;
+    EXPECT_EQ(rsup_.stabilize(qpu_,pudes_,hys_),diss_.qa(qpu_));
     // Initialized at (0,-pu), should dissipate if |pu| > pudes+hys
+    qpu_.pu = -2*pudes_;
+    EXPECT_EQ(rsup_.stabilize(qpu_,pudes_,hys_),diss_.qa(qpu_));
 
     // Swing up to (-qu,0), still dissipating 
+    qpu_.qu = -M_PI/2; qpu_.pu = 0;
+    EXPECT_EQ(rsup_.stabilize(qpu_,pudes_,hys_),diss_.qa(qpu_));
 
     // Pushed up through qu = -pi, dissipating even though pu > pudes
-    //
+     qpu_.qu = -M_PI; qpu_.pu = 2*pudes_;
+    EXPECT_EQ(rsup_.stabilize(qpu_,pudes_,hys_),diss_.qa(qpu_));
+   
     // Cross over to qu = pi, dissipating even though pu < pudes
+    qpu_.qu = M_PI; qpu_.pu = pudes_/2;
+    EXPECT_EQ(rsup_.stabilize(qpu_,pudes_,hys_),diss_.qa(qpu_));
 
     // Continue back to qu ~ 4deg, should dissipate when pu > pudes
+    qpu_.qu = 4*(M_PI/180); qpu_.pu = -pudes_-1.1*hys_;
+    EXPECT_EQ(rsup_.stabilize(qpu_,pudes_,hys_),diss_.qa(qpu_));
 
     // Sign of pudes and hys don't matter
+    EXPECT_EQ(rsup_.stabilize(qpu_,-pudes_,hys_),diss_.qa(qpu_));
+    EXPECT_EQ(rsup_.stabilize(qpu_,pudes_,-hys_),diss_.qa(qpu_));
+
 }
 
 TEST_F(RotationSupervisorTest, FIXED_QA)
 {
-    // TODO:
-    EXPECT_FALSE(true);
[…]

### 2021-04-27 sugar_bot/arduino/libraries/SUGAR_VNHC_Library/src/Supervisor.cpp
Removed iostream, which was accidentally in supervisor.cpp

@@ -15,8 +15,6 @@
 
 #include "include/Supervisor.h"
 #include <math.h>
-// TODO: remove
-#include <iostream>
 
 namespace SUGAR
 {

### 2021-04-27 sugar_bot/arduino/libraries/SUGAR_VNHC_Library/SUGAR_VNHC.h
including the supervisor header in SUGAR_VNHC.h

@@ -8,7 +8,7 @@
 * Department of Electrical and Computer Engineering
 * University of Toronto
 *
-* Last Modified: 06/Feb/20
+* Last Modified: 27/Apr/21
 * Last Editor:   Adan Moran-Macdonald
 * Description:   Header file to include to use this library in the arduino. It
 * simply includes all the relevant headers here.
@@ -21,6 +21,7 @@
 #include "include/AcrobotDynamics.h"
 #include "include/Phase.h"
 #include "include/VNHC.h"
+#include "include/Supervisor.h"
 
 #endif
 /* vim: set tw=80 ts=4 sw=4 sts=0 et ffs=unix : */

### 2021-04-28 latex/paper/main.tex
Finished intro, problem formulation, vnhc, and acrobot constraint sections. Need simulations, experiments, proof, and conclusion.

@@
 \thanks{Manuscript submitted for review on \today.}
-\thanks{A. Moran-MacDonald is with the Department of Electrical and Computer
-    Engineering, University of Toronto, Toronto, ON, Canada (e-mail:
-adan.moran@mail.utoronto.ca).}
-\thanks{M. Maggiore is with the Department of Electrical and Computer
-Engineering, University of Toronto, ON, Canada (e-mail:
-maggiore@control.utoronto.ca).}
+\thanks{A. Moran-MacDonald (e-mail: adan.moran@mail.utoronto.ca) and
+M. Maggiore (e-mail: maggiore@control.utoronto.ca) are with the Department of
+Electrical and Computer Engineering, University of Toronto, ON, Canada.}
 \thanks{X. Wang is with ??? (e-mail: ???).}
@@
 \section{Introduction}\label{sec:introduction}
-\IEEEPARstart{T}ODO: intro.
+
+In gymnastics terminology, a ``giant" is the motion a gymnast performs to
+achieve full rotations around a horizontal bar \cite{usagym_giant}. 
+A gymnast will begin by hanging at rest, then swing their legs
+appropriately to gain energy over time.
+The authors of \cite{pendulum_length_giant_gymnastics} modelled the gymnast as a
+variable length pendulum, and studied how the pendulum's length changes as a
+function of the gymnast's limb angle.
+Labeling the pendulum length by \(r\) and the gymnast's body orientation
+by \(\theta\), they observed experimentally that the value \(\dot{r}/r\) has
+the biggest impact on the magnitude of energy injection. 
+After testing several gymnasts under a variety of experimental conditions, 
+they discovered that the peak value of \(\dot{r}/r\) occured at the same fixed
+value of \(\dot{\theta}/\theta\) for all gymnasts.
+In other words, gymnasts appear to move their legs as a function of their body
+angle and velocity when performing giants; 
+doing so allows them to gain energy and rotate around the bar.
+
+While the simplest model of a gymnast is the variable-length pendulum, a
+more realistic model is the two-link acrobot (Figure \ref{fig:acrobot}).
+Here, the top link represents the torso, while the bottom link represents
+the legs. 
+The acrobot is actuated exclusively at the center joint (the hips).
+Controlling the acrobot is a nontrivial task because it is not feedback
+linearizable \cite{nonlinear_controllers_nonintegrable_acrobot}. 
+To solve the swingup problem, one might begin by designing a leg controller
+which provably injects energy into the acrobot, so that the resulting motion
+mimics that of a human performing a giant.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{acrobot_gymnast.png}
+    \caption{The two-link acrobot as a model for a gymnast.
+    Image modified from \cite{xingbo_thesis}.}
+    \label{fig:acrobot}
+\end{figure}
+
+Previous attempts at acrobot giant generation have involved
+trajectory tracking, partial feedback linearization, or other energy-based
+methods
+( see
+\cite{energy_pumping_robotic_swinging,swingup_giant_acrobot,dynamical_servo_acrobot_vc,control_giant_two_link_gymnastic_robot}
+).
+While all these approaches succeed at making the acrobot rotate around the
+bar, none of them use the results of \cite{pendulum_length_giant_gymnastics}.
+That is, none of these leg controllers track a function of the acrobot's body
+angle and velocity.
+In 2016, Wang attempted to design such a controller \cite{xingbo_thesis}
+but was unable to prove the acrobot would gain energy over time.
+His approach was a rudimentary version of a recent technique know as the method
+of virtual nonholonomic constraints.
+
+Virtual nonholonomic constraints (VNHCs) have been used for human-robot interaction
+\cite{vnhc_human_robot_cooperation,psd_based_vnhc_redundant_manipulator,haptic_vnhc},
+error-reduction on time-delayed systems \cite{vnhc_time_delay_teleop},
+and they have shown marked improvements to the field of bipedal locomotion 
+\cite{nhvc_dynamic_walking,
+hybrid_zero_dynamics_bipedal_nhvcs,output_nhvc_bipedal_control}.
+Indeed, they produce more robust walking motion in biped robots than
+other controllers which do not depend on velocity \cite{nhvc_incline_walking}.
+In particular, VNHCs may be capable of injecting and
+dissipating energy from a system in a robust manner, all while producing
+realistic biological motion.
+
+In this article, we will design a virtual nonholonomic constraint which provably
+injects energy into the acrobot through human-like giant motion.
 
@@
 \section{Problem Formulation}\label{sec:problem-formulation}
-%TODO: motivate injection/dissipation for energy regulation in an acrobot. Use
-%this to lead into VNHCs. e.g. represent torso as chain of a swing, pivot as
-%seat, and legs as human legs. Replicate fig3.3 from thesis but for acrobot: take
-%(qu,pu)-plane and show how a person moves their legs wrt theta, qu, or pu. Leave
-%the gymnast model until the end, or ignore it entirely.
-%
-%Alternatively, ff human movement journal says legs move wrt qu, we can use that
-%as a foundation.
-%Ideas: Human movement says appropriate regulation of pendulum allows energy to
-%be pumped into system. Pendulum length rapidly increased on downward swing, then
-%rapidly decreased on upward swing (I think these mean backward and forward).
-%They plot pendulum length rdot/r vs "body angle" theta to determine energy
-%injection, which hints at some relationship between (r,rdot) and (theta,thetad).
-%Indeed, the peak value of rdot/r did not occur at a fixed value of theta, nor
-%once anyone reached a particular speed; it occured at the same thetadot/theta for all
-%participants of the study. This perfectly motivates VNHCs for robots, because we
-%can shape (r,rtdot) as functions of (theta,thetad).
-%Start motivation from energy injection - how do humans inject energy? How does
-%acrobot model gymnast? use human movement to motivate tracking functions of both
-%position and velocity rather than time.
-
-In gymnastics terminology, a ``giant" is the motion a gymnast performs to
-achieve full rotations around a horizontal bar \cite{usagym_giant}. 
-A gymnast will begin by hanging at rest, then swing their legs
-appropriately to gain energy over time.
-The authors of \cite{pendulum_length_giant_gymnastics} modelled the gymnast as a
-variable length pendulum, and studied how the pendulum's length changes as a
-function of the gymnast's limb angle.
-Labeling the pendulum length by \(r\) and the gymnast's body orientation
-by \(\theta\), they observed experimentally that the value \(\dot{r}/r\) has
-the biggest impact on the magnitude of energy injection. 
-After testing several gymnasts under a variety of experimental conditions, 
-they discovered that the peak value of \(\dot{r}/r\) occured at the same fixed
-value of \(\dot{\theta}/\theta\) for all gymnasts.
-In other words, gymnasts appear to move their legs as a function of their body
-angle and velocity when performing giants; 
-doing so allows them to gain energy and rotate around the bar.
-
-While the simplest model of a gymnast is the variable-length pendulum, a
-more realistic model is the two-link acrobot (Figure \ref{fig:acrobot}).
-Here, the top link represents the torso, while the bottom link represents
-the legs. 
-The acrobot is actuated exclusively at the center joint (the hips).
-Controlling the acrobot is a nontrivial task because it is not feedback
-linearizable \cite{nonlinear_controllers_nonintegrable_acrobot}. 
-To solve the swingup problem, one might begin by designing a leg controller
-which provably injects energy into the acrobot, so that the resulting motion
-mimics that of a human performing a giant.
+We will use the simplified acrobot model in Figure
+\ref{fig:simple-acrobot-model}, where we assume the torso and leg rods are of
+equal length \(l\) with equal point masses \(m\) at the tips.
+A real gymnast cannot swing their legs in full circles, though they
+are usually flexible enough to raise them parallel to the floor;
+hence, we assume the leg angle \(q_a\) lies in \([-Q_a, Q_a]\) for some
+\(Q_a \in [\frac{\pi}{2}, \pi[\). 
+We also ignore any frictional forces.
 
@@
     \centering
-    \includegraphics[width=\linewidth]{acrobot_gymnast.png}
-    \caption{The two-link acrobot as a model for a gymnast.
-    Image modified from \cite{xingbo_thesis}.}
-    \label{fig:acrobot}
+    \includegraphics[width=0.8\linewidth]{simple_acrobot_model.pdf}
+    \caption{A simple acrobot has massless rods of equal length \(l\) and 
+    equal masses \(m\) at the tips.}
+    \label{fig:simple-acrobot-model}
 \end{figure}
 
-Previous attempts at acrobot giant generation have involved
-trajectory tracking, partial feedback linearization, or other energy-based
-methods
-( see
-\cite{energy_pumping_robotic_swinging,swingup_giant_acrobot,dynamical_servo_acrobot_vc,control_giant_two_link_gymnastic_robot}
-).
-While all these approaches succeed at making the acrobot rotate around the
-bar, none of them use the results of \cite{pendulum_length_giant_gymnastics}.
-That is, none of these leg controllers track a function of the acrobot's body
-angle and velocity.
-In 2016, Wang attempted to design such a controller \cite{xingbo_thesis}
-but was unable to prove the acrobot would gain energy over time.
-His approach was a rudimentary version of a recent technique know as the method
-of virtual nonholonomic constraints.
-
-Virtual nonholonomic constraints (VNHCs) have been used for human-robot interaction
-\cite{vnhc_human_robot_cooperation,psd_based_vnhc_redundant_manipulator,haptic_vnhc},
-error-reduction on time-delayed systems \cite{vnhc_time_delay_teleop},
-and they have shown marked improvements to the field of bipedal locomotion 
-\cite{nhvc_dynamic_walking,
-hybrid_zero_dynamics_bipedal_nhvcs,output_nhvc_bipedal_control}.
-Indeed, they produce more robust walking motion in biped robots than
-other controllers which do not depend on velocity \cite{nhvc_incline_walking}.
-In particular, VNHCs may be capable of injecting and
-dissipating energy from a system in a robust manner, all while producing
-realistic biological motion.
-
-In this article, we will design a virtual
-nonholonomic constraint which provably injects energy into the acrobot through 
-human-like giant motion.
-
-\textbf{MANFREDI: This section feels more like an intro than a problem
-formulation. Thoughts?}
+The acrobot has inertia matrix 
+\(M\), potential function \(V\) (with respect to
+the horizontal bar), and input matrix \(B\) given as follows:
+\begin{align}\label{eqn:acrobot-inertia}
+    M(q) &= \begin{bmatrix}
+        ml^2\left(3+2\cos(q_a)\right) & 
+        ml^2\left(1+\cos(q_a)\right) \\
+        ml^2\left(1+\cos(q_a)\right) &
+        ml^2
+    \end{bmatrix} 
+    , \\
+    \label{eqn:acrobot-potential}
+    V(q) &= -mgl\left(2\cos(q_u)+\cos(q_u+q_a)\right)
+    , \\
+    \label{eqn:acrobot-B}
+    B &= [0;1]
+    .
+\end{align}
+For reasons that will become clear in later sections of this article, we will
+use Hamiltonian mechanics to derive the dynamics of the acrobot. 
+For this we require the conjugate of momenta, \(p = (p_u,p_a) = M(q)\dot{q}\).
+The dynamics of the acrobot in \((q,p)\) coordinates are given in
+~\eqref{eqn:acrobot-hamiltonian}.
+For shorthand, we write \(c_u := \cos(q_u)\), \(c_a := \cos(q_a)\), and 
+\(c_{ua} := \cos(q_u + q_a)\); likewise, \(s_u := \sin(q_u)\), 
+\(s_a := \sin(q_a)\), and \(s_{ua} := \sin(q_u + q_a)\).
+\begin{align}\label{eqn:acrobot-hamiltonian}
+    \mathcal{H}(q,p) &= \frac{1}{2}p\tpose \Minv(q) p -
+    mgl\left(2 c_u + c_{ua}\right)
+    , \\
+     &\begin{cases}
+        \dot{q} = \Minv(q) p 
+        ,\\
+        \dot{p}_u = -mgl\left(2s_u + s_{ua}\right) 
+        ,\\
+        \dot{p}_a =-\frac{1}{2}p\tpose \nabla_{q_a}\Minv(q) p
+        - mgl s_{ua} + \tau.
+    \end{cases} \nonumber
+\end{align}
+The control input is a force \(\tau \in \R\) affecting only the dynamics of
+\(p_a\), representing a torque acting on the hip joint.
+
+Our goal is to design a smooth function \(f : \Sone \times \R\) such that the
+relation \(q_a = f(q_u,p_u)\) is a regular VNHC, where the constrained dynamics
+of the acrobot (under this constraint) gain energy on some set \(D\)
+in the sense of Definition \ref{defn:energy-gain}.
+We will formally define regularity in the next section.
+Essentially, if the constraint meets some regularity conditions then it can be
+stabilized via feedback.
+
+\begin{defn}\label{defn:energy-gain}
+    Let \(\mathcal{Q}\) be an
+    \(n\)-dimensional smooth generalized cylinder.
+    Let \(f : \mathcal{Q} \rightarrow \mathcal{Q}\times\R^n\) be a smooth vector
+    field and let \(D \subset M\) be open.
+    The system described by \(\dot{x} = f(x)\) 
+    \textit{gains energy on \(D\)} if, 
+    for all compact sets \(K \subset D\) and for almost every initial
+    condition \(x(0) \in K\), there exists \(T > 0\) such
+    that \(x(t) \notin K \, (\forall t > T)\).
+    The system \textit{loses energy on \(D\)} if it gains energy in
+    negative-time.
+\end{defn}
+
+Note that any system satisfying Definition \ref{defn:energy-gain}
+can have unstable equilibria on \(D\), but not limit cycles nor closed orbits.
 
@@
 
-\subsection{Simply Actuated Hamiltonian Systems}
+\subsection{Simply Actuated Hamiltonian Systems} \label{sec:vnhc-sah}
 Take a mechanical system modelled with generalized coordinates 
@@
 \eqref{eqn:hamiltonian-eom-general} expands to
-\begin{equation}\label{eqn:hamiltonian-full-dynamics}
+\begin{equation*}\label{eqn:hamiltonian-full-dynamics}
      \begin{cases}
@@
     \end{cases}
-\end{equation}
+\end{equation*}
+%\begin{align*}
+%    \dot{q} &= D\inv(q)p 
+%    ,\\
+%    \dot{p} &= -\frac{1}{2} (\Id{n} \otimes p\tpose) \nabla_q D\inv(q) p
+%        - \nabla_q P(q) + B(q) \tau
+%    .
+%\end{align*}
 
@@
 
-\subsection{Virtual Nonholonomic Constraints}
+\subsection{Virtual Nonholonomic Constraints}\label{sec:vnhc-vnhc}
 
@@
 
-We conclude this section by formalizing the notion of energy injection for
-VNHCs.
-
-\begin{defn}\label{defn:energy-gain}
-    Let \(\mathcal{Q}\) be an
-    \(n\)-dimensional smooth generalized cylinder.
-    Let \(f : \mathcal{Q} \rightarrow \mathcal{Q}\times\R^n\) be a smooth vector
-    field and let \(D \subset M\) be open.
-    The system described by \(\dot{x} = f(x)\) 
-    \textit{gains energy on \(D\)} if, 
-    for all compact sets \(K \subset D\) and for almost every initial
-    condition \(x(0) \in K\), there exists \(T > 0\) such
-    that \(x(t) \notin K \, (\forall t > T)\).
-    The system \textit{loses energy on \(D\)} if it gains energy in
-    negative-time.
-\end{defn}
-
-Any system satisfying Definition \ref{defn:energy-gain}
-can have unstable equilibria on \(D\), but not limit cycles nor closed orbits.
-The next definition ties this notion of energy gain to VNHCs.
+We conclude this section by formalizing the notion of energy
+injection/dissipation for VNHCs.
 
@@
     \textit{injects (dissipates) energy on \(D \subset \Gamma\)} if the
-    constrained dynamics gain (lose) energy everywhere on \(D\), except possibly
-    on a set of measure zero.
+    constrained dynamics gain (lose) energy everywhere on \(D\) according to
+    Definition \ref{defn:energy-gain}, except possibly on a set of measure
+    zero.
 \end{defn}
@@
 acrobot through giant-like motion.
-We will use the simplified acrobot model in Figure
-\ref{fig:simple-acrobot-model}, where we assume the torso and leg rods are of
-equal length \(l\) with equal point masses \(m\) at the tips.
-A real gymnast cannot swing their legs in full circles, though they
-are usually flexible enough to raise them parallel to the floor;
-hence, we assume the leg angle \(q_a\) lies in \([-Q_a, Q_a]\) for some
-\(Q_a \in [\frac{\pi}{2}, \pi[\). 
-We also ignore any frictional forces.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.7\linewidth]{simple_acrobot_model.pdf}
-    \caption{A simple acrobot has massless rods of equal length \(l\) and 
-    equal masses \(m\) at the tips.}
-    \label{fig:simple-acrobot-model}
-\end{figure}
-
-The acrobot has inertia matrix 
-\(M\), potential function \(V\) (with respect to
-the horizontal bar), and input matrix \(B\) given as follows:
-\begin{align}\label{eqn:acrobot-inertia}
-    M(q) &= \begin{bmatrix}
-        ml^2\left(3+2\cos(q_a)\right) & 
-        ml^2\left(1+\cos(q_a)\right) \\
-        ml^2\left(1+\cos(q_a)\right) &
-        ml^2
-    \end{bmatrix} 
-    , \\
-    \label{eqn:acrobot-potential}
-    V(q) &= -mgl\left(2\cos(q_u)+\cos(q_u+q_a)\right)
-    , \\
-    \label{eqn:acrobot-B}
-    B &= [0;1]
-    .
-\end{align}
-The conjugate of momenta is \(p = (p_u,p_a) = M(q)\dot{q}\).
-The dynamics of the acrobot in \((q,p)\) coordinates are given in
-~\eqref{eqn:acrobot-hamiltonian}, where
-for shorthand, we write \(c_u := \cos(q_u)\), \(c_a := \cos(q_a)\), and 
-\(c_{ua} := \cos(q_u + q_a)\); likewise, \(s_u := \sin(q_u)\), 
-\(s_a := \sin(q_a)\), and \(s_{ua} := \sin(q_u + q_a)\).
-\begin{align}\label{eqn:acrobot-hamiltonian}
+Recall that the acrobot in Figure \ref{fig:simple-acrobot-model} has dynamics given by
+~\eqref{eqn:acrobot-hamiltonian}, which are
+repeated here:
+\begin{align*}
     \mathcal{H}(q,p) &= \frac{1}{2}p\tpose \Minv(q) p -
@@
         \dot{p}_a =-\frac{1}{2}p\tpose \nabla_{q_a}\Minv(q) p
-        - mgl s_{ua} + \tau,
-    \end{cases} \nonumber
-\end{align}
-The control input is a force \(\tau \in \R\) affecting only the dynamics of
-\(p_a\), representing a torque acting on the hip joint.
-This means \((q,p)\) are simply actuated coordinates in the phase space
+        - mgl s_{ua} + \tau.
+    \end{cases}
+\end{align*}
+
+Since the control input \(\tau\) only affects the actuated momentum,
+\((q,p)\) are simply actuated coordinates in the phase space
 \(\mathcal{Q} \times \mathcal{P}\) where
@@
 fully by \((q_u,p_u) \in \SxR\).
-It is easy to show that the constrained dynamics are
-just the torso dynamics obtained when one swings the legs
-according to \eqref{eqn:acrobot-constraint}:
+It is easy to show that the constrained dynamics are just the torso dynamics
+obtained when one swings the legs according to \eqref{eqn:acrobot-constraint}.
+That is, the constrained dynamics are
 \begin{equation}\label{eqn:acrobot-constrained-dynamics}
@@
 
+The results of Theorem \ref{thm:acrobot-energy-stabilization} lead naturally to
+the idea of energy regulation: one could employ a supervisor which switches
+between injection and dissipation constraints to stabilize a desired energy
+level \(\bar{E}\).
+For instance, the supervisor could inject energy when \(E(q_u,p_u) < \bar{E}\) 
+and dissipate energy when \(E(q_u,p_u) > \bar{E}\), with some hysteresis to
+avoid infinite switching.
+As we shall see, such a simple supervisor will not produce human-like
+oscillations or rotations.
+However, it will allow the acrobot to stabilize some peak angle of oscillation
+\(\bar{q}_u \in \, ]0,\pi[\) so that \(E(\pm \bar{q}_u,0) = \bar{E}\),
+or some peak rotational momentum
+\(\bar{p}_u \in \, ]\sqrt{60m^2gl^3}, \bar{\rho}[\) where
+\(E(0,\pm \bar{p}_u) = \bar{E}\).
+
 %/========== Experiments ==========/%
+\section{Simulation Results}\label{sec:simulations}
+\textbf{TODO: simulation parameters in a table, show results for injection,
+dissipation, and regulation along with existing simulations. Ample commentary.}
+
 \section{Experimental Results}\label{sec:experiments}
+\textbf{TODO: 2 sections. a) is hardware description, paraphrasing Xingbo.
+    Should be brief but detailed (eg 2 arduinos, 3d printed, battery in the leg,
+    parameters of the system, etc). b) is
+experimental results, same as simulation section.}
 We tested our VNHC on the physical acrobot built by
@@
     \centering
-    \includegraphics[width=0.6\linewidth]{xingbo_acrobot.jpg}
+    \includegraphics[width=0.9\linewidth]{xingbo_acrobot.jpg}
     \caption{The acrobot built by Wang \cite{xingbo_thesis}.}
@@
     \centering
-    \includegraphics[width=0.7\linewidth]{acrobot_orbit.png}
+    \includegraphics[width=0.8\linewidth]{acrobot_orbit.png}
     \caption{A simulated orbit of Wang's acrobot.}
@@
     \centering
-    \includegraphics[width=0.7\linewidth]{acrobot_mc.png}
+    \includegraphics[width=0.8\linewidth]{acrobot_mc.png}
     \caption{Monte Carlo simulation results.}
@@
     \centering
-    \begin{subfigure}[t]{0.45\linewidth}
+    \begin{subfigure}[t]{0.49\linewidth}
         \includegraphics[width=\linewidth]{acrobot_unperturbed_orbit.png}
@@
     \end{subfigure}
-    \begin{subfigure}[t]{0.45\linewidth}
+    \begin{subfigure}[t]{0.49\linewidth}
         \includegraphics[width=\linewidth]{acrobot_stopped_orbit.png}
@@
     \centering
-    \begin{subfigure}[ht]{0.45\linewidth}
+    \begin{subfigure}[ht]{0.49\linewidth}
         \includegraphics[width=\linewidth]{acrobot_fpush_orbit.png}
@@
     \end{subfigure}
-    \begin{subfigure}[ht]{0.45\linewidth}
+    \begin{subfigure}[ht]{0.49\linewidth}
         \includegraphics[width=\linewidth]{acrobot_rpush_orbit.png}
@@
 In particular, energy injection through VNHCs appears to be robust against a
-variety of disturbances, including significant model mismatch.
+variety of disturbances, including significant model mismatch and sensor noise.
 
@@
 \section{Proof of Theorem \ref{thm:acrobot-energy-stabilization}}\label{sec:proof}
+\textbf{TODO: prove Thm 10}
 

### 2021-04-29 latex/paper/main.tex
Revamped the simulation and experimental sections, still lots to do

@@
 \usepackage{bm}
+\usepackage{booktabs}
 
@@
 \section{The Acrobot VNHC}\label{sec:acrobot}
-%specialize the theory of VNHCs to the acrobot. consider constraints which depend
-%only on pu, and show arctan vnhc and reduced dynamics. 
-%give definition of energy injection/dissipation. conclude with our
-%theorem. Make this section short and sweet.
-%
-%TODO: after writing S2, re-introduce what we want to do with acrobot (energy
-%injection for giant stabilization)
+
 Our goal in this article is to design a VNHC which injects energy into the
@@
 \section{Simulation Results}\label{sec:simulations}
-\textbf{TODO: simulation parameters in a table, show results for injection,
-dissipation, and regulation along with existing simulations. Ample commentary.}
-
-\section{Experimental Results}\label{sec:experiments}
-\textbf{TODO: 2 sections. a) is hardware description, paraphrasing Xingbo.
-    Should be brief but detailed (eg 2 arduinos, 3d printed, battery in the leg,
-    parameters of the system, etc). b) is
-experimental results, same as simulation section.}
-We tested our VNHC on the physical acrobot built by
-Xingbo Wang \cite{xingbo_thesis} (Figure \ref{fig:xingbo-acrobot}).
-The torso and legs on this robot have differing masses and lengths, so
-its inertia matrix and potential function are \textbf{not} described by 
-~\eqref{eqn:acrobot-inertia}-\eqref{eqn:acrobot-potential}.
+In Section \ref{sec:experiments} we will test our VNHC on the physical acrobot
+built by Xingbo Wang \cite{xingbo_thesis} (Figure \ref{fig:xingbo-acrobot}).
+To compare apples to apples, we will simulate our VNHC using the parameters of
+the physical system (Table \ref{tab:acrobot-parameters}).
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\linewidth]{acrobot_model.pdf}
+    \caption{The general acrobot model, represented by two weighted rods
+    differing in both length and mass.}%
+    \label{fig:acrobot-model}
+\end{figure}
+
+\begin{table}
+    \centering
+    \caption{Physical parameters for the real acrobot.}
+    \label{tab:acrobot-parameters}
+    \resizebox{\columnwidth}{!}{%
+    \begin{tabular}{ccccccccc}
+        \toprule
+        $m_u$ & $m_a$ & $l_u$ & $l_a$ & $l_{c_u}$ & $l_{c_a}$ & $J_u$ & $J_a$ & $g$ \\
+        \midrule
+        0.2112 & 0.1979 & % mass
+        0.148 & 0.145 & % total length
+        0.073 & 0.083 & % COM distance
+        0.00129 & 0.00075 & % moment of inertia
+        9.81 \\ % gravity
+        \bottomrule
+    \end{tabular}
+    } % resizebox
+\end{table}
+
+Note that these parameters do not correspond to our simplified acrobot in Figure
+\ref{fig:simple-acrobot-model}, but rather to the more general acrobot model in
+Figure \ref{fig:acrobot-model}.
+In particular, the inertia matrix and potential function are \textbf{not}
+described by ~\eqref{eqn:acrobot-inertia}-\eqref{eqn:acrobot-potential}.
+The correct \(M(q)\) and \(V(q)\) can be found in Wang's thesis
+\cite{xingbo_thesis}.
 If our VNHC injects energy into this acrobot, then we can infer that
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=0.9\linewidth]{xingbo_acrobot.jpg}
-    \caption{The acrobot built by Wang \cite{xingbo_thesis}.}
-    \label{fig:xingbo-acrobot}
-\end{figure}
-
-For our VNHC to inject energy, we know the control parameter \(I\) must be small
-(as per Theorem \ref{thm:acrobot-energy-stabilization}), but it must also be
-large enough to overcome friction in leg motor.
-Setting \(\bar{q}_a = 1\), we experimentally determined that \(I = 10\) is
-sufficient to ensure the actuator can fully rotate in its allowable range \(q_a
-\in \left[ -\frac{\pi}{2}, \frac{\pi}{2}\right]\). 
-This is the value of \(I\) we used in all simulations and experiments.
-
-\subsection{Simulations Results}
-Evaluating the true mechanical energy of Wang's acrobot at the VNHC
+Evaluating the true mechanical energy of this acrobot at the VNHC
 \(q_a = 0\) yields the energy of the nominal pendulum,
@@
 \]
-We simulated Wang's acrobot with initial condition
-\((q_u,p_u) = \left(\pi/32,0 \right)\) and plotted the resulting orbit
-in Figure \ref{fig:acrobot-orbit}.
+
+According to Theorem \ref{thm:acrobot-energy-stabilization}, the control
+parameter \(I\) must be small.
+While we could choose \(I\) arbitrarily small in simulations, we will eventually
+compare these simulations with the physical experiments where \(I\) must be
+large enough to overcome friction in leg motor.
+Setting \(\bar{q}_a = 1\), we experimentally determined that \(I = 10\) is a
+viable control value.
+This is the value of \(I\) we used in all simulations and experiments.
+
+\textbf{TODO: the injection and dissipation I is different, so we might not use
+I = 10 for dissipation.}
+
+\subsection{Energy Injection}
+
+To validate the energy injection property of our VNHC, we simulated the acrobot
+with initial condition \((q_u,p_u) = \left(\pi/32,0 \right)\) and plotted the
+resulting orbit in Figure \ref{fig:acrobot-orbit}.
 
@@
     \includegraphics[width=0.8\linewidth]{acrobot_orbit.png}
-    \caption{A simulated orbit of Wang's acrobot.}
+    \caption{A simulation of the acrobot gaining energy.}
     \label{fig:acrobot-orbit}
 \end{figure}
+
 To verify numerically that the acrobot would consistently achieve rotations, we
@@
 
-\subsection{Physical Experiments}
-
+\subsection{Energy Dissipation}
+
+\subsection{Energy Regulation}
+
+\section{Physical Experiments}\label{sec:experiments}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.9\linewidth]{xingbo_acrobot.jpg}
+    \caption{The acrobot built by Wang \cite{xingbo_thesis}.}
+    \label{fig:xingbo-acrobot}
+\end{figure}
+
+\subsection{Hardware Description}
+\textbf{TODO: paraphrase xingbo. 2 arduinos, 3d printed, battery in leg,
+parameters of system, etc}
+
+\subsection{Experimental Results}
+
+\textbf{TODO: rework this section. In simulation we showed that we gain energy
+continuously. Here we only gain energy up to speed 0.195 because of friction.}
 Wang's acrobot has an encoder which measures \(q_u\) and

### 2021-04-29 latex/paper/main.tex
First draft of simulation section, except for regulation.

@@
 \cite{xingbo_thesis}.
+\textbf{TODO: put them here if there is space}.
 If our VNHC injects energy into this acrobot, then we can infer that
@@
 
+\subsection{Energy Injection}
+
 According to Theorem \ref{thm:acrobot-energy-stabilization}, the control
-parameter \(I\) must be small.
-While we could choose \(I\) arbitrarily small in simulations, we will eventually
-compare these simulations with the physical experiments where \(I\) must be
-large enough to overcome friction in leg motor.
+parameter \(I\) must be small if one wants the VNHC to inject energy.
+While we could make \(I\) arbitrarily small in simulations, we will eventually
+implement this VNHC on a physical system where \(I\) must be large enough to
+overcome friction in leg motor.
 Setting \(\bar{q}_a = 1\), we experimentally determined that \(I = 10\) is a
-viable control value.
-This is the value of \(I\) we used in all simulations and experiments.
-
-\textbf{TODO: the injection and dissipation I is different, so we might not use
-I = 10 for dissipation.}
-
-\subsection{Energy Injection}
-
-To validate the energy injection property of our VNHC, we simulated the acrobot
-with initial condition \((q_u,p_u) = \left(\pi/32,0 \right)\) and plotted the
-resulting orbit in Figure \ref{fig:acrobot-orbit}.
-
-The set \(E_\pi := E(\pi,0)\) is outlined in black. 
+viable control value for energy injection.
+
+We simulated the acrobot with initial condition 
+\((q_u,p_u) = \left(\pi/32,0 \right)\) for \(30\) seconds and plotted the
+resulting orbit in Figure \ref{fig:acrobot-in-orbit}.
+
+The level set \(E_\pi\) with energy \(E(\pi,0)\) is outlined in black. 
 Due to the model mismatch, \(E_\pi\) is no longer the boundary between
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{acrobot_orbit.png}
+    \includegraphics[width=0.8\linewidth]{acrobot_in_orbit.png}
     \caption{A simulation of the acrobot gaining energy.}
-    \label{fig:acrobot-orbit}
+    \label{fig:acrobot-in-orbit}
 \end{figure}
@@
 
+According to Theorem \ref{thm:acrobot-energy-stabilization}, a negative value
+of \(I\) should dissipate energy from the acrobot.
+To test this, we initialized our acrobot with a rotation at the phase
+\((q_u,p_u) = (0,0.18)\),  set \(I = -10\), and simulated the constrained
+dynamics for \(30\) seconds.
+The resulting orbit is plotted in Figure \ref{fig:acrobot-diss-orbit}. 
+As is expected, the acrobot slows down over time.
+We again highlight the locations where the orbit crossed the set \(E_\pi\) in
+black, with the final crossing in red. 
+After this crossing, the acrobot ceased rotating and its oscillations decayed to
+zero.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\linewidth]{acrobot_diss_orbit.png}
+    \caption{A simulation of the acrobot dissipating energy.}
+    \label{fig:acrobot-diss-orbit}
+\end{figure}
+
 \subsection{Energy Regulation}
+
+To test how VNHCs can perform energy regulation, we employed a simple supervisor
+which toggles between injection and dissipation VNHCs to stabilize an
+orbit which crosses the \(q_u\)-axis at some desired angle (for oscillations) or
+the \(p_u\)-axis at some desired velocity (for rotations).
+Specifically, one chooses some appropriate energy value \(E_0 > 0\) and a hysteresis
+percentage \(\delta \in [0,1]\).
+The supervisor enables the injection VNHC if 
+\(E(q_u,p_u) < (1-\delta)E_0\), the dissipation VNHC if 
+\(E(q_u,p_u) > (1+\delta) E_0\), and it keeps the leg angle constant if
+\((1-\delta)E_0 \leq E(q_u,p_u) (1+\delta)\leq E_0\).
+
+First, we wanted the acrobot to oscillate with a maximum body angle 
+\(q_u = \pi/2\).
+The corresponding energy is \(E_0 = E(\pi/2,0)\), and we set our hysteresis to
+be \(\delta = 5\%\).
+\textbf{TODO: do the energy regulation experiment and see the results.}
+
+\subsection{Summary of Results}
+The simulation results in this section demonstrate the energy regulation
+capabilities of our VNHC.  
+In particular, Theorem \ref{thm:acrobot-energy-stabilization} appears to work
+even for acrobots whose limbs have differing masses and lengths.
+The energy regulation supervisor does not
 
@@
 \textbf{TODO: rework this section. In simulation we showed that we gain energy
-continuously. Here we only gain energy up to speed 0.195 because of friction.}
+everywhere. Here we only gain energy up to speed 0.195 because of friction.}
 Wang's acrobot has an encoder which measures \(q_u\) and
@@
 \subsection{Summary of Results} 
-These simulations and experiments demonstrate that VNHCs are excellent
-tools for injecting energy.
+These experiments demonstrate that VNHCs are excellent tools for injecting
+energy.
 In particular, energy injection through VNHCs appears to be robust against a
-variety of disturbances, including significant model mismatch and sensor noise.
+variety of disturbances, including significant model mismatch, friction, and
+sensor noise.
 

### 2021-04-29 matlab/thesis_sims/energyDissEvent.m
Updated the energy events because dissipation was going the wrong way, so now we need to update the image for dissipation too.

@@ -1,0 +1,25 @@
+%% Energy Level Events
+% Triggers an event in ode45 when the inputted energy E reaches a specific
+% value.
+% You use this in ode45 by setting the events function handle to
+% @(t,qp) energyEvent(t,qp,E,Eval).
+%
+% Inputs:
+% * t = current time value of simulation
+% * qp = current value of the phase on the (qu,pu)-plane
+% * E = function handle to the energy of the system
+% * Eval = energy value at which to trigger the event.
+
+function [value,isterminal,direction] = energyDissEvent(t,qp,E,Eval)
+    % Trigger when E(q,p) = Eval. The value increases as E -> Eval from
+    % below.
+    value = E(qp(1),qp(2))-Eval;
+    % Don't stop the integration, just record the event
+    isterminal = 0;
+    % Track only values where the energy is increasing past Eval, not those
+    % where you come back down. If you want it to track only decreasing,
+    % set direction to -1. If you want both increasing and decreasing, set
+    % it to 0.
+    direction = -1;
+end
+

### 2021-04-29 matlab/thesis_sims/energyInEvent.m

@@ -1,0 +1,25 @@
+%% Energy Level Events
+% Triggers an event in ode45 when the inputted energy E reaches a specific
+% value.
+% You use this in ode45 by setting the events function handle to
+% @(t,qp) energyEvent(t,qp,E,Eval).
+%
+% Inputs:
+% * t = current time value of simulation
+% * qp = current value of the phase on the (qu,pu)-plane
+% * E = function handle to the energy of the system
+% * Eval = energy value at which to trigger the event.
+
+function [value,isterminal,direction] = energyInEvent(t,qp,E,Eval)
+    % Trigger when E(q,p) = Eval. The value increases as E -> Eval from
+    % below.
+    value = E(qp(1),qp(2))-Eval;
+    % Don't stop the integration, just record the event
+    isterminal = 0;
+    % Track only values where the energy is increasing past Eval, not those
+    % where you come back down. If you want it to track only decreasing,
+    % set direction to -1. If you want both increasing and decreasing, set
+    % it to 0.
+    direction = 1;
+end
+

### 2021-05-10 latex/paper/main.tex
Wrote the rest of the simulation section, v1

@@
 The level set \(E_\pi\) with energy \(E(\pi,0)\) is outlined in black.
-Due to the model mismatch, \(E_\pi\) is no longer the boundary between
-rotations and oscillations.
-The oscillation domain is actually much larger: orbits rotate once they hit the
-\(p_u\)-axis at \(\abs{p_u} \approx 0.15\), while \(E_\pi\) intersects the
-\(p_u\)-axis at \(\abs{p_u} \approx 0.055\).
-Despite this, the VNHC still injects energy into the acrobot.
+Despite the change in acrobot model, this level set remains the boundary
+between oscillations and rotations for the nominal pendulum.
 The points where the orbit exits \(E_\pi\) are marked with black stars,
 with the final departure marked by a red star.
-After this final departure, the acrobot begins rotating and continues to gain
-energy over time.
+Interestingly, our choice of \(I\) is large enough that we observe significant
+differences between the nominal pendulum and the constrained dynamics:
+\(E_\pi\) intersects the \(p_u\)-axis at \(\abs{p_u} \approx 0.18\), yet the
+constrained acrobot begins rotating once it hits the
+\(p_u\)-axis at \(\abs{p_u} \approx 0.16\). 
+This indicates that higher values of \(I\) enable the acrobot to gain energy
+faster and begin rotating sooner, so long as the actuator does not saturate.
 
@@
 \] 
-and measured the how long it took to begin rotating.
+and measured how long it took to begin rotating.
 The results in Figure \ref{fig:acrobot-mc} show that
@@
 According to Theorem \ref{thm:acrobot-energy-stabilization}, a negative value
-of \(I\) should dissipate energy from the acrobot.
+of \(I\) will dissipate energy from the acrobot.
 To test this, we initialized our acrobot with a rotation at the phase
@@
 The resulting orbit is plotted in Figure \ref{fig:acrobot-diss-orbit}. 
-As is expected, the acrobot slows down over time.
+As expected, the acrobot slows down over time.
 We again highlight the locations where the orbit crossed the set \(E_\pi\) in
@@
 
-\subsection{Energy Regulation}
-
-To test how VNHCs can perform energy regulation, we employed a simple supervisor
-which toggles between injection and dissipation VNHCs to stabilize an
-orbit which crosses the \(q_u\)-axis at some desired angle (for oscillations) or
-the \(p_u\)-axis at some desired velocity (for rotations).
-Specifically, one chooses some appropriate energy value \(E_0 > 0\) and a hysteresis
-percentage \(\delta \in [0,1]\).
-The supervisor enables the injection VNHC if 
-\(E(q_u,p_u) < (1-\delta)E_0\), the dissipation VNHC if 
-\(E(q_u,p_u) > (1+\delta) E_0\), and it keeps the leg angle constant if
-\((1-\delta)E_0 \leq E(q_u,p_u) (1+\delta)\leq E_0\).
-
-First, we wanted the acrobot to oscillate with a maximum body angle 
-\(q_u = \pi/2\).
-The corresponding energy is \(E_0 = E(\pi/2,0)\), and we set our hysteresis to
-be \(\delta = 5\%\).
-\textbf{TODO: do the energy regulation experiment and see the results.}
+\subsection{Oscillation Regulation}
+
+Once can use a simple supervisor to stabilize oscillations by appropriately
+toggling between injection and dissipation VNHCs.
+Specifically, one chooses a desired oscillation angle 
+\(q_\text{des} \in \, ]0,\pi[\) and, to avoid infinite switching, a
+hysteresis value \(\delta \in [0,1]\) 
+Each time the orbit crosses the \(q_u\)-axis (i.e. \(p_u = 0\)) or 
+when \(|q_u| = \pi\), the supervisor changes which VNHC is enforced as follows:
+\begin{itemize}
+    \item If \(|q_u| < (1-\delta)q_\text{des}\), enable the injection VNHC.
+    \item If \(|q_u| > (1+\delta)q_\text{des}\), enable the dissipation VNHC.
+    \item If \((1-\delta)q_\text{des} \leq |q_u| \leq (1+\delta)q_\text{des}\),
+        keep the leg extended at \(q_a = 0\). This can be done continously since
+        \(q_a = 0\) when \(p_u = 0\).
+\end{itemize}
+
+Figure \ref{fig:acrobot-osc-reg} shows the supervisor stabilizing an
+oscillation with body angle \(q_\text{des} = \pi/2\) and a 
+\(5\%\) hysteresis, meaning \(\delta = 0.05\).
+The supervisor reevaluated its choice of VNHC at each black star;  
+the red circle corresponds to the part of the orbit where the supervisor kept
+the leg extended, because the oscillation was within tolerance of
+\(q_\text{des}\).
+
+In Figure \ref{fig:acrobot-osc-reg-in} the acrobot was initialized at 
+\((q_u,p_u) = (\pi/32,0)\); here the supervisor injected energy for most of the
+orbit.
+In Figure \ref{fig:acrobot-osc-reg-diss} the acrobot was initialized at the rotation
+\((q_u,p_u) = (0,0.19)\); here the supervisor dissipated energy. 
+In both cases, the supervisor stabilized the desired oscillation .
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[t]{0.49\linewidth}
+        \includegraphics[width=\linewidth]{acrobot_osc_orbit_1.png}
+        \caption{Stabilizing an oscillation from below.}
+        \label{fig:acrobot-osc-reg-in}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.49\linewidth}
+        \includegraphics[width=\linewidth]{acrobot_osc_orbit_2.png}
+        \caption{Stabilizing an oscillation from above.}
+        \label{fig:acrobot-osc-reg-diss}
+    \end{subfigure}
+    \caption{Using a supervisor to stabilize the oscillation with peak angle
+    \(q_\text{des} = \pi/2\).}
+    \label{fig:acrobot-osc-reg}
+\end{figure}
+
+\subsection{Rotation Regulation}
+Along with stabilizing oscillations, one can use a supervisor to stabilize
+rotations.
+In this case, one chooses a desired rotation rate \(p_\text{des} > 0\) and a
+hystersis value \(\delta \in [0,1]\).
+Each time the orbit crosses the \(p_u\)-axis (i.e. when \(q_u = 0\)), the
+supervisor changes which VNHC is enforced as follows:
+\begin{itemize}
+    \item If \(|p_u| < (1-\delta)p_\text{des}\), enable the injection VNHC.
+    \item If \(|p_u| > (1+\delta)p_\text{des}\), enable the dissipation VNHC.
+    \item If \((1-\delta)p_\text{des} \leq |p_u| \leq (1+\delta)p_\text{des}\),
+        extend the leg fully by setting \(q_a = 0\).
+        For these simulations, we assume this can be done instantaneously.
+\end{itemize}
+
+Rotation regulation is demonstrated in Figure \ref{fig:acrobot-rot-reg}, where
+the supervisor stabilizes \(p_\text{des} = 0.19\) with a \(2\%\) hysteresis 
+\(\delta = 0.02\).
+The supervisor changed its choice of VNHC at each black star until it finally extended
+the leg (shown in red).
+
+In Figure \ref{fig:acrobot-rot-reg-in} the acrobot was initialized at the
+oscillation \((q_u,p_u) = (\pi/2,0)\); here the supervisor injected energy until
+the orbit hit the \(p_u\)-axis near \(p_\text{des}\).
+In Figure \ref{fig:acrobot-rot-reg-diss} the acrobot was initialized at the
+faster rotation \((q_u,p_u) = (0, 0.23)\); the supervisor dissipated energy.
+In both cases, the desired rotation was stabilized correctly.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[t]{0.49\linewidth}
+        \includegraphics[width=\linewidth]{acrobot_rot_orbit_1.png}
+        \caption{Stabilizing a rotation from below.}
+        \label{fig:acrobot-rot-reg-in}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.49\linewidth}
+        \includegraphics[width=\linewidth]{acrobot_rot_orbit_2.png}
+        \caption{Stabilizing a rotation from above.}
+        \label{fig:acrobot-rot-reg-diss}
+    \end{subfigure}
+    \caption{Using a supervisor to stabilize the rotation with maximal momentum
+    \(p_\text{des} = 0.23\).}
+    \label{fig:acrobot-rot-reg}
+\end{figure}
 
@@
 capabilities of our VNHC.  
-In particular, Theorem \ref{thm:acrobot-energy-stabilization} appears to work
-even for acrobots whose limbs have differing masses and lengths.
-The energy regulation supervisor does not
+We were able to stabilize both oscillations and rotations by implementing a
+control supervisor which toggled between injection, dissipation, and
+leg-extension VNHCs.
+In particular, these simulations indicate that our VNHC
+works even for acrobots whose limbs have differing masses and lengths.
 

### 2021-05-10 latex/paper/main.tex
Modified the monte carlo description
Wrote the first draft of the hardware description

@@
 In Section \ref{sec:experiments} we will test our VNHC on the physical acrobot
-built by Xingbo Wang \cite{xingbo_thesis} (Figure \ref{fig:xingbo-acrobot}).
-To compare apples to apples, we will simulate our VNHC using the parameters of
-the physical system (Table \ref{tab:acrobot-parameters}).
+built by Xingbo Wang \cite{xingbo_thesis}.
+
+Note that this acrobt does not correspond to the simple acrobot in Figure
+\ref{fig:simple-acrobot-model}, but rather to the more general acrobot model in
+Figure \ref{fig:acrobot-model}.
+If we define 
+\begin{align*}
+    m_{11}(q) := &m_a l_u^2 + 2m_al_ul_{c_a}\cos(q_a) + m_al_{c_a}^2 +
+    m_ul_{c_u}^2 \\
+                 &+ J_u + J_a
+                 , \\
+    m_{12}(q) := &m_al_{c_a}^2 + m_al_ul_{c_a}\cos(q_a) + J_a
+    , \\
+    m_{22}(q) := &m_al_{c_a}^2 + J_a,
+\end{align*}
+then the inertia matrix and potential for the physical acrobot are
+\begin{align*}
+    M(q) &= \begin{bmatrix}
+        m_{11}(q) & m_{12}(q) \\
+        m_{12}(q) & m_{22}(q)
+    \end{bmatrix}
+    , \\
+    V(q) &= g\big(m_al_{c_a}(1-c_{ua})
+    + (m_al_u + m_ul_{c_u})(1-c_u)\big).
+\end{align*}
 
@@
 \end{figure}
+
+Evaluating the true mechanical energy of this acrobot at the VNHC
+\(q_a = 0\) yields the energy of the nominal pendulum,
+\[
+    E(q_u,p_u) \approx 396.5501 p_u^2 + 0.5997(1 - \cos(q_u))
+    .
+\]
+
+We will simulate our VNHC using the parameters of
+this physical robot, which can be found in Table \ref{tab:acrobot-parameters}.
+If our VNHC injects energy into this acrobot, then we can infer that
+VNHCs are robust to model mismatch.
 
@@
 
-Note that these parameters do not correspond to our simplified acrobot in Figure
-\ref{fig:simple-acrobot-model}, but rather to the more general acrobot model in
-Figure \ref{fig:acrobot-model}.
-In particular, the inertia matrix and potential function are \textbf{not}
-described by ~\eqref{eqn:acrobot-inertia}-\eqref{eqn:acrobot-potential}.
-The correct \(M(q)\) and \(V(q)\) can be found in Wang's thesis
-\cite{xingbo_thesis}.
-\textbf{TODO: put them here if there is space}.
-If our VNHC injects energy into this acrobot, then we can infer that
-VNHCs are robust to model mismatch.
-
-Evaluating the true mechanical energy of this acrobot at the VNHC
-\(q_a = 0\) yields the energy of the nominal pendulum,
-\[
-    E(q_u,p_u) \approx 396.5501 p_u^2 + 0.5997(1 - \cos(q_u))
-    .
-\]
-
 \subsection{Energy Injection}
@@
 The results in Figure \ref{fig:acrobot-mc} show that
-the acrobot always rotated within 10--35 seconds.
+the acrobot always rotated within 20--40 seconds.
 
@@
 
+\subsection{Hardware Description}
+In this section we will demonstrate that our VNHC is robust to friction, sensor
+noise, and other real-world considerations by testing it on the physical
+acrobot built by Xingbo Wang \cite{xingbo_thesis} 
+(Figure \ref{fig:xingbo-acrobot}.
+
 \begin{figure}
@@
 
-\subsection{Hardware Description}
-\textbf{TODO: paraphrase xingbo. 2 arduinos, 3d printed, battery in leg,
-parameters of system, etc}
+This acrobot is comprised of two 3D-printed links: a torso and a leg.
+The torso link houses an Arduino Nano microcontroller unit (MCU) which controls
+a Dynamixel RX24F servo motor between the torso and the leg.
+The MCU and the motor are powered by a 12V battery held in a compartment
+in the leg.
+
+The torso is rigidly attached to a metal bar, which is held up by two wooden
+posts.
+On the exterior of one post is a control box with a power switch and a second
+Arduino Nano 328.
+The purpose of this control box is to read measurements from a rotary
+encoder connected to the metal bar, and to transmit these measurements to the
+MCU.
+The two Arduinos communicate through wires attached to a slip ring on the metal
+bar, and the signals are transmitted via I2C.
+The control box also provides a USB interface which allows the user to read the
+data from the acrobot in real time.
+
+VNHC implementation requires full knowledge of the acrobot's phase. 
+The rotary encoder directly measures \(q_u\), and the Dynamixel servo
+motor provides measurements of \(q_a\).
+However, there are no sensors measuring the velocity \(\dot{q}\), which means we cannot
+directly evaluate \(p_u\) and \(p_a\).
+To resolve this issue, the MCU estimates \(\dot{q}\) using sequential measurements of \(q\)
+and a washout filter.
+We then compute \(p = M(q)\dot{q}\) for use in the VNHC controller.
+
+Communication between the Arduinos restricts the sampling rate of \(p\) to
+500Hz.
+This low sampling rate results in a noisy momentum signal, which suffers
+from noticeable phase lag.
+This also rate-limits the control signal to \(500\)Hz, which impacts any control
+implementation.
+
+Finally, The Dynamixel servo motor does not have a torque control mode; instead,
+we can assign the servo setpoint at iteration \(k \in \mathbb{Z}_{> 0}\)
+via \(q_a^{k} = \arctan(10 p_u^{k-1})\).
+Unfortunately, this negatively affects the stabilization to the constraint
+manifold because we are introducing timing errors from the servo's built-in PID
+controller.
+
+The system was given the name SUGAR: Simple Underactuated Gymnastics and
+Acrobatics Robot.
+Its dynamic parameters are outlined in Table \ref{tab:acrobot-parameters}.
 
@@
 everywhere. Here we only gain energy up to speed 0.195 because of friction.}
-Wang's acrobot has an encoder which measures \(q_u\) and
-\(\dot{q}_u\).
-Additionally, the leg actuator is a servo motor with a built-in PID controller
-which provides measurements of \(q_a\).
-To implement our VNHC, we estimated \(\dot{q}_a\) through sequential values of
-\(q_a\) and computed \(p_u\) through \(p_u = e_1^T M(q) \dot{q}\).
-We directly assigned the actuator setpoint at iteration \(k \in \mathbb{Z}_{> 0}\)
-via \(q_a^{k} = \arctan(10 p_u^{k-1})\).
-We then performed the following tests.
+We performed the following tests.
 

### 2021-05-11 latex/paper/main.tex
Updated the experimental results section. Everything is done its first draft except the conclusion and the proof of theorem 10.

@@
 avoid infinite switching.
-As we shall see, such a simple supervisor will not produce human-like
-oscillations or rotations.
-However, it will allow the acrobot to stabilize some peak angle of oscillation
-\(\bar{q}_u \in \, ]0,\pi[\) so that \(E(\pm \bar{q}_u,0) = \bar{E}\),
-or some peak rotational momentum
-\(\bar{p}_u \in \, ]\sqrt{60m^2gl^3}, \bar{\rho}[\) where
-\(E(0,\pm \bar{p}_u) = \bar{E}\).
 
@@
 
-\textbf{TODO: rework this section. In simulation we showed that we gain energy
-everywhere. Here we only gain energy up to speed 0.195 because of friction.}
+In Section \ref{sec:simulations} we showed that VNHCs can be used for energy
+injection, dissipation, and regulation. 
+We will now demonstrate that VNHC-based giant motion is robust to external
+disturbances by implementing our VNHC on the SUGAR system.
 We performed the following tests.
@@
     we initialized the acrobot at 
-    \((q_u,p_u) \approx \left(\frac{\pi}{8},0\right)\). 
-    The resulting orbit is shown in Figure \ref{fig:acrobot-unperturbed-orbit};
-    the acrobot is clearly gaining energy over time.
+    \((q_u,p_u) \approx \left(\pi/8,0\right)\). 
+    The resulting orbit in Figure \ref{fig:acrobot-unperturbed-orbit} shows
+    the acrobot clearly gaining energy over time.
+    Its motion looks similar to that of the energy injection simulation (Figure
+    \ref{fig:acrobot-in-orbit}), though the acrobot appears to stop gaining
+    energy once it reaches a rotation with energy \(E(0,0.195)\).
 
@@
     \centering
-    \begin{subfigure}[t]{0.49\linewidth}
-        \includegraphics[width=\linewidth]{acrobot_unperturbed_orbit.png}
-        \caption{The baseline test.}
+    \includegraphics[width=0.8\linewidth]{acrobot_unperturbed_orbit.png}
+    \caption{The physical acrobot's orbit during the baseline energy injection test.}
     \label{fig:acrobot-unperturbed-orbit}
-    \end{subfigure}
-    \begin{subfigure}[t]{0.49\linewidth}
-        \includegraphics[width=\linewidth]{acrobot_stopped_orbit.png}
-        \caption{Acrobot orbit before (blue) and after (red) stopping.}
-        \label{fig:acrobot-stopped-orbit}
-    \end{subfigure}
-    \caption{The physical acrobot's orbit during baseline and stop tests.}
 \end{figure}
@@
     The resulting orbit is shown in Figure \ref{fig:acrobot-stopped-orbit}.
-    The blue curves correspond to the orbit before the disturbance, while the red
-    spiral confirms that the acrobot begins oscillating after it is stopped.
+    The blue rotation curve corresponds to the orbit before the disturbance,
+    while the red spiral confirms that the acrobot begins oscillating after it
+    is stopped.  
     Despite the disturbance, it gains energy and eventually starts rotating
     again.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\linewidth]{acrobot_stopped_orbit.png}
+    \caption{The physical acrobot's orbit before (blue) and after (red) stopping.}
+    \label{fig:acrobot-stopped-orbit}
+\end{figure}
 
@@
     direction of motion.
-    The orbit in Figure \ref{fig:acrobot-fpush-orbit} shows that the acrobot speeds
-    up to rotate with energy \(E(0,-0.22)\), but then slows down until it reaches a
-    rotation with energy \(E(0,-0.195)\).
+    The orbit in Figure \ref{fig:acrobot-fpush-orbit} shows that the acrobot,
+    when pushed, rotates with energy \(E(0,-0.22)\), but then slows down until
+    it reaches a rotation with energy \(E(0,-0.195)\).
     We repeated this test by pushing the acrobot against its direction of motion.
     The orbit in Figure \ref{fig:acrobot-rpush-orbit} demonstrates that the acrobot
-    responds by readily changing direction, and quickly achieves its maximum speed
+    readily changes direction, and quickly achieves its maximum speed
     with energy \(E(0,0.195)\).
-
-    These push tests suggest that our VNHC injects energy into Wang's acrobot on
-    \(\mathcal{O}_2(0.195)\).
 \end{enumerate}
+
+In simulation, the acrobot was able to gain energy even when initialized
+with energy \(E(q_u,p_u) > E(0,0.195)\). 
+The baseline and push tests suggest that our VNHC injects energy into the SUGAR
+system only on \(\mathcal{O}_2(0.195)\).
+This difference between simulation and implementation is likely due to
+friction.
 
@@
 
+We had hoped to show experimental results for dissipation and
+oscillation/rotation regulation, but unfortunately our Dynamixel RX24F servo
+motor broke before we could perform these experiments.
+These motors are no longer in production at the time of writing, so the SUGAR
+system has been retired for the time being.
+
 \subsection{Summary of Results} 
-These experiments demonstrate that VNHCs are excellent tools for injecting
-energy.
-In particular, energy injection through VNHCs appears to be robust against a
-variety of disturbances, including significant model mismatch, friction, and
-sensor noise.
+We performed three tests on the SUGAR system: a baseline energy injection test,
+a stop test, and a push test.
+These experiments demonstrate that VNHC-based energy injection is robust to
+significant model mismatch, friction, sensor noise, discretized control
+implementation, rate-limited measurement and control signals, and dramatic
+external disturbances.
 

### 2021-05-12 latex/paper/main.tex
First edit of the simulation and experiment sections

@@
 
-The results of Theorem \ref{thm:acrobot-energy-stabilization} lead naturally to
-the idea of energy regulation: one could employ a supervisor which switches
-between injection and dissipation constraints to stabilize a desired energy
-level \(\bar{E}\).
-For instance, the supervisor could inject energy when \(E(q_u,p_u) < \bar{E}\) 
-and dissipate energy when \(E(q_u,p_u) > \bar{E}\), with some hysteresis to
-avoid infinite switching.
+One can use the results of Theorem \ref{thm:acrobot-energy-stabilization} 
+for energy regulation.
+For example, one might employ a supervisor which switches
+between injection and dissipation VNHCs to stabilize a desired energy
+level \(E_\text{des}\), with some hysteresis to avoid infinite switching.
+We will explore this idea in Section \ref{sec:simulations}.
 
@@
 built by Xingbo Wang \cite{xingbo_thesis}.
-
-Note that this acrobt does not correspond to the simple acrobot in Figure
-\ref{fig:simple-acrobot-model}, but rather to the more general acrobot model in
-Figure \ref{fig:acrobot-model}.
+This physical acrobot cannot be modelled by the simplified setup in Figure
+\ref{fig:simple-acrobot-model}, because the torso and leg links have 
+distributed mass. 
+We must therefore use the more general acrobot model in Figure
+\ref{fig:acrobot-model} to represent this system.
 If we define 
@@
 \end{align*}
-then the inertia matrix and potential for the physical acrobot are
-\begin{align*}
-    M(q) &= \begin{bmatrix}
+where \(J_u\) and \(J_a\) are the moments of inertia of the torso and leg links
+respectively, then the inertia matrix becomes
+\begin{equation*}
+    M(q) = \begin{bmatrix}
         m_{11}(q) & m_{12}(q) \\
@@
     \end{bmatrix}
-    , \\
-    V(q) &= g\big(m_al_{c_a}(1-c_{ua})
-    + (m_al_u + m_ul_{c_u})(1-c_u)\big).
-\end{align*}
+    ,
+\end{equation*}
+and the potential function is
+\begin{equation*}
+    V(q) = g\big(m_al_{c_a}(1-c_{ua})
+    + (m_al_u + m_ul_{c_u})(1-c_u)\big)
+    .
+\end{equation*}
 
@@
 
-Evaluating the true mechanical energy of this acrobot at the VNHC
-\(q_a = 0\) yields the energy of the nominal pendulum,
+The dynamic parameters of Wang's acrobot can be found in Table
+\ref{tab:acrobot-parameters}.
+Evaluating the mechanical energy of an acrobot with these parameters
+at the VNHC \(q_a = 0\) yields the mechanical energy of the nominal pendulum,
 \[
@@
 
-We will simulate our VNHC using the parameters of
-this physical robot, which can be found in Table \ref{tab:acrobot-parameters}.
-If our VNHC injects energy into this acrobot, then we can infer that
-VNHCs are robust to model mismatch.
+We will now simulate the effects of constraining Wang's acrobot with
+~\eqref{eqn:acrobot-constraint}, thereby demonstrating that VNHCs are robust to
+model mismatch.
+According to Theorem \ref{thm:acrobot-energy-stabilization}, the control
+parameter \(I\) must be ``small" for our VNHC to inject (or dissipate) energy
+into the simplified acrobot. 
+Unfortunately, the Theorem does not specify how small \(|I|\) must be;
+while we could make it arbitrarily small in simulations, we will eventually
+implement this VNHC on a physical testbed where \(|I|\) must be large enough to
+overcome friction in the leg motor.
+
+Setting \(\bar{q}_a = 1\), we experimentally determined that \(|I| = 10\) is a
+viable control parameter, so this is the value we will use for all simulations
+and experiments.
+In other words, our injection VNHC is ~\eqref{eqn:acrobot-constraint} with 
+\(I = 10\) while our dissipation VNHC is ~\eqref{eqn:acrobot-constraint} with 
+\(I = -10\).
 
@@
 
-According to Theorem \ref{thm:acrobot-energy-stabilization}, the control
-parameter \(I\) must be small if one wants the VNHC to inject energy.
-While we could make \(I\) arbitrarily small in simulations, we will eventually
-implement this VNHC on a physical system where \(I\) must be large enough to
-overcome friction in leg motor.
-Setting \(\bar{q}_a = 1\), we experimentally determined that \(I = 10\) is a
-viable control value for energy injection.
-
-We simulated the acrobot with initial condition 
-\((q_u,p_u) = \left(\pi/32,0 \right)\) for \(30\) seconds and plotted the
-resulting orbit in Figure \ref{fig:acrobot-in-orbit}.
+We simulated the acrobot constrained by our injection VNHC
+with initial condition \((q_u,p_u) = \left(\pi/32,0 \right)\) for \(30\)
+seconds, and plotted the resulting orbit in Figure \ref{fig:acrobot-in-orbit}.
 
@@
 To verify numerically that the acrobot would consistently achieve rotations, we
-ran a Monte-Carlo simulation \cite{montecarlo} where we initialized the acrobot
+ran a Monte-Carlo \cite{montecarlo} simulation where we initialized the acrobot
 randomly inside the sublevel set
@@
     \includegraphics[width=0.8\linewidth]{acrobot_mc.png}
-    \caption{Monte Carlo simulation results.}
+    \caption{Monte Carlo simulation for energy injection.}
     \label{fig:acrobot-mc}
@@
 
-According to Theorem \ref{thm:acrobot-energy-stabilization}, a negative value
-of \(I\) will dissipate energy from the acrobot.
-To test this, we initialized our acrobot with a rotation at the phase
-\((q_u,p_u) = (0,0.18)\),  set \(I = -10\), and simulated the constrained
-dynamics for \(30\) seconds.
+We initialized our acrobot with a rotation at \((q_u,p_u) = (0,0.18)\) and
+simulated the constrained dynamics (under the dissipation VNHC) for \(30\)
+seconds.
 The resulting orbit is plotted in Figure \ref{fig:acrobot-diss-orbit}. 
 As expected, the acrobot slows down over time.
-We again highlight the locations where the orbit crossed the set \(E_\pi\) in
-black, with the final crossing in red. 
+We highlight the locations where the orbit crossed the set \(E_\pi\) by
+black stars, with the final crossing in red.
 After this crossing, the acrobot ceased rotating and its oscillations decayed to
@@
 
-Once can use a simple supervisor to stabilize oscillations by appropriately
+Once can use a supervisor to stabilize oscillations by appropriately
 toggling between injection and dissipation VNHCs.
@@
 \(q_\text{des} \in \, ]0,\pi[\) and, to avoid infinite switching, a
-hysteresis value \(\delta \in [0,1]\) 
-Each time the orbit crosses the \(q_u\)-axis (i.e. \(p_u = 0\)) or 
-when \(|q_u| = \pi\), the supervisor changes which VNHC is enforced as follows:
+hysteresis value \(\delta \in [0,1]\).
+Each time the orbit crosses the \(q_u\)-axis (i.e. when \(p_u = 0\)) or 
+when \(|q_u| = \pi\), the supervisor does the following: 
 \begin{itemize}
@@
 The supervisor reevaluated its choice of VNHC at each black star;  
-the red circle corresponds to the part of the orbit where the supervisor kept
+the red contour corresponds to the part of the orbit where the supervisor kept
 the leg extended, because the oscillation was within tolerance of
 \(q_\text{des}\).
+The solid black line is the desired oscillation, and the dashed black
+lines show the hysteresis around that orbit.
 
@@
 \((q_u,p_u) = (0,0.19)\); here the supervisor dissipated energy. 
-In both cases, the supervisor stabilized the desired oscillation .
+In both cases, the supervisor stabilized the desired oscillation.
 
@@
 \subsection{Rotation Regulation}
-Along with stabilizing oscillations, one can use a supervisor to stabilize
-rotations.
+One can also use a supervisor to stabilize rotations.
 In this case, one chooses a desired rotation rate \(p_\text{des} > 0\) and a
-hystersis value \(\delta \in [0,1]\).
+hysteresis value \(\delta \in [0,1]\).
 Each time the orbit crosses the \(p_u\)-axis (i.e. when \(q_u = 0\)), the
@@
 \(\delta = 0.02\).
-The supervisor changed its choice of VNHC at each black star until it finally extended
-the leg (shown in red).
-
-In Figure \ref{fig:acrobot-rot-reg-in} the acrobot was initialized at the
-oscillation \((q_u,p_u) = (\pi/2,0)\); here the supervisor injected energy until
+The supervisor changed its choice of VNHC at each black star until it was within
+range of \(p_\text{des}\), at which point it extended the leg (shown in red).
+
+In Figure \ref{fig:acrobot-rot-reg-in} the acrobot was initialized at 
+\((q_u,p_u) = (\pi/2,0)\); here the supervisor injected energy until
 the orbit hit the \(p_u\)-axis near \(p_\text{des}\).
@@
 
+Note the difference in shape between the blue rotations of the dissipation VNHC
+and the red rotation of the nominal pendulum: the red one slows down much more
+near \(|q_u| = \pi\).
+This difference arises because of the size of \(I\):
+if \(|I|\) were smaller, the blue rotations would be more similar in shape to the
+red one because the constrained dynamics for the dissipation VNHC would be
+well approximated by the nominal pendulum.
+
 \subsection{Summary of Results}
@@
 acrobot built by Xingbo Wang \cite{xingbo_thesis} 
-(Figure \ref{fig:xingbo-acrobot}.
+(Figure \ref{fig:xingbo-acrobot}).
+This platform was given the name SUGAR, which stands for
+Simple Underactuated Gymnastics and Acrobatics Robot.
+Its dynamic parameters are outlined in Table \ref{tab:acrobot-parameters}.
 
@@
     \includegraphics[width=0.9\linewidth]{xingbo_acrobot.jpg}
-    \caption{The acrobot built by Wang \cite{xingbo_thesis}.}
+    \caption{SUGAR is the physical acrobot built by Wang \cite{xingbo_thesis}.}
     \label{fig:xingbo-acrobot}
@@
 
-This acrobot is comprised of two 3D-printed links: a torso and a leg.
+SUGAR is comprised of two 3D-printed links: a torso and a leg.
 The torso link houses an Arduino Nano microcontroller unit (MCU) which controls
@@
 
-VNHC implementation requires full knowledge of the acrobot's phase. 
 The rotary encoder directly measures \(q_u\), and the Dynamixel servo
@@
 
-The system was given the name SUGAR: Simple Underactuated Gymnastics and
-Acrobatics Robot.
-Its dynamic parameters are outlined in Table \ref{tab:acrobot-parameters}.
-
 \subsection{Experimental Results}
 
-In Section \ref{sec:simulations} we showed that VNHCs can be used for energy
-injection, dissipation, and regulation. 
-We will now demonstrate that VNHC-based giant motion is robust to external
-disturbances by implementing our VNHC on the SUGAR system.
-We performed the following tests.
+We performed the following tests on SUGAR.
 
@@
     The resulting orbit in Figure \ref{fig:acrobot-unperturbed-orbit} shows
-    the acrobot clearly gaining energy over time.
+    that SUGAR clearly gaining energy over time.
     Its motion looks similar to that of the energy injection simulation (Figure
-    \ref{fig:acrobot-in-orbit}), though the acrobot appears to stop gaining
-    energy once it reaches a rotation with energy \(E(0,0.195)\).
+    \ref{fig:acrobot-in-orbit}), though its energy gain ceases once it reaches a
+    rotation with energy \(E(0,0.195)\).
 
@@
     \includegraphics[width=0.8\linewidth]{acrobot_unperturbed_orbit.png}
-    \caption{The physical acrobot's orbit during the baseline energy injection test.}
+    \caption{SUGAR's orbit during the baseline energy injection test.}
     \label{fig:acrobot-unperturbed-orbit}
@@
 \item \textbf{Stop Test:}
-    we initialized the acrobot at 
+    we initialized SUGAR at 
     \((q_u,p_u) \approx \left(\pi,0\right)\), let it run for 15
@@
     \includegraphics[width=0.8\linewidth]{acrobot_stopped_orbit.png}
-    \caption{The physical acrobot's orbit before (blue) and after (red) stopping.}
+    \caption{SUGAR's orbit before (blue) and after (red) stopping.}
     \label{fig:acrobot-stopped-orbit}
@@
 \item \textbf{Push Test:}
-    to see how the acrobot responds when pushed, we allowed
-    the acrobot to rotate undisturbed for 15 seconds and then pushed it in its
+    to see how SUGAR responds when pushed, we allowed
+    it to rotate undisturbed for 15 seconds and then pushed it in its
     direction of motion.
@@
     it reaches a rotation with energy \(E(0,-0.195)\).
-    We repeated this test by pushing the acrobot against its direction of motion.
+    We repeated this test by pushing SUGAR against its direction of motion.
     The orbit in Figure \ref{fig:acrobot-rpush-orbit} demonstrates that the acrobot
@@
 with energy \(E(q_u,p_u) > E(0,0.195)\). 
-The baseline and push tests suggest that our VNHC injects energy into the SUGAR
-system only on \(\mathcal{O}_2(0.195)\).
+The baseline and push tests suggest that our VNHC injects energy into SUGAR
+only on \(\mathcal{O}_2(0.195)\).
 This difference between simulation and implementation is likely due to
@@
     \end{subfigure}
-    \caption{The acrobot orbit before (blue) and after (red) pushing.}
+    \caption{SUGAR's orbit before (blue) and after (red) pushing.}
 \end{figure}
 
-We had hoped to show experimental results for dissipation and
-oscillation/rotation regulation, but unfortunately our Dynamixel RX24F servo
+We had hoped to show more experimental results for dissipation and
+oscillation/rotation regulation, but unfortunately the Dynamixel RX24F servo
 motor broke before we could perform these experiments.
-These motors are no longer in production at the time of writing, so the SUGAR
-system has been retired for the time being.
+These motors are no longer in production at the time of writing this article, so
+SUGAR met an untimely end to its career.
 
 \subsection{Summary of Results} 
-We performed three tests on the SUGAR system: a baseline energy injection test,
+We performed three tests on SUGAR: a baseline energy injection test,
 a stop test, and a push test.

### 2021-05-12 latex/paper/main.tex
Wrote the conclusion and abstract

@@
 \begin{abstract}
-TODO: Abstract here.
+    In this article we study virtual nonholonomic constraints, which are
+    relations between the generalized coordinates and generalized momenta of a
+    mechanical system that can be enforced via feedback control.
+    We design a constraint which emulates gymnastics giant motion in an
+    acrobot, and prove rigorously that this constraint will inject or dissipate
+    energy.
+    This constraint is tested on a real-world acrobot, demonstrating highly
+    effective energy regulation properties and robustness to a variety of
+    disturbances.
 \end{abstract}
@@
 \begin{IEEEkeywords}
-TODO: keywords in alphabetical order, separated by commas.
+    energy regulation, virtual nonholonomic constraints, acrobot, gymnastics.
 \end{IEEEkeywords}
@@
 \section{Conclusion}\label{sec:conclusion}
-In this article we summarized the existing theory of virtual nonholonomic
-constraints (VNHCs) ........
-\textbf{TODO conclude nicely}
-In the end, we demonstrated that virtual nonholonomic constraints are capable of
+
+In this article we redeveloped the framework of virtual nonholonomic
+constraints for underactuated mechanical systems.
+We applied this framework towards the acrobot, and designed a constraint which
+emulates giant motion from gymnastics.
+We proved this constraint will inject energy into the simplified acrobot whose
+limbs are massless rods of equal length, with equal masses at the tips.
+Then we performed simulations of this VNHC, as well as physical experiments.
+These demonstrated that virtual nonholonomic constraints are capable of
 injecting and dissipating energy in a robust manner, all while producing

### 2021-05-13 latex/paper/main.tex
Final editing pass fo the paper, which is complete except for the proof of Thm 10. Sending to Manfredi.

@@
     We design a constraint which emulates gymnastics giant motion in an
-    acrobot, and prove rigorously that this constraint will inject or dissipate
+    acrobot, and rigorously prove that this constraint will inject or dissipate
     energy.
@@
 more realistic model is the two-link acrobot (Figure \ref{fig:acrobot}).
-Here, the top link represents the torso, while the bottom link represents
+Here, the top link represents the torso while the bottom link represents
 the legs. 
@@
 Indeed, they produce more robust walking motion in biped robots than
-other controllers which do not depend on velocity \cite{nhvc_incline_walking}.
+other virtual constraints which do not depend on velocity
+\cite{nhvc_incline_walking}.
 In particular, VNHCs may be capable of injecting and
@@
 We use the following notation and terminology in this article.
+The \(n \times n\) identity matrix is denoted \(\Id{n}\), and the \(n \times m\)
+matrix of zeros is denoted \(\Zmat{n\times m}\).
 A matrix \(A \in \R^{n \times m}\) is \textit{right semi-orthogonal} if
@@
 on top of \(B\). 
+Given \(\sigma_1,\ldots,\sigma_n \in \R\), we define 
+\(\diag{\sigma_1,\ldots,\sigma_n} \in \R^{n \times n}\) as the diagonal matrix
+whose value at row \(i\), column \(i\) is \(\sigma_i\).
 For \(T > 0\), the set of real numbers modulo \(T\) is denoted \(\Rt{T}\), with
@@
 \(Q_a \in [\frac{\pi}{2}, \pi[\). 
-We also ignore any frictional forces.
+We also ignore any dissipative forces.
 
@@
     Let \(f : \mathcal{Q} \rightarrow \mathcal{Q}\times\R^n\) be a smooth vector
-    field and let \(D \subset M\) be open.
+    field and let \(D \subset \mathcal{Q}\) be open.
     The system described by \(\dot{x} = f(x)\) 
@@
 
-It is easy to show using the matrix Kronecker product that
+Using the matrix Kronecker product, it is easy to show that
 \eqref{eqn:hamiltonian-eom-general} expands to
@@
 through state feedback.
-Horn et. al later extended their results
+Horn et. al later extended their results in
 \cite{hybrid_zero_dynamics_bipedal_nhvcs} to derive the constrained dynamics for
@@
 If any component of \(e_i\) has relative degree 1, we may not be able
-to stabilize \(\Gamma\) -- we can always guarantee \(e_i \to 0\), but not
+to stabilize \(\Gamma\): we can always guarantee \(e_i \to 0\), but not
 necessarily \(\dot{e}_i \to 0\).
@@
 controller
-\begin{equation}
+\begin{equation}\label{eqn:stabilizing-controller}
     \tau(q,p) = -H\inv(q,p)\left(E(q,p) + k_p e + k_d \dot{e}\right)
@@
 function of the unactuated variables.
-Intuitively then, the constrained dynamics should be parameterized by \((q_u, p_u)\).
+Regular constraints of this form always meet the mild conditions from
+\cite{vhcs_for_el_systems}, and hence we can stabilize the constraint manifold
+using ~\eqref{eqn:stabilizing-controller}.
+Since \(q_a\) is constrained to be a function of the unactuated phase variables,
+intuition tells us the constrained dynamics should be parameterized by 
+\((q_u, p_u)\).
 Unfortunately, \(\dot{q}_u\) depends on \(p_a\), and for general systems one
 cannot solve explicitly for \(p_a\) in terms of \((q_u,p_u)\) because
-the \(\dot{p}\) dynamics contains the coupling term 
+the \(\dot{p}\) dynamics contain the coupling term 
 \((\Id{n} \otimes p\tpose)\nabla_{q}M(q)p\).
@@
 We now introduce an assumption which allows us to solve for \(p_a\) as a
-function of \((q_u,p_u)\), which in turn allows us to find the constrained
-dynamics.
+function of \((q_u,p_u)\), which in turn allows us to explicitly solve 
+for the constrained dynamics.
 
@@
     where
-    \begin{align}\label{eqn:g-qpu}
-        &g(q_u,p_u) := 
-        \left(\mathcal{A}(q,p_u)[\Zmat{(n-k)\times k};\Id{k}] \right)\inv 
-        (dh_{p_u} \nabla_{q_u}V(q) \nonumber
+    \begin{multline}\label{eqn:g-qpu}
+        g(q,p_u) := \left( \mathcal{A}(q,p_u)
+        \begin{bmatrix} \Zmat{(n-k)\times k} \\ \Id{k} \end{bmatrix}\right)\inv 
+        \cdot
+        \Big(dh_{p_u}\nabla_{q_u}(V(q) -
         \\
-        &- \mathcal{A}(q,p_u)[\Id{(n-k)};\Zmat{k \times(n-k)}]p_u
-        \left.)\right|_{q_a = f(q_u,p_u)}
-        .
-    \end{align}
+        \mathcal{A}(q,p_u)
+        \begin{bmatrix} \Id{(n-k)} \\ \Zmat{k \times (n-k)} \end{bmatrix} p_u
+        \left.\Big)\right|_{q_a = f(q_u,p_u)}
+    \end{multline}
+
+%    \begin{align}\label{eqn:g-qpu}
+%        &g(q_u,p_u) := 
+%        \left(\mathcal{A}(q,p_u)[\Zmat{(n-k)\times k};\Id{k}] \right)\inv 
+%        (dh_{p_u} \nabla_{q_u}V(q) \nonumber
+%        \\
+%        &- \mathcal{A}(q,p_u)[\Id{(n-k)};\Zmat{k \times(n-k)}]p_u
+%        \left.)\right|_{q_a = f(q_u,p_u)}
+%        .
+%    \end{align}
 \end{thm}
@@
 assumption \textbf{H3} holds true.
-The only real distinction between this section and their work
-is that our constrained dynamics are explicit functions of the Hamiltonian phase
-coordinates \((q_u,p_u)\).
+The only real distinction between this section and their work is that we allow
+for higher degrees of underactuation, and our constrained dynamics are explicit
+functions of the Hamiltonian phase coordinates \((q_u,p_u)\).
 In fact, our constrained dynamics \eqref{eqn:qpu-dynamics} coincide with their
@@
 on a swing by pivoting the legs toward the direction of motion. 
-Since a person lifts their legs higher at higher speeds, the acrobot's legs should
+Since a person lifts their legs higher at faster speeds, the acrobot's legs should
 pivot to an angle proportional to the swing's speed.
@@
 It instead pivots the legs partially during rotations.
-However, the behaviour looks similar enough that this constraint will
+However, the behaviour is similar enough that this constraint will
 still inject energy into the acrobot.
@@
 
-One can use the results of Theorem \ref{thm:acrobot-energy-stabilization} 
-for energy regulation.
+One can apply the results of Theorem \ref{thm:acrobot-energy-stabilization} 
+towards energy regulation.
 For example, one might employ a supervisor which switches
@@
 \ref{fig:acrobot-model} to represent this system.
-If we define 
+Let us define 
 \begin{align*}
@@
 where \(J_u\) and \(J_a\) are the moments of inertia of the torso and leg links
-respectively, then the inertia matrix becomes
+respectively. 
+The general acrobot has inertia matrix
 \begin{equation*}
@@
 \end{equation*}
-and the potential function is
+and potential function
 \begin{equation*}
@@
 implement this VNHC on a physical testbed where \(|I|\) must be large enough to
-overcome friction in the leg motor.
+overcome friction.
 
@@
 black stars, with the final crossing in red.
-After this crossing, the acrobot ceased rotating and its oscillations decayed to
-zero.
+After this final crossing, the acrobot ceased rotating and its oscillations
+decayed to zero.
 
@@
 In Figure \ref{fig:acrobot-rot-reg-diss} the acrobot was initialized at the
-faster rotation \((q_u,p_u) = (0, 0.23)\); the supervisor dissipated energy.
+faster rotation \((q_u,p_u) = (0, 0.23)\); here the supervisor dissipated energy.
 In both cases, the desired rotation was stabilized correctly.
@@
 directly evaluate \(p_u\) and \(p_a\).
-To resolve this issue, the MCU estimates \(\dot{q}\) using sequential measurements of \(q\)
-and a washout filter.
+To resolve this issue, the MCU estimates \(\dot{q}\) by applying a washout
+filter to sequential measurements of \(q\).
 We then compute \(p = M(q)\dot{q}\) for use in the VNHC controller.
@@
 500Hz.
-This low sampling rate results in a noisy momentum signal, which suffers
+This low sampling rate results in a noisy momentum signal which suffers
 from noticeable phase lag.
@@
 we can assign the servo setpoint at iteration \(k \in \mathbb{Z}_{> 0}\)
-via \(q_a^{k} = \arctan(10 p_u^{k-1})\).
+via \(q_a^{k} = \arctan(I p_u^{k-1})\).
 Unfortunately, this negatively affects the stabilization to the constraint
@@
 
-We performed the following tests on SUGAR.
+We performed the following tests on SUGAR with the energy injection VNHC.
 
@@
 
-We had hoped to show more experimental results for dissipation and
-oscillation/rotation regulation, but unfortunately the Dynamixel RX24F servo
-motor broke before we could perform these experiments.
+We had hoped to show experimental results for dissipation and
+oscillation/rotation regulation. 
+Unfortunately the Dynamixel RX24F servo motor broke before we could perform
+these experiments.
 These motors are no longer in production at the time of writing this article, so
-SUGAR met an untimely end to its career.
+SUGAR has been retired from service.
 
@@
 limbs are massless rods of equal length, with equal masses at the tips.
-Then we performed simulations of this VNHC, as well as physical experiments.
+We then performed simulations and physical experiments of this VNHC.
 These demonstrated that virtual nonholonomic constraints are capable of

### 2021-05-14 latex/paper/main.tex
Made the draft a one column, double spaced format for Manfredi.

@@
 
-\documentclass[journal,twoside,web]{ieeecolor}
+% TODO: remove onecolumn and draftcls formatting
+\documentclass[journal,twoside,onecolumn,draftclsnofoot,web]{ieeecolor}
 
@@
 In this article we redeveloped the framework of virtual nonholonomic
-constraints for underactuated mechanical systems.
+constraints for underactuated mechanical systems with arbitrary degree of
+underactuation.
 We applied this framework towards the acrobot, and designed a constraint which

### 2021-05-20 latex/paper/main.tex
Major edits to all sections, excepting Sims and Experiments, as per Manfredi's suggestions.

@@
 \newtheorem{thm}{Theorem}% Uncomment: reset numbering at each chapter
-\newtheorem{prop}[thm]{Proposition} % Propositions depend on Theorem number
-\newtheorem{lemma}[thm]{Lemma} % Same with lemmas
-\newtheorem{defn}[thm]{Definition} % Definitions depend on theorem number
+\newtheorem{prop}{Proposition} % Propositions don't depend on Theorem number
+\newtheorem{lemma}{Lemma} % Same with lemmas
+\newtheorem{defn}{Definition} % Definitions do not depend on theorem number
 \newtheorem{assm}{Assumption} % Assumptions do not reset
@@
 \newcommand*{\pudmat}{(\Id{n-k} \otimes p\tpose)\nabla_{q_u}\Minv(q)p}
+
+\newcommand*{\etal}{\MakeLowercase{\textit{et al.}}}
 %/========== /Preamble ==========/%
@@
     \centering
+    \begin{subfigure}[t]{0.40\linewidth}
         \includegraphics[width=\linewidth]{acrobot_gymnast.png}
-    \caption{The two-link acrobot as a model for a gymnast.
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.48\linewidth}
+        \includegraphics[width=\linewidth]{simple_acrobot_model.pdf}
+    \end{subfigure}
+    \caption{A simplified two-link acrobot as a model for a gymnast.
     Image modified from \cite{xingbo_thesis}.}
@@
 but was unable to prove the acrobot would gain energy over time.
-His approach was a rudimentary version of a recent technique know as the method
-of virtual nonholonomic constraints.
+His approach was a preliminary version of a recent technique know as the method
+of virtual nonholonomic constraints \cite{hybrid_zero_dynamics_bipedal_nhvcs}.
 
@@
 In particular, VNHCs may be capable of injecting and
-dissipating energy from a system in a robust manner, all while producing
-realistic biological motion.
+dissipating energy from a system in a robust manner, all while producing realistic
+biological motion.
 
@@
 injects energy into the acrobot through human-like giant motion.
-
-\subsection{Notation}
+In particular, any acrobot constrained by this VNHC is guaranteed to perform
+rotations around the bar.
+Under suitable conditions on the acrobot's mass and length parameters, this VNHC
+also enables the acrobot to rotate with a desired speed.
+We will conclude by providing simulations which demonstrate a mechanism for
+energy regulation through VNHCS, along with experimental results which
+demonstrate the robustness of this behaviour to model uncertainty, sensor noise,
+and a variety of external disturbances.
+
+%%
+\textbf{Notation}:
+%%
 We use the following notation and terminology in this article.
@@
 We will use the simplified acrobot model in Figure
-\ref{fig:simple-acrobot-model}, where we assume the torso and leg rods are of
+\ref{fig:acrobot}, where we assume the torso and leg rods are of
 equal length \(l\) with equal point masses \(m\) at the tips.
+The acrobot's configuration is described in generalized coordinates
+\((q_u,q_a)\) on the configuration manifold 
+\(\mathcal{Q} = \Sone \times \Sone\), where \(q_u\) is unactuated and 
+\(q_a\) is actuated.
 A real gymnast cannot swing their legs in full circles, though they
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=0.8\linewidth]{simple_acrobot_model.pdf}
-    \caption{A simple acrobot has massless rods of equal length \(l\) and 
-    equal masses \(m\) at the tips.}
-    \label{fig:simple-acrobot-model}
-\end{figure}
+%\begin{figure}
+%    \centering
+%    \includegraphics[width=0.8\linewidth]{simple_acrobot_model.pdf}
+%    \caption{A simple acrobot has massless rods of equal length \(l\) and 
+%    equal masses \(m\) at the tips.}
+%    \label{fig:simple-acrobot-model}
+%\end{figure}
 
@@
 The dynamics of the acrobot in \((q,p)\) coordinates are given in
-~\eqref{eqn:acrobot-hamiltonian}.
+\eqref{eqn:acrobot-hamiltonian}.
 For shorthand, we write \(c_u := \cos(q_u)\), \(c_a := \cos(q_a)\), and 
@@
 
-Our goal is to design a smooth function \(f : \Sone \times \R\) such that the
-relation \(q_a = f(q_u,p_u)\) is a regular VNHC, where the constrained dynamics
-of the acrobot (under this constraint) gain energy on some set \(D\)
-in the sense of Definition \ref{defn:energy-gain}.
-We will formally define regularity in the next section.
-Essentially, if the constraint meets some regularity conditions then it can be
-stabilized via feedback.
-
 \begin{defn}\label{defn:energy-gain}
     Let \(\mathcal{Q}\) be an
-    \(n\)-dimensional smooth generalized cylinder.
-    Let \(f : \mathcal{Q} \rightarrow \mathcal{Q}\times\R^n\) be a smooth vector
+    \(n\)-dimensional smooth manifold. 
+    Let \(f : \mathcal{Q} \rightarrow T\mathcal{Q}\) be a smooth vector
     field and let \(D \subset \mathcal{Q}\) be open.
@@
     condition \(x(0) \in K\), there exists \(T > 0\) such
-    that \(x(t) \notin K \, (\forall t > T)\).
+    that \(x(t) \notin K\) for all \(t > T)\).
     The system \textit{loses energy on \(D\)} if it gains energy in
@@
 
+Our goal is to design a smooth function \(f : \Sone \times \R\) such that the
+relation \(q_a = f(q_u,p_u)\) can be enforced asymptotically via feedback
+control (in Section \ref{sec:vnhc} we call this a regular virtual
+nonholonomic constraint), and the dynamics of the acrobot when the relation
+holds gain energy on some set \(D\) in the sense of Definition
+\ref{defn:energy-gain}.
+
 Note that any system satisfying Definition \ref{defn:energy-gain}
@@
 
-Before embarking on our design problem, let us summarize the relevant theory of
-VNHCs.
-
 %/========== VNHC ==========/%
 \section{Theory of VNHCs}\label{sec:vnhc}
+Before embarking on our design problem, we begin by summarizing the relevant
+theory of virtual nonholonomic constraints for a class of mechanical systems we
+call ``simply actuated hamiltonian systems". 
+Note that the results we provide in Section \ref{sec:vnhc-vnhc} are not novel:
+they are a special case of the results in
+\cite{hybrid_zero_dynamics_bipedal_nhvcs}.
 
@@
 \(q = (q_1, \ldots, q_n)\) on a configuration manifold
-\(\mathcal{Q} = \Rt{T_1} \times \cdot \Rt{T_n}\), where
+\(\mathcal{Q} = \Rt{T_1} \times \cdots \times \Rt{T_n}\), where
 \(T_i = 2\pi\) if \(q_i\) is an angle and \(T_i = \infty\) if \(q_i\) is a
@@
 The above assumptions allow us to define a
-canonical coordinate transformation of ~\eqref{eqn:hamiltonian} 
+canonical coordinate transformation of \eqref{eqn:hamiltonian} 
 which decouples the input forces.
@@
 
-In light of Lemma \ref{lemma:B-orthogonal}, there is no loss of generality in
-assuming that the input matrix is left semi-orthogonal.
+In light of Lemma \ref{lemma:B-orthogonal} there is no loss of generality in
+assuming that the input matrix is left semi-orthogonal, which means that
+\(B\tpose\) is right semi-orthogonal.
 Now, let \(\mathbf{B} := [B^\perp; B\tpose]\).
@@
 
+We now take \(\mathcal{Q}\) to be embedded in \(\R^n\), so the coordinate
+transformation used the following theorem is well-defined.
+
 \begin{thm}\label{thm:simply-actuated}
-    Take the Hamiltonian system ~\eqref{eqn:hamiltonian} and suppose
+    Take the Hamiltonian system \eqref{eqn:hamiltonian} and suppose
     Assumptions \ref{assm:B-const} and \ref{assm:B-perp} hold.
@@
 We call the \((\tilde{q},\tilde{p})\) coordinates
-\textit{simply actuated coordinates}.
+\textit{simply actuated coordinates}, and we call system
+\eqref{eqn:simple-hamiltonian} a \textit{simply actuated hamiltonian system}.
 The first \((n-k)\) configuration variables in \(\tilde{q}\), labelled \(q_u\),
@@
 through state feedback.
-Horn et. al later extended their results in
+Horn \etal later extended their results in
 \cite{hybrid_zero_dynamics_bipedal_nhvcs} to derive the constrained dynamics for
 a certain class of mechanical systems.
-
 These researchers made use of the unactuated conjugate of momentum, but they
@@
 one.
-In this section we reformulate their ideas into the Hamiltonian framework and
-extend the definition of a virtual nonholonomic constraint to systems with
-higher degree of underactuation.
-This reformulation will provide us with better intuition and a more explicit form of the
-constrained dynamics.
+We will now present a special case of \cite{hybrid_zero_dynamics_bipedal_nhvcs}
+for simply actuated hamiltonian systems, so that the theory we apply to the
+acrobot is provided in its clearest form.
+After we finish this summary we will clarify the relationship between our
+material and that of \cite{hybrid_zero_dynamics_bipedal_nhvcs}.
 
@@
 
-We often want to stabilize a constraint within some neighbourhood of \(\Gamma\).
+In order to enforce the constraint \(h(q,p) = 0\), we want to asymptotically
+stabilize the set \(\Gamma\).
 To see when this is possible, let us define the error output \(e = h(q,p)\).
@@
 \begin{thm}\label{thm:vnhc-regularity}
-    A relation \(h(q,p) = 0\) for system ~\eqref{eqn:simple-hamiltonian}
+    A relation \(h(q,p) = 0\) for system \eqref{eqn:simple-hamiltonian}
     is a regular VNHC of order \(k\) if and only if 
@@
 Under additional mild conditions (see \cite{vhcs_for_el_systems}), a regular VNHC of
-order \(k\) can be stabilized by the output-linearizing phase-feedback
+order \(k\) can be stabilized by the output-linearizing state-feedback
 controller
@@
 \cite{vhcs_for_el_systems}, and hence we can stabilize the constraint manifold
-using ~\eqref{eqn:stabilizing-controller}.
-Since \(q_a\) is constrained to be a function of the unactuated phase variables,
+using \eqref{eqn:stabilizing-controller}.
+Since \(q_a\) is constrained to be a function of the unactuated variables,
 intuition tells us the constrained dynamics should be parameterized by 
@@
 \((\Id{n} \otimes p\tpose)\nabla_{q}M(q)p\). 
-
 We now introduce an assumption which allows us to solve for \(p_a\) as a
@@
 \begin{assm}\label{assm:inertially-actuated}
-The inertia matrix does not depend on the unactuated coordinates, so that 
+The inertia matrix does not depend on the unactuated coordinates, i.e.,
 \(\nabla_{q_u}M(q) = \Zmat{n(n-k) \times n}\).
@@
 \begin{thm}\label{thm:zero-dynamics}
-    Let \(\mathcal{H}\) be a mechanical system in simply actuated
-    coordinates satisfying Assumption \ref{assm:inertially-actuated}. 
-    Let \(h(q,p_u) = 0\) be a regular VNHC of order \(k\) with constraint
-    manifold \(\Gamma\). Suppose that on \(\Gamma\) one can solve for
-    \(q_a\) as a function \(q_a = f(q_u,p_u)\).
+    Let \(\mathcal{H}\) be a hamiltonian system in simply actuated form
+    \eqref{eqn:simple-hamiltonian} satisfying
+    Assumption \ref{assm:inertially-actuated}. 
+    Let \(h(q,p_u) = q_a - f(q_u,p_u)\) be a regular VNHC of order \(k\) with
+    constraint manifold \(\Gamma\). 
     Then the constrained dynamics are given by
@@
     \begin{multline}\label{eqn:g-qpu}
-        g(q,p_u) := \left( \mathcal{A}(q,p_u)
+        g(q_u,p_u) := \left( \mathcal{A}(q,p_u)
         \begin{bmatrix} \Zmat{(n-k)\times k} \\ \Id{k} \end{bmatrix}\right)\inv 
@@
 \begin{defn}\label{defn:energy-injection}
-    A regular VNHC \(h(q, p) = 0\) with constraint manifold \(\Gamma\)
+    A regular VNHC \(h(q,p_u) = q_a - f(q_u,p_u)\) with constraint manifold \(\Gamma\)
     \textit{injects (dissipates) energy on \(D \subset \Gamma\)} if the
@@
 
-\textbf{Comparison with existing literature}: Horn et.al. provide the constrained
+\textbf{Comparison with existing literature}: Horn \etal provide the constrained
 dynamics for VNHCs in \cite{nhvc_incline_walking}.
@@
 assumption \textbf{H3} holds true.
-The only real distinction between this section and their work is that we allow
-for higher degrees of underactuation, and our constrained dynamics are explicit
-functions of the Hamiltonian phase coordinates \((q_u,p_u)\).
-In fact, our constrained dynamics \eqref{eqn:qpu-dynamics} coincide with their
-system (17) when one chooses the special case \(\theta_1 = q_u\) and 
+The only real distinction between this section and their work is that
+our constrained dynamics are explicit functions of the Hamiltonian coordinates
+\((q_u,p_u)\).
+In fact, one can show that the constrained dynamics \eqref{eqn:qpu-dynamics}
+coincide with the constrained dynamics in
+\cite[Eqn. (17)]{hybrid_zero_dynamics_bipedal_nhvcs}
+when one chooses the special case \(\theta_1 = q_u\) and 
 \(\theta_2 = p_u\).
@@
 Our goal in this article is to design a VNHC which injects energy into the
-acrobot through giant-like motion.
-Recall that the acrobot in Figure \ref{fig:simple-acrobot-model} has dynamics given by
-~\eqref{eqn:acrobot-hamiltonian}, which are
-repeated here:
-\begin{align*}
-    \mathcal{H}(q,p) &= \frac{1}{2}p\tpose \Minv(q) p -
-    mgl\left(2 c_u + c_{ua}\right)
-    , \\
-     &\begin{cases}
+acrobot by means of a giant-like motion.
+Recall that the acrobot in Figure \ref{fig:acrobot} has dynamics given by
+\eqref{eqn:acrobot-hamiltonian}, repeated here for convenience{
+    \begin{equation*}
+     \begin{cases}
         \dot{q} = \Minv(q) p 
@@
     \end{cases}
-\end{align*}
+\end{equation*}
 
 Since the control input \(\tau\) only affects the actuated momentum,
-\((q,p)\) are simply actuated coordinates in the phase space
-\(\mathcal{Q} \times \mathcal{P}\) where
+the system above is already in simply actuated form.
+Its state space is \(\mathcal{Q} \times \mathcal{P}\) where
 \(\mathcal{Q} = \Sone \times \Sone\), and
@@
 We can therefore apply the theory from Section \ref{sec:vnhc} to design
-a VNHC of the form \(q_a = f(q_u,p_u)\). 
+a VNHC of the form \(q_a = f(q_u,p_u)\) (i.e., a VNHC 
+\(h(q,p_u) = q_a - f(q_u,p_u) = 0\)).
 Since we need the VNHC to be regular, the following proposition will be
@@
 \begin{prop}\label{prop:acrobot-fpu-regular}
-    A relation \(h(q,p) = q_a - f(p_u) = 0\) 
+    Any relation \(q_a = f(p_u)\) 
     with \(f \in C^2\left(\R; \Sone\right)\) is a regular
@@
 \begin{proof}
-    The regularity matrix for the acrobot evaluates to
+    The decoupling matrix for the acrobot evaluates to
     \(((1+c_a)\partial_{q_u}f(q_u,p_u) + (3+2c_a))/(ml^2(2-c_a^2))\).
     Since \(\partial_{q_u} f = 0\), this matrix is strictly positive for all values
-    of \(q_a\), and hence is full rank everywhere on the constraint manifold.
+    of \(q_a\), and hence is full rank \(1\) everywhere on the constraint manifold.
 \end{proof}
@@
 displayed in Figure \ref{fig:qa-arctan}.
-Here, \(\bar{q}_a \in \, ]0,2 Q_a/\pi]\) is constant and \(I \in \R\) is a fixed
-control parameter.
-
+Here, \(\bar{q}_a \in \, ]0,2 Q_a/\pi]\) and \(I \in \R\) are 
+design parameters.
 This constraint does not perfectly recreate giant motion, during which
@@
 Recall that \((q_u, p_u)\) denote the angle and momentum of the acrobot's torso.
-By Theorem \ref{thm:zero-dynamics}, the constrained acrobot is parameterized
-fully by \((q_u,p_u) \in \SxR\).
-It is easy to show that the constrained dynamics are just the torso dynamics
-obtained when one swings the legs according to \eqref{eqn:acrobot-constraint}.
-That is, the constrained dynamics are
+By Theorem \ref{thm:zero-dynamics}, the constrained dynamics arising from the
+VNHC \eqref{eqn:acrobot-constraint} are parameterized fully by 
+\((q_u,p_u) \in \SxR\).
+Here, \eqref{eqn:g-qpu} reduces to
+\begin{equation*}
+    g(q_u,p_u) = \frac{
+        (1+c_a)(1+I^2 p_u^2)p_u - m^2gl^3\bar{q}_a I (2-c_a^2)(2s_u + s_{ua})
+    }{ml^2(3+2c_a)(1+I^2 p_u^2)}
+    ,
+\end{equation*}
+and the constrained dynamics \eqref{eqn:qpu-dynamics} are
 \begin{equation}\label{eqn:acrobot-constrained-dynamics}
@@
 
-Suppose for a moment that \(I = 0\), i.e.~that the legs stay fully extended.
+Suppose for a moment that \(I = 0\) in \eqref{eqn:acrobot-constraint}, i.e.,
+that the legs stay fully extended.
 The acrobot becomes a nominal pendulum with two masses, whose total mechanical
@@
 
-When the pendulum is oscillating, its phase \((q_u,p_u)\) lies in the set
+When the pendulum is oscillating, its state \((q_u,p_u)\) lies in the set
 \begin{equation}\label{eqn:oscillation-domain}
@@
 \(\abs{p_u} < \bar{\rho}\) (for some chosen \(\bar{\rho}\)),
-the phase must lie inside the rotation domain
+\((q_u,p_u)\) must lie inside the rotation domain
 \begin{multline}\label{eqn:rotation-domain}
@@
 \begin{thm}\label{thm:acrobot-energy-stabilization}
-    Consider the acrobot with Hamiltonian dynamics as in ~\eqref{eqn:acrobot-hamiltonian}.
+    Consider the acrobot with Hamiltonian dynamics given in
+    \eqref{eqn:acrobot-hamiltonian}.
 \begin{enumerate}
-    \item For any \(m\), \(g\), \(l\), \(\bar{q}_a\), there exists
+    \item For positive \(m\), \(g\), \(l\), and \(\bar{q}_a\), there exists
         \(I^\star > 0\) such that, for all \(I \in \, ]0,I^\star]\), the VNHC
-        ~\eqref{eqn:acrobot-constraint} injects energy into the acrobot on
+        \eqref{eqn:acrobot-constraint} injects energy into the acrobot on
         \(\mathcal{O}_1\).
-        Moreover, almost every orbit will escape the closure of
+        Moreover, almost every orbit of the constrained dynamics 
+        \eqref{eqn:acrobot-constrained-dynamics} will escape the closure of
         \(\mathcal{O}_1\) in finite time.
@@
         \(I \in \, ]0,I^{\star\star}]\), the VNHC
-        ~\eqref{eqn:acrobot-constraint} injects energy into the acrobot on
+        \eqref{eqn:acrobot-constraint} injects energy into the acrobot on
         \(\mathcal{O}_2(\bar{\rho})\).
@@
 Theorem \ref{thm:acrobot-energy-stabilization} considers these sets separately.
-This separation is advantageous because the first result holds for any
+This separation is advantageous because the first result holds for any positive
 \(m\), \(g\), \(l\), and \(\bar{q}_a\). 
@@
 \ref{thm:acrobot-energy-stabilization} states that all acrobots constrained by
-~\eqref{eqn:acrobot-constraint} will gain enough energy to begin rotating around
+\eqref{eqn:acrobot-constraint} will gain enough energy to begin rotating around
 the bar.
@@
 
+\subsection{Energy Regulation}
 One can apply the results of Theorem \ref{thm:acrobot-energy-stabilization} 
 towards energy regulation.
-For example, one might employ a supervisor which switches
-between injection and dissipation VNHCs to stabilize a desired energy
-level \(E_\text{des}\), with some hysteresis to avoid infinite switching.
-We will explore this idea in Section \ref{sec:simulations}.
+
+\textbf{TODO: move energy regulation theory from sims to here}
+%For example, one might employ a supervisor which switches
+%between injection and dissipation VNHCs to stabilize a desired energy
+%level \(E_\text{des}\), with some hysteresis to avoid infinite switching.
+%We will explore this idea in Section \ref{sec:simulations}.
 
@@
 This physical acrobot cannot be modelled by the simplified setup in Figure
-\ref{fig:simple-acrobot-model}, because the torso and leg links have 
+\ref{fig:acrobot}, because the torso and leg links have 
 distributed mass. 
@@
 We will now simulate the effects of constraining Wang's acrobot with
-~\eqref{eqn:acrobot-constraint}, thereby demonstrating that VNHCs are robust to
+\eqref{eqn:acrobot-constraint}, thereby demonstrating that VNHCs are robust to
 model mismatch.
@@
 and experiments.
-In other words, our injection VNHC is ~\eqref{eqn:acrobot-constraint} with 
-\(I = 10\) while our dissipation VNHC is ~\eqref{eqn:acrobot-constraint} with 
+In other words, our injection VNHC is \eqref{eqn:acrobot-constraint} with 
+\(I = 10\) while our dissipation VNHC is \eqref{eqn:acrobot-constraint} with 
 \(I = -10\).

### 2021-05-24 latex/paper/main.tex
Made edits up to (but not including) page 9 of paper_v2

@@
 \(p_a\), representing a torque acting on the hip joint.
+Let us now define what it means for a mechanical system to gain or lose energy.
 
@@
 control (in Section \ref{sec:vnhc} we call this a regular virtual
-nonholonomic constraint), and the dynamics of the acrobot when the relation
-holds gain energy on some set \(D\) in the sense of Definition
+nonholonomic constraint).  
+We will further require that the dynamics of the acrobot, when the relation
+holds, gain or lose energy on some set \(D\) in the sense of Definition
 \ref{defn:energy-gain}.
@@
 
-\subsection{Energy Regulation}
+\subsection{Energy Regulation}\label{sec:energy-reg}
 One can apply the results of Theorem \ref{thm:acrobot-energy-stabilization} 
-towards energy regulation.
-
-\textbf{TODO: move energy regulation theory from sims to here}
-%For example, one might employ a supervisor which switches
-%between injection and dissipation VNHCs to stabilize a desired energy
-%level \(E_\text{des}\), with some hysteresis to avoid infinite switching.
-%We will explore this idea in Section \ref{sec:simulations}.
+towards energy regulation;
+that is, one can stabilize oscillations or rotations by appropriately toggling
+between injection and dissipation VNHCs.
+
+\textbf{Oscillation Regulation}: one chooses a desired oscillation angle 
+\(q_\text{des} \in \, ]0,\pi[\) and, to avoid infinite switching, a
+hysteresis value \(\delta \in [0,1]\).
+Each time the orbit crosses the \(q_u\)-axis (i.e. when \(p_u = 0\)) or 
+when \(|q_u| = \pi\), the supervisor does the following: 
+\begin{itemize}
+    \item If \(|q_u| < (1-\delta)q_\text{des}\), enable the injection VNHC.
+    \item If \(|q_u| > (1+\delta)q_\text{des}\), enable the dissipation VNHC.
+    \item If \((1-\delta)q_\text{des} \leq |q_u| \leq (1+\delta)q_\text{des}\),
+        keep the leg extended at \(q_a = 0\). This can be done continously since
+        \(q_a = 0\) when \(p_u = 0\).
+\end{itemize}
+
+\textbf{Rotation Regulation}: one chooses a desired rotation rate 
+\(p_\text{des} > 0\) and a hysteresis value \(\delta \in [0,1]\).
+Each time the orbit crosses the \(p_u\)-axis (i.e. when \(q_u = 0\)), the
+supervisor changes which VNHC is enforced as follows:
+\begin{itemize}
+    \item If \(|p_u| < (1-\delta)p_\text{des}\), enable the injection VNHC.
+    \item If \(|p_u| > (1+\delta)p_\text{des}\), enable the dissipation VNHC.
+    \item If \((1-\delta)p_\text{des} \leq |p_u| \leq (1+\delta)p_\text{des}\),
+        extend the leg fully by setting \(q_a = 0\).
+        For these simulations, we assume this can be done instantaneously.
+\end{itemize}
 
@@
 
-The dynamic parameters of Wang's acrobot can be found in Table
+Recall that we defined the sets \(\mathcal{O}_1\) and
+\(\mathcal{O}_2(\bar{\rho})\) used in Theorem \ref{thm:acrobot-energy-stabilization}
+via the mechanical energy \eqref{eqn:acrobot-energy} of the nominal
+pendulum associated with a simplified acrobot.
+We obtained this nominal pendulum by setting \(I = 0\) in
+\eqref{eqn:acrobot-constraint}.
+Recall that the boundary between oscillations and rotations of the nominal
+pendulum is obtained by finding the level set \(E_\pi\) of the nominal energy
+corresponding to the upright equilibrium \((q_u,p_u) = (\pi,0)\).
+
+We can extend this idea to find the nominal pendulum associated with Wang's
+acrobot, whose physical parameters are provided in Table
 \ref{tab:acrobot-parameters}.
-Evaluating the mechanical energy of an acrobot with these parameters
-at the VNHC \(q_a = 0\) yields the mechanical energy of the nominal pendulum,
+Setting \(I = 0\) in \eqref{eqn:acrobot-constraint} for this acrobot
+yields the following mechanical energy of the nominal pendulum:
 \[
@@
 \]
-
-We will now simulate the effects of constraining Wang's acrobot with
+Despite the change in acrobot model, the level set \(E_\pi\) with energy 
+\(E(\pi,0)\) remains the boundary between oscillations and rotations
+for the nominal pendulum.
+
+We will now simulate the effects of constraining Wang's acrobot with the VNHC
 \eqref{eqn:acrobot-constraint}, thereby demonstrating that VNHCs are robust to
@@
 The level set \(E_\pi\) with energy \(E(\pi,0)\) is outlined in black.
-Despite the change in acrobot model, this level set remains the boundary
-between oscillations and rotations for the nominal pendulum.
+Recall that this marks the boundary between oscillations and rotations of the
+nominal pendulum.
 The points where the orbit exits \(E_\pi\) are marked with black stars,
@@
 
-Once can use a supervisor to stabilize oscillations by appropriately
-toggling between injection and dissipation VNHCs.
-Specifically, one chooses a desired oscillation angle 
-\(q_\text{des} \in \, ]0,\pi[\) and, to avoid infinite switching, a
-hysteresis value \(\delta \in [0,1]\).
-Each time the orbit crosses the \(q_u\)-axis (i.e. when \(p_u = 0\)) or 
-when \(|q_u| = \pi\), the supervisor does the following: 
-\begin{itemize}
-    \item If \(|q_u| < (1-\delta)q_\text{des}\), enable the injection VNHC.
-    \item If \(|q_u| > (1+\delta)q_\text{des}\), enable the dissipation VNHC.
-    \item If \((1-\delta)q_\text{des} \leq |q_u| \leq (1+\delta)q_\text{des}\),
-        keep the leg extended at \(q_a = 0\). This can be done continously since
-        \(q_a = 0\) when \(p_u = 0\).
-\end{itemize}
-
-Figure \ref{fig:acrobot-osc-reg} shows the supervisor stabilizing an
+Recall from \ref{sec:energy-reg} that one can use a supervisor to stabilize
+oscillations by appropriately toggling between injection and dissipation VNHCs.
+Figure \ref{fig:acrobot-osc-reg} shows  stabilizing an
 oscillation with body angle \(q_\text{des} = \pi/2\) and a 
@@
 \subsection{Rotation Regulation}
-One can also use a supervisor to stabilize rotations.
-In this case, one chooses a desired rotation rate \(p_\text{des} > 0\) and a
-hysteresis value \(\delta \in [0,1]\).
-Each time the orbit crosses the \(p_u\)-axis (i.e. when \(q_u = 0\)), the
-supervisor changes which VNHC is enforced as follows:
-\begin{itemize}
-    \item If \(|p_u| < (1-\delta)p_\text{des}\), enable the injection VNHC.
-    \item If \(|p_u| > (1+\delta)p_\text{des}\), enable the dissipation VNHC.
-    \item If \((1-\delta)p_\text{des} \leq |p_u| \leq (1+\delta)p_\text{des}\),
-        extend the leg fully by setting \(q_a = 0\).
-        For these simulations, we assume this can be done instantaneously.
-\end{itemize}
-
-Rotation regulation is demonstrated in Figure \ref{fig:acrobot-rot-reg}, where
-the supervisor stabilizes \(p_\text{des} = 0.19\) with a \(2\%\) hysteresis 
-\(\delta = 0.02\).
+One can also use a supervisor to stabilize rotations through the mechanism
+described in Section \ref{sec:energy-reg}.
+Rotation regulation for the acrobot is demonstrated in 
+Figure \ref{fig:acrobot-rot-reg}, where the supervisor stabilizes 
+\(p_\text{des} = 0.19\) with a \(2\%\) hysteresis \(\delta = 0.02\).
 The supervisor changed its choice of VNHC at each black star until it was within

### 2021-05-27 latex/paper/main.tex
Finished editing the words that Manfredi wanted me to edit

@@
 \]
-Despite the change in acrobot model, the level set \(E_\pi\) with energy 
+Despite the change in model, the level set \(E_\pi\) with energy 
 \(E(\pi,0)\) remains the boundary between oscillations and rotations
-for the nominal pendulum.
+for this nominal pendulum.
 
@@
 and experiments.
-In other words, our injection VNHC is \eqref{eqn:acrobot-constraint} with 
-\(I = 10\) while our dissipation VNHC is \eqref{eqn:acrobot-constraint} with 
+In other words, our \textit{injection VNHC} is \eqref{eqn:acrobot-constraint} with 
+\(I = 10\) while our \textit{dissipation VNHC} is \eqref{eqn:acrobot-constraint} with 
 \(I = -10\).
@@
         $m_u$ & $m_a$ & $l_u$ & $l_a$ & $l_{c_u}$ & $l_{c_a}$ & $J_u$ & $J_a$ & $g$ \\
+        (kg) & (kg) & (m) & (m) & (m) & (m) & (kg\(\cdot\)m\(^2\)) &
+        (kg\(\cdot\)m\(^2\)) & (m/s\(^2\)) \\
         \midrule
@@
 
-We simulated the acrobot constrained by our injection VNHC
-with initial condition \((q_u,p_u) = \left(\pi/32,0 \right)\) for \(30\)
-seconds, and plotted the resulting orbit in Figure \ref{fig:acrobot-in-orbit}.
+In simulation, we stabilized the injection VNHC for the acrobot using the
+controller \eqref{eqn:stabilizing-controller}.
+We initialized the acrobot on the constraint manifold
+with initial condition \((q_u,p_u) = \left(\pi/32,0 \right)\) and simulated the
+constrained system for \(30\) seconds.
+The resulting orbit (on the constraint manifold) is plotted in Figure
+\ref{fig:acrobot-in-orbit}.
 
 The level set \(E_\pi\) with energy \(E(\pi,0)\) is outlined in black.
-Recall that this marks the boundary between oscillations and rotations of the
+Recall that this level set is the boundary between oscillations and rotations of the
 nominal pendulum.
@@
 
-We initialized our acrobot with a rotation at \((q_u,p_u) = (0,0.18)\) and
-simulated the constrained dynamics (under the dissipation VNHC) for \(30\)
-seconds.
-The resulting orbit is plotted in Figure \ref{fig:acrobot-diss-orbit}. 
+In simulation, we stabilized the dissipation VNHC and initialized the acrobot on
+the constraint manifold with a rotation \((q_u,p_u) = (0,0.18)\).
+We simulated the constrained system for \(30\) seconds and plotted the
+resulting orbit in Figure \ref{fig:acrobot-diss-orbit}. 
 As expected, the acrobot slows down over time.
@@
 Recall from \ref{sec:energy-reg} that one can use a supervisor to stabilize
-oscillations by appropriately toggling between injection and dissipation VNHCs.
-Figure \ref{fig:acrobot-osc-reg} shows  stabilizing an
+oscillations by appropriately toggling between injection and dissipation VNHCs
+whenever the orbit of the acrobot crosses the \(q_u\)-axis.
+Figure \ref{fig:acrobot-osc-reg} shows the supervisor stabilizing an
 oscillation with body angle \(q_\text{des} = \pi/2\) and a 
@@
     \caption{Using a supervisor to stabilize the oscillation with peak angle
-    \(q_\text{des} = \pi/2\).}
+    \(q_\text{des} = \pi/2\). The desired oscillation is depicted with a solid
+    black line.}
     \label{fig:acrobot-osc-reg}
@@
 One can also use a supervisor to stabilize rotations through the mechanism
-described in Section \ref{sec:energy-reg}.
+described in Section \ref{sec:energy-reg}, where the supervisor toggles between
+injection and dissipation VNHCs at each crossing of the \(p_u\)-axis.
 Rotation regulation for the acrobot is demonstrated in 
@@
 \(p_\text{des} = 0.19\) with a \(2\%\) hysteresis \(\delta = 0.02\).
-The supervisor changed its choice of VNHC at each black star until it was within
-range of \(p_\text{des}\), at which point it extended the leg (shown in red).
+The supervisor evaluated its choice of VNHC at each black star.
+Once it was within range of \(p_\text{des}\) it extended the legs completely,
+the orbit of which is shown in red.
 
 In Figure \ref{fig:acrobot-rot-reg-in} the acrobot was initialized at 
-\((q_u,p_u) = (\pi/2,0)\); here the supervisor injected energy until
-the orbit hit the \(p_u\)-axis near \(p_\text{des}\).
+the oscillation \((q_u,p_u) = (\pi/2,0)\); here the supervisor injected energy
+until the orbit hit the \(p_u\)-axis near \(p_\text{des}\).
 In Figure \ref{fig:acrobot-rot-reg-diss} the acrobot was initialized at the
-faster rotation \((q_u,p_u) = (0, 0.23)\); here the supervisor dissipated energy.
+(fast) rotation \((q_u,p_u) = (0, 0.23)\); here the supervisor dissipated energy.
 In both cases, the desired rotation was stabilized correctly.
@@
     \caption{Using a supervisor to stabilize the rotation with maximal momentum
-    \(p_\text{des} = 0.23\).}
+    \(p_\text{des} = 0.23\). The desired rotation is depicted with a solid black
+    line.}
     \label{fig:acrobot-rot-reg}
@@
     \item \textbf{Baseline Test:} 
-    we initialized the acrobot at 
+    we initialized SUGAR at 
     \((q_u,p_u) \approx \left(\pi/8,0\right)\). 
@@
     rotation with energy \(E(0,0.195)\).
+    The energy gain likely ceases because of friction at the pivot, which was
+    not modelled in simulation.
 
@@
     \includegraphics[width=0.8\linewidth]{acrobot_unperturbed_orbit.png}
-    \caption{SUGAR's orbit during the baseline energy injection test.}
+    \caption{Baseline Test: SUGAR's baseline energy injection orbit.}
     \label{fig:acrobot-unperturbed-orbit}
@@
 
-\item \textbf{Stop Test:}
+\item \textbf{Perturbation Test 1:}
     we initialized SUGAR at 
     \((q_u,p_u) \approx \left(\pi,0\right)\), let it run for 15
-    seconds, then stopped it at the bottom of its arc.
+    seconds, then introduced a rod as it passed through the bottom of
+    its arc.
+    This caused a collision which stopped SUGAR in its tracks, at which
+    point we immediately removed the rod so SUGAR could continue
+    unperturbed.
     The resulting orbit is shown in Figure \ref{fig:acrobot-stopped-orbit}.
     The blue rotation curve corresponds to the orbit before the disturbance,
-    while the red spiral confirms that the acrobot begins oscillating after it
-    is stopped.  
-    Despite the disturbance, it gains energy and eventually starts rotating
-    again.
+    while the red spiral confirms that SUGAR begins oscillating after it
+    was stopped.  
+    After the collision, SUGAR gains energy and eventually starts
+    rotating again.
 
@@
     \includegraphics[width=0.8\linewidth]{acrobot_stopped_orbit.png}
-    \caption{SUGAR's orbit before (blue) and after (red) stopping.}
+    \caption{Perturbation Test 1: SUGAR's orbit before (blue) and after (red) stopping.}
     \label{fig:acrobot-stopped-orbit}
@@
 
-\item \textbf{Push Test:}
+\item \textbf{Perturbation Test 2:}
     to see how SUGAR responds when pushed, we allowed
-    it to rotate undisturbed for 15 seconds and then pushed it in its
+    it to rotate unperturbed for 15 seconds and then pushed it in its
     direction of motion.
-    The orbit in Figure \ref{fig:acrobot-fpush-orbit} shows that the acrobot,
+    The orbit in Figure \ref{fig:acrobot-fpush-orbit} shows that SUGAR,
     when pushed, rotates with energy \(E(0,-0.22)\), but then slows down until
@@
     We repeated this test by pushing SUGAR against its direction of motion.
-    The orbit in Figure \ref{fig:acrobot-rpush-orbit} demonstrates that the acrobot
+    The orbit in Figure \ref{fig:acrobot-rpush-orbit} demonstrates that it
     readily changes direction, and quickly achieves its maximum speed
@@
 This difference between simulation and implementation is likely due to
-friction.
+friction, as well as timing errors incurred by the PID controller in the servo
+motor.
 
@@
     \end{subfigure}
-    \caption{SUGAR's orbit before (blue) and after (red) pushing.}
+    \caption{Perturbation Test 2: SUGAR's orbit before (blue) and after (red) pushing.}
 \end{figure}
@@
 
-In this article we redeveloped the framework of virtual nonholonomic
-constraints for underactuated mechanical systems with arbitrary degree of
-underactuation.
-We applied this framework towards the acrobot, and designed a constraint which
-emulates giant motion from gymnastics.
+In this article we apply the framework of virtual nonholonomic constraints
+towards the acrobot, and designed a constraint which emulates giant motion from
+gymnastics.
 We proved this constraint will inject energy into the simplified acrobot whose
 limbs are massless rods of equal length, with equal masses at the tips.
-We then performed simulations and physical experiments of this VNHC.
+We then performed simulations and physical experiments on a real acrobot robot.
 These demonstrated that virtual nonholonomic constraints are capable of
 injecting and dissipating energy in a robust manner, all while producing
-realistic biological motion.
+realistic biologically-inspired motion.
 

### 2021-05-27 matlab/thesis_sims/regulationEvent.m
Updated the acrobot osc/rot stabilization so the black lines are thicker.

@@ -11,16 +11,16 @@
 % * Eval = energy value at which to trigger the event. This will typically
 % be E(pi,0), but it can be any energy value.
 
-function [value,isterminal,direction] = energyEvent(t,qp,E,Eval)
+function [value,isterminal,direction] = regulationEvent(t,qp,E,Eval,d)
     % Trigger when E(q,p) = Eval. The value increases as E -> Eval from
     % below.
     value = E(qp(1),qp(2))-Eval;
-    % Don't stop the integration, just record the event
-    isterminal = 0;
+    % Stop the integration
+    isterminal = 1;
     % Track only values where the energy is increasing past Eval, not those
     % where you come back down. If you want it to track only decreasing,
     % set direction to -1. If you want both increasing and decreasing, set
     % it to 0.
-    direction = 1;
+    direction = d;
 end
 

### 2021-05-28 latex/paper/main.tex
Began writing the proof of Theorem 10

@@
 \textbf{TODO: prove Thm 10}
-
+The sets \(\mathcal{O}_1\) and \(\mathcal{O}_2(\bar{\rho}\) are sublevel sets of
+the nominal energy \(E(q_u,p_u)\) defined by \eqref{eqn:acrobot-energy}.
+To prove Theorem \ref{thm:acrobot-energy-stabilization}, we will find conditions
+under which the acrobot gains energy on an open ``energy annulus".
+That is, given two energy levels \(E_0 \geq 0\) and \(E_1 > E_0\), we define the
+energy annulus to be the domain
+\[
+    \mathcal{D} := \{(q_u,p_u) \in \SxR \mid \\
+    E_0 < E(q_u,p_u) < E_1\}
+    .
+\]
+The only condition we place on \(E_0\) and \(E_1\) is that they both correspond
+to oscillations or both to rotations; that is, if \(E_0 < E(\pi,0)\), then 
+\(E_1 \leq E(\pi,0)\).
+
+Here is the fundamental idea of the proof: for certain open subsets
+\(D \subset \mathcal{D}\), one can find a change of coordinates into 
+pseudo-polar coordinates adapted to level sets of \(E\).
+In other words, when \(I = 0\) in \eqref{eqn:acrobot-constraint}, one can find
+a diffeomorphism \(T(q_u,p_u) = (r,\theta)\), where \(r\)
+chooses a level set of \(E\) and \(\theta\) chooses a point on that level set.
+
+We will describe the explicit forms of these transformations later. 
+Thanks to the shape of a pendulum's orbits, we will show that (at \(I = 0\)) the
+dynamics of \((r,\theta)\) satisfy \(\dot{r} = 0\) and \(\dot{\theta} > 0\).
+By continuity of the constrained dynamics with respect to \(I\), we can find a
+small enough \(I > 0\) so that \(\dot{\theta} > 0\) for the acrobot.
+This property will enable us to perform a time-scaling transformation 
+\(t = t(\theta)\) to eliminate the dynamics of \(\theta\), and study how 
+the pseudo-radius \(r\) evolves as a function of the pseudo-angle \(\theta\),
+rather than as a function of time.
+If it increases ``on average" , then \(E\) must also be increasing ``on average"
+and the constrained dynamics must be gaining energy on \(D\).
+
+We formalize this idea in the following lemma.
+Then we will find the relevant subset \(D \subset \mathcal{D}\)
+and the coordinate transformations \((q_u,p_u) \mapsto (r,\theta)\) which
+satisfy the conditions of the lemma, thereby proving the acrobot gains energy on
+each set.
+
+\begin{lemma}
+    Consider the simplified acrobot \eqref{eqn:acrobot-hamiltonian} constrained
+    by the VNHC \eqref{eqn:acrobot-constraint}, whose constrained dynamics are
+    parameterized by \((q_u,p_u) \in \SxR\).
+    Let \(D \subset \SxR\) be open and 
+    let \(\underline{r}, \overline{r} \in \R_{> 0}\) with 
+    \(\underline{r} < \overline{r}\).
+    Suppose there exists a diffeomorphism 
+    \begin{align*}
+        T : D &\to ]\underline{r},\overline{r}[ \, \times \Sone \\
+        (q_u,p_u) &\mapsto (r,\theta)
+        ,
+    \end{align*}
+    satisfying \(\dot{r} = 0\) and \(\dot{\theta} > 0\) at \(I = 0\).
+    Let \(g(\theta,r,I) = \dot{r}/\dot{\theta}\) and assume that 
+    \((\partial g/\partial I)(\theta,r,0) = 0\).
+    Define \(b(\theta,r_0) = g(\theta,r_0,0)\) and assume there exists
+    \(\epsilon > 0\) such that 
+    \[
+        \int \limits_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon
+        ,
+    \]
+    for all \(r_0 \in \, ]\underline{r},\overline{r}[\).
+    Then there exists \(I^\star > 0\) small enough that the constrained dynamics
+    gain energy on \(D\) for all \(0 < I \leq I^\star\), and lose energy on \(D\)
+    for all \(-I^\star \leq I < 0\).
+\end{lemma}
 %/========== Conclusion ==========/%

### 2021-05-28 latex/paper/main.tex
Partway through the proof of Thm 10, still working out the best way to do this.

@@
 \section{Proof of Theorem \ref{thm:acrobot-energy-stabilization}}\label{sec:proof}
-\textbf{TODO: prove Thm 10}
-The sets \(\mathcal{O}_1\) and \(\mathcal{O}_2(\bar{\rho}\) are sublevel sets of
-the nominal energy \(E(q_u,p_u)\) defined by \eqref{eqn:acrobot-energy}.
+The sets \(\mathcal{O}_1\) and \(\mathcal{O}_2(\bar{\rho})\) are sublevel sets of
+the nominal pendulum's energy \(E(q_u,p_u)\) described in
+\eqref{eqn:acrobot-energy}.
 To prove Theorem \ref{thm:acrobot-energy-stabilization}, we will find conditions
@@
 energy annulus to be the domain
-\[
+\begin{equation*}
     \mathcal{D} := \{(q_u,p_u) \in \SxR \mid \\
@@
     .
-\]
+\end{equation*}
 The only condition we place on \(E_0\) and \(E_1\) is that they both correspond
@@
 We will describe the explicit forms of these transformations later. 
-Thanks to the shape of a pendulum's orbits, we will show that (at \(I = 0\)) the
-dynamics of \((r,\theta)\) satisfy \(\dot{r} = 0\) and \(\dot{\theta} > 0\).
+Importantly, we will show that the dynamics of \((r,\theta)\) satisfy 
+\(\dot{r} = 0\) and \(\dot{\theta} > 0\) for the nominal pendulum (i.e., when
+\(I = 0\)).
 By continuity of the constrained dynamics with respect to \(I\), we can find a
@@
 This property will enable us to perform a time-scaling transformation 
-\(t = t(\theta)\) to eliminate the dynamics of \(\theta\), and study how 
-the pseudo-radius \(r\) evolves as a function of the pseudo-angle \(\theta\),
-rather than as a function of time.
-If it increases ``on average" , then \(E\) must also be increasing ``on average"
-and the constrained dynamics must be gaining energy on \(D\).
+\(t = t(\theta)\), and study how the pseudo-radius \(r\) evolves as a function
+of the pseudo-angle \(\theta\), rather than as a function of time.
+If \(r(\theta)\) increases on average, then \(E\) must also be increasing
+on average and the constrained dynamics must be gaining energy on \(D\).
 
@@
     \begin{align*}
-        T : D &\to ]\underline{r},\overline{r}[ \, \times \Sone \\
+        T : D &\to T(D) = ]\underline{r},\overline{r}[ \, \times \Sone \\
         (q_u,p_u) &\mapsto (r,\theta)
@@
     \end{align*}
-    satisfying \(\dot{r} = 0\) and \(\dot{\theta} > 0\) at \(I = 0\).
+    satisfying \(\dot{r} = 0\) and \(\dot{\theta} > 0\) on \(T(D)\) when 
+    \(I = 0\).
     Let \(g(\theta,r,I) = \dot{r}/\dot{\theta}\) and assume that 
@@
 \end{lemma}
+\begin{proof}
+    Since \(\dot{\theta} > 0\) when \(I = 0\) and the constrained dynamics 
+    in \((r,\theta)\) coordinates are continuous in \(I\), there exists 
+    \(I_1 > 0\) small enough that \(\dot{\theta} > 0\) for all
+    \(I \in \, ]0,I_1]\).
+    Hence, we can perform a time reparameterization \(t = t(\theta)\) to get a
+    time-scaled pseudo-radius \(\hat{r}(\theta) := r(t(\theta))\).
+    This reduces the system \((\dot{r},\dot{\theta})\) into the scalar
+    time-varying ODE 
+    \begin{equation}\label{eqn:rhat-ode}
+        \begin{cases}
+            \hat{r}^\prime= \frac{\dot{r}}{\dot{\theta}} =
+            g(\theta,\hat{r},I)
+            , \\
+            \hat{r}(0) = r_0
+            ,
+        \end{cases}
+    \end{equation}
+    where the prime denotes differentiation by \(\theta\) and 
+    \(r_0 \in \, ]\underline{r},\overline{r}[\) is the initial condition.
+
+    Let \(\hat{r}(\theta,r_0,I)\) be the solution to \eqref{eqn:rhat-ode}.
+    Taking the Taylor expansion of this solution around \(I = 0\) we get
+    \(\hat{r}(\theta,r_0,I) = \hat{r}(\theta,r_0,0) + I\hat{r}_1(\theta,r_0,0) 
+    + R(\theta,r_0,I)\) where
+
+    \textbf{TODO: finish the proof}
+\end{proof}
+\textbf{TODO: Prove origin is unstable, then apply lemma to O1 and O2 and show
+that orbits go through Epi}
 %/========== Conclusion ==========/%

### 2021-05-31 latex/paper/main.tex
First draft of the proof of lemma, which is in the proof of the theorem

@@
 % TODO: remove onecolumn and draftcls formatting
-\documentclass[journal,twoside,onecolumn,draftclsnofoot,web]{ieeecolor}
+\documentclass[journal,twoside,web]{ieeecolor}
 
@@
 \end{equation}
-Finally, the Kronecker delta \(\delta_i^j = 1\) if \(i = j\) and \(0\)
+The Kronecker delta is \(\delta_i^j = 1\) if \(i = j\) and \(0\)
 otherwise.
+Finally, we say a function \(f(x,I)\) is \(O(I^2)\)
+if
+\[
+    \lim\limits_{\epsilon \to 0} \frac{f(x,I)}{I} = 0
+    .
+\]
 
@@
     Let \(D \subset \SxR\) be open and 
-    let \(\underline{r}, \overline{r} \in \R_{> 0}\) with 
-    \(\underline{r} < \overline{r}\).
+    let \(]\underline{r}, \overline{r}[ \, \subset \R_{> 0}\).
     Suppose there exists a diffeomorphism 
     \begin{align*}
-        T : D &\to T(D) = ]\underline{r},\overline{r}[ \, \times \Sone \\
+        T : D &\to ]\underline{r},\overline{r}[ \, \times \Sone \\
         (q_u,p_u) &\mapsto (r,\theta)
@@
     \end{align*}
-    satisfying \(\dot{r} = 0\) and \(\dot{\theta} > 0\) on \(T(D)\) when 
+    satisfying \(\dot{r} \equiv 0\) and \(\dot{\theta} > 0\) when
     \(I = 0\).
     Let \(g(\theta,r,I) = \dot{r}/\dot{\theta}\) and assume that 
-    \((\partial g/\partial I)(\theta,r,0) = 0\).
-    Define \(b(\theta,r_0) = g(\theta,r_0,0)\) and assume there exists
-    \(\epsilon > 0\) such that 
+    \((\partial g/\partial I)(\theta,r,0) \equiv 0\).
+    Define \(b(\theta,r_0) = (\partial g/\partial r)(\theta,r_0,0)\);
+    assume it is \(2\pi\)-periodic in \(\theta\) for fixed \(r_0\),
+    and that there exists \(\epsilon > 0\) such that 
     \[
@@
     Then there exists \(I^\star > 0\) small enough that the constrained dynamics
-    gain energy on \(D\) for all \(0 < I \leq I^\star\), and lose energy on \(D\)
-    for all \(-I^\star \leq I < 0\).
+    gain energy on \(D\) for all \(I \in \,]0, I^\star]\), and lose energy on \(D\)
+    for all \(I \in [-I^\star, 0[\).
 \end{lemma}
@@
     Taking the Taylor expansion of this solution around \(I = 0\) we get
-    \(\hat{r}(\theta,r_0,I) = \hat{r}(\theta,r_0,0) + I\hat{r}_1(\theta,r_0,0) 
-    + R(\theta,r_0,I)\) where
-
-    \textbf{TODO: finish the proof}
+    \(\hat{r}(\theta,r_0,I) = \hat{r}(\theta,r_0,0) + I\hat{r}_1(\theta,r_0) 
+    + R(\theta,r_0,I)\). 
+    Here, \(\hat{r}(\theta,r_0,0) \equiv r_0\) for all \(\theta\),
+    the function \(\hat{r}_1(\theta,r_0)\) is the solution to
+    \begin{equation}
+        \begin{cases}
+        \hat{r}_1^\prime = 
+        \frac{\partial g}{\partial \hat{r}}(\theta,r_0,0)\hat{r}_1 +
+        \frac{\partial g}{\partial I}(\theta,r_0,0)
+        , \\
+        \hat{r}_1(0) = 0
+        ,
+    \end{cases}
+    \end{equation}
+    and the remainder \(R(\theta,r_0,I)\) is \(O(I^2)\).
+
+    By assumption, \((\partial g/\partial I)(\theta,r_0,0) \equiv 0\) and
+    \(b(\theta,r_0) := (\partial g/\partial \hat{r})(\theta,r_0,0)\). 
+    Hence, one can compute
+    \[
+        \hat{r}_1(\theta,r_0) = \int\limits_0^\theta b(\sigma,r_0)d\sigma
+        .
+    \]
+    Since and \(\dot{\theta} > 0\) and \(b(\theta,r_0)\) is \(2\pi\)-periodic in
+    \(\theta\), there is a well-defined Poincar\'{e} map describing
+    how \(\hat{r}\) changes each time \(\theta\) increases by \(2\pi\);
+    this map is 
+    \begin{align*}
+        P : \, ]\underline{r},\overline{r}[ &\to \mathbb{R}_{\geq 0} \\
+        r_0 &\mapsto \hat{r}(2\pi,r_0,I)
+        .
+    \end{align*}
+    This map expands to \(P(r_0) = r_0 + IS(r_0) + R(2\pi,r_0,I)\), which
+    means we can understand the evolution of \(\hat{r}\) by studying
+    the dynamics of the discrete-time system
+    \begin{equation}\label{eqn:r-discrete}
+        r_{n+1} = r_n + IS(r_n) + R(2\pi,r_n,I)
+        ,
+    \end{equation}
+    with initial condition \(r_0\).
+    Here, \(r_n\) corresponds to the distance along the line segment
+    \(L := ]\underline{r},\overline{r}[ \, \times \{0\}\)
+    when an orbit of the constrained dynamics hits \(L\) for the
+    the \(n^\text{th}\) time, assuming the orbit was initialized at
+    \((r,\theta) = (r_0,0)\).
+
+    Proving the constrained dynamics gain energy on \(D\) is equivalent to
+    proving \(\hat{r}_{n}\) eventually reaches \(\overline{r}\).
+    Thanks to a perturbation analysis from \cite[Theorem 10.1]{khalil_nonlinear}, 
+    we know there exists some \(I_2 \in ]0,I_1]\) and \(\delta > 0\)
+    such that \(\left|R(2\pi,r_n,I)\right| \leq \delta I^2\) for all 
+    \(I \in [-I_2,I_2]\).
+    Taking this fact alongside the assumption that \(S(r_0)\) is uniformly
+    bounded below by \(\epsilon\), we can find a lower bound on
+    \eqref{eqn:r-discrete}:
+    \[
+        r_{n+1} \geq r_n + I\epsilon - \delta I_2^2
+        .
+    \]
+    Choosing \(I^\star \in ]0,I_2]\) small enough makes 
+    \(I\epsilon - \delta I_2^2 \geq \gamma > 0\) for all \(I \in ]0,I^\star]\).
+    Therefore, \(r_{n+1} \geq r_n + \gamma \geq r_0 + n\gamma\), which means all
+    solutions of \eqref{eqn:r-discrete} flow towards \(\overline{r}\).
+    We conclude that, for \(I \in ]0, I^\star]\), the constrained dynamics gain
+    energy on \(D\).
+    Changing the sign of \(I\) means \(r_{n+1} \leq r_n - n\gamma\), in which
+    case the constrained dynamics lose energy on \(D\).
 \end{proof}

### 2021-05-31 latex/paper/main.tex
Changed some sentence structure in the proof of the Theorem

@@
 energy is
-\begin{equation}\label{eqn:acrobot-energy}
+\begin{equation}\label{eqn:nominal-energy}
     E(q_u,p_u) := \frac{p_u^2}{10ml^2} + 3mgl(1 - \cos(q_u))
@@
 \(\mathcal{O}_2(\bar{\rho})\) used in Theorem \ref{thm:acrobot-energy-stabilization}
-via the mechanical energy \eqref{eqn:acrobot-energy} of the nominal
+via the mechanical energy \eqref{eqn:nominal-energy} of the nominal
 pendulum associated with a simplified acrobot.
@@
 \section{Proof of Theorem \ref{thm:acrobot-energy-stabilization}}\label{sec:proof}
-The sets \(\mathcal{O}_1\) and \(\mathcal{O}_2(\bar{\rho})\) are sublevel sets of
-the nominal pendulum's energy \(E(q_u,p_u)\) described in
-\eqref{eqn:acrobot-energy}.
-To prove Theorem \ref{thm:acrobot-energy-stabilization}, we will find conditions
-under which the acrobot gains energy on an open ``energy annulus".
-That is, given two energy levels \(E_0 \geq 0\) and \(E_1 > E_0\), we define the
-energy annulus to be the domain
-\begin{equation*}
-    \mathcal{D} := \{(q_u,p_u) \in \SxR \mid \\
-    E_0 < E(q_u,p_u) < E_1\}
-    .
-\end{equation*}
-The only condition we place on \(E_0\) and \(E_1\) is that they both correspond
-to oscillations or both to rotations; that is, if \(E_0 < E(\pi,0)\), then 
-\(E_1 \leq E(\pi,0)\).
-
-Here is the fundamental idea of the proof: for certain open subsets
-\(D \subset \mathcal{D}\), one can find a change of coordinates into 
-pseudo-polar coordinates adapted to level sets of \(E\).
+Theorem \ref{thm:acrobot-energy-stabilization} proves that the acrobot gains
+energy on \(\mathcal{O}_1\) and \(\mathcal{O}_2(\bar{\rho})\), which are
+sublevel sets of the nominal pendulum's energy \eqref{eqn:nominal-energy}.
+%To prove Theorem \ref{thm:acrobot-energy-stabilization}, we will find conditions
+%under which the acrobot gains energy on an open ``energy annulus",
+%\begin{equation*}
+%    \mathcal{D} := \{(q_u,p_u) \in \SxR \mid \\
+%    E_0 < E(q_u,p_u) < E_1\}
+%    ,
+%\end{equation*}
+%where \(E_1 > E_0\) are energy values, which we will choose later.
+
+Here is the fundamental idea of this proof: for certain subsets
+\(D \subset \SxR\), one can find a change of coordinates into 
+pseudo-polar coordinates adapted to level sets of \(E\) 
+(see Mohammadi \etal \cite{dynamic_vhcs_stabilize_closed_orbits}).
 In other words, when \(I = 0\) in \eqref{eqn:acrobot-constraint}, one can find
@@
 
-We will describe the explicit forms of these transformations later. 
-Importantly, we will show that the dynamics of \((r,\theta)\) satisfy 
-\(\dot{r} = 0\) and \(\dot{\theta} > 0\) for the nominal pendulum (i.e., when
-\(I = 0\)).
-By continuity of the constrained dynamics with respect to \(I\), we can find a
-small enough \(I > 0\) so that \(\dot{\theta} > 0\) for the acrobot.
+We will explicitly calculate these transformations later;
+what matters is that the dynamics of \((r,\theta)\) will satisfy 
+\(\dot{r} = 0\) and \(\dot{\theta} > 0\) when \(I = 0\).
+By continuity of the constrained dynamics with respect to \(I\), this implies 
+there exists a small enough value \(I > 0\) so that \(\dot{\theta} > 0\) for the
+acrobot.
 This property will enable us to perform a time-scaling transformation 
@@
 on average and the constrained dynamics must be gaining energy on \(D\).
-
-We formalize this idea in the following lemma.
-Then we will find the relevant subset \(D \subset \mathcal{D}\)
-and the coordinate transformations \((q_u,p_u) \mapsto (r,\theta)\) which
-satisfy the conditions of the lemma, thereby proving the acrobot gains energy on
-each set.
-
-\begin{lemma}
+This idea is formalized in the following lemma.
+
+\begin{lemma}\label{lemma:energy-regulation-D}
     Consider the simplified acrobot \eqref{eqn:acrobot-hamiltonian} constrained
@@
 \end{proof}
+
 \textbf{TODO: Prove origin is unstable, then apply lemma to O1 and O2 and show

### 2021-06-01 latex/paper/main.tex
Continued writing the proof of Theorem.

@@
     Let \(g(\theta,r,I) = \dot{r}/\dot{\theta}\) and assume that 
-    \((\partial g/\partial I)(\theta,r,0) \equiv 0\).
-    Define \(b(\theta,r_0) = (\partial g/\partial r)(\theta,r_0,0)\);
-    assume it is \(2\pi\)-periodic in \(\theta\) for fixed \(r_0\),
+    \((\partial g/\partial r)(\theta,r,0) \equiv 0\).
+    Define \(b(\theta,r_0) = (\partial g/\partial I)(\theta,r_0,0)\);
+    assume it is \(2\pi\)-periodic in \(\theta\) for any fixed \(r_0\),
     and that there exists \(\epsilon > 0\) such that 
     \[
-        \int \limits_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon
+        S(r_0) := \int \limits_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon
         ,
@@
     the function \(\hat{r}_1(\theta,r_0)\) is the solution to
-    \begin{equation}
+    \begin{equation}\label{eqn:r1-hat-prime}
         \begin{cases}
@@
 
-    By assumption, \((\partial g/\partial I)(\theta,r_0,0) \equiv 0\) and
-    \(b(\theta,r_0) := (\partial g/\partial \hat{r})(\theta,r_0,0)\). 
-    Hence, one can compute
+    Our assumption that \((\partial g/\partial r)(\theta,r_0,0) \equiv 0\)
+    implies that \((\partial g/\partial \hat{r})(\theta,r_0,0) \equiv 0\).
+    The solution of \ref{eqn:r1-hat-prime} can be obtained by quadrature, giving
     \[
         \hat{r}_1(\theta,r_0) = \int\limits_0^\theta b(\sigma,r_0)d\sigma
-        .
+        ,
     \]
+    with \(b(\theta,r_0) := (\partial g/\partial I)(\theta,r_0,0)\). 
     Since and \(\dot{\theta} > 0\) and \(b(\theta,r_0)\) is \(2\pi\)-periodic in
     \(\theta\), there is a well-defined Poincar\'{e} map describing
-    how \(\hat{r}\) changes each time \(\theta\) increases by \(2\pi\);
-    this map is 
+    how \(\hat{r}\) changes each time \(\theta\) increases by \(2\pi\).
+    This map is 
     \begin{align*}
@@
         r_0 &\mapsto \hat{r}(2\pi,r_0,I)
-        .
+        ,
     \end{align*}
-    This map expands to \(P(r_0) = r_0 + IS(r_0) + R(2\pi,r_0,I)\), which
-    means we can understand the evolution of \(\hat{r}\) by studying
+    which expands to \(P(r_0) = r_0 + IS(r_0) + R(2\pi,r_0,I)\).
+    Hence, we can understand the evolution of \(\hat{r}\) by studying
     the dynamics of the discrete-time system
@@
     \(L := ]\underline{r},\overline{r}[ \, \times \{0\}\)
-    when an orbit of the constrained dynamics hits \(L\) for the
-    the \(n^\text{th}\) time, assuming the orbit was initialized at
-    \((r,\theta) = (r_0,0)\).
+    when an orbit of \((r,\theta)\) hits \(L\) for the the \(n^\text{th}\) time,
+    assuming the orbit was initialized at \((r,\theta) = (r_0,0)\).
 
@@
     proving \(\hat{r}_{n}\) eventually reaches \(\overline{r}\).
-    Thanks to a perturbation analysis from \cite[Theorem 10.1]{khalil_nonlinear}, 
+    Thanks to \cite[Theorem 10.1]{khalil_nonlinear}, 
     we know there exists some \(I_2 \in ]0,I_1]\) and \(\delta > 0\)
@@
 
-\textbf{TODO: Prove origin is unstable, then apply lemma to O1 and O2 and show
-that orbits go through Epi}
+We will now prove the first part of Theorem
+\ref{thm:acrobot-energy-stabilization}.
+To apply Lemma \ref{lemma:energy-regulation-D}, we choose \(D\) to be the
+annulus \(\mathcal{O}_1 \backslash \{(0,0)\}\) consisting of all oscillations of
+the nominal pendulum.
+Using the procedure described by Mohammadi \etal 
+\cite{dynamic_vhcs_stabilize_closed_orbits} for finding pseudo-polar
+coordinates adapted to level sets of \(E\), we compute
+\begin{align}
+    \label{eqn:r-osc}
+    r &= \arccos\left(\cos(q_u) - \frac{p_u^2}{30m^2gl^3}\right)
+    , \\
+    \label{eqn:theta-osc}
+    \theta &=
+    \left.\arctan_2\left(-\sign{p_u}\sqrt{1-\frac{q_u^2}{r^2}},\frac{q_u}{r}\right)
+    \right|_{r = r(q_u,p_u)}
+    ,
+\end{align}
+where \(r \in \, ]0,\pi[\).
+We now verify these coordinates satisfy the requirements of 
+Lemma \ref{lemma:energy-regulation-D}.
+When \(I = 0\) we have \(\dot{r} = 0\) and
+\begin{equation} \label{eqn:theta-dot-nom-osc}
+    \dot{\theta} = \sqrt{\frac{6g}{5l}} 
+        \sqrt{\frac{\cos(r\cos(\theta)) - \cos(r)}
+            {r^2 \sin(\theta)^2}}
+    .
+\end{equation}
+Note that ~\eqref{eqn:theta-dot-nom-osc} has removable singularities at
+\(\theta \in \{0,\pi\}\).
+Taking the limit as \(\theta\) approaches these points yields
+\begin{equation*}
+    \lim\limits_{\theta \to 0}\dot{\theta} 
+    = \lim\limits_{\theta \to \pi} \dot{\theta}
+    = \sqrt{\frac{6g \sin(r)}{10lr}}
+    , 
+\end{equation*}
+which is smooth and well-defined for all
+\(r \in \, ]0,\pi[\).
+Taking these removable singularities into account,
+one can verify that \(\dot{\theta} > 0\) for every \(r \in \, ]0,\pi[\).
+Setting \(g(\theta,r,I) = \dot{r}/\dot{\theta}\), MATLAB symbolic computations
+confirm that \((\partial g/\partial r)(\theta,r,0) \equiv 0\) as required.
+Next, we compute
+\begin{equation*}
+    b(\theta,r_0) = 
+    \frac{\partial g}{\partial I}(\theta, r_0, 0) = K a(\theta,r_0)
+    ,
+\end{equation*}
+where
+\begin{align*}
+    K &:= \frac{\bar{q}_a \sqrt{30m^2g l^3}}{15}
+    , \\
+    a(\theta,r_0) &:= \frac{
+        r_0 |\sin(\theta)| \left(
+        5 c_{r_0} \cos(r_0 c_\theta) - 8 \cos(r_0c_\theta)^2 + 3
+    \right)
+    }{
+    \sin(r_0)\sqrt{\cos(r_0c_\theta) - c_{r_0}}
+    }
+    .
+\end{align*}
+Notice that \(a(\theta,r_0)\) does not depend on the physical parameters of the
+acrobot.
+
+\textbf{TODO: write about Qro, show the figure, apply lemma to D,
+    prove origin is unstable, show orbits will go through Epi, 
+    then derive the constrained dynamics for the rotation domain and apply lemma
+    there, concluding that O2 is O1 U Epi U rotation domain and is therefore
+    also gaining energy.}
+
 %/========== Conclusion ==========/%

### 2021-06-03 latex/paper/main.tex
Wrote the energy gain for O1, need to show exit from Epi and then do O2

@@
 energy on \(\mathcal{O}_1\) and \(\mathcal{O}_2(\bar{\rho})\), which are
-sublevel sets of the nominal pendulum's energy \eqref{eqn:nominal-energy}.
-%To prove Theorem \ref{thm:acrobot-energy-stabilization}, we will find conditions
-%under which the acrobot gains energy on an open ``energy annulus",
-%\begin{equation*}
-%    \mathcal{D} := \{(q_u,p_u) \in \SxR \mid \\
-%    E_0 < E(q_u,p_u) < E_1\}
-%    ,
-%\end{equation*}
-%where \(E_1 > E_0\) are energy values, which we will choose later.
+sublevel sets of the nominal pendulum's energy
+\(E(q_u,p_u)\) defined in \eqref{eqn:nominal-energy}.
 
@@
 (see Mohammadi \etal \cite{dynamic_vhcs_stabilize_closed_orbits}).
-In other words, when \(I = 0\) in \eqref{eqn:acrobot-constraint}, one can find
+That is, when \(I = 0\) in \eqref{eqn:acrobot-constraint}, one can find
 a diffeomorphism \(T(q_u,p_u) = (r,\theta)\), where \(r\)
 chooses a level set of \(E\) and \(\theta\) chooses a point on that level set.
-
-We will explicitly calculate these transformations later;
-what matters is that the dynamics of \((r,\theta)\) will satisfy 
+We will explicitly calculate these transformations later.
+What matters is that the dynamics of \((r,\theta)\) will satisfy 
 \(\dot{r} = 0\) and \(\dot{\theta} > 0\) when \(I = 0\).
@@
 of the pseudo-angle \(\theta\), rather than as a function of time.
-If \(r(\theta)\) increases on average, then \(E\) must also be increasing
-on average and the constrained dynamics must be gaining energy on \(D\).
+We will show that, for small enough \(I > 0\), \(r(\theta)\) increases on
+average.
+Thus, \(E\) must also be increasing on average and the constrained dynamics must
+be gaining energy on \(D\).
 This idea is formalized in the following lemma.
@@
     by the VNHC \eqref{eqn:acrobot-constraint}, whose constrained dynamics are
-    parameterized by \((q_u,p_u) \in \SxR\).
+    \eqref{eqn:acrobot-constrained-dynamics}.
     Let \(D \subset \SxR\) be open and 
@@
     \end{equation}
-    where the prime denotes differentiation by \(\theta\) and 
+    where \(\theta\) now represents time.
+    The prime denotes differentiation by \(\theta\) and 
     \(r_0 \in \, ]\underline{r},\overline{r}[\) is the initial condition.
@@
     + R(\theta,r_0,I)\). 
-    Here, \(\hat{r}(\theta,r_0,0) \equiv r_0\) for all \(\theta\),
-    the function \(\hat{r}_1(\theta,r_0)\) is the solution to
+    Here, \(\hat{r}(\theta,r_0,0) \equiv r_0\) for all \(\theta\) because
+    \(g(\theta,\hat{r},0) \equiv 0\).
+    The function \(\hat{r}_1(\theta,r_0)\) is the solution to
     \begin{equation}\label{eqn:r1-hat-prime}
@@
     implies that \((\partial g/\partial \hat{r})(\theta,r_0,0) \equiv 0\).
-    The solution of \ref{eqn:r1-hat-prime} can be obtained by quadrature, giving
+    The solution of \ref{eqn:r1-hat-prime} can therefore be obtained by
+    quadrature, giving
     \[
@@
     which expands to \(P(r_0) = r_0 + IS(r_0) + R(2\pi,r_0,I)\).
-    Hence, we can understand the evolution of \(\hat{r}\) by studying
+    In other words, we can understand the evolution of \(\hat{r}\) by studying
     the dynamics of the discrete-time system
@@
     when an orbit of \((r,\theta)\) hits \(L\) for the the \(n^\text{th}\) time,
-    assuming the orbit was initialized at \((r,\theta) = (r_0,0)\).
-
-    Proving the constrained dynamics gain energy on \(D\) is equivalent to
-    proving \(\hat{r}_{n}\) eventually reaches \(\overline{r}\).
-    Thanks to \cite[Theorem 10.1]{khalil_nonlinear}, 
+    assuming the acrobot was initialized at \((r,\theta) = (r_0,0)\).
+
+    If \(r_n \not\in \, ]\underline{r},\overline{r}[\), the orbit must have
+    departed \(D\) and \eqref{eqn:r-discrete} is no longer valid.
+    Proving the constrained dynamics gain energy on \(D\) is therefore
+    equivalent to finding \(N \in \mathbb{Z}_{> 0}\) where 
+    \(\hat{r}_{n} > \overline{r}\) for all \(n \geq N\).
+    Using a perturbation analysis (see Khalil 
+    \cite[Theorem 10.1]{khalil_nonlinear}),
     we know there exists some \(I_2 \in ]0,I_1]\) and \(\delta > 0\)
@@
     \(I \in [-I_2,I_2]\).
-    Taking this fact alongside the assumption that \(S(r_0)\) is uniformly
-    bounded below by \(\epsilon\), we can find a lower bound on
+    Taking this fact alongside the assumption that \(S(r_0) \geq \epsilon\),
+    we find a lower bound on \(r_{n+1}\):
     \eqref{eqn:r-discrete}:
@@
     \(I\epsilon - \delta I_2^2 \geq \gamma > 0\) for all \(I \in ]0,I^\star]\).
-    Therefore, \(r_{n+1} \geq r_n + \gamma \geq r_0 + n\gamma\), which means all
-    solutions of \eqref{eqn:r-discrete} flow towards \(\overline{r}\).
+    Therefore, \(r_{n} \geq r_{n-1} + \gamma \geq r_0 + n\gamma\), 
+    so all solutions of \eqref{eqn:r-discrete} eventually flow past
+    \(\overline{r}\).
     We conclude that, for \(I \in ]0, I^\star]\), the constrained dynamics gain
     energy on \(D\).
-    Changing the sign of \(I\) means \(r_{n+1} \leq r_n - n\gamma\), in which
-    case the constrained dynamics lose energy on \(D\).
+
+    Changing the sign of \(I\) changes the direction of all the
+    above inequalites, which implies \(r_{n} \leq r_0 - n\gamma\). 
+    In this case all orbits of \eqref{eqn:r-discrete} flow below
+    \(\underline{r}\), i.e., the constrained dynamics lose energy on \(D\).
 \end{proof}
 
+\subsection{Energy gain on \(\mathcal{O}_1\)}
 We will now prove the first part of Theorem
 \ref{thm:acrobot-energy-stabilization}.
-To apply Lemma \ref{lemma:energy-regulation-D}, we choose \(D\) to be the
+We choose \(D\) in Lemma \ref{lemma:energy-regulation-D} to be the
 annulus \(\mathcal{O}_1 \backslash \{(0,0)\}\) consisting of all oscillations of
@@
 Lemma \ref{lemma:energy-regulation-D}.
+
 When \(I = 0\) we have \(\dot{r} = 0\) and
@@
 Taking the limit as \(\theta\) approaches these points yields
-\begin{equation*}
+\begin{equation}\label{eqn:theta-dot-limit}
     \lim\limits_{\theta \to 0}\dot{\theta} 
@@
     , 
-\end{equation*}
+\end{equation}
 which is smooth and well-defined for all
 \(r \in \, ]0,\pi[\).
-Taking these removable singularities into account,
+From \eqref{eqn:theta-dot-nom-osc}-\eqref{eqn:theta-dot-limit},
 one can verify that \(\dot{\theta} > 0\) for every \(r \in \, ]0,\pi[\).
-Setting \(g(\theta,r,I) = \dot{r}/\dot{\theta}\), MATLAB symbolic computations
+Defining \(g(\theta,r,I) = \dot{r}/\dot{\theta}\), MATLAB symbolic computations
 confirm that \((\partial g/\partial r)(\theta,r,0) \equiv 0\) as required.
-Next, we compute
-\begin{equation*}
-    b(\theta,r_0) = 
-    \frac{\partial g}{\partial I}(\theta, r_0, 0) = K a(\theta,r_0)
-    ,
-\end{equation*}
+
+Next we compute
+\(b(\theta,r_0) = (\partial g/\partial I)(\theta, r_0, 0) = K a(\theta,r_0)\),
 where
@@
 \end{align*}
-Notice that \(a(\theta,r_0)\) does not depend on the physical parameters of the
-acrobot.
-
-\textbf{TODO: write about Qro, show the figure, apply lemma to D,
-    prove origin is unstable, show orbits will go through Epi, 
+Notice that \(K\) is a positive constant which depends only on
+\(m\),\(g\),\(l\), and \(\bar{q}_a\).
+This means that there exists \(\epsilon > 0\) such that 
+\(S(r_0) := \int_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon\)
+if and only if \(Q(r_0) := \int_0^{2\pi} a(\sigma,r_0) d\sigma\) is strictly
+positive for all \(r_0 \in \, ]0,\pi[\).
+Since \(a(\theta,r_0)\) is adimensional, we can numerically compute \(Q(r_0)\)
+once for all acrobots.
+The plot of \(Q(r_0)\) when \(r_0 \in [10^{-10}, \pi - 10^{-3}]\) is shown in
+Figure \ref{fig:acrobot-Q}.
+We see that it is strictly positive and monotonically increasing, with an
+asymptote at \(r_0 = \pi\).
+Simulations with smaller \(r_0\) result in an infinite integral error due to a
+division by zero, yet we believe the \(Q(r_0)\) is in fact positive for all 
+\(r_0 \in \, ]0,\pi[\).
+
+\begin{figure}
+    \centering
+    % TODO: Show this plot
+    %\includegraphics[width=0.8\linewidth]{}
+    \caption{The plot of \(Q(r_0)\).}
+    \label{fig:acrobot-Q}
+\end{figure}
+
+Finally, \(a(\theta,r_0)\) is clearly \(2\pi\)-periodic in \(\theta\),
+which means all conditions of Lemma \ref{lemma:energy-regulation-D}
+are satisfied. 
+Therefore, the acrobot is gaining energy everywhere on
+\(D = \mathcal{O}_1 \backslash \{(0,0)\}\).
+The origin must be a repeller (otherwise Lemma
+\ref{lemma:energy-regulation-D} would fail), so we conclude that the acrobot
+is gaining energy on \(\mathcal{O}_1\).
+
+\textbf{TODO: show orbits will go through Epi, 
     then derive the constrained dynamics for the rotation domain and apply lemma

### 2021-06-04 latex/paper/main.tex
Finished writing proof of theorem, need to edit now.

@@
 where \(r \in \, ]0,\pi[\).
+This has inverse
+\begin{align*}
+    q_u &= r\cos(\theta)
+    , \\
+    p_u &= -\sign{\sin(\theta)}
+    \sqrt{30m^2gl^3(\cos(rc_\theta - c_r)}
+    .
+\end{align*}
 We now verify these coordinates satisfy the requirements of 
@@
 From \eqref{eqn:theta-dot-nom-osc}-\eqref{eqn:theta-dot-limit},
-one can verify that \(\dot{\theta} > 0\) for every \(r \in \, ]0,\pi[\).
+one can verify that \(\dot{\theta} > 0\) when \(I = 0\), for every 
+\(r \in \, ]0,\pi[\).
 Defining \(g(\theta,r,I) = \dot{r}/\dot{\theta}\), MATLAB symbolic computations
@@
 
-\textbf{TODO: show orbits will go through Epi, 
-    then derive the constrained dynamics for the rotation domain and apply lemma
-    there, concluding that O2 is O1 U Epi U rotation domain and is therefore
-    also gaining energy.}
+Before moving on to the second part of Theorem
+\ref{thm:acrobot-energy-stabilization}, we must first confirm that any acrobot
+constrained by \eqref{eqn:acrobot-constraint} will eventually start rotating.
+The definition of energy gain states that all orbits initialized in
+\(\mathcal{O}_1\) will escape compact subsets of \(\mathcal{O}_1\) in finite time. 
+This is not enough to prove that all orbits will escape the boundary of
+\(\mathcal{O}_1\).
+
+Let \(E_\pi\) be the level set with energy \(E(\pi,0)\), 
+which forms the boundary of \(\mathcal{O}_1\). 
+Define \(\bar{\mathcal{O}}_1 = \mathcal{O}_1 \cup E_\pi\) to be the closure of
+\(\mathcal{O}_1\).
+An orbit of the acrobot will only begin rotating if it escapes
+\(\bar{\mathcal{O}}_1\) by crossing through \(E_\pi\), and eventually remains
+outside this set.
+It is possible for orbits starting in \(\mathcal{O}_1\) to always approach
+\(E_\pi\) without ever crossing into the rotation zone.
+We will prove this does not happen in general.
+
+To begin, let us analyze the upright equilibrium.
+Taking the Jacobian of the constrained dynamics
+~\eqref{eqn:acrobot-constrained-dynamics} at \((q_u,p_u) = (\pi,0)\) yields
+\[
+    J = \begin{bmatrix}
+        -\frac{6mgl\bar{q}_aI}{5} & \frac{1 - 2m^2gl^3\bar{q}_a^2 I^2}{5ml^2} \\
+        3mgl & mgl\bar{q}_aI
+    \end{bmatrix}
+    ,
+\]
+which has characteristic polynomial
+\(\det\left(\lambda \Id{2} - J\right)
+= \lambda^2 + 1/5 mgl\bar{q}_a I \lambda - 3g\).
+By Descartes' rule of signs, this polynomial has one root with positive real
+part. 
+The equilibrium \((\pi,0)\) is therefore unstable, so the stable manifold \(\Pi^+\) of
+initial conditions converging to \((\pi,0)\) is one-dimensional, and hence is
+of measure zero in \(\SxR\).
+
+Using \(x(t) := (q_u(t),p_u(t))\) as shorthand, let
+\(x(0) \in \mathcal{O}_1\) be a nonzero initial condition of the acrobot.
+Suppose by way of contradiction that the orbit \(x(\R)\) is confined within 
+\(\bar{\mathcal{O}}_1\), and does not exit through \(E_\pi\) into the rotation
+zone.
+Since \(\bar{\mathcal{O}}_1\) is compact, the Birkhoff Theorem \cite{birkhoff}
+implies:
+\begin{itemize}
+    \item The positive limit set \(L_+\) of \(x(t)\) is non-empty, compact, and
+        invariant.
+    \item The solution \(x(t)\) asymptotically tends to \(L_+\).
+\end{itemize}
+Since the acrobot gains energy on \(\mathcal{O}_1\), the positive limit set of
+\(x(t)\) must be the largest invariant subset of \(E_\pi\). 
+From our discussion on pseudo-polar coordinates (in particular the extension
+\ref{eqn:theta-dot-limit} of \(\dot{\theta}\) to \(\bar{\mathcal{O}}_1\)),
+we know that \(\dot{\theta} \geq 0\) on \(E_\pi\) with equality if and only if 
+\(\theta \in \{0,\pi\}\).
+Taking the derivative of \(E(q_u,p_u)\) at the \(p_u\)-axis yields
+\(\dot{E}(0,p_u) = g/(5l) \sin(q_a)p_u\).
+This is non-zero everywhere except at the origin, which means it is non-zero on
+\(E_\pi\); this implies \(E_\pi\) is not invariant, so the positive limit set of 
+\(x(t)\) must be the set \(\{(\pi,0)\}\). Hence, \(x(t)\) converges to 
+\((\pi,0)\), so \(x(0) \in \Pi^+\).
+Since \(\Pi^+\) is a set of measure zero in \(\SxR\), almost every orbit
+initialized in \(\mathcal{O}_1\) must escape through \(E_\pi\) into the rotation
+domain.
+
+Escaping once into the rotation domain does not guarantee the acrobot
+orbit will never return to \(\mathcal{O}_1\).
+Indeed, the orbit could escape through \(E_\pi\) and return to \(\mathcal{O}_1\)
+several times.
+However, almost every orbit must eventually remain outside the closure of 
+\(\mathcal{O}_1\). 
+If not, the positive limit set would once again be
+\(E_\pi\); by the same argument as above, that orbit would be in \(\Pi^+\).
+
+We conclude that all orbits beginning in \(\mathcal{O}_1\) will, in finite time,
+escape the closure of \(\mathcal{O}_1\) and remain in the rotation domain
+forever after.
+
+\subsection{Energy Gain on \(\mathcal{O}_2(\bar{\rho})\)}
+Since \(\mathcal{O}_2(\bar{\rho}) = \bar{\mathcal{O}}_1 \cup \mathcal{R}\)
+(where \(\mathcal{R}\) is the rotation domain defined in
+\eqref{eqn:rotation-domain}), and since all orbits escape 
+\(\bar{\mathcal{O}_1}\) in finite time, we need simply prove energy gain on
+\(\mathcal{R}\) to complete the proof of Theorem
+\ref{thm:acrobot-energy-stabilization}.
+We separate \(\mathcal{R}\) into its two distinct connected components:
+\(\mathcal{R}^+\), the set of rotations with \(p_u > 0\); 
+and \(\mathcal{R}^-\), the set of rotations with \(p_u < 0\).
+We take the domain of interest for Lemma \ref{lemma:energy-regulation-D} to be
+\(D \in \{\mathcal{R}^+, \mathcal{R}^-\}\).
+
+We now apply the method described in
+\cite{dynamic_vhcs_stabilize_closed_orbits} to find pseudo-polar coordinates
+adapted to rotations.
+These coordinates are 
+\begin{align}
+    \label{eqn:r-rot}
+    r &= \sqrt{p_u^2 - 30m^2gl^3(1-c_u)}
+    , \\
+    \label{eqn:theta-rot}
+    \theta &= \pm q_u
+\end{align}
+where \(r \in \, ]\sqrt{60m^2gl^3}, \bar{\rho}[\) and \(\pm\) denotes
+whether \(D=\mathcal{R}^+\) or \(D=\mathcal{R}^-\).
+The inverse of this transformation is
+\begin{align*}
+    q_u &= \pm \theta
+    , \\
+    p_u = \pm \sqrt{r^2+30m^2gl^3(1-c_\theta)}
+    .
+\end{align*}
+
+When we compute the constrained dynamics in
+\((r,\theta)\) coordinates and set \(I = 0\), MATLAB symbolic computations
+reveal that \(\dot{r} = 0\) and
+\(\dot{\theta} = 1/(5ml^2) \sqrt{r^2 - 30m^2gl^3(1-c_\theta)}\).
+Regardless of \(D\), we have \(\dot{\theta} > 0\) when \(I = 0\) as required.
+Defining \(g(\theta,r,I) = \dot{r}/\dot{\theta}\), MATLAB symbolic computations
+confirm that \((\partial g/\partial r)(\theta,r,0) \equiv 0\).
+The function \(b(\theta,r_0) = (\partial g/\partial I)(\theta,r_0,0)\) is given
+in the statement of Theorem \ref{thm:acrobot-energy-stabilization}.
+The theorem also assumes that \(S(r_0) \geq \epsilon\).
+Thus, all requirements of Lemma \ref{lemma:energy-regulation-D} are satisfied,
+so there exists some \(I_2^\star(\bar{\rho})\) whereby the acrobot gains/loses
+energy on \(\mathcal{R}^+\) and \(\mathcal{R}^-\).
+
+Let \(I_1^\star\) be the value returned by Lemma \ref{lemma:energy-regulation-D}
+for \(\mathcal{O}_1\).
+We conclude that the acrobot gains/loses energy on \(\mathcal{O}_2(\bar{\rho})\)
+for \(|I| \leq I^{\star\star}\), where 
+\(I^{\star\star} = \min \{I_1^\star, I_2^\star(\bar{\rho})\}\).
+This completes the proof of Theorem \ref{thm:acrobot-energy-stabilization}.
 

### 2021-06-04 latex/paper/main.tex
Forgot about showing the origin is a repeller / stable, started doing that now.

@@
 \ref{thm:acrobot-energy-stabilization}.
-We choose \(D\) in Lemma \ref{lemma:energy-regulation-D} to be the
-annulus \(\mathcal{O}_1 \backslash \{(0,0)\}\) consisting of all oscillations of
-the nominal pendulum.
+We begin by studying the punctured open set
+\(\mathcal{O}_1 \backslash \{(0,0)\}\) consisting of all (nonzero) oscillations
+of the nominal pendulum.
 Using the procedure described by Mohammadi \etal 
@@
 Notice that \(K\) is a positive constant which depends only on
-\(m\),\(g\),\(l\), and \(\bar{q}_a\).
+\(m\),\(g\),\(l\), and \(\bar{q}_a\), and \(a(\theta,r_0)\) is \(2\pi\)-periodic
+in \(\theta\).
 This means that there exists \(\epsilon > 0\) such that 
@@
 \(r_0 \in \, ]0,\pi[\).
-
 \begin{figure}
@@
 
-Finally, \(a(\theta,r_0)\) is clearly \(2\pi\)-periodic in \(\theta\),
-which means all conditions of Lemma \ref{lemma:energy-regulation-D}
-are satisfied. 
-Therefore, the acrobot is gaining energy everywhere on
-\(D = \mathcal{O}_1 \backslash \{(0,0)\}\).
-The origin must be a repeller (otherwise Lemma
-\ref{lemma:energy-regulation-D} would fail), so we conclude that the acrobot
-is gaining energy on \(\mathcal{O}_1\).
+While \(Q(r_0) \not \geq \epsilon\) on \(]0,\pi[\), it is certainly bounded
+below on \(]\omega,\pi[\) for each fixed \(\omega > 0\).
+This means all conditions of Lemma \ref{lemma:energy-regulation-D}
+are satisfied on the set 
+\[
+    D_\omega = \mathcal{O}_1 \backslash 
+    \{(q_u,p_u) \in \SxR \mid E(q_u,p_u) \leq E(\omega,0)\}
+    .
+\]
+Therefore, for each \(\omega \in \, ]0,\pi[\) there exists some \(I_1^\star\)
+such that the acrobot gains/loses energy everywhere on \(D_\omega\).
+To prove energy gain on all of \(\mathcal{O}_1\), we will show the origin is a
+repeller if \(I > 0\), and that it is an asymptotically stable equilibrium
+if \(I < 0\).
+Recall the Poincar\'{e} section used in the proof of Lemama
+\ref{lemma:energy-regulation-D}.
+\textbf{TODO: Show the origin is a repeller/stable eqm via the poincare map}.
 

### 2021-06-05 latex/paper/main.tex
First draft of proof of Thm is finished, need to edit now.

@@
     This map is 
-    \begin{align*}
+    \begin{align}\label{eqn:poincare-map}
         P : \, ]\underline{r},\overline{r}[ &\to \mathbb{R}_{\geq 0} \\
@@
         ,
-    \end{align*}
+    \end{align}
     which expands to \(P(r_0) = r_0 + IS(r_0) + R(2\pi,r_0,I)\).
@@
 This means that there exists \(\epsilon > 0\) such that 
-\(S(r_0) := \int_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon\)
-if and only if \(Q(r_0) := \int_0^{2\pi} a(\sigma,r_0) d\sigma\) is strictly
-positive for all \(r_0 \in \, ]0,\pi[\).
+\(S(r_0) := \int_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon\) at \(r_0\)
+if and only if there exists \(\delta > 0\) such that 
+\(Q(r_0) := \int_0^{2\pi} a(\sigma,r_0) d\sigma \geq \delta\).
 Since \(a(\theta,r_0)\) is adimensional, we can numerically compute \(Q(r_0)\)
@@
 
-While \(Q(r_0) \not \geq \epsilon\) on \(]0,\pi[\), it is certainly bounded
-below on \(]\omega,\pi[\) for each fixed \(\omega > 0\).
+Unfortunately there is no \(\delta\) such that \(Q(r_0) \geq \delta\)
+everywhere on \(]0,\pi[\), though this \(\delta\) does exist 
+on \(]\omega,\pi[\) for each fixed \(\omega > 0\).
 This means all conditions of Lemma \ref{lemma:energy-regulation-D}
@@
 \]
-Therefore, for each \(\omega \in \, ]0,\pi[\) there exists some \(I_1^\star\)
-such that the acrobot gains/loses energy everywhere on \(D_\omega\).
+Therefore, for each \(\omega \in \, ]0,\pi[\) there exists some 
+\(I_1(\omega) > 0\) such that the acrobot gains/loses energy everywhere on
+\(D_\omega\).
 To prove energy gain on all of \(\mathcal{O}_1\), we will show the origin is a
-repeller if \(I > 0\), and that it is an asymptotically stable equilibrium
-if \(I < 0\).
-Recall the Poincar\'{e} section used in the proof of Lemama
-\ref{lemma:energy-regulation-D}.
-\textbf{TODO: Show the origin is a repeller/stable eqm via the poincare map}.
+repeller so that all orbits near the origin flow away from the origin into
+\(D_\omega\) (for small enough \(\omega\)) and continue gaining energy
+thereafter.
+
+Recall the discrete-time pseudo-radius dynamics \eqref{eqn:r-discrete} from the
+proof of Lemma \ref{lemma:energy-regulation-D}. 
+To prove the origin in the \((q_u,p_u)\)-plane is a repeller, we need simply
+prove the linearization of \eqref{eqn:r-discrete} is greater than \(1\) at 
+\(r_0 = 0\).
+Linearizing these discrete-time dynamics is equivalent to linearizing the
+Poincar\'{e} map \eqref{eqn:poincare-map} to get
+\[
+    P^\prime(0) = 1 + I K Q^\prime(0) + R^\prime(2\pi,0,I)
+    ,
+\]
+where prime denotes differentiation with respect to \(r\).
+MATLAB symbolic computations reveal that
+\begin{equation*}
+    \lim \limits_{r_0 \to 0^+}
+    \pdiff{a}{r}(\theta,r_0) 
+    = -\frac{\sqrt{2}}{2} \left(11\sin(\theta)^2 - 6\right)
+    ,
+\end{equation*}
+which means
+\begin{align*}
+    Q^\prime(0) = \int \limits_0^\pi \pdiff{a}{\mu}(\sigma,0) d\sigma
+    = \frac{\pi}{2\sqrt{2}} > 0
+    .
+\end{align*}
+Since the remainder term \(R\) is \(O(I^2)\), its partial derivative 
+\(R^\prime\) is also \(O(I^2)\).
+Hence, it can be written in the form
+\(I^2 \tilde{R}(I)\) where \(\tilde{R}(I)\) is smooth and zero at \(I = 0\).
+Thus, there exists \(I_0 > 0\) such that 
+\(I_0 K Q^\prime(0) + (I_0)^2 \tilde{R}(I) > 0\), which means
+\(P^\prime(0) > 1\) for all \(I \in \, ]0, I_0]\).
+We conclude that the origin is a repeller, which means there exists some
+(unknown) \(\mu > 0\) where the interval \(]0,\mu]\) is negatively
+invariant for \eqref{eqn:r-discrete}, and all solutions starting in this
+interval flow through \(r = \mu\).
+This in turn implies that all orbits starting in \(\mathcal{O}_1\) near the
+origin will gain energy up to \(E(\mu,0)\).
+We know there exists \(I_1(\mu)\) so that the acrobot gains energy on \(D_\mu\).
+Taking \(I^\star = \min\{I_0,I_1(\mu)\}\) means the acrobot will gain energy on
+\(\mathcal{O}_1\) for each \(I \in \, ]0,I^\star]\).
+If \(I < 0\), then \(P^\prime(0) < 1\) which implies the origin is an
+asymptotically stable equilibrium.
+Since the acrobot now loses energy on \(D_\mu\), all orbits in \(\mathcal{O}_1\)
+lose energy and converge to the origin for \(I \in [-I^\star,0[\).
 

### 2021-06-05 latex/paper/main.tex
First editing pass of the proof of the theorem

@@
 \section{Proof of Theorem \ref{thm:acrobot-energy-stabilization}}\label{sec:proof}
-Theorem \ref{thm:acrobot-energy-stabilization} proves that the acrobot gains
-energy on \(\mathcal{O}_1\) and \(\mathcal{O}_2(\bar{\rho})\), which are
-sublevel sets of the nominal pendulum's energy
-\(E(q_u,p_u)\) defined in \eqref{eqn:nominal-energy}.
-
-Here is the fundamental idea of this proof: for certain subsets
+In this section we prove Theorem \ref{thm:acrobot-energy-stabilization},
+which claims that the acrobot gains energy on \(\mathcal{O}_1\) and
+\(\mathcal{O}_2(\bar{\rho})\), which are sublevel sets of the nominal pendulum's
+energy \(E(q_u,p_u)\) defined in \eqref{eqn:nominal-energy}.
+
+Here is the summary of this proof: for certain subsets
 \(D \subset \SxR\), one can find a change of coordinates into 
-pseudo-polar coordinates adapted to level sets of \(E\) 
+pseudo-polar coordinates adapted to level sets of \(E(q_u,p_u)\)
 (see Mohammadi \etal \cite{dynamic_vhcs_stabilize_closed_orbits}).
 That is, when \(I = 0\) in \eqref{eqn:acrobot-constraint}, one can find
-a diffeomorphism \(T(q_u,p_u) = (r,\theta)\), where \(r\)
+a diffeomorphism \((q_u,p_u) \leftrightarrow (r,\theta)\), where \(r\)
 chooses a level set of \(E\) and \(\theta\) chooses a point on that level set.
-We will explicitly calculate these transformations later.
-What matters is that the dynamics of \((r,\theta)\) will satisfy 
+The acrobot's dynamics in \((r,\theta)\) coordinates will satisfy 
 \(\dot{r} = 0\) and \(\dot{\theta} > 0\) when \(I = 0\).
-By continuity of the constrained dynamics with respect to \(I\), this implies 
-there exists a small enough value \(I > 0\) so that \(\dot{\theta} > 0\) for the
-acrobot.
-This property will enable us to perform a time-scaling transformation 
-\(t = t(\theta)\), and study how the pseudo-radius \(r\) evolves as a function
-of the pseudo-angle \(\theta\), rather than as a function of time.
-We will show that, for small enough \(I > 0\), \(r(\theta)\) increases on
-average.
-Thus, \(E\) must also be increasing on average and the constrained dynamics must
-be gaining energy on \(D\).
+By continuity of the constrained dynamics with respect to \(I\),
+we maintain \(\dot{\theta} > 0\) for small non-zero values of \(I\).
+This will enable us to perform a time-scaling transformation 
+\(t = t(\theta)\), and study how the acrobot's pseudo-radius \(r\) evolves as a
+function of the pseudo-angle \(\theta\), rather than as a function of time.
+We will formally show that \(r(\theta)\) increases ``on average", which means
+\(E\) is also be increasing on average.
+This means the constrained dynamics must be gaining energy on \(D\).
 This idea is formalized in the following lemma.
@@
     \((\partial g/\partial r)(\theta,r,0) \equiv 0\).
-    Define \(b(\theta,r_0) = (\partial g/\partial I)(\theta,r_0,0)\);
-    assume it is \(2\pi\)-periodic in \(\theta\) for any fixed \(r_0\),
+    Assume \(b(\theta,r_0) = (\partial g/\partial I)(\theta,r_0,0)\)
+    is \(2\pi\)-periodic in \(\theta\) for any fixed \(r_0\),
     and that there exists \(\epsilon > 0\) such that 
@@
     \(I_1 > 0\) small enough that \(\dot{\theta} > 0\) for all
-    \(I \in \, ]0,I_1]\).
+    \(I \in \, [-I_1,I_1]\).
     Hence, we can perform a time reparameterization \(t = t(\theta)\) to get a
@@
     + R(\theta,r_0,I)\). 
-    Here, \(\hat{r}(\theta,r_0,0) \equiv r_0\) for all \(\theta\) because
+    Note that \(\hat{r}(\theta,r_0,0) \equiv r_0\) for all \(\theta\) because
     \(g(\theta,\hat{r},0) \equiv 0\).
@@
         \frac{\partial g}{\partial \hat{r}}(\theta,r_0,0)\hat{r}_1 +
-        \frac{\partial g}{\partial I}(\theta,r_0,0)
+        b(\theta,r_0)
+        %\frac{\partial g}{\partial I}(\theta,r_0,0)
         , \\
@@
     \end{equation}
-    and the remainder \(R(\theta,r_0,I)\) is \(O(I^2)\).
+    and the remainder term \(R(\theta,r_0,I)\) is \(O(I^2)\).
 
     Our assumption that \((\partial g/\partial r)(\theta,r_0,0) \equiv 0\)
-    implies that \((\partial g/\partial \hat{r})(\theta,r_0,0) \equiv 0\).
-    The solution of \ref{eqn:r1-hat-prime} can therefore be obtained by
+    means that \((\partial g/\partial \hat{r})(\theta,r_0,0) \equiv 0\).
+    The solution of \eqref{eqn:r1-hat-prime} can therefore be obtained by
     quadrature, giving
@@
         \hat{r}_1(\theta,r_0) = \int\limits_0^\theta b(\sigma,r_0)d\sigma
-        ,
+        .
     \]
-    with \(b(\theta,r_0) := (\partial g/\partial I)(\theta,r_0,0)\). 
+%    with \(b(\theta,r_0) := (\partial g/\partial I)(\theta,r_0,0)\). 
     Since and \(\dot{\theta} > 0\) and \(b(\theta,r_0)\) is \(2\pi\)-periodic in
@@
 
-    If \(r_n \not\in \, ]\underline{r},\overline{r}[\), the orbit must have
-    departed \(D\) and \eqref{eqn:r-discrete} is no longer valid.
+    If \(r_n \not\in \, ]\underline{r},\overline{r}[\), the orbit 
+    in \((q_u,p_u)\) coordinates must have
+    departed \(D\). % and \eqref{eqn:r-discrete} is no longer valid.
     Proving the constrained dynamics gain energy on \(D\) is therefore
     equivalent to finding \(N \in \mathbb{Z}_{> 0}\) where 
-    \(\hat{r}_{n} > \overline{r}\) for all \(n \geq N\).
+    \(\hat{r}_{n} \geq \overline{r}\) for all \(n \geq N\).
     Using a perturbation analysis (see Khalil 
     \cite[Theorem 10.1]{khalil_nonlinear}),
-    we know there exists some \(I_2 \in ]0,I_1]\) and \(\delta > 0\)
-    such that \(\left|R(2\pi,r_n,I)\right| \leq \delta I^2\) for all 
+    we know there exists some \(I_2 \in ]0,I_1]\) and \(\kappa > 0\)
+    such that \(\left|R(2\pi,r_n,I)\right| \leq \kappa I^2\) for all 
     \(I \in [-I_2,I_2]\).
@@
     we find a lower bound on \(r_{n+1}\):
-    \eqref{eqn:r-discrete}:
     \[
-        r_{n+1} \geq r_n + I\epsilon - \delta I_2^2
+        r_{n+1} \geq r_n + I\epsilon - \kappa I_2^2
         .
     \]
-    Choosing \(I^\star \in ]0,I_2]\) small enough makes 
-    \(I\epsilon - \delta I_2^2 \geq \gamma > 0\) for all \(I \in ]0,I^\star]\).
-    Therefore, \(r_{n} \geq r_{n-1} + \gamma \geq r_0 + n\gamma\), 
+    Choosing \(\gamma > 0\) and \(I^\star \in \, ]0,I_2]\) small enough makes 
+    \(I\epsilon - \delta I_2^2 \geq \gamma\) for all \(I \in ]0,I^\star]\).
+    Therefore, \(r_{n} \geq r_0 + n\gamma\), 
     so all solutions of \eqref{eqn:r-discrete} eventually flow past
     \(\overline{r}\).
-    We conclude that, for \(I \in ]0, I^\star]\), the constrained dynamics gain
+    We conclude that, for \(I \in \, ]0, I^\star]\), the constrained dynamics gain
     energy on \(D\).
@@
     Changing the sign of \(I\) changes the direction of all the
-    above inequalites, which implies \(r_{n} \leq r_0 - n\gamma\). 
+    above inequalites, so that \(r_{n} \leq r_0 - n\gamma\). 
     In this case all orbits of \eqref{eqn:r-discrete} flow below
@@
 
-\subsection{Energy gain on \(\mathcal{O}_1\)}
-We will now prove the first part of Theorem
-\ref{thm:acrobot-energy-stabilization}.
-We begin by studying the punctured open set
-\(\mathcal{O}_1 \backslash \{(0,0)\}\) consisting of all (nonzero) oscillations
-of the nominal pendulum.
-Using the procedure described by Mohammadi \etal 
-\cite{dynamic_vhcs_stabilize_closed_orbits} for finding pseudo-polar
-coordinates adapted to level sets of \(E\), we compute
+To prove Theorem \ref{thm:acrobot-energy-stabilization}, we will explicitly
+compute the pseudo-radius coordinates for \(\mathcal{O}_1\) and
+\(\mathcal{O}_2(\bar{\rho})\) and apply Lemma \ref{lemma:energy-regulation-D}.
+
+\subsection{Energy Gain on \(\mathcal{O}_1\)}\label{sec:proof-o1}
+To prove the first part of Theorem
+\ref{thm:acrobot-energy-stabilization},
+we begin by studying the punctured open set
+\(\mathcal{O}_1 \backslash \{(0,0)\} \subset \SxR\) consisting of all
+oscillations of the nominal pendulum.
+Using the procedure from \cite{dynamic_vhcs_stabilize_closed_orbits} for finding
+pseudo-polar coordinates adapted to level sets of \(E\), we compute
 \begin{align}
@@
 \begin{align*}
-    q_u &= r\cos(\theta)
+    q_u &= rc_\theta
     , \\
-    p_u &= -\sign{\sin(\theta)}
-    \sqrt{30m^2gl^3(\cos(rc_\theta - c_r)}
-    .
+    p_u &= -\sign{s_\theta}
+    \sqrt{30m^2gl^3\cos(rc_\theta - c_r)}
+    ,
 \end{align*}
+where for shorthand we write \(c_\theta := \cos(\theta)\), \(s_\theta :=
+\sin(\theta)\), \(c_r := \cos(r)\), and \(s_r := \sin(r)\).
 We now verify these coordinates satisfy the requirements of 
@@
     \dot{\theta} = \sqrt{\frac{6g}{5l}} 
-        \sqrt{\frac{\cos(r\cos(\theta)) - \cos(r)}
-            {r^2 \sin(\theta)^2}}
+        \sqrt{\frac{\cos(r c_\theta) - c_r}
+            {r^2 s_\theta^2}}
     .
@@
     = \lim\limits_{\theta \to \pi} \dot{\theta}
-    = \sqrt{\frac{6g \sin(r)}{10lr}}
+    = \sqrt{\frac{6g s_r}{10lr}}
     , 
@@
     a(\theta,r_0) &:= \frac{
-        r_0 |\sin(\theta)| \left(
+        r_0 |s_\theta| \left(
         5 c_{r_0} \cos(r_0 c_\theta) - 8 \cos(r_0c_\theta)^2 + 3
@@
     }{
-    \sin(r_0)\sqrt{\cos(r_0c_\theta) - c_{r_0}}
+    s_{r_0}\sqrt{\cos(r_0c_\theta) - c_{r_0}}
     }
@@
 Notice that \(K\) is a positive constant which depends only on
-\(m\),\(g\),\(l\), and \(\bar{q}_a\), and \(a(\theta,r_0)\) is \(2\pi\)-periodic
-in \(\theta\).
+\(m\),\(g\),\(l\), and \(\bar{q}_a\), while \(a(\theta,r_0)\) is adimensional
+and \(2\pi\)-periodic in \(\theta\).
 This means that there exists \(\epsilon > 0\) such that 
-\(S(r_0) := \int_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon\) at \(r_0\)
+\(S(r_0) := \int_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon\)
 if and only if there exists \(\delta > 0\) such that 
@@
 Unfortunately there is no \(\delta\) such that \(Q(r_0) \geq \delta\)
-everywhere on \(]0,\pi[\), though this \(\delta\) does exist 
+everywhere on \(]0,\pi[\), though such a \(\delta\) does exist 
 on \(]\omega,\pi[\) for each fixed \(\omega > 0\).
@@
 Linearizing these discrete-time dynamics is equivalent to linearizing the
-Poincar\'{e} map \eqref{eqn:poincare-map} to get
+Poincar\'{e} map \eqref{eqn:poincare-map}, which yields 
 \[
@@
 Thus, there exists \(I_0 > 0\) such that 
-\(I_0 K Q^\prime(0) + (I_0)^2 \tilde{R}(I) > 0\), which means
+\(I_0 K Q^\prime(0) + (I_0)^2 \tilde{R}(I_0) > 0\), which means
 \(P^\prime(0) > 1\) for all \(I \in \, ]0, I_0]\).
@@
 
+We have succeeded in proving that any acrobot constrained by 
+our VNHC \eqref{eqn:acrobot-constraint} gains/loses energy on \(\mathcal{O}_1\).
 Before moving on to the second part of Theorem
-\ref{thm:acrobot-energy-stabilization}, we must first confirm that any acrobot
-constrained by \eqref{eqn:acrobot-constraint} will eventually start rotating.
+\ref{thm:acrobot-energy-stabilization}, we must first confirm that the
+acrobot will eventually start rotating.
 The definition of energy gain states that all orbits initialized in
@@
 which has characteristic polynomial
-\(\det\left(\lambda \Id{2} - J\right)
-= \lambda^2 + 1/5 mgl\bar{q}_a I \lambda - 3g\).
+\[
+    \det\left(\lambda \Id{2} - J\right)
+    = \lambda^2 + \frac{mgl\bar{q}_a I}{5} \lambda - 3g\)
+    .
+\]
 By Descartes' rule of signs, this polynomial has one root with positive real
@@
 From our discussion on pseudo-polar coordinates (in particular the extension
-\ref{eqn:theta-dot-limit} of \(\dot{\theta}\) to \(\bar{\mathcal{O}}_1\)),
+\eqref{eqn:theta-dot-limit} of \(\dot{\theta}\) to \(\bar{\mathcal{O}}_1\)),
 we know that \(\dot{\theta} \geq 0\) on \(E_\pi\) with equality if and only if 
@@
     , \\
-    p_u = \pm \sqrt{r^2+30m^2gl^3(1-c_\theta)}
+    p_u &= \pm \sqrt{r^2+30m^2gl^3(1-c_\theta)}
     .
@@
 reveal that \(\dot{r} = 0\) and
-\(\dot{\theta} = 1/(5ml^2) \sqrt{r^2 - 30m^2gl^3(1-c_\theta)}\).
-Regardless of \(D\), we have \(\dot{\theta} > 0\) when \(I = 0\) as required.
+\[
+\dot{\theta} = \frac{1}{5ml^2} \sqrt{r^2 - 30m^2gl^3(1-c_\theta)}
+.
+\]
+We thus have \(\dot{\theta} > 0\) when \(I = 0\) as required.
 Defining \(g(\theta,r,I) = \dot{r}/\dot{\theta}\), MATLAB symbolic computations
-confirm that \((\partial g/\partial r)(\theta,r,0) \equiv 0\).
+also confirm that \((\partial g/\partial r)(\theta,r,0) \equiv 0\).
 The function \(b(\theta,r_0) = (\partial g/\partial I)(\theta,r_0,0)\) is given
-in the statement of Theorem \ref{thm:acrobot-energy-stabilization}.
-The theorem also assumes that \(S(r_0) \geq \epsilon\).
+in the statement of Theorem \ref{thm:acrobot-energy-stabilization}, and by
+assumption we have \(S(r_0) \geq \epsilon\) on
+\(]\sqrt{60m^2gl^3},\bar{\rho}[\).
 Thus, all requirements of Lemma \ref{lemma:energy-regulation-D} are satisfied,
-so there exists some \(I_2^\star(\bar{\rho})\) whereby the acrobot gains/loses
-energy on \(\mathcal{R}^+\) and \(\mathcal{R}^-\).
-
-Let \(I_1^\star\) be the value returned by Lemma \ref{lemma:energy-regulation-D}
-for \(\mathcal{O}_1\).
-We conclude that the acrobot gains/loses energy on \(\mathcal{O}_2(\bar{\rho})\)
-for \(|I| \leq I^{\star\star}\), where 
-\(I^{\star\star} = \min \{I_1^\star, I_2^\star(\bar{\rho})\}\).
+so there exists a common \(I_2(\bar{\rho})\) whereby the acrobot gains/loses
+energy on both \(\mathcal{R}^+\) and \(\mathcal{R}^-\).
+
+Let \(I_1^\star\) be the parameter we found for \(\mathcal{O}_1\) in 
+Section \ref{sec:proof-o1}.
+Taking \(I^{\star\star} = \min \{I_1^\star, I_2(\bar{\rho})\}\),
+the acrobot gains/loses energy on \(\mathcal{O}_2(\bar{\rho})\)
+when \(|I| \leq I^{\star\star}\).
 This completes the proof of Theorem \ref{thm:acrobot-energy-stabilization}.

### 2021-06-05 latex/paper/main.tex
Round 2 edit of the whole paper, got to page 5.

@@
 % TODO: remove onecolumn and draftcls formatting
-\documentclass[journal,twoside,web]{ieeecolor}
+\documentclass[journal,twoside,web, onecolumn, draftcls]{ieeecolor}
 
@@
 
-\newcommand*{\etal}{\MakeLowercase{\textit{et al.}}}
+\newcommand*{\etal}{\MakeLowercase{\textit{et al.~}}}
 %/========== /Preamble ==========/%
@@
     energy.
-    This constraint is tested on a real-world acrobot, demonstrating highly
-    effective energy regulation properties and robustness to a variety of
-    disturbances.
+    This constraint is tested in simulation and on a real-world acrobot,
+    demonstrating highly effective energy regulation properties and robustness
+    to a variety of disturbances.
 \end{abstract}
@@
 In particular, VNHCs may be capable of injecting and
-dissipating energy from a system in a robust manner, all while producing realistic
-biological motion.
+dissipating energy from a system in a robust manner, all while producing
+realistic biological motion.
 
@@
 rotations around the bar.
-Under suitable conditions on the acrobot's mass and length parameters, this VNHC
-also enables the acrobot to rotate with a desired speed.
-We will conclude by providing simulations which demonstrate a mechanism for
-energy regulation through VNHCS, along with experimental results which
-demonstrate the robustness of this behaviour to model uncertainty, sensor noise,
-and a variety of external disturbances.
+Under suitable conditions on the acrobot's physical parameters, this VNHC
+also enables the acrobot to rotate around the bar a desired speed.
+We will provide simulations to show how one can use VNHCs to regulate energy,
+along with experimental results which demonstrate the robustness of this
+behaviour to model uncertainty, sensor noise, and a variety of external
+disturbances.
 
@@
 otherwise.
-Finally, we say a function \(f(x,I)\) is \(O(I^2)\)
+Finally, we say a function \(R(I)\) is \(O(I^2)\)
 if
 \[
-    \lim\limits_{\epsilon \to 0} \frac{f(x,I)}{I} = 0
+    \lim\limits_{I \to 0} \frac{R(I)}{I} = 0
     .
@@
 Our goal is to design a smooth function \(f : \Sone \times \R\) such that the
-relation \(q_a = f(q_u,p_u)\) can be enforced asymptotically via feedback
-control (in Section \ref{sec:vnhc} we call this a regular virtual
-nonholonomic constraint).  
+relation \(q_a = f(q_u,p_u)\) for system \eqref{eqn:acrobot-hamiltonian}
+can be enforced asymptotically via feedback control (in Section \ref{sec:vnhc}
+we call this a regular virtual nonholonomic constraint).  
 We will further require that the dynamics of the acrobot, when the relation
@@
 \section{Theory of VNHCs}\label{sec:vnhc}
-Before embarking on our design problem, we begin by summarizing the relevant
+Before embarking on our design problem, we must summarize the relevant
 theory of virtual nonholonomic constraints for a class of mechanical systems we
@@
 
-We now take \(\mathcal{Q}\) to be embedded in \(\R^n\), so the coordinate
-transformation used the following theorem is well-defined.
+We temporarily take \(\mathcal{Q}\) to be embedded in \(\R^N\), so the
+coordinate transformation used the following theorem is well-defined.
 
@@
     \det\left(\lambda \Id{2} - J\right)
-    = \lambda^2 + \frac{mgl\bar{q}_a I}{5} \lambda - 3g\)
+    = \lambda^2 + \frac{mgl\bar{q}_a I}{5} \lambda - 3g
     .

### 2021-06-07 latex/paper/main.tex
Full edit of everything except theorem proof, and edit of first part of proof

@@
 angle and velocity.
-In 2016, Wang attempted to design such a controller \cite{xingbo_thesis}
-but was unable to prove the acrobot would gain energy over time.
+In 2016, Wang designed a controller \cite{xingbo_thesis} which tracked the
+body angle and an \textit{estimate} of the velocity, but not the velocity itself.
 His approach was a preliminary version of a recent technique know as the method
@@
 Under suitable conditions on the acrobot's physical parameters, this VNHC
-also enables the acrobot to rotate around the bar a desired speed.
+also enables the acrobot to rotate around the bar at a desired speed.
 We will provide simulations to show how one can use VNHCs to regulate energy,
@@
 derivatives, 
-\(\nabla_xA := [\pdiff{A}{x_1};\ldots;\pdiff{A}{x_m}] \in \R^{nm \times n}\).
+\(\nabla_xA := [\partial A/\partial x_1; \cdots; \partial A/\partial x_m] \in
+\R^{nm \times n}\).
+%\(\nabla_xA := [\pdiff{A}{x_1};\ldots;\pdiff{A}{x_m}] \in \R^{nm \times n}\).
 Given two matrices \(A \in \R^{n \times m}\) and \(B \in \R^{r \times s}\), the
-Kronecker product \(A \otimes B \in \R^{nr \times ms}\) (see \cite{kronprod}) is
-the matrix
+Kronecker product (see \cite{kronprod}) is the matrix  
+\(A \otimes B \in \R^{nr \times ms}\)  defined as
 \begin{equation}\label{eqn:kronprod}
@@
 \end{equation}
-The Kronecker delta is \(\delta_i^j = 1\) if \(i = j\) and \(0\)
+The Kronecker delta \(\delta_i^j\) is \(1\) if \(i = j\) and \(0\)
 otherwise.
 Finally, we say a function \(R(I)\) is \(O(I^2)\)
-if
-\[
-    \lim\limits_{I \to 0} \frac{R(I)}{I} = 0
-    .
-\]
+if \(\lim_{I \to 0} R(I)/I = 0\).
 
@@
 We also ignore any dissipative forces.
-
-%\begin{figure}
-%    \centering
-%    \includegraphics[width=0.8\linewidth]{simple_acrobot_model.pdf}
-%    \caption{A simple acrobot has massless rods of equal length \(l\) and 
-%    equal masses \(m\) at the tips.}
-%    \label{fig:simple-acrobot-model}
-%\end{figure}
 
@@
     condition \(x(0) \in K\), there exists \(T > 0\) such
-    that \(x(t) \notin K\) for all \(t > T)\).
+    that \(x(t) \notin K\) for all \(t > T\).
     The system \textit{loses energy on \(D\)} if it gains energy in
@@
 
-We temporarily take \(\mathcal{Q}\) to be embedded in \(\R^N\), so the
-coordinate transformation used the following theorem is well-defined.
+We temporarily take \(\mathcal{Q}\) to be embedded in \(\R^N\) so that the
+transformation \(\tilde{q} = \mathbf{B}q\) is well-defined.
 
@@
     \(dh_{p_a} = \Zmat{k \times k}\) 
-    and
-    \[
-        \rank{\left(\mathcal{A}(q,p_u) - dh_{p_u}\mathcal{M}(q,p)\right)\simpleB} = k
-         ,
-    \]
-    everywhere on the constraint manifold \(\Gamma\).
+    and the decoupling matrix
+    \begin{equation}\label{eqn:decoupling-matrix}
+        \left(\mathcal{A}(q,p_u) - dh_{p_u}\mathcal{M}(q,p)\right)\simpleB
+         ,
+     \end{equation}
+    is full rank \(k\) everywhere on the constraint manifold \(\Gamma\).
 \end{thm}
@@
     \( \ddot{e} = (\star) - 
-        dh_{p_u}\left(1/2~d/dt~\left(\mathcal{M}(q,p)p\right)\right) 
+        dh_{p_u}\left(1/2~d/dt\left(\mathcal{M}(q,p)p\right)\right) 
         + \mathcal{A}(q,p_u)[\Zmat{(n-k)\times k};\Id{k}] \tau\),
@@
         \cdot
-        \Big(dh_{p_u}\nabla_{q_u}(V(q) -
+        \Big(dh_{p_u}\nabla_{q_u}V(q) -
         \\
@@
         \left.\Big)\right|_{q_a = f(q_u,p_u)}
+        .
     \end{multline}
@@
     + \mathcal{A}(q,p_u)[\Id{n-k};\Zmat{k \times (n-k)}] p_u\).
-    Since \(h(q,p_u)\) is regular, \(\mathcal{A}(q,p_u)\) is invertible.
+    Since \(h(q,p_u)\) is regular,
+    \(\mathcal{A}(q,p_u)[\Zmat{(n-k)\times k}; \Id{k}]\) is invertible.
     Taking \(e = \dot{e} = 0\), solving for \(p_a\), and setting 
@@
 Recall that the acrobot in Figure \ref{fig:acrobot} has dynamics given by
-\eqref{eqn:acrobot-hamiltonian}, repeated here for convenience{
+\eqref{eqn:acrobot-hamiltonian}, repeated here for convenience:
     \begin{equation*}
@@
 \begin{proof}
-    The decoupling matrix for the acrobot evaluates to
+    The decoupling matrix \eqref{eqn:decoupling-matrix} for the acrobot
+    evaluates to
     \(((1+c_a)\partial_{q_u}f(q_u,p_u) + (3+2c_a))/(ml^2(2-c_a^2))\).
@@
 When the pendulum is rotating with bounded momentum
-\(\abs{p_u} < \bar{\rho}\) (for some chosen \(\bar{\rho}\)),
+\(\abs{p_u} < \rho\) (for some chosen \(\rho\)),
 \((q_u,p_u)\) must lie inside the rotation domain
@@
 \begin{equation}\label{eqn:o-rhobar}
-    \mathcal{O}_2(\bar{\rho}) := \left\{(q_u,p_u) \in \SxR
-        \mid E(q_u,p_u) < E(0,\bar{\rho}) \right\}
+    \mathcal{O}_2(\rho) := \left\{(q_u,p_u) \in \SxR
+        \mid E(q_u,p_u) < E(0,\rho) \right\}
     ,
@@
 shown in Figure \ref{fig:acrobot-o2}.
-If the VNHC injects energy on \(\mathcal{O}_2(\bar{\rho})\) for some 
+If the VNHC injects energy on \(\mathcal{O}_2(\rho)\) for some 
 \(I \neq 0\), then the acrobot must necessarily swing up, begin rotating, 
-and eventually rotate with a momentum of at least \(\bar{\rho}\).
+and eventually rotate with a momentum of at least \(\rho\).
 
@@
         \includegraphics[width=\linewidth]{acrobot_omega.pdf}
-        \caption{The set \(\mathcal{O}_2(\bar{\rho})\). Orbits starting in this
-            set are guaranteed to reach the level set \(E(0,\bar{\rho})\) of the
+        \caption{The set \(\mathcal{O}_2(\rho)\). Orbits starting in this
+            set are guaranteed to reach the level set \(E(0,\rho)\) of the
             nominal pendulum.}
@@
     \end{subfigure}
-    \caption{The sets on which the acrobot gains energy, according to Theorem
-        \ref{thm:acrobot-energy-stabilization}.}
+    \caption{The sets on which the acrobot gains energy, according to
+        Theorem \ref{thm:acrobot-energy-stabilization}.}
 \end{figure}
@@
 Unfortunately, our VNHC does not always inject energy on \(\mathcal{O}_1\) and
-\(\mathcal{O}_2(\bar{\rho})\).
+\(\mathcal{O}_2(\rho)\).
 If \(I\) is too large, the leg controller saturates and the body oscillates
@@
 \begin{enumerate}
-    \item For positive \(m\), \(g\), \(l\), and \(\bar{q}_a\), there exists
+    \item For each \(m, g, l, \bar{q}_a > 0\), there exists
         \(I^\star > 0\) such that, for all \(I \in \, ]0,I^\star]\), the VNHC
@@
         \(S(\rho_0) := \int \limits_{0}^{2\pi} b(\sigma,\rho_0)d\sigma\).
-        Fix \(\bar{\rho} > \sqrt{60m^2gl^3}\) and
+        Fix \(\rho > \sqrt{60m^2gl^3}\) and
         suppose there exists \(\epsilon > 0\) so that 
         \(S(\rho_0) \geq \epsilon\) for all 
-        \(\rho_0 \in \, \left]\sqrt{60m^2gl^3}, \bar{\rho}\right]\).
+        \(\rho_0 \in \, \left]\sqrt{60m^2gl^3}, \rho\right]\).
         Then there exists \(I^{\star\star} \in \, ]0, I^\star]\) such that, for all 
@@
         \eqref{eqn:acrobot-constraint} injects energy into the acrobot on
-        \(\mathcal{O}_2(\bar{\rho})\).
+        \(\mathcal{O}_2(\rho)\).
         If instead \(I \in [-I^{\star\star},0[\), the VNHC dissipates energy.
@@
 
-Notice that \(\mathcal{O}_1 \subset \mathcal{O}_2(\bar{\rho})\), yet
+Notice that \(\mathcal{O}_1 \subset \mathcal{O}_2(\rho)\), yet
 Theorem \ref{thm:acrobot-energy-stabilization} considers these sets separately.
@@
 For the acrobot to achieve giants with energy
-\(E(0,\bar{\rho})\), it must satisfy the assumption on the integral
+\(E(0,\rho)\), it must satisfy the assumption on the integral
 of \(b(\beta,\rho_0)\). 
@@
 If the assumption holds, there is some control value \(I\)
-(which depends on \(\bar{\rho}\)) for which the acrobot will 
-achieve rotations with a momentum of at least \(\bar{\rho}\).
+(which depends on \(\rho\)) for which the acrobot will 
+achieve rotations with a momentum of at least \(\rho\).
 
@@
 hysteresis value \(\delta \in [0,1]\).
-Each time the orbit crosses the \(q_u\)-axis (i.e. when \(p_u = 0\)) or 
+Each time the orbit crosses the \(q_u\)-axis (i.e., when \(p_u = 0\)) or 
 when \(|q_u| = \pi\), the supervisor does the following: 
@@
 \end{itemize}
+Here one must be careful to choose \(\delta\) so that
+\((1+\delta)q_\text{des} < \pi\), otherwise a rotating acrobot will not
+dissipate energy.
 
@@
 \end{itemize}
+While this method is titled ``rotation regulation", it is entirely possible to
+choose \(p_\text{des} < \sqrt{60m^2gl^3}\);
+this would stabilize an oscillation whose momentum at the bottom of the swing
+arc is \(p_\text{des}\).
 
@@
 \section{Simulation Results}\label{sec:simulations}
-In Section \ref{sec:experiments} we will test our VNHC on the physical acrobot
-built by Xingbo Wang \cite{xingbo_thesis}.
-This physical acrobot cannot be modelled by the simplified setup in Figure
+In Section \ref{sec:experiments} we will test our VNHC on a physical acrobot.
+%built by Xingbo Wang \cite{xingbo_thesis}.
+This acrobot cannot be modelled by the simplified setup in Figure
 \ref{fig:acrobot}, because the torso and leg links have 
@@
 Recall that we defined the sets \(\mathcal{O}_1\) and
-\(\mathcal{O}_2(\bar{\rho})\) used in Theorem \ref{thm:acrobot-energy-stabilization}
+\(\mathcal{O}_2(\rho)\) used in Theorem \ref{thm:acrobot-energy-stabilization}
 via the mechanical energy \eqref{eqn:nominal-energy} of the nominal
@@
 \eqref{eqn:acrobot-constraint}.
-Recall that the boundary between oscillations and rotations of the nominal
+The boundary between oscillations and rotations of the nominal
 pendulum is obtained by finding the level set \(E_\pi\) of the nominal energy
@@
 
-We can extend this idea to find the nominal pendulum associated with Wang's
-acrobot, whose physical parameters are provided in Table
+We can extend this idea to find the nominal pendulum associated with our
+physical acrobot, whose parameters are provided in Table
 \ref{tab:acrobot-parameters}.
@@
 
-We will now simulate the effects of constraining Wang's acrobot with the VNHC
+We will now simulate the effects of constraining our real acrobot with the VNHC
 \eqref{eqn:acrobot-constraint}, thereby demonstrating that VNHCs are robust to
@@
 constrained system for \(30\) seconds.
-The resulting orbit (on the constraint manifold) is plotted in Figure
+The resulting orbit is plotted in Figure
 \ref{fig:acrobot-in-orbit}.
@@
 
-Recall from \ref{sec:energy-reg} that one can use a supervisor to stabilize
+Recall from Section \ref{sec:energy-reg} that one can use a supervisor to stabilize
 oscillations by appropriately toggling between injection and dissipation VNHCs
@@
 In this section we will demonstrate that our VNHC is robust to friction, sensor
-noise, and other real-world considerations by testing it on the physical
-acrobot built by Xingbo Wang \cite{xingbo_thesis} 
-(Figure \ref{fig:xingbo-acrobot}).
-This platform was given the name SUGAR, which stands for
+noise, and other real-world considerations by testing it on our physical
+acrobot (Figure \ref{fig:xingbo-acrobot}).
+This platform is called SUGAR, which stands for
 Simple Underactuated Gymnastics and Acrobatics Robot.
@@
 SUGAR is comprised of two 3D-printed links: a torso and a leg.
-The torso link houses an Arduino Nano microcontroller unit (MCU) which controls
+The torso houses an Arduino Nano microcontroller unit (MCU) which controls
 a Dynamixel RX24F servo motor between the torso and the leg.
@@
 via \(q_a^{k} = \arctan(I p_u^{k-1})\).
-Unfortunately, this negatively affects the stabilization to the constraint
+This negatively affects the stabilization to the constraint
 manifold because we are introducing timing errors from the servo's built-in PID
@@
 which claims that the acrobot gains energy on \(\mathcal{O}_1\) and
-\(\mathcal{O}_2(\bar{\rho})\), which are sublevel sets of the nominal pendulum's
+\(\mathcal{O}_2(\rho)\), which are sublevel sets of the nominal pendulum's
 energy \(E(q_u,p_u)\) defined in \eqref{eqn:nominal-energy}.
@@
 pseudo-polar coordinates adapted to level sets of \(E(q_u,p_u)\)
-(see Mohammadi \etal \cite{dynamic_vhcs_stabilize_closed_orbits}).
+(see the general procedure devised by Mohammadi \etal 
+\cite{dynamic_vhcs_stabilize_closed_orbits}).
 That is, when \(I = 0\) in \eqref{eqn:acrobot-constraint}, one can find
-a diffeomorphism \((q_u,p_u) \leftrightarrow (r,\theta)\), where \(r\)
+a diffeomorphism \(T(q_u,p_u) =  (r,\theta)\), where \(r\)
 chooses a level set of \(E\) and \(\theta\) chooses a point on that level set.
@@
 function of the pseudo-angle \(\theta\), rather than as a function of time.
-We will formally show that \(r(\theta)\) increases ``on average", which means
-\(E\) is also be increasing on average.
-This means the constrained dynamics must be gaining energy on \(D\).
-This idea is formalized in the following lemma.
+We will formally show that \(r(\theta)\) increases on average, which means
+\(E\) also increases on average.
+This in turn implies the constrained dynamics are gaining energy on \(D\).
+We formalize these ideas in the following lemma.
 
@@
     \end{equation}
-    where \(\theta\) now represents time.
-    The prime denotes differentiation by \(\theta\) and 
+    where \(\theta\) now represents time,
+    the prime denotes differentiation by \(\theta\), and
     \(r_0 \in \, ]\underline{r},\overline{r}[\) is the initial condition.
@@
     \cite[Theorem 10.1]{khalil_nonlinear}),
-    we know there exists some \(I_2 \in ]0,I_1]\) and \(\kappa > 0\)
+    we know there exists some \(I_2 \in \, ]0,I_1]\) and \(\kappa > 0\)
     such that \(\left|R(2\pi,r_n,I)\right| \leq \kappa I^2\) for all 
@@
     \[
-        r_{n+1} \geq r_n + I\epsilon - \kappa I_2^2
+        r_{n+1} \geq r_n + I\epsilon - \kappa I^2
         .
@@
     Choosing \(\gamma > 0\) and \(I^\star \in \, ]0,I_2]\) small enough makes 
-    \(I\epsilon - \delta I_2^2 \geq \gamma\) for all \(I \in ]0,I^\star]\).
+    \(I\epsilon - \delta I^2 \geq \gamma\) for all \(I \in \, ]0,I^\star]\).
     Therefore, \(r_{n} \geq r_0 + n\gamma\), 
@@
 compute the pseudo-radius coordinates for \(\mathcal{O}_1\) and
-\(\mathcal{O}_2(\bar{\rho})\) and apply Lemma \ref{lemma:energy-regulation-D}.
+\(\mathcal{O}_2(\rho)\) and apply Lemma \ref{lemma:energy-regulation-D}.
 
@@
 \end{align}
-where \(r \in \, ]0,\pi[\).
+where \((r,\theta) \in \, ]0,\pi[ \times \Sone\).
 This has inverse
@@
 From \eqref{eqn:theta-dot-nom-osc}-\eqref{eqn:theta-dot-limit},
-one can verify that \(\dot{\theta} > 0\) when \(I = 0\), for every 
-\(r \in \, ]0,\pi[\).
+one can verify that \(\dot{\theta} > 0\) when \(I = 0\).
 Defining \(g(\theta,r,I) = \dot{r}/\dot{\theta}\), MATLAB symbolic computations
@@
 Notice that \(K\) is a positive constant which depends only on
-\(m\),\(g\),\(l\), and \(\bar{q}_a\), while \(a(\theta,r_0)\) is adimensional
+\(m\), \(g\), \(l\), and \(\bar{q}_a\), while \(a(\theta,r_0)\) is adimensional
 and \(2\pi\)-periodic in \(\theta\).
@@
     \centering
-    % TODO: Show this plot
-    %\includegraphics[width=0.8\linewidth]{}
+    \includegraphics[width=0.8\linewidth]{Qro.png}
     \caption{The plot of \(Q(r_0)\).}
@@
 This means all conditions of Lemma \ref{lemma:energy-regulation-D}
-are satisfied on the set 
+are satisfied on the annulus
 \[
@@
 \begin{align*}
-    Q^\prime(0) = \int \limits_0^\pi \pdiff{a}{\mu}(\sigma,0) d\sigma
+    Q^\prime(0) = \int \limits_0^\pi \pdiff{a}{r}(\sigma,0) d\sigma
     = \frac{\pi}{2\sqrt{2}} > 0
@@
 Thus, there exists \(I_0 > 0\) such that 
-\(I_0 K Q^\prime(0) + (I_0)^2 \tilde{R}(I_0) > 0\), which means
-\(P^\prime(0) > 1\) for all \(I \in \, ]0, I_0]\).
-We conclude that the origin is a repeller, which means there exists some
+\(I K Q^\prime(0) + I^2 \tilde{R}(I) > 0\) for all \(I \in \, ]0,I_0]\),
+which implies \(P^\prime(0) > 1\).
+The origin is therefore a repeller, which means there exists some
 (unknown) \(\mu > 0\) where the interval \(]0,\mu]\) is negatively
@@
 origin will gain energy up to \(E(\mu,0)\).
-We know there exists \(I_1(\mu)\) so that the acrobot gains energy on \(D_\mu\).
-Taking \(I^\star = \min\{I_0,I_1(\mu)\}\) means the acrobot will gain energy on
+
+Since there exists \(I_1(\mu)\) so that the acrobot gains energy on \(D_\mu\),
+taking \(I^\star = \min\{I_0,I_1(\mu)\}\) means the acrobot will gain energy on
 \(\mathcal{O}_1\) for each \(I \in \, ]0,I^\star]\).
@@
 asymptotically stable equilibrium.
-Since the acrobot now loses energy on \(D_\mu\), all orbits in \(\mathcal{O}_1\)
-lose energy and converge to the origin for \(I \in [-I^\star,0[\).
+The acrobot now loses energy on \(D_\mu\), and all orbits in \(\mathcal{O}_1\)
+lose energy for \(I \in [-I^\star,0[\).
 
@@
 \ref{thm:acrobot-energy-stabilization}, we must first confirm that the
-acrobot will eventually start rotating.
-The definition of energy gain states that all orbits initialized in
-\(\mathcal{O}_1\) will escape compact subsets of \(\mathcal{O}_1\) in finite time. 
-This is not enough to prove that all orbits will escape the boundary of
-\(\mathcal{O}_1\).
+acrobot will eventually start rotating:
+the definition of energy gain states that all orbits initialized in
+\(\mathcal{O}_1\) will escape compact subsets of \(\mathcal{O}_1\) in finite
+time, which is not enough to conclude that all orbits will escape the boundary
+of \(\mathcal{O}_1\).
 
@@
 
-To begin, let us analyze the upright equilibrium.
 Taking the Jacobian of the constrained dynamics
-~\eqref{eqn:acrobot-constrained-dynamics} at \((q_u,p_u) = (\pi,0)\) yields
+~\eqref{eqn:acrobot-constrained-dynamics} at the upright equilibrium 
+\((q_u,p_u) = (\pi,0)\) yields
 \[
@@
 \(x(t)\) must be the largest invariant subset of \(E_\pi\). 
-From our discussion on pseudo-polar coordinates (in particular the extension
-\eqref{eqn:theta-dot-limit} of \(\dot{\theta}\) to \(\bar{\mathcal{O}}_1\)),
-we know that \(\dot{\theta} \geq 0\) on \(E_\pi\) with equality if and only if 
+Taking the limit as \(r \to \pi\) of
+\eqref{eqn:theta-dot-nom-osc}--\eqref{eqn:theta-dot-limit},
+we get \(\dot{\theta} \geq 0\) on \(E_\pi\) with equality if and only if
 \(\theta \in \{0,\pi\}\).
+Thus, if \(E_\pi\) is not invariant, the largest invariant subset is the upright
+equilibrium \((q_u,p_u) = (\pi,0)\).
+
 Taking the derivative of \(E(q_u,p_u)\) at the \(p_u\)-axis yields
@@
 
-We conclude that all orbits beginning in \(\mathcal{O}_1\) will, in finite time,
-escape the closure of \(\mathcal{O}_1\) and remain in the rotation domain
-forever after.
-
-\subsection{Energy Gain on \(\mathcal{O}_2(\bar{\rho})\)}
-Since \(\mathcal{O}_2(\bar{\rho}) = \bar{\mathcal{O}}_1 \cup \mathcal{R}\)
-(where \(\mathcal{R}\) is the rotation domain defined in
+We conclude that almost all orbits beginning in \(\mathcal{O}_1\) will, in
+finite time, escape the closure of \(\mathcal{O}_1\) and remain in the rotation
+domain forever after.
+
+\subsection{Energy Gain on \(\mathcal{O}_2(\rho)\)}
+Since \(\mathcal{O}_2(\rho) = \bar{\mathcal{O}}_1 \cup \mathcal{R}(\rho)\)
+(where \(\mathcal{R}(\rho)\) is the rotation domain defined in
 \eqref{eqn:rotation-domain}), and since all orbits escape 
 \(\bar{\mathcal{O}_1}\) in finite time, we need simply prove energy gain on
-\(\mathcal{R}\) to complete the proof of Theorem
+\(\mathcal{R}(\rho)\) to complete the proof of Theorem
 \ref{thm:acrobot-energy-stabilization}.
-We separate \(\mathcal{R}\) into its two distinct connected components:
-\(\mathcal{R}^+\), the set of rotations with \(p_u > 0\); 
-and \(\mathcal{R}^-\), the set of rotations with \(p_u < 0\).
+We separate \(\mathcal{R}(\rho)\) into its two distinct connected components:
+\(\mathcal{R}^+(\rho)\), the set of rotations with \(p_u > 0\); 
+and \(\mathcal{R}^-(\rho)\), the set of rotations with \(p_u < 0\).
 We take the domain of interest for Lemma \ref{lemma:energy-regulation-D} to be
-\(D \in \{\mathcal{R}^+, \mathcal{R}^-\}\).
+\(D \in \{\mathcal{R}^+(\rho), \mathcal{R}^-(\rho)\}\).
 
@@
 \end{align}
-where \(r \in \, ]\sqrt{60m^2gl^3}, \bar{\rho}[\) and \(\pm\) denotes
-whether \(D=\mathcal{R}^+\) or \(D=\mathcal{R}^-\).
+where \(r \in \, ]\sqrt{60m^2gl^3}, \rho[\) and \(\pm\) denotes
+whether \(D=\mathcal{R}^+(\rho)\) or \(D=\mathcal{R}^-(\rho)\).
 The inverse of this transformation is
@@
 assumption we have \(S(r_0) \geq \epsilon\) on
-\(]\sqrt{60m^2gl^3},\bar{\rho}[\).
+\(]\sqrt{60m^2gl^3},\rho[\).
 Thus, all requirements of Lemma \ref{lemma:energy-regulation-D} are satisfied,
-so there exists a common \(I_2(\bar{\rho})\) whereby the acrobot gains/loses
+so there exists a common \(I_2(\rho)\) whereby the acrobot gains/loses
 energy on both \(\mathcal{R}^+\) and \(\mathcal{R}^-\).
@@
 Section \ref{sec:proof-o1}.
-Taking \(I^{\star\star} = \min \{I_1^\star, I_2(\bar{\rho})\}\),
-the acrobot gains/loses energy on \(\mathcal{O}_2(\bar{\rho})\)
+Taking \(I^{\star\star} = \min \{I_1^\star, I_2(\rho)\}\),
+the acrobot gains/loses energy on \(\mathcal{O}_2(\rho)\)
 when \(|I| \leq I^{\star\star}\).
@@
 
-In this article we apply the framework of virtual nonholonomic constraints
-towards the acrobot, and designed a constraint which emulates giant motion from
+In this article we applied the framework of virtual nonholonomic constraints
+on the acrobot, and designed a constraint which emulates giant motion from
 gymnastics.

### 2021-06-07 latex/paper/main.tex
First full editing pass of the paper

@@
 \(\bar{\mathcal{O}_1}\) in finite time, we need simply prove energy gain on
-\(\mathcal{R}(\rho)\) to complete the proof of Theorem
-\ref{thm:acrobot-energy-stabilization}.
+\(\mathcal{R}(\rho)\) to prove energy gain on \(\mathcal{O}_2(\rho)\).
 We separate \(\mathcal{R}(\rho)\) into its two distinct connected components:
@@
     \label{eqn:r-rot}
-    r &= \sqrt{p_u^2 - 30m^2gl^3(1-c_u)}
+    r &= \sqrt{p_u^2 + 30m^2gl^3(1-c_u)}
     , \\
@@
 \end{align}
-where \(r \in \, ]\sqrt{60m^2gl^3}, \rho[\) and \(\pm\) denotes
-whether \(D=\mathcal{R}^+(\rho)\) or \(D=\mathcal{R}^-(\rho)\).
+where \((r,\theta) \in \, ]\sqrt{60m^2gl^3}, \rho[ \times \Sone\).
 The inverse of this transformation is
@@
     , \\
-    p_u &= \pm \sqrt{r^2+30m^2gl^3(1-c_\theta)}
+    p_u &= \pm \sqrt{r^2-30m^2gl^3(1-c_\theta)}
     .
 \end{align*}
-
-When we compute the constrained dynamics in
-\((r,\theta)\) coordinates and set \(I = 0\), MATLAB symbolic computations
-reveal that \(\dot{r} = 0\) and
+In this context, \(\pm\) denotes whether \(D=\mathcal{R}^+(\rho)\) or
+\(D=\mathcal{R}^-(\rho)\).
+
+In computing the constrained dynamics in
+\((r,\theta)\) coordinates and setting \(I = 0\), 
+we find that \(\dot{r} \equiv 0\) and
 \[
 \dot{\theta} = \frac{1}{5ml^2} \sqrt{r^2 - 30m^2gl^3(1-c_\theta)}
-.
+,
 \]
-We thus have \(\dot{\theta} > 0\) when \(I = 0\) as required.
-Defining \(g(\theta,r,I) = \dot{r}/\dot{\theta}\), MATLAB symbolic computations
+which is strictly positive as required.
+Defining \(g(\theta,r,I) = \dot{r}/\dot{\theta}\), symbolic computations
 also confirm that \((\partial g/\partial r)(\theta,r,0) \equiv 0\).
@@
 \(]\sqrt{60m^2gl^3},\rho[\).
-Thus, all requirements of Lemma \ref{lemma:energy-regulation-D} are satisfied,
-so there exists a common \(I_2(\rho)\) whereby the acrobot gains/loses
-energy on both \(\mathcal{R}^+\) and \(\mathcal{R}^-\).
-
-Let \(I_1^\star\) be the parameter we found for \(\mathcal{O}_1\) in 
+
+All requirements of Lemma \ref{lemma:energy-regulation-D} are satisfied,
+so there exists a control parameter \(I_2(\rho) > 0\) whereby the acrobot
+gains energy on both \(\mathcal{R}^+\) and \(\mathcal{R}^-\) for
+each \(I \in \, ]0,I_2(\rho)]\).
+Let \(I_1^\star\) be the value we found for \(\mathcal{O}_1\) in 
 Section \ref{sec:proof-o1}.
@@
 the acrobot gains/loses energy on \(\mathcal{O}_2(\rho)\)
-when \(|I| \leq I^{\star\star}\).
+for each \(I \in \,]0, I^{\star\star}]\).
+If instead \(I \in [-I^{\star\star},0[\), the acrobot dissipates energy on
+\(\mathcal{O}_2(\rho)\).
 This completes the proof of Theorem \ref{thm:acrobot-energy-stabilization}.

### 2021-06-07 latex/paper/main.tex
Edit to send to Manfredi, which is still a paragraph too long.

@@
 % TODO: remove onecolumn and draftcls formatting
-\documentclass[journal,twoside,web, onecolumn, draftcls]{ieeecolor}
+\documentclass[journal,twoside,web, onecolumn,draftcls]{ieeecolor}
 
@@
 In this section we prove Theorem \ref{thm:acrobot-energy-stabilization},
-which claims that the acrobot gains energy on \(\mathcal{O}_1\) and
-\(\mathcal{O}_2(\rho)\), which are sublevel sets of the nominal pendulum's
-energy \(E(q_u,p_u)\) defined in \eqref{eqn:nominal-energy}.
-
-Here is the summary of this proof: for certain subsets
-\(D \subset \SxR\), one can find a change of coordinates into 
-pseudo-polar coordinates adapted to level sets of \(E(q_u,p_u)\)
-(see the general procedure devised by Mohammadi \etal 
-\cite{dynamic_vhcs_stabilize_closed_orbits}).
-That is, when \(I = 0\) in \eqref{eqn:acrobot-constraint}, one can find
-a diffeomorphism \(T(q_u,p_u) =  (r,\theta)\), where \(r\)
-chooses a level set of \(E\) and \(\theta\) chooses a point on that level set.
-The acrobot's dynamics in \((r,\theta)\) coordinates will satisfy 
-\(\dot{r} = 0\) and \(\dot{\theta} > 0\) when \(I = 0\).
-By continuity of the constrained dynamics with respect to \(I\),
-we maintain \(\dot{\theta} > 0\) for small non-zero values of \(I\).
-This will enable us to perform a time-scaling transformation 
-\(t = t(\theta)\), and study how the acrobot's pseudo-radius \(r\) evolves as a
-function of the pseudo-angle \(\theta\), rather than as a function of time.
-We will formally show that \(r(\theta)\) increases on average, which means
-\(E\) also increases on average.
-This in turn implies the constrained dynamics are gaining energy on \(D\).
-We formalize these ideas in the following lemma.
+which claims the acrobot gains/loses energy on \(\mathcal{O}_1\) and
+\(\mathcal{O}_2(\rho)\).
+These are the sublevel sets \eqref{eqn:oscillation-domain} and
+\eqref{eqn:o-rhobar} of the nominal pendulum's energy \(E(q_u,p_u)\) defined in
+\eqref{eqn:nominal-energy}.
+
+Here is fundamental idea behind this proof:
+using a procedure devised by Mohammadi \etal
+\cite{dynamic_vhcs_stabilize_closed_orbits}, on certain subsets 
+\(D \subset \SxR\) one can find a diffeomorphism \(T(q_u,p_u) = (r,\theta)\)
+into pseudo-polar coordinates adapted to level sets of \(E(q_u,p_u)\).
+That is, when \(I = 0\) in \eqref{eqn:acrobot-constraint}, the pseudo-radius
+\(r\) chooses a level set of \(E\) and the pseudo-angle \(\theta\) chooses a
+point on that level set.
+We can formally show that \(r\) increases ``on average", which means \(E\)
+also increases on average and the constrained dynamics are gaining energy on
+\(D\).
+We formalize this idea in the following lemma.
+
+%Here is the summary of this proof: for certain subsets
+%\(D \subset \SxR\), one can find a change of coordinates into 
+%pseudo-polar coordinates adapted to level sets of \(E(q_u,p_u)\)
+%(see the general procedure devised by Mohammadi \etal 
+%\cite{dynamic_vhcs_stabilize_closed_orbits}).
+%That is, when \(I = 0\) in \eqref{eqn:acrobot-constraint}, one can find
+%a diffeomorphism \(T(q_u,p_u) =  (r,\theta)\), where \(r\)
+%chooses a level set of \(E\) and \(\theta\) chooses a point on that level set.
+%The acrobot's dynamics in \((r,\theta)\) coordinates will satisfy 
+%\(\dot{r} = 0\) and \(\dot{\theta} > 0\) when \(I = 0\).
+%By continuity of the constrained dynamics with respect to \(I\),
+%we maintain \(\dot{\theta} > 0\) for small non-zero values of \(I\).
+%This will enable us to perform a time-scaling transformation 
+%\(t = t(\theta)\), and study how the acrobot's pseudo-radius \(r\) evolves as a
+%function of the pseudo-angle \(\theta\), rather than as a function of time.
+%We will formally show that \(r(\theta)\) increases on average, which means
+%\(E\) also increases on average.
+%This in turn implies the constrained dynamics are gaining energy on \(D\).
+%We formalize these ideas in the following lemma.
 
@@
     \begin{align*}
-        T : D &\to ]\underline{r},\overline{r}[ \, \times \Sone \\
+        T : D &\to ]\underline{r},\overline{r}[ \, \times \Sone
+         \\
         (q_u,p_u) &\mapsto (r,\theta)
@@
     \end{align*}
+    into pseudo-polar coordinates adapted to level sets of \(E(q_u,p_u)\),
     satisfying \(\dot{r} \equiv 0\) and \(\dot{\theta} > 0\) when
@@
     and that there exists \(\epsilon > 0\) such that 
-    \[
-        S(r_0) := \int \limits_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon
-        ,
-    \]
+    \(S(r_0) := \int_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon\)
+%    \[
+%        S(r_0) := \int \limits_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon
+%        ,
+%    \]
     for all \(r_0 \in \, ]\underline{r},\overline{r}[\).
@@
         \begin{cases}
-            \hat{r}^\prime= \frac{\dot{r}}{\dot{\theta}} =
+            \hat{r}^\prime= \left.\frac{\dot{r}}{\dot{\theta}}\right|_{r =
+                \hat{r}(\theta)} = 
             g(\theta,\hat{r},I)
@@
     + R(\theta,r_0,I)\). 
-    Note that \(\hat{r}(\theta,r_0,0) \equiv r_0\) for all \(\theta\) because
+    Note that \(\hat{r}(\theta,r_0,0) \equiv r_0\) because
     \(g(\theta,\hat{r},0) \equiv 0\).
@@
     quadrature, giving
-    \[
-        \hat{r}_1(\theta,r_0) = \int\limits_0^\theta b(\sigma,r_0)d\sigma
-        .
-    \]
+    \(\hat{r}_1(\theta,r_0) = \int_0^\theta b(\sigma,r_0)d\sigma\).
+%    \[
+%        \hat{r}_1(\theta,r_0) = \int\limits_0^\theta b(\sigma,r_0)d\sigma
+%        .
+%    \]
 %    with \(b(\theta,r_0) := (\partial g/\partial I)(\theta,r_0,0)\). 
@@
     \begin{equation}\label{eqn:r-discrete}
-        r_{n+1} = r_n + IS(r_n) + R(2\pi,r_n,I)
+        r_{n+1} = P(r_n) =  r_n + IS(r_n) + R(2\pi,r_n,I)
         ,
@@
     Taking this fact alongside the assumption that \(S(r_0) \geq \epsilon\),
-    we find a lower bound on \(r_{n+1}\):
-    \[
-        r_{n+1} \geq r_n + I\epsilon - \kappa I^2
-        .
-    \]
+    we find a lower bound \(r_{n+1} \geq r_n + I\epsilon - \kappa I^2\).
     Choosing \(\gamma > 0\) and \(I^\star \in \, ]0,I_2]\) small enough makes 
-    \(I\epsilon - \delta I^2 \geq \gamma\) for all \(I \in \, ]0,I^\star]\).
+    \(I\epsilon - \kappa I^2 \geq \gamma\) for all \(I \in \, ]0,I^\star]\).
     Therefore, \(r_{n} \geq r_0 + n\gamma\), 
@@
 
-To prove Theorem \ref{thm:acrobot-energy-stabilization}, we will explicitly
-compute the pseudo-radius coordinates for \(\mathcal{O}_1\) and
-\(\mathcal{O}_2(\rho)\) and apply Lemma \ref{lemma:energy-regulation-D}.
-
 \subsection{Energy Gain on \(\mathcal{O}_1\)}\label{sec:proof-o1}
 To prove the first part of Theorem
-\ref{thm:acrobot-energy-stabilization},
-we begin by studying the punctured open set
-\(\mathcal{O}_1 \backslash \{(0,0)\} \subset \SxR\) consisting of all
-oscillations of the nominal pendulum.
-Using the procedure from \cite{dynamic_vhcs_stabilize_closed_orbits} for finding
-pseudo-polar coordinates adapted to level sets of \(E\), we compute
+\ref{thm:acrobot-energy-stabilization}, we begin by using the procedure from
+\cite{dynamic_vhcs_stabilize_closed_orbits} to find 
+pseudo-polar coordinates (adapted to level sets of \(E\)) on
+the punctured open set \(\mathcal{O}_1 \backslash \{(0,0)\} \subset \SxR\).
+The resulting coordinate transformation is
 \begin{align}
@@
 where \((r,\theta) \in \, ]0,\pi[ \times \Sone\).
-This has inverse
+One can verify that the inverse to this transformation is
 \begin{align*}
@@
 \sin(\theta)\), \(c_r := \cos(r)\), and \(s_r := \sin(r)\).
-We now verify these coordinates satisfy the requirements of 
+We now verify these pseudo-polar coordinates satisfy the requirements of 
 Lemma \ref{lemma:energy-regulation-D}.
@@
 \(r \in \, ]0,\pi[\).
-From \eqref{eqn:theta-dot-nom-osc}-\eqref{eqn:theta-dot-limit},
+From \eqref{eqn:theta-dot-nom-osc}--\eqref{eqn:theta-dot-limit},
 one can verify that \(\dot{\theta} > 0\) when \(I = 0\).
@@
 Next we compute
-\(b(\theta,r_0) = (\partial g/\partial I)(\theta, r_0, 0) = K a(\theta,r_0)\),
+\(b(\theta,r_0) := (\partial g/\partial I)(\theta, r_0, 0) = K a(\theta,r_0)\),
 where
@@
 
-Unfortunately there is no \(\delta\) such that \(Q(r_0) \geq \delta\)
+Unfortunately there is no \(\delta > 0\) such that \(Q(r_0) \geq \delta\)
 everywhere on \(]0,\pi[\), though such a \(\delta\) does exist 
@@
 proof of Lemma \ref{lemma:energy-regulation-D}. 
-To prove the origin in the \((q_u,p_u)\)-plane is a repeller, we need simply
+To prove the origin is a repeller, we need simply
 prove the linearization of \eqref{eqn:r-discrete} is greater than \(1\) at 
 \(r_0 = 0\). 
-Linearizing these discrete-time dynamics is equivalent to linearizing the
-Poincar\'{e} map \eqref{eqn:poincare-map}, which yields 
+This linearization yields
 \[
@@
 \]
-where prime denotes differentiation with respect to \(r\).
+where the prime now denotes differentiation with respect to \(r\).
 MATLAB symbolic computations reveal that
-\begin{equation*}
-    \lim \limits_{r_0 \to 0^+}
-    \pdiff{a}{r}(\theta,r_0) 
-    = -\frac{\sqrt{2}}{2} \left(11\sin(\theta)^2 - 6\right)
-    ,
-\end{equation*}
-which means
+%\begin{equation*}
+%    \lim \limits_{r_0 \to 0^+}
+%    \pdiff{a}{r}(\theta,r_0) 
+%    = -\frac{\sqrt{2}}{2} \left(11\sin(\theta)^2 - 6\right)
+%    ,
+%\end{equation*}
+%which means
 \begin{align*}
-    Q^\prime(0) = \int \limits_0^\pi \pdiff{a}{r}(\sigma,0) d\sigma
+    Q^\prime(0) = \int \limits_0^\pi 
+    \lim \limits_{r_0 \to 0^+}\pdiff{a}{r}(\sigma,r_0) d\sigma
     = \frac{\pi}{2\sqrt{2}} > 0
@@
 \end{align*}
-Since the remainder term \(R\) is \(O(I^2)\), its partial derivative 
-\(R^\prime\) is also \(O(I^2)\).
+Since \(R(2\pi,0,I)\) is \(O(I^2)\), so is
+\(R^\prime(2\pi,0,I)\).
 Hence, it can be written in the form
@@
 The origin is therefore a repeller, which means there exists some
-(unknown) \(\mu > 0\) where the interval \(]0,\mu]\) is negatively
-invariant for \eqref{eqn:r-discrete}, and all solutions starting in this
+(unknown) \(\mu > 0\) where \(]0,\mu]\) is negatively
+invariant for \eqref{eqn:r-discrete}, and all solutions in this
 interval flow through \(r = \mu\).
-This in turn implies that all orbits starting in \(\mathcal{O}_1\) near the
+Hence, all orbits in \(\mathcal{O}_1\) near the
 origin will gain energy up to \(E(\mu,0)\).
@@
 \(\mathcal{O}_1\) for each \(I \in \, ]0,I^\star]\).
-If \(I < 0\), then \(P^\prime(0) < 1\) which implies the origin is an
+If \(I < 0\), then \(P^\prime(0) < 1\) and the origin is an
 asymptotically stable equilibrium.
-The acrobot now loses energy on \(D_\mu\), and all orbits in \(\mathcal{O}_1\)
-lose energy for \(I \in [-I^\star,0[\).
-
-We have succeeded in proving that any acrobot constrained by 
+Since the acrobot now loses energy on \(D_\mu\), it also loses energy 
+on \(\mathcal{O}_1\) for \(I \in [-I^\star,0[\).
+
+We have proved that any acrobot constrained by 
 our VNHC \eqref{eqn:acrobot-constraint} gains/loses energy on \(\mathcal{O}_1\).
-Before moving on to the second part of Theorem
-\ref{thm:acrobot-energy-stabilization}, we must first confirm that the
-acrobot will eventually start rotating:
+We must now confirm that the acrobot will start rotating:
 the definition of energy gain states that all orbits initialized in
 \(\mathcal{O}_1\) will escape compact subsets of \(\mathcal{O}_1\) in finite
-time, which is not enough to conclude that all orbits will escape the boundary
-of \(\mathcal{O}_1\).
+time, which is not enough to conclude that these orbits will also escape the
+boundary of \(\mathcal{O}_1\) in finite time.
 
@@
 outside this set.
-It is possible for orbits starting in \(\mathcal{O}_1\) to always approach
-\(E_\pi\) without ever crossing into the rotation zone.
-We will prove this does not happen in general.
 
@@
 \]
-By Descartes' rule of signs, this polynomial has one root with positive real
-part. 
-The equilibrium \((\pi,0)\) is therefore unstable, so the stable manifold \(\Pi^+\) of
-initial conditions converging to \((\pi,0)\) is one-dimensional, and hence is
-of measure zero in \(\SxR\).
+By Descartes' rule of signs, this polynomial always has exactly one root with
+positive real part. 
+The equilibrium \((\pi,0)\) is therefore unstable, so the stable manifold
+\(\Pi^+\) of initial conditions converging to \((\pi,0)\) is one-dimensional,
+and hence is of measure zero in \(\SxR\).
 
@@
 \(\theta \in \{0,\pi\}\).
-Thus, if \(E_\pi\) is not invariant, the largest invariant subset is the upright
-equilibrium \((q_u,p_u) = (\pi,0)\).
+Thus, the largest invariant subset of \(E_\pi\) is either the upright
+equilibrium \(\{(\pi,0)\}\) or \(E_\pi\) itself.
 
@@
 \(\dot{E}(0,p_u) = g/(5l) \sin(q_a)p_u\).
-This is non-zero everywhere except at the origin, which means it is non-zero on
-\(E_\pi\); this implies \(E_\pi\) is not invariant, so the positive limit set of 
-\(x(t)\) must be the set \(\{(\pi,0)\}\). Hence, \(x(t)\) converges to 
-\((\pi,0)\), so \(x(0) \in \Pi^+\).
+This is non-zero on \(E_\pi\), so \(E_\pi\) is not invariant and the positive
+limit set of \(x(t)\) must be the set \(\{(\pi,0)\}\). 
+Hence, \(x(t)\) converges to \((\pi,0)\), so \(x(0) \in \Pi^+\).
 Since \(\Pi^+\) is a set of measure zero in \(\SxR\), almost every orbit
-initialized in \(\mathcal{O}_1\) must escape through \(E_\pi\) into the rotation
-domain.
-
-Escaping once into the rotation domain does not guarantee the acrobot
-orbit will never return to \(\mathcal{O}_1\).
-Indeed, the orbit could escape through \(E_\pi\) and return to \(\mathcal{O}_1\)
-several times.
-However, almost every orbit must eventually remain outside the closure of 
-\(\mathcal{O}_1\). 
+initialized in \(\mathcal{O}_1\) must escape at least once through \(E_\pi\)
+into the rotation domain.
+Moreover, almost every orbit must eventually remain outside
+\(\bar{\mathcal{O}}_1\).
 If not, the positive limit set would once again be
@@
     \label{eqn:theta-rot}
-    \theta &= \pm q_u
+    \theta &= \alpha(D) q_u
 \end{align}
-where \((r,\theta) \in \, ]\sqrt{60m^2gl^3}, \rho[ \times \Sone\).
+where \((r,\theta) \in \, ]\sqrt{60m^2gl^3}, \rho[ \times \Sone\),
+and \(\alpha(D) = 1\) if \(D = \mathcal{R}^+(\rho)\)
+or \(-1\) if \(D = \mathcal{R}^-(\rho)\).
 The inverse of this transformation is
 \begin{align*}
-    q_u &= \pm \theta
+    q_u &= \alpha(D) \theta
     , \\
-    p_u &= \pm \sqrt{r^2-30m^2gl^3(1-c_\theta)}
+    p_u &= \alpha(D) \sqrt{r^2-30m^2gl^3(1-c_\theta)}
     .
 \end{align*}
-In this context, \(\pm\) denotes whether \(D=\mathcal{R}^+(\rho)\) or
-\(D=\mathcal{R}^-(\rho)\).
 
@@
 so there exists a control parameter \(I_2(\rho) > 0\) whereby the acrobot
-gains energy on both \(\mathcal{R}^+\) and \(\mathcal{R}^-\) for
+gains energy on both \(\mathcal{R}^+(\rho)\) and \(\mathcal{R}^-(\rho)\) for
 each \(I \in \, ]0,I_2(\rho)]\).
 Let \(I_1^\star\) be the value we found for \(\mathcal{O}_1\) in 
-Section \ref{sec:proof-o1}.
-Taking \(I^{\star\star} = \min \{I_1^\star, I_2(\rho)\}\),
-the acrobot gains/loses energy on \(\mathcal{O}_2(\rho)\)
+Section \ref{sec:proof-o1}, and let
+\(I^{\star\star} = \min \{I_1^\star, I_2(\rho)\}\).
+Then the acrobot gains/loses energy on \(\mathcal{O}_2(\rho)\)
 for each \(I \in \,]0, I^{\star\star}]\).
@@
 \(\mathcal{O}_2(\rho)\).
+
 This completes the proof of Theorem \ref{thm:acrobot-energy-stabilization}.

### 2021-06-10 latex/paper/main.tex
Began new edits using paper_v3, currently describing rotations and oscillations.

@@
 the legs. 
-The acrobot is actuated exclusively at the center joint (the hips).
+The acrobot is actuated exclusively at the center joint (the hip).
 Controlling the acrobot is a nontrivial task because it is not feedback
@@
 
-In this article, we will design a virtual nonholonomic constraint which provably
+In this article, we design a virtual nonholonomic constraint which provably
 injects energy into the acrobot through human-like giant motion.
@@
 \end{defn}
-
+The definition above applies to vector fields that are a not necessarily
+Hamiltonian, and therefore might not have an associated notion of ``energy". 
+The terminology " gain energy" or "lose energy" is to be interpreted loosely in
+this context.
 Our goal is to design a smooth function \(f : \Sone \times \R\) such that the
@@
 \ref{defn:energy-gain}. 
-
 Note that any system satisfying Definition \ref{defn:energy-gain}
@@
 call ``simply actuated hamiltonian systems". 
-Note that the results we provide in Section \ref{sec:vnhc-vnhc} are not novel:
+The results we provide in Section \ref{sec:vnhc-vnhc} are not novel:
 they are a special case of the results in
@@
 \begin{assm}\label{assm:B-const}
-    The input matrix \(B(q) \equiv B \in \R^{n\times k}\) is constant,
-    full rank, and \(k < n\).
+    The input matrix \(B(q) \equiv B \in \R^{n\times k}\) is constant
+    and has full rank \(k < n\).
 \end{assm}
@@
 In particular, if \(k = (n-1)\), the constraint manifold is \textit{always}
-2-dimensional and its dynamics can be plotted on a plane.
+2-dimensional and its dynamics can be displayed on a plane.
 
@@
      \end{equation}
-    is full rank \(k\) everywhere on the constraint manifold \(\Gamma\).
+    is invertible everywhere on the constraint manifold \(\Gamma\).
 \end{thm}
@@
 %/========== Acrobot ==========/%
-\section{The Acrobot VNHC}\label{sec:acrobot}
+\section{The Proposed Acrobot VNHC}\label{sec:acrobot}
 
@@
     with \(f \in C^2\left(\R; \Sone\right)\) is a regular
-    VNHC of order 1 for the acrobot.
+    VNHC of order 1 for the acrobot in \eqref{eqn:acrobot-hamiltonian}.
 \end{prop}
@@
 
+\subsection{Energy Injection and Dissipation with the Proposed VNHC}
 Suppose for a moment that \(I = 0\) in \eqref{eqn:acrobot-constraint}, i.e.,
@@
 without gaining energy.
-Choosing \(I\) small enough guarantees the legs will synchronize
+However, choosing \(I\) small enough guarantees the legs will synchronize
 properly with the body, and the acrobot will begin rotating around the bar.
@@
         \(\mathcal{O}_1\) in finite time.
-        If instead \(I \in [-I^\star,0[\), the VNHC dissipates energy.
+        If instead \(I \in [-I^\star,0[\), the VNHC dissipates energy
+        on \(\mathcal{O}_1\).
     \item Let \(C = m^2gl^3\bar{q}_a\) and 
@@
         \(\mathcal{O}_2(\rho)\).
-        If instead \(I \in [-I^{\star\star},0[\), the VNHC dissipates energy.
+        If instead \(I \in [-I^{\star\star},0[\), the VNHC dissipates energy
+        on \(\mathcal{O}_2(\rho)\).
 \end{enumerate}
@@
 One can apply the results of Theorem \ref{thm:acrobot-energy-stabilization} 
-towards energy regulation;
-that is, one can stabilize oscillations or rotations by appropriately toggling
-between injection and dissipation VNHCs.
-
-\textbf{Oscillation Regulation}: one chooses a desired oscillation angle 
-\(q_\text{des} \in \, ]0,\pi[\) and, to avoid infinite switching, a
-hysteresis value \(\delta \in [0,1]\).
-Each time the orbit crosses the \(q_u\)-axis (i.e., when \(p_u = 0\)) or 
-when \(|q_u| = \pi\), the supervisor does the following: 
-\begin{itemize}
-    \item If \(|q_u| < (1-\delta)q_\text{des}\), enable the injection VNHC.
-    \item If \(|q_u| > (1+\delta)q_\text{des}\), enable the dissipation VNHC.
-    \item If \((1-\delta)q_\text{des} \leq |q_u| \leq (1+\delta)q_\text{des}\),
-        keep the leg extended at \(q_a = 0\). This can be done continously since
-        \(q_a = 0\) when \(p_u = 0\).
-\end{itemize}
-Here one must be careful to choose \(\delta\) so that
-\((1+\delta)q_\text{des} < \pi\), otherwise a rotating acrobot will not
-dissipate energy.
+towards energy regulation.
+That is, given a level set of \eqref{eqn:nominal-energy} contained in 
+\(\mathcal{O}_2(\rho)\), one can stabilize a neighbourhood of this level set by
+appropriately toggling between injection and dissipation VNHCs, which can be
+achieved by changing the sign of \(I\) in \eqref{eqn:acrobot-constraint}.
+
+In the rest of this article we distinguish between ``rotations" and
+``oscillations".
+
+%We say the acrobot is \textit{rotating} with \textit{rotation rate} \(p_0\)
+%whenever \((q_u(t),p_u(t))\) crosses the \(p_u\)-axis at \(|p_u| = p_0\), 
+%and \(p_u(t)\) has not changed sign since the previous time the orbit crossed
+%the \(p_u\)-axis.
+We say the acrobot is \textit{rotating} whenever \((q_u(t),p_u(t))\) crosses the 
+\(p_u\)-axis, and \(p_u(t)\) has not changed sign since the previous time the
+orbit crossed the \(p_u\)-axis.
+The value of \(|p_u|\) at this crossing is the \textit{rotation rate}. 
+If \(I = 0\) in \eqref{eqn:acrobot-constraint}, rotations are
+closed periodic orbits of \((q_u(t),p_u(t))\) which are entirely contained in
+\(\mathcal{R}(\rho)\), the rotation domain defined in
+\eqref{eqn:rotation-domain}.
+In this case, each rotation is contained in a level set of
+\eqref{eqn:nominal-energy}, which is uniquely determined by its rotation rate.
+\textbf{TODO: state that if abs(qu) = pi and pu neq 0, the acrobot must necessarily
+rotate}
+
+We say the acrobot is \textit{oscillating} if it is not rotating.
+Since \(p_u(t)\) changes sign, an oscillating acrobot must necessarily cross the
+\(q_u\)-axis at some point \(q \in \, ]-\pi,\pi[\).
+The value of \(|q|\) at this crossing is called the \textit{oscillation angle}.
+If \(I = 0\) in \eqref{eqn:acrobot-constraint}, oscillations are closed periodic
+orbits of \((q_u(t),p_u(t))\) which are entirely contained in \(\mathcal{O}_1\).
+In this case, each oscillation is contained in a level set of
+\eqref{eqn:nominal-energy} which is uniquely determined by its oscillation
+angle.
 
@@
 \end{itemize}
-While this method is titled ``rotation regulation", it is entirely possible to
-choose \(p_\text{des} < \sqrt{60m^2gl^3}\);
-this would stabilize an oscillation whose momentum at the bottom of the swing
-arc is \(p_\text{des}\).
+This method is entitled ``rotation regulation", which implies that one must choose
+\(p_\text{des}\) and \(\delta\) so that 
+\((1-\delta) p_\text{des} > \sqrt{60m^2gl^3}\).
+that way, when \(q_a = 0\), the resulting orbit is a rotation.
+In fact, it is entirely possible to choose \(p_\text{des}\) and \(\delta\) so
+that \((1+\delta)p_\text{des} \leq \sqrt{60m^2gl^3}\).
+Rather than stabilizing a rotation, the above method would stabilize an
+oscillation whose momentum at the bottom of the swing arc is \(p_\text{des}\).
+
+The problem with rotation regulation is that it assumes the supervisor can
+instantaneously transition from an injection/dissipation VNHC (with non-zero \(I\))
+to extending the leg by setting \(q_a = 0\).
+This is impossible because the transition is happening on the \(p_u\)-axis.
+If the supervisor changed the VNHC on the \(q_u\)-axis instead, 
+this transition to \(q_a = 0\) can be done continuously.
+Hence, we now present the preferred method for stabilizing oscillations.
+
+\textbf{Oscillation Regulation}: one chooses a desired oscillation angle 
+\(q_\text{des} \in \, ]0,\pi[\) and, to avoid infinite switching, a
+hysteresis value \(\delta \in [0,1]\).
+Each time the orbit crosses the \(q_u\)-axis (i.e., when \(p_u = 0\)) or 
+whenever the acrobot begins rotating (i.e., when \(|q_u| = \pi\) and \(p_u
+\neq 0\)), the supervisor does the following: 
+\begin{itemize}
+    \item If \(|q_u| < (1-\delta)q_\text{des}\), enable the injection VNHC by
+        setting \(I > 0\) in the VNHC \eqref{eqn:acrobot-constraint}.
+    \item If \(|q_u| > (1+\delta)q_\text{des}\), enable the dissipation VNHC by
+        setting \(I < 0\) in the VNHC \eqref{eqn:acrobot-constraint}.
+    \item If \((1-\delta)q_\text{des} \leq |q_u| \leq (1+\delta)q_\text{des}\),
+        keep the leg extended at \(q_a = 0\). This can be done continously since
+        \(q_a = 0\) when \(p_u = 0\).
+\end{itemize}
+Here one must be careful to choose \(\delta\) so that
+\((1+\delta)q_\text{des} < \pi\), so t
 

### 2021-06-11 latex/paper/main.tex
Final edit as per meeting with Manfredi, he will now rewrite the proof

--- 2021-06-07 latex/paper/main.tex
@@
 % TODO: remove onecolumn and draftcls formatting
-\documentclass[journal,twoside,web, onecolumn,draftcls]{ieeecolor}
+\documentclass[journal,twoside,web, twocolumn,draftcls]{ieeecolor}
 
@@
 %---------- Header ----------% 
-\newcommand*{\Title}{VNHC and Acrobot Title}
+\newcommand*{\Title}{From gymnastics to virtual nonholonomic constraints: energy
+injection, dissipation, and regulation for the acrobot}
 \markboth{\journalname, VOL. XX, NO. XX, XXXX 2021}
@@
 
+The definition above applies to vector fields that are a not necessarily
+Hamiltonian, and therefore might not have an associated notion of ``energy". 
+The terms ``gain energy" and ``lose energy" are to be interpreted loosely in
+this context.
 Our goal is to design a smooth function \(f : \Sone \times \R\) such that the
@@
 \begin{assm}\label{assm:B-const}
-    The input matrix \(B(q) \equiv B \in \R^{n\times k}\) is constant,
-    full rank, and \(k < n\).
+    The input matrix \(B(q) \equiv B \in \R^{n\times k}\) is constant
+    and has full rank \(k < n\).
 \end{assm}
 
-\begin{assm}\label{assm:B-perp}
-    There exists a right semi-orthogonal matrix 
-    \(B^\perp \in \R^{(n-k)\times n}\)
-    which is a left-annihilator for \(B\). 
-\end{assm}
-
-Note that Assumption \ref{assm:B-perp} requires the rows of \(B^\perp\) be unit vectors
-that are mutually orthogonal. 
-When \(k = (n-1)\), Assumption \ref{assm:B-perp} can be removed because it is
-automatically implied by Assumption \ref{assm:B-const}.
-
-The above assumptions allow us to define a
+%\begin{assm}\label{assm:B-perp}
+%    There exists a right semi-orthogonal matrix 
+%    \(B^\perp \in \R^{(n-k)\times n}\)
+%    which is a left-annihilator for \(B\). 
+%\end{assm}
+%
+%Note that Assumption \ref{assm:B-perp} requires the rows of \(B^\perp\) be unit vectors
+%that are mutually orthogonal. 
+%When \(k = (n-1)\), Assumption \ref{assm:B-perp} can be removed because it is
+%automatically implied by Assumption \ref{assm:B-const}.
+%
+When \(\mathcal{Q} = \R^n\), the above assumption allows us to define a
 canonical coordinate transformation of \eqref{eqn:hamiltonian} 
@@
 \(B\tpose\) is right semi-orthogonal.
-Now, let \(\mathbf{B} := [B^\perp; B\tpose]\).
-Since \(B^\perp\) is a left annihilator of \(B\) and both \(B^\perp\) and
-\(B\tpose\) are right semi-orthogonal, one can easily show that \(\mathbf{B}\) is
-an orthogonal matrix.
-
-We temporarily take \(\mathcal{Q}\) to be embedded in \(\R^N\) so that the
-transformation \(\tilde{q} = \mathbf{B}q\) is well-defined.
+Now, let \(\mathbf{B} := [B^\perp; B\tpose]\) where 
+\(B^\perp \in \R^{(n-k)\times k}\) is a left annihilator of \(B\), i.e.,
+\(B^\perp B = \Zmat{(n-k) \times k}\).
+Since \(B\) is constant, such a \(B^\perp\) exists and
+\(\mathbf{B}\) is invertible.
+
+The following theorem requires that \(\mathcal{Q} = \R^n\) so that the
+transformation \(\tilde{q} = \mathbf{B}q\) is well defined.
 
 \begin{thm}\label{thm:simply-actuated}
-    Take the Hamiltonian system \eqref{eqn:hamiltonian} and suppose
-    Assumptions \ref{assm:B-const} and \ref{assm:B-perp} hold.
+    Take the Hamiltonian system \eqref{eqn:hamiltonian} with configuration
+    manifold \(\mathcal{Q} = \R^n\) and suppose
+    Assumption \ref{assm:B-const} holds.
     The coordinate transformation
-    \(\left(\tilde{q} = \mathbf{B}q, \tilde{p} = \mathbf{B}p\right)\)
+    \(\left(\tilde{q} = (\mathbf{B}\tpose)\inv q, \tilde{p} = \mathbf{B}p\right)\)
     is a canonical transformation and the resulting dynamics are given by 
@@
     \(\Minv(\tilde{q}) := 
-    \mathbf{B}D^{-1}(\mathbf{B}\tpose \tilde{q})\mathbf{B}\tpose\)
+    (\mathbf{B}\tpose)\inv D^{-1}(\mathbf{B}\tpose \tilde{q})\mathbf{B}\inv\)
     and
@@
     and \([\tilde{p}_i,\tilde{p}_j]\) are both zero.
-    Then, since \(\mathbf{B}\) is orthogonal, 
-    \([\tilde{p}_i, \tilde{q}_j] = (\mathbf{B}_i)\tpose (\mathbf{B}\tpose)_j
+    Then, one can show that
+    \([\tilde{p}_i, \tilde{q}_j] = (\mathbf{B}\tpose)\inv_i (\mathbf{B}\tpose)_j
         = \delta_i^j\).
@@
     and the new Hamiltonian is
-    \(\mathcal{H}(\mathbf{B}\tpose \tilde{q}, \mathbf{B}\tpose \tilde{p})\).
+    \(\mathcal{H}(\mathbf{B}\tpose \tilde{q}, \mathbf{B}\inv \tilde{p})\).
     Finally, since \(\dot{\tilde{p}} = \mathbf{B} \dot{p}\), the input
@@
 We call the \((\tilde{q},\tilde{p})\) coordinates
-\textit{simply actuated coordinates}, and we call system
-\eqref{eqn:simple-hamiltonian} a \textit{simply actuated hamiltonian system}.
+\textit{simply actuated coordinates}, and we call any Hamiltonian system
+whose input matrix is \([\Zmat{(n-k)\times k}; \Id{k}]\) a 
+\textit{simply actuated hamiltonian system}.
 The first \((n-k)\) configuration variables in \(\tilde{q}\), labelled \(q_u\),
@@
 and \textit{actuated momenta}, respectively.
+
+In this article we focus on the acrobot which, despite having a configuration
+manifold which is not \(\R^n\), is already a simply actuated Hamiltonian system.
 
@@
 that is, one can stabilize oscillations or rotations by appropriately toggling
-between injection and dissipation VNHCs.
+between injection and dissipation VNHCs, which can be achieved by changing the
+sign of \(I\) in \eqref{eqn:acrobot-constraint}.
+The acrobot is \textit{oscillating} when it swings back and forth without
+performing a full revolution around the pivot point.
+Oscillations correspond to \(p_u\) changing sign periodically, and the peak
+torso angle (i.e., the point where an orbit crosses the \(q_u\)-axis)
+is called the \textit{oscillation angle}.
+Conversely, the acrobot is \textit{rotating} when its torso is performing full
+revolutions around the pivot, which corresponds mathematically to \(p_u\)
+maintaining the same sign.
+The momentum achieved by the acrobot at the bottom of the swing
+(i.e., the point where an orbit of rotation crosses the \(p_u\)-axis) is
+called the \textit{rotation rate}.
+
+\textbf{Rotation Regulation}: one chooses a desired rotation rate 
+\(p_\text{des} > 0\) and a hysteresis value \(\delta \in [0,1]\).
+Each time the orbit crosses the \(p_u\)-axis (i.e. when \(q_u = 0\)), the
+supervisor changes which VNHC is enforced as follows:
+\begin{itemize}
+    \item If \(|p_u| < (1-\delta)p_\text{des}\), enable the injection VNHC by
+        setting \(I > 0\) in \eqref{eqn:acrobot-constraint}.
+    \item If \(|p_u| > (1+\delta)p_\text{des}\), enable the dissipation VNHC by
+        replacing \(I > 0\) with \(-I < 0\) in \eqref{eqn:acrobot-constraint}.
+    \item If \((1-\delta)p_\text{des} \leq |p_u| \leq (1+\delta)p_\text{des}\),
+        extend the leg fully by setting \(q_a = 0\).
+        For these simulations, we assume this can be done instantaneously,
+        though in practice this is impossible.
+\end{itemize}
+Note that all orbits of the acrobot must cross the \(p_u\) axis, 
+so this method will stabilize a rotation even if the acrobot is initialized in
+\(\mathcal{O}_1\).
+Also, the above supervisor is designed to regulate a rotation rate, but it
+cannot regulate the rotation direction.
+
+The term ``rotation regulation" implies that one must choose
+\(p_\text{des}\) and \(\delta\) so that 
+\((1-\delta) p_\text{des} > \sqrt{60m^2gl^3}\).
+In fact, it is entirely possible to choose \(p_\text{des}\) and \(\delta\) so
+that \((1+\delta)p_\text{des} \leq \sqrt{60m^2gl^3}\).
+Rather than stabilizing a rotation, the supervisor would then stabilize an
+oscillation whose momentum at the bottom of the swing arc is \(p_\text{des}\).
+However, there is a preferred method for stabilizing oscillations
+which we outline next.
 
@@
 \(q_\text{des} \in \, ]0,\pi[\) and, to avoid infinite switching, a
-hysteresis value \(\delta \in [0,1]\).
-Each time the orbit crosses the \(q_u\)-axis (i.e., when \(p_u = 0\)) or 
-when \(|q_u| = \pi\), the supervisor does the following: 
+hysteresis value \(\delta \in [0,\pi - q_\text{des}]\).
+An orbit in the \((q_u,p_u)\)-plane will either cross the \(q_u\) axis if the
+orbit corresponds to a rocking motion, or it will cross the line \(|q_u| = \pi\)
+if the orbit corresponds to a full revolution.
+When either of these occur, the supervisor does the following: 
 \begin{itemize}
@@
 \end{itemize}
-Here one must be careful to choose \(\delta\) so that
-\((1+\delta)q_\text{des} < \pi\), otherwise a rotating acrobot will not
-dissipate energy.
-
-\textbf{Rotation Regulation}: one chooses a desired rotation rate 
-\(p_\text{des} > 0\) and a hysteresis value \(\delta \in [0,1]\).
-Each time the orbit crosses the \(p_u\)-axis (i.e. when \(q_u = 0\)), the
-supervisor changes which VNHC is enforced as follows:
-\begin{itemize}
-    \item If \(|p_u| < (1-\delta)p_\text{des}\), enable the injection VNHC.
-    \item If \(|p_u| > (1+\delta)p_\text{des}\), enable the dissipation VNHC.
-    \item If \((1-\delta)p_\text{des} \leq |p_u| \leq (1+\delta)p_\text{des}\),
-        extend the leg fully by setting \(q_a = 0\).
-        For these simulations, we assume this can be done instantaneously.
-\end{itemize}
-While this method is titled ``rotation regulation", it is entirely possible to
-choose \(p_\text{des} < \sqrt{60m^2gl^3}\);
-this would stabilize an oscillation whose momentum at the bottom of the swing
-arc is \(p_\text{des}\).
+Note that by our choice of \(\delta\), if the supervisor kicks in when 
+\(|q_u| = \pi\) (i.e., when the robot is rotating) then the supervisor will
+automatically enable the dissipation VNHC.
 
@@
 nominal pendulum.
-The points where the orbit exits \(E_\pi\) are marked with black stars,
-with the final departure marked by a red star.
+The points where the orbit exits \(E_\pi\) are marked with black asterisks,
+with the final departure marked by a red asterisk.
 Interestingly, our choice of \(I\) is large enough that we observe significant
@@
 We highlight the locations where the orbit crossed the set \(E_\pi\) by
-black stars, with the final crossing in red.
+black asterisks, with the final crossing in red.
 After this final crossing, the acrobot ceased rotating and its oscillations
@@
 \(5\%\) hysteresis, meaning \(\delta = 0.05\).
-The supervisor reevaluated its choice of VNHC at each black star;  
+The supervisor reevaluated its choice of VNHC at each black asterisk;  
 the red contour corresponds to the part of the orbit where the supervisor kept
@@
 In Figure \ref{fig:acrobot-osc-reg-in} the acrobot was initialized at 
-\((q_u,p_u) = (\pi/32,0)\); here the supervisor injected energy for most of the
-orbit.
+\((q_u,p_u) = (\pi/32,0)\); here the supervisor injected energy to stabilize the
+desired orbit.
 In Figure \ref{fig:acrobot-osc-reg-diss} the acrobot was initialized at the rotation
-\((q_u,p_u) = (0,0.19)\); here the supervisor dissipated energy. 
-In both cases, the supervisor stabilized the desired oscillation.
+\((q_u,p_u) = (0,0.19)\); note that the supervisor is first enabled at the line
+\(|q_u| = \pi\) and dissipates energy. Eventually the acrobot begins a rocking
+motion and the supervisor is enable again each time the acrobot hits the
+\(q_u\)-axis. 
+The supervisor continues to dissipate energy until it stabilizes the desired
+oscillation.
 
@@
 \(p_\text{des} = 0.19\) with a \(2\%\) hysteresis \(\delta = 0.02\).
-The supervisor evaluated its choice of VNHC at each black star.
+The supervisor evaluated its choice of VNHC at each black asterisk.
 Once it was within range of \(p_\text{des}\) it extended the legs completely,
@@
 Note the difference in shape between the blue rotations of the dissipation VNHC
-and the red rotation of the nominal pendulum: the red one slows down much more
+and the red rotation of the nominal pendulum in Figure
+\ref{fig:acrobot-rot-reg}: the red one slows down much more
 near \(|q_u| = \pi\).
@@
 
-We had hoped to show experimental results for dissipation and
-oscillation/rotation regulation. 
-Unfortunately the Dynamixel RX24F servo motor broke before we could perform
-these experiments.
-These motors are no longer in production at the time of writing this article, so
-SUGAR has been retired from service.
-
 \subsection{Summary of Results} 
@@
 Here is fundamental idea behind this proof:
-using a procedure devised by Mohammadi \etal
-\cite{dynamic_vhcs_stabilize_closed_orbits}, on certain subsets 
-\(D \subset \SxR\) one can find a diffeomorphism \(T(q_u,p_u) = (r,\theta)\)
+By means of a diffeomorphism \(T(q_u,p_u) = (r,\theta)\) we transform the
+constrained dynamics \eqref{eqn:acrobot-constrained-dynamics} 
 into pseudo-polar coordinates adapted to level sets of \(E(q_u,p_u)\).
@@
     \end{align*}
-    into pseudo-polar coordinates adapted to level sets of \(E(q_u,p_u)\),
     satisfying \(\dot{r} \equiv 0\) and \(\dot{\theta} > 0\) when
@@
 To prove the first part of Theorem
-\ref{thm:acrobot-energy-stabilization}, we begin by using the procedure from
-\cite{dynamic_vhcs_stabilize_closed_orbits} to find 
+\ref{thm:acrobot-energy-stabilization}, we adapt
+\cite[Eqn. (9)]{dynamic_vhcs_stabilize_closed_orbits} to find 
 pseudo-polar coordinates (adapted to level sets of \(E\)) on
@@
 one can verify that \(\dot{\theta} > 0\) when \(I = 0\).
-Defining \(g(\theta,r,I) = \dot{r}/\dot{\theta}\), MATLAB symbolic computations
-confirm that \((\partial g/\partial r)(\theta,r,0) \equiv 0\) as required.
+Defining \(g(\theta,r,I) = \dot{r}/\dot{\theta}\), one can show that
+\((\partial g/\partial r)(\theta,r,0) \equiv 0\) as required.
 
@@
 where the prime now denotes differentiation with respect to \(r\).
-MATLAB symbolic computations reveal that
+One can show that
 %\begin{equation*}
@@
 
-We now apply the method described in
-\cite{dynamic_vhcs_stabilize_closed_orbits} to find pseudo-polar coordinates
+We now apply
+\cite[Eqn. (8)]{dynamic_vhcs_stabilize_closed_orbits} to find pseudo-polar coordinates
 adapted to rotations.

### 2021-08-23 latex/paper/main.tex
Attempting to fix proof, made lots of edits but it still needs work.

@@
 often have vector relative degree \(\{2,\ldots,2\}\).
-Indeed, we shall now provide a characterization of regularity which shows that
-regular constraints cannot use the actuated momentum at all.
+Indeed, we shall now prove that regular constraints cannot depend on the
+actuated momentum.
 
@@
     \begin{equation}\label{eqn:decoupling-matrix}
-        \left(\mathcal{A}(q,p_u) - dh_{p_u}\mathcal{M}(q,p)\right)\simpleB
+        \big(\mathcal{A}(q,p_u) - dh_{p_u}\mathcal{M}(q,p)\big)\simpleB
          ,
@@
        \ddot{e} = (\star) +
-       \left(\mathcal{A}(q,p_u) - dh_{p_u}\mathcal{M}(q,p)\right) \simpleB \tau
+       \big(\mathcal{A}(q,p_u) - dh_{p_u}\mathcal{M}(q,p)\big) \simpleB \tau
         ,
@@
 Here, \eqref{eqn:g-qpu} reduces to
-\begin{equation*}
+\begin{multline*}
     g(q_u,p_u) = \frac{
-        (1+c_a)(1+I^2 p_u^2)p_u - m^2gl^3\bar{q}_a I (2-c_a^2)(2s_u + s_{ua})
-    }{ml^2(3+2c_a)(1+I^2 p_u^2)}
-    ,
-\end{equation*}
+    (1+c_a)p_u}{ml^2(3+2c_a)}
+    - \frac{mgl\bar{q}_a I (2-c_a^2)(2s_u + s_{ua})
+    }{(3+2c_a)(1+I^2 p_u^2)}
+    ,
+\end{multline*}
 and the constrained dynamics \eqref{eqn:qpu-dynamics} are
@@
 \(\mathcal{O}_1\), and the constrained dynamics escape
-\(\mathcal{O}_1\) in finite time, then the acrobot is guaranteed to perform
+\(\mathcal{O}_1\) and hit the vertical line \(|q_u| = \pi\)
+in finite time, then the acrobot is guaranteed to perform
 giant-like motion and begin rotating around the bar.
@@
         \(\mathcal{O}_1\).
-        Moreover, almost every orbit of the constrained dynamics 
-        \eqref{eqn:acrobot-constrained-dynamics} will escape the closure of
-        \(\mathcal{O}_1\) in finite time.
+        Moreover, every orbit of the constrained dynamics 
+        \eqref{eqn:acrobot-constrained-dynamics} will reach the line
+        \(|q_u| = \pi\) in finite time.
         If instead \(I \in [-I^\star,0[\), the VNHC dissipates energy
@@
 \(q_\text{des} \in \, ]0,\pi[\) and, to avoid infinite switching, a
-hysteresis value \(\delta \in [0,\pi - q_\text{des}]\).
+hysteresis value \(\delta \in [0,\pi/q_\text{des} - 1]\).
 An orbit in the \((q_u,p_u)\)-plane will either cross the \(q_u\) axis if the
@@
 differences between the nominal pendulum and the constrained dynamics:
-\(E_\pi\) intersects the \(p_u\)-axis at \(\abs{p_u} \approx 0.18\), yet the
+\(E_\pi\) intersects the \(p_u\)-axis at \(\abs{p_u} \approx 0.17\), yet the
 constrained acrobot begins rotating once it hits the
@@
 
-Communication between the Arduinos restricts the sampling rate of \(p\) to
+The communication speed between Arduinos restricts the sampling rate of \(p\) to
 500Hz.
@@
     \((q_u,p_u) \approx \left(\pi,0\right)\), let it run for 15
-    seconds, then introduced a rod as it passed through the bottom of
+    seconds, then introduced a rod as SUGAR passed through the bottom of
     its arc.
@@
 \item \textbf{Perturbation Test 2:}
-    to see how SUGAR responds when pushed, we allowed
+    to see how SUGAR would respond when pushed, we allowed
     it to rotate unperturbed for 15 seconds and then pushed it in its
@@
     We repeated this test by pushing SUGAR against its direction of motion.
-    The orbit in Figure \ref{fig:acrobot-rpush-orbit} demonstrates that it
+    The orbit in Figure \ref{fig:acrobot-rpush-orbit} demonstrates that SUGAR
     readily changes direction, and quickly achieves its maximum speed
@@
 Here is fundamental idea behind this proof:
-By means of a diffeomorphism \(T(q_u,p_u) = (r,\theta)\) we transform the
-constrained dynamics \eqref{eqn:acrobot-constrained-dynamics} 
-into pseudo-polar coordinates adapted to level sets of \(E(q_u,p_u)\).
-That is, when \(I = 0\) in \eqref{eqn:acrobot-constraint}, the pseudo-radius
+for each of \(\mathcal{O}_1\) or \(\mathcal{O}_2(\rho)\),
+we can find a diffeomorphism \(T(q_u,p_u) = (r,\theta)\) into pseudo-polar
+coordinates adapted to level sets of \(E(q_u,p_u)\);
+that is, when \(I = 0\) in \eqref{eqn:acrobot-constraint}, the pseudo-radius
 \(r\) chooses a level set of \(E\) and the pseudo-angle \(\theta\) chooses a
 point on that level set.
-We can formally show that \(r\) increases ``on average", which means \(E\)
-also increases on average and the constrained dynamics are gaining energy on
-\(D\).
-We formalize this idea in the following lemma.
+If the constrained dynamics in \((r,\theta)\) coordinates flow in such a way
+that \(r\) increases ``on average", then \(E\) must also increase ``on average",
+and the original constrained dynamics \eqref{eqn:acrobot-constrained-dynamics} 
+must be gaining energy on the domain of the diffeomorphism.
+
+The following lemma provides a set of conditions on the \((r,\theta)\) dynamics
+under which energy will increase on a domain \(D\); later in this section, we
+will assign \(D = \mathcal{O}_1\) or \(D = \mathcal{O}_2(\rho)\), find an
+appropriate transformation \(T\), and apply this lemma to prove the claims in
+Theorem \ref{thm:acrobot-energy-stabilization}.
 
@@
     \end{align*}
-    satisfying \(\dot{r} \equiv 0\) and \(\dot{\theta} > 0\) when
-    \(I = 0\).
-    Let \(g(\theta,r,I) = \dot{r}/\dot{\theta}\) and assume that 
-    \((\partial g/\partial r)(\theta,r,0) \equiv 0\).
-    Assume \(b(\theta,r_0) = (\partial g/\partial I)(\theta,r_0,0)\)
-    is \(2\pi\)-periodic in \(\theta\) for any fixed \(r_0\),
-    and that there exists \(\epsilon > 0\) such that 
+    yielding the new constrained dynamics \(\dot{r} = f_r(\theta,r,I)\),
+    \(\dot{\theta} = f_\theta(\theta,r,I)\).
+    Suppose that \(f_r(\theta,r,0) \equiv 0\) and \(f_\theta(\theta,r,0) > 0\).
+    Let \(g(\theta,r,I) = f_r(\theta,r,I)/f_\theta(\theta,r,I)\), and assume that 
+    \((\partial g/\partial r)(\theta,r,0) \equiv 0\) and
+    \(b(\theta,r) := (\partial g/\partial I)(\theta,r,0)\)
+    is \(2\pi\)-periodic in \(\theta\).
+    Assume also that there exists \(\epsilon > 0\) such that 
     \(S(r_0) := \int_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon\)
@@
         \begin{cases}
-            \hat{r}^\prime= \left.\frac{\dot{r}}{\dot{\theta}}\right|_{r =
-                \hat{r}(\theta)} = 
+            \hat{r}^\prime= \left.\frac{dr}{dt} \frac{dt}{d\theta}\right|_{r =
+                \hat{r}} = 
             g(\theta,\hat{r},I)
@@
     \end{equation}
-    where \(\theta\) now represents time,
-    the prime denotes differentiation by \(\theta\), and
+    with state \(\hat{r}\).
+    Here \(\theta\) represents time,
+    prime denotes differentiation by \(\theta\), and
     \(r_0 \in \, ]\underline{r},\overline{r}[\) is the initial condition.
@@
 
-    Our assumption that \((\partial g/\partial r)(\theta,r_0,0) \equiv 0\)
-    means that \((\partial g/\partial \hat{r})(\theta,r_0,0) \equiv 0\).
+    Since \((\partial g/\partial r)(\theta,r_0,0) \equiv 0\), we get
+    \((\partial g/\partial \hat{r})(\theta,r_0,0) \equiv 0\).
     The solution of \eqref{eqn:r1-hat-prime} can therefore be obtained by
@@
 %    with \(b(\theta,r_0) := (\partial g/\partial I)(\theta,r_0,0)\). 
-    Since and \(\dot{\theta} > 0\) and \(b(\theta,r_0)\) is \(2\pi\)-periodic in
+    Since \(b(\theta,r_0)\) is \(2\pi\)-periodic in
     \(\theta\), there is a well-defined Poincar\'{e} map describing
@@
 
-    If \(r_n \not\in \, ]\underline{r},\overline{r}[\), the orbit 
-    in \((q_u,p_u)\) coordinates must have
-    departed \(D\). % and \eqref{eqn:r-discrete} is no longer valid.
-    Proving the constrained dynamics gain energy on \(D\) is therefore
-    equivalent to finding \(N \in \mathbb{Z}_{> 0}\) where 
-    \(\hat{r}_{n} \geq \overline{r}\) for all \(n \geq N\).
+    If at any point in discrete time \(r_n\) exits \(]\underline{r},\overline{r}[\), the orbit 
+    in \((q_u,p_u)\) coordinates must have departed \(D\).
+    Hence, proving the constrained dynamics gain energy on \(D\) is
+    equivalent to finding \(N \in \mathbb{Z}_{> 0}\) such that 
+    \(r_{n} \geq \overline{r}\) for all \(n \geq N\).
     Using a perturbation analysis (see Khalil 
@@
     we know there exists some \(I_2 \in \, ]0,I_1]\) and \(\kappa > 0\)
-    such that \(\left|R(2\pi,r_n,I)\right| \leq \kappa I^2\) for all 
-    \(I \in [-I_2,I_2]\).
+    such that \(\left|R(\theta,r,I)\right| \leq \kappa I^2\) for all 
+    \(\theta \in [0,2\pi]\),
+    \(r \in \, ]\underline{r},\overline{r}[\),
+    and \(I \in [-I_2,I_2]\).
     Taking this fact alongside the assumption that \(S(r_0) \geq \epsilon\),
@@
 one can verify that \(\dot{\theta} > 0\) when \(I = 0\).
-Defining \(g(\theta,r,I) = \dot{r}/\dot{\theta}\), one can show that
-\((\partial g/\partial r)(\theta,r,0) \equiv 0\) as required.
+Defining \(g(\theta,r,I)\) as in Lemma \ref{lemma:energy-regulation-D}, one can
+verify that \((\partial g/\partial r)(\theta,r,0) \equiv 0\) as required.
 
 Next we compute
-\(b(\theta,r_0) := (\partial g/\partial I)(\theta, r_0, 0) = K a(\theta,r_0)\),
+\(b(\theta,r_0) = K a(\theta,r_0)\),
 where
@@
 In this article we applied the framework of virtual nonholonomic constraints
-on the acrobot, and designed a constraint which emulates giant motion from
+to the acrobot, and designed a constraint which emulates giant motion from
 gymnastics.
@@
 These demonstrated that virtual nonholonomic constraints are capable of
-injecting and dissipating energy in a robust manner, all while producing
+injecting and dissipating energy in a robust manner, while producing
 realistic biologically-inspired motion.

### 2021-08-24 latex/paper/main.tex
Started writing the segment on achieving rotations in the proof.

@@
 our VNHC \eqref{eqn:acrobot-constraint} gains/loses energy on \(\mathcal{O}_1\).
-We must now confirm that the acrobot will start rotating:
-the definition of energy gain states that all orbits initialized in
+We must now confirm that the acrobot will begin rotating, 
+i.e., that for \(I > 0\) almost every orbit initialized in \(\mathcal{O}_1\)
+will reach the line \(|q_u| = \pi\).
+Since the definition of energy gain only guarantees that orbits initialized in
 \(\mathcal{O}_1\) will escape compact subsets of \(\mathcal{O}_1\) in finite
-time, which is not enough to conclude that these orbits will also escape the
-boundary of \(\mathcal{O}_1\) in finite time.
-
-Let \(E_\pi\) be the level set with energy \(E(\pi,0)\), 
-which forms the boundary of \(\mathcal{O}_1\). 
+time, we must first prove that almost every orbit in \(\mathcal{O}_1\) will
+escape \(\mathcal{O}_1\) through its boundary, the level set \(E_\pi\) with
+energy \(E(\pi,0)\).
+
 Define \(\bar{\mathcal{O}}_1 = \mathcal{O}_1 \cup E_\pi\) to be the closure of
@@
 outside this set.
-
 Taking the Jacobian of the constrained dynamics
@@
 zone.
-Since \(\bar{\mathcal{O}}_1\) is compact, the Birkhoff Theorem \cite{birkhoff}
-implies:
-\begin{itemize}
-    \item The positive limit set \(L_+\) of \(x(t)\) is non-empty, compact, and
-        invariant.
-    \item The solution \(x(t)\) asymptotically tends to \(L_+\).
-\end{itemize}
-Since the acrobot gains energy on \(\mathcal{O}_1\), the positive limit set of
-\(x(t)\) must be the largest invariant subset of \(E_\pi\). 
+%Since \(\bar{\mathcal{O}}_1\) is compact, the Birkhoff Theorem \cite{birkhoff}
+%implies:
+%\begin{itemize}
+%    \item The positive limit set \(L_+\) of \(x(t)\) is non-empty, compact, and
+%        invariant.
+%    \item The solution \(x(t)\) asymptotically tends to \(L_+\).
+%\end{itemize}
+Since the acrobot gains energy on \(\mathcal{O}_1\) and \(\bar{\mathcal{O}}_1\)
+is compact, the Birkhoff Theorem \cite{birkhoff} implies that
+the solution \(x(t)\) asymptotically tends to the positive limit set, which is
+the largest invariant subset of \(E_\pi\).
 Taking the limit as \(r \to \pi\) of
@@
 into the rotation domain.
-Moreover, almost every orbit must eventually remain outside
-\(\bar{\mathcal{O}}_1\).
-If not, the positive limit set would once again be
-\(E_\pi\); by the same argument as above, that orbit would be in \(\Pi^+\).
+
+Using this fact, we now prove that almost every orbit in \(\mathcal{O}_1\) will
+cross the line \(|q_u| = \pi\).
+Let \(x(0) \in \mathcal{O}_1 \backslah \Pi^+\) and suppose that the orbit
+\(x([0,\infty[)\) does not hit the line \(|q_u| = \pi\).
+This orbit must remain in the compact set \(\Sone \times [-K,K]\) for some 
+\(K > 0\).
+Indeed, suppose otherwise.
+Then for arbitrary \(K > 0\), there exists some \(t_K > 0\) such that
+\(|p_u(t_K)| = K\), and there exists a \(t_{K/2}\) so that
+\(|p_u(t)| >= K/2\) for \(t \in [t_K, t_K + t_{K/2}]\).
+Looking at the constrained dynamics \eqref{eqn:acrobot-constrained-dynamics} in
+this timespan, we see that \(\dot{p}_u \geq -3mgl\) and
+\(\dot{q}_u \geq c_1 p_u + c_2 \geq c_1 K/2 + c_2\) for some constants
+\(c_1, c_2 > 0\).
+Picking \(K\) large enough guarantees that \(q_u\) covers a distance of \(2\pi\)
+over this time interval, and so the orbit hits the line \(|q_u| = \pi\)
+(thus providing a contradiction).
+
+Thus, the orbit \(x(t)\) remains within the compact set \(\Sone \times [-K,K]\).
+By the Birkhoff theorem, \(x(t)\) approaches a positive limit set, which is
+either a closed orbit or the equilibrium \((\pi,0)\).
+
+TODO: closed orbit case
 
 We conclude that almost all orbits beginning in \(\mathcal{O}_1\) will, in
-finite time, escape the closure of \(\mathcal{O}_1\) and remain in the rotation
-domain forever after.
+finite time, escape the closure of \(\mathcal{O}_1\) and begin rotating by
+crossing the vertical line \(|q_u| = \pi\).
 

### 2021-08-25 latex/paper/main.tex
Updated the proof, needs way more editing but it's now complete.

@@
     we find a lower bound \(r_{n+1} \geq r_n + I\epsilon - \kappa I^2\).
-    Choosing \(\gamma > 0\) and \(I^\star \in \, ]0,I_2]\) small enough makes 
-    \(I\epsilon - \kappa I^2 \geq \gamma\) for all \(I \in \, ]0,I^\star]\).
+    Let \(0 < I^\star < \min\{I_2, \epsilon/\kappa\}\).
+    Then for each \(I \in ]0,I^\star]\), the value 
+    \(\gamma := I\epsilon - \kappa I^2\) is strictly positive, so
+    \(P(r_0) \geq r_0 + \gamma\).
     Therefore, \(r_{n} \geq r_0 + n\gamma\), 
-    so all solutions of \eqref{eqn:r-discrete} eventually flow past
+    and all solutions of \eqref{eqn:r-discrete} eventually flow past
     \(\overline{r}\).
@@
 Hence, all orbits in \(\mathcal{O}_1\) near the
-origin will gain energy up to \(E(\mu,0)\).
+origin will gain energy up to (and including) \(E(\mu,0)\).
 
@@
 cross the line \(|q_u| = \pi\).
-Let \(x(0) \in \mathcal{O}_1 \backslah \Pi^+\) and suppose that the orbit
+Let \(x(0) \in \mathcal{O}_1\) and suppose that the orbit
 \(x([0,\infty[)\) does not hit the line \(|q_u| = \pi\).
-This orbit must remain in the compact set \(\Sone \times [-K,K]\) for some 
-\(K > 0\).
-Indeed, suppose otherwise.
+We claim that this orbit must remain in the compact set 
+\(\Sone \times [-K,K]\), for some \(K > 0\).
+By contradiction, suppose otherwise.
 Then for arbitrary \(K > 0\), there exists some \(t_K > 0\) such that
-\(|p_u(t_K)| = K\), and there exists a \(t_{K/2}\) so that
-\(|p_u(t)| >= K/2\) for \(t \in [t_K, t_K + t_{K/2}]\).
-Looking at the constrained dynamics \eqref{eqn:acrobot-constrained-dynamics} in
-this timespan, we see that \(\dot{p}_u \geq -3mgl\) and
-\(\dot{q}_u \geq c_1 p_u + c_2 \geq c_1 K/2 + c_2\) for some constants
-\(c_1, c_2 > 0\).
-Picking \(K\) large enough guarantees that \(q_u\) covers a distance of \(2\pi\)
-over this time interval, and so the orbit hits the line \(|q_u| = \pi\)
-(thus providing a contradiction).
-
-Thus, the orbit \(x(t)\) remains within the compact set \(\Sone \times [-K,K]\).
-By the Birkhoff theorem, \(x(t)\) approaches a positive limit set, which is
-either a closed orbit or the equilibrium \((\pi,0)\).
-
-TODO: closed orbit case
-
-We conclude that almost all orbits beginning in \(\mathcal{O}_1\) will, in
-finite time, escape the closure of \(\mathcal{O}_1\) and begin rotating by
-crossing the vertical line \(|q_u| = \pi\).
+\(|p_u(t_K)| = K\). 
+Looking at the constrained dynamics \eqref{eqn:acrobot-constrained-dynamics},
+we see that \(\dot{p}_u \geq -3mgl\) and
+\(\dot{q}_u \geq c_1 p_u + c_2\) for some constants \(c_1,c_2 > 0\).
+This implies there exists a \(t_{1/2} > 0\) (independent of \(K\)) 
+where
+\(|p_u(t)| >= K/2\) for \(t \in [t_K, t_K + t_{1/2}]\),
+and \(\dot{q}_u \geq c_1 K/2 + c_2\) on this interval.
+Picking \(K\) large enough guarantees \(q_u(t)\) covers a distance of \(2\pi\)
+over this time interval, and so the orbit must hit the line \(|q_u| = \pi\).
+Contradiction.
+
+Thus, the orbit must \(x(t)\) remain within the compact set 
+\(\Sone \times [-K,K]\).
+By the Birkhoff theorem, \(x(t)\) approaches a positive limit set \(L^+\), 
+which is either an invariant closed orbit which crosse the \(q_u\)-axis, 
+or the set \(\{(\pi,0)\}\).
+By contradiction, suppose \(L^+\) is a closed orbit which 
+hits the \(q_u\)-axis at \(\bar{q} < \pi\).
+Since \(]0,\mu]\) is negatively invariant for \eqref{eqn:r-discrete}, 
+we must have \(\bar{q} \in ]\mu,\pi[\).
+The Poincar\'{e} map \eqref{eqn:poincare-map} is smooth on \(]0,\pi[\), and its
+image contains \(\bar{q}\), so \(P^{-1}(\bar{q}) \neq \emptyset\).
+Take \(q^\prime \in P^{-1}(\bar{q})\). 
+If \(q^\prime \in ]0,\mu]\), then \(q^\prime < \bar{q}\).
+If \(q^\prime > \mu\), then we know 
+from the proof of Lemma \ref{lemma:energy-regulation-D} that
+\(P(q^\prime) = \bar{q} \geq q^\prime + \gamma\) for some \(\gamma > 0\), 
+so \(q^\prime \leq \bar{q} - \gamma < \bar{q}\).
+Since \(q^\prime < \bar{q}\), the orbit starting at \(q^\prime\) stays within
+\(\mathcal{O}_1\) and intersects the closed orbit at \(\bar{q}\).
+This violates uniqueness of solutions for the constrained dynamics,
+so \(L^+\) must be \(\{(\pi,0)\}\), which in turn implies that 
+\(x(0) \in \Pi^+\).
+
+We conclude that almost all orbits beginning in \(\mathcal{O}_1\) will
+escape the closure of \(\mathcal{O}_1\) and, in finite time, 
+begin rotating by crossing the vertical line \(|q_u| = \pi\).
 

### 2021-08-26 latex/paper/main.tex
Full edit of the proof. Needs one or two more passes, and we need to remove over one column of material.

@@
 Here is fundamental idea behind this proof:
-for each of \(\mathcal{O}_1\) or \(\mathcal{O}_2(\rho)\),
+for each of \(\mathcal{O}_1\) and \(\mathcal{O}_2(\rho)\),
 we can find a diffeomorphism \(T(q_u,p_u) = (r,\theta)\) into pseudo-polar
@@
 under which energy will increase on a domain \(D\); later in this section, we
-will assign \(D = \mathcal{O}_1\) or \(D = \mathcal{O}_2(\rho)\), find an
-appropriate transformation \(T\), and apply this lemma to prove the claims in
+will assign \(D \subset \mathcal{O}_1\) or \(D \subset \mathcal{O}_2(\rho)\),
+find the diffeomorphism \(T\), and apply this lemma to prove the claims in
 Theorem \ref{thm:acrobot-energy-stabilization}.
@@
     Let \(g(\theta,r,I) = f_r(\theta,r,I)/f_\theta(\theta,r,I)\), and assume that 
-    \((\partial g/\partial r)(\theta,r,0) \equiv 0\) and
-    \(b(\theta,r) := (\partial g/\partial I)(\theta,r,0)\)
-    is \(2\pi\)-periodic in \(\theta\).
-    Assume also that there exists \(\epsilon > 0\) such that 
+    \((\partial g/\partial r)(\theta,r,0) \equiv 0\).
+    Assume also that \(b(\theta,r) := (\partial g/\partial I)(\theta,r,0)\)
+    is \(2\pi\)-periodic in \(\theta\),
+    and that there exists \(\epsilon > 0\) such that 
     \(S(r_0) := \int_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon\)
@@
 \begin{proof}
-    Since \(\dot{\theta} > 0\) when \(I = 0\) and the constrained dynamics 
-    in \((r,\theta)\) coordinates are continuous in \(I\), there exists 
+    Since \(f_\theta(\theta,r,0) > 0\) and \(f_\theta\)
+    is continuous in \(I\), there exists 
     \(I_1 > 0\) small enough that \(\dot{\theta} > 0\) for all
@@
     \end{equation}
-    with state \(\hat{r}\).
-    Here \(\theta\) represents time,
+    with state \(\hat{r}\), where now
+    \(\theta\) represents time,
     prime denotes differentiation by \(\theta\), and
@@
     + R(\theta,r_0,I)\). 
-    Note that \(\hat{r}(\theta,r_0,0) \equiv r_0\) because
-    \(g(\theta,\hat{r},0) \equiv 0\).
-    The function \(\hat{r}_1(\theta,r_0)\) is the solution to
+    Here, \(\hat{r}(\theta,r_0,0) \equiv r_0\) because
+    \(g(\theta,\hat{r},0) \equiv 0\),
+    the function \(\hat{r}_1(\theta,r_0)\) is the solution to
     \begin{equation}\label{eqn:r1-hat-prime}
@@
 
-    Since \((\partial g/\partial r)(\theta,r_0,0) \equiv 0\), we get
-    \((\partial g/\partial \hat{r})(\theta,r_0,0) \equiv 0\).
-    The solution of \eqref{eqn:r1-hat-prime} can therefore be obtained by
+    Since \((\partial g/\partial \hat{r})(\theta,r_0,0) \equiv 0\),
+    the solution of \eqref{eqn:r1-hat-prime} can be obtained by
     quadrature, giving
@@
 %    with \(b(\theta,r_0) := (\partial g/\partial I)(\theta,r_0,0)\). 
-    Since \(b(\theta,r_0)\) is \(2\pi\)-periodic in
-    \(\theta\), there is a well-defined Poincar\'{e} map describing
+    Recall that \(b(\theta,r_0)\) is \(2\pi\)-periodic in
+    \(\theta\), so there is a well-defined Poincar\'{e} map describing
     how \(\hat{r}\) changes each time \(\theta\) increases by \(2\pi\).
@@
     with initial condition \(r_0\).
-    Here, \(r_n\) corresponds to the distance along the line segment
-    \(L := ]\underline{r},\overline{r}[ \, \times \{0\}\)
-    when an orbit of \((r,\theta)\) hits \(L\) for the the \(n^\text{th}\) time,
-    assuming the acrobot was initialized at \((r,\theta) = (r_0,0)\).
+%    Here, \(r_n\) corresponds to the distance along the line segment
+%    \(L := ]\underline{r},\overline{r}[ \, \times \{0\}\)
+%    when an orbit of \((r,\theta)\) hits \(L\) for the the \(n^\text{th}\) time,
+%    assuming the acrobot was initialized at \((r,\theta) = (r_0,0)\).
 
@@
     \(\overline{r}\).
+
     We conclude that, for \(I \in \, ]0, I^\star]\), the constrained dynamics gain
     energy on \(D\).
-
-    Changing the sign of \(I\) changes the direction of all the
+    Changing the sign of \(I\) changes the direction of the
     above inequalites, so that \(r_{n} \leq r_0 - n\gamma\). 
@@
 To prove the first part of Theorem
-\ref{thm:acrobot-energy-stabilization}, we adapt
+\ref{thm:acrobot-energy-stabilization}, we use
 \cite[Eqn. (9)]{dynamic_vhcs_stabilize_closed_orbits} to find 
-pseudo-polar coordinates (adapted to level sets of \(E\)) on
-the punctured open set \(\mathcal{O}_1 \backslash \{(0,0)\} \subset \SxR\).
+pseudo-polar coordinates adapted to level sets of \(E\) on
+\(\mathcal{O}_1 \backslash \{(0,0)\}\).
 The resulting coordinate transformation is
@@
 Lemma \ref{lemma:energy-regulation-D}.
-
-When \(I = 0\) we have \(\dot{r} = 0\) and
+We omit the expressions for \(f_r(\theta,r,I)\) and \(f_\theta(\theta,r,I)\), as
+they are too long for print.
+
+It is straightforward to show that
+\(f_r(\theta,r,0) \equiv 0\)
+and
 \begin{equation} \label{eqn:theta-dot-nom-osc}
-    \dot{\theta} = \sqrt{\frac{6g}{5l}} 
+    f_\theta(\theta,r,0) = \sqrt{\frac{6g}{5l}} 
         \sqrt{\frac{\cos(r c_\theta) - c_r}
@@
 \begin{equation}\label{eqn:theta-dot-limit}
-    \lim\limits_{\theta \to 0}\dot{\theta} 
-    = \lim\limits_{\theta \to \pi} \dot{\theta}
+    \lim\limits_{\theta \to 0}f_\theta(\theta,r,0)
+    = \lim\limits_{\theta \to \pi} f_\theta(\theta,r,0)
     = \sqrt{\frac{6g s_r}{10lr}}
@@
 From \eqref{eqn:theta-dot-nom-osc}--\eqref{eqn:theta-dot-limit},
-one can verify that \(\dot{\theta} > 0\) when \(I = 0\).
+one can verify that \(f_\theta(\theta,r,0) > 0\) as required.
 Defining \(g(\theta,r,I)\) as in Lemma \ref{lemma:energy-regulation-D}, one can
-verify that \((\partial g/\partial r)(\theta,r,0) \equiv 0\) as required.
-
-Next we compute
+also verify that \((\partial g/\partial r)(\theta,r,0) \equiv 0\).
+
+Symbolic computations reveal that
 \(b(\theta,r_0) = K a(\theta,r_0)\),
@@
 and \(2\pi\)-periodic in \(\theta\).
-This means that there exists \(\epsilon > 0\) such that 
+This means the integral of \(a(\theta,r_0)\) can be computed once for all
+acrobots; 
+hence, there exists \(\epsilon > 0\) such that 
 \(S(r_0) := \int_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon\)
 if and only if there exists \(\delta > 0\) such that 
-\(Q(r_0) := \int_0^{2\pi} a(\sigma,r_0) d\sigma \geq \delta\).
-Since \(a(\theta,r_0)\) is adimensional, we can numerically compute \(Q(r_0)\)
-once for all acrobots.
-The plot of \(Q(r_0)\) when \(r_0 \in [10^{-10}, \pi - 10^{-3}]\) is shown in
+\(Q(r_0) := \int_0^{2\pi} a(\sigma,r_0) d\sigma \geq \delta\)
+for every \(r_0 \in ]0,\pi[\).
+
+Unfortunately, this latter statement is not true.
+The plot of \(Q(r_0)\) for \(r_0 \in [10^{-10}, \pi - 10^{-3}]\) is shown in
 Figure \ref{fig:acrobot-Q}.
 We see that it is strictly positive and monotonically increasing, with an
-asymptote at \(r_0 = \pi\).
-Simulations with smaller \(r_0\) result in an infinite integral error due to a
-division by zero, yet we believe the \(Q(r_0)\) is in fact positive for all 
-\(r_0 \in \, ]0,\pi[\).
+asymptote at \(r_0 = \pi\), but \(Q(r_0) \to 0\) as \(r_0 \to 0\).
+
 \begin{figure}
@@
 
-Unfortunately there is no \(\delta > 0\) such that \(Q(r_0) \geq \delta\)
-everywhere on \(]0,\pi[\), though such a \(\delta\) does exist 
-on \(]\omega,\pi[\) for each fixed \(\omega > 0\).
-This means all conditions of Lemma \ref{lemma:energy-regulation-D}
-are satisfied on the annulus
-\[
-    D_\omega = \mathcal{O}_1 \backslash 
-    \{(q_u,p_u) \in \SxR \mid E(q_u,p_u) \leq E(\omega,0)\}
-    .
-\]
-Therefore, for each \(\omega \in \, ]0,\pi[\) there exists some 
-\(I_1(\omega) > 0\) such that the acrobot gains/loses energy everywhere on
-\(D_\omega\).
-To prove energy gain on all of \(\mathcal{O}_1\), we will show the origin is a
-repeller so that all orbits near the origin flow away from the origin into
-\(D_\omega\) (for small enough \(\omega\)) and continue gaining energy
-thereafter.
-
-Recall the discrete-time pseudo-radius dynamics \eqref{eqn:r-discrete} from the
-proof of Lemma \ref{lemma:energy-regulation-D}. 
-To prove the origin is a repeller, we need simply
-prove the linearization of \eqref{eqn:r-discrete} is greater than \(1\) at 
-\(r_0 = 0\). 
-This linearization yields
+Thankfully, we can still use Lemma \ref{lemma:energy-regulation-D} on an
+appropriate subset of \(\mathcal{O}_1\).
+To find this subset, first recall the discrete-time pseudo-radius dynamics
+\eqref{eqn:r-discrete} from the proof of Lemma \ref{lemma:energy-regulation-D}.
+Linearizing \eqref{eqn:r-discrete} at \(r_0 = 0\) yields
 \[
@@
 One can show that
-%\begin{equation*}
-%    \lim \limits_{r_0 \to 0^+}
-%    \pdiff{a}{r}(\theta,r_0) 
-%    = -\frac{\sqrt{2}}{2} \left(11\sin(\theta)^2 - 6\right)
-%    ,
-%\end{equation*}
-%which means
 \begin{align*}
@@
 Hence, it can be written in the form
-\(I^2 \tilde{R}(I)\) where \(\tilde{R}(I)\) is smooth and zero at \(I = 0\).
+\(R^\prime(2\pi,0,I) = I^2 \tilde{R}(I)\) where \(\tilde{R}(I)\) is smooth and
+zero at \(I = 0\).
 Thus, there exists \(I_0 > 0\) such that 
@@
 which implies \(P^\prime(0) > 1\).
-The origin is therefore a repeller, which means there exists some
+The origin is now a repeller, which means there exists some
 (unknown) \(\mu > 0\) where \(]0,\mu]\) is negatively
@@
 interval flow through \(r = \mu\).
-Hence, all orbits in \(\mathcal{O}_1\) near the
-origin will gain energy up to (and including) \(E(\mu,0)\).
-
-Since there exists \(I_1(\mu)\) so that the acrobot gains energy on \(D_\mu\),
-taking \(I^\star = \min\{I_0,I_1(\mu)\}\) means the acrobot will gain energy on
-\(\mathcal{O}_1\) for each \(I \in \, ]0,I^\star]\).
-If \(I < 0\), then \(P^\prime(0) < 1\) and the origin is an
-asymptotically stable equilibrium.
-Since the acrobot now loses energy on \(D_\mu\), it also loses energy 
-on \(\mathcal{O}_1\) for \(I \in [-I^\star,0[\).
-
-We have proved that any acrobot constrained by 
-our VNHC \eqref{eqn:acrobot-constraint} gains/loses energy on \(\mathcal{O}_1\).
-We must now confirm that the acrobot will begin rotating, 
-i.e., that for \(I > 0\) almost every orbit initialized in \(\mathcal{O}_1\)
-will reach the line \(|q_u| = \pi\).
-Since the definition of energy gain only guarantees that orbits initialized in
+If \(I < 0\), we have \(P^\prime(0) < 1\) and the origin is an
+asymptotically stable equilibrium for \eqref{eqn:r-discrete} instead.
+
+Let \(E_{\leq\mu}\) be the sublevel set of the energy \(E(\mu,0)\).
+Since there exists \(\delta_\mu > 0\) such that \(Q(r_0) \geq \delta_\mu\) 
+for \(r_0 \in [\mu,\pi[\), all conditions of Lemma
+\ref{lemma:energy-regulation-D} are valid on the annulus
+\(D_\mu = \mathcal{O}_1 \backslash E_{\leq\mu}\). 
+Hence, there exists some \(I_1(\mu) > 0\) such that the acrobot gains/loses
+energy everywhere on \(D_\mu\).
+Taking \(I^\star = \min\{I_0,I_1(\mu)\}\), the acrobot will gain energy on
+\(\mathcal{O}_1\) for \(I \in \, ]0,I^\star]\) and lose energy 
+for \(I \in [-I^\star,0[\).
+
+\subsection{All acrobots will rotate}
+
+We have proven that any acrobot constrained by 
+our VNHC \eqref{eqn:acrobot-constraint} will gain (or lose) energy on
+\(\mathcal{O}_1\).
+We must now prove that every acrobot will begin rotating, 
+i.e., that when \(I \in \, ]0,I^\star]\) almost every orbit initialized in
+\(\mathcal{O}_1\) will reach the line \(|q_u| = \pi\).
+
+The definition of energy gain only guarantees that orbits initialized in
 \(\mathcal{O}_1\) will escape compact subsets of \(\mathcal{O}_1\) in finite
-time, we must first prove that almost every orbit in \(\mathcal{O}_1\) will
-escape \(\mathcal{O}_1\) through its boundary, the level set \(E_\pi\) with
-energy \(E(\pi,0)\).
-
-Define \(\bar{\mathcal{O}}_1 = \mathcal{O}_1 \cup E_\pi\) to be the closure of
-\(\mathcal{O}_1\).
-An orbit of the acrobot will only begin rotating if it escapes
-\(\bar{\mathcal{O}}_1\) by crossing through \(E_\pi\), and eventually remains
-outside this set.
+time. 
+For an orbit of the acrobot to reach the line \(|q_u| = \pi\), it must first
+escape the closure \(\bar{\mathcal{O}}_1 = \mathcal{O}_1 \cup E_\pi\) through
+its boundary, the level set \(E_\pi\) with energy level \(E(\pi,0)\).
+
 Taking the Jacobian of the constrained dynamics
@@
 \]
-By Descartes' rule of signs, this polynomial always has exactly one root with
-positive real part. 
+This polynomial always has exactly one root with positive real part. 
 The equilibrium \((\pi,0)\) is therefore unstable, so the stable manifold
@@
 \(\Sone \times [-K,K]\), for some \(K > 0\).
-By contradiction, suppose otherwise.
-Then for arbitrary \(K > 0\), there exists some \(t_K > 0\) such that
+Suppose otherwise;
+then for arbitrary \(K > 0\), there exists some \(t_K > 0\) such that
 \(|p_u(t_K)| = K\). 
@@
 where
-\(|p_u(t)| >= K/2\) for \(t \in [t_K, t_K + t_{1/2}]\),
-and \(\dot{q}_u \geq c_1 K/2 + c_2\) on this interval.
+\(|p_u(t)| >= K/2\) 
+and \(\dot{q}_u \geq c_1 K/2 + c_2\) 
+for \(t \in [t_K, t_K + t_{1/2}]\).
 Picking \(K\) large enough guarantees \(q_u(t)\) covers a distance of \(2\pi\)
-over this time interval, and so the orbit must hit the line \(|q_u| = \pi\).
-Contradiction.
-
-Thus, the orbit must \(x(t)\) remain within the compact set 
-\(\Sone \times [-K,K]\).
+over this time interval, so the orbit would hit the line \(|q_u| = \pi\).
+
+If the orbit \(x(t)\) does not hit the line \(|q_u| = \pi\), we now know it must
+remain within the compact set \(\Sone \times [-K,K]\).
 By the Birkhoff theorem, \(x(t)\) approaches a positive limit set \(L^+\), 
-which is either an invariant closed orbit which crosse the \(q_u\)-axis, 
-or the set \(\{(\pi,0)\}\).
-By contradiction, suppose \(L^+\) is a closed orbit which 
-hits the \(q_u\)-axis at \(\bar{q} < \pi\).
+which is either the upright equilibrium \(\{(\pi,0)\}\), or
+a closed orbit not fully contained in \(\mathcal{O}_1\)
+which crosses the positive \(q_u\)-axis at some point \(\bar{q} < \pi\).
+Suppose \(L^+\) is the latter.
 Since \(]0,\mu]\) is negatively invariant for \eqref{eqn:r-discrete}, 
-we must have \(\bar{q} \in ]\mu,\pi[\).
+we must have \(\bar{q} > \mu\).
 The Poincar\'{e} map \eqref{eqn:poincare-map} is smooth on \(]0,\pi[\), and its
@@
 Take \(q^\prime \in P^{-1}(\bar{q})\). 
-If \(q^\prime \in ]0,\mu]\), then \(q^\prime < \bar{q}\).
-If \(q^\prime > \mu\), then we know 
+If \(q^\prime \in \, ]0,\mu]\), then \(q^\prime < \bar{q}\);
+if \(q^\prime > \mu\), then we know 
 from the proof of Lemma \ref{lemma:energy-regulation-D} that
@@
 so \(q^\prime \leq \bar{q} - \gamma < \bar{q}\).
-Since \(q^\prime < \bar{q}\), the orbit starting at \(q^\prime\) stays within
-\(\mathcal{O}_1\) and intersects the closed orbit at \(\bar{q}\).
+In either case, \(q^\prime < \bar{q}\), so the orbit starting at \(q^\prime\)
+stays within \(\mathcal{O}_1\) and intersects the closed orbit at \(\bar{q}\).
 This violates uniqueness of solutions for the constrained dynamics,
@@
 We conclude that almost all orbits beginning in \(\mathcal{O}_1\) will
-escape the closure of \(\mathcal{O}_1\) and, in finite time, 
-begin rotating by crossing the vertical line \(|q_u| = \pi\).
+escape the closure of \(\mathcal{O}_1\) and begin rotating by crossing the
+vertical line \(|q_u| = \pi\).
 
@@
 (where \(\mathcal{R}(\rho)\) is the rotation domain defined in
-\eqref{eqn:rotation-domain}), and since all orbits escape 
-\(\bar{\mathcal{O}_1}\) in finite time, we need simply prove energy gain on
+\eqref{eqn:rotation-domain}), and since all orbits in \(\mathcal{O}_1\) will
+begin rotating, we need simply prove energy gain on
 \(\mathcal{R}(\rho)\) to prove energy gain on \(\mathcal{O}_2(\rho)\).
-We separate \(\mathcal{R}(\rho)\) into its two distinct connected components:
+We separate \(\mathcal{R}(\rho)\) into its two separate connected components:
 \(\mathcal{R}^+(\rho)\), the set of rotations with \(p_u > 0\); 
@@
 We take the domain of interest for Lemma \ref{lemma:energy-regulation-D} to be
-\(D \in \{\mathcal{R}^+(\rho), \mathcal{R}^-(\rho)\}\).
+either of these rotation sets.
 
@@
     \label{eqn:theta-rot}
-    \theta &= \alpha(D) q_u
+    \theta &= \sign{p_u} q_u
 \end{align}
-where \((r,\theta) \in \, ]\sqrt{60m^2gl^3}, \rho[ \times \Sone\),
-and \(\alpha(D) = 1\) if \(D = \mathcal{R}^+(\rho)\)
-or \(-1\) if \(D = \mathcal{R}^-(\rho)\).
-The inverse of this transformation is
+with \(]\underline{r},\overline{r}[ \, = \, ]\sqrt{60m^2gl^3}, \rho[\).
+The inverse transformation is
 \begin{align*}
-    q_u &= \alpha(D) \theta
+    q_u &= \pm \theta
     , \\
-    p_u &= \alpha(D) \sqrt{r^2-30m^2gl^3(1-c_\theta)}
-    .
+    p_u &= \pm \sqrt{r^2-30m^2gl^3(1-c_\theta)}
+    ,
 \end{align*}
-
-In computing the constrained dynamics in
-\((r,\theta)\) coordinates and setting \(I = 0\), 
-we find that \(\dot{r} \equiv 0\) and
+where the \(\pm\) is \(+1\) on \(\mathcal{R}^+(\rho)\), or \(-1\) on
+\(\mathcal{R}^-(\rho)\).
+
+After computing the constrained dynamics in
+\((r,\theta)\) coordinates, one can verify that
+\(f_r(\theta,r,0) \equiv 0\) and
 \[
-\dot{\theta} = \frac{1}{5ml^2} \sqrt{r^2 - 30m^2gl^3(1-c_\theta)}
+    f_\theta(\theta,r,0) = \frac{1}{5ml^2} \sqrt{r^2 - 30m^2gl^3(1-c_\theta)}
 ,
@@
 which is strictly positive as required.
-Defining \(g(\theta,r,I) = \dot{r}/\dot{\theta}\), symbolic computations
+Symbolic computations
 also confirm that \((\partial g/\partial r)(\theta,r,0) \equiv 0\).
@@
 in the statement of Theorem \ref{thm:acrobot-energy-stabilization}, and by
-assumption we have \(S(r_0) \geq \epsilon\) on
-\(]\sqrt{60m^2gl^3},\rho[\).
-
-All requirements of Lemma \ref{lemma:energy-regulation-D} are satisfied,
-so there exists a control parameter \(I_2(\rho) > 0\) whereby the acrobot
-gains energy on both \(\mathcal{R}^+(\rho)\) and \(\mathcal{R}^-(\rho)\) for
-each \(I \in \, ]0,I_2(\rho)]\).
-Let \(I_1^\star\) be the value we found for \(\mathcal{O}_1\) in 
+assumption we have \(S(r_0) \geq \epsilon\).
+
+Thus, all requirements of Lemma \ref{lemma:energy-regulation-D} are satisfied,
+so there exists \(I_2(\rho) > 0\) where the acrobot
+gains energy for \(I \in \, ]0,I_2(\rho)]\)
+(on both \(\mathcal{R}^+(\rho)\) and \(\mathcal{R}^-(\rho)\)).
+Let \(I^\star\) be the value we found for \(\mathcal{O}_1\) in 
 Section \ref{sec:proof-o1}, and let
-\(I^{\star\star} = \min \{I_1^\star, I_2(\rho)\}\).
-Then the acrobot gains/loses energy on \(\mathcal{O}_2(\rho)\)
-for each \(I \in \,]0, I^{\star\star}]\).
-If instead \(I \in [-I^{\star\star},0[\), the acrobot dissipates energy on
-\(\mathcal{O}_2(\rho)\).
+\(I^{\star\star} = \min \{I^\star, I_2(\rho)\}\).
+Then the acrobot gains energy on \(\mathcal{O}_2(\rho)\)
+for \(I \in \,]0, I^{\star\star}]\),
+and dissipates energy for \(I \in [-I^{\star\star},0[\).
 

### 2021-08-27 latex/paper/main.tex
Second pass of the first part of the proof of Theorem 4

@@
 In this section we prove Theorem \ref{thm:acrobot-energy-stabilization},
-which claims the acrobot gains/loses energy on \(\mathcal{O}_1\) and
-\(\mathcal{O}_2(\rho)\).
-These are the sublevel sets \eqref{eqn:oscillation-domain} and
-\eqref{eqn:o-rhobar} of the nominal pendulum's energy \(E(q_u,p_u)\) defined in
-\eqref{eqn:nominal-energy}.
-
-Here is fundamental idea behind this proof:
+which claims the acrobot constrained by the VNHC \eqref{eqn:acrobot-constraint}
+will gain or lose energy on \(\mathcal{O}_1\) and \(\mathcal{O}_2(\rho)\) for
+``small enough" values of \(I\).
+%These are the sublevel sets \eqref{eqn:oscillation-domain} and
+%\eqref{eqn:o-rhobar} of the nominal pendulum's energy \(E(q_u,p_u)\) defined in
+%\eqref{eqn:nominal-energy}.
+
+Here is intuition behind this proof:
 for each of \(\mathcal{O}_1\) and \(\mathcal{O}_2(\rho)\),
-we can find a diffeomorphism \(T(q_u,p_u) = (r,\theta)\) into pseudo-polar
-coordinates adapted to level sets of \(E(q_u,p_u)\);
+we can find a diffeomorphism from \((q_u,p_u)\) into pseudo-polar
+coordinates \((r,\theta)\) adapted to level sets of \(E(q_u,p_u)\);
 that is, when \(I = 0\) in \eqref{eqn:acrobot-constraint}, the pseudo-radius
@@
 point on that level set.
-If the constrained dynamics in \((r,\theta)\) coordinates flow in such a way
-that \(r\) increases ``on average", then \(E\) must also increase ``on average",
-and the original constrained dynamics \eqref{eqn:acrobot-constrained-dynamics} 
-must be gaining energy on the domain of the diffeomorphism.
-
-The following lemma provides a set of conditions on the \((r,\theta)\) dynamics
-under which energy will increase on a domain \(D\); later in this section, we
-will assign \(D \subset \mathcal{O}_1\) or \(D \subset \mathcal{O}_2(\rho)\),
-find the diffeomorphism \(T\), and apply this lemma to prove the claims in
-Theorem \ref{thm:acrobot-energy-stabilization}.
+We will show that \(r(t)\) increases ``on average", 
+which implies \(E\) also increases ``on average",
+and that the original constrained dynamics \eqref{eqn:acrobot-constrained-dynamics} 
+are gaining energy on the domain of the diffeomorphism.
+
+The following lemma provides a set of conditions on the \((r,\theta)\)-dynamics
+under which energy will increase on a domain \(D \subset \SxR\); later in this section, we
+will assign \(D \subset \mathcal{O}_1\) or \(D \subset \mathcal{O}_2(\rho)\)
+and apply this lemma to prove the claims in Theorem
+\ref{thm:acrobot-energy-stabilization}.
 
@@
         P : \, ]\underline{r},\overline{r}[ &\to \mathbb{R}_{\geq 0} \\
-        r_0 &\mapsto \hat{r}(2\pi,r_0,I)
+        r_0 &\mapsto \hat{r}(2\pi,r_0,I) \nonumber
         ,
@@
 
-    If at any point in discrete time \(r_n\) exits \(]\underline{r},\overline{r}[\), the orbit 
-    in \((q_u,p_u)\) coordinates must have departed \(D\).
+    If \(r_{n+1}\in \, ]\underline{r},\overline{r}[\), the orbit 
+    \((q_u(t),p_u(t))\) starting at \(T^{-1}(r_n,0)\) remained in \(D\) as
+    \(\theta\) increased by \(2\pi\) and
+    landed at \(T^{-1}(r_{n+1},2\pi)\); otherwise, this orbit exited
+    \(D\)  before \(\theta\) reached \(2\pi\).
     Hence, proving the constrained dynamics gain energy on \(D\) is
@@
     Let \(0 < I^\star < \min\{I_2, \epsilon/\kappa\}\).
-    Then for each \(I \in ]0,I^\star]\), the value 
+    Then for each \(I \in \, ]0,I^\star]\), the value 
     \(\gamma := I\epsilon - \kappa I^2\) is strictly positive, so
@@
 To prove the first part of Theorem
-\ref{thm:acrobot-energy-stabilization}, we use
+\ref{thm:acrobot-energy-stabilization}, we apply
 \cite[Eqn. (9)]{dynamic_vhcs_stabilize_closed_orbits} to find 
@@
 Lemma \ref{lemma:energy-regulation-D}.
-We omit the expressions for \(f_r(\theta,r,I)\) and \(f_\theta(\theta,r,I)\), as
-they are too long for print.
+Note: we omit the expressions for \(f_r(\theta,r,I)\) and
+\(f_\theta(\theta,r,I)\) as they are too long for print.
 
@@
 \end{equation}
-Note that ~\eqref{eqn:theta-dot-nom-osc} has removable singularities at
+Observe that ~\eqref{eqn:theta-dot-nom-osc} has removable singularities at
 \(\theta \in \{0,\pi\}\).
@@
 \(Q(r_0) := \int_0^{2\pi} a(\sigma,r_0) d\sigma \geq \delta\)
-for every \(r_0 \in ]0,\pi[\).
+for every \(r_0 \in \,]0,\pi[\).
 
@@
 \(I K Q^\prime(0) + I^2 \tilde{R}(I) > 0\) for all \(I \in \, ]0,I_0]\),
-which implies \(P^\prime(0) > 1\).
-The origin is now a repeller, which means there exists some
+which implies \(P^\prime(0) > 1\) for all \(I \in \, ]0,I_0]\).
+This means the origin is a repeller, so there exists some
 (unknown) \(\mu > 0\) where \(]0,\mu]\) is negatively
@@
 If \(I < 0\), we have \(P^\prime(0) < 1\) and the origin is an
-asymptotically stable equilibrium for \eqref{eqn:r-discrete} instead.
+asymptotically stable equilibrium for \eqref{eqn:r-discrete}.
 

### 2021-08-27 latex/paper/main.tex
Full editing pass of the proof. It is still way too long and needs modifications because some parts feel wrong.

@@
 
-\subsection{All acrobots will rotate}
+\subsection{All Acrobots Rotate}
 
@@
 Let \(x(0) \in \mathcal{O}_1\) and suppose that the orbit
-\(x([0,\infty[)\) does not hit the line \(|q_u| = \pi\).
+\(x(t)\) does not hit the line \(|q_u| = \pi\).
 We claim that this orbit must remain in the compact set 
@@
 \(|p_u(t_K)| = K\). 
-Looking at the constrained dynamics \eqref{eqn:acrobot-constrained-dynamics},
-we see that \(\dot{p}_u \geq -3mgl\) and
-\(\dot{q}_u \geq c_1 p_u + c_2\) for some constants \(c_1,c_2 > 0\).
+From \eqref{eqn:acrobot-constrained-dynamics},
+we see that \(-3mgl \leq \dot{p}_u \leq 3mgl\) and
+\(\dot{q}_u \geq c_1 p_u - c_2\) for some constants \(c_1,c_2 > 0\).
 This implies there exists a \(t_{1/2} > 0\) (independent of \(K\)) 
@@
 \(|p_u(t)| >= K/2\) 
-and \(\dot{q}_u \geq c_1 K/2 + c_2\) 
+and \(|\dot{q}_u(t)| \geq c_1 K/2 - c_2\) 
 for \(t \in [t_K, t_K + t_{1/2}]\).
@@
 
-If the orbit \(x(t)\) does not hit the line \(|q_u| = \pi\), we now know it must
-remain within the compact set \(\Sone \times [-K,K]\).
+We now know that if \(x(t)\) does not hit the line \(|q_u| = \pi\),
+it must remain within the compact set \(\Sone \times [-K,K]\).
 By the Birkhoff theorem, \(x(t)\) approaches a positive limit set \(L^+\), 
 which is either the upright equilibrium \(\{(\pi,0)\}\), or
-a closed orbit not fully contained in \(\mathcal{O}_1\)
+a closed orbit (not fully contained in \(\mathcal{O}_1\))
 which crosses the positive \(q_u\)-axis at some point \(\bar{q} < \pi\).

### 2021-09-10 latex/nserc/introduction.tex
Modified the main text based on OGS_2021, the paper, and my thesis.

--- 2020-12-25 latex/introduction.tex
@@
 Even if the robot could perfectly track a time-based trajectory, an external
-disturbance may affect where squatting and standing occurs to the point that
+disturbance may affect where squatting and standing occur to the point that
 the swing actually slows down.
@@
 \citet{nhvc_dynamic_walking} in 2015, though there are references to more
-primitive versions going back as early as the the year 2000
+primitive versions going back as early as the year 2000
 \cite{vnhc_human_robot_cooperation}.

### 2021-09-10 latex/nserc/main.tex
Deleting extraneous data and just writing the nserc proposal
Modified the main text based on OGS_2021, the paper, and my thesis.

--- 2021-04-09 latex/ogs/main.tex
@@
 
+%---------- Add name to Header ----------%
+\usepackage{atbegshi,picture}
+\AtBeginShipout{\AtBeginShipoutUpperLeft{%
+        \put(\dimexpr\paperwidth-5.5cm\relax,-1.2cm){\makebox{\setlength{\fboxrule}{0pt}\framebox{Adan
+        Moran-MacDonald}}}%
+}}
+
 %/========== Main Document ==========/%
@@
 %---------- Title -----------%
-\title{Ontario Graduate Scholarship: Plan of Study}
-\author{Adan Moran-MacDonald}
+\title{Energy Regulation for Biologically-inspired Robotics}
+\author{}
 \date{}
 \maketitle
+
+\vspace{-1cm} % Move the following text up
+\textbf{What is your area of research? What sort of problems do you want to
+    address? WHy is the area important, and why are the applications important?
+    What is special about your approach, and how is it different from
+    existing/competing approaches? What is a sample application of the outcomes
+of your thesis?}
 
@@
 
-When a gymnast on a horizontal bar wants to swing up to do backflips, they
-typically perform a technique called a ``giant".
-This technique involves the gymnast swinging their legs based on their current
-speed and position, thereby generating enough momentum to rotate around the bar
+When a gymnast wants to do backflips around a horizontal bar, they start by
+performing a technique called a ``giant" \cite{usagym_giant}. 
+The gymnast swings their legs based on their current speed and position, thereby
+generating enough momentum to rotate around the bar.
+
+Now imagine the gymnast is a robot, and their creator is teaching them to act
+like a human.
+If the roboticist had studied classical control theory, they would plot a
+human gymnast's hip angle as a trajectory over time and tell the robot to synchronize
+its legs with this trajectory.
+While this approach often works for simple systems, it is not great for
+biologically-inspired robots: it can cause timing delays, it is not robust to
+many external disturbances, and it is \textit{not what humans do}.
+After all, human gymnasts do not have an internal stopwatch telling them when to
+extend or retract their legs.
+Instead, existing research suggests that human gymnasts actually move their legs
+as a function of their body angle and velocity
 \cite{pendulum_length_giant_gymnastics}.
-
-Standard practice in robotics begins with finding a trajectory over time
-replicating a desired behaviour, then using the robot's actuators to track this
-time-based trajectory. 
-While this often works for simple systems, it can cause timing delays, it is not
-robust to external disturbances, and it is not what humans do.
-A more natural approach is to constrain the robot's actuators according to
-some function of speed and position.
+Rather than using time-based motion, a clever roboticist would attempt to
+emulate this more natural human behaviour.
+% by driveing robot's legs according to some function of its torso speed and position.
 This type of behaviour is well described by the method of virtual constraints.
@@
 velocity (e.g. cars cannot slide sideways).
-One can design a specific constraint for a robot and enforce it using the
-actuators.
+One can design a constraint they want their robot to satisfy and enforce it
+using the robot's actuators.
 Since this constraint is not enforced by physics, it is known as a ``virtual"
-constraint
-\cite{vhcs_for_el_systems}.
+constraint \cite{vhcs_for_el_systems}.
 Virtual holonomic constraints (VHCs) have been used to control walking robots
@@
 behaviour in robotic systems.
-In my master's thesis, I demonstrated that VNHCs can also inject energy into
-robots:
+In my master's thesis, I demonstrated that VNHCs can inject/dissipate energy
+into/from robots:
 I designed a VNHC which enabled a gymnastics robot to gain energy
@@
 First, we will find mathematical conditions under which VNHCs are guaranteed to
-inject/dissipate energy in abstract mechanical systems.
-Then we will design a means of safely transitioning between two different VNHCs.
-Finally, we will test the theory on a gymnastics robot by designing a gymnastics
+inject/dissipate energy in arbitrary mechanical systems.
+Then we will design a means of transitioning between two different VNHCs,
+enabling us to regulate energy in a provably safe manner.
+Finally, we will test the theory on two robots: a gymnastics robot and a robotic
+monkey.
+For the gymnastics robot (which is already built), we will design a gymnastics
 routine which is performed solely by transitioning between different virtual
 constraints.
+After this, we will build the robotic monkey and design virtual constraints
+which allow it to swing on monkey bars, just like real monkeys swing through
+trees.
 

### 2021-09-10 latex/nserc/paper.tex
Modified the main text based on OGS_2021, the paper, and my thesis.

--- 2021-08-27 latex/paper/main.tex
@@
 \(q_a\) is actuated.
-A real gymnast cannot swing their legs in full circles, though they
-are usually flexible enough to raise them parallel to the floor;
-hence, we assume the leg angle \(q_a\) lies in \([-Q_a, Q_a]\) for some
-\(Q_a \in [\frac{\pi}{2}, \pi[\). 
-We also ignore any dissipative forces.
+We ignore dissipative forces in this model.
 
@@
 \end{align}
-For reasons that will become clear in later sections of this article, we will
+For reasons that will become clear later in this article, we
 use Hamiltonian mechanics to derive the dynamics of the acrobot. 
@@
 The following theorem requires that \(\mathcal{Q} = \R^n\) so that the
-transformation \(\tilde{q} = \mathbf{B}q\) is well defined.
+coordinate transformation is well defined.
 
@@
 always upright. 
-Imagine further that the swing has no seat at all, allowing the person to extend
+Imagine further that the swing has no seat, allowing the person to extend
 their legs beneath them. 
@@
 pivot to an angle proportional to the swing's speed.
+A real gymnast cannot swing their legs in full circles, though they
+are usually flexible enough to raise them parallel to the floor;
+hence, the VNHC must restrict the leg angle \(q_a\) 
+to lie in \([-Q_a, Q_a]\) for some \(Q_a \in [\frac{\pi}{2}, \pi[\). 
 Because the direction of motion is entirely determined by \(p_u\), 
@@
 design parameters.
+
 This constraint does not perfectly recreate giant motion, during which
@@
         extend the leg fully by setting \(q_a = 0\).
-        For these simulations, we assume this can be done instantaneously,
-        though in practice this is impossible.
+        In simulation we assume this can be done instantaneously,
+        though in practice this takes time.
 \end{itemize}
-Note that all orbits of the acrobot must cross the \(p_u\) axis, 
+All orbits of the acrobot must cross the \(p_u\) axis, 
 so this method will stabilize a rotation even if the acrobot is initialized in
 \(\mathcal{O}_1\).
-Also, the above supervisor is designed to regulate a rotation rate, but it
+While the above supervisor is designed to regulate a rotation rate, it
 cannot regulate the rotation direction.
@@
 Setting \(I = 0\) in \eqref{eqn:acrobot-constraint} for this acrobot
-yields the following mechanical energy of the nominal pendulum:
+yields the mechanical energy of the nominal pendulum:
 \[
@@
 
-Finally, The Dynamixel servo motor does not have a torque control mode; instead,
+Finally, the Dynamixel servo motor does not have a torque control mode; instead,
 we can assign the servo setpoint at iteration \(k \in \mathbb{Z}_{> 0}\)
@@
     Suppose that \(f_r(\theta,r,0) \equiv 0\) and \(f_\theta(\theta,r,0) > 0\).
-    Let \(g(\theta,r,I) = f_r(\theta,r,I)/f_\theta(\theta,r,I)\), and assume that 
+    Let \(g(\theta,r,I) := f_r(\theta,r,I)/f_\theta(\theta,r,I)\), and assume that 
     \((\partial g/\partial r)(\theta,r,0) \equiv 0\).
@@
     Let \(\hat{r}(\theta,r_0,I)\) be the solution to \eqref{eqn:rhat-ode}.
-    Taking the Taylor expansion of this solution around \(I = 0\) we get
+    Taking the Taylor expansion of this solution around \(I = 0\) yields
     \(\hat{r}(\theta,r_0,I) = \hat{r}(\theta,r_0,0) + I\hat{r}_1(\theta,r_0) 
@@
     \(g(\theta,\hat{r},0) \equiv 0\),
-    the function \(\hat{r}_1(\theta,r_0)\) is the solution to
+    the function \(\hat{r}_1(\theta,r_0)\) is the solution to the linear
+    time-varying ODE
     \begin{equation}\label{eqn:r1-hat-prime}
@@
 Hence, it can be written in the form
-\(R^\prime(2\pi,0,I) = I^2 \tilde{R}(I)\) where \(\tilde{R}(I)\) is smooth and
-zero at \(I = 0\).
+\(R^\prime(2\pi,0,I) = I^2 \tilde{R}(I)\) where \(\tilde{R}(I)\) is smooth,
+bounded, and zero at \(I = 0\).
 Thus, there exists \(I_0 > 0\) such that 

### 2021-09-10 latex/nserc/contributions.tex
Added the contributions part of the application.

--- 2021-04-09 latex/ogs/main.tex
@@
 %---------- Title -----------%
-\title{Ontario Graduate Scholarship: Plan of Study}
-\author{Adan Moran-MacDonald}
-\date{}
-\maketitle
+\subsection*{Part I: Contributions to Research and Development}
+\subsubsection*{d. Non-peer-reviewed Contributions}
+\textbf{Moran-MacDonald, A.}, \& Alajaji, F. (2018). The R\'{e}nyi Deterministic
+Information Bottleneck and Geometric Clustering. Unpublished technical report.
 
-%---------- Main Content ----------%
-\begin{large}
-\textbf{Background}
-\end{large}
+\subsection*{Part II: Most Significant Contributions to Research \& Development}
 
-When a gymnast on a horizontal bar wants to swing up to do backflips, they
-typically perform a technique called a ``giant".
-This technique involves the gymnast swinging their legs based on their current
-speed and position, thereby generating enough momentum to rotate around the bar
-\cite{pendulum_length_giant_gymnastics}.
+\subsection*{Part III: Applicant's Statement}
+\subsubsection*{Research Experience}
 
-Standard practice in robotics begins with finding a trajectory over time
-replicating a desired behaviour, then using the robot's actuators to track this
-time-based trajectory. 
-While this often works for simple systems, it can cause timing delays, it is not
-robust to external disturbances, and it is not what humans do.
-A more natural approach is to constrain the robot's actuators according to
-some function of speed and position.
-This type of behaviour is well described by the method of virtual constraints.
+\subsubsection*{Relevant Activities}
 
-There are two types of constraints on a mechanical system:
-\textit{holonomic} constraints restrict position (e.g. a snake robot can slither
-but not fly),
-while \textit{nonholonomic} constraints restrict position and
-velocity (e.g. cars cannot slide sideways).
-One can design a specific constraint for a robot and enforce it using the
-actuators.
-Since this constraint is not enforced by physics, it is known as a ``virtual"
-constraint
-\cite{vhcs_for_el_systems}.
-Virtual holonomic constraints (VHCs) have been used to control walking robots
-\cite{stable-walking}, autonomous bicycles \cite{bicycle},
-gymnastics robots \cite{xingbo-thesis}, and snake robots \cite{snake-robot}.
-Virtual nonholonomic constraints (VNHCs) have been used for human-robot
-cooperation \cite{vnhc-human-robot-coop} and for improving robustness of walking
-robots \cite{vnhc-biped-robot,hybrid_zero_dynamics_bipedal_nhvcs}, among other
-applications.
-
-\begin{large} \textbf{Proposal} \end{large}
-
-Virtual constraints have proven useful for generating biologically-inspired
-behaviour in robotic systems.
-In my master's thesis, I demonstrated that VNHCs can also inject energy into
-robots:
-I designed a VNHC which enabled a gymnastics robot to gain energy
-and perform backflips on a horizontal bar \cite{my-thesis}.
-Using VNHCs for \textit{energy injection} is beneficial because it allows robots to
-safely increase their momentum.
-Likewise, designing VNHCs for \textit{energy dissipation} allows robots to
-safely slow down in a realistic and reliable manner.
-
-We propose to study the energy injection and dissipation properties of VNHCs.
-First, we will find mathematical conditions under which VNHCs are guaranteed to
-inject/dissipate energy in abstract mechanical systems.
-Then we will design a means of safely transitioning between two different VNHCs.
-Finally, we will test the theory on a gymnastics robot by designing a gymnastics
-routine which is performed solely by transitioning between different virtual
-constraints.
-
-The advancements in mathematics from this research will bring improvements to
-the control of autonomous systems. 
-The ability to safely transition between complex constraints by adding or
-removing energy will allow for more expressive motion, and may become a standard
-technique for controlling biologically inspired robots.
-
-%---------- Bibliography ----------%
-\printbibliography
 \end{document}

### 2021-09-14 latex/nserc/proposal.tex
First completed editing pass of the proposal, with updated bibliography and extra background info.

%! TEX program = lualatex

\documentclass[a4paper,12pt]{article}

%/========== Preamble ==========/%
%----------- Set Main Font ----------%
\usepackage{times} % Times New Roman

%----------- Choose Margins ----------%
\usepackage[a4paper,margin=1.87cm]{geometry}

%---------- Bibliography Style ----------%
\usepackage[style=ieee,backend=biber]{biblatex}
\addbibresource{bib.bib}

%----------- Math Packages -----------%
\usepackage{mathtools}

%---------- Remove indenting from paragraphs ----------%
% Also makes each new paragraph start with an vertical space.
\usepackage[parfill]{parskip}

%---------- Add name to Header ----------%
\usepackage{atbegshi,picture}
\AtBeginShipout{\AtBeginShipoutUpperLeft{%
        \put(\dimexpr\paperwidth-5.5cm\relax,-1.2cm){\makebox{\setlength{\fboxrule}{0pt}\framebox{Adan
        Moran-MacDonald}}}%
}}

%/========== Main Document ==========/%
\begin{document}

%---------- Title -----------%
\title{Energy Regulation for Biologically-inspired Robotics}
\author{}
\date{}
\maketitle

\vspace{-1cm} % Move the following text up
%\textbf{What is your area of research? What sort of problems do you want to
%    address? WHy is the area important, and why are the applications important?
%    What is special about your approach, and how is it different from
%    existing/competing approaches? What is a sample application of the outcomes
%of your thesis?}
%\textbf{Detailed but concise. Provide background to position your research
%    within current knowledge of the field. State significance of research to the
%    field. Outline approach to be taken (citing literature). Clearly state
%differences btwn masters and this work.}

%---------- Main Content ----------%
\begin{large}
\textbf{Background}
\end{large}

When a gymnast wants to do backflips around a horizontal bar, they start by
performing a technique called a ``giant" \cite{usagym_giant}. 
The gymnast swings their legs based on their current speed and position, thereby
generating enough momentum to rotate around the bar.

Now imagine the gymnast is a robot, and their creator is teaching them to
perform giants like a human.
If the roboticist had studied classical control theory, they would plot a human
gymnast's leg angle as a trajectory over time and tell the robot to synchronize
its legs with this trajectory.
While this is the standard approach in robotics, it is not appropriate for
biologically-inspired robots: it is susceptible to timing delays, it is not
robust to external disturbances, and it is \textit{not what humans do}.
After all, human gymnasts do not have an internal stopwatch telling them when to
move their legs.
%If our robot were pushed mid-swing, its legs would continue along the
%time-based trajectory but the leg motion would desynchronize with
%the swinging of the torso, making the robot slow down.
Instead, existing research suggests that human gymnasts actually move their legs
as a function of their body angle and velocity
\cite{pendulum_length_giant_gymnastics}. 
Rather than using time-based motion, a clever roboticist would attempt to
emulate this natural human behaviour, which is well described by the method
of virtual constraints.

There are two types of constraints on a mechanical system:
\textit{holonomic} constraints restrict position (e.g. a snake robot can slither
on the ground but cannot fly),
while \textit{nonholonomic} constraints restrict both position and
velocity (e.g. an autonomous car with normal wheels cannot slide sideways).
It is often possible to use a robot's actuators to enforce a desired constraint
which is engineered to achieve a safety or motion planning goal.
Since this constraint is enforced by the robot's actuators, and not by the
physics of the system, it is known as a ``virtual" constraint
\cite{vhcs_for_el_systems}.
Virtual holonomic constraints (VHCs) have been used to control walking robots
\cite{stable_walking}, autonomous bicycles \cite{bicycle},
helicopters \cite{helicopter}, and snake robots \cite{snake_robot},
among other applications.
Unfortunately, VHCs cannot adequately recreate many animal behaviours.
For example, they cannot perfectly recreate the giant motion of gymnastics 
because they do not incorporate information about the gymnast's velocity
\cite{xingbo_thesis}.
This is where virtual nonholonomic constraints (VNHCs) are most useful.

The modern concept of VNHCs was described by
Griffin and Grizzle \cite{nhvc_dynamic_walking} in 2015, though there are
references to preliminary versions going back as early as the year 2000
\cite{vnhc_human_robot_coop}.
VNHCs have been of most notable use in bipedal walking robots,
where they show marked improvements in the robustness of walking gaits
when compared to previous control techniques
\cite{nhvc_gait_optimization,output_nhvc_bipedal_control}.
They have also been used for error-reduction in time-delayed teleoperation
\cite{vnhc_time_delay_teleop} and in the field of human-robot interaction
\cite{psd_based_vnhc_redundant_manipulator,haptic_vnhc}.
Horn \textit{et.~al.}~\cite{hybrid_zero_dynamics_bipedal_nhvcs} derived the 
equations of motion for robots constrained by VNHCs, and they used VNHCs to
improve the gait of walking robots on a variable-slope terrain
\cite{nhvc_incline_walking}.

\begin{large} \textbf{Proposal} \end{large}

Virtual constraints have proven useful for generating biologically-realistic
behaviour in robotic systems.
In my master's thesis I showed that VNHCs allow for \textit{energy regulation},
i.e., they can inject energy into (or dissipate energy from) a certain class of
mechanical systems \cite{my-thesis}.
One such system is a gymnastics robot called the
\textit{acrobot}, for which I designed a VNHC that generates giant-like motion. 
This VNHC enabled the acrobot to gain energy and perform backflips on a
horizontal bar.

Using VNHCs for \textit{energy injection} is beneficial because it allows robots to
safely increase their momentum.
Likewise, designing VNHCs for \textit{energy dissipation} allows robots to
safely slow down in a realistic and reliable manner. 
In this PhD, I propose to further study the energy regulation properties of VNHCs.
First, we will find mathematical conditions under which VNHCs are guaranteed to
inject or dissipate energy in arbitrary fixed-base mechanical systems.
Then we will design a means of transitioning between two different VNHCs,
enabling us to regulate energy in a provably safe manner.
We will experimentally verify the results of this theory on the acrobot by
designing a gymnastics routine which is performed solely by transitioning
between different virtual constraints.
Finally, we will extend the theory to robots which can walk or drive, 
and we will build a robotic monkey which uses VNHCs to swing on monkey bars just
like a real monkey would swing through trees.

The advancements in mathematics from this research will bring improvements
to the control of all autonomous systems. 
The ability to safely transition between complex constraints by adding or
removing energy will allow for more expressive motion, and may become a standard
technique for controlling biologically inspired robots.

\newpage
%---------- Bibliography ----------%
\printbibliography
\end{document}

% vim: set tw=80 ts=4 sw=4 sts=0 et ffs=unix :

### 2021-09-14 latex/nserc/proposal.tex
Minor edit to proposal.

@@
 behaviour in robotic systems.
-In my master's thesis I showed that VNHCs allow for \textit{energy regulation},
-i.e., they can inject energy into (or dissipate energy from) a certain class of
-mechanical systems \cite{my-thesis}.
-One such system is a gymnastics robot called the
+In my master's thesis I showed that VNHCs also allow for 
+\textit{energy regulation}, i.e., they can inject energy into (or dissipate
+energy from) a certain class of mechanical systems \cite{my-thesis}.
+One such mechanical system is a gymnastics robot called the
 \textit{acrobot}, for which I designed a VNHC that generates giant-like motion. 
-This VNHC enabled the acrobot to gain energy and perform backflips on a
+This VNHC enables the acrobot to gain energy and perform backflips on a
 horizontal bar.
@@
 Using VNHCs for \textit{energy injection} is beneficial because it allows robots to
-safely increase their momentum.
+safely and reliably increase their momentum.
 Likewise, designing VNHCs for \textit{energy dissipation} allows robots to
-safely slow down in a realistic and reliable manner. 
+safely and realistically slow down.
 In this PhD, I propose to further study the energy regulation properties of VNHCs.
 First, we will find mathematical conditions under which VNHCs are guaranteed to
-inject or dissipate energy in arbitrary fixed-base mechanical systems.
+inject or dissipate energy in fixed-base robots.
 Then we will design a means of transitioning between two different VNHCs,
-enabling us to regulate energy in a provably safe manner.
+enabling us to regulate a robot's energy in a provably safe manner.
 We will experimentally verify the results of this theory on the acrobot by

### 2021-09-14 matlab/thesis_sims/regulationEvent.m
Changes made in a different branch

@@ -11,12 +11,12 @@
 % * fval = value at which to trigger the event.
 % * d = direction: +1 for increasing, -1 for decreasing, 0 for either.
 
-function [value,isterminal,direction] = energyEvent(t,qp,E,Eval)
-    % Trigger when E(q,p) = Eval. The value increases as E -> Eval from
+function [value,isterminal,direction] = regulationEvent(t,qp,f,fval,d)
+    % Trigger when f(q,p) = fval. The value increases as f -> fval from
     % below.
-    value = E(qp(1),qp(2))-Eval;
-    % Don't stop the integration, just record the event
-    isterminal = 0;
+    value = f(qp(1),qp(2))-fval;
+    % Stop the integration
+    isterminal = 1;
     % Track only values where the energy is increasing past Eval, not those
     % where you come back down. If you want it to track only decreasing,
     % set direction to -1. If you want both increasing and decreasing, set

### 2021-09-15 latex/nserc/contributions.tex
Wrote about the RDIB project.

@@
 \subsubsection*{d. Non-peer-reviewed Contributions}
-\textbf{Moran-MacDonald, A.}, \& Alajaji, F. (2018). The R\'{e}nyi Deterministic
-Information Bottleneck and Geometric Clustering. Unpublished technical report.
+\textbf{Moran-MacDonald, A.} \& Alajaji, F. (2018). The R\'{e}nyi Deterministic
+Information Bottleneck and Geometric Clustering. NSERC summer research project,
+Queen's University.
 
 \subsection*{Part II: Most Significant Contributions to Research \& Development}
+\textit{Describe your role in research, clarify your contribution, provide
+    details on significance of tech report, indicate collaborations, discuss
+relevance to engineering practice (if appropriate)}
+``The R\'{e}nyi Deterministic Information Bottleneck and Geometric
+Clustering" is a technical report in which we develop a new
+information-theoretic tool for clustering datapoints.
+I wrote the report in its entirety, and I produced the theoretical results with
+the help and advice of my supervisor, Fady Alajaji.
+In the report, I summarized the existing theory on the
+\textit{information bottleneck} (IB) and its variant, the \textit{deterministic
+information bottleneck} (DIB), which are tools from information theory that
+have been touted as a rigorous foundation for Deep Neural Networks. 
+The DIB in particular is useful for clustering data.
+Its equations are defined using an information-theoretic concept called
+``Shannon's Mutual Information" (SMI).
+In this report, Alajaji suggested we extend the equations by using a more
+general concept called ``R\'{e}nyi's Mutual Information" to get the aptly-named
+\textit{R\'{e}nyi's deterministic information bottleneck} (RDIB).
+I compared the efficacy of the DIB and RDIB for clustering data, and we found
+that the RDIB is more robust when the probability distribution of the data model 
+does not match the true distirbution which generated the data.
+This report was used by two undergraduate thesis groups at Queen's University in
+2019, and it motivated further published research [1].
+
+[1] J.J. Weng, F. Alajaji and T. Linder, \textit{An information bottleneck
+problem with R\'{e}nyi's Entropy}, Proceedings of the IEEE International
+Symposium on Information Theory, Melbourne, Australia, July 2021.
 
@@
 \subsubsection*{Research Experience}
+\textit{Describe scientific and engineering abilities you gained in past
+    research experience, including projects, honours thesis, coop reports. If
+    you have work experience, discuss relevance of that to your proposed field
+of study and what benefits you gained.}
 
 \subsubsection*{Relevant Activities}
+\textit{Describe professional, academic, extracurricular activities that
+    demonstrate communication, interpersonal and leadership skills, e.g.
+    teaching, mentoring, managing projects, volunteer work, committee work,
+participation in organizations and clubs, or work experience}.
 

### 2021-09-15 latex/nserc/contributions.tex
Started the Part III extracurriculars

@@
 of study and what benefits you gained.}
+Masters Research skills
+Undergrad thesis research skills
+Neptec skills
 
@@
 
+Over the course of my career, I have enjoyed several extracurricular roles which
+developed my communication, interpersonal, and leadership skills.
+What follows is a brief description of some of these roles, along with a summary
+of what I learned.
+\begin{itemize}
+\item \textbf{Lionel Massey Fund Co-chair, Massey College, Sep. 2021 -- Present}.
+    I am one of six co-chairs for the Lionel Massey Fund (LMF) -- the social
+    committee for Massey College. Thus far I have dramatically improved my
+    communication and leadership skills by organizing 13 orientation events for
+    all 150 members of the college. I have learned to communicate problems
+    early, and to ask people what they want or expect so I know how to best
+    serve their needs.
+
+\item \textbf{Teaching Assistant positions, 2017 -- Present}.
+    I have been a teaching assistant for 14 engineering courses in the past 5 years.
+    Some of these were in-person, some were online, but in both cases I always
+    learned a new ways to make the material more engaging.
+    I focus most on attaining student participation, and I actively learn each
+    student's name to make them feel comfortable asking questions. When teaching
+    pre-recorded material, I use many examples and make the lectures as easily
+    digestible as possible to account for the difficulties of online
+    interactions.
+
+\item \textbf{Jiu-Jitsu Instruction}. 
+    I am an instructor at the University of Toronto Jiu-Jitsu club. I teach
+    people to be confident in what their bodies can accomplish while fostering a
+    safe, friendly atmosphere that encourages learning through action.
+    This role has refined my communication and interpersonal skills:
+    it has taught me to create learning plans for my students;
+    break down difficult concepts into manageable chunks; 
+    and to reward mistakes, because the best way to learn is to try, fail, and
+    then try again anyway.
+
+\item \textbf{Radar Developer}
+    TODO: managed the project, led the software design, and made sure everyone
+    was learning and was participating so they would have a feeling of agency
+    over the project.
+
+    TODO: These need to be edited for more than just teaching, but also to show
+    what I learned and how I showed leadership and communication.
+\item \textbf{Queen’s University Satellite Design Team}. 
+    I taught C++ to the team’s software manager and other developers. I worked
+    with the manager to create good design patterns for the satellite’s
+    software. I also taught three first-year students how to use information
+    theory to improve one of the algorithms used on the satellite.
+
+\item \textbf{Resource Manager at Queen’s University}. 
+    I created new workbooks which intuitively explained difficult concepts for 9
+    first-year engineering courses. The workshops were attended by over 200
+    first-year students.
+    TODO: might be unnecessary
+
+\item \textbf{Head Programming mentor for FIRST Team 2809}. 
+    I mentored high school students in Object Oriented programming and robotics
+    software. Together we programmed a robot in under 6 weeks and placed 3rd at
+    the Ontario provincial competition in 2014.
+    TODO: Learned to let go of pride, give power to students, and communicate
+    with other teams in a high-stress scenario.
+
+I take pride in my teaching abilities. I always look for teaching opportunities
+and strive to become a better educator by reducing knowledge to its
+fundamentals. It is one of my goals in life to teach and learn as much as
+possible; the variety of roles I have taken on are a demonstration of this. By
+teaching APS106 for the third summer in a row, I aim to give students confidence
+in their programming abilities while improving my own teaching skills.
+
 \end{document}

### 2021-09-16 latex/nserc/contributions.tex
Edits and started filling out most of Part III

--- 2021-09-10 latex/nserc/contributions.tex
@@
 \subsubsection*{d. Non-peer-reviewed Contributions}
-\textbf{Moran-MacDonald, A.}, \& Alajaji, F. (2018). The R\'{e}nyi Deterministic
-Information Bottleneck and Geometric Clustering. Unpublished technical report.
+\textbf{Moran-MacDonald, A.} \& Alajaji, F. (2018). The R\'{e}nyi Deterministic
+Information Bottleneck and Geometric Clustering. NSERC summer research project,
+Queen's University.
 
 \subsection*{Part II: Most Significant Contributions to Research \& Development}
+\textit{Describe your role in research, clarify your contribution, provide
+    details on significance of tech report, indicate collaborations, discuss
+relevance to engineering practice (if appropriate)}
+``The R\'{e}nyi Deterministic Information Bottleneck and Geometric
+Clustering" is a technical report about a new information-theoretic tool for
+clustering datapoints.
+My contribution was substantial: I produced the theoretical and simulation
+results (with lots of advice from my supervisor, Fady Alajaji), and then I wrote
+the report in its entirety.
+%In the report, I summarized the existing theory on the
+%\textit{information bottleneck} (IB) and its variant, the \textit{deterministic
+%information bottleneck} (DIB), which are tools from information theory that
+%have been touted as a rigorous foundation for Deep Neural Networks. 
+%The DIB in particular is useful for clustering data.
+%Its equations are defined using an information-theoretic concept called
+%``Shannon's Mutual Information" (SMI).
+%In this report, Alajaji suggested we extend the equations by using a more
+%general concept called ``R\'{e}nyi's Mutual Information" to get the aptly-named
+%\textit{R\'{e}nyi's deterministic information bottleneck} (RDIB).
+%I compared the efficacy of the DIB and RDIB for clustering data, and we found
+%that the RDIB is more robust when the probability distribution of the data model 
+%does not match the true distirbution which generated the data.
+This report was used by two undergraduate groups at Queen's University for their
+honours theses in 2019. 
+It also motivated further research, which was published in [1].
+
+[1] J.J. Weng, F. Alajaji and T. Linder, \textit{An information bottleneck
+problem with R\'{e}nyi's Entropy}, Proceedings of the IEEE International
+Symposium on Information Theory, Melbourne, Australia, July 2021.
 
@@
 \subsubsection*{Research Experience}
+\textit{Describe scientific and engineering abilities you gained in past
+    research experience, including projects, honours thesis, coop reports. If
+    you have work experience, discuss relevance of that to your proposed field
+of study and what benefits you gained.}
+The following list outlines the scientific and engineering abilities I gained in
+my research and professional roles.
+\begin{itemize}
+    \item \textbf{MASc Research, University of Toronto, 2018 -- 2021}
+        TODO.
+        Research skills: time management, critical thinking, coming up with
+        well-formulated research question, tenacity and creativity
+        (always ``attacking" the problem from multiple directions), dedication
+        to solving a complex problem, explaining
+        my work both orally and in writing, presentation skills, 
+        literature reviews and how to read papers
+        Engineering skills: nonlinear control techniques, implementation of
+        control systems on hardware
+
+    \item \textbf{Radar Developer, 2018}
+        TODO:
+        Managing a software project, implementing software for a hardware
+        system, validating documentation (i.e. never trust it blindly).
+
+    \item \textbf{Undergraduate Thesis, Queen's University, 2017 -- 2018}
+        TODO:
+        First time doing a research project. Collaborating with fellow
+        researchers. Coming up with an application for existing research. Basic
+        literature review skills, report writing. Code optimization (turning
+        equations into matrix form for speed, parallelization, C++ vs MATLAB),
+        creating a GUI for users.
+
+    \item \textbf{Software Developer, Neptec Technologies Crop, 2016 -- 2017}
+        My experience at Neptec was incredibly beneficial.
+        I worked on several embedded software projects, including a control
+        systems project. The exposure to various areas of electrical and
+        computer engineering is what convinced me to pursue research in the
+        first place.
+        I learned to work on large software projects, where multiple people work
+        together on the same topic.
+        I also became extremely proficient at software development, a skill
+        which has served me well in my research experiments on physical robots.
+
+\end{itemize}
 
 \subsubsection*{Relevant Activities}
+\textit{Describe professional, academic, extracurricular activities that
+    demonstrate communication, interpersonal and leadership skills, e.g.
+    teaching, mentoring, managing projects, volunteer work, committee work,
+participation in organizations and clubs, or work experience}.
+
+Over the course of my career, I have enjoyed several extracurricular roles which
+developed my communication, interpersonal, and leadership skills.
+What follows is a brief description of some of these roles, along with a summary
+of what I learned.
+\begin{itemize}
+\item \textbf{Lionel Massey Fund Co-chair, Massey College, September 2021 -- Present}.
+    I am one of six co-chairs for the Lionel Massey Fund (LMF) -- the social
+    committee for Massey College, a graduate college at the University of
+    Toronto. Thus far I have dramatically improved my
+    communication and leadership skills by managing 13 orientation events for
+    150 members of the college. I have learned to explicitly communicate my needs
+    with others, and to ask questions to understand their needs so we all succeed.
+
+\item \textbf{Teaching Assistant, 2017 -- Present}.
+    I have been a teaching assistant for 14 engineering courses in the past 5 years.
+    Each course has taught me better teaching skills, along with new ways of
+    making the material more engaging.
+    I focus most on attaining high student participation, and I find
+    commonalities between myself and students so they feel comfortable asking
+    questions.
+
+\item \textbf{Jiu-Jitsu Instructor, 2019 -- Present}. 
+    I am an instructor at the University of Toronto Jiu-Jitsu club. I teach
+    people martial arts while fostering a safe, friendly atmosphere that
+    encourages learning through action.
+    This role has refined my communication and interpersonal skills:
+    it has taught me to create learning plans for my students;
+    break down difficult concepts into manageable chunks; 
+    and reward mistakes, because the best way to learn is to try, fail, and
+    to always try again.
+
+\item \textbf{Science Payload Designer, Queen’s University Satellite Design
+    Team, 2017 -- 2018}. 
+    In this role, I taught C++ and information theory to the team’s software
+    developers. I also learned project management skills, because I organized
+    the satellite codebase but took on a supportive role for my peers (rather
+    than writing the software myself).
+
+\item \textbf{Head Programming Mentor, FIRST Team 2809, 2013 -- 2015}. 
+    I mentored high school students in Object Oriented programming for robotics
+    software. Together we programmed a robot in under 6 weeks and placed 3rd at
+    the Ontario provincial competition in 2014.
+    In this role I learned to give my students agency over the project, and to
+    communicate with people from all over the world in high-stress situations.
+
+\end{itemize}
+
+I take pride in my teaching, communication, and leadership abilities.
+Obtaining an NSERC scholarship will provide me with the financial liberty to
+participate in more volunteer roles, where I can better myself while giving back to my
+community.
 

### 2021-09-17 latex/nserc/contributions.tex
Edited Part III

@@
     I am one of six co-chairs for the Lionel Massey Fund (LMF) -- the social
-    committee for Massey College, a graduate college at the University of
-    Toronto. Thus far I have dramatically improved my
-    communication and leadership skills by managing 13 orientation events for
-    150 members of the college. I have learned to explicitly communicate my needs
-    with others, and to ask questions to understand their needs so we all succeed.
+    committee for Massey College, a graduate residence at the University of
+    Toronto.
+    In managing 13 orientation events for 150 members of the college, I have
+    learned to explicitly communicate my needs with others, and to ask pointed
+    questions to understand their needs.
 
@@
     I have been a teaching assistant for 14 engineering courses in the past 5 years.
-    Each course has taught me better teaching skills, along with new ways of
-    making the material more engaging.
-    I focus most on attaining high student participation, and I find
-    commonalities between myself and students so they feel comfortable asking
-    questions.
+    Each course I teach improves my communication skills, and I've learned
+    many ways of making course material fun and engaging.
+    I focus most on student participation by finding making them feel
+    comfortable asking questions.
 
@@
     break down difficult concepts into manageable chunks; 
-    and reward mistakes, because the best way to learn is to try, fail, and
-    to always try again.
+    and to reward mistakes, because the best way to learn is to try, fail, and
+    always try again.
 
@@
     In this role, I taught C++ and information theory to the team’s software
-    developers. I also learned project management skills, because I organized
-    the satellite codebase but took on a supportive role for my peers (rather
-    than writing the software myself).
+    developers. I learned project management skills by organizing
+    the satellite codebase, and I took on a supportive role for my peers rather
+    than writing the software myself.
 
@@
     the Ontario provincial competition in 2014.
-    In this role I learned to give my students agency over the project, and to
-    communicate with people from all over the world in high-stress situations.
+    In this role I learned to lead my students by giving them agency over the
+    project, and to communicate with people from all over the world in
+    high-stress situations.
 

### 2021-09-17 latex/nserc/contributions.tex
First draft of the contributions document

@@
     \item \textbf{MASc Research, University of Toronto, 2018 -- 2021}
-        TODO.
-        Research skills: time management, critical thinking, coming up with
-        well-formulated research question, tenacity and creativity
-        (always ``attacking" the problem from multiple directions), dedication
-        to solving a complex problem, explaining
-        my work both orally and in writing, presentation skills, 
-        literature reviews and how to read papers
-        Engineering skills: nonlinear control techniques, implementation of
-        control systems on hardware
+        In my master's degree, I learned several important research skills.
+        My supervisor taught me to formulate research questions
+        correctly; I learned the tenacity and dedication required to solve a
+        complex problem; I found creative solutions by attacked problems from
+        multiple directions; I learned to read papers, perform a literature
+        review, think critically about the literature, and to apply existing
+        results to my research; I improved my time management skills, practiced
+        explaining my work (both orally and in writing), and improved my
+        presentation skills for the thesis defense. Finally, I learned to write
+        research papers and a thesis, which is important for communicating my
+        work to the engineering community.
+        I also learned several engineering skills, including: advanced control
+        techniques; simulating physical systems in MATLAB; and implementing
+        control systems in hardware.
 
-    \item \textbf{Radar Developer, 2018}
-        TODO:
-        Managing a software project, implementing software for a hardware
-        system, validating documentation (i.e. never trust it blindly).
+    \item \textbf{Radar Developer, University of Toronto, 2018}
+        As part of a course project, I learned useful engineering skills by
+        writing software for a radar device.
+        First, I improved my project management skills by keeping our group
+        meetings organized, and by separating the tasks in such a way
+        that everyone was learning something from the project.
+        I learned to read hardware documentation; to program a hardware
+        device by implementing the documentation in software;
+        and to validate the claims of the hardware capabilities for myself,
+        i.e., to never blindly trust documentation.
 
     \item \textbf{Undergraduate Thesis, Queen's University, 2017 -- 2018}
-        TODO:
-        First time doing a research project. Collaborating with fellow
-        researchers. Coming up with an application for existing research. Basic
-        literature review skills, report writing. Code optimization (turning
-        equations into matrix form for speed, parallelization, C++ vs MATLAB),
-        creating a GUI for users.
+        My undergraduate thesis was my first time doing a research project.
+        I learned to do collaborative application-based research; that is, I
+        worked with my peers to find an application for existing theoretical
+        results. During this project I also learned to perform a basic
+        literature review, and to manage deadlines with a supervisor.
+        In terms of engineering skills, I learned to write reports in LaTeX,
+        and my thesis group worked together on code optimization. We learned to
+        convert equations into matrix form to make use of hardware speed; we
+        parallelized our code to take advantage of multithreading; we ported
+        MATLAB code to C++ to achieve faster algorithms; and we created a GUI so
+        we could test our code quickly and easily.
 

### 2021-09-19 latex/nserc/contributions.tex
Major edits of contributions.

--- 2021-09-17 latex/nserc/contributions.tex
@@
 \subsection*{Part II: Most Significant Contributions to Research \& Development}
-\textit{Describe your role in research, clarify your contribution, provide
-    details on significance of tech report, indicate collaborations, discuss
-relevance to engineering practice (if appropriate)}
+%\textit{Describe your role in research, clarify your contribution, provide
+%    details on significance of tech report, indicate collaborations, discuss
+%relevance to engineering practice (if appropriate)}
 ``The R\'{e}nyi Deterministic Information Bottleneck and Geometric
@@
 My contribution was substantial: I produced the theoretical and simulation
-results (with lots of advice from my supervisor, Fady Alajaji), and then I wrote
-the report in its entirety.
-%In the report, I summarized the existing theory on the
-%\textit{information bottleneck} (IB) and its variant, the \textit{deterministic
-%information bottleneck} (DIB), which are tools from information theory that
-%have been touted as a rigorous foundation for Deep Neural Networks. 
-%The DIB in particular is useful for clustering data.
-%Its equations are defined using an information-theoretic concept called
-%``Shannon's Mutual Information" (SMI).
-%In this report, Alajaji suggested we extend the equations by using a more
-%general concept called ``R\'{e}nyi's Mutual Information" to get the aptly-named
-%\textit{R\'{e}nyi's deterministic information bottleneck} (RDIB).
-%I compared the efficacy of the DIB and RDIB for clustering data, and we found
-%that the RDIB is more robust when the probability distribution of the data model 
-%does not match the true distirbution which generated the data.
+results (with an abundance of advice from my supervisor, Fady Alajaji), and I
+wrote the report in its entirety.
 This report was used by two undergraduate groups at Queen's University for their
-honours theses in 2019. 
-It also motivated further research, which was published in [1].
+honours theses in 2019, and it motivated further research [1].
 
@@
 \subsubsection*{Research Experience}
-\textit{Describe scientific and engineering abilities you gained in past
-    research experience, including projects, honours thesis, coop reports. If
-    you have work experience, discuss relevance of that to your proposed field
-of study and what benefits you gained.}
+%\textit{Describe scientific and engineering abilities you gained in past
+%    research experience, including projects, honours thesis, coop reports. If
+%    you have work experience, discuss relevance of that to your proposed field
+%of study and what benefits you gained.}
 The following list outlines the scientific and engineering abilities I gained in
@@
 \begin{itemize}
-    \item \textbf{MASc Research, University of Toronto, 2018 -- 2021}
-        TODO.
-        Research skills: time management, critical thinking, coming up with
-        well-formulated research question, tenacity and creativity
-        (always ``attacking" the problem from multiple directions), dedication
-        to solving a complex problem, explaining
-        my work both orally and in writing, presentation skills, 
-        literature reviews and how to read papers
-        Engineering skills: nonlinear control techniques, implementation of
-        control systems on hardware
+    \item \textbf{MASc Research, University of Toronto, 2018 -- 2021}.
+        In my master's degree, I learned several important research skills:
+        my supervisor taught me to correctly formulate research questions;
+        I learned the tenacity and dedication required to solve a
+        complex problem; I developed my creativity by attacking problems from
+        multiple directions; I learned to read papers, perform a literature
+        review, think critically about the literature, and apply existing
+        results to my research; I improved my time management skills, practiced
+        explaining my work (both orally and in writing), and improved my
+        presentation skills for the thesis defense; and I learned to write
+        research papers and a thesis, which are important for communicating my
+        work to the engineering community.
+        I also learned several engineering skills, including advanced control
+        techniques; simulation of physical systems in MATLAB; and implementation
+        of control systems in hardware.
 
-    \item \textbf{Radar Developer, 2018}
-        TODO:
-        Managing a software project, implementing software for a hardware
-        system, validating documentation (i.e. never trust it blindly).
+    \item \textbf{Radar Developer, University of Toronto, 2018}.
+        I wrote software for a radar device as part of a group project in one of
+        my graduate courses.
+        This project taught me several engineering skills.
+        First, I improved my project management skills by keeping our group
+        meetings organized, and by separating the tasks in such a way
+        that everyone was learning something from the project.
+        I learned to read hardware documentation and to wire a device from
+        schematics.
+        I wrote networking code in Python to interface with the
+        device, and validatet the hardware's capabilities for myself by
+        creating appropriate testing environments. 
+        All these skills will be beneficial for my PhD research when I build a
+        robot.
 
-    \item \textbf{Undergraduate Thesis, Queen's University, 2017 -- 2018}
-        TODO:
-        First time doing a research project. Collaborating with fellow
-        researchers. Coming up with an application for existing research. Basic
-        literature review skills, report writing. Code optimization (turning
-        equations into matrix form for speed, parallelization, C++ vs MATLAB),
-        creating a GUI for users.
+    \item \textbf{Undergraduate Thesis, Queen's University, 2017 -- 2018}.
+        For my undergraduate thesis, I worked with my peers to find an
+        application for existing theoretical results. 
+        During this project I also learned to perform a basic literature review,
+        and to manage deadlines with a supervisor.
+        In terms of engineering skills, I learned to write reports in LaTeX
+        and my thesis group worked together on code optimization: we learned to
+        convert equations into matrix form to take advantage of hardware
+        optimization; we parallelized our code so our algorithm would run faster
+        through multithreading; we ported MATLAB code to C++ to achieve even
+        faster results; and we created a GUI so we could test our code quickly
+        and easily.
 
-    \item \textbf{Software Developer, Neptec Technologies Crop, 2016 -- 2017}
-        My experience at Neptec was incredibly beneficial.
-        I worked on several embedded software projects, including a control
-        systems project. The exposure to various areas of electrical and
-        computer engineering is what convinced me to pursue research in the
-        first place.
-        I learned to work on large software projects, where multiple people work
-        together on the same topic.
-        I also became extremely proficient at software development, a skill
-        which has served me well in my research experiments on physical robots.
+    \item \textbf{Software Developer, Neptec Technologies Corp., 2016 -- 2017}.
+        My experience at Neptec was integral to my career, as it is what drove
+        me to pursue research in robotics.
+        Working at Neptec exposed me to several areas of electrical and computer
+        engineering.
+        I worked on several embedded software projects, my favourite of which
+        was a control systems project.
+        I loved this project, and I decided to do a master's (and a PhD) so I
+        could learn as much about control as possible.
+        I also learned to work on large software projects, where multiple people work
+        together on the same topic; 
+        and I became extremely proficient at software development, a skill which
+        has served me well in my research experiments on physical robots.
 
@@
 \subsubsection*{Relevant Activities}
-\textit{Describe professional, academic, extracurricular activities that
-    demonstrate communication, interpersonal and leadership skills, e.g.
-    teaching, mentoring, managing projects, volunteer work, committee work,
-participation in organizations and clubs, or work experience}.
+%\textit{Describe professional, academic, extracurricular activities that
+%    demonstrate communication, interpersonal and leadership skills, e.g.
+%    teaching, mentoring, managing projects, volunteer work, committee work,
+%participation in organizations and clubs, or work experience}.
 
@@
 developed my communication, interpersonal, and leadership skills.
-What follows is a brief description of some of these roles, along with a summary
-of what I learned.
+What follows is a brief description of some of these roles.
 \begin{itemize}
 \item \textbf{Lionel Massey Fund Co-chair, Massey College, September 2021 -- Present}.
-    I am one of six co-chairs for the Lionel Massey Fund (LMF) -- the social
-    committee for Massey College, a graduate residence at the University of
-    Toronto.
-    In managing 13 orientation events for 150 members of the college, I have
-    learned to explicitly communicate my needs with others, and to ask pointed
-    questions to understand their needs.
+    I am one of six co-chairs for the Lionel Massey Fund, which is the events
+    committee for Massey College (an interdisciplinary college at the University
+    of Toronto).
+    I have already organized 13 events for members of the college, and will
+    continue to organize biweekly events throughout the year.
+    This role has taught me the logistics and leadership skills required to
+    manage large events for 150 people.
+    I have also learned to make events accessible to everyone by asking
+    questions so I can understand the needs of all attendees.
 
@@
     I have been a teaching assistant for 14 engineering courses in the past 5 years.
-    Each course I teach improves my communication skills, and I've learned
-    many ways of making course material fun and engaging.
-    I focus most on student participation by finding making them feel
-    comfortable asking questions.
+    Each course I teach improves my communication skills, and each tutorial gives
+    me the opportunity to make the course material more fun and engaging.
+    I focus most on attaining high levels of student participation by being
+    kind and non-judgemental, which makes students feel
+    comfortable enough to ask whatever question crosses their mind.
 
@@
     I am an instructor at the University of Toronto Jiu-Jitsu club. I teach
-    people martial arts while fostering a safe, friendly atmosphere that
+    martial arts while fostering a safe, friendly atmosphere that
     encourages learning through action.
@@
     break down difficult concepts into manageable chunks; 
-    and to reward mistakes, because the best way to learn is to try, fail, and
-    always try again.
+    and to reward students for trying their best, even if they make a mistake.
 
@@
     developers. I learned project management skills by organizing
-    the satellite codebase, and I took on a supportive role for my peers rather
-    than writing the software myself.
+    the satellite codebase and handling code reviews, and I took on a supportive
+    and mentorship role for my peers rather than writing software myself.
 
@@
 
-I take pride in my teaching, communication, and leadership abilities.
+I take pride in my teaching, communication, and leadership abilities, yet I
+remain aware that these can always be improved.
 Obtaining an NSERC scholarship will provide me with the financial liberty to

### 2021-09-19 latex/nserc/proposal.tex
Major edits of contributions.

@@
 between different virtual constraints.
-Finally, we will extend the theory to robots which can walk or drive, 
+Finally, we will extend the theory to mobile animal-like robots,
 and we will build a robotic monkey which uses VNHCs to swing on monkey bars just

### 2021-09-19 latex/nserc/contributions.tex
More edits, now contributions is too long. Needs to be shortened

@@
 clustering datapoints.
-My contribution was substantial: I produced the theoretical and simulation
-results (with an abundance of advice from my supervisor, Fady Alajaji), and I
-wrote the report in its entirety.
+I contributed to this report by producing all the theoretical and simulation
+results, though I could not have done this without the abundant advice of my
+supervisor, Fady Alajaji. 
+I also wrote the report in its entirety.
 This report was used by two undergraduate groups at Queen's University for their
@@
         presentation skills for the thesis defense; and I learned to write
-        research papers and a thesis, which are important for communicating my
+        research papers and a thesis, which are important tools for communicating my
         work to the engineering community.
         I also learned several engineering skills, including advanced control
-        techniques; simulation of physical systems in MATLAB; and implementation
-        of control systems in hardware.
+        techniques, simulation of physical systems in MATLAB, and implementation
+        of control systems on a physical robot.
 
     \item \textbf{Radar Developer, University of Toronto, 2018}.
-        I wrote software for a radar device as part of a group project in one of
-        my graduate courses.
+        In the group project for a graduate course, I worked with three other
+        students to write software which reads data from a radar device.
         This project taught me several engineering skills.
@@
 
-\item \textbf{Teaching Assistant, 2017 -- Present}.
-    I have been a teaching assistant for 14 engineering courses in the past 5 years.
-    Each course I teach improves my communication skills, and each tutorial gives
-    me the opportunity to make the course material more fun and engaging.
-    I focus most on attaining high levels of student participation by being
-    kind and non-judgemental, which makes students feel
-    comfortable enough to ask whatever question crosses their mind.
-
 \item \textbf{Jiu-Jitsu Instructor, 2019 -- Present}. 
@@
     and to reward students for trying their best, even if they make a mistake.
+
+\item \textbf{Teaching Assistant, 2017 -- Present}.
+    I have been a teaching assistant for 14 engineering courses in the past 5 years.
+    Each course I teach improves my communication skills, and each tutorial gives
+    me the opportunity to make the course material more fun and engaging.
+    I focus most on attaining high levels of student participation by being
+    kind and non-judgemental, which makes students feel
+    comfortable enough to ask whatever question crosses their mind.
 

### 2021-09-19 latex/nserc/contributions.tex
Editing the contributions again, still not done and too long.

@@
 
-Over the course of my career, I have enjoyed several extracurricular roles which
-developed my communication, interpersonal, and leadership skills.
+Over the course of my academic career, I have enjoyed several extracurricular
+roles which developed my communication, interpersonal, and leadership skills.
 What follows is a brief description of some of these roles.
@@
     This role has taught me the logistics and leadership skills required to
-    manage large events for 150 people.
-    I have also learned to make events accessible to everyone by asking
-    questions so I can understand the needs of all attendees.
+    manage large events for up to 60 people.
+    I have also learned to make events accessible to everyone, and to ask
+    questions so I can meet the needs of all attendees.
 

### 2021-09-20 latex/nserc/proposal.tex
Wrote a better intro for the proposal based on darby's advice

@@
 %---------- Title -----------%
-\title{Energy Regulation for Biologically-inspired Robotics}
+\title{Energy Regulation for Biologically-Inspired Robotics}
 \author{}
@@
 
-\vspace{-1cm} % Move the following text up
+\vspace{-1.2cm} % Move the following text up
 %\textbf{What is your area of research? What sort of problems do you want to
@@
 %---------- Main Content ----------%
+A robotic monkey swings through trees in a Mexican jungle, capturing valuable
+wildlife data without disturbing the local environment. 
+In the nearby city, a robotic police dog runs through the streets, weaving through
+pedestrians' legs to catch a thief.
+It passes a traveling circus group, where robot gymnasts perform elaborate
+routines in front of an awestruck audience.
+This world sounds like science fiction, but it is truly within sight;
+our research will enable roboticists to accurately recreate human and animal
+behaviour in their robots.
+The future we imagined will become a present-day reality.
+
 \begin{large}
@@
 
-When a gymnast wants to do backflips around a horizontal bar, they start by
-performing a technique called a ``giant" \cite{usagym_giant}. 
-The gymnast swings their legs based on their current speed and position, thereby
-generating enough momentum to rotate around the bar.
+Before can explain how these humanoid and animal-like robots will work, we
+must first understand how humans and animals move. 
+Take the example of a human gymnast who  wants to do backflips around a
+horizontal bar. 
+The gymnast starts by swinging their legs based on their current speed and
+position in a technique called a ``giant" \cite{usagym_giant}. 
+This allows them to generate enough momentum to rotate around the bar.
 
@@
 
+\newpage
+
 \begin{large} \textbf{Proposal} \end{large}
 
-Virtual constraints have proven useful for generating biologically-realistic
+Virtual constraints have proven useful for generating biologically realistic
 behaviour in robotic systems.
@@
 removing energy will allow for more expressive motion, and may become a standard
-technique for controlling biologically inspired robots.
+technique for controlling biologically-inspired robots.
 

### 2021-09-20 latex/nserc/contributions.tex
Updated the contributions to be better.

--- 2021-09-19 latex/nserc/contributions.tex
@@
 I contributed to this report by producing all the theoretical and simulation
-results, though I could not have done this without the abundant advice of my
-supervisor, Fady Alajaji. 
+results, though I received an abundance of advice from my supervisor, Fady
+Alajaji. 
 I also wrote the report in its entirety.
@@
         I wrote networking code in Python to interface with the
-        device, and validatet the hardware's capabilities for myself by
-        creating appropriate testing environments. 
+        device, and validated the hardware's capabilities for myself by
+        creating appropriate test environments. 
         All these skills will be beneficial for my PhD research when I build a
-        robot.
+        robotic monkey.
 
     \item \textbf{Undergraduate Thesis, Queen's University, 2017 -- 2018}.
-        For my undergraduate thesis, I worked with my peers to find an
-        application for existing theoretical results. 
-        During this project I also learned to perform a basic literature review,
+        For my undergraduate thesis, I worked with my peers to design an image
+        filtering algorithm by applying theoretical results from information theory.
+        During this project I learned to perform a basic literature review,
         and to manage deadlines with a supervisor.
-        In terms of engineering skills, I learned to write reports in LaTeX
+        I also learned to write reports in LaTeX,
         and my thesis group worked together on code optimization: we learned to
@@
     \item \textbf{Software Developer, Neptec Technologies Corp., 2016 -- 2017}.
-        My experience at Neptec was integral to my career, as it is what drove
+        My experience at Neptec was integral to my career because it convinced
         me to pursue research in robotics.
-        Working at Neptec exposed me to several areas of electrical and computer
+        Working at Neptec exposed me to many areas of electrical and computer
         engineering.
@@
         was a control systems project.
-        I loved this project, and I decided to do a master's (and a PhD) so I
+        After completing this project, I decided to do a master's (and a PhD) so I
         could learn as much about control as possible.
-        I also learned to work on large software projects, where multiple people work
-        together on the same topic; 
+        At Neptec I also learned to work on large software projects, where multiple people work
+        together on the same code; 
         and I became extremely proficient at software development, a skill which
@@
 
-Over the course of my career, I have enjoyed several extracurricular roles which
-developed my communication, interpersonal, and leadership skills.
-What follows is a brief description of some of these roles.
+Over the course of my career, I have enjoyed several extracurricular
+roles which developed my communication, interpersonal, and leadership skills.
+Here is a brief description of some of these roles.
 \begin{itemize}
@@
     I have already organized 13 events for members of the college, and will
-    continue to organize biweekly events throughout the year.
+    continue to organize events throughout the year.
     This role has taught me the logistics and leadership skills required to
-    manage large events for 150 people.
-    I have also learned to make events accessible to everyone by asking
-    questions so I can understand the needs of all attendees.
+    manage large events for up to 60 people.
+    I have also learned to make events accessible to everyone, and to
+    communicate with attendees so I can meet their needs.
 
@@
 \item \textbf{Science Payload Designer, Queen’s University Satellite Design
-    Team, 2017 -- 2018}. 
-    In this role, I taught C++ and information theory to the team’s software
-    developers. I learned project management skills by organizing
-    the satellite codebase and handling code reviews, and I took on a supportive
-    and mentorship role for my peers rather than writing software myself.
+    Team (QSAT), 2017 -- 2018}. 
+    I took on a mentorship role at QSAT, where I taught C++, Git, and information
+    theory to the team’s software developers. I helped them organize
+    the satellite codebase, managed all code reviews, and provided advice for my
+    peers rather than writing software myself.
 
-\item \textbf{Head Programming Mentor, FIRST Team 2809, 2013 -- 2015}. 
-    I mentored high school students in Object Oriented programming for robotics
-    software. Together we programmed a robot in under 6 weeks and placed 3rd at
-    the Ontario provincial competition in 2014.
+\item \textbf{Head Programming Mentor, FRC Team 2809, 2013 -- 2015}. 
+    In this role, I mentored high school students in Object Oriented programming
+    for robotics software. I taught students to program a robot in under 6
+    weeks, which earned us 3rd place at the Ontario provincial competition in
+    2014.
     In this role I learned to lead my students by giving them agency over the

### 2021-09-20 latex/nserc/contributions.tex
Editing pass, two more passes before submission

@@
 clustering datapoints.
-I contributed to this report by producing all the theoretical and simulation
-results, though I received an abundance of advice from my supervisor, Fady
-Alajaji. 
-I also wrote the report in its entirety.
+I wrote this report in its entirety, and I contributed all the theoretical and
+simulation results (with abundant advice from my supervisor, Fady Alajaji).
 This report was used by two undergraduate groups at Queen's University for their
-honours theses in 2019, and it motivated further research [1].
+honours theses in 2019, and it motivated further research in [1].
 
@@
         students to write software which reads data from a radar device.
-        This project taught me several engineering skills.
-        First, I improved my project management skills by keeping our group
-        meetings organized, and by separating the tasks in such a way
-        that everyone was learning something from the project.
+        This project improved my project management skills, as I kept our group
+        meetings organized and separated the project tasks in such a way
+        that everyone was challenged appropriately.
         I learned to read hardware documentation and to wire a device from
@@
         creating appropriate test environments. 
-        All these skills will be beneficial for my PhD research when I build a
-        robotic monkey.
+        All these skills will be beneficial for my PhD research, where I will
+        collaborate with labmates to build a robotic monkey.
 
@@
         I also learned to write reports in LaTeX,
-        and my thesis group worked together on code optimization: we learned to
+        and my group worked together on code optimization: we learned to
         convert equations into matrix form to take advantage of hardware
@@
         could learn as much about control as possible.
-        At Neptec I also learned to work on large software projects, where multiple people work
-        together on the same code; 
+        At Neptec I also learned to work on large software projects,
         and I became extremely proficient at software development, a skill which
-        has served me well in my research experiments on physical robots.
+        has served me well in all aspects of my research.
 
@@
     kind and non-judgemental, which makes students feel
-    comfortable enough to ask whatever question crosses their mind.
+    comfortable enough to ask more questions.
 
@@
 Obtaining an NSERC scholarship will provide me with the financial liberty to
-participate in more volunteer roles, where I can better myself while giving back to my
+participate in more volunteer roles, where I can improve my skills while giving back to my
 community.

### 2021-09-20 latex/nserc/proposal.tex
Editing pass, two more passes before submission

@@
 %---------- Main Content ----------%
-A robotic monkey swings through trees in a Mexican jungle, capturing valuable
+A robotic monkey swings through jungle trees, capturing valuable
 wildlife data without disturbing the local environment. 
-In the nearby city, a robotic police dog runs through the streets, weaving through
+In a nearby city, a robotic police dog runs through the streets, weaving through
 pedestrians' legs to catch a thief.
@@
 This world sounds like science fiction, but it is truly within sight;
-our research will enable roboticists to accurately recreate human and animal
-behaviour in their robots.
-The future we imagined will become a present-day reality.
+the research we propose below will enable roboticists to accurately recreate
+human and animal behaviour in their robots.
+This world we imagined can become a present-day reality.
 

### 2021-09-20 latex/nserc/proposal.tex
Final version of proposal, ready to submit.

@@
 routines in front of an awestruck audience.
-This world sounds like science fiction, but it is truly within sight;
+This world sounds like science fiction, but it is truly within sight:
 the research we propose below will enable roboticists to accurately recreate
@@
 
-Before can explain how these humanoid and animal-like robots will work, we
+Before we can explain how these humanoid and animal-like robots will work, we
 must first understand how humans and animals move. 
@@
 horizontal bar. 
-The gymnast starts by swinging their legs based on their current speed and
-position in a technique called a ``giant" \cite{usagym_giant}. 
+The gymnast starts by swinging their legs in a technique called a ``giant"
+\cite{usagym_giant}, during which they move their legs based on their current
+position and velocity \cite{pendulum_length_giant_gymnastics}. 
 This allows them to generate enough momentum to rotate around the bar.
@@
 and we will build a robotic monkey which uses VNHCs to swing on monkey bars just
-like a real monkey would swing through trees.
+like a real monkey would swing through jungle trees.
 
@@
 %---------- Bibliography ----------%
+\pagenumbering{arabic}
 \printbibliography

### 2021-09-20 latex/nserc/contributions.tex
Almost-final editing pass of contributions

@@
         was a control systems project.
-        After completing this project, I decided to do a master's (and a PhD) so I
+        After completing this work, I decided to do a master's (and a PhD) so I
         could learn as much about control as possible.
-        At Neptec I also learned to work on large software projects,
+        At Neptec I also learned to work with large codebases,
         and I became extremely proficient at software development, a skill which
@@
     I took on a mentorship role at QSAT, where I taught C++, Git, and information
-    theory to the team’s software developers. I helped them organize
+    theory to the team's software developers. I helped them organize
     the satellite codebase, managed all code reviews, and provided advice for my

### 2021-09-20 latex/nserc/contributions.tex
Final version of contributions.

@@
     2014.
-    In this role I learned to lead my students by giving them agency over the
-    project, and to communicate with people from all over the world in
-    high-stress situations.
+    In this role I became a better leader by listening to my students' feedback,
+    and I learned to communicate effectively with people from all over the world
+    in high-stress situations.
 

### 2021-09-20 latex/nserc/contributions.tex
Updated the contributions ot be letter paper, not A4 paper

@@
 
-\documentclass[a4paper,12pt]{article}
+\documentclass[letterpaper,12pt]{article}
 
@@
 %----------- Choose Margins ----------%
-\usepackage[a4paper,margin=1.87cm]{geometry}
+\usepackage[letterpaper,margin=1.87cm]{geometry}
 
@@
 
-\item \textbf{Science Payload Designer, Queen’s University Satellite Design
-    Team (QSAT), 2017 -- 2018}. 
-    I took on a mentorship role at QSAT, where I taught C++, Git, and information
-    theory to the team's software developers. I helped them organize
-    the satellite codebase, managed all code reviews, and provided advice for my
-    peers rather than writing software myself.
+%\item \textbf{Science Payload Designer, Queen’s University Satellite Design
+%    Team (QSAT), 2017 -- 2018}. 
+%    I took on a mentorship role at QSAT, where I taught C++, Git, and information
+%    theory to the team's software developers. I helped them organize
+%    the satellite codebase, managed all code reviews, and provided advice for my
+%    peers rather than writing software myself.
 

### 2021-09-20 latex/nserc/proposal.tex
Made proposal in letter paper size, not A4 size
Final update of proposal, thanks to Rushay

@@
 %----------- Choose Margins ----------%
-\usepackage[a4paper,margin=1.87cm]{geometry}
+\usepackage[margin=1.87cm]{geometry}
 
@@
 
-\vspace{-1.2cm} % Move the following text up
+\vspace{-1.5cm} % Move the following text up
 %\textbf{What is your area of research? What sort of problems do you want to
@@
 There are two types of constraints on a mechanical system:
-\textit{holonomic} constraints restrict position (e.g. a snake robot can slither
-on the ground but cannot fly),
-while \textit{nonholonomic} constraints restrict both position and
-velocity (e.g. an autonomous car with normal wheels cannot slide sideways).
+\textit{holonomic} constraints restrict position (e.g. an autonomous car must
+remain on the ground),
+while \textit{nonholonomic} constraints restrict position and
+velocity (e.g.~an autonomous car with normal wheels cannot slide sideways).
 It is often possible to use a robot's actuators to enforce a desired constraint
@@
 \cite{xingbo_thesis}.
-This is where virtual nonholonomic constraints (VNHCs) are most useful.
+This is where virtual nonholonomic constraints (VNHCs) become useful.
 
@@
 \cite{nhvc_incline_walking}.
-
-\newpage
 
@@
 like a real monkey would swing through jungle trees.
+This robot monkey is representative of a variety of robotic systems, and will be
+a useful platform on which to test our mathematical theory.
 

### 2021-10-06 latex/paper/main.tex
Revision fixing theorem statement and proof, plus other smaller changes.

--- 2021-08-26 latex/paper/main.tex
@@
 \usepackage{amsmath}
+%\usepackage{amsthm}
 \usepackage{amsfonts}
@@
 \usepackage{cite}
-
+\usepackage{xspace}
+\usepackage{enumerate}
 %---------- More packages ----------%
@@
 %---------- Bibliography Style ----------%
-\bibliographystyle{IEEEtran}
+\bibliographystyle{IEEEtranS}
 
@@
 \newtheorem{assm}{Assumption} % Assumptions do not reset 
-
+\newtheorem{rem}{Remark}
 %---------- Special Commands ----------%
@@
 \DeclareMathOperator{\Diag}{diag}
+\DeclareMathOperator{\interior}{int}
 \newcommand*{\rank}[1]{\Rank\left(#1\right)}
@@
 \newcommand*{\R}{\mathbb{R}}
+\renewcommand*{\Re}{\mathbb{R}}
 \newcommand*{\n}{\mathbf{n}}
@@
 \newcommand*{\etal}{\MakeLowercase{\textit{et al.~}}}
+\newcommand*{\cU}{\mathcal{U}}
+\newcommand*{\cV}{\mathcal{V}}
+\newcommand*{\cO}{\mathcal{O}}
+\newcommand*{\cK}{\mathcal{K}}
+\newcommand*{\cR}{\mathcal{R}}
+\newcommand*{\cT}{\mathcal{T}}
+\newcommand*{\cI}{\mathcal{I}}
+\newcommand*{\cP}{\mathcal{P}}
+\newcommand*{\cQ}{\mathcal{Q}}
+\newcommand*{\cX}{\mathcal{X}}
+\newcommand*{\rone}{r}%{r_1}
+\newcommand*{\rtwo}{r}%{r_2}
+\newcommand*{\ri}{r}%{r_i}
+\newcommand*{\thetaone}{\theta}%{\theta_1}
+\newcommand*{\thetatwo}{\theta}%{\theta_2}
+\newcommand*{\thetai}{\theta}%{\theta_i}
+\newcommand*{\Ze}{\mathbb{Z}}
+\newcommand*{\Ne}{\mathbb{N}}
+\newcommand*{\eqdef}{\vcentcolon=}
+
+\newcommand*{\vhc}{\textsc{vhc}\xspace}
+\newcommand*{\vhcs}{\textsc{vhc}s\xspace}
+\newcommand*{\vnhc}{\textsc{vnhc}\xspace}
+\newcommand*{\vnhcs}{\textsc{vnhc}s\xspace}
+
 %/========== /Preamble ==========/%
@@
 \begin{abstract}
-    In this article we study virtual nonholonomic constraints, which are
-    relations between the generalized coordinates and generalized momenta of a
+    In this article we study virtual nonholonomic constraints, 
+    relations between the generalized coordinates and momenta of a
     mechanical system that can be enforced via feedback control.
     We design a constraint which emulates gymnastics giant motion in an
-    acrobot, and rigorously prove that this constraint will inject or dissipate
-    energy.
-    This constraint is tested in simulation and on a real-world acrobot,
+    acrobot, and prove that this constraint can inject or dissipate
+    energy based on the sign of a design parameter.
+    The proposed constraint is tested both in simulation and experimentally on a real-world acrobot,
     demonstrating highly effective energy regulation properties and robustness
@@
 function of the gymnast's limb angle.
-Labeling the pendulum length by \(r\) and the gymnast's body orientation
+Labelling the pendulum length by \(r\) and the gymnast's body orientation
 by \(\theta\), they observed experimentally that the value \(\dot{r}/r\) has
@@
 After testing several gymnasts under a variety of experimental conditions, 
-they discovered that the peak value of \(\dot{r}/r\) occured at the same fixed
+they discovered that the peak value of \(\dot{r}/r\) occurred at the same fixed
 value of \(\dot{\theta}/\theta\) for all gymnasts.
@@
 the legs. 
-The acrobot is actuated exclusively at the center joint (the hip).
-Controlling the acrobot is a nontrivial task because it is not feedback
-linearizable \cite{nonlinear_controllers_nonintegrable_acrobot}. 
+The acrobot is actuated exclusively at the centre joint (the hip).
 To solve the swingup problem, one might begin by designing a leg controller
@@
 
-Virtual nonholonomic constraints (VNHCs) have been used for human-robot interaction
+Virtual nonholonomic constraints (\vnhcs) have been used for human-robot interaction
 \cite{vnhc_human_robot_cooperation,psd_based_vnhc_redundant_manipulator,haptic_vnhc},
@@
 \cite{nhvc_incline_walking}.
-In particular, VNHCs may be capable of injecting and
+In particular, \vnhcs may be capable of injecting and
 dissipating energy from a system in a robust manner, all while producing
@@
 injects energy into the acrobot through human-like giant motion.
-In particular, any acrobot constrained by this VNHC is guaranteed to perform
+In particular, any acrobot constrained by this \vnhc is guaranteed to perform
 rotations around the bar.
-Under suitable conditions on the acrobot's physical parameters, this VNHC
+Under suitable conditions on the acrobot's physical parameters, this \vnhc
 also enables the acrobot to rotate around the bar at a desired speed.
-We will provide simulations to show how one can use VNHCs to regulate energy,
+We will provide simulations to show how one can use \vnhcs to regulate energy,
 along with experimental results which demonstrate the robustness of this
@@
 For \(T > 0\), the set of real numbers modulo \(T\) is denoted \(\Rt{T}\), with
-\(\Rt{\infty} := \R\).
+\(\Rt{\infty} \eqdef \R\). %The interior of a set $S$ is denoted $\interior S$.
 The gradient of a matrix-valued function 
@@
 derivatives, 
-\(\nabla_xA := [\partial A/\partial x_1; \cdots; \partial A/\partial x_m] \in
+\(\nabla_xA \eqdef [\partial A/\partial x_1; \cdots; \partial A/\partial x_m] \in
 \R^{nm \times n}\).
-%\(\nabla_xA := [\pdiff{A}{x_1};\ldots;\pdiff{A}{x_m}] \in \R^{nm \times n}\).
+%\(\nabla_xA \eqdef [\pdiff{A}{x_1};\ldots;\pdiff{A}{x_m}] \in \R^{nm \times n}\).
 Given two matrices \(A \in \R^{n \times m}\) and \(B \in \R^{r \times s}\), the
@@
 \begin{equation}\label{eqn:poisson-bracket}
-    [f,g] := \sum \limits_{i=1}^n \pdiff{f}{p_i}\pdiff{g}{q_i} - 
+    [f,g] \eqdef \sum \limits_{i=1}^n \pdiff{f}{p_i}\pdiff{g}{q_i} - 
         \pdiff{f}{q_i}\pdiff{g}{p_i}
@@
 otherwise.
-Finally, we say a function \(R(I)\) is \(O(I^2)\)
-if \(\lim_{I \to 0} R(I)/I = 0\).
+Finally, we say a function \(\Delta(I)\) is \(O(I^2)\)
+if \(\lim_{I \to 0} \Delta(I)/I = 0\).
 
@@
 \eqref{eqn:acrobot-hamiltonian}.
-For shorthand, we write \(c_u := \cos(q_u)\), \(c_a := \cos(q_a)\), and 
-\(c_{ua} := \cos(q_u + q_a)\); likewise, \(s_u := \sin(q_u)\), 
-\(s_a := \sin(q_a)\), and \(s_{ua} := \sin(q_u + q_a)\).
+For shorthand, we write \(c_u \eqdef \cos(q_u)\), \(c_a \eqdef \cos(q_a)\), and 
+\(c_{ua} \eqdef \cos(q_u + q_a)\); likewise, \(s_u \eqdef \sin(q_u)\), 
+\(s_a \eqdef \sin(q_a)\), and \(s_{ua} \eqdef \sin(q_u + q_a)\).
 \begin{align}\label{eqn:acrobot-hamiltonian}
@@
 \(p_a\), representing a torque acting on the hip joint.
-Let us now define what it means for a mechanical system to gain or lose energy.
-
-\begin{defn}\label{defn:energy-gain}
-    Let \(\mathcal{Q}\) be an
-    \(n\)-dimensional smooth manifold. 
-    Let \(f : \mathcal{Q} \rightarrow T\mathcal{Q}\) be a smooth vector
-    field and let \(D \subset \mathcal{Q}\) be open.
-    The system described by \(\dot{x} = f(x)\) 
-    \textit{gains energy on \(D\)} if, 
-    for all compact sets \(K \subset D\) and for almost every initial
-    condition \(x(0) \in K\), there exists \(T > 0\) such
-    that \(x(t) \notin K\) for all \(t > T\).
-    The system \textit{loses energy on \(D\)} if it gains energy in
-    negative-time.
-\end{defn}
-
-The definition above applies to vector fields that are a not necessarily
-Hamiltonian, and therefore might not have an associated notion of ``energy". 
-The terms ``gain energy" and ``lose energy" are to be interpreted loosely in
-this context.
-Our goal is to design a smooth function \(f : \Sone \times \R\) such that the
-relation \(q_a = f(q_u,p_u)\) for system \eqref{eqn:acrobot-hamiltonian}
-can be enforced asymptotically via feedback control (in Section \ref{sec:vnhc}
-we call this a regular virtual nonholonomic constraint).  
-We will further require that the dynamics of the acrobot, when the relation
-holds, gain or lose energy on some set \(D\) in the sense of Definition
-\ref{defn:energy-gain}. 
-Note that any system satisfying Definition \ref{defn:energy-gain}
-can have unstable equilibria on \(D\), but not limit cycles nor closed orbits.
-
-%/========== VNHC ==========/%
-\section{Theory of VNHCs}\label{sec:vnhc}
-Before embarking on our design problem, we must summarize the relevant
+
+Our goal is to design a smooth function \(f : \Sone \times \R\) such that the relation \(q_a = f(q_u,p_u)\) for system \eqref{eqn:acrobot-hamiltonian} can be enforced asymptotically via feedback control (in Section \ref{sec:vnhc} we call this a  virtual nonholonomic constraint).
+We will further require that the dynamics of the acrobot, when the relation holds, gain or lose energy in a sense that will be defined precisely in Section~\ref{sec:energy-inject-diss}.
+
+%/========== \vnhc ==========/%
+\section{Preliminaries on \vnhcs}\label{sec:vnhc}
+Before embarking on the design problem, we must summarize the relevant
 theory of virtual nonholonomic constraints for a class of mechanical systems we
-call ``simply actuated hamiltonian systems". 
+call ``simply actuated Hamiltonian systems". 
 The results we provide in Section \ref{sec:vnhc-vnhc} are not novel:
@@
 is smooth and positive definite for all \(q \in \mathcal{Q}\).
-The \textit{conjugate of momentum} to \(q\) is the vector
-\(p := \partial\mathcal{L}/\partial\dot{q} = D(q) \dot{q} \in \R^n\).
+The \textit{conjugate momentum} to \(q\) is the vector
+\(p \eqdef \partial\mathcal{L}/\partial\dot{q} = D(q) \dot{q} \in \R^n\).
 As per \cite{landau_mechanics}, 
@@
 
-In light of Lemma \ref{lemma:B-orthogonal} there is no loss of generality in
-assuming that the input matrix is left semi-orthogonal, which means that
+In light of Lemma \ref{lemma:B-orthogonal}, there is no loss of generality in
+assuming that the input matrix \(B\) is left semi-orthogonal, which means that
 \(B\tpose\) is right semi-orthogonal.
-Now, let \(\mathbf{B} := [B^\perp; B\tpose]\) where 
-\(B^\perp \in \R^{(n-k)\times k}\) is a left annihilator of \(B\), i.e.,
+Now, let \(\mathbf{B} \eqdef [B^\perp; B\tpose]\) where 
+\(B^\perp \in \R^{(n-k)\times k}\) is a full-rank left annihilator of \(B\), i.e.,
 \(B^\perp B = \Zmat{(n-k) \times k}\).
@@
     where 
-    \(\Minv(\tilde{q}) := 
+    \(\Minv(\tilde{q}) \eqdef 
     (\mathbf{B}\tpose)\inv D^{-1}(\mathbf{B}\tpose \tilde{q})\mathbf{B}\inv\)
     and
-    \(V(\tilde{q}) := P(\mathbf{B}\tpose \tilde{q})\).
+    \(V(\tilde{q}) \eqdef P(\mathbf{B}\tpose \tilde{q})\).
 \end{thm}
@@
 whose input matrix is \([\Zmat{(n-k)\times k}; \Id{k}]\) a 
-\textit{simply actuated hamiltonian system}.
+\textit{simply actuated Hamiltonian system}.
 The first \((n-k)\) configuration variables in \(\tilde{q}\), labelled \(q_u\),
@@
 
-Griffin and Grizzle \cite{nhvc_dynamic_walking} were the first to define
-relative degree two nonholonomic constraints which can be enforced
-through state feedback.
-Horn \etal later extended their results in
-\cite{hybrid_zero_dynamics_bipedal_nhvcs} to derive the constrained dynamics for
-a certain class of mechanical systems.
-These researchers made use of the unactuated conjugate of momentum, but they
-developed their results in the Lagrangian framework.
-In particular, they focused on Lagrangian systems with degree of underactuation
-one.
-We will now present a special case of \cite{hybrid_zero_dynamics_bipedal_nhvcs}
-for simply actuated hamiltonian systems, so that the theory we apply to the
-acrobot is provided in its clearest form.
-After we finish this summary we will clarify the relationship between our
-material and that of \cite{hybrid_zero_dynamics_bipedal_nhvcs}.
-
-For the rest of this section we take the system of inquiry to be a
-Hamiltonian mechanical system in simply actuated coordinates, as in
-\eqref{eqn:simple-hamiltonian}.
+Griffin and Grizzle \cite{nhvc_dynamic_walking} were the first to define relative degree two nonholonomic constraints which can be enforced through state feedback.
+Horn \etal later extended their results in \cite{hybrid_zero_dynamics_bipedal_nhvcs} to derive the constrained dynamics for a certain class of mechanical systems.
+These researchers made use of the unactuated conjugate momentum, but they developed their results in the Lagrangian framework.
+In particular, they focused on Lagrangian systems with degree of underactuation one.
+We will now present a special case of \cite{hybrid_zero_dynamics_bipedal_nhvcs} for simply actuated Hamiltonian systems, so that the theory we apply to the acrobot is provided in its clearest form.
+After we finish this summary we will clarify the relationship between our material and that of \cite{hybrid_zero_dynamics_bipedal_nhvcs}.
+
+The rest of this section refers to the simply actuated system \eqref{eqn:simple-hamiltonian}.
 For simplicity of notation, we relabel \((\tilde{q},\tilde{p})\) to \((q,p)\).
+We assume henceforth that $(q,p) \in \cQ \times \Re^n$.
 
 \begin{defn}\label{defn:vnhc}
-    A \textit{virtual nonholonomic constraint} (VNHC) \textit{of order \(k\)} is a
-    relation \(h(q,p) = 0\) where \(h : \mathcal{Q}\times\R^n \rightarrow \R^k\) is
-    \(C^2\), \(\rank{\left[ dh_q,\, dh_p \right]} = k\) for all 
-    \((q,p) \in h\inv(0)\), and there exists a feedback controller \(\tau(q,p)\)
-    rendering the \textit{constraint manifold} \(\Gamma\) invariant,
-    where
+    A \textit{virtual nonholonomic constraint} (\vnhc) \textit{of order \(k\)} is a     relation \(h(q,p) = 0\) where \(h : \mathcal{Q}\times\R^n \rightarrow \R^k\) is     \(C^2\), \(\rank{\left[ dh_q,\, dh_p \right]} = k\) for all      \((q,p) \in h\inv(0)\), and there exists a feedback controller \(\tau(q,p)\)     rendering the \textit{constraint manifold} \(\Gamma\) invariant,     where
     \begin{equation}
@@
 
-The constraint manifold is a \(2(n-k)\)-dimensional
-closed embedded submanifold of \(\mathcal{Q} \times \R^n\).
-A VNHC thereby allows us to study a reduced-order model of the system: it reduces
-the original \(2n\) differential equations to \(2(n-k)\) equations.
-In particular, if \(k = (n-1)\), the constraint manifold is \textit{always}
-2-dimensional and its dynamics can be displayed on a plane.
-
-In order to enforce the constraint \(h(q,p) = 0\), we want to asymptotically
-stabilize the set \(\Gamma\).
+The constraint manifold is a \(2(n-k)\)-dimensional closed embedded submanifold of \(\mathcal{Q} \times \R^n\).
+A \vnhc thereby allows us to study a reduced-order model of the system: it reduces the original \(2n\) differential equations to \(2(n-k)\) equations.
+In particular, if the mechanical system has degree of underactuation one, i.e., \(k = (n-1)\), the constraint manifold is \textit{always} two-dimensional.
+
+In order to enforce the constraint \(h(q,p) = 0\), we want to asymptotically stabilize the set \(\Gamma\).
 To see when this is possible, let us define the error output \(e = h(q,p)\).
-If any component of \(e_i\) has relative degree 1, we may not be able
-to stabilize \(\Gamma\): we can always guarantee \(e_i \to 0\), but not
-necessarily \(\dot{e}_i \to 0\).
-It is for this reason that we define the following special type of VNHC.
+If any component of \(e_i\) has relative degree 1, we may not be able to stabilize \(\Gamma\): we can always guarantee \(e_i \to 0\), but not necessarily \(\dot{e}_i \to 0\).
+It is for this reason that we define the following special type of \vnhc.
 
 \begin{defn}
-    A VNHC \(h(q,p) = 0\) of order \(k\) is \textit{regular} if the output 
+    A \vnhc \(h(q,p) = 0\) of order \(k\) is \textit{regular} if the output 
     \(e = h(q,p)\) is of relative degree \(\{2,2.\ldots,2\}\) everywhere on the
@@
 \cite{nhvc_dynamic_walking,hybrid_zero_dynamics_bipedal_nhvcs}
-observed that relations which use only the unactuated conjugate of momentum
+observed that relations which use only the unactuated conjugate momentum
 often have vector relative degree \(\{2,\ldots,2\}\).
@@
 \begin{align}
-    \mathcal{A}(q,p_u) &:= dh_q(q,p_u) \Minv(q) 
+    \mathcal{A}(q,p_u) &\eqdef dh_q(q,p_u) \Minv(q) 
         ,\\
-    \mathcal{M}(q,p) &:= (\Id{n-k} \otimes p\tpose)\nabla_{q_u}\Minv(q) 
+    \mathcal{M}(q,p) &\eqdef (\Id{n-k} \otimes p\tpose)\nabla_{q_u}\Minv(q) 
     .
@@
     A relation \(h(q,p) = 0\) for system \eqref{eqn:simple-hamiltonian}
-    is a regular VNHC of order \(k\) if and only if 
+    is a regular \vnhc of order \(k\) if and only if 
     \(dh_{p_a} = \Zmat{k \times k}\) 
@@
     \begin{equation}\label{eqn:decoupling-matrix}
-        \big(\mathcal{A}(q,p_u) - dh_{p_u}\mathcal{M}(q,p)\big)\simpleB
+        H(q,p) \eqdef \big(\mathcal{A}(q,p_u) - dh_{p_u}\mathcal{M}(q,p)\big)\simpleB
          ,
@@
     If \(dh_{p_a} \neq \Zmat{k\times k}\) for some \((q,p) \in \Gamma\), 
-    then \(\tau\) appears in \(\dot{e}\) and the VNHC is not of relative degree
+    then \(\tau\) appears in \(\dot{e}\) and the \vnhc is not of relative degree
     \(\{2,\ldots,2\}\). Suppose now that \(dh_{p_a} = \Zmat{k\times k}\).
@@
         \Id{k}]\tau\).
-    Hence,
-    \[
-       \ddot{e} = (\star) +
-       \big(\mathcal{A}(q,p_u) - dh_{p_u}\mathcal{M}(q,p)\big) \simpleB \tau
-        ,
-    \]
-    which we write as \( \ddot{e} = E(q,p) + H(q,p)\tau\) for appropriate \(E\)
-    and \(H\).
-    From the definition of regularity, the VNHC \(h\) is regular 
+    Hence, $\ddot e$ has the form \( \ddot{e} = E(q,p) + H(q,p)\tau\) for appropriate \(E\).
+    From the definition of regularity, the \vnhc \(h\) is regular 
     when \(e\) is of relative degree \(\{2,\ldots,2\}\), which is true 
-    if and only if the matrix premultiplying \(\tau\) is nonsingular, and hence
-    that \(H\) is invertible. This proves the theorem.
+    if and only if the matrix $H(q,p)$ premultiplying \(\tau\) is nonsingular. This proves the theorem.
 \end{proof}
 
-Under additional mild conditions (see \cite{vhcs_for_el_systems}), a regular VNHC of
-order \(k\) can be stabilized by the output-linearizing state-feedback
+Under additional mild conditions (see \cite{vhcs_for_el_systems}), a regular \vnhc of
+order \(k\) can be stabilized by the input-output feedback linearizing
 controller
@@
 where \(k_p, k_d > 0\) are control parameters which can be tuned on the
-resulting error dynamices \(\ddot{e} = -k_p e - k_d \dot{e}\).
+resulting error dynamics \(\ddot{e} = -k_p e - k_d \dot{e}\).
 
@@
 \begin{thm}\label{thm:zero-dynamics}
-    Let \(\mathcal{H}\) be a hamiltonian system in simply actuated form
+    Let \(\mathcal{H}\) be a Hamiltonian system in simply actuated form
     \eqref{eqn:simple-hamiltonian} satisfying
     Assumption \ref{assm:inertially-actuated}. 
-    Let \(h(q,p_u) = q_a - f(q_u,p_u)\) be a regular VNHC of order \(k\) with
+    Let \(h(q,p_u) = q_a - f(q_u,p_u)\) be a regular \vnhc of order \(k\) with
     constraint manifold \(\Gamma\). 
@@
     \begin{multline}\label{eqn:g-qpu}
-        g(q_u,p_u) := \left( \mathcal{A}(q,p_u)
+        g(q_u,p_u) \eqdef \left( \mathcal{A}(q,p_u)
         \begin{bmatrix} \Zmat{(n-k)\times k} \\ \Id{k} \end{bmatrix}\right)\inv 
@@
 %    \begin{align}\label{eqn:g-qpu}
-%        &g(q_u,p_u) := 
+%        &g(q_u,p_u) \eqdef 
 %        \left(\mathcal{A}(q,p_u)[\Zmat{(n-k)\times k};\Id{k}] \right)\inv 
@@
 
-We conclude this section by formalizing the notion of energy
-injection/dissipation for VNHCs.
-
-\begin{defn}\label{defn:energy-injection}
-    A regular VNHC \(h(q,p_u) = q_a - f(q_u,p_u)\) with constraint manifold \(\Gamma\)
-    \textit{injects (dissipates) energy on \(D \subset \Gamma\)} if the
-    constrained dynamics gain (lose) energy everywhere on \(D\) according to
-    Definition \ref{defn:energy-gain}, except possibly on a set of measure
-    zero.
-\end{defn}
+%% We conclude this section by formalizing the notion of energy
+%% injection/dissipation for \vnhcs.
+
+%% \begin{defn}\label{defn:energy-injection}
+%%     A regular \vnhc \(h(q,p_u) = q_a - f(q_u,p_u)\) with constraint manifold \(\Gamma\)
+%%     \textit{injects (dissipates) energy on \(D \subset \Gamma\)} if the
+%%     constrained dynamics~\eqref{eqn:qpu-dynamics} gain (lose) energy everywhere on \(D\) according to
+%%     Definition \ref{defn:energy-gain}.
+%% \end{defn}
 
 \textbf{Comparison with existing literature}: Horn \etal provide the constrained
-dynamics for VNHCs in \cite{nhvc_incline_walking}.
+dynamics for \vnhcs in \cite{nhvc_incline_walking}.
 Their assumption \textbf{H2} is what we call regularity, and our requirement
@@
 This explicit representation will be beneficial when we apply the theory of
-VNHCs to the acrobot.
+\vnhcs to the acrobot.
 
 %/========== Acrobot ==========/%
-\section{The Proposed Acrobot VNHC}\label{sec:acrobot}
-
-Our goal in this article is to design a VNHC which injects energy into the
+\section{The Proposed Acrobot \vnhc}\label{sec:acrobot}
+
+The goal in this article is to design a \vnhc which injects energy into the
 acrobot by means of a giant-like motion.
@@
 the system above is already in simply actuated form.
-Its state space is \(\mathcal{Q} \times \mathcal{P}\) where
-\(\mathcal{Q} = \Sone \times \Sone\), and
-\(\mathcal{P} = \R \times \R\).
+Its state space is \(\Sone \times \Re \times \Re \times \Re\).
 We can therefore apply the theory from Section \ref{sec:vnhc} to design
-a VNHC of the form \(q_a = f(q_u,p_u)\) (i.e., a VNHC 
+a \vnhc of the form \(q_a = f(q_u,p_u)\) (i.e., a \vnhc 
 \(h(q,p_u) = q_a - f(q_u,p_u) = 0\)).
-Since we need the VNHC to be regular, the following proposition will be
+Since we need the \vnhc to be regular, the following proposition will be
 useful.
@@
     with \(f \in C^2\left(\R; \Sone\right)\) is a regular
-    VNHC of order 1 for the acrobot in \eqref{eqn:acrobot-hamiltonian}.
+    \vnhc of order 1 for the acrobot in \eqref{eqn:acrobot-hamiltonian}.
 \end{prop}
@@
     \(((1+c_a)\partial_{q_u}f(q_u,p_u) + (3+2c_a))/(ml^2(2-c_a^2))\).
-    Since \(\partial_{q_u} f = 0\), this matrix is strictly positive for all values
+    Since \(\partial_{q_u} f = 0\), this function is positive for all values
     of \(q_a\), and hence is full rank \(1\) everywhere on the constraint manifold.
@@
 
-To design our VNHC, we begin by examining a person on a
+To design the \vnhc, we begin by examining a person on a
 seated swing.
-The person extends their legs when the swing moves forwards, and retracts their
-legs when the swing moves backwards.
+The person extends their legs when the swing moves forward, and retracts their
+legs when the swing moves backward.
 As the swing gains speed, the person leans their body while
 extending their legs higher, thereby shortening the distance
-from their center of mass to the pivot and adding more energy to the swing
+from their centre of mass to the pivot and adding more energy to the swing
 \cite{how_to_pump_a_swing}.
@@
 pivot to an angle proportional to the swing's speed.
+A real gymnast cannot swing their legs in full circles, though they
+are usually flexible enough to raise them parallel to the floor;
+hence, the \vnhc must restrict the leg angle \(q_a\) 
+to lie in \([-Q_a, Q_a]\) for some \(Q_a \in [\frac{\pi}{2}, \pi[\). 
 Because the direction of motion is entirely determined by \(p_u\), 
-one VNHC which emulates this process is \(q_a = \bar{q}_a\arctan( I p_u)\),
+one \vnhc which emulates this process is \(q_a = \bar{q}_a\arctan( I p_u)\),
 displayed in Figure \ref{fig:qa-arctan}.
@@
 still inject energy into the acrobot.
-Our final VNHC is
+The final \vnhc is
 \begin{equation}\label{eqn:acrobot-constraint}
@@
 By Theorem \ref{thm:zero-dynamics}, the constrained dynamics arising from the
-VNHC \eqref{eqn:acrobot-constraint} are parameterized fully by 
+\vnhc \eqref{eqn:acrobot-constraint} are parameterized fully by 
 \((q_u,p_u) \in \SxR\).
@@
 subject to \(q_a = \bar{q}_a \arctan(I p_u)\).
-
-\subsection{Energy Injection and Dissipation with the Proposed VNHC}
+In what follows, we denote by $x=(q_u,p_u)$ the state of the constrained dynamics~\eqref{eqn:acrobot-constrained-dynamics}, and by $x(t,x_0)$ the solution of~\eqref{eqn:acrobot-constrained-dynamics} at time $t$ with initial condition $x(0)=x_0$.
+We let $x([a,b],x_0)$ denote the set $\{x(t,x_0): t \in [a,b]\}$.
+
+\subsection{Energy Injection and Dissipation with the Proposed \vnhc}\label{sec:energy-inject-diss}
 Suppose for a moment that \(I = 0\) in \eqref{eqn:acrobot-constraint}, i.e.,
@@
 \begin{equation}\label{eqn:nominal-energy}
-    E(q_u,p_u) := \frac{p_u^2}{10ml^2} + 3mgl(1 - \cos(q_u))
+    E(q_u,p_u) \eqdef \frac{p_u^2}{10ml^2} + 3mgl(1 - \cos(q_u))
     .
 \end{equation}
-The upright equilibrium of this pendulum is located at \((q_u,p_u) = (\pi,0)\).
-Imagine the pendulum hits the bottom of the swing arc with momentum 
-\(p_u \neq 0\). 
-To reach the upright equilibrium, this momentum must be
-\(p_u = \pm\sqrt{60m^2gl^3}\) because \(E(\pi,0) = E(0,\pm\sqrt{60m^2gl^3})\).
-If the momentum is smaller in magnitude, the acrobot will oscillate; 
-if it is larger, the pendulum will rotate around the bar.
-
-When the pendulum is oscillating, its state \((q_u,p_u)\) lies in the set
-\begin{equation}\label{eqn:oscillation-domain}
-    \mathcal{O}_1 := \left\{(q_u,p_u) \in \SxR 
-    \mid E(q_u,p_u) < E(\pi,0) \right\}
-    ,
-\end{equation}
-which is shown in Figure \ref{fig:acrobot-oscillation-domain}.
-If for some \(I \neq 0\) our VNHC injects energy into the acrobot on 
-\(\mathcal{O}_1\), and the constrained dynamics escape
-\(\mathcal{O}_1\) and hit the vertical line \(|q_u| = \pi\)
-in finite time, then the acrobot is guaranteed to perform
-giant-like motion and begin rotating around the bar.
-
-When the pendulum is rotating with bounded momentum
-\(\abs{p_u} < \rho\) (for some chosen \(\rho\)),
-\((q_u,p_u)\) must lie inside the rotation domain
-\begin{multline}\label{eqn:rotation-domain}
-    \mathcal{R}(\rho) := \left\{
-        (q_u,p_u) \in \SxR \mid\right.
-        \\
-        \left.E(\pi,0) < E(q_u,p_u) < E(0,\rho)
-    \right\}
-    .
-\end{multline}
-Connecting the regions \eqref{eqn:oscillation-domain} and
-\eqref{eqn:rotation-domain} yields the set
-\begin{equation}\label{eqn:o-rhobar}
-    \mathcal{O}_2(\rho) := \left\{(q_u,p_u) \in \SxR
-        \mid E(q_u,p_u) < E(0,\rho) \right\}
-    ,
-\end{equation}
-shown in Figure \ref{fig:acrobot-o2}.
-If the VNHC injects energy on \(\mathcal{O}_2(\rho)\) for some 
-\(I \neq 0\), then the acrobot must necessarily swing up, begin rotating, 
-and eventually rotate with a momentum of at least \(\rho\).
-
-\begin{figure}
+The function $E(q_u,p_u)$ is a first integral for the constrained dynamics~\eqref{eqn:acrobot-constrained-dynamics}, i.e., each solution of the constrained dynamics is confined to a level set of $E$, and each orbit not containing equilibria is a connected component of a level set of $E$.
+Given scalars $R$ and $R_1<R_2$, and letting $\# \in \{$ ``$<$''$,$''$\leq$''$,$''$>$''$,$''$\geq$''$\}$, we define the following sets (see Figure~\ref{fig:domains}):
+\begin{equation}\label{eq:E_sets}
+\begin{aligned}
+E_R &\eqdef \{(q_u,p_u): E(q_u,p_u) = R\}, \\
+E_{\# R} &\eqdef \{(q_u,p_u): E(q_u,p_u) \#R\}, \\
+E_{[R_1,R_2]} &\eqdef \{(q_u,p_u): R_1 \leq E(q_u,p_u) \leq R_2\}.
+\end{aligned}
+\end{equation}
+%
+%
+%
+\begin{figure}[htb]
     \centering
     \begin{subfigure}[t]{0.45\linewidth}
-        \includegraphics[width=\linewidth]{acrobot_oscillation_domain.pdf}
-        \caption{The set \(\mathcal{O}_1\). An orbit starting in this set (blue)
-            will pass through the level set \(E(\pi,0)\) of the nominal pendulum.}
+        \includegraphics[width=\linewidth]{acrobot_EleqR.pdf}
+        \caption{The set $E_{\leq R}$ with $0 < R < \bar R$ and an oscillation gaining energy.} 
         \label{fig:acrobot-oscillation-domain}
@@
     \begin{subfigure}[t]{0.45\linewidth}
-        \includegraphics[width=\linewidth]{acrobot_omega.pdf}
-        \caption{The set \(\mathcal{O}_2(\rho)\). Orbits starting in this
-            set are guaranteed to reach the level set \(E(0,\rho)\) of the
-            nominal pendulum.}
+        \includegraphics[width=\linewidth]{acrobot_ER1R2.pdf}
+        \caption{The set $E_{[R_1,R_2]}$, with $\bar R < R_1 < R_2$, and a rotation gaining energy.}
             \label{fig:acrobot-o2}
     \end{subfigure}
-    \caption{The sets on which the acrobot gains energy, according to
-        Theorem \ref{thm:acrobot-energy-stabilization}.}
+    \caption{Oscillations and rotations gaining energy. 
+}
+\label{fig:domains}
 \end{figure}
-
-Unfortunately, our VNHC does not always inject energy on \(\mathcal{O}_1\) and
-\(\mathcal{O}_2(\rho)\).
-If \(I\) is too large, the leg controller saturates and the body oscillates
-without gaining energy.
-However, choosing \(I\) small enough guarantees the legs will synchronize
-properly with the body, and the acrobot will begin rotating around the bar.
-The following theorem provides conditions under which such an \(I\) exists.
-
-\begin{thm}\label{thm:acrobot-energy-stabilization}
-    Consider the acrobot with Hamiltonian dynamics given in
-    \eqref{eqn:acrobot-hamiltonian}.
-\begin{enumerate}
-    \item For each \(m, g, l, \bar{q}_a > 0\), there exists
-        \(I^\star > 0\) such that, for all \(I \in \, ]0,I^\star]\), the VNHC
-        \eqref{eqn:acrobot-constraint} injects energy into the acrobot on
-        \(\mathcal{O}_1\).
-        Moreover, every orbit of the constrained dynamics 
-        \eqref{eqn:acrobot-constrained-dynamics} will reach the line
-        \(|q_u| = \pi\) in finite time.
-        If instead \(I \in [-I^\star,0[\), the VNHC dissipates energy
-        on \(\mathcal{O}_1\).
-    \item Let \(C = m^2gl^3\bar{q}_a\) and 
-        define \(b : \SxR_{> 0} \rightarrow \R\) by
+%
+%
+%
+
+Letting $\bar R \eqdef E(\pi,0)$, solutions initialized in the interior of the set $E_{\leq \bar R}$ correspond to {\em oscillations} of the pendulum, i.e., rocking motions where the pendulum does not perform full revolutions.
+On the other hand, for each $R>\bar R$, solutions initialized in the interior of the set $E_{[\bar R,R]}$ correspond to {\em rotations} of the pendulum, i.e., full revolutions around the pivot point.
+
+Now consider the case $I\neq 0$, i.e., when the acrobot is not fully extended.
+The constrained dynamics~\eqref{eqn:acrobot-constrained-dynamics} are no longer Hamiltonian, and their solutions are no longer confined to level sets of $E$.
+In what follows, we will show that the orbits of the constrained dynamics roughly speaking gain energy or lose energy depending on the sign of the design parameter $I$.
+To be more precise in describing this energy injection or dissipation property, we next define what is an oscillation or a rotation gaining or losing energy.
+Intuitively, oscillations gaining energy correspond to rocking motions of the torso with increasing amplitude, while rotations gaining energy correspond to full revolutions with increasing speed.
+Figure~\ref{fig:domains} illustrates these concepts.
+
+We define two Poincar\'e sections,
 \[
-        b(\beta,\rho_0) := 
+\begin{aligned}
+\cP_o & \eqdef \{(q_u,p_u) \in E_{<\bar R}: q_u >0, p_u=0\} \\
+\cP_r & \eqdef \{(q_u,p_u) \in E_{> \bar R}: q_u=0\}.
+\end{aligned}
+\]
+Consider a solution $x(t,x_0)$ of the constrained dynamics~\eqref{eqn:acrobot-constrained-dynamics}.
+We say that $x(t,x_0)$ is an {\em oscillation over $[0,\bar t]$} if $x([0,\bar t],x_0) \cap \cP_o$ is a discrete set, i.e., a set of isolated points.
+Similarly, $x(t,x_0)$ is a {\em rotation over $[0,\bar t]$} if $x([0,\bar t],x_0) \cap \cP_r$ is a discrete set.
+The isolated points mentioned above represent consecutive crossings of the Poincar\'e sections $\cP_o$ and $\cP_r$ by the solution $x(t,x_0)$ at times $t^i_o$, $t^i_r$, where $t^i_{o,r} < t^{i+1}_{o,r}$.
+
+Letting $x^i_o(x_0) \eqdef x(t^i_o,x_0)$ and $x^i_r(x_0) \eqdef x(t^i_r,x_0)$, we say that {\em an oscillation over $[0,\bar t]$ gains energy (respectively, loses energy)} if the sequence $\{\|x^i_o(x_0)\|\}$ is monotonically increasing (respectively, monotonically decreasing) with $i$.
+Similarly, {\em a rotation over $[0,\bar t]$ gains energy (respectively, loses energy)} if the sequence $\{\|x^i_r(x_0)\|\}$ is monotonically increasing (respectively, monotonically decreasing) with $i$.
+
+The next theorem states that if the design parameter $I$ in the \vnhc~\eqref{eqn:acrobot-constraint} is chosen small enough, then the constrained dynamics have oscillations gaining or losing energy, depending on the sign of $I$.
+
+\begin{thm}[oscillations gaining/losing energy]\label{thm:acrobot-oscillations}
+ Consider the constrained dynamics in~\eqref{eqn:acrobot-constrained-dynamics}.
+For each $R_1<R_2 \in \mathopen] 0, \bar R \mathclose[$, there exists \(I^\star > 0\) such that for each \(I \in \, ]0,I^\star]\) (respectively, $I \in [-I^\star,0\mathclose[$) and $x_0 \in E_{[R_1,R_2]}$, there exists $\bar t\geq 0$ such that $x(\bar t,x_0) \in E_{R_2}$ (respectively, $x(\bar t, x_0) \in E_{R_1}$), and $x(t,x_0)$ is an oscillation gaining energy (respectively, losing energy) over $[0, \bar t]$.
+\end{thm}
+
+\begin{proof}
+ See Section~\ref{sec:proof}.
+\end{proof}
+
+The next result concerns rotations.  In preparation for the theorem statement, we define 
+    \[
+        b(r,\theta) \eqdef 
         \frac{5C \left(
-        \frac{C}{\bar{q}_a}\left(18s_\beta^2 + 30c_\beta(1 - c_\beta)\right)
-            - c_\beta\rho_0^2
+        \frac{C}{\bar{q}_a}\left(18s_\theta^2 + 30c_\theta(1 - c_\theta)\right)
+            - c_\theta r^2
         \right)}{
-        |\rho_0|\sqrt{\rho_0^2 - 30m^2gl^3(1 - c_\beta)}
-        }
-        .
+        | r|\sqrt{ r^2 - 30m^2gl^3(1 - c_\theta)}
+        },
 \]
-        Define 
-        \(S(\rho_0) := \int \limits_{0}^{2\pi} b(\sigma,\rho_0)d\sigma\).
-        Fix \(\rho > \sqrt{60m^2gl^3}\) and
-        suppose there exists \(\epsilon > 0\) so that 
-        \(S(\rho_0) \geq \epsilon\) for all 
-        \(\rho_0 \in \, \left]\sqrt{60m^2gl^3}, \rho\right]\).
-        Then there exists \(I^{\star\star} \in \, ]0, I^\star]\) such that, for all 
-        \(I \in \, ]0,I^{\star\star}]\), the VNHC
-        \eqref{eqn:acrobot-constraint} injects energy into the acrobot on
-        \(\mathcal{O}_2(\rho)\).
-        If instead \(I \in [-I^{\star\star},0[\), the VNHC dissipates energy
-        on \(\mathcal{O}_2(\rho)\).
-\end{enumerate}
+with \(C = m^2gl^3\bar{q}_a\), and 
+        \(S( r) \eqdef \int \limits_{0}^{2\pi} b(r,\theta)d\theta\).
+
+\begin{thm}[rotations gaining/losing energy]\label{thm:acrobot-rotations} Consider the constrained dynamics in~\eqref{eqn:acrobot-constrained-dynamics}, and fix $R_1, R_2$ such that \(\bar R<R_1<R_2\).
+Suppose there exists \(\varepsilon > 0\) such that \(S( r) \geq \varepsilon\) for all \( r \in \, [(10 m l^2 R_1)^{1/2}, (10 m l^2 R_2)^{1/2}]\).
+Then there exists \(I^{\star}>0\) such that, for each \(I \in \, ]0,I^{\star}]\) (respectively, $I \in [-I^\star,0\mathclose[$) and $x_0 \in E_{[R_1,R_2]}$, there exists $\bar t \geq 0$ such that $x(\bar t,x_0) \in E_{R_2}$  (respectively, $x(\bar t, x_0) \in E_{R_1}$), and $x(t,x_0)$ is a rotation gaining energy (respectively, losing energy) over $[0, \bar t]$.
 \end{thm}
 \begin{proof}
-    See Section \ref{sec:proof}.
+See Section~\ref{sec:proof}.
 \end{proof}
 
-Notice that \(\mathcal{O}_1 \subset \mathcal{O}_2(\rho)\), yet
-Theorem \ref{thm:acrobot-energy-stabilization} considers these sets separately.
-This separation is advantageous because the first result holds for any positive
-\(m\), \(g\), \(l\), and \(\bar{q}_a\). 
-In other words, the first result of Theorem
-\ref{thm:acrobot-energy-stabilization} states that all acrobots constrained by
-\eqref{eqn:acrobot-constraint} will gain enough energy to begin rotating around
-the bar.
+\begin{rem}
+Theorem~\ref{thm:acrobot-oscillations} concerning oscillations gai\-ning/lo\-sing energy states that for any set of physical parameters, the acrobot constrained by
+\eqref{eqn:acrobot-constraint} will gain enough energy to exit in finite time a level set of the function $E$ arbitrarily close to the boundary $E_{\bar R}$ separating oscillations from rotations, provided the parameter $I$ in~\eqref{eqn:acrobot-constraint} is chosen small enough.
+The result is semiglobal relative to the set $E_{<\bar R}$ in that energy gain can be achieved on any compact subset $E_{[R_1,R_2]}$ of $E_{<\bar R}$ by a suitably small $I$.
+%% Moreover, after this finite time, as long as the solution is contained in the set $E_{< \bar R}$, then its intersections with the Poincar\'e section $\cP_o$ will occur outside of $E_{<R}$.
+Roughly speaking, this means the oscillations will retain the energy they have acquired.
+Vice versa, for small {\em negative} $I$, the oscillations will lose energy on any compact subset $E_{[R_1,R_2]}$ of $E_{<\bar R}$.
+
+Theorem~\ref{thm:acrobot-rotations} concerning rotations relies on an assumption on the function $S(r)$ which depends on the physical parameters of the acrobot, as well as the design parameter $\bar q_a$.
+The result is otherwise analogous to Theorem~\ref{thm:acrobot-oscillations}.
+\end{rem}
+
+\begin{rem}\label{rem:thm:caveat}
+The theorems above state that under certain conditions, if the design parameter $I$ is chosen small enough then the acrobot will exhibit oscillations and rotations gaining energy.
+The oscillations will exit a compact subset of $E_{<\bar R}$ in finite time, while the rotations will exit a compact subset of $E_{>\bar R}$.
+Will the oscillations gaining energy eventually turn into rotations gaining energy?
+The theorem does not answer this question, but extensive simulations and physical experiments presented below suggest that the answer is yes.
+\end{rem}
+%
+%
+%% The property that each orbit of the constrained dynamics reaches the lines $|q_u|=\pi$ in finite time allows us, in the proof of the theorem, to use the conclusion of the first part to prove the second part.
+%
+%
 %In the worst case, the acrobot will at least perform a swing-up routine to reach
@@
 
-For the acrobot to achieve giants with energy
-\(E(0,\rho)\), it must satisfy the assumption on the integral
-of \(b(\beta,\rho_0)\). 
-The value of this integral depends on the acrobot's physical parameters.
-If the assumption holds, there is some control value \(I\)
-(which depends on \(\rho\)) for which the acrobot will 
-achieve rotations with a momentum of at least \(\rho\).
+%% For the acrobot to achieve giants with energy
+%% \(E(0,\rho)\), it must satisfy the assumption on the integral
+%% of \(b(\theta,\rho_0)\). 
+%% The value of this integral depends on the acrobot's physical parameters.
+%% If the assumption holds, there is some control value \(I\)
+%% (which depends on \(\rho\)) for which the acrobot will 
+%% achieve rotations with a momentum of at least \(\rho\).
 
 \subsection{Energy Regulation}\label{sec:energy-reg}
-One can apply the results of Theorem \ref{thm:acrobot-energy-stabilization} 
+One can apply the results of Theorems~\ref{thm:acrobot-oscillations} and~\ref{thm:acrobot-rotations}
 towards energy regulation; 
 that is, one can stabilize oscillations or rotations by appropriately toggling
-between injection and dissipation VNHCs, which can be achieved by changing the
+between injection and dissipation \vnhcs, which can be achieved by changing the
 sign of \(I\) in \eqref{eqn:acrobot-constraint}.
-The acrobot is \textit{oscillating} when it swings back and forth without
-performing a full revolution around the pivot point.
-Oscillations correspond to \(p_u\) changing sign periodically, and the peak
-torso angle (i.e., the point where an orbit crosses the \(q_u\)-axis)
-is called the \textit{oscillation angle}.
-Conversely, the acrobot is \textit{rotating} when its torso is performing full
-revolutions around the pivot, which corresponds mathematically to \(p_u\)
-maintaining the same sign.
-The momentum achieved by the acrobot at the bottom of the swing
-(i.e., the point where an orbit of rotation crosses the \(p_u\)-axis) is
-called the \textit{rotation rate}.
-
-\textbf{Rotation Regulation}: one chooses a desired rotation rate 
-\(p_\text{des} > 0\) and a hysteresis value \(\delta \in [0,1]\).
-Each time the orbit crosses the \(p_u\)-axis (i.e. when \(q_u = 0\)), the
-supervisor changes which VNHC is enforced as follows:
+
+\textbf{Rotation Regulation}: choose a desired rotation rate 
+\(p_\text{des} > 0\) and a hysteresis value \(\delta \in [0,1]\) such that \((1-\delta) p_\text{des} > \sqrt{60m^2gl^3}\). Each time the orbit crosses the \(p_u\)-axis (i.e. when \(q_u = 0\)), the
+supervisor changes which \vnhc is enforced as follows:
 \begin{itemize}
-    \item If \(|p_u| < (1-\delta)p_\text{des}\), enable the injection VNHC by
+    \item If \(|p_u| < (1-\delta)p_\text{des}\), enable the injection \vnhc by
         setting \(I > 0\) in \eqref{eqn:acrobot-constraint}.
-    \item If \(|p_u| > (1+\delta)p_\text{des}\), enable the dissipation VNHC by
+    \item If \(|p_u| > (1+\delta)p_\text{des}\), enable the dissipation \vnhc by
         replacing \(I > 0\) with \(-I < 0\) in \eqref{eqn:acrobot-constraint}.
@@
         extend the leg fully by setting \(q_a = 0\).
-        For these simulations, we assume this can be done instantaneously,
-        though in practice this is impossible.
+        In simulation we assume this can be done instantaneously,
+        though in practice this takes time.
 \end{itemize}
-Note that all orbits of the acrobot must cross the \(p_u\) axis, 
-so this method will stabilize a rotation even if the acrobot is initialized in
-\(\mathcal{O}_1\).
-Also, the above supervisor is designed to regulate a rotation rate, but it
-cannot regulate the rotation direction.
-
-The term ``rotation regulation" implies that one must choose
-\(p_\text{des}\) and \(\delta\) so that 
-\((1-\delta) p_\text{des} > \sqrt{60m^2gl^3}\).
-In fact, it is entirely possible to choose \(p_\text{des}\) and \(\delta\) so
-that \((1+\delta)p_\text{des} \leq \sqrt{60m^2gl^3}\).
-Rather than stabilizing a rotation, the supervisor would then stabilize an
-oscillation whose momentum at the bottom of the swing arc is \(p_\text{des}\).
-However, there is a preferred method for stabilizing oscillations
-which we outline next.
-
-\textbf{Oscillation Regulation}: one chooses a desired oscillation angle 
-\(q_\text{des} \in \, ]0,\pi[\) and, to avoid infinite switching, a
-hysteresis value \(\delta \in [0,\pi/q_\text{des} - 1]\).
-An orbit in the \((q_u,p_u)\)-plane will either cross the \(q_u\) axis if the
-orbit corresponds to a rocking motion, or it will cross the line \(|q_u| = \pi\)
-if the orbit corresponds to a full revolution.
+
+All orbits of the acrobot initialized in $E_{< \bar R}$ must cross the \(p_u\) axis.
+The procedure above exploits this observation to stabilize a rotation even if the acrobot is initialized in $E_{< \bar R}$, although in this case we do not offer a theoretical guarantee (see Remark~\ref{rem:thm:caveat}).
+While the above supervisor is designed to regulate a rotation rate, it
+does not impose the rotation direction.
+
+%The term ``rotation regulation" implies that one must choose
+%\(p_\text{des}\) and \(\delta\) so that 
+%\((1-\delta) p_\text{des} > \sqrt{60m^2gl^3}\).
+%% In fact, it is entirely possible to choose \(p_\text{des}\) and \(\delta\) so
+%% that \((1+\delta)p_\text{des} \leq \sqrt{60m^2gl^3}\).
+%% Rather than stabilizing a rotation, the supervisor would then stabilize an
+%% oscillation whose momentum at the bottom of the swing arc is \(p_\text{des}\).
+%% However, there is a preferred method for stabilizing oscillations
+%% which we outline next.
+
+\textbf{Oscillation Regulation}: Choose a desired oscillation angle  \(q_\text{des} \in \, ]0,\pi[\) and, to avoid infinite switching, a hysteresis value \(\delta \in [0,\pi/q_\text{des} - 1]\).
+An orbit in the \((q_u,p_u)\)-plane will either cross the \(q_u\) axis if the orbit corresponds to a rocking motion, or it will cross the line \(|q_u| = \pi\) if the orbit corresponds to a full revolution.
 When either of these occur, the supervisor does the following:
 \begin{itemize}
-    \item If \(|q_u| < (1-\delta)q_\text{des}\), enable the injection VNHC.
-    \item If \(|q_u| > (1+\delta)q_\text{des}\), enable the dissipation VNHC.
+    \item If \(|q_u| < (1-\delta)q_\text{des}\), enable the injection \vnhc.
+    \item If \(|q_u| > (1+\delta)q_\text{des}\), enable the dissipation \vnhc.
     \item If \((1-\delta)q_\text{des} \leq |q_u| \leq (1+\delta)q_\text{des}\),
-        keep the leg extended at \(q_a = 0\). This can be done continously since
+        keep the leg extended at \(q_a = 0\). This can be done continuously since
         \(q_a = 0\) when \(p_u = 0\).
 \end{itemize}
-Note that by our choice of \(\delta\), if the supervisor kicks in when 
-\(|q_u| = \pi\) (i.e., when the robot is rotating) then the supervisor will
-automatically enable the dissipation VNHC.
+Note that by the choice of \(\delta\), if the supervisor kicks in when  \(|q_u| = \pi\) (i.e., when the robot is rotating) then the supervisor will automatically enable the dissipation \vnhc.
 
@@
 \section{Simulation Results}\label{sec:simulations}
-In Section \ref{sec:experiments} we will test our VNHC on a physical acrobot.
-%built by Xingbo Wang \cite{xingbo_thesis}.
-This acrobot cannot be modelled by the simplified setup in Figure
-\ref{fig:acrobot}, because the torso and leg links have 
-distributed mass. 
-We must therefore use the more general acrobot model in Figure
-\ref{fig:acrobot-model} to represent this system.
-Let us define 
+In Section \ref{sec:experiments} we will test the \vnhc~\eqref{eqn:acrobot-constraint} on a physical acrobot.
+While the acrobot model used for the theoretical development assumes point-masses and links of equal lengths, the physical acrobot has distributed mass and unequal lengths.
+In preparation for the experiments, in this section we simulate the proposed \vnhc controller on a more accurate acrobot model, depicted in Figure~\ref{fig:acrobot-model}, with distributed mass and unequal link lengths.
+Let
 \begin{align*}
-    m_{11}(q) := &m_a l_u^2 + 2m_al_ul_{c_a}\cos(q_a) + m_al_{c_a}^2 +
+    m_{11}(q) \eqdef &m_a l_u^2 + 2m_al_ul_{c_a}\cos(q_a) + m_al_{c_a}^2 +
     m_ul_{c_u}^2 \\
@@
                  , \\
-    m_{12}(q) := &m_al_{c_a}^2 + m_al_ul_{c_a}\cos(q_a) + J_a
+    m_{12}(q) \eqdef &m_al_{c_a}^2 + m_al_ul_{c_a}\cos(q_a) + J_a
     , \\
-    m_{22}(q) := &m_al_{c_a}^2 + J_a,
+    m_{22}(q) \eqdef &m_al_{c_a}^2 + J_a,
 \end{align*}
 where \(J_u\) and \(J_a\) are the moments of inertia of the torso and leg links
-respectively. 
-The general acrobot has inertia matrix
+respectively, and all other parameters are illustrated in Figure~\ref{fig:acrobot-model}. The distributed mass acrobot has inertia matrix
 \begin{equation*}
@@
     \includegraphics[width=0.8\linewidth]{acrobot_model.pdf}
-    \caption{The general acrobot model, represented by two weighted rods
+    \caption{The distributed mass acrobot model, represented by two weighted rods
     differing in both length and mass.}%
@@
 
-Recall that we defined the sets \(\mathcal{O}_1\) and
-\(\mathcal{O}_2(\rho)\) used in Theorem \ref{thm:acrobot-energy-stabilization}
-via the mechanical energy \eqref{eqn:nominal-energy} of the nominal
-pendulum associated with a simplified acrobot.
-We obtained this nominal pendulum by setting \(I = 0\) in
-\eqref{eqn:acrobot-constraint}.
-The boundary between oscillations and rotations of the nominal
-pendulum is obtained by finding the level set \(E_\pi\) of the nominal energy
-corresponding to the upright equilibrium \((q_u,p_u) = (\pi,0)\).
-
-We can extend this idea to find the nominal pendulum associated with our
-physical acrobot, whose parameters are provided in Table
-\ref{tab:acrobot-parameters}.
-Setting \(I = 0\) in \eqref{eqn:acrobot-constraint} for this acrobot
-yields the following mechanical energy of the nominal pendulum:
+Theorems~\ref{thm:acrobot-oscillations} and~\ref{thm:acrobot-rotations} make use of the mechanical energy \eqref{eqn:nominal-energy} of the nominal pendulum obtained by setting \(I = 0\) in \eqref{eqn:acrobot-constraint}.
+Using the parameters in Table~\ref{tab:acrobot-parameters} and the inertia matrix and potential function given above, we get the mechanical energy of the nominal pendulum with $I=0$:
 \[
@@
 \]
-Despite the change in model, the level set \(E_\pi\) with energy 
-\(E(\pi,0)\) remains the boundary between oscillations and rotations
-for this nominal pendulum.
-
-We will now simulate the effects of constraining our real acrobot with the VNHC
-\eqref{eqn:acrobot-constraint}, thereby demonstrating that VNHCs are robust to
-model mismatch.
-According to Theorem \ref{thm:acrobot-energy-stabilization}, the control
-parameter \(I\) must be ``small" for our VNHC to inject (or dissipate) energy
-into the simplified acrobot. 
-Unfortunately, the Theorem does not specify how small \(|I|\) must be;
-while we could make it arbitrarily small in simulations, we will eventually
-implement this VNHC on a physical testbed where \(|I|\) must be large enough to
-overcome friction.
-
-Setting \(\bar{q}_a = 1\), we experimentally determined that \(|I| = 10\) is a
-viable control parameter, so this is the value we will use for all simulations
-and experiments.
-In other words, our \textit{injection VNHC} is \eqref{eqn:acrobot-constraint} with 
-\(I = 10\) while our \textit{dissipation VNHC} is \eqref{eqn:acrobot-constraint} with 
-\(I = -10\).
+This is the function we use in interpreting the simulation results presented next.
+
+We will now simulate the effects of constraining the physical acrobot with the \vnhc \eqref{eqn:acrobot-constraint}, thereby demonstrating that \vnhcs are robust to model mismatch.
+According to Theorems~\ref{thm:acrobot-oscillations} and~\ref{thm:acrobot-rotations}, the control parameter \(I\) must be ``small" for the \vnhc~\eqref{eqn:acrobot-constraint} to inject (or dissipate) energy into the simplified acrobot.
+The Theorem does not specify how small \(|I|\) must be; while we could make it arbitrarily small in simulations, we will eventually implement this \vnhc on a physical testbed where \(|I|\) must be large enough to overcome friction.
+
+Setting \(\bar{q}_a = 1\), we experimentally determined that \(|I| = 10\) is a viable control parameter, so this is the value we will use for all simulations and experiments.
+In other words, the \textit{injection \vnhc} is \eqref{eqn:acrobot-constraint} with  \(I = 10\) while the \textit{dissipation \vnhc} is \eqref{eqn:acrobot-constraint} with  \(I = -10\).
 
@@
 
-In simulation, we stabilized the injection VNHC for the acrobot using the
+In simulation, we stabilized the injection \vnhc for the acrobot using the
 controller \eqref{eqn:stabilizing-controller}.
@@
 
-The level set \(E_\pi\) with energy \(E(\pi,0)\) is outlined in black.
+The level set \(E_{\bar R}\), with \(\bar R=E(\pi,0)\), is outlined in black.
 Recall that this level set is the boundary between oscillations and rotations of the
 nominal pendulum.
-The points where the orbit exits \(E_\pi\) are marked with black asterisks,
+The points where the orbit exits \(E_{\bar R}\) are marked with black asterisks,
 with the final departure marked by a red asterisk.
-Interestingly, our choice of \(I\) is large enough that we observe significant
+Interestingly, the choice of \(I\) is large enough that we observe significant
 differences between the nominal pendulum and the constrained dynamics:
-\(E_\pi\) intersects the \(p_u\)-axis at \(\abs{p_u} \approx 0.17\), yet the
+\(E_{\bar R}\) intersects the \(p_u\)-axis at \(\abs{p_u} \approx 0.17\), yet the
 constrained acrobot begins rotating once it hits the
@@
 
-In simulation, we stabilized the dissipation VNHC and initialized the acrobot on
+In simulation, we stabilized the dissipation \vnhc and initialized the acrobot on
 the constraint manifold with a rotation \((q_u,p_u) = (0,0.18)\).
@@
 As expected, the acrobot slows down over time.
-We highlight the locations where the orbit crossed the set \(E_\pi\) by
+We highlight the locations where the orbit crossed the set \(E_{\bar R}\) by
 black asterisks, with the final crossing in red.
@@
 Recall from Section \ref{sec:energy-reg} that one can use a supervisor to stabilize
-oscillations by appropriately toggling between injection and dissipation VNHCs
+oscillations by appropriately toggling between injection and dissipation \vnhcs
 whenever the orbit of the acrobot crosses the \(q_u\)-axis.
@@
 \(5\%\) hysteresis, meaning \(\delta = 0.05\).
-The supervisor reevaluated its choice of VNHC at each black asterisk;  
+The supervisor reevaluated its choice of \vnhc at each black asterisk;  
 the red contour corresponds to the part of the orbit where the supervisor kept
@@
 \(|q_u| = \pi\) and dissipates energy. Eventually the acrobot begins a rocking
-motion and the supervisor is enable again each time the acrobot hits the
+motion and the supervisor is enabled again each time the acrobot hits the
 \(q_u\)-axis. 
@@
 described in Section \ref{sec:energy-reg}, where the supervisor toggles between
-injection and dissipation VNHCs at each crossing of the \(p_u\)-axis.
+injection and dissipation \vnhcs at each crossing of the \(p_u\)-axis.
 Rotation regulation for the acrobot is demonstrated in 
@@
 \(p_\text{des} = 0.19\) with a \(2\%\) hysteresis \(\delta = 0.02\).
-The supervisor evaluated its choice of VNHC at each black asterisk.
+The supervisor evaluated its choice of \vnhc at each black asterisk.
 Once it was within range of \(p_\text{des}\) it extended the legs completely,
@@
 
-Note the difference in shape between the blue rotations of the dissipation VNHC
+Note the difference in shape between the blue rotations of the dissipation \vnhc
 and the red rotation of the nominal pendulum in Figure
@@
 if \(|I|\) were smaller, the blue rotations would be more similar in shape to the
-red one because the constrained dynamics for the dissipation VNHC would be
+red one because the constrained dynamics for the dissipation \vnhc would be
 well approximated by the nominal pendulum.
@@
 The simulation results in this section demonstrate the energy regulation
-capabilities of our VNHC.  
+capabilities of the proposed \vnhc.  
 We were able to stabilize both oscillations and rotations by implementing a
 control supervisor which toggled between injection, dissipation, and
-leg-extension VNHCs.
-In particular, these simulations indicate that our VNHC
+leg-extension \vnhcs.
+In particular, these simulations indicate that the proposed \vnhc
 works even for acrobots whose limbs have differing masses and lengths.
@@
 \subsection{Hardware Description}
-In this section we will demonstrate that our VNHC is robust to friction, sensor
-noise, and other real-world considerations by testing it on our physical
-acrobot (Figure \ref{fig:xingbo-acrobot}).
+In this section we will demonstrate that the proposed \vnhc is robust to friction, sensor
+noise, and other real-world considerations by testing it on the physical
+acrobot depicted in Figure \ref{fig:xingbo-acrobot}.
 This platform is called SUGAR, which stands for
@@
 filter to sequential measurements of \(q\).
-We then compute \(p = M(q)\dot{q}\) for use in the VNHC controller.
+We then compute \(p = M(q)\dot{q}\) for use in the \vnhc controller.
 
@@
 
-We performed the following tests on SUGAR with the energy injection VNHC.
+We performed the following tests on SUGAR with the energy injection \vnhc.
 
@@
     \includegraphics[width=0.8\linewidth]{acrobot_unperturbed_orbit.png}
-    \caption{Baseline Test: SUGAR's baseline energy injection orbit.}
+    \caption{Baseline Test: SUGAR's baseline energy injection orbit. A video of this experiment is available at {\footnotesize https://play.library.utoronto.ca/watch/6703e828acdcd38058ca884d13660a88}}
     \label{fig:acrobot-unperturbed-orbit}
@@
     \includegraphics[width=0.8\linewidth]{acrobot_stopped_orbit.png}
-    \caption{Perturbation Test 1: SUGAR's orbit before (blue) and after (red) stopping.}
+    \caption{Perturbation Test 1: SUGAR's orbit before (blue) and after (red) stopping.
+A video of this experiment is available at {\footnotesize https://play.library.utoronto.ca/watch/164a21f0a52d4d56ef60bf1911545e0e}}
     \label{fig:acrobot-stopped-orbit}
@@
 with energy \(E(q_u,p_u) > E(0,0.195)\). 
-The baseline and push tests suggest that our VNHC injects energy into SUGAR
-only on \(\mathcal{O}_2(0.195)\).
+The baseline and push tests suggest that the \vnhc injects energy into SUGAR
+only on \(\{E(q_u,p_u) \leq E(0,0.195)\}\).
 This difference between simulation and implementation is likely due to
@@
         \includegraphics[width=\linewidth]{acrobot_fpush_orbit.png}
-        \caption{The forwards push test.}
+        \caption{The forward push test.}
         \label{fig:acrobot-fpush-orbit}
@@
     \end{subfigure}
-    \caption{Perturbation Test 2: SUGAR's orbit before (blue) and after (red) pushing.}
+    \caption{Perturbation Test 2: SUGAR's orbit before (blue) and after (red) pushing.
+A video of this experiment is available at {\footnotesize https://play.library.utoronto.ca/watch/b13d4dc9f6daefa7f4d56dbf1b1bd482.}}
 \end{figure}
@@
 a stop test, and a push test.
-These experiments demonstrate that VNHC-based energy injection is robust to
+These experiments demonstrate that \vnhc-based energy injection is robust to
 significant model mismatch, friction, sensor noise, discretized control
@@
 %/========== Proof ==========/%
-\section{Proof of Theorem \ref{thm:acrobot-energy-stabilization}}\label{sec:proof}
-In this section we prove Theorem \ref{thm:acrobot-energy-stabilization},
-which claims the acrobot gains/loses energy on \(\mathcal{O}_1\) and
-\(\mathcal{O}_2(\rho)\).
-These are the sublevel sets \eqref{eqn:oscillation-domain} and
-\eqref{eqn:o-rhobar} of the nominal pendulum's energy \(E(q_u,p_u)\) defined in
-\eqref{eqn:nominal-energy}.
-
-Here is fundamental idea behind this proof:
-for each of \(\mathcal{O}_1\) and \(\mathcal{O}_2(\rho)\),
-we can find a diffeomorphism \(T(q_u,p_u) = (r,\theta)\) into pseudo-polar
-coordinates adapted to level sets of \(E(q_u,p_u)\);
-that is, when \(I = 0\) in \eqref{eqn:acrobot-constraint}, the pseudo-radius
-\(r\) chooses a level set of \(E\) and the pseudo-angle \(\theta\) chooses a
-point on that level set.
-If the constrained dynamics in \((r,\theta)\) coordinates flow in such a way
-that \(r\) increases ``on average", then \(E\) must also increase ``on average",
-and the original constrained dynamics \eqref{eqn:acrobot-constrained-dynamics} 
-must be gaining energy on the domain of the diffeomorphism.
-
-The following lemma provides a set of conditions on the \((r,\theta)\) dynamics
-under which energy will increase on a domain \(D\); later in this section, we
-will assign \(D \subset \mathcal{O}_1\) or \(D \subset \mathcal{O}_2(\rho)\),
-find the diffeomorphism \(T\), and apply this lemma to prove the claims in
-Theorem \ref{thm:acrobot-energy-stabilization}.
-
-%Here is the summary of this proof: for certain subsets
-%\(D \subset \SxR\), one can find a change of coordinates into 
-%pseudo-polar coordinates adapted to level sets of \(E(q_u,p_u)\)
-%(see the general procedure devised by Mohammadi \etal 
-%\cite{dynamic_vhcs_stabilize_closed_orbits}).
-%That is, when \(I = 0\) in \eqref{eqn:acrobot-constraint}, one can find
-%a diffeomorphism \(T(q_u,p_u) =  (r,\theta)\), where \(r\)
-%chooses a level set of \(E\) and \(\theta\) chooses a point on that level set.
-%The acrobot's dynamics in \((r,\theta)\) coordinates will satisfy 
-%\(\dot{r} = 0\) and \(\dot{\theta} > 0\) when \(I = 0\).
-%By continuity of the constrained dynamics with respect to \(I\),
-%we maintain \(\dot{\theta} > 0\) for small non-zero values of \(I\).
-%This will enable us to perform a time-scaling transformation 
-%\(t = t(\theta)\), and study how the acrobot's pseudo-radius \(r\) evolves as a
-%function of the pseudo-angle \(\theta\), rather than as a function of time.
-%We will formally show that \(r(\theta)\) increases on average, which means
-%\(E\) also increases on average.
-%This in turn implies the constrained dynamics are gaining energy on \(D\).
-%We formalize these ideas in the following lemma.
-
-\begin{lemma}\label{lemma:energy-regulation-D}
-    Consider the simplified acrobot \eqref{eqn:acrobot-hamiltonian} constrained
-    by the VNHC \eqref{eqn:acrobot-constraint}, whose constrained dynamics are
-    \eqref{eqn:acrobot-constrained-dynamics}.
-    Let \(D \subset \SxR\) be open and 
-    let \(]\underline{r}, \overline{r}[ \, \subset \R_{> 0}\).
-    Suppose there exists a diffeomorphism 
-    \begin{align*}
-        T : D &\to ]\underline{r},\overline{r}[ \, \times \Sone
-         \\
-        (q_u,p_u) &\mapsto (r,\theta)
-        ,
-    \end{align*}
-    yielding the new constrained dynamics \(\dot{r} = f_r(\theta,r,I)\),
-    \(\dot{\theta} = f_\theta(\theta,r,I)\).
-    Suppose that \(f_r(\theta,r,0) \equiv 0\) and \(f_\theta(\theta,r,0) > 0\).
-    Let \(g(\theta,r,I) = f_r(\theta,r,I)/f_\theta(\theta,r,I)\), and assume that 
-    \((\partial g/\partial r)(\theta,r,0) \equiv 0\).
-    Assume also that \(b(\theta,r) := (\partial g/\partial I)(\theta,r,0)\)
-    is \(2\pi\)-periodic in \(\theta\),
-    and that there exists \(\epsilon > 0\) such that 
-    \(S(r_0) := \int_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon\)
-%    \[
-%        S(r_0) := \int \limits_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon
-%        ,
-%    \]
-    for all \(r_0 \in \, ]\underline{r},\overline{r}[\).
-    Then there exists \(I^\star > 0\) small enough that the constrained dynamics
-    gain energy on \(D\) for all \(I \in \,]0, I^\star]\), and lose energy on \(D\)
-    for all \(I \in [-I^\star, 0[\).
+\section{Proofs of Theorems~\ref{thm:acrobot-oscillations} and~\ref{thm:acrobot-rotations}}\label{sec:proof}
+
+The proofs of the theorems rely on the next lemma.
+
+\begin{lemma}\label{lem:gain_lose}Let $\cU \subset \Sone \times \Re$ be an open set and $R_2 > R_1 >0$ be such that a connected component of $E_{[R_1,R_2]}$ is contained in $\cU$.
+Let $T : \cU \to \cV \eqdef \mathopen]\underline{r},\overline{r}[ \times \Sone$, $(q_u,p_u) \mapsto (r,\theta)$, be a diffeomorphism such that
+\begin{equation}\label{eq:T_property}
+\begin{aligned}
+&\big(\forall R \in [R_1,R_2]\big) \big(\exists s \in \mathopen]\underline{r},\overline{r}\mathclose[\big) \\
+& T(E_R) = \{(r,\theta) \in \cV: r = s \},
+\end{aligned}
+\end{equation}
+and the map $R \mapsto s$ is monotonically increasing.
+Suppose that $T: \cU \to \cV$ transforms system~\eqref{eqn:acrobot-constrained-dynamics} into
+\begin{equation}
+\label{eq:r_theta_system}
+\begin{aligned}
+\dot r &= f_r(r,\theta,I) \\
+\dot \theta &= f_\theta(r,\theta,I),
+\end{aligned}
+\end{equation}
+where $f_r$, $f_\theta$ and $g \eqdef f_r / f_\theta$ satisfy, for every $(r,\theta) \in \cU$,
+\begin{subequations}\label{eq:lemma_assumptions}
+\begin{align}
+& f_r(r,\theta,0) = 0, \ f_\theta(r,\theta,0) >0, \label{eq:lemma_assumptions:1}\\
+& \partial_r g(r,\theta,0) =0, \label{eq:lemma:assumptions:2}\\
+& \int_0^{2\pi} \partial_I g(r,\theta,I)\big|_{I=0} d \theta >0. \label{eq:lemma:assumptions:3} 
+\end{align}
+\end{subequations}
+Then there exists $I^\star>0$ such that for each $I \in \mathopen]0,I^\star]$ (respectively, $I \in [-I^\star,0\mathclose[$) and every $x_0 \in E_{[R_1,R_2]} \cap \cU$, the solution $x(t,x_0)$ enjoys the following properties:
+
+\begin{enumerate}[(a)]
+\item There exists $\bar t \geq 0$ such that $x(\bar t,x_0) \in E_{R_2}$ (respectively, $E_{R_1}$).
+
+\item If $R_2 < \bar R$, $x(t,x_0)$ is an oscillation gaining (respectively, losing) energy over $[0,\bar t]$.
+If $R_1 > \bar R$, $x(t,x_0)$ is a rotation gaining (respectively, losing) energy over $[0,\bar t]$.
+\end{enumerate}
 \end{lemma}
+
 \begin{proof}
-    Since \(f_\theta(\theta,r,0) > 0\) and \(f_\theta\)
-    is continuous in \(I\), there exists 
-    \(I_1 > 0\) small enough that \(\dot{\theta} > 0\) for all
-    \(I \in \, [-I_1,I_1]\).
-    Hence, we can perform a time reparameterization \(t = t(\theta)\) to get a
-    time-scaled pseudo-radius \(\hat{r}(\theta) := r(t(\theta))\).
-    This reduces the system \((\dot{r},\dot{\theta})\) into the scalar
-    time-varying ODE
+By property~\eqref{eq:T_property}, there exist $r_1,r_2 \in \mathopen]\underline{r},\overline{r}[$ such that $T(E_{[R_1,R_2]}) = [r_1,r_2]\times \Sone$.
+Moreover, $T(E_{R_2}) = \{r_2 \} \times \Sone$.
+Define $K \eqdef [r_1,r_2]\times \Sone$, a compact subset of $\cV$, and
+\begin{equation}\label{eq:b_defn}
+b(r,\theta) \eqdef \partial_I g(r,\theta,I)\big|_{I=0}.
+\end{equation}
+By properties~\eqref{eq:lemma_assumptions:1} and~\eqref{eq:lemma:assumptions:3} and since $f_r, f_\theta$ are continuous and $K$ is compact, there exist $I_1,\varepsilon_1,\varepsilon_2>0$ such that for each $(r,\theta) \in K$ and $I \in [-I_1,I_1]$,
+\begin{subequations}
+\label{eq:properties:new_coords}
+\begin{align}
+&f_r(r,\theta,I) > \varepsilon_1,\label{eq:properties:new_coords:1}\\
+& \int_0^{2\pi} b(r,\theta) d \theta > \varepsilon_2.\label{eq:properties:new_coords:2}
+\end{align}
+\end{subequations}
+Property~\eqref{eq:properties:new_coords:1} implies that for $I \in [-I_1,I_1]$ the first component of any solution $(r(t),\theta(t))$ contained in $K$ is a strictly monotonic function of $t$, and therefore $r(t)$  can be expressed as $r(t) = \hat r(\theta(t))$, where $\hat r(\theta)$ is a solution of
     \begin{equation}\label{eqn:rhat-ode}
-        \begin{cases}
-            \hat{r}^\prime= \left.\frac{dr}{dt} \frac{dt}{d\theta}\right|_{r =
-                \hat{r}} = 
-            g(\theta,\hat{r},I)
-            , \\
-            \hat{r}(0) = r_0
-            ,
-        \end{cases}
-    \end{equation}
-    with state \(\hat{r}\), where now
-    \(\theta\) represents time,
-    prime denotes differentiation by \(\theta\), and
-    \(r_0 \in \, ]\underline{r},\overline{r}[\) is the initial condition.
-
-    Let \(\hat{r}(\theta,r_0,I)\) be the solution to \eqref{eqn:rhat-ode}.
-    Taking the Taylor expansion of this solution around \(I = 0\) we get
-    \(\hat{r}(\theta,r_0,I) = \hat{r}(\theta,r_0,0) + I\hat{r}_1(\theta,r_0) 
-    + R(\theta,r_0,I)\). 
-    Here, \(\hat{r}(\theta,r_0,0) \equiv r_0\) because
-    \(g(\theta,\hat{r},0) \equiv 0\),
-    the function \(\hat{r}_1(\theta,r_0)\) is the solution to
+            \frac{d \hat r}{d \theta} = g(\hat r,\theta,I) = \frac{f_r(\hat r,\theta,I)}{f_\theta(\hat r,\theta,I)}. %\\
+    \end{equation}
+Now let \(\hat{r}(\theta,r_0,I)\) be the solution of \eqref{eqn:rhat-ode} with initial condition $r(0)=r_0$ and $I\in [-I_1,I_1]$. The Taylor expansion of $\hat{r}(\theta,r_0,I)$ around \(I = 0\) is
+\begin{equation}\label{eq:Taylor_exp}
+\begin{aligned}
+\hat{r}(\theta,r_0,I) &= \hat{r}(\theta,r_0,0) + I\hat{r}_1(\theta,r_0) 
+    + \Delta(\theta,r_0,I) \\
+%
+&=r_0 + I\hat{r}_1(\theta,r_0) + \Delta(r_0,\theta,I),
+\end{aligned}
+\end{equation}
+where the function $\Delta(r,\theta,I)$ is continuous and $I \mapsto \Delta(r,\theta,I)$ is $O(I^2)$.
+The second identity follows from the fact (due to~\eqref{eq:lemma_assumptions:1}) that \(g(\hat{r},\theta,0) \equiv 0\).
+Following~\cite[Chapter 10]{khalil_nonlinear}, the function \(\hat{r}_1(\theta,r_0)\) is the solution to the linear time-varying ODE
     \begin{equation}\label{eqn:r1-hat-prime}
-        \begin{cases}
-        \hat{r}_1^\prime = 
-        \frac{\partial g}{\partial \hat{r}}(\theta,r_0,0)\hat{r}_1 +
-        b(\theta,r_0)
-        %\frac{\partial g}{\partial I}(\theta,r_0,0)
-        , \\
-        \hat{r}_1(0) = 0
-        ,
-    \end{cases}
-    \end{equation}
-    and the remainder term \(R(\theta,r_0,I)\) is \(O(I^2)\).
-
-    Since \((\partial g/\partial \hat{r})(\theta,r_0,0) \equiv 0\),
-    the solution of \eqref{eqn:r1-hat-prime} can be obtained by
-    quadrature, giving
-    \(\hat{r}_1(\theta,r_0) = \int_0^\theta b(\sigma,r_0)d\sigma\).
-%    \[
-%        \hat{r}_1(\theta,r_0) = \int\limits_0^\theta b(\sigma,r_0)d\sigma
-%        .
-%    \]
-%    with \(b(\theta,r_0) := (\partial g/\partial I)(\theta,r_0,0)\). 
-    Recall that \(b(\theta,r_0)\) is \(2\pi\)-periodic in
-    \(\theta\), so there is a well-defined Poincar\'{e} map describing
-    how \(\hat{r}\) changes each time \(\theta\) increases by \(2\pi\).
-    This map is 
-    \begin{align}\label{eqn:poincare-map}
-        P : \, ]\underline{r},\overline{r}[ &\to \mathbb{R}_{\geq 0} \\
-        r_0 &\mapsto \hat{r}(2\pi,r_0,I)
-        ,
+\begin{aligned}         
+\frac{d \hat{r}_1}{d \theta}&=
+        \partial_{r} g(r_0,\theta,0)\hat{r}_1 +
+        b(r_0,\theta) \\
+&=b(r_0,\theta),
+\end{aligned}
+     \end{equation}
+with initial condition $\hat r_1(0)=0$. The second identity in~\eqref{eqn:r1-hat-prime} follows from~\eqref{eq:lemma:assumptions:2}. We can solve \eqref{eqn:r1-hat-prime} by quadrature, giving
+\begin{equation}
+\label{eq:ronehat_solution}
+\hat{r}_1(\theta,r_0) = \int_0^\theta b(r_0,\theta)d\theta.
+\end{equation}
+Now consider the Poincar\'e section $\cP = \{(r,\theta) \in \cV: \theta =0\}$ defined in the lemma statement.
+Using~\eqref{eq:Taylor_exp}, and~\eqref{eq:ronehat_solution}, the Poincar\'e map on $\cP$ is $P: \cP \to \Re_{>0}$,
+\begin{equation}\label{eq:Poincare_map:expansion}
+r \mapsto r + I \int_0^{2\pi} b(r,\theta) d \theta + \Delta(r,2 \pi,I).
+\end{equation}
+Recall that $(r,I) \mapsto \Delta(r,2 \pi,I)$ is continuous and $O(I^2)$ for each fixed $r$.
+Since the set $[r_1,r_2] \times [-I_1,I_1]$ is compact, there exists $\kappa>0$ such that $| \Delta(r,2\pi,I)| \leq \kappa I^2$ for all  $(r,I) \in [r_1,r_2] \times [-I_1,I_1]$.
+Using this inequality and property~\eqref{eq:properties:new_coords:2} in~\eqref{eq:Poincare_map:expansion}, we get
+\[
+\begin{aligned}
+\big(\forall I \in \mathopen]0,I_1]\big) & \ P(r) > r + I \varepsilon_2 - \kappa I^2 \\
+\big(\forall I \in [-I_1,0\mathclose[\big) & \ P(r) < r - |I| \varepsilon_2 + \kappa I^2,
+\end{aligned}
+\]
+for all $r \in [r_1,r_2]$. Pick $I^\star \in \mathopen]0, \min\{I_1, \varepsilon_2/\kappa\} \mathclose[$ and let $\gamma \eqdef |I|\varepsilon_2 - \kappa I^2>0$. Then, 
+\begin{subequations}\label{eq:poincare_inequality}
+\begin{align}
+(\forall r \in [r_1,r_2]) (\forall I \in \mathopen]0,I^\star]) & \ P(r) > r + \gamma \label{eq:poincare_inequality:1}\\
+(\forall r \in [r_1,r_2]) (\forall I \in [-I^\star,0\mathclose[) & \ P(r) < r - \gamma. \label{eq:poincare_inequality:2}
 \end{align}
-    which expands to \(P(r_0) = r_0 + IS(r_0) + R(2\pi,r_0,I)\).
-    In other words, we can understand the evolution of \(\hat{r}\) by studying
-    the dynamics of the discrete-time system
-    \begin{equation}\label{eqn:r-discrete}
-        r_{n+1} = P(r_n) =  r_n + IS(r_n) + R(2\pi,r_n,I)
-        ,
-    \end{equation}
-    with initial condition \(r_0\).
-%    Here, \(r_n\) corresponds to the distance along the line segment
-%    \(L := ]\underline{r},\overline{r}[ \, \times \{0\}\)
-%    when an orbit of \((r,\theta)\) hits \(L\) for the the \(n^\text{th}\) time,
-%    assuming the acrobot was initialized at \((r,\theta) = (r_0,0)\).
-
-    If at any point in discrete time \(r_n\) exits \(]\underline{r},\overline{r}[\), the orbit 
-    in \((q_u,p_u)\) coordinates must have departed \(D\).
-    Hence, proving the constrained dynamics gain energy on \(D\) is
-    equivalent to finding \(N \in \mathbb{Z}_{> 0}\) such that 
-    \(r_{n} \geq \overline{r}\) for all \(n \geq N\).
-    Using a perturbation analysis (see Khalil 
-    \cite[Theorem 10.1]{khalil_nonlinear}),
-    we know there exists some \(I_2 \in \, ]0,I_1]\) and \(\kappa > 0\)
-    such that \(\left|R(\theta,r,I)\right| \leq \kappa I^2\) for all 
-    \(\theta \in [0,2\pi]\),
-    \(r \in \, ]\underline{r},\overline{r}[\),
-    and \(I \in [-I_2,I_2]\).
-    Taking this fact alongside the assumption that \(S(r_0) \geq \epsilon\),
-    we find a lower bound \(r_{n+1} \geq r_n + I\epsilon - \kappa I^2\).
-    Let \(0 < I^\star < \min\{I_2, \epsilon/\kappa\}\).
-    Then for each \(I \in ]0,I^\star]\), the value 
-    \(\gamma := I\epsilon - \kappa I^2\) is strictly positive, so
-    \(P(r_0) \geq r_0 + \gamma\).
-    Therefore, \(r_{n} \geq r_0 + n\gamma\), 
-    and all solutions of \eqref{eqn:r-discrete} eventually flow past
-    \(\overline{r}\).
-
-    We conclude that, for \(I \in \, ]0, I^\star]\), the constrained dynamics gain
-    energy on \(D\).
-    Changing the sign of \(I\) changes the direction of the
-    above inequalites, so that \(r_{n} \leq r_0 - n\gamma\). 
-    In this case all orbits of \eqref{eqn:r-discrete} flow below
-    \(\underline{r}\), i.e., the constrained dynamics lose energy on \(D\).
+\end{subequations}
+Returning to the solution $\hat r(\theta,r_0,I)$ of~\eqref{eqn:rhat-ode}, property~\eqref{eq:poincare_inequality:1} implies that for each $(r_0,I) \in [r_1,r_2] \times \mathopen]0,I^\star]$, there exists an integer $\bar k \geq 0$ such that the sequence $\{r(2 \pi k, r_0,I)\}_{k =1,\ldots,\bar k}$ is monotonically increasing and $r(2 \pi \bar k,r_0,I) \geq r_2$.
+Vice versa, property~\eqref{eq:poincare_inequality:2} implies that for each $(r_0,I) \in [r_1,r_2] \times [-I^\star,0\mathclose[$, there exists an integer $\bar k \geq 0$ such that the sequence $\{r(2 \pi k, r_0,I)\}_{k =1,\ldots,\bar k}$ is monotonically decreasing and $r(2 \pi \bar k,r_0,I) \leq r_1$.
+
+Now we return to system~\eqref{eq:r_theta_system} with state $(r,\theta)$, and let $(r(t),\theta(t))$ be a solution initialized in $K$, with $0< |I| \leq I^\star$. 
+For $I>0$ (respectively, $I<0$), let $\bar t \geq 0$ be the first time such that $( r(\bar t), \theta(\bar t)) \in \{r_2\} \times \Sone$ (respectively, $( r(\bar t), \theta(\bar t)) \in \{r_1\} \times \Sone$).
+If the solution never intersects $\{r_2\} \times \Sone$, we set $\bar t = \infty$.
+
+By property~\eqref{eq:properties:new_coords:1}, $\dot \theta > \varepsilon_1>0$, implying that the intersection of $( r([0,\bar t]),\theta([0,\bar t]) )$ with $\cP$ is a discrete set, and property~\eqref{eq:poincare_inequality} implies that the intersection points in $( r([0,\bar t]),\theta([0,\bar t]) ) \cap \cP$ form a monotonically increasing (respectively, monotonically decreasing) sequence crossing the boundary $\{r_2\} \times \Sone$ (respectively, $\{r_1\} \times \Sone$) in finite time.
+This in turn implies that $\bar t< \infty$.
+
+Returning to system~\eqref{eqn:acrobot-constrained-dynamics} in original coordinates, we have shown that for each $x_0 \in E_{[R_1,R_2]}\cap \cU$ and $0< |I| \leq I^\star$, there exists $\bar t\geq 0$ such that the solution $x(t,x_0)$ enjoys these properties:
+\begin{itemize}
+\item For $I>0$, $x(\bar t,x_0) \in E_{R_2}$, and for $I<0$, $x(\bar t,x_0) \in E_{R_1}$.
+\item For $I>0$, $x(t,x_0)$ is an oscillation or a rotation gaining energy over $[0, \bar t]$, depending on the values of $R_1$ and $R_2$.
+For $I<0$, $x(t,x_0)$ is an oscillation or a rotation losing energy over $[0, \bar t]$.
+\end{itemize}
+This concludes the proof of Lemma~\ref{lem:gain_lose}.
 \end{proof}
 
-\subsection{Energy Gain on \(\mathcal{O}_1\)}\label{sec:proof-o1}
-To prove the first part of Theorem
-\ref{thm:acrobot-energy-stabilization}, we use
-\cite[Eqn. (9)]{dynamic_vhcs_stabilize_closed_orbits} to find 
-pseudo-polar coordinates adapted to level sets of \(E\) on
-\(\mathcal{O}_1 \backslash \{(0,0)\}\).
-The resulting coordinate transformation is
-\begin{align}
-    \label{eqn:r-osc}
-    r &= \arccos\left(\cos(q_u) - \frac{p_u^2}{30m^2gl^3}\right)
-    , \\
-    \label{eqn:theta-osc}
-    \theta &=
-    \left.\arctan_2\left(-\sign{p_u}\sqrt{1-\frac{q_u^2}{r^2}},\frac{q_u}{r}\right)
-    \right|_{r = r(q_u,p_u)}
-    ,
-\end{align}
-where \((r,\theta) \in \, ]0,\pi[ \times \Sone\).
-One can verify that the inverse to this transformation is
-\begin{align*}
-    q_u &= rc_\theta
-    , \\
-    p_u &= -\sign{s_\theta}
-    \sqrt{30m^2gl^3\cos(rc_\theta - c_r)}
-    ,
-\end{align*}
-where for shorthand we write \(c_\theta := \cos(\theta)\), \(s_\theta :=
-\sin(\theta)\), \(c_r := \cos(r)\), and \(s_r := \sin(r)\).
-We now verify these pseudo-polar coordinates satisfy the requirements of 
-Lemma \ref{lemma:energy-regulation-D}.
-We omit the expressions for \(f_r(\theta,r,I)\) and \(f_\theta(\theta,r,I)\), as
-they are too long for print.
-
-It is straightforward to show that
-\(f_r(\theta,r,0) \equiv 0\)
+\subsection{Proof of Theorem~\ref{thm:acrobot-oscillations}}
+
+We prove the theorem by producing a coordinate transformation $T : \cU \to \cV$ meeting the assumptions of Lemma~\ref{lem:gain_lose}.
+The transformation is adapted from~\cite{dynamic_vhcs_stabilize_closed_orbits}.
+
+Define $\cU \eqdef E_{<\bar R} \setminus \{0\}$ and $\cV \eqdef \mathopen]0,\pi\mathclose[$, and consider the coordinate transformation\footnote{Despite the presence of a $\text{sgn}(\cdot)$ function, the coordinate transformation $T_1$ is actually smooth.} $T: \cU \to \cV$ defined as
+\begin{equation}\label{eq:T1}
+\begin{aligned}
+T: & \, (q_u,p_u) \mapsto (r,\theta)  \\
+%
+r &= \arccos\left(\cos(q_u) - \frac{p_u^2}{30m^2gl^3}\right) \\
+%
+\theta &=\left.\arctan_2\left(-\sign{p_u}\sqrt{1-\frac{q_u^2}{r^2}},\frac{q_u}{r}\right)
+    \right|_{r = \rone(q_u,p_u)},
+\end{aligned}
+\end{equation}
+whose inverse is
+\[
+\begin{aligned}
+T^{-1}: & \, (r,\theta) \mapsto (q_u,p_u)\\
+%
+q_u &= r \cos(\theta) \\
+% 
+p_u &=-\sign{\sin(\theta)} 
+        \sqrt{30m^2gl^3\cos(r\cos(\theta) - \cos(r))}.
+\end{aligned}
+\]
+It is easily seen that for each $R \in \mathopen ]0,\bar R\mathclose[$, $T(E_R) = \{(r,\theta)\in \cV : r = \acos(1 - R/(3 m g l))\}$, and the function $R \mapsto \acos(1 - R/(3 m g l))$ is monotonically increasing.
+The diffeomorphism $T: \cU \to \cV$ therefore satisfies the first requirement of Lemma~\ref{lem:gain_lose} for any choice of $0< R_1 < R_2 < \bar R$.
+
+In $(r,\theta)$ coordinates, $T$ maps system~\eqref{eqn:acrobot-constrained-dynamics} into a system of the form~\eqref{eq:r_theta_system}, where is straightforward to show that \(f_r(\theta,r,0) \equiv 0\)
 and
@@
         \sqrt{\frac{\cos(r c_\theta) - c_r}
-            {r^2 s_\theta^2}}
-    .
-\end{equation}
-Note that ~\eqref{eqn:theta-dot-nom-osc} has removable singularities at
-\(\theta \in \{0,\pi\}\).
-Taking the limit as \(\theta\) approaches these points yields
+            {r^2 s_\theta^2}}.
+\end{equation}
+In the above we used the notation \(c_\theta \eqdef \cos(\theta)\), \(s_\theta \eqdef \sin(\theta)\), \(c_r \eqdef \cos(r)\), and \(s_r \eqdef \sin(r)\).
+We note that~\eqref{eqn:theta-dot-nom-osc} has removable singularities at \(\theta \in \{0,\pi\}\), since taking the limits as \(\theta\to 0,\pi\) gives
 \begin{equation}\label{eqn:theta-dot-limit}
@@
 \end{equation}
-which is smooth and well-defined for all
-\(r \in \, ]0,\pi[\).
-From \eqref{eqn:theta-dot-nom-osc}--\eqref{eqn:theta-dot-limit},
-one can verify that \(f_\theta(\theta,r,0) > 0\) as required.
-Defining \(g(\theta,r,I)\) as in Lemma \ref{lemma:energy-regulation-D}, one can
-also verify that \((\partial g/\partial r)(\theta,r,0) \equiv 0\).
-
-Symbolic computations reveal that
-\(b(\theta,r_0) = K a(\theta,r_0)\),
-where
+and these limits smooth and well-defined for all \(r \in \, ]0,\pi[\).
+From \eqref{eqn:theta-dot-nom-osc}--\eqref{eqn:theta-dot-limit}, one can verify that \(f_\theta(\theta,r,0) > 0\) for all $(\theta,r) \in \cV$, and thus assumption~\eqref{eq:lemma_assumptions:1} holds.
+Letting \(g= f_r/f_\theta\), one can also verify that \((\partial g/\partial \hat r)(\theta,\hat r,0) \equiv 0\), and assumption~\eqref{eq:lemma:assumptions:2} holds.
+Symbolic computations reveal that \(\partial_I g(r,\theta,I) \big|_{I=0} = L a(r,\theta)\), where
 \begin{align*}
-    K &:= \frac{\bar{q}_a \sqrt{30m^2g l^3}}{15}
+    L &\eqdef \frac{\bar{q}_a \sqrt{30m^2g l^3}}{15}
     , \\
-    a(\theta,r_0) &:= \frac{
-        r_0 |s_\theta| \left(
-        5 c_{r_0} \cos(r_0 c_\theta) - 8 \cos(r_0c_\theta)^2 + 3
+    a(r,\theta) &\eqdef \frac{
+        r |s_\theta| \left(
+        5 c_{r} \cos(r c_\theta) - 8 \cos(r c_\theta)^2 + 3
     \right)
     }{
-    s_{r_0}\sqrt{\cos(r_0c_\theta) - c_{r_0}}
+    s_{r}\sqrt{\cos(r c_\theta) - c_{r}}
     }
@@
 \end{align*}
-Notice that \(K\) is a positive constant which depends only on
-\(m\), \(g\), \(l\), and \(\bar{q}_a\), while \(a(\theta,r_0)\) is adimensional
-and \(2\pi\)-periodic in \(\theta\).
-This means the integral of \(a(\theta,r_0)\) can be computed once for all
-acrobots; 
-hence, there exists \(\epsilon > 0\) such that 
-\(S(r_0) := \int_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon\)
-if and only if there exists \(\delta > 0\) such that 
-\(Q(r_0) := \int_0^{2\pi} a(\sigma,r_0) d\sigma \geq \delta\)
-for every \(r_0 \in ]0,\pi[\).
-
-Unfortunately, this latter statement is not true.
-The plot of \(Q(r_0)\) for \(r_0 \in [10^{-10}, \pi - 10^{-3}]\) is shown in
-Figure \ref{fig:acrobot-Q}.
-We see that it is strictly positive and monotonically increasing, with an
-asymptote at \(r_0 = \pi\), but \(Q(r_0) \to 0\) as \(r_0 \to 0\).
-
-\begin{figure}
+%
+\begin{figure}[htb]
     \centering
     \includegraphics[width=0.8\linewidth]{Qro.png}
-    \caption{The plot of \(Q(r_0)\).}
+    \caption{The graph of \(r \mapsto \int_0^{2\pi} a(r,\theta) d\theta\).}
     \label{fig:acrobot-Q}
 \end{figure}
-
-Thankfully, we can still use Lemma \ref{lemma:energy-regulation-D} on an
-appropriate subset of \(\mathcal{O}_1\).
-To find this subset, first recall the discrete-time pseudo-radius dynamics
-\eqref{eqn:r-discrete} from the proof of Lemma \ref{lemma:energy-regulation-D}.
-Linearizing \eqref{eqn:r-discrete} at \(r_0 = 0\) yields
+%
+Notice that \(L\) is a positive constant which depends only on \(m\), \(g\), \(l\), and \(\bar{q}_a\), while \(a(r,\theta)\) is adimensional. We have
 \[
-    P^\prime(0) = 1 + I K Q^\prime(0) + R^\prime(2\pi,0,I)
-    ,
+\int_0^{2\pi} \partial_I g(r,\theta,I) \big|_{I=0} d \theta =L \int_0^{2\pi} a(r,\theta) d\theta.
 \]
-where the prime now denotes differentiation with respect to \(r\).
-One can show that
-\begin{align*}
-    Q^\prime(0) = \int \limits_0^\pi 
-    \lim \limits_{r_0 \to 0^+}\pdiff{a}{r}(\sigma,r_0) d\sigma
-    = \frac{\pi}{2\sqrt{2}} > 0
-    .
-\end{align*}
-Since \(R(2\pi,0,I)\) is \(O(I^2)\), so is
-\(R^\prime(2\pi,0,I)\).
-Hence, it can be written in the form
-\(R^\prime(2\pi,0,I) = I^2 \tilde{R}(I)\) where \(\tilde{R}(I)\) is smooth and
-zero at \(I = 0\).
-Thus, there exists \(I_0 > 0\) such that 
-\(I K Q^\prime(0) + I^2 \tilde{R}(I) > 0\) for all \(I \in \, ]0,I_0]\),
-which implies \(P^\prime(0) > 1\).
-The origin is now a repeller, which means there exists some
-(unknown) \(\mu > 0\) where \(]0,\mu]\) is negatively
-invariant for \eqref{eqn:r-discrete}, and all solutions in this
-interval flow through \(r = \mu\).
-If \(I < 0\), we have \(P^\prime(0) < 1\) and the origin is an
-asymptotically stable equilibrium for \eqref{eqn:r-discrete} instead.
-
-Let \(E_{\leq\mu}\) be the sublevel set of the energy \(E(\mu,0)\).
-Since there exists \(\delta_\mu > 0\) such that \(Q(r_0) \geq \delta_\mu\) 
-for \(r_0 \in [\mu,\pi[\), all conditions of Lemma
-\ref{lemma:energy-regulation-D} are valid on the annulus
-\(D_\mu = \mathcal{O}_1 \backslash E_{\leq\mu}\). 
-Hence, there exists some \(I_1(\mu) > 0\) such that the acrobot gains/loses
-energy everywhere on \(D_\mu\).
-Taking \(I^\star = \min\{I_0,I_1(\mu)\}\), the acrobot will gain energy on
-\(\mathcal{O}_1\) for \(I \in \, ]0,I^\star]\) and lose energy 
-for \(I \in [-I^\star,0[\).
-
-\subsection{All acrobots will rotate}
-
-We have proven that any acrobot constrained by 
-our VNHC \eqref{eqn:acrobot-constraint} will gain (or lose) energy on
-\(\mathcal{O}_1\).
-We must now prove that every acrobot will begin rotating, 
-i.e., that when \(I \in \, ]0,I^\star]\) almost every orbit initialized in
-\(\mathcal{O}_1\) will reach the line \(|q_u| = \pi\).
-
-The definition of energy gain only guarantees that orbits initialized in
-\(\mathcal{O}_1\) will escape compact subsets of \(\mathcal{O}_1\) in finite
-time. 
-For an orbit of the acrobot to reach the line \(|q_u| = \pi\), it must first
-escape the closure \(\bar{\mathcal{O}}_1 = \mathcal{O}_1 \cup E_\pi\) through
-its boundary, the level set \(E_\pi\) with energy level \(E(\pi,0)\).
-
-Taking the Jacobian of the constrained dynamics
-~\eqref{eqn:acrobot-constrained-dynamics} at the upright equilibrium 
-\((q_u,p_u) = (\pi,0)\) yields
+The graph of \(\int_0^{2\pi} a(r,\theta) d\theta\), shown in
+Figure \ref{fig:acrobot-Q}, indicates that for each $r \in \mathopen] 0,\pi\mathclose[$, the function $r \mapsto \int_0^{2\pi} a(r,\theta) d\theta$ is positive, and assumption~\eqref{eq:lemma:assumptions:3} holds.
+Thus the hypotheses of Lemma~\ref{lem:gain_lose} are satisfied, proving the theorem.
+\hfill\QED
+
+\subsection{Proof of Theorem~\ref{thm:acrobot-rotations}}
+
+We use again Lemma~\ref{lem:gain_lose}, this time producing two coordinate transformations, adapted from~\cite{dynamic_vhcs_stabilize_closed_orbits}, whose domains are the two connected components of the set $E_{>\bar R}$.
+
+Define $\cU^+ \eqdef \big(E_{>\bar R}\big) \cap \{ p_u >0\}$, and $\cU^- \eqdef \big(E_{>\bar R}\big) \cap \{ p_u <0\}$.
+Let $\cV \eqdef \mathopen] (10 m l^2 \bar R)^{1/2}, \infty \mathclose[ \times \Sone$, and consider the coordinate transformations $T^+ : \cU^+ \to \cV$ and $T^- : \cU^- \to\cV$ defined as
+\begin{equation}\label{eq:T2}
+\begin{aligned}
+T^\pm: & \, (q_u,p_u) \mapsto (r,\theta) \\
+%
+r &= \sqrt{p_u^2 + 30m^2gl^3(1-c_u)} \\
+%
+\theta &= \pm q_u,
+\end{aligned}
+\end{equation}
+whose inverses are
 \[
-    J = \begin{bmatrix}
-        -\frac{6mgl\bar{q}_aI}{5} & \frac{1 - 2m^2gl^3\bar{q}_a^2 I^2}{5ml^2} \\
-        3mgl & mgl\bar{q}_aI
-    \end{bmatrix}
-    ,
+\begin{aligned}
+(T^\pm)^{-1}: & \, (r,\theta) \mapsto (q_u,p_u)\\
+%
+q_u &= \pm \theta \\
+% 
+p_u &= \pm \sqrt{r^2-30m^2gl^3(1-c_\theta)}.
+\end{aligned}
 \]
-which has characteristic polynomial
-\[
-    \det\left(\lambda \Id{2} - J\right)
-    = \lambda^2 + \frac{mgl\bar{q}_a I}{5} \lambda - 3g
-    .
-\]
-This polynomial always has exactly one root with positive real part. 
-The equilibrium \((\pi,0)\) is therefore unstable, so the stable manifold
-\(\Pi^+\) of initial conditions converging to \((\pi,0)\) is one-dimensional,
-and hence is of measure zero in \(\SxR\).
-
-Using \(x(t) := (q_u(t),p_u(t))\) as shorthand, let
-\(x(0) \in \mathcal{O}_1\) be a nonzero initial condition of the acrobot.
-Suppose by way of contradiction that the orbit \(x(\R)\) is confined within 
-\(\bar{\mathcal{O}}_1\), and does not exit through \(E_\pi\) into the rotation
-zone.
-%Since \(\bar{\mathcal{O}}_1\) is compact, the Birkhoff Theorem \cite{birkhoff}
-%implies:
-%\begin{itemize}
-%    \item The positive limit set \(L_+\) of \(x(t)\) is non-empty, compact, and
-%        invariant.
-%    \item The solution \(x(t)\) asymptotically tends to \(L_+\).
-%\end{itemize}
-Since the acrobot gains energy on \(\mathcal{O}_1\) and \(\bar{\mathcal{O}}_1\)
-is compact, the Birkhoff Theorem \cite{birkhoff} implies that
-the solution \(x(t)\) asymptotically tends to the positive limit set, which is
-the largest invariant subset of \(E_\pi\).
-Taking the limit as \(r \to \pi\) of
-\eqref{eqn:theta-dot-nom-osc}--\eqref{eqn:theta-dot-limit},
-we get \(\dot{\theta} \geq 0\) on \(E_\pi\) with equality if and only if
-\(\theta \in \{0,\pi\}\).
-Thus, the largest invariant subset of \(E_\pi\) is either the upright
-equilibrium \(\{(\pi,0)\}\) or \(E_\pi\) itself.
-
-Taking the derivative of \(E(q_u,p_u)\) at the \(p_u\)-axis yields
-\(\dot{E}(0,p_u) = g/(5l) \sin(q_a)p_u\).
-This is non-zero on \(E_\pi\), so \(E_\pi\) is not invariant and the positive
-limit set of \(x(t)\) must be the set \(\{(\pi,0)\}\). 
-Hence, \(x(t)\) converges to \((\pi,0)\), so \(x(0) \in \Pi^+\).
-Since \(\Pi^+\) is a set of measure zero in \(\SxR\), almost every orbit
-initialized in \(\mathcal{O}_1\) must escape at least once through \(E_\pi\)
-into the rotation domain.
-
-Using this fact, we now prove that almost every orbit in \(\mathcal{O}_1\) will
-cross the line \(|q_u| = \pi\).
-Let \(x(0) \in \mathcal{O}_1\) and suppose that the orbit
-\(x([0,\infty[)\) does not hit the line \(|q_u| = \pi\).
-We claim that this orbit must remain in the compact set 
-\(\Sone \times [-K,K]\), for some \(K > 0\).
-Suppose otherwise;
-then for arbitrary \(K > 0\), there exists some \(t_K > 0\) such that
-\(|p_u(t_K)| = K\). 
-Looking at the constrained dynamics \eqref{eqn:acrobot-constrained-dynamics},
-we see that \(\dot{p}_u \geq -3mgl\) and
-\(\dot{q}_u \geq c_1 p_u + c_2\) for some constants \(c_1,c_2 > 0\).
-This implies there exists a \(t_{1/2} > 0\) (independent of \(K\)) 
-where
-\(|p_u(t)| >= K/2\) 
-and \(\dot{q}_u \geq c_1 K/2 + c_2\) 
-for \(t \in [t_K, t_K + t_{1/2}]\).
-Picking \(K\) large enough guarantees \(q_u(t)\) covers a distance of \(2\pi\)
-over this time interval, so the orbit would hit the line \(|q_u| = \pi\).
-
-If the orbit \(x(t)\) does not hit the line \(|q_u| = \pi\), we now know it must
-remain within the compact set \(\Sone \times [-K,K]\).
-By the Birkhoff theorem, \(x(t)\) approaches a positive limit set \(L^+\), 
-which is either the upright equilibrium \(\{(\pi,0)\}\), or
-a closed orbit not fully contained in \(\mathcal{O}_1\)
-which crosses the positive \(q_u\)-axis at some point \(\bar{q} < \pi\).
-Suppose \(L^+\) is the latter.
-Since \(]0,\mu]\) is negatively invariant for \eqref{eqn:r-discrete}, 
-we must have \(\bar{q} > \mu\).
-The Poincar\'{e} map \eqref{eqn:poincare-map} is smooth on \(]0,\pi[\), and its
-image contains \(\bar{q}\), so \(P^{-1}(\bar{q}) \neq \emptyset\).
-Take \(q^\prime \in P^{-1}(\bar{q})\). 
-If \(q^\prime \in \, ]0,\mu]\), then \(q^\prime < \bar{q}\);
-if \(q^\prime > \mu\), then we know 
-from the proof of Lemma \ref{lemma:energy-regulation-D} that
-\(P(q^\prime) = \bar{q} \geq q^\prime + \gamma\) for some \(\gamma > 0\), 
-so \(q^\prime \leq \bar{q} - \gamma < \bar{q}\).
-In either case, \(q^\prime < \bar{q}\), so the orbit starting at \(q^\prime\)
-stays within \(\mathcal{O}_1\) and intersects the closed orbit at \(\bar{q}\).
-This violates uniqueness of solutions for the constrained dynamics,
-so \(L^+\) must be \(\{(\pi,0)\}\), which in turn implies that 
-\(x(0) \in \Pi^+\).
-
-We conclude that almost all orbits beginning in \(\mathcal{O}_1\) will
-escape the closure of \(\mathcal{O}_1\) and begin rotating by crossing the
-vertical line \(|q_u| = \pi\).
-
-\subsection{Energy Gain on \(\mathcal{O}_2(\rho)\)}
-Since \(\mathcal{O}_2(\rho) = \bar{\mathcal{O}}_1 \cup \mathcal{R}(\rho)\)
-(where \(\mathcal{R}(\rho)\) is the rotation domain defined in
-\eqref{eqn:rotation-domain}), and since all orbits in \(\mathcal{O}_1\) will
-begin rotating, we need simply prove energy gain on
-\(\mathcal{R}(\rho)\) to prove energy gain on \(\mathcal{O}_2(\rho)\).
-We separate \(\mathcal{R}(\rho)\) into its two separate connected components:
-\(\mathcal{R}^+(\rho)\), the set of rotations with \(p_u > 0\); 
-and \(\mathcal{R}^-(\rho)\), the set of rotations with \(p_u < 0\).
-We take the domain of interest for Lemma \ref{lemma:energy-regulation-D} to be
-either of these rotation sets.
-
-We now apply
-\cite[Eqn. (8)]{dynamic_vhcs_stabilize_closed_orbits} to find pseudo-polar coordinates
-adapted to rotations.
-These coordinates are 
-\begin{align}
-    \label{eqn:r-rot}
-    r &= \sqrt{p_u^2 + 30m^2gl^3(1-c_u)}
-    , \\
-    \label{eqn:theta-rot}
-    \theta &= \sign{p_u} q_u
-\end{align}
-with \(]\underline{r},\overline{r}[ \, = \, ]\sqrt{60m^2gl^3}, \rho[\).
-The inverse transformation is
-\begin{align*}
-    q_u &= \pm \theta
-    , \\
-    p_u &= \pm \sqrt{r^2-30m^2gl^3(1-c_\theta)}
-    ,
-\end{align*}
-where the \(\pm\) is \(+1\) on \(\mathcal{R}^+(\rho)\), or \(-1\) on
-\(\mathcal{R}^-(\rho)\).
-
-After computing the constrained dynamics in
-\((r,\theta)\) coordinates, one can verify that
+For each $R> \bar R$, it is easily seen that $T^\pm(E_R) = \{(r,\theta) \in \cV : r = (10 m l^2 R)^{1/2}\}$, and the function $R \mapsto (10 m l^2 R)^{1/2}$ is monotonically increasing.
+Both $T^+$ and $T^-$, therefore, satisfy the first requirement of Lemma~\ref{lem:gain_lose} for any $\bar R < R_1 < R_2$.
+
+In $(r,\theta)$ coordinates, both $T^+$ and $T^-$ map system~\eqref{eqn:acrobot-constrained-dynamics} to a system of the form~\eqref{eq:r_theta_system}, where one can verify that
 \(f_r(\theta,r,0) \equiv 0\) and
@@
 \]
-which is strictly positive as required.
-Symbolic computations
-also confirm that \((\partial g/\partial r)(\theta,r,0) \equiv 0\).
-The function \(b(\theta,r_0) = (\partial g/\partial I)(\theta,r_0,0)\) is given
-in the statement of Theorem \ref{thm:acrobot-energy-stabilization}, and by
-assumption we have \(S(r_0) \geq \epsilon\).
-
-Thus, all requirements of Lemma \ref{lemma:energy-regulation-D} are satisfied,
-so there exists \(I_2(\rho) > 0\) where the acrobot
-gains energy for \(I \in \, ]0,I_2(\rho)]\)
-(on both \(\mathcal{R}^+(\rho)\) and \(\mathcal{R}^-(\rho)\)).
-Let \(I^\star\) be the value we found for \(\mathcal{O}_1\) in 
-Section \ref{sec:proof-o1}, and let
-\(I^{\star\star} = \min \{I^\star, I_2(\rho)\}\).
-Then the acrobot gains energy on \(\mathcal{O}_2(\rho)\)
-for \(I \in \,]0, I^{\star\star}]\),
-and dissipates energy for \(I \in [-I^{\star\star},0[\).
-
-This completes the proof of Theorem \ref{thm:acrobot-energy-stabilization}.
+which is positive so assumption~\eqref{eq:lemma_assumptions:1} holds.
+Letting $g = f_r / f _\theta$, one can verify that \((\partial g/\partial r)(\theta,r,0) \equiv 0\), and assumption~\eqref{eq:lemma:assumptions:2} holds.
+The function \(b(r,\theta) = (\partial g/\partial I)(r,\theta,0)\) is given in the statement of Theorem~\ref{thm:acrobot-rotations}, and by assumption we have that condition~\eqref{eq:lemma:assumptions:3} of Lemma~\ref{lem:gain_lose} holds.
+Theorem~\ref{thm:acrobot-rotations} now follows from  Lemma~\ref{lem:gain_lose} and the fact that $\cU^+ \cup \cU^- = E_{>\bar R}$.
+\hfill \QED
 
@@
 
-In this article we applied the framework of virtual nonholonomic constraints
-to the acrobot, and designed a constraint which emulates giant motion from
-gymnastics.
-We proved this constraint will inject energy into the simplified acrobot whose
-limbs are massless rods of equal length, with equal masses at the tips.
-We then performed simulations and physical experiments on a real acrobot robot.
-These demonstrated that virtual nonholonomic constraints are capable of
-injecting and dissipating energy in a robust manner, while producing
-realistic biologically-inspired motion.
+In this article we applied the framework of virtual nonholonomic constraints to the acrobot, and designed a \vnhc emulating a giant motion from gymnastics.
+Our theoretical analysis applies to a simplified acrobot with point-mass limbs of equal length and mass, but the simulation and experimental results show the validity of the approach for a real acrobot.
+The main theoretical results of the paper, Theorems~\ref{thm:acrobot-oscillations} and~\ref{thm:acrobot-rotations}, assert that the proposed \vnhc makes the dynamics of the acrobot on the constrained manifold gain or lose energy.
+There are two theoretical points that were unexplored in this paper.
+First, our analysis only considers initial conditions on the constraint manifold.
+For initial conditions off the manifold, there will be a transient whose effects we have not analyzed theoretically.
+Second, as pointed out in Remark~\ref{rem:thm:caveat}, we have shown energy gain/loss of the constrained dynamics on compact subsets $E_{[R_1,R_2]}$ of $E_{<\bar R}$ and $E_{>\bar R}$, so both oscillations and rotations of the acrobot will gain/lose energy.
+We have not given theoretical guarantees that oscillations gaining energy eventually turn into rotations gaining energy.
+Both our simulations and the physical experiments suggest that our theoretical predictions continue to hold in the face of transients in enforcing the constraint.
+Moreover, oscillations gaining energy turn into rotations gaining energy.
 

### 2021-10-12 latex/paper/main.tex
Final edits of the paper to send to Manfredi and Xingbo

@@
 \usepackage{generic}
+\usepackage{hyperref}
 %/========== Preamble ==========/%
@@
 \begin{abstract}
-    In this article we study virtual nonholonomic constraints, 
+    In this article we study virtual nonholonomic constraints, which are
     relations between the generalized coordinates and momenta of a
@@
 angle and velocity.
-In 2016, Wang designed a controller \cite{xingbo_thesis} which tracked the
-body angle and an \textit{estimate} of the velocity, but not the velocity itself.
+In 2016, Wang designed a controller which tracked the
+body angle and an \textit{estimate} of the velocity, but not the velocity itself
+\cite{xingbo_thesis}.
 His approach was a preliminary version of a recent technique know as the method
-of virtual nonholonomic constraints \cite{hybrid_zero_dynamics_bipedal_nhvcs}.
+of virtual nonholonomic constraints.
 
@@
 injects energy into the acrobot through human-like giant motion.
-In particular, any acrobot constrained by this \vnhc is guaranteed to perform
-rotations around the bar.
-Under suitable conditions on the acrobot's physical parameters, this \vnhc
-also enables the acrobot to rotate around the bar at a desired speed.
+%In particular, any acrobot constrained by this \vnhc is guaranteed to perform
+%rotations around the bar.
+%Under suitable conditions on the acrobot's physical parameters, this \vnhc
+%also enables the acrobot to rotate around the bar at a desired speed.
 We will provide simulations to show how one can use \vnhcs to regulate energy,
@@
 
-Our goal is to design a smooth function \(f : \Sone \times \R\) such that the relation \(q_a = f(q_u,p_u)\) for system \eqref{eqn:acrobot-hamiltonian} can be enforced asymptotically via feedback control (in Section \ref{sec:vnhc} we call this a  virtual nonholonomic constraint).
-We will further require that the dynamics of the acrobot, when the relation holds, gain or lose energy in a sense that will be defined precisely in Section~\ref{sec:energy-inject-diss}.
+Our goal is to design a smooth function \(f : \Sone \times \R \to \Sone\) such that the
+relation \(q_a = f(q_u,p_u)\) for system \eqref{eqn:acrobot-hamiltonian} can be
+enforced asymptotically via feedback control (in Section \ref{sec:vnhc} we call
+this a  virtual nonholonomic constraint).   We will further require that the
+dynamics of the acrobot, when the relation holds, gain or lose energy in a sense
+that will be defined precisely in Section~\ref{sec:energy-inject-diss}.
 
@@
 the system above is already in simply actuated form.
-Its state space is \(\Sone \times \Re \times \Re \times \Re\).
+Its state space is \(\Sone \times \Sone \times \Re \times \Re\).
 We can therefore apply the theory from Section \ref{sec:vnhc} to design
@@
 \end{equation}
-The function $E(q_u,p_u)$ is a first integral for the constrained dynamics~\eqref{eqn:acrobot-constrained-dynamics}, i.e., each solution of the constrained dynamics is confined to a level set of $E$, and each orbit not containing equilibria is a connected component of a level set of $E$.
-Given scalars $R$ and $R_1<R_2$, and letting $\# \in \{$ ``$<$''$,$''$\leq$''$,$''$>$''$,$''$\geq$''$\}$, we define the following sets (see Figure~\ref{fig:domains}):
+The function $E(q_u,p_u)$ is a first integral for the constrained
+dynamics~\eqref{eqn:acrobot-constrained-dynamics}, i.e., each solution of the
+constrained dynamics is confined to a level set of $E$, and each orbit not
+containing equilibria is a connected component of a level set of $E$. Given
+scalars $R$ and $R_1<R_2$, and letting $\# \in \{$
+``$<$''$,$``$\leq$''$,$``$>$''$,$``$\geq$''$\}$, we define the following sets
+(see Figure~\ref{fig:domains}):
 \begin{equation}\label{eq:E_sets}
@@
 
-Letting $\bar R \eqdef E(\pi,0)$, solutions initialized in the interior of the set $E_{\leq \bar R}$ correspond to {\em oscillations} of the pendulum, i.e., rocking motions where the pendulum does not perform full revolutions.
-On the other hand, for each $R>\bar R$, solutions initialized in the interior of the set $E_{[\bar R,R]}$ correspond to {\em rotations} of the pendulum, i.e., full revolutions around the pivot point.
+Letting $\bar R \eqdef E(\pi,0)$, solutions initialized in the interior of the
+set $E_{\leq \bar R}$ correspond to {\em oscillations} of the pendulum, i.e.,
+rocking motions where the pendulum does not perform full revolutions. On the
+other hand, for each $R>\bar R$, solutions initialized in the interior of the
+set $E_{[\bar R,R]}$ correspond to {\em rotations} of the pendulum, i.e., full
+revolutions around the pivot point.
 
 Now consider the case $I\neq 0$, i.e., when the acrobot is not fully extended.
-The constrained dynamics~\eqref{eqn:acrobot-constrained-dynamics} are no longer Hamiltonian, and their solutions are no longer confined to level sets of $E$.
-In what follows, we will show that the orbits of the constrained dynamics roughly speaking gain energy or lose energy depending on the sign of the design parameter $I$.
-To be more precise in describing this energy injection or dissipation property, we next define what is an oscillation or a rotation gaining or losing energy.
-Intuitively, oscillations gaining energy correspond to rocking motions of the torso with increasing amplitude, while rotations gaining energy correspond to full revolutions with increasing speed.
+The constrained dynamics~\eqref{eqn:acrobot-constrained-dynamics} are no longer
+Hamiltonian, and their solutions are no longer confined to level sets of $E$. In
+what follows, we will show that the orbits of the constrained dynamics roughly
+speaking gain energy or lose energy depending on the sign of the design
+parameter $I$. To be more precise in describing this energy injection or
+dissipation property, we next define what is an oscillation or a rotation
+gaining or losing energy. Intuitively, oscillations gaining energy correspond to
+rocking motions of the torso with increasing amplitude, while rotations gaining
+energy correspond to full revolutions with increasing speed.
 Figure~\ref{fig:domains} illustrates these concepts.
@@
 
-\begin{thm}[rotations gaining/losing energy]\label{thm:acrobot-rotations} Consider the constrained dynamics in~\eqref{eqn:acrobot-constrained-dynamics}, and fix $R_1, R_2$ such that \(\bar R<R_1<R_2\).
-Suppose there exists \(\varepsilon > 0\) such that \(S( r) \geq \varepsilon\) for all \( r \in \, [(10 m l^2 R_1)^{1/2}, (10 m l^2 R_2)^{1/2}]\).
-Then there exists \(I^{\star}>0\) such that, for each \(I \in \, ]0,I^{\star}]\) (respectively, $I \in [-I^\star,0\mathclose[$) and $x_0 \in E_{[R_1,R_2]}$, there exists $\bar t \geq 0$ such that $x(\bar t,x_0) \in E_{R_2}$  (respectively, $x(\bar t, x_0) \in E_{R_1}$), and $x(t,x_0)$ is a rotation gaining energy (respectively, losing energy) over $[0, \bar t]$.
+\begin{thm}[rotations gaining/losing energy]\label{thm:acrobot-rotations}
+    Consider the constrained dynamics
+    in~\eqref{eqn:acrobot-constrained-dynamics}, and fix $R_1, R_2$ such that
+    \(\bar R<R_1<R_2\). Suppose there exists \(\varepsilon > 0\) such that \(S(
+    r) \geq \varepsilon\) for all \( r \in \, [(10 m l^2 R_1)^{1/2}, (10 m l^2
+    R_2)^{1/2}]\). Then there exists \(I^{\star}>0\) such that, for each \(I \in
+    \, ]0,I^{\star}]\) (respectively, $I \in [-I^\star,0\mathclose[$) and $x_0
+    \in E_{[R_1,R_2]}$, there exists $\bar t \geq 0$ such that $x(\bar t,x_0)
+    \in E_{R_2}$  (respectively, $x(\bar t, x_0) \in E_{R_1}$), and $x(t,x_0)$
+    is a rotation gaining energy (respectively, losing energy) over $[0, \bar
+    t]$. 
 \end{thm}
@@
 \begin{rem}
-Theorem~\ref{thm:acrobot-oscillations} concerning oscillations gai\-ning/lo\-sing energy states that for any set of physical parameters, the acrobot constrained by
-\eqref{eqn:acrobot-constraint} will gain enough energy to exit in finite time a level set of the function $E$ arbitrarily close to the boundary $E_{\bar R}$ separating oscillations from rotations, provided the parameter $I$ in~\eqref{eqn:acrobot-constraint} is chosen small enough.
-The result is semiglobal relative to the set $E_{<\bar R}$ in that energy gain can be achieved on any compact subset $E_{[R_1,R_2]}$ of $E_{<\bar R}$ by a suitably small $I$.
+Theorem~\ref{thm:acrobot-oscillations} concerning oscillations
+gai\-ning/lo\-sing energy states that for any set of physical parameters, the
+acrobot constrained by
+\eqref{eqn:acrobot-constraint} will gain enough energy to exit (in finite time) a
+level set of the function $E$ arbitrarily close to the boundary $E_{\bar R}$
+separating oscillations from rotations, provided the parameter $I$
+in~\eqref{eqn:acrobot-constraint} is chosen small enough. The result is
+semiglobal relative to the set $E_{<\bar R}$ in that energy gain can be achieved
+on any compact subset $E_{[R_1,R_2]}$ of $E_{<\bar R}$ by a suitably small $I$.
 %% Moreover, after this finite time, as long as the solution is contained in the set $E_{< \bar R}$, then its intersections with the Poincar\'e section $\cP_o$ will occur outside of $E_{<R}$.
@@
 
-We will now simulate the effects of constraining the physical acrobot with the \vnhc \eqref{eqn:acrobot-constraint}, thereby demonstrating that \vnhcs are robust to model mismatch.
-According to Theorems~\ref{thm:acrobot-oscillations} and~\ref{thm:acrobot-rotations}, the control parameter \(I\) must be ``small" for the \vnhc~\eqref{eqn:acrobot-constraint} to inject (or dissipate) energy into the simplified acrobot.
-The Theorem does not specify how small \(|I|\) must be; while we could make it arbitrarily small in simulations, we will eventually implement this \vnhc on a physical testbed where \(|I|\) must be large enough to overcome friction.
+We will now simulate the effects of constraining the physical acrobot with the
+\vnhc \eqref{eqn:acrobot-constraint}, thereby demonstrating that \vnhcs are
+robust to model mismatch. According to Theorems~\ref{thm:acrobot-oscillations}
+and~\ref{thm:acrobot-rotations}, the control parameter \(I\) must be ``small"
+for the \vnhc~\eqref{eqn:acrobot-constraint} to inject (or dissipate) energy
+into the simplified acrobot. The theorems do not specify how small \(|I|\) must
+be; while we could make it arbitrarily small in simulations, we will eventually
+implement this \vnhc on a physical testbed where \(|I|\) must be large enough to
+overcome friction.
 
@@
     \includegraphics[width=0.8\linewidth]{acrobot_unperturbed_orbit.png}
-    \caption{Baseline Test: SUGAR's baseline energy injection orbit. A video of this experiment is available at {\footnotesize https://play.library.utoronto.ca/watch/6703e828acdcd38058ca884d13660a88}}
+    \caption{Baseline Test: SUGAR's baseline energy injection orbit. A video of this experiment is available at 
+    {\footnotesize
+\url{https://play.library.utoronto.ca/watch/6703e828acdcd38058ca884d13660a88}}.}
     \label{fig:acrobot-unperturbed-orbit}
@@
     \includegraphics[width=0.8\linewidth]{acrobot_stopped_orbit.png}
-    \caption{Perturbation Test 1: SUGAR's orbit before (blue) and after (red) stopping.
-A video of this experiment is available at {\footnotesize https://play.library.utoronto.ca/watch/164a21f0a52d4d56ef60bf1911545e0e}}
+    \caption{Perturbation Test 1: SUGAR's orbit before (blue) and after (red)
+    stopping. A video of this experiment is available at {\footnotesize
+\url{https://play.library.utoronto.ca/watch/164a21f0a52d4d56ef60bf1911545e0e}}.}
     \label{fig:acrobot-stopped-orbit}
@@
     \end{subfigure}
-    \caption{Perturbation Test 2: SUGAR's orbit before (blue) and after (red) pushing.
-A video of this experiment is available at {\footnotesize https://play.library.utoronto.ca/watch/b13d4dc9f6daefa7f4d56dbf1b1bd482.}}
+    \caption{Perturbation Test 2: SUGAR's orbit before (blue) and after (red)
+    pushing. A video of this experiment is available at 
+    {\footnotesize\url{https://play.library.utoronto.ca/watch/b13d4dc9f6daefa7f4d56dbf1b1bd482}.}
+    }
 \end{figure}
@@
 \end{equation}
-By properties~\eqref{eq:lemma_assumptions:1} and~\eqref{eq:lemma:assumptions:3} and since $f_r, f_\theta$ are continuous and $K$ is compact, there exist $I_1,\varepsilon_1,\varepsilon_2>0$ such that for each $(r,\theta) \in K$ and $I \in [-I_1,I_1]$,
+By properties~\eqref{eq:lemma_assumptions:1} and~\eqref{eq:lemma:assumptions:3}
+and since $f_r, f_\theta$ are continuous and $K$ is compact, there exist
+$I_1,\varepsilon_1,\varepsilon_2>0$ such that for each $(r,\theta) \in K$ and $I
+\in [-I_1,I_1]$,
 \begin{subequations}
@@
 \begin{align}
-&f_r(r,\theta,I) > \varepsilon_1,\label{eq:properties:new_coords:1}\\
+&f_\theta(r,\theta,I) > \varepsilon_1,\label{eq:properties:new_coords:1}\\
 & \int_0^{2\pi} b(r,\theta) d \theta > \varepsilon_2.\label{eq:properties:new_coords:2}
@@
 \end{subequations}
-Property~\eqref{eq:properties:new_coords:1} implies that for $I \in [-I_1,I_1]$ the first component of any solution $(r(t),\theta(t))$ contained in $K$ is a strictly monotonic function of $t$, and therefore $r(t)$  can be expressed as $r(t) = \hat r(\theta(t))$, where $\hat r(\theta)$ is a solution of
+Property~\eqref{eq:properties:new_coords:1} implies that for $I \in [-I_1,I_1]$
+the second component of any solution $(r(t),\theta(t))$ contained in $K$ is a
+strictly monotonic function of $t$, and therefore $r(t)$  can be expressed as
+$r(t) = \hat r(\theta(t))$, where $\hat r(\theta)$ is a solution of
     \begin{equation}\label{eqn:rhat-ode}
@@
 \hat{r}(\theta,r_0,I) &= \hat{r}(\theta,r_0,0) + I\hat{r}_1(\theta,r_0) 
-    + \Delta(\theta,r_0,I) \\
+    + \Delta(r_0,\theta,I) \\
 %
@@
 \end{equation}
-where the function $\Delta(r,\theta,I)$ is continuous and $I \mapsto \Delta(r,\theta,I)$ is $O(I^2)$.
+where the function $\Delta(r,\theta,I)$ is continuous and $I \mapsto
+\Delta(r,\theta,I)$ is $O(I^2)$.
 The second identity follows from the fact (due to~\eqref{eq:lemma_assumptions:1}) that \(g(\hat{r},\theta,0) \equiv 0\).
@@
 \end{equation}
-Now consider the Poincar\'e section $\cP = \{(r,\theta) \in \cV: \theta =0\}$ defined in the lemma statement.
-Using~\eqref{eq:Taylor_exp}, and~\eqref{eq:ronehat_solution}, the Poincar\'e map on $\cP$ is $P: \cP \to \Re_{>0}$,
+Now consider the Poincar\'e section $\cP = \{(r,\theta) \in \cV: \theta =0\}$. Using~\eqref{eq:Taylor_exp},
+and~\eqref{eq:ronehat_solution}, the Poincar\'e map on $\cP$ is $P: \cP \to
+\Re_{>0}$, 
 \begin{equation}\label{eq:Poincare_map:expansion}
@@
 
-Define $\cU \eqdef E_{<\bar R} \setminus \{0\}$ and $\cV \eqdef \mathopen]0,\pi\mathclose[$, and consider the coordinate transformation\footnote{Despite the presence of a $\text{sgn}(\cdot)$ function, the coordinate transformation $T_1$ is actually smooth.} $T: \cU \to \cV$ defined as
+Define $\cU \eqdef E_{<\bar R} \setminus \{0\}$ and $\cV \eqdef
+\mathopen]0,\pi\mathclose[ \, \times \Sone$, and consider the coordinate
+transformation\footnote{Despite the presence of a $\text{sgn}(\cdot)$ function,
+this coordinate transformation $T$ is actually smooth.} $T: \cU \to \cV$
+defined as
 \begin{equation}\label{eq:T1}
@@
 \]
-It is easily seen that for each $R \in \mathopen ]0,\bar R\mathclose[$, $T(E_R) = \{(r,\theta)\in \cV : r = \acos(1 - R/(3 m g l))\}$, and the function $R \mapsto \acos(1 - R/(3 m g l))$ is monotonically increasing.
-The diffeomorphism $T: \cU \to \cV$ therefore satisfies the first requirement of Lemma~\ref{lem:gain_lose} for any choice of $0< R_1 < R_2 < \bar R$.
-
-In $(r,\theta)$ coordinates, $T$ maps system~\eqref{eqn:acrobot-constrained-dynamics} into a system of the form~\eqref{eq:r_theta_system}, where is straightforward to show that \(f_r(\theta,r,0) \equiv 0\)
+It is easily seen that for each $R \in \mathopen ]0,\bar R\mathclose[$, 
+$T(E_R) = \{(r,\theta)\in \cV : r = \arccos(1 - R/(3 m g l))\}$, and the function
+$R \mapsto \arccos(1 - R/(3 m g l))$ is monotonically increasing. The
+diffeomorphism
+$T: \cU \to \cV$ therefore satisfies the first requirement of
+Lemma~\ref{lem:gain_lose} for any choice of $0< R_1 < R_2 < \bar R$.
+
+In $(r,\theta)$ coordinates, $T$ maps
+system~\eqref{eqn:acrobot-constrained-dynamics} into a system of the
+form~\eqref{eq:r_theta_system}, where is straightforward to show that
+\(f_r(r,\theta,0) \equiv 0\)
 and
 \begin{equation} \label{eqn:theta-dot-nom-osc}
-    f_\theta(\theta,r,0) = \sqrt{\frac{6g}{5l}} 
+    f_\theta(r,\theta,0) = \sqrt{\frac{6g}{5l}} 
         \sqrt{\frac{\cos(r c_\theta) - c_r}
@@
 \begin{equation}\label{eqn:theta-dot-limit}
-    \lim\limits_{\theta \to 0}f_\theta(\theta,r,0)
-    = \lim\limits_{\theta \to \pi} f_\theta(\theta,r,0)
+    \lim\limits_{\theta \to 0}f_\theta(r,\theta,0)
+    = \lim\limits_{\theta \to \pi} f_\theta(r,\theta,0)
     = \sqrt{\frac{6g s_r}{10lr}}
@@
 \end{equation}
-and these limits smooth and well-defined for all \(r \in \, ]0,\pi[\).
-From \eqref{eqn:theta-dot-nom-osc}--\eqref{eqn:theta-dot-limit}, one can verify that \(f_\theta(\theta,r,0) > 0\) for all $(\theta,r) \in \cV$, and thus assumption~\eqref{eq:lemma_assumptions:1} holds.
-Letting \(g= f_r/f_\theta\), one can also verify that \((\partial g/\partial \hat r)(\theta,\hat r,0) \equiv 0\), and assumption~\eqref{eq:lemma:assumptions:2} holds.
-Symbolic computations reveal that \(\partial_I g(r,\theta,I) \big|_{I=0} = L a(r,\theta)\), where
+and these limits are smooth and well-defined for all \(r \in \, ]0,\pi[\). From
+\eqref{eqn:theta-dot-nom-osc}--\eqref{eqn:theta-dot-limit}, one can verify that
+\(f_\theta(r,\theta,0) > 0\) for all $(r,\theta) \in \cV$, and thus
+assumption~\eqref{eq:lemma_assumptions:1} holds. Letting \(g= f_r/f_\theta\),
+one can also verify that \(\partial_r g(r,\theta,0) \equiv
+0\), and assumption~\eqref{eq:lemma:assumptions:2} holds. Symbolic computations
+reveal that \(\partial_I g(r,\theta,I) \big|_{I=0} = L a(r,\theta)\), where
 \begin{align*}
@@
 
-In $(r,\theta)$ coordinates, both $T^+$ and $T^-$ map system~\eqref{eqn:acrobot-constrained-dynamics} to a system of the form~\eqref{eq:r_theta_system}, where one can verify that
-\(f_r(\theta,r,0) \equiv 0\) and
+In $(r,\theta)$ coordinates, both $T^+$ and $T^-$ map
+system~\eqref{eqn:acrobot-constrained-dynamics} to a system of the
+form~\eqref{eq:r_theta_system}, where one can verify that
+\(f_r(r,\theta,0) \equiv 0\) and
 \[
- f_\theta(\theta,r,0) = \frac{1}{5ml^2} \sqrt{r^2 - 30m^2gl^3(1-c_\theta)}
+ f_\theta(r,\theta,0) = \frac{1}{5ml^2} \sqrt{r^2 - 30m^2gl^3(1-c_\theta)}
 ,
 \]
-which is positive so assumption~\eqref{eq:lemma_assumptions:1} holds.
-Letting $g = f_r / f _\theta$, one can verify that \((\partial g/\partial r)(\theta,r,0) \equiv 0\), and assumption~\eqref{eq:lemma:assumptions:2} holds.
-The function \(b(r,\theta) = (\partial g/\partial I)(r,\theta,0)\) is given in the statement of Theorem~\ref{thm:acrobot-rotations}, and by assumption we have that condition~\eqref{eq:lemma:assumptions:3} of Lemma~\ref{lem:gain_lose} holds.
-Theorem~\ref{thm:acrobot-rotations} now follows from  Lemma~\ref{lem:gain_lose} and the fact that $\cU^+ \cup \cU^- = E_{>\bar R}$.
+which is positive so assumption~\eqref{eq:lemma_assumptions:1} holds. Letting $g
+= f_r / f _\theta$, one can verify that \(\partial_r g (r,\theta,0)
+\equiv 0\), and assumption~\eqref{eq:lemma:assumptions:2} holds. The function
+\(b(r,\theta) = \partial_I g(r,\theta,0)\) is given in the statement
+of Theorem~\ref{thm:acrobot-rotations}, and by assumption we have that
+condition~\eqref{eq:lemma:assumptions:3} of Lemma~\ref{lem:gain_lose} holds.
+Theorem~\ref{thm:acrobot-rotations} now follows from  Lemma~\ref{lem:gain_lose}
+and the fact that $\cU^+ \cup \cU^- = E_{>\bar R}$.
 \hfill \QED
@@
 
-In this article we applied the framework of virtual nonholonomic constraints to the acrobot, and designed a \vnhc emulating a giant motion from gymnastics.
-Our theoretical analysis applies to a simplified acrobot with point-mass limbs of equal length and mass, but the simulation and experimental results show the validity of the approach for a real acrobot.
-The main theoretical results of the paper, Theorems~\ref{thm:acrobot-oscillations} and~\ref{thm:acrobot-rotations}, assert that the proposed \vnhc makes the dynamics of the acrobot on the constrained manifold gain or lose energy.
-There are two theoretical points that were unexplored in this paper.
-First, our analysis only considers initial conditions on the constraint manifold.
-For initial conditions off the manifold, there will be a transient whose effects we have not analyzed theoretically.
-Second, as pointed out in Remark~\ref{rem:thm:caveat}, we have shown energy gain/loss of the constrained dynamics on compact subsets $E_{[R_1,R_2]}$ of $E_{<\bar R}$ and $E_{>\bar R}$, so both oscillations and rotations of the acrobot will gain/lose energy.
-We have not given theoretical guarantees that oscillations gaining energy eventually turn into rotations gaining energy.
-Both our simulations and the physical experiments suggest that our theoretical predictions continue to hold in the face of transients in enforcing the constraint.
-Moreover, oscillations gaining energy turn into rotations gaining energy.
+In this article we applied the framework of virtual nonholonomic constraints to
+the acrobot, and designed a \vnhc emulating giant motion from gymnastics. Our
+theoretical analysis applies to a simplified acrobot with point-mass limbs of
+equal length and mass, but the simulation and experimental results show the
+validity of the approach for a real acrobot.  The main theoretical results of
+the paper, Theorems~\ref{thm:acrobot-oscillations}
+and~\ref{thm:acrobot-rotations}, assert that the proposed \vnhc makes the
+dynamics of the acrobot on the constrained manifold gain or lose energy. There
+are two theoretical points that were unexplored in this paper. First, our
+analysis only considers initial conditions on the constraint manifold. For
+initial conditions off the manifold, there will be a transient whose effects we
+have not analyzed theoretically. Second, as pointed out in
+Remark~\ref{rem:thm:caveat}, we have shown energy gain/loss of the constrained
+dynamics on compact subsets $E_{[R_1,R_2]}$ of $E_{<\bar R}$ and $E_{>\bar R}$,
+so both oscillations and rotations of the acrobot will gain/lose energy. We have
+not given theoretical guarantees that oscillations gaining energy eventually
+turn into rotations gaining energy.
+Both our simulations and the physical experiments suggest that our theoretical
+predictions continue to hold in the face of transients in enforcing the
+constraint, and that oscillations gaining energy will turn into rotations
+gaining energy.
 

### 2021-12-08 latex/paper/main.tex
Updated the paper to include the new paper I found which uses VNHCs of the form h = qa - f(qu,pu)

@@
 \thanks{Manuscript submitted for review on \today.}
-\thanks{A. Moran-MacDonald (e-mail: adan.moran@mail.utoronto.ca) and
-M. Maggiore (e-mail: maggiore@control.utoronto.ca) are with the Department of
+\thanks{A. Moran-MacDonald (e-mail: adan.moran@mail.utoronto.ca),
+M. Maggiore (e-mail: maggiore@control.utoronto.ca), and
+X. Wang (e-mail: xingbo.wang@utoronto.ca) are with the Department of
 Electrical and Computer Engineering, University of Toronto, ON, Canada.}
-\thanks{X. Wang is with ??? (e-mail: ???).}
 } %/author
@@
 
-Griffin and Grizzle \cite{nhvc_dynamic_walking} were the first to define relative degree two nonholonomic constraints which can be enforced through state feedback.
-Horn \etal later extended their results in \cite{hybrid_zero_dynamics_bipedal_nhvcs} to derive the constrained dynamics for a certain class of mechanical systems.
-These researchers made use of the unactuated conjugate momentum, but they developed their results in the Lagrangian framework.
-In particular, they focused on Lagrangian systems with degree of underactuation one.
-We will now present a special case of \cite{hybrid_zero_dynamics_bipedal_nhvcs} for simply actuated Hamiltonian systems, so that the theory we apply to the acrobot is provided in its clearest form.
-After we finish this summary we will clarify the relationship between our material and that of \cite{hybrid_zero_dynamics_bipedal_nhvcs}.
-
-The rest of this section refers to the simply actuated system \eqref{eqn:simple-hamiltonian}.
-For simplicity of notation, we relabel \((\tilde{q},\tilde{p})\) to \((q,p)\).
-We assume henceforth that $(q,p) \in \cQ \times \Re^n$.
+Griffin and Grizzle \cite{nhvc_dynamic_walking} were the first to define
+relative degree two nonholonomic constraints which can be enforced through state
+feedback. Horn \etal extended their results in
+\cite{hybrid_zero_dynamics_bipedal_nhvcs} to derive the constrained dynamics for
+a certain class of mechanical systems. 
+Hamed and Ames \cite{nonholonomic_hybrid_zero_dynamics} then found an explicit
+representation of these constrained dynamics for virtual
+nonholonomic constraints with a specific form.
+All these researchers made use of the
+unactuated conjugate momentum, but they developed their results in the
+Lagrangian framework. In particular, they focused on Lagrangian systems with
+degree of underactuation one. We will now present a special case of
+\cite{hybrid_zero_dynamics_bipedal_nhvcs} for simply actuated Hamiltonian
+systems, so that the theory we apply to the acrobot is provided in its clearest
+form. After we finish this summary we will clarify the relationship between our
+material and that of 
+\cite{hybrid_zero_dynamics_bipedal_nhvcs,nonholonomic_hybrid_zero_dynamics}.
+
+The rest of this section refers to the simply actuated system
+\eqref{eqn:simple-hamiltonian}. For simplicity of notation, we relabel
+\((\tilde{q},\tilde{p})\) to \((q,p)\). We assume henceforth that 
+$(q,p) \in \cQ \times \Re^n$.
 
 \begin{defn}\label{defn:vnhc}
-    A \textit{virtual nonholonomic constraint} (\vnhc) \textit{of order \(k\)} is a     relation \(h(q,p) = 0\) where \(h : \mathcal{Q}\times\R^n \rightarrow \R^k\) is     \(C^2\), \(\rank{\left[ dh_q,\, dh_p \right]} = k\) for all      \((q,p) \in h\inv(0)\), and there exists a feedback controller \(\tau(q,p)\)     rendering the \textit{constraint manifold} \(\Gamma\) invariant,     where
+    A \textit{virtual nonholonomic constraint} (\vnhc) \textit{of order \(k\)}
+    is a relation \(h(q,p) = 0\) where 
+    \(h : \mathcal{Q}\times\R^n \rightarrow \R^k\) is \(C^2\),
+    \(\rank{\left[ dh_q,\, dh_p \right]} = k\) for all 
+    \((q,p) \in h\inv(0)\), and there exists a feedback
+    controller \(\tau(q,p)\) rendering the \textit{constraint manifold}
+    \(\Gamma\) invariant, where
     \begin{equation}
@@
 
-The constraint manifold is a \(2(n-k)\)-dimensional closed embedded submanifold of \(\mathcal{Q} \times \R^n\).
-A \vnhc thereby allows us to study a reduced-order model of the system: it reduces the original \(2n\) differential equations to \(2(n-k)\) equations.
-In particular, if the mechanical system has degree of underactuation one, i.e., \(k = (n-1)\), the constraint manifold is \textit{always} two-dimensional.
-
-In order to enforce the constraint \(h(q,p) = 0\), we want to asymptotically stabilize the set \(\Gamma\).
-To see when this is possible, let us define the error output \(e = h(q,p)\).
-If any component of \(e_i\) has relative degree 1, we may not be able to stabilize \(\Gamma\): we can always guarantee \(e_i \to 0\), but not necessarily \(\dot{e}_i \to 0\).
-It is for this reason that we define the following special type of \vnhc.
+The constraint manifold is a \(2(n-k)\)-dimensional closed embedded submanifold
+of \(\mathcal{Q} \times \R^n\). A \vnhc thereby allows us to study a
+reduced-order model of the system: it reduces the original \(2n\) differential
+equations to \(2(n-k)\) equations. In particular, if the mechanical system has
+degree of underactuation one, i.e., \(k = (n-1)\), the constraint manifold is
+\textit{always} two-dimensional.
+
+In order to enforce the constraint \(h(q,p) = 0\), we want to asymptotically
+stabilize the set \(\Gamma\). To see when this is possible, let us define the
+error output \(e = h(q,p)\). If any component of \(e_i\) has relative degree 1,
+we may not be able to stabilize \(\Gamma\): we can always guarantee 
+\(e_i \to 0\), but not necessarily \(\dot{e}_i \to 0\). It is for this reason
+that we define the following special type of \vnhc.
 
@@
 assumption \textbf{H3} holds true.
-The only real distinction between this section and their work is that
-our constrained dynamics are explicit functions of the Hamiltonian coordinates
-\((q_u,p_u)\).
+Constraints of this form are used by Hamed and Ames in
+\cite{nonholonomic_hybrid_zero_dynamics}.
+%The only real distinction between this section and their work is that
+%our constrained dynamics are explicit functions of the Hamiltonian coordinates
+%\((q_u,p_u)\).
 In fact, one can show that the constrained dynamics \eqref{eqn:qpu-dynamics}
-coincide with the constrained dynamics in
+are exactly the constrained dynamics in
+\cite[Eqn. (9)]{nonholonomic_hybrid_zero_dynamics}, which coincide with 
 \cite[Eqn. (17)]{hybrid_zero_dynamics_bipedal_nhvcs}
@@
 \(\theta_2 = p_u\).
-This explicit representation will be beneficial when we apply the theory of
-\vnhcs to the acrobot.
+The only real distinction between this section and the work of
+\cite{nonholonomic_hybrid_zero_dynamics,hybrid_zero_dynamics_bipedal_nhvcs} is
+that our constrained dynamics are derived in the Hamiltonian framework. 
+%This explicit representation will be beneficial when we apply the theory of
+%\vnhcs to the acrobot.
 

### 2023-06-22 1.SUBMISSION/vnhcs.tex
Changed file structure in preparation for the revision

--- 2020-12-27 latex/vnhcs.tex
@@
 \end{defn}
-Using the kronecker product, one can collect the derivative of momentum into
+Using the Kronecker product, one can collect the derivative of momentum into
 the vector form
@@
     \(\pdiff{Q_i}{p_m} = \pdiff{P_i}{q_m} = 0\) for all 
-    \(i,m \in \n\).
+    \(i,m \in \{1,\ldots,n\}\).
     Hence, 
@@
     Assume that the input matrix \(B\) is
-    left semi-orthogonal, \ie, \(B\tpose B = I_k\). 
+    left semi-orthogonal, \ie \(B\tpose B = I_k\). 
 \end{assm}

### 2023-06-22 sugar_bot/arduino/libraries/SUGAR_VNHC_Library/src/AcrobotDynamics.cpp

@@ -1,188 +1,0 @@
-/*******************************************************************************
-* File:	        AcrobotDynamics.cpp
-*		 
-* Author:       Adan Moran-MacDonald
-* Created:      17/Jan/20
-*
-* Systems Control Group
-* Department of Electrical and Computer Engineering
-* University of Toronto
-*
-* Last Modified: 17/Jan/20
-* Last Editor:   Adan Moran-MacDonald
-* Description:   Implements AcrobotDynamics.h
-*******************************************************************************/
-
-#include "include/AcrobotDynamics.h"
-
-namespace SUGAR
-{
-
-///////////////////
-// Matrix Object //
-///////////////////
-Matrix2::Matrix2(double a11, double a12, double a21, double a22)
-{
-    matrix_[0][0] = a11;
-    matrix_[0][1] = a12;
-    matrix_[1][0] = a21;
-    matrix_[1][1] = a22;
-}
-
-auto Matrix2::at(unsigned int i, unsigned int j) const -> double
-{
-    if(i <= 0 || i > 2 || j <= 0 || j > 2)
-    {
-        return NAN;
-    }
-    // Arrays are 0-based, while our row/col is 1-based
-    return matrix_[i-1][j-1];
-}
-
-////////////////////
-// AcrobotInertia //
-////////////////////
-
-AcrobotInertia::AcrobotInertia(double m, double l)
-: AcrobotInertia(
-    m, m,
-    l, l,
-    l, l,
-    0, 0)
-{}
-
-AcrobotInertia::AcrobotInertia(
-double mt, double ml,
-double dt, double dl,
-double lt, double ll,
-double Jt, double Jl
-)
-:   mt_(mt),
-    ml_(ml),
-    dt_(dt),
-    dl_(dl),
-    lt_(lt),
-    ll_(ll),
-    Jt_(Jt),
-    Jl_(Jl),
-    JlJt_(Jl*Jt),
-    Jlmldt2_(Jl*ml*dt*dt),
-    Jtmlll2_(Jt*ml*ll*ll),
-    Jlmtlt2_(Jl*mt*lt*lt),
-    ll2lt2mlmt_(ll*ll*lt*lt*ml*mt),
-    mlll2_(ml*ll*ll),
-    mldt2_(ml*dt*dt),
[…]

### 2023-06-22 sugar_bot/arduino/libraries/SUGAR_VNHC_Library/src/Phase.cpp

@@ -1,62 +1,0 @@
-/*******************************************************************************
-* File:	        Phase.cpp
-*		 
-* Author:       Adan Moran-MacDonald
-* Created:      14/Jan/20
-*
-* Systems Control Group
-* Department of Electrical and Computer Engineering
-* University of Toronto
-*
-* Last Modified: 14/Jan/20
-* Last Editor:   Adan Moran-MacDonald
-* Description:   Source file for the Phase library
-*******************************************************************************/
-
-#include "include/Phase.h"
-
-namespace SUGAR
-{
-
-//////////////////
-// Constructors //
-//////////////////
-
-Phase::Phase()
-: qu(0)		    ,
-  qa(0)		    ,
-  pu(0)		    ,
-  pa(0)		    ,
-  E(0) 		    
-{}
-
-Phase::Phase(const AcrobotInertia& M, const State& state)
-: Phase()
-{
-    // Set the State
-    qu = state.psi;
-    qa = state.alpha;
-
-    //  Set the conjugate of momenta
-    //  p = M(q)*[dpsi; dalpha]
-    auto Mq = M.at(state);
-    pu = Mq.at(1,1)*state.dpsi + Mq.at(1,2)*state.dalpha;
-    pa = Mq.at(2,1)*state.dpsi + Mq.at(2,2)*state.dalpha;
-
-    // Set the energy
-    E = state.E; 
-}
-
-////////////////////
-// Public Methods //
-////////////////////
-
-auto Phase::unactuatedPhase() -> UnactuatedPhase
-{
-    UnactuatedPhase qpu;
-    qpu.qu = qu; qpu.pu = pu;
-    return qpu;
-}
-
-}; // namespace SUGAR
-/* vim: set tw=80 ts=4 sw=4 sts=0 et ffs=unix : */

### 2023-06-22 sugar_bot/arduino/libraries/SUGAR_VNHC_Library/src/VNHC.cpp

@@ -1,210 +1,0 @@
-/*******************************************************************************
-* File:	        VNHC.cpp
-*		 
-* Author:       Adan Moran-MacDonald
-* Created:      04/Feb/20
-*
-* Systems Control Group
-* Department of Electrical and Computer Engineering
-* University of Toronto
-*
-* Last Modified: 04/Feb/20
-* Last Editor:   Adan Moran-MacDonald
-* Description:   Implements VNHC.h, the virutal nonholonomic constraint code
-*******************************************************************************/
-
-#include "include/VNHC.h"
-// For tanh, cosh, sin, cos, etc
-#include <math.h>
-
-namespace SUGAR
-{
-/////////////////
-// AcrobotVNHC //
-/////////////////
-
-AcrobotVNHC::AcrobotVNHC(const Acrobot& acrobot)
-: acrobot_(acrobot),
-  qmax_(M_PI)
-{}
-
-AcrobotVNHC::AcrobotVNHC(const Acrobot& acrobot, ActuatorLimit qmax)
-: acrobot_(acrobot)
-{
-    // Limit ourselves to have an actuator limit within -PI to PI, since this is
-    // an acrobot after all.
-    if(qmax < 0)
-    {
-        qmax = 0;
-    }
-    else if(qmax > M_PI)
-    {
-        qmax = M_PI;
-    }
-    qmax_ = qmax;
-}
-
-AcrobotVNHC::~AcrobotVNHC()
-{}
-
-auto AcrobotVNHC::pa(const UnactuatedPhase& qpu) const -> double
-{
-    // Compute pa with the equation from VNHC research.
-    // This equation is of the form
-    // pa = (dhq*Minv*e1)^{-1} * (dhpu*dP/dqu - dhq*Minv*e1*pu)
-    
-    // The acrobot inertia matrix depends on qa, so take qa = f(qu,pu) and
-    // use that in the configuration state.
-    Configuration q;
-    q.psi = qpu.qu; q.alpha = qa(qpu);
-    auto Minv = acrobot_.M().inverseAt(q);
-
-    // Now compute dhq*Minv, which is a row vector with two components
-    auto dhqu = dqu(qpu);
[…]

### 2023-06-27 2.REVISION/REPLY_REVIEWERS/reply_reviewers.tex
Revised paper, resubmitted to the IEEE TCST. Also two folders containing literature and the reply to reviewers.

\documentclass[11pt,letter]{article}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amstext,pdfsync}
\usepackage[dvips]{color}
\usepackage{tcolorbox}
\usepackage{mathpazo}
\usepackage{amssymb,xspace,enumerate}
\definecolor{LineColor}{rgb}{0.4,0.4,0.4}

%--------------------------------------------------%
\definecolor{shadecolor}{gray}{.9}
\definecolor{darkblue}{rgb}{0.0,0.0,0.3}
\newenvironment{myquote}{%
\bigskip%
\begin{tcolorbox}[]\footnotesize}{%
\end{tcolorbox}%			
\bigskip}

%\headsep 0 true cm
%\textwidth 15.0 true cm
%\textheight 24.0 true cm
%\topmargin 0 true cm
%\oddsidemargin 0 true cm
%\evensidemargin 0 true cm
%\headsep 0 true cm
\usepackage{geometry}

%\setlength{\parindent}{2ex}

\newcommand*{\vhc}{\textsc{vhc}\xspace}
\newcommand*{\vhcs}{\textsc{vhc}s\xspace}
\newcommand*{\vcg}{\textsc{vcg}\xspace}
\newcommand*{\vcgs}{\textsc{vcg}s\xspace}
\newcommand*{\R}{\mathbb{R}}
\newcommand*{\Q}{\mathcal{Q}}
\newcommand*{\X}{\mathcal{X}}
\newcommand*{\specone}{\textsc{spec 1}\xspace}
\newcommand*{\spectwo}{\textsc{spec 2}\xspace}
\newcommand{\NC}{\textsc{nlc}\xspace}
\DeclareMathOperator{\Image}{Im}

\newcommand{\drawline}
{   \hspace*{-\marginparsep minus \marginparwidth}
	\begin{minipage}[t]{\textwidth+\marginparwidth+\marginparsep}%
		{\large \bfseries {} \hfill {}}\\[-0.15\baselineskip]%
		\textcolor{LineColor}{\rule{\columnwidth}{2 pt}}%}
	\end{minipage}}

\newcommand{\myver}{\vspace{0.5cm}}

\newcommand{\mybox}[1]{{\color{blue}\fbox{#1}}}

%\include{Bibtex_DistTeleop}

\title{\vspace{-1cm} {\Large \bf \color{darkblue}{Reply to the Associate
      Editor and Reviewers}\\[0.6cm] \large {\bf From gymnastics to virtual nonholonomic constraints: energy
injection, dissipation, and regulation for the acrobot
      }\\[0.2cm] {(Revision of  IEEE TCST 22-0089)\\ A.~Moran-MacDonald, M.~Maggiore, X.~Wang}}} \author{}
\date{\today}

\begin{document}
\vspace{-4cm}
\maketitle

%\drawline
\pagenumbering{gobble}

\indent We sincerely thank the Associate Editor for collecting two thoughtful reviews and for the 
constructive remarks on this submission. We have revised the paper addressing the reviewers'
comments. Below are detailed replies to the Associate Editor and the reviewers.
\pagenumbering{arabic}
%---------------------------------------------------------------------------%
%\vspace{-1cm}
\subsection*{\textcolor{darkblue}{Reply to the Associate Editor}} 
\vspace{0cm}

\begin{myquote}
We have received only two reviews for this paper, which are - however -
sufficiently informative for a recommendation  to be  issued. Overall,
the reviewers praise the high quality of the paper, but request that a
sharper focus is placed on its contribution with respect not only on
the literature on virtual nonholonomic constraints, but the  landscape
of energy-based nonlinear control. While this may appear a daunting
task, the authors are invited to put their work into perspective as
suggested by reviewer no.2.  The remaining issues appear to be well
within the authors' reach.  
\end{myquote}

In this revision, we have significantly expanded the introduction to include a \emph{detailed} overview of the literature on orbital stabilization mentioned by reviewer 2, contrasting our ideas to existing ones. We have also included a new subsection titled ``Contributions of this article'' in which we better showcase the main results of the paper.  We have also addressed all other issues raised by the two reviewers.

\newpage
%---------------------------------------------------------------%
\vspace{0.2cm}
\subsection*{\textcolor{darkblue}{Reply to Reviewer 2}} 
\vspace{0.2cm}

We wish to sincerely thank the reviewer for the constructive comments,
which we have  addressed in this revision. Below are
detailed replies to the reviewer's remarks.

\begin{myquote}
The problem of designing a feedback control for the two-link acrobot
system through virtual nonholonomic constraints is addressed in this
paper. \((\cdots)\) 
The paper is technically correct and the proofs of Theorem 4 and
Theorem 5 are mathematically rigorous and consistent with the
statements.
\end{myquote}

We thank the reviewer for this assessment.

\begin{myquote}
However, my major concern is that neither the benefit nor
the novelty of considering stabilization by assigning virtual
nonholonomic constraints are clearly articulated. The literature on how
to design nonlinear controllers, possibly passive, for achieving either
swing-up and orbital stabilization or swing-up and asymptotic
stabilization is extensive. In terms of the acrobot, most of the
existing approaches rely on the construction of Casimir functions
(conserved quantities) to help shape the energy of the closed-loop
system via Control-by-Interconnection (CbI) as in [R9], [R8], the
potential energy shaping of the closed-loop system via Interconnection
and Damping Assignment (IDA-PBC) as in [R1], [R2], [R3], [R7], orbital
stabilization via Immersion and Invariance (I\&I) as in [R4], [R5], and
finally orbital stabilization via energy shaping and with the injection
of both pumping and damping as in [R6]. As an application-based
research paper, I would have expected an application of established
research findings and a comparison with other existing approaches. In
addition to showing the effectiveness of implementing virtual
nonholonomic constraints, it would be valuable to discuss the benefit
of the method when compared with the approaches mentioned beforehand. 
\end{myquote}

The reviewer makes a good point, and we completely agree that not placing our paper in a wider context was an omission. This revision \emph{significantly expands} the literature treatment and placed our work in a wider context. The introduction has been significantly expanded with a new subsection titled ``Related literature'' in which we review other techniques for orbital stabilization: I\&I, IDA, VHC-based, and the Fantoni-Lozano approach. We also stress the differences between these approaches and the ideas of this submission.

\begin{myquote}
Furthermore, the following issues should be addressed:

1.	Please define q before the Hamiltonian function given in by the
equations (6). I assume that the generalized coordinate q is composed
of (qu, qa), however, the structure of q is not defined anywhere in the
manuscript.
\end{myquote}

Done, and thank you for raising this point.

\begin{myquote}
 2.	Please enumerate the Hamiltonian function H(q,p) in (6) with
(6a) and the dynamics associated with H(q, p) with (6b). It is quite
uncommon to write "The dynamics of the acrobot in (q,p) coordinates are
given in (6)î. I would suggest, instead, writing îThe dynamics of the
acrobot with Hamiltonian function (6a) are given in (6b)î.  
\end{myquote}

Done, and thank you.

\begin{myquote}
3.	The manuscript deals with the acrobot model, which yields a
Hamiltonian system described by an inhomogeneous differential equation
with a constant input matrix. What is the benefit of introducing the
general Hamiltonian system as in (8) with a nonconstant input matrix B
and then assuming B constant? For ease of presentation, I would suggest
considering directly the inhomogeneous Hamiltonian system used in the
paper.  
\end{myquote}

Done. As suggested by the reviewer, we now directly consider a model with \(B\) constant.

\begin{myquote}
4.	If Lemma 1 and Theorem 1 are recalled from another paper,
please cite the paper which contains the statements and proofs, and
remove the proofs from this paper. Otherwise, if Lemma 1 and Theorem 1
are original results or enhance some other previous findings, then keep
the proofs but emphasise the originality of the results.  
\end{myquote}

We did not take Lemma 1 and Theorem 1 from the literature. However,
the results are relatively obvious and for this reason we do not
showcase them. We clarified this point in a footnote.

\begin{myquote}
5.	Please restate Definition 2. The expression îthe output e =
h(q,p) is of relative degree 2, 2, ..., 2î is unusual. The output is
not of relative degree, rather the system associated with the output e
= h(q,p) has a relative degree. In addition, relative degree 2, 2, ...,
2 is unclear. I think the authors wanted to say that the system has
relative degree 2 for each e\_i.
\end{myquote}

Fixed, thank you. The correct term is \emph{vector relative degree}. We also added a reference to the book by Isidori where this notion is defined.

\begin{myquote}
6.	From Definition 2 results that, if you define the output as e =
h(q,p), where h satisfies the relation h(q,p) = 0, you are looking for
a special characterization of the zero dynamics of your closed-loop
system. This approach reminds me of the construction of the invariant
manifold used to construct a stabilizer through I\&I [R5]. Is there any
relation between the two?  
\end{myquote}

In this revision we have given a detailed comparison between the  I\&I technique and the approach proposed in this paper. Both approaches involve the stabilization of a submanifold of the state space on which one has ``desirable'' behaviour. While in the I\&I literature related to orbital stabilization one assumes that the dynamics on the submanifold are those of an oscillator, in this paper we show that the dynamics on the constraint manifold induced by the proposed VNHC are ``expansive'' or ``contractive'' in that orbits gain energy or lose energy, depending on the sign of a design parameter. In particular the dynamics on our constraint manifold do not contain closed orbits. 
\begin{myquote}
  7.	The output e = h(q,p), which is here assumed relative degree 2,
rules out the most common passive outputs because the relative degree
of the passive output is at most 1. Hence, the virtual nonholonomic
constraints seem to be weak compared to the IDA-PBC approach in [R6].
Please, clarify their relationship. Moreover, the original results in
Section IV.A also seems to have much in common with the
pumping-and-damping technique in [R6].
\end{myquote}

Having relative degree two rather than one does not imply any weakness per se. There is indeed a philosophical  similarity between pumping and damping and the ideas of this submission. This is clarified in the revised introduction. Despite the philosophical similarity, the two techniques are very different and we point this out.

\begin{myquote}
8.	In the sentence before equation (14), what type of stability do
you refer to? 
\end{myquote}

We were referring to asymptotic stability of the constraint manifold. We have modified the language to make this clear.

\begin{myquote}
9.	As I have already mentioned, the subsection îComparison with
existing literatureî only discusses previous work on virtual
nonholonomic constraints and a comparison with other consolidated
methods such as I\&I, CbI, and IDA-PBC is missing. Moreover, a
comparison should be given also in terms of simulations showing the
effectiveness of the virtual nonholonomic constraints when compared
with the existing methods.   
\end{myquote}

We have addressed this point in great detail. We have not however included a simulation comparison because for it to be meaningful, the comparison would require  detailed discussions and an overview of the various techniques  which are beyond the scope of this paper. We hope that the detailed comparison provided in the introduction will at least partially address the reviewer's concern.

\begin{myquote}
10.	Some general comments: the introduction should be improved by
motivating and properly discussing the main results achieved within the
paper; it would be preferable to conclude the introduction with the
organization of the paper; the reading of the article would be easier
if the problem formulation is presented after the preliminary section.  
\end{myquote}

We agree. We have added two new subsections titled ``Contributions of this paper'' and ``Paper organization'' in which we do exactly what the reviewer recommended.

\newpage
%---------------------------------------------------------------------------------%
\vspace{0.2cm}
\subsection*{\textcolor{darkblue}{Reply to Reviewer 4}}
\vspace{0.2cm}

We wish to sincerely thank the reviewer for the constructive comments. Below are
detailed replies to the reviewer's remarks.

\begin{myquote}
My main concern: The topic of virtual (nonlinear) nonholonomic
constraints has its roots in the works of Grizzle and Griffin in CDC
2015. Since then that notion was used and employed in a very
interesting context as the one in the present paper. Nevertheless, a
mathematical foundation of the subject as happens with virtual
holonomic constraints (done by the second author of the paper under
review) was not provided yet. Some effort has been done recently to
formulate virtual nonholonomic constraints from a geometric framework
in the linear (https://arxiv.org/abs/2207.01299) and affine cases
(https://arxiv.org/abs/2301.03890) that I guess they merit to be
mentioned in the literature review of the paper.
\end{myquote}

Thank you very much for flagging this paper. We are now citing it and discussing it both in the introduction and in Section IIIB on virtual nonholonomic constraints.

\begin{myquote}
  My recommendation is based on the next aspects of the paper:

(1)The paper is mathematically correct and self-contained.

(2)The paper is accessible to a wide range of researchers in control
and robotics

(3)The paper demonstrates the theory with real-life experimental.

(4) The experiments are understandable for a wide range of researchers
in control and robotics

(5) The paper is original and the topic is on current research.

I recommend acceptance of the manuscript.
\end{myquote}

We sincerely appreciate the positive assessment of our work.
\end{document}

### 2023-06-27 2.REVISION/main.tex
Revised paper, resubmitted to the IEEE TCST. Also two folders containing literature and the reply to reviewers.

--- 2021-12-08 latex/paper/main.tex
@@
 \usepackage{hyperref}
-%/========== Preamble ==========/%
+\hypersetup{
+    colorlinks,
+    citecolor=red,
+    filecolor=black,
+    linkcolor=blue,
+    urlcolor=black
+  }
+% /========== Preamble ==========/%
 % Packages required by ieeecolor
@@
 \usepackage{enumerate}
+\usepackage{xcolor}
 %---------- More packages ----------%
@@
 
+\definecolor{MM-blue}{HTML}{2883E6}
+\newenvironment{MM}[1]{{\color{MM-blue}#1}}{}
+
 %/========== /Preamble ==========/%
@@
 \author{Adan Moran-MacDonald, \IEEEmembership{Member, IEEE}, Manfredi Maggiore,
-\IEEEmembership{Member, IEEE}*, and Xingbo Wang
+\IEEEmembership{Senior Member, IEEE}*, and Xingbo Wang
 \thanks{Manuscript submitted for review on \today.}
@@
     \begin{subfigure}[t]{0.40\linewidth}
-        \includegraphics[width=\linewidth]{acrobot_gymnast.png}
+        \includegraphics[width=\linewidth]{FIGURES/acrobot_gymnast.png}
     \end{subfigure}
@@
     \begin{subfigure}[t]{0.48\linewidth}
-        \includegraphics[width=\linewidth]{simple_acrobot_model.pdf}
+        \includegraphics[width=\linewidth]{FIGURES/simple_acrobot_model.pdf}
     \end{subfigure}
@@
 methods
-( see
+(see
 \cite{energy_pumping_robotic_swinging,swingup_giant_acrobot,dynamical_servo_acrobot_vc,control_giant_two_link_gymnastic_robot}
@@
 dissipating energy from a system in a robust manner, all while producing
-realistic biological motion.
-
-In this article, we design a virtual nonholonomic constraint which provably
-injects energy into the acrobot through human-like giant motion.
-%In particular, any acrobot constrained by this \vnhc is guaranteed to perform
-%rotations around the bar.
-%Under suitable conditions on the acrobot's physical parameters, this \vnhc
-%also enables the acrobot to rotate around the bar at a desired speed.
-We will provide simulations to show how one can use \vnhcs to regulate energy,
-along with experimental results which demonstrate the robustness of this
-behaviour to model uncertainty, sensor noise, and a variety of external
+realistic biological motion. \MM{The recent paper~\cite{SimStraBloCol23} initiates a systematic geometric investigation of \vnhcs that are affine functions of the configuration velocities.}
+
+\MM{\textbf{Contributions of this article:}}
+In this article, we design a \vnhc which provably
+injects \MM{or dissipates} energy into the acrobot \MM{depending on the sign of a design parameter. The proposed constraint produces a} human-like giant motion.
+\MM{The main result of the paper is presented in Theorems~\ref{thm:acrobot-oscillations} and~\ref{thm:acrobot-rotations}.
+Theorem~\ref{thm:acrobot-oscillations} shows that once the virtual constraint has been enforced, for any nonzero ``low-energy'' initial condition, the proposed controller injects energy in the acrobot making it exit, in finite time, any compact subset of a certain region of the state space where the acrobot oscillates about its pivot point without performing full rotations.
+Theorem~\ref{thm:acrobot-rotations} handles the case when the acrobot is initialized in the complement of the above region, and it shows that under a mild assumption on the physical parameters, the acrobot  is guaranteed to keep increasing its energy while performing full rotations about the pivot point.
+By changing the sign of the design parameter, the results of Theorems~\ref{thm:acrobot-oscillations} and~\ref{thm:acrobot-rotations} hold for energy dissipation.
+By combining these energy injection/damping properties, we then propose an elementary switching logic to attain approximate energy regulation.
+The results of this paper are demonstrated both in simulation and experimentally.
+The experimental results, in particular,} demonstrate the robustness of \MM{the controller against} model uncertainty, sensor noise, and a variety of external
 disturbances.
+
+\MM{\textbf{Related literature:} While the focus of this paper is on energy injection and dissipation, a closely related concept is that of orbital stabilization to a particular energy level set.
+The  problem of orbital stabilization has been investigated extensively and we mention four main approaches.
+
+  One approach, applicable chiefly to underactuated mechanical systems with degree of underactuation one, relies on virtual holonomic constraints to create a target closed orbit and some mechanism to stabilize this orbit.
+In~\cite{ShiPerWit05}, the mechanical system is linearized around the orbit and a linear time-varying controller is designed to stabilize it.
+In~\cite{dynamic_vhcs_stabilize_closed_orbits}, the constraint is embedded in a one-parameter family, and a dynamic controller is designed to adapt the parameter so as to stabilize the target orbit.
+
+  A second approach to orbital stabilization, developed in~\cite{OrtYiRomAst20} and applicable to general control-affine systems, is an adaptation of the immersion and invariance (I\&I) technique of~\cite{AstOrt03}.
+In this approach, one seeks a controlled invariant submanifold of the state space on which the system has desired oscillator dynamics, and designs a feedback controller to stabilize this submanifold.
+The result is a controller that stabilizes \emph{some} closed orbit on the manifold, depending on the initial condition.
+This is in contrast to the approaches in~\cite{ShiPerWit05,dynamic_vhcs_stabilize_closed_orbits} that are able to fix the closed orbit to be stabilized.
+
+  A third approach, developed in~\cite{YiOrtWuZha20}, is an adaptation of the passivity-based interconnection and damping assignment (IDA) technique of~\cite{OrtSpoGomBlan02} and is applicable to underactuated port-Hamiltonian systems.
+In~\cite{YiOrtWuZha20}, the authors assume that the target closed orbit is contained in a two-dimensional plane and present two ways of stabilizing it.
+The first way relies on a modification of energy shaping via IDA in which the component of the shaped energy on the plane has a minimum at the target closed orbit so that its graph resembles a Mexican sombrero.
+The second way represents the closed orbit as an energy level set and assigns the damping matrix in such a way that the controller pumps  energy when the energy is below the target level and damps energy otherwise.
+
+  A final approach, developed in~\cite{FanLoz02}, stabilizes homoclinic orbits of pendulum-like systems using energy methods and Lyapunov-based passivity techniques.
+The stabilization of a homoclinic orbit allows the designer to inject energy in the mechanical system, and it typically yields large domains of attraction.
+
+ We now place the contributions of this paper in the context of the literature just reviewed.
+It is important to bear in mind that the ideas put forth here are limited in scope to the acrobot.
+They are nonetheless notable in that they explore a mechanism to pump or dissipate energy  that is guaranteed to work in a large region of the state space.
+Theorems~\ref{thm:acrobot-oscillations} and~\ref{thm:acrobot-rotations} make this precise.
+In contrast, many the approaches found in the literature and reviewed  above yield only local asymptotic stabilization of a closed orbit.
+When it comes to the acrobot example, the employment of \vnhcs allows one to incorporate knowledge about giant maneuvers in gymnastic into the control design, something that cannot be easily achieved with existing techniques.
+
+ There are similarities between the virtual constraint approach and the I\&I technique for orbital stabilization  proposed in~\cite{OrtYiRomAst20} in that both approaches seeks a controlled invariant submanifold on which the closed-loop system exhibit desired dynamics.
+But while in~\cite{OrtYiRomAst20} one wishes to obtain oscillator dynamics on the submanifold, in this paper we wish to obtain expansive or contractive dynamics (depending on the sign of a design parameter) which rules out oscillator dynamics.
+
+In this paper, like in~\cite{YiOrtWuZha20}, we design a controller for energy pumping/damping, but the energy pumping/dumping mechanism proposed here does not arise from any form of energy shaping.
+In particular, it does not arise from an assignment of the damping matrix of the closed-loop system.
+
+Finally, while in~\cite{FanLoz02} the emphasis is typically the stabilization of a specific energy level set corresponding to the homoclinic orbit of an unstable equilibrium, the technique used in this paper to inject energy does not involve the stabilization of any homoclinic orbit, and as such it does not limit the energy level that can be attained.
+  }
+
+  \MM{\textbf{Paper organization:}
+In Section~\ref{sec:problem-formulation} we review the acrobot model and formulate the problem investigated in this paper.
+Section~\ref{sec:vnhc} presents a class of simply actuated Hamiltonian systems and preliminary notions on \vnhcs.
+In particular, Theorem~\ref{thm:vnhc-regularity} in this section characterizes the constrained dynamics arising from a class of \vnhcs.
+This result is then compared with existing literature on the subject.
+Section~\ref{sec:acrobot} presents the proposed \vnhc for the acrobot and the two main results of the paper, Theorems~\ref{thm:acrobot-oscillations} and~\ref{thm:acrobot-rotations} mentioned earlier.
+Sections~\ref{sec:simulations} and~\ref{sec:experiments} present simulations and experimental results.
+Section~\ref{sec:proof} presents the proof of Theorems~\ref{thm:acrobot-oscillations} and~\ref{thm:acrobot-rotations}, and Section~\ref{sec:conclusion} presents concluding remarks.
+}
 
@@
 The acrobot's configuration is described in generalized coordinates
-\((q_u,q_a)\) on the configuration manifold 
+\MM{\(q=(q_u,q_a)\)} on the configuration manifold 
 \(\mathcal{Q} = \Sone \times \Sone\), where \(q_u\) is unactuated and 
@@
 We ignore dissipative forces in this model.
+
+\MM{In what follows, we denote by \(c_u \eqdef \cos(q_u)\), \(c_a \eqdef \cos(q_a)\), and 
+\(c_{ua} \eqdef \cos(q_u + q_a)\); likewise, \(s_u \eqdef \sin(q_u)\), 
+\(s_a \eqdef \sin(q_a)\), and \(s_{ua} \eqdef \sin(q_u + q_a)\).
+}
 
@@
     M(q) &= \begin{bmatrix}
-        ml^2\left(3+2\cos(q_a)\right) & 
-        ml^2\left(1+\cos(q_a)\right) \\
-        ml^2\left(1+\cos(q_a)\right) &
+        ml^2\left(3+2 c_a\right) & 
+        ml^2\left(1+ c_a \right) \\
+        ml^2\left(1+ c_a\right) &
         ml^2
@@
     \label{eqn:acrobot-potential}
-    V(q) &= -mgl\left(2\cos(q_u)+\cos(q_u+q_a)\right)
+    V(q) &= -mgl\left(2 c_u + c_{ua}\right)
     , \\
@@
 use Hamiltonian mechanics to derive the dynamics of the acrobot. 
-For this we require the conjugate of momenta, \(p = (p_u,p_a) = M(q)\dot{q}\).
-The dynamics of the acrobot in \((q,p)\) coordinates are given in
-\eqref{eqn:acrobot-hamiltonian}.
-For shorthand, we write \(c_u \eqdef \cos(q_u)\), \(c_a \eqdef \cos(q_a)\), and 
-\(c_{ua} \eqdef \cos(q_u + q_a)\); likewise, \(s_u \eqdef \sin(q_u)\), 
-\(s_a \eqdef \sin(q_a)\), and \(s_{ua} \eqdef \sin(q_u + q_a)\).
-\begin{align}\label{eqn:acrobot-hamiltonian}
-    \mathcal{H}(q,p) &= \frac{1}{2}p\tpose \Minv(q) p -
-    mgl\left(2 c_u + c_{ua}\right)
-    , \\
-     &\begin{cases}
-        \dot{q} = \Minv(q) p 
-        ,\\
-        \dot{p}_u = -mgl\left(2s_u + s_{ua}\right) 
-        ,\\
-        \dot{p}_a =-\frac{1}{2}p\tpose \nabla_{q_a}\Minv(q) p
-        - mgl s_{ua} + \tau.
-    \end{cases} \nonumber
-\end{align}
-The control input is a force \(\tau \in \R\) affecting only the dynamics of
-\(p_a\), representing a torque acting on the hip joint.
+\MM{To this end, recalling the partitioning  \(q= (q_u,q_a)\), we define the conjugate of momenta, \(p = (p_u,p_a) = M(q)\dot{q}\) and the Hamiltonian function
+\begin{equation}\label{eqn:acrobot-hamiltonian}
+    \mathcal{H}(q,p) = \frac{1}{2}p\tpose \Minv(q) p -
+   mgl\left(2 c_u + c_{ua}\right).
+  \end{equation}
+The dynamics of the
+acrobot with Hamiltonian function~\eqref{eqn:acrobot-hamiltonian} are
+\begin{equation}
+  \label{eq:acrobot:model}
+  \begin{aligned}
+&    \dot{q} = \Minv(q) p,\\
+&    \dot{p}_u = -mgl\left(2s_u + s_{ua}\right),\\
+&     \dot{p}_a =-\frac{1}{2}p\tpose \nabla_{q_a}\Minv(q) p - m g l s_{ua} + \tau,
+  \end{aligned}
+\end{equation}
+where \(\tau\), the control input, is the hip joint torque affecting only the dynamics of
+\(p_a\).}
 
@@
 Before embarking on the design problem, we must summarize the relevant
-theory of virtual nonholonomic constraints for a class of mechanical systems we
+theory of \vnhcs for a class of mechanical systems we
 call ``simply actuated Hamiltonian systems". 
-The results we provide in Section \ref{sec:vnhc-vnhc} are not novel:
-they are a special case of the results in
+\MM{Although the results in Section \ref{sec:vnhc-vnhc} do not appear in the literature in the same form, we will show that} they are a special case of results that appeared in
 \cite{hybrid_zero_dynamics_bipedal_nhvcs}.
@@
 \subsection{Simply Actuated Hamiltonian Systems} \label{sec:vnhc-sah}
-Take a mechanical system modelled with generalized coordinates 
+\MM{Consider} a mechanical system modelled with generalized coordinates 
 \(q = (q_1, \ldots, q_n)\) on a configuration manifold
@@
         , \\
-        \dot{p} = -\nabla_q\mathcal{H} + B(q) \tau
+        \dot{p} = -\nabla_q\mathcal{H} + \MM{B} \, \tau
         ,
@@
 where \(\tau \in \R^k\) is a vector of generalized input forces and the input
-matrix \(B : \mathcal{Q} \rightarrow \R^{n \times k}\) is full rank for all 
-\(q \in \mathcal{Q}\).
+matrix \MM{\(B \in \R^{n \times k}\) is assumed to be constant and of full rank \(k\).}
 If \(k < n\), we say the system is \textit{underactuated} with degree of
@@
         \dot{p} = -\frac{1}{2} (\Id{n} \otimes p\tpose) \nabla_q D\inv(q) p
-        - \nabla_q P(q) + B(q) \tau
+        - \nabla_q P(q) + \MM{B} \,\tau
         . \\
@@
 
-Because \(\tau\) is transformed by \(B(q)\), it is not obvious how any
+Because \(\tau\) is transformed by \(\MM{B}\), it is not obvious how any
 particular input force \(\tau_i\) affects the system.
-As a first step in addressing this problem, we make the following assumptions.
-
-\begin{assm}\label{assm:B-const}
-    The input matrix \(B(q) \equiv B \in \R^{n\times k}\) is constant
-    and has full rank \(k < n\).
-\end{assm}
-
-%\begin{assm}\label{assm:B-perp}
-%    There exists a right semi-orthogonal matrix 
-%    \(B^\perp \in \R^{(n-k)\times n}\)
-%    which is a left-annihilator for \(B\). 
-%\end{assm}
-%
-%Note that Assumption \ref{assm:B-perp} requires the rows of \(B^\perp\) be unit vectors
-%that are mutually orthogonal. 
-%When \(k = (n-1)\), Assumption \ref{assm:B-perp} can be removed because it is
-%automatically implied by Assumption \ref{assm:B-const}.
-%
-When \(\mathcal{Q} = \R^n\), the above assumption allows us to define a
+% As a first step in addressing this problem, we make the following assumptions.
+
+% \begin{assm}\label{assm:B-const}
+%     The input matrix \(B(q) \equiv B \in \R^{n\times k}\) is constant
+%     and has full rank \(k < n\).
+% \end{assm}
+%
+When \(\mathcal{Q} = \R^n\), \MM{one may} define a
 canonical coordinate transformation of \eqref{eqn:hamiltonian} 
 which decouples the input forces.
-To define this transformation we will make use of the following lemma.
+To define this transformation we will make use of the following lemma\footnote{\MM{Lemma~\ref{lemma:B-orthogonal} and Theorem~\ref{thm:simply-actuated} are obvious, but we include their proofs for completeness.}}.
 
 \begin{lemma}\label{lemma:B-orthogonal}
-    Suppose Assumption \ref{assm:B-const} holds. Then
-    there exists a nonsingular matrix \(\hat{T} \in \R^{k \times k}\) 
+There exists a nonsingular matrix \(\hat{T} \in \R^{k \times k}\) 
     so that the regular feedback transformation 
@@
 \begin{thm}\label{thm:simply-actuated}
-    Take the Hamiltonian system \eqref{eqn:hamiltonian} with configuration
-    manifold \(\mathcal{Q} = \R^n\) and suppose
-    Assumption \ref{assm:B-const} holds.
-    The coordinate transformation
-    \(\left(\tilde{q} = (\mathbf{B}\tpose)\inv q, \tilde{p} = \mathbf{B}p\right)\)
-    is a canonical transformation and the resulting dynamics are given by 
-    \begin{gather}\label{eqn:simple-hamiltonian}
+\MM{For} the Hamiltonian system \eqref{eqn:hamiltonian} with configuration
+    manifold \(\mathcal{Q} = \R^n\)\MM{,
+    the} coordinate transformation
+    \((\tilde q,\tilde p) = \left(\mathbf{B}\tpose)\inv q,\mathbf{B}p\right)\)
+    is a canonical transformation. \MM{ the} resulting \MM{Hamiltonian function is 
+    \begin{equation}
+    \label{eqn:simple-hamiltonian}
         \mathcal{H}(\tilde{q},\tilde{p}) = 
-        \frac{1}{2} \tilde{p}\tpose \Minv(\tilde{q}) \tilde{p} + V(\tilde{q})
-        , \\
-       \begin{cases}
-           \dot{\tilde{q}} = \Minv(\tilde{q})\tilde{p}
-           , \\
-           \dot{\tilde{p}} = -\frac{1}{2} (\Id{n} \otimes \tilde{p}\tpose)
+        \frac{1}{2} \tilde{p}\tpose \Minv(\tilde{q}) \tilde{p} + V(\tilde{q}),
+      \end{equation}
+      and the associated Hamiltonian dynamics are
+      \begin{equation}
+        \label{eq:Hamiltonian_system:simply_actuated}
+          \begin{aligned}
+            &           \dot{\tilde{q}} = \Minv(\tilde{q})\tilde{p}, \\
+            &           \dot{\tilde{p}} = -\frac{1}{2} (\Id{n} \otimes \tilde{p}\tpose)
               \nabla_{\tilde{q}} \Minv(\tilde{q}) \tilde{p} \\
-           \phantom{---} - \nabla_{\tilde{q}} V(\tilde{q}) + \simpleB \tau
-            ,
-        \end{cases} \nonumber
-    \end{gather}
+            &          \phantom{---} - \nabla_{\tilde{q}} V(\tilde{q}) + \simpleB \tau,
+          \end{aligned}
+  \end{equation}}
   where 
@@
 \textit{simply actuated coordinates}, and we call any Hamiltonian system
-whose input matrix is \([\Zmat{(n-k)\times k}; \Id{k}]\) a 
+\MM{of the form~\eqref{eq:Hamiltonian_system:simply_actuated}} a 
 \textit{simply actuated Hamiltonian system}.
@@
 Hamed and Ames \cite{nonholonomic_hybrid_zero_dynamics} then found an explicit
-representation of these constrained dynamics for virtual
-nonholonomic constraints with a specific form.
+representation of these constrained dynamics for \vnhcs with a specific form.
 All these researchers made use of the
@@
 Lagrangian framework. In particular, they focused on Lagrangian systems with
-degree of underactuation one. We will now present a special case of
+degree of underactuation one.
+\MM{The recent paper~\cite{SimStraBloCol23} has initiated a systematic geometric investigation of \vnhcs that are affine functions of the configuration velocities.
+The ones used in this paper are nonlinear functions of the configuration velocities.}
+
+We will now present a special case of
 \cite{hybrid_zero_dynamics_bipedal_nhvcs} for simply actuated Hamiltonian
@@
 \begin{defn}
-    A \vnhc \(h(q,p) = 0\) of order \(k\) is \textit{regular} if the output 
-    \(e = h(q,p)\) is of relative degree \(\{2,2.\ldots,2\}\) everywhere on the
+    A \vnhc \(h(q,p) = 0\) of order \(k\) is \textit{regular} if \MM{system~\eqref{eq:Hamiltonian_system:simply_actuated}} with output 
+    \(e = h(q,p)\)  \MM{has vector} relative degree \(\{2,2.\ldots,2\}\) everywhere on the
     constraint manifold \(\Gamma\).
@@
 
+\MM{The reader is referred to~\cite{Isi95} for the definition of vector relative degree.}
 The authors of
@@
 
-Under additional mild conditions (see \cite{vhcs_for_el_systems}), a regular \vnhc of
-order \(k\) can be stabilized by the input-output feedback linearizing
+Under additional mild conditions (see \cite{vhcs_for_el_systems}), \MM{the constraint manifold associated with } a regular \vnhc of
+order \(k\) can be \MM{asymptotically} stabilized by the input-output feedback linearizing
 controller
@@
 
-\textbf{Comparison with existing literature}: Horn \etal provide the constrained
+\textbf{Comparison with existing \MM{virtual constraint} literature}: Horn \etal provide the constrained
 dynamics for \vnhcs in \cite{nhvc_incline_walking}.
@@
     \centering
-    \includegraphics[width=0.7\linewidth]{qa_arctan.pdf}
+    \includegraphics[width=0.7\linewidth]{FIGURES/qa_arctan.pdf}
     \caption{The acrobot constraint \(q_a = \bar{q}_a \arctan(I p_u)\).}
@@
     \begin{subfigure}[t]{0.45\linewidth}
-        \includegraphics[width=\linewidth]{acrobot_EleqR.pdf}
+        \includegraphics[width=\linewidth]{FIGURES/acrobot_EleqR.pdf}
         \caption{The set $E_{\leq R}$ with $0 < R < \bar R$ and an oscillation gaining energy.} 
@@
     \begin{subfigure}[t]{0.45\linewidth}
-        \includegraphics[width=\linewidth]{acrobot_ER1R2.pdf}
+        \includegraphics[width=\linewidth]{FIGURES/acrobot_ER1R2.pdf}
         \caption{The set $E_{[R_1,R_2]}$, with $\bar R < R_1 < R_2$, and a rotation gaining energy.}
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{acrobot_model.pdf}
+    \includegraphics[width=0.8\linewidth]{FIGURES/acrobot_model.pdf}
     \caption{The distributed mass acrobot model, represented by two weighted rods
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{acrobot_in_orbit.png}
+    \includegraphics[width=0.8\linewidth]{FIGURES/acrobot_in_orbit.png}
     \caption{A simulation of the acrobot gaining energy.}
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{acrobot_mc.png}
+    \includegraphics[width=0.8\linewidth]{FIGURES/acrobot_mc.png}
     \caption{Monte Carlo simulation for energy injection.}
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{acrobot_diss_orbit.png}
+    \includegraphics[width=0.8\linewidth]{FIGURES/acrobot_diss_orbit.png}
     \caption{A simulation of the acrobot dissipating energy.}
@@
     \begin{subfigure}[t]{0.49\linewidth}
-        \includegraphics[width=\linewidth]{acrobot_osc_orbit_1.png}
+        \includegraphics[width=\linewidth]{FIGURES/acrobot_osc_orbit_1.png}
         \caption{Stabilizing an oscillation from below.}
@@
     \begin{subfigure}[t]{0.49\linewidth}
-        \includegraphics[width=\linewidth]{acrobot_osc_orbit_2.png}
+        \includegraphics[width=\linewidth]{FIGURES/acrobot_osc_orbit_2.png}
         \caption{Stabilizing an oscillation from above.}
@@
     \begin{subfigure}[t]{0.49\linewidth}
-        \includegraphics[width=\linewidth]{acrobot_rot_orbit_1.png}
+        \includegraphics[width=\linewidth]{FIGURES/acrobot_rot_orbit_1.png}
         \caption{Stabilizing a rotation from below.}
@@
     \begin{subfigure}[t]{0.49\linewidth}
-        \includegraphics[width=\linewidth]{acrobot_rot_orbit_2.png}
+        \includegraphics[width=\linewidth]{FIGURES/acrobot_rot_orbit_2.png}
         \caption{Stabilizing a rotation from above.}
@@
     \centering
-    \includegraphics[width=0.9\linewidth]{xingbo_acrobot.jpg}
+    \includegraphics[width=0.9\linewidth]{FIGURES/xingbo_acrobot.jpg}
     \caption{SUGAR is the physical acrobot built by Wang \cite{xingbo_thesis}.}
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{acrobot_unperturbed_orbit.png}
+    \includegraphics[width=0.8\linewidth]{FIGURES/acrobot_unperturbed_orbit.png}
     \caption{Baseline Test: SUGAR's baseline energy injection orbit. A video of this experiment is available at 
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{acrobot_stopped_orbit.png}
+    \includegraphics[width=0.8\linewidth]{FIGURES/acrobot_stopped_orbit.png}
     \caption{Perturbation Test 1: SUGAR's orbit before (blue) and after (red)
@@
     \begin{subfigure}[ht]{0.49\linewidth}
-        \includegraphics[width=\linewidth]{acrobot_fpush_orbit.png}
+        \includegraphics[width=\linewidth]{FIGURES/acrobot_fpush_orbit.png}
         \caption{The forward push test.}
@@
     \begin{subfigure}[ht]{0.49\linewidth}
-        \includegraphics[width=\linewidth]{acrobot_rpush_orbit.png}
+        \includegraphics[width=\linewidth]{FIGURES/acrobot_rpush_orbit.png}
         \caption{The reverse push test.}
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{Qro.png}
+    \includegraphics[width=0.8\linewidth]{FIGURES/Qro.png}
     \caption{The graph of \(r \mapsto \int_0^{2\pi} a(r,\theta) d\theta\).}

### 2023-07-07 3.FINAL_VERSION/BIOS/maggiore.tex
Final, accepted version sent to the journal

Manfredi Maggiore received the "Laurea" degree in Electronic
Engineering in 1996 from the University of Genoa and the PhD degree in
Electrical Engineering from the Ohio State University, USA, in
2000. Since 2000 he has been with the Edward S. Rogers Sr. Department
of Electrical and Computer Engineering, University of Toronto, Canada,
where he is currently Professor. He has been a visiting Professor at
the University of Bologna (2007-2008), and the Laboratoire des Signaux
et Syst\`emes, Ecole CentraleSup\'elec (2015-2016). His research
focuses on mathematical nonlinear control theory, and relies on
methods from dynamical systems theory and differential geometry.

### 2023-07-07 3.FINAL_VERSION/BIOS/moran-macdonald.tex
Final, accepted version sent to the journal

Adan Moran-MacDonald graduated from Queen's University, Canada, in 2018,
with a BASc. in Mathematics Engineering and a subspecialty in Systems and Robotics.
He then attained an MASc. in Electrical and Computer Engineering under the 
tutelage of Manfredi Maggiore at the University of Toronto, Canada, in 2021, 
with a focus on Systems Control.
Adan's academic research interests lie in the field of geometric nonlinear control
theory and its applications to biologically-inspired robotics.
He is now an Automation Innovator at Artificial Futures in Toronto, Canada,
where he specializes in research and development for human-centric AI, robotics,
and automation applications.

### 2023-07-07 3.FINAL_VERSION/main.tex
Final, accepted version sent to the journal

--- 2023-06-27 2.REVISION/main.tex
@@
 % TODO: remove onecolumn and draftcls formatting
-\documentclass[journal,twoside,web, twocolumn,draftcls]{ieeecolor}
+\documentclass[journal,twoside,web, twocolumn]{ieeecolor}
 
 \usepackage{generic}
-\usepackage{hyperref}
+\usepackage[hypertexnames=false]{hyperref}
 \hypersetup{
@@
 \usepackage{enumerate}
-\usepackage{xcolor}
+%\usepackage{xcolor}
 %---------- More packages ----------%
@@
 
-\definecolor{MM-blue}{HTML}{2883E6}
-\newenvironment{MM}[1]{{\color{MM-blue}#1}}{}
-
 %/========== /Preamble ==========/%
@@
 \author{Adan Moran-MacDonald, \IEEEmembership{Member, IEEE}, Manfredi Maggiore,
-\IEEEmembership{Senior Member, IEEE}*, and Xingbo Wang
-\thanks{Manuscript submitted for review on \today.}
+\IEEEmembership{Senior Member, IEEE}, and Xingbo Wang
 \thanks{A. Moran-MacDonald (e-mail: adan.moran@mail.utoronto.ca),
@@
 X. Wang (e-mail: xingbo.wang@utoronto.ca) are with the Department of
-Electrical and Computer Engineering, University of Toronto, ON, Canada.}
+Electrical and Computer Engineering, University of Toronto, ON, Canada. This research was supported by the Natural Sciences and Engineering Research Council of Canada (NSERC).}
 } %/author
@@
 dissipating energy from a system in a robust manner, all while producing
-realistic biological motion. \MM{The recent paper~\cite{SimStraBloCol23} initiates a systematic geometric investigation of \vnhcs that are affine functions of the configuration velocities.}
-
-\MM{\textbf{Contributions of this article:}}
+realistic biological motion. The recent paper~\cite{SimStraBloCol23} initiates a systematic geometric investigation of \vnhcs that are affine functions of the configuration velocities.
+
+\textbf{Contributions of this article:}
 In this article, we design a \vnhc which provably
-injects \MM{or dissipates} energy into the acrobot \MM{depending on the sign of a design parameter. The proposed constraint produces a} human-like giant motion.
-\MM{The main result of the paper is presented in Theorems~\ref{thm:acrobot-oscillations} and~\ref{thm:acrobot-rotations}.
+injects or dissipates energy into the acrobot depending on the sign of a design parameter. The proposed constraint produces a human-like giant motion.
+The main result of the paper is presented in Theorems~\ref{thm:acrobot-oscillations} and~\ref{thm:acrobot-rotations}.
 Theorem~\ref{thm:acrobot-oscillations} shows that once the virtual constraint has been enforced, for any nonzero ``low-energy'' initial condition, the proposed controller injects energy in the acrobot making it exit, in finite time, any compact subset of a certain region of the state space where the acrobot oscillates about its pivot point without performing full rotations.
@@
 The results of this paper are demonstrated both in simulation and experimentally.
-The experimental results, in particular,} demonstrate the robustness of \MM{the controller against} model uncertainty, sensor noise, and a variety of external
+The experimental results, in particular, demonstrate the robustness of the controller against model uncertainty, sensor noise, and a variety of external
 disturbances.
 
-\MM{\textbf{Related literature:} While the focus of this paper is on energy injection and dissipation, a closely related concept is that of orbital stabilization to a particular energy level set.
+\textbf{Related literature:} While the focus of this paper is on energy injection and dissipation, a closely related concept is that of orbital stabilization to a particular energy level set.
 The  problem of orbital stabilization has been investigated extensively and we mention four main approaches.
@@
 
- There are similarities between the virtual constraint approach and the I\&I technique for orbital stabilization  proposed in~\cite{OrtYiRomAst20} in that both approaches seeks a controlled invariant submanifold on which the closed-loop system exhibit desired dynamics.
+ There are similarities between the virtual constraint approach and the I\&I technique for orbital stabilization  proposed in~\cite{OrtYiRomAst20} in that both approaches seek a controlled invariant submanifold on which the closed-loop system exhibits desired dynamics.
 But while in~\cite{OrtYiRomAst20} one wishes to obtain oscillator dynamics on the submanifold, in this paper we wish to obtain expansive or contractive dynamics (depending on the sign of a design parameter) which rules out oscillator dynamics.
 
-In this paper, like in~\cite{YiOrtWuZha20}, we design a controller for energy pumping/damping, but the energy pumping/dumping mechanism proposed here does not arise from any form of energy shaping.
+In this paper, like in~\cite{YiOrtWuZha20}, we design a controller for energy pumping/damping, but the energy pumping/damping mechanism proposed here does not arise from any form of energy shaping.
 In particular, it does not arise from an assignment of the damping matrix of the closed-loop system.
@@
 Finally, while in~\cite{FanLoz02} the emphasis is typically the stabilization of a specific energy level set corresponding to the homoclinic orbit of an unstable equilibrium, the technique used in this paper to inject energy does not involve the stabilization of any homoclinic orbit, and as such it does not limit the energy level that can be attained.
-  }
-
-  \MM{\textbf{Paper organization:}
+
+\textbf{Paper organization:}
 In Section~\ref{sec:problem-formulation} we review the acrobot model and formulate the problem investigated in this paper.
@@
 Section~\ref{sec:proof} presents the proof of Theorems~\ref{thm:acrobot-oscillations} and~\ref{thm:acrobot-rotations}, and Section~\ref{sec:conclusion} presents concluding remarks.
-}
 
@@
 \section{Problem Formulation}\label{sec:problem-formulation}
-We will use the simplified acrobot model in Figure
-\ref{fig:acrobot}, where we assume the torso and leg rods are of
-equal length \(l\) with equal point masses \(m\) at the tips.
-The acrobot's configuration is described in generalized coordinates
-\MM{\(q=(q_u,q_a)\)} on the configuration manifold 
-\(\mathcal{Q} = \Sone \times \Sone\), where \(q_u\) is unactuated and 
-\(q_a\) is actuated.
+We will use the simplified acrobot model in Figure \ref{fig:acrobot}, where we assume the torso and leg rods are of equal length \(l\) with equal point masses \(m\) at the tips.
+The acrobot's configuration is described in generalized coordinates \(q=(q_u,q_a)\) on the configuration manifold \(\mathcal{Q} = \Sone \times \Sone\), where \(q_u\) is unactuated and  \(q_a\) is actuated.
 We ignore dissipative forces in this model.
 
-\MM{In what follows, we denote by \(c_u \eqdef \cos(q_u)\), \(c_a \eqdef \cos(q_a)\), and 
-\(c_{ua} \eqdef \cos(q_u + q_a)\); likewise, \(s_u \eqdef \sin(q_u)\), 
-\(s_a \eqdef \sin(q_a)\), and \(s_{ua} \eqdef \sin(q_u + q_a)\).
-}
-
-The acrobot has inertia matrix 
-\(M\), potential function \(V\) (with respect to
-the horizontal bar), and input matrix \(B\) given as follows:
+In what follows, we denote by \(c_u \eqdef \cos(q_u)\), \(c_a \eqdef \cos(q_a)\), and \(c_{ua} \eqdef \cos(q_u + q_a)\); likewise, \(s_u \eqdef \sin(q_u)\), \(s_a \eqdef \sin(q_a)\), and \(s_{ua} \eqdef \sin(q_u + q_a)\).
+
+The acrobot has inertia matrix \(M\), potential function \(V\) (with respect to the horizontal bar), and input matrix \(B\) given as follows:
 \begin{align}\label{eqn:acrobot-inertia}
@@
 use Hamiltonian mechanics to derive the dynamics of the acrobot. 
-\MM{To this end, recalling the partitioning  \(q= (q_u,q_a)\), we define the conjugate of momenta, \(p = (p_u,p_a) = M(q)\dot{q}\) and the Hamiltonian function
+To this end, recalling the partitioning  \(q= (q_u,q_a)\), we define the conjugate of momenta, \(p = (p_u,p_a) = M(q)\dot{q}\) and the Hamiltonian function
 \begin{equation}\label{eqn:acrobot-hamiltonian}
@@
 where \(\tau\), the control input, is the hip joint torque affecting only the dynamics of
-\(p_a\).}
+\(p_a\).
 
@@
 \section{Preliminaries on \vnhcs}\label{sec:vnhc}
-Before embarking on the design problem, we must summarize the relevant
-theory of \vnhcs for a class of mechanical systems we
-call ``simply actuated Hamiltonian systems". 
-\MM{Although the results in Section \ref{sec:vnhc-vnhc} do not appear in the literature in the same form, we will show that} they are a special case of results that appeared in
-\cite{hybrid_zero_dynamics_bipedal_nhvcs}.
+Before embarking on the design problem, we must summarize the relevant theory of \vnhcs for a class of mechanical systems we call ``simply actuated Hamiltonian systems".
+Although the results in Section \ref{sec:vnhc-vnhc} do not appear in the literature in the same form, we will show that they are a special case of results that appeared in \cite{hybrid_zero_dynamics_bipedal_nhvcs}.
 
 \subsection{Simply Actuated Hamiltonian Systems} \label{sec:vnhc-sah}
-\MM{Consider} a mechanical system modelled with generalized coordinates 
-\(q = (q_1, \ldots, q_n)\) on a configuration manifold
-\(\mathcal{Q} = \Rt{T_1} \times \cdots \times \Rt{T_n}\), where
-\(T_i = 2\pi\) if \(q_i\) is an angle and \(T_i = \infty\) if \(q_i\) is a
-displacement. The corresponding generalized velocities are 
-\(\dot{q} = (\dot{q}_1,\ldots,\dot{q}_n) \in \R^n\).
-
-Suppose this system has Lagrangian
-\(\mathcal{L}(q,\dot{q}) = 1/2~\dot{q}^T D(q) \dot{q} - P(q)\),
-where the potential energy 
-\(P : \mathcal{Q} \rightarrow \mathbb{R}\) 
-is smooth, and the inertia matrix 
-\(D : \mathcal{Q} \rightarrow \mathbb{R}^{n \times n}\)
-is smooth and positive definite for all \(q \in \mathcal{Q}\).
-The \textit{conjugate momentum} to \(q\) is the vector
-\(p \eqdef \partial\mathcal{L}/\partial\dot{q} = D(q) \dot{q} \in \R^n\).
-As per \cite{landau_mechanics}, 
-the \textit{Hamiltonian} of the system in \((q,p)\) coordinates
-is
+Consider a mechanical system modelled with generalized coordinates  \(q = (q_1, \ldots, q_n)\) on a configuration manifold \(\mathcal{Q} = \Rt{T_1} \times \cdots \times \Rt{T_n}\), where \(T_i = 2\pi\) if \(q_i\) is an angle and \(T_i = \infty\) if \(q_i\) is a displacement.
+The corresponding generalized velocities are \(\dot{q} = (\dot{q}_1,\ldots,\dot{q}_n) \in \R^n\).
+
+Suppose this system has Lagrangian \(\mathcal{L}(q,\dot{q}) = 1/2~\dot{q}^T D(q) \dot{q} - P(q)\), where the potential energy  \(P : \mathcal{Q} \rightarrow \mathbb{R}\)  is smooth, and the inertia matrix  \(D : \mathcal{Q} \rightarrow \mathbb{R}^{n \times n}\) is smooth and positive definite for all \(q \in \mathcal{Q}\).
+The \textit{conjugate momentum} to \(q\) is the vector \(p \eqdef \partial\mathcal{L}/\partial\dot{q} = D(q) \dot{q} \in \R^n\).
+As per \cite{landau_mechanics},  the \textit{Hamiltonian} of the system in \((q,p)\) coordinates is
 \begin{equation}\label{eqn:hamiltonian}
@@
         , \\
-        \dot{p} = -\nabla_q\mathcal{H} + \MM{B} \, \tau
+        \dot{p} = -\nabla_q\mathcal{H} + B \, \tau
         ,
@@
 \end{equation}
-where \(\tau \in \R^k\) is a vector of generalized input forces and the input
-matrix \MM{\(B \in \R^{n \times k}\) is assumed to be constant and of full rank \(k\).}
-If \(k < n\), we say the system is \textit{underactuated} with degree of
-underactuation \((n-k)\).
+where \(\tau \in \R^k\) is a vector of generalized input forces and the input matrix \(B \in \R^{n \times k}\) is assumed to be constant and of full rank \(k\).
+If \(k < n\), we say the system is \textit{underactuated} with degree of underactuation \((n-k)\).
 
@@
         \dot{p} = -\frac{1}{2} (\Id{n} \otimes p\tpose) \nabla_q D\inv(q) p
-        - \nabla_q P(q) + \MM{B} \,\tau
+        - \nabla_q P(q) + B \,\tau
         . \\
@@
 
-Because \(\tau\) is transformed by \(\MM{B}\), it is not obvious how any
-particular input force \(\tau_i\) affects the system.
-% As a first step in addressing this problem, we make the following assumptions.
-
-% \begin{assm}\label{assm:B-const}
-%     The input matrix \(B(q) \equiv B \in \R^{n\times k}\) is constant
-%     and has full rank \(k < n\).
-% \end{assm}
-%
-When \(\mathcal{Q} = \R^n\), \MM{one may} define a
-canonical coordinate transformation of \eqref{eqn:hamiltonian} 
-which decouples the input forces.
-To define this transformation we will make use of the following lemma\footnote{\MM{Lemma~\ref{lemma:B-orthogonal} and Theorem~\ref{thm:simply-actuated} are obvious, but we include their proofs for completeness.}}.
+Because \(\tau\) is transformed by \(B\), it is not obvious how any particular input force \(\tau_i\) affects the system.
+
+When \(\mathcal{Q} = \R^n\), one may define a canonical coordinate transformation of \eqref{eqn:hamiltonian} which decouples the input forces.
+To define this transformation we will make use of the following lemma\footnote{Lemma~\ref{lemma:B-orthogonal} and Theorem~\ref{thm:simply-actuated} are obvious, but we include their proofs for completeness.}.
 
@@
 \begin{thm}\label{thm:simply-actuated}
-\MM{For} the Hamiltonian system \eqref{eqn:hamiltonian} with configuration
-    manifold \(\mathcal{Q} = \R^n\)\MM{,
-    the} coordinate transformation
-    \((\tilde q,\tilde p) = \left(\mathbf{B}\tpose)\inv q,\mathbf{B}p\right)\)
-    is a canonical transformation. \MM{ the} resulting \MM{Hamiltonian function is 
+For the Hamiltonian system \eqref{eqn:hamiltonian} with configuration manifold \(\mathcal{Q} = \R^n\), the coordinate transformation \((\tilde q,\tilde p) = \left(\mathbf{B}\tpose)\inv q,\mathbf{B}p\right)\) is a canonical transformation.
+The resulting Hamiltonian function is
     \begin{equation}
@@
           \end{aligned}
-  \end{equation}}
+  \end{equation}
   where 
@@
 
-We call the \((\tilde{q},\tilde{p})\) coordinates
-\textit{simply actuated coordinates}, and we call any Hamiltonian system
-\MM{of the form~\eqref{eq:Hamiltonian_system:simply_actuated}} a 
-\textit{simply actuated Hamiltonian system}.
-The first \((n-k)\) configuration variables in \(\tilde{q}\), labelled \(q_u\),
-are the \textit{unactuated coordinates}; 
-the remaining \(k\) configuration variables, labelled \(q_a\), are the
-\textit{actuated coordinates}.
-The corresponding \((p_u, p_a)\) in \(\tilde{p}\) are the \textit{unactuated}
-and \textit{actuated momenta}, respectively.
+We call the \((\tilde{q},\tilde{p})\) coordinates \textit{simply actuated coordinates}, and we call any Hamiltonian system of the form~\eqref{eq:Hamiltonian_system:simply_actuated} a  \textit{simply actuated Hamiltonian system}.
+The first \((n-k)\) configuration variables in \(\tilde{q}\), labelled \(q_u\), are the \textit{unactuated coordinates};  the remaining \(k\) configuration variables, labelled \(q_a\), are the \textit{actuated coordinates}.
+The corresponding \((p_u, p_a)\) in \(\tilde{p}\) are the \textit{unactuated} and \textit{actuated momenta}, respectively.
 
@@
 
-Griffin and Grizzle \cite{nhvc_dynamic_walking} were the first to define
-relative degree two nonholonomic constraints which can be enforced through state
-feedback. Horn \etal extended their results in
-\cite{hybrid_zero_dynamics_bipedal_nhvcs} to derive the constrained dynamics for
-a certain class of mechanical systems. 
-Hamed and Ames \cite{nonholonomic_hybrid_zero_dynamics} then found an explicit
-representation of these constrained dynamics for \vnhcs with a specific form.
-All these researchers made use of the
-unactuated conjugate momentum, but they developed their results in the
-Lagrangian framework. In particular, they focused on Lagrangian systems with
-degree of underactuation one.
-\MM{The recent paper~\cite{SimStraBloCol23} has initiated a systematic geometric investigation of \vnhcs that are affine functions of the configuration velocities.
-The ones used in this paper are nonlinear functions of the configuration velocities.}
+Griffin and Grizzle \cite{nhvc_dynamic_walking} were the first to define relative degree two nonholonomic constraints which can be enforced through state feedback.
+Horn \etal extended their results in \cite{hybrid_zero_dynamics_bipedal_nhvcs} to derive the constrained dynamics for a certain class of mechanical systems.
+Hamed and Ames \cite{nonholonomic_hybrid_zero_dynamics} then found an explicit representation of these constrained dynamics for \vnhcs with a specific form.
+All these researchers made use of the unactuated conjugate momentum, but they developed their results in the Lagrangian framework.
+In particular, they focused on Lagrangian systems with degree of underactuation one.
+The recent paper~\cite{SimStraBloCol23} has initiated a systematic geometric investigation of \vnhcs that are affine functions of the configuration velocities.
+The ones used in this paper are nonlinear functions of the configuration velocities.
 
@@
 \begin{equation}
-        \Gamma = \left\{(q,p) \mid h(q,p) = 0, dh_q \dot{q} + dh_p \dot{p} = 0\right\}
-        .
-    \end{equation}
+        \Gamma = \left\{(q,p) \mid h(q,p) = 0, dh_q \dot{q} + dh_p \dot{p} = 0\right\}.    \end{equation}
 \end{defn}
@@
 \begin{defn}
-    A \vnhc \(h(q,p) = 0\) of order \(k\) is \textit{regular} if \MM{system~\eqref{eq:Hamiltonian_system:simply_actuated}} with output 
-    \(e = h(q,p)\)  \MM{has vector} relative degree \(\{2,2.\ldots,2\}\) everywhere on the
-    constraint manifold \(\Gamma\).
+A \vnhc \(h(q,p) = 0\) of order \(k\) is \textit{regular} if system~\eqref{eq:Hamiltonian_system:simply_actuated} with output \(e = h(q,p)\) has vector relative degree \(\{2,2.\ldots,2\}\) everywhere on the constraint manifold \(\Gamma\).
 \end{defn}
 
-\MM{The reader is referred to~\cite{Isi95} for the definition of vector relative degree.}
-The authors of
-\cite{nhvc_dynamic_walking,hybrid_zero_dynamics_bipedal_nhvcs}
-observed that relations which use only the unactuated conjugate momentum
-often have vector relative degree \(\{2,\ldots,2\}\).
-Indeed, we shall now prove that regular constraints cannot depend on the
-actuated momentum.
+The reader is referred to~\cite{Isi95} for the definition of vector relative degree.
+The authors of \cite{nhvc_dynamic_walking,hybrid_zero_dynamics_bipedal_nhvcs} observed that relations which use only the unactuated conjugate momentum often have vector relative degree \(\{2,\ldots,2\}\).
+Indeed, we shall now prove that regular constraints cannot depend on the actuated momentum.
 
@@
 
-Under additional mild conditions (see \cite{vhcs_for_el_systems}), \MM{the constraint manifold associated with } a regular \vnhc of
-order \(k\) can be \MM{asymptotically} stabilized by the input-output feedback linearizing
-controller
+Under additional mild conditions (see \cite{vhcs_for_el_systems}), the constraint manifold associated with a regular \vnhc of order \(k\) can be asymptotically stabilized by the input-output feedback linearizing controller
 \begin{equation}\label{eqn:stabilizing-controller}
@@
 
-\textbf{Comparison with existing \MM{virtual constraint} literature}: Horn \etal provide the constrained
+\textbf{Comparison with existing virtual constraint literature}: Horn \etal provide the constrained
 dynamics for \vnhcs in \cite{nhvc_incline_walking}.
